### 2012-10-18 inst/examples/gaussian-process-basics.Rmd
added script with basic gaussian process example of regression

`ro comment=NA, message=FALSE, cache.path="gpb/", tidy=FALSE or`

```{r include=FALSE}
sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
short_sha <- gsub("(^.{10}).*", "\\1", sha)
date <- format(Sys.time(), "%Y-%m-%d")
fig.path = paste("../../../assets/figures/", date, "-", short_sha, "-", sep="")
opts_chunk$set(fig.path=fig.path)
theme_publish <- theme_set(theme_bw(14))
theme_publish <- theme_update(
        # panel.grid.major=theme_blank(),panel.grid.minor=theme_blank(),
        legend.title=theme_blank(), 
        panel.background = theme_rect(fill = "transparent",colour = NA), 
        plot.background = theme_rect(fill = "transparent",colour = NA))
```

Basic regression in Gaussian processes  
---------------------------------------

Working out coded examples for basic Gaussian process regression using R.  I've just read through the first few chapters of [Rasmussen & Williams (2006)](http://www.GaussianProcess.org/gpml), this implements the examples discussed in Chapter 2.1-2.5.

<!-- Also motivated by [James Keirstead](http://www.jameskeirstead.ca/r/gaussian-process-regression-with-r/), with cleaner code.  -->

Required R libraries (for multivariate normal, also for plotting):

```{r results="hide"}
require(MASS)
require(reshape2)
require(ggplot2)
```

Set a seed for repeatable plots

```{r}
set.seed(12345)
```

Define the points at which we want to compute the function values (x values of the prediction points or test points), and the scale parameter for the covariance function $\ell=1$

```{r}
x_predict <- seq(-5,5,len=50)
l <- 1
```

We will use the squared exponential (also called radial basis or Gaussian, though it is not this that gives Gaussian process it's name; any covariance function would do) as the covariance function, $\operatorname{cov}(X_i, X_j) = \exp\left(-\frac{(X_i - X_j)^2}{2 \ell ^ 2}\right)$

```{r}
SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
cov <- function(X, Y) outer(X, Y, SE, l)
```

```{r}
COV <- cov(x_predict, x_predict)
```
Generate a number of functions from the process

```{r}
values <- mvrnorm(3, rep(0, length=length(x_predict)), COV)
```

Reshape the data into long (tidy) form, listing x value, y value, and sample number
```{r}
dat <- data.frame(x=x_predict, t(values))
dat <- melt(dat, id="x")
head(dat)
```

Plot the result

```{r}
fig2a <- ggplot(dat,aes(x=x,y=value)) +
  geom_rect(xmin=-Inf, xmax=Inf, ymin=-2, ymax=2, fill="grey80") +
  geom_line(aes(group=variable)) +   theme_bw() +
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
  xlab("input, x")
fig2a
```

### Posterior distribution given the data

Imagine we have some data,

```{r}
obs <- data.frame(x = c(-4, -3, -1,  0,  2),
                  y = c(-2,  0,  1,  2, -1))
```

In general we aren't interested in drawing from the prior, but want to include information from the data as well.  We want the joint distribution of the observed values and the prior is:

$$\begin{pmatrix} y_{\textrm{obs}} \\ y_{\textrm{pred}} \end{pmatrix} \sim \mathcal{N}\left( \mathbf{0}, \begin{bmatrix} cov(X_o,X_o) & cov(X_o, X_p) \\ cov(X_p,X_o) & cov(X_p, X_p) \end{bmatrix} \right)$$

## No observation noise

Assuming the data are known without error and conditioning on the data, and given $x \sim \mathcal{N}(0, cov(X_o, X_o))$, then the conditional probability of observing our data is easily solved by exploiting the nice properties of Gaussians,

$$x|y \sim \mathcal{N}\left(E,C)$$
$$E = cov(X_p, X_o) cov(X_o,X_o)^{-1} y$$
$$C= cov(X_p, X_p) - cov(X_p, X_o) cov(X_o,X_o)^{-1} cov(X_o, X_p) \right)$$

(We use `solve(M)` which with no second argument will simply invert the matrix `M`, but should use the Cholsky decomposition instead for better numerical stability)

```{r}
cov_xx_inv <- solve(cov(obs$x, obs$x))
Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% obs$y
Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)
```

Now lets take 3 random samples from the posterior distribution,
```{r}
values <- mvrnorm(3, Ef, Cf)
```

and plot our solution (mean, 2 standard deviations, and the ranom samples.)

```{r}
dat <- data.frame(x=x_predict, t(values))
dat <- melt(dat, id="x")

fig2b <- ggplot(dat,aes(x=x,y=value)) +
  geom_ribbon(data=NULL, 
              aes(x=x_predict, y=Ef, ymin=(Ef-2*sqrt(diag(Cf))), ymax=(Ef+2*sqrt(diag(Cf)))),
              fill="grey80") +
  geom_line(aes(color=variable)) + #REPLICATES
  geom_line(data=NULL,aes(x=x_predict,y=Ef), size=1) + #MEAN
  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
  xlab("input, x")
fig2b
```

Additive noise
--------------

In general the model may have process error, and rather than observe the deterministic map $f(x)$ we only observe $y = f(x) + \varepsilon$.  Let us assume for the moment that $\varepsilon$ are independent, normally distributed random variables with variance $\sigma_n^2$.

```{r}
sigma.n <- 0.25
cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma.n^2 * diag(1, length(obs$x)))
Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% obs$y
Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)
```

Now lets take 3 random samples from the posterior distribution,
```{r}
values <- mvrnorm(3, Ef, Cf)
```

and plot

```{r}
dat <- data.frame(x=x_predict, t(values))
dat <- melt(dat, id="x")

fig2c <- ggplot(dat,aes(x=x,y=value)) +
  geom_ribbon(data=NULL, 
              aes(x=x_predict, y=Ef, ymin=(Ef-2*sqrt(diag(Cf))), ymax=(Ef+2*sqrt(diag(Cf)))),
              fill="grey80") + # Var
  geom_line(aes(color=variable)) + #REPLICATES
  geom_line(data=NULL,aes(x=x_predict,y=Ef), size=1) + #MEAN
  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
  xlab("input, x")
fig2c
```

Note that unlike the previous case, the posterior no longer collapses completely around the neighborhood of the test points.

We can also compute the likelihood (and marginal likelihood over the prior) of the data directly from the inferred multivariate normal distribution, which can allow us to tune the hyperparameters such as the characteristic length scale $\ell$ and the observation noise $\sigma_n$.  The most obvious approach would be to do so by maximum likelihood, giving point estimates of the hyper-parameters, though presumably we could be Bayesian about these as well.

### 2012-10-23 inst/examples/comparing-methods.Rmd
Example comparing a few different appoaches to the GP calculations, differ by some (possibly numerical?) error...

```{r}
require(MASS)
require(ggplot2)
require(kernlab)
```

Parameterization-specific

```{r}
X <- seq(-5,5,len=50)
obs <- data.frame(x = c(-4, -3, -1,  0,  2),
                  y = c(-2,  0,  1,  2, -1))
l <- 1
sigma_n <- 0.8
```

Radial basis function/Gaussian kernel:

```{r}
  SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
  cov <- function(X, Y) outer(X, Y, SE, l) 
```

Cholksy method

```{r}
  n <- length(obs$x)
  K <- cov(obs$x, obs$x)
  I <-  diag(1, n)

  L <- chol(K + sigma_n^2 * I)
  alpha <- solve(t(L), solve(L, obs$y))
  k_star <- cov(obs$x, X)
  Y <- t(k_star) %*% alpha
  v <- solve(L, k_star)
  Var <- cov(X,X) - t(v) %*% v
  loglik <- -.5 * t(obs$y) %*% alpha - sum(log(diag(L))) - n * log(2 * pi) / 2
```

Direct method 
```{r}
  cov_xx_inv <- solve(K + sigma_n^2*I)
  Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
  Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X)
```

```{r}
gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), fit=FALSE, scaled=FALSE, var=0.8)
y_p <- predict(gp, X)
```

Things that should be equivelent but aren't quite:

```{r}
ggplot(data.frame(x=X, Ef=Ef, Y=Y, y_p = y_p))+ geom_point(aes(x,Ef), col='red') + geom_point(aes(x,Y)) + geom_line(aes(x,y_p))
````

```{r}
cov_xx_inv %*% obs$y
alpha
alpha(gp)
```

### 2012-10-23 inst/examples/comparing-methods.Rmd
examples

@@
+`ro comment=NA, message=FALSE, cache.path="compare/", tidy=FALSE or`
+
+```{r include=FALSE}
+sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
+short_sha <- gsub("(^.{10}).*", "\\1", sha)
+date <- format(Sys.time(), "%Y-%m-%d")
+fig.path = paste("../../../assets/figures/", date, "-", short_sha, "-", sep="")
+opts_chunk$set(fig.path=fig.path)
+theme_publish <- theme_set(theme_bw(14))
+theme_publish <- theme_update(
+        # panel.grid.major=theme_blank(),panel.grid.minor=theme_blank(),
+        legend.title=theme_blank(), 
+        panel.background = theme_rect(fill = "transparent",colour = NA), 
+        plot.background = theme_rect(fill = "transparent",colour = NA))
+```
+
 ```{r}

### 2012-10-23 R/gp_fit.R
beginning to create basic functions for common tasks

@@ -1,0 +1,30 @@
+
+
+gp_fit <- function(){
+  
+  ## Parameterization-specific
+  SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
+  cov <- function(X, Y) outer(X, Y, SE, l) 
+   
+  
+  n <- length(obs$x)
+  K <- cov(obs$x, obs$x)
+  I <-  diag(1, n)
+  
+  
+  ## Cholksy method
+  L <- chol(K + sigma_n^2 * I)
+  alpha <- solve(t(L), solve(L, obs$y))
+  k_star <- cov(obs$x, X)
+  Y <- t(k_star) %*% alpha
+  v <- solve(L, k_star)
+  Var <- cov(X,X) - t(v) %*% v
+  loglik <- -.5 * t(obs$y) %*% alpha - sum(log(diag(L))) - n * log(2 * pi) / 2
+  
+  ## Direct method 
+  cov_xx_inv <- solve(K + sigma_n^2*I)
+  Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
+  Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X)
+  
+  return(list(mu = Y, Sigma = Var, loglik=loglik, Ef = Ef, Cf = Cf)
+  }

### 2012-10-24 inst/examples/compare-cholsky.Rmd
Examples comparing cholsky method, and simulated stock dynamics
Something seems to be going wrong with the Cholsky method.  Results
match fine over certain function spaces but poorly on others.  Funny,
since this is supposed to be numerically *more* stable.

Examples implemented that infer stock dynamics from a stochastic
Beverton Holt and stochastic May (1979) bistable model.  Results
look reasonable, though still need to utilize tuning for the
hyperparameters.

Also added basic package documentation and updated the gp_fit.R
function.  Function currently supports both cholksy and direct metho

`ro comment=NA, message=FALSE, tidy=FALSE or`

## Compare Cholsky and Direct methods

```{r}
gp_fit_direct <- function(obs, X, pars=c(sigma_n=1, l=1)){

  sigma_n <- pars["sigma_n"]
  l <- pars["l"]

  ## Parameterization-specific
  SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
  cov <- function(X, Y) outer(X, Y, SE, l)

  n <- length(obs$x)
  K <- cov(obs$x, obs$x)
  I <-  diag(1, n)

  ## Direct method 
  cov_xx_inv <- solve(K + sigma_n ^ 2 * I)
  Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
  Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X)
  llik <- -.5 * t(obs$y) %*% cov_xx_inv %*% obs$y - 0.5 * log(det(cov_xx_inv)) - n * log(2 * pi) / 2

  out <- list(mu = Ef, var = Cf, loglik=llik)
  class(out) = "gp"
  out
}
```

```{r}
gp_fit_cholsky <- function(obs, X, pars=c(sigma_n=1, l=1)){

  sigma_n <- pars["sigma_n"]
  l <- pars["l"]

  ## Parameterization-specific
  SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
  cov <- function(X, Y) outer(X, Y, SE, l)

  n <- length(obs$x)
  K <- cov(obs$x, obs$x)
  I <-  diag(1, n)

  ## Cholksy method -- Consider implementing in C
  L <- chol(K + sigma_n ^ 2 * I)
  alpha <- solve(t(L), solve(L, obs$y))
  loglik <- -.5 * t(obs$y) %*% alpha - sum(log(diag(L))) - n * log(2 * pi) / 2
  k_star <- cov(obs$x, X)
  mu <- t(k_star) %*% alpha
  v <- solve(L, k_star)
  var <- diag( cov(X,X) - t(v) %*% v )

  ## Pt by pt following Rassmussen & WIlliams, but there's no reason to do it this way
#  tmp <- lapply(X, function(x_i){  
#    k_star <- cov(obs$x, x_i)
#    Y <- t(k_star) %*% alpha
#    v <- solve(L, k_star)
#    Var <- cov(x_i,x_i) - t(v) %*% v
#    list(Y=Y, Var=Var)
#  })
#  mu <- sapply(tmp,`[[`, "Y")
#  var <- sapply(tmp,`[[`, "Var")

  out <- list(mu = mu, var = var, loglik=loglik)
  class(out) = "gp"
  out
}
```

Run the comparison

```{r}
X <- seq(-5,5,len=50)
obs <- data.frame(x = c(-4, -3, -1,  0,  2),
                  y = c(-2,  0,  1,  2, -1))
system.time(gp <- gp_fit_cholsky(obs, X, c(sigma_n=0.8, l=1)))
system.time(gp_d <- gp_fit_direct(obs, X, c(sigma_n=0.8, l=1)))
```

Plot the results

```{r}
require(reshape2)
df <- data.frame(x=X, y=gp$mu, ymin=(gp$mu-2*sqrt(gp$var)), ymax=(gp$mu+2*sqrt(gp$var)))
df2 <- data.frame(x=X, y=gp_d$mu, ymin=(gp_d$mu-2*sqrt(diag(gp_d$var))), ymax=(gp_d$mu+2*sqrt(diag(gp_d$var))))
df2 <- melt(df2, id="x")
require(ggplot2)
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") + 
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) + 
  geom_line(data=df2, aes(x,value,color=variable))
```

Looks fine for this data, but not for Reed example:

```{r libraries}
require(pdgControl)
require(ggplot2)
f <- BevHolt
p <- c(1.2,.02)
K <- (p[1]-1)/p[2] 
z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
x_grid = seq(0, 1.5 * K, length=100)
T <- 40
sigma_g <- 0.01
x <- numeric(T)
x[1] <- 1
for(t in 1:(T-1))
  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
plot(x)
```

Predict the function over the target grid

```{r}
obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
X <- x_grid
system.time(gp <- gp_fit_cholsky(obs, X, c(sigma_n=0.8, l=1)))
system.time(gp_d <- gp_fit_direct(obs, X, c(sigma_n=0.8, l=1)))
```

Plot the results

```{r}
require(reshape2)
df <- data.frame(x=X, y=gp$mu, ymin=(gp$mu-2*sqrt(gp$var)), ymax=(gp$mu+2*sqrt(gp$var)))
df2 <- data.frame(x=X, y=gp_d$mu, ymin=(gp_d$mu-2*sqrt(diag(gp_d$var))), ymax=(gp_d$mu+2*sqrt(diag(gp_d$var))))
df2 <- melt(df2, id="x")
require(ggplot2)
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") + 
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) + 
  geom_line(data=df2, aes(x,value,color=variable))
```

### 2012-10-24 inst/examples/reed-example.Rmd
Examples comparing cholsky method, and simulated stock dynamics
Something seems to be going wrong with the Cholsky method.  Results
match fine over certain function spaces but poorly on others.  Funny,
since this is supposed to be numerically *more* stable.

Examples implemented that infer stock dynamics from a stochastic
Beverton Holt and stochastic May (1979) bistable model.  Results
look reasonable, though still need to utilize tuning for the
hyperparameters.

Also added basic package documentation and updated the gp_fit.R
function.  Function currently supports both cholksy and direct metho

Reed (1979) with Gaussian processes 
========================================================

```{r libraries}
require(pdgControl)
require(ggplot2)
```

Simulate some training data under a stochastic growth function with standard parameterization,

``` {r stateeqn}
f <- BevHolt
p <- c(1.5,.05)
K <- (p[1]-1)/p[2] 
```

Noise function 
```{r}
z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
```

Parameter definitions

```{r}
x_grid = seq(0, 1.5 * K, length=100)
T <- 40
sigma_g <- 0.1
x <- numeric(T)
x[1] <- 1
```

Simulation

```{r}
for(t in 1:(T-1))
  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)

plot(x)
```

Predict the function over the target grid

```{r}
obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
X <- x_grid
library(nonparametricbayes)
gp <- gp_fit(obs, X, c(sigma_n=.05, l=10))
```

Gaussian Process inference from this model.  True model shown in red.

```{r}
df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
true <- data.frame(x=X, y=sapply(X,f, 0, p))
require(ggplot2)
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
  geom_line(data=true, aes(x,y), col='red', lty=2)
```

## Another example using the May model

```{r}
f <- May
p <- c(r = .75, k = 10, a=1.2, H=1, Q = 3)
K <- 8 # approx
```

Model dynamics look like this:

```{r}
birth <- function(x) p["r"] * (1-  x / p["k"])
death <- function(x) p["a"] * x ^ (p["Q"] - 1) / (x ^ p["Q"] + p["H"])
df <- data.frame(x = x_grid, b = sapply(x_grid, birth), d = sapply(x_grid, death))
ggplot(df) + geom_line(aes(x, b), col = "blue") + geom_line(aes(x,d), col = "red")
````

Simulation

```{r}
x[1] = 2.5
for(t in 1:(T-1))
  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
plot(x)
```

Predict the function over the target grid

```{r}
obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
X <- x_grid
library(nonparametricbayes)
gp <- gp_fit(obs, X, c(sigma_n=.05, l=10))
```

Gaussian Process inference from this model

```{r}
df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
true <- data.frame(x=X, y=sapply(X,f, 0, p))
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
  geom_line(data=true, aes(x,y), col='red', lty=2)
```

### 2012-10-24 R/gp_fit.R

@@ -1,6 +1,28 @@
+#' gp_fit fits a Gaussian process to the observed data
+#' 
+#' API not yet stable! Arguments to this function will probably be updated to be much more generic
+#' @param obs a data frame of observations with columns obs$x and obs$y
+#' @param X the desired points over which to predict
+#' @param pars a named numeric specifying "sigma_n" for the (additive) and "l" for the covariance length scale
+#' @return a list with items "mu", the expected Y values at X (mean of the posterior Gaussian process), 
+#' Sigma, the covariance matrix for the posterior Gaussian process, and 
+#' "loglik", the log likelihood of observering the given data under the process, marginalized over the prior
+#' @details so far treates the prior as mean 0 and covariance given by cov(X)
+#' @export
+#' @examples 
+#' X <- seq(-5,5,len=50)
+#' obs <- data.frame(x = c(-4, -3, -1,  0,  2),
+#'                   y = c(-2,  0,  1,  2, -1))
+#' gp <- gp_fit(obs, X, c(sigma_n=0.8, l=1))
+#' 
+#' df <- data.frame(x=X, y=gp$mu, ymin=(gp$mu-2*sqrt((gp$var))), ymax=(gp$mu+2*sqrt((gp$var))))
+#' require(ggplot2)
+#' ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") + geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y))
 
-
-gp_fit <- function(){
+gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1)){
+  
+  sigma_n <- pars["sigma_n"]
+  l <- pars["l"]
   
   ## Parameterization-specific
   SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
@@ -12,19 +34,31 @@ gp_fit <- function(){
   I <-  diag(1, n)
   
   
-  ## Cholksy method
-  L <- chol(K + sigma_n^2 * I)
+  ## Cholksy method -- Consider implementing in C
+  L <- chol(K + sigma_n ^ 2 * I)
   alpha <- solve(t(L), solve(L, obs$y))
[…]

### 2012-10-24 inst/examples/reed-example.Rmd
Different choice of hyperparameters
Varying the hyperparameters has little direct impact around the data,
though with lower correlation lengths the influence of the prior
where there is no data becomes obvious.

@@
 require(ggplot2)
+opts_knit$set(upload.fun = socialR::flickr.url)
 ```
@@
 library(nonparametricbayes)
-gp <- gp_fit(obs, X, c(sigma_n=.05, l=10))
+gp <- gp_fit(obs, X, c(sigma_n=1, l=1))
 ```
@@
 X <- x_grid
-library(nonparametricbayes)
-gp <- gp_fit(obs, X, c(sigma_n=.05, l=10))
+gp <- gp_fit(obs, X, c(sigma_n=1, l=1))
 ```
@@
 ```
+
+## Further steps
+
+* is the posterior GP capturing the process noise?
+* Need to add optimization over hyperparameters

### 2012-10-24 inst/examples/reed-example.Rmd
estimating scaling length by mle -- way over?

@@
 
-## Further steps
+## Simple optimization over hyperparameters:
 
-* is the posterior GP capturing the process noise?
-* Need to add optimization over hyperparameters
+```{r}
+minusloglik <- function(par){
+  gp <- gp_fit(obs, X, par)
+  -gp$llik
+}
+par <- c(sigma_n=1, l=1)
+minusloglik(par)
+o <- optim(par, minusloglik)
+o
+hyperpars <- o$par
+```
+
+Yikes.  let's try 1-D optimization:
+
+```{r}
+minusloglik <- function(par){
+  gp <- gp_fit(obs, X, c(sigma_n=1, l=par))
+  -gp$llik
+}
+minusloglik(1)
+o <- optimize(minusloglik, c(0,100))
+o
+hyperpars <- c(sigma_n=1, l=o$objective)
+```
+
+Plot the optimal hyperparameter solution:
+
+```{r}
+gp <- gp_fit(obs, X, hyperpars)
+df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
+true <- data.frame(x=X, y=sapply(X,f, 0, p))
+ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
+  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
+  geom_line(data=true, aes(x,y), col='red', lty=2)
+```

### 2012-10-24 inst/examples/reed-example.Rmd
getting useless boundary solution when maximizing likelihood over sigma_n ... hmm

@@
 minusloglik <- function(par){
-  gp <- gp_fit(obs, X, c(sigma_n=1, l=par))
+  gp <- gp_fit(obs, X, c(sigma_n=par, l=1))
   -gp$llik
@@
 minusloglik(1)
-o <- optimize(minusloglik, c(0,100))
+o <- optimize(minusloglik, c(0,150))
 o
-hyperpars <- c(sigma_n=1, l=o$objective)
+hyperpars <- c(sigma_n=o$minimum, l=1)
 ```

### 2012-10-25 inst/examples/testing-gp-sdp.Rmd
testing out gp sdp, not sure why I do not valid policy functions

Quick trial SDP approaches with GP function
========================================================

```{r libraries}
require(pdgControl)
require(ggplot2)
opts_knit$set(upload.fun = socialR::flickr.url)
```

### Beverton-Holt function

Simulate some training data under a stochastic growth function with standard parameterization,

``` {r stateeqn}
f <- BevHolt
p <- c(1.5,.05)
K <- (p[1]-1)/p[2] 
```

Noise function 
```{r}
z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
```

Parameter definitions

```{r}
x_grid = seq(0, 1.5 * K, length=100)
T <- 40
sigma_g <- 0.1
x <- numeric(T)
x[1] <- 1
```

Simulation

```{r}
for(t in 1:(T-1))
  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
```

Predict the function over the target grid

```{r}
obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
X <- x_grid
library(nonparametricbayes)
gp <- gp_fit(obs, X, c(sigma_n=1, l=1))
```

Gaussian Process inference from this model.  True model shown in red.

```{r}
df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
true <- data.frame(x=X, y=sapply(X,f, 0, p))
require(ggplot2)
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
  geom_line(data=true, aes(x,y), col='red', lty=2)
```

## Stochastic Dynamic programming solution based on the posterior Gaussian process:

Define a transition matrix $F$ from the Gaussian process, giving the probability of going from state $x_t$ to $x_{t+1}$.
We already have the Gaussian process mean and variance predicted for each point $x$ on our grid, so this is simply:

```{r}
V <- sqrt(diag(gp$Cf))
F <- sapply(x_grid, function(x) dnorm(x, gp$Ef, V))
F <- t(apply(F,1,function(x) x/sum(x))) # normalize
h_grid <- x_grid
```

True $f(x)$

```{r}
mu <- sapply(x_grid, f, 0, p)
F_true <- sapply(x_grid, function(x) dnorm(x, mu, sigma_g))
F_true <- t(apply(F_true,1,function(x) x/sum(x))) # normalize
```

Somewhat silly way to adjust the matrix by harvest level:
```{r}
n <- length(h_grid)
per_harvest <- function(F,n)
  lapply(0:(n-1), function(i){
  d <- diag(1, nrow=(n-i))
  top <- matrix(0, nrow=i, ncol=n)
  top[,1] <- 1
  side <- matrix(0, nrow=(n-i), ncol=i)
  out <- F %*% rbind(top, cbind(d,side))
  t(apply(out,1,function(x) x/sum(x))) # normalize
})

```

True F is okay but not quite right:

```{r}
matrices <- per_harvest(F_true,n)
opt <- find_dp_optim(matrices, x_grid, h_grid, 20, 0, profit, delta=.01)
plot(opt$D[,1])
```

whoops: gp inferred F isn't working:

```{r}
matrices <- per_harvest(F,n)
opt <- find_dp_optim(matrices, x_grid, h_grid, 20, 0, profit, delta=.01)
plot(opt$D[,1])
```

Old-school calculation method:

```{r}
profit <- profit_harvest(price=1, c0 = 0, c1=0)
pdfn <- function(P, s) dlnorm(P, 0, s)
matrices <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g, pdfn)
opt <- find_dp_optim(matrices, x_grid=x_grid, h_grid=h_grid, OptTime=20, xT=0, profit=profit, delta=.05)
plot(opt$D[,1])
```

### 2012-10-25 inst/examples/reed-example.Rmd
Reed example showing basic use of a GP in SDP

--- 2012-10-24 inst/examples/reed-example.Rmd
@@
 
+### Beverton-Holt function
+
 Simulate some training data under a stochastic growth function with standard parameterization,
@@
 
-Noise function 
-```{r}
-z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
-```
-
 Parameter definitions
@@
 ```{r}
-x_grid = seq(0, 1.5 * K, length=100)
+x_grid = seq(0, 1.5 * K, length=101)
 T <- 40
@@
 
+Noise function, profit function
+```{r}
+z_g <- function() rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
+profit <- profit_harvest(1,0,0)
+```
+
 Simulation
@@
 for(t in 1:(T-1))
-  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
-
-plot(x)
+  x[t+1] = z_g() * f(x[t], h=0, p=p)
 ```
@@
 library(nonparametricbayes)
-gp <- gp_fit(obs, X, c(sigma_n=1, l=1))
+gp <- gp_fit(obs, X, c(sigma_n=.5, l=1.5))
 ```
@@
 
-## Another example using the May model
+## Stochastic Dynamic programming solution based on the posterior Gaussian process:
 
 ```{r}
-f <- May
-p <- c(r = .75, k = 10, a=1.2, H=1, Q = 3)
-K <- 8 # approx
+rownorm <- function(M) t(apply(M, 1, function(x) x/sum(x)))
+h_grid <- x_grid
 ```
 
-Model dynamics look like this:
+Define a transition matrix $F$ from the Gaussian process, giving the probability of going from state $x_t$ to $x_{t+1}$.
+We already have the Gaussian process mean and variance predicted for each point $x$ on our grid, so this is simply:
 
 ```{r}
-birth <- function(x) p["r"] * (1-  x / p["k"])
-death <- function(x) p["a"] * x ^ (p["Q"] - 1) / (x ^ p["Q"] + p["H"])
-df <- data.frame(x = x_grid, b = sapply(x_grid, birth), d = sapply(x_grid, death))
-ggplot(df) + geom_line(aes(x, b), col = "blue") + geom_line(aes(x,d), col = "red")
+V <- sqrt(diag(gp$Cf))
+matrices_gp <- lapply(h_grid, function(h){
+  F <- sapply(x_grid, function(x) dnorm(x, gp$Ef-h, V))
+  F <- rownorm(F)
+})
+```
+
+True $f(x)$
+
+```{r}
+matrices_true <- lapply(h_grid, function(h){
+  mu <- sapply(x_grid, f, h, p)
+  F_true <- sapply(x_grid, function(x) dnorm(x, mu, sigma_g))
+  F_true <- rownorm(F_true)
+})
+```
+
+Calculate the policy function using the true F:
+
+```{r}
+opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
+```
+
+Calculate using the inferred GP
+
+```{r}
+opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
+```
+
+```{r}
+require(reshape2)
+policies <- melt(data.frame(stock=x_grid, GP = x_grid[opt_gp$D[,1]], Exact = x_grid[opt_true$D[,1]]), id="stock")
+q1 <- ggplot(policies, aes(stock, stock - value, color=variable)) + geom_point() + xlab("stock size") + ylab("escapement") 
+q1
+```
+
+We can see what happens when we attempt to manage a stock using this:
+
+``` {r othernoise}
+z_g <- function() rlnorm(1,0, sigma_g)
 ````
 
-Simulation
+``` {r  simulate }
+set.seed(1)
+sim_gp <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_gp$D, z_g, profit=profit)
+set.seed(1)
+sim_true <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_true$D, z_g, profit=profit)
+````
+
+``` {r  simplot}
+df <- data.frame(time = sim_gp$time, stock_gp = sim_gp$fishstock, stock_true = sim_true$fishstock, harvest_gp = sim_gp$harvest, havest_true = sim_true$harvest)
+df <- melt(df, id="time")
+ggplot(df) + geom_line(aes(time,value, color=variable))
+````
+
+Total Profits
 
 ```{r}
-x[1] = 2.5
-for(t in 1:(T-1))
-  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
-plot(x)
+sum(sim_gp$profit)
+sum(sim_true$profit)
 ```
-
-Predict the function over the target grid
-
-```{r}
-obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
-X <- x_grid
-gp <- gp_fit(obs, X, c(sigma_n=1, l=1))
-```
-
-Gaussian Process inference from this model
-
-```{r}
-df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
-true <- data.frame(x=X, y=sapply(X,f, 0, p))
-ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
-  geom_line(data=true, aes(x,y), col='red', lty=2)
-```
-
-## Further steps
-
-* is the posterior GP capturing the process noise?
-* Need to add optimization over hyperparameters

### 2012-10-25 inst/examples/may79-example.Rmd
Code for running the Reed example with May's bistable model instead of a BH model

GP Example using the May (1979) bistable model
==============================================

Use my custom libraries and the `ggplot` plotting tools

```{r libraries}
require(nonparametric-bayes)
require(pdgControl)
require(ggplot2)
opts_knit$set(upload.fun = socialR::flickr.url)
```

```{r}
f <- May
p <- c(r = .75, k = 10, a=1.3, H=1, Q = 3)
K <- 8 # approx
```

Model dynamics look like this:

```{r}
birth <- function(x) p["r"] * (1-  x / p["k"])
death <- function(x) p["a"] * x ^ (p["Q"] - 1) / (x ^ p["Q"] + p["H"])
df <- data.frame(x = x_grid, b = sapply(x_grid, birth), d = sapply(x_grid, death))
ggplot(df) + geom_line(aes(x, b), col = "blue") + geom_line(aes(x,d), col = "red")
````

Noise function, profit function
```{r}
z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
profit <- profit_harvest(1,0,0)
```

Parameter definitions

```{r}
x_grid = seq(0, 1.5 * K, length=101)
T <- 40
sigma_g <- 0.1
x <- numeric(T)
x[1] <- 1
```

Simulation

```{r}
x[1] = 2.5
for(t in 1:(T-1))
  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
plot(x)
```

Predict the function over the target grid

```{r}
obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
X <- x_grid
gp <- gp_fit(obs, X, c(sigma_n=1, l=1))
```

Gaussian Process inference from this model

```{r}
df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
true <- data.frame(x=X, y=sapply(X,f, 0, p))
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
  geom_line(data=true, aes(x,y), col='red', lty=2)
```

### 2012-10-26 inst/examples/testing-gp-sdp.Rmd
exploring hyperparameter estimation, etc

@@
 Quick trial SDP approaches with GP function
-========================================================
+==============================================
 
@@
 require(pdgControl)
+require(nonparametricbayes)
 require(ggplot2)
@@
 for(t in 1:(T-1))
-  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
+  x[t+1] = z_g() * f(x[t], h=0, p=p)
 ```
@@
 
-## Stochastic Dynamic programming solution based on the posterior Gaussian process:
-
-Define a transition matrix $F$ from the Gaussian process, giving the probability of going from state $x_t$ to $x_{t+1}$.
-We already have the Gaussian process mean and variance predicted for each point $x$ on our grid, so this is simply:
+## Optimization over hyperparameters:
 
 ```{r}
-V <- sqrt(diag(gp$Cf))
-F <- sapply(x_grid, function(x) dnorm(x, gp$Ef, V))
-F <- t(apply(F,1,function(x) x/sum(x))) # normalize
-h_grid <- x_grid
+minusloglik <- function(par){
+  gp <- gp_fit(obs, X, par)
+  -gp$llik
+}
+par <- c(sigma_n=1, l=1)
+minusloglik(par)
+o <- optim(par, minusloglik)
+o
+hyperpars <- o$par
 ```
 
-True $f(x)$
+**Yikes.**
+
+let's try 1-D optimization:
 
 ```{r}
-mu <- sapply(x_grid, f, 0, p)
-F_true <- sapply(x_grid, function(x) dnorm(x, mu, sigma_g))
-F_true <- t(apply(F_true,1,function(x) x/sum(x))) # normalize
+minusloglik <- function(par){
+  gp <- gp_fit(obs, X, c(sigma_n=par, l=1))
+  -gp$llik
+}
+minusloglik(1)
+o <- optimize(minusloglik, c(0,150))
+o
+hyperpars <- c(sigma_n=o$minimum, l=1)
+```
+
+**Yikes, that shouldn't happen either**
+
+Plot the optimal hyperparameter solution:
+
+```{r}
+gp <- gp_fit(obs, X, hyperpars)
+df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
+true <- data.frame(x=X, y=sapply(X,f, 0, p))
+ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
+  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
+  geom_line(data=true, aes(x,y), col='red', lty=2)
+```
+
+### Misc ways to calc harvest matrices
+
+Old-school calculation method:
+
+```{r}
+profit <- profit_harvest(price=1, c0 = 0, c1=0)
+pdfn <- function(P, s) dlnorm(P, 0, s)
+sdp_mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g, pdfn)
+opt <- find_dp_optim(sdp_mat, x_grid=x_grid, h_grid=h_grid, OptTime=20, xT=0, profit=profit, delta=.05)
+plot(opt$D[,1])
 ```
@@
 Somewhat silly way to adjust the matrix by harvest level:
+
 ```{r}
 n <- length(h_grid)
-per_harvest <- function(F,n)
+silly <- function(F,n)
   lapply(0:(n-1), function(i){
@@
 })
-
 ```
 
-True F is okay but not quite right:
+binning function, not quite right:
 
 ```{r}
-matrices <- per_harvest(F_true,n)
-opt <- find_dp_optim(matrices, x_grid, h_grid, 20, 0, profit, delta=.01)
-plot(opt$D[,1])
+bin <- function(x, grid_centers){ 
+  match <- which.min(abs(grid_centers - x)) 
+  out <- integer(length(grid_centers))
+  out[match] <- 1
+  out
+  }
+per_harvest <- function(F,n){
+  lapply(h_grid, function(h){
+    t(sapply(x_grid, function(x) bin(x - h, x_grid)))
+  })
+}
 ```
-
-whoops: gp inferred F isn't working:
-
-```{r}
-matrices <- per_harvest(F,n)
-opt <- find_dp_optim(matrices, x_grid, h_grid, 20, 0, profit, delta=.01)
-plot(opt$D[,1])
-```
-
-Old-school calculation method:
-
-```{r}
-profit <- profit_harvest(price=1, c0 = 0, c1=0)
-pdfn <- function(P, s) dlnorm(P, 0, s)
-matrices <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g, pdfn)
-opt <- find_dp_optim(matrices, x_grid=x_grid, h_grid=h_grid, OptTime=20, xT=0, profit=profit, delta=.05)
-plot(opt$D[,1])
-```

### 2012-11-09 inst/examples/comparing-methods.Rmd
updated examples using sequential approach

@@
 X <- seq(-5,5,len=50)
-obs <- data.frame(x = c(-4, -3, -1,  0,  2),
-                  y = c(-2,  0,  1,  2, -1))
+obs <- data.frame(x = c(-4, -3, -1,  0,  2, 3),
+                  y = c(-2,  0,  1,  2, -1, -1))
+#x <- -5:5
+#obs <- data.frame(x = -5:5, y = sin(x) + rnorm(length(x),sd=.1))
 l <- 1
@@
 Direct method
-```{r}
+
+``` {r}
   cov_xx_inv <- solve(K + sigma_n^2*I)
@@
 
-```{r}
+  Direct sequential method, avoids matrix inverse instability
+``` {r}
+ef <- numeric(length(X))
+cf <- matrix(0, nrow=length(X), ncol=length(X))
+for(i in 1:length(obs$x)){
+  ef <- ef + cov(X, obs$x[i]) * obs$y[i] / as.numeric(cov(obs$x[i],obs$x[i]) + sigma_n^2)
+  cf <- cf + cov(X, X) - cov(X, obs$x[i])  %*% cov(obs$x[i], X) / as.numeric(cov(obs$x[i],obs$x[i]) + sigma_n^2)
+}
+```
+
+``` {r}
+ggplot(data.frame(x=X, Ef=Ef, ef=ef))+ geom_point(aes(x,Ef), col='red') + geom_line(aes(x,ef))
+````
+
+kernlab method
+
+``` {r}
 gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), fit=FALSE, scaled=FALSE, var=0.8)
@@
 
-Things that should be equivelent but aren't quite:
+mlegp method
 
-```{r}
-ggplot(data.frame(x=X, Ef=Ef, Y=Y, y_p = y_p))+ geom_point(aes(x,Ef), col='red') + geom_point(aes(x,Y)) + geom_line(aes(x,y_p))
+``` {r}
+require(mlegp)
+out <- mlegp(obs$x, obs$y, nugget = 1)
+y_m <- predict(out, as.matrix(X))
+```
+
+Compare these results:
+
+``` {r}
+require(reshape2)
+df <- data.frame(x = X, Ef = Ef, Y = Y, y_p = y_p, ef = ef, y_m = y_m)
+df <- melt(df, id = "x")
+ggplot(df)+ geom_jitter(aes(x, value, color = variable)) + geom_point(data = obs, aes(x,y))
 ````
+
+Compare alpha from Cholesky and alpha from `kernlab` to $K(x,x)^{-1} \vec y$
 

### 2012-11-09 inst/examples/gaussian-process-basics.Rmd
updated examples using sequential approach

@@
 
-$$x|y \sim \mathcal{N}\left(E,C)$$
+$$x|y \sim \mathcal{N}(E,C)$$
 $$E = cov(X_p, X_o) cov(X_o,X_o)^{-1} y$$
-$$C= cov(X_p, X_p) - cov(X_p, X_o) cov(X_o,X_o)^{-1} cov(X_o, X_p) \right)$$
+$$C= cov(X_p, X_p) - cov(X_p, X_o) cov(X_o,X_o)^{-1} cov(X_o, X_p)$$
 
@@
 ```{r}
-sigma.n <- 0.25
+sigma.n <- 0.3
 cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma.n^2 * diag(1, length(obs$x)))
@@
 dat <- melt(dat, id="x")
+Ey <- data.frame(x=x_predict, y=(Ef), ymin=(Ef-2*sqrt(diag(Cf))), ymax=(Ef+2*sqrt(diag(Cf))))
 
 fig2c <- ggplot(dat,aes(x=x,y=value)) +
-  geom_ribbon(data=NULL, 
-              aes(x=x_predict, y=Ef, ymin=(Ef-2*sqrt(diag(Cf))), ymax=(Ef+2*sqrt(diag(Cf)))),
-              fill="grey80") + # Var
+  geom_ribbon(data=Ey, aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
   geom_line(aes(color=variable)) + #REPLICATES
-  geom_line(data=NULL,aes(x=x_predict,y=Ef), size=1) + #MEAN
+  geom_line(data=Ey,aes(x=x,y=y), size=1) + #MEAN
   geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
@@
   xlab("input, x") 
-fig2c
+fig2c + opts(panel.background = theme_rect(fill = "transparent"), 
+        plot.background = theme_rect(fill = "transparent"))
 ```
@@
 The most obvious approach would be to do so by maximum likelihood, giving point estimates of the hyper-parameters, though presumably we could be Bayesian about these as well.
+
+<!-- Continuing... 2012-10-22 -->
+
+The likelihood is given by 
+$$\log(p(y | X)) = -\tfrac{1}{2} \mathbf{y}^T (K + \sigma_n^2 \mathbf{I})^{-1} y  - \tfrac{1}{2} \log\left| K + \sigma_n^2 \mathbf{I} \right| - \tfrac{n}{2}\log 2 \pi$$
+
+Which is:
+
+```{r}
+
+#' observation pairs, obs$x, obs$x, 
+#' SE covariance between two observations 
+minusloglik <- function(pars){
+
+  l <- pars[1]
+  sigma.n <- pars[2]
+
+  cov <- function(X, Y) outer(X, Y, SE, l)
+  I <- diag(1, length(obs$x))
+  K <- cov(obs$x, obs$x)
+
+  0.5 * t(obs$y) %*% solve(K + sigma.n^2 * I) %*% obs$y +
+    log(det(K + sigma.n^2*I)) +
+    length(obs$y) * log(2 * pi) / 2
+  }
+```
+
+```{r}
+pars <- c(l=1, sigma.n=1)
+o <- optim(pars, minusloglik)
+o$par
+```
+
+## A comparison to built-in R utilities
+
+Some basic Gaussian process tools are provided in the `kernlab` package,
+
+```{r}
+library(kernlab)
+```
+
+through the `gausspr` function,
+
+```{r}
+gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), fit=FALSE, scaled=FALSE, var=0.8)
+```
+
+Where `rbfdot` indicates the Gaussian (Radial basis function) kernel we use above, with parameter `sigma=1/(2l^2)` (note that gausspr defines the kernel parameter differently, see:
+
+```{r}
+gp@kernelf@.Data
+```
+note this takes vectors `x`, `y` and returns a scalar, not the covaraiance matrix).
+The function will try to estimate hyperparameters such as `sigma` if they are not given unless `fit=FALSE`.
+The data is also scaled to 0 mean, unit variance unless `scaled=FALSE` is set.
+
+We can then use the estimated process to infer values on a specified grid through `predict`, e.g. for the example above,
+
+```{r}
+y_p <- predict(gp, x_predict)
+dat2 <- data.frame(x = x_predict, variable = "Y_p", value = y_p)
+```
+
+```{r}
+ggplot(dat2, aes(x, value)) + geom_point() + geom_line(data=Ey,aes(x=x,y=y), size=1) 
+```
+
+### Fitting hyperparameters by maximum likelihood
+
+If we do not specify `fit=false` then `gausspr` will estimate the hyperparamters by maximum likelihood:
+
+```{r}
+gp_fit <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), scaled=FALSE, var=0.8)
+```

### 2012-11-09 R/gp_fit.R

@@ -34,28 +34,55 @@ gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1)){
   I <-  diag(1, n)
   
   
-  ## Cholksy method -- Consider implementing in C
-  L <- chol(K + sigma_n ^ 2 * I)
-  alpha <- solve(t(L), solve(L, obs$y))
-  loglik <- -.5 * t(obs$y) %*% alpha - sum(log(diag(L))) - n * log(2 * pi) / 2
-  tmp <- lapply(X, function(x_i){  ## Must be a clever way to vectorize this...
-    k_star <- cov(obs$x, x_i)
-    Y <- t(k_star) %*% alpha
+  ## Cholesky not-exactly-sequential method -- Consider implementing in C
+  if(0){
+    L <- chol(K + sigma_n ^ 2 * I)
+    alpha <- solve(t(L), solve(L, obs$y))
+    loglik <- -.5 * t(obs$y) %*% alpha - sum(log(diag(L))) - n * log(2 * pi) / 2
+    tmp <- lapply(X, function(x_i){  ## Must be a clever way to vectorize this...
+      k_star <- cov(obs$x, x_i)
+      Y <- t(k_star) %*% alpha
+      v <- solve(L, k_star)
+      Var <- cov(x_i,x_i) - t(v) %*% v
+      list(Y=Y, Var=Var)
+    })
+    mu <- sapply(tmp,`[[`, "Y")
+    var <- sapply(tmp,`[[`, "Var")
+  }
+  
+  ## Cholesky simultaneous method.  
+  if(0){
+    L <- chol(K + sigma_n ^ 2 * I)
+    alpha <- solve(t(L), solve(L, obs$y))
+    loglik <- -.5 * t(obs$y) %*% alpha - sum(log(diag(L))) - n * log(2 * pi) / 2
+    k_star <- cov(obs$x, X)
+    mu <- t(k_star) %*% alpha
     v <- solve(L, k_star)
-    Var <- cov(x_i,x_i) - t(v) %*% v
-    list(Y=Y, Var=Var)
-  })
-  mu <- sapply(tmp,`[[`, "Y")
-  var <- sapply(tmp,`[[`, "Var")
+    var <- diag( cov(X,X) - t(v) %*% v )
+  }
   
   ## Direct method 
[…]

### 2012-11-09 inst/examples/testing-gp-sdp.Rmd
updated examples using sequential approach
show figures

@@
 require(ggplot2)
-opts_knit$set(upload.fun = socialR::flickr.url)
+opts_knit$set(upload.fun = socialR::notebook.url)
 ```
@@
 
+```{r}
+#escapement <- Vectorize(function(i,j) pmax(0, x_grid[i] - h_grid[D[j,(Tmax-t+1)]]))
+#S <- outer(1:n_x, 1:n_h, escapement)
+```
+
 Old-school calculation method:

### 2012-11-09 inst/examples/testing-gp-sdp.Rmd
remove old misc ways to calc harvest matrix

@@
 ```
-
-### Misc ways to calc harvest matrices
-
-```{r}
-#escapement <- Vectorize(function(i,j) pmax(0, x_grid[i] - h_grid[D[j,(Tmax-t+1)]]))
-#S <- outer(1:n_x, 1:n_h, escapement)
-```
-
-Old-school calculation method:
-
-```{r}
-profit <- profit_harvest(price=1, c0 = 0, c1=0)
-pdfn <- function(P, s) dlnorm(P, 0, s)
-sdp_mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g, pdfn)
-opt <- find_dp_optim(sdp_mat, x_grid=x_grid, h_grid=h_grid, OptTime=20, xT=0, profit=profit, delta=.05)
-plot(opt$D[,1])
-```
-
-Somewhat silly way to adjust the matrix by harvest level:
-
-```{r}
-n <- length(h_grid)
-silly <- function(F,n)
-  lapply(0:(n-1), function(i){
-  d <- diag(1, nrow=(n-i))
-  top <- matrix(0, nrow=i, ncol=n)
-  top[,1] <- 1
-  side <- matrix(0, nrow=(n-i), ncol=i)
-  out <- F %*% rbind(top, cbind(d,side))
-  t(apply(out,1,function(x) x/sum(x))) # normalize
-})
-```
-
-binning function, not quite right:
-
-```{r}
-bin <- function(x, grid_centers){ 
-  match <- which.min(abs(grid_centers - x)) 
-  out <- integer(length(grid_centers))
-  out[match] <- 1
-  out
-  }
-per_harvest <- function(F,n){
-  lapply(h_grid, function(h){
-    t(sapply(x_grid, function(x) bin(x - h, x_grid)))
-  })
-}
-```

### 2012-11-09 inst/examples/compare-cholesky.Rmd
updated comparing-methods plots

--- 2012-10-24 inst/examples/compare-cholsky.Rmd
@@
 
-## Compare Cholsky and Direct methods
+## Compare Cholesky and Direct methods
+```{r settings}
+opts_knit$set(upload.fun = socialR::notebook.url)
+```
 
@@
 
-  ## Cholksy method -- Consider implementing in C
+  ## Cholesky method -- Consider implementing in C
   L <- chol(K + sigma_n ^ 2 * I)
@@
 
-  ## Pt by pt following Rassmussen & WIlliams, but there's no reason to do it this way
+## Pt by pt following Rassmussen & Williams, but there's no reason to do it this way
 #  tmp <- lapply(X, function(x_i){  

### 2012-11-09 inst/examples/comparing-methods.Rmd
updated comparing-methods plots

@@
 `ro comment=NA, message=FALSE, cache.path="compare/", tidy=FALSE or`
-
-```{r include=FALSE}
-sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
-short_sha <- gsub("(^.{10}).*", "\\1", sha)
-date <- format(Sys.time(), "%Y-%m-%d")
-fig.path = paste("../../../assets/figures/", date, "-", short_sha, "-", sep="")
-opts_chunk$set(fig.path=fig.path)
-theme_publish <- theme_set(theme_bw(14))
-theme_publish <- theme_update(
-        # panel.grid.major=theme_blank(),panel.grid.minor=theme_blank(),
-        legend.title=theme_blank(), 
-        panel.background = theme_rect(fill = "transparent",colour = NA), 
-        plot.background = theme_rect(fill = "transparent",colour = NA))
-```
 
@@
 require(kernlab)
+opts_knit$set(upload.fun = socialR::notebook.url)
 ```
@@
 
-Cholksy method
+Cholesky method
 

### 2012-11-09 inst/examples/hyperparameter-optimization.Rmd
hyperparameter optimization testing now as a separate file

--- 2012-11-09 inst/examples/testing-gp-sdp.Rmd
@@
-Quick trial SDP approaches with GP function
-==============================================
+## Optimization over hyperparameters:
 
-```{r libraries}
-require(pdgControl)
-require(nonparametricbayes)
-require(ggplot2)
-opts_knit$set(upload.fun = socialR::notebook.url)
+```{r results="hide"}
+knit("beverton_holt_data.Rmd")
 ```
-
-### Beverton-Holt function
-
-Simulate some training data under a stochastic growth function with standard parameterization,
-
-``` {r stateeqn}
-f <- BevHolt
-p <- c(1.5,.05)
-K <- (p[1]-1)/p[2] 
-```
-
-Parameter definitions
-
-```{r}
-x_grid = seq(0, 1.5 * K, length=101)
-T <- 40
-sigma_g <- 0.1
-x <- numeric(T)
-x[1] <- 1
-```
-
-Noise function, profit function
-```{r}
-z_g <- function() rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
-profit <- profit_harvest(1,0,0)
-```
-
-Simulation
-
-```{r}
-for(t in 1:(T-1))
-  x[t+1] = z_g() * f(x[t], h=0, p=p)
-```
-
-Predict the function over the target grid
-
-```{r}
-obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
-X <- x_grid
-library(nonparametricbayes)
-gp <- gp_fit(obs, X, c(sigma_n=.5, l=1.5))
-```
-
-Gaussian Process inference from this model.  True model shown in red.
-
-```{r}
-df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
-true <- data.frame(x=X, y=sapply(X,f, 0, p))
-require(ggplot2)
-ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
-  geom_line(data=true, aes(x,y), col='red', lty=2)
-```
-
-## Optimization over hyperparameters:
 

### 2012-11-09 inst/examples/testing-gp-sdp.Rmd
Testing sequential mode on BH data

@@
 
-```{r libraries}
-require(pdgControl)
-require(nonparametricbayes)
-require(ggplot2)
+``` {r results="hide"}
+knit("beverton_holt_data.Rmd")
 opts_knit$set(upload.fun = socialR::notebook.url)
@@
 
-### Beverton-Holt function
+Radial basis function/Gaussian kernel:
 
-Simulate some training data under a stochastic growth function with standard parameterization,
-
-``` {r stateeqn}
-f <- BevHolt
-p <- c(1.5,.05)
-K <- (p[1]-1)/p[2] 
+```{r}
+  SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
+  cov <- function(X, Y) outer(X, Y, SE, l) 
 ```
 
-Parameter definitions
+Cholesky method
 
 ```{r}
-x_grid = seq(0, 1.5 * K, length=101)
-T <- 40
-sigma_g <- 0.1
-x <- numeric(T)
-x[1] <- 1
+  n <- length(obs$x)
+  K <- cov(obs$x, obs$x)
+  I <-  diag(1, n)
+
+  L <- chol(K + sigma_n^2 * I)
+  alpha <- solve(t(L), solve(L, obs$y))
+  k_star <- cov(obs$x, X)
+  Y <- t(k_star) %*% alpha
+  v <- solve(L, k_star)
+  Var <- cov(X,X) - t(v) %*% v
+  loglik <- -.5 * t(obs$y) %*% alpha - sum(log(diag(L))) - n * log(2 * pi) / 2
 ```
 
-Noise function, profit function
-```{r}
-z_g <- function() rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
-profit <- profit_harvest(1,0,0)
+Direct method
+
+``` {r}
+  cov_xx_inv <- solve(K + sigma_n^2*I)
+  Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
+  Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X)
 ```
 
-Simulation
-
-```{r}
-for(t in 1:(T-1))
-  x[t+1] = z_g() * f(x[t], h=0, p=p)
+  Direct sequential method, avoids matrix inverse instability
+``` {r}
+ef <- numeric(length(X))
+cf <- matrix(0, nrow=length(X), ncol=length(X))
+for(i in 1:length(obs$x)){
+  ef <- ef + cov(X, obs$x[i]) * obs$y[i] / as.numeric(cov(obs$x[i],obs$x[i]) + sigma_n^2)
+  cf <- cf + cov(X, X) - cov(X, obs$x[i])  %*% cov(obs$x[i], X) / as.numeric(cov(obs$x[i],obs$x[i]) + sigma_n^2)
+}
 ```
 
-Predict the function over the target grid
+``` {r}
+ggplot(data.frame(x=X, Ef=Ef, ef=ef))+ geom_point(aes(x,Ef), col='red') + geom_line(aes(x,ef))
+````
 
-```{r}
-obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
-X <- x_grid
-library(nonparametricbayes)
-gp <- gp_fit(obs, X, c(sigma_n=.5, l=1.5))
+kernlab method
+
+``` {r}
+gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), fit=FALSE, scaled=FALSE, var=0.8)
+y_p <- predict(gp, X)
 ```
 
-Gaussian Process inference from this model.  True model shown in red.
+mlegp method
 
-```{r}
-df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
-true <- data.frame(x=X, y=sapply(X,f, 0, p))
-require(ggplot2)
-ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
-  geom_line(data=true, aes(x,y), col='red', lty=2)
+``` {r}
+require(mlegp)
+out <- mlegp(obs$x, obs$y, nugget = 1)
+y_m <- predict(out, as.matrix(X))
 ```
 
-## Optimization over hyperparameters:
+Compare these results:
 
-```{r}
-minusloglik <- function(par){
-  gp <- gp_fit(obs, X, par)
-  -gp$llik
-}
-par <- c(sigma_n=1, l=1)
-minusloglik(par)
-o <- optim(par, minusloglik)
-o
-hyperpars <- o$par
-```
-
-**Yikes.**
-
-let's try 1-D optimization:
-
-```{r}
-minusloglik <- function(par){
-  gp <- gp_fit(obs, X, c(sigma_n=par, l=1))
-  -gp$llik
-}
-minusloglik(1)
-o <- optimize(minusloglik, c(0,150))
-o
-hyperpars <- c(sigma_n=o$minimum, l=1)
-```
-
-**Yikes, that shouldn't happen either**
-
-Plot the optimal hyperparameter solution:
-
-```{r}
-gp <- gp_fit(obs, X, hyperpars)
-df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
-true <- data.frame(x=X, y=sapply(X,f, 0, p))
-ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
-  geom_line(data=true, aes(x,y), col='red', lty=2)
-```
+``` {r}
+require(reshape2)
+df <- data.frame(x = X, Ef = Ef, Y = Y, y_p = y_p, ef = ef, y_m = y_m)
+df <- melt(df, id = "x")
+ggplot(df)+ geom_jitter(aes(x, value, color = variable)) + geom_point(data = obs, aes(x,y))
+````

### 2012-11-09 inst/examples/testing-gp-sdp.Rmd
label methods explicitly

@@
 require(reshape2)
-df <- data.frame(x = X, Ef = Ef, Y = Y, y_p = y_p, ef = ef, y_m = y_m)
+df <- data.frame(x = X, direct = Ef, Cholesky = Y, kernlab = y_p, sequential = ef, mlegp = y_m)
 df <- melt(df, id = "x")

### 2012-11-15 inst/examples/gaussian-process-basics.Rmd
add plot to basics-example

@@
 ```{r include=FALSE}
-sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
-short_sha <- gsub("(^.{10}).*", "\\1", sha)
-date <- format(Sys.time(), "%Y-%m-%d")
-fig.path = paste("../../../assets/figures/", date, "-", short_sha, "-", sep="")
-opts_chunk$set(fig.path=fig.path)
-theme_publish <- theme_set(theme_bw(14))
-theme_publish <- theme_update(
-        # panel.grid.major=theme_blank(),panel.grid.minor=theme_blank(),
-        legend.title=theme_blank(), 
-        panel.background = theme_rect(fill = "transparent",colour = NA), 
-        plot.background = theme_rect(fill = "transparent",colour = NA))
+opts_knit$set(upload.fun = socialR::notebook.url)
 ```

### 2012-11-16 inst/examples/hyperparameter-optimization.Rmd
example hyperparameter optimization not working. check likelihood?
Updated hyperparameter-optimization example
Shows that we can estimate l appropriately but fails to get sigma_n; instead estimating the upper boundary.

@@
 ## Optimization over hyperparameters:
+
+```{r results="hide"}
+opts_knit$set(upload.fun = socialR::notebook.url)
+```
+
+Static definition of gp function
+
+```{r}
+gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1)){
+  sigma_n <- pars["sigma_n"]
+  l <- pars["l"]
+
+  SE <- function(Xi,Xj, l=l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
+  cov <- function(X, Y) outer(X, Y, SE, l)
+
+  n <- length(obs$x)
+  K <- cov(obs$x, obs$x)
+  I <-  diag(1, n)
+
+    cov_xx_inv <- solve(K + sigma_n ^ 2 * I)
+    Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
+    Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X)
+    llik <- -.5 * t(obs$y) %*% cov_xx_inv %*% obs$y - 0.5 * log(det(cov_xx_inv)) - n * log(2 * pi) / 2
+
+  out <- list(Ef = Ef, Cf = Cf, llik = llik, obs=obs, X=X, pars=pars, K=K, inv=cov_xx_inv)
+  class(out) = "gpfit"
+  out
+}
+
+plot.gpfit <- function(gp){
+  dat <- data.frame(x=gp$X, y=(gp$Ef), ymin=(gp$Ef-2*sqrt(diag(gp$Cf))), ymax=(gp$Ef+2*sqrt(diag(gp$Cf))))
+  ggplot(dat) +
+  geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") +
+  geom_line(aes(x=x,y=y), size=1) + 
+  geom_point(data=gp$obs,aes(x=x,y=y)) + 
+  labs(title=paste("llik =", prettyNum(gp$llik)))
+}
+```
+
+Estimate a GP from data with given hyperparameters, then redraw data from an actual multivariate normal.  Presumably optimal hyperparameters might be near the given values then.
+
+```{r}
+require(MASS)
+X <- seq(-5,5,len=20)
+obs <- data.frame(x = c(-4, -3, -1,  0,  2, 3),
+                  y = c(-2,  0,  1,  2, -1, -1))
+
+fit <- gp_fit(obs, X, pars=c("sigma_n"=.3, "l"=.6))
+gp_dat <- mvrnorm(1, fit$Ef, fit$Cf)
+obs <- data.frame(x=X, y=gp_dat)
+X <- seq(-5,5,len=50)
+```
+
+Example plot
+
+```{r}
+plot.gpfit(gp_fit(obs, X, pars=c("sigma_n"=.3, "l"=.6)))
+```
+
+estimates non-boundary solution of `l`:
+
+```{r}
+minusloglik <- function(par){
+  gp <- gp_fit(obs, X, c(sigma_n=.3, l=par))
+  -gp$llik
+}
+minusloglik(1)
+o <- optimize(minusloglik, c(0,40))
+o
+
+```
+
+But makes `sigma_n` arbitrarily large (boundary solution).
+
+```{r}
+minusloglik <- function(par){
+  gp <- gp_fit(obs, X, c(sigma_n=par, l=.6))
+  -gp$llik
+}
+minusloglik(1)
+o <- optimize(minusloglik, c(0,40))
+o
+
+```
+
+```{r}
+gp_k <- gausspr(obs$x, obs$y, kernel="rbfdot")
+gp_k
+```
+
+Convert to `l`, `sigma=1/(2l^2)`
+
+```{r}
+1/sqrt(2*as.numeric(gp_k@kernelf@kpar))
+```
+
+Weirdly inconsistent?
+
+```{r}
+gp_k <- gausspr(obs$x, obs$y, kernel="rbfdot")
+1/sqrt(2*as.numeric(gp_k@kernelf@kpar))
+gp_k <- gausspr(obs$x, obs$y, kernel="rbfdot")
+1/sqrt(2*as.numeric(gp_k@kernelf@kpar))
+```
+
+### a second example, using the beverton holt data:
 
@@
 minusloglik <- function(par){
+  if(any(par < 0))
+    Inf
+  else {
   gp <- gp_fit(obs, X, par)
@@
   }
+}
 par <- c(sigma_n=1, l=1)
 minusloglik(par)
-o <- optim(par, minusloglik)
+#o <- optim(par, minusloglik)
+o <- optim(par, minusloglik, method="L", lower=c(0,0), upper=c(20,20))
 o
@@
 minusloglik <- function(par){
-  gp <- gp_fit(obs, X, c(sigma_n=par, l=1))
+  gp <- gp_fit(obs, X, c(sigma_n=par, l=5.7))
   -gp$llik
@@
 minusloglik(1)
-o <- optimize(minusloglik, c(0,150))
+o <- optimize(minusloglik, c(0,50))
 o

### 2012-11-16 R/gp_fit.R
Stick with direct method.
Current version of sequential method is not correct.

@@ -25,9 +25,8 @@ gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1)){
   l <- pars["l"]
   
   ## Parameterization-specific
-  SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
+  SE <- function(Xi,Xj, l=l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
   cov <- function(X, Y) outer(X, Y, SE, l) 
-   
   
   n <- length(obs$x)
   K <- cov(obs$x, obs$x)
@@ -62,13 +61,14 @@ gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1)){
   }
   
   ## Direct method 
-  if(1){
+  {
     cov_xx_inv <- solve(K + sigma_n ^ 2 * I)
     Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
     Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X)
     llik <- -.5 * t(obs$y) %*% cov_xx_inv %*% obs$y - 0.5 * log(det(cov_xx_inv)) - n * log(2 * pi) / 2
   }
   
+  if(1){
   ## Direct sequential method, avoids matrix inverse instability
   ef <- numeric(length(X))
   cf <- matrix(0, nrow=length(X), ncol=length(X))
@@ -79,13 +79,20 @@ gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1)){
     cf <- cf + cov(X, X) - cov(X, obs$x[i])  %*% cov(obs$x[i], X) / S
     llik <- llik - 0.5 * obs$y[i] ^ 2 /  S - 0.5 * log(S) - n * log(2 * pi) / 2
   }
+  }
   
-  
-  out <- list(Ef = ef, Cf = cf, llik = llik)
+  out <- list(Ef = Ef, Cf = Cf, llik = llik, obs=obs, X=X, K=K, pars)
   class(out) = "gpfit"
   out
   }
 
 
-
+plot.gpfit <- function(gp){
+  dat <- data.frame(x=gp$X, y=(gp$Ef), ymin=(gp$Ef-2*sqrt(diag(gp$Cf))), ymax=(gp$Ef+2*sqrt(diag(gp$Cf))))
+  ggplot(dat) +
[…]

### 2012-11-16 inst/examples/sequential-method.Rmd
Attempting to write the sequential method algorithm.
Not correct at the moment, need to understand the recursion properly.

`ro comment=NA, message=FALSE, cache.path="compare/", tidy=FALSE or`

## Sequential method

```{r}
require(MASS)
require(ggplot2)
require(kernlab)
opts_knit$set(upload.fun = socialR::notebook.url)
```

Parameterization-specific

```{r}
X <- 1
obs <- data.frame(x = c(-3, -1, 3),
                  y = c(0,  1, -1))
#x <- -5:5
#obs <- data.frame(x = -5:5, y = sin(x) + rnorm(length(x),sd=.1))
l <- 1
sigma_n <- 0.
```

Radial basis function/Gaussian kernel:

```{r}
  SE <- function(Xi,Xj, l=1) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
  cov <- function(X, Y) outer(X, Y, SE, l) 
```

### Cholesky method

```{r}
  n <- length(obs$x)
  K <- cov(obs$x, obs$x)
  I <-  diag(1, n)

  L <- chol(K + sigma_n^2 * I)
  alpha <- solve(t(L), solve(L, obs$y))
  k_star <- cov(obs$x, X)
  Y <- t(k_star) %*% alpha
  v <- solve(L, k_star)
  Var <- cov(X,X) - t(v) %*% v
  loglik <- -.5 * t(obs$y) %*% alpha - sum(log(diag(L))) - n * log(2 * pi) / 2
```

### Direct method

``` {r}
  cov_xx_inv <- solve(K + sigma_n^2*I)
  Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
  Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X)
```

### Direct sequential method, avoids matrix inverse instability

``` {r}
ef <- numeric(length(X))
cf <- matrix(0, nrow=length(X), ncol=length(X))

A <- as.numeric( cov(obs$x[1], obs$x[1]) )
mu <- obs$y[1]

for(i in 2:length(obs$x)){
  mu <- obs$y[i] + cov(obs$x[i], obs$x[1:(i-1)]) %*% (obs$y[1:(i-1)] - mu) / (A + sigma_n^2)
  A <- as.numeric( cov(obs$x[i], obs$x[i]) - cov(obs$x[i], obs$x[1:(i-1)]) %*% cov(obs$x[1:(i-1)], obs$x[i]) / (A + sigma_n^2) )
}

ef <- cov(X, obs$x) %*% (obs$y - mu) / A
cf <- cov(X,X) - cov(X, obs$x) %*% cov(obs$x, X) / A

```

See if this iteration scheme is correct?

```{r}
V1 <- cov(x[1], x[1])
V2 <- cov(x[2], x[2]) - cov(x[2], x[1]) %*% solve(cov(x[1], x[1])) %*% cov(x[1], x[2])
V2b <- cov(x[2], x[2]) - cov(x[2], x[1]) %*% cov(x[1], x[2]) / V1
V3 <- cov(x[3], x[3]) - cov(x[3], x[1:2]) %*% solve(cov(x[1:2], x[1:2])) %*% cov(x[1:2], x[3])
V3b <- cov(x[3], x[3]) - cov(x[3], x[1:2]) %*% cov(x[1:2], x[3]) / V2
c(V3, V3b)
```

### 2012-11-16 inst/examples/comparing-methods.Rmd
Attempted to fix the sequential method example.
Still incorrect.  See sequential-method.Rmd

@@
 l <- 1
-sigma_n <- 0.8
+sigma_n <- 0.
 ```
@@
 ```{r}
-  SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
+  SE <- function(Xi,Xj, l=1) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
   cov <- function(X, Y) outer(X, Y, SE, l) 
@@
 
-Cholesky method
+### Cholesky method
 
@@
 
-Direct method
+### Direct method
 
@@
 
-  Direct sequential method, avoids matrix inverse instability
+### Direct sequential method, avoids matrix inverse instability
+
 ``` {r}
@@
 cf <- matrix(0, nrow=length(X), ncol=length(X))
-for(i in 1:length(obs$x)){
-  ef <- ef + cov(X, obs$x[i]) * obs$y[i] / as.numeric(cov(obs$x[i],obs$x[i]) + sigma_n^2)
-  cf <- cf + cov(X, X) - cov(X, obs$x[i])  %*% cov(obs$x[i], X) / as.numeric(cov(obs$x[i],obs$x[i]) + sigma_n^2)
+
+A <- as.numeric( cov(obs$x[1], obs$x[1]) )
+mu <- obs$y[1]
+
+for(i in 2:length(obs$x)){
+  mu <- obs$y[i] + cov(obs$x[i], obs$x[1:(i-1)]) %*% (obs$y[1:(i-1)] - mu) / (A + sigma_n^2)
+  A <- as.numeric( cov(obs$x[i], obs$x[i]) - cov(obs$x[i], obs$x[1:(i-1)]) %*% cov(obs$x[1:(i-1)], obs$x[i]) / (A + sigma_n^2) )
 }
+
+ef <- cov(X, obs$x) %*% (obs$y - mu) / A
+cf <- cov(X,X) - cov(X, obs$x) %*% cov(obs$x, X) / A
+
 ```
@@
 
-kernlab method
+### kernlab method
 
@@
 
-mlegp method
+These are the same as `y_p`, explicitly showing the $E_y = K(x, x_i) \alpha$ multiplication
+
+``` {r}
+y_mult <-  kernelMult(kernelf(gp), X, xmatrix(gp), as.matrix(alpha(gp))) 
+y_se <- sapply(X, function(x) cov(x, obs$x) %*% alpha(gp))
+rbind(as.numeric(y_p), as.numeric(y_mult), as.numeric(y_se))[,1:10]
+```
+
+Which demonstrates the kernel trick `alpha = cov_xx_inv %*% obs$y`
+
+### mlegp method
 
@@
 require(reshape2)
-df <- data.frame(x = X, Ef = Ef, Y = Y, y_p = y_p, ef = ef, y_m = y_m)
+df <- data.frame(x = X, direct = Ef, Cholesky = Y, kernlab = y_p, sequential = ef)
 df <- melt(df, id = "x")

### 2012-11-16 inst/examples/question.Rmd
Question on how to get the covariance matrix back from the kernlab gausspr method.

`ro comment=NA, message=FALSE, cache.path="gpb/", tidy=FALSE or`

```{r include=FALSE}
opts_knit$set(upload.fun = socialR::notebook.url)
```

To provide a minimal working example for comparison: If I were to do this manually (following [Rasmussen & Williams (2006)](http://www.GaussianProcess.org/gpml) Chapter 2), I would do:

Consider we have the observed `x,y` points and `x` values where we desire predicted `y` values:

```{r}
obs <- data.frame(x = c(-4, -3, -1,  0,  2),
                  y = c(-2,  0,  1,  2, -1))
x_predict <- seq(-5,5,len=50)
```

Use a radial basis kernel:

```{r}
SE <- function(Xi,Xj, l=1) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
cov <- function(X, Y) outer(X, Y, SE)
```

Calculate mean and variance:

```{r}
sigma.n <- 0.3
cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma.n^2 * diag(1, length(obs$x)))
Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% obs$y
Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)
```

## kernlab

I think I see how I get the equivalent expected values in kernlab:

```{r}
library(kernlab)
gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), fit=FALSE, scaled=FALSE, var=0.8)
Ef_kernlab <- predict(gp, x_predict)
```

but I don't see how I get the associated covariances?  Perhaps I have missed something in the documentation?

There are many cases where it would be nice to have access to the resulting Gaussian process, such as in generating plots as in Rasmussen and Williams:

```{r}
require(ggplot2)
dat <- data.frame(x=x_predict, y=(Ef), ymin=(Ef-2*sqrt(diag(Cf))), ymax=(Ef+2*sqrt(diag(Cf))))    
ggplot(dat) +
  geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
  geom_line(aes(x=x,y=y), size=1) + #MEAN
  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") + xlab("input, x") 
```

### 2012-11-17 inst/examples/question.Rmd
Illustrate the direct approach to calculating the covariance matrix from the kernlab package.

@@
                   y = c(-2,  0,  1,  2, -1))
-x_predict <- seq(-5,5,len=50)
+X <- seq(-5,5,len=50)
+```
+
+```{r results="hide"}
+knit("beverton_holt_data.Rmd")
 ```
@@
 ```{r}
-sigma.n <- 0.3
-cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma.n^2 * diag(1, length(obs$x)))
-Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% obs$y
-Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)
+sigma_n <- 0.3
+cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma_n^2 * diag(1, length(obs$x)))
+Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
+system.time(Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X))
+```
+
+Plot
+
+```{r}
+require(ggplot2)
+dat <- data.frame(x=X, y=(Ef), ymin=(Ef-2*sqrt(diag(Cf))), ymax=(Ef+2*sqrt(diag(Cf))))    
+ggplot(dat) +
+  geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
+  geom_line(aes(x=x,y=y), size=1) + #MEAN
+  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
+  scale_y_continuous(name="output, f(x)") + xlab("input, x") 
 ```
@@
 
+rename hyperparameters
+
+```{r}
+l = 1
+lengthscale = 1/(2*l^2)
+var = sigma_n^2
+```
+
 ```{r}
 library(kernlab)
-gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), fit=FALSE, scaled=FALSE, var=0.8)
-Ef_kernlab <- predict(gp, x_predict)
+#gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=0.5), fit=FALSE, scaled=FALSE, var=.09)
+gp <- gausspr(obs$x, obs$y, kernel="rbfdot", scaled=FALSE, var=.09)
+Ef_k <- predict(gp, X)
 ```
 
-but I don't see how I get the associated covariances?  Perhaps I have missed something in the documentation?
+Manually get the covariance.  First we compute $K(x, x)$
 
-There are many cases where it would be nice to have access to the resulting Gaussian process, such as in generating plots as in Rasmussen and Williams:
+```{r}
+K <- kernelMatrix(kernelf(gp), xmatrix(gp))
+```
+
+Then compute $(K(x,x) + \sigma_n^2 \mathbb{I})^{-1}$,
+
+```{r}
+Inv <- solve(K + diag(rep(gp@kcall$var, length = length(xmatrix(gp)))))
+```
+
+The covariance is $K(x_*, x_*) - K(x_*, x)(K(x, x) - \sigma_n^2\mathbb{I})^{-1}K(x,x_*)$.
+
+```{r}
+system.time(Cf_k <- kernelMatrix(kernelf(gp), as.matrix(X)) - kernelMatrix(kernelf(gp), as.matrix(X), xmatrix(gp)) %*% Inv %*% kernelMatrix(kernelf(gp), xmatrix(gp), as.matrix(X)))
+```
+
+Identical to computing the $K(x_*, x)(K(x, x) - \sigma_n^2\mathbb{I})^{-1}$ part first, multiplying the off-diagonal matrix block by the inverse
+
+```{r}
+system.time(Cf_k2 <- kernelMatrix(kernelf(gp), as.matrix(X)) - kernelMult(kernelf(gp), as.matrix(X), xmatrix(gp),  Inv) %*% kernelMatrix(kernelf(gp), xmatrix(gp), as.matrix(X)))
+```
+
+which is also faster.
+
+compare the two in plots
 
@@
 require(ggplot2)
-dat <- data.frame(x=x_predict, y=(Ef), ymin=(Ef-2*sqrt(diag(Cf))), ymax=(Ef+2*sqrt(diag(Cf))))    
-ggplot(dat) +
-  geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
-  geom_line(aes(x=x,y=y), size=1) + #MEAN
+dat_k <- data.frame(x=X, y=(Ef_k), ymin=(Ef_k-2*sqrt(diag(Cf_k))), ymax=(Ef_k+2*sqrt(diag(Cf_k))))    
+ggplot(dat_k) +
+  geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill=rgb(0,0,1,.5)) + # Var
+  geom_line(aes(x=x,y=y), size=1, col="blue") + #MEAN
   geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
-  scale_y_continuous(lim=c(-3,3), name="output, f(x)") + xlab("input, x") 
+  geom_ribbon(data=dat, aes(x=x,y=y, ymin=ymin, ymax=ymax), fill=rgb(1,0,0,.5)) + # Var
+  geom_line(data=dat, aes(x=x,y=y), size=1, col="red") + #MEAN
+  scale_y_continuous(name="output, f(x)") + xlab("input, x") 
 ```

### 2012-11-17 inst/examples/reed-example.Rmd
updated examples using sequential approach
SDP using kernlab

@@
 
-```{r libraries}
-require(pdgControl)
-require(ggplot2)
-opts_knit$set(upload.fun = socialR::flickr.url)
+```{r include=FALSE}
+opts_knit$set(upload.fun = socialR::notebook.url)
 ```
 
-### Beverton-Holt function
-
-Simulate some training data under a stochastic growth function with standard parameterization,
-
-``` {r stateeqn}
-f <- BevHolt
-p <- c(1.5,.05)
-K <- (p[1]-1)/p[2] 
+```{r}
+require(pdgControl)
+require(nonparametricbayes)
+require(ggplot2)
 ```
 
-Parameter definitions
-
-```{r}
-x_grid = seq(0, 1.5 * K, length=101)
-T <- 40
-sigma_g <- 0.1
-x <- numeric(T)
-x[1] <- 1
-```
-
-Noise function, profit function
-```{r}
-z_g <- function() rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
-profit <- profit_harvest(1,0,0)
-```
-
-Simulation
-
-```{r}
-for(t in 1:(T-1))
-  x[t+1] = z_g() * f(x[t], h=0, p=p)
+```{r results="hide"}
+knit("beverton_holt_data.Rmd")
 ```
@@
 ```{r}
-obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
-X <- x_grid
 library(nonparametricbayes)
-gp <- gp_fit(obs, X, c(sigma_n=.5, l=1.5))
+gp <- gp_fit(obs, X, c(sigma_n=1, l=1.5))
+plot.gpfit(gp)
 ```
 
-Gaussian Process inference from this model.  True model shown in red.
+Solution using `kernlab`
 
 ```{r}
-df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
-true <- data.frame(x=X, y=sapply(X,f, 0, p))
+library(kernlab)
+#gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=0.5), fit=FALSE, scaled=FALSE, var=.09)
+gp <- gausspr(obs$x, obs$y, kernel="rbfdot", scaled=FALSE, var=.01)
+Ef <- predict(gp, X)
+```
+
+Manually get the covariance, $K(x_*, x_*) - K(x_*, x)(K(x, x) - \sigma_n^2\mathbb{I})^{-1}K(x,x_*)$.
+
+```{r}
+var <- gp@kcall$var
+var <- 9
+Inv <- solve(kernelMatrix(kernelf(gp), xmatrix(gp)) + diag(rep(var, length = length(xmatrix(gp)))))
+Cf <- kernelMatrix(kernelf(gp), as.matrix(X)) - 
+  kernelMult(kernelf(gp), as.matrix(X), xmatrix(gp),  Inv) %*% 
+  kernelMatrix(kernelf(gp), xmatrix(gp), as.matrix(X))
+```
+
+```{r}
 require(ggplot2)
-ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
-  geom_line(data=true, aes(x,y), col='red', lty=2)
+dat <- data.frame(x=X, y=(Ef), ymin=(Ef-2*sqrt(diag(Cf))), ymax=(Ef+2*sqrt(diag(Cf))))    
+ggplot(dat) + 
+  geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") +
+  geom_line(aes(x=x,y=y), size=1) +
+  geom_point(data=obs,aes(x=x,y=y)) + 
+  scale_y_continuous(name="output, f(x)") + xlab("input, x") 
 ```
@@
 ```{r}
-V <- sqrt(diag(gp$Cf))
+V <- sqrt(diag(Cf))
 matrices_gp <- lapply(h_grid, function(h){
-  F <- sapply(x_grid, function(x) dnorm(x, gp$Ef-h, V))
+  F <- sapply(x_grid, function(x) dnorm(x, Ef-h, V))
   F <- rownorm(F)

### 2012-11-17 R/gp_fit.R
plot function added for gp_fit routine

@@ -87,6 +87,12 @@ gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1)){
   }
 
 
+#' Plot a Gaussian Process
+#' 
+#' Plot the mean (line) and 2 standard deviations (ribbon)
+#' @param gp a gpfit object from \code{\link{gp_fit}}
+#' @return a ggplot object of the plot
+#' @export
 plot.gpfit <- function(gp){
   dat <- data.frame(x=gp$X, y=(gp$Ef), ymin=(gp$Ef-2*sqrt(diag(gp$Cf))), ymax=(gp$Ef+2*sqrt(diag(gp$Cf))))
   ggplot(dat) +

### 2012-11-28 inst/examples/reed-example.Rmd
basic GP example in BH system, compared to Reed Optimum

@@
 
-```{r include=FALSE}
-opts_knit$set(upload.fun = socialR::notebook.url)
-```
-
-```{r}
+```{r graphing-options, include=FALSE}
 require(pdgControl)
@@
 require(ggplot2)
+opts_knit$set(upload.fun = socialR::notebook.url)
+opts_chunk$set(dev.args=list(bg="transparent"), comment=NA)
+theme_set(theme_bw())
+theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
+             plot.background = element_rect(fill = "transparent",colour = NA))
 ```
@@
 #gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=0.5), fit=FALSE, scaled=FALSE, var=.09)
-gp <- gausspr(obs$x, obs$y, kernel="rbfdot", scaled=FALSE, var=.01)
+gp <- gausspr(obs$x, obs$y, kernel="rbfdot", scaled=FALSE, var=sigma_g^2)
 Ef <- predict(gp, X)
@@
 
-```{r}
+```{r policy_plot}
 require(reshape2)
 policies <- melt(data.frame(stock=x_grid, GP = x_grid[opt_gp$D[,1]], Exact = x_grid[opt_true$D[,1]]), id="stock")
-q1 <- ggplot(policies, aes(stock, stock - value, color=variable)) + geom_point() + xlab("stock size") + ylab("escapement") 
-q1
+policy_plot <- ggplot(policies, aes(stock, stock - value, color=variable)) + geom_point() + xlab("stock size") + ylab("escapement") 
+policy_plot
 ```
@@
 df <- melt(df, id="time")
-ggplot(df) + geom_line(aes(time,value, color=variable))
+simplot <- ggplot(df) + geom_line(aes(time,value, color=variable))
+simplot
 ````

### 2012-11-28 inst/examples/reed-example.Rmd
Compare GP to Reed optimum for BH (successful run this time)

@@
 #gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=0.5), fit=FALSE, scaled=FALSE, var=.09)
-gp <- gausspr(obs$x, obs$y, kernel="rbfdot", scaled=FALSE, var=sigma_g^2)
+gp <- gausspr(obs$x, obs$y, kernel="rbfdot", scaled=FALSE, var = sigma_g^2)
 Ef <- predict(gp, X)

### 2012-11-28 inst/examples/reed-example.Rmd
Double the actual variance leads to terminal overfishing

--- 2012-11-28 inst/examples/reed-example.Rmd
@@
 #gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=0.5), fit=FALSE, scaled=FALSE, var=.09)
-gp <- gausspr(obs$x, obs$y, kernel="rbfdot", scaled=FALSE, var=sigma_g^2)
+gp <- gausspr(obs$x, obs$y, kernel="rbfdot", scaled=FALSE, var = 2 * sigma_g^2)
 Ef <- predict(gp, X)

### 2012-11-28 inst/examples/reed-example.Rmd
functionalized methods into R/
Different GP fitting routines can all be called with gp_fit and plotted with plot.gpfit.

transition matrices for GP can be estimated with gp_transition_matrices.

See example use in reed-example.R

--- 2012-11-28 inst/examples/reed-example.Rmd
@@
 require(ggplot2)
+library(kernlab)
 opts_knit$set(upload.fun = socialR::notebook.url)
-opts_chunk$set(dev.args=list(bg="transparent"), comment=NA)
+opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE)
 theme_set(theme_bw())
@@
 knit("beverton_holt_data.Rmd")
+true <- data.frame(x = X, y = sapply(X, f, 0, p))
 ```
@@
 ```{r}
-library(nonparametricbayes)
-gp <- gp_fit(obs, X, c(sigma_n=1, l=1.5))
-plot.gpfit(gp)
-```
-
-Solution using `kernlab`
-
-```{r}
-library(kernlab)
-#gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=0.5), fit=FALSE, scaled=FALSE, var=.09)
-gp <- gausspr(obs$x, obs$y, kernel="rbfdot", scaled=FALSE, var=sigma_g^2)
-Ef <- predict(gp, X)
-```
-
-Manually get the covariance, $K(x_*, x_*) - K(x_*, x)(K(x, x) - \sigma_n^2\mathbb{I})^{-1}K(x,x_*)$.
-
-```{r}
-var <- gp@kcall$var
-var <- 9
-Inv <- solve(kernelMatrix(kernelf(gp), xmatrix(gp)) + diag(rep(var, length = length(xmatrix(gp)))))
-Cf <- kernelMatrix(kernelf(gp), as.matrix(X)) - 
-  kernelMult(kernelf(gp), as.matrix(X), xmatrix(gp),  Inv) %*% 
-  kernelMatrix(kernelf(gp), xmatrix(gp), as.matrix(X))
-```
-
-```{r}
-require(ggplot2)
-dat <- data.frame(x=X, y=(Ef), ymin=(Ef-2*sqrt(diag(Cf))), ymax=(Ef+2*sqrt(diag(Cf))))    
-ggplot(dat) + 
-  geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") +
-  geom_line(aes(x=x,y=y), size=1) +
-  geom_point(data=obs,aes(x=x,y=y)) + 
-  scale_y_continuous(name="output, f(x)") + xlab("input, x") 
+gp <- gp_fit(obs, X, c(sigma_n=.2, l=10))
+plot.gpfit(gp, true)
 ```
@@
 ```{r}
-rownorm <- function(M) t(apply(M, 1, function(x) x/sum(x)))
-h_grid <- x_grid
-```
-
-Define a transition matrix $F$ from the Gaussian process, giving the probability of going from state $x_t$ to $x_{t+1}$.
-We already have the Gaussian process mean and variance predicted for each point $x$ on our grid, so this is simply:
-
-```{r}
-V <- sqrt(diag(Cf))
-matrices_gp <- lapply(h_grid, function(h){
-  F <- sapply(x_grid, function(x) dnorm(x, Ef-h, V))
-  F <- rownorm(F)
-})
-```
-
-True $f(x)$
-
-```{r}
-matrices_true <- lapply(h_grid, function(h){
-  mu <- sapply(x_grid, f, h, p)
-  F_true <- sapply(x_grid, function(x) dnorm(x, mu, sigma_g))
-  F_true <- rownorm(F_true)
-})
-```
-
-Calculate the policy function using the true F:
-
-```{r}
+matrices_gp <- gp_transition_matrix(gp)
+matrices_true <- f_transition_matrix(f, p, x_grid)
 opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
-```
-
-Calculate using the inferred GP
-
-```{r}
 opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
@@
 policies <- melt(data.frame(stock=x_grid, GP = x_grid[opt_gp$D[,1]], Exact = x_grid[opt_true$D[,1]]), id="stock")
-policy_plot <- ggplot(policies, aes(stock, stock - value, color=variable)) + geom_point() + xlab("stock size") + ylab("escapement") 
+policy_plot <- ggplot(policies, aes(stock, stock - value, color=variable)) +
+  geom_point() + xlab("stock size") + ylab("escapement") 
 policy_plot

### 2012-11-28 R/gp_fit.R

@@ -4,6 +4,8 @@
 #' @param obs a data frame of observations with columns obs$x and obs$y
 #' @param X the desired points over which to predict
 #' @param pars a named numeric specifying "sigma_n" for the (additive) and "l" for the covariance length scale
+#' @param method select the method to use
+#' @param out_var optionally force kernlab method to generate Cf using a different variance than sigma_n^2 assumed in the fit
 #' @return a list with items "mu", the expected Y values at X (mean of the posterior Gaussian process), 
 #' Sigma, the covariance matrix for the posterior Gaussian process, and 
 #' "loglik", the log likelihood of observering the given data under the process, marginalized over the prior
@@ -19,38 +21,22 @@
 #' require(ggplot2)
 #' ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") + geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y))
 
-gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1)){
+gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1), method=c("direct", "sequential", "kernlab", "cholesky"), fit = true, out_var=NULL, ...){
   
+  method <- match.arg(method)
   sigma_n <- pars["sigma_n"]
   l <- pars["l"]
   
   ## Parameterization-specific
   SE <- function(Xi,Xj, l=l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
   cov <- function(X, Y) outer(X, Y, SE, l) 
-  
   n <- length(obs$x)
   K <- cov(obs$x, obs$x)
   I <-  diag(1, n)
   
   
-  ## Cholesky not-exactly-sequential method -- Consider implementing in C
-  if(0){
-    L <- chol(K + sigma_n ^ 2 * I)
-    alpha <- solve(t(L), solve(L, obs$y))
-    loglik <- -.5 * t(obs$y) %*% alpha - sum(log(diag(L))) - n * log(2 * pi) / 2
[…]

### 2012-11-28 R/gp_plot.R
additional functions (from previous commit).

@@ -1,0 +1,19 @@
+
+#' Plot a Gaussian Process
+#' 
+#' Plot the mean (line) and 2 standard deviations (ribbon)
+#' @param gp a gpfit object from \code{\link{gp_fit}}
+#' @param true an optional data.frame with columns "x" and "y" containing target curve.  
+#' @return a ggplot object of the plot
+#' @export
+plot.gpfit <- function(gp, true=NULL){
+  dat <- data.frame(x=gp$X, y=(gp$Ef), ymin=(gp$Ef-2*sqrt(diag(gp$Cf))), ymax=(gp$Ef+2*sqrt(diag(gp$Cf))))
+  p <- ggplot(dat) +
+    geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
+    geom_line(aes(x=x,y=y), size=1) + #MEAN
+    geom_point(data=gp$obs,aes(x=x,y=y)) + 
+    labs(title=paste("llik =", prettyNum(gp$llik)))
+  if(!is.null(true))
+    p <- p + geom_line(data = true, aes(x, y), col = "red", lty = 2) 
+  p
+}

### 2012-11-28 R/gp_trasition_matrix.R

@@ -1,0 +1,39 @@
+  
+#' Determine the transition matrix given a Guassian process
+#' 
+#' @param gp a gpfit type object, a list with the GP mean gp$Ef, Covariance gp$Cf on grid gp$X
+#' @param h_grid the harvest grid over which to compute.  Assumed to match the stock grid gp$X if not given.
+#' @return a list of matrices corresponding to the state transition matrix at each harvest level h
+#' @export
+gp_transition_matrix <- function(gp, h_grid=NULL){
+  # consider taking x_grid, Ef, & Cf as arguments instead?
+  rownorm <- function(M) t(apply(M, 1, function(x) x/sum(x)))
+  x_grid <- gp$X
+  if(is.null(h_grid))
+    h_grid <- x_grid
+  V <- sqrt(diag(gp$Cf))
+  matrices_gp <- lapply(h_grid, function(h){
+    F_ <- sapply(x_grid, function(x) dnorm(x, gp$Ef-h, V))
+    F_ <- rownorm(F_)
+  })
+  matrices_gp
+}
+
+#' calculate transition matrix based on parameterized function
+#' 
+#' @param f the transtion function, a function of state x, harvest h, and parameters p
+#' @param p the parameters p of the transition function
+#' @param x_grid the grid over which to solve 
+#' @param h_grid the harvest grid over which to compute.  Assumed to match the stock grid gp$X if not given.
+#' @return a list of matrices corresponding to the state transition matrix at each harvest level h
+#' @export
+f_transition_matrix <- function(f, p, x_grid, h_grid=NULL){
+  if(is.null(h_grid))
+    h_grid <- x_grid
+  matrices_F <- lapply(h_grid, function(h){
+    mu <- sapply(x_grid, f, h, p)
+    F_true <- sapply(x_grid, function(x) dnorm(x, mu, sigma_g))
+    F_true <- rownorm(F_true)
+  })
+  matrices_F
+}

### 2012-11-28 R/gp_trasition_matrix.R
added helper function to internal namespace (bugfix)

@@ -7,7 +7,6 @@
 #' @export
 gp_transition_matrix <- function(gp, h_grid=NULL){
   # consider taking x_grid, Ef, & Cf as arguments instead?
-  rownorm <- function(M) t(apply(M, 1, function(x) x/sum(x)))
   x_grid <- gp$X
   if(is.null(h_grid))
     h_grid <- x_grid
@@ -37,3 +36,7 @@ f_transition_matrix <- function(f, p, x_grid, h_grid=NULL){
   })
   matrices_F
 }
+
+
+# internal helper function
+rownorm <- function(M) t(apply(M, 1, function(x) x/sum(x)))

### 2012-11-28 inst/examples/reed-example.Rmd
Example direct method (using package functions from this repo) with preset hyperparameters.

@@
 require(nonparametricbayes)
+require(reshape2)
 require(ggplot2)
@@
 ```{r}
-gp <- gp_fit(obs, X, c(sigma_n=.2, l=10))
+gp <- gp_fit(obs, X, c(sigma_n=.1, l=8))
 plot.gpfit(gp, true)
@@
 ```{r}
+h_grid <- x_grid
 matrices_gp <- gp_transition_matrix(gp)
-matrices_true <- f_transition_matrix(f, p, x_grid)
+matrices_true <- f_transition_matrix(f, p, x_grid, h_grid)
 opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
@@
 ```{r policy_plot}
-require(reshape2)
 policies <- melt(data.frame(stock=x_grid, GP = x_grid[opt_gp$D[,1]], Exact = x_grid[opt_true$D[,1]]), id="stock")

### 2012-11-28 R/gp_fit.R
fixed typo

@@ -21,7 +21,7 @@
 #' require(ggplot2)
 #' ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") + geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y))
 
-gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1), method=c("direct", "sequential", "kernlab", "cholesky"), fit = true, out_var=NULL, ...){
+gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1), method=c("direct", "sequential", "kernlab", "cholesky"), fit = TRUE, out_var=NULL, ...){
   
   method <- match.arg(method)
   sigma_n <- pars["sigma_n"]

### 2012-11-28 R/gp_fit.R
fixed typo and add likelihood to kern method

@@ -26,6 +26,7 @@ gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1), method=c("direct", "sequentia
   method <- match.arg(method)
   sigma_n <- pars["sigma_n"]
   l <- pars["l"]
+  llik <- NA 
   
   ## Parameterization-specific
   SE <- function(Xi,Xj, l=l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
@@ -91,7 +92,7 @@ gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1), method=c("direct", "sequentia
   ## Kernlab method
   else if(method=="kernlab"){
   var <- sigma_n^2
-  if(fit)
+  if(!fit)
     gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), fit=FALSE, scaled=FALSE, var=var) # fixed length-scale
   else
     gp <- gausspr(obs$x, obs$y, kernel="rbfdot", scaled=FALSE, var = var)
@@ -105,6 +106,8 @@ gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1), method=c("direct", "sequentia
   Cf <- kernelMatrix(kernelf(gp), as.matrix(X)) - 
     kernelMult(kernelf(gp), as.matrix(X), xmatrix(gp),  Inv) %*% 
     kernelMatrix(kernelf(gp), xmatrix(gp), as.matrix(X))
+  llik <- -.5 * t(obs$y) %*% Inv %*% obs$y - 0.5 * log(det(Inv)) - n * log(2 * pi) / 2
+  
   }
   
   

### 2012-11-29 R/gp_fit.R
Reed example using tau not set to 1 and conditioning on going through (0,0)

@@ -21,7 +21,7 @@
 #' require(ggplot2)
 #' ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") + geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y))
 
-gp_fit <- function(obs, X, pars=c(sigma_n=1, tau=1, l=1), method=c("direct", "sequential", "kernlab", "cholesky"), fit = TRUE, out_var=NULL, ...){
+gp_fit <- function(obs, X, pars=c(sigma_n=1, tau=1, l=1), method=c("direct", "sequential", "kernlab", "cholesky", "conditional"), fit = TRUE, out_var=NULL, ...){
   
   method <- match.arg(method)
   sigma_n <- pars["sigma_n"]
@@ -56,6 +56,20 @@ gp_fit <- function(obs, X, pars=c(sigma_n=1, tau=1, l=1), method=c("direct", "se
     llik <- -.5 * t(obs$y) %*% cov_xx_inv %*% obs$y - 0.5 * log(det(cov_xx_inv)) - n * log(2 * pi) / 2
   }
   
+  
+  
+  ## Direct method, conditional on passing through 0,0 (quick hack)
+  else if(method=="conditional"){
+    obs <- data.frame(x=c(0, obs$x), y = c(0, obs$y))
+    n <- length(obs$x)
+    K <- cov(obs$x, obs$x)
+    sigmaI <-  sigma_n ^ 2 * diag(1, n)
+    sigmaI[1,1] <- 1e-5 # no noise about zero.  
+    cov_xx_inv <- solve(K + sigmaI)
+    Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
+    Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X)
+    llik <- -.5 * t(obs$y) %*% cov_xx_inv %*% obs$y - 0.5 * log(det(cov_xx_inv)) - n * log(2 * pi) / 2
+  }
   
   ### Sequential method, using function recursion
   else if(method=="sequential"){
@@ -110,12 +124,16 @@ gp_fit <- function(obs, X, pars=c(sigma_n=1, tau=1, l=1), method=c("direct", "se
[…]

### 2012-11-29 inst/examples/comparing-methods.Rmd
basic example comparing methods

@@
 require(kernlab)
+require(nonparametricbayes)
 opts_knit$set(upload.fun = socialR::notebook.url)
+opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE)
+theme_set(theme_bw())
+theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
+             plot.background = element_rect(fill = "transparent",colour = NA))
 ```
@@
                   y = c(-2,  0,  1,  2, -1, -1))
-#x <- -5:5
-#obs <- data.frame(x = -5:5, y = sin(x) + rnorm(length(x),sd=.1))
 l <- 1
-sigma_n <- 0.
+sigma_n <- 0.1
 ```
 
-Radial basis function/Gaussian kernel:
-
-```{r}
-  SE <- function(Xi,Xj, l=1) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
-  cov <- function(X, Y) outer(X, Y, SE, l) 
+```{r results="hide", eval=FALSE, include=FALSE}
+#knit("beverton_holt_data.Rmd")
+#true <- data.frame(x = X, y = sapply(X, f, 0, p))
 ```
@@
 ```{r}
+  SE <- function(Xi,Xj, l=1) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
+  cov <- function(X, Y) outer(X, Y, SE, l) 
   n <- length(obs$x)
@@
   I <-  diag(1, n)
-
   L <- chol(K + sigma_n^2 * I)
-  alpha <- solve(t(L), solve(L, obs$y))
+  alpha <- forwardsolve(t(L), backsolve(L, obs$y)) # see http://stat.ethz.ch/R-manual/R-patched/library/base/html/chol2inv.html
   k_star <- cov(obs$x, X)
   Y <- t(k_star) %*% alpha
-  v <- solve(L, k_star)
+  v <- backsolve(L, k_star)
   Var <- cov(X,X) - t(v) %*% v
@@
 ``` {r}
-  cov_xx_inv <- solve(K + sigma_n^2*I)
-  Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
-  Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X)
+  d <- gp_fit(obs, X, c(sigma_n=sigma_n, l=l, tau=1), "direct")
+  Ef <- d$Ef
+  Cf <- d$Cf
 ```
@@
 ``` {r}
-ef <- numeric(length(X))
-cf <- matrix(0, nrow=length(X), ncol=length(X))
-
-A <- as.numeric( cov(obs$x[1], obs$x[1]) )
-mu <- obs$y[1]
-
-for(i in 2:length(obs$x)){
-  mu <- obs$y[i] + cov(obs$x[i], obs$x[1:(i-1)]) %*% (obs$y[1:(i-1)] - mu) / (A + sigma_n^2)
-  A <- as.numeric( cov(obs$x[i], obs$x[i]) - cov(obs$x[i], obs$x[1:(i-1)]) %*% cov(obs$x[1:(i-1)], obs$x[i]) / (A + sigma_n^2) )
-}
-
-ef <- cov(X, obs$x) %*% (obs$y - mu) / A
-cf <- cov(X,X) - cov(X, obs$x) %*% cov(obs$x, X) / A
+s <- gp_fit(obs, X, c(sigma_n=sigma_n, l=l, tau=1), "sequential")
+ef <- s$Ef
+cf <- s$Cf
 
@@
 ``` {r}
-gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), fit=FALSE, scaled=FALSE, var=0.8)
-y_p <- predict(gp, X)
-```
-
-These are the same as `y_p`, explicitly showing the $E_y = K(x, x_i) \alpha$ multiplication
-
-``` {r}
-y_mult <-  kernelMult(kernelf(gp), X, xmatrix(gp), as.matrix(alpha(gp))) 
-y_se <- sapply(X, function(x) cov(x, obs$x) %*% alpha(gp))
-rbind(as.numeric(y_p), as.numeric(y_mult), as.numeric(y_se))[,1:10]
-```
-
-Which demonstrates the kernel trick `alpha = cov_xx_inv %*% obs$y`
-
-### mlegp method
-
-``` {r}
-require(mlegp)
-out <- mlegp(obs$x, obs$y, nugget = 1)
-y_m <- predict(out, as.matrix(X))
+k <- gp_fit(obs, X, c(sigma_n=sigma_n, l=l, tau=1), "sequential")
+y_p <- k$Ef
 ```
@@
 ````
-
-Compare alpha from Cholesky and alpha from `kernlab` to $K(x,x)^{-1} \vec y$
-
-```{r}
-cov_xx_inv %*% obs$y
-alpha
-alpha(gp)
-```

### 2012-11-29 R/gp_fit.R

@@ -71,7 +71,7 @@ gp_fit <- function(obs, X, pars=c(sigma_n=1, tau=1, l=1), method=c("direct", "se
     llik <- -.5 * t(obs$y) %*% cov_xx_inv %*% obs$y - 0.5 * log(det(cov_xx_inv)) - n * log(2 * pi) / 2
   }
   
-  ### Sequential method, using function recursion
+  ### Sequential method, using function recursion (terribly slow for more than a few data points)
   else if(method=="sequential"){
     mmult <- function(x,y){
       if(length(x) == 1){

### 2012-11-29 inst/examples/singular-inversion.Rmd
seeing if I get singular matrices in the inversion

testing singular matrix condition

```{r}
require(ggplot2)
require(kernlab)
require(nonparametricbayes)
opts_knit$set(upload.fun = socialR::notebook.url)
opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE)
theme_set(theme_bw())
theme_update(panel.background = element_rect(fill = "transparent",colour = NA))
```

Parameterization-specific

```{r}
X <- seq(0,1, length=1000)
obs <- data.frame(x = X,
                  y = sin(2*pi*10*X)+.3*rnorm(length(X)))
```

### Direct method

``` {r}
d <- gp_fit(obs, X, c(sigma_n=1, l=.0001, tau=.3), "direct")
plot.gpfit(d)
```

``` {r}
d <- gp_fit(obs, X, c(sigma_n=1, l=10, tau=.3), "direct")
plot.gpfit(d)
```

### 2012-11-29 inst/examples/singular-inversion.Rmd
print determinant value

@@
 
+Fit with reasonable choice of length scale `l`
+
 ``` {r}
-d <- gp_fit(obs, X, c(sigma_n=1, l=.0001, tau=.3), "direct")
+d <- gp_fit(obs, X, c(sigma_n=.5, l=.02, tau=.3), "direct")
+det(d$K + .5^2 * diag(1, length(X)))
 plot.gpfit(d)
@@
 
+Fit with long length scales.
+
 ``` {r}
-d <- gp_fit(obs, X, c(sigma_n=1, l=10, tau=.3), "direct")
+d <- gp_fit(obs, X, c(sigma_n=.5, l=10, tau=.3), "direct")
+det(d$K + .5^2 * diag(1, length(X)))
 plot.gpfit(d)

### 2012-11-30 inst/examples/mcmc-gp.Rmd
Reed example using tgp, MCMC estimated hyperparameters #6 #3

``` {r libs, include=FALSE}
rm(list=ls())   
require(nonparametricbayes)
require(reshape2)
require(ggplot2)
require(data.table)
require(tgp)
```

```{r graphing-options, include=FALSE}
opts_knit$set(upload.fun = socialR::notebook.url)
opts_chunk$set(dev.args=list(bg="transparent"), tidy=FALSE, comment=NA)
theme_notebook <- theme_grey() + 
                  theme(plot.background = element_rect(fill = "transparent",colour = NA),
                        panel.grid.minor=element_line(colour="transparent"))
theme_set(theme_notebook)
```

```{r results="hide"}
knit("beverton_holt_data.Rmd")
true <- data.frame(x = X, y = sapply(X, f, 0, p))
```

```{r}
gp <- bgp(X=obs$x, XX=X, Z=obs$y)
```

Extract the posterior Gaussian process mean and the $\pm 2$ standard deviations over the predicted grid from the fit:

```{r}
center <- tgp:::tgp.choose.center(gp, "mean")
Z.mean <- center$Z
X_vals <- center$X[, 1]
o <- order(X_vals)
Zb.q1 <- Z.mean + 1.96 * sqrt(c(gp$Zp.ks2, gp$ZZ.ks2))
Zb.q2 <- Z.mean - 1.96 * sqrt(c(gp$Zp.ks2, gp$ZZ.ks2))
V <- c(gp$Zp.ks2, gp$ZZ.ks2)[o]
dat <- data.frame(x=X_vals[o], y=Z.mean[o], ymin=Zb.q1[o], ymax=Zb.q2[o])
```

Plot the posterior Gaussian Process:

```{r gp-plot}
  p1 <- ggplot(dat) +
    geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
    geom_line(aes(x=x,y=y), size=1) + #MEAN
    geom_point(data=gp$obs,aes(x=x,y=y)) + 
    labs(title=paste("llik =", prettyNum(gp$llik)))
  if(!is.null(true))
    p1 <- p1 + geom_line(data = true, aes(x, y), col = "red", lty = 2) 
p1 + theme_notebook
```

Find the actual optimum policy from the true parametric model

```{r}
x_grid <- dat$x
h_grid <- x_grid
matrices_true <- f_transition_matrix(f, p, x_grid, h_grid)
opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
```

Estimate a policy from the Gaussian Process:

```{r}
rownorm <- function(M) t(apply(M, 1, function(x) x/sum(x)))
matrices_gp <- lapply(h_grid, function(h){
    S <- dat$y - h
    F_ <- t(sapply(1:length(S), function(i){
      if(S[i]>0) {
        out <- dlnorm(x_grid/S[i], 0, V[i])
      } else {
        out <- numeric(length(x_grid))
        out[1] <- 1
        out
      }
    }))
    F_ <- rownorm(F_)
  })
opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
```

```{r}
matrices_true[[1]][1:10,1:10]
matrices_gp[[1]][1:10,1:10]

```

```{r policy_plot}
policies <- melt(data.frame(stock=x_grid, GP = x_grid[opt_gp$D[,1]], Exact = x_grid[opt_true$D[,1]]), id="stock")
policy_plot <- ggplot(policies, aes(stock, stock - value, color=variable)) +
  geom_point() + xlab("stock size") + ylab("escapement") 
policy_plot + theme_notebook
```

We can see what happens when we attempt to manage a stock using this:

``` {r othernoise}
z_g <- function() rlnorm(1,0, sigma_g)
````

``` {r  simulate }
set.seed(1)
sim_gp <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_gp$D, z_g, profit=profit)
set.seed(1)
sim_true <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_true$D, z_g, profit=profit)
````

``` {r  simplot}
df <- data.frame(time = sim_gp$time, stock_gp = sim_gp$fishstock, stock_true = sim_true$fishstock, harvest_gp = sim_gp$harvest, havest_true = sim_true$harvest)
df <- melt(df, id="time")
simplot <- ggplot(df) + geom_line(aes(time,value, color=variable))
simplot + theme_notebook
````

Total Profits

```{r}
sum(sim_gp$profit)
sum(sim_true$profit)
```

### 2012-11-30 R/gp_trasition_matrix.R

@@ -12,7 +12,16 @@ gp_transition_matrix <- function(gp, h_grid=NULL){
     h_grid <- x_grid
   V <- sqrt(diag(gp$Cf))
   matrices_gp <- lapply(h_grid, function(h){
-    F_ <- sapply(x_grid, function(x) dnorm(x, gp$Ef-h, V))
+    S <- gp$Ef - h
+    F_ <- t(sapply(1:length(S), function(i){
+      if(S[i]>0) {
+        out <- dlnorm(x_grid/S[i], 0, V[i])
+      } else {
+        out <- numeric(length(x_grid))
+        out[1] <- 1
+        out
+      }
+    }))
     F_ <- rownorm(F_)
   })
   matrices_gp
@@ -29,11 +38,27 @@ gp_transition_matrix <- function(gp, h_grid=NULL){
 f_transition_matrix <- function(f, p, x_grid, h_grid=NULL){
   if(is.null(h_grid))
     h_grid <- x_grid
+  
   matrices_F <- lapply(h_grid, function(h){
     mu <- sapply(x_grid, f, h, p)
-    F_true <- sapply(x_grid, function(x) dnorm(x, mu, sigma_g))
+    F_true <- t(sapply(mu, function(m){
+      if(m>0) {
+        out <- dlnorm(x_grid/m, 0, sigma_g)
+      } else {
+        out <- numeric(length(x_grid))
+        out[1] <- 1
+        out
+      }
+    }))
     F_true <- rownorm(F_true)
   })
+  
+#  matrices_F <- lapply(h_grid, function(h){
+#    mu <- sapply(x_grid, f, h, p)
+#    F_true <- sapply(x_grid, function(x) dlnorm(x/mu, 0, sigma_g))
+#    F_true <- rownorm(F_true)
+#  })
+  
   matrices_F
 }
 

### 2012-12-05 inst/examples/comparing-methods.Rmd
illustration of configuring some of the parameters and accessing the correct output in tgp #6

@@
 theme_set(theme_bw())
-theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
-             plot.background = element_rect(fill = "transparent",colour = NA))
+theme_update(panel.background = element_rect(fill = "transparent",colour = NA))
 ```

### 2012-12-05 inst/examples/mcmc-gp.Rmd
illustration of configuring some of the parameters and accessing the correct output in tgp #6

@@
 ```{r results="hide"}
-knit("beverton_holt_data.Rmd")
-true <- data.frame(x = X, y = sapply(X, f, 0, p))
+knit("~/Documents/code/nonparametric-bayes/inst/examples/beverton_holt_data.Rmd")
+true <- data.frame(x = x_grid, y = sapply(x_grid, f, 0, p))
+XX <- as.matrix(x_grid)
+X <- as.matrix(obs$x)
+Z <- as.matrix(obs$y)
+```
+
+Normalize data to the unit (hyper)cube.
+
+```{r, include=FALSE, eval=FALSE}
+unit_norm <- function(x, M=NULL, m=NULL, invert=FALSE){
+  if(is.null(M)){
+    M <- max(x) 
+    m <- min(x)
+    message(paste("Max M = ", M, "min m = ", m))
+  }
+
+  if(!invert) {
+    (x-m)/(M-m)
+  } else {
+    (M-m)*x+m
+  }
+}
+XX = unit_norm(XX)
+X = unit_norm(X)
+Z = unit_norm(Y)
 ```
@@
 ```{r}
-gp <- bgp(X=obs$x, XX=X, Z=obs$y)
+gp <- bgp(X=X, XX=XX, Z=Z, meanfn="constant", bprior="b0", BTE=c(1000,6000,2), m0r1=FALSE, verb=4, corr="exp", trace=TRUE, s2.p=c(5,10), tau2.p=c(5,10), s2.lam="fixed", tau2.lam="fixed")
 ```
+
+Take a look at the trace
+
+```{r}
+names(gp$trace$XX[[1]])
+dim(gp$trace$XX[[1]])
+gp$trace$XX[[1]][1,]
+gp$trace$XX[[1]][2500,]
+````
+
+* `index` is the sampling point (from `BTE`, we see we sample starting at step 1000 and ending at step 6000, recording every 2 steps, so there are $(T-B)/E = 2500$ index points).
+
+* `lambda`
+* `s2`
+* `tau2`
+* `beta0`
+* `nug`
+* `d`
 
@@
 ```{r}
-center <- tgp:::tgp.choose.center(gp, "mean")
-Z.mean <- center$Z
-X_vals <- center$X[, 1]
-o <- order(X_vals)
-Zb.q1 <- Z.mean + 1.96 * sqrt(c(gp$Zp.ks2, gp$ZZ.ks2))
-Zb.q2 <- Z.mean - 1.96 * sqrt(c(gp$Zp.ks2, gp$ZZ.ks2))
-V <- c(gp$Zp.ks2, gp$ZZ.ks2)[o]
-dat <- data.frame(x=X_vals[o], y=Z.mean[o], ymin=Zb.q1[o], ymax=Zb.q2[o])
+
+dat <- data.frame(x    = gp$XX[[1]], 
+                 y    = gp$ZZ.km, 
+                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
+                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
+
 ```
@@
 p1 + theme_notebook
+```
+
+is mean ZZ.mean or ZZ.km?
+is the var I want ZZ.s2 or ZZ.ks2?)  I want the kreiging values (km, ks2).
+The others slightly don't agree with the above.
+We do this by comparing to the plotted option (which handles the rescaling(?) from `choose.center` first).
+
+```{r}
+center <- tgp:::tgp.choose.center(gp, "mean")
+Z.mean <- center$Z
+X_vals <- center$X[, 1]
+o <- order(X_vals)
+Zb.q1 <- Z.mean + 1.96 * sqrt(c(gp$Zp.ks2, gp$ZZ.ks2))
+Zb.q2 <- Z.mean - 1.96 * sqrt(c(gp$Zp.ks2, gp$ZZ.ks2))
+V <- c(gp$Zp.ks2, gp$ZZ.ks2)[o]
+df <- data.frame(x=X_vals[o], y=Z.mean[o], ymin=Zb.q1[o], ymax=Zb.q2[o])
+
+raw <- data.frame(x    = gp$XX[[1]], 
+                 y    = gp$ZZ.mean, 
+                 ymin = gp$ZZ.mean - 1.96 * sqrt(gp$ZZ.s2), 
+                 ymax = gp$ZZ.mean + 1.96 * sqrt(gp$ZZ.s2))
+
+p1 + geom_ribbon(aes(x,y, ymin=ymin, ymax=ymax), fill="red", alpha=.1, data=df) + geom_line(aes(x,y), col="red", data=df)  + geom_ribbon(aes(x,y, ymin=ymin, ymax=ymax), fill="blue", alpha=.1, data=raw) + geom_line(aes(x,y), col="blue", data=raw)
 ```

### 2012-12-05 inst/examples/mcmc-gp.Rmd
Notes on `tgp` package MCMC

--- 2012-11-30 inst/examples/mcmc-gp.Rmd
@@
 ```{r results="hide"}
-knit("beverton_holt_data.Rmd")
-true <- data.frame(x = X, y = sapply(X, f, 0, p))
+knit("~/Documents/code/nonparametric-bayes/inst/examples/beverton_holt_data.Rmd")
+true <- data.frame(x = x_grid, y = sapply(x_grid, f, 0, p))
+XX <- as.matrix(x_grid)
+X <- as.matrix(obs$x)
+Z <- as.matrix(obs$y)
+```
+
+## Call to `tgp` function `bgp()`
+
+Tweak the hyperparameter prior's parameters (`s2.p`) to influence the GP inferred.
+
+```{r}
+gp <- bgp(X=X, XX=XX, Z=Z, meanfn="constant", bprior="b0", BTE=c(1000,6000,2), m0r1=TRUE, verb=4, corr="exp", trace=TRUE, s2.p=c(5,10), tau2.p=c(5,10), s2.lam="fixed", tau2.lam="fixed")
+```
+
+## parsing the function call
+
+* `X` is the observed X values (or matrix of appropriate dimension)
+* `XX` the desired predicted grid. (will also predict on X)
+* `Z` is observed Y values (or matrix of appropriate dimension)
+* `m0r1` means scale data to mean zero and unit range
+* `bprior` is the prior distribution, defaulting to `bfalt`.
+`b0` is a Gaussian.
+This prior is for the Gaussian process and is not to be confused with the hyperpriors describing the distribution of various parameters.
+* `BTE` Burn-in B steps, Terminate after T steps, sample Every E steps. 
+* `verb` Very verbose output
+* Other options (for adaptive/query learning, etc, igored for now.  Focus on priors).  
+* The priors for the hyperparameters are given by `s2.p`, `tau.p` etc (both are Inverse Gammas).  Their parameters can vary hiearchically in general, but we hold them fixed, `s2.lam = 'fixed'`, etc.)
+
+## Hyperparameters
+
+Hyperparameter distributions returned in the trace:
+
+```{r}
+names(gp$trace$XX[[1]][2500,])
+```
+
+* `s2` is $\sigma^2$ in Gramacy and $\tau$ in Rassmussen and Williams, is the scale of the noise. Comes from an inverse gamma prior.  Set by inverse-gamma parameters `a0` and `g0`.
+
+* `tau2` Also comes from an inverse gamma prior. It is also set by inverse-gamma parameters `a0` and `g0`.
+
+$$Z | \beta, \sigma^2, K \sim N_n(\mathbf{F} \beta, \sigma^2 \mathbf{K}) $$
+$$ \beta | \sigma^2, \tau^2, \mathbf{W}, \beta_0 ~ N_m(\beta_0, \sigma^2 \tau^2 \mathbf{W} )  $$
+$$\beta_0 | N_m(\mathbf{\mu}, \mathbf{B})$$
+
+The parameters in this run are constant across the trace points (because hyperparameters are fixed).
+
+```{r}
+gp$trace$XX[[1]][2500,]
+gp$trace$XX[[3]][2500,]
 ```
@@
 ```{r}
-gp <- bgp(X=obs$x, XX=X, Z=obs$y)
+pars <- melt(gp$trace$XX[[1]], id = "index")
+ggplot(pars, aes(index, value)) + geom_line() + facet_wrap(~variable, scales="free_y")
+````
+
+It is possible to calculate summary statistics and check out the distribution of these fellows.
+
+```{r}
+mean(gp$trace$XX[[1]][,"nug"])
+ggplot(pars, aes(value)) + geom_histogram() + facet_wrap(~variable, scales="free")
 ```
+
+* We keep the heirachical priors fixed in the `tau.lm`, `s2.lm`, etc.
+`nug` fixed by default.
+Note this results in the the priors from which these are sampled are fixed: e.g.
+`gp$trace$hier$s2.a0`, `gp$trace$hier$s2.g0`
+
+* `nug.p` This is the "nugget," the measurement error.
+Comes from a mixture of gamma prior parameter (initial values) for the nugget parameter `c(a1,g1,a2,g2)` where `g1` and `g2` are scale (`1/rate`) parameters.
+_Default reduces to simple exponential prior._ Specifying `nug.p = 0` fixes the nugget parameter to the "starting" value in `gd[1]`, i.e., it is excluded from the MCMC
+
+* `index` is the sampling point (from `BTE`, we see we sample starting at step 1000 and ending at step 6000, recording every 2 steps, so there are $(T-B)/E = 2500$ index points).  
+* `lambda` possibly a mixing parameter in the importance sampler temperature??
+* `beta0` No mention in the manual
+* `d` The parameter for priors using the hierachical exponential distribution for the parameters `a1`, `a2`, `g1`, `g2`.  
+* `b` ??
+
+* Where are the hyperparameters for the correlation function (e.g. the length-scale for the Gaussian??)
+
+## Extracting the estimated process
+
+(The **MAP**, maximum a posteriori expected mean and variance).
 
@@
 ```{r}
-center <- tgp:::tgp.choose.center(gp, "mean")
-Z.mean <- center$Z
-X_vals <- center$X[, 1]
-o <- order(X_vals)
-Zb.q1 <- Z.mean + 1.96 * sqrt(c(gp$Zp.ks2, gp$ZZ.ks2))
-Zb.q2 <- Z.mean - 1.96 * sqrt(c(gp$Zp.ks2, gp$ZZ.ks2))
-V <- c(gp$Zp.ks2, gp$ZZ.ks2)[o]
-dat <- data.frame(x=X_vals[o], y=Z.mean[o], ymin=Zb.q1[o], ymax=Zb.q2[o])
+V <- gp$ZZ.ks2
+dat <- data.frame(x   = gp$XX[[1]], 
+                 y    = gp$ZZ.km, 
+                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
+                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
 ```

### 2012-12-07 inst/examples/gaussian-process-basics.Rmd
cleaner plots, remove kernlab comparison

@@
 opts_knit$set(upload.fun = socialR::notebook.url)
+theme_update(plot.background = element_rect(fill = "transparent",colour = NA))
 ```
@@
   geom_rect(xmin=-Inf, xmax=Inf, ymin=-2, ymax=2, fill="grey80") +
-  geom_line(aes(group=variable)) +   theme_bw() +
+  geom_line(aes(group=variable)) +  
   scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
@@
   xlab("input, x") 
-fig2c + opts(panel.background = theme_rect(fill = "transparent"), 
-        plot.background = theme_rect(fill = "transparent"))
+fig2c
 ```
@@
 ```
-
-## A comparison to built-in R utilities
-
-Some basic Gaussian process tools are provided in the `kernlab` package,
-
-```{r}
-library(kernlab)
-```
-
-through the `gausspr` function,
-
-```{r}
-gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), fit=FALSE, scaled=FALSE, var=0.8)
-```
-
-Where `rbfdot` indicates the Gaussian (Radial basis function) kernel we use above, with parameter `sigma=1/(2l^2)` (note that gausspr defines the kernel parameter differently, see:
-
-```{r}
-gp@kernelf@.Data
-```
-note this takes vectors `x`, `y` and returns a scalar, not the covaraiance matrix).
-The function will try to estimate hyperparameters such as `sigma` if they are not given unless `fit=FALSE`.
-The data is also scaled to 0 mean, unit variance unless `scaled=FALSE` is set.
-
-We can then use the estimated process to infer values on a specified grid through `predict`, e.g. for the example above,
-
-```{r}
-y_p <- predict(gp, x_predict)
-dat2 <- data.frame(x = x_predict, variable = "Y_p", value = y_p)
-```
-
-```{r}
-ggplot(dat2, aes(x, value)) + geom_point() + geom_line(data=Ey,aes(x=x,y=y), size=1) 
-```
-
-### Fitting hyperparameters by maximum likelihood
-
-If we do not specify `fit=false` then `gausspr` will estimate the hyperparamters by maximum likelihood:
-
-```{r}
-gp_fit <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), scaled=FALSE, var=0.8)
-```

### 2012-12-08 inst/examples/mcmc-gp.Rmd
Updated notes on tgp package

--- 2012-11-30 inst/examples/mcmc-gp.Rmd
@@
-``` {r libs, include=FALSE}
-rm(list=ls())   
-require(nonparametricbayes)
-require(reshape2)
-require(ggplot2)
-require(data.table)
-require(tgp)
+```{r include=FALSE}
+opts_knit$set(upload.fun = socialR::notebook.url)
+opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE, message=FALSE)
+theme_update(plot.background = element_rect(fill = "transparent",colour = NA))
 ```
 
-```{r graphing-options, include=FALSE}
-opts_knit$set(upload.fun = socialR::notebook.url)
-opts_chunk$set(dev.args=list(bg="transparent"), tidy=FALSE, comment=NA)
-theme_notebook <- theme_grey() + 
-                  theme(plot.background = element_rect(fill = "transparent",colour = NA),
-                        panel.grid.minor=element_line(colour="transparent"))
-theme_set(theme_notebook)
+I am trying to understand the interface for tgp method, but much of it is not well documented (yes, despite reading through the package R manual, two nice vignettes, and a the PhD Thesis describing it.)
+
+Just to get some consistent notation down: consider the case of a Gaussian process with a Gaussian/radial basis function kernel, parameterized as:
+
+$$K(X, X') = \sigma^2 e^{\frac{(X-X')^2}{d}}$$
+
+And observations from $Z = f(X) + \varepsilon$, for which we seek to approximate $f$ as a (multivariate) Gaussian process, $Z | X ~ N(\mu, C)$.
+The covariance matrix $C$ is given by our kernel $K$, conditioned on our observations; the mean $\mu$ is given by a constant or linear model, likewise conditioned on our observations.
+If the multivariate normal of observed and predicted points is,
+
+$$\begin{pmatrix} y_{\textrm{obs}} \ y_{\textrm{pred}} \end{pmatrix} \sim \mathcal{N}\left( \mathbf{0}, \begin{bmatrix} cov(X_o,X_o) & cov(X_o, X_p) \ cov(X_p,X_o) & cov(X_p, X_p) \end{bmatrix} \right)$$
+
+the conditional probability is
+
+$$x|y \sim \mathcal{N}(E,C)$$ $$E = cov(X_p, X_o) (cov(X_o,X_o) + \varepsilon \mathbb{I}) ^{-1} y$$ $$C= cov(X_p, X_p) - cov(X_p, X_o) (cov(X_o,X_o)+ \varepsilon \mathbb{I} )^{-1} cov(X_o, X_p)$$
+
+Our hyperparameters are $d$, $\sigma^2$, and $\varepsilon$, for which we need priors.  We'll use the inverse gamma,
+
+$$P(x; a, g) = \frac{g^a}{\Gamma(a)} x^{-a - 1}\exp\left(-\frac{g}{x}\right)$$
+
+The distribution arises as the marginal posterior distribution for the unknown variance of a normal distribution if an uninformative prior is used; and as a useful conjugate prior if a less uninformative prior is preferred.
+However, it is common among Bayesians to consider an alternative parametrization of the normal distribution in terms of the precision, defined as the reciprocal of the variance, which allows the gamma distribution to be used directly as a conjugate prior.
+
+* $X$ is `X`, $Z$ is `Z`
+
+* $\sigma^2$ is the parameter `s2` in the `tgp` package, coming from inverse gamma prior with parameters $a = a_0$ and $g = g_0$, which are set by `s2.p = c(a_0, g_0)`.
+Optionally, $a_0$ and $g_0$ can be taken from an exponential prior $\lambda e^{-\lambda x}$, in which the value of $\lambda$ is given by `s2.lam = c(lambda_a, lambda_g)`.
+Setting `s2.lam = "fixed"` "turns off" the use of a hyperprior.
+
+* $d$ I believe is `d`, which Grammarcy calls the range parameter (other terms include the "lengthscale").
+I am not sure what either `d` or `s2` correspond to if the covariance function `corr` is something other than the radial basis function.
+I'm also not clear how to set the correlation function to a strictly Gaussian covariance -- `corr` is set to the `exp` family, which leaves the power as a parameter `p_0`:
+
+$$K(X, X') =  e^{\frac{(X-X')^p_0}{d}}$$
+
+(from equation 4 in the Vignette (1)).
+I do not see a prior for `p_0` or any mention of it.
+Nor can I tell what is the functional form for it's prior (`d.p`) (which is not an inverse gamma?)?
+Not clear why it's prior gets defined with four parameters, `c(a1, g1, a2, g2)`, instead of two.
+Form the manual:
+
+> Mixture of gamma prior parameter (initial values) for the range parameter(s) c(a1,g1,a2,g2) where g1 and g2 are scale (1/rate) parameters.
+
+As before, these parameters  `c(a1, g1, a2, g2)` can be drawn from an exponential distribution hyperprior whose $\lambda$ values are given by `d.lam`, or turned off.
+
+* $\varepsilon$ is `nug`.  It seems to have same shape prior as $d$, specified by `c(a1,g1,a2,g2)`, and optional hyperprior as well.  Additionally it can be fixed to a constant:
+
+> specifying nug.p = 0 fixes the nugget parameter to the starting value in gd[1], i.e., it is excluded from the MCMC
+
+But what is `gd[1]` and where is it defined?
+
+If the mean is not fixed at zero but is instead a linear model, we need hyperparameters for it as well.
+The key parameter there is the slope $\beta$.
+The functional form for it's prior is specified as `bprior`, which takes values `bflat`, `b0`, `bmzt`, or `bmznot`, but I can find no precise statement of the functional forms of these priors (e.g.
+I see that `b0` is Gaussian prior and `bflat` an improper or flat prior, but beyond that), nor do I see where their shape parameters are set.
+
+The manual suggests that we need to specify start values for the hyperparameters for the MCMC (presumably we could have just drawn them from the prior? we burn in anyhow...)  The are given in
+
+> `start =  c(0.5,0.1,1.0,1.0)` starting values for range `d`, nugget `g`, `s2`, and `tau2`
+
+Actually specifiying start values throws an error though, it appears the documentation is out of date and the start parameter has been removed.
+
+* I'm still not clear on the role of `tau2`, whether it's unique to the "treed" partioning or whether I've misunderstood $sigma^2$.  Grammacy phrases the model as:
+
+$$Z | \beta, \sigma^2, K \sim N_n(\mathbf{F} \beta, \sigma^2 \mathbf{K}) $$
+$$ \beta | \sigma^2, \tau^2, \mathbf{W}, \beta_0 ~ N_m(\beta_0, \sigma^2 \tau^2 \mathbf{W} )  $$
+$$\beta_0 | N_m(\mathbf{\mu}, \mathbf{B})$$
+
+(I guess I haven't parsed the second two equations)
+
+## Example comparison
+
+So, how now would I make a call to `tgp::bgp` that would give me the equivalent output of, say, a direct calculation of the above model (given fixed hyperparameters)?
+Say, given some observed data pairs $x,y$ and predictions desired on a grid $X$,
+
+```{r}
+X = seq(-5, 5, length=50)
+x = c(-4, -3, -1,  0,  2)
+y = c(-2,  0,  1,  2, -1)
 ```
 
-```{r results="hide"}
-knit("beverton_holt_data.Rmd")
-true <- data.frame(x = X, y = sapply(X, f, 0, p))
+Manually I could just do:
+
+```{r}
+d = .5; epsilon = .1; sigma = 1; #fixed hyperparamaters
+SE <- function(Xi,Xj, d) sigma * exp(-0.5 * (Xi - Xj) ^ 2 / d ^ 2)
+cov <- function(X, Y) outer(X, Y, SE, d) 
+cov_xx_inv <- solve(cov(x, x) + epsilon * diag(1, length(x)))
+Ef <- cov(X, x) %*% cov_xx_inv %*% y
+Cf <- cov(X, X) - cov(X, x)  %*% cov_xx_inv %*% cov(x, X)
 ```
 
+And `Ef` and `Cf` are the estimated (krieging) mean and covariance.
+
+I might try something like
+
 ```{r}
-gp <- bgp(X=obs$x, XX=X, Z=obs$y)
+library(tgp)
+gp <- bgp(X=x, XX=X, Z=y, verb=0,
+          meanfn="constant", bprior="bflat", BTE=c(1,2,1), m0r1=FALSE, 
+          corr="exp", trace=TRUE, beta = 0,
+          s2.p = c(25,10), d.p = c(50, 10, 10, 10), nug.p = c(50, 10, 10, 10),
+          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", 
+          tau2.lam = "fixed", tau2.p=c(5,10))
 ```
@@
 ```{r}
-center <- tgp:::tgp.choose.center(gp, "mean")
-Z.mean <- center$Z
-X_vals <- center$X[, 1]
-o <- order(X_vals)
-Zb.q1 <- Z.mean + 1.96 * sqrt(c(gp$Zp.ks2, gp$ZZ.ks2))
-Zb.q2 <- Z.mean - 1.96 * sqrt(c(gp$Zp.ks2, gp$ZZ.ks2))
-V <- c(gp$Zp.ks2, gp$ZZ.ks2)[o]
-dat <- data.frame(x=X_vals[o], y=Z.mean[o], ymin=Zb.q1[o], ymax=Zb.q2[o])
-```
-
-Plot the posterior Gaussian Process:
-
-```{r gp-plot}
-  p1 <- ggplot(dat) +
-    geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
-    geom_line(aes(x=x,y=y), size=1) + #MEAN
-    geom_point(data=gp$obs,aes(x=x,y=y)) + 
-    labs(title=paste("llik =", prettyNum(gp$llik)))
-  if(!is.null(true))
-    p1 <- p1 + geom_line(data = true, aes(x, y), col = "red", lty = 2) 
-p1 + theme_notebook
-```
-
-Find the actual optimum policy from the true parametric model
-
-```{r}
-x_grid <- dat$x
-h_grid <- x_grid
-matrices_true <- f_transition_matrix(f, p, x_grid, h_grid)
-opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
-```
-
-Estimate a policy from the Gaussian Process:
-
-```{r}
-rownorm <- function(M) t(apply(M, 1, function(x) x/sum(x)))
-matrices_gp <- lapply(h_grid, function(h){
-    S <- dat$y - h
-    F_ <- t(sapply(1:length(S), function(i){
-      if(S[i]>0) {
-        out <- dlnorm(x_grid/S[i], 0, V[i])
-      } else {
-        out <- numeric(length(x_grid))
-        out[1] <- 1
-        out
-      }
-    }))
-    F_ <- rownorm(F_)
-  })
-opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
-```
-
-```{r}
-matrices_true[[1]][1:10,1:10]
-matrices_gp[[1]][1:10,1:10]
+V <- gp$ZZ.ks2
+tgp_dat <- data.frame(x   = gp$XX[[1]], 
+                  y   = gp$ZZ.km, 
+                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
+                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
 
@@
 
-```{r policy_plot}
-policies <- melt(data.frame(stock=x_grid, GP = x_grid[opt_gp$D[,1]], Exact = x_grid[opt_true$D[,1]]), id="stock")
-policy_plot <- ggplot(policies, aes(stock, stock - value, color=variable)) +
-  geom_point() + xlab("stock size") + ylab("escapement") 
-policy_plot + theme_notebook
+and likewise for our manual data:
+
+```{r}
+manual_dat <- data.frame(x=X, y=Ef, ymin=(Ef - 1.96 * sqrt(diag(Cf))), ymax=(Ef + 2 * sqrt(diag(Cf))))
 ```
 
-We can see what happens when we attempt to manage a stock using this:
+Compare the GP posteriors:
 
-``` {r othernoise}
-z_g <- function() rlnorm(1,0, sigma_g)
-````
-
-``` {r  simulate }
-set.seed(1)
-sim_gp <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_gp$D, z_g, profit=profit)
-set.seed(1)
-sim_true <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_true$D, z_g, profit=profit)
-````
-
-``` {r  simplot}
-df <- data.frame(time = sim_gp$time, stock_gp = sim_gp$fishstock, stock_true = sim_true$fishstock, harvest_gp = sim_gp$harvest, havest_true = sim_true$harvest)
-df <- melt(df, id="time")
-simplot <- ggplot(df) + geom_line(aes(time,value, color=variable))
-simplot + theme_notebook
-````
-
-Total Profits
-
-```{r}
-sum(sim_gp$profit)
-sum(sim_true$profit)
+```{r gp-plot}
+library(ggplot2)
+ggplot(tgp_dat) +
+    geom_ribbon(aes(x, y, ymin = ymin, ymax = ymax), fill="red", alpha = .1) + # Var
+    geom_line(aes(x, y), col="red") + # mean
+    geom_point(data = data.frame(x = x, y = y), aes(x, y)) + # raw data
+    geom_ribbon(dat = manual_dat, aes(x, y, ymin = ymin, ymax = ymax), fill = "blue", alpha = .1) + # Var
+    geom_line(dat = manual_dat, aes(x, y), col = "blue")  + #MEAN    
+    theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA))
 ```

### 2012-12-08 inst/examples/myer-example.Rmd
A preliminary example of the GP outperforming a parametric model with error in parameter estimate

```{r libraries}
require(pdgControl)
require(nonparametricbayes)
require(reshape2)
require(ggplot2)
require(data.table)
library(kernlab)
```

```{r graphing-options, include=FALSE}
opts_knit$set(upload.fun = socialR::notebook.url)
opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE)
theme_set(theme_bw())
theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
             plot.background = element_rect(fill = "transparent",colour = NA))
```

```{r}
f <- Myer_harvest
pars <- c(1, 2, 6) 
p <- pars # shorthand 
K <- p[1] * p[3] / 2 + sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2
xT <- p[1] * p[3] / 2 - sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 # allee threshold
e_star <- (p[1] * sqrt(p[3]) - 2) / 2 ## Bifurcation point 
```

Noise function, profit function
```{r}
z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
profit <- profit_harvest(1,0,0)
```

Parameter definitions

```{r}
x_grid = seq(0, 1.5 * K, length=101)
T <- 40
sigma_g <- 0.2
x <- numeric(T)
x[1] <- 1
seed <- 123
```

Simulation

```{r}
x[1] = 2.5
for(t in 1:(T-1))
  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
plot(x)
```

Predict the function over the target grid

```{r}
obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
X <- x_grid
gp <- gp_fit(obs, X, c(sigma_n=.9, tau = 1., l=.1), "kernlab")

```

Gaussian Process inference from this model

```{r}
df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
true <- data.frame(x=X, y=sapply(X,f, 0, p))
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
  geom_line(data=true, aes(x,y), col='red', lty=2)
```

## Stochastic Dynamic programming solution based on the posterior Gaussian process:

```{r}
h_grid <- x_grid
```

GP Solution

```{r}
matrices_gp <- gp_transition_matrix(gp)
opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
```

Optimal solution

```{r}
matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
```

Optimal parametric solution in which a parameter is estimated with error

```{r}
p_est = c(1.5,2,6)
matrices_estimated <- f_transition_matrix(f, p_est, x_grid, h_grid, sigma_g=sigma_g)
opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, 20, 0, profit, delta=.01)
```

```{r policy_plot}
policies <- melt(data.frame(stock=x_grid, 
                            GP = x_grid[opt_gp$D[,1]], 
                            Exact = x_grid[opt_true$D[,1]],
                            Approx = x_grid[opt_estimated$D[,1]]),
                  id="stock")

policy_plot <- ggplot(policies, aes(stock, stock - value, color=variable)) +
  geom_point() + xlab("stock size") + ylab("escapement") 
policy_plot
```

We can see what happens when we attempt to manage a stock using this:

``` {r othernoise}
z_g <- function() rlnorm(1,0, sigma_g)
````

``` {r  simulate }
set.seed(seed)
sim_gp <- lapply(1:20, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_gp$D, z_g, profit=profit))
set.seed(seed)
sim_true <- lapply(1:20, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_true$D, z_g, profit=profit))
set.seed(seed)
sim_est <- lapply(1:20, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_estimated$D, z_g, profit=profit))
````

```{r}
dat <- list(est = sim_est, gp = sim_gp, true = sim_true)
dat <- melt(dat, id=names(dat[[1]][[1]]))
dt <- data.table(dat)
setnames(dt, c("L1", "L2"),  c("method", "reps")) 
```

And plot the resulting stock dynamcis across the replicates [^1]

``` {r  simplot}
ggplot(dt) + geom_line(aes(time,fishstock, group=reps, color=method)) + facet_wrap(~method)
````

[^1]: Not sure why that doesn't plot correctly as overlay when I drop the `facet_wrap` option from the command.

Total Profits

```{r}
c( gp = sum(sim_gp$profit), true = sum(sim_true$profit), est = sum(sim_est$profit))
```

### 2012-12-08 inst/examples/may79-example.Rmd
example of resulting policy and simulation of GP vs true and "estimated" optimum.

@@
 ```{r libraries}
-require(nonparametric-bayes)
 require(pdgControl)
+require(nonparametricbayes)
+require(reshape2)
 require(ggplot2)
-opts_knit$set(upload.fun = socialR::flickr.url)
+require(data.table)
+library(kernlab)
+```
+
+```{r graphing-options, include=FALSE}
+opts_knit$set(upload.fun = socialR::notebook.url)
+opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE)
+theme_set(theme_bw())
+theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
+             plot.background = element_rect(fill = "transparent",colour = NA))
 ```
@@
 T <- 40
-sigma_g <- 0.1
+sigma_g <- 0.3
 x <- numeric(T)
@@
 X <- x_grid
-gp <- gp_fit(obs, X, c(sigma_n=1, l=1))
+gp <- gp_fit(obs, X, c(sigma_n=.9, tau = 1., l=.1), "kernlab")
+
 ```
@@
 ```
+
+## Stochastic Dynamic programming solution based on the posterior Gaussian process:
+
+```{r}
+h_grid <- x_grid
+```
+
+GP Solution
+
+```{r}
+matrices_gp <- gp_transition_matrix(gp)
+opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
+```
+
+Optimal solution
+
+```{r}
+matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
+opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
+```
+
+Optimal parametric solution in which a parameter is estimated with error
+
+```{r}
+matrices_estimated <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g=0.01)
+opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, 20, 0, profit, delta=.01)
+```
+
+```{r policy_plot}
+policies <- melt(data.frame(stock=x_grid, 
+                            GP = x_grid[opt_gp$D[,1]], 
+                            Exact = x_grid[opt_true$D[,1]],
+                            Approx = x_grid[opt_estimated$D[,1]]),
+                  id="stock")
+
+policy_plot <- ggplot(policies, aes(stock, stock - value, color=variable)) +
+  geom_point() + xlab("stock size") + ylab("escapement") 
+policy_plot
+```
+
+We can see what happens when we attempt to manage a stock using this:
+
+``` {r othernoise}
+z_g <- function() rlnorm(1,0, sigma_g)
+````
+
+``` {r  simulate }
+set.seed(1)
+sim_gp <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_gp$D, z_g, profit=profit)
+set.seed(1)
+sim_true <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_true$D, z_g, profit=profit)
+set.seed(1)
+sim_est <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_estimated$D, z_g, profit=profit)
+````
+
+```{r}
+dat <- list(est = sim_est, gp = sim_gp, true = sim_true)
+dat <- melt(dat, id=names(dat[[1]]))
+dt <- data.table(dat)
+setnames(dt, "L1", "method") 
+```
+
+``` {r  simplot}
+ggplot(dt) + geom_line(aes(time,fishstock, color=method))
+ggplot(dt) + geom_line(aes(time,harvest, color=method))
+````
+
+Total Profits
+
+```{r}
+c( gp = sum(sim_gp$profit), true = sum(sim_true$profit), est = sum(sim_est$profit))
+```

### 2012-12-10 inst/examples/reed-example.Rmd
Identical to previous reed-example commit but with kernlab engine (not fit)
as before but with kernlab fit (underestimates lengthscale)
Reed example using tau not set to 1 and conditioning on going through (0,0)
parameterized calc of transition matrix can take sigma_g explicitly

--- 2012-11-28 inst/examples/reed-example.Rmd
@@
 require(nonparametricbayes)
+require(reshape2)
 require(ggplot2)
+require(data.table)
 library(kernlab)
@@
 ```{r results="hide"}
-knit("beverton_holt_data.Rmd")
+knit("~/Documents/code/nonparametric-bayes/inst/examples/beverton_holt_data.Rmd")
 true <- data.frame(x = X, y = sapply(X, f, 0, p))
@@
 ```{r}
-gp <- gp_fit(obs, X, c(sigma_n=.2, l=10))
+gp <- gp_fit(obs, X, c(sigma_n=2, tau=10, l=8), "conditional")
 plot.gpfit(gp, true)
@@
 ```{r}
+h_grid <- x_grid
+```
+
+GP Solution
+
+```{r}
 matrices_gp <- gp_transition_matrix(gp)
-matrices_true <- f_transition_matrix(f, p, x_grid)
-opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
 opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
@@
 
+Optimal solution
+
+```{r}
+matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
+opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
+```
+
+Optimal parametric solution in which a parameter is estimated with error
+
+```{r}
+k <- 15
+A <- 1.5
+B <- (A - 1)/k
+matrices_estimated <- f_transition_matrix(f, c(A, B), x_grid, h_grid, sigma_g)
+opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, 20, 0, profit, delta=.01)
+```
+
 ```{r policy_plot}
-require(reshape2)
-policies <- melt(data.frame(stock=x_grid, GP = x_grid[opt_gp$D[,1]], Exact = x_grid[opt_true$D[,1]]), id="stock")
+policies <- melt(data.frame(stock=x_grid, 
+                            GP = x_grid[opt_gp$D[,1]], 
+                            Exact = x_grid[opt_true$D[,1]],
+                            Approx = x_grid[opt_estimated$D[,1]]),
+                  id="stock")
+
 policy_plot <- ggplot(policies, aes(stock, stock - value, color=variable)) +
@@
 sim_true <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_true$D, z_g, profit=profit)
+set.seed(1)
+sim_est <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_estimated$D, z_g, profit=profit)
 ````
 
+```{r}
+dat <- list(est = sim_est, gp = sim_gp, true = sim_true)
+dat <- melt(dat, id=names(dat[[1]]))
+dt <- data.table(dat)
+setnames(dt, "L1", "method") 
+```
+
 ``` {r  simplot}
-df <- data.frame(time = sim_gp$time, stock_gp = sim_gp$fishstock, stock_true = sim_true$fishstock, harvest_gp = sim_gp$harvest, havest_true = sim_true$harvest)
-df <- melt(df, id="time")
-simplot <- ggplot(df) + geom_line(aes(time,value, color=variable))
-simplot
+ggplot(dt) + geom_line(aes(time,fishstock, color=method))
+ggplot(dt) + geom_line(aes(time,harvest, color=method))
 ````
@@
 ```{r}
-sum(sim_gp$profit)
-sum(sim_true$profit)
+c( gp = sum(sim_gp$profit), true = sum(sim_true$profit), est = sum(sim_est$profit))
 ```

### 2012-12-10 R/gp_trasition_matrix.R

@@ -33,9 +33,10 @@ gp_transition_matrix <- function(gp, h_grid=NULL){
 #' @param p the parameters p of the transition function
 #' @param x_grid the grid over which to solve 
 #' @param h_grid the harvest grid over which to compute.  Assumed to match the stock grid gp$X if not given.
+#' @param sigma_g the scale of the noise
 #' @return a list of matrices corresponding to the state transition matrix at each harvest level h
 #' @export
-f_transition_matrix <- function(f, p, x_grid, h_grid=NULL){
+f_transition_matrix <- function(f, p, x_grid, h_grid = NULL, sigma_g){
   if(is.null(h_grid))
     h_grid <- x_grid
   

### 2012-12-10 inst/examples/mcmc-gp.Rmd
trouble-shooting priors

@@
 theme_update(plot.background = element_rect(fill = "transparent",colour = NA))
+```
+
+```{r libraries}
+library(MCMCpack)
+library(tgp)
+library(reshape2)
+library(ggplot2)
 ```
@@
 
-I might try something like
+## Priors
 
 ```{r}
-library(tgp)
+s2.p <- c(50,50)
+tau2.p <- c(20,1)
+d.p = c(10, 1/0.01, 10, 1/0.01)
+nug.p = c(10, 1/0.01, 10, 1/0.01)
+```
+
+Define as random samples (avoids having to specify range)
+
+```{r}
+s2_rprior <- function(x) rinvgamma(x, s2.p[1], s2.p[2])
+tau2_rprior <- function(x) rinvgamma(x, tau2.p[1], tau2.p[2])
+d_rprior <- function(x) rgamma(x, d.p[1], scale = d.p[2]) + rgamma(x, d.p[3], scale = d.p[4])
+nug_rprior <- function(x) rgamma(x, nug.p[1], scale = nug.p[2]) + rgamma(x, nug.p[3], scale = nug.p[4])
+beta0_rprior <- function(x, tau) rnorm(x, 0, tau)
+n = 1000
+rpriors <- data.frame(s2 = s2_rprior(n), tau2 = tau2_rprior(n), beta0 = beta0_rprior(n, 1), nug = nug_rprior(n), d = d_rprior(n))
+rpriors <- melt(rpriors)
+ggplot(rpriors) + geom_histogram(aes(value)) + facet_wrap(~variable, scale="free")
+````
+
+Define as curves
+
+```{r}
+s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
+tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
+d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
+nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
+beta0_prior <- function(x, tau) dnorm(x, 0, tau)
+
+xx <- seq(.0001, 5, length.out=100)
+priors <- data.frame(x = xx, s2 = s2_prior(xx), tau2 = tau2_prior(xx), beta0 = beta0_prior(xx, 1), nug = nug_prior(xx), d = d_prior(xx))
+priors <- melt(priors, id="x")
+ggplot(priors) + geom_line(aes(x, value)) + facet_wrap(~variable, scale="free")
+```
+
+```{r}
 gp <- bgp(X=x, XX=X, Z=y, verb=0,
-          meanfn="constant", bprior="bflat", BTE=c(1,2,1), m0r1=FALSE, 
+          meanfn="constant", bprior="b0", BTE=c(1,2000,1), m0r1=FALSE, 
           corr="exp", trace=TRUE, beta = 0,
-          s2.p = c(25,10), d.p = c(50, 10, 10, 10), nug.p = c(50, 10, 10, 10),
+          s2.p = s2.p, d.p = d.p, nug.p = nug.p,
           s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", 
-          tau2.lam = "fixed", tau2.p=c(5,10))
+          tau2.lam = "fixed", tau2.p = tau2.p)
 ```
@@
                  ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
-
 ```
@@
 ```{r gp-plot}
-library(ggplot2)
 ggplot(tgp_dat) +
@@
 ```
+
+Look at the actual parameter estimates from the trace
+
+```{r}
+gp$trace$XX[[1]][1999,]
+```
+
+Posteriors
+
+```{r}
+require(reshape2)
+hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
+posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
+ggplot(posteriors) + geom_histogram(aes(value)) + facet_wrap(~variable, scales="free")
+
+```
+
+## Compare to priors with `scale` misplaced as `rate` in `d`, `nug`
+
+Define as random samples (avoids having to specify range)
+
+```{r}
+s2_rprior <- function(x) rinvgamma(x, s2.p[1], s2.p[2])
+tau2_rprior <- function(x) rinvgamma(x, tau2.p[1], tau2.p[2])
+d_rprior <- function(x) rgamma(x, d.p[1], d.p[2]) + rgamma(x, d.p[3], d.p[4])
+nug_rprior <- function(x) rgamma(x, nug.p[1], nug.p[2]) + rgamma(x, nug.p[3], nug.p[4])
+beta0_rprior <- function(x, tau) rnorm(x, 0, tau)
+n = 1000
+rpriors <- data.frame(s2 = s2_rprior(n), tau2 = tau2_rprior(n), beta0 = beta0_rprior(n, 1), nug = nug_rprior(n), d = d_rprior(n))
+rpriors <- melt(rpriors)
+ggplot(rpriors) + geom_histogram(aes(value)) + facet_wrap(~variable, scale="free")
+````
+
+```{r}
+s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
+tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
+d_prior <- function(x) dgamma(x, d.p[1], d.p[2]) + dgamma(x, d.p[3], d.p[4])
+nug_prior <- function(x) dgamma(x, nug.p[1], nug.p[2]) + dgamma(x, nug.p[3], nug.p[4])
+beta0_prior <- function(x, tau) dnorm(x, 0, tau)
+
+xx <- seq(.0001, 2500, length.out=1000)
+priors <- data.frame(x = xx, s2 = s2_prior(xx), tau2 = tau2_prior(xx), beta0 = beta0_prior(xx, 1), nug = nug_prior(xx), d = d_prior(xx))
+priors <- melt(priors, id="x")
+ggplot(priors) + geom_line(aes(x, value)) + facet_wrap(~variable, scale="free")
+```

### 2012-12-10 inst/examples/tgp-priors.Rmd
Potential bug in hyperparameter priors for nug and d

--- 2012-12-05 inst/examples/mcmc-gp.Rmd
@@
-``` {r libs, include=FALSE}
-rm(list=ls())   
-require(nonparametricbayes)
-require(reshape2)
-require(ggplot2)
-require(data.table)
-require(tgp)
+Exploring the behavior and ifluence of the prior in the `tgp` package.
+It appears there may be a bug (or rather incorrect documentation) in how the mixed Gamma priors are defined for the nugget `nug` and range / length-scale `d`.
+
+```{r include=FALSE}
+opts_knit$set(upload.fun = socialR::notebook.url)
+opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE, message=FALSE)
+theme_update(plot.background = element_rect(fill = "transparent",colour = NA))
 ```
 
-```{r graphing-options, include=FALSE}
-opts_knit$set(upload.fun = socialR::notebook.url)
-opts_chunk$set(dev.args=list(bg="transparent"), tidy=FALSE, comment=NA)
-theme_notebook <- theme_grey() + 
-                  theme(plot.background = element_rect(fill = "transparent",colour = NA),
-                        panel.grid.minor=element_line(colour="transparent"))
-theme_set(theme_notebook)
+```{r libraries}
+library(MCMCpack)
+library(tgp)
+library(reshape2)
+library(ggplot2)
 ```
 
-```{r results="hide"}
-knit("~/Documents/code/nonparametric-bayes/inst/examples/beverton_holt_data.Rmd")
-true <- data.frame(x = x_grid, y = sapply(x_grid, f, 0, p))
-XX <- as.matrix(x_grid)
-X <- as.matrix(obs$x)
-Z <- as.matrix(obs$y)
+Consider some very simple data:
+
+```{r}
+X = seq(-5, 5, length=50)
+x = c(-4, -3, -1,  0,  2)
+y = c(-2,  0,  1,  2, -1)
 ```
 
-## Call to `tgp` function `bgp()`
+## Priors
 
-Tweak the hyperparameter prior's parameters (`s2.p`) to influence the GP inferred.
+We choose prior parameters that give us very strong priors around particular values.
 
 ```{r}
-gp <- bgp(X=X, XX=XX, Z=Z, meanfn="constant", bprior="b0", BTE=c(1000,6000,2), m0r1=TRUE, verb=4, corr="exp", trace=TRUE, s2.p=c(5,10), tau2.p=c(5,10), s2.lam="fixed", tau2.lam="fixed")
+s2.p <- c(50,50) # InvGamma, approaches Gaussian approaching delta-fn around 1
+tau2.p <- c(20,1) # InvGamma, approaches exponential approaching delta-fn around 0
+d.p = c(10, 0.01, 10, 0.01) # Mixed Gammas, approaches Gaussian around 0.1, with var .001
+nug.p = c(10, 0.1, 10, 0.1) # Mixed Gammas, approaches Gaussian around 1, with var .1
 ```
 
-## parsing the function call
-
-* `X` is the observed X values (or matrix of appropriate dimension)
-* `XX` the desired predicted grid. (will also predict on X)
-* `Z` is observed Y values (or matrix of appropriate dimension)
-* `m0r1` means scale data to mean zero and unit range
-* `bprior` is the prior distribution, defaulting to `bfalt`.
-`b0` is a Gaussian.
-This prior is for the Gaussian process and is not to be confused with the hyperpriors describing the distribution of various parameters.
-* `BTE` Burn-in B steps, Terminate after T steps, sample Every E steps. 
-* `verb` Very verbose output
-* Other options (for adaptive/query learning, etc, igored for now.  Focus on priors).  
-* The priors for the hyperparameters are given by `s2.p`, `tau.p` etc (both are Inverse Gammas).  Their parameters can vary hiearchically in general, but we hold them fixed, `s2.lam = 'fixed'`, etc.)
-
-## Hyperparameters
-
-Hyperparameter distributions returned in the trace:
+Visualize the priors on lengthscale and nugget:
 
 ```{r}
-names(gp$trace$XX[[1]][2500,])
+d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
+nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
+xx <- seq(.0001, 2, length.out=100)
+priors <- data.frame(x = xx, nug = nug_prior(xx), d = d_prior(xx))
+priors <- melt(priors, id="x")
+ggplot(priors) + geom_line(aes(x, value)) + facet_wrap(~variable, scale="free")
 ```
 
-* `s2` is $\sigma^2$ in Gramacy and $\tau$ in Rassmussen and Williams, is the scale of the noise. Comes from an inverse gamma prior.  Set by inverse-gamma parameters `a0` and `g0`.
-
-* `tau2` Also comes from an inverse gamma prior. It is also set by inverse-gamma parameters `a0` and `g0`.
-
-$$Z | \beta, \sigma^2, K \sim N_n(\mathbf{F} \beta, \sigma^2 \mathbf{K}) $$
-$$ \beta | \sigma^2, \tau^2, \mathbf{W}, \beta_0 ~ N_m(\beta_0, \sigma^2 \tau^2 \mathbf{W} )  $$
-$$\beta_0 | N_m(\mathbf{\mu}, \mathbf{B})$$
-
-The parameters in this run are constant across the trace points (because hyperparameters are fixed).
+Solve the gp with these priors:
 
 ```{r}
-gp$trace$XX[[1]][2500,]
-gp$trace$XX[[3]][2500,]
+gp <- bgp(X=x, XX=X, Z=y, verb=0,
+          meanfn="constant", bprior="b0", BTE=c(1,2000,1), m0r1=FALSE, 
+          corr="exp", trace=TRUE, beta = 0,
+          s2.p = s2.p, d.p = d.p, nug.p = nug.p,
+          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", 
+          tau2.lam = "fixed", tau2.p = tau2.p)
 ```
-
-```{r}
-pars <- melt(gp$trace$XX[[1]], id = "index")
-ggplot(pars, aes(index, value)) + geom_line() + facet_wrap(~variable, scales="free_y")
-````
-
-It is possible to calculate summary statistics and check out the distribution of these fellows.
-
-```{r}
-mean(gp$trace$XX[[1]][,"nug"])
-ggplot(pars, aes(value)) + geom_histogram() + facet_wrap(~variable, scales="free")
-```
-
-* We keep the heirachical priors fixed in the `tau.lm`, `s2.lm`, etc.
-`nug` fixed by default.
-Note this results in the the priors from which these are sampled are fixed: e.g.
-`gp$trace$hier$s2.a0`, `gp$trace$hier$s2.g0`
-
-* `nug.p` This is the "nugget," the measurement error.
-Comes from a mixture of gamma prior parameter (initial values) for the nugget parameter `c(a1,g1,a2,g2)` where `g1` and `g2` are scale (`1/rate`) parameters.
-_Default reduces to simple exponential prior._ Specifying `nug.p = 0` fixes the nugget parameter to the "starting" value in `gd[1]`, i.e., it is excluded from the MCMC
-
-* `index` is the sampling point (from `BTE`, we see we sample starting at step 1000 and ending at step 6000, recording every 2 steps, so there are $(T-B)/E = 2500$ index points).  
-* `lambda` possibly a mixing parameter in the importance sampler temperature??
-* `beta0` No mention in the manual
-* `d` The parameter for priors using the hierachical exponential distribution for the parameters `a1`, `a2`, `g1`, `g2`.  
-* `b` ??
-
-* Where are the hyperparameters for the correlation function (e.g. the length-scale for the Gaussian??)
-
-## Extracting the estimated process
-
-(The **MAP**, maximum a posteriori expected mean and variance).
 
@@
 V <- gp$ZZ.ks2
-dat <- data.frame(x   = gp$XX[[1]], 
+tgp_dat <- data.frame(x   = gp$XX[[1]], 
                   y   = gp$ZZ.km, 
@@
 
-Plot the posterior Gaussian Process:
+For comparison, let us manually estimate the GP with hyperparameters fixed at the mean value of the priors.
+
+```{r}
+d = .1; epsilon = .1; sigma = 1; #fixed hyperparamaters
+SE <- function(Xi,Xj, d) sigma * exp(- (Xi - Xj) ^ 2 / d)
+cov <- function(X, Y) outer(X, Y, SE, d) 
+cov_xx_inv <- solve(cov(x, x) + epsilon * diag(1, length(x)))
+Ef <- cov(X, x) %*% cov_xx_inv %*% y
+Cf <- cov(X, X) - cov(X, x)  %*% cov_xx_inv %*% cov(x, X)
+manual_dat <- data.frame(x = X, 
+                         y = Ef, 
+                         ymin = (Ef - 1.96 * sqrt(diag(Cf))), 
+                         ymax = (Ef + 1.96 * sqrt(diag(Cf))))
+```
+
+Compare the GP posteriors in a plot:
 
 ```{r gp-plot}
-  p1 <- ggplot(dat) +
-    geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
-    geom_line(aes(x=x,y=y), size=1) + #MEAN
-    geom_point(data=gp$obs,aes(x=x,y=y)) + 
-    labs(title=paste("llik =", prettyNum(gp$llik)))
-  if(!is.null(true))
-    p1 <- p1 + geom_line(data = true, aes(x, y), col = "red", lty = 2) 
-p1 + theme_notebook
+ggplot(tgp_dat) +
+    geom_ribbon(aes(x, y, ymin = ymin, ymax = ymax), fill="red", alpha = .1) + # Var
+    geom_line(aes(x, y), col="red") + # mean
+    geom_point(data = data.frame(x = x, y = y), aes(x, y)) + # raw data
+    geom_ribbon(dat = manual_dat, aes(x, y, ymin = ymin, ymax = ymax), fill = "blue", alpha = .1) + # Var
+    geom_line(dat = manual_dat, aes(x, y), col = "blue")  + #MEAN    
+    theme_bw() + theme(plot.background = element_rect(fill = "transparent",colour = NA))
 ```
 
-Find the actual optimum policy from the true parametric model
+Wait a minute.
+The Bayesian estimate of the lengthscale seems way too long, given our strongly informative prior.
+Likewise the measurement uncertainty appears to have induced a rather broad distribution.
+Let's look at the posteriors:
 
 ```{r}
-x_grid <- dat$x
-h_grid <- x_grid
-matrices_true <- f_transition_matrix(f, p, x_grid, h_grid)
-opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
-```
-
-Estimate a policy from the Gaussian Process:
-
-```{r}
-rownorm <- function(M) t(apply(M, 1, function(x) x/sum(x)))
-matrices_gp <- lapply(h_grid, function(h){
-    S <- dat$y - h
-    F_ <- t(sapply(1:length(S), function(i){
-      if(S[i]>0) {
-        out <- dlnorm(x_grid/S[i], 0, V[i])
-      } else {
-        out <- numeric(length(x_grid))
-        out[1] <- 1
-        out
-      }
-    }))
-    F_ <- rownorm(F_)
-  })
-opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
-```
-
-```{r}
-matrices_true[[1]][1:10,1:10]
-matrices_gp[[1]][1:10,1:10]
+require(reshape2)
+hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
+posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
+ggplot(posteriors) + geom_histogram(aes(value)) + facet_wrap(~variable, scales="free")
 
@@
 
-```{r policy_plot}
-policies <- melt(data.frame(stock=x_grid, GP = x_grid[opt_gp$D[,1]], Exact = x_grid[opt_true$D[,1]]), id="stock")
-policy_plot <- ggplot(policies, aes(stock, stock - value, color=variable)) +
-  geom_point() + xlab("stock size") + ylab("escapement") 
-policy_plot + theme_notebook
+Hmm...
+posteriors indeed show a huge value for the length scale `d`, and for the nugget `nug`.
+Looks supiciously like the mixed Gamma prior when the scale parameter is used as the rate parameter (it's inverse):
+
+## Compare to priors with `scale` misplaced as `rate` in `d`, `nug`
+
+```{r density_prior}
+d_prior <- function(x) dgamma(x, d.p[1], d.p[2]) + dgamma(x, d.p[3], d.p[4])
+nug_prior <- function(x) dgamma(x, nug.p[1], nug.p[2]) + dgamma(x, nug.p[3], nug.p[4])
+xx <- seq(.0001, 2500, length.out=1000)
+priors <- data.frame(x = xx, nug = nug_prior(xx), d = d_prior(xx))
+priors <- melt(priors, id="x")
+ggplot(priors) + geom_line(aes(x, value)) + facet_wrap(~variable, scale="free")
 ```
-
-We can see what happens when we attempt to manage a stock using this:
-
-``` {r othernoise}
-z_g <- function() rlnorm(1,0, sigma_g)
-````
-
-``` {r  simulate }
-set.seed(1)
-sim_gp <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_gp$D, z_g, profit=profit)
-set.seed(1)
-sim_true <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_true$D, z_g, profit=profit)
-````
-
-``` {r  simplot}
-df <- data.frame(time = sim_gp$time, stock_gp = sim_gp$fishstock, stock_true = sim_true$fishstock, harvest_gp = sim_gp$harvest, havest_true = sim_true$harvest)
-df <- melt(df, id="time")
-simplot <- ggplot(df) + geom_line(aes(time,value, color=variable))
-simplot + theme_notebook
-````
-
-Total Profits
-
-```{r}
-sum(sim_gp$profit)
-sum(sim_true$profit)
-```

### 2012-12-11 inst/examples/may79-example.Rmd
GP that results in scorched earth fishing strategy

@@
 ```{r libraries}
-require(pdgControl)
-require(nonparametricbayes)
-require(reshape2)
-require(ggplot2)
-require(data.table)
+library(pdgControl)
+library(nonparametricbayes)
+library(reshape2)
+library(ggplot2)
+library(data.table)
 library(kernlab)
+library(tgp)
 ```
@@
 
+State-space model
+
 ```{r}
@@
 K <- 8 # approx
+```
+
+Parameters
+
+```{r}
+x_grid = seq(0, 1.5 * K, length=101)
+T <- 40
+sigma_g <- 0.3
+x <- numeric(T)
+x[1] <- 1
 ```
@@
 
-Parameter definitions
-
-```{r}
-x_grid = seq(0, 1.5 * K, length=101)
-T <- 40
-sigma_g <- 0.3
-x <- numeric(T)
-x[1] <- 1
-```
-
 Simulation
@@
 X <- x_grid
-gp <- gp_fit(obs, X, c(sigma_n=.9, tau = 1., l=.1), "kernlab")
-
 ```
@@
 ```{r}
-df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
+gp <- bgp(X=obs$x, XX=X, Z=obs$y, verb=0,
+          meanfn="linear", bprior="b0", BTE=c(2000,6000,2), m0r1=FALSE, 
+          corr="exp", trace=TRUE, beta = c(0,0),
+          s2.p = c(50,50), d.p = c(10, 1/0.01, 10, 1/0.01), nug.p = c(10, 1/0.01, 10, 1/0.01),
+          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", 
+          tau2.lam = "fixed", tau2.p = c(50,1))
+```
+
+Extract the posterior Gaussian process mean and the $\pm 2$ standard deviations over the predicted grid from the fit:
+
+```{r}
+V <- gp$ZZ.ks2
+Ef = gp$ZZ.km
+tgp_dat <- data.frame(x   = gp$XX[[1]], 
+                  y   = gp$ZZ.km, 
+                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
+                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
+```
+
+```{r}
 true <- data.frame(x=X, y=sapply(X,f, 0, p))
-ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
+ggplot(tgp_dat)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
   geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
@@
 ```{r}
-matrices_gp <- gp_transition_matrix(gp)
+matrices_gp <- gp_transition_matrix(Ef, sqrt(V), x_grid, h_grid)
 opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)

### 2012-12-11 R/gp_trasition_matrix.R
transition matrix method for GP takes Ef, Cf explicitly now.

@@ -1,21 +1,20 @@
   
 #' Determine the transition matrix given a Guassian process
 #' 
-#' @param gp a gpfit type object, a list with the GP mean gp$Ef, Covariance gp$Cf on grid gp$X
-#' @param h_grid the harvest grid over which to compute.  Assumed to match the stock grid gp$X if not given.
+#' @param Ef GP mean
+#' @param Sf GP standard deviations (sqrt(diag(Cf))
+#' @param x_grid x grid over which to define transition matrices
+#' @param h_grid the harvest grid over which to compute.  Assumed to match the stock grid x_grid if not given.
 #' @return a list of matrices corresponding to the state transition matrix at each harvest level h
 #' @export
-gp_transition_matrix <- function(gp, h_grid=NULL){
-  # consider taking x_grid, Ef, & Cf as arguments instead?
-  x_grid <- gp$X
+gp_transition_matrix <- function(Ef, Sf, x_grid, h_grid=NULL){  
   if(is.null(h_grid))
     h_grid <- x_grid
-  V <- sqrt(diag(gp$Cf))
   matrices_gp <- lapply(h_grid, function(h){
-    S <- gp$Ef - h
+    S <- Ef - h
     F_ <- t(sapply(1:length(S), function(i){
       if(S[i]>0) {
-        out <- dlnorm(x_grid/S[i], 0, V[i])
+        out <- dlnorm(x_grid/S[i], 0, Sf[i])
       } else {
         out <- numeric(length(x_grid))
         out[1] <- 1

### 2012-12-12 inst/examples/multidimensional.Rmd
a simple multi-demensional example (no policy function yet) #7

multi-dimensional dynamics example
========================================================

```{r libraries}
library(reshape2)
library(ggplot2)
library(data.table)
library(kernlab)
library(tgp)
```

```{r graphing-options, include=FALSE}
opts_knit$set(upload.fun = socialR::notebook.url)
opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE)
theme_set(theme_bw())
theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
             plot.background = element_rect(fill = "transparent",colour = NA))
```

$$\begin{align}
\frac{dM}{dt} = aMC - \frac{(g-h) M}{M+T} + \gamma M T \\
\frac{dC}{dt} = rTC - dC - a M C 
\end{align}$$

```{r}              
mumby <- function(x, h, p= c(a = 0.1, g = .5, gamma = 0.8, r = 1, d = 0.44)){
  M <- x[1] # Macroalgae
  C <- x[2] # Corals
  a <- p[1] # algal growth rate
  g <- p[2] # grazing rate (reduced by harvesting)
  gamma <- p[3] # algal colinization of dead tufts
  r <- p[4] # Coral growth rate
  d <- p[5] # Coral death rate
  dt <- 0.025 # 
  M_t <- M * exp( dt * (a * M * C - (g - h) * M / (M + (1-M-C)) + gamma * M * (1-M-C)) )
  C_t <- C * exp( dt * (r * (1-M-C) * C - d * C - a * M * C) )
  c(M_t, C_t)
}

```

```{r}
Tmax <- 2500
X <- matrix(NA, nrow=Tmax, ncol=2)
colnames(X) = c("M", "C")
```

```{r}
X[1,] = c(0.9,0.9)
for(i in 1:(Tmax-1))
  X[i+1,] = mumby(X[i,],0)
dat <- data.frame(time = 1:Tmax, X)
dat <- dat[seq(1,2500,by=25),]
```

```{r}
ggplot(dat) + geom_line(aes(time, M), col="red") + geom_line(aes(time, C), col="blue")
```

```{r phase-space}
inits <- as.matrix(expand.grid(seq(0.05,0.6,length=11), seq(0.05,0.6, length=11)))
df <- lapply(1:round(length(inits[,1])), function(j){
  X[1,] = inits[j,]
  for(i in 1:(Tmax-1))
    X[i+1,] = mumby(X[i,],0.2)
  data.frame(time = 1:Tmax, X)
  })
df2 <- melt(df, id=c("time", "M", "C"))
ggplot(df2) + geom_line(aes(M, C, group=L1)) + 
  geom_point(data = subset(df2, time==1), aes(M, C, group=L1), col="red")
```

Noisy data example

```{r}
sigma_g = 0.1
X[1,] = c(0.9,0.9)
for(i in 1:(Tmax-1))
  X[i+1,] = rlnorm(1,0, sigma_g) * mumby(X[i,],0)
dat <- data.frame(time = 1:Tmax, X)
dat <- dat[seq(1,2500,by=25),]
```

Multi-dimensional GP, `tgp` method, based on time-delay data:

```{r delay-equations}
n <- length(dat[,1])
obs <- data.frame(x1=dat[2:(n-1),2], x2 = dat[1:(n-2),2], y=dat[3:n,2])
x_grid <- seq(0,1,length.out=100)
```

```{r}
require(rgl)
plot3d(dat$time, dat$M, dat$C)
plot3d(obs$x1, obs$x2, obs$y)
```

```{r}
gp <- bgp(X=cbind(obs$x1, obs$x2), XX=cbind(x_grid,x_grid), Z=cbind(obs$y), verb=0,
          meanfn="linear", bprior="b0", BTE=c(2000,6000,2), m0r1=FALSE,
          corr="exp", trace=TRUE, beta = c(0,0,0),
          s2.p = c(50,50), d.p = c(10, 1/0.01, 10, 1/0.01), nug.p = c(10, 1/0.01, 10, 1/0.01),
          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", 
          tau2.lam = "fixed", tau2.p = c(50,1))
```

```{r}
plot(gp, layout="surf")
```

Extract the posterior Gaussian process mean and the $\pm 2$ standard deviations over the predicted grid from the fit:

```{r}
V <- gp$ZZ.ks2
Ef = gp$ZZ.km
tgp_dat <- data.frame(x1   = gp$XX[[1]], 
                      x2 =  gp$XX[[2]],
                  y   = gp$ZZ.km, 
                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
```

```{r}
ggplot(tgp_dat)  + geom_ribbon(aes(x1,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x1,y)) + geom_point(data=obs, aes(x1,y)) 
```

```{r}
ggplot(tgp_dat)  + geom_ribbon(aes(x2,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x2,y)) + geom_point(data=obs, aes(x2,y)) 
```

`kernlab` method

### 2012-12-12 inst/examples/multidimensional.Rmd
another example (includes some data in the higher range) #7

@@
 
-```{r phase-space}
+```{r phase-space, cache=TRUE, cache.path="phasespace/"}
 inits <- as.matrix(expand.grid(seq(0.05,0.6,length=11), seq(0.05,0.6, length=11)))
@@
 
-```{r}
+```{r x1}
 ggplot(tgp_dat)  + geom_ribbon(aes(x1,y,ymin=ymin,ymax=ymax), fill="gray80") +
@@
 
-```{r}
+```{r x2}
 ggplot(tgp_dat)  + geom_ribbon(aes(x2,y,ymin=ymin,ymax=ymax), fill="gray80") +

### 2012-12-12 inst/examples/may79-example.Rmd
Exploring GP behavior and origin of deviations

@@
 T <- 40
-sigma_g <- 0.3
+sigma_g <- 0.6
 x <- numeric(T)
@@
 z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
-profit <- profit_harvest(1,0,0)
+#profit <- profit_harvest(1,0,0)
+
+profit = function(x,h) pmin(x, h)
 ```
@@
 
-GP Solution
+## Understanding the GP Solution
 
 ```{r}
-matrices_gp <- gp_transition_matrix(Ef, sqrt(V), x_grid, h_grid)
+h <- x_grid[1]
+X <- numeric(length(x_grid))
+X[18] = 1
+S <- Ef-h
+ F_ <- t(sapply(1:length(S), function(i){
+      if(S[i]>0) {
+        out <- dnorm(x_grid, S[i], sqrt(V[i]))
+#        negatives <- sum(dnorm(seq(-K,0, length=50), S[i], sqrt(V[i])))
+#        out[1] <- out[1] + negatives
+        out <- out/sum(out)
+      } else {
+        out <- numeric(length(x_grid))
+        out[1] <- 1
+        out
+      }
+    }))
+xt1 <- X %*% F_
+
+xt10 <- xt1
+for(s in 1:20)
+  xt10 <- xt10 %*% F_
+
+qplot(x_grid, xt1[1,])
+```
+
+whereas
+
+```{r}
+rownorm <- function(M) t(apply(M, 1, function(x) x/sum(x)))
+
+mu <- sapply(x_grid, f, h, p)
+F_true <- t(sapply(mu, function(m){
+      if(m>0) {
+        out <- dlnorm(x_grid/m, 0, sigma_g)
+
+      } else {
+        out <- numeric(length(x_grid))
+        out[1] <- 1
+        out
+      }
+    }))
+    F_true <- rownorm(F_true)
+F_true <- rownorm(F_true)
+yt1 <- X %*% F_true
+
+yt10 <- yt1
+for(s in 1:10)
+  yt10 <- yt10 %*% F_true
+qplot(x_grid, yt10[1,])
+```
+
+```{r}
+transition <- melt(data.frame(x = x_grid, gp = xt1[1,], parametric = yt1[1,]), id="x")
+ggplot(transition) + geom_point(aes(x,value, col=variable))
+```
+
+```{r}
+matrices_gp <- gp_transition_matrix(Ef, V, x_grid, h_grid)
 opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
+```
+
+```{r}
+F <- matrices_gp
+Tmax <- 20
+    n_x <- length(x_grid)
+    n_h <- length(h_grid)
+    D <- matrix(NA, nrow=n_x, ncol=Tmax)  
+    P <-  outer(x_grid, h_grid, profit)   
+    V <- P
+    for(t in 1:Tmax){
+      D[,(Tmax-t+1)] <- apply(V, 1, which.max) 
+      v_t <- apply(V, 1, max) # vector of current values 
+      V <- sapply(1:n_h, function(j) # updated value matrix
+        P[,j] + (1-delta) * F[[j]] %*% v_t)
+    }
 ```

### 2012-12-13 R/gp_trasition_matrix.R
gp transition matrix filled out on basis of normally distributed errors

@@ -2,19 +2,20 @@
 #' Determine the transition matrix given a Guassian process
 #' 
 #' @param Ef GP mean
-#' @param Sf GP standard deviations (sqrt(diag(Cf))
+#' @param V GP vector of variances
 #' @param x_grid x grid over which to define transition matrices
 #' @param h_grid the harvest grid over which to compute.  Assumed to match the stock grid x_grid if not given.
 #' @return a list of matrices corresponding to the state transition matrix at each harvest level h
 #' @export
-gp_transition_matrix <- function(Ef, Sf, x_grid, h_grid=NULL){  
+gp_transition_matrix <- function(Ef, V, x_grid, h_grid=NULL){  
   if(is.null(h_grid))
     h_grid <- x_grid
   matrices_gp <- lapply(h_grid, function(h){
     S <- Ef - h
     F_ <- t(sapply(1:length(S), function(i){
       if(S[i]>0) {
-        out <- dlnorm(x_grid/S[i], 0, Sf[i])
+        out <- dnorm(x_grid, S[i], sqrt(V[i]))  
+        out <- out/sum(out)  # handled by rownorm already
       } else {
         out <- numeric(length(x_grid))
         out[1] <- 1
@@ -44,6 +45,7 @@ f_transition_matrix <- function(f, p, x_grid, h_grid = NULL, sigma_g){
     F_true <- t(sapply(mu, function(m){
       if(m>0) {
         out <- dlnorm(x_grid/m, 0, sigma_g)
+        
       } else {
         out <- numeric(length(x_grid))
         out[1] <- 1

### 2012-12-13 inst/examples/reed-example.Rmd
transition matrix method for GP takes Ef, Cf explicitly now.
basic examples managing a beverton holt system
clear example where GP performs nearly optimally while ricker dynamics in place of BH perform poorly.

@@
 require(data.table)
+require(tgp)
 library(kernlab)
@@
 
+Parameters
+
+```{r}
+x_grid <- seq(0, 1.5 * K, length=101)
+sigma_g <- 0.01
+delta <- 0.01
+OptTime = 20
+reward = numeric(length(x_grid))
+xT = x_grid[4]
+reward = profit(x_grid[101],x_grid[101]) + 1/(1-delta)^OptTime # reward for more than 
+```
+
+Noise function, profit function
+```{r}
+z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
+#profit <- profit_harvest(1,0,0)
+profit = function(x,h) pmin(x, h)
+```
+
+Simulation
+
+```{r}
+Tobs <- 100
+x <- numeric(Tobs)
+x[1] <- x_grid[2]
+for(t in 1:(Tobs-1))
+  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
+plot(x)
+```
+
+Very crude estimate of parameter values
+
+```{r}
+K = mean(x)
+r = 0.5 * sigma_g^2 / var(x)
+
+#cor <- acf(x)$acf[,,1]
+#m <- which(cor < 0)[1] - 1
+#l <- lm(y ~ x, data=data.frame(x = 1:6, y = log(cor[1:6])))
+#r <- - as.numeric(l$coefficients[2] / 2)
+
+f_alt <- Ricker
+p_alt <- p
+```
+
 Predict the function over the target grid
@@
 ```{r}
-gp <- gp_fit(obs, X, c(sigma_n=2, tau=10, l=8), "conditional")
-plot.gpfit(gp, true)
+obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
+X <- x_grid
+```
+
+Gaussian Process inference from this model
+
+```{r}
+gp <- bgp(X=obs$x, XX=X, Z=obs$y, verb=0,
+          meanfn="linear", bprior="b0", BTE=c(2000,6000,2), m0r1=FALSE, 
+          corr="exp", trace=TRUE, beta = c(0,0),
+          s2.p = c(50,50), d.p = c(10, 1/0.01, 10, 1/0.01), nug.p = c(10, 1/0.01, 10, 1/0.01),
+          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", 
+          tau2.lam = "fixed", tau2.p = c(50,1))
+```
+
+Extract the posterior Gaussian process mean and the $\pm 2$ standard deviations over the predicted grid from the fit:
+
+```{r}
+V <- gp$ZZ.ks2
+Ef = gp$ZZ.km
+tgp_dat <- data.frame(x   = gp$XX[[1]], 
+                  y   = gp$ZZ.km, 
+                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
+                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
+```
+
+```{r}
+true <- data.frame(x=X, y=sapply(X,f, 0, p))
+ggplot(tgp_dat)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
+  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
+  geom_line(data=true, aes(x,y), col='red', lty=2)
 ```
@@
 
-GP Solution
+## Understanding the GP Solution
 
 ```{r}
-matrices_gp <- gp_transition_matrix(gp)
-opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
+h <- x_grid[1]
+X <- numeric(length(x_grid))
+X[18] = 1
+S <- Ef-h
+ F_ <- t(sapply(1:length(S), function(i){
+      if(S[i]>0) {
+        out <- dnorm(x_grid, S[i], sqrt(V[i]))
+#        negatives <- sum(dnorm(seq(-K,0, length=50), S[i], sqrt(V[i])))
+#        out[1] <- out[1] + negatives
+        out <- out/sum(out)
+      } else {
+        out <- numeric(length(x_grid))
+        out[1] <- 1
+        out
+      }
+    }))
+xt1 <- X %*% F_
+
+xt10 <- xt1
+for(s in 1:20)
+  xt10 <- xt10 %*% F_
+
+qplot(x_grid, xt10[1,]) + geom_point(aes(y=xt1[1,]), col="grey")
+
+```
+
+whereas
+
+```{r}
+rownorm <- function(M) t(apply(M, 1, function(x) x/sum(x)))
+
+mu <- sapply(x_grid, f, h, p)
+F_true <- t(sapply(mu, function(m){
+      if(m>0) {
+        out <- dlnorm(x_grid/m, 0, sigma_g)
+
+      } else {
+        out <- numeric(length(x_grid))
+        out[1] <- 1
+        out
+      }
+    }))
+    F_true <- rownorm(F_true)
+F_true <- rownorm(F_true)
+yt1 <- X %*% F_true
+
+yt10 <- yt1
+for(s in 1:20)
+  yt10 <- yt10 %*% F_true
+qplot(x_grid, yt10[1,]) + geom_point(aes(y=yt1[1,]), col="grey")
+```
+
+```{r}
+transition <- melt(data.frame(x = x_grid, gp = xt1[1,], parametric = yt1[1,]), id="x")
+ggplot(transition) + geom_point(aes(x,value, col=variable))
+```
+
+```{r}
+matrices_gp <- gp_transition_matrix(Ef, V, x_grid, h_grid)
+opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=1e6)
+```
+
+```{r}
+F <- matrices_gp
+Tmax <- 20
+    n_x <- length(x_grid)
+    n_h <- length(h_grid)
+    D <- matrix(NA, nrow=n_x, ncol=Tmax)  
+    P <-  outer(x_grid, h_grid, profit)   
+    V <- P
+    for(t in 1:Tmax){
+      D[,(Tmax-t+1)] <- apply(V, 1, which.max) 
+      v_t <- apply(V, 1, max) # vector of current values 
+      V <- sapply(1:n_h, function(j) # updated value matrix
+        P[,j] + (1-delta) * F[[j]] %*% v_t)
+    }
 ```
@@
 matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
+opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
 ```
@@
 ```{r}
-k <- 15
-A <- 1.5
-B <- (A - 1)/k
-matrices_estimated <- f_transition_matrix(f, c(A, B), x_grid, h_grid, sigma_g)
-opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, 20, 0, profit, delta=.01)
+matrices_estimated <- f_transition_matrix(f_alt, p_alt, x_grid, h_grid, sigma_g)
+opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
 ```
@@
 
-``` {r  simplot}
+``` {r  sim-fish}
 ggplot(dt) + geom_line(aes(time,fishstock, color=method))
+```
+
+```{r sim-harvest}
 ggplot(dt) + geom_line(aes(time,harvest, color=method))

### 2012-12-13 inst/examples/may79-example.Rmd
clear example where GP performs nearly optimally while ricker dynamics in place of BH perform poorly.

@@
 f <- May
-p <- c(r = .75, k = 10, a=1.3, H=1, Q = 3)
-K <- 8 # approx
+p <- c(r = .75, k = 10, a=1.5, H=1, Q = 3)
+K <- 7.5 # approx
 ```
@@
 ```{r}
-x_grid = seq(0, 1.5 * K, length=101)
-T <- 40
-sigma_g <- 0.6
-x <- numeric(T)
-x[1] <- 1
+x_grid <- seq(0, 1.5 * K, length=101)
+sigma_g <- 0.3
+delta <- 0.01
+OptTime = 20
+reward = numeric(length(x_grid))
+xT = x_grid[4]
+reward = profit(x_grid[101],x_grid[101]) + 1/(1-delta)^OptTime # reward for more than 
 ```
@@
 #profit <- profit_harvest(1,0,0)
-
 profit = function(x,h) pmin(x, h)
@@
 ```{r}
-x[1] = 2.5
-for(t in 1:(T-1))
+Tobs <- 100
+x <- numeric(Tobs)
+x[1] <- K
+for(t in 1:(Tobs-1))
   x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
 plot(x)
+```
+
+Very crude estimate of parameter values
+
+```{r}
+K = mean(x)
+r = 0.5 * sigma_g^2 / var
+
+#cor <- acf(x)$acf[,,1]
+#m <- which(cor < 0)[1] - 1
+#l <- lm(y ~ x, data=data.frame(x = 1:6, y = log(cor[1:6])))
+#r <- - as.numeric(l$coefficients[2] / 2)
+
+f_alt <- Ricker
+p_alt <- p
 ```
@@
 
-qplot(x_grid, xt1[1,])
+qplot(x_grid, xt10[1,]) + geom_point(aes(y=xt1[1,]), col="grey")
+
 ```
@@
 yt10 <- yt1
-for(s in 1:10)
+for(s in 1:20)
   yt10 <- yt10 %*% F_true
-qplot(x_grid, yt10[1,])
+qplot(x_grid, yt10[1,]) + geom_point(aes(y=yt1[1,]), col="grey")
 ```
@@
 matrices_gp <- gp_transition_matrix(Ef, V, x_grid, h_grid)
-opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
+opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=1e6)
 ```
@@
 ```{r}
-matrices_estimated <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g=0.01)
-opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, 20, 0, profit, delta=.01)
+matrices_estimated <- f_transition_matrix(f_alt, p_alt, x_grid, h_grid, sigma_g)
+opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
 ```

### 2012-12-13 inst/examples/reed-example.Rmd
clear example where GP performs nearly optimally while ricker dynamics in place of BH perform poorly.

@@
 require(data.table)
-require(tgp)
 library(kernlab)
@@
 K = mean(x)
-r = 0.5 * sigma_g^2 / var(x)
+r = 0.5 * sigma_g^2 / var
 

### 2012-12-15 inst/examples/myer-example.Rmd
Myers example of GP peformance with adequate data

@@
-```{r libraries}
-require(pdgControl)
-require(nonparametricbayes)
-require(reshape2)
-require(ggplot2)
-require(data.table)
-library(kernlab)
+```{r set-options, echo = FALSE, cache = FALSE, external = TRUE}
+opts_chunk$set(external = TRUE, cache = TRUE, cache.path = "myers-cache/")
+read_chunk('gaussian-process-control.R')
+library(knitcitations)
+```
+
+```{r libraries, include=FALSE}
 ```
@@
 ```{r graphing-options, include=FALSE}
-opts_knit$set(upload.fun = socialR::notebook.url)
-opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE)
-theme_set(theme_bw())
-theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
-             plot.background = element_rect(fill = "transparent",colour = NA))
 ```
 
-```{r}
-f <- Myer_harvest
-pars <- c(1, 2, 6) 
-p <- pars # shorthand 
-K <- p[1] * p[3] / 2 + sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2
-xT <- p[1] * p[3] / 2 - sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 # allee threshold
-e_star <- (p[1] * sqrt(p[3]) - 2) / 2 ## Bifurcation point 
+```{r Myer}
 ```
 
-Noise function, profit function
-```{r}
-z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
-profit <- profit_harvest(1,0,0)
+We use the model of `r citet("10.1126/science.269.5227.1106")`.
+
+```{r sdp-pars}
 ```
 
-Parameter definitions
+With parameters `r p`.
 
-```{r}
-x_grid = seq(0, 1.5 * K, length=101)
-T <- 40
-sigma_g <- 0.2
-x <- numeric(T)
-x[1] <- 1
-seed <- 123
+```{r sim-obs}
 ```
 
-Simulation
-
-```{r}
-x[1] = 2.5
-for(t in 1:(T-1))
-  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
-plot(x)
+```{r lag-data}
 ```
 
-Predict the function over the target grid
-
-```{r}
-obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
-X <- x_grid
-gp <- gp_fit(obs, X, c(sigma_n=.9, tau = 1., l=.1), "kernlab")
-
+```{r par-est}
 ```
 
-Gaussian Process inference from this model
-
-```{r}
-df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
-true <- data.frame(x=X, y=sapply(X,f, 0, p))
-ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
-  geom_line(data=true, aes(x,y), col='red', lty=2)
+```{r gp-fit, dependson="lag-data"}
 ```
 
-## Stochastic Dynamic programming solution based on the posterior Gaussian process:
+We fit a Gaussian process with
 
-```{r}
-h_grid <- x_grid
+```{r gp-data, dependson="gp-fit"}
 ```
 
-GP Solution
-
-```{r}
-matrices_gp <- gp_transition_matrix(gp)
-opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
+```{r gp-plot, dependson="gp-data"}
 ```
 
-Optimal solution
+## The transition matrix of the inferred process
 
-```{r}
-matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
+```{r gp-F-sim, dependson="gp-data"}
 ```
 
-Optimal parametric solution in which a parameter is estimated with error
+```{r par-F-sim}
+```
 
-```{r}
-p_est = c(1.5,2,6)
-matrices_estimated <- f_transition_matrix(f, p_est, x_grid, h_grid, sigma_g=sigma_g)
-opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, 20, 0, profit, delta=.01)
+```{r F-sim-plot}
+```
+
+```{r gp-opt, dependson="gp-data"}
+```
+
+```{r true-opt}
+```
+
+```{r est-opt}
 ```
@@
 ```{r policy_plot}
-policies <- melt(data.frame(stock=x_grid, 
-                            GP = x_grid[opt_gp$D[,1]], 
-                            Exact = x_grid[opt_true$D[,1]],
-                            Approx = x_grid[opt_estimated$D[,1]]),
-                  id="stock")
-
-policy_plot <- ggplot(policies, aes(stock, stock - value, color=variable)) +
-  geom_point() + xlab("stock size") + ylab("escapement") 
-policy_plot
 ```
 
-We can see what happens when we attempt to manage a stock using this:
-
-``` {r othernoise}
-z_g <- function() rlnorm(1,0, sigma_g)
-````
-
-``` {r  simulate }
-set.seed(seed)
-sim_gp <- lapply(1:20, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_gp$D, z_g, profit=profit))
-set.seed(seed)
-sim_true <- lapply(1:20, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_true$D, z_g, profit=profit))
-set.seed(seed)
-sim_est <- lapply(1:20, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_estimated$D, z_g, profit=profit))
-````
-
-```{r}
-dat <- list(est = sim_est, gp = sim_gp, true = sim_true)
-dat <- melt(dat, id=names(dat[[1]][[1]]))
-dt <- data.table(dat)
-setnames(dt, c("L1", "L2"),  c("method", "reps")) 
+```{r othernoise}
 ```
 
-And plot the resulting stock dynamcis across the replicates [^1]
+```{r simulate}
+```
 
-``` {r  simplot}
-ggplot(dt) + geom_line(aes(time,fishstock, group=reps, color=method)) + facet_wrap(~method)
-````
+```{r tidy}
+```
 
-[^1]: Not sure why that doesn't plot correctly as overlay when I drop the `facet_wrap` option from the command.
+```{r sim-fish}
+```
 
-Total Profits
+```{r sim-harvest}
+```
 
-```{r}
-c( gp = sum(sim_gp$profit), true = sum(sim_true$profit), est = sum(sim_est$profit))
+```{r costs}
 ```
+
+```{r echo=FALSE, results="as.is"}
+bibliography("html")
+```

### 2012-12-15 inst/examples/gaussian-process-control.R

@@ -1,29 +1,39 @@
 
+## @knitr libraries
+require(pdgControl)
+require(nonparametricbayes)
+require(reshape2)
+require(ggplot2)
+require(data.table)
+require(tgp)
+library(kernlab)
 
 
 
+## @knitr graphing-options
+opts_knit$set(upload.fun = socialR::notebook.url)
+opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE)
+theme_set(theme_bw())
+theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
+             plot.background = element_rect(fill = "transparent",colour = NA))
 
 
-
-
-
-
-
-```r
+## @knitr Myer
 f <- Myer_harvest
 pars <- c(1, 2, 6) 
 p <- pars # shorthand 
 K <- p[1] * p[3] / 2 + sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2
 allee <- p[1] * p[3] / 2 - sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 # allee threshold
 e_star <- (p[1] * sqrt(p[3]) - 2) / 2 ## Bifurcation point 
-```
 
 
-We use the model of Myers _et. al._ (1995).
+## @knitr true-data
+true <- data.frame(x = X, y = sapply(X, f, 0, p))
 
 
 
-```r
+
+## @knitr sdp-pars
 sigma_g <- 0.02
 z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
 x_grid <- seq(0, 1.5 * K, length=101)
@@ -38,140 +48,98 @@ reward = profit(x_grid[length(x_grid)], x_grid[length(x_grid)]) + 1 / (1 - delta
 ## above Allee threshold for Allee model, 
 ## and near zero for BH or Ricker models
 x_0_observed <- allee + x_grid[1] 
-```
 
 
-With parameters `1, 2, 6`. 
 
-
-```r
+## @knitr sim-obs
 Tobs <- 100
 x <- numeric(Tobs)
 x[1] <- x_0_observed
 for(t in 1:(Tobs-1))
   x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
 plot(x)
-```
 
[…]

### 2012-12-15 inst/examples/reed-example.Rmd
bh example

--- 2012-12-10 inst/examples/reed-example.Rmd
@@
-Reed (1979) with Gaussian processes 
-========================================================
+```{r set-options, echo = FALSE, cache = FALSE, external = TRUE}
+opts_chunk$set(external = TRUE, cache = TRUE, cache.path = "bevholt-cache/")
+read_chunk('gaussian-process-control.R')
+library(knitcitations)
+```
+
+```{r libraries, include=FALSE}
+```
 
 ```{r graphing-options, include=FALSE}
-require(pdgControl)
-require(nonparametricbayes)
-require(reshape2)
-require(ggplot2)
-require(data.table)
-library(kernlab)
-opts_knit$set(upload.fun = socialR::notebook.url)
-opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE)
-theme_set(theme_bw())
-theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
-             plot.background = element_rect(fill = "transparent",colour = NA))
 ```
 
-```{r results="hide"}
-knit("~/Documents/code/nonparametric-bayes/inst/examples/beverton_holt_data.Rmd")
-true <- data.frame(x = X, y = sapply(X, f, 0, p))
+```{r BevHolt}
 ```
 
-Predict the function over the target grid
+We use the model of `r citet("10.1126/science.269.5227.1106")`.
 
-```{r}
-gp <- gp_fit(obs, X, c(sigma_n=2, tau=10, l=8), "conditional")
-plot.gpfit(gp, true)
+```{r sdp-pars-bh}
+sigma_g <- 0.02
+z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
+x_grid <- seq(0, 1.5 * K, length=101)
+h_grid <- x_grid
+profit = function(x,h) pmin(x, h)
+delta <- 0.01
+OptTime = 20
+xT = 0
+reward = profit(x_grid[length(x_grid)], x_grid[length(x_grid)]) + 1 / (1 - delta) ^ OptTime 
+## x_0_observed is starting condition for simulation of the observed data.  
+## It should be in preferred state for bistable model, 
+## above Allee threshold for Allee model, 
+## and near zero for BH or Ricker models
+x_0_observed <- x_grid[2] 
 ```
 
-## Stochastic Dynamic programming solution based on the posterior Gaussian process:
+With parameters `r p`.
 
-```{r}
-h_grid <- x_grid
+```{r sim-obs}
 ```
 
-GP Solution
-
-```{r}
-matrices_gp <- gp_transition_matrix(gp)
-opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
+```{r lag-data}
 ```
 
-Optimal solution
-
-```{r}
-matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
+```{r par-est}
 ```
 
-Optimal parametric solution in which a parameter is estimated with error
+```{r gp-fit, dependson="lag-data"}
+```
 
-```{r}
-k <- 15
-A <- 1.5
-B <- (A - 1)/k
-matrices_estimated <- f_transition_matrix(f, c(A, B), x_grid, h_grid, sigma_g)
-opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, 20, 0, profit, delta=.01)
+We fit a Gaussian process with
+
+```{r gp-data, dependson="gp-fit"}
+```
+
+```{r gp-plot, dependson="gp-data"}
+```
+
+## The transition matrix of the inferred process
+
+```{r persistence-test}
+```
+
+```{r gp-F-sim, dependson=c("persistence-test", "gp-data")}
+```
+
+```{r par-F-sim, dependson="persistence-test"}
+```
+
+```{r F-sim-plot, dependson=c("gp-F-sim", "par-F-sim")}
+```
+
+```{r est-F-sim, dependson="persistence-test"}
+```
+
+```{r gp-opt, dependson="gp-data"}
+```
+
+```{r true-opt}
+```
+
+```{r est-opt}
 ```
@@
 ```{r policy_plot}
-policies <- melt(data.frame(stock=x_grid, 
-                            GP = x_grid[opt_gp$D[,1]], 
-                            Exact = x_grid[opt_true$D[,1]],
-                            Approx = x_grid[opt_estimated$D[,1]]),
-                  id="stock")
-
-policy_plot <- ggplot(policies, aes(stock, stock - value, color=variable)) +
-  geom_point() + xlab("stock size") + ylab("escapement") 
-policy_plot
 ```
 
-We can see what happens when we attempt to manage a stock using this:
-
-``` {r othernoise}
-z_g <- function() rlnorm(1,0, sigma_g)
-````
-
-``` {r  simulate }
-set.seed(1)
-sim_gp <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_gp$D, z_g, profit=profit)
-set.seed(1)
-sim_true <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_true$D, z_g, profit=profit)
-set.seed(1)
-sim_est <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_estimated$D, z_g, profit=profit)
-````
-
-```{r}
-dat <- list(est = sim_est, gp = sim_gp, true = sim_true)
-dat <- melt(dat, id=names(dat[[1]]))
-dt <- data.table(dat)
-setnames(dt, "L1", "method") 
+```{r othernoise}
 ```
 
-``` {r  simplot}
-ggplot(dt) + geom_line(aes(time,fishstock, color=method))
-ggplot(dt) + geom_line(aes(time,harvest, color=method))
-````
+```{r simulate}
+```
 
-Total Profits
+```{r tidy}
+```
 
-```{r}
-c( gp = sum(sim_gp$profit), true = sum(sim_true$profit), est = sum(sim_est$profit))
+```{r sim-fish}
 ```
+
+```{r sim-harvest}
+```
+
+```{r costs}
+```
+
+```{r echo=FALSE, results="asis"}
+bibliography("html")
+```

### 2012-12-15 inst/examples/gaussian-process-control.R

@@ -18,6 +18,16 @@ theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
              plot.background = element_rect(fill = "transparent",colour = NA))
 
 
+
+
+
+## @knitr BevHolt
+f <- BevHolt
+p <- c(1.5,.05)
+K <- (p[1]-1)/p[2] 
+
+
+
 ## @knitr Myer
 f <- Myer_harvest
 pars <- c(1, 2, 6) 
@@ -47,7 +57,7 @@ reward = profit(x_grid[length(x_grid)], x_grid[length(x_grid)]) + 1 / (1 - delta
 ## It should be in preferred state for bistable model, 
 ## above Allee threshold for Allee model, 
 ## and near zero for BH or Ricker models
-x_0_observed <- allee + x_grid[1] 
+x_0_observed <- allee + x_grid[2] 
 
 
 
@@ -121,6 +131,16 @@ qplot(x_grid, yt10[1,]) + geom_point(aes(y=yt1[1,]), col="grey")
 ## @knitr F-sim-plot
 transition <- melt(data.frame(x = x_grid, gp = xt1[1,], parametric = yt1[1,]), id="x")
 ggplot(transition) + geom_point(aes(x,value, col=variable))
+
+
+
+## @knitr est-F-sim
+F_est <- par_F(h, f_alt, p_alt, x_grid, sigma_g)
+zt1 <- X %*% F_true
+zt10 <- zt1
+for(s in 1:OptTime)
+  zt10 <- zt10 %*% F_true
+qplot(x_grid, zt10[1,]) + geom_point(aes(y=zt1[1,]), col="grey")
 
 
 

### 2012-12-15 inst/examples/myer-example.Rmd
updated run shows differences in viability analysis, but GP still does optimally
Readme txt populated
Improved example of Allee model
* Compares to a Ricker estimated by MLE (#8)
* Considers fewer non-stationary observations, but include 0,0 as observation (#13)
* Includes scrap value (#10), though this does not guarentee no fishing under non-persistence estimation

@@
 ```{r set-options, echo = FALSE, cache = FALSE, external = TRUE}
-opts_chunk$set(external = TRUE, cache = TRUE, cache.path = "myers-cache/")
+opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "myers-cache/")
 read_chunk('gaussian-process-control.R')
@@
 ```
-
 ```{r graphing-options, include=FALSE}
@@
 
+We simulate data under this model, starting from a size of `r x_0_observed`.
+
 ```{r lag-data}
 ```
+
+We consider the observations as ordered pairs of observations of current stock size $x_t$ and observed stock in the following year, $x_{t+1}$.
+We add the pseudo-observation of $0,0$.
+Alternatively we could condition strictly on solutions passing through the origin, though in practice the weaker assumption is often sufficient.
 
@@
 
-```{r gp-F-sim, dependson="gp-data"}
+```{r persistence-test}
 ```
 
-```{r par-F-sim}
+```{r gp-F-sim, dependson=c("persistence-test", "gp-data")}
 ```
 
-```{r F-sim-plot}
+```{r par-F-sim, dependson="persistence-test"}
+```
+
+```{r F-sim-plot, dependson=c("gp-F-sim", "par-F-sim")}
 ```
@@
 
-```{r echo=FALSE, results="as.is"}
+```{r echo=FALSE, results="asis"}
 bibliography("html")

### 2012-12-15 inst/examples/gaussian-process-control.R

@@ -44,7 +44,7 @@ true <- data.frame(x = X, y = sapply(X, f, 0, p))
 
 
 ## @knitr sdp-pars
-sigma_g <- 0.02
+sigma_g <- 0.04
 z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
 x_grid <- seq(0, 1.5 * K, length=101)
 h_grid <- x_grid
@@ -57,7 +57,7 @@ reward = profit(x_grid[length(x_grid)], x_grid[length(x_grid)]) + 1 / (1 - delta
 ## It should be in preferred state for bistable model, 
 ## above Allee threshold for Allee model, 
 ## and near zero for BH or Ricker models
-x_0_observed <- allee + x_grid[2] 
+x_0_observed <- allee + x_grid[15] 
 
 
 
@@ -69,14 +69,17 @@ for(t in 1:(Tobs-1))
   x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
 plot(x)
 
-## @knitr lag-data
-obs <- data.frame(x=x[1:(Tobs-1)],y=x[2:Tobs])
+  ## @knitr lag-data
+obs <- data.frame(x=c(0,x[1:(Tobs-1)]),y=c(0,x[2:Tobs]))
 
 ## @knitr par-est
-K = mean(x)
-r = 0.5 * sigma_g^2 / var(x)
+estf <- function(p){
+  mu <- log(obs$x) + p["r"]*(1-obs$x/p["K"])
+  -sum(dlnorm(obs$y, mu, p["s"]), log=TRUE)
+}
+o <- optim(par = c(r=1,K=1,s=1), estf, method="L", lower=c(1e-3,1e-3,1e-3))
 f_alt <- Ricker
-p_alt <- p
+p_alt <- c(o$pars$r, o$pars$K)
 
 ## @knitr gp-fit
 gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,

### 2012-12-15 inst/examples/reed-example.Rmd
Underlying BH model compared to estimated GP and Ricker.

@@
 ```{r set-options, echo = FALSE, cache = FALSE, external = TRUE}
-opts_chunk$set(external = TRUE, cache = TRUE, cache.path = "bevholt-cache/")
+opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "bevholt-cache/")
 read_chunk('gaussian-process-control.R')
@@
 
-We use the model of `r citet("10.1126/science.269.5227.1106")`.
+We use the model of
 
@@
 ```
+
+Estimates a Ricker curve with parameters $r =$ `r p_alt[1]` and $K =$ `r p_alt[2]`
 

### 2012-12-15 inst/examples/may79-example.Rmd
May model in which GP inference fails to persist, kills stock.

--- 2012-12-11 inst/examples/may79-example.Rmd
@@
 
-Use my custom libraries and the `ggplot` plotting tools
-
-```{r libraries}
-library(pdgControl)
-library(nonparametricbayes)
-library(reshape2)
-library(ggplot2)
-library(data.table)
-library(kernlab)
-library(tgp)
+```{r set-options, echo = FALSE, cache = FALSE, external = TRUE}
+opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "may-cache/")
+read_chunk('gaussian-process-control.R')
+library(knitcitations)
 ```
 
+```{r libraries, include=FALSE}
+```
 ```{r graphing-options, include=FALSE}
-opts_knit$set(upload.fun = socialR::notebook.url)
-opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE)
-theme_set(theme_bw())
-theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
-             plot.background = element_rect(fill = "transparent",colour = NA))
 ```
 
-State-space model
-
-```{r}
-f <- May
-p <- c(r = .75, k = 10, a=1.3, H=1, Q = 3)
-K <- 8 # approx
+```{r May}
 ```
 
-Parameters
+We use the model of `r citet("10.1126/science.205.4403.267")`.
 
-```{r}
-x_grid = seq(0, 1.5 * K, length=101)
-T <- 40
-sigma_g <- 0.3
-x <- numeric(T)
-x[1] <- 1
+```{r sdp-pars}
 ```
 
-Model dynamics look like this:
+With parameters `r p`.
 
-```{r}
-birth <- function(x) p["r"] * (1-  x / p["k"])
-death <- function(x) p["a"] * x ^ (p["Q"] - 1) / (x ^ p["Q"] + p["H"])
-df <- data.frame(x = x_grid, b = sapply(x_grid, birth), d = sapply(x_grid, death))
-ggplot(df) + geom_line(aes(x, b), col = "blue") + geom_line(aes(x,d), col = "red")
-````
-
-Noise function, profit function
-```{r}
-z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
-profit <- profit_harvest(1,0,0)
+```{r x0-obs}
+xT <- x_grid[2]
+x_0_observed <- x_grid[60]
 ```
 
-Simulation
-
-```{r}
-x[1] = 2.5
-for(t in 1:(T-1))
-  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
-plot(x)
+```{r sim-obs}
 ```
 
-Predict the function over the target grid
+We simulate data under this model, starting from a size of `r x_0_observed`.
 
-```{r}
-obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
-X <- x_grid
+```{r lag-data}
 ```
 
-Gaussian Process inference from this model
+We consider the observations as ordered pairs of observations of current stock size $x_t$ and observed stock in the following year, $x_{t+1}$.
+We add the pseudo-observation of $0,0$.
+Alternatively we could condition strictly on solutions passing through the origin, though in practice the weaker assumption is often sufficient.
 
-```{r}
-gp <- bgp(X=obs$x, XX=X, Z=obs$y, verb=0,
-          meanfn="linear", bprior="b0", BTE=c(2000,6000,2), m0r1=FALSE, 
-          corr="exp", trace=TRUE, beta = c(0,0),
-          s2.p = c(50,50), d.p = c(10, 1/0.01, 10, 1/0.01), nug.p = c(10, 1/0.01, 10, 1/0.01),
-          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", 
-          tau2.lam = "fixed", tau2.p = c(50,1))
+```{r par-est}
 ```
 
-Extract the posterior Gaussian process mean and the $\pm 2$ standard deviations over the predicted grid from the fit:
-
-```{r}
-V <- gp$ZZ.ks2
-Ef = gp$ZZ.km
-tgp_dat <- data.frame(x   = gp$XX[[1]], 
-                  y   = gp$ZZ.km, 
-                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
-                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
+```{r gp-fit, dependson="lag-data"}
 ```
 
-```{r}
-true <- data.frame(x=X, y=sapply(X,f, 0, p))
-ggplot(tgp_dat)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
-  geom_line(data=true, aes(x,y), col='red', lty=2)
+We fit a Gaussian process with
+
+```{r gp-data, dependson="gp-fit"}
 ```
 
-## Stochastic Dynamic programming solution based on the posterior Gaussian process:
-
-```{r}
-h_grid <- x_grid
+```{r gp-plot, dependson="gp-data"}
 ```
 
-GP Solution
+## The transition matrix of the inferred process
 
-```{r}
-matrices_gp <- gp_transition_matrix(Ef, sqrt(V), x_grid, h_grid)
-opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
+```{r persistence-test}
 ```
 
-Optimal solution
-
-```{r}
-matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
+```{r gp-F-sim, dependson=c("persistence-test", "gp-data")}
 ```
 
-Optimal parametric solution in which a parameter is estimated with error
+```{r par-F-sim, dependson="persistence-test"}
+```
 
-```{r}
-matrices_estimated <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g=0.01)
-opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, 20, 0, profit, delta=.01)
+```{r F-sim-plot, dependson=c("gp-F-sim", "par-F-sim")}
+```
+
+```{r gp-opt, dependson="gp-data"}
+```
+
+```{r true-opt}
+```
+
+```{r est-opt}
 ```

### 2012-12-16 inst/examples/reed-example.Rmd
Updated example Ricker approximation vs GP approx to BH

@@
 
-We use the model of
+We use a Beverton-Holt model to drive the underlying dynamics, with parameters $A =$ `r p[1]` and $B =$ `r p[2]`.
 
@@
 
-With parameters `r p`.
+We consider stochastic growth driven by a lognormal noise process, $X_{t+1} = z_g f(X_t)$, where $f$ is the stock recruitment curve and $z_g$ a lognormal shock with $\sigma_g$ = `r sigma_g`.
 

### 2012-12-19 inst/examples/reed-example.Rmd
Parameteric Ricker model underestimates growth and thus underperforms relative to BH (true) or GP

@@
-```{r set-options, echo = FALSE, cache = FALSE, external = TRUE}
+```{r set-options, include=FALSE, echo = FALSE, cache = FALSE, external = TRUE}
 opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "bevholt-cache/")
@@
 ```
+
+Simulate data
 

### 2012-12-19 inst/examples/ricker-oscillate-example.Rmd
oscillatory example

--- 2012-12-15 inst/examples/myer-example.Rmd
@@
-```{r set-options, echo = FALSE, cache = FALSE, external = TRUE}
+```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
 opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "myers-cache/")
@@
 
-```{r Myer}
+```{r Ricker}
 ```
 
-We use the model of `r citet("10.1126/science.269.5227.1106")`.
+We use the Ricker model
 
@@
 
-With parameters `r p`.
+with parameters `r p`.
+
+```{r x0-obs}
+x_0_observed <- K
+xT <- 0
+set.seed(1111)
+```
 
@@
 
+```{r stationary_policy_only}
+```
+
 ```{r simulate}

### 2012-12-19 inst/examples/ricker-oscillate-example.Rmd
no so very good on the oscillation data

@@
 
-```{r sdp-pars}
+```{r sdp-pars-ricker}
+sigma_g <- 0.05
+z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
+x_grid <- seq(0, 2.5 * K, length=101)
+h_grid <- x_grid
+profit = function(x,h) pmin(x, h)
+delta <- 0.01
+OptTime = 20
+reward = profit(x_grid[length(x_grid)], x_grid[length(x_grid)]) + 1 / (1 - delta) ^ OptTime 
+xT <- 0
 ```

### 2012-12-19 inst/examples/myer-example.Rmd
Myers model as estimated from stationary data
base simulations off of stationary policy only

@@
 With parameters `r p`.
+
+```{r x0-obs}
+x_0_observed <- allee + x_grid[15]
+xT <- 0
+set.seed(1111)
+```
 

### 2012-12-19 inst/examples/gaussian-process-control.R

@@ -42,6 +42,10 @@ allee <- p[1] * p[3] / 2 - sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 # allee thre
 e_star <- (p[1] * sqrt(p[3]) - 2) / 2 ## Bifurcation point 
 
 
+## @knitr Ricker
+f <- Ricker
+p <- c(3.5, 10) 
+K <- 10
 
 
 ## @knitr May
@@ -83,7 +87,6 @@ x_0_observed <- K
 Tobs <- 30
 x <- numeric(Tobs)
 x[1] <- x_0_observed
-set.seed(123)
 for(t in 1:(Tobs-1))
   x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
 plot(x)
@@ -208,7 +211,7 @@ policy_plot
 z_g = function() rlnorm(1, 0, sigma_g)
 z_m = function() 1+(2*runif(1, 0,  1)-1) * 0.1
 
-### @knitr stationary_policy_only
+## @knitr stationary_policy_only
 m <- sapply(1:OptTime, function(i) opt_gp$D[,1])
 opt_gp$D <- m
 mm <- sapply(1:OptTime, function(i) opt_true$D[,1])

### 2012-12-19 inst/examples/gaussian-process-control.R
estimate BH as parametric model, poor example in Ricker

@@ -44,8 +44,8 @@ e_star <- (p[1] * sqrt(p[3]) - 2) / 2 ## Bifurcation point
 
 ## @knitr Ricker
 f <- Ricker
-p <- c(3.5, 10) 
-K <- 10
+p <- c(2.5, 10) 
+K <- 12.5
 
 
 ## @knitr May
@@ -103,6 +103,19 @@ o <- optim(par = c(r=1,K=mean(x),s=1), estf, method="L", lower=c(1e-3,1e-3,1e-3)
 f_alt <- Ricker
 p_alt <- c(o$par['r'], o$par['K'])
 sigma_g_alt <- o$par['s']
+
+
+
+## @knitr par-est-BH
+estf <- function(p){
+  mu <- obs$x * p["A"] / (1+ obs$x * p["B"])
+  -sum(dlnorm(obs$y, log(mu), p["s"]), log=TRUE)
+}
+o <- optim(par = c(A=2, B=1/mean(x),s=1), estf, method="L", lower=c(1e-3,1e-3,1e-3))
+f_alt <- BevHolt
+p_alt <- c(A=as.numeric(o$par['A']), B=as.numeric(o$par['B']))
+sigma_g_alt <- o$par['s']
+
 
 ## @knitr gp-fit
 gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,

### 2012-12-19 inst/examples/gaussian-process-control.R
lets try greater measurement error

@@ -63,6 +63,7 @@ ggplot(df) + geom_line(aes(x, b), col = "blue") + geom_line(aes(x,d), col = "red
 
 ## @knitr sdp-pars
 sigma_g <- 0.05
+sigma_m <- 0.2
 z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
 x_grid <- seq(0, 1.5 * K, length=101)
 h_grid <- x_grid
@@ -222,7 +223,7 @@ policy_plot
 
 ## @knitr othernoise
 z_g = function() rlnorm(1, 0, sigma_g)
-z_m = function() 1+(2*runif(1, 0,  1)-1) * 0.1
+z_m = function() 1+(2*runif(1, 0,  1)-1) * sigma_m
 
 ## @knitr stationary_policy_only
 m <- sapply(1:OptTime, function(i) opt_gp$D[,1])

### 2012-12-19 inst/examples/gaussian-process-control.R
Ah, good. With a bit more data, a very nice example of how GP can avoid problems of a structurally inaccurate parametric approach (Myers vs Ricker).

@@ -85,7 +85,7 @@ x_0_observed <- K
 
 
 ## @knitr sim-obs
-Tobs <- 30
+Tobs <- 50
 x <- numeric(Tobs)
 x[1] <- x_0_observed
 for(t in 1:(Tobs-1))

### 2012-12-19 inst/doc/nonparametric-bayes.Rmd
time to start writing

Structural Uncertainty 
========================================================

Problem of structural uncertainty in managing ecological systems.  Motivate via the widepread concern of tipping points, more general concerns of complex/nonlinear dynamics.

Typical approaches using parametric models.  motivation/strengths (mechanstic underpinnings, computationally simple/potentially analytic solutions.  Parametric structure makes it difficult to represent uncertainty accurately in areas where data is not available.

### 2012-12-20 inst/examples/myers-exploration.Rmd
trying methods for looping over replicates efficiently

```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "myers-cache/", warning=FALSE)
read_chunk('gaussian-process-control.R')
library(knitcitations)
```

```{r libraries, include=FALSE}
```
```{r graphing-options, include=FALSE}
```

Fixed priors on hyperparameters, fixed model type.

```{r gp-priors}
```

```{r sdp-pars-fixed}
profit = function(x,h) pmin(x, h)
delta <- 0.01
OptTime = 20
reward = 0
xT <- 0
z_g = function() rlnorm(1, 0, sigma_g)
z_m = function() 1+(2*runif(1, 0,  1)-1) * sigma_m
```

```{r Myer-explore}
f <- Myer_harvest
pars <- c(1, 2, 4.5) 
p <- pars # shorthand 
K <- p[1] * p[3] / 2 + sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2
allee <- p[1] * p[3] / 2 - sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 # allee threshold
e_star <- (p[1] * sqrt(p[3]) - 2) / 2 ## Bifurcation point 
```

```{r sdp-pars-explore}
sigma_g <- 0.05
sigma_m <- 0.2
x_grid <- seq(0, 1.5 * K, length=101)
h_grid <- x_grid
```

With parameters `r p`.

```{r }
yields <- 
  lapply(1:3, function(j){
  source("single-gp-fit.R")
  yield
})
```

```{r}
yields <- melt(yields, id=c("method", "V1", "sd"))
yields         
````

```{r echo=FALSE, results="asis"}
bibliography("html")
```

### 2012-12-20 inst/examples/single-gp-fit.R

@@ -1,0 +1,143 @@
+x_0_observed <- allee + x_grid[5]
+xT <- 0
+seed <- round(runif(1) * 1e6)
+seed
+set.seed(seed)
+
+## @knitr sim-obs
+Tobs <- 50
+x <- numeric(Tobs)
+x[1] <- x_0_observed
+for(t in 1:(Tobs-1))
+  x[t+1] = z_g() * f(x[t], h=0, p=p)
+plot(x)
+
+## @knitr lag-data
+obs <- data.frame(x=c(0,x[1:(Tobs-1)]),y=c(0,x[2:Tobs]))
+
+## @knitr par-est
+estf <- function(p){
+  mu <- log(obs$x) + p["r"]*(1-obs$x/p["K"])
+  -sum(dlnorm(obs$y, mu, p["s"]), log=TRUE)
+}
+o <- optim(par = c(r=1,K=mean(x),s=1), estf, method="L", lower=c(1e-3,1e-3,1e-3))
+f_alt <- Ricker
+p_alt <- c(o$par['r'], o$par['K'])
+sigma_g_alt <- o$par['s']
+
+
+## @knitr gp-fit
+gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
+          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
+          m0r1=FALSE, corr="exp", trace=TRUE, 
+          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
+          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")
+
+
+## @knitr gp-data
+V <- gp$ZZ.ks2
+Ef = gp$ZZ.km
+tgp_dat <- data.frame(x   = gp$XX[[1]], 
+                      y   = gp$ZZ.km, 
+                      ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
+                      ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
+
+
+## @knitr gp-plot
+true <- sapply(x_grid, f, 0, p)
+est <- sapply(x_grid, f_alt, 0, p_alt)
+models <- data.frame(x=x_grid, GP=tgp_dat$y, Parametric=est, True=true)
+models <- melt(models, id="x")
+names(models) <- c("x", "method", "value")
[…]

### 2012-12-20 inst/examples/myers-exploration.Rmd
avoid external source call, print plots

@@
   lapply(1:3, function(j){
-  source("single-gp-fit.R")
+
+x_0_observed <- allee + x_grid[5]
+seed <- round(runif(1) * 1e6)
+set.seed(seed)
+print(seed)
+
+## @knitr sim-obs
+Tobs <- 50
+x <- numeric(Tobs)
+x[1] <- x_0_observed
+for(t in 1:(Tobs-1))
+  x[t+1] = z_g() * f(x[t], h=0, p=p)
+qplot(1:Tobs, x)
+
+## @knitr lag-data
+obs <- data.frame(x=c(0,x[1:(Tobs-1)]),y=c(0,x[2:Tobs]))
+
+## @knitr par-est
+estf <- function(p){
+  mu <- log(obs$x) + p["r"]*(1-obs$x/p["K"])
+  -sum(dlnorm(obs$y, mu, p["s"]), log=TRUE)
+}
+o <- optim(par = c(r=1,K=mean(x),s=1), estf, method="L", lower=c(1e-3,1e-3,1e-3))
+f_alt <- Ricker
+p_alt <- c(o$par['r'], o$par['K'])
+sigma_g_alt <- o$par['s']
+
+## @knitr gp-fit
+gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
+          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
+          m0r1=FALSE, corr="exp", trace=TRUE, 
+          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
+          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")
+
+## @knitr gp-data
+V <- gp$ZZ.ks2
+Ef = gp$ZZ.km
+tgp_dat <- data.frame(x   = gp$XX[[1]], 
+                      y   = gp$ZZ.km, 
+                      ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
+                      ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
+
+## @knitr gp-plot
+true <- sapply(x_grid, f, 0, p)
+est <- sapply(x_grid, f_alt, 0, p_alt)
+models <- data.frame(x=x_grid, GP=tgp_dat$y, Parametric=est, True=true)
+models <- melt(models, id="x")
+names(models) <- c("x", "method", "value")
+gp_plot <- ggplot(tgp_dat)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
+  geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
+  geom_point(data=obs, aes(x,y), alpha=0.8) + 
+  xlab(expression(X[t])) + ylab(expression(X[t+1])) +
+  scale_colour_manual(values=cbPalette)
+print(gp_plot)
+
+## @knitr gp-posteriors
+hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
+posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
+priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
+prior_curves <- ddply(posteriors, "variable", function(dd){
+  grid <- seq(min(dd$value), max(dd$value), length = 100)
+  data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
+})
+posteriors_plot <- ggplot(posteriors) + 
+  geom_histogram(aes(x=value, y=..density..), lwd=2) +
+  geom_line(data=prior_curves, aes(x=value, y=density), col="red", lwd=2) +
+  facet_wrap(~ variable, scale="free")
+#print(posteriors_plot)
+
+## @knitr gp-opt
+matrices_gp <- gp_transition_matrix(Ef, V, x_grid, h_grid)
+opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=reward)
+
+## @knitr true-opt
+matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
+opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
+
+## @knitr est-opt
+matrices_estimated <- f_transition_matrix(f_alt, p_alt, x_grid, h_grid, sigma_g_alt)
+opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
+
+## @knitr policy_plot
+policies <- melt(data.frame(stock=x_grid, 
+                            GP = x_grid[opt_gp$D[,1]], 
+                            Parametric = x_grid[opt_estimated$D[,1]],
+                            True = x_grid[opt_true$D[,1]]),
+                 id="stock")
+names(policies) <- c("stock", "method", "value")
+policy_plot <- ggplot(policies, aes(stock, stock - value, color=method)) +
+  geom_line(lwd=2, alpha=0.8) + 
+  xlab("stock size") + ylab("escapement")  +
+  scale_colour_manual(values=cbPalette)
+print(policy_plot)
+
+## @knitr stationary_policy_only
+m <- sapply(1:OptTime, function(i) opt_gp$D[,1])
+opt_gp$D <- m
+mm <- sapply(1:OptTime, function(i) opt_true$D[,1])
+opt_true$D <- mm
+mmm <- sapply(1:OptTime, function(i) opt_estimated$D[,1])
+opt_estimated$D <- mmm
+
+## @knitr simulate
+set.seed(1)
+sim_gp <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_gp$D, z_g, profit=profit))
+set.seed(1)
+sim_true <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_true$D, z_g, profit=profit))
+set.seed(1)
+sim_est <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_estimated$D, z_g, profit=profit))
+
+## @knitr tidy
+dat <- list(GP = sim_gp, Parametric = sim_est, True = sim_true)
+dat <- melt(dat, id=names(dat[[1]][[1]]))
+dt <- data.table(dat)
+setnames(dt, c("L1", "L2"), c("method", "reps"))
+
+## @knitr sim-fish
+fish_plot <- ggplot(dt) + 
+  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
+  scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))
+print(fish_plot)
+
+## @knitr sim-harvest
+harvest_plot <- ggplot(dt) +
+  geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
+  scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))
+print(harvest_plot)
+
+## @knitr costs
+profits <- dt[, sum(profit), by = c("reps", "method")]
+means <- profits[, mean(V1), by = method]
+sds <- profits[, sd(V1), by = method]
+yield <- cbind(means, sd = sds$V1)
+yield
+
   yield

### 2012-12-20 inst/examples/myers-exploration.Rmd
30 replicates with full graphs

@@
 
+```{r parallel}
+require(snowfall)
+sfInit(cpu=2)
+sfExportAll()
+sfLibrary(pdgControl)
+sfLibrary(nonparametricbayes)
+sfLibrary(reshape2)
+sfLibrary(ggplot2)
+sfLibrary(data.table)
+sfLibrary(tgp)
+sfLibrary(kernlab)
+sfLibrary(MCMCpack)
+sfLibrary(plyr)
+```
+
 ```{r }
 yields <- 
-  lapply(1:3, function(j){
+  sfLapply(1:20, function(j){
 
@@
   geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
-  scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))
+  scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))+
+  ggtitle(paste("rep", j, "seed", seed))
 print(fish_plot)
@@
   geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
-  scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))
+  scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1))) +
+  ggtitle(paste("rep", j, "seed", seed))
 print(harvest_plot)

### 2012-12-20 inst/examples/myers-exploration.Rmd
unique seeds for parallel computing

@@
 ```{r }
+seed <- round(runif(20) * 1e6)
+seed
 yields <- 
-  sfLapply(1:20, function(j){
+  sfLapply(1:length(seed), 
+           function(j){
 
 x_0_observed <- allee + x_grid[5]
-seed <- round(runif(1) * 1e6)
-set.seed(seed)
-print(seed)
+set.seed(seed[j])
 
@@
   x[t+1] = z_g() * f(x[t], h=0, p=p)
-qplot(1:Tobs, x)
+data_plot <- qplot(1:Tobs, x) + ggtitle(paste("rep", j, "seed", seed))
+print(data_plot)
 
@@
   xlab(expression(X[t])) + ylab(expression(X[t+1])) +
-  scale_colour_manual(values=cbPalette)
+  scale_colour_manual(values=cbPalette) +
+  ggtitle(paste("rep", j, "seed", seed))
 print(gp_plot)
@@
   xlab("stock size") + ylab("escapement")  +
-  scale_colour_manual(values=cbPalette)
+  scale_colour_manual(values=cbPalette) +
+  ggtitle(paste("rep", j, "seed", seed))
 print(policy_plot)

### 2012-12-20 inst/examples/single-gp-fit.R

@@ -1,143 +1,0 @@
-x_0_observed <- allee + x_grid[5]
-xT <- 0
-seed <- round(runif(1) * 1e6)
-seed
-set.seed(seed)
-
-## @knitr sim-obs
-Tobs <- 50
-x <- numeric(Tobs)
-x[1] <- x_0_observed
-for(t in 1:(Tobs-1))
-  x[t+1] = z_g() * f(x[t], h=0, p=p)
-plot(x)
-
-## @knitr lag-data
-obs <- data.frame(x=c(0,x[1:(Tobs-1)]),y=c(0,x[2:Tobs]))
-
-## @knitr par-est
-estf <- function(p){
-  mu <- log(obs$x) + p["r"]*(1-obs$x/p["K"])
-  -sum(dlnorm(obs$y, mu, p["s"]), log=TRUE)
-}
-o <- optim(par = c(r=1,K=mean(x),s=1), estf, method="L", lower=c(1e-3,1e-3,1e-3))
-f_alt <- Ricker
-p_alt <- c(o$par['r'], o$par['K'])
-sigma_g_alt <- o$par['s']
-
-
-## @knitr gp-fit
-gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
-          m0r1=FALSE, corr="exp", trace=TRUE, 
-          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
-          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")
-
-
-## @knitr gp-data
-V <- gp$ZZ.ks2
-Ef = gp$ZZ.km
-tgp_dat <- data.frame(x   = gp$XX[[1]], 
-                      y   = gp$ZZ.km, 
-                      ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
-                      ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
-
-
-## @knitr gp-plot
-true <- sapply(x_grid, f, 0, p)
-est <- sapply(x_grid, f_alt, 0, p_alt)
-models <- data.frame(x=x_grid, GP=tgp_dat$y, Parametric=est, True=true)
-models <- melt(models, id="x")
-names(models) <- c("x", "method", "value")
[…]

### 2012-12-20 inst/examples/myers-exploration.Rmd
cpu=8 for running on one.ucdavis

@@
 require(snowfall)
-sfInit(cpu=2)
+sfInit(cpu=8)
 sfExportAll()

### 2012-12-21 inst/examples/myers-exploration.Rmd
corrected iterations

--- 2012-12-20 inst/examples/myers-exploration.Rmd
@@
 
-```{r parallel}
+```{r}
+
+sim_obs <- function(Xo, z_g, f, p, Tobs = 40, seed = 1){
+  x <- numeric(Tobs)
+  x[1] <- Xo
+  set.seed(seed)
+  for(t in 1:(Tobs-1))
+    x[t+1] = z_g() * f(x[t], h=0, p=p)
+  data_plot <- qplot(1:Tobs, x) + ggtitle(paste("seed", seed))
+  print(data_plot)
+  obs <- data.frame(x=c(0,x[1:(Tobs-1)]),y=c(0,x[2:Tobs]))
+}
+
+par_est <- function(obs){
+  estf <- function(p){
+    mu <- log(obs$x) + p["r"]*(1-obs$x/p["K"])
+    llik <- -sum(dlnorm(obs$y, mu, p["s"]), log=TRUE)
+    if(!is.numeric(llik) | is.nan(llik) | !(llik < Inf)){
+      warning("possible error in llik")
+      llik <- 1e30
+    }
+    llik
+  }
+  o <- optim(par = c(r=1,K=mean(obs$x),s=1), estf, method="L", lower=c(1e-3,1e-3,1e-3))
+  f_alt <- Ricker
+  p_alt <- c(o$par['r'], o$par['K'])
+  sigma_g_alt <- o$par['s']
+  list(f_alt = f_alt, p_alt = p_alt, sigma_g_alt = sigma_g_alt)
+}
+optimal_policy <- function(gp, f, f_alt, p, p_alt, x_grid, h_grid, sigma_g, sigma_g_alt, delta, xT, profit, reward, OptTime){
+  matrices_gp <- gp_transition_matrix(gp$ZZ.ks2, gp$ZZ.km, x_grid, h_grid)
+  opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=reward)
+  matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
+  opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
+  matrices_estimated <- f_transition_matrix(f_alt, p_alt, x_grid, h_grid, sigma_g_alt)
+  opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
+  list(gp_D = opt_gp$D[,1], true_D = opt_true$D[,1], est_D = opt_estimated$D[,1])
+}
+
+simulate_opt <- function(OPT, f, p, x_grid, h_grid, K, z_g, profit){
+  gp_D <- sapply(1:OptTime, function(i) OPT$gp_D)
+  true_D <- sapply(1:OptTime, function(i) OPT$true_D)
+  est_D <- sapply(1:OptTime, function(i) OPT$est_D)
+  set.seed(1)
+  sim_gp <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, gp_D, z_g, profit=profit))
+  set.seed(1)
+  sim_true <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, true_D, z_g, profit=profit))
+  set.seed(1)
+  sim_est <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, est_D, z_g, profit=profit))
+  dat <- list(GP = sim_gp, Parametric = sim_est, True = sim_true)
+  dat <- melt(dat, id=names(dat[[1]][[1]]))
+  dt <- data.table(dat)
+  setnames(dt, c("L1", "L2"), c("method", "reps")) 
+  dt
+}
+
+sim_plots <- function(dt, seed=1){
+  fish_plot <- ggplot(dt) + 
+    geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
+    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))+
+    ggtitle(paste("seed", seed))
+  print(fish_plot)
+  harvest_plot <- ggplot(dt) +
+    geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
+    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1))) +
+    ggtitle(paste("seed", seed))
+  print(harvest_plot)
+}
+
+profits_stats <- function(dt){
+  profits <- dt[, sum(profit), by = c("reps", "method")]
+  means <- profits[, mean(V1), by = method]
+  sds <- profits[, sd(V1), by = method]
+  yield <- cbind(means, sd = sds$V1)
+  yield
+}
+```
+
+```{r parallel, message=FALSE, warning=FALSE, results='hide', eval=FALSE, include=FALSE, echo=FALSE}
 require(snowfall)
-sfInit(cpu=2)
+sfInit(par=TRUE, cpu=8)
 sfExportAll()
@@
 ```{r }
-yields <- 
-  sfLapply(1:20, function(j){
-
-x_0_observed <- allee + x_grid[5]
-seed <- round(runif(1) * 1e6)
-set.seed(seed)
-print(seed)
-
-## @knitr sim-obs
-Tobs <- 50
-x <- numeric(Tobs)
-x[1] <- x_0_observed
-for(t in 1:(Tobs-1))
-  x[t+1] = z_g() * f(x[t], h=0, p=p)
-qplot(1:Tobs, x)
-
-## @knitr lag-data
-obs <- data.frame(x=c(0,x[1:(Tobs-1)]),y=c(0,x[2:Tobs]))
-
-## @knitr par-est
-estf <- function(p){
-  mu <- log(obs$x) + p["r"]*(1-obs$x/p["K"])
-  -sum(dlnorm(obs$y, mu, p["s"]), log=TRUE)
-}
-o <- optim(par = c(r=1,K=mean(x),s=1), estf, method="L", lower=c(1e-3,1e-3,1e-3))
-f_alt <- Ricker
-p_alt <- c(o$par['r'], o$par['K'])
-sigma_g_alt <- o$par['s']
-
-## @knitr gp-fit
+seed <- round(runif(2) * 1e6)
+seed
+yields <- lapply(seed, 
+function(seed_i){
+  Xo <- allee + x_grid[5]
+  obs <- sim_obs(Xo, z_g, f, p, Tobs=40, seed = seed_i)
+  est <- par_est(obs)
   gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
-          m0r1=FALSE, corr="exp", trace=TRUE, 
+          meanfn="constant", bprior="b0", BTE=c(2000,6000,2),
+          m0r1=FALSE, corr="exp", trace=FALSE, 
           beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
           s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
-
-## @knitr gp-data
-V <- gp$ZZ.ks2
-Ef = gp$ZZ.km
-tgp_dat <- data.frame(x   = gp$XX[[1]], 
-                      y   = gp$ZZ.km, 
-                      ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
-                      ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
-
-## @knitr gp-plot
-true <- sapply(x_grid, f, 0, p)
-est <- sapply(x_grid, f_alt, 0, p_alt)
-models <- data.frame(x=x_grid, GP=tgp_dat$y, Parametric=est, True=true)
-models <- melt(models, id="x")
-names(models) <- c("x", "method", "value")
-gp_plot <- ggplot(tgp_dat)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-  geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
-  geom_point(data=obs, aes(x,y), alpha=0.8) + 
-  xlab(expression(X[t])) + ylab(expression(X[t+1])) +
-  scale_colour_manual(values=cbPalette)
-print(gp_plot)
-
-## @knitr gp-posteriors
-hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
-posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
-priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
-prior_curves <- ddply(posteriors, "variable", function(dd){
-  grid <- seq(min(dd$value), max(dd$value), length = 100)
-  data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
-})
-posteriors_plot <- ggplot(posteriors) + 
-  geom_histogram(aes(x=value, y=..density..), lwd=2) +
-  geom_line(data=prior_curves, aes(x=value, y=density), col="red", lwd=2) +
-  facet_wrap(~ variable, scale="free")
-#print(posteriors_plot)
-
-## @knitr gp-opt
-matrices_gp <- gp_transition_matrix(Ef, V, x_grid, h_grid)
-opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=reward)
-
-## @knitr true-opt
-matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
-
-## @knitr est-opt
-matrices_estimated <- f_transition_matrix(f_alt, p_alt, x_grid, h_grid, sigma_g_alt)
-opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
-
-## @knitr policy_plot
-policies <- melt(data.frame(stock=x_grid, 
-                            GP = x_grid[opt_gp$D[,1]], 
-                            Parametric = x_grid[opt_estimated$D[,1]],
-                            True = x_grid[opt_true$D[,1]]),
-                 id="stock")
-names(policies) <- c("stock", "method", "value")
-policy_plot <- ggplot(policies, aes(stock, stock - value, color=method)) +
-  geom_line(lwd=2, alpha=0.8) + 
-  xlab("stock size") + ylab("escapement")  +
-  scale_colour_manual(values=cbPalette)
-print(policy_plot)
-
-## @knitr stationary_policy_only
-m <- sapply(1:OptTime, function(i) opt_gp$D[,1])
-opt_gp$D <- m
-mm <- sapply(1:OptTime, function(i) opt_true$D[,1])
-opt_true$D <- mm
-mmm <- sapply(1:OptTime, function(i) opt_estimated$D[,1])
-opt_estimated$D <- mmm
-
-## @knitr simulate
-set.seed(1)
-sim_gp <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_gp$D, z_g, profit=profit))
-set.seed(1)
-sim_true <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_true$D, z_g, profit=profit))
-set.seed(1)
-sim_est <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, opt_estimated$D, z_g, profit=profit))
-
-## @knitr tidy
-dat <- list(GP = sim_gp, Parametric = sim_est, True = sim_true)
-dat <- melt(dat, id=names(dat[[1]][[1]]))
-dt <- data.table(dat)
-setnames(dt, c("L1", "L2"), c("method", "reps"))
-
-## @knitr sim-fish
-fish_plot <- ggplot(dt) + 
-  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
-  scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))+
-  ggtitle(paste("rep", j, "seed", seed))
-print(fish_plot)
-
-## @knitr sim-harvest
-harvest_plot <- ggplot(dt) +
-  geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
-  scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1))) +
-  ggtitle(paste("rep", j, "seed", seed))
-print(harvest_plot)
-
-## @knitr costs
-profits <- dt[, sum(profit), by = c("reps", "method")]
-means <- profits[, mean(V1), by = method]
-sds <- profits[, sd(V1), by = method]
-yield <- cbind(means, sd = sds$V1)
-yield
-
-  yield
+  OPT <- optimal_policy(gp, f, est$f_alt, p, est$p_alt, x_grid, h_grid, sigma_g, est$sigma_g_alt, delta, xT, profit, reward, OptTime)
+  dt <- simulate_opt(OPT, f, p, x_grid, h_grid, K, z_g, profit)
+  sim_plots(dt, seed=seed_i)
+  profits_stats(dt)
 })

### 2012-12-21 inst/examples/myers-exploration.Rmd
more iterations

@@
 
-```{r parallel, message=FALSE, warning=FALSE, results='hide', eval=FALSE, include=FALSE, echo=FALSE}
+```{r parallel, message=FALSE, warning=FALSE, results='hide'}
 require(snowfall)
@@
 ```{r }
-seed <- round(runif(2) * 1e6)
+seed <- round(runif(16) * 1e6)
 seed
@@
   Xo <- allee + x_grid[5]
-  obs <- sim_obs(Xo, z_g, f, p, Tobs=40, seed = seed_i)
+  obs <- sim_obs(Xo, z_g, f, p, Tobs=50, seed = seed_i)
   est <- par_est(obs)

### 2012-12-21 inst/examples/myers-exploration.Rmd
fixed bug (had Ef and V in gp_transition_matrix)

@@
 f <- Myer_harvest
-pars <- c(1, 2, 4.5) 
-p <- pars # shorthand 
+p <- c(1, 2, 4.5) 
 K <- p[1] * p[3] / 2 + sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2
@@
   obs <- data.frame(x=c(0,x[1:(Tobs-1)]),y=c(0,x[2:Tobs]))
+  obs
 }
@@
 }
+
 optimal_policy <- function(gp, f, f_alt, p, p_alt, x_grid, h_grid, sigma_g, sigma_g_alt, delta, xT, profit, reward, OptTime){
-  matrices_gp <- gp_transition_matrix(gp$ZZ.ks2, gp$ZZ.km, x_grid, h_grid)
+  matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.ks2, x_grid, h_grid)
   opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=reward)
@@
 }
+
+gp_plot <- function(gp, f, p, f_alt, p_alt, x_grid, obs){
+  tgp_dat <- 
+  data.frame(  x = gp$XX[[1]], 
+               y = gp$ZZ.km, 
+             ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
+             ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
+  true <- sapply(x_grid, f, 0, p)
+  est <- sapply(x_grid, f_alt, 0, p_alt)
+  models <- data.frame(x=x_grid, GP=tgp_dat$y, Parametric=est, True=true)
+  models <- melt(models, id="x")
+  names(models) <- c("x", "method", "value")
+  plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
+    geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
+    geom_point(data=obs, aes(x,y), alpha=0.8) + 
+    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
+    scale_colour_manual(values=cbPalette)
+  print(plot_gp)
+}
+
 ```
@@
 seed <- round(runif(16) * 1e6)
+seed[1] <- 1
 seed
@@
   Xo <- allee + x_grid[5]
-  obs <- sim_obs(Xo, z_g, f, p, Tobs=50, seed = seed_i)
+  obs <- sim_obs(Xo, z_g, f, p, Tobs=40, seed = seed_i)
   est <- par_est(obs)
   gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(2000,6000,2),
+          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
           m0r1=FALSE, corr="exp", trace=FALSE, 
@@
           s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
+  gp_plot(gp, f, p, est$f_alt, est$p_alt, x_grid, obs)
   OPT <- optimal_policy(gp, f, est$f_alt, p, est$p_alt, x_grid, h_grid, sigma_g, est$sigma_g_alt, delta, xT, profit, reward, OptTime)

### 2012-12-21 inst/examples/myers-exploration.Rmd
shorter OptTime and lower initial state do not much change the fraction that fail to persist

@@
 delta <- 0.01
-OptTime = 20
+OptTime = 10  # stationarity with unstable models is tricky thing
 reward = 0
@@
 
-simulate_opt <- function(OPT, f, p, x_grid, h_grid, K, z_g, profit){
+simulate_opt <- function(OPT, f, p, x_grid, h_grid, x0, z_g, profit){
   gp_D <- sapply(1:OptTime, function(i) OPT$gp_D)
@@
   set.seed(1)
-  sim_gp <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, gp_D, z_g, profit=profit))
+  sim_gp <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, x0, gp_D, z_g, profit=profit))
   set.seed(1)
-  sim_true <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, true_D, z_g, profit=profit))
+  sim_true <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, x0, true_D, z_g, profit=profit))
   set.seed(1)
-  sim_est <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, est_D, z_g, profit=profit))
+  sim_est <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, x0, est_D, z_g, profit=profit))
   dat <- list(GP = sim_gp, Parametric = sim_est, True = sim_true)
@@
 
-gp_plot <- function(gp, f, p, f_alt, p_alt, x_grid, obs){
+gp_plot <- function(gp, f, p, f_alt, p_alt, x_grid, obs, seed){
   tgp_dat <- 
@@
     xlab(expression(X[t])) + ylab(expression(X[t+1])) +
-    scale_colour_manual(values=cbPalette)
+    scale_colour_manual(values=cbPalette) +
+    ggtitle(paste("seed", seed))
   print(plot_gp)
@@
 ```{r }
-seed <- round(runif(16) * 1e6)
+seed <- round(runif(32) * 1e6)
 seed[1] <- 1
@@
 function(seed_i){
-  Xo <- allee + x_grid[5]
+  Xo <- allee + x_grid[5] # observations start from
+  x0 <- 2 # simulation under policy starts from
   obs <- sim_obs(Xo, z_g, f, p, Tobs=40, seed = seed_i)
@@
           s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
-  gp_plot(gp, f, p, est$f_alt, est$p_alt, x_grid, obs)
+  gp_plot(gp, f, p, est$f_alt, est$p_alt, x_grid, obs, seed_i)
   OPT <- optimal_policy(gp, f, est$f_alt, p, est$p_alt, x_grid, h_grid, sigma_g, est$sigma_g_alt, delta, xT, profit, reward, OptTime)
-  dt <- simulate_opt(OPT, f, p, x_grid, h_grid, K, z_g, profit)
+  dt <- simulate_opt(OPT, f, p, x_grid, h_grid, x0, z_g, profit)
   sim_plots(dt, seed=seed_i)
@@
 yields <- melt(yields, id=c("method", "V1", "sd"))
+ggplot(yields) + geom_histogram(aes(V1)) + facet_wrap(~method, scale='free')
 yields

### 2012-12-21 inst/examples/myers-exploration.Rmd
harvest, and then escaped population recruits; comparable

@@
 delta <- 0.01
-OptTime = 10  # stationarity with unstable models is tricky thing
+OptTime = 20  # stationarity with unstable models is tricky thing
 reward = 0
@@
 ```{r Myer-explore}
-f <- Myer_harvest
+f <- Myers
 p <- c(1, 2, 4.5) 

### 2012-12-21 inst/examples/myers-exploration.Rmd
higher K value results in better persistence in general, not always outperforming parametric model though.

@@
 f <- Myers
-p <- c(1, 2, 4.5) 
+p <- c(1, 2, 6) 
 K <- p[1] * p[3] / 2 + sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2

### 2012-12-21 inst/examples/myers-exploration.Rmd
delta 1.5, r 1.3 K 4.5

@@
 f <- Myers
-p <- c(1, 2, 6) 
-K <- p[1] * p[3] / 2 + sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2
-allee <- p[1] * p[3] / 2 - sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 # allee threshold
-e_star <- (p[1] * sqrt(p[3]) - 2) / 2 ## Bifurcation point 
+p <- c(1.3, 1.5, 4.5) 
+K <- 3.5
+allee <- 1
+# valid only for p[2] = 2
+#K <- p[1] * p[3] / 2 + sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 #
+#allee <- p[1] * p[3] / 2 - sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 # allee threshold
+#e_star <- (p[1] * sqrt(p[3]) - 2) / 2 ## Bifurcation point 
 ```

### 2012-12-21 inst/examples/myers-exploration.Rmd
strange Ricker Allee example -- even true model doesn't harvest successfully

--- 2012-12-21 inst/examples/myers-exploration.Rmd
@@
 ```{r Myer-explore}
-f <- Myers
-p <- c(1, 2, 6) 
-K <- p[1] * p[3] / 2 + sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2
-allee <- p[1] * p[3] / 2 - sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 # allee threshold
-e_star <- (p[1] * sqrt(p[3]) - 2) / 2 ## Bifurcation point 
+f <- RickerAllee
+p <- c(1.5, 10, 5) 
+K <- 10
+allee <- 5
+# valid only for p[2] = 2
+#K <- p[1] * p[3] / 2 + sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 #
+#allee <- p[1] * p[3] / 2 - sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 # allee threshold
+#e_star <- (p[1] * sqrt(p[3]) - 2) / 2 ## Bifurcation point 
 ```

### 2012-12-21 inst/examples/myers-exploration.Rmd
Ah here we go! with a decent allee effect model, the results are much clearer!

@@
 function(seed_i){
-  Xo <- allee + x_grid[5] # observations start from
-  x0 <- 2 # simulation under policy starts from
+  Xo <- allee + x_grid[10] # observations start from
+  x0 <- Xo # simulation under policy starts from
   obs <- sim_obs(Xo, z_g, f, p, Tobs=40, seed = seed_i)

### 2012-12-21 inst/examples/myers-exploration.Rmd
less successful with higher noise level
GP performs poorly with weak noise prior (nugget) when not conditioned on 0,0

--- 2012-12-21 inst/examples/myers-exploration.Rmd
@@
 opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "myers-cache/", warning=FALSE)
-read_chunk('gaussian-process-control.R')
+#read_chunk('gaussian-process-control.R')
 library(knitcitations)
@@
 ```{r libraries, include=FALSE}
+require(pdgControl)
+require(nonparametricbayes)
+require(reshape2)
+require(ggplot2)
+require(data.table)
+require(tgp)
+library(kernlab)
+require(MCMCpack)
+require(plyr)
+
 ```
+
 ```{r graphing-options, include=FALSE}
+opts_knit$set(upload.fun = socialR::notebook.url)
+#opts_chunk$set(dev = 'Cairo_pdf', dev.args=list(""))
+opts_chunk$set(dev="png", dev.args=list(bg="transparent"))
+opts_chunk$set(comment=NA, tidy=FALSE)
+theme_set(theme_bw(base_size=16))
+theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
+             plot.background = element_rect(fill = "transparent",colour = NA))
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 ```
@@
 ```{r gp-priors}
+s2.p <- c(50,50)  # inverse gamma mean is 
+tau2.p <- c(20,1)
+d.p = c(10, 1/0.01, 10, 1/0.01)
+nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
+s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
+tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
+d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
+nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
+beta0_prior <- function(x, tau) dnorm(x, 0, tau)
+beta = c(0)
+priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
 ```
@@
 ```{r sdp-pars-explore}
-sigma_g <- 0.05
+sigma_g <- 0.1
 sigma_m <- 0.2
@@
 
+posteriors_plot <- function(gp, priors){
+  hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
+  posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
+  prior_curves <- ddply(posteriors, "variable", function(dd){
+    grid <- seq(min(dd$value), max(dd$value), length = 100)
+    data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
+  })
+  plot_posteriors <- ggplot(posteriors) + 
+    #geom_density(aes(value), lwd=2) +
+    geom_histogram(aes(x=value, y=..density..), alpha=0.7) +
+    geom_line(data=prior_curves, aes(x=value, y=density), col="red", lwd=2) +
+    facet_wrap(~ variable, scale="free")
+  print(plot_posteriors)
+}
+
 ```
@@
   gp_plot(gp, f, p, est$f_alt, est$p_alt, x_grid, obs, seed_i)
+#  posteriors_plot(gp, priors) # needs trace=TRUE!
   OPT <- optimal_policy(gp, f, est$f_alt, p, est$p_alt, x_grid, h_grid, sigma_g, est$sigma_g_alt, delta, xT, profit, reward, OptTime)

### 2012-12-21 inst/examples/myers-exploration.Rmd
small noise. works well.

@@
 d.p = c(10, 1/0.01, 10, 1/0.01)
-nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
+nug.p = c(10, 1/0.01, 10, 1/0.01) # gamma mean
 s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
@@
 ```{r sdp-pars-explore}
-sigma_g <- 0.1
-sigma_m <- 0.2
+sigma_g <- 0.01
+sigma_m <- 0.02
 x_grid <- seq(0, 1.5 * K, length=101)

### 2012-12-23 inst/examples/allee.Rmd
example allee model, against estimated with correct structure

--- 2012-12-15 inst/examples/myer-example.Rmd
@@
-```{r set-options, echo = FALSE, cache = FALSE, external = TRUE}
-opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "myers-cache/")
+```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
+opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "allee-cache/", warning=FALSE)
 read_chunk('gaussian-process-control.R')
@@
 
-```{r Myer}
+```{r RickerAllee}
 ```
 
-We use the model of `r citet("10.1126/science.269.5227.1106")`.
+We use the model of Allen
 
@@
 
-```{r sim-obs}
+```{r x0-obs}
+x_0_observed <- K
+xT <- 0
+set.seed(1)
+```
+
+```{r sim-with-harvest}
 ```
@@
 We simulate data under this model, starting from a size of `r x_0_observed`.
-
-```{r lag-data}
-```
 
@@
 
-```{r par-est}
+```{r MLE-RickerAllee}
 ```
 
-```{r gp-fit, dependson="lag-data"}
+```{r gp-priors}
+```
+
+```{r gp-fit}
 ```
@@
 ```{r gp-plot, dependson="gp-data"}
+```
+
+```{r gp-posteriors}
 ```

### 2012-12-24 inst/examples/gaussian-process-control.R
offline runs on allee

@@ -13,7 +13,7 @@ require(plyr)
 
 
 ## @knitr graphing-options
-opts_knit$set(upload.fun = socialR::notebook.url)
+opts_knit$set(upload.fun = function(file) socialR::notebook.url(file, cp=TRUE, sync=FALSE))
 #opts_chunk$set(dev = 'Cairo_pdf', dev.args=list(""))
 opts_chunk$set(dev="png", dev.args=list(bg="transparent"))
 opts_chunk$set(comment=NA, tidy=FALSE)
@@ -21,7 +21,10 @@ theme_set(theme_bw(base_size=16))
 theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
              plot.background = element_rect(fill = "transparent",colour = NA))
 cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-
+#sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
+#short_sha <- gsub("(^.{10}).*", "\\1", sha)
+#date <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S")
+#opts_chunk$set(fig.path = paste("figure/", date, "-", short_sha, "-", sep=""))
 
 
 
@@ -64,9 +67,11 @@ ggplot(df) + geom_line(aes(x, b), col = "blue") + geom_line(aes(x,d), col = "red
 
 ## @knitr RickerAllee
 f <- RickerAllee
-p <- c(r = 1.5, C=5, K = 10)
-K <- p[3]
-allee <- p[2]
+p <- c(r = 3, K = 10, C=4)
+K <- p[2]
+allee <- p[3]
+
+
 
 ## @knitr sdp-pars
 sigma_g <- 0.05
@@ -76,7 +81,7 @@ x_grid <- seq(0, 1.5 * K, length=101)
 h_grid <- x_grid
 profit = function(x,h) pmin(x, h)
 delta <- 0.01
-OptTime = 20
+OptTime = 50 # should be long enough that profit is greater sustainably harvesting
 reward = profit(x_grid[length(x_grid)], x_grid[length(x_grid)]) + 1 / (1 - delta) ^ OptTime 
[…]

### 2012-12-24 inst/examples/myers-exploration.Rmd

@@
 f <- RickerAllee
-p <- c(1.5, 10, 5) 
+p <- c(1.1, 10, 5) 
 K <- 10
@@
 ```{r sdp-pars-explore}
-sigma_g <- 0.01
+sigma_g <- 0.05
 sigma_m <- 0.02

### 2012-12-24 R/gp-replicate-fns.R
additional externalized functions

@@ -34,6 +34,28 @@ sim_obs <- function(Xo, z_g, f, p, Tobs = 40, seed = 1, nz = 10, harvest = sort(
 #' @param obs the observed data, two columns giving x_t, x_t+1 respectively
 #' @return a list with the f given, the MLE estimated parameters, and estimated noise level
 #' @export 
+par_est_allee <- function(obs, f, p){
+  estf <- function(p){ 
+    mu <- f(obs$x,0,p)
+    -sum(dlnorm(obs$y, log(mu), p["s"]), log=TRUE)
+  }
+  par = c(p[1] - 1, 
+          p[2] - 1, 
+          p[3] + 2, 
+          s = sigma_g + abs(rnorm(1,0,.1)))
+  o <- optim(par, estf, method="L", lower=c(1e-3,1e-3,1e-3, 1e-3))
+  f_alt <- f
+  p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
+  sigma_g_alt <- as.numeric(o$par[4])
+  list(f_alt = f_alt, p_alt = p_alt, sigma_g_alt = sigma_g_alt)
+}
+
+
+
+#' MLE estimate of parameters given by function f
+#' @param obs the observed data, two columns giving x_t, x_t+1 respectively
+#' @return a list with the f given, the MLE estimated parameters, and estimated noise level
+#' @export 
 par_est <- function(obs){
   estf <- function(p){
     mu <- log(obs$x) + p["r"]*(1-obs$x/p["K"])
@@ -55,14 +77,16 @@ par_est <- function(obs){
 #' helper function to determine the optimal policies of each model
 #' @import pdgControl
 #' @export
-optimal_policy <- function(gp, f, f_alt, p, p_alt, x_grid, h_grid, sigma_g, sigma_g_alt, delta, xT, profit, reward, OptTime){
+optimal_policy <- function(gp, f, f_est, f_alt, p, p_est, p_alt, x_grid, h_grid, sigma_g, sigma_g_est, sigma_g_alt, delta, xT, profit, reward, OptTime){
   matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.ks2, x_grid, h_grid)
   opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=reward)
   matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
[…]

### 2012-12-24 inst/examples/sensitivity.Rmd
updated for trial run on one.ucdavis.edu

--- 2012-12-24 inst/examples/myers-exploration.Rmd
@@
 
-```{r Myer-explore}
+```{r RickerAllee-exp}
 f <- RickerAllee
@@
 allee <- 5
-# valid only for p[2] = 2
-#K <- p[1] * p[3] / 2 + sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 #
-#allee <- p[1] * p[3] / 2 - sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 # allee threshold
-#e_star <- (p[1] * sqrt(p[3]) - 2) / 2 ## Bifurcation point 
 ```
@@
 ```{r}
-
-sim_obs <- function(Xo, z_g, f, p, Tobs = 40, seed = 1){
-  x <- numeric(Tobs)
-  x[1] <- Xo
-  set.seed(seed)
-  for(t in 1:(Tobs-1))
-    x[t+1] = z_g() * f(x[t], h=0, p=p)
-  data_plot <- qplot(1:Tobs, x) + ggtitle(paste("seed", seed))
-  print(data_plot)
-  obs <- data.frame(x=c(0,x[1:(Tobs-1)]),y=c(0,x[2:Tobs]))
-  obs
-}
-
-par_est <- function(obs){
-  estf <- function(p){
-    mu <- log(obs$x) + p["r"]*(1-obs$x/p["K"])
-    llik <- -sum(dlnorm(obs$y, mu, p["s"]), log=TRUE)
-    if(!is.numeric(llik) | is.nan(llik) | !(llik < Inf)){
-      warning("possible error in llik")
-      llik <- 1e30
-    }
-    llik
-  }
-  o <- optim(par = c(r=1,K=mean(obs$x),s=1), estf, method="L", lower=c(1e-3,1e-3,1e-3))
-  f_alt <- Ricker
-  p_alt <- c(o$par['r'], o$par['K'])
-  sigma_g_alt <- o$par['s']
-  list(f_alt = f_alt, p_alt = p_alt, sigma_g_alt = sigma_g_alt)
-}
-
-optimal_policy <- function(gp, f, f_alt, p, p_alt, x_grid, h_grid, sigma_g, sigma_g_alt, delta, xT, profit, reward, OptTime){
-  matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.ks2, x_grid, h_grid)
-  opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=reward)
-  matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-  opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
-  matrices_estimated <- f_transition_matrix(f_alt, p_alt, x_grid, h_grid, sigma_g_alt)
-  opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
-  list(gp_D = opt_gp$D[,1], true_D = opt_true$D[,1], est_D = opt_estimated$D[,1])
-}
-
-simulate_opt <- function(OPT, f, p, x_grid, h_grid, x0, z_g, profit){
-  gp_D <- sapply(1:OptTime, function(i) OPT$gp_D)
-  true_D <- sapply(1:OptTime, function(i) OPT$true_D)
-  est_D <- sapply(1:OptTime, function(i) OPT$est_D)
-  set.seed(1)
-  sim_gp <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, x0, gp_D, z_g, profit=profit))
-  set.seed(1)
-  sim_true <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, x0, true_D, z_g, profit=profit))
-  set.seed(1)
-  sim_est <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, x0, est_D, z_g, profit=profit))
-  dat <- list(GP = sim_gp, Parametric = sim_est, True = sim_true)
-  dat <- melt(dat, id=names(dat[[1]][[1]]))
-  dt <- data.table(dat)
-  setnames(dt, c("L1", "L2"), c("method", "reps")) 
-  dt
-}
-
-sim_plots <- function(dt, seed=1){
-  fish_plot <- ggplot(dt) + 
-    geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
-    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))+
-    ggtitle(paste("seed", seed))
-  print(fish_plot)
-  harvest_plot <- ggplot(dt) +
-    geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
-    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1))) +
-    ggtitle(paste("seed", seed))
-  print(harvest_plot)
-}
-
-profits_stats <- function(dt){
-  profits <- dt[, sum(profit), by = c("reps", "method")]
-  means <- profits[, mean(V1), by = method]
-  sds <- profits[, sd(V1), by = method]
-  yield <- cbind(means, sd = sds$V1)
-  yield
-}
-
-gp_plot <- function(gp, f, p, f_alt, p_alt, x_grid, obs, seed){
-  tgp_dat <- 
-  data.frame(  x = gp$XX[[1]], 
-               y = gp$ZZ.km, 
-             ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
-             ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
-  true <- sapply(x_grid, f, 0, p)
-  est <- sapply(x_grid, f_alt, 0, p_alt)
-  models <- data.frame(x=x_grid, GP=tgp_dat$y, Parametric=est, True=true)
-  models <- melt(models, id="x")
-  names(models) <- c("x", "method", "value")
-  plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-    geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
-    geom_point(data=obs, aes(x,y), alpha=0.8) + 
-    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
-    scale_colour_manual(values=cbPalette) +
-    ggtitle(paste("seed", seed))
-  print(plot_gp)
-}
-
-posteriors_plot <- function(gp, priors){
-  hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
-  posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
-  prior_curves <- ddply(posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100)
-    data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
-  })
-  plot_posteriors <- ggplot(posteriors) + 
-    #geom_density(aes(value), lwd=2) +
-    geom_histogram(aes(x=value, y=..density..), alpha=0.7) +
-    geom_line(data=prior_curves, aes(x=value, y=density), col="red", lwd=2) +
-    facet_wrap(~ variable, scale="free")
-  print(plot_posteriors)
-}
 
@@
   obs <- sim_obs(Xo, z_g, f, p, Tobs=40, seed = seed_i)
-  est <- par_est(obs)
+  alt <- par_est(obs)
+  est <- par_est_allee(obs, f, p)
   gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
@@
 #  posteriors_plot(gp, priors) # needs trace=TRUE!
-  OPT <- optimal_policy(gp, f, est$f_alt, p, est$p_alt, x_grid, h_grid, sigma_g, est$sigma_g_alt, delta, xT, profit, reward, OptTime)
+  OPT <- optimal_policy(gp, f, est$f_alt, alt$f_alt,
+                        p, est$p_alt, alt$p_alt,
+                        x_grid, h_grid, sigma_g, 
+                        est$sigma_g_alt, alt$sigma_g_alt, 
+                        delta, xT, profit, reward, OptTime)
   dt <- simulate_opt(OPT, f, p, x_grid, h_grid, x0, z_g, profit)

### 2012-12-24 inst/examples/sensitivity.Rmd
update plotting routines to compare against true (RickerAllee), alt (Ricker), and est (RickerAllee)

@@
 
-```{r}
-
-```
-
 ```{r parallel, message=FALSE, warning=FALSE, results='hide'}
@@
 ```{r }
-seed <- round(runif(32) * 1e6)
+seed <- round(runif(2) * 1e6)
 seed[1] <- 1
@@
           s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
-  gp_plot(gp, f, p, est$f_alt, est$p_alt, x_grid, obs, seed_i)
+  gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
 #  posteriors_plot(gp, priors) # needs trace=TRUE!
-  OPT <- optimal_policy(gp, f, est$f_alt, alt$f_alt,
-                        p, est$p_alt, alt$p_alt,
+  OPT <- optimal_policy(gp, f, est$f, alt$f,
+                        p, est$p, alt$p,
                         x_grid, h_grid, sigma_g, 
-                        est$sigma_g_alt, alt$sigma_g_alt, 
+                        est$sigma_g, alt$sigma_g, 
                         delta, xT, profit, reward, OptTime)
+  plot_policies(x_grid, OPT$gp_D, OPT$est_D, opt$true_D, opt$alt_D)
   dt <- simulate_opt(OPT, f, p, x_grid, h_grid, x0, z_g, profit)

### 2012-12-24 R/gp-replicate-fns.R

@@ -13,7 +13,7 @@ cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2",
 #' @param harvest a sequence of harvest values that occur durning the simulation
 #' @details varying harvest values allow the system to explore the state space, making for better training data.
 #' @export
-sim_obs <- function(Xo, z_g, f, p, Tobs = 40, seed = 1, nz = 10, harvest = sort(rep(seq(1, 2, length=5), 8))){
+sim_obs <- function(Xo, z_g, f, p, Tobs = 40, seed = 1, nz = 10, harvest = sort(rep(seq(0, 1.2, length=5), 8))){
   x <- numeric(Tobs)
   x[1] <- Xo
   set.seed(seed)
@@ -47,7 +47,7 @@ par_est_allee <- function(obs, f, p){
   f_alt <- f
   p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
   sigma_g_alt <- as.numeric(o$par[4])
-  list(f_alt = f_alt, p_alt = p_alt, sigma_g_alt = sigma_g_alt)
+  list(f = f_alt, p = p_alt, sigma_g = sigma_g_alt)
 }
 
 
@@ -86,23 +86,28 @@ optimal_policy <- function(gp, f, f_est, f_alt, p, p_est, p_alt, x_grid, h_grid,
   opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
   matrices_alt <- f_transition_matrix(f_alt, p_alt, x_grid, h_grid, sigma_g_alt)
   opt_alt <- find_dp_optim(matrices_alt, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
-  list(gp_D = opt_gp$D[,1], true_D = opt_true$D[,1], est_D = opt_estimated$D[,1], alt_D = opt_alt$D[,1])
+  
+  gp_D <- sapply(1:OptTime, function(i) opt_gp$D[,1])
+  true_D <- sapply(1:OptTime, function(i) opt_true$D[,1])
+  est_D <- sapply(1:OptTime, function(i) opt_estimated$D[,1])
+  alt_D <- sapply(1:OptTime, function(i) opt_alt$D[,1])
+  
[…]

### 2012-12-24 inst/examples/sensitivity.Rmd
fixed minor typos

--- 2012-12-24 inst/examples/sensitivity.Rmd
@@
 function(seed_i){
-  Xo <- allee + x_grid[10] # observations start from
+  Xo <- K # observations start from
   x0 <- Xo # simulation under policy starts from
@@
           s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
-  gp_plot(gp, f, p, est$f_alt, est$p_alt, x_grid, obs, seed_i)
+  gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
 #  posteriors_plot(gp, priors) # needs trace=TRUE!
-  OPT <- optimal_policy(gp, f, est$f_alt, alt$f_alt,
-                        p, est$p_alt, alt$p_alt,
+  OPT <- optimal_policy(gp, f, est$f, alt$f,
+                        p, est$p, alt$p,
                         x_grid, h_grid, sigma_g, 
-                        est$sigma_g_alt, alt$sigma_g_alt, 
+                        est$sigma_g, alt$sigma_g, 
                         delta, xT, profit, reward, OptTime)
+  plot_policies(x_grid, OPT$gp_D, OPT$est_D, OPT$true_D, OPT$alt_D)
   dt <- simulate_opt(OPT, f, p, x_grid, h_grid, x0, z_g, profit)

### 2012-12-24 R/gp-replicate-fns.R

@@ -92,7 +92,7 @@ optimal_policy <- function(gp, f, f_est, f_alt, p, p_est, p_alt, x_grid, h_grid,
   est_D <- sapply(1:OptTime, function(i) opt_estimated$D[,1])
   alt_D <- sapply(1:OptTime, function(i) opt_alt$D[,1])
   
-  list(gp_D = gp_D, true_D = true_D, est_D = est_D, , alt_D = alt_D)
+  list(gp_D = gp_D, true_D = true_D, est_D = est_D, alt_D = alt_D)
 }
 
 #' helper function to simulate the optimal policies of each model
@@ -120,7 +120,7 @@ simulate_opt <- function(OPT, f, p, x_grid, h_grid, x0, z_g, profit){
 #' @import ggplot2
 #' @return plots of stock and harvest dynamics over time and replicates
 #' @export
-sim_plots <- function(dt, seed=1){
+sim_plots <- function(dt, seed=1, harvest=FALSE){
   fish_plot <- ggplot(dt) + 
     geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
     scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))+
@@ -130,7 +130,8 @@ sim_plots <- function(dt, seed=1){
     geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
     scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1))) +
     ggtitle(paste("seed", seed))
-  print(harvest_plot)
+  if(harvest)
+    print(harvest_plot)
 }
 
 #' function to provide summary stats or the profits

### 2012-12-24 inst/examples/allee.Rmd
different parameters
a new puzzle

@@
 ```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
-opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "allee-cache/", warning=FALSE)
-read_chunk('gaussian-process-control.R')
+opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "allee/", warning=FALSE)
 library(knitcitations)
@@
 ```{r libraries, include=FALSE}
-```
-```{r graphing-options, include=FALSE}
+require(pdgControl)
+require(nonparametricbayes)
+require(reshape2)
+require(ggplot2)
+require(data.table)
+require(tgp)
+library(kernlab)
+require(MCMCpack)
+require(plyr)
+
 ```
 
-```{r RickerAllee}
+```{r graphing-options, include=FALSE}
+opts_knit$set(upload.fun = socialR::notebook.url)
+#opts_chunk$set(dev = 'Cairo_pdf', dev.args=list(""))
+opts_chunk$set(dev="png", dev.args=list(bg="transparent"))
+opts_chunk$set(comment=NA, tidy=FALSE)
+theme_set(theme_bw(base_size=16))
+theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
+             plot.background = element_rect(fill = "transparent",colour = NA))
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 ```
 
-We use the model of Allen
+Fixed priors on hyperparameters, fixed model type.
 
-```{r sdp-pars}
+```{r gp-priors}
+s2.p <- c(50,50)  # inverse gamma mean is 
+tau2.p <- c(20,1)
+d.p = c(10, 1/0.01, 10, 1/0.01)
+nug.p = c(10, 1/0.01, 10, 1/0.01) # gamma mean
+s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
+tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
+d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
+nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
+beta0_prior <- function(x, tau) dnorm(x, 0, tau)
+beta = c(0)
+priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
+```
+
+```{r sdp-pars-fixed}
+profit = function(x,h) pmin(x, h)
+delta <- 0.01
+OptTime = 20  # stationarity with unstable models is tricky thing
+reward = 0
+xT <- 0
+z_g = function() rlnorm(1, 0, sigma_g)
+z_m = function() 1+(2*runif(1, 0,  1)-1) * sigma_m
+```
+
+```{r RickerAllee-exp}
+f <- RickerAllee
+p <- c(1.1, 10, 5) 
+K <- 10
+allee <- 5
+```
+
+```{r sdp-pars-explore}
+sigma_g <- 0.05
+sigma_m <- 0.02
+x_grid <- seq(0, 1.5 * K, length=101)
+h_grid <- x_grid
 ```
@@
 
-```{r x0-obs}
-x_0_observed <- K
-xT <- 0
-set.seed(1)
+```{r }
+seed_i <- 1
+
+  Xo <- K # observations start from
+  x0 <- Xo # simulation under policy starts from
+  obs <- sim_obs(Xo, z_g, f, p, Tobs=40, seed = seed_i)
+  alt <- par_est(obs)
+  est <- par_est_allee(obs, f, p)
+  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
+          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
+          m0r1=FALSE, corr="exp", trace=FALSE, 
+          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
+          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
+  gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
+#  posteriors_plot(gp, priors) # needs trace=TRUE!
+  OPT <- optimal_policy(gp, f, est$f, alt$f,
+                        p, est$p, alt$p,
+                        x_grid, h_grid, sigma_g, 
+                        est$sigma_g, alt$sigma_g, 
+                        delta, xT, profit, reward, OptTime)
+  plot_policies(x_grid, OPT$gp_D, OPT$est_D, OPT$true_D, OPT$alt_D)
+  dt <- simulate_opt(OPT, f, p, x_grid, h_grid, x0, z_g, profit)
+  sim_plots(dt, seed=seed_i)
+  profits_stats(dt)
 ```
 
-```{r sim-with-harvest}
+```{r}
+matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.vark, x_grid, h_grid) # 
+opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=reward)
 ```
 
-We simulate data under this model, starting from a size of `r x_0_observed`.
-
-We consider the observations as ordered pairs of observations of current stock size $x_t$ and observed stock in the following year, $x_{t+1}$.
-We add the pseudo-observation of $0,0$.
-Alternatively we could condition strictly on solutions passing through the origin, though in practice the weaker assumption is often sufficient.
-
-```{r MLE-RickerAllee}
+```{r}
+matrices_same <- gp_transition_matrix(sapply(x_grid, f, 0, p), gp$ZZ.vark, x_grid, h_grid) 
+opt_same <- find_dp_optim(matrices_same, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
 ```
 
-```{r gp-priors}
+```{r}
+matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
+opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
 ```
 
-```{r gp-fit}
-```
-
-We fit a Gaussian process with
-
-```{r gp-data, dependson="gp-fit"}
-```
-
-```{r gp-plot, dependson="gp-data"}
-```
-
-```{r gp-posteriors}
-```
-
-## The transition matrix of the inferred process
-
-```{r persistence-test}
-```
-
-```{r gp-F-sim, dependson=c("persistence-test", "gp-data")}
-```
-
-```{r par-F-sim, dependson="persistence-test"}
-```
-
-```{r F-sim-plot, dependson=c("gp-F-sim", "par-F-sim")}
-```
-
-```{r gp-opt, dependson="gp-data"}
-```
-
-```{r true-opt}
-```
-
-```{r est-opt}
-```
-
-```{r policy_plot}
-```
-
-```{r othernoise}
-```
-
-```{r stationary_policy_only}
-```
-
-```{r simulate}
-```
-
-```{r tidy}
-```
-
-```{r sim-fish}
-```
-
-```{r sim-harvest}
-```
-
-```{r costs}
+```{r}
+dd <- melt(data.frame(x=x_grid, gp= opt_gp$D[,1], same=opt_same$D[,1], true=opt_true$D[,1]), id="x")
+ggplot(dd) + geom_point(aes(x, value, col=variable))
 ```

### 2012-12-24 R/gp-replicate-fns.R

@@ -13,7 +13,7 @@ cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2",
 #' @param harvest a sequence of harvest values that occur durning the simulation
 #' @details varying harvest values allow the system to explore the state space, making for better training data.
 #' @export
-sim_obs <- function(Xo, z_g, f, p, Tobs = 40, seed = 1, nz = 10, harvest = sort(rep(seq(0, 1.2, length=5), 8))){
+sim_obs <- function(Xo, z_g, f, p, Tobs = 40, seed = 1, nz = 10, harvest = sort(rep(seq(0, 1.8, length=5), 8))){
   x <- numeric(Tobs)
   x[1] <- Xo
   set.seed(seed)

### 2012-12-24 R/gp_trasition_matrix.R

@@ -10,34 +10,24 @@
 gp_transition_matrix <- function(Ef, V, x_grid, h_grid=NULL){  
   if(is.null(h_grid))
     h_grid <- x_grid
-  lapply(h_grid, gp_F, Ef, V, x_grid)
+  lapply(h_grid, gp_calc_F, Ef, V, x_grid)
 }
 
-#' Determine the transition matrix given a Guassian process
-#' 
-#' @param Ef GP mean
-#' @param V GP vector of variances
-#' @param x_grid x grid over which to define transition matrices
-#' @return a list of matrices corresponding to the state transition matrix at each harvest level h
-#' @export
-gp_F <- function(h, Ef, V, x_grid){
-  S <- Ef - h
-  F_ <- t(sapply(1:length(S), function(i){
-    if(S[i]>0) {
-      out <- dnorm(x_grid, S[i], sqrt(V[i]))  
-      #        negatives <- sum(dnorm(seq(-K,0, length=50), S[i], sqrt(V[i])))
-      #        out[1] <- out[1] + negatives
-      # out <- out/sum(out)  # handled by rownorm already
-    } else {
+
+
+gp_calc_F <- function(h, Ef, V, x_grid){
+  mu <- Ef - h
+  F_ <- t(sapply(1:length(mu), function(i){    
+    if(snap_to_grid(mu[i],x_grid) < x_grid[2]){ # 
       out <- numeric(length(x_grid))
       out[1] <- 1
       out
+    } else {
+      out <- dlnorm(x_grid/mu[i], 0, V[i])
     }
   }))
   F_ <- rownorm(F_)
 }
-
-
 
 #' Determine the transition matrix given a Guassian process
 #' 
@@ -73,21 +63,10 @@ gp_F <- function(h, Ef, V, x_grid){
 #' @return a list of matrices corresponding to the state transition matrix at each harvest level h
 #' @export
 f_transition_matrix <- function(f, p, x_grid, h_grid = NULL, sigma_g){
-  
   if(is.null(h_grid))
     h_grid <- x_grid
-  matrices_F <- lapply(h_grid, par_F, f, p, x_grid, sigma_g)
-  
-#  matrices_F <- lapply(h_grid, function(h){
-#    mu <- sapply(x_grid, f, h, p)
-#    F_true <- sapply(x_grid, function(x) dlnorm(x/mu, 0, sigma_g))
-#    F_true <- rownorm(F_true)
-#  })
-  
[…]

### 2012-12-24 inst/examples/gaussian-process-control.R

@@ -280,7 +280,7 @@ qplot(x_grid, zt10[1,]) + geom_point(aes(y=zt1[1,]), col="grey")
 
 
 ## @knitr gp-opt
-matrices_gp <- gp_transition_matrix(Ef, .01*V, x_grid, h_grid)
+matrices_gp <- gp_transition_matrix(Ef, V, x_grid, h_grid) # 
 opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=reward)
 
 ## @knitr true-opt

### 2012-12-26 inst/examples/allee.Rmd
corrected the policy calculation to correspond to f(x-h) rather than f(x) - h

@@
 ```{r graphing-options, include=FALSE}
-opts_knit$set(upload.fun = socialR::notebook.url)
+#opts_knit$set(upload.fun = socialR::notebook.url)
 #opts_chunk$set(dev = 'Cairo_pdf', dev.args=list(""))
@@
 cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
+short_sha <- gsub("(^.{10}).*", "\\1", sha)
+date <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S")
+opts_chunk$set(fig.path = paste("figure/", date, "-", short_sha, "-", sep=""))
+
 ```
@@
 ```{r}
-matrices_same <- gp_transition_matrix(sapply(x_grid, f, 0, p), gp$ZZ.vark, x_grid, h_grid) 
+matrices_same <- gp_transition_matrix(sapply(x_grid, f, 0, p),rep(sigma_g, length(gp$ZZ.vark)), x_grid, h_grid) 
 opt_same <- find_dp_optim(matrices_same, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)

### 2012-12-26 R/gp_trasition_matrix.R

@@ -16,7 +16,11 @@ gp_transition_matrix <- function(Ef, V, x_grid, h_grid=NULL){
 
 
 gp_calc_F <- function(h, Ef, V, x_grid){
-  mu <- Ef - h
+  j <- which.min(abs(x_grid - h))
+  indices <- 1:length(Ef)
+  shift <- pmax(indices-j,1)
+  mu <- Ef[shift]
+  mu[shift == 1] <- 0
   F_ <- t(sapply(1:length(mu), function(i){    
     if(snap_to_grid(mu[i],x_grid) < x_grid[2]){ # 
       out <- numeric(length(x_grid))
@@ -24,30 +28,6 @@ gp_calc_F <- function(h, Ef, V, x_grid){
       out
     } else {
       out <- dlnorm(x_grid/mu[i], 0, V[i])
-    }
-  }))
-  F_ <- rownorm(F_)
-}
-
-#' Determine the transition matrix given a Guassian process
-#' 
-#' @param Ef GP mean
-#' @param V GP vector of variances
-#' @param x_grid x grid over which to define transition matrices
-#' @return a list of matrices corresponding to the state transition matrix at each harvest level h
-#' @export
-gp_F <- function(h, Ef, V, x_grid){
-  S <- Ef - h
-  F_ <- t(sapply(1:length(S), function(i){
-    if(S[i]>0) {
-      out <- dnorm(x_grid, S[i], sqrt(V[i]))  
-      #        negatives <- sum(dnorm(seq(-K,0, length=50), S[i], sqrt(V[i])))
-      #        out[1] <- out[1] + negatives
-      # out <- out/sum(out)  # handled by rownorm already
-    } else {
-      out <- numeric(length(x_grid))
-      out[1] <- 1
-      out
     }
   }))
   F_ <- rownorm(F_)

### 2012-12-27 inst/examples/sensitivity.Rmd
updated local run

@@
 require(plyr)
-
 ```
@@
 cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
+short_sha <- gsub("(^.{10}).*", "\\1", sha)
+date <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S")
+opts_chunk$set(fig.path = paste("figure/", date, "-", short_sha, "-", sep=""))
 ```

### 2012-12-27 inst/examples/allee.Rmd
example without harvest in observation stage

--- 2012-12-24 inst/examples/allee.Rmd
@@
 cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+#sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
+#short_sha <- gsub("(^.{10}).*", "\\1", sha)
+#date <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S")
+#opts_chunk$set(fig.path = paste("figure/", date, "-", short_sha, "-", sep=""))
+
 ```
@@
 sigma_g <- 0.05
-sigma_m <- 0.02
+sigma_m <- 0.0
 x_grid <- seq(0, 1.5 * K, length=101)
@@
   x0 <- Xo # simulation under policy starts from
-  obs <- sim_obs(Xo, z_g, f, p, Tobs=40, seed = seed_i)
+  obs <- sim_obs(Xo, z_g, f, p, Tobs=40, nz=1, 
+                 harvest = sort(rep(seq(0, .01, length=8), 5)), seed = seed_i)
   alt <- par_est(obs)
@@
 ```{r}
-matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.vark, x_grid, h_grid) # 
-opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=reward)
-```
-
-```{r}
-matrices_same <- gp_transition_matrix(sapply(x_grid, f, 0, p), gp$ZZ.vark, x_grid, h_grid) 
-opt_same <- find_dp_optim(matrices_same, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
-```
-
-```{r}
-matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
-```
-
-```{r}
-dd <- melt(data.frame(x=x_grid, gp= opt_gp$D[,1], same=opt_same$D[,1], true=opt_true$D[,1]), id="x")
-ggplot(dd) + geom_point(aes(x, value, col=variable))
+est
+alt
 ```

### 2012-12-27 inst/examples/allee.Rmd
fixed observed data (needed to pass harvests individually in sim)

@@
 seed_i <- 1
-
   Xo <- K # observations start from
   x0 <- Xo # simulation under policy starts from
-  obs <- sim_obs(Xo, z_g, f, p, Tobs=40, nz=1, 
-                 harvest = sort(rep(seq(0, .01, length=8), 5)), seed = seed_i)
+  obs <- sim_obs(Xo, z_g, f, p, Tobs=35, nz=1, 
+                 harvest = sort(rep(seq(0, .9, length=7), 5)), seed = seed_i)
+
   alt <- par_est(obs)

### 2012-12-27 inst/examples/sensitivity.Rmd
fixed observed data (needed to pass harvests individually in sim)

@@
 ```{r graphing-options, include=FALSE}
-#opts_knit$set(upload.fun = socialR::notebook.url)
+opts_knit$set(upload.fun = socialR::notebook.url)
 #opts_chunk$set(dev = 'Cairo_pdf', dev.args=list(""))
@@
 cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
-short_sha <- gsub("(^.{10}).*", "\\1", sha)
-date <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S")
-opts_chunk$set(fig.path = paste("figure/", date, "-", short_sha, "-", sep=""))
+#sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
+#short_sha <- gsub("(^.{10}).*", "\\1", sha)
+#date <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S")
+#opts_chunk$set(fig.path = paste("figure/", date, "-", short_sha, "-", sep=""))
 ```
@@
 ```{r sdp-pars-explore}
-sigma_g <- 0.05
-sigma_m <- 0.02
+sigma_g <- 0.1
+sigma_m <- 0.0
 x_grid <- seq(0, 1.5 * K, length=101)

### 2012-12-27 R/gp-replicate-fns.R

@@ -13,12 +13,13 @@ cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2",
 #' @param harvest a sequence of harvest values that occur durning the simulation
 #' @details varying harvest values allow the system to explore the state space, making for better training data.
 #' @export
-sim_obs <- function(Xo, z_g, f, p, Tobs = 40, seed = 1, nz = 10, harvest = sort(rep(seq(0, 1.8, length=5), 8))){
+sim_obs <- function(Xo, z_g, f, p, Tobs = 35, seed = 1, nz = 10, 
+                    harvest = sort(rep(seq(0, .9, length=7), 5))){
   x <- numeric(Tobs)
   x[1] <- Xo
   set.seed(seed)
   for(t in 1:(Tobs-1))
-    x[t+1] = z_g() * f(x[t], h=harvest, p=p)
+    x[t+1] = z_g() * f(x[t], h=harvest[t], p=p)
   plot(x)
   obs <- data.frame(x = c(rep(0,nz), 
                           pmax(rep(0,Tobs-1), x[1:(Tobs-1)] -

### 2012-12-27 inst/examples/allee.Rmd
exploring different harvest values, initial conditions on noise, initial guesses for the MLE solvers

@@
   obs <- sim_obs(Xo, z_g, f, p, Tobs=35, nz=1, 
-                 harvest = sort(rep(seq(0, .9, length=7), 5)), seed = seed_i)
+                 harvest = sort(rep(seq(0, .8, length=7), 5)), seed = seed_i)
 

### 2012-12-27 inst/examples/sensitivity.Rmd
exploring different harvest values, initial conditions on noise, initial guesses for the MLE solvers

@@
   x0 <- Xo # simulation under policy starts from
-  obs <- sim_obs(Xo, z_g, f, p, Tobs=40, seed = seed_i)
+  obs <- sim_obs(Xo, z_g, f, p, Tobs=35, nz= 1, 
+                 harvest = sort(rep(seq(0, .8, length=7), 5)), seed = seed_i)
   alt <- par_est(obs)

### 2012-12-27 R/gp-replicate-fns.R

@@ -14,7 +14,7 @@ cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2",
 #' @details varying harvest values allow the system to explore the state space, making for better training data.
 #' @export
 sim_obs <- function(Xo, z_g, f, p, Tobs = 35, seed = 1, nz = 10, 
-                    harvest = sort(rep(seq(0, .9, length=7), 5))){
+                    harvest = sort(rep(seq(0, .8, length=7), 5))){
   x <- numeric(Tobs)
   x[1] <- Xo
   set.seed(seed)
@@ -40,7 +40,7 @@ par_est_allee <- function(obs, f, p){
     mu <- f(obs$x,0,p)
     -sum(dlnorm(obs$y, log(mu), p["s"]), log=TRUE)
   }
-  par = c(p[1] - 1, 
+  par = c(p[1] + 1, 
           p[2] - 1, 
           p[3] + 2, 
           s = sigma_g + abs(rnorm(1,0,.1)))

### 2012-12-27 inst/examples/allee.Rmd
take initial values for MLE guesses

@@
                  harvest = sort(rep(seq(0, .8, length=7), 5)), seed = seed_i)
-
-  alt <- par_est(obs)
-  est <- par_est_allee(obs, f, p)
+  alt <- par_est(obs,  init = c(r=1.5, K=mean(obs$x), s=1))
+  est <- par_est_allee(obs, f, p,  init = c(p[1]+1, p[2]-1, p[3]-2, 
+                                   s = sigma_g + abs(rnorm(1,0,.1))))
   gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,

### 2012-12-27 inst/examples/sensitivity.Rmd
take initial values for MLE guesses

--- 2012-12-27 inst/examples/sensitivity.Rmd
@@
   x0 <- Xo # simulation under policy starts from
-  obs <- sim_obs(Xo, z_g, f, p, Tobs=40, seed = seed_i)
-  alt <- par_est(obs)
-  est <- par_est_allee(obs, f, p)
+  obs <- sim_obs(Xo, z_g, f, p, Tobs=35, nz= 1, 
+                 harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
+  alt <- par_est(obs,  init = c(r=1.5, K=mean(obs$x), s=1))
+  est <- par_est_allee(obs, f, p,  init = c(p[1]+1, p[2]-1, p[3]-2, 
+                                   s = sigma_g + abs(rnorm(1,0,.1))))
   gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,

### 2012-12-27 inst/examples/allee.Rmd
local runs

--- 2012-12-27 inst/examples/allee.Rmd
@@
   x0 <- Xo # simulation under policy starts from
+```
+
+```{r}
   obs <- sim_obs(Xo, z_g, f, p, Tobs=35, nz=1, 
-                 harvest = sort(rep(seq(0, .8, length=7), 5)), seed = seed_i)
+                 harvest = sort(rep(seq(0, .7, length=7), 5)), seed = seed_i)
+```
 
+```{r}
   alt <- par_est(obs)
   est <- par_est_allee(obs, f, p)
+```
+
+```{r}
   gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
           meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
-          m0r1=FALSE, corr="exp", trace=FALSE, 
+          m0r1=FALSE, corr="exp", trace=TRUE, 
           beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
@@
   gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
-#  posteriors_plot(gp, priors) # needs trace=TRUE!
+```
+
+```{r}
+  posteriors_plot(gp, priors) # needs trace=TRUE!
+```
+
+```{r}
   OPT <- optimal_policy(gp, f, est$f, alt$f,
@@
   plot_policies(x_grid, OPT$gp_D, OPT$est_D, OPT$true_D, OPT$alt_D)
+```
+
+```{r}
 dt <- simulate_opt(OPT, f, p, x_grid, h_grid, x0, z_g, profit)
@@
 ```{r}
-est
-alt
+est$par
+alt$par
 ```

### 2012-12-27 inst/examples/sensitivity.Rmd
small noise example

--- 2012-12-27 inst/examples/sensitivity.Rmd
@@
 delta <- 0.01
-OptTime = 20  # stationarity with unstable models is tricky thing
+OptTime = 50  # stationarity with unstable models is tricky thing
 reward = 0
@@
 ```{r sdp-pars-explore}
-sigma_g <- 0.1
+sigma_g <- 0.05
 sigma_m <- 0.0
@@
 With parameters `r p`.
-
-```{r parallel, message=FALSE, warning=FALSE, results='hide'}
-require(snowfall)
-sfInit(par=TRUE, cpu=8)
-sfExportAll()
-sfLibrary(pdgControl)
-sfLibrary(nonparametricbayes)
-sfLibrary(reshape2)
-sfLibrary(ggplot2)
-sfLibrary(data.table)
-sfLibrary(tgp)
-sfLibrary(kernlab)
-sfLibrary(MCMCpack)
-sfLibrary(plyr)
-```
 
@@
   x0 <- Xo # simulation under policy starts from
-  obs <- sim_obs(Xo, z_g, f, p, Tobs=40, seed = seed_i)
-  alt <- par_est(obs)
-  est <- par_est_allee(obs, f, p)
+  obs <- sim_obs(Xo, z_g, f, p, Tobs=35, nz= 1, 
+                 harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
+
+  alt <- par_est(obs,  init = c(r=1.5, K=mean(obs$x), s=1))
+  est <- par_est_allee(obs, f, p,  init = c(p[1]+1, p[2]-1, p[3]-2, 
+                                   s = sigma_g + abs(rnorm(1,0,.1))))
+
   gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
@@
   gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
+
 #  posteriors_plot(gp, priors) # needs trace=TRUE!
+
   OPT <- optimal_policy(gp, f, est$f, alt$f,

### 2012-12-27 inst/examples/allee.Rmd
Large noise case handled correctly by using GP's parametric form for the noise (Gaussian vs lognormal)

@@
 ```{r gp-priors}
+#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
 s2.p <- c(50,50)  # inverse gamma mean is 
@@
 d.p = c(10, 1/0.01, 10, 1/0.01)
-nug.p = c(10, 1/0.01, 10, 1/0.01) # gamma mean
+nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
 s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
@@
 ```{r sdp-pars-explore}
-sigma_g <- 0.05
+sigma_g <- 0.1
 sigma_m <- 0.0
@@
 ```{r}
-  obs <- sim_obs(Xo, z_g, f, p, Tobs=35, nz=1, 
-                 harvest = sort(rep(seq(0, .7, length=7), 5)), seed = seed_i)
+  obs <- sim_obs(Xo, z_g, f, p, Tobs=35, nz=15, 
+                 harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
 ```
@@
 ```{r}
-  alt <- par_est(obs)
-  est <- par_est_allee(obs, f, p)
+  alt <- par_est(obs,  init = c(r=1.5, K=mean(obs$x), s=1))
+  est <- par_est_allee(obs, f, p,  init = c(p[1]+1, p[2]-1, p[3]-2, 
+                                   s = sigma_g + abs(rnorm(1,0,.1))))
 ```

### 2012-12-27 R/gp_trasition_matrix.R

@@ -27,7 +27,7 @@ gp_calc_F <- function(h, Ef, V, x_grid){
       out[1] <- 1
       out
     } else {
-      out <- dlnorm(x_grid/mu[i], 0, V[i])
+      out <- dnorm(x_grid, mu[i], V[i])
     }
   }))
   F_ <- rownorm(F_)

### 2012-12-28 inst/examples/allee.Rmd
Example of Ricker-Allee in oscillation regime, shows non-trivial dynamics, GP does very well (closes #22)

@@
 #inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
-s2.p <- c(50,50)  # inverse gamma mean is 
-tau2.p <- c(20,1)
-d.p = c(10, 1/0.01, 10, 1/0.01)
+s2.p <- c(5,5)  
+tau2.p <- c(5,1)
+d.p = c(10, 1/0.1, 10, 1/0.1)
 nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
@@
 xT <- 0
-z_g = function() rlnorm(1, 0, sigma_g)
-z_m = function() 1+(2*runif(1, 0,  1)-1) * sigma_m
+
 ```
@@
 f <- RickerAllee
-p <- c(1.1, 10, 5) 
+# c(5, 10, 5) is 2-cycle, c(5.5, 10, 5) is 6 cycle, 5.3 is about 4
+p <- c(5.3, 10, 5) 
 K <- 10
@@
 ```{r sdp-pars-explore}
-sigma_g <- 0.1
+sigma_g <- 0.001
 sigma_m <- 0.0
+z_g = function() rlnorm(1, 0, sigma_g)
+z_m = function() 1+(2*runif(1, 0,  1)-1) * sigma_m
 x_grid <- seq(0, 1.5 * K, length=101)
@@
 ```{r}
-  alt <- par_est(obs,  init = c(r=1.5, K=mean(obs$x), s=1))
-  est <- par_est_allee(obs, f, p,  init = c(p[1]+1, p[2]-1, p[3]-2, 
-                                   s = sigma_g + abs(rnorm(1,0,.1))))
+  alt <- par_est(obs,  init = c(r=2, K=mean(obs$x), s=1))
+  est <- par_est_allee(obs, f, p,  init = c(2, 6, 2, s = sigma_g))
 ```
+
+Which estimates a Ricker model with $r =$ `r alt$p[1]`, $K =$ `r alt$p[2]`, and the Allen allee model with $r =$ `r est$p[1]`, $K =$ `r est$p[2]` and $C =$ `r est$p[3]`.
 
@@
                         x_grid, h_grid, sigma_g, 
-                        est$sigma_g, alt$sigma_g, 
+                        sigma_g, sigma_g, # est$sigma_g, alt$sigma_g, but those ests are poor
                         delta, xT, profit, reward, OptTime)
@@
 
-```{r}
-est$par
-alt$par
-```
-
 ```{r echo=FALSE, results="asis"}

### 2012-12-28 inst/examples/allee.Rmd
in oscillatory regime with non-negligible noise.  again GP performs nearly optimally while MLE methods suffer greatly

@@
 ```{r sdp-pars-explore}
-sigma_g <- 0.001
+sigma_g <- 0.05
 sigma_m <- 0.0

### 2012-12-28 inst/examples/sensitivity.Rmd
more relaxed priors

--- 2012-12-27 inst/examples/sensitivity.Rmd
@@
 ```{r gp-priors}
-s2.p <- c(50,50)  # inverse gamma mean is 
-tau2.p <- c(20,1)
-d.p = c(10, 1/0.01, 10, 1/0.01)
-nug.p = c(10, 1/0.01, 10, 1/0.01) # gamma mean
+#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
+s2.p <- c(5,5)  
+tau2.p <- c(5,1)
+d.p = c(10, 1/0.1, 10, 1/0.1)
+nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
 s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
@@
 f <- RickerAllee
-p <- c(1.1, 10, 5) 
+p <- c(1.5, 10, 5) 
 K <- 10
@@
 ```{r sdp-pars-explore}
-sigma_g <- 0.05
+sigma_g <- 0.1
 sigma_m <- 0.0
+z_g = function() rlnorm(1, 0, sigma_g)
+z_m = function() 1+(2*runif(1, 0,  1)-1) * sigma_m
 x_grid <- seq(0, 1.5 * K, length=101)
@@
   x0 <- Xo # simulation under policy starts from
-  obs <- sim_obs(Xo, z_g, f, p, Tobs=35, nz= 1, 
+  obs <- sim_obs(Xo, z_g, f, p, Tobs=35, nz= 10, 
                  harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)

### 2012-12-28 inst/doc/nonparametric-bayes.Rmd
writing writing

@@
-Structural Uncertainty 
-========================================================
+Non-parametric approaches to optimal policy are more robust
+===========================================================
 
-Problem of structural uncertainty in managing ecological systems.  Motivate via the widepread concern of tipping points, more general concerns of complex/nonlinear dynamics.
+Carl Boettiger, Steve Munch, Marc Mangel
+
+# Introduction
+
+Problem of structural uncertainty in managing ecological systems.  Motivate via the widespread concern of tipping points, more general concerns of complex/nonlinear dynamics.
 
 Typical approaches using parametric models.
-motivation/strengths (mechanstic underpinnings, computationally simple/potentially analytic solutions.
+motivation/strengths (mechanistic underpinnings, computationally simple/potentially analytic solutions.
 Parametric structure makes it difficult to represent uncertainty accurately in areas where data is not available.
+
+# Methods
+
+```{r image-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
+opts_chunk$set(cache = TRUE, cache.path = "writeup/", warning = FALSE, message = FALSE,
+               comment = NA, tidy = FALSE, echo = FALSE)
+opts_knit$set(upload.fun = socialR::notebook.url)
+#opts_chunk$set(dev = 'Cairo_pdf', dev.args=list(""))
+opts_chunk$set(dev="png", dev.args=list(bg="transparent"))
+```
+```{r libraries, include=FALSE}
+library(pdgControl)
+library(nonparametricbayes)
+library(reshape2)
+library(ggplot2)
+library(data.table)
+library(tgp)
+library(MCMCpack)
+library(plyr)
+library(knitcitations)
+```
+```{r plotting-options, echo=FALSE, include=FALSE}
+theme_set(theme_bw(base_size=16))
+theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
+             plot.background = element_rect(fill = "transparent",colour = NA))
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+#sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
+#short_sha <- gsub("(^.{10}).*", "\\1", sha)
+#date <- format(Sys.time(), "%Y-%m-%d__%H-%M-%S")
+#opts_chunk$set(fig.path = paste("figure/", date, "-", short_sha, "-", sep=""))
+```
+
+### Discussion of state equation
+```{r stateeq}
+f <- RickerAllee
+# c(5, 10, 5) is 2-cycle, c(5.5, 10, 5) is 6 cycle, 5.3 is about 4
+p <- c(2, 10, 5) 
+K <- 10
+allee <- 5
+```
+
+```{r sdp-pars, dependson="stateeq"}
+sigma_g <- 0.05
+sigma_m <- 0.0
+z_g <- function() rlnorm(1, 0, sigma_g)
+z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
+x_grid <- seq(0, 1.5 * K, length=101)
+h_grid <- x_grid
+profit <- function(x,h) pmin(x, h)
+delta <- 0.01
+OptTime <- 20  # stationarity with unstable models is tricky thing
+reward <- 0
+xT <- 0
+seed_i <- 1
+Xo <- K # observations start from
+x0 <- Xo # simulation under policy starts from
+```
+
+with parameters $r =$ `r p[1]`, $K =$ `r p[2]` and $C =$ `r p[3]`.
+
+### Discussion of training data
+```{r obs, dependson="sdp-pars"}
+  obs <- sim_obs(Xo, z_g, f, p, Tobs=35, nz=15, 
+                 harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
+```
+
+### Discussion of maximum likelihood estimated models
+```{r mle, dependson="obs"}
+alt <- par_est(obs,  init = c(r = p[1], 
+                              K = mean(obs$x[obs$x>0]), 
+                              s = sigma_g))
+est <- par_est_allee(obs, f, p,  
+                     init = c(r = p[1] + rnorm(1), 
+                              K = p[2] + rnorm(1), 
+                              C = p[3] + rnorm(1), 
+                              s = sigma_g))
+```
+
+Which estimates a Ricker model with $r =$ `r alt$p[1]`, $K =$ `r alt$p[2]`, and the Allen Allele model with $r =$ `r est$p[1]`, $K =$ `r est$p[2]` and $C =$ `r est$p[3]`.
+
+### (Brief) Discussion of GP inference
+
+The inference of the Gaussian process Following @Munch2005
+
+```{r gp-priors}
+#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
+s2.p <- c(5,5)  
+tau2.p <- c(5,1)
+d.p = c(10, 1/0.1, 10, 1/0.1)
+nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
+s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
+tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
+d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
+nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
+beta0_prior <- function(x, tau) dnorm(x, 0, tau)
+beta = c(0)
+priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
+```
+
+```{r gp, dependson=c("obs", "gp-priors")}
+  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
+          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
+          m0r1=FALSE, corr="exp", trace=TRUE, 
+          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
+          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
+```
+
+### (More thorough) Discussion of the dynamic programming solution
+
+```{r opt, dependson=c("gp", "mle")}
+  OPT <- optimal_policy(gp, f, est$f, alt$f,
+                        p, est$p, alt$p,
+                        x_grid, h_grid, sigma_g, 
+                        sigma_g, sigma_g, # est$sigma_g, alt$sigma_g, but those ests are poor
+                        delta, xT, profit, reward, OptTime)
+```
+
+### Discussion on how we compare performance of policies
+
+```{r sim, dependson="opt"}
+dt <- simulate_opt(OPT, f, p, x_grid, h_grid, x0, z_g, profit)
+```
+
+* Replicate stochastic simulations 
+* Sensitivity analysis
+
+# Results
+
+## Figure 1:
+
+Inferred Gaussian Process compared to the true and parametric models.
+
+```{r gp_plot, dependson=c("gp", "mle")}
+gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
+```
+
+## Figure 2:
+
+```{r policies_plot, dependson="opt"}
+  plot_policies(x_grid, OPT$gp_D, OPT$est_D, OPT$true_D, OPT$alt_D)
+```
+
+## Figure 3:
+
+```{r sim_plot, dependson="sim", fig.cap="**Figure 3**: Gaussian process inference outperforms parametric estimates.
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."}
+ggplot(dt) + 
+    geom_line(aes(time, fishstock, 
+                  group=interaction(reps, method), 
+                  color=method), alpha=.1) +
+    scale_colour_manual(values=cbPalette, 
+                        guide = guide_legend(override.aes = list(alpha = 1)))
+```
+
+# Discussion / Conclusion
+
+* Non-parametric methods have received far too little attention in ecological modeling efforts that are aimed at improved conservation planning and decision making support.
+
+* Importance of non-parametric approaches in conservation planning / resource management / decision theory.
+
+* Decision-theoretic tools such as optimal control calculations rely on robust _forecasting_ more strongly than they rely on accurate _mechanistic_ relationships.
+
+* Adapting a non-parametric approach requires modification of existing methods for decision theory.
+We have illustrated how this might be done in the context of stochastic dynamic programming, opening the door for substantial further research into how these applications might be improved.
+
+```{r echo=FALSE, results="asis"}
+#bibliography("html")
+```
+
+# Appendix / Supplementary Materials
+
+## MCMC posterior distributions and convergence analysis
+
+```{r posteriors, dependson="gp"}
+  posteriors_plot(gp, priors) # needs trace=TRUE!
+```
+
+## Tables of nuisance parameters, sensitivity analysis
+
+## Code, "Research Compendium"

### 2012-12-28 inst/doc/nonparametric-bayes.Rmd
writing writing

@@
 Carl Boettiger, Steve Munch, Marc Mangel
+
+# Abstract
 
@@
 
-# Methods
+### Discussion of previous decision theory literature on the vulnerability of parametric methods
+
+* Parametric vs structural uncertainty
+* Alternative approaches: POMDP, etc.
+
+# Approach and Methods
 
@@
 
+To illustrate (the pitfalls of decision theory based on parametric assumptions and compare the performance of a non-parametric approach) these challenges, we use a simple parametric model for a single species [derived from fist principles by @Allen].
+This model contains an Allee effect, or level below which the population is not self-sustaining and shrinks to zero [@Courchamp].
+Model containing such "Tipping point" phenomena have become of particular interest ...
+
+$$\begin{align}
+X_{t+1} &= Z_t f(S_t) \\
+S_t &= X_t - h_t \\
+f(S_t) &= e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
+\end{align}$$
+
 ```{r sdp-pars, dependson="stateeq"}
@@
 x0 <- Xo # simulation under policy starts from
+Tobs <- 35
 ```
 
-with parameters $r =$ `r p[1]`, $K =$ `r p[2]` and $C =$ `r p[3]`.
+Where $Z_t$ is multiplicative noise function with mean 1, representing stochastic growth.
+We will consider log-normal noise with shape parameter $\sigma_g$.
+We start with an example in which the parameters are $r =$ `r p[1]`, $K =$ `r p[2]`, $C =$ `r p[3]`, and  $\sigma_g =$ `r sigma_g`.
 
 ### Discussion of training data
+
+Both parametric and nonparametric approaches will require some training data on which to base their model of the process.
+We generate the training data under the model described in Eq 1 for `r Tobs` timesteps, under a known but not necessarily optimal sequence of harvest intensities, $h_t$.
+For simplicity we imagine a fishery that started from zero harvest pressure and has been gradually increasing the harvest.
+(Motivation, alternatives, stationarity, examples without a stable node (limit-cycle models), examples based on observations near a stable node alone, and why that isn't impossible).
+
 ```{r obs, dependson="sdp-pars"}
-  obs <- sim_obs(Xo, z_g, f, p, Tobs=35, nz=15, 
+  obs <- sim_obs(Xo, z_g, f, p, Tobs=Tobs, nz=15, 
                  harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
@@
 ```
+
+We estimate two parametric models from the data using a maximum likelihood approach.
+The first model is structurally identical to the true model (Eq 1), differing only in that it's parameters are estimated from the observed data rather than given.
+The alternative model is the Ricker model, which is structurally similar and commonly assumed
+
+(MLE models will assume the noise is log-normal, which it is in the simulation).
 
@@
 
-### (More thorough) Discussion of the dynamic programming solution
+###  Discussion of the dynamic programming solution
+
+(More thorough, but general-audience targeted.  Technical details and code provided in appendices).
 
@@
 
-```{r gp_plot, dependson=c("gp", "mle")}
+```{r gp_plot, dependson=c("gp", "mle"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
+Graph shows the expected value for the function $f$ under each model.
+Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.
+The training data is also shown as black points.
+(The GP is conditioned on 0,0, shown as a pseudo-data point).
+"}
 gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
@@
 
-```{r policies_plot, dependson="opt"}
+```{r policies_plot, dependson="opt", fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
+Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
   plot_policies(x_grid, OPT$gp_D, OPT$est_D, OPT$true_D, OPT$alt_D)
@@
 
+## Figure 4:
+
+Sensitivity analysis.  Distribution of yield over stochastic realizations.
+
 # Discussion / Conclusion
@@
 
+* Anticipate improved relative performance in higher dimensional examples
+
+* Constant escapement
+
+## Future directions
+
+* Multiple species
+* Online learning
+* Multiple step-ahead predictions
+* Additional uncertainties
+* Improving inference of optimal policy from the GP
+
 ```{r echo=FALSE, results="asis"}
@@
 
-```{r posteriors, dependson="gp"}
+```{r posteriors, dependson="gp", fig.cap = "Histogram of posterior distributions for the estimated Gaussian Process shown in Figure 1.  Prior distributions overlaid."}
   posteriors_plot(gp, priors) # needs trace=TRUE!
@@
 
+ @Gramacy2005
+
 ## Tables of nuisance parameters, sensitivity analysis
 
-## Code, "Research Compendium"
+### List of hyperparameters, prior distributions and their parameters
+
+## Reproducible code, "Research Compendium"

### 2012-12-31 inst/doc/nonparametric-bayes.Rmd
working draft

@@
 ===========================================================
+
+```{r knit-settings, cache=FALSE}
+      options(xtable.print.comment=FALSE)
+      options(xtable.type = 'latex', table.placement="H")
+      opts_chunk$set(warning=FALSE, message=FALSE, comment=NA, tidy=FALSE,
+                     echo=FALSE)
+      opts_knit$set(progress = TRUE, verbose = TRUE)
+      opts_chunk$set(dev = 'Cairo_pdf', fig.width=5.5, fig.height=4,
+                     cache.path = 'cache-pdf/', cache=TRUE)
+```
 
@@
 
-Problem of structural uncertainty in managing ecological systems.  Motivate via the widespread concern of tipping points, more general concerns of complex/nonlinear dynamics.
+The problem of structural uncertainty in managing ecological systems.
 
-Typical approaches using parametric models.
-motivation/strengths (mechanistic underpinnings, computationally simple/potentially analytic solutions.
-Parametric structure makes it difficult to represent uncertainty accurately in areas where data is not available.
+* Most management recommendations from the ecological literature / management policies based on (motivated by) parametric models.  _Preference to pitch towards policy or theoretical literature?_
 
-### Discussion of previous decision theory literature on the vulnerability of parametric methods
+* Background discussion on the importance/success of parametric modeling; (e.g. @Levins1966, etc., up through @Geritz2011a)
 
-* Parametric vs structural uncertainty
-* Alternative approaches: POMDP, etc.
+* Background on the concerns of structural uncertainty -- we don't have the right models.  (Also: measurement uncertainty, parameter uncertainty, unobserved states, boundary conditions, etc.)
+
+* Hierarchical Bayesian approach has provided a natural way to address these from a statistical standpoint.  Successes and challenges from the parametric route: e.g. @Cressie2009.
+
+* Among pitfalls of these approaches: particularly difficult to apply in management context.
+Optimization-based (Decision-theoretic) approaches to ecological management need to be precisely parameterized about everything, even the uncertainty itself [@Polasky2011] (discussion of @Brozovic2011 SDP example in context of threshold system would perhaps be useful as well).
+
+* Management goals / decision-theoretic approaches need accurate prediction over relevant (short?) timescales more than accurate (but incomplete or noisy estimated) mechanisms.
+
+* Nonparametric (machine-learning?) approaches may offer the benefit of the hierarchical Bayesian approach without the practical and computational limitations of their parametric kin.
+Non-parametric models are flexible enough to take maximum advantage of the data available, while being appropriately ambiguous about the dynamics of a system in regions of parameter space that have been poorly or never sampled.
+
+* Nonparametric approaches are beginning to appear more frequently in ecological and conservation literature (Species distribution models/maxent), including the Gaussian process based approach used here [@Munch2005].
+However, such approaches have yet to be applied to the decision-theoretic framework that could guide management decisions, where we expect them to excel for several reasons (1) the ability to make accurate forecasts by more closely approximating the underlying process where data is available (2) remaining appropriately ambiguous where data is not available (3) remaining computationally simple enough to avoid some pitfalls common to hierarchical parametric approaches.
 
@@
 
-```{r image-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
-opts_chunk$set(cache = TRUE, cache.path = "writeup/", warning = FALSE, message = FALSE,
-               comment = NA, tidy = FALSE, echo = FALSE)
-opts_knit$set(upload.fun = socialR::notebook.url)
-#opts_chunk$set(dev = 'Cairo_pdf', dev.args=list(""))
-opts_chunk$set(dev="png", dev.args=list(bg="transparent"))
-```
 ```{r libraries, include=FALSE}
@@
 ```{r plotting-options, echo=FALSE, include=FALSE}
-theme_set(theme_bw(base_size=16))
+theme_set(theme_bw(base_size=10))
 theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
@@
 ### Discussion of state equation
+
 ```{r stateeq}
@@
 
-To illustrate (the pitfalls of decision theory based on parametric assumptions and compare the performance of a non-parametric approach) these challenges, we use a simple parametric model for a single species [derived from fist principles by @Allen].
-This model contains an Allee effect, or level below which the population is not self-sustaining and shrinks to zero [@Courchamp].
-Model containing such "Tipping point" phenomena have become of particular interest ...
+Concerns over the potential for tipping points in ecological dynamics [@Scheffer2001] highlight the dangers of uncertainty in ecological management and pose a substantial challenge to existing decision-theoretic approaches [@Brozovic2011].
+To compare the performance of nonparametric and parametric approaches in an example that is easy to conceptualize, we will focus on a simple parametric model for a single species [derived from fist principles by @Allen2005a] as our underlying "reality".
 
-$$\begin{align}
+\begin{align}
 X_{t+1} &= Z_t f(S_t) \\
@@
 f(S_t) &= e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
-\end{align}$$
+\end{align}
+
+As a low-dimensional system completely described by three parameters, this scenario should if anything be favorable to a parametric-based approach.
+This model contains an Allee effect, or tipping point, below which the population is not self-sustaining and shrinks to zero [@Courchamp2008].
 
@@
 
+### Brief discussion on choice of model parameters, nuisance parameters
+
 Where $Z_t$ is multiplicative noise function with mean 1, representing stochastic growth.
@@
 Both parametric and nonparametric approaches will require some training data on which to base their model of the process.
-We generate the training data under the model described in Eq 1 for `r Tobs` timesteps, under a known but not necessarily optimal sequence of harvest intensities, $h_t$.
+We generate the training data under the model described in Eq 1 for `r Tobs` time steps, under a known but not necessarily optimal sequence of harvest intensities, $h_t$.
 For simplicity we imagine a fishery that started from zero harvest pressure and has been gradually increasing the harvest.
@@
 ### Discussion of maximum likelihood estimated models
+
 ```{r mle, dependson="obs"}
@@
 est <- par_est_allee(obs, f, p,  
-                     init = c(r = p[1] + rnorm(1), 
-                              K = p[2] + rnorm(1), 
-                              C = p[3] + rnorm(1), 
+                     init = c(r = p[1] + 1, 
+                              K = p[2] + 2, 
+                              C = p[3] + 2, 
                               s = sigma_g))
@@
 
-The inference of the Gaussian process Following @Munch2005
+* The use of Gaussian processes for inference in dynamical systems [introduced by @Kocijan2005]
+
+* Gaussian processes in ecological literature [@Munch2005]
+
+* Our methodology (e.g. following @Munch2005).
 
@@
 
+Outside the ecological community, Gaussian processes have been introduced into optimization literature, but as an approximation to the value function rather than to underlying dynamics [@Deisenroth2009].
+
 ```{r opt, dependson=c("gp", "mle")}
@@
 * Replicate stochastic simulations 
-* Sensitivity analysis
+* Sensitivity analysis (Figure 4).
 
@@
 
-Inferred Gaussian Process compared to the true and parametric models.
+_Shows the inferred Gaussian Process compared to the true and parametric models.  Refer to the appendix for details on the GP posteriors, etc._
 
-```{r gp_plot, dependson=c("gp", "mle"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
+```{r gp_plot, dependson=c("gp", "mle", "plotting-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
 Graph shows the expected value for the function $f$ under each model.
@@
 "}
-gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
+tgp_dat <- 
+    data.frame(  x = gp$XX[[1]], 
+                 y = gp$ZZ.km, 
+                 ymin = gp$ZZ.km - 2 * sqrt(gp$ZZ.ks2), 
+                 ymax = gp$ZZ.km + 2 * sqrt(gp$ZZ.ks2),
+                 ymin2 = gp$ZZ.mean - 2 * sqrt(gp$ZZ.vark), 
+                 ymax2 = gp$ZZ.mean + 2 * sqrt(gp$ZZ.vark))
+  true <- sapply(x_grid, f, 0, p)
+  alt_mean <- sapply(x_grid, alt$f, 0, alt$p)
+  est_mean <- sapply(x_grid, est$f, 0, est$p)
+  models <- data.frame(x=x_grid, GP=tgp_dat$y, 
+                       Parametric=est_mean, 
+                       True=true, 
+                       Structural=alt_mean)
+  models <- melt(models, id="x")
+  names(models) <- c("x", "method", "value")
+
+ggplot(tgp_dat) + 
+  geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
+  geom_ribbon(aes(x,y,ymin=ymin2,ymax=ymax2), fill="gray60") +
+  geom_line(data=models, aes(x, value, col=method), alpha=0.8, lwd=1) + 
+  geom_point(data=obs, aes(x,y)) + 
+  xlab(expression(X[t])) + ylab(expression(X[t+1])) +
+  scale_colour_manual(values=cbPalette) 
 ```
@@
 
-```{r policies_plot, dependson="opt", fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
+_The take-home message, showing that the GP is closest to the optimal strategy, while the parametric methods are less accurate.
+Visualizing the policy may be more useful for the technical reader, the general audience may prefer Figure 3 showing all replicates of the population collapse under the parametric model and not under the GP._
+
+```{r policies_plot, dependson=c("opt", "plotting-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
 Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
-  plot_policies(x_grid, OPT$gp_D, OPT$est_D, OPT$true_D, OPT$alt_D)
+policies <- 
+  melt(data.frame(stock=x_grid, 
+                  GP = x_grid[OPT$gp_D], 
+                  Parametric = x_grid[OPT$est_D],
+                  True = x_grid[OPT$true_D],
+                  Structural = x_grid[OPT$alt_D]),
+       id="stock")
+names(policies) <- c("stock", "method", "value")
+
+ggplot(policies, aes(stock, stock - value, color=method)) +
+    geom_line(alpha=0.7, lwd=1) + 
+    xlab("stock size") + ylab("escapement")  +
+    scale_colour_manual(values=cbPalette)
 ```
@@
 
-```{r sim_plot, dependson="sim", fig.cap="**Figure 3**: Gaussian process inference outperforms parametric estimates.
+_Figure 3 is a less abstract and more visceral visualization of the take-home message, with the structurally inaccurate model leading universally to a collapse of the fishery and very few profits, while the Gaussian process performs nearly optimally.
+The parametric approach even with the correct underlying structure does not perform optimally, choosing in this case to under-fish (may need to show harvest dynamics since that is not clear from the figure!
+Also isn't general, sometimes does optimally, sometimes over-fishes.
+Perhaps need to show more examples.)  May need to show profits too?_
+
+```{r sim_plot, dependson=c("sim", "plotting-options"), fig.cap="Gaussian process inference outperforms parametric estimates.
 Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."}
@@
 
-Sensitivity analysis.  Distribution of yield over stochastic realizations.
+_Shows the sensitivity analysis.
+A histogram of distribution of yield over stochastic realizations, showing that the qualitative results do not depend on the stochastic realization of the training data here, or on the parameters of the underlying model, though quantitative differences are visible._
 
@@
 
-* Constant escapement
+* Discuss constant escapement in model, in policies.
+
+* Limitations of this comparison: Are the maximum-likelihood solutions a straw man?
+
+* Discussion of alternative related approaches: POMDP/MOMDP,
 
@@
 * Multiple step-ahead predictions
-* Additional uncertainties
+* Explicitly accomidating additional uncertainties
 * Improving inference of optimal policy from the GP
@@
 
-```{r posteriors, dependson="gp", fig.cap = "Histogram of posterior distributions for the estimated Gaussian Process shown in Figure 1.  Prior distributions overlaid."}
-  posteriors_plot(gp, priors) # needs trace=TRUE!
+```{r posteriors, dependson=c("gp", "plotting-options"), fig.cap = "Histogram of posterior distributions for the estimated Gaussian Process shown in Figure 1.  Prior distributions overlaid."}
+hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
+posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
+  prior_curves <- ddply(posteriors, "variable", function(dd){
+  grid <- seq(min(dd$value), max(dd$value), length = 100)
+  data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
+})
+
+ggplot(posteriors) + 
+    geom_histogram(aes(x=value, y=..density..), alpha=0.7) +
+    geom_line(data=prior_curves, aes(x=value, y=density), col="red") +
+    facet_wrap(~ variable, scale="free")
 ```
@@
 
-### List of hyperparameters, prior distributions and their parameters
+### List of hyper-parameters, prior distributions and their parameters
 

### 2013-02-06 inst/doc/nonparametric-bayes.Rmd
minor edits

@@
 ```
-
-Carl Boettiger, Steve Munch, Marc Mangel
-
-# Abstract
-
-# Introduction
-
-The problem of structural uncertainty in managing ecological systems.
-
-* Most management recommendations from the ecological literature / management policies based on (motivated by) parametric models.  _Preference to pitch towards policy or theoretical literature?_
-
-* Background discussion on the importance/success of parametric modeling; (e.g. @Levins1966, etc., up through @Geritz2011a)
-
-* Background on the concerns of structural uncertainty -- we don't have the right models.  (Also: measurement uncertainty, parameter uncertainty, unobserved states, boundary conditions, etc.)
-
-* Hierarchical Bayesian approach has provided a natural way to address these from a statistical standpoint.  Successes and challenges from the parametric route: e.g. @Cressie2009.
-
-* Among pitfalls of these approaches: particularly difficult to apply in management context.
-Optimization-based (Decision-theoretic) approaches to ecological management need to be precisely parameterized about everything, even the uncertainty itself [@Polasky2011] (discussion of @Brozovic2011 SDP example in context of threshold system would perhaps be useful as well).
-
-* Management goals / decision-theoretic approaches need accurate prediction over relevant (short?) timescales more than accurate (but incomplete or noisy estimated) mechanisms.
-
-* Nonparametric (machine-learning?) approaches may offer the benefit of the hierarchical Bayesian approach without the practical and computational limitations of their parametric kin.
-Non-parametric models are flexible enough to take maximum advantage of the data available, while being appropriately ambiguous about the dynamics of a system in regions of parameter space that have been poorly or never sampled.
-
-* Nonparametric approaches are beginning to appear more frequently in ecological and conservation literature (Species distribution models/maxent), including the Gaussian process based approach used here [@Munch2005].
-However, such approaches have yet to be applied to the decision-theoretic framework that could guide management decisions, where we expect them to excel for several reasons (1) the ability to make accurate forecasts by more closely approximating the underlying process where data is available (2) remaining appropriately ambiguous where data is not available (3) remaining computationally simple enough to avoid some pitfalls common to hierarchical parametric approaches.
-
-# Approach and Methods
 
@@
 cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-#sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
-#short_sha <- gsub("(^.{10}).*", "\\1", sha)
-#date <- format(Sys.time(), "%Y-%m-%d__%H-%M-%S")
-#opts_chunk$set(fig.path = paste("figure/", date, "-", short_sha, "-", sep=""))
 ```
 
-### Discussion of state equation
+Carl Boettiger, Marc Mangel, Steve Munch
 
-```{r stateeq}
-f <- RickerAllee
-# c(5, 10, 5) is 2-cycle, c(5.5, 10, 5) is 6 cycle, 5.3 is about 4
-p <- c(2, 10, 5) 
-K <- 10
-allee <- 5
-```
+# Abstract
 
-Concerns over the potential for tipping points in ecological dynamics [@Scheffer2001] highlight the dangers of uncertainty in ecological management and pose a substantial challenge to existing decision-theoretic approaches [@Brozovic2011].
-To compare the performance of nonparametric and parametric approaches in an example that is easy to conceptualize, we will focus on a simple parametric model for a single species [derived from fist principles by @Allen2005a] as our underlying "reality".
+# Introduction
 
-\begin{align}
-X_{t+1} &= Z_t f(S_t) \\
-S_t &= X_t - h_t \\
-f(S_t) &= e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
-\end{align}
+The problem of structural uncertainty in managing ecological systems.
 
-As a low-dimensional system completely described by three parameters, this scenario should if anything be favorable to a parametric-based approach.
-This model contains an Allee effect, or tipping point, below which the population is not self-sustaining and shrinks to zero [@Courchamp2008].
+* Most management recommendations from the ecological literature
+/ management policies based on (motivated by) parametric models.
+_Preference to pitch towards policy or theoretical literature?_
+
+* Background discussion on the importance/success of parametric modeling;
+(e.g. @Levins1966, etc., up through @Geritz2011a)
+
+* Background on the concerns of structural uncertainty -- we don't have
+the right models.  (Also: measurement uncertainty, parameter uncertainty,
+unobserved states, boundary conditions, etc.)
+
+* Hierarchical Bayesian approach has provided a natural way to address
+these from a statistical standpoint.  Successes and challenges from the
+parametric route: e.g. @Cressie2009.
+
+* Among pitfalls of these approaches: particularly difficult to apply in
+management context. Optimization-based (Decision-theoretic) approaches to
+ecological management need to be precisely parameterized about everything,
+even the uncertainty itself [@Polasky2011] (discussion of @Brozovic2011
+SDP example in context of threshold system would perhaps be useful
+as well).
+
+* Management goals / decision-theoretic approaches need accurate
+prediction over relevant (short?) timescales more than accurate (but
+incomplete or noisy estimated) mechanisms.
+
+* Nonparametric (machine-learning?) approaches may offer the benefit of
+the hierarchical Bayesian approach without the practical and computational
+limitations of their parametric kin.  Non-parametric models are flexible
+enough to take maximum advantage of the data available, while being
+appropriately ambiguous about the dynamics of a system in regions of
+parameter space that have been poorly or never sampled.
+
+* Nonparametric approaches are beginning to appear more frequently
+in ecological and conservation literature (Species distribution
+models/maxent), including the Gaussian process based approach used here
+[@Munch2005].  However, such approaches have yet to be applied to the
+decision-theoretic framework that could guide management decisions,
+where we expect them to excel for several reasons (1) the ability to
+make accurate forecasts by more closely approximating the underlying
+process where data is available (2) remaining appropriately ambiguous
+where data is not available (3) remaining computationally simple enough
+to avoid some pitfalls common to hierarchical parametric approaches.
+
+# Approach and Methods
 
@@
 
+### (Brief) Discussion of GP inference
+
+* The use of Gaussian processes for inference in dynamical systems [introduced by @Kocijan2005]
+
+* Gaussian processes in ecological literature [@Munch2005]
+
+* Our methodology (e.g. following @Munch2005).
+
+###  Discussion of the dynamic programming solution
+
+_(More thorough, but general-audience targeted.  Technical details and code provided in appendices)._
+
+The fishery management problem over an infinite time horizon can be stated as:
+
+\begin{align}
+& \max_{ \{h_t\} \geq 0 } \mathbf{E} \lbrace \sum_0^\infty \delta^t \Pi(h_t) \rbrace \\
+& \mathrm{s.t.}  \\
+ & X_t = Z_t f\left(S_{t-1}\right) \\
+ & S_t = X_t - h_t \\
+ & X_t  \geq 0 
+\end{align}
+
+Where $\mathbf{E}$ is the expectation operator, $\delta$ the discount rate, $\Pi(h_t)$ the profit expected from a harvest of $h_t$, and other terms as in Eq.
+(1).
+For simplicity, we have assumed that profits depend only on the chosen harvest; simplifying further we will usually consider profits to be proportional to harvest, $\Pi(h_t) = h_t$.
+
+Once the posterior Gaussian process (GP) has been estimated [e.g.
+see @Munch2005], it is necessary to adapt it in place of the parametric equation for the stochastic dynamic programming (SDP) solution [see @Mangel1988 for a detailed description of parametric SDP methods] to the optimal policy.
+The essense of the idea is straight forward -- we will use the estimated GP in place of the parametric growth function to determine the stochastic transition matrix on which the SDP calculations are based.
+
+The posterior Gaussian process is completely defined by the expected value and covariance matrix at a defined set of training points.
+For simplicty we will consider a these points to fall on a discrete, uniform grid $x$ of `r length(x_grid)` points from `r min(x_grid)` to `r max(x_grid)` (1.5 times the positive equilibrium $K$).
+Again to keep things simple we will use this same grid discritization for the parametric approach.
+Other options for choosing the grid points, including collocation methods and functional basis expansion (or even using Guassian processes in place of the discrete optimization; an entirely different context in which GP can be used in SDP, see [@Deisenroth2009]) could also be considered.
+
+The transition matrix $\mathbf{F}$ is thus an `r length(x_grid)` by `r length(x_grid)` matrix for which the ${i,j}$ entry gives the probability of transitioning into state $x_i$ given that the system is in state $x_j$ in the previous timestep.
+To generate the transition matrix based on the posterior GP, we need only the expected values at each grid point and the corresponding variances (the diagonal of the covariance matrix), as shown in Figure 1.
+Given the mean at each gridpoint as the length `r length(x_grid)` vector $E$ and variance $V$, the probability of transitioning from state $x_i$ to state $x_j$ is simply $\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)$, where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and variance $\sigma^2$.
+Strictly speaking, the transition probability should be calculated by integrating the normal density over the bin of width $\Delta$ centered at $x_j$.
+For a sufficiently fine grid that $f(x_j) \approx f(x_j + \Delta)$, it is sufficient to calculate the density at $x_j$ and then row-normalize the transition matrix.
+
+## Pseudocode for the determining the transtion matrix from the GP
+```r
+for(h in h_grid)
+  F_h = for(x_j in grid)
+          for(i in 1:N) 
+            dnorm(x_j, mu[i]-h, V[i])
+```
+
+A transition matrix for each of the parametric models $f$ is calculated using the log-normal density with mean $f(x_i)$ and log-variance as estimated by maximum likelihood.
+From the discrete transition matrix we may write down the Bellman recursion defining the the stochastic dynamic programming iteration:
+
+\begin{equation}
+V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
+\end{equation}
+
+where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$ is control (harvest level) chosen.
+Numerically, the maximization is accomplished as follows.
+Consider the set of possible control values to be the discrete `r length(h_grid)` values corresponding the the grid of stock sizes.
+Then for each $h_t$ there is a corresponding transition matrix $\mathbf{F}_h$ determined as described above but with mean $\mu = x_j - h_t$.
+Let $\vec{V_t}$ be the vector whose $i$th element corresponds to the value of having stock $x_i$ at time $t$.
+Then let $\Pi_h$ be the vector whose $i$th element indicates the profit from harvesting at intensity $h_t$ given a population $x_i$ (e.g.
+$\max(x_i, h_t)$ since one cannot harvest more fish then the current population size).
+Then the Bellman recursion can be given in matrix form as
+
+$$V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)$$
+
+where the sum is element by element and the expectation is computed by the matrix multiplication $\mathbf{F} V_{t+1}$.
+
+### Pseudocode for the Bellman iteration
+
+```r
+ V1 <- sapply(1:length(h_grid), function(h){
+      delta * F[[h]] %*% V +  profit(x_grid, h_grid[h]) 
+    })
+    # find havest, h that gives the maximum value
+    out <- sapply(1:gridsize, function(j){
+      value <- max(V1[j,], na.rm = T) # each col is a diff h, max over these
+      index <- which.max(V1[j,])  # store index so we can recover h's 
+      c(value, index) # returns both profit value & index of optimal h.  
+    })
+    # Sets V[t+1] = max_h V[t] at each possible state value, x
+    V <- out[1,]                        # The new value-to-go
+    D[,OptTime-time+1] <- out[2,]       # The index positions
+```
+
+_Currently this shows the literal R code, should be adapted_
+
+### Discussion on how we compare performance of policies
+
+* Replicate stochastic simulations 
+* Sensitivity analysis (Figure 4).
+
+### Discussion of state equation
+
+Concerns over the potential for tipping points in ecological dynamics [@Scheffer2001] highlight the dangers of uncertainty in ecological management and pose a substantial challenge to existing decision-theoretic approaches [@Brozovic2011].
+To compare the performance of nonparametric and parametric approaches in an example that is easy to conceptualize, we will focus on a simple parametric model for a single species [derived from fist principles by @Allen2005a] as our underlying "reality".
+
+\begin{align}
+X_{t+1} &= Z_t f(S_t) \\
+S_t &= X_t - h_t \\
+f(S_t) &= e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
+\end{align}
+
+As a low-dimensional system completely described by three parameters, this scenario should if anything be favorable to a parametric-based approach.
+This model contains an Allee effect, or tipping point, below which the population is not self-sustaining and shrinks to zero [@Courchamp2008].
+
+```{r stateeq}
+f <- RickerAllee
+p <- c(2, 10, 5) 
+K <- 10
+allee <- 5
+```
+
 ### Brief discussion on choice of model parameters, nuisance parameters
@@
 
-Both parametric and nonparametric approaches will require some training data on which to base their model of the process.
-We generate the training data under the model described in Eq 1 for `r Tobs` time steps, under a known but not necessarily optimal sequence of harvest intensities, $h_t$.
-For simplicity we imagine a fishery that started from zero harvest pressure and has been gradually increasing the harvest.
-(Motivation, alternatives, stationarity, examples without a stable node (limit-cycle models), examples based on observations near a stable node alone, and why that isn't impossible).
+Both parametric and nonparametric approaches will require some training
+data on which to base their model of the process.  We generate the
+training data under the model described in Eq 1 for `r Tobs` time
+steps, under a known but not necessarily optimal sequence of harvest
+intensities, $h_t$.  For simplicity we imagine a fishery that started
+from zero harvest pressure and has been gradually increasing the harvest.
+(Motivation, alternatives, stationarity, examples without a stable node
+(limit-cycle models), examples based on observations near a stable node
+alone, and why that isn't impossible).
+
+Results
+====================================================
+
+### Discussion of maximum likelihood estimated models
+
+We estimate two parametric models from the data using a maximum likelihood
+approach.  The first model is structurally identical to the true model
+(Eq 1), differing only in that it's parameters are estimated from the
+observed data rather than given.  The alternative model is the Ricker
+model, which is structurally similar and commonly assumed
+
+(MLE models will assume the noise is log-normal, which it is in the
+simulation).
+
+Which estimates a Ricker model with $r =$ `r alt$p[1]`, $K =$ `r alt$p[2]`, and the Allen Allele model with $r =$ `r est$p[1]`, $K =$ `r est$p[2]` and $C =$ `r est$p[3]`.
 
@@
 ```
-
-### Discussion of maximum likelihood estimated models
 
@@
 ```
-
-We estimate two parametric models from the data using a maximum likelihood approach.
-The first model is structurally identical to the true model (Eq 1), differing only in that it's parameters are estimated from the observed data rather than given.
-The alternative model is the Ricker model, which is structurally similar and commonly assumed
-
-(MLE models will assume the noise is log-normal, which it is in the simulation).
-
-Which estimates a Ricker model with $r =$ `r alt$p[1]`, $K =$ `r alt$p[2]`, and the Allen Allele model with $r =$ `r est$p[1]`, $K =$ `r est$p[2]` and $C =$ `r est$p[3]`.
-
-### (Brief) Discussion of GP inference
-
-* The use of Gaussian processes for inference in dynamical systems [introduced by @Kocijan2005]
-
-* Gaussian processes in ecological literature [@Munch2005]
-
-* Our methodology (e.g. following @Munch2005).
 
@@
 
-###  Discussion of the dynamic programming solution
-
-(More thorough, but general-audience targeted.  Technical details and code provided in appendices).
-
-Outside the ecological community, Gaussian processes have been introduced into optimization literature, but as an approximation to the value function rather than to underlying dynamics [@Deisenroth2009].
-
 ```{r opt, dependson=c("gp", "mle")}
@@
 
-### Discussion on how we compare performance of policies
-
 ```{r sim, dependson="opt"}
@@
 ```
-
-* Replicate stochastic simulations 
-* Sensitivity analysis (Figure 4).
-
-# Results
 

### 2013-02-06 inst/doc/nonparametric-bayes.tex
additional helper files

\documentclass[author-year, review]{elsarticle} %review=doublespace preprint=single 5p=2 column
\usepackage{amsmath, amsfonts, amssymb}  % extended mathematics
% My package additions
\usepackage[hyphens]{url}
\usepackage{lineno} % add 
%\linenumbers % turns line numbering on 
\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables

%% Redefines the elsarticle footer
\makeatletter
\def\ps@pprintTitle{%
 \let\@oddhead\@empty
 \let\@evenhead\@empty
 \def\@oddfoot{\it \hfill\today}%
 \let\@evenfoot\@oddfoot}
\makeatother

% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in

\usepackage{microtype}
\usepackage{fancyhdr}
\pagestyle{fancy}
\pagenumbering{arabic}

\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

%%% Syntax Highlighting for code  %%%
%%% Adapted from knitr book %%% 
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{enumerate}
\usepackage{ctable}
\usepackage{float}

% This is needed because raggedright in table elements redefines \\:
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash
\usepackage[normalem]{ulem}
\newcommand{\textsubscr}[1]{\ensuremath{_{\scriptsize\textrm{#1}}}}

% Configure hyperlinks package
\usepackage[breaklinks=true,linktocpage,pdftitle={},pdfauthor={},colorlinks]{hyperref}
\hypersetup{breaklinks=true, pdfborder={0 0 0}}

% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

\VerbatimFootnotes % allows verbatim text in footnotes

% Pandoc header

\begin{document}
\begin{frontmatter}
  \title{}
  \author[cstar]{Carl Boettiger\corref{cor1}}
  \author[noaa]{Stephan B. Munch}
  \author[cstar]{Marc Mangel}
  \ead{cboettig@ucsc.edu}
  \cortext[cor1]{Corresponding author, cboettig@ucsc.edu}
  \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA 95064, USA}
  \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}
 \end{frontmatter}

\section{Non-parametric approaches to optimal policy are more robust}

Carl Boettiger, Marc Mangel, Steve Munch

\section{Abstract}

\section{Introduction}

The problem of structural uncertainty in managing ecological systems.

\begin{itemize}
\item
  Most management recommendations from the ecological literature /
  management policies based on (motivated by) parametric models.
  \emph{Preference to pitch towards policy or theoretical literature?}
\item
  Background discussion on the importance/success of parametric
  modeling; (e.g. Levins (1966), etc., up through Geritz and Kisdi
  (2011))
\item
  Background on the concerns of structural uncertainty -- we don't have
  the right models. (Also: measurement uncertainty, parameter
  uncertainty, unobserved states, boundary conditions, etc.)
\item
  Hierarchical Bayesian approach has provided a natural way to address
  these from a statistical standpoint. Successes and challenges from the
  parametric route: e.g. Cressie et al. (2009).
\item
  Among pitfalls of these approaches: particularly difficult to apply in
  management context. Optimization-based (Decision-theoretic) approaches
  to ecological management need to be precisely parameterized about
  everything, even the uncertainty itself (Polasky et al. 2011)
  (discussion of Brozović and Schlenker (2011) SDP example in context of
  threshold system would perhaps be useful as well).
\item
  Management goals / decision-theoretic approaches need accurate
  prediction over relevant (short?) timescales more than accurate (but
  incomplete or noisy estimated) mechanisms.
\item
  Nonparametric (machine-learning?) approaches may offer the benefit of
  the hierarchical Bayesian approach without the practical and
  computational limitations of their parametric kin. Non-parametric
  models are flexible enough to take maximum advantage of the data
  available, while being appropriately ambiguous about the dynamics of a
  system in regions of parameter space that have been poorly or never
  sampled.
\item
  Nonparametric approaches are beginning to appear more frequently in
  ecological and conservation literature (Species distribution
  models/maxent), including the Gaussian process based approach used
  here (Munch et al. 2005). However, such approaches have yet to be
  applied to the decision-theoretic framework that could guide
  management decisions, where we expect them to excel for several
  reasons (1) the ability to make accurate forecasts by more closely
  approximating the underlying process where data is available (2)
  remaining appropriately ambiguous where data is not available (3)
  remaining computationally simple enough to avoid some pitfalls common
  to hierarchical parametric approaches.
\end{itemize}

\section{Approach and Methods}

\subsubsection{Discussion of state equation}

Concerns over the potential for tipping points in ecological dynamics
(Scheffer et al. 2001) highlight the dangers of uncertainty in
ecological management and pose a substantial challenge to existing
decision-theoretic approaches (Brozović and Schlenker 2011). To compare
the performance of nonparametric and parametric approaches in an example
that is easy to conceptualize, we will focus on a simple parametric
model for a single species (derived from fist principles by Allen et al.
2005) as our underlying ``reality''.

\begin{align}
X_{t+1} &= Z_t f(S_t) \\
S_t &= X_t - h_t \\
f(S_t) &= e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
\end{align}

As a low-dimensional system completely described by three parameters,
this scenario should if anything be favorable to a parametric-based
approach. This model contains an Allee effect, or tipping point, below
which the population is not self-sustaining and shrinks to zero
(Courchamp, Berec, and Gascoigne 2008).

\subsubsection{Brief discussion on choice of model parameters, nuisance
parameters}

Where $Z_t$ is multiplicative noise function with mean 1, representing
stochastic growth. We will consider log-normal noise with shape
parameter $\sigma_g$. We start with an example in which the parameters
are $r =$ \texttt{2}, $K =$ \texttt{10}, $C =$ \texttt{5}, and
$\sigma_g =$ \texttt{0.05}.

\subsubsection{Discussion of training data}

Both parametric and nonparametric approaches will require some training
data on which to base their model of the process. We generate the
training data under the model described in Eq 1 for \texttt{35} time
steps, under a known but not necessarily optimal sequence of harvest
intensities, $h_t$. For simplicity we imagine a fishery that started
from zero harvest pressure and has been gradually increasing the
harvest. (Motivation, alternatives, stationarity, examples without a
stable node (limit-cycle models), examples based on observations near a
stable node alone, and why that isn't impossible).

\subsubsection{Discussion of maximum likelihood estimated models}

We estimate two parametric models from the data using a maximum
likelihood approach. The first model is structurally identical to the
true model (Eq 1), differing only in that it's parameters are estimated
from the observed data rather than given. The alternative model is the
Ricker model, which is structurally similar and commonly assumed

(MLE models will assume the noise is log-normal, which it is in the
simulation).

Which estimates a Ricker model with $r =$ \texttt{1.8501}, $K =$
\texttt{9.8091}, and the Allen Allele model with $r =$ \texttt{2.8079},
$K =$ \texttt{11.8235} and $C =$ \texttt{7.2159}.

\subsubsection{(Brief) Discussion of GP inference}

\begin{itemize}
\item
  The use of Gaussian processes for inference in dynamical systems
  {[}introduced by @Kocijan2005{]}
\item
  Gaussian processes in ecological literature (Munch et al. 2005)
\item
  Our methodology (e.g.~following Munch et al. (2005)).
\end{itemize}

\subsubsection{Discussion of the dynamic programming solution}

\emph{(More thorough, but general-audience targeted. Technical details
and code provided in appendices).}

The fishery management problem over an infinite time horizon can be
stated as:

\begin{align}
& \max_{ \{h_t\} \geq 0 } \mathbf{E} \lbrace \sum_0^\infty \delta^t \Pi(h_t) \rbrace \\
& \mathrm{s.t.}  \\
 & X_t = Z_t f\left(S_{t-1}\right) \\
 & S_t = X_t - h_t \\
 & X_t  \geq 0 
\end{align}

Where $\mathbf{E}$ is the expectation operator, $\delta$ the discount
rate, $\Pi(h_t)$ the profit expected from a harvest of $h_t$, and other
terms as in Eq. (1). For simplicity, we have assumed that profits depend
only on the chosen harvest; simplifying further we will usually consider
profits to be proportional to harvest, $\Pi(h_t) = h_t$.

Once the posterior Gaussian process (GP) has been estimated (e.g.~see
Munch et al. 2005), it is necessary to adapt it in place of the
parametric equation for the stochastic dynamic programming (SDP)
solution (see Mangel and Clark 1988 for a detailed description of
parametric SDP methods) to the optimal policy. The essense of the idea
is straight forward -- we will use the estimated GP in place of the
parametric growth function to determine the stochastic transition matrix
on which the SDP calculations are based.

The posterior Gaussian process is completely defined by the expected
value and covariance matrix at a defined set of training points. For
simplicty we will consider a these points to fall on a discrete, uniform
grid $x$ of \texttt{101} points from \texttt{0} to \texttt{15} (1.5
times the positive equilibrium $K$). Again to keep things simple we will
use this same grid discritization for the parametric approach. Other
options for choosing the grid points, including collocation methods and
functional basis expansion (or even using Guassian processes in place of
the discrete optimization; an entirely different context in which GP can
be used in SDP, see {[}@Deisenroth2009{]}) could also be considered.

The transition matrix $\mathbf{F}$ is thus an \texttt{101} by
\texttt{101} matrix for which the ${i,j}$ entry gives the probability of
transitioning into state $x_i$ given that the system is in state $x_j$
in the previous timestep. To generate the transition matrix based on the
posterior GP, we need only the expected values at each grid point and
the corresponding variances (the diagonal of the covariance matrix), as
shown in Figure 1. Given the mean at each gridpoint as the length
\texttt{101} vector $E$ and variance $V$, the probability of
transitioning from state $x_i$ to state $x_j$ is simply
$\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)$, where
$\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
variance $\sigma^2$. Strictly speaking, the transition probability
should be calculated by integrating the normal density over the bin of
width $\Delta$ centered at $x_j$. For a sufficiently fine grid that
$f(x_j) \approx f(x_j + \Delta)$, it is sufficient to calculate the
density at $x_j$ and then row-normalize the transition matrix.

\subsection{Pseudocode for the determining the transtion matrix from the
GP}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{for(h in h_grid)}
  \NormalTok{F_h = for(x_j in grid)}
          \NormalTok{for(i in }\DecValTok{1}\NormalTok{:N) }
            \KeywordTok{dnorm}\NormalTok{(x_j, mu[i]-h, V[i])}
\end{Highlighting}
\end{Shaded}

A transition matrix for each of the parametric models $f$ is calculated
using the log-normal density with mean $f(x_i)$ and log-variance as
estimated by maximum likelihood. From the discrete transition matrix we
may write down the Bellman recursion defining the the stochastic dynamic
programming iteration:

\begin{equation}
V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
\end{equation}

where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$ is
control (harvest level) chosen. Numerically, the maximization is
accomplished as follows. Consider the set of possible control values to
be the discrete \texttt{101} values corresponding the the grid of stock
sizes. Then for each $h_t$ there is a corresponding transition matrix
$\mathbf{F}_h$ determined as described above but with mean
$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element
corresponds to the value of having stock $x_i$ at time $t$. Then let
$\Pi_h$ be the vector whose $i$th element indicates the profit from
harvesting at intensity $h_t$ given a population $x_i$ (e.g.
$\max(x_i, h_t)$ since one cannot harvest more fish then the current
population size). Then the Bellman recursion can be given in matrix form
as

\[V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)\]

where the sum is element by element and the expectation is computed by
the matrix multiplication $\mathbf{F} V_{t+1}$.

\subsection{Pseudocode for the Bellman iteration}

\begin{Shaded}
\begin{Highlighting}[]
 \NormalTok{V1 <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
      \NormalTok{delta * F[[h]] %*% V +  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
    \NormalTok{\})}
    \CommentTok{# find havest, h that gives the maximum value}
    \NormalTok{out <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
      \NormalTok{value <- }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
      \NormalTok{index <- }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
      \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
    \NormalTok{\})}
    \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
    \NormalTok{V <- out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <- out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
\end{Highlighting}
\end{Shaded}

\emph{Currently this shows the literal R code, should be adapted}

\subsubsection{Discussion on how we compare performance of policies}

\begin{itemize}
\item
  Replicate stochastic simulations
\item
  Sensitivity analysis (Figure 4).
\end{itemize}

\section{Results}

\subsection{Figure 1:}

\emph{Shows the inferred Gaussian Process compared to the true and
parametric models. Refer to the appendix for details on the GP
posteriors, etc.}

\begin{figure}[htbp]
\centering
\includegraphics{figure/gp_plot.pdf}
\caption{Graph of the inferred Gaussian process compared to the true
process and maximum-likelihood estimated process. Graph shows the
expected value for the function $f$ under each model. Two standard
deviations from the estimated Gaussian process covariance with (light
grey) and without (darker grey) measurement error are also shown. The
training data is also shown as black points. (The GP is conditioned on
0,0, shown as a pseudo-data point).}
\end{figure}

\subsection{Figure 2:}

\emph{The take-home message, showing that the GP is closest to the
optimal strategy, while the parametric methods are less accurate.
Visualizing the policy may be more useful for the technical reader, the
general audience may prefer Figure 3 showing all replicates of the
population collapse under the parametric model and not under the GP.}

\begin{figure}[htbp]
\centering
\includegraphics{figure/policies_plot.pdf}
\caption{The steady-state optimal policy (infinite boundary) calculated
under each model. Policies are shown in terms of target escapement,
$S_t$, as under models such as this a constant escapement policy is
expected to be optimal (Reed 1979).}
\end{figure}

\subsection{Figure 3:}

\emph{Figure 3 is a less abstract and more visceral visualization of the
take-home message, with the structurally inaccurate model leading
universally to a collapse of the fishery and very few profits, while the
Gaussian process performs nearly optimally. The parametric approach even
with the correct underlying structure does not perform optimally,
choosing in this case to under-fish (may need to show harvest dynamics
since that is not clear from the figure! Also isn't general, sometimes
does optimally, sometimes over-fishes. Perhaps need to show more
examples.) May need to show profits too?}

\begin{figure}[htbp]
\centering
\includegraphics{figure/sim_plot.pdf}
\caption{Gaussian process inference outperforms parametric estimates.
Shown are 100 replicate simulations of the stock dynamics (eq 1) under
the policies derived from each of the estimated models, as well as the
policy based on the exact underlying model.}
\end{figure}

\subsection{Figure 4:}

\emph{Shows the sensitivity analysis. A histogram of distribution of
yield over stochastic realizations, showing that the qualitative results
do not depend on the stochastic realization of the training data here,
or on the parameters of the underlying model, though quantitative
differences are visible.}

\section{Discussion / Conclusion}

\begin{itemize}
\item
  Non-parametric methods have received far too little attention in
  ecological modeling efforts that are aimed at improved conservation
  planning and decision making support.
\item
  Importance of non-parametric approaches in conservation planning /
  resource management / decision theory.
\item
  Decision-theoretic tools such as optimal control calculations rely on
  robust \emph{forecasting} more strongly than they rely on accurate
  \emph{mechanistic} relationships.
\item
  Adapting a non-parametric approach requires modification of existing
  methods for decision theory. We have illustrated how this might be
  done in the context of stochastic dynamic programming, opening the
  door for substantial further research into how these applications
  might be improved.
\item
  Anticipate improved relative performance in higher dimensional
  examples
\item
  Discuss constant escapement in model, in policies.
\item
  Limitations of this comparison: Are the maximum-likelihood solutions a
  straw man?
\item
  Discussion of alternative related approaches: POMDP/MOMDP,
\end{itemize}

\subsection{Future directions}

\begin{itemize}
\item
  Multiple species
\item
  Online learning
\item
  Multiple step-ahead predictions
\item
  Explicitly accomidating additional uncertainties
\item
  Improving inference of optimal policy from the GP
\end{itemize}

\section{Appendix / Supplementary Materials}

\subsection{MCMC posterior distributions and convergence analysis}

\begin{figure}[htbp]
\centering
\includegraphics{figure/posteriors.pdf}
\caption{Histogram of posterior distributions for the estimated Gaussian
Process shown in Figure 1. Prior distributions overlaid.}
\end{figure}

@Gramacy2005

\subsection{Tables of nuisance parameters, sensitivity analysis}

\subsubsection{List of hyper-parameters, prior distributions and their
parameters}

\subsection{Reproducible code, ``Research Compendium''}

Allen, Linda J. S., Jesse F. Fagan, Göran Högnäs, and Henrik Fagerholm.
2005. ``Population extinction in discrete-time stochastic population
models with an Allee effect.'' \emph{Journal of Difference Equations and
Applications} 11 (apr): 273--293. doi:10.1080/10236190412331335373.

Brozović, Nicholas, and Wolfram Schlenker. 2011. ``Optimal management of
an ecosystem with an unknown threshold.'' \emph{Ecological Economics}
(jan): 1--14. doi:10.1016/j.ecolecon.2010.10.001.

Courchamp, Franck, Ludek Berec, and Joanna Gascoigne. 2008. \emph{Allee
Effects in Ecology and Conservation}. Oxford University Press, USA.

Cressie, Noel, Catherine a Calder, James S. Clark, Jay M. Ver Hoef, and
Christopher K. Wikle. 2009. ``Accounting for uncertainty in ecological
analysis: the strengths and limitations of hierarchical statistical
modeling.'' \emph{Ecological Applications} 19 (apr): 553--70.

Geritz, Stefan a. H., and Eva Kisdi. 2011. ``Mathematical ecology: why
mechanistic models?.'' \emph{Journal of mathematical biology} (dec).
doi:10.1007/s00285-011-0496-3.

Levins, Richard. 1966. ``The strategy of model building in population
biology.'' \emph{American Scientist} 54: 421--431.

Mangel, Marc, and Colin W. Clark. 1988. \emph{Dynamic Modeling in
Behavioral Ecology}. Princeton: Princeton University Press.

Munch, Stephan B., Melissa L. Snover, George M. Watters, and Marc
Mangel. 2005. ``A unified treatment of top-down and bottom-up control of
reproduction in populations.'' \emph{Ecology Letters} 8 (may): 691--695.
doi:10.1111/j.1461-0248.2005.00766.x.

Polasky, Stephen, Stephen R. Carpenter, Carl Folke, and Bonnie Keeler.
2011. ``Decision-making under great uncertainty: environmental
management in an era of global change.'' \emph{Trends in ecology \&
evolution} (may): 1--7. doi:10.1016/j.tree.2011.04.007.

Reed, William J. 1979. ``Optimal escapement levels in stochastic and
deterministic harvesting models.'' \emph{Journal of Environmental
Economics and Management} 6 (dec): 350--363.
doi:10.1016/0095-0696(79)90014-7.

Scheffer, Marten, Stephen R. Carpenter, J. A. Foley, C. Folke, and B.
Walker. 2001. ``Catastrophic shifts in ecosystems.'' \emph{Nature} 413
(oct): 591--6. doi:10.1038/35098000.

\bibliography{}

\end{document}

### 2013-02-13 inst/doc/nonparametric-bayes.Rmd
writing

@@
 
+```{r stateeq}
+f <- RickerAllee
+p <- c(2, 10, 5) 
+K <- 10
+allee <- 5
+```
+
 ```{r sdp-pars, dependson="stateeq"}
@@
 
-Where $\mathbf{E}$ is the expectation operator, $\delta$ the discount rate, $\Pi(h_t)$ the profit expected from a harvest of $h_t$, and other terms as in Eq.
-(1).
-For simplicity, we have assumed that profits depend only on the chosen harvest; simplifying further we will usually consider profits to be proportional to harvest, $\Pi(h_t) = h_t$.
+Where $\mathbf{E}$ is the expectation operator, $\delta$ the discount
+rate, $\Pi(h_t)$ the profit expected from a harvest of $h_t$, and other
+terms as in Eq. (1).  For simplicity, we have assumed that profits depend
+only on the chosen harvest; simplifying further we will usually consider
+profits to be proportional to harvest, $\Pi(h_t) = h_t$.
 
-Once the posterior Gaussian process (GP) has been estimated [e.g.
-see @Munch2005], it is necessary to adapt it in place of the parametric equation for the stochastic dynamic programming (SDP) solution [see @Mangel1988 for a detailed description of parametric SDP methods] to the optimal policy.
-The essense of the idea is straight forward -- we will use the estimated GP in place of the parametric growth function to determine the stochastic transition matrix on which the SDP calculations are based.
+Once the posterior Gaussian process (GP) has been estimated [e.g. see
+@Munch2005], it is necessary to adapt it in place of the parametric
+equation for the stochastic dynamic programming (SDP) solution [see
+@Mangel1988 for a detailed description of parametric SDP methods] to the
+optimal policy. The essense of the idea is straight forward -- we will use
+the estimated GP in place of the parametric growth function to determine
+the stochastic transition matrix on which the SDP calculations are based.
 
-The posterior Gaussian process is completely defined by the expected value and covariance matrix at a defined set of training points.
-For simplicty we will consider a these points to fall on a discrete, uniform grid $x$ of `r length(x_grid)` points from `r min(x_grid)` to `r max(x_grid)` (1.5 times the positive equilibrium $K$).
-Again to keep things simple we will use this same grid discritization for the parametric approach.
-Other options for choosing the grid points, including collocation methods and functional basis expansion (or even using Guassian processes in place of the discrete optimization; an entirely different context in which GP can be used in SDP, see [@Deisenroth2009]) could also be considered.
+The posterior Gaussian process is completely defined by the expected value
+and covariance matrix at a defined set of training points.  For simplicty
+we will consider a these points to fall on a discrete, uniform grid $x$
+of `r length(x_grid)` points from `r min(x_grid)` to `r max(x_grid)`
+(1.5 times the positive equilibrium $K$).  Again to keep things simple
+we will use this same grid discritization for the parametric approach.
+Other options for choosing the grid points, including collocation methods
+and functional basis expansion (or even using Guassian processes in place
+of the discrete optimization; an entirely different context in which GP
+can be used in SDP, see [@Deisenroth2009]) could also be considered.
 
-The transition matrix $\mathbf{F}$ is thus an `r length(x_grid)` by `r length(x_grid)` matrix for which the ${i,j}$ entry gives the probability of transitioning into state $x_i$ given that the system is in state $x_j$ in the previous timestep.
-To generate the transition matrix based on the posterior GP, we need only the expected values at each grid point and the corresponding variances (the diagonal of the covariance matrix), as shown in Figure 1.
-Given the mean at each gridpoint as the length `r length(x_grid)` vector $E$ and variance $V$, the probability of transitioning from state $x_i$ to state $x_j$ is simply $\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)$, where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and variance $\sigma^2$.
-Strictly speaking, the transition probability should be calculated by integrating the normal density over the bin of width $\Delta$ centered at $x_j$.
-For a sufficiently fine grid that $f(x_j) \approx f(x_j + \Delta)$, it is sufficient to calculate the density at $x_j$ and then row-normalize the transition matrix.
+The transition matrix $\mathbf{F}$ is thus an `r length(x_grid)`
+by `r length(x_grid)` matrix for which the ${i,j}$ entry gives the
+probability of transitioning into state $x_i$ given that the system is
+in state $x_j$ in the previous timestep.  To generate the transition
+matrix based on the posterior GP, we need only the expected values
+at each grid point and the corresponding variances (the diagonal of
+the covariance matrix), as shown in Figure 1.  Given the mean at each
+gridpoint as the length `r length(x_grid)` vector $E$ and variance $V$,
+the probability of transitioning from state $x_i$ to state $x_j$ is
+simply $\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)$,
+where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
+variance $\sigma^2$.  Strictly speaking, the transition probability should
+be calculated by integrating the normal density over the bin of width
+$\Delta$ centered at $x_j$.  For a sufficiently fine grid that $f(x_j)
+\approx f(x_j + \Delta)$, it is sufficient to calculate the density at
+$x_j$ and then row-normalize the transition matrix.
 
 ## Pseudocode for the determining the transtion matrix from the GP
+
 ```r
@@
 
-A transition matrix for each of the parametric models $f$ is calculated using the log-normal density with mean $f(x_i)$ and log-variance as estimated by maximum likelihood.
-From the discrete transition matrix we may write down the Bellman recursion defining the the stochastic dynamic programming iteration:
+A transition matrix for each of the parametric models $f$ is calculated
+using the log-normal density with mean $f(x_i)$ and log-variance as
+estimated by maximum likelihood.  From the discrete transition matrix we
+may write down the Bellman recursion defining the the stochastic dynamic
+programming iteration:
 
@@
 
-where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$ is control (harvest level) chosen.
-Numerically, the maximization is accomplished as follows.
-Consider the set of possible control values to be the discrete `r length(h_grid)` values corresponding the the grid of stock sizes.
-Then for each $h_t$ there is a corresponding transition matrix $\mathbf{F}_h$ determined as described above but with mean $\mu = x_j - h_t$.
-Let $\vec{V_t}$ be the vector whose $i$th element corresponds to the value of having stock $x_i$ at time $t$.
-Then let $\Pi_h$ be the vector whose $i$th element indicates the profit from harvesting at intensity $h_t$ given a population $x_i$ (e.g.
-$\max(x_i, h_t)$ since one cannot harvest more fish then the current population size).
+where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$
+is control (harvest level) chosen. Numerically, the maximization is
+accomplished as follows. Consider the set of possible control values to
+be the discrete `r length(h_grid)` values corresponding the the grid of
+stock sizes.  Then for each $h_t$ there is a corresponding transition
+matrix $\mathbf{F}_h$ determined as described above but with mean 
+$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element corresponds
+to the value of having stock $x_i$ at time $t$.  Then let $\Pi_h$ be
+the vector whose $i$th element indicates the profit from harvesting
+at intensity $h_t$ given a population $x_i$ (e.g. $\max(x_i, h_t)$
+since one cannot harvest more fish then the current population size).
 Then the Bellman recursion can be given in matrix form as
@@
 
-### Discussion of state equation
+## Example in a bistable system
 
-Concerns over the potential for tipping points in ecological dynamics [@Scheffer2001] highlight the dangers of uncertainty in ecological management and pose a substantial challenge to existing decision-theoretic approaches [@Brozovic2011].
-To compare the performance of nonparametric and parametric approaches in an example that is easy to conceptualize, we will focus on a simple parametric model for a single species [derived from fist principles by @Allen2005a] as our underlying "reality".
+Concerns over the potential for tipping points in ecological dynamics
+[@Scheffer2001] highlight the dangers of uncertainty in ecological
+management and pose a substantial challenge to existing decision-theoretic
+approaches [@Brozovic2011].  To compare the performance of nonparametric
+and parametric approaches in an example that is easy to conceptualize,
+we will focus on a simple parametric model for a single species [derived
+from fist principles by @Allen2005a] as our underlying "reality".
 
@@
 
-As a low-dimensional system completely described by three parameters, this scenario should if anything be favorable to a parametric-based approach.
-This model contains an Allee effect, or tipping point, below which the population is not self-sustaining and shrinks to zero [@Courchamp2008].
+Where $Z_t$ is multiplicative noise function with mean 1, representing
+stochastic growth. We will consider log-normal noise with shape parameter
+$\sigma_g$.  We start with an example in which the parameters are $r =$
+`r p[1]`, $K =$ `r p[2]`, $C =$ `r p[3]`, and  $\sigma_g =$ `r sigma_g`.
 
-```{r stateeq}
-f <- RickerAllee
-p <- c(2, 10, 5) 
-K <- 10
-allee <- 5
-```
+As a low-dimensional system completely described by three parameters, this
+scenario should if anything be favorable to a parametric-based approach.
+This model contains an Allee effect, or tipping point, below which the
+population is not self-sustaining and shrinks to zero [@Courchamp2008].
 
-### Brief discussion on choice of model parameters, nuisance parameters
+## Sample training data
 
-Where $Z_t$ is multiplicative noise function with mean 1, representing stochastic growth.
-We will consider log-normal noise with shape parameter $\sigma_g$.
-We start with an example in which the parameters are $r =$ `r p[1]`, $K =$ `r p[2]`, $C =$ `r p[3]`, and  $\sigma_g =$ `r sigma_g`.
-
-### Discussion of training data
-
-Both parametric and nonparametric approaches will require some training
+Both parametric and nonparametric approaches will require training
 data on which to base their model of the process.  We generate the
@@
 from zero harvest pressure and has been gradually increasing the harvest.
+
+<!--Should we include any emprical examples? -->
+Using data simulated from a specified model rather than empirical data
+permits the comparison against the true underlying dynamics, setting 
+a bar for the optimal performance possible.
+
 (Motivation, alternatives, stationarity, examples without a stable node
@@
 observed data rather than given.  The alternative model is the Ricker
-model, which is structurally similar and commonly assumed
-
-(MLE models will assume the noise is log-normal, which it is in the
-simulation).
-
-Which estimates a Ricker model with $r =$ `r alt$p[1]`, $K =$ `r alt$p[2]`, and the Allen Allele model with $r =$ `r est$p[1]`, $K =$ `r est$p[2]` and $C =$ `r est$p[3]`.
+model, which is structurally similar and commonly used in for such data.
 
@@
 ```
+
+(MLE models will assume the noise is log-normal, which it is in the
+simulation).
+
+Which estimates a Ricker model with $r =$ `r alt$p[1]`, $K =$ 
+`r alt$p[2]`, and the Allen Allele model with $r =$ `r est$p[1]`, $K =$
+`r est$p[2]` and $C =$ `r est$p[3]`.
 
@@
 
-_Shows the inferred Gaussian Process compared to the true and parametric models.  Refer to the appendix for details on the GP posteriors, etc._
+_Shows the inferred Gaussian Process compared to the true and parametric
+models.  Refer to the appendix for details on the GP posteriors, etc._
 
@@
 
-_The take-home message, showing that the GP is closest to the optimal strategy, while the parametric methods are less accurate.
-Visualizing the policy may be more useful for the technical reader, the general audience may prefer Figure 3 showing all replicates of the population collapse under the parametric model and not under the GP._
+_The take-home message, showing that the GP is closest to the optimal
+strategy, while the parametric methods are less accurate.  Visualizing the
+policy may be more useful for the technical reader, the general audience
+may prefer Figure 3 showing all replicates of the population collapse
+under the parametric model and not under the GP._
 
@@
 
-_Figure 3 is a less abstract and more visceral visualization of the take-home message, with the structurally inaccurate model leading universally to a collapse of the fishery and very few profits, while the Gaussian process performs nearly optimally.
-The parametric approach even with the correct underlying structure does not perform optimally, choosing in this case to under-fish (may need to show harvest dynamics since that is not clear from the figure!
-Also isn't general, sometimes does optimally, sometimes over-fishes.
-Perhaps need to show more examples.)  May need to show profits too?_
+_Figure 3 is a less abstract and more visceral visualization of the
+take-home message, with the structurally inaccurate model leading
+universally to a collapse of the fishery and very few profits, while the
+Gaussian process performs nearly optimally.  The parametric approach even
+with the correct underlying structure does not perform optimally, choosing
+in this case to under-fish (may need to show harvest dynamics since
+that is not clear from the figure! Also isn't general, sometimes does
+optimally, sometimes over-fishes.  Perhaps need to show more examples.)
+May need to show profits too?_
 
@@
 
-_Shows the sensitivity analysis.
-A histogram of distribution of yield over stochastic realizations, showing that the qualitative results do not depend on the stochastic realization of the training data here, or on the parameters of the underlying model, though quantitative differences are visible._
+_Shows the sensitivity analysis.  A histogram of distribution of yield
+over stochastic realizations, showing that the qualitative results do
+not depend on the stochastic realization of the training data here, or on
+the parameters of the underlying model, though quantitative differences
+are visible._
 
@@
 
-* Non-parametric methods have received far too little attention in ecological modeling efforts that are aimed at improved conservation planning and decision making support.
+* Non-parametric methods have received far too little attention in
+ecological modeling efforts that are aimed at improved conservation
+planning and decision making support.
 
-* Importance of non-parametric approaches in conservation planning / resource management / decision theory.
+* Importance of non-parametric approaches in conservation planning /
+resource management / decision theory.
 
-* Decision-theoretic tools such as optimal control calculations rely on robust _forecasting_ more strongly than they rely on accurate _mechanistic_ relationships.
+* Decision-theoretic tools such as optimal control calculations rely
+on robust _forecasting_ more strongly than they rely on accurate
+_mechanistic_ relationships.
 
-* Adapting a non-parametric approach requires modification of existing methods for decision theory.
-We have illustrated how this might be done in the context of stochastic dynamic programming, opening the door for substantial further research into how these applications might be improved.
+* Adapting a non-parametric approach requires modification of existing
+methods for decision theory.  We have illustrated how this might be
+done in the context of stochastic dynamic programming, opening the
+door for substantial further research into how these applications might
+be improved.
 
@@
 
-* Limitations of this comparison: Are the maximum-likelihood solutions a straw man?
+* Limitations of this comparison: Are the maximum-likelihood solutions
+a straw man?
 
@@
 
-* Multiple species
-* Online learning
-* Multiple step-ahead predictions
-* Explicitly accomidating additional uncertainties
-* Improving inference of optimal policy from the GP
+* Multiple species * Online learning * Multiple step-ahead predictions *
+Explicitly accomidating additional uncertainties * Improving inference
+of optimal policy from the GP
 
-```{r echo=FALSE, results="asis"}
-#bibliography("html")
-```
+```{r echo=FALSE, results="asis"} #bibliography("html") ```
 

### 2013-02-13 inst/doc/nonparametric-bayes.tex
writing

@@
 
-\subsubsection{Discussion of state equation}
-
-Concerns over the potential for tipping points in ecological dynamics
-(Scheffer et al. 2001) highlight the dangers of uncertainty in
-ecological management and pose a substantial challenge to existing
-decision-theoretic approaches (Brozović and Schlenker 2011). To compare
-the performance of nonparametric and parametric approaches in an example
-that is easy to conceptualize, we will focus on a simple parametric
-model for a single species (derived from fist principles by Allen et al.
-2005) as our underlying ``reality''.
-
-\begin{align}
-X_{t+1} &= Z_t f(S_t) \\
-S_t &= X_t - h_t \\
-f(S_t) &= e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
-\end{align}
-
-As a low-dimensional system completely described by three parameters,
-this scenario should if anything be favorable to a parametric-based
-approach. This model contains an Allee effect, or tipping point, below
-which the population is not self-sustaining and shrinks to zero
-(Courchamp, Berec, and Gascoigne 2008).
-
-\subsubsection{Brief discussion on choice of model parameters, nuisance
-parameters}
-
-Where $Z_t$ is multiplicative noise function with mean 1, representing
-stochastic growth. We will consider log-normal noise with shape
-parameter $\sigma_g$. We start with an example in which the parameters
-are $r =$ \texttt{2}, $K =$ \texttt{10}, $C =$ \texttt{5}, and
-$\sigma_g =$ \texttt{0.05}.
-
-\subsubsection{Discussion of training data}
-
-Both parametric and nonparametric approaches will require some training
-data on which to base their model of the process. We generate the
-training data under the model described in Eq 1 for \texttt{35} time
-steps, under a known but not necessarily optimal sequence of harvest
-intensities, $h_t$. For simplicity we imagine a fishery that started
-from zero harvest pressure and has been gradually increasing the
-harvest. (Motivation, alternatives, stationarity, examples without a
-stable node (limit-cycle models), examples based on observations near a
-stable node alone, and why that isn't impossible).
-
-\subsubsection{Discussion of maximum likelihood estimated models}
-
-We estimate two parametric models from the data using a maximum
-likelihood approach. The first model is structurally identical to the
-true model (Eq 1), differing only in that it's parameters are estimated
-from the observed data rather than given. The alternative model is the
-Ricker model, which is structurally similar and commonly assumed
-
-(MLE models will assume the noise is log-normal, which it is in the
-simulation).
-
-Which estimates a Ricker model with $r =$ \texttt{1.8501}, $K =$
-\texttt{9.8091}, and the Allen Allele model with $r =$ \texttt{2.8079},
-$K =$ \texttt{11.8235} and $C =$ \texttt{7.2159}.
-
 \subsubsection{(Brief) Discussion of GP inference}
@@
 simplicty we will consider a these points to fall on a discrete, uniform
-grid $x$ of \texttt{101} points from \texttt{0} to \texttt{15} (1.5
-times the positive equilibrium $K$). Again to keep things simple we will
-use this same grid discritization for the parametric approach. Other
-options for choosing the grid points, including collocation methods and
-functional basis expansion (or even using Guassian processes in place of
-the discrete optimization; an entirely different context in which GP can
-be used in SDP, see {[}@Deisenroth2009{]}) could also be considered.
+grid $x$ of 101 points from 0 to 15 (1.5 times the positive equilibrium
+$K$). Again to keep things simple we will use this same grid
+discritization for the parametric approach. Other options for choosing
+the grid points, including collocation methods and functional basis
+expansion (or even using Guassian processes in place of the discrete
+optimization; an entirely different context in which GP can be used in
+SDP, see {[}@Deisenroth2009{]}) could also be considered.
 
-The transition matrix $\mathbf{F}$ is thus an \texttt{101} by
-\texttt{101} matrix for which the ${i,j}$ entry gives the probability of
-transitioning into state $x_i$ given that the system is in state $x_j$
-in the previous timestep. To generate the transition matrix based on the
-posterior GP, we need only the expected values at each grid point and
-the corresponding variances (the diagonal of the covariance matrix), as
-shown in Figure 1. Given the mean at each gridpoint as the length
-\texttt{101} vector $E$ and variance $V$, the probability of
-transitioning from state $x_i$ to state $x_j$ is simply
-$\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)$, where
-$\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
+The transition matrix $\mathbf{F}$ is thus an 101 by 101 matrix for
+which the ${i,j}$ entry gives the probability of transitioning into
+state $x_i$ given that the system is in state $x_j$ in the previous
+timestep. To generate the transition matrix based on the posterior GP,
+we need only the expected values at each grid point and the
+corresponding variances (the diagonal of the covariance matrix), as
+shown in Figure 1. Given the mean at each gridpoint as the length 101
+vector $E$ and variance $V$, the probability of transitioning from state
+$x_i$ to state $x_j$ is simply
+$\mathcal{N}\left(x_j \vert  \mu = E_i, \sigma = \sqrt{V_i}\right)$,
+where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
 variance $\sigma^2$. Strictly speaking, the transition probability
@@
 \NormalTok{for(h in h_grid)}
-  \NormalTok{F_h = for(x_j in grid)}
+  \NormalTok{F_h =}\StringTok{ }\NormalTok{for(x_j in grid)}
           \NormalTok{for(i in }\DecValTok{1}\NormalTok{:N) }
@@
 accomplished as follows. Consider the set of possible control values to
-be the discrete \texttt{101} values corresponding the the grid of stock
-sizes. Then for each $h_t$ there is a corresponding transition matrix
+be the discrete 101 values corresponding the the grid of stock sizes.
+Then for each $h_t$ there is a corresponding transition matrix
 $\mathbf{F}_h$ determined as described above but with mean
@@
 
-\subsection{Pseudocode for the Bellman iteration}
+\subsubsection{Pseudocode for the Bellman iteration}
 
@@
 \begin{Highlighting}[]
- \NormalTok{V1 <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
-      \NormalTok{delta * F[[h]] %*% V +  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
+ \NormalTok{V1 <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
+      \NormalTok{delta *}\StringTok{ }\NormalTok{F[[h]] %*%}\StringTok{ }\NormalTok{V +}\StringTok{  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
     \NormalTok{\})}
     \CommentTok{# find havest, h that gives the maximum value}
-    \NormalTok{out <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
-      \NormalTok{value <- }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
-      \NormalTok{index <- }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
+    \NormalTok{out <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
+      \NormalTok{value <-}\StringTok{ }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
+      \NormalTok{index <-}\StringTok{ }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
       \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
@@
     \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
-    \NormalTok{V <- out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
-    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <- out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
+    \NormalTok{V <-}\StringTok{ }\NormalTok{out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
+    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <-}\StringTok{ }\NormalTok{out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
 \end{Highlighting}
@@
 \begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
 \item
@@
 
+\subsubsection{Example in a bistable system}
+
+Concerns over the potential for tipping points in ecological dynamics
+(Scheffer et al. 2001) highlight the dangers of uncertainty in
+ecological management and pose a substantial challenge to existing
+decision-theoretic approaches (Brozović and Schlenker 2011). To compare
+the performance of nonparametric and parametric approaches in an example
+that is easy to conceptualize, we will focus on a simple parametric
+model for a single species (derived from fist principles by Allen et al.
+2005) as our underlying ``reality''.
+
+\begin{align}
+X_{t+1} &= Z_t f(S_t) \\
+S_t &= X_t - h_t \\
+f(S_t) &= e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
+\end{align}
+
+Where $Z_t$ is multiplicative noise function with mean 1, representing
+stochastic growth. We will consider log-normal noise with shape
+parameter $\sigma_g$. We start with an example in which the parameters
+are $r =$ 2, $K =$ 10, $C =$ 5, and $\sigma_g =$ 0.05.
+
+As a low-dimensional system completely described by three parameters,
+this scenario should if anything be favorable to a parametric-based
+approach. This model contains an Allee effect, or tipping point, below
+which the population is not self-sustaining and shrinks to zero
+(Courchamp, Berec, and Gascoigne 2008).
+
+\subsubsection{Discussion of training data}
+
+Both parametric and nonparametric approaches will require some training
+data on which to base their model of the process. We generate the
+training data under the model described in Eq 1 for 35 time steps, under
+a known but not necessarily optimal sequence of harvest intensities,
+$h_t$. For simplicity we imagine a fishery that started from zero
+harvest pressure and has been gradually increasing the harvest.
+(Motivation, alternatives, stationarity, examples without a stable node
+(limit-cycle models), examples based on observations near a stable node
+alone, and why that isn't impossible).
+
 \section{Results}
+
+\subsubsection{Discussion of maximum likelihood estimated models}
+
+We estimate two parametric models from the data using a maximum
+likelihood approach. The first model is structurally identical to the
+true model (Eq 1), differing only in that it's parameters are estimated
+from the observed data rather than given. The alternative model is the
+Ricker model, which is structurally similar and commonly assumed
+
+(MLE models will assume the noise is log-normal, which it is in the
+simulation).
+
+Which estimates a Ricker model with $r =$
+
+\begin{verbatim}
+
+Error in eval(expr, envir, enclos) : object 'alt' not found
+\end{verbatim}
+
+, $K =$
+
+\begin{verbatim}
+
+Error in eval(expr, envir, enclos) : object 'alt' not found
+\end{verbatim}
+
+, and the Allen Allele model with $r =$
+
+\begin{verbatim}
+
+Error in eval(expr, envir, enclos) : object 'est' not found
+\end{verbatim}
+
+, $K =$
+
+\begin{verbatim}
+
+Error in eval(expr, envir, enclos) : object 'est' not found
+\end{verbatim}
+
+and $C =$
+
+\begin{verbatim}
+
+Error in eval(expr, envir, enclos) : object 'est' not found
+\end{verbatim}
+
+.
 
@@
 \begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
 \item
-  Multiple species
-\item
-  Online learning
-\item
-  Multiple step-ahead predictions
-\item
-  Explicitly accomidating additional uncertainties
-\item
-  Improving inference of optimal policy from the GP
+  Multiple species * Online learning * Multiple step-ahead predictions *
+  Explicitly accomidating additional uncertainties * Improving inference
+  of optimal policy from the GP
 \end{itemize}
+
+\texttt{\{r echo=FALSE, results="asis"\} \#bibliography("html")}
 
@@
 models with an Allee effect.'' \emph{Journal of Difference Equations and
-Applications} 11 (apr): 273--293. doi:10.1080/10236190412331335373.
+Applications} 11 (4-5) (apr): 273--293.
+doi:10.1080/10236190412331335373.
 
@@
 analysis: the strengths and limitations of hierarchical statistical
-modeling.'' \emph{Ecological Applications} 19 (apr): 553--70.
+modeling.'' \emph{Ecological Applications} 19 (3) (apr): 553--70.
 
 Geritz, Stefan a. H., and Eva Kisdi. 2011. ``Mathematical ecology: why
-mechanistic models?.'' \emph{Journal of mathematical biology} (dec).
+mechanistic models?'' \emph{Journal of mathematical biology} (dec).
 doi:10.1007/s00285-011-0496-3.
@@
 Levins, Richard. 1966. ``The strategy of model building in population
-biology.'' \emph{American Scientist} 54: 421--431.
+biology.'' \emph{American Scientist} 54 (4): 421--431.
 
@@
 Mangel. 2005. ``A unified treatment of top-down and bottom-up control of
-reproduction in populations.'' \emph{Ecology Letters} 8 (may): 691--695.
-doi:10.1111/j.1461-0248.2005.00766.x.
+reproduction in populations.'' \emph{Ecology Letters} 8 (7) (may):
+691--695. doi:10.1111/j.1461-0248.2005.00766.x.
 
@@
 deterministic harvesting models.'' \emph{Journal of Environmental
-Economics and Management} 6 (dec): 350--363.
+Economics and Management} 6 (4) (dec): 350--363.
 doi:10.1016/0095-0696(79)90014-7.
@@
 Walker. 2001. ``Catastrophic shifts in ecosystems.'' \emph{Nature} 413
-(oct): 591--6. doi:10.1038/35098000.
+(6856) (oct): 591--6. doi:10.1038/35098000.
 

### 2013-02-18 inst/doc/nonparametric-bayes.Rmd
reorganize content

@@
-Non-parametric approaches to optimal policy are more robust
-===========================================================
+% Carl Boettiger; Marc Mangel; Steve Munch
+% Non-parametric approaches to optimal policy are more robust
 
@@
 
-Carl Boettiger, Marc Mangel, Steve Munch
+Abstract
+=======================================================================
 
-# Abstract
-
-# Introduction
+Introduction
+=======================================================================
 
@@
 
-# Approach and Methods
+Approach and Methods
+=====================================================================
 
@@
 
-# Discussion / Conclusion
+Discussion / Conclusion
+==================================================================
 
@@
 
-## Future directions
+Future directions
+--------------------------------------------------------------------
 
@@
 
-# Appendix / Supplementary Materials
+Acknowledgments
+================
 
-## MCMC posterior distributions and convergence analysis
+This work was partially supported by the Center for Stock Assessment Research, a partnership between the University of California Santa Cruz and the Fisheries Ecology Division, Southwest Fisheries Science Center, Santa Cruz, CA and by NSF grant EF-0924195 to MM.
+
+Appendix / Supplementary Materials
+==================================
+
+MCMC posterior distributions and convergence analysis
+----------------------------------------------------------------------------
 

### 2013-03-06 inst/doc/nonparametric-bayes.Rmd
notes for marc's query

@@
-% Carl Boettiger; Marc Mangel; Steve Munch
 % Non-parametric approaches to optimal policy are more robust
@@
 
+1. Classical approaches are to fix relationship of stock recruitment.  
+2. Parameter uncertainty
+3. Gaussian processes for model uncertainty
+
 The problem of structural uncertainty in managing ecological systems.
+
+(Big literature here of course, primary approaches have been using
+parametric uncertainty and model uncertainty (e.g. belief distribution over models 
+I believe. Worth distinguishing our problem from the more trivially 
+sized problems in which the action space is discrete and consists
+of 2 or three possible choices, e.g. much of the recent Possingham et al work.)
 
@@
 
-### (Brief) Discussion of GP inference
+### Background on Gaussian Process inference
 
-* The use of Gaussian processes for inference in dynamical systems [introduced by @Kocijan2005]
+The use of Gaussian process regression (known as Kreging in the geospatial literature) to formulate a predictive model is relatively new in the context of modeling dynamical systems [@Kocijan2005] and introduced in the ecological modeling and fisheries management by @Munch2005.
+
+_Is the following  discription useful or too elementary?_
+
+The essense of the Gaussian process approach can be captured in the following thought experiment: An exhaustive parametric approach to the challenge of structural uncertainty might proceed by writing down all possible functional forms for the underlying dynamical system with all possible parameter values for each form, and then consider searching over this huge space to select the most likely model and parameters; or using a Bayesian approach, assign priors to each of these possible models and infer the posterior distribution of possible models.
+The Gaussian process approach can be thought of as a computationally efficient approximation to this approach.
+Gaussian processes represent a large class of models that can be though of as capturing or reasonably approximating the set of models in this collection.
+By modeling at the level of the process, rather than the level of parametric equation, we can more concisely capture the possible behavior of these curves.
+In place of a parametric model of the dynamical system, the Gaussian Process approach postulates a prior distribution of (n-dimensional) curves that can be though of as approximations to a large range of possible
+
+_Do we need more specifics on Gaussian process as an approximation to parametric models?
+Discussion of Gaussian process vs other machine learning / forecasting approaches that have less clear statistical foundations?
+If so, does this all belong in the discussion?_
+
+The Gaussian process is defined by a covariance kernel ...
+_discussion of our (gaussian) kernel, blah blah how specific here vs appendix?_.
+By requiring our kernel to follow a generic functional form, we can compactly describe the Gaussian process using only a handful of parameters (Table 1)
+
+A more thorough introduction to the formulation and use of Gaussian processes can be found in @Rasmussen2006.
+
+parameter       interpretation
+---------       -------------- 
+$\sigma^2$      The process noise (from the kernel)
+$\tau^2$        Variance around the mean
+$\beta$_0       The mean is given by a linear model of slope $\beta$ 
+$d$             The length-scale of the covariance function
+$n$             The observation error
+
+: Table of parameters for the Gaussian process
+
+Rather than estimate values of these parameters directly, we take a hierarchical approach of placing prior distributions on each.
+Following @Gramarcy2005 we use a Gaussian prior on $\beta_0$, an inverse gamma prior on $\sigma^2$ and $\tau^2$, a gamma prior on the observation noise $n$, and exponential prior on the length-scale $d$.
 
@@
 ###  Discussion of the dynamic programming solution
-
-_(More thorough, but general-audience targeted.  Technical details and code provided in appendices)._
 

### 2013-03-06 inst/doc/nonparametric-bayes.tex
notes for marc's query

@@
 \documentclass[author-year, review]{elsarticle} %review=doublespace preprint=single 5p=2 column
-\usepackage{amsmath, amsfonts, amssymb}  % extended mathematics
-% My package additions
+%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
 \usepackage[hyphens]{url}
@@
 \usepackage{booktabs} % book-quality tables
-
 %% Redefines the elsarticle footer
@@
 \topmargin -0.5in
+%%%%%%%%%%%%%%%% end my additions to header
 
-\usepackage{microtype}
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\pagenumbering{arabic}
-
-\usepackage{listings}
-\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}
-
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-
-%%% Syntax Highlighting for code  %%%
-%%% Adapted from knitr book %%% 
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \usepackage{fontspec}
+  \ifxetex
+    \usepackage{xltxtra,xunicode}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage{color}
 \usepackage{fancyvrb}
+\DefineShortVerb[commandchars=\\\{\}]{\|}
 \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
@@
 \newcommand{\NormalTok}[1]{{#1}}
-\usepackage{enumerate}
-\usepackage{ctable}
-\usepackage{float}
-
-% This is needed because raggedright in table elements redefines \\:
-\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
-\let\PBS=\PreserveBackslash
-\usepackage[normalem]{ulem}
-\newcommand{\textsubscr}[1]{\ensuremath{_{\scriptsize\textrm{#1}}}}
-
-% Configure hyperlinks package
-\usepackage[breaklinks=true,linktocpage,pdftitle={},pdfauthor={},colorlinks]{hyperref}
-\hypersetup{breaklinks=true, pdfborder={0 0 0}}
-
+\usepackage{longtable}
+\usepackage{graphicx}
+% We will generate all images so they have a width \maxwidth. This means
+% that they will get their normal width if they fit onto the page, but
+% are scaled down if they would overflow the margins.
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
+\else\Gin@nat@width\fi}
+\makeatother
+\let\Oldincludegraphics\includegraphics
+\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={Non-parametric approaches to optimal policy are more robust},
+            colorlinks=true,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{0}
 % Pandoc toggle for numbering sections (defaults to be off)
 \setcounter{secnumdepth}{0}
-
-\VerbatimFootnotes % allows verbatim text in footnotes
-
 % Pandoc header
@@
 \begin{frontmatter}
-  \title{}
+  \title{Non-parametric approaches to optimal policy are more robust}
   \author[cstar]{Carl Boettiger\corref{cor1}}
@@
 
-\section{Non-parametric approaches to optimal policy are more robust}
-
-Carl Boettiger, Marc Mangel, Steve Munch
-
 \section{Abstract}
@@
 The problem of structural uncertainty in managing ecological systems.
+
+(Big literature here of course, primary approaches have been using
+parametric uncertainty and model uncertainty (e.g.~belief distribution
+over models I believe. Worth distinguishing our problem from the more
+trivially sized problems in which the action space is discrete and
+consists of 2 or three possible choices, e.g.~much of the recent
+Possingham et al work.)
 
@@
 
-\subsubsection{(Brief) Discussion of GP inference}
+\subsubsection{Background on Gaussian Process inference}
+
+The use of Gaussian process regression (known as Kreging in the
+geospatial literature) to formulate a predictive model is relatively new
+in the context of modeling dynamical systems {[}@Kocijan2005{]} and
+introduced in the ecological modeling and fisheries management by Munch
+et al. (2005).
+
+\emph{Is the following discription useful or too elementary?}
+
+The essense of the Gaussian process approach can be captured in the
+following thought experiment: An exhaustive parametric approach to the
+challenge of structural uncertainty might proceed by writing down all
+possible functional forms for the underlying dynamical system with all
+possible parameter values for each form, and then consider searching
+over this huge space to select the most likely model and parameters; or
+using a Bayesian approach, assign priors to each of these possible
+models and infer the posterior distribution of possible models. The
+Gaussian process approach can be thought of as a computationally
+efficient approximation to this approach. Gaussian processes represent a
+large class of models that can be though of as capturing or reasonably
+approximating the set of models in this collection. By modeling at the
+level of the process, rather than the level of parametric equation, we
+can more concisely capture the possible behavior of these curves. In
+place of a parametric model of the dynamical system, the Gaussian
+Process approach postulates a prior distribution of (n-dimensional)
+curves that can be though of as approximations to a large range of
+possible
+
+\emph{Do we need more specifics on Gaussian process as an approximation
+to parametric models? Discussion of Gaussian process vs other machine
+learning / forecasting approaches that have less clear statistical
+foundations? If so, does this all belong in the discussion?}
+
+The Gaussian process is defined by a covariance kernel \ldots{}
+\emph{discussion of our (gaussian) kernel, blah blah how specific here
+vs appendix?}. By requiring our kernel to follow a generic functional
+form, we can compactly describe the Gaussian process using only a
+handful of parameters (Table 1)
+
+A more thorough introduction to the formulation and use of Gaussian
+processes can be found in @Rasmussen2006.
+
+\begin{longtable}[c]{@{}ll@{}}
+\hline\noalign{\medskip}
+parameter & interpretation
+\\\noalign{\medskip}
+\hline\noalign{\medskip}
+$\sigma^2$ & The process noise (from the kernel)
+\\\noalign{\medskip}
+$\tau^2$ & Variance around the mean
+\\\noalign{\medskip}
+$\beta$\_0 & The mean is given by a linear model of slope $\beta$
+\\\noalign{\medskip}
+$d$ & The length-scale of the covariance function
+\\\noalign{\medskip}
+$n$ & The observation error
+\\\noalign{\medskip}
+\hline
+\noalign{\medskip}
+\caption{Table of parameters for the Gaussian process}
+\end{longtable}
+
+Rather than estimate values of these parameters directly, we take a
+hierarchical approach of placing prior distributions on each. Following
+@Gramarcy2005 we use a Gaussian prior on $\beta_0$, an inverse gamma
+prior on $\sigma^2$ and $\tau^2$, a gamma prior on the observation noise
+$n$, and exponential prior on the length-scale $d$.
 
 \begin{itemize}
-\item
-  The use of Gaussian processes for inference in dynamical systems
-  {[}introduced by @Kocijan2005{]}
 \item
@@
 \subsubsection{Discussion of the dynamic programming solution}
-
-\emph{(More thorough, but general-audience targeted. Technical details
-and code provided in appendices).}
 
@@
 
-\subsubsection{Example in a bistable system}
+\subsection{Example in a bistable system}
 
@@
 
-\subsubsection{Discussion of training data}
+\subsection{Sample training data}
 
-Both parametric and nonparametric approaches will require some training
-data on which to base their model of the process. We generate the
-training data under the model described in Eq 1 for 35 time steps, under
-a known but not necessarily optimal sequence of harvest intensities,
-$h_t$. For simplicity we imagine a fishery that started from zero
-harvest pressure and has been gradually increasing the harvest.
+Both parametric and nonparametric approaches will require training data
+on which to base their model of the process. We generate the training
+data under the model described in Eq 1 for 35 time steps, under a known
+but not necessarily optimal sequence of harvest intensities, $h_t$. For
+simplicity we imagine a fishery that started from zero harvest pressure
+and has been gradually increasing the harvest.
+
+Using data simulated from a specified model rather than empirical data
+permits the comparison against the true underlying dynamics, setting a
+bar for the optimal performance possible.
+
 (Motivation, alternatives, stationarity, examples without a stable node
@@
 from the observed data rather than given. The alternative model is the
-Ricker model, which is structurally similar and commonly assumed
+Ricker model, which is structurally similar and commonly used in for
+such data.
 
@@
 
-Which estimates a Ricker model with $r =$
-
-\begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'alt' not found
-\end{verbatim}
-
-, $K =$
-
-\begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'alt' not found
-\end{verbatim}
-
-, and the Allen Allele model with $r =$
-
-\begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'est' not found
-\end{verbatim}
-
-, $K =$
-
-\begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'est' not found
-\end{verbatim}
-
-and $C =$
-
-\begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'est' not found
-\end{verbatim}
-
-.
+Which estimates a Ricker model with $r =$ 1.8501, $K =$ 9.8091, and the
+Allen Allele model with $r =$ 2.8079, $K =$ 11.8235 and $C =$ 7.2159.
 
@@
 
+\section{Acknowledgments}
+
+This work was partially supported by the Center for Stock Assessment
+Research, a partnership between the University of California Santa Cruz
+and the Fisheries Ecology Division, Southwest Fisheries Science Center,
+Santa Cruz, CA and by NSF~grant EF-0924195 to MM.
+
 \section{Appendix / Supplementary Materials}
@@
 
-\bibliography{}
-
 \end{document}

### 2013-03-13 inst/examples/adp-intro.Rmd
adp intro

## Introductory examples in approximate dynamic programming

```{r  setup, echo=FALSE, cache=FALSE}
require(multipleuncertainty)
require(reshape2)
require(ggplot2)
require(data.table)
````

```{r plotting, include=FALSE}
opts_knit$set(upload.fun = socialR::flickr.url)
opts_chunk$set(dev.args=list(bg="transparent"), 
               cache=TRUE, cache.path = "adp/",
               comment=NA, message=FALSE, warning=FALSE)
```

### Alogorithm 1

Based on Powell 2006, page 97.

```{r}
a <- 1.5
b <- .5
f <- function(x, z) z * a * x / (x + b)
```

We begin with a simulation method $X_{t+1} = f(X_t, Z_t)$.  For illustration, let us consider $f(X_t, Z_t) = Z_t \frac{a X_t}{b + X_t}$ with a = `r a` and b = `r b`.  We define a statespace $S$

```{r}
S <- seq(0, 1, length=10) 
```

as a uniform grid of `r length(S)` points from `r min(S)` to `r max(S)`.  We also need a value function on the state space, $C_t(S_t)$. For simplicity, we set the price of harvest at unity and the cost of harvesting at zero, so that $C_t(S_t, x_t) = \min(x_t, S_t)$.  
($C_t$ is sometimes denoted $\mathbb{\Pi}$).  We also need an action space $\chi_t$ of possible harvest values.  Again for simplicity we assume that harvest can be set to any possible state size, $\chi_t \equiv S_t$,

```r
chi <- S
```

```{r}
T <- 10
N <- 10
```

The approximate dynamic programming algorithm will perform a finite number $N$ = `r N$ iterations over a window of time $T$ =`r T` in our example.  The algorithm can be described as follows:

- **Step 0**
  - Initialize some $\tilde V_t^0(S_t)$ for all states $S_t$ 
  where the superscripts denote iterations in the forward approximation.  As we know absolutely nothing yet to base our initial guess on, we just arbitrarily set this to zero.

```{r}
V <- numeric(length(S))
```

  - Choose some initial state $S_0^1$
  We start at some initial state for $n = 1$ (superscript) and $t = 0$ (subscript). The choice of initial condition may come from the problem itself, otherwise we choose something arbitrarily.

```{r}
S_0 <- 0.5
```

  - Set $n = 1$

- **Step 1**: Choose a sample path, $\omega^n$

```r
Z <- lnorm(T)
```

- ** Step 2**: For $t = 0, 1, 2, \ldots, T$, do:

  - Solve:

$$V_t(S_t) = \max_{x_t \in \chi_t} \left(C(S_t, x_t) + \gamma \sum_{s^{\prime} \in mathcal{S}} \mathbb{P}(s^{\prime} | S_t^n, x_t) V_{t+1}^{n-1} s^{\prime} \right)

Let's start with $t=0$, $n=1$ and fix an $x_0$ from the set of $\chi$ (allowing the action space to be the same in each period, we can omit the subscript on $\chi$) to get started.  We first compute $C(S_0, x_0)$. $S_0 = S_0^1$ which we fixed in step 0b arbitrarily at `r S_0`.  Taking $x_0$ as the smallest harvest, $\min(\chi)$ = `r min(chi)` and evaluating $C(S_0,X_0) = \min(S_0, X_0)$ gives us `r min(S_0, min(chi))`, rather trivially.

The next terms depend on the value $\tilde V^0_1(s^{\prime})$ for all $s^{\prime} \in S$, which we have no idea about.  Fortunately we have assumed a value for each of these in step 0a.

We must also come up with some values for the probability $\mathbb{P}(s^{\prime} | S_1^0, x_1)$ for each state, given our current state $S_1^0$ and considered action $x_1$.  This is more straight forward, since it is determined by our one-step transition function (without simulation - recall that the single step transition is given exactly).

To do so, we evaluate the argument for each value in our action space, $x_t \in \chi_t$,

```{r}
s <- S_0

sapply(0:T, function(t){
  max(
    arg <- sapply(chi, function(x){
      min(s, x) + 
    c(x.n_t = which.max(arg), v.n_t = max(arg))

```

### 2013-03-19 inst/doc/beyond-mechanistic-models.Rmd
add code to beyond-mechanistic-models. adapt text into manuscript form

```{r}
read_chunk("nonparametric-bayes.R")
```
```{r knit-settings, echo=FALSE, include=FALSE, cache=FALSE}
```
```{r libraries, echo=FALSE, include=FALSE}
```
```{r plotting-options, echo=FALSE, include=FALSE}
```
```{r stateeq}
```
```{r sdp-pars, dependson="stateeq"}
```
```{r obs, dependson="sdp-pars"}
```
```{r mle, dependson="obs"}
```
```{r gp-priors}
```
```{r gp, dependson=c("obs", "gp-priors")}
```
```{r opt, dependson=c("gp", "mle")}
```
```{r sim, dependson="opt"}
```

Process-based models built on modern machine learning approaches 
offer a powerful and underappreciated approach for ecological
forecasting and management.  We present a novel example of how
one such example - non-parametric Gaussian processes, can be adapted
to the context of management and decision-making.

While effective machine learning algorithms have speckled the literature, 
most theoretical ecologists in particular have been wary of these 'black-box'
approaches in favor of simple, mechanistic models.  While a blind application
of machine learning approaches would indeed be dangerous grounds to stake 
either ecological understanding or management practice, such concerns are not
grounds to reject such methods out of hand.  Rather, further attention is 
warranted to identify the circumstances in which machine-learning approaches
will most likely outperform existing more mechanistic approaches, or help
crack difficult management questions where ecological or dynamical complexity
has caused traditional approaches to founder.

Here we illustrate how a machine learning model can be implemented in place
of simple mechanistic model in a decision theory (optimal control) problem,
and highlight the features of this system that make the nonparametric approach
particularly effective.

Mechanistic models have long been the gold standard of theoretical
modeling in ecology (e.g. see @Geritz2012 or @Cuddington2013).  Only by
understanding the processes involved can we make reliable long term
predictions and build an knowledge of cause and effect that guides the
hypotheses we make, the data we collect, and the management decisions
we make.  Process-based models, whether expressed in the language of
mathematics or English, identify the connection between mosquitoes and
the spread of malaria, or greenhouses gases and climate change, guiding
our approach to understand and manage these threats. Despite this central
importance, we argue here that ecologists would do well to give greater
attention to the role non-mechanistic models can play in ecological
management and decision making.  The value of these approaches is
greatest in a context where decisions are made over short time horizons,
and updated as new data becomes available.

## The alternative to process-based models: pattern-based modeling

Before we proceed further it would be useful to define our terms and
lay some greater context for the discussion. Throughout, we will distinguish
"mechanistic" or "process-based" models from models that are merely "correlative" 
or "pattern-based". Historically pattern-based modeling meant regression (usually linear
regression) and resided in methodology rooted in statistics departments
and was the primary focus of empirical ecologists, while mechanistic
modeling meant dynamical systems (ODEs, PDEs, SDEs and their discrete
kin), residing in methodology from the mathematics department and was
the primary focus of theoretical ecologists.

We believe these historical divides continue to color much of the
literature today, with theorists more skeptical than empiricists of
statistical methodology and vice versa.  Meanwhile, the ground underneath
has shifted.  Originally pattern-matching approaches could be critiqued
on the grounds of their simplicity (not everything is linear) while
mechanistic approaches could be critiqued on their tenuous connections
to data -- model parameter values such as death rates, birth rates,
etc. would be estimated in advance and then stuck into the model, rather
than estimated in the context of the model itself.  Today both of these
critiques are outdated. Computational power and hierarchical statistical
methods (particularly approximate likelihood or simulation techniques such
as particle filters and approximate Bayesian computing) have brought the
inference richer dynamical systems into their fold, while pattern-based
modeling has spawned an entirely new approach under the banner of machine
learning that has divided the statistics community (See @Bremman2001,
"The Two Cultures").   Machine learning models can represent almost
arbitrarily complex patterns and incorporate learning and decision making
strategies that have earned them the name 'artificial intelligence.'
Yet at the same time they can be far less transparent their linear
regression predecessors, making both astounding successes and startling
biases and failures like their biological namesake. More ecologists,
particularly theorists and modelers, would do well to learn both how to
take advantage of these strengths and recognize their weaknesses.

A simple example comparing the performance of the two modeling approaches
in a quantitative decision problem will help to introduce these issues.

Mechanistic modeling emphasizes the importance of capturing the correct
gross properties of a system over the tracking minute fluctuations.
For instance, in selecting and parameterizing model of a population
of conservation concern, we may be most interested in getting the
long-term behavior correct -- such as identifying if the dynamics support
persistence of the population -- rather than worrying how well they
reflect the year-to-year fluctuations.  We would certainly have good
reason to prefer such a model over alternatives which are irreconcilable
to the most basic biological processes, such as unbounded growth, or
growth curves that do not pass through the origin in a closed system. So
it may come as a surprise to realize that such obviously wrong models
can perform as well or even better than reasonable mechanistic models
in guiding ecological management and decision making.

```{r gp_plot, dependson=c("gp", "mle", "plotting-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.  Graph shows the expected value for the function $f$ under each model.  Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.  The training data is also shown as black points.  (The GP is conditioned on 0,0, shown as a pseudo-data point). "}
```
```{r policies_plot, dependson=c("opt", "plotting-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.  Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."
}
```
```{r sim_plot, dependson=c("sim", "plotting-options"), fig.cap="Gaussian process inference outperforms parametric estimates. Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."
}
```

An example of such a comparison is shown in Figure 1, in which the optimal
management decision is determined by stochastic dynamic programming
algorithm using the biologically (a) plausible and (b) the implausible
model of the population growth.

Three elements contribute to the biologically implausible phenomenological
model performing better in this context:

### 1. Relevant state space

The dynamics occur over a range of state-space in which the biologically
implausible model performs as well or better than the more plausible
model.  This is the most obvious and immediate reason why the shortcomings
that appear to make the model implausible do not make it useless.
This effect is enhanced in our example because the management actions help
drive the system towards rather than away from this region of state-space.

### 2. Predictive accuracy

This application relies only on the predictive accuracy of the model,
not an interpretation of the parameters.  Predictive accuracy is not the
goal of all modeling, as ecologists have been observing for as long as
they made models (perhaps none more memorably than @Levins1969).

Mechanistic modeling is at its most powerful not when it is used to
directly forecast future states but when it provides an understanding
of how to approach a problem.  SIR-type models from the epidemiological
literature are a good example.  While the simplest SIR models have little
predictive power over the outbreak intensity or timing at a particular
location, they provide a powerful understanding of the spread of an
infection in terms of a single, biologically meaningful parameter: $R_0$,
the basic reproductive number.  From the model, it becomes clear that
management need not vaccinate every member of the population to stop
the spread, but rather it suffices to vaccinate a sufficient fraction
of the population to reduce $R_0$ below 1.

### 3. Time scale for new data

In the sequential decision making problem we considered, we are
presented with new data after each action.  The relevant timescale
for the prediction is thus not the long-term dynamics, which would be
wildly divergent, but the dynamics over this much shorter interval.
While ecologists may be hesitant to base continual management on a model
with obviously inaccurate long-term behavior, engineers tend to consider
the problem in frequency space and gravitate to the opposite position --
a good control model should prioritize high-frequency accuracy over low
frequency accuracy.  The differences in intuition may arise from the
timescales at which each profession can typically adjust their control
variables -- much faster for a control system of a chemical plant than
state policy for a natural resource.  Still, the lesson is clear: when
facing repeated management decisions over a short timescale, such as
setting annual harvests of a fishery, it may be more valuable to use a
machine learning algorithm that makes accurate year-ahead predictions
that capture some of the high-frequency fluctuations that appear only
as noise in a mechanistic model of the long-term population dynamics.

## Advantages of non-mechanistic models

1. Can better use all of the data available.

Incorporating various sources of information into mechanistic models
can be an immensely difficult due to the increased complexity involved.
Only thanks to tandem advances in increasing computational power and
hierarchical statistical methodology have we been able to tackle such
intuitively important complexity (and the potentially new available data
that accompanies it) such as spatial distribution, heterogeneities of
space, time, and individuals, to shift to ecosystem-based approaches
from single-species based approaches.  Without the need to formulate
mechanisms, many modern machine learning algorithms can leverage potential
information from all available sources of data directly.  The algorithms
can recognize unanticipated or subtle patterns in large data sets
that enable more accurate predictions than mechanistic models that are
formulated at a more macroscopic level.  Such approaches depend critically

2. Better express ambiguity in regions where data is not available

One of the greatest strengths of mechanistic models is their greatest weakness as well.

## Better the devil you know?  The dangers of pattern-based approaches

When all we have identified is a pattern rather than a mechanism, there is
little way to guarantee that the pattern will not change in the future.
Few examples illustrate this danger more broadly than the sub-prime
mortgage crisis of 2008.  Simplifying greatly, the unquenchable demand for
mortgage backed securities had shifted the underlying dynamics on which
the phenomenological models were based.  Historical data indicated that
mortgage-backed securities involved little risk. As demand exhausted
the supply of traditional mortgages, it became profitable to invent
new mortgages that could be sold to higher risk customers, such as
those without proof of income.  The pattern-based models did not have
enough time or data to learn about the higher default rate occurring as
a result, but remained anchored too in the old patterns where default
rates were low.  While a mechanistic formulation may have anticipated the
resulting higher risk, the pattern based model must learn by seeing it.
The economy still struggles to pay off that lesson.

Of course mechanistic models may also fail to predict sudden
structural changes.  It is all too easy to overlook some slowly changing
environmental variable, particularly if temporal changes in a parameter
must be modeled explicitly if they are to be included at all.  Machine
learning approaches are frequently less rigid in their assumptions, but
can also less transparent about them.  This creates a second weakness
in the machine learning approach.

Mechanistic models are not free from this danger either. In fact they can
be more susceptible to them for the very same reasons.  A mechanistic
model has the ability to more convincingly produce out-of-sample
predictions or extrapolation.  This may allow us to determine that a
virus will spread or an endangered species will go extinct even before
we have observed

Dangers of biased estimation.

Parametric models are frequently used phenomenologically, rather than
being derived by plausible underlying mechanisms.  @Geritz2012 condemn
this practice on the basis of the common misinterpretations that result.
These mistakes arise from ignoring or misinterpreting the biological
meaning of the model parameters. The use of delay equations is a
frequent offender -- for instance, one is hard-pressed to describe a
mechanistic model consistent with Hutchinson's 1948 delay-logistic.
Assuming either scramble competition or interference competition gives
rise to two alternative delay equations, both structurally different
from the original.

Three elements contribute to this.

A process-based understanding helps us formulate hypotheses, guide
experimental design and data collection, predict outcomes, and influence
management decisions.  The alternative to a processed-based model is  a
correlative one. Statical models machine learning.

> since the late 1980s there has been some consensus amongst ecologists
that management decisions are best guided by models which are grounded
in ecological theory, and which strike a balance between too much or too
little detail describing the relevant processes (e.g., DeAngelis 1988,
Starfield 1997, Jackson et al. 2000, Carpenter 2003, Nelson et al. 2008)

machine learning algorithms

Prediction without

Example of Statistical model vs process-based model.

@Geritz2012 @Cuddington2013

Managing by rule-of-thumb

Managing by algorithm

```{r posteriors, dependson=c("gp", "plotting-options"), fig.cap = "Histogram of posterior distributions for the estimated Gaussian Process shown in Figure 1.  Prior distributions overlaid."}
```

### 2013-03-21 inst/doc/beyond-mechanistic-models.Rmd
pomp example

@@
-```{r}
+% Beyond Mechanistic Models: The case for machine learning in ecological management
+
+Abstract
+============
+
+<!-- Run R code for analysis, to be called by figures -->
+```{r, echo=FALSE, cache=FALSE, include=FALSE}
+options(xtable.print.comment=FALSE)
+options(xtable.type = 'latex', table.placement="H")
+opts_chunk$set(warning=FALSE, message=FALSE, comment=NA,
+               tidy=FALSE, echo=FALSE)
+opts_knit$set(progress = TRUE, verbose = TRUE)
+opts_chunk$set(dev = 'Cairo_pdf', fig.width=5.5, fig.height=4,
+ cache.path = 'cache-pdf/', cache=TRUE, external=TRUE)
 read_chunk("nonparametric-bayes.R")
@@
 ```
+
+Introduction
+===============================================================
 
@@
 
-Before we proceed further it would be useful to define our terms and
-lay some greater context for the discussion. Throughout, we will distinguish
-"mechanistic" or "process-based" models from models that are merely "correlative" 
-or "pattern-based". Historically pattern-based modeling meant regression (usually linear
-regression) and resided in methodology rooted in statistics departments
-and was the primary focus of empirical ecologists, while mechanistic
-modeling meant dynamical systems (ODEs, PDEs, SDEs and their discrete
-kin), residing in methodology from the mathematics department and was
-the primary focus of theoretical ecologists.
+Before we proceed further it would be useful to define our terms and lay
+some greater context for the discussion. Throughout, we will distinguish
+"mechanistic" or "process-based" models from models that are merely
+"correlative" or "pattern-based". Historically pattern-based modeling
+meant regression (usually linear regression) and resided in methodology
+rooted in statistics departments and was the primary focus of empirical
+ecologists, while mechanistic modeling meant dynamical systems (ODEs,
+PDEs, SDEs and their discrete kin), residing in methodology from
+the mathematics department and was the primary focus of theoretical
+ecologists.
 
@@
 
-A simple example comparing the performance of the two modeling approaches
-in a quantitative decision problem will help to introduce these issues.
-
 Mechanistic modeling emphasizes the importance of capturing the correct
-gross properties of a system over the tracking minute fluctuations.
+gross properties of a system over tracking minute fluctuations.
 For instance, in selecting and parameterizing model of a population
@@
 in guiding ecological management and decision making.
+
+A simple example comparing the performance of the two modeling approaches
+in a quantitative decision problem will help to introduce these issues.
+
+Models & Methods
+===========================================================================
+
+<!-- Include example of a machine learning prediction that involves biologically
+inaccurate predictions such as unbounded growth or growth at zero and yet still
+performs well (by fitting well in the region of interest?)  -->
 
@@
 model of the population growth.
+
+Discussion
+===========================================================================
 
@@
 
-Three elements contribute to this.
-
-A process-based understanding helps us formulate hypotheses, guide
-experimental design and data collection, predict outcomes, and influence
-management decisions.  The alternative to a processed-based model is  a
-correlative one. Statical models machine learning.
-
 > since the late 1980s there has been some consensus amongst ecologists
@@
 Starfield 1997, Jackson et al. 2000, Carpenter 2003, Nelson et al. 2008)
-
-machine learning algorithms
-
-Prediction without
-
-Example of Statistical model vs process-based model.
 

### 2013-03-21 inst/doc/nonparametric-bayes.Rmd
some writing
pomp example

@@
 
-* Most management recommendations from the ecological literature
-/ management policies based on (motivated by) parametric models.
-_Preference to pitch towards policy or theoretical literature?_
+Most management recommendations from the ecological literature are based on
+(or at least motivated by) parametric models. Though in principle these models
+reflect mechanistic underpinnings of biological interactions involved, in practice
+real mechanisms are often unknown and the true dynamics too complex to be captured
+by simple models [@refs]. While simple mechanistic models can nevertheless
+provide imporant insights into possible dynamics -- for instance, demonstrating
+that a vaccine does not have to be 100% effective to eliminate the transmission of a virus [@Kermack1921] -- such approaches are not well-suited for use in 
+forecasting future dynamics upon which outcomes management policy can be based.  
+Despite a long history of work emphasizing the difference between modeling 
+for understanding (generic or strategic models) and modeling for prediction (precise, specific, or tactical models) [e.g. @Levins1966], simple parametric models 
+continue to find application in predictive management contexts.  The management
+of marine fisheries is a prime example, in which simple, mechanistically 
+motivated models of stock-recruitment relationships such as Ricker or Beverton-Holt
+curves are systematically fit to data and used in solving decision-theoretic
+or optimal control problems determining resource management policies such
+as maximum sustainable yield [@refs].
 
-* Background discussion on the importance/success of parametric modeling;
-(e.g. @Levins1966, etc., up through @Geritz2011a)
+Stochastic dynamic programming @Mangel1982
 
-* Background on the concerns of structural uncertainty -- we don't have
-the right models.  (Also: measurement uncertainty, parameter uncertainty,
-unobserved states, boundary conditions, etc.)
+To address structural uncertainty in the processes involved, such mechanistic models
+can be adapted to accomidate uncertainty in the parameters or consider a set of 
+alternative models simultaneously (belief SDP)
 
-* Hierarchical Bayesian approach has provided a natural way to address
-these from a statistical standpoint.  Successes and challenges from the
-parametric route: e.g. @Cressie2009.
-
-* Among pitfalls of these approaches: particularly difficult to apply in
-management context. Optimization-based (Decision-theoretic) approaches to
-ecological management need to be precisely parameterized about everything,
-even the uncertainty itself [@Polasky2011] (discussion of @Brozovic2011
-SDP example in context of threshold system would perhaps be useful
-as well).
+Background on the concerns of structural uncertainty -- we don't have
+the right models.  Further unknonws such as measurement uncertainty,
+parameter uncertainty, unobserved states, knowledge of boundary
+conditions, etc. further compound the issue.  Though a hierarchical
+Bayesian approach provides a natural way to address these from a
+statistical standpoint, formulating reasonable parametric descriptions
+of each form of uncertainty is a challenging task in itself, let alone
+the computational difficulties of solving such a system. @Cressie2009
+provides a good account of the successess and challenges of the approach. 
+Applying these approaches in the management context of sequential decision
+making, in which forecasts must be obtained over a range of possible actions
+and updated regularly as new information arrives makes such an approach 
+less feasible still.
 
@@
 
-* Nonparametric approaches are beginning to appear more frequently
-in ecological and conservation literature (Species distribution
-models/maxent), including the Gaussian process based approach used here
-[@Munch2005].  However, such approaches have yet to be applied to the
+Though machine learning approaches have begun to appear in the
+ecological and conservation literature (Species distribution models),
+including the Gaussian process based approach used here [@Munch2005],
+they remain unfamiliar and untrusted approaches for most ecologists.
+Machine learning approaches represent an essentially pattern-based
+rather than process-based approach, raising the same skepticism from most
+theoretical ecologists that they hold for older correlative methods such
+as linear regression, while the complexity of the techniques has barred
+their adoption in more empirical audiences. <!-- Whoa! rampant speculation
+alert! -->
+
+.  However, such approaches have yet to be applied to the
 decision-theoretic framework that could guide management decisions,
@@
 The use of Gaussian process regression (known as Kreging in the geospatial literature) to formulate a predictive model is relatively new in the context of modeling dynamical systems [@Kocijan2005] and introduced in the ecological modeling and fisheries management by @Munch2005.
-
-_Is the following  discription useful or too elementary?_
+An accessible and thorough introduction to the formulation and use of Gaussian processes can be found in @Rasmussen2006.
 
@@
 By modeling at the level of the process, rather than the level of parametric equation, we can more concisely capture the possible behavior of these curves.
-In place of a parametric model of the dynamical system, the Gaussian Process approach postulates a prior distribution of (n-dimensional) curves that can be though of as approximations to a large range of possible
+In place of a parametric model of the dynamical system, the Gaussian Process approach postulates a prior distribution of (n-dimensional) curves that can be though of as approximations to a range of possible (parametric) models that might describe the data.
+The GP allows us to consider a set of possible curves simultaneously (Figure 1).
+<!-- Figure of curves drawn from the posterior density?
+-->
 
-_Do we need more specifics on Gaussian process as an approximation to parametric models?
+<!-- Do we need more specifics on Gaussian process as an approximation to parametric models?
 Discussion of Gaussian process vs other machine learning / forecasting approaches that have less clear statistical foundations?
-If so, does this all belong in the discussion?_
+If so, does this all belong in the discussion?
+-->
 
-The Gaussian process is defined by a covariance kernel ...
-_discussion of our (gaussian) kernel, blah blah how specific here vs appendix?_.
-By requiring our kernel to follow a generic functional form, we can compactly describe the Gaussian process using only a handful of parameters (Table 1)
-
-A more thorough introduction to the formulation and use of Gaussian processes can be found in @Rasmussen2006.
-
-parameter       interpretation
----------       -------------- 
-$\sigma^2$      The process noise (from the kernel)
-$\tau^2$        Variance around the mean
-$\beta$_0       The mean is given by a linear model of slope $\beta$ 
-$d$             The length-scale of the covariance function
-$n$             The observation error
-
-: Table of parameters for the Gaussian process
-
-Rather than estimate values of these parameters directly, we take a hierarchical approach of placing prior distributions on each.
-Following @Gramarcy2005 we use a Gaussian prior on $\beta_0$, an inverse gamma prior on $\sigma^2$ and $\tau^2$, a gamma prior on the observation noise $n$, and exponential prior on the length-scale $d$.
-
-* Gaussian processes in ecological literature [@Munch2005]
-
-* Our methodology (e.g. following @Munch2005).
-
-###  Discussion of the dynamic programming solution
-
-The fishery management problem over an infinite time horizon can be stated as:
-
-\begin{align}
-& \max_{ \{h_t\} \geq 0 } \mathbf{E} \lbrace \sum_0^\infty \delta^t \Pi(h_t) \rbrace \\
-& \mathrm{s.t.}  \\
- & X_t = Z_t f\left(S_{t-1}\right) \\
- & S_t = X_t - h_t \\
- & X_t  \geq 0 
-\end{align}
-
-Where $\mathbf{E}$ is the expectation operator, $\delta$ the discount
-rate, $\Pi(h_t)$ the profit expected from a harvest of $h_t$, and other
-terms as in Eq. (1).  For simplicity, we have assumed that profits depend
-only on the chosen harvest; simplifying further we will usually consider
-profits to be proportional to harvest, $\Pi(h_t) = h_t$.
-
-Once the posterior Gaussian process (GP) has been estimated [e.g. see
-@Munch2005], it is necessary to adapt it in place of the parametric
-equation for the stochastic dynamic programming (SDP) solution [see
-@Mangel1988 for a detailed description of parametric SDP methods] to the
-optimal policy. The essense of the idea is straight forward -- we will use
-the estimated GP in place of the parametric growth function to determine
-the stochastic transition matrix on which the SDP calculations are based.
-
-The posterior Gaussian process is completely defined by the expected value
-and covariance matrix at a defined set of training points.  For simplicty
-we will consider a these points to fall on a discrete, uniform grid $x$
-of `r length(x_grid)` points from `r min(x_grid)` to `r max(x_grid)`
-(1.5 times the positive equilibrium $K$).  Again to keep things simple
-we will use this same grid discritization for the parametric approach.
-Other options for choosing the grid points, including collocation methods
-and functional basis expansion (or even using Guassian processes in place
-of the discrete optimization; an entirely different context in which GP
-can be used in SDP, see [@Deisenroth2009]) could also be considered.
-
-The transition matrix $\mathbf{F}$ is thus an `r length(x_grid)`
-by `r length(x_grid)` matrix for which the ${i,j}$ entry gives the
-probability of transitioning into state $x_i$ given that the system is
-in state $x_j$ in the previous timestep.  To generate the transition
-matrix based on the posterior GP, we need only the expected values
-at each grid point and the corresponding variances (the diagonal of
-the covariance matrix), as shown in Figure 1.  Given the mean at each
-gridpoint as the length `r length(x_grid)` vector $E$ and variance $V$,
-the probability of transitioning from state $x_i$ to state $x_j$ is
-simply $\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)$,
-where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
-variance $\sigma^2$.  Strictly speaking, the transition probability should
-be calculated by integrating the normal density over the bin of width
-$\Delta$ centered at $x_j$.  For a sufficiently fine grid that $f(x_j)
-\approx f(x_j + \Delta)$, it is sufficient to calculate the density at
-$x_j$ and then row-normalize the transition matrix.
-
-## Pseudocode for the determining the transtion matrix from the GP
-
-```r
-for(h in h_grid)
-  F_h = for(x_j in grid)
-          for(i in 1:N) 
-            dnorm(x_j, mu[i]-h, V[i])
-```
-
-A transition matrix for each of the parametric models $f$ is calculated
-using the log-normal density with mean $f(x_i)$ and log-variance as
-estimated by maximum likelihood.  From the discrete transition matrix we
-may write down the Bellman recursion defining the the stochastic dynamic
-programming iteration:
-
-\begin{equation}
-V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
-\end{equation}
-
-where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$
-is control (harvest level) chosen. Numerically, the maximization is
-accomplished as follows. Consider the set of possible control values to
-be the discrete `r length(h_grid)` values corresponding the the grid of
-stock sizes.  Then for each $h_t$ there is a corresponding transition
-matrix $\mathbf{F}_h$ determined as described above but with mean 
-$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element corresponds
-to the value of having stock $x_i$ at time $t$.  Then let $\Pi_h$ be
-the vector whose $i$th element indicates the profit from harvesting
-at intensity $h_t$ given a population $x_i$ (e.g. $\max(x_i, h_t)$
-since one cannot harvest more fish then the current population size).
-Then the Bellman recursion can be given in matrix form as
-
-$$V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)$$
-
-where the sum is element by element and the expectation is computed by the matrix multiplication $\mathbf{F} V_{t+1}$.
-
-### Pseudocode for the Bellman iteration
-
-```r
- V1 <- sapply(1:length(h_grid), function(h){
-      delta * F[[h]] %*% V +  profit(x_grid, h_grid[h]) 
-    })
-    # find havest, h that gives the maximum value
-    out <- sapply(1:gridsize, function(j){
-      value <- max(V1[j,], na.rm = T) # each col is a diff h, max over these
-      index <- which.max(V1[j,])  # store index so we can recover h's 
-      c(value, index) # returns both profit value & index of optimal h.  
-    })
-    # Sets V[t+1] = max_h V[t] at each possible state value, x
-    V <- out[1,]                        # The new value-to-go
-    D[,OptTime-time+1] <- out[2,]       # The index positions
-```
-
-_Currently this shows the literal R code, should be adapted_
+### The optimal control problem
 
@@
 
+### The Guassian process
+
+The Gaussian process is defined by a covariance kernel.
+By requiring our kernel to follow a generic functional form, we can compactly describe the Gaussian process using only a handful of parameters (Table 1)
+
+parameter       interpretation
+---------       -------------- 
+$\sigma^2$      The process noise (from the kernel)
+$\tau^2$        Variance around the mean
+$\beta$_0       The mean is given by a linear model of slope $\beta$ 
+$d$             The length-scale of the covariance function
+$n$             The observation error
+
+: Table of parameters for the Gaussian process
+
+Rather than estimate values of these parameters directly, we take a hierarchical approach of placing prior distributions on each.
+Following @Gramarcy2005 we use a Gaussian prior on $\beta_0$, an inverse gamma prior on $\sigma^2$ and $\tau^2$, a gamma prior on the observation noise $n$, and exponential prior on the length-scale $d$.
+
+###  Formulating a dynamic programming solution
+
+The fishery management problem over an infinite time horizon can be stated as:
+
+\begin{align}
+& \max_{ \{h_t\} \geq 0 } \mathbf{E} \lbrace \sum_0^\infty \delta^t \Pi(h_t) \rbrace \\
+& \mathrm{s.t.}  \\
+ & X_t = Z_t f\left(S_{t-1}\right) \\
+ & S_t = X_t - h_t \\
+ & X_t  \geq 0 
+\end{align}
+
+Where $\mathbf{E}$ is the expectation operator, $\delta$ the discount
+rate, $\Pi(h_t)$ the profit expected from a harvest of $h_t$, and other
+terms as in Eq. (1).  For simplicity, we have assumed that profits depend
+only on the chosen harvest; simplifying further we will usually consider
+profits to be proportional to harvest, $\Pi(h_t) = h_t$.
+
+Once the posterior Gaussian process (GP) has been estimated [e.g. see
+@Munch2005], it is necessary to adapt it in place of the parametric
+equation for the stochastic dynamic programming (SDP) solution [see
+@Mangel1988 for a detailed description of parametric SDP methods] to the
+optimal policy. The essense of the idea is straight forward -- we will use
+the estimated GP in place of the parametric growth function to determine
+the stochastic transition matrix on which the SDP calculations are based.
+
+The posterior Gaussian process is completely defined by the expected value
+and covariance matrix at a defined set of training points.  For simplicty
+we will consider a these points to fall on a discrete, uniform grid $x$
+of `r length(x_grid)` points from `r min(x_grid)` to `r max(x_grid)`
+(1.5 times the positive equilibrium $K$).  Again to keep things simple
+we will use this same grid discritization for the parametric approach.
+Other options for choosing the grid points, including collocation methods
+and functional basis expansion (or even using Guassian processes in place
+of the discrete optimization; an entirely different context in which GP
+can be used in SDP, see [@Deisenroth2009]) could also be considered.
+
+The transition matrix $\mathbf{F}$ is thus an `r length(x_grid)`
+by `r length(x_grid)` matrix for which the ${i,j}$ entry gives the
+probability of transitioning into state $x_i$ given that the system is
+in state $x_j$ in the previous timestep.  To generate the transition
+matrix based on the posterior GP, we need only the expected values
+at each grid point and the corresponding variances (the diagonal of
+the covariance matrix), as shown in Figure 1.  Given the mean at each
+gridpoint as the length `r length(x_grid)` vector $E$ and variance $V$,
+the probability of transitioning from state $x_i$ to state $x_j$ is
+simply $\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)$,
+where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
+variance $\sigma^2$.  Strictly speaking, the transition probability should
+be calculated by integrating the normal density over the bin of width
+$\Delta$ centered at $x_j$.  For a sufficiently fine grid that $f(x_j)
+\approx f(x_j + \Delta)$, it is sufficient to calculate the density at
+$x_j$ and then row-normalize the transition matrix.
+
+## Pseudocode for the determining the transtion matrix from the GP
+
+```r
+for(h in h_grid)
+  F_h = for(x_j in grid)
+          for(i in 1:N) 
+            dnorm(x_j, mu[i]-h, V[i])
+```
+
+A transition matrix for each of the parametric models $f$ is calculated
+using the log-normal density with mean $f(x_i)$ and log-variance as
+estimated by maximum likelihood.  From the discrete transition matrix we
+may write down the Bellman recursion defining the the stochastic dynamic
+programming iteration:
+
+\begin{equation}
+V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
+\end{equation}
+
+where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$
+is control (harvest level) chosen. Numerically, the maximization is
+accomplished as follows. Consider the set of possible control values to
+be the discrete `r length(h_grid)` values corresponding the the grid of
+stock sizes.  Then for each $h_t$ there is a corresponding transition
+matrix $\mathbf{F}_h$ determined as described above but with mean 
+$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element corresponds
+to the value of having stock $x_i$ at time $t$.  Then let $\Pi_h$ be
+the vector whose $i$th element indicates the profit from harvesting
+at intensity $h_t$ given a population $x_i$ (e.g. $\max(x_i, h_t)$
+since one cannot harvest more fish then the current population size).
+Then the Bellman recursion can be given in matrix form as
+
+$$V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)$$
+
+where the sum is element by element and the expectation is computed by the matrix multiplication $\mathbf{F} V_{t+1}$.
+
+### Pseudocode for the Bellman iteration
+
+```r
+ V1 <- sapply(1:length(h_grid), function(h){
+      delta * F[[h]] %*% V +  profit(x_grid, h_grid[h]) 
+    })
+    # find havest, h that gives the maximum value
+    out <- sapply(1:gridsize, function(j){
+      value <- max(V1[j,], na.rm = T) # each col is a diff h, max over these
+      index <- which.max(V1[j,])  # store index so we can recover h's 
+      c(value, index) # returns both profit value & index of optimal h.  
+    })
+    # Sets V[t+1] = max_h V[t] at each possible state value, x
+    V <- out[1,]                        # The new value-to-go
+    D[,OptTime-time+1] <- out[2,]       # The index positions
+```
+
+_Currently this shows the literal R code, should be adapted_
+
 MCMC posterior distributions and convergence analysis

### 2013-03-21 inst/examples/pomp-example.Rmd
pomp example

```{r}
library(pomp)
```

```{r}
process.sim <- function(x, t, params, delta.t, ...){
# r = 0.75, K = 10, Q = 3, H = 1, a = 1.3 
  r <- params["r"]
  K <- params["K"]
  Q <- params["Q"]
  H <- params["H"]
  a <- params["a"]
  sigma <- params["sigma"]
  X <- x["X"]
  epsilon <- exp(rlnorm(n = 1, mean = 0, sd = sigma))

  c(X = unname(
    X * exp(r * delta.t * (1 - X / K) - a * X ^ (Q - 1) / (X ^ Q + H ^ Q)) * epsilon)) 
}
```

```{r}
measure.sim <- function(x, t, params, ...){
  tau <- params["tau"]
  X <- x["X"]
  y <- c(Y = unname(rlnorm(n = 1, meanlog = log(X), sd=tau)))
}
```

Create a container of class `pomp`

```{r}
may <- pomp(data = data.frame(time = 1:100, Y = NA), times = "time", rprocess = discrete.time.sim(step.fun = process.sim, delta.t = 1), rmeasure = measure.sim, t0 = 0)
```

Parameters and an initial condition

```{r}
theta <- c(r = 0.75, K = 10, Q = 3, H = 1, a = 1.3, 
           sigma = 0.1, tau = 0.1, X.0 = 5) 
```
```

Simulate

```{r}
dat <- simulate(may, params=theta, obs=TRUE, states=TRUE)
library(reshape2)
library(ggplot2)
dat <- melt(dat)[3:5]
names(dat) = c("time", "value", "variable")
ggplot(dat, aes(time, value, lty=variable)) + geom_line()
```

Note that simulate returns an instance of our object, with the data column filled in.  This is not true if we give it arguments such as state or obs = TRUE.

Measurement density

```{r}
measure.dens <- function(y, x, t, params, log, ...) {
    tau <- params["tau"]
    ## state at time t:
    X <- x["X"]
    ## observation at time t:
    Y <- y["Y"]
    ## compute the likelihood of Y|X,tau
    dlnorm(x = Y, meanlog = log(X), sdlog = tau, log = log)
}
```

Add to our container

```{r}
may <- pomp(may, dmeasure = measure.dens)
```

Optionally, work in the transformed variable space instead by providing parameter and inverse parameter transforms.

```{r}
may  <- pomp(may, parameter.transform = function(params, 
    ...) {
    exp(params)
}, parameter.inv.transform = function(params, ...) {
    log(params)
})
```

Run the particle filter

```{r}
pf <- pfilter(may, params = theta, Np = 1000)
logLik(pf)
```

```{r}

```

### 2013-03-21 inst/examples/pomp-example.Rmd
pomp example
need to combine versions agressively here...

@@
 library(pomp)
+opts_knit$set(upload.fun = socialR::flickr.url)
+options(flickr_tags = "nonparametric-bayes")
 ```
@@
 theta <- c(r = 0.75, K = 10, Q = 3, H = 1, a = 1.3, 
-           sigma = 0.1, tau = 0.1, X.0 = 5) 
+           sigma = 0.2, tau = 0.1, X.0 = 5) 
 ```
@@
 library(ggplot2)
-dat <- melt(dat)[3:5]
-names(dat) = c("time", "value", "variable")
-ggplot(dat, aes(time, value, lty=variable)) + geom_line()
+dt <- melt(dat)[3:5]
+names(dt) = c("time", "value", "variable")
+ggplot(dt, aes(time, value, lty=variable)) + geom_line()
 ```
 
-Note that simulate returns an instance of our object, with the data column filled in.  This is not true if we give it arguments such as state or obs = TRUE.
+Note that simulate returns an instance of our object, with the data column filled in.
+This is not true if we give it arguments such as state or obs = TRUE as above.
+However, when we query for observed and true states in this way, the simulated data is not stored in our object.
+We must assign it with `may <- pomp(data = ...` where data is a data frame with columns "time" and "Y".
+This is much simpler if we use the default return:
+
+```{r}
+may <- simulate(may, params=theta)
+```
 
@@
 ```{r}
-measure.dens <- function(y, x, t, params, log, ...) {
+measure.density <- function(y, x, t, params, log, ...) {
     tau <- params["tau"]
@@
     ## compute the likelihood of Y|X,tau
-    dlnorm(x = Y, meanlog = log(X), sdlog = tau, log = log)
+    f <- dlnorm(x = Y, meanlog = log(X), sdlog = tau, log = log)
+
 }
@@
 ```{r}
-may <- pomp(may, dmeasure = measure.dens)
+may <- pomp(may, dmeasure = measure.density)
 ```
@@
 
-Run the particle filter
+Run the particle filter.  Note that this evaluates the likelihood of our simulated data (stored in the object) at the given (true) parameters.
 
@@
 
+Here we go, estimate the model by particle filter.  This'll be slow...
+
 ```{r}
+theta.guess <- theta + abs(rnorm(length(theta)))
+estpars <- names(theta) # which pars should be estimated. lets try and get them all
+rw.sd <- theta/(50*theta) # a named vector of random walk values. This sets it to 0.02 for all pars
+system.time(mf <- 
+  mif(may, Nmif = 100, # Number of iterations
+      start = theta.guess, # Starting value of parameters
+      transform = TRUE, # work in transformed parameter space (e.g. log to avoid negative par values)
+      pars = estpars,  # which parameters to estimate (others fixed at starting value)
+      rw.sd = rw.sd, # random walk width
+      Np = 2000, # Number of particles
+      var.factor = 4, # scaling factor for random walk sd
+      ic.lag = 10, # fixed lag
+      cooling.factor = 0.999, # exponential cooling
+      max.fail = 10) # internal tolerance
+)
+```
 
+```{r}
+compare.mif(mf)
 ```
+
+```{r}
+coef(mf)
+logLik(mf)
+pf_est <- pfilter(may, params = theta.guess, Np = 1000)
+logLik(pf_est) # initial loglik
+```
+
+These functions apply to replicated mifs above.
+
+```{r}
+theta.true <- coef(may)
+theta.mif <- apply(sapply(mf,coef),1,mean)
+loglik.mif <- replicate(n=10,logLik(pfilter(mf[[1]],params=theta.mif,Np=10000)))
+bl <- mean(loglik.mif)
+loglik.mif.est <- bl+log(mean(exp(loglik.mif-bl)))
+loglik.mif.se <- sd(exp(loglik.mif-bl))/exp(loglik.mif.est-bl)
+```

### 2013-03-21 inst/examples/pomp-example.Rmd
pomp example updated with iterated filtering

@@
            sigma = 0.2, tau = 0.1, X.0 = 5) 
-```
 ```
@@
 ```{r}
-theta.guess <- theta + abs(rnorm(length(theta)))
-estpars <- names(theta) # which pars should be estimated. lets try and get them all
-rw.sd <- theta/(50*theta) # a named vector of random walk values. This sets it to 0.02 for all pars
+theta.guess <- theta + c(abs(rnorm(length(theta)-1)),0)
+estpars <- names(theta[1:7]) # which pars should be estimated. lets try and get them all
+rw.sd <- theta[1:7]/(50*theta[1:7]) # a named vector of random walk values. This sets it to 0.02 for all pars
 system.time(mf <- 
-  mif(may, Nmif = 100, # Number of iterations
+  mif(may, Nmif = 10, # Number of iterations
       start = theta.guess, # Starting value of parameters
-      transform = TRUE, # work in transformed parameter space (e.g. log to avoid negative par values)
+      transform = TRUE, # work in transformed parameter space (e.g. logs avoid neg par values)
       pars = estpars,  # which parameters to estimate (others fixed at starting value)
       rw.sd = rw.sd, # random walk width
-      Np = 2000, # Number of particles
+      Np = 200, # Number of particles
       var.factor = 4, # scaling factor for random walk sd

### 2013-03-21 inst/doc/nonparametric-bayes.Rmd
writing writing

@@
 
-1. Classical approaches are to fix relationship of stock recruitment.  
-2. Parameter uncertainty
-3. Gaussian processes for model uncertainty
-
-The problem of structural uncertainty in managing ecological systems.
-
-(Big literature here of course, primary approaches have been using
-parametric uncertainty and model uncertainty (e.g. belief distribution over models 
-I believe. Worth distinguishing our problem from the more trivially 
-sized problems in which the action space is discrete and consists
-of 2 or three possible choices, e.g. much of the recent Possingham et al work.)
-
 Most management recommendations from the ecological literature are based on
@@
 
+<!-- Better defintion of focal problems here .... -->
 Stochastic dynamic programming @Mangel1982
@@
 
-Background on the concerns of structural uncertainty -- we don't have
-the right models.  Further unknonws such as measurement uncertainty,
-parameter uncertainty, unobserved states, knowledge of boundary
-conditions, etc. further compound the issue.  Though a hierarchical
-Bayesian approach provides a natural way to address these from a
-statistical standpoint, formulating reasonable parametric descriptions
-of each form of uncertainty is a challenging task in itself, let alone
-the computational difficulties of solving such a system. @Cressie2009
-provides a good account of the successess and challenges of the approach. 
-Applying these approaches in the management context of sequential decision
-making, in which forecasts must be obtained over a range of possible actions
-and updated regularly as new information arrives makes such an approach 
-less feasible still.
+Structural uncertainty.
 
-* Management goals / decision-theoretic approaches need accurate
-prediction over relevant (short?) timescales more than accurate (but
-incomplete or noisy estimated) mechanisms.
-
-* Nonparametric (machine-learning?) approaches may offer the benefit of
-the hierarchical Bayesian approach without the practical and computational
-limitations of their parametric kin.  Non-parametric models are flexible
-enough to take maximum advantage of the data available, while being
-appropriately ambiguous about the dynamics of a system in regions of
-parameter space that have been poorly or never sampled.
+Further unknowns such as measurement uncertainty, parameter uncertainty,
+unobserved states, knowledge of boundary conditions, etc. further
+compound the issue.  Though a hierarchical Bayesian approach provides a
+natural way to address these from a statistical standpoint, formulating
+reasonable parametric descriptions of each form of uncertainty is a
+challenging task in itself, let alone the computational difficulties
+of solving such a system. @Cressie2009 provides a good account of the
+successess and challenges of the approach.  Applying these approaches in
+the management context of sequential decision making, in which forecasts
+must be obtained over a range of possible actions and updated regularly
+as new information arrives makes such an approach less feasible still.
 
@@
 
-.  However, such approaches have yet to be applied to the
-decision-theoretic framework that could guide management decisions,
-where we expect them to excel for several reasons (1) the ability to
-make accurate forecasts by more closely approximating the underlying
-process where data is available (2) remaining appropriately ambiguous
-where data is not available (3) remaining computationally simple enough
-to avoid some pitfalls common to hierarchical parametric approaches.
+<!-- Whoa, isn't machine learning used to make decisions all the time,
+ eg, in engineering and Artificial Intelligence?  -->
+
+A more immediate barrier to their adoption is the absence of a framework
+for applying machine learning approaches to resource management problems.
+Traditional approaches to optimal control (Pontryagin's principle, stochastic
+dynamic programming) rely on knowledge of the state equation, usually described
+by a simple parametric model. Here we illustrate how a stochastic dynamic
+programming algorithm can alternatively be driven by the predictions from
+a Gaussian process -- a machine learning approximation to the state dynamics.  
+<!-- Okay, perhaps that's novel, but it's pretty trivial.  Isn't it
+obvious to everyon that it's trivial? -->
+
+<!-- paraphrase advantages of machine learning -->
+
+Management goals / decision-theoretic approaches need accurate prediction
+over relevant (short?) timescales more than reasonable long-term
+mechanisms.  Machine-learning approaches may offer the benefit of the
+hierarchical Bayesian approach without the practical and computational
+limitations of their parametric kin.  Non-parametric models are flexible
+enough to take maximum advantage of the data available, while being
+appropriately ambiguous about the dynamics of a system in regions of
+parameter space that have been poorly or never sampled.
 
@@
 
-The use of Gaussian process regression (known as Kreging in the geospatial literature) to formulate a predictive model is relatively new in the context of modeling dynamical systems [@Kocijan2005] and introduced in the ecological modeling and fisheries management by @Munch2005.
-An accessible and thorough introduction to the formulation and use of Gaussian processes can be found in @Rasmussen2006.
+The use of Gaussian process regression (known as Kreging in the geospatial
+literature) to formulate a predictive model is relatively new in the
+context of modeling dynamical systems [@Kocijan2005] and introduced
+in the ecological modeling and fisheries management by @Munch2005.
+An accessible and thorough introduction to the formulation and use of
+Gaussian processes can be found in @Rasmussen2006.
 
-The essense of the Gaussian process approach can be captured in the following thought experiment: An exhaustive parametric approach to the challenge of structural uncertainty might proceed by writing down all possible functional forms for the underlying dynamical system with all possible parameter values for each form, and then consider searching over this huge space to select the most likely model and parameters; or using a Bayesian approach, assign priors to each of these possible models and infer the posterior distribution of possible models.
-The Gaussian process approach can be thought of as a computationally efficient approximation to this approach.
-Gaussian processes represent a large class of models that can be though of as capturing or reasonably approximating the set of models in this collection.
-By modeling at the level of the process, rather than the level of parametric equation, we can more concisely capture the possible behavior of these curves.
-In place of a parametric model of the dynamical system, the Gaussian Process approach postulates a prior distribution of (n-dimensional) curves that can be though of as approximations to a range of possible (parametric) models that might describe the data.
-The GP allows us to consider a set of possible curves simultaneously (Figure 1).
-<!-- Figure of curves drawn from the posterior density?
--->
+The essense of the Gaussian process approach can be captured in the
+following thought experiment: An exhaustive parametric approach to the
+challenge of structural uncertainty might proceed by writing down all
+possible functional forms for the underlying dynamical system with all
+possible parameter values for each form, and then consider searching
+over this huge space to select the most likely model and parameters;
+or using a Bayesian approach, assign priors to each of these possible
+models and infer the posterior distribution of possible models. The
+Gaussian process approach can be thought of as a computationally
+efficient approximation to this approach. Gaussian processes represent
+a large class of models that can be though of as capturing or reasonably
+approximating the set of models in this collection.  By modeling at the
+level of the process, rather than the level of parametric equation,
+we can more concisely capture the possible behavior of these curves.
+In place of a parametric model of the dynamical system, the Gaussian
+Process approach postulates a prior distribution of (n-dimensional)
+curves that can be though of as approximations to a range of possible
+(parametric) models that might describe the data. The GP allows us
+to consider a set of possible curves simultaneously.  
+<!-- Figure 1 include curves drawn from the posterior density?  -->
 
-<!-- Do we need more specifics on Gaussian process as an approximation to parametric models?
-Discussion of Gaussian process vs other machine learning / forecasting approaches that have less clear statistical foundations?
-If so, does this all belong in the discussion?
--->
+<!-- Do we need more specifics on Gaussian process as an approximation
+to parametric models? Discussion of Gaussian process vs other machine
+learning / forecasting approaches that have less clear statistical
+foundations?  If so, does this all belong in the discussion? -->
 
@@
 
-* Non-parametric methods have received far too little attention in
+Non-parametric methods have received far too little attention in
 ecological modeling efforts that are aimed at improved conservation
 planning and decision making support.
-
-* Importance of non-parametric approaches in conservation planning /
-resource management / decision theory.
-
-* Decision-theoretic tools such as optimal control calculations rely
-on robust _forecasting_ more strongly than they rely on accurate
-_mechanistic_ relationships.
 
@@
 
-* Anticipate improved relative performance in higher dimensional examples
+Three elements contribute to the machine learning model performing as well
+or better than the mechanisitic models in this context.
 
+### 1. Relevant state space
+
+The dynamics occur over a range of state-space in which the machine learning
+model closely matches the predictions of y
+implausible model performs as well or better than the more plausible
+model.
+
+This is the most obvious and immediate reason why the shortcomings
+that appear to make the model implausible do not make it useless.
+This effect is enhanced in our example because the management actions help
+drive the system towards rather than away from this region of state-space.
+
+### 2. Predictive accuracy
+
+This application relies only on the predictive accuracy of the model,
+not an interpretation of the parameters.
+
+<!-- wow, run-on tangent? -->
+Predictive accuracy is not the goal of all modeling, as ecologists
+have been observing for as long as they made models (perhaps none
+more memorably than @Levins1969).  Mechanistic modeling is at its
+most powerful not when it is used to directly forecast future states
+but when it provides an understanding of how to approach a problem.
+SIR-type models from the epidemiological literature are a good example.
+While the simplest SIR models have little predictive power over the
+outbreak intensity or timing at a particular location, they provide a
+powerful understanding of the spread of an infection in terms of a single,
+biologically meaningful parameter: $R_0$, the basic reproductive number.
+From the model, it becomes clear that management need not vaccinate every
+member of the population to stop the spread, but rather it suffices to
+vaccinate a sufficient fraction of the population to reduce $R_0$ below 1.
+
+### 3. Time scale for new data
+
+In the sequential decision making problem we considered, we are
+presented with new data after each action.  The relevant timescale
+for the prediction is thus not the long-term dynamics, which would be
+wildly divergent, but the dynamics over this much shorter interval.
+While ecologists may be hesitant to base continual management on a model
+with obviously inaccurate long-term behavior, engineers tend to consider
+the problem in frequency space and gravitate to the opposite position --
+a good control model should prioritize high-frequency accuracy over low
+frequency accuracy.  The differences in intuition may arise from the
+timescales at which each profession can typically adjust their control
+variables -- much faster for a control system of a chemical plant than
+state policy for a natural resource.  Still, the lesson is clear: when
+facing repeated management decisions over a short timescale, such as
+setting annual harvests of a fishery, it may be more valuable to use a
+machine learning algorithm that makes accurate year-ahead predictions
+that capture some of the high-frequency fluctuations that appear only
+as noise in a mechanistic model of the long-term population dynamics.
+
+2. Better express ambiguity in regions where data is not available
+
+One of the greatest strengths of mechanistic models is their greatest weakness as well.
+
+## Future directions
+
+<!-- Jargony and vague -->
+While we have highlighted certain generic of this problem that allow the
+nonparametric approach to excel -- short timescales between successive
+observations and actions, the accuracy in the appropriate region of state
+space, the ability to express uncertainty outside the observed data --
+there are equally several aspects for which the nonparametric approach is
+at a relative disadvantage.
+
+<!-- Model complexity -->
+In this simulated example, the underlying
+dynamics are truly governed by a simple parametric model, allowing
+the parametric approaches to be more accurate.  Similarly, because the
+dynamics are  one-dimensional dynamics and lead to  stable nodes (rather
+than other attractors such as limit-cycles resulting in oscillations),
+the training data provides relatively limited information about the
+dynamics.  For these reasons, we anticipate that in higher-dimensional
+examples characteristic of ecosystem management problems that the machine
+learning approach will prove even more valuable.
+
+<!-- Data complexity. Perhaps too far out of scope... -->
+The machine learning approach is also better suited to complex and
+disparate data.  Incorporating various sources of information into
+mechanistic models can be an immensely difficult due to the increased
+complexity involved.  Only thanks to tandem advances in increasing
+computational power and hierarchical statistical methodology have we
+been able to tackle such intuitively important complexity (and the
+potentially new available data that accompanies it) such as spatial
+distribution, heterogeneities of space, time, and individuals, to shift
+to ecosystem-based approaches from single-species based approaches.
+Without the need to formulate mechanisms, many modern machine learning
+algorithms can leverage potential information from all available sources
+of data directly.  The algorithms can recognize unanticipated or subtle
+patterns in large data sets that enable more accurate predictions than
+mechanistic models that are formulated at a more macroscopic level.
+
+<!-- 
 * Discuss constant escapement in model, in policies.
@@
 
-* Discussion of alternative related approaches: POMDP/MOMDP,
+* Discussion of alternative related approaches: POMDP/MOMDP
+-->
 
-Future directions
 --------------------------------------------------------------------

### 2013-03-21 inst/doc/nonparametric-bayes.Rmd
long way to go

@@
 
-Most management recommendations from the ecological literature are based on
-(or at least motivated by) parametric models. Though in principle these models
-reflect mechanistic underpinnings of biological interactions involved, in practice
-real mechanisms are often unknown and the true dynamics too complex to be captured
-by simple models [@refs]. While simple mechanistic models can nevertheless
-provide imporant insights into possible dynamics -- for instance, demonstrating
-that a vaccine does not have to be 100% effective to eliminate the transmission of a virus [@Kermack1921] -- such approaches are not well-suited for use in 
-forecasting future dynamics upon which outcomes management policy can be based.  
-Despite a long history of work emphasizing the difference between modeling 
-for understanding (generic or strategic models) and modeling for prediction (precise, specific, or tactical models) [e.g. @Levins1966], simple parametric models 
-continue to find application in predictive management contexts.  The management
-of marine fisheries is a prime example, in which simple, mechanistically 
-motivated models of stock-recruitment relationships such as Ricker or Beverton-Holt
-curves are systematically fit to data and used in solving decision-theoretic
-or optimal control problems determining resource management policies such
-as maximum sustainable yield [@refs].
+Process-based models built on modern machine learning approaches 
+offer a powerful and underappreciated approach for ecological
+forecasting and management.  We present a novel example of how
+one such example - non-parametric Gaussian processes, can be adapted
+to the context of management and decision-making.
+
+While effective machine learning algorithms have speckled the literature, 
+most theoretical ecologists in particular have been wary of these 'black-box'
+approaches in favor of simple, mechanistic models.  While a blind application
+of machine learning approaches would indeed be dangerous grounds to stake 
+either ecological understanding or management practice, such concerns are not
+grounds to reject such methods out of hand.  Rather, further attention is 
+warranted to identify the circumstances in which machine-learning approaches
+will most likely outperform existing more mechanistic approaches, or help
+crack difficult management questions where ecological or dynamical complexity
+has caused traditional approaches to founder.
+
+Here we illustrate how a machine learning model can be implemented in place
+of simple mechanistic model in a decision theory (optimal control) problem,
+and highlight the features of this system that make the nonparametric approach
+particularly effective.
+
+<!-- Um, Intro take 2, here we go... -->
+
+Mechanistic models have long been the gold standard of theoretical
+modeling in ecology (e.g. see @Geritz2012 or @Cuddington2013).  Only by
+understanding the processes involved can we make reliable long term
+predictions and build an knowledge of cause and effect that guides the
+hypotheses we make, the data we collect, and the management decisions
+we make.  Process-based models, whether expressed in the language of
+mathematics or English, identify the connection between mosquitoes and
+the spread of malaria, or greenhouses gases and climate change, guiding
+our approach to understand and manage these threats. Despite this central
+importance, we argue here that ecologists would do well to give greater
+attention to the role non-mechanistic models can play in ecological
+management and decision making.  The value of these approaches is
+greatest in a context where decisions are made over short time horizons,
+and updated as new data becomes available.
+
+## The alternative to process-based models: pattern-based modeling
+
+Before we proceed further it would be useful to define our terms and lay
+some greater context for the discussion. Throughout, we will distinguish
+"mechanistic" or "process-based" models from models that are merely
+"correlative" or "pattern-based". Historically pattern-based modeling
+meant regression (usually linear regression) and resided in methodology
+rooted in statistics departments and was the primary focus of empirical
+ecologists, while mechanistic modeling meant dynamical systems (ODEs,
+PDEs, SDEs and their discrete kin), residing in methodology from
+the mathematics department and was the primary focus of theoretical
+ecologists.
+
+We believe these historical divides continue to color much of the
+literature today, with theorists more skeptical than empiricists of
+statistical methodology and vice versa.  Meanwhile, the ground underneath
+has shifted.  Originally pattern-matching approaches could be critiqued
+on the grounds of their simplicity (not everything is linear) while
+mechanistic approaches could be critiqued on their tenuous connections
+to data -- model parameter values such as death rates, birth rates,
+etc. would be estimated in advance and then stuck into the model, rather
+than estimated in the context of the model itself.  Today both of these
+critiques are outdated. Computational power and hierarchical statistical
+methods (particularly approximate likelihood or simulation techniques such
+as particle filters and approximate Bayesian computing) have brought the
+inference richer dynamical systems into their fold, while pattern-based
+modeling has spawned an entirely new approach under the banner of machine
+learning that has divided the statistics community (See @Bremman2001,
+"The Two Cultures").   Machine learning models can represent almost
+arbitrarily complex patterns and incorporate learning and decision making
+strategies that have earned them the name 'artificial intelligence.'
+Yet at the same time they can be far less transparent their linear
+regression predecessors, making both astounding successes and startling
+biases and failures like their biological namesake. More ecologists,
+particularly theorists and modelers, would do well to learn both how to
+take advantage of these strengths and recognize their weaknesses.
+
+Mechanistic modeling emphasizes the importance of capturing the correct
+gross properties of a system over tracking minute fluctuations.
+For instance, in selecting and parameterizing model of a population
+of conservation concern, we may be most interested in getting the
+long-term behavior correct -- such as identifying if the dynamics support
+persistence of the population -- rather than worrying how well they
+reflect the year-to-year fluctuations.  We would certainly have good
+reason to prefer such a model over alternatives which are irreconcilable
+to the most basic biological processes, such as unbounded growth, or
+growth curves that do not pass through the origin in a closed system. So
+it may come as a surprise to realize that such obviously wrong models
+can perform as well or even better than reasonable mechanistic models
+in guiding ecological management and decision making.
+
+<!--- Um, Intro take 3? here we go again... -->
+
+Most management recommendations from the ecological literature are based
+on (or at least motivated by) parametric models. Though in principle
+these models reflect mechanistic underpinnings of biological interactions
+involved, in practice real mechanisms are often unknown and the true
+dynamics too complex to be captured by simple models [@refs]. While
+simple mechanistic models can nevertheless provide imporant insights
+into possible dynamics -- for instance, demonstrating that a vaccine
+does not have to be 100% effective to eliminate the transmission of a
+virus [@Kermack1921] -- such approaches are not well-suited for use in
+forecasting future dynamics upon which outcomes management policy can
+be based.  Despite a long history of work emphasizing the difference
+between modeling for understanding (generic or strategic models)
+and modeling for prediction (precise, specific, or tactical models)
+[e.g. @Levins1966], simple parametric models continue to find application
+in predictive management contexts.  The management of marine fisheries
+is a prime example, in which simple, mechanistically motivated models of
+stock-recruitment relationships such as Ricker or Beverton-Holt curves
+are systematically fit to data and used in solving decision-theoretic
+or optimal control problems determining resource management policies
+such as maximum sustainable yield [@refs].
 
@@
 alternative models simultaneously (belief SDP)
-
-Structural uncertainty.
 
@@
 
-* Multiple species * Online learning * Multiple step-ahead predictions *
-Explicitly accomidating additional uncertainties * Improving inference
-of optimal policy from the GP
+* Multiple species 
+* Online learning 
+* Multiple step-ahead predictions 
+* Explicitly accomidating additional uncertainties 
+* Improving inference of optimal policy from the GP
 
-```{r echo=FALSE, results="asis"} #bibliography("html") ```
+```{r echo=FALSE, results="asis"}
+#bibliography()
+```
 

### 2013-03-21 inst/doc/nonparametric-bayes.tex
long way to go

@@
 
-The problem of structural uncertainty in managing ecological systems.
+Process-based models built on modern machine learning approaches offer a
+powerful and underappreciated approach for ecological forecasting and
+management. We present a novel example of how one such example -
+non-parametric Gaussian processes, can be adapted to the context of
+management and decision-making.
 
-(Big literature here of course, primary approaches have been using
-parametric uncertainty and model uncertainty (e.g.~belief distribution
-over models I believe. Worth distinguishing our problem from the more
-trivially sized problems in which the action space is discrete and
-consists of 2 or three possible choices, e.g.~much of the recent
-Possingham et al work.)
+While effective machine learning algorithms have speckled the
+literature, most theoretical ecologists in particular have been wary of
+these `black-box' approaches in favor of simple, mechanistic models.
+While a blind application of machine learning approaches would indeed be
+dangerous grounds to stake either ecological understanding or management
+practice, such concerns are not grounds to reject such methods out of
+hand. Rather, further attention is warranted to identify the
+circumstances in which machine-learning approaches will most likely
+outperform existing more mechanistic approaches, or help crack difficult
+management questions where ecological or dynamical complexity has caused
+traditional approaches to founder.
 
-\begin{itemize}
-\item
-  Most management recommendations from the ecological literature /
-  management policies based on (motivated by) parametric models.
-  \emph{Preference to pitch towards policy or theoretical literature?}
-\item
-  Background discussion on the importance/success of parametric
-  modeling; (e.g. Levins (1966), etc., up through Geritz and Kisdi
-  (2011))
-\item
-  Background on the concerns of structural uncertainty -- we don't have
-  the right models. (Also: measurement uncertainty, parameter
-  uncertainty, unobserved states, boundary conditions, etc.)
-\item
-  Hierarchical Bayesian approach has provided a natural way to address
-  these from a statistical standpoint. Successes and challenges from the
-  parametric route: e.g. Cressie et al. (2009).
-\item
-  Among pitfalls of these approaches: particularly difficult to apply in
-  management context. Optimization-based (Decision-theoretic) approaches
-  to ecological management need to be precisely parameterized about
-  everything, even the uncertainty itself (Polasky et al. 2011)
-  (discussion of Brozović and Schlenker (2011) SDP example in context of
-  threshold system would perhaps be useful as well).
-\item
+Here we illustrate how a machine learning model can be implemented in
+place of simple mechanistic model in a decision theory (optimal control)
+problem, and highlight the features of this system that make the
+nonparametric approach particularly effective.
+
+Mechanistic models have long been the gold standard of theoretical
+modeling in ecology (e.g.~see @Geritz2012 or @Cuddington2013). Only by
+understanding the processes involved can we make reliable long term
+predictions and build an knowledge of cause and effect that guides the
+hypotheses we make, the data we collect, and the management decisions we
+make. Process-based models, whether expressed in the language of
+mathematics or English, identify the connection between mosquitoes and
+the spread of malaria, or greenhouses gases and climate change, guiding
+our approach to understand and manage these threats. Despite this
+central importance, we argue here that ecologists would do well to give
+greater attention to the role non-mechanistic models can play in
+ecological management and decision making. The value of these approaches
+is greatest in a context where decisions are made over short time
+horizons, and updated as new data becomes available.
+
+\subsection{The alternative to process-based models: pattern-based
+modeling}
+
+Before we proceed further it would be useful to define our terms and lay
+some greater context for the discussion. Throughout, we will distinguish
+``mechanistic'' or ``process-based'' models from models that are merely
+``correlative'' or ``pattern-based''. Historically pattern-based
+modeling meant regression (usually linear regression) and resided in
+methodology rooted in statistics departments and was the primary focus
+of empirical ecologists, while mechanistic modeling meant dynamical
+systems (ODEs, PDEs, SDEs and their discrete kin), residing in
+methodology from the mathematics department and was the primary focus of
+theoretical ecologists.
+
+We believe these historical divides continue to color much of the
+literature today, with theorists more skeptical than empiricists of
+statistical methodology and vice versa. Meanwhile, the ground underneath
+has shifted. Originally pattern-matching approaches could be critiqued
+on the grounds of their simplicity (not everything is linear) while
+mechanistic approaches could be critiqued on their tenuous connections
+to data -- model parameter values such as death rates, birth rates, etc.
+would be estimated in advance and then stuck into the model, rather than
+estimated in the context of the model itself. Today both of these
+critiques are outdated. Computational power and hierarchical statistical
+methods (particularly approximate likelihood or simulation techniques
+such as particle filters and approximate Bayesian computing) have
+brought the inference richer dynamical systems into their fold, while
+pattern-based modeling has spawned an entirely new approach under the
+banner of machine learning that has divided the statistics community
+(See @Bremman2001, ``The Two Cultures''). Machine learning models can
+represent almost arbitrarily complex patterns and incorporate learning
+and decision making strategies that have earned them the name
+`artificial intelligence.' Yet at the same time they can be far less
+transparent their linear regression predecessors, making both astounding
+successes and startling biases and failures like their biological
+namesake. More ecologists, particularly theorists and modelers, would do
+well to learn both how to take advantage of these strengths and
+recognize their weaknesses.
+
+Mechanistic modeling emphasizes the importance of capturing the correct
+gross properties of a system over tracking minute fluctuations. For
+instance, in selecting and parameterizing model of a population of
+conservation concern, we may be most interested in getting the long-term
+behavior correct -- such as identifying if the dynamics support
+persistence of the population -- rather than worrying how well they
+reflect the year-to-year fluctuations. We would certainly have good
+reason to prefer such a model over alternatives which are irreconcilable
+to the most basic biological processes, such as unbounded growth, or
+growth curves that do not pass through the origin in a closed system. So
+it may come as a surprise to realize that such obviously wrong models
+can perform as well or even better than reasonable mechanistic models in
+guiding ecological management and decision making.
+
+Most management recommendations from the ecological literature are based
+on (or at least motivated by) parametric models. Though in principle
+these models reflect mechanistic underpinnings of biological
+interactions involved, in practice real mechanisms are often unknown and
+the true dynamics too complex to be captured by simple models
+{[}@refs{]}. While simple mechanistic models can nevertheless provide
+imporant insights into possible dynamics -- for instance, demonstrating
+that a vaccine does not have to be 100\% effective to eliminate the
+transmission of a virus {[}@Kermack1921{]} -- such approaches are not
+well-suited for use in forecasting future dynamics upon which outcomes
+management policy can be based. Despite a long history of work
+emphasizing the difference between modeling for understanding (generic
+or strategic models) and modeling for prediction (precise, specific, or
+tactical models) (e.g. Levins 1966), simple parametric models continue
+to find application in predictive management contexts. The management of
+marine fisheries is a prime example, in which simple, mechanistically
+motivated models of stock-recruitment relationships such as Ricker or
+Beverton-Holt curves are systematically fit to data and used in solving
+decision-theoretic or optimal control problems determining resource
+management policies such as maximum sustainable yield {[}@refs{]}.
+
+Stochastic dynamic programming @Mangel1982
+
+To address structural uncertainty in the processes involved, such
+mechanistic models can be adapted to accomidate uncertainty in the
+parameters or consider a set of alternative models simultaneously
+(belief SDP)
+
+Further unknowns such as measurement uncertainty, parameter uncertainty,
+unobserved states, knowledge of boundary conditions, etc. further
+compound the issue. Though a hierarchical Bayesian approach provides a
+natural way to address these from a statistical standpoint, formulating
+reasonable parametric descriptions of each form of uncertainty is a
+challenging task in itself, let alone the computational difficulties of
+solving such a system. Cressie et al. (2009) provides a good account of
+the successess and challenges of the approach. Applying these approaches
+in the management context of sequential decision making, in which
+forecasts must be obtained over a range of possible actions and updated
+regularly as new information arrives makes such an approach less
+feasible still.
+
+Though machine learning approaches have begun to appear in the
+ecological and conservation literature (Species distribution models),
+including the Gaussian process based approach used here (Munch et al.
+2005), they remain unfamiliar and untrusted approaches for most
+ecologists. Machine learning approaches represent an essentially
+pattern-based rather than process-based approach, raising the same
+skepticism from most theoretical ecologists that they hold for older
+correlative methods such as linear regression, while the complexity of
+the techniques has barred their adoption in more empirical audiences.
+
+A more immediate barrier to their adoption is the absence of a framework
+for applying machine learning approaches to resource management
+problems. Traditional approaches to optimal control (Pontryagin's
+principle, stochastic dynamic programming) rely on knowledge of the
+state equation, usually described by a simple parametric model. Here we
+illustrate how a stochastic dynamic programming algorithm can
+alternatively be driven by the predictions from a Gaussian process -- a
+machine learning approximation to the state dynamics.\\
+
 Management goals / decision-theoretic approaches need accurate
-  prediction over relevant (short?) timescales more than accurate (but
-  incomplete or noisy estimated) mechanisms.
-\item
-  Nonparametric (machine-learning?) approaches may offer the benefit of
-  the hierarchical Bayesian approach without the practical and
-  computational limitations of their parametric kin. Non-parametric
-  models are flexible enough to take maximum advantage of the data
-  available, while being appropriately ambiguous about the dynamics of a
-  system in regions of parameter space that have been poorly or never
-  sampled.
-\item
-  Nonparametric approaches are beginning to appear more frequently in
-  ecological and conservation literature (Species distribution
-  models/maxent), including the Gaussian process based approach used
-  here (Munch et al. 2005). However, such approaches have yet to be
-  applied to the decision-theoretic framework that could guide
-  management decisions, where we expect them to excel for several
-  reasons (1) the ability to make accurate forecasts by more closely
-  approximating the underlying process where data is available (2)
-  remaining appropriately ambiguous where data is not available (3)
-  remaining computationally simple enough to avoid some pitfalls common
-  to hierarchical parametric approaches.
-\end{itemize}
+prediction over relevant (short?) timescales more than reasonable
+long-term mechanisms. Machine-learning approaches may offer the benefit
+of the hierarchical Bayesian approach without the practical and
+computational limitations of their parametric kin. Non-parametric models
+are flexible enough to take maximum advantage of the data available,
+while being appropriately ambiguous about the dynamics of a system in
+regions of parameter space that have been poorly or never sampled.
 
@@
 introduced in the ecological modeling and fisheries management by Munch
-et al. (2005).
-
-\emph{Is the following discription useful or too elementary?}
+et al. (2005). An accessible and thorough introduction to the
+formulation and use of Gaussian processes can be found in
+@Rasmussen2006.
 
@@
 Process approach postulates a prior distribution of (n-dimensional)
-curves that can be though of as approximations to a large range of
-possible
+curves that can be though of as approximations to a range of possible
+(parametric) models that might describe the data. The GP allows us to
+consider a set of possible curves simultaneously.\\
 
-\emph{Do we need more specifics on Gaussian process as an approximation
-to parametric models? Discussion of Gaussian process vs other machine
-learning / forecasting approaches that have less clear statistical
-foundations? If so, does this all belong in the discussion?}
-
-The Gaussian process is defined by a covariance kernel \ldots{}
-\emph{discussion of our (gaussian) kernel, blah blah how specific here
-vs appendix?}. By requiring our kernel to follow a generic functional
-form, we can compactly describe the Gaussian process using only a
-handful of parameters (Table 1)
-
-A more thorough introduction to the formulation and use of Gaussian
-processes can be found in @Rasmussen2006.
-
-\begin{longtable}[c]{@{}ll@{}}
-\hline\noalign{\medskip}
-parameter & interpretation
-\\\noalign{\medskip}
-\hline\noalign{\medskip}
-$\sigma^2$ & The process noise (from the kernel)
-\\\noalign{\medskip}
-$\tau^2$ & Variance around the mean
-\\\noalign{\medskip}
-$\beta$\_0 & The mean is given by a linear model of slope $\beta$
-\\\noalign{\medskip}
-$d$ & The length-scale of the covariance function
-\\\noalign{\medskip}
-$n$ & The observation error
-\\\noalign{\medskip}
-\hline
-\noalign{\medskip}
-\caption{Table of parameters for the Gaussian process}
-\end{longtable}
-
-Rather than estimate values of these parameters directly, we take a
-hierarchical approach of placing prior distributions on each. Following
-@Gramarcy2005 we use a Gaussian prior on $\beta_0$, an inverse gamma
-prior on $\sigma^2$ and $\tau^2$, a gamma prior on the observation noise
-$n$, and exponential prior on the length-scale $d$.
-
-\begin{itemize}
-\item
-  Gaussian processes in ecological literature (Munch et al. 2005)
-\item
-  Our methodology (e.g.~following Munch et al. (2005)).
-\end{itemize}
-
-\subsubsection{Discussion of the dynamic programming solution}
-
-The fishery management problem over an infinite time horizon can be
-stated as:
-
-\begin{align}
-& \max_{ \{h_t\} \geq 0 } \mathbf{E} \lbrace \sum_0^\infty \delta^t \Pi(h_t) \rbrace \\
-& \mathrm{s.t.}  \\
- & X_t = Z_t f\left(S_{t-1}\right) \\
- & S_t = X_t - h_t \\
- & X_t  \geq 0 
-\end{align}
-
-Where $\mathbf{E}$ is the expectation operator, $\delta$ the discount
-rate, $\Pi(h_t)$ the profit expected from a harvest of $h_t$, and other
-terms as in Eq. (1). For simplicity, we have assumed that profits depend
-only on the chosen harvest; simplifying further we will usually consider
-profits to be proportional to harvest, $\Pi(h_t) = h_t$.
-
-Once the posterior Gaussian process (GP) has been estimated (e.g.~see
-Munch et al. 2005), it is necessary to adapt it in place of the
-parametric equation for the stochastic dynamic programming (SDP)
-solution (see Mangel and Clark 1988 for a detailed description of
-parametric SDP methods) to the optimal policy. The essense of the idea
-is straight forward -- we will use the estimated GP in place of the
-parametric growth function to determine the stochastic transition matrix
-on which the SDP calculations are based.
-
-The posterior Gaussian process is completely defined by the expected
-value and covariance matrix at a defined set of training points. For
-simplicty we will consider a these points to fall on a discrete, uniform
-grid $x$ of 101 points from 0 to 15 (1.5 times the positive equilibrium
-$K$). Again to keep things simple we will use this same grid
-discritization for the parametric approach. Other options for choosing
-the grid points, including collocation methods and functional basis
-expansion (or even using Guassian processes in place of the discrete
-optimization; an entirely different context in which GP can be used in
-SDP, see {[}@Deisenroth2009{]}) could also be considered.
-
-The transition matrix $\mathbf{F}$ is thus an 101 by 101 matrix for
-which the ${i,j}$ entry gives the probability of transitioning into
-state $x_i$ given that the system is in state $x_j$ in the previous
-timestep. To generate the transition matrix based on the posterior GP,
-we need only the expected values at each grid point and the
-corresponding variances (the diagonal of the covariance matrix), as
-shown in Figure 1. Given the mean at each gridpoint as the length 101
-vector $E$ and variance $V$, the probability of transitioning from state
-$x_i$ to state $x_j$ is simply
-$\mathcal{N}\left(x_j \vert  \mu = E_i, \sigma = \sqrt{V_i}\right)$,
-where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
-variance $\sigma^2$. Strictly speaking, the transition probability
-should be calculated by integrating the normal density over the bin of
-width $\Delta$ centered at $x_j$. For a sufficiently fine grid that
-$f(x_j) \approx f(x_j + \Delta)$, it is sufficient to calculate the
-density at $x_j$ and then row-normalize the transition matrix.
-
-\subsection{Pseudocode for the determining the transtion matrix from the
-GP}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{for(h in h_grid)}
-  \NormalTok{F_h =}\StringTok{ }\NormalTok{for(x_j in grid)}
-          \NormalTok{for(i in }\DecValTok{1}\NormalTok{:N) }
-            \KeywordTok{dnorm}\NormalTok{(x_j, mu[i]-h, V[i])}
-\end{Highlighting}
-\end{Shaded}
-
-A transition matrix for each of the parametric models $f$ is calculated
-using the log-normal density with mean $f(x_i)$ and log-variance as
-estimated by maximum likelihood. From the discrete transition matrix we
-may write down the Bellman recursion defining the the stochastic dynamic
-programming iteration:
-
-\begin{equation}
-V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
-\end{equation}
-
-where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$ is
-control (harvest level) chosen. Numerically, the maximization is
-accomplished as follows. Consider the set of possible control values to
-be the discrete 101 values corresponding the the grid of stock sizes.
-Then for each $h_t$ there is a corresponding transition matrix
-$\mathbf{F}_h$ determined as described above but with mean
-$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element
-corresponds to the value of having stock $x_i$ at time $t$. Then let
-$\Pi_h$ be the vector whose $i$th element indicates the profit from
-harvesting at intensity $h_t$ given a population $x_i$ (e.g.
-$\max(x_i, h_t)$ since one cannot harvest more fish then the current
-population size). Then the Bellman recursion can be given in matrix form
-as
-
-\[V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)\]
-
-where the sum is element by element and the expectation is computed by
-the matrix multiplication $\mathbf{F} V_{t+1}$.
-
-\subsubsection{Pseudocode for the Bellman iteration}
-
-\begin{Shaded}
-\begin{Highlighting}[]
- \NormalTok{V1 <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
-      \NormalTok{delta *}\StringTok{ }\NormalTok{F[[h]] %*%}\StringTok{ }\NormalTok{V +}\StringTok{  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
-    \NormalTok{\})}
-    \CommentTok{# find havest, h that gives the maximum value}
-    \NormalTok{out <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
-      \NormalTok{value <-}\StringTok{ }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
-      \NormalTok{index <-}\StringTok{ }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
-      \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
-    \NormalTok{\})}
-    \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
-    \NormalTok{V <-}\StringTok{ }\NormalTok{out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
-    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <-}\StringTok{ }\NormalTok{out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
-\end{Highlighting}
-\end{Shaded}
-
-\emph{Currently this shows the literal R code, should be adapted}
+\subsubsection{The optimal control problem}
 
@@
 
-\begin{itemize}
-\item
 Non-parametric methods have received far too little attention in
@@
 planning and decision making support.
-\item
-  Importance of non-parametric approaches in conservation planning /
-  resource management / decision theory.
-\item
-  Decision-theoretic tools such as optimal control calculations rely on
-  robust \emph{forecasting} more strongly than they rely on accurate
-  \emph{mechanistic} relationships.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
 \item
@@
   might be improved.
-\item
-  Anticipate improved relative performance in higher dimensional
-  examples
-\item
-  Discuss constant escapement in model, in policies.
-\item
-  Limitations of this comparison: Are the maximum-likelihood solutions a
-  straw man?
-\item
-  Discussion of alternative related approaches: POMDP/MOMDP,
 \end{itemize}
 
+Three elements contribute to the machine learning model performing as
+well or better than the mechanisitic models in this context.
+
+\subsubsection{1. Relevant state space}
+
+The dynamics occur over a range of state-space in which the machine
+learning model closely matches the predictions of y implausible model
+performs as well or better than the more plausible model.
+
+This is the most obvious and immediate reason why the shortcomings that
+appear to make the model implausible do not make it useless. This effect
+is enhanced in our example because the management actions help drive the
+system towards rather than away from this region of state-space.
+
+\subsubsection{2. Predictive accuracy}
+
+This application relies only on the predictive accuracy of the model,
+not an interpretation of the parameters.
+
+Predictive accuracy is not the goal of all modeling, as ecologists have
+been observing for as long as they made models (perhaps none more
+memorably than @Levins1969). Mechanistic modeling is at its most
+powerful not when it is used to directly forecast future states but when
+it provides an understanding of how to approach a problem. SIR-type
+models from the epidemiological literature are a good example. While the
+simplest SIR models have little predictive power over the outbreak
+intensity or timing at a particular location, they provide a powerful
+understanding of the spread of an infection in terms of a single,
+biologically meaningful parameter: $R_0$, the basic reproductive number.
+From the model, it becomes clear that management need not vaccinate
+every member of the population to stop the spread, but rather it
+suffices to vaccinate a sufficient fraction of the population to reduce
+$R_0$ below 1.
+
+\subsubsection{3. Time scale for new data}
+
+In the sequential decision making problem we considered, we are
+presented with new data after each action. The relevant timescale for
+the prediction is thus not the long-term dynamics, which would be wildly
+divergent, but the dynamics over this much shorter interval. While
+ecologists may be hesitant to base continual management on a model with
+obviously inaccurate long-term behavior, engineers tend to consider the
+problem in frequency space and gravitate to the opposite position -- a
+good control model should prioritize high-frequency accuracy over low
+frequency accuracy. The differences in intuition may arise from the
+timescales at which each profession can typically adjust their control
+variables -- much faster for a control system of a chemical plant than
+state policy for a natural resource. Still, the lesson is clear: when
+facing repeated management decisions over a short timescale, such as
+setting annual harvests of a fishery, it may be more valuable to use a
+machine learning algorithm that makes accurate year-ahead predictions
+that capture some of the high-frequency fluctuations that appear only as
+noise in a mechanistic model of the long-term population dynamics.
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Better express ambiguity in regions where data is not available
+\end{enumerate}
+
+One of the greatest strengths of mechanistic models is their greatest
+weakness as well.
+
 \subsection{Future directions}
+
+While we have highlighted certain generic of this problem that allow the
+nonparametric approach to excel -- short timescales between successive
+observations and actions, the accuracy in the appropriate region of
+state space, the ability to express uncertainty outside the observed
+data -- there are equally several aspects for which the nonparametric
+approach is at a relative disadvantage.
+
+In this simulated example, the underlying dynamics are truly governed by
+a simple parametric model, allowing the parametric approaches to be more
+accurate. Similarly, because the dynamics are one-dimensional dynamics
+and lead to stable nodes (rather than other attractors such as
+limit-cycles resulting in oscillations), the training data provides
+relatively limited information about the dynamics. For these reasons, we
+anticipate that in higher-dimensional examples characteristic of
+ecosystem management problems that the machine learning approach will
+prove even more valuable.
+
+The machine learning approach is also better suited to complex and
+disparate data. Incorporating various sources of information into
+mechanistic models can be an immensely difficult due to the increased
+complexity involved. Only thanks to tandem advances in increasing
+computational power and hierarchical statistical methodology have we
+been able to tackle such intuitively important complexity (and the
+potentially new available data that accompanies it) such as spatial
+distribution, heterogeneities of space, time, and individuals, to shift
+to ecosystem-based approaches from single-species based approaches.
+Without the need to formulate mechanisms, many modern machine learning
+algorithms can leverage potential information from all available sources
+of data directly. The algorithms can recognize unanticipated or subtle
+patterns in large data sets that enable more accurate predictions than
+mechanistic models that are formulated at a more macroscopic level.
+
+\begin{center}\rule{3in}{0.4pt}\end{center}
 
@@
 \section{Appendix / Supplementary Materials}
+
+\subsubsection{The Guassian process}
+
+The Gaussian process is defined by a covariance kernel. By requiring our
+kernel to follow a generic functional form, we can compactly describe
+the Gaussian process using only a handful of parameters (Table 1)
+
+\begin{longtable}[c]{@{}ll@{}}
+\hline\noalign{\medskip}
+parameter & interpretation
+\\\noalign{\medskip}
+\hline\noalign{\medskip}
+$\sigma^2$ & The process noise (from the kernel)
+\\\noalign{\medskip}
+$\tau^2$ & Variance around the mean
+\\\noalign{\medskip}
+$\beta$\_0 & The mean is given by a linear model of slope $\beta$
+\\\noalign{\medskip}
+$d$ & The length-scale of the covariance function
+\\\noalign{\medskip}
+$n$ & The observation error
+\\\noalign{\medskip}
+\hline
+\noalign{\medskip}
+\caption{Table of parameters for the Gaussian process}
+\end{longtable}
+
+Rather than estimate values of these parameters directly, we take a
+hierarchical approach of placing prior distributions on each. Following
+@Gramarcy2005 we use a Gaussian prior on $\beta_0$, an inverse gamma
+prior on $\sigma^2$ and $\tau^2$, a gamma prior on the observation noise
+$n$, and exponential prior on the length-scale $d$.
+
+\subsubsection{Formulating a dynamic programming solution}
+
+The fishery management problem over an infinite time horizon can be
+stated as:
+
+\begin{align}
+& \max_{ \{h_t\} \geq 0 } \mathbf{E} \lbrace \sum_0^\infty \delta^t \Pi(h_t) \rbrace \\
+& \mathrm{s.t.}  \\
+ & X_t = Z_t f\left(S_{t-1}\right) \\
+ & S_t = X_t - h_t \\
+ & X_t  \geq 0 
+\end{align}
+
+Where $\mathbf{E}$ is the expectation operator, $\delta$ the discount
+rate, $\Pi(h_t)$ the profit expected from a harvest of $h_t$, and other
+terms as in Eq. (1). For simplicity, we have assumed that profits depend
+only on the chosen harvest; simplifying further we will usually consider
+profits to be proportional to harvest, $\Pi(h_t) = h_t$.
+
+Once the posterior Gaussian process (GP) has been estimated (e.g.~see
+Munch et al. 2005), it is necessary to adapt it in place of the
+parametric equation for the stochastic dynamic programming (SDP)
+solution (see Mangel and Clark 1988 for a detailed description of
+parametric SDP methods) to the optimal policy. The essense of the idea
+is straight forward -- we will use the estimated GP in place of the
+parametric growth function to determine the stochastic transition matrix
+on which the SDP calculations are based.
+
+The posterior Gaussian process is completely defined by the expected
+value and covariance matrix at a defined set of training points. For
+simplicty we will consider a these points to fall on a discrete, uniform
+grid $x$ of 101 points from 0 to 15 (1.5 times the positive equilibrium
+$K$). Again to keep things simple we will use this same grid
+discritization for the parametric approach. Other options for choosing
+the grid points, including collocation methods and functional basis
+expansion (or even using Guassian processes in place of the discrete
+optimization; an entirely different context in which GP can be used in
+SDP, see {[}@Deisenroth2009{]}) could also be considered.
+
+The transition matrix $\mathbf{F}$ is thus an 101 by 101 matrix for
+which the ${i,j}$ entry gives the probability of transitioning into
+state $x_i$ given that the system is in state $x_j$ in the previous
+timestep. To generate the transition matrix based on the posterior GP,
+we need only the expected values at each grid point and the
+corresponding variances (the diagonal of the covariance matrix), as
+shown in Figure 1. Given the mean at each gridpoint as the length 101
+vector $E$ and variance $V$, the probability of transitioning from state
+$x_i$ to state $x_j$ is simply
+$\mathcal{N}\left(x_j \vert  \mu = E_i, \sigma = \sqrt{V_i}\right)$,
+where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
+variance $\sigma^2$. Strictly speaking, the transition probability
+should be calculated by integrating the normal density over the bin of
+width $\Delta$ centered at $x_j$. For a sufficiently fine grid that
+$f(x_j) \approx f(x_j + \Delta)$, it is sufficient to calculate the
+density at $x_j$ and then row-normalize the transition matrix.
+
+\subsection{Pseudocode for the determining the transtion matrix from the
+GP}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{for(h in h_grid)}
+  \NormalTok{F_h =}\StringTok{ }\NormalTok{for(x_j in grid)}
+          \NormalTok{for(i in }\DecValTok{1}\NormalTok{:N) }
+            \KeywordTok{dnorm}\NormalTok{(x_j, mu[i]-h, V[i])}
+\end{Highlighting}
+\end{Shaded}
+
+A transition matrix for each of the parametric models $f$ is calculated
+using the log-normal density with mean $f(x_i)$ and log-variance as
+estimated by maximum likelihood. From the discrete transition matrix we
+may write down the Bellman recursion defining the the stochastic dynamic
+programming iteration:
+
+\begin{equation}
+V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
+\end{equation}
+
+where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$ is
+control (harvest level) chosen. Numerically, the maximization is
+accomplished as follows. Consider the set of possible control values to
+be the discrete 101 values corresponding the the grid of stock sizes.
+Then for each $h_t$ there is a corresponding transition matrix
+$\mathbf{F}_h$ determined as described above but with mean
+$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element
+corresponds to the value of having stock $x_i$ at time $t$. Then let
+$\Pi_h$ be the vector whose $i$th element indicates the profit from
+harvesting at intensity $h_t$ given a population $x_i$ (e.g.
+$\max(x_i, h_t)$ since one cannot harvest more fish then the current
+population size). Then the Bellman recursion can be given in matrix form
+as
+
+\[V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)\]
+
+where the sum is element by element and the expectation is computed by
+the matrix multiplication $\mathbf{F} V_{t+1}$.
+
+\subsubsection{Pseudocode for the Bellman iteration}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+ \NormalTok{V1 <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
+      \NormalTok{delta *}\StringTok{ }\NormalTok{F[[h]] %*%}\StringTok{ }\NormalTok{V +}\StringTok{  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
+    \NormalTok{\})}
+    \CommentTok{# find havest, h that gives the maximum value}
+    \NormalTok{out <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
+      \NormalTok{value <-}\StringTok{ }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
+      \NormalTok{index <-}\StringTok{ }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
+      \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
+    \NormalTok{\})}
+    \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
+    \NormalTok{V <-}\StringTok{ }\NormalTok{out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
+    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <-}\StringTok{ }\NormalTok{out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Currently this shows the literal R code, should be adapted}
 
@@
 
-Geritz, Stefan a. H., and Eva Kisdi. 2011. ``Mathematical ecology: why
-mechanistic models?'' \emph{Journal of mathematical biology} (dec).
-doi:10.1007/s00285-011-0496-3.
-
 Levins, Richard. 1966. ``The strategy of model building in population
@@
 
-Polasky, Stephen, Stephen R. Carpenter, Carl Folke, and Bonnie Keeler.
-2011. ``Decision-making under great uncertainty: environmental
-management in an era of global change.'' \emph{Trends in ecology \&
-evolution} (may): 1--7. doi:10.1016/j.tree.2011.04.007.
-
 Reed, William J. 1979. ``Optimal escapement levels in stochastic and

### 2013-03-21 inst/doc/nonparametric-bayes.tex
updates to pomp during our discussion today

@@
   \author[cstar]{Carl Boettiger\corref{cor1}}
+  \author[cstar]{Marc Mangel}
   \author[noaa]{Stephan B. Munch}
-  \author[cstar]{Marc Mangel}
   \ead{cboettig@ucsc.edu}
@@
 \section{Introduction}
-
-Process-based models built on modern machine learning approaches offer a
-powerful and underappreciated approach for ecological forecasting and
-management. We present a novel example of how one such example -
-non-parametric Gaussian processes, can be adapted to the context of
-management and decision-making.
-
-While effective machine learning algorithms have speckled the
-literature, most theoretical ecologists in particular have been wary of
-these `black-box' approaches in favor of simple, mechanistic models.
-While a blind application of machine learning approaches would indeed be
-dangerous grounds to stake either ecological understanding or management
-practice, such concerns are not grounds to reject such methods out of
-hand. Rather, further attention is warranted to identify the
-circumstances in which machine-learning approaches will most likely
-outperform existing more mechanistic approaches, or help crack difficult
-management questions where ecological or dynamical complexity has caused
-traditional approaches to founder.
-
-Here we illustrate how a machine learning model can be implemented in
-place of simple mechanistic model in a decision theory (optimal control)
-problem, and highlight the features of this system that make the
-nonparametric approach particularly effective.
-
-Mechanistic models have long been the gold standard of theoretical
-modeling in ecology (e.g.~see @Geritz2012 or @Cuddington2013). Only by
-understanding the processes involved can we make reliable long term
-predictions and build an knowledge of cause and effect that guides the
-hypotheses we make, the data we collect, and the management decisions we
-make. Process-based models, whether expressed in the language of
-mathematics or English, identify the connection between mosquitoes and
-the spread of malaria, or greenhouses gases and climate change, guiding
-our approach to understand and manage these threats. Despite this
-central importance, we argue here that ecologists would do well to give
-greater attention to the role non-mechanistic models can play in
-ecological management and decision making. The value of these approaches
-is greatest in a context where decisions are made over short time
-horizons, and updated as new data becomes available.
-
-\subsection{The alternative to process-based models: pattern-based
-modeling}
-
-Before we proceed further it would be useful to define our terms and lay
-some greater context for the discussion. Throughout, we will distinguish
-``mechanistic'' or ``process-based'' models from models that are merely
-``correlative'' or ``pattern-based''. Historically pattern-based
-modeling meant regression (usually linear regression) and resided in
-methodology rooted in statistics departments and was the primary focus
-of empirical ecologists, while mechanistic modeling meant dynamical
-systems (ODEs, PDEs, SDEs and their discrete kin), residing in
-methodology from the mathematics department and was the primary focus of
-theoretical ecologists.
-
-We believe these historical divides continue to color much of the
-literature today, with theorists more skeptical than empiricists of
-statistical methodology and vice versa. Meanwhile, the ground underneath
-has shifted. Originally pattern-matching approaches could be critiqued
-on the grounds of their simplicity (not everything is linear) while
-mechanistic approaches could be critiqued on their tenuous connections
-to data -- model parameter values such as death rates, birth rates, etc.
-would be estimated in advance and then stuck into the model, rather than
-estimated in the context of the model itself. Today both of these
-critiques are outdated. Computational power and hierarchical statistical
-methods (particularly approximate likelihood or simulation techniques
-such as particle filters and approximate Bayesian computing) have
-brought the inference richer dynamical systems into their fold, while
-pattern-based modeling has spawned an entirely new approach under the
-banner of machine learning that has divided the statistics community
-(See @Bremman2001, ``The Two Cultures''). Machine learning models can
-represent almost arbitrarily complex patterns and incorporate learning
-and decision making strategies that have earned them the name
-`artificial intelligence.' Yet at the same time they can be far less
-transparent their linear regression predecessors, making both astounding
-successes and startling biases and failures like their biological
-namesake. More ecologists, particularly theorists and modelers, would do
-well to learn both how to take advantage of these strengths and
-recognize their weaknesses.
-
-Mechanistic modeling emphasizes the importance of capturing the correct
-gross properties of a system over tracking minute fluctuations. For
-instance, in selecting and parameterizing model of a population of
-conservation concern, we may be most interested in getting the long-term
-behavior correct -- such as identifying if the dynamics support
-persistence of the population -- rather than worrying how well they
-reflect the year-to-year fluctuations. We would certainly have good
-reason to prefer such a model over alternatives which are irreconcilable
-to the most basic biological processes, such as unbounded growth, or
-growth curves that do not pass through the origin in a closed system. So
-it may come as a surprise to realize that such obviously wrong models
-can perform as well or even better than reasonable mechanistic models in
-guiding ecological management and decision making.
 
@@
 \item
-  Multiple species * Online learning * Multiple step-ahead predictions *
-  Explicitly accomidating additional uncertainties * Improving inference
-  of optimal policy from the GP
+  Multiple species
+\item
+  Online learning
+\item
+  Multiple step-ahead predictions
+\item
+  Explicitly accomidating additional uncertainties
+\item
+  Improving inference of optimal policy from the GP
 \end{itemize}
-
-\texttt{\{r echo=FALSE, results="asis"\} \#bibliography("html")}
 

### 2013-03-27 inst/doc/nonparametric-bayes.Rmd
updates to pomp during our discussion today
writing writing

--- 2013-03-21 inst/doc/nonparametric-bayes.Rmd
@@
 
-Most management recommendations from the ecological literature are based on
-(or at least motivated by) parametric models. Though in principle these models
-reflect mechanistic underpinnings of biological interactions involved, in practice
-real mechanisms are often unknown and the true dynamics too complex to be captured
-by simple models [@refs]. While simple mechanistic models can nevertheless
-provide imporant insights into possible dynamics -- for instance, demonstrating
-that a vaccine does not have to be 100% effective to eliminate the transmission of a virus [@Kermack1921] -- such approaches are not well-suited for use in 
-forecasting future dynamics upon which outcomes management policy can be based.  
-Despite a long history of work emphasizing the difference between modeling 
-for understanding (generic or strategic models) and modeling for prediction (precise, specific, or tactical models) [e.g. @Levins1966], simple parametric models 
-continue to find application in predictive management contexts.  The management
-of marine fisheries is a prime example, in which simple, mechanistically 
-motivated models of stock-recruitment relationships such as Ricker or Beverton-Holt
-curves are systematically fit to data and used in solving decision-theoretic
-or optimal control problems determining resource management policies such
-as maximum sustainable yield [@refs].
+Most management recommendations from the ecological literature are based
+on (or at least motivated by) parametric models. Though in principle
+these models reflect mechanistic underpinnings of biological interactions
+involved, in practice real mechanisms are often unknown and the true
+dynamics too complex to be captured by simple models [@refs]. While
+simple mechanistic models can nevertheless provide imporant insights
+into possible dynamics -- for instance, demonstrating that a vaccine
+does not have to be 100% effective to eliminate the transmission of a
+virus [@Kermack1921] -- such approaches are not well-suited for use in
+forecasting outcomes of potential management options.  Non-parametric 
+approaches offer a more flexible alternative that can both more accurately
+reflect the data available while also representing greater uncertainty
+in areas (of state-space) where data is lacking.  Ecological research
+and management strategy should pay closer attention to the opportunities 
+and challenges nonparametric modeling can offer.
 
-<!-- Better defintion of focal problems here .... -->
-Stochastic dynamic programming @Mangel1982
+### Quantitative vs Qualitative Decisions
 
-To address structural uncertainty in the processes involved, such mechanistic models
-can be adapted to accomidate uncertainty in the parameters or consider a set of 
-alternative models simultaneously (belief SDP)
+In this paper, we consider those ecological management problems in which
+a mathematical (or computational) model is used to quantitatively inform
+decision-making by allowing a manager to compare to the expected consequences
+of potential management actions (or policies).  We distinguish this from the
+solely qualitative use of a model, in which models are used to represent and
+compare hypotheses of different mechanisms that would lead to qualitatively 
+different actions.
 
-Structural uncertainty.
+We argue that while simple mechanistically motivated models
+may be best for the latter case [@Cunningham2013; @Geritz2012], such models 
+can be not only inaccurate but misleading in quantitative decision making.
+Nonparametric models can more realistically represent uncertainties outside 
+of observed data while also better capturing the dynamics in the region
+observed.  Advances in the theory and computational implementations of 
+nonparametric methods make them ripe for such applications.  We use the classic
+problem of optimal harvest of a marine fishery to illustrate how the 
+nonparametric approach of Gaussian processes can be applied.
 
+We imagine the manager seeks to maximize some known socio-economic value function, 
+given their knowledge of the system. 
+In particular, we will focus on the case in which the manager faces a series of decisions,
+such as setting a harvest quota for a marine fishery each year based on observations
+of the stock size the year before.
+
+### Uncertainties: the classical approach
+
+The ecological dynamics of most systems of management interest are 
+typically both complex and unknown [@refs].  Despite this, quantitative policy
+optimization in these systems has almost exclusively been based on
+simple, mechanistically motivated parametric models.
+There are several reasons for this.  First, limited available data almost 
+always percludes any accurate estimate of more complicated models.  Second,
+the computational demands of determining an optimal stategy in a sequential
+decision-theory problem suffer from the "curse of dimensionality" that makes  
+it difficult to consider more complicated models, states, or actions.  Unfortunately,
+these challenges also make it particularly difficult to reflect the true uncertainty
+either in the parameters of the given models (parametric uncertainty) or in 
+the structure of the models themselves (structural uncertainty).  
 Further unknowns such as measurement uncertainty, parameter uncertainty,
@@
 
+An active literature and growing computational power over the past several decades have
+only marginally improved this situation.  Parametric or structural uncertainty
+can be introduced only by increasing the state space to reflect a distribution
+of possible parameters [@refs] or the degree of belief in each of a set of possible models [@refs]
+Frequently, the space of possible actions must then be reduced or the algorithms adjusted
+by approximations to keep computations feasible.
+
 Though machine learning approaches have begun to appear in the
@@
 they remain unfamiliar and untrusted approaches for most ecologists.
-Machine learning approaches represent an essentially pattern-based
-rather than process-based approach, raising the same skepticism from most
-theoretical ecologists that they hold for older correlative methods such
-as linear regression, while the complexity of the techniques has barred
-their adoption in more empirical audiences. <!-- Whoa! rampant speculation
-alert! -->
-
-<!-- Whoa, isn't machine learning used to make decisions all the time,
- eg, in engineering and Artificial Intelligence?  -->
-
-A more immediate barrier to their adoption is the absence of a framework
+One potential barrier to their adoption is the absence of a framework
 for applying machine learning approaches to resource management problems.
@@
 a Gaussian process -- a machine learning approximation to the state dynamics.
+
 <!-- Okay, perhaps that's novel, but it's pretty trivial.  Isn't it
@@
 
-```{r stateeq}
-f <- RickerAllee
-p <- c(2, 10, 5) 
-K <- 10
-allee <- 5
-```
+### The fisheries management problem
 
-```{r sdp-pars, dependson="stateeq"}
-sigma_g <- 0.05
-sigma_m <- 0.0
-z_g <- function() rlnorm(1, 0, sigma_g)
-z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
-x_grid <- seq(0, 1.5 * K, length=101)
-h_grid <- x_grid
-profit <- function(x,h) pmin(x, h)
-delta <- 0.01
-OptTime <- 20  # stationarity with unstable models is tricky thing
-reward <- 0
-xT <- 0
-seed_i <- 1
-Xo <- K # observations start from
-x0 <- Xo # simulation under policy starts from
-Tobs <- 35
-```
+In our example, we focus on the problem in which a manager must set 
+the harvest level for a marine fishery each year to maximize the net
+present value of the resource, given an estimated stock size from the year
+before. Rich data and global concerns have made marine fisheries the crucible for much
+of the founding work [@Gordon1954; @Reed1979; @May1979; @Ludwig1982]
+in managing ecosystems under uncertainty.  Global trends [@Worm2006]
+and controversy [@Hilborn2007; @Worm2009] have made understanding these
+challenges all the more pressing.
 
-### Background on Gaussian Process inference
+To permit comparisons against a theoretical optimum we will consider
+data on the stock dynamics simulated from a simple parametric model
+in which recruitment of the fish stock $X_{t+1}$ in the following year
+is a stochastic process governed by a function $f$ of the current 
+stock $X_t$, selected harvest policy $h_t$, and noise process $Z$,
+
+$$X_{t+1} = Z_t f(X_t, h_t) $$
+
+Given parameters for the function $f$ and probability distribution $Z$, 
+along with a given economic model determining the  price/profit $\Pi(X_t, h_t)$
+realized in a given year given a choice of harvest $h_t$ and observed stock $X_t$.
+This problem can be solved exactly for discretized values of stock $X$ and policy $h$
+using stochastic dynamic programming [@Mangel1982].  
+Problems of this sort underpin much marine fisheries management today.
+
+A crux of this approach is correctly specifying the functional form of $f$,
+along with its parameters.  The standard approach [@refs] uses one of a 
+handful of common parametric models representing the stock-recruitment
+relationship, usually after estimating the model parameters from any 
+available existing data. Though modifications of this approach mentioned in the 
+introduction can permit additional sources of uncertainty such as measurement 
+error in the stock assessment, implementation errors in the harvest policy, [@Clark1986, @Roughgarden1996, @Sethi2005]
+uncertainty in parameters [@Mangel1985, @Schaupaugh2013] or model structure [@Williams2001, @Athanassoglou2012]
+
+Here, we compare this approach to our alternative
+that uses a Gaussian Process in place of the a given stock recruitment curve.
+
+### The Non-parametric Bayesian alternative for stock-recruitment curves
 
@@
 
+Following [@Munch2005], we will estimate Gaussian processes from
+
 <!-- Do we need more specifics on Gaussian process as an approximation
@@
 foundations?  If so, does this all belong in the discussion? -->
+
+###
+
+```{r stateeq}
+f <- RickerAllee
+p <- c(2, 10, 5) 
+K <- 10
+allee <- 5
+```
+
+```{r sdp-pars, dependson="stateeq"}
+sigma_g <- 0.05
+sigma_m <- 0.0
+z_g <- function() rlnorm(1, 0, sigma_g)
+z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
+x_grid <- seq(0, 1.5 * K, length=101)
+h_grid <- x_grid
+profit <- function(x,h) pmin(x, h)
+delta <- 0.01
+OptTime <- 20  # stationarity with unstable models is tricky thing
+reward <- 0
+xT <- 0
+seed_i <- 1
+Xo <- K # observations start from
+x0 <- Xo # simulation under policy starts from
+Tobs <- 35
+```
+
+### Background on Gaussian Process inference
 

### 2013-03-29 inst/doc/nonparametric-bayes.Rmd
automate bibliographic extraction

@@
 
+```{r stateeq}
+f <- RickerAllee
+p <- c(2, 10, 5) 
+K <- 10
+allee <- 5
+```
+
+```{r sdp-pars, dependson="stateeq"}
+sigma_g <- 0.05
+sigma_m <- 0.0
+z_g <- function() rlnorm(1, 0, sigma_g)
+z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
+x_grid <- seq(0, 1.5 * K, length=101)
+h_grid <- x_grid
+profit <- function(x,h) pmin(x, h)
+delta <- 0.01
+OptTime <- 20  # stationarity with unstable models is tricky thing
+reward <- 0
+xT <- 0
+seed_i <- 1
+Xo <- K # observations start from
+x0 <- Xo # simulation under policy starts from
+Tobs <- 35
+```
+
 Abstract
@@
 does not have to be 100% effective to eliminate the transmission of a
-virus [@Kermack1921] -- such approaches are not well-suited for use in
+virus  -- such approaches are not well-suited for use in
 forecasting outcomes of potential management options.  Non-parametric 
@@
 We argue that while simple mechanistically motivated models
-may be best for the latter case [@Cunningham2013; @Geritz2012], such models 
+may be best for the latter case [@Cuddington2013; @Geritz2012], such models 
 can be not only inaccurate but misleading in quantitative decision making.
@@
 
-### The fisheries management problem
+### The optimal control problem in fisheries management
 
@@
 This problem can be solved exactly for discretized values of stock $X$ and policy $h$
-using stochastic dynamic programming [@Mangel1982].  
+using stochastic dynamic programming (SDP) [@Mangel1982].  
 Problems of this sort underpin much marine fisheries management today.
@@
 error in the stock assessment, implementation errors in the harvest policy, [@Clark1986, @Roughgarden1996, @Sethi2005]
-uncertainty in parameters [@Mangel1985, @Schaupaugh2013] or model structure [@Williams2001, @Athanassoglou2012]
+uncertainty in parameters [@Mangel1985, @Schaupaugh2013] or model structure [@Williams2001, @Athanassoglou2012],
+we focus on the simplest example to illustrate the comparison, where only a 
+single form of uncertainty (the growth shocks $Z_t$) is present in the underlying
+model, making additional modes unnecessary. 
+<!--Add comment / apology on single species dynamics -->
 
-Here, we compare this approach to our alternative
-that uses a Gaussian Process in place of the a given stock recruitment curve.
+We compare this approach to our alternative that uses a Gaussian Process (GP)
+in place of the a given stock recruitment curve.  We briefly revisit
+a description of the Gaussian Process approach that has appeared in 
+recent literature before describing how we take the important step of 
+adapting the GP to the decision theory problem that can still be solved
+using SDP.
 
@@
 
-The use of Gaussian process regression (known as Kreging in the geospatial
+The use of Gaussian process (GP) regression (known as Kreging in the geospatial
 literature) to formulate a predictive model is relatively new in the
@@
 An accessible and thorough introduction to the formulation and use of
-Gaussian processes can be found in @Rasmussen2006.
+GPs can be found in @Rasmussen2006.
 
-The essense of the Gaussian process approach can be captured in the
+The essense of the GP approach can be captured in the
 following thought experiment: An exhaustive parametric approach to the
@@
 models and infer the posterior distribution of possible models. The
-Gaussian process approach can be thought of as a computationally
-efficient approximation to this approach. Gaussian processes represent
+GP approach can be thought of as a computationally
+efficient approximation to this approach. GPs represent
 a large class of models that can be though of as capturing or reasonably
@@
 we can more concisely capture the possible behavior of these curves.
-In place of a parametric model of the dynamical system, the Gaussian
-Process approach postulates a prior distribution of (n-dimensional)
+In place of a parametric model of the dynamical system, the GP 
+approach postulates a prior distribution of (n-dimensional)
 curves that can be though of as approximations to a range of possible
@@
 <!-- Figure 1 include curves drawn from the posterior density?  -->
-
-Following [@Munch2005], we will estimate Gaussian processes from
 
@@
 
-###
+### Background on Gaussian Process inference
 
-```{r stateeq}
-f <- RickerAllee
-p <- c(2, 10, 5) 
-K <- 10
-allee <- 5
+Once the posterior Gaussian process (GP) has been estimated [e.g. see
+@Munch2005], it is necessary to adapt it in place of the parametric
+equation for the stochastic dynamic programming (SDP) solution [see
+@Mangel1988 for a detailed description of parametric SDP methods] to the
+optimal policy. The essense of the idea is straight forward -- we will use
+the estimated GP in place of the parametric growth function to determine
+the stochastic transition matrix on which the SDP calculations are based.
+
+The posterior Gaussian process is completely defined by the expected value
+and covariance matrix at a defined set of training points.  For simplicty
+we will consider a these points to fall on a discrete, uniform grid $x$
+of `r length(x_grid)` points from `r min(x_grid)` to `r max(x_grid)`
+(1.5 times the positive equilibrium $K$).  Again to keep things simple
+we will use this same grid discritization for the parametric approach.
+Other options for choosing the grid points, including collocation methods
+and functional basis expansion (or even using Guassian processes in place
+of the discrete optimization; an entirely different context in which GP
+can be used in SDP, see [@Deisenroth2009]) could also be considered.
+
+The transition matrix $\mathbf{F}$ is thus an `r length(x_grid)`
+by `r length(x_grid)` matrix for which the ${i,j}$ entry gives the
+probability of transitioning into state $x_i$ given that the system is
+in state $x_j$ in the previous timestep.  To generate the transition
+matrix based on the posterior GP, we need only the expected values
+at each grid point and the corresponding variances (the diagonal of
+the covariance matrix), as shown in Figure 1.  Given the mean at each
+gridpoint as the length `r length(x_grid)` vector $E$ and variance $V$,
+the probability of transitioning from state $x_i$ to state $x_j$ is
+simply $\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)$,
+where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
+variance $\sigma^2$.  Strictly speaking, the transition probability should
+be calculated by integrating the normal density over the bin of width
+$\Delta$ centered at $x_j$.  For a sufficiently fine grid that $f(x_j)
+\approx f(x_j + \Delta)$, it is sufficient to calculate the density at
+$x_j$ and then row-normalize the transition matrix.
+
+## Pseudocode for the determining the transtion matrix from the GP
+
+```r
+for(h in h_grid)
+  F_h = for(x_j in grid)
+          for(i in 1:N) 
+            dnorm(x_j, mu[i]-h, V[i])
 ```
 
-```{r sdp-pars, dependson="stateeq"}
-sigma_g <- 0.05
-sigma_m <- 0.0
-z_g <- function() rlnorm(1, 0, sigma_g)
-z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
-x_grid <- seq(0, 1.5 * K, length=101)
-h_grid <- x_grid
-profit <- function(x,h) pmin(x, h)
-delta <- 0.01
-OptTime <- 20  # stationarity with unstable models is tricky thing
-reward <- 0
-xT <- 0
-seed_i <- 1
-Xo <- K # observations start from
-x0 <- Xo # simulation under policy starts from
-Tobs <- 35
+A transition matrix for each of the parametric models $f$ is calculated
+using the log-normal density with mean $f(x_i)$ and log-variance as
+estimated by maximum likelihood.  From the discrete transition matrix we
+may write down the Bellman recursion defining the the stochastic dynamic
+programming iteration:
+
+\begin{equation}
+V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
+\end{equation}
+
+where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$
+is control (harvest level) chosen. Numerically, the maximization is
+accomplished as follows. Consider the set of possible control values to
+be the discrete `r length(h_grid)` values corresponding the the grid of
+stock sizes.  Then for each $h_t$ there is a corresponding transition
+matrix $\mathbf{F}_h$ determined as described above but with mean 
+$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element corresponds
+to the value of having stock $x_i$ at time $t$.  Then let $\Pi_h$ be
+the vector whose $i$th element indicates the profit from harvesting
+at intensity $h_t$ given a population $x_i$ (e.g. $\max(x_i, h_t)$
+since one cannot harvest more fish then the current population size).
+Then the Bellman recursion can be given in matrix form as
+
+$$V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)$$
+
+where the sum is element by element and the expectation is computed by the matrix multiplication $\mathbf{F} V_{t+1}$.
+
+### Pseudocode for the Bellman iteration
+
+```r
+ V1 <- sapply(1:length(h_grid), function(h){
+      delta * F[[h]] %*% V +  profit(x_grid, h_grid[h]) 
+    })
+    # find havest, h that gives the maximum value
+    out <- sapply(1:gridsize, function(j){
+      value <- max(V1[j,], na.rm = T) # each col is a diff h, max over these
+      index <- which.max(V1[j,])  # store index so we can recover h's 
+      c(value, index) # returns both profit value & index of optimal h.  
+    })
+    # Sets V[t+1] = max_h V[t] at each possible state value, x
+    V <- out[1,]                        # The new value-to-go
+    D[,OptTime-time+1] <- out[2,]       # The index positions
 ```
 
-### Background on Gaussian Process inference
+_Currently this shows the literal R code, should be adapted_
 

### 2013-03-29 inst/doc/nonparametric-bayes.tex
automate bibliographic extraction

@@
   \author[cstar]{Carl Boettiger\corref{cor1}}
+  \cortext[cor1]{Corresponding author}
+  \ead{cboettig@ucsc.edu}
   \author[cstar]{Marc Mangel}
   \author[noaa]{Stephan B. Munch}
-  \ead{cboettig@ucsc.edu}
-  \cortext[cor1]{Corresponding author, cboettig@ucsc.edu}
   \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA 95064, USA}
@@
 that a vaccine does not have to be 100\% effective to eliminate the
-transmission of a virus {[}@Kermack1921{]} -- such approaches are not
-well-suited for use in forecasting future dynamics upon which outcomes
-management policy can be based. Despite a long history of work
-emphasizing the difference between modeling for understanding (generic
-or strategic models) and modeling for prediction (precise, specific, or
-tactical models) (e.g. Levins 1966), simple parametric models continue
-to find application in predictive management contexts. The management of
-marine fisheries is a prime example, in which simple, mechanistically
-motivated models of stock-recruitment relationships such as Ricker or
-Beverton-Holt curves are systematically fit to data and used in solving
-decision-theoretic or optimal control problems determining resource
-management policies such as maximum sustainable yield {[}@refs{]}.
+transmission of a virus -- such approaches are not well-suited for use
+in forecasting outcomes of potential management options. Non-parametric
+approaches offer a more flexible alternative that can both more
+accurately reflect the data available while also representing greater
+uncertainty in areas (of state-space) where data is lacking. Ecological
+research and management strategy should pay closer attention to the
+opportunities and challenges nonparametric modeling can offer.
 
-Stochastic dynamic programming @Mangel1982
+\subsubsection{Quantitative vs Qualitative Decisions}
 
-To address structural uncertainty in the processes involved, such
-mechanistic models can be adapted to accomidate uncertainty in the
-parameters or consider a set of alternative models simultaneously
-(belief SDP)
+In this paper, we consider those ecological management problems in which
+a mathematical (or computational) model is used to quantitatively inform
+decision-making by allowing a manager to compare to the expected
+consequences of potential management actions (or policies). We
+distinguish this from the solely qualitative use of a model, in which
+models are used to represent and compare hypotheses of different
+mechanisms that would lead to qualitatively different actions.
 
-Further unknowns such as measurement uncertainty, parameter uncertainty,
+We argue that while simple mechanistically motivated models may be best
+for the latter case (Cuddington, Fortin, and Gerber 2013; Geritz and
+Kisdi 2012), such models can be not only inaccurate but misleading in
+quantitative decision making. Nonparametric models can more
+realistically represent uncertainties outside of observed data while
+also better capturing the dynamics in the region observed. Advances in
+the theory and computational implementations of nonparametric methods
+make them ripe for such applications. We use the classic problem of
+optimal harvest of a marine fishery to illustrate how the nonparametric
+approach of Gaussian processes can be applied.
+
+We imagine the manager seeks to maximize some known socio-economic value
+function, given their knowledge of the system. In particular, we will
+focus on the case in which the manager faces a series of decisions, such
+as setting a harvest quota for a marine fishery each year based on
+observations of the stock size the year before.
+
+\subsubsection{Uncertainties: the classical approach}
+
+The ecological dynamics of most systems of management interest are
+typically both complex and unknown {[}@refs{]}. Despite this,
+quantitative policy optimization in these systems has almost exclusively
+been based on simple, mechanistically motivated parametric models. There
+are several reasons for this. First, limited available data almost
+always percludes any accurate estimate of more complicated models.
+Second, the computational demands of determining an optimal stategy in a
+sequential decision-theory problem suffer from the ``curse of
+dimensionality'' that makes\\it difficult to consider more complicated
+models, states, or actions. Unfortunately, these challenges also make it
+particularly difficult to reflect the true uncertainty either in the
+parameters of the given models (parametric uncertainty) or in the
+structure of the models themselves (structural uncertainty).\\Further
+unknowns such as measurement uncertainty, parameter uncertainty,
 unobserved states, knowledge of boundary conditions, etc. further
@@
 
+An active literature and growing computational power over the past
+several decades have only marginally improved this situation. Parametric
+or structural uncertainty can be introduced only by increasing the state
+space to reflect a distribution of possible parameters {[}@refs{]} or
+the degree of belief in each of a set of possible models {[}@refs{]}
+Frequently, the space of possible actions must then be reduced or the
+algorithms adjusted by approximations to keep computations feasible.
+
 Though machine learning approaches have begun to appear in the
@@
 2005), they remain unfamiliar and untrusted approaches for most
-ecologists. Machine learning approaches represent an essentially
-pattern-based rather than process-based approach, raising the same
-skepticism from most theoretical ecologists that they hold for older
-correlative methods such as linear regression, while the complexity of
-the techniques has barred their adoption in more empirical audiences.
-
-A more immediate barrier to their adoption is the absence of a framework
-for applying machine learning approaches to resource management
-problems. Traditional approaches to optimal control (Pontryagin's
-principle, stochastic dynamic programming) rely on knowledge of the
-state equation, usually described by a simple parametric model. Here we
-illustrate how a stochastic dynamic programming algorithm can
-alternatively be driven by the predictions from a Gaussian process -- a
-machine learning approximation to the state dynamics.\\
+ecologists. One potential barrier to their adoption is the absence of a
+framework for applying machine learning approaches to resource
+management problems. Traditional approaches to optimal control
+(Pontryagin's principle, stochastic dynamic programming) rely on
+knowledge of the state equation, usually described by a simple
+parametric model. Here we illustrate how a stochastic dynamic
+programming algorithm can alternatively be driven by the predictions
+from a Gaussian process -- a machine learning approximation to the state
+dynamics.
 
@@
 
+\subsubsection{The optimal control problem in fisheries management}
+
+In our example, we focus on the problem in which a manager must set the
+harvest level for a marine fishery each year to maximize the net present
+value of the resource, given an estimated stock size from the year
+before. Rich data and global concerns have made marine fisheries the
+crucible for much of the founding work (Gordon 1954; Reed 1979; May et
+al. 1979; Ludwig and Walters 1982) in managing ecosystems under
+uncertainty. Global trends (Worm et al. 2006) and controversy (Hilborn
+2007; Worm et al. 2009) have made understanding these challenges all the
+more pressing.
+
+To permit comparisons against a theoretical optimum we will consider
+data on the stock dynamics simulated from a simple parametric model in
+which recruitment of the fish stock $X_{t+1}$ in the following year is a
+stochastic process governed by a function $f$ of the current stock
+$X_t$, selected harvest policy $h_t$, and noise process $Z$,
+
+\[X_{t+1} = Z_t f(X_t, h_t) \]
+
+Given parameters for the function $f$ and probability distribution $Z$,
+along with a given economic model determining the price/profit
+$\Pi(X_t, h_t)$ realized in a given year given a choice of harvest $h_t$
+and observed stock $X_t$. This problem can be solved exactly for
+discretized values of stock $X$ and policy $h$ using stochastic dynamic
+programming (SDP) {[}@Mangel1982{]}.\\Problems of this sort underpin
+much marine fisheries management today.
+
+A crux of this approach is correctly specifying the functional form of
+$f$, along with its parameters. The standard approach {[}@refs{]} uses
+one of a handful of common parametric models representing the
+stock-recruitment relationship, usually after estimating the model
+parameters from any available existing data. Though modifications of
+this approach mentioned in the introduction can permit additional
+sources of uncertainty such as measurement error in the stock
+assessment, implementation errors in the harvest policy, (Clark and
+Kirkwood 1986, , ) uncertainty in parameters (Mangel and Clark 1985,
+@Schaupaugh2013) or model structure (B. K. Williams 2001, ), we focus on
+the simplest example to illustrate the comparison, where only a single
+form of uncertainty (the growth shocks $Z_t$) is present in the
+underlying model, making additional modes unnecessary.
+
+We compare this approach to our alternative that uses a Gaussian Process
+(GP) in place of the a given stock recruitment curve. We briefly revisit
+a description of the Gaussian Process approach that has appeared in
+recent literature before describing how we take the important step of
+adapting the GP to the decision theory problem that can still be solved
+using SDP.
+
+\subsubsection{The Non-parametric Bayesian alternative for
+stock-recruitment curves}
+
+The use of Gaussian process (GP) regression (known as Kreging in the
+geospatial literature) to formulate a predictive model is relatively new
+in the context of modeling dynamical systems (Kocijan et al. 2005) and
+introduced in the ecological modeling and fisheries management by Munch
+et al. (2005). An accessible and thorough introduction to the
+formulation and use of GPs can be found in C. E. Rasmussen and Williams
+(2006).
+
+The essense of the GP approach can be captured in the following thought
+experiment: An exhaustive parametric approach to the challenge of
+structural uncertainty might proceed by writing down all possible
+functional forms for the underlying dynamical system with all possible
+parameter values for each form, and then consider searching over this
+huge space to select the most likely model and parameters; or using a
+Bayesian approach, assign priors to each of these possible models and
+infer the posterior distribution of possible models. The GP approach can
+be thought of as a computationally efficient approximation to this
+approach. GPs represent a large class of models that can be though of as
+capturing or reasonably approximating the set of models in this
+collection. By modeling at the level of the process, rather than the
+level of parametric equation, we can more concisely capture the possible
+behavior of these curves. In place of a parametric model of the
+dynamical system, the GP approach postulates a prior distribution of
+(n-dimensional) curves that can be though of as approximations to a
+range of possible (parametric) models that might describe the data. The
+GP allows us to consider a set of possible curves simultaneously.\\
+
 \subsubsection{Background on Gaussian Process inference}
 
-The use of Gaussian process regression (known as Kreging in the
-geospatial literature) to formulate a predictive model is relatively new
-in the context of modeling dynamical systems {[}@Kocijan2005{]} and
-introduced in the ecological modeling and fisheries management by Munch
-et al. (2005). An accessible and thorough introduction to the
-formulation and use of Gaussian processes can be found in
-@Rasmussen2006.
+Once the posterior Gaussian process (GP) has been estimated (e.g.~see
+Munch et al. 2005), it is necessary to adapt it in place of the
+parametric equation for the stochastic dynamic programming (SDP)
+solution (see Mangel and Clark 1988 for a detailed description of
+parametric SDP methods) to the optimal policy. The essense of the idea
+is straight forward -- we will use the estimated GP in place of the
+parametric growth function to determine the stochastic transition matrix
+on which the SDP calculations are based.
 
-The essense of the Gaussian process approach can be captured in the
-following thought experiment: An exhaustive parametric approach to the
-challenge of structural uncertainty might proceed by writing down all
-possible functional forms for the underlying dynamical system with all
-possible parameter values for each form, and then consider searching
-over this huge space to select the most likely model and parameters; or
-using a Bayesian approach, assign priors to each of these possible
-models and infer the posterior distribution of possible models. The
-Gaussian process approach can be thought of as a computationally
-efficient approximation to this approach. Gaussian processes represent a
-large class of models that can be though of as capturing or reasonably
-approximating the set of models in this collection. By modeling at the
-level of the process, rather than the level of parametric equation, we
-can more concisely capture the possible behavior of these curves. In
-place of a parametric model of the dynamical system, the Gaussian
-Process approach postulates a prior distribution of (n-dimensional)
-curves that can be though of as approximations to a range of possible
-(parametric) models that might describe the data. The GP allows us to
-consider a set of possible curves simultaneously.\\
+The posterior Gaussian process is completely defined by the expected
+value and covariance matrix at a defined set of training points. For
+simplicty we will consider a these points to fall on a discrete, uniform
+grid $x$ of 101 points from 0 to 15 (1.5 times the positive equilibrium
+$K$). Again to keep things simple we will use this same grid
+discritization for the parametric approach. Other options for choosing
+the grid points, including collocation methods and functional basis
+expansion (or even using Guassian processes in place of the discrete
+optimization; an entirely different context in which GP can be used in
+SDP, see (Deisenroth, Rasmussen, and Peters 2009)) could also be
+considered.
+
+The transition matrix $\mathbf{F}$ is thus an 101 by 101 matrix for
+which the ${i,j}$ entry gives the probability of transitioning into
+state $x_i$ given that the system is in state $x_j$ in the previous
+timestep. To generate the transition matrix based on the posterior GP,
+we need only the expected values at each grid point and the
+corresponding variances (the diagonal of the covariance matrix), as
+shown in Figure 1. Given the mean at each gridpoint as the length 101
+vector $E$ and variance $V$, the probability of transitioning from state
+$x_i$ to state $x_j$ is simply
+$\mathcal{N}\left(x_j \vert  \mu = E_i, \sigma = \sqrt{V_i}\right)$,
+where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
+variance $\sigma^2$. Strictly speaking, the transition probability
+should be calculated by integrating the normal density over the bin of
+width $\Delta$ centered at $x_j$. For a sufficiently fine grid that
+$f(x_j) \approx f(x_j + \Delta)$, it is sufficient to calculate the
+density at $x_j$ and then row-normalize the transition matrix.
+
+\subsection{Pseudocode for the determining the transtion matrix from the
+GP}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{for(h in h_grid)}
+  \NormalTok{F_h =}\StringTok{ }\NormalTok{for(x_j in grid)}
+          \NormalTok{for(i in }\DecValTok{1}\NormalTok{:N) }
+            \KeywordTok{dnorm}\NormalTok{(x_j, mu[i]-h, V[i])}
+\end{Highlighting}
+\end{Shaded}
+
+A transition matrix for each of the parametric models $f$ is calculated
+using the log-normal density with mean $f(x_i)$ and log-variance as
+estimated by maximum likelihood. From the discrete transition matrix we
+may write down the Bellman recursion defining the the stochastic dynamic
+programming iteration:
+
+\begin{equation}
+V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
+\end{equation}
+
+where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$ is
+control (harvest level) chosen. Numerically, the maximization is
+accomplished as follows. Consider the set of possible control values to
+be the discrete 101 values corresponding the the grid of stock sizes.
+Then for each $h_t$ there is a corresponding transition matrix
+$\mathbf{F}_h$ determined as described above but with mean
+$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element
+corresponds to the value of having stock $x_i$ at time $t$. Then let
+$\Pi_h$ be the vector whose $i$th element indicates the profit from
+harvesting at intensity $h_t$ given a population $x_i$ (e.g.
+$\max(x_i, h_t)$ since one cannot harvest more fish then the current
+population size). Then the Bellman recursion can be given in matrix form
+as
+
+\[V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)\]
+
+where the sum is element by element and the expectation is computed by
+the matrix multiplication $\mathbf{F} V_{t+1}$.
+
+\subsubsection{Pseudocode for the Bellman iteration}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+ \NormalTok{V1 <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
+      \NormalTok{delta *}\StringTok{ }\NormalTok{F[[h]] %*%}\StringTok{ }\NormalTok{V +}\StringTok{  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
+    \NormalTok{\})}
+    \CommentTok{# find havest, h that gives the maximum value}
+    \NormalTok{out <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
+      \NormalTok{value <-}\StringTok{ }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
+      \NormalTok{index <-}\StringTok{ }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
+      \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
+    \NormalTok{\})}
+    \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
+    \NormalTok{V <-}\StringTok{ }\NormalTok{out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
+    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <-}\StringTok{ }\NormalTok{out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
+\end{Highlighting}
+\end{Shaded}
+
+\emph{Currently this shows the literal R code, should be adapted}
 
@@
 been observing for as long as they made models (perhaps none more
-memorably than @Levins1969). Mechanistic modeling is at its most
+memorably than Levins (1969)). Mechanistic modeling is at its most
 powerful not when it is used to directly forecast future states but when
@@
 optimization; an entirely different context in which GP can be used in
-SDP, see {[}@Deisenroth2009{]}) could also be considered.
+SDP, see (Deisenroth, Rasmussen, and Peters 2009)) could also be
+considered.
 
@@
 
-@Gramacy2005
+Gramacy (2005)
 
@@
 doi:10.1080/10236190412331335373.
+\url{http://www.tandfonline.com/doi/abs/10.1080/10236190412331335373}.
+
+Athanassoglou, Stergios, and Anastasios Xepapadeas. 2012. ``Pollution
+control with uncertain stock dynamics: When, and how, to be
+precautious.'' \emph{Journal of Environmental Economics and Management}
+63 (3) (may): 304--320. doi:10.1016/j.jeem.2011.11.001.
+\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069611001409}.
 
@@
 (jan): 1--14. doi:10.1016/j.ecolecon.2010.10.001.
+\url{http://linkinghub.elsevier.com/retrieve/pii/S0921800910004167}.
+
+Clark, Colin W., and Geoffrey P. Kirkwood. 1986. ``On uncertain
+renewable resource stocks: Optimal harvest policies and the value of
+stock surveys.'' \emph{Journal of Environmental Economics and
+Management} 13 (3) (sep): 235--244. doi:10.1016/0095-0696(86)90024-0.
+\url{http://linkinghub.elsevier.com/retrieve/pii/0095069686900240}.
 
@@
 Effects in Ecology and Conservation}. Oxford University Press, USA.
+\url{http://www.amazon.com/Effects-Ecology-Conservation-Franck-Courchamp/dp/0198570309}.
 
@@
 modeling.'' \emph{Ecological Applications} 19 (3) (apr): 553--70.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/19425416}.
 
-Levins, Richard. 1966. ``The strategy of model building in population
-biology.'' \emph{American Scientist} 54 (4): 421--431.
+Cuddington, Kim M., M. J. Fortin, and L. R. Gerber. 2013.
+``Process-based models are required to manage ecological systems in a
+changing world.'' \emph{Ecosphere} 4 (February): 1--12.
+doi:10.1890/ES12-00178.1.
+\url{http://www.esajournals.org/doi/abs/10.1890/ES12-00178.1}.
 
-Mangel, Marc, and Colin W. Clark. 1988. \emph{Dynamic Modeling in
-Behavioral Ecology}. Princeton: Princeton University Press.
+Deisenroth, Marc Peter, Carl Edward Rasmussen, and Jan Peters. 2009.
+``Gaussian process dynamic programming.'' \emph{Neurocomputing} 72 (7-9)
+(mar): 1508--1524. doi:10.1016/j.neucom.2008.12.019.
+\url{http://linkinghub.elsevier.com/retrieve/pii/S0925231209000162}.
+
+Geritz, Stefan a. H., and Eva Kisdi. 2012. ``Mathematical ecology: why
+mechanistic models?'' \emph{Journal of mathematical biology} 65 (6-7)
+(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.
+
+Gordon, H. S. 1954. ``The economic theory of a common-property resource:
+the fishery.'' \emph{The Journal of Political Economy} 62 (2): 124--142.
+\url{http://www.jstor.org/stable/10.2307/1825571}.
+
+Gramacy, Robert B. 2005. ``BAYESIAN TREED GAUSSIAN PROCESS MODELS.''
+
+Hilborn, Ray. 2007. ``Reinterpreting the State of Fisheries and their
+Management.'' \emph{Ecosystems} 10 (8) (oct): 1362--1369.
+doi:10.1007/s10021-007-9100-5.
+\url{http://www.springerlink.com/index/10.1007/s10021-007-9100-5}.
+
+Kocijan, Juš, Agathe Girard, Bla\textbackslash{}vz Banko, and Roderick
+Murray-Smith. 2005. ``Dynamic systems identification with Gaussian
+processes.'' \emph{Mathematical and Computer Modelling of Dynamical
+Systems} 11 (4) (dec): 411--424. doi:10.1080/13873950500068567.
+\url{http://www.tandfonline.com/doi/abs/10.1080/13873950500068567}.
+
+Levins, Richard. 1969. ``Some demographic and genetic consequences of
+environmental heterogeneity for biological control.'' \emph{Bulletin of
+the Entomological Society of America} 15 (3): 237--240.
+\url{http://www.ingentaconnect.com/content/esa/besa/1969/00000015/00000003/art00009}.
+
+Ludwig, Donald, and Carl J. Walters. 1982. ``Optimal harvesting with
+imprecise parameter estimates.'' \emph{Ecological Modelling} 14 (3-4)
+(jan): 273--292. doi:10.1016/0304-3800(82)90023-0.
+\url{http://linkinghub.elsevier.com/retrieve/pii/0304380082900230}.
+
+Mangel, Marc, and Colin W. Clark. 1985. \emph{Decision and control in
+uncertain resource systems, Volume 172 (Mathematics in Science and
+Engineering)}. Academic Press.
+\url{http://www.amazon.com/Decision-uncertain-resource-Mathematics-Engineering/dp/0124687202}.
+
+---------. 1988. \emph{Dynamic Modeling in Behavioral Ecology}.
+Princeton: Princeton University Press.
+
+May, Robert M., John R. Beddington, Colin W. Clark, Sidney J. Holt, and
+R. M. Laws. 1979. ``Management of multispecies fisheries.''
+\emph{Science (New York, N.Y.)} 205 (4403) (jul): 267--77.
+doi:10.1126/science.205.4403.267.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/17747032}.
 
@@
 691--695. doi:10.1111/j.1461-0248.2005.00766.x.
+\url{http://doi.wiley.com/10.1111/j.1461-0248.2005.00766.x}.
+
+Rasmussen, C. E., and C. K. I. Williams. 2006. \emph{Gaussian Processes
+for Machine Learning}. Ed. Thomas Dietterich. Boston: MIT Press,.
+\url{www.GaussianProcess.org/gpml}.
 
@@
 doi:10.1016/0095-0696(79)90014-7.
+\href{http://www.sciencedirect.com/science/article/pii/0095069679900147 http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}{http://www.sciencedirect.com/science/article/pii/0095069679900147
+http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}.
+
+Roughgarden, Joan E., and F. Smith. 1996. ``Why fisheries collapse and
+what to do about it.'' \emph{Proceedings of the National Academy of
+Sciences of the United States of America} 93 (10): 5078.
+\url{http://www.pnas.org/content/93/10/5078.short}.
 
@@
 (6856) (oct): 591--6. doi:10.1038/35098000.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/11595939}.
+
+Sethi, Gautam, Christopher Costello, Anthony Fisher, Michael Hanemann,
+and Larry Karp. 2005. ``Fishery management under multiple uncertainty.''
+\emph{Journal of Environmental Economics and Management} 50 (2) (sep):
+300--318. doi:10.1016/j.jeem.2004.11.005.
+\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069605000057}.
+
+Williams, Byron K. 2001. ``Uncertainty , learning , and the optimal
+management of wildlife.'' \emph{Environmental and Ecological Statistics}
+8: 269--288. doi:10.1023/A:1011395725123.
+
+Worm, Boris, Edward B. Barbier, Nicola Beaumont, J. Emmett Duffy, Carl
+Folke, Benjamin S. Halpern, Jeremy B. C. Jackson, et al. 2006. ``Impacts
+of biodiversity loss on ocean ecosystem services.'' \emph{Science (New
+York, N.Y.)} 314 (5800) (nov): 787--90. doi:10.1126/science.1132294.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/17082450}.
+
+Worm, Boris, Ray Hilborn, Julia K. Baum, Trevor A. Branch, Jeremy S.
+Collie, Christopher Costello, Michael J. Fogarty, et al. 2009.
+``Rebuilding global fisheries.'' \emph{Science (New York, N.Y.)} 325
+(5940) (jul): 578--85. doi:10.1126/science.1173146.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/19644114}.
 

### 2013-04-01 inst/doc/nonparametric-bayes.Rmd
build drafts, update html support

@@
 The ecological dynamics of most systems of management interest are
-typically both complex and unknown [@refs].  Despite this, quantitative policy
-optimization in these systems has almost exclusively been based on
-simple, mechanistically motivated parametric models.
-There are several reasons for this.  First, limited available data almost 
-always percludes any accurate estimate of more complicated models.  Second,
-the computational demands of determining an optimal stategy in a sequential
-decision-theory problem suffer from the "curse of dimensionality" that makes  
-it difficult to consider more complicated models, states, or actions.  Unfortunately,
-these challenges also make it particularly difficult to reflect the true uncertainty
-either in the parameters of the given models (parametric uncertainty) or in 
-the structure of the models themselves (structural uncertainty).  
-Further unknowns such as measurement uncertainty, parameter uncertainty,
-unobserved states, knowledge of boundary conditions, etc. further
-compound the issue.  Though a hierarchical Bayesian approach provides a
-natural way to address these from a statistical standpoint, formulating
-reasonable parametric  descriptions of each form of uncertainty is a
-challenging task in itself, let alone the computational difficulties
-of solving such a system. @Cressie2009 provides a good account of the
-successess and challenges of the approach.  Applying these approaches in
-the management context of sequential decision making, in which forecasts
-must be obtained over a range of possible actions and updated regularly
-as new information arrives makes such an approach less feasible still.
+typically both complex and unknown [@refs].  Despite this, quantitative
+policy optimization in these systems has almost exclusively been based
+on simple, mechanistically motivated parametric models.  There are
+several reasons for this.  First, limited available data almost always
+percludes any accurate estimate of more complicated models.  Second, the
+computational demands of determining an optimal stategy in a sequential
+decision-theory problem suffer from the "curse of dimensionality"
+that makes it difficult to consider more complicated models, states,
+or actions.  Unfortunately, these challenges also make it particularly
+difficult to reflect the true uncertainty either in the parameters of
+the given models (parametric uncertainty) or in the structure of the
+models themselves (structural uncertainty).  Further unknowns such as
+measurement uncertainty, parameter uncertainty, unobserved states,
+knowledge of boundary conditions, etc. further compound the issue.
+Though a hierarchical Bayesian approach provides a natural way to
+address these from a statistical standpoint, formulating reasonable
+parametric  descriptions of each form of uncertainty is a challenging
+task in itself, let alone the computational difficulties of solving such
+a system. @Cressie2009 provides a good account of the successess and
+challenges of the approach.  Applying these approaches in the management
+context of sequential decision making, in which forecasts must be obtained
+over a range of possible actions and updated regularly as new information
+arrives makes such an approach less feasible still.
 
@@
 
+<span style="color:#007FFF;"> ***This part is overstated and unnecessary***
 Though machine learning approaches have begun to appear in the
-ecological and conservation literature (Species distribution models),
+ecological and conservation literature (e.g. in species distribution models),
 including the Gaussian process based approach used here [@Munch2005],
-they remain unfamiliar and untrusted approaches for most ecologists.
+overall machine learning remains an unfamiliar and somewhat untrusted approach for many ecologists.
 One potential barrier to their adoption is the absence of a framework
-for applying machine learning approaches to resource management problems.
+for applying machine learning approaches to resource management problems. </span>
+
+<span style="color:#007FFF;">***transition sentences!?***</span>
+
 Traditional approaches to optimal control (Pontryagin's principle, stochastic
@@
 
-<!-- Okay, perhaps that's novel, but it's pretty trivial.  Isn't it
-obvious to everyon that it's trivial? -->
+<span style="color:#007FFF;">***Okay, perhaps that's novel, but it's pretty trivial.  Isn't it
+obvious to everyon that it's trivial?***</span>
 
-<!-- paraphrase advantages of machine learning -->
+<span style="color:#007FFF;">***paraphrase advantages of machine learning***</span>
 
@@
 This problem can be solved exactly for discretized values of stock $X$ and policy $h$
-using stochastic dynamic programming (SDP) [@Mangel1982].  
-Problems of this sort underpin much marine fisheries management today.
+using stochastic dynamic programming (SDP) [@Mangel1985]. Problems of this sort underpin much marine fisheries management today.
 
@@
 introduction can permit additional sources of uncertainty such as measurement 
-error in the stock assessment, implementation errors in the harvest policy, [@Clark1986, @Roughgarden1996, @Sethi2005]
-uncertainty in parameters [@Mangel1985, @Schaupaugh2013] or model structure [@Williams2001, @Athanassoglou2012],
+error in the stock assessment, implementation errors in the harvest policy, [@Clark1986; @Roughgarden1996; @Sethi2005]
+uncertainty in parameters [@Mangel1985; @Schapaugh2013] or model structure [@Williams2001; @Athanassoglou2012],
 we focus on the simplest example to illustrate the comparison, where only a 
@@
 model, making additional modes unnecessary. 
-<!--Add comment / apology on single species dynamics -->
+<span style="color:#007FFF;">***Add comment / apology on single species dynamics?*** </span>
 
@@
 
-The use of Gaussian process (GP) regression (known as Kreging in the geospatial
+The use of Gaussian process (GP) regression (or "kreging" in the geospatial
 literature) to formulate a predictive model is relatively new in the
@@
 to consider a set of possible curves simultaneously.  
-<!-- Figure 1 include curves drawn from the posterior density?  -->
+<span style="color:#007FFF;"> ***Figure 1 include curves drawn from the posterior density?*** </span>
 
-<!-- Do we need more specifics on Gaussian process as an approximation
+<span style="color:#007FFF;">***Do we need more specifics on Gaussian process as an approximation
 to parametric models? Discussion of Gaussian process vs other machine
 learning / forecasting approaches that have less clear statistical
-foundations?  If so, does this all belong in the discussion? -->
+foundations?  If so, does this all belong in the discussion?***</span>
 
@@
 
-<!--Should we include any emprical examples? -->
+<span style="color:#007FFF;">***Should we include any emprical examples?***</span>
 Using data simulated from a specified model rather than empirical data
@@
 
-<!-- wow, run-on tangent? -->
+<span style="color:#007FFF;"> wow, run-on tangent? </span>
 Predictive accuracy is not the goal of all modeling, as ecologists
@@
 
-<!-- Jargony and vague -->
+<span style="color:#007FFF;"> Jargony and vague </span>
 While we have highlighted certain generic of this problem that allow the
@@
 
-<!-- Model complexity -->
+<span style="color:#007FFF;"> Model complexity </span>
 In this simulated example, the underlying
@@
 
-<!-- Data complexity. Perhaps too far out of scope... -->
+<span style="color:#007FFF;"> Data complexity. Perhaps too far out of scope... </span>
 The machine learning approach is also better suited to complex and
@@
 
-<!-- 
+<span style="color:#007FFF;"> 
 * Discuss constant escapement in model, in policies.
@@
 * Discussion of alternative related approaches: POMDP/MOMDP
--->
+</span>
 

### 2013-04-01 inst/doc/nonparametric-bayes.tex
build drafts, update html support

@@
 sequential decision-theory problem suffer from the ``curse of
-dimensionality'' that makes\\it difficult to consider more complicated
+dimensionality'' that makes it difficult to consider more complicated
 models, states, or actions. Unfortunately, these challenges also make it
@@
 parameters of the given models (parametric uncertainty) or in the
-structure of the models themselves (structural uncertainty).\\Further
+structure of the models themselves (structural uncertainty). Further
 unknowns such as measurement uncertainty, parameter uncertainty,
@@
 
-Though machine learning approaches have begun to appear in the
-ecological and conservation literature (Species distribution models),
-including the Gaussian process based approach used here (Munch et al.
-2005), they remain unfamiliar and untrusted approaches for most
-ecologists. One potential barrier to their adoption is the absence of a
-framework for applying machine learning approaches to resource
-management problems. Traditional approaches to optimal control
-(Pontryagin's principle, stochastic dynamic programming) rely on
-knowledge of the state equation, usually described by a simple
-parametric model. Here we illustrate how a stochastic dynamic
-programming algorithm can alternatively be driven by the predictions
-from a Gaussian process -- a machine learning approximation to the state
-dynamics.
+ \textbf{This part is overstated and unnecessary} Though machine
+learning approaches have begun to appear in the ecological and
+conservation literature (e.g.~in species distribution models), including
+the Gaussian process based approach used here (Munch et al. 2005),
+overall machine learning remains an unfamiliar and somewhat untrusted
+approach for many ecologists. One potential barrier to their adoption is
+the absence of a framework for applying machine learning approaches to
+resource management problems.
+
+ transition sentences!?
+
+Traditional approaches to optimal control (Pontryagin's principle,
+stochastic dynamic programming) rely on knowledge of the state equation,
+usually described by a simple parametric model. Here we illustrate how a
+stochastic dynamic programming algorithm can alternatively be driven by
+the predictions from a Gaussian process -- a machine learning
+approximation to the state dynamics.
+
+ Okay, perhaps that's novel, but it's pretty trivial. Isn't it obvious
+to everyon that it's trivial?
+
+ paraphrase advantages of machine learning
 
@@
 discretized values of stock $X$ and policy $h$ using stochastic dynamic
-programming (SDP) {[}@Mangel1982{]}.\\Problems of this sort underpin
-much marine fisheries management today.
+programming (SDP) (Mangel and Clark 1985). Problems of this sort
+underpin much marine fisheries management today.
 
@@
 assessment, implementation errors in the harvest policy, (Clark and
-Kirkwood 1986, , ) uncertainty in parameters (Mangel and Clark 1985,
-@Schaupaugh2013) or model structure (B. K. Williams 2001, ), we focus on
-the simplest example to illustrate the comparison, where only a single
-form of uncertainty (the growth shocks $Z_t$) is present in the
-underlying model, making additional modes unnecessary.
+Kirkwood 1986; Roughgarden and Smith 1996; Sethi et al. 2005)
+uncertainty in parameters (Mangel and Clark 1985; Schapaugh and Tyre
+2013) or model structure (B. K. Williams 2001; Athanassoglou and
+Xepapadeas 2012), we focus on the simplest example to illustrate the
+comparison, where only a single form of uncertainty (the growth shocks
+$Z_t$) is present in the underlying model, making additional modes
+unnecessary. Add comment / apology on single species dynamics
 
@@
 
-The use of Gaussian process (GP) regression (known as Kreging in the
+The use of Gaussian process (GP) regression (or ``kreging'' in the
 geospatial literature) to formulate a predictive model is relatively new
@@
 GP allows us to consider a set of possible curves simultaneously.\\
+Figure 1 include curves drawn from the posterior density?
+
+ Do we need more specifics on Gaussian process as an approximation to
+parametric models? Discussion of Gaussian process vs other machine
+learning / forecasting approaches that have less clear statistical
+foundations? If so, does this all belong in the discussion?
 
@@
 
-Using data simulated from a specified model rather than empirical data
-permits the comparison against the true underlying dynamics, setting a
-bar for the optimal performance possible.
+Should we include any emprical examples? Using data simulated from a
+specified model rather than empirical data permits the comparison
+against the true underlying dynamics, setting a bar for the optimal
+performance possible.
 
@@
 
-Predictive accuracy is not the goal of all modeling, as ecologists have
-been observing for as long as they made models (perhaps none more
-memorably than Levins (1969)). Mechanistic modeling is at its most
-powerful not when it is used to directly forecast future states but when
-it provides an understanding of how to approach a problem. SIR-type
-models from the epidemiological literature are a good example. While the
-simplest SIR models have little predictive power over the outbreak
-intensity or timing at a particular location, they provide a powerful
-understanding of the spread of an infection in terms of a single,
-biologically meaningful parameter: $R_0$, the basic reproductive number.
-From the model, it becomes clear that management need not vaccinate
-every member of the population to stop the spread, but rather it
-suffices to vaccinate a sufficient fraction of the population to reduce
-$R_0$ below 1.
+ wow, run-on tangent? Predictive accuracy is not the goal of all
+modeling, as ecologists have been observing for as long as they made
+models (perhaps none more memorably than Levins (1969)). Mechanistic
+modeling is at its most powerful not when it is used to directly
+forecast future states but when it provides an understanding of how to
+approach a problem. SIR-type models from the epidemiological literature
+are a good example. While the simplest SIR models have little predictive
+power over the outbreak intensity or timing at a particular location,
+they provide a powerful understanding of the spread of an infection in
+terms of a single, biologically meaningful parameter: $R_0$, the basic
+reproductive number. From the model, it becomes clear that management
+need not vaccinate every member of the population to stop the spread,
+but rather it suffices to vaccinate a sufficient fraction of the
+population to reduce $R_0$ below 1.
 
@@
 
-While we have highlighted certain generic of this problem that allow the
-nonparametric approach to excel -- short timescales between successive
-observations and actions, the accuracy in the appropriate region of
-state space, the ability to express uncertainty outside the observed
-data -- there are equally several aspects for which the nonparametric
-approach is at a relative disadvantage.
-
-In this simulated example, the underlying dynamics are truly governed by
-a simple parametric model, allowing the parametric approaches to be more
-accurate. Similarly, because the dynamics are one-dimensional dynamics
-and lead to stable nodes (rather than other attractors such as
-limit-cycles resulting in oscillations), the training data provides
-relatively limited information about the dynamics. For these reasons, we
-anticipate that in higher-dimensional examples characteristic of
-ecosystem management problems that the machine learning approach will
-prove even more valuable.
-
-The machine learning approach is also better suited to complex and
-disparate data. Incorporating various sources of information into
-mechanistic models can be an immensely difficult due to the increased
-complexity involved. Only thanks to tandem advances in increasing
-computational power and hierarchical statistical methodology have we
-been able to tackle such intuitively important complexity (and the
-potentially new available data that accompanies it) such as spatial
-distribution, heterogeneities of space, time, and individuals, to shift
-to ecosystem-based approaches from single-species based approaches.
-Without the need to formulate mechanisms, many modern machine learning
-algorithms can leverage potential information from all available sources
-of data directly. The algorithms can recognize unanticipated or subtle
-patterns in large data sets that enable more accurate predictions than
-mechanistic models that are formulated at a more macroscopic level.
+ Jargony and vague While we have highlighted certain generic of this
+problem that allow the nonparametric approach to excel -- short
+timescales between successive observations and actions, the accuracy in
+the appropriate region of state space, the ability to express
+uncertainty outside the observed data -- there are equally several
+aspects for which the nonparametric approach is at a relative
+disadvantage.
+
+ Model complexity In this simulated example, the underlying dynamics are
+truly governed by a simple parametric model, allowing the parametric
+approaches to be more accurate. Similarly, because the dynamics are
+one-dimensional dynamics and lead to stable nodes (rather than other
+attractors such as limit-cycles resulting in oscillations), the training
+data provides relatively limited information about the dynamics. For
+these reasons, we anticipate that in higher-dimensional examples
+characteristic of ecosystem management problems that the machine
+learning approach will prove even more valuable.
+
+ Data complexity. Perhaps too far out of scope\ldots{} The machine
+learning approach is also better suited to complex and disparate data.
+Incorporating various sources of information into mechanistic models can
+be an immensely difficult due to the increased complexity involved. Only
+thanks to tandem advances in increasing computational power and
+hierarchical statistical methodology have we been able to tackle such
+intuitively important complexity (and the potentially new available data
+that accompanies it) such as spatial distribution, heterogeneities of
+space, time, and individuals, to shift to ecosystem-based approaches
+from single-species based approaches. Without the need to formulate
+mechanisms, many modern machine learning algorithms can leverage
+potential information from all available sources of data directly. The
+algorithms can recognize unanticipated or subtle patterns in large data
+sets that enable more accurate predictions than mechanistic models that
+are formulated at a more macroscopic level.
+
+ * Discuss constant escapement in model, in policies.
+
+\begin{itemize}
+\item
+  Limitations of this comparison: Are the maximum-likelihood solutions a
+  straw man?
+\item
+  Discussion of alternative related approaches: POMDP/MOMDP 
+\end{itemize}
 
@@
 
+Schapaugh, Adam W., and Andrew J. Tyre. 2013. ``Accounting for
+parametric uncertainty in Markov decision processes.'' \emph{Ecological
+Modelling} 254 (apr): 15--21. doi:10.1016/j.ecolmodel.2013.01.003.
+\url{http://linkinghub.elsevier.com/retrieve/pii/S0304380013000306}.
+
 Scheffer, Marten, Stephen R. Carpenter, J. A. Foley, C. Folke, and B.

### 2013-04-02 inst/doc/nonparametric-bayes.Rmd
writing writing

@@
 
-```{r knit-settings, cache=FALSE}
-options(xtable.print.comment=FALSE)
-options(xtable.type = 'latex', table.placement="H")
-opts_chunk$set(warning=FALSE, message=FALSE, comment=NA, tidy=FALSE,
- echo=FALSE)
-opts_knit$set(progress = TRUE, verbose = TRUE)
-opts_chunk$set(dev = 'Cairo_pdf', fig.width=5.5, fig.height=4,
- cache.path = 'cache-pdf/', cache=TRUE)
-```
+```{r libraries, include=FALSE}
+## Dargh these are ignored if not placed here!?
+options(xtable.type = 'latex', table.placement="H", xtable.print.comment=FALSE)
+opts_chunk$set(cache.path = 'cache-pdf/', cache=TRUE)
+opts_chunk$set(dev = 'Cairo_pdf', fig.width=5.5, fig.height=4)
 
-```{r libraries, include=FALSE}
 library(pdgControl)
@@
 ```
+
 ```{r plotting-options, echo=FALSE, include=FALSE}
+# This stuff should be handled by knit script(?) (load ggplot2 first though)!
 theme_set(theme_bw(base_size=10))
@@
 Tobs <- 35
+```
+
+```{r obs, dependson="sdp-pars"}
+obs <- sim_obs(Xo, z_g, f, p, Tobs=Tobs, nz=1, 
+                harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
+```
+
+```{r mle, dependson="obs"}
+alt <- par_est(obs,  init = c(r = p[1], 
+                              K = mean(obs$x[obs$x>0]), 
+                              s = sigma_g))
+est <- par_est_allee(obs, f, p,  
+                     init = c(r = p[1] + 1, 
+                              K = p[2] + 2, 
+                              C = p[3] + 2, 
+                              s = sigma_g))
+```
+
+```{r gp-priors}
+#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
+s2.p <- c(5,5)  
+tau2.p <- c(5,1)
+d.p = c(10, 1/0.1, 10, 1/0.1)
+nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
+s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
+tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
+d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
+nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
+beta0_prior <- function(x, tau) dnorm(x, 0, tau)
+beta = c(0)
+priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
+```
+
+```{r gp, dependson=c("obs", "gp-priors")}
+  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
+          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
+          m0r1=FALSE, corr="exp", trace=TRUE, 
+          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
+          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
+```
+
+```{r opt, dependson=c("gp", "mle")}
+  OPT <- optimal_policy(gp, f, est$f, alt$f,
+                        p, est$p, alt$p,
+                        x_grid, h_grid, sigma_g, 
+                        sigma_g, sigma_g, # est$sigma_g, alt$sigma_g, but those ests are poor
+                        delta, xT, profit, reward, OptTime)
+```
+
+```{r sim, dependson="opt"}
+dt <- simulate_opt(OPT, f, p, x_grid, h_grid, x0, z_g, profit)
 ```
@@
 the stochastic transition matrix on which the SDP calculations are based.
+The SDP is solved in a discritized state space -- both the continuously 
+valued population densities $X$ and harvest quotas $h$ are first mapped to
+a bounded, discrete grid.  (For simplicity we will consider a uniform grid,
+though for either parametric or GP-based SDP it is often advantageous to 
+use a non-uniform discretization such as a basis function representation,
+e.g. see [@Deisenroth2009]).
 
-The posterior Gaussian process is completely defined by the expected value
-and covariance matrix at a defined set of training points.  For simplicty
-we will consider a these points to fall on a discrete, uniform grid $x$
-of `r length(x_grid)` points from `r min(x_grid)` to `r max(x_grid)`
-(1.5 times the positive equilibrium $K$).  Again to keep things simple
-we will use this same grid discritization for the parametric approach.
-Other options for choosing the grid points, including collocation methods
-and functional basis expansion (or even using Guassian processes in place
-of the discrete optimization; an entirely different context in which GP
-can be used in SDP, see [@Deisenroth2009]) could also be considered.
-
-The transition matrix $\mathbf{F}$ is thus an `r length(x_grid)`
-by `r length(x_grid)` matrix for which the ${i,j}$ entry gives the
-probability of transitioning into state $x_i$ given that the system is
-in state $x_j$ in the previous timestep.  To generate the transition
+The SDP approach then computes a transition matrix, $\mathbf{F}$.  We 
+demonstrate that calculation is just as straight forward based on the GP
+as it is in the classical context using the parametric model.  The 
+${i,j}$ of the transition matrix $F$ entry gives the probability of transitioning into 
+state $x_i$ given that the system is in state $x_j$ in the previous 
+timestep.  To generate the transition
 matrix based on the posterior GP, we need only the expected values
 at each grid point and the corresponding variances (the diagonal of
-the covariance matrix), as shown in Figure 1.  Given the mean at each
-gridpoint as the length `r length(x_grid)` vector $E$ and variance $V$,
-the probability of transitioning from state $x_i$ to state $x_j$ is
-simply $\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)$,
+the covariance matrix), as shown in Figure 1.  Given the mean of the 
+GP posterior at each gridpoint as the vector $E$ and variance at that
+point as vector $V$, the probability of transitioning from state $x_i$ to state $x_j$ is
+
+$$\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)$$
+
 where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
@@
 \approx f(x_j + \Delta)$, it is sufficient to calculate the density at
-$x_j$ and then row-normalize the transition matrix.
+$x_j$ and then row-normalize the transition matrix. The process
+can then be repeated for each possible discrete value of our control 
+variable, (harvest $h$).
 
-## Pseudocode for the determining the transtion matrix from the GP
+**Pseudocode for the determining the transtion matrix from the GP**
 
@@
 
-A transition matrix for each of the parametric models $f$ is calculated
-using the log-normal density with mean $f(x_i)$ and log-variance as
-estimated by maximum likelihood.  From the discrete transition matrix we
-may write down the Bellman recursion defining the the stochastic dynamic
-programming iteration:
+Using the discrete transition matrix we may write down the Bellman
+recursion defining the the stochastic dynamic programming iteration:
 
@@
 accomplished as follows. Consider the set of possible control values to
-be the discrete `r length(h_grid)` values corresponding the the grid of
+be the discrete values corresponding the the grid of
 stock sizes.  Then for each $h_t$ there is a corresponding transition
@@
 
-_Currently this shows the literal R code, should be adapted_
-
-### The optimal control problem
-
-### Discussion on how we compare performance of policies
-
-* Replicate stochastic simulations 
-* Sensitivity analysis (Figure 4).
+This completes the algorithm adapting the GP to the sequential decision-making 
+problem through SDP, which we believe has not yet been demonstrated in the
+literature.  We provide an R package implementation of this, along with
+the Gaussian process inference, in the supplemental materials.
 
@@
 management and pose a substantial challenge to existing decision-theoretic
-approaches [@Brozovic2011].  To compare the performance of nonparametric
-and parametric approaches in an example that is easy to conceptualize,
-we will focus on a simple parametric model for a single species [derived
+approaches [@Brozovic2011].  Because intervention is often too late 
+after a tipping point has been crossed (but see @Hughes2013), management
+is most often concerned with avoiding potentially catostrophic tipping
+points before any data is available at or following a transition that
+would more clearly reveal these regime shift dyanmics [e.g. @Bestelmeyer2012].
+
+<!-- Explain the connection to alternative stable states -->
+
+To illustrate the value of the non-parametric Bayesian approach to management,
+we focus on example of a system containing such a tipping point whose dynamics
+can still be described by a simple, one-dimensional parametric model.  
+We will focus on a simple parametric model for a single species [derived
 from fist principles by @Allen2005a] as our underlying "reality".
@@
 
-## Sample training data
-
 Both parametric and nonparametric approaches will require training
@@
 
-<span style="color:#007FFF;">***Should we include any emprical examples?***</span>
+<!-- things that we could hit upon in discussion:
 Using data simulated from a specified model rather than empirical data
@@
 alone, and why that isn't impossible).
+-->
 
-Results
-====================================================
+<span style="color:#007FFF;">***Should we include any emprical examples?***</span>
 
-### Discussion of maximum likelihood estimated models
+### Estimating parametric models
 
@@
 
-```{r obs, dependson="sdp-pars"}
-  obs <- sim_obs(Xo, z_g, f, p, Tobs=Tobs, nz=15, 
-                 harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
-```
+The maximum likelihood estimates (MLE) will assume that the process noise $Z_t$
+is log-normal, as in the simulation itself.  In reality the noise structure might 
+differ from this assumption, making this a best-case scenario for the MLE approach.  
+The GP must also make an assumption about the noise structure -- in this case we 
+assume additve noise, though multiplicative noise would be more accurate.  This
+will illustrate the flexibility of the GP in fitting different noise structures
+even without knowledge of the correct assumptions.
 
-```{r mle, dependson="obs"}
-alt <- par_est(obs,  init = c(r = p[1], 
-                              K = mean(obs$x[obs$x>0]), 
-                              s = sigma_g))
-est <- par_est_allee(obs, f, p,  
-                     init = c(r = p[1] + 1, 
-                              K = p[2] + 2, 
-                              C = p[3] + 2, 
-                              s = sigma_g))
-```
+The GP estimate is Bayesian approach which requires specification of
+prior distributions for the hyperparameters [see Appendix for details,
+or @Rassmussen1996, @Munch2005 for background], as given in Table 1.
+Then given the data, both the GP and the MLE parametric models can
+be estimated as described above, as we implement in R in the code
+provided in supplementary materials.
 
-(MLE models will assume the noise is log-normal, which it is in the
-simulation).
+Results
+====================================================
 
-Which estimates a Ricker model with $r =$ `r alt$p[1]`, $K =$ 
-`r alt$p[2]`, and the Allen Allele model with $r =$ `r est$p[1]`, $K =$
-`r est$p[2]` and $C =$ `r est$p[3]`.
+Figure 1 shows the  inferred Gaussian Process compared to the true and the MLE parametric
+models. The grey shading shows the uncertainty estimated by the GP.  Unlike the MLE estimates,
+the GP uncertainty changes depending on where the data is available, allowing for a much 
+greater level of uncertainty at low stock sizes.  While it would be natural (and straight forward)
+to condition the GP on passing through the origin (0,0) (see appendix), the estimate shown
+here is based only on the observed data.
 
-```{r gp-priors}
-#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
-s2.p <- c(5,5)  
-tau2.p <- c(5,1)
-d.p = c(10, 1/0.1, 10, 1/0.1)
-nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
-s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
-tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
-d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
-nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
-beta0_prior <- function(x, tau) dnorm(x, 0, tau)
-beta = c(0)
-priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
-```
+<!-- Would it be useful to show the uncertainty in the other models? -->
 
-```{r gp, dependson=c("obs", "gp-priors")}
-  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
-          m0r1=FALSE, corr="exp", trace=TRUE, 
-          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
-          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
-```
-
-```{r opt, dependson=c("gp", "mle")}
-  OPT <- optimal_policy(gp, f, est$f, alt$f,
-                        p, est$p, alt$p,
-                        x_grid, h_grid, sigma_g, 
-                        sigma_g, sigma_g, # est$sigma_g, alt$sigma_g, but those ests are poor
-                        delta, xT, profit, reward, OptTime)
-```
-
-```{r sim, dependson="opt"}
-dt <- simulate_opt(OPT, f, p, x_grid, h_grid, x0, z_g, profit)
-```
-
-## Figure 1:
-
-_Shows the inferred Gaussian Process compared to the true and parametric
-models.  Refer to the appendix for details on the GP posteriors, etc._
+<!-- with (light grey) and without (darker grey) measurement error -->
 
@@
 Graph shows the expected value for the function $f$ under each model.
-Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.
+Two standard deviations from the estimated Gaussian process covariance (grey shading)  is also shown.
 The training data is also shown as black points.
-(The GP is conditioned on 0,0, shown as a pseudo-data point).
 "}
@@
 ggplot(tgp_dat) + 
-  geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-  geom_ribbon(aes(x,y,ymin=ymin2,ymax=ymax2), fill="gray60") +
+#  geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +    # measurement error + process error
+  geom_ribbon(aes(x,y,ymin=ymin2,ymax=ymax2), fill="gray60") + # process error only
   geom_line(data=models, aes(x, value, col=method), alpha=0.8, lwd=1) + 
@@
 are visible._
+
+### Discussion on how we compare performance of policies
+
+* Replicate stochastic simulations 
+* Sensitivity analysis (Figure 4).
 
@@
 $\tau^2$        Variance around the mean
-$\beta$_0       The mean is given by a linear model of slope $\beta$ 
+$\beta_0$       The mean is given by a linear model of slope $\beta$ 
 $d$             The length-scale of the covariance function

### 2013-04-08 inst/examples/allee.Rmd
another allee example altering MLE initial condition to attempt better likelihood estimates.  GP still outperforms.
update example

@@
 ```{r graphing-options, include=FALSE}
-opts_knit$set(upload.fun = socialR::notebook.url)
+opts_knit$set(upload.fun = socialR::flickr.url)
 #opts_chunk$set(dev = 'Cairo_pdf', dev.args=list(""))
@@
 # c(5, 10, 5) is 2-cycle, c(5.5, 10, 5) is 6 cycle, 5.3 is about 4
-p <- c(5.3, 10, 5) 
+p <- c(2, 10, 5) 
 K <- 10
@@
 ```{r}
-  alt <- par_est(obs,  init = c(r=2, K=mean(obs$x), s=1))
-  est <- par_est_allee(obs, f, p,  init = c(2, 6, 2, s = sigma_g))
+  alt <- par_est(obs,  init = c(r=p[1], K=mean(obs$x), s=sigma_g))
+  est <- par_est_allee(obs, f, p,  init = c(2, mean(obs$x), 2, s = sigma_g))
 ```
@@
 ```
-
-```{r echo=FALSE, results="asis"}
-bibliography("html")
-```

### 2013-04-08 inst/doc/nonparametric-bayes.tex
writing writing -- draft sent to Marc & Steve

@@
 {[}@refs{]}. While simple mechanistic models can nevertheless provide
-imporant insights into possible dynamics -- for instance, demonstrating
+important insights into possible dynamics -- for instance, demonstrating
 that a vaccine does not have to be 100\% effective to eliminate the
@@
 are several reasons for this. First, limited available data almost
-always percludes any accurate estimate of more complicated models.
-Second, the computational demands of determining an optimal stategy in a
-sequential decision-theory problem suffer from the ``curse of
+always precludes any accurate estimate of more complicated models.
+Second, the computational demands of determining an optimal strategy in
+a sequential decision-theory problem suffer from the ``curse of
 dimensionality'' that makes it difficult to consider more complicated
@@
 solving such a system. Cressie et al. (2009) provides a good account of
-the successess and challenges of the approach. Applying these approaches
+the successes and challenges of the approach. Applying these approaches
 in the management context of sequential decision making, in which
@@
 
- \textbf{This part is overstated and unnecessary} Though machine
+ \textbf{\emph{This part is overstated and unnecessary}} Though machine
 learning approaches have begun to appear in the ecological and
@@
 
- transition sentences!?
+\textbf{\emph{transition sentences!?}}
 
@@
 
- Okay, perhaps that's novel, but it's pretty trivial. Isn't it obvious
-to everyon that it's trivial?
-
- paraphrase advantages of machine learning
+\textbf{\emph{Okay, perhaps that's novel, but it's pretty trivial. Isn't
+it obvious to everyone that it's trivial?}}
+
+\textbf{\emph{paraphrase advantages of machine learning}}
 
@@
 $Z_t$) is present in the underlying model, making additional modes
-unnecessary. Add comment / apology on single species dynamics
+unnecessary. \textbf{\emph{Add comment / apology on single species
+dynamics?}}
 
@@
 
-The essense of the GP approach can be captured in the following thought
+The essence of the GP approach can be captured in the following thought
 experiment: An exhaustive parametric approach to the challenge of
@@
 GP allows us to consider a set of possible curves simultaneously.\\
-Figure 1 include curves drawn from the posterior density?
-
- Do we need more specifics on Gaussian process as an approximation to
-parametric models? Discussion of Gaussian process vs other machine
-learning / forecasting approaches that have less clear statistical
-foundations? If so, does this all belong in the discussion?
+\textbf{\emph{Figure 1 include curves drawn from the posterior
+density?}}
+
+\textbf{\emph{Do we need more specifics on Gaussian process as an
+approximation to parametric models? Discussion of Gaussian process vs
+other machine learning / forecasting approaches that have less clear
+statistical foundations? If so, does this all belong in the
+discussion?}}
 
@@
 solution (see Mangel and Clark 1988 for a detailed description of
-parametric SDP methods) to the optimal policy. The essense of the idea
+parametric SDP methods) to the optimal policy. The essence of the idea
 is straight forward -- we will use the estimated GP in place of the
 parametric growth function to determine the stochastic transition matrix
-on which the SDP calculations are based.
-
-The posterior Gaussian process is completely defined by the expected
-value and covariance matrix at a defined set of training points. For
-simplicty we will consider a these points to fall on a discrete, uniform
-grid $x$ of 101 points from 0 to 15 (1.5 times the positive equilibrium
-$K$). Again to keep things simple we will use this same grid
-discritization for the parametric approach. Other options for choosing
-the grid points, including collocation methods and functional basis
-expansion (or even using Guassian processes in place of the discrete
-optimization; an entirely different context in which GP can be used in
-SDP, see (Deisenroth, Rasmussen, and Peters 2009)) could also be
-considered.
-
-The transition matrix $\mathbf{F}$ is thus an 101 by 101 matrix for
-which the ${i,j}$ entry gives the probability of transitioning into
-state $x_i$ given that the system is in state $x_j$ in the previous
-timestep. To generate the transition matrix based on the posterior GP,
-we need only the expected values at each grid point and the
-corresponding variances (the diagonal of the covariance matrix), as
-shown in Figure 1. Given the mean at each gridpoint as the length 101
-vector $E$ and variance $V$, the probability of transitioning from state
-$x_i$ to state $x_j$ is simply
-$\mathcal{N}\left(x_j \vert  \mu = E_i, \sigma = \sqrt{V_i}\right)$,
+on which the SDP calculations are based. The SDP is solved in a
+discretized state space -- both the continuously valued population
+densities $X$ and harvest quotas $h$ are first mapped to a bounded,
+discrete grid. (For simplicity we will consider a uniform grid, though
+for either parametric or GP-based SDP it is often advantageous to use a
+non-uniform discretization such as a basis function representation,
+e.g.~see (Deisenroth, Rasmussen, and Peters 2009)).
+
+The SDP approach then computes a transition matrix, $\mathbf{F}$. We
+demonstrate that calculation is just as straight forward based on the GP
+as it is in the classical context using the parametric model. The
+${i,j}$ of the transition matrix $F$ entry gives the probability of
+transitioning into state $x_i$ given that the system is in state $x_j$
+in the previous time-step. To generate the transition matrix based on
+the posterior GP, we need only the expected values at each grid point
+and the corresponding variances (the diagonal of the covariance matrix),
+as shown in Figure 1. Given the mean of the GP posterior at each
+grid-point as the vector $E$ and variance at that point as vector $V$,
+the probability of transitioning from state $x_i$ to state $x_j$ is
+
+\[\mathcal{N}\left(x_j \vert  \mu = E_i, \sigma = \sqrt{V_i}\right)\]
+
 where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
@@
 $f(x_j) \approx f(x_j + \Delta)$, it is sufficient to calculate the
-density at $x_j$ and then row-normalize the transition matrix.
-
-\subsection{Pseudocode for the determining the transtion matrix from the
+density at $x_j$ and then row-normalize the transition matrix. The
+process can then be repeated for each possible discrete value of our
+control variable, (harvest $h$).
+
+\textbf{Pseudocode for the determining the transition matrix from the
 GP}
@@
 
-A transition matrix for each of the parametric models $f$ is calculated
-using the log-normal density with mean $f(x_i)$ and log-variance as
-estimated by maximum likelihood. From the discrete transition matrix we
-may write down the Bellman recursion defining the the stochastic dynamic
-programming iteration:
+Using the discrete transition matrix we may write down the Bellman
+recursion defining the stochastic dynamic programming iteration:
 
@@
 accomplished as follows. Consider the set of possible control values to
-be the discrete 101 values corresponding the the grid of stock sizes.
-Then for each $h_t$ there is a corresponding transition matrix
-$\mathbf{F}_h$ determined as described above but with mean
-$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element
-corresponds to the value of having stock $x_i$ at time $t$. Then let
-$\Pi_h$ be the vector whose $i$th element indicates the profit from
-harvesting at intensity $h_t$ given a population $x_i$ (e.g.
-$\max(x_i, h_t)$ since one cannot harvest more fish then the current
-population size). Then the Bellman recursion can be given in matrix form
-as
+be the discrete values corresponding the grid of stock sizes. Then for
+each $h_t$ there is a corresponding transition matrix $\mathbf{F}_h$
+determined as described above but with mean $\mu = x_j - h_t$. Let
+$\vec{V_t}$ be the vector whose $i$th element corresponds to the value
+of having stock $x_i$ at time $t$. Then let $\Pi_h$ be the vector whose
+$i$th element indicates the profit from harvesting at intensity $h_t$
+given a population $x_i$ (e.g. $\max(x_i, h_t)$ since one cannot harvest
+more fish then the current population size). Then the Bellman recursion
+can be given in matrix form as
 
@@
 
-\emph{Currently this shows the literal R code, should be adapted}
-
-\subsubsection{The optimal control problem}
-
-\subsubsection{Discussion on how we compare performance of policies}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Replicate stochastic simulations
-\item
-  Sensitivity analysis (Figure 4).
-\end{itemize}
+This completes the algorithm adapting the GP to the sequential
+decision-making problem through SDP, which we believe has not yet been
+demonstrated in the literature. We provide an R package implementation
+of this, along with the Gaussian process inference, in the supplemental
+materials.
 
@@
 ecological management and pose a substantial challenge to existing
-decision-theoretic approaches (Brozović and Schlenker 2011). To compare
-the performance of nonparametric and parametric approaches in an example
-that is easy to conceptualize, we will focus on a simple parametric
-model for a single species (derived from fist principles by Allen et al.
-2005) as our underlying ``reality''.
+decision-theoretic approaches (Brozović and Schlenker 2011). Because
+intervention is often too late after a tipping point has been crossed
+(but see Hughes et al. (2013)), management is most often concerned with
+avoiding potentially catostrophic tipping points before any data is
+available at or following a transition that would more clearly reveal
+these regime shift dynamics (e.g. Bestelmeyer et al. 2012).
+
+To illustrate the value of the non-parametric Bayesian approach to
+management, we focus on example of a system containing such a tipping
+point whose dynamics can still be described by a simple, one-dimensional
+parametric model.\\We will focus on a simple parametric model for a
+single species (derived from fist principles by Allen et al. 2005) as
+our underlying ``reality''.
 
@@
 (Courchamp, Berec, and Gascoigne 2008).
-
-\subsection{Sample training data}
 
@@
 
-Should we include any emprical examples? Using data simulated from a
-specified model rather than empirical data permits the comparison
-against the true underlying dynamics, setting a bar for the optimal
-performance possible.
-
-(Motivation, alternatives, stationarity, examples without a stable node
-(limit-cycle models), examples based on observations near a stable node
-alone, and why that isn't impossible).
-
-\section{Results}
-
-\subsubsection{Discussion of maximum likelihood estimated models}
+\textbf{\emph{Should we include any emprical examples?}}
+
+\subsubsection{Estimating parametric models}
 
@@
 
-(MLE models will assume the noise is log-normal, which it is in the
-simulation).
-
-Which estimates a Ricker model with $r =$ 1.8501, $K =$ 9.8091, and the
-Allen Allele model with $r =$ 2.8079, $K =$ 11.8235 and $C =$ 7.2159.
-
-\subsection{Figure 1:}
-
-\emph{Shows the inferred Gaussian Process compared to the true and
-parametric models. Refer to the appendix for details on the GP
-posteriors, etc.}
+The maximum likelihood estimates (MLE) will assume that the process
+noise $Z_t$ is log-normal, as in the simulation itself. In reality the
+noise structure might differ from this assumption, making this a
+best-case scenario for the MLE approach.\\The GP must also make an
+assumption about the noise structure -- in this case we assume additive
+noise, though multiplicative noise would be more accurate. This will
+illustrate the flexibility of the GP in fitting different noise
+structures even without knowledge of the correct assumptions.
+
+The GP estimate is Bayesian approach which requires specification of
+prior distributions for the hyper-parameters (see Appendix for details,
+or @Rassmussen1996, Munch et al. 2005 for background), as given in Table
+1. Then given the data, both the GP and the MLE parametric models can be
+estimated as described above, as we implement in R in the code provided
+in supplementary materials.
+
+\section{Results}
+
+Figure 1 shows the inferred Gaussian Process compared to the true and
+the MLE parametric models, predicting the relationship between observed
+stock size (x-axis) to the stock size following recruitment the
+following year. The gray shading shows the uncertainty estimated by the
+GP. Unlike the MLE estimates, the GP uncertainty changes depending on
+where the data is available, allowing for a much greater level of
+uncertainty at low stock sizes. While it would be natural (and straight
+forward) to condition the GP on passing through the origin (0,0) (see
+appendix), the estimate shown here is based only on the observed data.
+The observed data from which each model is estimated is also shown. The
+observations come from only a limited region of state space
+corresponding to unharvested or weakly harvested system. No observations
+occur at the theoretical optimum (7.5 units, see Figure 2) or near the
+tipping point (5 units).
 
@@
 \centering
-\includegraphics{figure/gp_plot.pdf}
+\includegraphics{figure/gp_plot.png}
 \caption{Graph of the inferred Gaussian process compared to the true
@@
 expected value for the function $f$ under each model. Two standard
-deviations from the estimated Gaussian process covariance with (light
-grey) and without (darker grey) measurement error are also shown. The
-training data is also shown as black points. (The GP is conditioned on
-0,0, shown as a pseudo-data point).}
+deviations from the estimated Gaussian process covariance (grey shading)
+is also shown. The training data is also shown as black points.}
 \end{figure}
 
-\subsection{Figure 2:}
-
-\emph{The take-home message, showing that the GP is closest to the
-optimal strategy, while the parametric methods are less accurate.
-Visualizing the policy may be more useful for the technical reader, the
-general audience may prefer Figure 3 showing all replicates of the
-population collapse under the parametric model and not under the GP.}
+The resulting optimal management strategy based on each of the inferred
+models is shown in Figure 2, against the optimal strategy given the true
+underlying dynamics. Policies are shown in terms of target escapement,
+$S_t$. Under models such as this a constant escapement policy is
+expected to be optimal (Reed 1979), whereby population levels below a
+certain size $S$ are unharvested, while above that size the harvest
+strategy aims to return the population to $S$, resulting in the
+hockey-stick shaped policies shown. Only the GP policy deviates slightly
+from this. The MLE parametric model corresponding to parametric
+uncertainty alone (gold line labeled ``Parametric'', that is, using the
+same structural form as the underlying simulation) nevertheless
+overestimates the carrying capacity of the stock and therefore
+underestimates the harvest rates, always waiting for the stock to rise
+further than is actually possible. The Ricker model (green line, showing
+``Structural'' inaccuracies) does not foresee the tipping point dynamics
+reflected in the concavity of the other three models (Figure 1), and
+selects a much more aggressive harvesting strategy.
 
@@
 \centering
-\includegraphics{figure/policies_plot.pdf}
+\includegraphics{figure/policies_plot.png}
 \caption{The steady-state optimal policy (infinite boundary) calculated
-under each model. Policies are shown in terms of target escapement,
-$S_t$, as under models such as this a constant escapement policy is
-expected to be optimal (Reed 1979).}
+under each model.}
 \end{figure}
 
-\subsection{Figure 3:}
-
-\emph{Figure 3 is a less abstract and more visceral visualization of the
-take-home message, with the structurally inaccurate model leading
-universally to a collapse of the fishery and very few profits, while the
-Gaussian process performs nearly optimally. The parametric approach even
-with the correct underlying structure does not perform optimally,
-choosing in this case to under-fish (may need to show harvest dynamics
-since that is not clear from the figure! Also isn't general, sometimes
-does optimally, sometimes over-fishes. Perhaps need to show more
-examples.) May need to show profits too?}
+The consequences of managing 100 replicate realizations of the simulated
+fishery under each of the policies estimated is shown in Figure 3. As
+expected from the policy curves, the structurally correct model
+(``Parametric'' uncertainty, gold) underharvests, leaving the stock to
+vary around it's unfished optimum. The GP overharvests the population
+slightly relative to the optimal value. The structurally incorrect
+Ricker model (``Structural'', green line) overharvests the population
+passed the tipping point consistently, resulting in the immediate crash
+of the stock and thus derives minimal profits.
+
+\textbf{\emph{(may need to show harvest dynamics since that is not clear
+from the figure! Also isn't general, sometimes does optimally, sometimes
+over-fishes. Perhaps need to show more examples.) May need to show
+profits too?}}
 
@@
 \centering
-\includegraphics{figure/sim_plot.pdf}
+\includegraphics{figure/sim_plot.png}
 \caption{Gaussian process inference outperforms parametric estimates.
@@
 
-\subsection{Figure 4:}
-
-\emph{Shows the sensitivity analysis. A histogram of distribution of
-yield over stochastic realizations, showing that the qualitative results
-do not depend on the stochastic realization of the training data here,
-or on the parameters of the underlying model, though quantitative
-differences are visible.}
-
-\section{Discussion / Conclusion}
-
-Non-parametric methods have received far too little attention in
-ecological modeling efforts that are aimed at improved conservation
-planning and decision making support.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Adapting a non-parametric approach requires modification of existing
-  methods for decision theory. We have illustrated how this might be
-  done in the context of stochastic dynamic programming, opening the
-  door for substantial further research into how these applications
+The results shown in Figures 1-3 are not unique to the simulated data or
+models chosen here, but arises across a range of parameter values and
+simulations as shown in the supplemental figures. The results across
+this range can most easily be compared by the relative differences in
+net present value realized by each of the approaches, as shown in Figure
+4. The Gaussian Process most consistently realizes a value close to the
+optimal solution, and importantly avoids ever driving the system across
+the tipping point, which results in the near-zero value cases in the
+parametric models.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/figure4.png}
+\caption{Histograms of the realized net present value of the fishery
+over a range of simulated data and resulting parameter estimates. For
+each data set, the three models are estimated as described above. Values
+plotted are the averages of a given policy over 100 replicate
+simulations. Details and code provided in the supplement.}
+\end{figure}
+
+\section{Discussion}
+
+Non-parametric Bayesian methods have received far too little attention
+in ecological modeling efforts that are aimed at improved conservation
+planning and decision making support. Such approaches may be
+particularly useful when the available data is restricted to a limited
+area of state-space, which can lead parametric models to underestimate
+the uncertainty in dynamics at population levels (states) which have not
+been observed. One reason for the relative absence of nonparametric
+approaches in the natural resource management context may be the lack of
+existing approaches for adapting the non-parametric Bayesian models
+previously proposed (Munch et al. 2005) to a decision-theoretic
+framework. Adapting a non-parametric approach requires modification of
+existing methods for decision theory. We have illustrated how this might
+be done for a classic stochastic dynamic programming problem, opening
+the door for substantial further research into how these applications
 might be improved.
-\end{itemize}
-
-Three elements contribute to the machine learning model performing as
-well or better than the mechanisitic models in this context.
-
-\subsubsection{1. Relevant state space}
-
-The dynamics occur over a range of state-space in which the machine
-learning model closely matches the predictions of y implausible model
-performs as well or better than the more plausible model.
-
-This is the most obvious and immediate reason why the shortcomings that
-appear to make the model implausible do not make it useless. This effect
-is enhanced in our example because the management actions help drive the
-system towards rather than away from this region of state-space.
-
-\subsubsection{2. Predictive accuracy}
-
-This application relies only on the predictive accuracy of the model,
-not an interpretation of the parameters.
-
- wow, run-on tangent? Predictive accuracy is not the goal of all
-modeling, as ecologists have been observing for as long as they made
-models (perhaps none more memorably than Levins (1969)). Mechanistic
-modeling is at its most powerful not when it is used to directly
-forecast future states but when it provides an understanding of how to
-approach a problem. SIR-type models from the epidemiological literature
-are a good example. While the simplest SIR models have little predictive
-power over the outbreak intensity or timing at a particular location,
-they provide a powerful understanding of the spread of an infection in
-terms of a single, biologically meaningful parameter: $R_0$, the basic
-reproductive number. From the model, it becomes clear that management
-need not vaccinate every member of the population to stop the spread,
-but rather it suffices to vaccinate a sufficient fraction of the
-population to reduce $R_0$ below 1.
+
+While non-parametric Bayesian approaches will not always be preferable
+to simple mechanistic models, we highlight three aspects of the problem
+consider here that make these methods particularly valuable. These
+aspects are common to many conservation decision making problems, which
+thus merit greater use of non-parametric machine learning approaches
+that can best take advantage of them.
+
+\subsubsection{1. Large uncertainty where the data is poor}
+
+The parametric models perform worst when they propose a management
+strategy outside the range of the observed data. The non-parametric
+Bayesian approach, in contrast, allows a predictive model that expresses
+a great deal of uncertainty about the probable dynamics outside the
+observed range, while retaining very good predictive accuracy in the
+range observed. The management policy dictated by the GP balance this
+uncertainty against the immediate value of the harvest, and act to
+stabilize the population dynamics in a region of state space in which
+the predictions can be reliably reflected by the data.
+
+\subsubsection{2. Predictive accuracy where data is good}
+
+While expressing larger uncertainty outside the observed data, the GP
+can also provide a better fit with smaller uncertainty inside the range
+of the observed data. This arises from the greater flexibility of the
+Gaussian process, which describes a large family of possible
+curves.\\While in a parametric context this over-fitting would be more
+worrisome -- a high-degree polynomial could fit the data even better --
+those concerns are driven by the resulting parametric fit outside the
+data, which may involve wild oscillations unsupported by the data. As we
+have seen in \#1, the GP is less vulnerable to such unjustified
+predictions outside the data, and is meanwhile free to benefit from the
+greater fit where the data is available.
 
@@
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Better express ambiguity in regions where data is not available
-\end{enumerate}
-
-One of the greatest strengths of mechanistic models is their greatest
-weakness as well.
-
 \subsection{Future directions}
@@
 uncertainty outside the observed data -- there are equally several
-aspects for which the nonparametric approach is at a relative
-disadvantage.
-
- Model complexity In this simulated example, the underlying dynamics are
-truly governed by a simple parametric model, allowing the parametric
-approaches to be more accurate. Similarly, because the dynamics are
-one-dimensional dynamics and lead to stable nodes (rather than other
-attractors such as limit-cycles resulting in oscillations), the training
-data provides relatively limited information about the dynamics. For
-these reasons, we anticipate that in higher-dimensional examples
-characteristic of ecosystem management problems that the machine
-learning approach will prove even more valuable.
-
- Data complexity. Perhaps too far out of scope\ldots{} The machine
-learning approach is also better suited to complex and disparate data.
+aspects for which the nonparametric Bayesian approach is nevertheless at
+a relative disadvantage. In this simulated example, the underlying
+dynamics are truly governed by a simple parametric model, allowing the
+parametric approaches to be more accurate. Similarly, because the
+dynamics are one-dimensional dynamics and lead to stable nodes (rather
+than other attractors such as limit-cycles resulting in oscillations),
+the training data provides relatively limited information about the
+dynamics. For these reasons, we anticipate that in higher-dimensional
+examples characteristic of ecosystem management problems that the
+machine learning approach will prove even more valuable.
+
+ Data complexity. Perhaps too far out of scope\ldots{} The nonparametric
+Bayesian approach is also better suited to complex and disparate data.
 Incorporating various sources of information into mechanistic models can
@@
 
- * Discuss constant escapement in model, in policies.
+\section{Acknowledgments}
+
+This work was partially supported by the Center for Stock Assessment
+Research, a partnership between the University of California Santa Cruz
+and the Fisheries Ecology Division, Southwest Fisheries Science Center,
+Santa Cruz, CA and by NSF~grant EF-0924195 to MM.
+
+ \textbf{\emph{Possible additional discussion material?}}
 
 \begin{itemize}
+\item
+  Discuss constant escapement in model, in policies.
 \item
@@
   Discussion of alternative related approaches: POMDP/MOMDP
-\end{itemize}
-
-\begin{center}\rule{3in}{0.4pt}\end{center}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Multiple species
 \item
@@
 
-\section{Acknowledgments}
-
-This work was partially supported by the Center for Stock Assessment
-Research, a partnership between the University of California Santa Cruz
-and the Fisheries Ecology Division, Southwest Fisheries Science Center,
-Santa Cruz, CA and by NSF~grant EF-0924195 to MM.
+\begin{center}\rule{3in}{0.4pt}\end{center}
 
@@
 
-\subsubsection{The Guassian process}
+\subsubsection{The Gaussian process}
 
@@
 \\\noalign{\medskip}
-$\beta$\_0 & The mean is given by a linear model of slope $\beta$
+$\beta_0$ & The mean is given by a linear model of slope $\beta$
 \\\noalign{\medskip}
@@
 value and covariance matrix at a defined set of training points. For
-simplicty we will consider a these points to fall on a discrete, uniform
-grid $x$ of 101 points from 0 to 15 (1.5 times the positive equilibrium
-$K$). Again to keep things simple we will use this same grid
-discritization for the parametric approach. Other options for choosing
+simplicity we will consider a these points to fall on a discrete,
+uniform grid $x$ of 101 points from 0 to 15 (1.5 times the positive
+equilibrium $K$). Again to keep things simple we will use this same grid
+discretization for the parametric approach. Other options for choosing
 the grid points, including collocation methods and functional basis
-expansion (or even using Guassian processes in place of the discrete
+expansion (or even using Gaussian processes in place of the discrete
 optimization; an entirely different context in which GP can be used in
@@
 considered.
-
-The transition matrix $\mathbf{F}$ is thus an 101 by 101 matrix for
-which the ${i,j}$ entry gives the probability of transitioning into
-state $x_i$ given that the system is in state $x_j$ in the previous
-timestep. To generate the transition matrix based on the posterior GP,
-we need only the expected values at each grid point and the
-corresponding variances (the diagonal of the covariance matrix), as
-shown in Figure 1. Given the mean at each gridpoint as the length 101
-vector $E$ and variance $V$, the probability of transitioning from state
-$x_i$ to state $x_j$ is simply
-$\mathcal{N}\left(x_j \vert  \mu = E_i, \sigma = \sqrt{V_i}\right)$,
-where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
-variance $\sigma^2$. Strictly speaking, the transition probability
-should be calculated by integrating the normal density over the bin of
-width $\Delta$ centered at $x_j$. For a sufficiently fine grid that
-$f(x_j) \approx f(x_j + \Delta)$, it is sufficient to calculate the
-density at $x_j$ and then row-normalize the transition matrix.
 
@@
 
+\textbf{\emph{Can we get this figure into the main text with appropriate
+discussion without getting too technical?}}
+
 \begin{figure}[htbp]
 \centering
-\includegraphics{figure/posteriors.pdf}
+\includegraphics{figure/posteriors.png}
 \caption{Histogram of posterior distributions for the estimated Gaussian
@@
 
+Bestelmeyer, Brandon T., Michael C. Duniway, Darren K. James, Laura M.
+Burkett, and Kris M. Havstad. 2012. ``A test of critical thresholds and
+their indicators in a desertification-prone ecosystem: more resilience
+than we thought.'' Ed. Katharine Suding. \emph{Ecology Letters} (dec).
+doi:10.1111/ele.12045. \url{http://doi.wiley.com/10.1111/ele.12045}.
+
 Brozović, Nicholas, and Wolfram Schlenker. 2011. ``Optimal management of
@@
 
+Hughes, Terry P., Cristina Linares, Vasilis Dakos, Ingrid a van de
+Leemput, and Egbert H. van Nes. 2013. ``Living dangerously on borrowed
+time during slow, unrecognized regime shifts.'' \emph{Trends in ecology
+\& evolution} 28 (3) (mar): 149--55. doi:10.1016/j.tree.2012.08.022.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/22995893}.
+
 Kocijan, Juš, Agathe Girard, Bla\textbackslash{}vz Banko, and Roderick
@@
 \url{http://www.tandfonline.com/doi/abs/10.1080/13873950500068567}.
-
-Levins, Richard. 1969. ``Some demographic and genetic consequences of
-environmental heterogeneity for biological control.'' \emph{Bulletin of
-the Entomological Society of America} 15 (3): 237--240.
-\url{http://www.ingentaconnect.com/content/esa/besa/1969/00000015/00000003/art00009}.
 

### 2013-04-08 R/gp-replicate-fns.R
updated examples to use value iteration of infinite time horizon, addressing #26

@@ -81,37 +81,39 @@ par_est <- function(obs,  init = c(r=1.5, K=mean(obs$x), s=1)){
 #' @import pdgControl
 #' @export
 optimal_policy <- function(gp, f, f_est, f_alt, p, p_est, p_alt, x_grid, h_grid, sigma_g, sigma_g_est, sigma_g_alt, delta, xT, profit, reward, OptTime){
+  MaxT = 1000
   matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.ks2, x_grid, h_grid)
-  opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=reward)
+  opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
   matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-  opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
+  opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
   matrices_estimated <- f_transition_matrix(f_est, p_est, x_grid, h_grid, sigma_g_est)
-  opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
+  opt_estimated <- value_iteration(matrices_estimated, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
   matrices_alt <- f_transition_matrix(f_alt, p_alt, x_grid, h_grid, sigma_g_alt)
-  opt_alt <- find_dp_optim(matrices_alt, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
+  opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
   
-  gp_D <- sapply(1:OptTime, function(i) opt_gp$D[,1])
-  true_D <- sapply(1:OptTime, function(i) opt_true$D[,1])
-  est_D <- sapply(1:OptTime, function(i) opt_estimated$D[,1])
-  alt_D <- sapply(1:OptTime, function(i) opt_alt$D[,1])
   
[…]

### 2013-04-08 inst/doc/beyond-mechanistic-models.Rmd
writing writing

@@
 ```
+
 ```{r policies_plot, dependson=c("opt", "plotting-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."
-}
+Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
 ```
+
 ```{r sim_plot, dependson=c("sim", "plotting-options"), fig.cap="Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."
-}
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."}
 ```
@@
 implausible model performs as well or better than the more plausible
-model.  This is the most obvious and immediate reason why the shortcomings
+model.
+
+This is the most obvious and immediate reason why the shortcomings
 that appear to make the model implausible do not make it useless.

### 2013-04-08 inst/doc/nonparametric-bayes.tex

--- 2013-04-01 inst/doc/nonparametric-bayes.tex
@@
 
- Okay, perhaps that's novel, but it's pretty trivial. Isn't it obvious
-to everyon that it's trivial?
-
- paraphrase advantages of machine learning
+\textbf{\emph{Okay, perhaps that's novel, but it's pretty trivial. Isn't
+it obvious to everyon that it's trivial?}}
+
+\textbf{\emph{paraphrase advantages of machine learning}}
 
@@
 parametric growth function to determine the stochastic transition matrix
-on which the SDP calculations are based.
-
-The posterior Gaussian process is completely defined by the expected
-value and covariance matrix at a defined set of training points. For
-simplicty we will consider a these points to fall on a discrete, uniform
-grid $x$ of 101 points from 0 to 15 (1.5 times the positive equilibrium
-$K$). Again to keep things simple we will use this same grid
-discritization for the parametric approach. Other options for choosing
-the grid points, including collocation methods and functional basis
-expansion (or even using Guassian processes in place of the discrete
-optimization; an entirely different context in which GP can be used in
-SDP, see (Deisenroth, Rasmussen, and Peters 2009)) could also be
-considered.
-
-The transition matrix $\mathbf{F}$ is thus an 101 by 101 matrix for
-which the ${i,j}$ entry gives the probability of transitioning into
-state $x_i$ given that the system is in state $x_j$ in the previous
-timestep. To generate the transition matrix based on the posterior GP,
-we need only the expected values at each grid point and the
-corresponding variances (the diagonal of the covariance matrix), as
-shown in Figure 1. Given the mean at each gridpoint as the length 101
-vector $E$ and variance $V$, the probability of transitioning from state
-$x_i$ to state $x_j$ is simply
-$\mathcal{N}\left(x_j \vert  \mu = E_i, \sigma = \sqrt{V_i}\right)$,
+on which the SDP calculations are based. The SDP is solved in a
+discritized state space -- both the continuously valued population
+densities $X$ and harvest quotas $h$ are first mapped to a bounded,
+discrete grid. (For simplicity we will consider a uniform grid, though
+for either parametric or GP-based SDP it is often advantageous to use a
+non-uniform discretization such as a basis function representation,
+e.g.~see (Deisenroth, Rasmussen, and Peters 2009)).
+
+The SDP approach then computes a transition matrix, $\mathbf{F}$. We
+demonstrate that calculation is just as straight forward based on the GP
+as it is in the classical context using the parametric model. The
+${i,j}$ of the transition matrix $F$ entry gives the probability of
+transitioning into state $x_i$ given that the system is in state $x_j$
+in the previous timestep. To generate the transition matrix based on the
+posterior GP, we need only the expected values at each grid point and
+the corresponding variances (the diagonal of the covariance matrix), as
+shown in Figure 1. Given the mean of the GP posterior at each gridpoint
+as the vector $E$ and variance at that point as vector $V$, the
+probability of transitioning from state $x_i$ to state $x_j$ is
+
+\[\mathcal{N}\left(x_j \vert  \mu = E_i, \sigma = \sqrt{V_i}\right)\]
+
 where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
@@
 $f(x_j) \approx f(x_j + \Delta)$, it is sufficient to calculate the
-density at $x_j$ and then row-normalize the transition matrix.
-
-\subsection{Pseudocode for the determining the transtion matrix from the
-GP}
+density at $x_j$ and then row-normalize the transition matrix. The
+process can then be repeated for each possible discrete value of our
+control variable, (harvest $h$).
+
+\textbf{Pseudocode for the determining the transtion matrix from the GP}
 
@@
 
-A transition matrix for each of the parametric models $f$ is calculated
-using the log-normal density with mean $f(x_i)$ and log-variance as
-estimated by maximum likelihood. From the discrete transition matrix we
-may write down the Bellman recursion defining the the stochastic dynamic
-programming iteration:
+Using the discrete transition matrix we may write down the Bellman
+recursion defining the the stochastic dynamic programming iteration:
 
@@
 accomplished as follows. Consider the set of possible control values to
-be the discrete 101 values corresponding the the grid of stock sizes.
-Then for each $h_t$ there is a corresponding transition matrix
-$\mathbf{F}_h$ determined as described above but with mean
-$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element
-corresponds to the value of having stock $x_i$ at time $t$. Then let
-$\Pi_h$ be the vector whose $i$th element indicates the profit from
-harvesting at intensity $h_t$ given a population $x_i$ (e.g.
-$\max(x_i, h_t)$ since one cannot harvest more fish then the current
-population size). Then the Bellman recursion can be given in matrix form
-as
+be the discrete values corresponding the the grid of stock sizes. Then
+for each $h_t$ there is a corresponding transition matrix $\mathbf{F}_h$
+determined as described above but with mean $\mu = x_j - h_t$. Let
+$\vec{V_t}$ be the vector whose $i$th element corresponds to the value
+of having stock $x_i$ at time $t$. Then let $\Pi_h$ be the vector whose
+$i$th element indicates the profit from harvesting at intensity $h_t$
+given a population $x_i$ (e.g. $\max(x_i, h_t)$ since one cannot harvest
+more fish then the current population size). Then the Bellman recursion
+can be given in matrix form as
 
@@
 ecological management and pose a substantial challenge to existing
-decision-theoretic approaches (Brozović and Schlenker 2011). To compare
-the performance of nonparametric and parametric approaches in an example
-that is easy to conceptualize, we will focus on a simple parametric
-model for a single species (derived from fist principles by Allen et al.
-2005) as our underlying ``reality''.
+decision-theoretic approaches (Brozović and Schlenker 2011). Because
+intervention is often too late after a tipping point has been crossed
+(but see @Hughes2013), management is most often concerned with avoiding
+potentially catostrophic tipping points before any data is available at
+or following a transition that would more clearly reveal these regime
+shift dyanmics {[}e.g. @Bestelmeyer2012{]}.
+
+To illustrate the value of the non-parametric Bayesian approach to
+management, we focus on example of a system containing such a tipping
+point whose dynamics can still be described by a simple, one-dimensional
+parametric model.\\We will focus on a simple parametric model for a
+single species (derived from fist principles by Allen et al. 2005) as
+our underlying ``reality''.
 
@@
 
-(MLE models will assume the noise is log-normal, which it is in the
-simulation).
-
-Which estimates a Ricker model with $r =$ 1.8501, $K =$ 9.8091, and the
-Allen Allele model with $r =$ 2.8079, $K =$ 11.8235 and $C =$ 7.2159.
-
-\subsection{Figure 1:}
-
-\emph{Shows the inferred Gaussian Process compared to the true and
-parametric models. Refer to the appendix for details on the GP
-posteriors, etc.}
+The maximum likelihood estimates (MLE) will assume that the process
+noise $Z_t$ is log-normal, as in the simulation itself. In reality the
+noise structure might differ from this assumption, making this a
+best-case scenario for the MLE approach.\\The GP must also make an
+assumption about the noise structure -- in this case we assume additve
+noise, though multiplicative noise would be more accurate. This will
+illustrate the flexibility of the GP in fitting different noise
+structures even without knowledge of the correct assumptions.
+
+The GP estimate is Bayesian approach which requires specification of
+prior distributions for the hyperparameters (see Appendix for details,
+or @Rassmussen1996, Munch et al. 2005 for background), as given in Table
+1. Then given the data, both the GP and the MLE parametric models can be
+estimated as described above, as we implement in R in the code provided
+in supplementary materials.
+
+\section{Results}
+
+Figure 1 shows the inferred Gaussian Process compared to the true and
+the MLE parametric models. The grey shading shows the uncertainty
+estimated by the GP. Unlike the MLE estimates, the GP uncertainty
+changes depending on where the data is available, allowing for a much
+greater level of uncertainty at low stock sizes. While it would be
+natural (and straight forward) to condition the GP on passing through
+the origin (0,0) (see appendix), the estimate shown here is based only
+on the observed data.
 
@@
 expected value for the function $f$ under each model. Two standard
-deviations from the estimated Gaussian process covariance with (light
-grey) and without (darker grey) measurement error are also shown. The
-training data is also shown as black points. (The GP is conditioned on
-0,0, shown as a pseudo-data point).}
+deviations from the estimated Gaussian process covariance (grey shading)
+is also shown. The training data is also shown as black points.}
 \end{figure}
@@
 \centering
-\includegraphics{figure/policies_plot.pdf}
+\includegraphics{figure/policies_plot.png}
 \caption{The steady-state optimal policy (infinite boundary) calculated
@@
 differences are visible.}
+
+\subsubsection{Discussion on how we compare performance of policies}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Replicate stochastic simulations
+\item
+  Sensitivity analysis (Figure 4).
+\end{itemize}
 
@@
 \centering
-\includegraphics{figure/posteriors.pdf}
+\includegraphics{figure/posteriors.png}
 \caption{Histogram of posterior distributions for the estimated Gaussian

### 2013-04-08 inst/examples/adp-intro.Rmd

@@
                cache=TRUE, cache.path = "adp/",
-               comment=NA, message=FALSE, warning=FALSE)
+               comment=NA, message=FALSE, warning=FALSE,
+               tidy=FALSE)
 ```
@@
 
-```{r}
-a <- 1.5
-b <- .5
-f <- function(x, z) z * a * x / (x + b)
+``` {r bevholt}
+f <- function(x, h, p){
+    A <- p[1] 
+    B <- p[2] 
+    s <- pmax(x-h, 0)
+    A * s/(1 + B * s)
+}
+pars <- c(1.5, 0.05)
+K <- (pars[1] - 1)/pars[2]
 ```
 
-We begin with a simulation method $X_{t+1} = f(X_t, Z_t)$.  For illustration, let us consider $f(X_t, Z_t) = Z_t \frac{a X_t}{b + X_t}$ with a = `r a` and b = `r b`.  We define a statespace $S$
+We begin with a simulation method $X_{t+1} = f(X_t, Z_t)$.
+For illustration, let us consider $f(X_t, Z_t) = Z_t \frac{a X_t}{b + X_t}$ with a = `r pars[1]` and b = `r pars[2]`.
+We define a statespace $S$
 
@@
 
-```r
+```{r}
 chi <- S
@@
 
-The approximate dynamic programming algorithm will perform a finite number $N$ = `r N$ iterations over a window of time $T$ =`r T` in our example.  The algorithm can be described as follows:
+The approximate dynamic programming algorithm will perform a finite number $N$ = `r N` iterations over a window of time $T$ =`r T` in our example.  The algorithm can be described as follows:
 
@@
 
-```r
-Z <- lnorm(T)
+```{r}
+sigma <- 0.2
+Z <- rlnorm(T, 0, sigma)
 ```
@@
 
-$$V_t(S_t) = \max_{x_t \in \chi_t} \left(C(S_t, x_t) + \gamma \sum_{s^{\prime} \in mathcal{S}} \mathbb{P}(s^{\prime} | S_t^n, x_t) V_{t+1}^{n-1} s^{\prime} \right)
+$$V_t(S_t) = \max_{x_t \in \chi_t} \left(C(S_t, x_t) + \gamma \sum_{s^{\prime} \in mathcal{S}} \mathbb{P}(s^{\prime} | S_t^n, x_t) V_{t+1}^{n-1} s^{\prime} \right)$$
 
@@
 
-```{r}
+```r
 s <- S_0
+C <- function(S, X) pmin(S, X)
 
-sapply(0:T, function(t){
-  max(
-    arg <- sapply(chi, function(x){
-      min(s, x) + 
-    c(x.n_t = which.max(arg), v.n_t = max(arg))
+arg <- sapply(chi, function(x) C(s, x) + f(S, x, pars) %*% V)
+c(x_nt = which.max(arg), v_nt = max(arg))
+
+## Update V
 

### 2013-04-08 inst/doc/nonparametric-bayes.Rmd

@@
 
-```{r libraries, include=FALSE}
-## Dargh these are ignored if not placed here!?
-options(xtable.type = 'latex', table.placement="H", xtable.print.comment=FALSE)
-opts_chunk$set(cache.path = 'cache-pdf/', cache=TRUE)
-opts_chunk$set(dev = 'Cairo_pdf', fig.width=5.5, fig.height=4)
-
+```{r libraries, include=FALSE, cache=FALSE}
 library(pdgControl)
@@
 
-```{r plotting-options, echo=FALSE, include=FALSE}
+```{r plotting-options, echo=FALSE, include=FALSE, cache=TRUE}
+## Dargh these are ignored if not placed here!?
+options(xtable.type = 'latex', table.placement="H", xtable.print.comment=FALSE)
+opts_chunk$set(cache.path = 'cache-pdf/', cache=TRUE)
+opts_chunk$set(dev = 'Cairo_pdf', fig.width=5.5, fig.height=4)
+
 # This stuff should be handled by knit script(?) (load ggplot2 first though)!
@@
 dynamics too complex to be captured by simple models [@refs]. While
-simple mechanistic models can nevertheless provide imporant insights
+simple mechanistic models can nevertheless provide important insights
 into possible dynamics -- for instance, demonstrating that a vaccine
@@
 several reasons for this.  First, limited available data almost always
-percludes any accurate estimate of more complicated models.  Second, the
-computational demands of determining an optimal stategy in a sequential
+precludes any accurate estimate of more complicated models.  Second, the
+computational demands of determining an optimal strategy in a sequential
 decision-theory problem suffer from the "curse of dimensionality"
@@
 task in itself, let alone the computational difficulties of solving such
-a system. @Cressie2009 provides a good account of the successess and
+a system. @Cressie2009 provides a good account of the successes and
 challenges of the approach.  Applying these approaches in the management
@@
 <span style="color:#007FFF;">***Okay, perhaps that's novel, but it's pretty trivial.  Isn't it
-obvious to everyon that it's trivial?***</span>
+obvious to everyone that it's trivial?***</span>
 
@@
 
-The essense of the GP approach can be captured in the
+The essence of the GP approach can be captured in the
 following thought experiment: An exhaustive parametric approach to the
@@
 @Mangel1988 for a detailed description of parametric SDP methods] to the
-optimal policy. The essense of the idea is straight forward -- we will use
+optimal policy. The essence of the idea is straight forward -- we will use
 the estimated GP in place of the parametric growth function to determine
 the stochastic transition matrix on which the SDP calculations are based.
-The SDP is solved in a discritized state space -- both the continuously 
+The SDP is solved in a discretized state space -- both the continuously 
 valued population densities $X$ and harvest quotas $h$ are first mapped to
@@
 state $x_i$ given that the system is in state $x_j$ in the previous 
-timestep.  To generate the transition
+time-step.  To generate the transition
 matrix based on the posterior GP, we need only the expected values
@@
 the covariance matrix), as shown in Figure 1.  Given the mean of the 
-GP posterior at each gridpoint as the vector $E$ and variance at that
+GP posterior at each grid-point as the vector $E$ and variance at that
 point as vector $V$, the probability of transitioning from state $x_i$ to state $x_j$ is
@@
 
-**Pseudocode for the determining the transtion matrix from the GP**
+**Pseudocode for the determining the transition matrix from the GP**
 
@@
 Using the discrete transition matrix we may write down the Bellman
-recursion defining the the stochastic dynamic programming iteration:
+recursion defining the stochastic dynamic programming iteration:
 
@@
 accomplished as follows. Consider the set of possible control values to
-be the discrete values corresponding the the grid of
+be the discrete values corresponding the grid of
 stock sizes.  Then for each $h_t$ there is a corresponding transition
@@
 points before any data is available at or following a transition that
-would more clearly reveal these regime shift dyanmics [e.g. @Bestelmeyer2012].
+would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].
 
@@
 The GP must also make an assumption about the noise structure -- in this case we 
-assume additve noise, though multiplicative noise would be more accurate.  This
+assume additive noise, though multiplicative noise would be more accurate.  This
 will illustrate the flexibility of the GP in fitting different noise structures
@@
 The GP estimate is Bayesian approach which requires specification of
-prior distributions for the hyperparameters [see Appendix for details,
+prior distributions for the hyper-parameters [see Appendix for details,
 or @Rassmussen1996, @Munch2005 for background], as given in Table 1.
@@
 
-Figure 1 shows the  inferred Gaussian Process compared to the true and the MLE parametric
-models. The grey shading shows the uncertainty estimated by the GP.  Unlike the MLE estimates,
-the GP uncertainty changes depending on where the data is available, allowing for a much 
-greater level of uncertainty at low stock sizes.  While it would be natural (and straight forward)
-to condition the GP on passing through the origin (0,0) (see appendix), the estimate shown
-here is based only on the observed data.
+Figure 1 shows the  inferred Gaussian Process compared to the true and the
+MLE parametric models, predicting the relationship between observed stock
+size (x-axis) to the stock size following recruitment the following year.
+The gray shading shows the uncertainty estimated by the GP.  Unlike the
+MLE estimates, the GP uncertainty changes depending on where the data is
+available, allowing for a much greater level of uncertainty at low stock
+sizes.  While it would be natural (and straight forward) to condition
+the GP on passing through the origin (0,0) (see appendix), the estimate
+shown here is based only on the observed data. The observed data from
+which each model is estimated is also shown.  The observations come
+from only a limited region of state space corresponding to unharvested
+or weakly harvested system.  No observations occur at the theoretical
+optimum (7.5 units, see Figure 2) or near the tipping point (5 units).
 
@@
 
-## Figure 2:
+The resulting optimal management strategy based on each of the inferred
+models is shown in Figure 2, against the optimal strategy given the
+true underlying dynamics.  Policies are shown in terms of target
+escapement, $S_t$.  Under models such as this a constant escapement
+policy is expected to be optimal [@Reed1979], whereby population levels
+below a certain size $S$ are unharvested, while above that size the harvest
+strategy aims to return the population to $S$, resulting in the hockey-stick
+shaped policies shown.  Only the GP policy deviates slightly from this.  The MLE parametric model
+corresponding to parametric uncertainty alone (gold line labeled "Parametric", 
+that is, using the same structural form as the underlying simulation) nevertheless overestimates 
+the carrying capacity of the stock and therefore underestimates the harvest
+rates, always waiting for the stock to rise further than is actually possible.
+The Ricker model (green line, showing "Structural" inaccuracies) does not 
+foresee the tipping point dynamics reflected in the concavity of the other three
+models (Figure 1), and selects a much more aggressive harvesting strategy.
 
-_The take-home message, showing that the GP is closest to the optimal
-strategy, while the parametric methods are less accurate.  Visualizing the
-policy may be more useful for the technical reader, the general audience
-may prefer Figure 3 showing all replicates of the population collapse
-under the parametric model and not under the GP._
-
-```{r policies_plot, dependson=c("opt", "plotting-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
+```{r policies_plot, dependson=c("opt", "plotting-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.  "}
 policies <- 
@@
 
-## Figure 3:
+The consequences of managing 100 replicate realizations of the 
+simulated fishery under each of the policies estimated is shown in Figure 3.  As expected
+from the policy curves, the structurally correct model ("Parametric" uncertainty, gold) 
+underharvests, leaving the stock to vary around it's unfished optimum.  The GP overharvests
+the population slightly relative to the optimal value.  The structurally
+incorrect Ricker model ("Structural", green line) overharvests the population passed
+the tipping point consistently, resulting in the immediate crash of the stock and 
+thus derives minimal profits.
 
-_Figure 3 is a less abstract and more visceral visualization of the
-take-home message, with the structurally inaccurate model leading
-universally to a collapse of the fishery and very few profits, while the
-Gaussian process performs nearly optimally.  The parametric approach even
-with the correct underlying structure does not perform optimally, choosing
-in this case to under-fish (may need to show harvest dynamics since
+***(may need to show harvest dynamics since
 that is not clear from the figure! Also isn't general, sometimes does
 optimally, sometimes over-fishes.  Perhaps need to show more examples.)
-May need to show profits too?_
+May need to show profits too?***
 
@@
                   color=method), alpha=.1) +
-    scale_colour_manual(values=cbPalette, 
+    scale_colour_manual(values=c("GP" = cbPalette[1], "Parametric" = cbPalette[2], "True" = cbPalette[3], "Structural" = cbPalette[4]),  
                         guide = guide_legend(override.aes = list(alpha = 1)))
@@
 
-## Figure 4:
+The results shown in Figures 1-3 are not unique to the simulated data or models chosen
+here, but arises across a range of parameter values and simulations as shown in the 
+supplemental figures.  The results across this range can most easily be compared 
+by the relative differences in net present value realized by each of the approaches,
+as shown in Figure 4.  The Gaussian Process most consistently realizes a value 
+close to the optimal solution, and importantly avoids ever driving the system across
+the tipping point, which results in the near-zero value cases in the parametric models.
 
-_Shows the sensitivity analysis.  A histogram of distribution of yield
-over stochastic realizations, showing that the qualitative results do
-not depend on the stochastic realization of the training data here, or on
-the parameters of the underlying model, though quantitative differences
-are visible._
+![Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
+For each data set, the three models are estimated as described above.
+Values plotted are the averages of a given policy over 100 replicate simulations.
+Details and code provided in the supplement.](figure/figure4.png)
 
-### Discussion on how we compare performance of policies
-
-* Replicate stochastic simulations 
-* Sensitivity analysis (Figure 4).
-
-Discussion / Conclusion
+Discussion 
 ==================================================================
 
-Non-parametric methods have received far too little attention in
-ecological modeling efforts that are aimed at improved conservation
-planning and decision making support.
-
-* Adapting a non-parametric approach requires modification of existing
+Non-parametric Bayesian methods have received far too little attention
+in ecological modeling efforts that are aimed at improved conservation
+planning and decision making support.  Such approaches may be particularly
+useful when the available data is restricted to a limited area of
+state-space, which can lead parametric models to underestimate the
+uncertainty in dynamics at population levels (states) which have not
+been observed.  One reason for the relative absence of nonparametric
+approaches in the natural resource management context may be the lack
+of existing approaches for adapting the non-parametric Bayesian models
+previously proposed [@Munch2005] to a decision-theoretic framework.
+Adapting a non-parametric approach requires modification of existing
 methods for decision theory.  We have illustrated how this might be
-done in the context of stochastic dynamic programming, opening the
+done for a classic stochastic dynamic programming problem, opening the
 door for substantial further research into how these applications might
@@
 
-Three elements contribute to the machine learning model performing as well
-or better than the mechanisitic models in this context.
+While non-parametric Bayesian approaches will not always be preferable 
+to simple mechanistic models, we highlight three aspects of the problem
+consider here that make these methods particularly valuable.  These aspects
+are common to many conservation decision making problems, which thus merit
+greater use of non-parametric machine learning approaches that can best 
+take advantage of them.
 
-### 1. Relevant state space
+### 1. Large uncertainty where the data is poor
 
-The dynamics occur over a range of state-space in which the machine learning
-model closely matches the predictions of y
-implausible model performs as well or better than the more plausible
-model.
+The parametric models perform worst when they propose a management strategy
+outside the range of the observed data. The non-parametric Bayesian approach, 
+in contrast, allows a predictive model that expresses a great deal of uncertainty
+about the probable dynamics outside the observed range, while retaining very
+good predictive accuracy in the range observed.  The management policy 
+dictated by the GP balance this uncertainty against
+the immediate value of the harvest, and act to stabilize the population 
+dynamics in a region of state space in which the predictions can be 
+reliably reflected by the data.
 
-This is the most obvious and immediate reason why the shortcomings
-that appear to make the model implausible do not make it useless.
-This effect is enhanced in our example because the management actions help
-drive the system towards rather than away from this region of state-space.
+### 2. Predictive accuracy where data is good
 
-### 2. Predictive accuracy
+While expressing larger uncertainty outside the observed data, the GP
+can also provide a better fit with smaller uncertainty inside the range
+of the observed data. This arises from the greater flexibility of the 
+Gaussian process, which describes a large family of possible curves.  
+While in a parametric context this over-fitting would be more worrisome --
+a high-degree polynomial could fit the data even better -- those 
+concerns are driven by the resulting parametric fit outside the data,
+which may involve wild oscillations unsupported by the data.  As we 
+have seen in #1, the GP is less vulnerable to such unjustified predictions
+outside the data, and is meanwhile free to benefit from the greater 
+fit where the data is available.
 
-This application relies only on the predictive accuracy of the model,
-not an interpretation of the parameters.
-
-<span style="color:#007FFF;"> wow, run-on tangent? </span>
+<!--
+<span style="color:#007FFF;"> ***wow, run-on tangent here?*** </span>
 Predictive accuracy is not the goal of all modeling, as ecologists
@@
 vaccinate a sufficient fraction of the population to reduce $R_0$ below 1.
+-->
 
@@
 
-2. Better express ambiguity in regions where data is not available
-
-One of the greatest strengths of mechanistic models is their greatest weakness as well.
-
 ## Future directions
@@
 space, the ability to express uncertainty outside the observed data --
-there are equally several aspects for which the nonparametric approach is
-at a relative disadvantage.
-
-<span style="color:#007FFF;"> Model complexity </span>
-In this simulated example, the underlying
+there are equally several aspects for which the nonparametric Bayesian approach is
+nevertheless at a relative disadvantage. In this simulated example, the underlying
 dynamics are truly governed by a simple parametric model, allowing
@@
 <span style="color:#007FFF;"> Data complexity. Perhaps too far out of scope... </span>
-The machine learning approach is also better suited to complex and
+The nonparametric Bayesian approach is also better suited to complex and
 disparate data.  Incorporating various sources of information into
@@
 
-<span style="color:#007FFF;"> 
-* Discuss constant escapement in model, in policies.
-
-* Limitations of this comparison: Are the maximum-likelihood solutions
-a straw man?
-
-* Discussion of alternative related approaches: POMDP/MOMDP
-</span>
-
---------------------------------------------------------------------
-
-* Multiple species 
-* Online learning 
-* Multiple step-ahead predictions 
-* Explicitly accomidating additional uncertainties 
-* Improving inference of optimal policy from the GP
-
-```{r echo=FALSE, results="asis"}
-#bibliography()
-```
-
 Acknowledgments
@@
 
+<span style="color:#007FFF;"> 
+***Possible additional discussion material?***
+
+* Discuss constant escapement in model, in policies.
+* Limitations of this comparison: Are the maximum-likelihood solutions
+a straw man?
+* Discussion of alternative related approaches: POMDP/MOMDP
+
+* Online learning 
+* Multiple step-ahead predictions 
+* Explicitly accomidating additional uncertainties 
+* Improving inference of optimal policy from the GP
+</span>
+
+--------------------------------------------------------------------
+
 Appendix / Supplementary Materials
@@
 
-### The Guassian process
+### The Gaussian process
 
@@
 The posterior Gaussian process is completely defined by the expected value
-and covariance matrix at a defined set of training points.  For simplicty
+and covariance matrix at a defined set of training points.  For simplicity
 we will consider a these points to fall on a discrete, uniform grid $x$
@@
 (1.5 times the positive equilibrium $K$).  Again to keep things simple
-we will use this same grid discritization for the parametric approach.
+we will use this same grid discretization for the parametric approach.
 Other options for choosing the grid points, including collocation methods
-and functional basis expansion (or even using Guassian processes in place
+and functional basis expansion (or even using Gaussian processes in place
 of the discrete optimization; an entirely different context in which GP
 can be used in SDP, see [@Deisenroth2009]) could also be considered.
-
-The transition matrix $\mathbf{F}$ is thus an `r length(x_grid)`
-by `r length(x_grid)` matrix for which the ${i,j}$ entry gives the
-probability of transitioning into state $x_i$ given that the system is
-in state $x_j$ in the previous timestep.  To generate the transition
-matrix based on the posterior GP, we need only the expected values
-at each grid point and the corresponding variances (the diagonal of
-the covariance matrix), as shown in Figure 1.  Given the mean at each
-gridpoint as the length `r length(x_grid)` vector $E$ and variance $V$,
-the probability of transitioning from state $x_i$ to state $x_j$ is
-simply $\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)$,
-where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
-variance $\sigma^2$.  Strictly speaking, the transition probability should
-be calculated by integrating the normal density over the bin of width
-$\Delta$ centered at $x_j$.  For a sufficiently fine grid that $f(x_j)
-\approx f(x_j + \Delta)$, it is sufficient to calculate the density at
-$x_j$ and then row-normalize the transition matrix.
 
@@
 
+***Can we get this figure into the main text with appropriate discussion without getting too technical?***
+
 ```{r posteriors, dependson=c("gp", "plotting-options"), fig.cap = "Histogram of posterior distributions for the estimated Gaussian Process shown in Figure 1.  Prior distributions overlaid."}

### 2013-04-12 inst/examples/BUGS/allen.Rmd
Jags example added

```{r libraries, include=FALSE, cache=FALSE}
library(pdgControl)
library(nonparametricbayes)
library(reshape2)
library(ggplot2)
library(data.table)
library(tgp)
library(MCMCpack)
library(plyr)
library(knitcitations)
```

```{r plotting-options, echo=FALSE, include=FALSE, cache=TRUE}
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=10))
theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
             plot.background = element_rect(fill = "transparent",colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
f <- RickerAllee
p <- c(2, 10, 5) 
K <- 10
allee <- 5
```

```{r sdp-pars, dependson="stateeq"}
sigma_g <- 0.05
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
x_grid <- seq(0, 1.5 * K, length=101)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 20  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
seed_i <- 1
Xo <- K # observations start from
x0 <- Xo # simulation under policy starts from
Tobs <- 35
```

```{r obs, dependson="sdp-pars"}
obs <- sim_obs(Xo, z_g, f, p, Tobs=Tobs, nz=1, 
                harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
```

```{r mle, dependson="obs"}
alt <- par_est(obs,  init = c(r = p[1], 
                              K = mean(obs$x[obs$x>0]), 
                              s = sigma_g))
est <- par_est_allee(obs, f, p,  
                     init = c(r = p[1] + 1, 
                              K = p[2] + 2, 
                              C = p[3] + 2, 
                              s = sigma_g))
```

```{r}
library(R2jags)
```

```{r}
init_p = est$p
names(init_p) = c("r0", "K", "theta")
y <- obs$y[-1] 
N=length(y);
```

```{r}
jags.data <- list("N","y")
n.chains = 1
n.iter = 20000
n.burnin = floor(n.iter/2)
n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
```

## Gamma priors on precision

```{r}
jags.params=c("K","logr0","logtheta","iR","iQ"); # Don't need to save "x"
jags.inits <- function(){
  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]),"iQ"=1/0.05,"iR"=1/0.1,"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345)
time<-system.time(       
  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="bugmodel-GammaPrior.txt")
)         
time <- unname(time["elapsed"]);
```

```{r}
library(emdbook)
library(coda)
tfit_jags_m <- as.mcmc.bugs(jagsfit$BUGSoutput)
print(xyplot(tfit_jags_m))
print(densityplot(tfit_jags_m))
```

```{r}
mcmc <- as.mcmc(jagsfit)
mcmcall <- mcmc[,-2]
mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4],mcmcall[,5])
## get parameter estimates (and transform inverse variances back to SDs)
coef <- apply(mcmcall,2,mean)
coef[2:3]=log(sqrt(1/coef[2:3]));
coef[4:5]=exp(coef[2:3]);

coef.median<-apply(mcmcall,2,median)
coef.median[2:3]=log(sqrt(1/coef.median[2:3]));
coef.median[4:5]=exp(coef.median[4:5]);

dens<-apply(mcmcall,2,density)
coef.mode<-sapply(dens,function(x) x$x[which.max(x$y)] )
coef.mode[2:3]=log(sqrt(1/coef.mode[2:3]));
coef.mode[4:5]=exp(coef.mode[4:5]);

ci.mc <- t(apply(mcmcall,2,quantile, probs=c(0.025,0.975)))
## OBS remember that upper and lower are reversed!
ci.mc[2:3,2:1]=log(sqrt(1/ci.mc[2:3,]));
ci.mc[4:5,]=exp(ci.mc[4:5,]);

##reorder the stuff (K,Q,R,r0,theta)=>(theta r0 K Q R)
myorder=c(5,4,1,2,3);
coef=coef[myorder];
coef.median=coef.median[myorder];
coef.mode=coef.mode[myorder];
ci.mc=ci.mc[myorder,];

convinfo=list(effsize=effectiveSize(mcmc), geweke=geweke.diag(mcmc))       ## MCMC convergence info

```

### 2013-04-12 R/gp-replicate-fns.R

@@ -135,7 +135,8 @@ sim_plots <- function(dt, seed=1, harvest=FALSE){
   print(fish_plot)
   harvest_plot <- ggplot(dt) +
     geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
-    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1))) +
+    scale_colour_manual(values=c("GP" = cbPalette[1], "Parametric" = cbPalette[2], "True" = cbPalette[3], "Structural" = cbPalette[4]),  
+                        guide = guide_legend(override.aes = list(alpha = 1))) +
     ggtitle(paste("seed", seed))
   if(harvest)
     print(harvest_plot)

### 2013-04-19 inst/examples/BUGS/allen.Rmd
parametric bayes intial example

@@
 mcmcall <- mcmc[,-2]
+who <- colnames(mcmcall)
 mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4],mcmcall[,5])
-## get parameter estimates (and transform inverse variances back to SDs)
-coef <- apply(mcmcall,2,mean)
-coef[2:3]=log(sqrt(1/coef[2:3]));
-coef[4:5]=exp(coef[2:3]);
+colnames(mcmcall) <- who
+```
 
-coef.median<-apply(mcmcall,2,median)
-coef.median[2:3]=log(sqrt(1/coef.median[2:3]));
-coef.median[4:5]=exp(coef.median[4:5]);
+```{r}
+theta <- exp(mcmcall[,"logtheta"])
+theta_dist <- hist(theta, freq=FALSE)
+theta_dist$mids
+theta_dist$density
+delta <- theta_dist$mids[2] - theta_dist$mids[1]
+```
 
-dens<-apply(mcmcall,2,density)
-coef.mode<-sapply(dens,function(x) x$x[which.max(x$y)] )
-coef.mode[2:3]=log(sqrt(1/coef.mode[2:3]));
-coef.mode[4:5]=exp(coef.mode[4:5]);
+evaluating the value function given $f$ fixed at each of `theta_dist$mids` multiplied by  `theta_dist$density * delta` and summed over each $\theta$ is going to be slow for a single parameter but simply unrealistic for higher dimensions.
+Instead we will rely on Monte Carlo sampling of parameter values from their posteriors,
 
-ci.mc <- t(apply(mcmcall,2,quantile, probs=c(0.025,0.975)))
-## OBS remember that upper and lower are reversed!
-ci.mc[2:3,2:1]=log(sqrt(1/ci.mc[2:3,]));
-ci.mc[4:5,]=exp(ci.mc[4:5,]);
-
-##reorder the stuff (K,Q,R,r0,theta)=>(theta r0 K Q R)
-myorder=c(5,4,1,2,3);
-coef=coef[myorder];
-coef.median=coef.median[myorder];
-coef.mode=coef.mode[myorder];
-ci.mc=ci.mc[myorder,];
-
-convinfo=list(effsize=effectiveSize(mcmc), geweke=geweke.diag(mcmc))       ## MCMC convergence info
+```{r}
+mc_n <- 100
+thetas <- sample(theta, mc_n, replace=T)
 

### 2013-04-19 inst/examples/BUGS/par_uncertainty_sdp.Rmd
parametric bayes intial example

```{r}
f_transition_matrix <- function(f, p, x_grid, h_grid, sigma_g, pardist){
  lapply(h_grid, par_F, f, p, x_grid, sigma_g, pardist)
}

par_F <- function(h, f, p, x_grid, sigma_g, pardist, n_mc = 100){

  # Set up monte carlo sampling 
  d <- dim(pardist)
  indices <- round(runif(n_mc,1, d[1]))

  # compute mean with monte carlo samples
  mu <- sapply(indices, function(i) { # Cols are parameter samples
    p <- unname(pardist[i,c(4,1,5)])          # Only, we need to add the measurement and process noise  
    unname(sapply(x_grid, f, h, p))   # rows are possible values of x_t
  })
  mu <- data.frame(t(mu)) # Let's have cols as possible x_t, rows as parameters, and type as data.frame

  F_true <- 
    sapply(mu, function(cur_state){    # For each x_t

      bypar <- sapply(cur_state, function(m){ # For each parameter value
        if(snap_to_grid(m,x_grid) < x_grid[2]){ # 
          out <- numeric(length(x_grid))
          out[1] <- 1
          out
        } else {
          out <- dlnorm(x_grid/m, 0, sigma_g) # should be using the estimate process noise!
        }})  
      ave_over_pars <- apply(bypar, 1, sum) # collapse by weighted average over possible parameters    
      ave_over_pars / sum(ave_over_pars)
      })

  F_true <- t(F_true)

}

snap_to_grid <- function(x, grid) sapply(x, function(x) grid[which.min(abs(grid - x))])

# internal helper function
rownorm <- function(M) 
  t(apply(M, 1, function(x){ 
    if(sum(x)>0){
      x/sum(x)
    } else {
      out <- numeric(length(x))
      out[1] <- 1
      out
    }
  }))
```

```{r}
knit("allen.Rmd")

pardist <- mcmcall
#pardist[,1] = p[2] + rnorm(100, 0, 0.000001)
#pardist[,4] = p[1] + rnorm(100, 0, 0.000001)
#pardist[,5] = p[3] + rnorm(100, 0, 0.000001)

sdp = f_transition_matrix(f, p, x_grid, h_grid, sigma_g, pardist)
s_opt <- value_iteration(sdp, x_grid, h_grid, OptTime=1000, xT, profit, delta)
#opt <- find_dp_optim(sdp, x_grid, h_grid, OptTime, xT, profit, delta, reward=0)

SDP_Mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g)
pars_fixed <- value_iteration(SDP_Mat, x_grid, h_grid, OptTime=1000, xT, profit, delta)

require(reshape2)
policies <- melt(data.frame(stock=x_grid, pars.uncert = x_grid[s_opt$D], pars.fixed = x_grid[pars_fixed$D]), id="stock")

ggplot(policies, aes(stock, value, color=variable)) + geom_point(alpha=.5) + xlab("stock size") + ylab("harvest")

ggplot(policies, aes(stock, stock - value, color=variable)) + geom_line(alpha=1) + xlab("stock size") + ylab("escapement") 
```

### 2013-04-20 inst/examples/BUGS/2013-04-19-bayesian-parametric-uncertainty-sdp.Rmd
without uncertainty invariance parameters

--- 2013-04-19 inst/examples/BUGS/par_uncertainty_sdp.Rmd
@@
+---
+layout: post
+category: ecology
+tags:
+- nonparametric-bayes
+
+---
+
+Adapting the parametric uncertainty represented by the posterior distributions of the Bayesian estimate (see [earlier notes](http://carlboettiger.info/2013/04/12/parametric-bayesian-example.html)) to the stochastic dynamic programming solution for the optimal policy.
+Simply requires evaluating the expectation over the distribution, but is computationally intensive given the spread and the three parameters.
+
 ```{r}
@@
 
+Run the Bayesian analysis to obtain posterior distributions for parameters.
+
+```{r, echo=TRUE, include=FALSE}
+knit("allen.Rmd")
+```
+
+## Test case using perturbed parameters
+
+First, sanity test. Use the correct parameter values (slightly perturbed).
+
 ```{r}
-knit("allen.Rmd")
+pardist <- mcmcall
+pardist[,1] = p[2] + rnorm(100, 0, 0.000001)
+pardist[,4] = p[1] + rnorm(100, 0, 0.000001)
+pardist[,5] = p[3] + rnorm(100, 0, 0.000001)
+```
 
-pardist <- mcmcall
-#pardist[,1] = p[2] + rnorm(100, 0, 0.000001)
-#pardist[,4] = p[1] + rnorm(100, 0, 0.000001)
-#pardist[,5] = p[3] + rnorm(100, 0, 0.000001)
+Compute optimal policy
 
+```{r}
 sdp = f_transition_matrix(f, p, x_grid, h_grid, sigma_g, pardist)
 s_opt <- value_iteration(sdp, x_grid, h_grid, OptTime=1000, xT, profit, delta)
-#opt <- find_dp_optim(sdp, x_grid, h_grid, OptTime, xT, profit, delta, reward=0)
+```
 
+Compare to the case without parameter uncertainty (growth noise only)
+
+```{r}
 SDP_Mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g)
 pars_fixed <- value_iteration(SDP_Mat, x_grid, h_grid, OptTime=1000, xT, profit, delta)
+```
 
+Plot results
+
+```{r}
 require(reshape2)
 policies <- melt(data.frame(stock=x_grid, pars.uncert = x_grid[s_opt$D], pars.fixed = x_grid[pars_fixed$D]), id="stock")
-
-ggplot(policies, aes(stock, value, color=variable)) + geom_point(alpha=.5) + xlab("stock size") + ylab("harvest")
-
 ggplot(policies, aes(stock, stock - value, color=variable)) + geom_line(alpha=1) + xlab("stock size") + ylab("escapement") 
 ```
+
+## Using actual estimates
+
+```{r}
+pardist <- mcmcall
+```
+
+Transform parameters back
+
+```{r}
+pardist[,4] = exp(pardist[,4])
+pardist[,5] = exp(pardist[,5])
+```
+
+Compute optimal policy
+
+```{r}
+sdp = f_transition_matrix(f, p, x_grid, h_grid, sigma_g, pardist)
+s_opt <- value_iteration(sdp, x_grid, h_grid, OptTime=1000, xT, profit, delta)
+```
+
+Compare to the case without parameter uncertainty (growth noise only)
+
+```{r}
+SDP_Mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g)
+pars_fixed <- value_iteration(SDP_Mat, x_grid, h_grid, OptTime=1000, xT, profit, delta)
+```
+
+Plot results
+
+```{r}
+require(reshape2)
+policies <- melt(data.frame(stock=x_grid, pars.uncert = x_grid[s_opt$D], pars.fixed = x_grid[pars_fixed$D]), id="stock")
+ggplot(policies, aes(stock, stock - value, color=variable)) + geom_line(alpha=1) + xlab("stock size") + ylab("escapement") 
+```
+
+## To Do
+
+So far this is just a proof of principle example.
+
+* Needs to be adjusted to account for uncertainty in the estimates of the noise processes as well. 
+* Needs to be written as generic and documented functions, add to nonparametric-bayes package routines.

### 2013-04-20 inst/examples/BUGS/allen.Rmd
without uncertainty invariance parameters

@@
 who <- colnames(mcmcall)
+who 
 mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4],mcmcall[,5])

### 2013-04-25 inst/examples/allee.Rmd
tidying code

--- 2013-04-08 inst/examples/allee.Rmd
@@
 cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-#sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
-#short_sha <- gsub("(^.{10}).*", "\\1", sha)
-#date <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S")
-#opts_chunk$set(fig.path = paste("figure/", date, "-", short_sha, "-", sep=""))
-
 ```
@@
 ```{r}
-dt <- simulate_opt(OPT, f, p, x_grid, h_grid, x0, z_g, profit)
+dt <- simulate_opt(OPT, f, p, x_grid, h_grid, x0, z_g, profit, OptTime=OptTime)
 sim_plots(dt, seed=seed_i)

### 2013-04-25 inst/examples/sensitivity.Rmd
tidying code

--- 2013-04-08 inst/examples/sensitivity.Rmd
@@
 cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-#sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
-#short_sha <- gsub("(^.{10}).*", "\\1", sha)
-#date <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S")
-#opts_chunk$set(fig.path = paste("figure/", date, "-", short_sha, "-", sep=""))
 ```

### 2013-04-25 inst/examples/BUGS/2013-04-19-bayesian-parametric-uncertainty-sdp.Rmd
updated examples

@@
 ---
+
+```{r libs, include=FALSE, cache=FALSE}
+library(pdgControl)
+library(nonparametricbayes)
+library(reshape2)
+library(ggplot2)
+library(knitcitations)
+```
+
+```{r setupall, echo=FALSE}
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE)
+opts_knit$set(upload.fun = socialR::flickr.url)
+theme_set(theme_bw(base_size=10))
+theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
+             plot.background = element_rect(fill = "transparent",colour = NA))
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+```
 
@@
 
-  # compute mean with monte carlo samples
-  mu <- sapply(indices, function(i) { # Cols are parameter samples
-    p <- unname(pardist[i,c(4,1,5)])          # Only, we need to add the measurement and process noise  
-    unname(sapply(x_grid, f, h, p))   # rows are possible values of x_t
-  })
-  mu <- data.frame(t(mu)) # Let's have cols as possible x_t, rows as parameters, and type as data.frame
+  # Compute the matrix
+  F_true <- 
+    sapply(x_grid, function(x_t){           # For each x_t
 
-  F_true <- 
-    sapply(mu, function(cur_state){    # For each x_t
+      bypar <- sapply(indices, function(i){ # For each parameter value
+        p <- unname(pardist[i,c(4,1,5)])    # parameters for the mean, at current sample 
+        mu <- f(x_t,h,p)
+        est_sigma_g <- pardist[i,2]         # Variance parameter
 
-      bypar <- sapply(cur_state, function(m){ # For each parameter value
-        if(snap_to_grid(m,x_grid) < x_grid[2]){ # 
+        if(snap_to_grid(mu,x_grid) < x_grid[2]){ # handle the degenerate case 
           out <- numeric(length(x_grid))
@@
         } else {
-          out <- dlnorm(x_grid/m, 0, sigma_g) # should be using the estimate process noise!
-        }})  
+          out <- dlnorm(x_grid/mu, 0, est_sigma_g) 
+        }
+      })  
       ave_over_pars <- apply(bypar, 1, sum) # collapse by weighted average over possible parameters    
@@
       })
-
   F_true <- t(F_true)
@@
 pardist[,4] = p[1] + rnorm(100, 0, 0.000001)
+pardist[,2] = sigma_g + rnorm(100, 0, 0.000001)
 pardist[,5] = p[3] + rnorm(100, 0, 0.000001)

### 2013-04-25 inst/examples/BUGS/allen.Rmd
updated examples

@@
 ```{r plotting-options, echo=FALSE, include=FALSE, cache=TRUE}
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE)
 opts_knit$set(upload.fun = socialR::flickr.url)
@@
 
+### Model and parameters
+
 ```{r}
@@
 p <- c(2, 10, 5) 
-K <- 10
-allee <- 5
+K <- p[2]
+allee <- p[3]
 ```
@@
 
+### Sample Data
+
 ```{r obs, dependson="sdp-pars"}
-obs <- sim_obs(Xo, z_g, f, p, Tobs=Tobs, nz=1, 
+obs <- sim_obs(Xo, z_g, f, p, Tobs=Tobs, nz=15, 
                 harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
 ```
+
+## Maximum Likelihood
 
@@
 ```
+
+## Non-parametric Bayes
+
+```{r gp-priors}
+#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
+s2.p <- c(5,5)  
+tau2.p <- c(5,1)
+d.p = c(10, 1/0.1, 10, 1/0.1)
+nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
+s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
+tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
+d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
+nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
+beta0_prior <- function(x, tau) dnorm(x, 0, tau)
+beta = c(0)
+priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
+```
+
+```{r}
+  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
+          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
+          m0r1=FALSE, corr="exp", trace=TRUE, 
+          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
+          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
+  gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
+```
+
+```{r}
+  posteriors_plot(gp, priors) # needs trace=TRUE!
+```
+
+#### Parametric Bayes
 
@@
 ```{r}
-jags.params=c("K","logr0","logtheta","iR","iQ"); # Don't need to save "x"
+jags.params=c("K","logr0","logtheta","iR","iQ") # Don't need to save "x"
 jags.inits <- function(){
@@
 ```
+
+#### Convergence diagnostics for parametric bayes
 
@@
 ```{r}
-theta <- exp(mcmcall[,"logtheta"])
-theta_dist <- hist(theta, freq=FALSE)
-theta_dist$mids
-theta_dist$density
-delta <- theta_dist$mids[2] - theta_dist$mids[1]
+MaxT = 1000
+matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.ks2, x_grid, h_grid)
+opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
+
+matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
+opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+
+matrices_estimated <- f_transition_matrix(est$f, est$p, x_grid, h_grid, est$sigma_g)
+opt_estimated <- value_iteration(matrices_estimated, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+
+matrices_alt <- f_transition_matrix(alt$f, alt$p, x_grid, h_grid, alt$sigma_g)
+opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+
+pardist <- mcmcall
+pardist[,4] = exp(pardist[,4])
+pardist[,5] = exp(pardist[,5])
+
+matrices_par_bayes <- parameter_uncertainty_SDP(f, p, x_grid, h_grid, pardist)
+opt_par_bayes <- value_iteration(matrices_par_bayes, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+
+OPT = list(gp_D = opt_gp$D, true_D = opt_true$D, est_D = opt_estimated$D, alt_D = opt_alt$D, par_bayes = opt_par_bayes$D)
 ```
 
-evaluating the value function given $f$ fixed at each of `theta_dist$mids` multiplied by  `theta_dist$density * delta` and summed over each $\theta$ is going to be slow for a single parameter but simply unrealistic for higher dimensions.
-Instead we will rely on Monte Carlo sampling of parameter values from their posteriors,
+```{r}
+policies <- melt(data.frame(stock=x_grid, 
+                            GP = x_grid[opt_gp$D], 
+                            MLE = x_grid[opt_estimated$D],
+                            True = x_grid[opt_true$D],
+                            Ricker_MLE = x_grid[opt_alt$D],
+                            Parametric.Bayes = x_grid[opt_par_bayes$D]),
+                   id="stock")
+names(policies) <- c("stock", "method", "value")
+policy_plot <- ggplot(policies, aes(stock, stock - value, color=method)) +
+  geom_line(lwd=1.2, alpha=0.8) + 
+  xlab("stock size") + ylab("escapement")  +
+  scale_colour_manual(values=cbPalette)
+```
 
 ```{r}
-mc_n <- 100
-thetas <- sample(theta, mc_n, replace=T)
+sims <- lapply(OPT, function(D){
+  set.seed(1)
+  lapply(1:100, function(i) 
+    ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
+  )
+})
+
+dat <- melt(sims, id=names(sims[[1]][[1]]))
+dt <- data.table(dat)
+setnames(dt, c("L1", "L2"), c("method", "reps")) 
+setkey(dt, method) # change the ordering
+
+ggplot(dt) + 
+    geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
+    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))
+
+ggplot(dt) + 
+    geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
+    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))
 

### 2013-04-25 inst/examples/BUGS/par_uncertainty_sdp.Rmd
updated examples

@@
-```{r}
-f_transition_matrix <- function(f, p, x_grid, h_grid, sigma_g, pardist){
-  lapply(h_grid, par_F, f, p, x_grid, sigma_g, pardist)
-}
+Run the Bayesian analysis to obtain posterior distributions for parameters.
 
-par_F <- function(h, f, p, x_grid, sigma_g, pardist, n_mc = 100){
-
-  # Set up monte carlo sampling 
-  d <- dim(pardist)
-  indices <- round(runif(n_mc,1, d[1]))
-
-  # compute mean with monte carlo samples
-  mu <- sapply(indices, function(i) { # Cols are parameter samples
-    p <- unname(pardist[i,c(4,1,5)])          # Only, we need to add the measurement and process noise  
-    unname(sapply(x_grid, f, h, p))   # rows are possible values of x_t
-  })
-  mu <- data.frame(t(mu)) # Let's have cols as possible x_t, rows as parameters, and type as data.frame
-
-  F_true <- 
-    sapply(mu, function(cur_state){    # For each x_t
-
-      bypar <- sapply(cur_state, function(m){ # For each parameter value
-        if(snap_to_grid(m,x_grid) < x_grid[2]){ # 
-          out <- numeric(length(x_grid))
-          out[1] <- 1
-          out
-        } else {
-          out <- dlnorm(x_grid/m, 0, sigma_g) # should be using the estimate process noise!
-        }})  
-      ave_over_pars <- apply(bypar, 1, sum) # collapse by weighted average over possible parameters    
-      ave_over_pars / sum(ave_over_pars)
-      })
-
-  F_true <- t(F_true)
-
-}
-
-snap_to_grid <- function(x, grid) sapply(x, function(x) grid[which.min(abs(grid - x))])
-
-# internal helper function
-rownorm <- function(M) 
-  t(apply(M, 1, function(x){ 
-    if(sum(x)>0){
-      x/sum(x)
-    } else {
-      out <- numeric(length(x))
-      out[1] <- 1
-      out
-    }
-  }))
+```{r, echo=TRUE, include=FALSE}
+knit("allen.Rmd")
 ```
@@
 ```{r}
-knit("allen.Rmd")
+pardist <- mcmcall
+```
 
-pardist <- mcmcall
-#pardist[,1] = p[2] + rnorm(100, 0, 0.000001)
-#pardist[,4] = p[1] + rnorm(100, 0, 0.000001)
-#pardist[,5] = p[3] + rnorm(100, 0, 0.000001)
+Transform parameters back
 
+```{r}
+pardist[,4] = exp(pardist[,4])
+pardist[,5] = exp(pardist[,5])
+```
+
+Compute optimal policy
+
+```{r}
 sdp = f_transition_matrix(f, p, x_grid, h_grid, sigma_g, pardist)
 s_opt <- value_iteration(sdp, x_grid, h_grid, OptTime=1000, xT, profit, delta)
-#opt <- find_dp_optim(sdp, x_grid, h_grid, OptTime, xT, profit, delta, reward=0)
+```
 
+Compare to the case without parameter uncertainty (growth noise only)
+
+```{r}
 SDP_Mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g)
 pars_fixed <- value_iteration(SDP_Mat, x_grid, h_grid, OptTime=1000, xT, profit, delta)
+```
 
+Plot results
+
+```{r}
 require(reshape2)
 policies <- melt(data.frame(stock=x_grid, pars.uncert = x_grid[s_opt$D], pars.fixed = x_grid[pars_fixed$D]), id="stock")
-
-ggplot(policies, aes(stock, value, color=variable)) + geom_point(alpha=.5) + xlab("stock size") + ylab("harvest")
-
 ggplot(policies, aes(stock, stock - value, color=variable)) + geom_line(alpha=1) + xlab("stock size") + ylab("escapement") 

### 2013-04-26 inst/examples/BUGS/allen.Rmd
different seed, replicate

@@
 
-```{r plotting-options, echo=FALSE, include=FALSE, cache=TRUE}
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE)
+```{r plotting-options}
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE)
 opts_knit$set(upload.fun = socialR::flickr.url)
@@
 xT <- 0
-seed_i <- 1
+seed_i <- 111
 Xo <- K # observations start from
@@
 ```{r}
-init_p = est$p
+# a bit unfair to start with the correct values, but anyhow...
+init_p = p
 names(init_p) = c("r0", "K", "theta")

### 2013-04-26 inst/examples/BUGS/allen.Rmd
parametric bayes initialized with true parameter values still performs poorly.

@@
 ```{r}
-  posteriors_plot(gp, priors) # needs trace=TRUE!
+hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
+posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
+ggplot(posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free")
+
+prior_curves <- ddply(posteriors, "variable", function(dd){
+    grid <- seq(min(dd$value), max(dd$value), length = 100)
+    data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
+})
+
+ggplot(posteriors) + geom_histogram(aes(x=value, y=..density..), alpha=0.7) +
+    geom_line(data=prior_curves, aes(x=value, y=density), col="red", lwd=2) +
+    facet_wrap(~ variable, scale="free")
 ```
@@
 print(densityplot(tfit_jags_m))
+
+par_posteriors <- as.data.frame(tfit_jags_m)
 ```

### 2013-04-27 inst/examples/BUGS/allen.Rmd
example updated

--- 2013-04-26 inst/examples/BUGS/allen.Rmd
@@
-```{r libraries, include=FALSE, cache=FALSE}
-library(pdgControl)
-library(nonparametricbayes)
-library(reshape2)
-library(ggplot2)
-library(data.table)
-library(tgp)
-library(MCMCpack)
-library(plyr)
-library(knitcitations)
+# Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach
+
+Load necessary libraries,
+
+```{r libraries, cache=FALSE}
+library(pdgControl) # custom SDP functions
+library(nonparametricbayes) # custom 
+library(reshape2) # data manipulation
+library(data.table) # data manipulation
+library(plyr) # data manipulation
+library(ggplot2) # plotting 
+library(tgp) # Gaussian Processes
+library(MCMCpack) # Markov Chain Monte Carlo tools 
+library(R2jags) # Markov Chain Monte Carlo tools 
+library(emdbook) # Markov Chain Monte Carlo tools 
+library(coda) # Markov Chain Monte Carlo tools 
 ```
 
+Plotting and knitr options, (can generally be ignored)
+
 ```{r plotting-options}
+library(knitcitations)
 opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE)
@@
 
+Uses the model derived in `r citet("")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
+
 ```{r}
@@
 ```
+
+Various parameters defining noise dynamics, grid, and policy costs.
 
@@
   gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
+          meanfn="constant", bprior="b0", BTE=c(2000,20000,2),
           m0r1=FALSE, corr="exp", trace=TRUE, 
@@
           s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
-  gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
 ```
@@
 ```{r}
-  posteriors_plot(gp, priors) # needs trace=TRUE!
+hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
+posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
+
+# Traces 
+ggplot(posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
+
+prior_curves <- ddply(posteriors, "variable", function(dd){
+    grid <- seq(min(dd$value), max(dd$value), length = 100)
+    data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
+})
+
+# 
+ggplot(posteriors) + geom_histogram(aes(x=value, y=..density..), alpha=0.7) +
+    geom_line(data=prior_curves, aes(x=value, y=density), col="red", lwd=2) +
+    facet_wrap(~ variable, scale="free")
+
+ggplot(posteriors, aes(value)) + stat_density(geom="path", position="identity", alpha=0.7) +
+  geom_line(data=prior_curves, aes(x=value, y=density), col="red") + 
+  facet_wrap(~ variable, scale="free", ncol=2)
 ```
@@
 
-```{r}
-library(R2jags)
-```
+We initiate the MCMC chain (`init_p`) using the true values of the parameters `p` from the simulation.
+While impossible in real data, this gives the parametric Bayesian approach the best chance at succeeding.
+`y` is the timeseries (recall `obs` has the $x_t$, $x_{t+1}$ pairs)
 
@@
 
+We'll be using the JAGS Gibbs sampler, a recent open source BUGS implementation with an R interface that works on most platforms.  We initialize the usual MCMC parameters; see `?jags` for details.
+
 ```{r}
@@
 
-## Gamma priors on precision
+The actual model is defined in a `model.file` that contains an R function that is automatically translated into BUGS code by *R2WinBUGS*.
+The file defines the priors and the model, as seen when read in here
+
+```{r}
+cat(readLines(con="bugmodel-GammaPrior.txt"), sep="\n")
+```
+
+We define which parameters to keep track of, and set the initial values of parameters in the transformed space used by the MCMC.
+We use logarithms to maintain strictly positive values of parameters where appropriate.
+Because our priors on the noise parameters are inverse gamma distributed.
 
@@
 ```{r}
-library(emdbook)
-library(coda)
-tfit_jags_m <- as.mcmc.bugs(jagsfit$BUGSoutput)
-print(xyplot(tfit_jags_m))
-print(densityplot(tfit_jags_m))
+jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
+par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
+
+# Traces
+ggplot(par_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
+
+# posterior distributiosn
+ggplot(par_posteriors, aes(value)) + stat_density(geom="path", position="identity", alpha=0.7) +
+  facet_wrap(~ variable, scale="free", ncol=2)
 ```
@@
 ```
+
+#### Phase-space diagram of the expected dynamics
+
+```{r}
+  gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
+```
+
+## Optimal policies by value iteration
+
+Compute the optimal policy under each model using stochastic dynamic programming.
 
@@
 pardist <- mcmcall
-pardist[,4] = exp(pardist[,4])
+pardist[,4] = exp(pardist[,4]) # transform model parameters back first
 pardist[,5] = exp(pardist[,5])
@@
 
-OPT = list(gp_D = opt_gp$D, true_D = opt_true$D, est_D = opt_estimated$D, alt_D = opt_alt$D, par_bayes = opt_par_bayes$D)
+OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker.MLE = opt_alt$D, Parametric.Bayes = opt_par_bayes$D)
+colorkey=cbPalette
+names(colorkey) = names(OPT) 
 ```
 
+## Graph of the optimal policies
+
 ```{r}
-policies <- melt(data.frame(stock=x_grid, 
-                            GP = x_grid[opt_gp$D], 
-                            MLE = x_grid[opt_estimated$D],
-                            True = x_grid[opt_true$D],
-                            Ricker_MLE = x_grid[opt_alt$D],
-                            Parametric.Bayes = x_grid[opt_par_bayes$D]),
-                   id="stock")
+policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
 names(policies) <- c("stock", "method", "value")
-policy_plot <- ggplot(policies, aes(stock, stock - value, color=method)) +
-  geom_line(lwd=1.2, alpha=0.8) + 
-  xlab("stock size") + ylab("escapement")  +
-  scale_colour_manual(values=cbPalette)
+ggplot(policies, aes(stock, stock - value, color=method)) +
+  geom_line(lwd=1.2, alpha=0.8) + xlab("stock size") + ylab("escapement")  +
+  scale_colour_manual(values=colorkey)
 ```
 
+## Simulate 100 realizations managed under each of the policies
+
 ```{r}
+
 sims <- lapply(OPT, function(D){
@@
 setnames(dt, c("L1", "L2"), c("method", "reps")) 
-setkey(dt, method) # change the ordering
+# Legend in original ordering please, not alphabetical: 
+dt$method = factor(dt$method, ordered=TRUE, levels=names(OPT))
 
@@
   geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
-    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
 
@@
   geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
-    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
 

### 2013-04-27 R/gp-replicate-fns.R

@@ -130,7 +130,9 @@ simulate_opt <- function(OPT, f, p, x_grid, h_grid, x0, z_g, profit, OptTime){
 sim_plots <- function(dt, seed=1, harvest=FALSE){
   fish_plot <- ggplot(dt) + 
     geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
-    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))+
+    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1))) +
+    scale_colour_manual(values=c("GP" = cbPalette[1], "Parametric" = cbPalette[2], "True" = cbPalette[3], "Structural" = cbPalette[4]),  
+                        guide = guide_legend(override.aes = list(alpha = 1))) +
     ggtitle(paste("seed", seed))
   print(fish_plot)
   harvest_plot <- ggplot(dt) +

### 2013-04-29 inst/examples/BUGS/allen.Rmd
uniform prior on variances

@@
 
-```{r libraries, cache=FALSE}
+```{r libraries, cache=FALSE, message=FALSE, warning=FALSE}
 library(pdgControl) # custom SDP functions
@@
 
-Uses the model derived in `r citet("")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
+Uses the model derived in `r citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
 
@@
 ```{r}
-cat(readLines(con="bugmodel-GammaPrior.txt"), sep="\n")
+cat(readLines(con="bugmodel-UPrior.txt"), sep="\n")
 ```
@@
 jags.params=c("K","logr0","logtheta","iR","iQ") # Don't need to save "x"
+#jags.inits <- function(){ list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]),"iQ"=1/0.05,"iR"=1/0.1,"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123) }
+
+jags.params=c("K","logr0","logtheta","stdQ", "stdR")
 jags.inits <- function(){
-  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]),"iQ"=1/0.05,"iR"=1/0.1,"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
 }
+
 set.seed(12345)
+
 time<-system.time(       
   jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
-                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="bugmodel-GammaPrior.txt")
+                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="bugmodel-UPrior.txt")
 )         
@@
 colnames(mcmcall) <- who
+
+pardist <- mcmcall
+pardist[,4] = exp(pardist[,4]) # transform model parameters back first
+pardist[,5] = exp(pardist[,5])
+
+bayes_coef <- apply(mcmcall,2,mean)
+bayes_pars <- c(bayes_coef[4], bayes_coef[1], bayes_coef[5])
+
 ```
 
-#### Phase-space diagram of the expected dynamics
+### Phase-space diagram of the expected dynamics
 
 ```{r}
-  gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
+# Summarize the GP model
+tgp_dat <- 
+    data.frame(  x = gp$XX[[1]], 
+                 y = gp$ZZ.km, 
+                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
+                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2),
+                 ymin2 = gp$ZZ.mean - 1.96 * sqrt(gp$ZZ.vark), 
+                 ymax2 = gp$ZZ.mean + 1.96 * sqrt(gp$ZZ.vark))
+
+  true_means <- sapply(x_grid, f, 0, p)
+  alt_means <- sapply(x_grid, alt$f, 0, alt$p)
+  est_means <- sapply(x_grid, est$f, 0, est$p)
+  par_bayes_means <- sapply(x_grid, est$f, 0, bayes_pars)
+
+models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
+                     MLE=est_means, Ricker.MLE=alt_means, 
+                     Parametric.Bayes = par_bayes_means)
+
+  models <- melt(models, id="x")
+  names(models) <- c("x", "method", "value")
+
+  plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
+    geom_ribbon(aes(x,y,ymin=ymin2,ymax=ymax2), fill="gray60") +
+    geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
+    geom_point(data=obs, aes(x,y), alpha=0.8) + 
+    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
+    scale_colour_manual(values=cbPalette) 
+  print(plot_gp)
 ```
@@
 opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-pardist <- mcmcall
-pardist[,4] = exp(pardist[,4]) # transform model parameters back first
-pardist[,5] = exp(pardist[,5])
 

### 2013-04-29 inst/examples/BUGS/allen.Rmd
updating uniform priors example

@@
 # Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach
-
-Load necessary libraries,
-
-```{r libraries, cache=FALSE, message=FALSE, warning=FALSE}
-library(pdgControl) # custom SDP functions
-library(nonparametricbayes) # custom 
-library(reshape2) # data manipulation
-library(data.table) # data manipulation
-library(plyr) # data manipulation
-library(ggplot2) # plotting 
-library(tgp) # Gaussian Processes
-library(MCMCpack) # Markov Chain Monte Carlo tools 
-library(R2jags) # Markov Chain Monte Carlo tools 
-library(emdbook) # Markov Chain Monte Carlo tools 
-library(coda) # Markov Chain Monte Carlo tools 
-```
 
@@
 ```{r plotting-options}
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA)
+
 library(knitcitations)
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE)
+
+library(ggplot2) # plotting 
 opts_knit$set(upload.fun = socialR::flickr.url)
 theme_set(theme_bw(base_size=10))
-theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
-             plot.background = element_rect(fill = "transparent",colour = NA))
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
+             plot.background = element_rect(fill = "transparent", colour = NA))
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+```
+Load necessary libraries,
+
+```{r libraries, cache=FALSE, message=FALSE, warning=FALSE}
+library(nonparametricbayes) # loads the rest as dependencies
 ```
@@
 
+Estimate the Gaussian Process (nonparametric Bayesian fit)
+
 ```{r}
   gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(2000,20000,2),
+          meanfn="constant", bprior="b0", BTE=c(2000,10000,10),
           m0r1=FALSE, corr="exp", trace=TRUE, 
@@
 ```
+
+Show traces and posteriors against priors
 
@@
 
+## A list of functions has been defined in "priors", 
+## Here we just make sure to compute priors over the range of values returned by the posteriors
 prior_curves <- ddply(posteriors, "variable", function(dd){
@@
 })
+```
 
-# 
-ggplot(posteriors) + geom_histogram(aes(x=value, y=..density..), alpha=0.7) +
-    geom_line(data=prior_curves, aes(x=value, y=density), col="red", lwd=2) +
-    facet_wrap(~ variable, scale="free")
+(Note that I keep thinking this should be done with `stat_function`, but cannot `facet_wrap` across different functions, see [SO/1376967](http://stackoverflow.com/questions/1376967/using-stat-function-and-facet-wrap-together-in-ggplot2-in-r)
 
-ggplot(posteriors, aes(value)) + stat_density(geom="path", position="identity", alpha=0.7) +
+```{r}
+# Posteriors (easier to read than histograms)
+ggplot(posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
   geom_line(data=prior_curves, aes(x=value, y=density), col="red") + 
@@
 jags.params=c("K","logr0","logtheta","iR","iQ") # Don't need to save "x"
-#jags.inits <- function(){ list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]),"iQ"=1/0.05,"iR"=1/0.1,"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123) }
-
+# Uniform priors on standard deviation terms
 jags.params=c("K","logr0","logtheta","stdQ", "stdR")
@@
 
-time<-system.time(       
+time_jags <- system.time(       
   jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
@@
 
-# posterior distributiosn
-ggplot(par_posteriors, aes(value)) + stat_density(geom="path", position="identity", alpha=0.7) +
+## priors (untransformed variables)
+K_prior <- function(x) dunif(x, 1, 22000)
+logr_prior <- function(x) dunif(x, -4, 2)
+logtheta_prior <- function(x) dunif(x, -4, 2)
+stdQ_prior <- function(x) dunif(0, 100)
+stdR_prior <- function(x) dunif(0, 100)
+
+par_priors <- list(K = K_prior, logr0 = logr_prior, logtheta = logtheta_prior, 
+     stdQ = stdQ_prior, stdR = stdR_prior, deviance = function(x) 0)
+
+par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
+    grid <- seq(min(dd$value), max(dd$value), length = 100)
+    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
+})
+
+# posterior distributions
+ggplot(par_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
   facet_wrap(~ variable, scale="free", ncol=2)
@@
 ```{r}
+# um, cleaner if we were just be using the long form, par_posterior
 mcmc <- as.mcmc(jagsfit)
@@
 pardist <- mcmcall
-pardist[,4] = exp(pardist[,4]) # transform model parameters back first
-pardist[,5] = exp(pardist[,5])
+pardist[,2] = exp(pardist[,2]) # transform model parameters back first
+pardist[,3] = exp(pardist[,3])
 
 bayes_coef <- apply(mcmcall,2,mean)
-bayes_pars <- c(bayes_coef[4], bayes_coef[1], bayes_coef[5])
-
+bayes_pars <- c(bayes_coef[2], bayes_coef[1], bayes_coef[3])
+bayes_pars
 ```

### 2013-04-29 inst/examples/BUGS/allen.Rmd
runs for farm

@@
   gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(2000,10000,10),
-          m0r1=FALSE, corr="exp", trace=TRUE, 
+          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
+          m0r1=FALSE, corr="exp", trace=FALSE, 
           beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
@@
 
-```{r}
+```{r, eval=FALSE}
 hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
@@
 })
-```
 
-(Note that I keep thinking this should be done with `stat_function`, but cannot `facet_wrap` across different functions, see [SO/1376967](http://stackoverflow.com/questions/1376967/using-stat-function-and-facet-wrap-together-in-ggplot2-in-r)
-
-```{r}
 # Posteriors (easier to read than histograms)
@@
 ```
+
+(Note that I keep thinking this should be done with `stat_function`, but cannot `facet_wrap` across different functions, see [SO/1376967](http://stackoverflow.com/questions/1376967/using-stat-function-and-facet-wrap-together-in-ggplot2-in-r)
 
@@
 )         
-time <- unname(time["elapsed"]);
+time_jags <- unname(time_jags["elapsed"]);
 ```
@@
 
-bayes_coef <- apply(mcmcall,2,mean)
+bayes_coef <- apply(pardist,2,mean)
 bayes_pars <- c(bayes_coef[2], bayes_coef[1], bayes_coef[3])

### 2013-04-29 inst/examples/BUGS/allen.Rmd
mend
runs for farm

@@
 
-library(knitcitations)
+#library(knitcitations)
 
 library(ggplot2) # plotting 
-opts_knit$set(upload.fun = socialR::flickr.url)
+#opts_knit$set(upload.fun = socialR::flickr.url)
 theme_set(theme_bw(base_size=10))
@@
 
-Uses the model derived in `r citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
+Uses the model derived in `citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
 

### 2013-04-30 inst/examples/BUGS/allen.Rmd
exploring comparison cases

--- 2013-04-29 inst/examples/BUGS/allen.Rmd
@@
 # Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach
+
+```{r runsetup}
+setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
+```
 
@@
 f <- RickerAllee
-p <- c(2, 10, 5) 
+p <- c(.2, 7.5, 5)
 K <- p[2]
@@
 ```{r obs, dependson="sdp-pars"}
-obs <- sim_obs(Xo, z_g, f, p, Tobs=Tobs, nz=15, 
-                harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
+  #harvest <- sort(rep(seq(0, .5, length=7), 5))
+  x <- numeric(Tobs)
+  x[1] <- 6
+  nz <- 1
+  for(t in 1:(Tobs-1))
+    x[t+1] = z_g() * f(x[t], h=0, p=p)
+  obs <- data.frame(x = c(rep(0,nz), 
+                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
+                    y = c(rep(0,nz), 
+                          x[2:Tobs]))
 ```
@@
 ```{r}
-  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
-          m0r1=FALSE, corr="exp", trace=FALSE, 
+gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, zcov=TRUE, verb=0,
+          meanfn="constant", bprior="b0", BTE=c(20,16000,2),
+          m0r1=FALSE, corr="exp", trace=TRUE, 
           beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
@@
 
-#### Parametric Bayes
+### Parametric Bayes
 
@@
 names(init_p) = c("r0", "K", "theta")
-y <- obs$y[-1] 
+y <- obs$x[-1] 
 N=length(y);
@@
 n.chains = 1
-n.iter = 20000
-n.burnin = floor(n.iter/2)
+n.iter = 40000
+n.burnin = floor(10000)
 n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
@@
 ```{r}
-jags.params=c("K","logr0","logtheta","iR","iQ") # Don't need to save "x"
 # Uniform priors on standard deviation terms
@@
 ## priors (untransformed variables)
-K_prior <- function(x) dunif(x, 1, 22000)
-logr_prior <- function(x) dunif(x, -4, 2)
-logtheta_prior <- function(x) dunif(x, -4, 2)
-stdQ_prior <- function(x) dunif(0, 100)
-stdR_prior <- function(x) dunif(0, 100)
+K_prior <- function(x) dunif(x, 0.01, 20)
+logr_prior <- function(x) dunif(x, -5, 5)
+logtheta_prior <- function(x) dunif(x, -5, 5)
+stdQ_prior <- function(x) dunif(x, 0.001, 100)
+stdR_prior <- function(x) dunif(x, 0.001, 100)
 
-par_priors <- list(K = K_prior, logr0 = logr_prior, logtheta = logtheta_prior, 
-     stdQ = stdQ_prior, stdR = stdR_prior, deviance = function(x) 0)
+par_priors <- list(K = K_prior, deviance = function(x) 0 * x, logr0 = logr_prior, logtheta = logtheta_prior, stdQ = stdQ_prior, stdR = stdR_prior)
 
@@
   facet_wrap(~ variable, scale="free", ncol=2)
+
 ```
@@
 bayes_coef <- apply(pardist,2,mean)
-bayes_pars <- c(bayes_coef[2], bayes_coef[1], bayes_coef[3])
+bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3]))
 bayes_pars
+```
+
+```{r}
+par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)
 ```
@@
   est_means <- sapply(x_grid, est$f, 0, est$p)
-  par_bayes_means <- sapply(x_grid, est$f, 0, bayes_pars)
+  par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)
 
@@
                      Parametric.Bayes = par_bayes_means)
+
+models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, MLE=est_means, Ricker.MLE=alt_means, Parametric.Bayes = par_bayes_means)
 
@@
 MaxT = 1000
-matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.ks2, x_grid, h_grid)
+matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.vark, x_grid, h_grid)
 opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
@@
 names(policies) <- c("stock", "method", "value")
+
 ggplot(policies, aes(stock, stock - value, color=method)) +
@@
 
-ggplot(dt) + 
-  geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
-
 ```

### 2013-04-30 inst/examples/BUGS/nonparametric.Rmd
exploring comparison cases

--- 2013-04-29 inst/examples/BUGS/allen.Rmd
@@
-# Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach
-
-Plotting and knitr options, (can generally be ignored)
-
-```{r plotting-options}
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA)
-
-#library(knitcitations)
-
-library(ggplot2) # plotting 
-#opts_knit$set(upload.fun = socialR::flickr.url)
-theme_set(theme_bw(base_size=10))
-theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
-             plot.background = element_rect(fill = "transparent", colour = NA))
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
-               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-```
-Load necessary libraries,
-
-```{r libraries, cache=FALSE, message=FALSE, warning=FALSE}
-library(nonparametricbayes) # loads the rest as dependencies
-```
-
-### Model and parameters
-
-Uses the model derived in `citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
-
-```{r}
-f <- RickerAllee
-p <- c(2, 10, 5) 
-K <- p[2]
-allee <- p[3]
-```
-
-Various parameters defining noise dynamics, grid, and policy costs.
-
-```{r sdp-pars, dependson="stateeq"}
-sigma_g <- 0.05
-sigma_m <- 0.0
-z_g <- function() rlnorm(1, 0, sigma_g)
-z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
-x_grid <- seq(0, 1.5 * K, length=101)
-h_grid <- x_grid
-profit <- function(x,h) pmin(x, h)
-delta <- 0.01
-OptTime <- 20  # stationarity with unstable models is tricky thing
-reward <- 0
-xT <- 0
-seed_i <- 111
-Xo <- K # observations start from
-x0 <- Xo # simulation under policy starts from
-Tobs <- 35
-```
-
-### Sample Data
-
-```{r obs, dependson="sdp-pars"}
-obs <- sim_obs(Xo, z_g, f, p, Tobs=Tobs, nz=15, 
-                harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
-```
-
-## Maximum Likelihood
-
-```{r mle, dependson="obs"}
-alt <- par_est(obs,  init = c(r = p[1], 
-                              K = mean(obs$x[obs$x>0]), 
-                              s = sigma_g))
-est <- par_est_allee(obs, f, p,  
-                     init = c(r = p[1] + 1, 
-                              K = p[2] + 2, 
-                              C = p[3] + 2, 
-                              s = sigma_g))
-```
-
 ## Non-parametric Bayes
@@
   gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
-          m0r1=FALSE, corr="exp", trace=FALSE, 
+          meanfn="constant", bprior="b0", BTE=c(20,16000,2),
+          m0r1=FALSE, corr="exp", trace=F, 
           beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
@@
 (Note that I keep thinking this should be done with `stat_function`, but cannot `facet_wrap` across different functions, see [SO/1376967](http://stackoverflow.com/questions/1376967/using-stat-function-and-facet-wrap-together-in-ggplot2-in-r)
-
-#### Parametric Bayes
-
-We initiate the MCMC chain (`init_p`) using the true values of the parameters `p` from the simulation.
-While impossible in real data, this gives the parametric Bayesian approach the best chance at succeeding.
-`y` is the timeseries (recall `obs` has the $x_t$, $x_{t+1}$ pairs)
-
-```{r}
-# a bit unfair to start with the correct values, but anyhow...
-init_p = p
-names(init_p) = c("r0", "K", "theta")
-y <- obs$y[-1] 
-N=length(y);
-```
-
-We'll be using the JAGS Gibbs sampler, a recent open source BUGS implementation with an R interface that works on most platforms.  We initialize the usual MCMC parameters; see `?jags` for details.
-
-```{r}
-jags.data <- list("N","y")
-n.chains = 1
-n.iter = 20000
-n.burnin = floor(n.iter/2)
-n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
-```
-
-The actual model is defined in a `model.file` that contains an R function that is automatically translated into BUGS code by *R2WinBUGS*.
-The file defines the priors and the model, as seen when read in here
-
-```{r}
-cat(readLines(con="bugmodel-UPrior.txt"), sep="\n")
-```
-
-We define which parameters to keep track of, and set the initial values of parameters in the transformed space used by the MCMC.
-We use logarithms to maintain strictly positive values of parameters where appropriate.
-Because our priors on the noise parameters are inverse gamma distributed.
-
-```{r}
-jags.params=c("K","logr0","logtheta","iR","iQ") # Don't need to save "x"
-# Uniform priors on standard deviation terms
-jags.params=c("K","logr0","logtheta","stdQ", "stdR")
-jags.inits <- function(){
-  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
-}
-
-set.seed(12345)
-
-time_jags <- system.time(       
-  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
-                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="bugmodel-UPrior.txt")
-)         
-time_jags <- unname(time_jags["elapsed"]);
-```
-
-#### Convergence diagnostics for parametric bayes
-
-```{r}
-jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
-
-# Traces
-ggplot(par_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
-
-## priors (untransformed variables)
-K_prior <- function(x) dunif(x, 1, 22000)
-logr_prior <- function(x) dunif(x, -4, 2)
-logtheta_prior <- function(x) dunif(x, -4, 2)
-stdQ_prior <- function(x) dunif(0, 100)
-stdR_prior <- function(x) dunif(0, 100)
-
-par_priors <- list(K = K_prior, logr0 = logr_prior, logtheta = logtheta_prior, 
-     stdQ = stdQ_prior, stdR = stdR_prior, deviance = function(x) 0)
-
-par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100)
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-
-# posterior distributions
-ggplot(par_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=2)
-```
-
-```{r}
-# um, cleaner if we were just be using the long form, par_posterior
-mcmc <- as.mcmc(jagsfit)
-mcmcall <- mcmc[,-2]
-who <- colnames(mcmcall)
-who 
-mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4],mcmcall[,5])
-colnames(mcmcall) <- who
-
-pardist <- mcmcall
-pardist[,2] = exp(pardist[,2]) # transform model parameters back first
-pardist[,3] = exp(pardist[,3])
-
-bayes_coef <- apply(pardist,2,mean)
-bayes_pars <- c(bayes_coef[2], bayes_coef[1], bayes_coef[3])
-bayes_pars
-```
-
-### Phase-space diagram of the expected dynamics
-
-```{r}
-# Summarize the GP model
-tgp_dat <- 
-    data.frame(  x = gp$XX[[1]], 
-                 y = gp$ZZ.km, 
-                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
-                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2),
-                 ymin2 = gp$ZZ.mean - 1.96 * sqrt(gp$ZZ.vark), 
-                 ymax2 = gp$ZZ.mean + 1.96 * sqrt(gp$ZZ.vark))
-
-  true_means <- sapply(x_grid, f, 0, p)
-  alt_means <- sapply(x_grid, alt$f, 0, alt$p)
-  est_means <- sapply(x_grid, est$f, 0, est$p)
-  par_bayes_means <- sapply(x_grid, est$f, 0, bayes_pars)
-
-models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
-                     MLE=est_means, Ricker.MLE=alt_means, 
-                     Parametric.Bayes = par_bayes_means)
-
-  models <- melt(models, id="x")
-  names(models) <- c("x", "method", "value")
-
-  plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-    geom_ribbon(aes(x,y,ymin=ymin2,ymax=ymax2), fill="gray60") +
-    geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
-    geom_point(data=obs, aes(x,y), alpha=0.8) + 
-    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
-    scale_colour_manual(values=cbPalette) 
-  print(plot_gp)
-```
-
-## Optimal policies by value iteration
-
-Compute the optimal policy under each model using stochastic dynamic programming.
-
-```{r}
-MaxT = 1000
-matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.ks2, x_grid, h_grid)
-opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
-
-matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-matrices_estimated <- f_transition_matrix(est$f, est$p, x_grid, h_grid, est$sigma_g)
-opt_estimated <- value_iteration(matrices_estimated, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-matrices_alt <- f_transition_matrix(alt$f, alt$p, x_grid, h_grid, alt$sigma_g)
-opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-matrices_par_bayes <- parameter_uncertainty_SDP(f, p, x_grid, h_grid, pardist)
-opt_par_bayes <- value_iteration(matrices_par_bayes, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker.MLE = opt_alt$D, Parametric.Bayes = opt_par_bayes$D)
-colorkey=cbPalette
-names(colorkey) = names(OPT) 
-```
-
-## Graph of the optimal policies
-
-```{r}
-policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
-names(policies) <- c("stock", "method", "value")
-ggplot(policies, aes(stock, stock - value, color=method)) +
-  geom_line(lwd=1.2, alpha=0.8) + xlab("stock size") + ylab("escapement")  +
-  scale_colour_manual(values=colorkey)
-```
-
-## Simulate 100 realizations managed under each of the policies
-
-```{r}
-
-sims <- lapply(OPT, function(D){
-  set.seed(1)
-  lapply(1:100, function(i) 
-    ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
-  )
-})
-
-dat <- melt(sims, id=names(sims[[1]][[1]]))
-dt <- data.table(dat)
-setnames(dt, c("L1", "L2"), c("method", "reps")) 
-# Legend in original ordering please, not alphabetical: 
-dt$method = factor(dt$method, ordered=TRUE, levels=names(OPT))
-
-ggplot(dt) + 
-  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
-
-ggplot(dt) + 
-  geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
-
-```

### 2013-04-30 inst/examples/BUGS/parametric.Rmd
exploring comparison cases

--- 2013-04-30 inst/examples/BUGS/allen.Rmd
@@
-# Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach
-
-```{r runsetup}
-setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
-```
-
-Plotting and knitr options, (can generally be ignored)
-
-```{r plotting-options}
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA)
-
-library(knitcitations)
-
-library(ggplot2) # plotting 
-opts_knit$set(upload.fun = socialR::flickr.url)
-theme_set(theme_bw(base_size=10))
-theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
-             plot.background = element_rect(fill = "transparent", colour = NA))
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
-               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-```
-Load necessary libraries,
-
-```{r libraries, cache=FALSE, message=FALSE, warning=FALSE}
-library(nonparametricbayes) # loads the rest as dependencies
-```
-
-### Model and parameters
-
-Uses the model derived in `citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
-
-```{r}
-f <- RickerAllee
-p <- c(.2, 7.5, 5)
-K <- p[2]
-allee <- p[3]
-```
-
-Various parameters defining noise dynamics, grid, and policy costs.
-
-```{r sdp-pars, dependson="stateeq"}
-sigma_g <- 0.05
-sigma_m <- 0.0
-z_g <- function() rlnorm(1, 0, sigma_g)
-z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
-x_grid <- seq(0, 1.5 * K, length=101)
-h_grid <- x_grid
-profit <- function(x,h) pmin(x, h)
-delta <- 0.01
-OptTime <- 20  # stationarity with unstable models is tricky thing
-reward <- 0
-xT <- 0
-seed_i <- 111
-Xo <- K # observations start from
-x0 <- Xo # simulation under policy starts from
-Tobs <- 35
-```
-
-### Sample Data
-
-```{r obs, dependson="sdp-pars"}
-  #harvest <- sort(rep(seq(0, .5, length=7), 5))
-  x <- numeric(Tobs)
-  x[1] <- 6
-  nz <- 1
-  for(t in 1:(Tobs-1))
-    x[t+1] = z_g() * f(x[t], h=0, p=p)
-  obs <- data.frame(x = c(rep(0,nz), 
-                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
-                    y = c(rep(0,nz), 
-                          x[2:Tobs]))
-```
-
-## Maximum Likelihood
-
-```{r mle, dependson="obs"}
-alt <- par_est(obs,  init = c(r = p[1], 
-                              K = mean(obs$x[obs$x>0]), 
-                              s = sigma_g))
-est <- par_est_allee(obs, f, p,  
-                     init = c(r = p[1] + 1, 
-                              K = p[2] + 2, 
-                              C = p[3] + 2, 
-                              s = sigma_g))
-```
-
-## Non-parametric Bayes
-
-```{r gp-priors}
-#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
-s2.p <- c(5,5)  
-tau2.p <- c(5,1)
-d.p = c(10, 1/0.1, 10, 1/0.1)
-nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
-s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
-tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
-d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
-nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
-beta0_prior <- function(x, tau) dnorm(x, 0, tau)
-beta = c(0)
-priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
-```
-
-Estimate the Gaussian Process (nonparametric Bayesian fit)
-
-```{r}
-gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, zcov=TRUE, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(20,16000,2),
-          m0r1=FALSE, corr="exp", trace=TRUE, 
-          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
-          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
-```
-
-Show traces and posteriors against priors
-
-```{r, eval=FALSE}
-hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
-posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
-
-# Traces 
-ggplot(posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
-
-## A list of functions has been defined in "priors", 
-## Here we just make sure to compute priors over the range of values returned by the posteriors
-prior_curves <- ddply(posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100)
-    data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
-})
-
-# Posteriors (easier to read than histograms)
-ggplot(posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=prior_curves, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=2)
-```
-
-(Note that I keep thinking this should be done with `stat_function`, but cannot `facet_wrap` across different functions, see [SO/1376967](http://stackoverflow.com/questions/1376967/using-stat-function-and-facet-wrap-together-in-ggplot2-in-r)
-
-### Parametric Bayes
-
 We initiate the MCMC chain (`init_p`) using the true values of the parameters `p` from the simulation.
@@
 ```
-
-### Phase-space diagram of the expected dynamics
-
-```{r}
-# Summarize the GP model
-tgp_dat <- 
-    data.frame(  x = gp$XX[[1]], 
-                 y = gp$ZZ.km, 
-                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
-                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2),
-                 ymin2 = gp$ZZ.mean - 1.96 * sqrt(gp$ZZ.vark), 
-                 ymax2 = gp$ZZ.mean + 1.96 * sqrt(gp$ZZ.vark))
-
-  true_means <- sapply(x_grid, f, 0, p)
-  alt_means <- sapply(x_grid, alt$f, 0, alt$p)
-  est_means <- sapply(x_grid, est$f, 0, est$p)
-  par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)
-
-models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
-                     MLE=est_means, Ricker.MLE=alt_means, 
-                     Parametric.Bayes = par_bayes_means)
-
-models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, MLE=est_means, Ricker.MLE=alt_means, Parametric.Bayes = par_bayes_means)
-
-models <- melt(models, id="x")
-names(models) <- c("x", "method", "value")
-
-plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-    geom_ribbon(aes(x,y,ymin=ymin2,ymax=ymax2), fill="gray60") +
-    geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
-    geom_point(data=obs, aes(x,y), alpha=0.8) + 
-    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
-    scale_colour_manual(values=cbPalette) 
-print(plot_gp)
-```
-
-## Optimal policies by value iteration
-
-Compute the optimal policy under each model using stochastic dynamic programming.
-
-```{r}
-MaxT = 1000
-matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.vark, x_grid, h_grid)
-opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
-
-matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-matrices_estimated <- f_transition_matrix(est$f, est$p, x_grid, h_grid, est$sigma_g)
-opt_estimated <- value_iteration(matrices_estimated, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-matrices_alt <- f_transition_matrix(alt$f, alt$p, x_grid, h_grid, alt$sigma_g)
-opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-matrices_par_bayes <- parameter_uncertainty_SDP(f, p, x_grid, h_grid, pardist)
-opt_par_bayes <- value_iteration(matrices_par_bayes, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker.MLE = opt_alt$D, Parametric.Bayes = opt_par_bayes$D)
-colorkey=cbPalette
-names(colorkey) = names(OPT) 
-```
-
-## Graph of the optimal policies
-
-```{r}
-policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
-names(policies) <- c("stock", "method", "value")
-
-ggplot(policies, aes(stock, stock - value, color=method)) +
-  geom_line(lwd=1.2, alpha=0.8) + xlab("stock size") + ylab("escapement")  +
-  scale_colour_manual(values=colorkey)
-```
-
-## Simulate 100 realizations managed under each of the policies
-
-```{r}
-
-sims <- lapply(OPT, function(D){
-  set.seed(1)
-  lapply(1:100, function(i) 
-    ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
-  )
-})
-
-dat <- melt(sims, id=names(sims[[1]][[1]]))
-dt <- data.table(dat)
-setnames(dt, c("L1", "L2"), c("method", "reps")) 
-# Legend in original ordering please, not alphabetical: 
-dt$method = factor(dt$method, ordered=TRUE, levels=names(OPT))
-
-ggplot(dt) + 
-  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
-
-```

### 2013-04-30 inst/examples/BUGS/setup.Rmd
exploring comparison cases

--- 2013-04-30 inst/examples/BUGS/allen.Rmd
@@
-# Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach
-
-```{r runsetup}
-setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
-```
-
 Plotting and knitr options, (can generally be ignored)
@@
 ```
-
-## Maximum Likelihood
-
-```{r mle, dependson="obs"}
-alt <- par_est(obs,  init = c(r = p[1], 
-                              K = mean(obs$x[obs$x>0]), 
-                              s = sigma_g))
-est <- par_est_allee(obs, f, p,  
-                     init = c(r = p[1] + 1, 
-                              K = p[2] + 2, 
-                              C = p[3] + 2, 
-                              s = sigma_g))
-```
-
-## Non-parametric Bayes
-
-```{r gp-priors}
-#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
-s2.p <- c(5,5)  
-tau2.p <- c(5,1)
-d.p = c(10, 1/0.1, 10, 1/0.1)
-nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
-s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
-tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
-d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
-nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
-beta0_prior <- function(x, tau) dnorm(x, 0, tau)
-beta = c(0)
-priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
-```
-
-Estimate the Gaussian Process (nonparametric Bayesian fit)
-
-```{r}
-gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, zcov=TRUE, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(20,16000,2),
-          m0r1=FALSE, corr="exp", trace=TRUE, 
-          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
-          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
-```
-
-Show traces and posteriors against priors
-
-```{r, eval=FALSE}
-hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
-posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
-
-# Traces 
-ggplot(posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
-
-## A list of functions has been defined in "priors", 
-## Here we just make sure to compute priors over the range of values returned by the posteriors
-prior_curves <- ddply(posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100)
-    data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
-})
-
-# Posteriors (easier to read than histograms)
-ggplot(posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=prior_curves, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=2)
-```
-
-(Note that I keep thinking this should be done with `stat_function`, but cannot `facet_wrap` across different functions, see [SO/1376967](http://stackoverflow.com/questions/1376967/using-stat-function-and-facet-wrap-together-in-ggplot2-in-r)
-
-### Parametric Bayes
-
-We initiate the MCMC chain (`init_p`) using the true values of the parameters `p` from the simulation.
-While impossible in real data, this gives the parametric Bayesian approach the best chance at succeeding.
-`y` is the timeseries (recall `obs` has the $x_t$, $x_{t+1}$ pairs)
-
-```{r}
-# a bit unfair to start with the correct values, but anyhow...
-init_p = p
-names(init_p) = c("r0", "K", "theta")
-y <- obs$x[-1] 
-N=length(y);
-```
-
-We'll be using the JAGS Gibbs sampler, a recent open source BUGS implementation with an R interface that works on most platforms.  We initialize the usual MCMC parameters; see `?jags` for details.
-
-```{r}
-jags.data <- list("N","y")
-n.chains = 1
-n.iter = 40000
-n.burnin = floor(10000)
-n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
-```
-
-The actual model is defined in a `model.file` that contains an R function that is automatically translated into BUGS code by *R2WinBUGS*.
-The file defines the priors and the model, as seen when read in here
-
-```{r}
-cat(readLines(con="bugmodel-UPrior.txt"), sep="\n")
-```
-
-We define which parameters to keep track of, and set the initial values of parameters in the transformed space used by the MCMC.
-We use logarithms to maintain strictly positive values of parameters where appropriate.
-Because our priors on the noise parameters are inverse gamma distributed.
-
-```{r}
-# Uniform priors on standard deviation terms
-jags.params=c("K","logr0","logtheta","stdQ", "stdR")
-jags.inits <- function(){
-  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
-}
-
-set.seed(12345)
-
-time_jags <- system.time(       
-  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
-                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="bugmodel-UPrior.txt")
-)         
-time_jags <- unname(time_jags["elapsed"]);
-```
-
-#### Convergence diagnostics for parametric bayes
-
-```{r}
-jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
-
-# Traces
-ggplot(par_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
-
-## priors (untransformed variables)
-K_prior <- function(x) dunif(x, 0.01, 20)
-logr_prior <- function(x) dunif(x, -5, 5)
-logtheta_prior <- function(x) dunif(x, -5, 5)
-stdQ_prior <- function(x) dunif(x, 0.001, 100)
-stdR_prior <- function(x) dunif(x, 0.001, 100)
-
-par_priors <- list(K = K_prior, deviance = function(x) 0 * x, logr0 = logr_prior, logtheta = logtheta_prior, stdQ = stdQ_prior, stdR = stdR_prior)
-
-par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100) 
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-
-# posterior distributions
-ggplot(par_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=2)
-
-```
-
-```{r}
-# um, cleaner if we were just be using the long form, par_posterior
-mcmc <- as.mcmc(jagsfit)
-mcmcall <- mcmc[,-2]
-who <- colnames(mcmcall)
-who 
-mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4],mcmcall[,5])
-colnames(mcmcall) <- who
-
-pardist <- mcmcall
-pardist[,2] = exp(pardist[,2]) # transform model parameters back first
-pardist[,3] = exp(pardist[,3])
-
-bayes_coef <- apply(pardist,2,mean)
-bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3]))
-bayes_pars
-```
-
-```{r}
-par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)
-```
-
-### Phase-space diagram of the expected dynamics
-
-```{r}
-# Summarize the GP model
-tgp_dat <- 
-    data.frame(  x = gp$XX[[1]], 
-                 y = gp$ZZ.km, 
-                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
-                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2),
-                 ymin2 = gp$ZZ.mean - 1.96 * sqrt(gp$ZZ.vark), 
-                 ymax2 = gp$ZZ.mean + 1.96 * sqrt(gp$ZZ.vark))
-
-  true_means <- sapply(x_grid, f, 0, p)
-  alt_means <- sapply(x_grid, alt$f, 0, alt$p)
-  est_means <- sapply(x_grid, est$f, 0, est$p)
-  par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)
-
-models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
-                     MLE=est_means, Ricker.MLE=alt_means, 
-                     Parametric.Bayes = par_bayes_means)
-
-models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, MLE=est_means, Ricker.MLE=alt_means, Parametric.Bayes = par_bayes_means)
-
-models <- melt(models, id="x")
-names(models) <- c("x", "method", "value")
-
-plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-    geom_ribbon(aes(x,y,ymin=ymin2,ymax=ymax2), fill="gray60") +
-    geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
-    geom_point(data=obs, aes(x,y), alpha=0.8) + 
-    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
-    scale_colour_manual(values=cbPalette) 
-print(plot_gp)
-```
-
-## Optimal policies by value iteration
-
-Compute the optimal policy under each model using stochastic dynamic programming.
-
-```{r}
-MaxT = 1000
-matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.vark, x_grid, h_grid)
-opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
-
-matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-matrices_estimated <- f_transition_matrix(est$f, est$p, x_grid, h_grid, est$sigma_g)
-opt_estimated <- value_iteration(matrices_estimated, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-matrices_alt <- f_transition_matrix(alt$f, alt$p, x_grid, h_grid, alt$sigma_g)
-opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-matrices_par_bayes <- parameter_uncertainty_SDP(f, p, x_grid, h_grid, pardist)
-opt_par_bayes <- value_iteration(matrices_par_bayes, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker.MLE = opt_alt$D, Parametric.Bayes = opt_par_bayes$D)
-colorkey=cbPalette
-names(colorkey) = names(OPT) 
-```
-
-## Graph of the optimal policies
-
-```{r}
-policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
-names(policies) <- c("stock", "method", "value")
-
-ggplot(policies, aes(stock, stock - value, color=method)) +
-  geom_line(lwd=1.2, alpha=0.8) + xlab("stock size") + ylab("escapement")  +
-  scale_colour_manual(values=colorkey)
-```
-
-## Simulate 100 realizations managed under each of the policies
-
-```{r}
-
-sims <- lapply(OPT, function(D){
-  set.seed(1)
-  lapply(1:100, function(i) 
-    ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
-  )
-})
-
-dat <- melt(sims, id=names(sims[[1]][[1]]))
-dt <- data.table(dat)
-setnames(dt, c("L1", "L2"), c("method", "reps")) 
-# Legend in original ordering please, not alphabetical: 
-dt$method = factor(dt$method, ordered=TRUE, levels=names(OPT))
-
-ggplot(dt) + 
-  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
-
-```

### 2013-04-30 inst/examples/gaussian-process-basics.Rmd
include bayesian mcmc example

@@
-`ro comment=NA, message=FALSE, cache.path="gpb/", tidy=FALSE or`
-
-```{r include=FALSE}
-opts_knit$set(upload.fun = socialR::notebook.url)
-theme_update(plot.background = element_rect(fill = "transparent",colour = NA))
+```{r plotting-options}
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA)
+library(ggplot2) # plotting 
+opts_knit$set(upload.fun = socialR::flickr.url)
+theme_set(theme_bw(base_size=10))
+theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
+             plot.background = element_rect(fill = "transparent", colour = NA))
 ```
@@
 require(reshape2)
-require(ggplot2)
 ```
@@
 ```
+
+```{r}
+lpriors <- function(pars){
+   d.p <- c(5, 5)
+  s2.p <- c(5, 5)
+
+  prior <- unname(
+    dgamma(exp(pars[1]), d.p[1], scale = d.p[2]) *
+    dgamma(exp(pars[2]), s2.p[1], s2.p[2]) 
+#    dunif(exp(pars[1]), 0, 100) * 
+#    dunif(exp(pars[2]), 0, 100)
+  )
+
+  log(prior)
+}
+
+posterior <- function(pars, x, y){
+
+  l <- exp(pars[1])
+  sigma.n <- exp(pars[2])
+
+  cov <- function(X, Y) outer(X, Y, SE, l)
+  I <- diag(1, length(x))
+  K <- cov(x, x)
+
+  loglik <- - 0.5 * t(y) %*% solve(K + sigma.n^2 * I) %*% y -
+    log(det(K + sigma.n^2*I)) -
+    length(y) * log(2 * pi) / 2
+
+  loglik + lpriors(pars)
+}
+
+```
+
+```{r}
+posterior(log(pars), obs$x, obs$y)
+posterior(log(o$par), obs$x, obs$y)
+```
+
+```{r}
+require(mcmc)
+n <- 1e4
+out <- metrop(posterior, log(pars), n, x = obs$x, y = obs$y)
+out$accept
+```
+
+```{r}
+postdist <- cbind(index=1:n, as.data.frame(exp(out$batch)))
+names(postdist) <- c("index", names(pars))
+df <- melt(postdist, id="index")
+# TRACES
+ggplot(df) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
+
+ggplot(df) + geom_line(aes(index, log(value))) + facet_wrap(~ variable, scale="free", ncol=1)
+```
+
+```{r}
+
+d.p <- c(5, 5)
+s2.p <- c(5, 5)
+
+s2_prior <- function(x) dgamma(x, s2.p[1], s2.p[2])
+d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2])
+prior_fns <- list(l = d_prior, sigma.n = s2_prior)
+
+require(plyr)
+prior_curves <- ddply(df, "variable", function(dd){
+    grid <- seq(min(dd$value), max(dd$value), length = 100)
+    data.frame(value = grid, density = prior_fns[[dd$variable[1]]](grid))
+})
+
+# Posteriors (easier to read than histograms)
+ggplot(df, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+  geom_line(data=prior_curves, aes(x=value, y=density), col="red") + 
+  facet_wrap(~ variable, scale="free", ncol=2)
+```
+
+```{r}
+prior_curves <- ddply(df, "variable", function(dd){
+    grid <- seq(-4, 4, length = 100)
+    data.frame(value = grid, density = prior_fns[[dd$variable[1]]](grid))
+})
+
+ggplot(prior_curves, aes(value, density)) + 
+  geom_line(data=prior_curves, col="red") + 
+  facet_wrap(~ variable, ncol=1, scale="free_y")
+```

### 2013-04-30 inst/examples/BUGS/parametric.Rmd
breaking out sections from allen.Rmd for individual testing.

@@
   stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
+  #geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
   facet_wrap(~ variable, scale="free", ncol=2)

### 2013-04-30 inst/examples/BUGS/setup.Rmd
breaking out sections from allen.Rmd for individual testing.

@@
                           x[2:Tobs]))
+write.csv(obs, "obs.txt")
 ```

### 2013-05-01 inst/examples/BUGS/allen.Rmd
Example using custom GP routine.
In this case, both MLE methods fail, while the GP does fine and the parametric Bayesian method performs best.

@@
 f <- RickerAllee
-p <- c(.2, 7.5, 5)
+p <- c(2, 8, 5)
 K <- p[2]
@@
 delta <- 0.01
-OptTime <- 20  # stationarity with unstable models is tricky thing
+OptTime <- 30  # stationarity with unstable models is tricky thing
 reward <- 0
@@
 s2.p <- c(5,5)  
-tau2.p <- c(5,1)
-d.p = c(10, 1/0.1, 10, 1/0.1)
-nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
-s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
-tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
-d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
-nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
-beta0_prior <- function(x, tau) dnorm(x, 0, tau)
-beta = c(0)
-priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
+d.p = c(10, 1/0.1)
 ```
@@
 ```{r}
-gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, zcov=TRUE, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(20,16000,2),
-          m0r1=FALSE, corr="exp", trace=TRUE, 
-          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
-          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
+gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
+gp_dat <- gp_predict(gp, x_grid)
+
 ```
@@
 
-```{r, eval=FALSE}
-hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
-posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
-
-# Traces 
-ggplot(posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
-
-## A list of functions has been defined in "priors", 
-## Here we just make sure to compute priors over the range of values returned by the posteriors
-prior_curves <- ddply(posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100)
-    data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
-})
-
-# Posteriors (easier to read than histograms)
-ggplot(posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=prior_curves, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=2)
+```{r}
+plots <- summary_gp_mcmc(gp)
+plots[[1]]
+plots[[2]]
 ```
 
-(Note that I keep thinking this should be done with `stat_function`, but cannot `facet_wrap` across different functions, see [SO/1376967](http://stackoverflow.com/questions/1376967/using-stat-function-and-facet-wrap-together-in-ggplot2-in-r)
+```{r}
+# Summarize the GP model
+tgp_dat <- 
+    data.frame(  x = x_grid, 
+                 y = gp_dat$E_Ef, 
+                 ymin = gp_dat$E_Ef - 2 * sqrt(gp_dat$E_Vf), 
+                 ymax = gp_dat$E_Ef + 2 * sqrt(gp_dat$E_Vf) )
+```
 
@@
 ## priors (untransformed variables)
-K_prior <- function(x) dunif(x, 0.01, 20)
-logr_prior <- function(x) dunif(x, -5, 5)
-logtheta_prior <- function(x) dunif(x, -5, 5)
+K_prior <- function(x) dunif(x, 0.01, 40)
+logr_prior <- function(x) dunif(x, -6, 6)
+logtheta_prior <- function(x) dunif(x, -6, 6)
 stdQ_prior <- function(x) dunif(x, 0.001, 100)
@@
 ```{r}
-# Summarize the GP model
-tgp_dat <- 
-    data.frame(  x = gp$XX[[1]], 
-                 y = gp$ZZ.km, 
-                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
-                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2),
-                 ymin2 = gp$ZZ.mean - 1.96 * sqrt(gp$ZZ.vark), 
-                 ymax2 = gp$ZZ.mean + 1.96 * sqrt(gp$ZZ.vark))
 
@@
 
-models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, MLE=est_means, Ricker.MLE=alt_means, Parametric.Bayes = par_bayes_means)
-
 models <- melt(models, id="x")
@@
 plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-    geom_ribbon(aes(x,y,ymin=ymin2,ymax=ymax2), fill="gray60") +
     geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
@@
 MaxT = 1000
-matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.vark, x_grid, h_grid)
+matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)
 opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)

### 2013-05-01 R/gp_mcmc.R
package updates with the custom gp method

@@ -1,0 +1,148 @@
+#' Basic regression in Gaussian processes  
+#' 
+#' 
+#' @param x Observed x values, (vector or matrix with columns for each dimension of data)
+#' @param y Vector of observed y values in the training data
+#' @param init_pars the initial guesses for lengthscale l and process noise sigma_n
+#' @param n iterations of the metropolis algorithm
+#' @details Currently assumes the covariance function.  By default we will use 
+#' the squared exponential (also called radial basis or Gaussian, 
+#' though it is not this that gives Gaussian process it's name; 
+#' any covariance function would do) as the covariance function, 
+#' $\operatorname{cov}(X_i, X_j) = \exp\left(-\frac{(X_i - X_j)^2}{2 \ell ^ 2}\right)$
+#' @return the MCMC output, as constructed by \link{metrop} from the mcmc package.
+#' @export
+#' @import mcmc  
+gp_mcmc <- function(x, y, init_pars = c(l=1, sigma.n=1), n = 1e4, d.p = c(5,5), s2.p = c(5,5)){
+
+  lpriors <- function(pars){
+    prior <- unname(
+      dgamma(exp(pars[1]), d.p[1], scale = d.p[2]) *
+      dgamma(exp(pars[2]), s2.p[1], s2.p[2]) 
+    )
+    log(prior)
+  }
+  posterior <- function(pars, x, y){ 
+    l <- exp(pars[1])
+    sigma.n <- exp(pars[2])
+    
+    cov <- function(X, Y) outer(X, Y, SE, l)
+    I <- diag(1, length(x))
+    K <- cov(x, x) 
+    
+    loglik <- - 0.5 * t(y) %*% solve(K + sigma.n^2 * I) %*% y -
+      log(det(K + sigma.n^2*I)) -
+      length(y) * log(2 * pi) / 2
+    
+    loglik + lpriors(pars)
+  }
+  
+  out <- metrop(posterior, log(init_pars), n, x = obs$x, y = obs$y)  
+  out$d.p <- d.p
+  out$s2.p <- s2.p
+  
+  out
+}
+
+#' predict the expected values and posterior distributions of the Gaussian Process
+#' 
+#' @param x_predict the values at which we desire predictions
+#' @import reshape2
+#' @export
[…]

### 2013-05-01 R/gp_trasition_matrix.R

@@ -10,12 +10,26 @@
 gp_transition_matrix <- function(Ef, V, x_grid, h_grid=NULL){  
   if(is.null(h_grid))
     h_grid <- x_grid
+  
+  if(is.matrix(Ef)){ # loop over posteriors
+    n_samples <- dim(Ef)[2]
+    loop_over_h <- function(h){
+      F_s <- lapply(1:n_samples, function(i){
+        gp_calc_F(h, Ef[,i], V[,i], x_grid)
+      })
+      # Average across matrices in list, and then normalize
+    }
+    lapply(h_grid, loop_over_h)  
+  } else { 
   lapply(h_grid, gp_calc_F, Ef, V, x_grid)
+  }
 }
 
 
 
 gp_calc_F <- function(h, Ef, V, x_grid){
+  
+  
   j <- which.min(abs(x_grid - h))
   indices <- 1:length(Ef)
   shift <- pmax(indices-j,1)

### 2013-05-01 inst/examples/gp_mcmc_example.Rmd
Test examples demonstrating the custom mcmc GP

```{r plotting-options, message=FALSE, warning=FALSE, include=FALSE}
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA)
library(ggplot2) # plotting 
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=10))
theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
             plot.background = element_rect(fill = "transparent", colour = NA))

library(nonparametricbayes)
```

Basic regression in Gaussian processes  
---------------------------------------

Working out coded examples for basic Gaussian process regression using R.  I've just read through the first few chapters of [Rasmussen & Williams (2006)](http://www.GaussianProcess.org/gpml), this implements the examples discussed in Chapter 2.1-2.5.

<!-- Also motivated by [James Keirstead](http://www.jameskeirstead.ca/r/gaussian-process-regression-with-r/), with cleaner code.  -->

```{r}
x_predict <- seq(-5,5,len=50)
obs <- data.frame(x = c(-4, -3, -1,  0,  2),
                  y = c(-2,  0,  1,  2, -1))
```

```{r}
gp <- gp_mcmc(obs$x, obs$y, d.p = c(1,1), s2.p = c(1,1))
```

```{r}
summary_gp_mcmc(gp)
```

```{r}
gp_dat <- gp_predict(gp, x_predict)
```

```{r}
tgp_dat <- 
    data.frame(  x = x_predict, 
                 y = gp_dat$E_Ef, 
                 ymin = gp_dat$E_Ef - 2 * sqrt(gp_dat$E_Vf), 
                 ymax = gp_dat$E_Ef + 2 * sqrt(gp_dat$E_Vf) )

ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
    geom_line(aes(x, y), lwd=2, alpha=0.8) + 
    geom_point(data=obs, aes(x,y), alpha=0.8) + 
    xlab(expression(X[t])) + ylab(expression(X[t+1])) 
```

### 2013-05-01 inst/examples/BUGS/nonparametric.Rmd
adapted to the custom version

@@
 ## Non-parametric Bayes
+
+```{r}
+setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
+knit("setup.Rmd")
+```
 
@@
 ```{r}
-  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(20,16000,2),
-          m0r1=FALSE, corr="exp", trace=F, 
-          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
-          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
+gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
+gp_dat <- gp_predict(gp, x_grid)
+
 ```
@@
 
-```{r, eval=FALSE}
-hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
-posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
-
-# Traces 
-ggplot(posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
-
-## A list of functions has been defined in "priors", 
-## Here we just make sure to compute priors over the range of values returned by the posteriors
-prior_curves <- ddply(posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100)
-    data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
-})
-
-# Posteriors (easier to read than histograms)
-ggplot(posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=prior_curves, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=2)
+```{r}
+plots <- summary_gp_mcmc(gp)
+plots[[1]]
+plots[[2]]
 ```
 
-(Note that I keep thinking this should be done with `stat_function`, but cannot `facet_wrap` across different functions, see [SO/1376967](http://stackoverflow.com/questions/1376967/using-stat-function-and-facet-wrap-together-in-ggplot2-in-r)
+```{r}
+# Summarize the GP model
+tgp_dat <- 
+    data.frame(  x = x_grid, 
+                 y = gp_dat$E_Ef, 
+                 ymin = gp_dat$E_Ef - 2 * sqrt(gp_dat$E_Vf), 
+                 ymax = gp_dat$E_Ef + 2 * sqrt(gp_dat$E_Vf) )
+
+  true_means <- sapply(x_grid, f, 0, p)
+  alt_means <- sapply(x_grid, alt$f, 0, alt$p)
+  est_means <- sapply(x_grid, est$f, 0, est$p)
+  par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)
+
+models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
+                     MLE=est_means, Ricker.MLE=alt_means, 
+                     Parametric.Bayes = par_bayes_means)
+
+models <- melt(models, id="x")
+names(models) <- c("x", "method", "value")
+
+plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
+    geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
+    geom_point(data=obs, aes(x,y), alpha=0.8) + 
+    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
+    scale_colour_manual(values=cbPalette) 
+print(plot_gp)
+
+```

### 2013-05-01 inst/examples/BUGS/allen.Rmd

@@
 z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
-x_grid <- seq(0, 1.5 * K, length=101)
+x_grid <- seq(0, 1.5 * K, length=50)
 h_grid <- x_grid
@@
 
-Compute the optimal policy under each model using stochastic dynamic programming.
+Compute the optimal policy under each model using stochastic dynamic programming. We begin with the policy based on the GP model,
 
@@
 MaxT = 1000
-matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)
+# uses expected values from GP, instead of integrating over posterior
+# matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)
+
+# Integrate over posteriors 
+matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid)
+
+# Solve the SDP using the GP-derived transition matrix
 opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
+```
 
+Determine the optimal policy based on the true and MLE models
+
+```{r}
 matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
@@
 opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+```
 
+Determine the optimal policy based on parametric Bayesian model
+
+```{r}
 matrices_par_bayes <- parameter_uncertainty_SDP(f, p, x_grid, h_grid, pardist)
 opt_par_bayes <- value_iteration(matrices_par_bayes, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+```
 
+Assemble the data
+
+```{r}
 OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker.MLE = opt_alt$D, Parametric.Bayes = opt_par_bayes$D)

### 2013-05-01 inst/examples/gaussian-process-basics.Rmd

@@
 ```
-
-```{r}
-prior_curves <- ddply(df, "variable", function(dd){
-    grid <- seq(-4, 4, length = 100)
-    data.frame(value = grid, density = prior_fns[[dd$variable[1]]](grid))
-})
-
-ggplot(prior_curves, aes(value, density)) + 
-  geom_line(data=prior_curves, col="red") + 
-  facet_wrap(~ variable, ncol=1, scale="free_y")
-```

### 2013-05-01 inst/examples/BUGS/parametric-vs-nonparametric.Rmd
run means example as separate script

--- 2013-05-01 inst/examples/BUGS/allen.Rmd
@@
 # uses expected values from GP, instead of integrating over posterior
-# matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)
+matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)
 
 # Integrate over posteriors 
-matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid)
+#matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid)
 

### 2013-05-02 inst/examples/BUGS/parametric-vs-nonparametric.Rmd
varying parameters

@@
 ```{r sdp-pars, dependson="stateeq"}
-sigma_g <- 0.05
+sigma_g <- 0.1
 sigma_m <- 0.0

### 2013-05-02 inst/examples/BUGS/allen.Rmd
replicate run

@@
 gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
-gp_dat <- gp_predict(gp, x_grid)
-
+gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)
 ```
@@
 
+## Parametric Bayes based on the structurally wrong model
+
+We initiate the MCMC chain (`init_p`) using the true values of the parameters `p` from the simulation.
+While impossible in real data, this gives the parametric Bayesian approach the best chance at succeeding.
+`y` is the timeseries (recall `obs` has the $x_t$, $x_{t+1}$ pairs)
+
+```{r}
+# a bit unfair to start with the correct values, but anyhow...
+init_p = p
+names(init_p) = c("r0", "K")
+y <- obs$x[-1] 
+N=length(y);
+```
+
+We'll be using the JAGS Gibbs sampler, a recent open source BUGS implementation with an R interface that works on most platforms.  We initialize the usual MCMC parameters; see `?jags` for details.
+
+```{r}
+jags.data <- list("N","y")
+n.chains = 1
+n.iter = 40000
+n.burnin = floor(10000)
+n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
+```
+
+The actual model is defined in a `model.file` that contains an R function that is automatically translated into BUGS code by *R2WinBUGS*.
+The file defines the priors and the model, as seen when read in here
+
+```{r}
+cat(readLines(con="ricker-UPrior.txt"), sep="\n")
+```
+
+We define which parameters to keep track of, and set the initial values of parameters in the transformed space used by the MCMC.
+We use logarithms to maintain strictly positive values of parameters where appropriate.
+Because our priors on the noise parameters are inverse gamma distributed.
+
+```{r}
+# Uniform priors on standard deviation terms
+jags.params=c("K","logr0", "stdQ", "stdR")
+jags.inits <- function(){
+  list("K"=init_p["K"],"logr0"=log(init_p["r0"]), "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+}
+
+set.seed(12345)
+
+time_jags <- system.time(       
+  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
+                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="ricker-UPrior.txt")
+)         
+time_jags <- unname(time_jags["elapsed"]);
+```
+
+#### Convergence diagnostics for parametric bayes
+
+```{r}
+jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
+par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
+
+# Traces
+ggplot(par_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
+
+## priors (untransformed variables)
+K_prior <- function(x) dunif(x, 0.01, 40)
+logr_prior <- function(x) dunif(x, -6, 6)
+stdQ_prior <- function(x) dunif(x, 0.001, 100)
+stdR_prior <- function(x) dunif(x, 0.001, 100)
+
+par_priors <- list(K = K_prior, deviance = function(x) 0 * x, logr0 = logr_prior, stdQ = stdQ_prior, stdR = stdR_prior)
+
+par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
+    grid <- seq(min(dd$value), max(dd$value), length = 100) 
+    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
+})
+
+# posterior distributions
+ggplot(par_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
+  facet_wrap(~ variable, scale="free", ncol=2)
+
+```
+
+```{r}
+# um, cleaner if we were just be using the long form, par_posterior
+mcmc <- as.mcmc(jagsfit)
+mcmcall <- mcmc[,-2]
+who <- colnames(mcmcall)
+who 
+mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4])
+colnames(mcmcall) <- who
+```
+
+```{r}
+ricker_pardist <- mcmcall
+ricker_pardist[,2] = exp(ricker_pardist[,2]) # transform model parameters back first
+
+bayes_coef <- apply(ricker_pardist,2,mean)
+ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
+ricker_bayes_pars
+```
+
 ### Phase-space diagram of the expected dynamics
@@
   true_means <- sapply(x_grid, f, 0, p)
-  alt_means <- sapply(x_grid, alt$f, 0, alt$p)
+  alt_means <- sapply(x_grid, alt$f, 0, ricker_bayes_pars[c(1,2)])
   est_means <- sapply(x_grid, est$f, 0, est$p)
@@
 models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
-                     MLE=est_means, Ricker.MLE=alt_means, 
+                     MLE=est_means, Ricker=alt_means, 
                      Parametric.Bayes = par_bayes_means)
@@
 # uses expected values from GP, instead of integrating over posterior
-# matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)
+#matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)
 
@@
 
-matrices_alt <- f_transition_matrix(alt$f, alt$p, x_grid, h_grid, alt$sigma_g)
-opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
 ```
@@
 
+Bayesian Ricker
+
+```{r}
+matrices_alt <- parameter_uncertainty_SDP(alt$f, ricker_bayes_pars[1,2], x_grid, h_grid, ricker_pardist)
+opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+```
+
 Assemble the data
@@
 ```{r}
-OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker.MLE = opt_alt$D, Parametric.Bayes = opt_par_bayes$D)
+OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_alt$D, Parametric.Bayes = opt_par_bayes$D)
 colorkey=cbPalette

### 2013-05-02 inst/examples/BUGS/allen.Rmd
replicate run

@@
 x0 <- Xo # simulation under policy starts from
-Tobs <- 35
+Tobs <- 30
 ```
@@
   x <- numeric(Tobs)
-  x[1] <- 6
+  x[1] <- 5.5
   nz <- 1

### 2013-05-03 inst/doc/nonparametric-bayes.Rmd
draft for ERE

@@
-% Non-parametric approaches to optimal policy are more robust
-
-```{r libraries, include=FALSE, cache=FALSE}
-library(pdgControl)
-library(nonparametricbayes)
-library(reshape2)
-library(ggplot2)
-library(data.table)
-library(tgp)
-library(MCMCpack)
-library(plyr)
-library(knitcitations)
-```
-
-```{r plotting-options, echo=FALSE, include=FALSE, cache=TRUE}
-## Dargh these are ignored if not placed here!?
-options(xtable.type = 'latex', table.placement="H", xtable.print.comment=FALSE)
-opts_chunk$set(cache.path = 'cache-pdf/', cache=TRUE)
-opts_chunk$set(dev = 'Cairo_pdf', fig.width=5.5, fig.height=4)
-
-# This stuff should be handled by knit script(?) (load ggplot2 first though)!
-theme_set(theme_bw(base_size=10))
-theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
-             plot.background = element_rect(fill = "transparent",colour = NA))
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-```
-
-```{r stateeq}
-f <- RickerAllee
-p <- c(2, 10, 5) 
-K <- 10
-allee <- 5
-```
-
-```{r sdp-pars, dependson="stateeq"}
-sigma_g <- 0.05
-sigma_m <- 0.0
-z_g <- function() rlnorm(1, 0, sigma_g)
-z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
-x_grid <- seq(0, 1.5 * K, length=101)
-h_grid <- x_grid
-profit <- function(x,h) pmin(x, h)
-delta <- 0.01
-OptTime <- 20  # stationarity with unstable models is tricky thing
-reward <- 0
-xT <- 0
-seed_i <- 1
-Xo <- K # observations start from
-x0 <- Xo # simulation under policy starts from
-Tobs <- 35
-```
-
-```{r obs, dependson="sdp-pars"}
-obs <- sim_obs(Xo, z_g, f, p, Tobs=Tobs, nz=1, 
-                harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
-```
-
-```{r mle, dependson="obs"}
-alt <- par_est(obs,  init = c(r = p[1], 
-                              K = mean(obs$x[obs$x>0]), 
-                              s = sigma_g))
-est <- par_est_allee(obs, f, p,  
-                     init = c(r = p[1] + 1, 
-                              K = p[2] + 2, 
-                              C = p[3] + 2, 
-                              s = sigma_g))
-```
-
-```{r gp-priors}
-#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
-s2.p <- c(5,5)  
-tau2.p <- c(5,1)
-d.p = c(10, 1/0.1, 10, 1/0.1)
-nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
-s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
-tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
-d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
-nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
-beta0_prior <- function(x, tau) dnorm(x, 0, tau)
-beta = c(0)
-priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
-```
-
-```{r gp, dependson=c("obs", "gp-priors")}
-  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
-          m0r1=FALSE, corr="exp", trace=TRUE, 
-          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
-          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
-```
-
-```{r opt, dependson=c("gp", "mle")}
-  OPT <- optimal_policy(gp, f, est$f, alt$f,
-                        p, est$p, alt$p,
-                        x_grid, h_grid, sigma_g, 
-                        sigma_g, sigma_g, # est$sigma_g, alt$sigma_g, but those ests are poor
-                        delta, xT, profit, reward, OptTime)
-```
-
-```{r sim, dependson="opt"}
-dt <- simulate_opt(OPT, f, p, x_grid, h_grid, x0, z_g, profit)
-```
+% Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach
 
@@
 =======================================================================
+
+Decision-theoretic methods often rely on simple parametric models of
+ecological dynamics to compare the value of a potential sequence of
+actions. Unfortunately, such simple models rarely capture the complexity
+or uncertainty found in most real ecosystems. Non-parametric Bayesian
+methods offer a promising statistical approach for predictive modeling
+of ecological dynamics in regions of state space where the data is
+adequate, while at the same time offering more flexible patterns with
+greater uncertainty outside the observed data. This contrasts from simple
+parametric models which provide relatively constant level of uncertainty
+in regions with and without adequate data. The consequence of such
+misplaced confidence outside the data can lead to highly undesirable
+results that may be avoided with the more flexible non-parametric
+Bayesian approach.
 
@@
 involved, in practice real mechanisms are often unknown and the true
-dynamics too complex to be captured by simple models [@refs]. While
+dynamics too complex to be captured by simple models. While
 simple mechanistic models can nevertheless provide important insights
@@
 
-We argue that while simple mechanistically motivated models
+Though simple mechanistically motivated models
 may be best for the latter case [@Cuddington2013; @Geritz2012], such models 
-can be not only inaccurate but misleading in quantitative decision making.
-Nonparametric models can more realistically represent uncertainties outside 
+can be not only inaccurate but misleading in quantitative decision making. This 
+arises for two reasons:
+
+1. We do not know what the correct parameter values are for the models.
+1. We do not know what the correct models are for ecological systems.
+
+The complexity of ecological interactions and a lack of data contribute
+greatly to both of the problems. This concern is particularly acute
+in case of ecological tipping points [@Barnosky2012; @Scheffer2001],
+which arise from feedbacks common in ecological systems and can lead to a
+sudden catastrophic transition to an undesirable state.  In general we do
+not know where such tipping points are unless we have already observed
+the transition, in which case it is frequently too late to respond.
+Though there may exist early warning signals for certain kinds of these
+transitions that are driven by slow changes [@Scheffer2009], we do not
+know when, where, or how to apply them to the decision making context
+more generally [@Boettiger2013].
+
+In addition to facing the potentially grave consequences of such transitions,
+if they do exist in a system of interest, this generally means that we lack
+data in the region near and beyond a possible tipping point.  As a consequence,
+our models are least accurate just where we need them to be most accurate.  
+Unfortunately, parametric modeling approaches are not well-suited to this 
+scenario, in which good data is avialble over only part of the relevant 
+state-space.  Necessary assumptions about model structure in any parametric
+approach can lead such models to perform very poorly, even when accounting
+for parameter uncertainty.
+
+Non-parametric models can more realistically represent uncertainties outside 
 of observed data while also better capturing the dynamics in the region
@@
 problem of optimal harvest of a marine fishery to illustrate how the 
-nonparametric approach of Gaussian processes can be applied.
+nonparametric approach of Gaussian processes can be applied.  We will 
+compare Bayesian implementations of both nonparametric and parametric 
+models, which best allow us to capture the uncertainty of model estimation
+in either case and permits a more natural comparison between approaches.
 
-We imagine the manager seeks to maximize some known socio-economic value function, 
-given their knowledge of the system. 
-In particular, we will focus on the case in which the manager faces a series of decisions,
-such as setting a harvest quota for a marine fishery each year based on observations
-of the stock size the year before.
+### Terminology: "Non-parametric"
+
+The term non-parametric to describe modeling approaches such as Gaussian
+processes is a common source of confusion, notwithstanding the fact that
+the model is still specified by parameters.  Some literature has introduced
+the term "semi-parametric" in tacit acknowledgment of this, which no doubt
+only contributes to the confusion.  The problem is further exacerbated by
+the several meanings assigned to the term in statistics: (a) that the method 
+does not assume some particular probability distribution (e.g. a non-parametric bootstrap)
+or (b) that the method does not assume a fixed structure to the model.  Our 
+use is that of (b), in contrast to the classical approaches that do.  The 
+difference is clearest by way of example.
 
@@
 The ecological dynamics of most systems of management interest are
-typically both complex and unknown [@refs].  Despite this, quantitative
+typically both complex and unknown.  Despite this, quantitative
 policy optimization in these systems has almost exclusively been based
@@
 
-An active literature and growing computational power over the past several decades have
-only marginally improved this situation.  Parametric or structural uncertainty
-can be introduced only by increasing the state space to reflect a distribution
-of possible parameters [@refs] or the degree of belief in each of a set of possible models [@refs]
-Frequently, the space of possible actions must then be reduced or the algorithms adjusted
-by approximations to keep computations feasible.
+<!-- cut?? -->
 
-<span style="color:#007FFF;"> ***This part is overstated and unnecessary***
-Though machine learning approaches have begun to appear in the
-ecological and conservation literature (e.g. in species distribution models),
-including the Gaussian process based approach used here [@Munch2005],
-overall machine learning remains an unfamiliar and somewhat untrusted approach for many ecologists.
-One potential barrier to their adoption is the absence of a framework
-for applying machine learning approaches to resource management problems. </span>
+An active literature and growing computational power over the past several
+decades have only marginally improved this situation.  Parametric or
+structural uncertainty can be introduced only by increasing the state
+space to reflect a distribution of possible parameters or
+the degree of belief in each of a set of possible models Frequently,
+the space of possible actions must then be reduced or the algorithms
+adjusted by approximations to keep computations feasible.
 
-<span style="color:#007FFF;">***transition sentences!?***</span>
+<!-- cut?? -->
 
@@
 programming algorithm can alternatively be driven by the predictions from
-a Gaussian process -- a machine learning approximation to the state dynamics.
-
-<span style="color:#007FFF;">***Okay, perhaps that's novel, but it's pretty trivial.  Isn't it
-obvious to everyone that it's trivial?***</span>
-
-<span style="color:#007FFF;">***paraphrase advantages of machine learning***</span>
-
-Management goals / decision-theoretic approaches need accurate prediction
-over relevant (short?) timescales more than reasonable long-term
-mechanisms.  Machine-learning approaches may offer the benefit of the
-hierarchical Bayesian approach without the practical and computational
-limitations of their parametric kin.  Non-parametric models are flexible
-enough to take maximum advantage of the data available, while being
-appropriately ambiguous about the dynamics of a system in regions of
-parameter space that have been poorly or never sampled.
+a Gaussian process.
 
@@
 =====================================================================
+
+## Underlying Model
+
+Concerns over the potential for tipping points in ecological dynamics
+[@Scheffer2001] highlight the dangers of uncertainty in ecological
+management and pose a substantial challenge to existing decision-theoretic
+approaches [@Brozovic2011].  Because intervention is often too late 
+after a tipping point has been crossed (but see @Hughes2013), management
+is most often concerned with avoiding potentially catastrophic tipping
+points before any data is available at or following a transition that
+would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].
+
+To illustrate the value of the non-parametric Bayesian approach to management,
+we focus on example of a system containing such a tipping point whose dynamics
+can still be described by a simple, one-dimensional parametric model.  
+We will focus on a simple parametric model for a single species [derived
+from fist principles by @Allen2005a] as our underlying "reality".
+
+\begin{align}
+X_{t+1} &= Z_t f(S_t) \\
+S_t &= X_t - h_t \\
+f(S_t) &= e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
+\end{align}
+
+Where $Z_t$ is multiplicative noise function with mean 1, representing
+stochastic growth. We will consider log-normal noise with shape parameter
+$\sigma_g$.  We start with an example in which the parameters are $r =2$,
+$K=8$, $C=5$ and $\sigma_g = 0.1$.
+<!-- ` r p[1]`, $K =$ ` r p[2]`, $C =$ ` r p[3]`, and  $\sigma_g =$ ` r sigma_g`. -->
+
+As a low-dimensional system completely described by three parameters, this
+scenario should if anything be favorable to a parametric-based approach.
+This model contains an Allee effect, or tipping point, below which the
+population is not self-sustaining and shrinks to zero [@Courchamp2008].
+
+Both parametric and nonparametric approaches will require training
+data on which to base their model of the process.  We generate the
+training data under the model described in Eq 1 for $T_{\textrm{obs}}=40$ time
+steps, under a known but not necessarily optimal sequence of harvest
+intensities, $h_t$.  For simplicity we imagine a fishery that started
+from zero harvest pressure and has been gradually increasing the harvest.
+
+<!-- things that we could hit upon in discussion:
+Using data simulated from a specified model rather than empirical data
+permits the comparison against the true underlying dynamics, setting 
+a bar for the optimal performance possible.
+
+(Motivation, alternatives, stationarity, examples without a stable node
+(limit-cycle models), examples based on observations near a stable node
+alone, and why that isn't impossible).
+-->
+
+<span style="color:#007FFF;">***Should we include any emprical examples?***</span>
 
@@
 Given parameters for the function $f$ and probability distribution $Z$,
-along with a given economic model determining the  price/profit $\Pi(X_t, h_t)$
-realized in a given year given a choice of harvest $h_t$ and observed stock $X_t$.
-This problem can be solved exactly for discretized values of stock $X$ and policy $h$
-using stochastic dynamic programming (SDP) [@Mangel1985]. Problems of this sort underpin much marine fisheries management today.
+along with a given economic model determining the  price/profit $\Pi(X_t,
+h_t)$ realized in a given year given a choice of harvest $h_t$ and
+observed stock $X_t$.  This problem can be solved exactly for discretized
+values of stock $X$ and policy $h$ using stochastic dynamic programming
+(SDP) [@Mangel1985]. Problems of this sort underpin much marine fisheries
+management today.
 
 A crux of this approach is correctly specifying the functional form of $f$,
-along with its parameters.  The standard approach [@refs] uses one of a 
+along with its parameters.  The standard approach uses one of a 
 handful of common parametric models representing the stock-recruitment
 relationship, usually after estimating the model parameters from any 
-available existing data. Though modifications of this approach mentioned in the 
-introduction can permit additional sources of uncertainty such as measurement 
-error in the stock assessment, implementation errors in the harvest policy, [@Clark1986; @Roughgarden1996; @Sethi2005]
-uncertainty in parameters [@Mangel1985; @Schapaugh2013] or model structure [@Williams2001; @Athanassoglou2012],
-we focus on the simplest example to illustrate the comparison, where only a 
-single form of uncertainty (the growth shocks $Z_t$) is present in the underlying
-model, making additional modes unnecessary. 
-<span style="color:#007FFF;">***Add comment / apology on single species dynamics?*** </span>
+available existing data. Uncertainty in the parameter estimates can 
+be estimated and integrated over to determine the optimal policy under
+under uncertainty [@Mangel1985; @Schapaugh2013]. Uncertainty in the model
+structure itself can only be addressed in this approach by hypothesizing 
+alternative model structures, and then performing some model choice or
+model averaging  [@Williams2001; @Athanassoglou2012]. For simplicity 
+of the comparison we consider only the case of a structurally correct
+and structurally incorrect model, and estimate Bayesian posterior distributions
+of the parameters.  We will also compare to a simpler maximum likelihood
+estimate of the parameters that ignores the issue of parameter uncertainty.
 
-We compare this approach to our alternative that uses a Gaussian Process (GP)
-in place of the a given stock recruitment curve.  We briefly revisit
+We compare these parametric approaches to our alternative that uses a Gaussian Process (GP)
+in place of a given stock recruitment curve.  We briefly revisit
 a description of the Gaussian Process approach that has appeared in 
@@
 to consider a set of possible curves simultaneously.
-<span style="color:#007FFF;"> ***Figure 1 include curves drawn from the posterior density?*** </span>
-
-<span style="color:#007FFF;">***Do we need more specifics on Gaussian process as an approximation
-to parametric models? Discussion of Gaussian process vs other machine
-learning / forecasting approaches that have less clear statistical
-foundations?  If so, does this all belong in the discussion?***</span>
 
@@
 
-## Example in a bistable system
-
-Concerns over the potential for tipping points in ecological dynamics
-[@Scheffer2001] highlight the dangers of uncertainty in ecological
-management and pose a substantial challenge to existing decision-theoretic
-approaches [@Brozovic2011].  Because intervention is often too late 
-after a tipping point has been crossed (but see @Hughes2013), management
-is most often concerned with avoiding potentially catostrophic tipping
-points before any data is available at or following a transition that
-would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].
-
-<!-- Explain the connection to alternative stable states -->
-
-To illustrate the value of the non-parametric Bayesian approach to management,
-we focus on example of a system containing such a tipping point whose dynamics
-can still be described by a simple, one-dimensional parametric model.  
-We will focus on a simple parametric model for a single species [derived
-from fist principles by @Allen2005a] as our underlying "reality".
-
-\begin{align}
-X_{t+1} &= Z_t f(S_t) \\
-S_t &= X_t - h_t \\
-f(S_t) &= e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
-\end{align}
-
-Where $Z_t$ is multiplicative noise function with mean 1, representing
-stochastic growth. We will consider log-normal noise with shape parameter
-$\sigma_g$.  We start with an example in which the parameters are $r =$
-`r p[1]`, $K =$ `r p[2]`, $C =$ `r p[3]`, and  $\sigma_g =$ `r sigma_g`.
-
-As a low-dimensional system completely described by three parameters, this
-scenario should if anything be favorable to a parametric-based approach.
-This model contains an Allee effect, or tipping point, below which the
-population is not self-sustaining and shrinks to zero [@Courchamp2008].
-
-Both parametric and nonparametric approaches will require training
-data on which to base their model of the process.  We generate the
-training data under the model described in Eq 1 for `r Tobs` time
-steps, under a known but not necessarily optimal sequence of harvest
-intensities, $h_t$.  For simplicity we imagine a fishery that started
-from zero harvest pressure and has been gradually increasing the harvest.
-
-<!-- things that we could hit upon in discussion:
-Using data simulated from a specified model rather than empirical data
-permits the comparison against the true underlying dynamics, setting 
-a bar for the optimal performance possible.
-
-(Motivation, alternatives, stationarity, examples without a stable node
-(limit-cycle models), examples based on observations near a stable node
-alone, and why that isn't impossible).
--->
-
-<span style="color:#007FFF;">***Should we include any emprical examples?***</span>
-
 ### Estimating parametric models
 
-We estimate two parametric models from the data using a maximum likelihood
-approach.  The first model is structurally identical to the true model
-(Eq 1), differing only in that it's parameters are estimated from the
-observed data rather than given.  The alternative model is the Ricker
-model, which is structurally similar and commonly used in for such data.
+We estimate posterior distributions for two parametric models: one using the structurally correct
+model as given in Eq (1), which we refer to as the "Parametric Bayes" model,
+and another using the familiar Ricker model, using a Gibbs sampler
+as described (with source code) in the appendix).  In addition we 
+estimate the parameters of the structurally correct model by maximum
+likelihood.
 
-The maximum likelihood estimates (MLE) will assume that the process noise $Z_t$
-is log-normal, as in the simulation itself.  In reality the noise structure might 
-differ from this assumption, making this a best-case scenario for the MLE approach.  
-The GP must also make an assumption about the noise structure -- in this case we 
-assume additive noise, though multiplicative noise would be more accurate.  This
-will illustrate the flexibility of the GP in fitting different noise structures
-even without knowledge of the correct assumptions.
-
-The GP estimate is Bayesian approach which requires specification of
-prior distributions for the hyper-parameters [see Appendix for details,
-or @Rassmussen1996, @Munch2005 for background], as given in Table 1.
-Then given the data, both the GP and the MLE parametric models can
-be estimated as described above, as we implement in R in the code
-provided in supplementary materials.
+The posterior distribution for the hyper-parameters of the Gaussian 
+process model are estimated by Metropolis-Hastings algorithm, again with
+details and code provided in the Appendix.  @Rasmussen2006 provides
+an excellent general introduction to Gaussian Processes and @Munch2005 
+first discusses their application in this context.
 
@@
 
-Figure 1 shows the  inferred Gaussian Process compared to the true and the
-MLE parametric models, predicting the relationship between observed stock
-size (x-axis) to the stock size following recruitment the following year.
-The gray shading shows the uncertainty estimated by the GP.  Unlike the
-MLE estimates, the GP uncertainty changes depending on where the data is
-available, allowing for a much greater level of uncertainty at low stock
-sizes.  While it would be natural (and straight forward) to condition
+![ Mean dynamics predicted by each model.](Figure1.pdf)
+
+Figure 1 shows the mean inferred state space dynamics of each model
+relative to the true model used to generate the data, predicting the
+relationship between observed stock size (x-axis) to the stock size
+after recruitment the following year.  All models except the MLE model
+estimate a distribution around the means shown here, and all models 
+estimate a level of process noise, which is independent of the state
+value (x).  Note that in contrast to the  other models shown, the mean
+Gaussian process corresponds to a distribution of curves - as indicated
+by the gray band - which itself has a mean shown in black.  Note that 
+this mean GP is thus more certain of the dynamics in the region where
+data is available then where it is not.
+
+While it would be natural (and straight forward) to condition
 the GP on passing through the origin (0,0) (see appendix), the estimate
@@
 or weakly harvested system.  No observations occur at the theoretical
-optimum (7.5 units, see Figure 2) or near the tipping point (5 units).
+optimum or near the tipping point.
 
-<!-- Would it be useful to show the uncertainty in the other models? -->
-
-<!-- with (light grey) and without (darker grey) measurement error -->
-
-```{r gp_plot, dependson=c("gp", "mle", "plotting-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
-Graph shows the expected value for the function $f$ under each model.
-Two standard deviations from the estimated Gaussian process covariance (grey shading)  is also shown.
-The training data is also shown as black points.
-"}
-tgp_dat <- 
-    data.frame(  x = gp$XX[[1]], 
-                 y = gp$ZZ.km, 
-                 ymin = gp$ZZ.km - 2 * sqrt(gp$ZZ.ks2), 
-                 ymax = gp$ZZ.km + 2 * sqrt(gp$ZZ.ks2),
-                 ymin2 = gp$ZZ.mean - 2 * sqrt(gp$ZZ.vark), 
-                 ymax2 = gp$ZZ.mean + 2 * sqrt(gp$ZZ.vark))
-  true <- sapply(x_grid, f, 0, p)
-  alt_mean <- sapply(x_grid, alt$f, 0, alt$p)
-  est_mean <- sapply(x_grid, est$f, 0, est$p)
-  models <- data.frame(x=x_grid, GP=tgp_dat$y, 
-                       Parametric=est_mean, 
-                       True=true, 
-                       Structural=alt_mean)
-  models <- melt(models, id="x")
-  names(models) <- c("x", "method", "value")
-
-ggplot(tgp_dat) + 
-#  geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +    # measurement error + process error
-  geom_ribbon(aes(x,y,ymin=ymin2,ymax=ymax2), fill="gray60") + # process error only
-  geom_line(data=models, aes(x, value, col=method), alpha=0.8, lwd=1) + 
-  geom_point(data=obs, aes(x,y)) + 
-  xlab(expression(X[t])) + ylab(expression(X[t+1])) +
-  scale_colour_manual(values=cbPalette) 
-```
+![ Steady-state optimal policies by each model.](Figure2.pdf)
 
@@
 strategy aims to return the population to $S$, resulting in the hockey-stick
-shaped policies shown.  Only the GP policy deviates slightly from this.  The MLE parametric model
-corresponding to parametric uncertainty alone (gold line labeled "Parametric", 
-that is, using the same structural form as the underlying simulation) nevertheless overestimates 
-the carrying capacity of the stock and therefore underestimates the harvest
-rates, always waiting for the stock to rise further than is actually possible.
-The Ricker model (green line, showing "Structural" inaccuracies) does not 
-foresee the tipping point dynamics reflected in the concavity of the other three
-models (Figure 1), and selects a much more aggressive harvesting strategy.
+shaped policies shown.
 
-```{r policies_plot, dependson=c("opt", "plotting-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.  "}
-policies <- 
-  melt(data.frame(stock=x_grid, 
-                  GP = x_grid[OPT$gp_D], 
-                  Parametric = x_grid[OPT$est_D],
-                  True = x_grid[OPT$true_D],
-                  Structural = x_grid[OPT$alt_D]),
-       id="stock")
-names(policies) <- c("stock", "method", "value")
-
-ggplot(policies, aes(stock, stock - value, color=method)) +
-    geom_line(alpha=0.7, lwd=1) + 
-    xlab("stock size") + ylab("escapement")  +
-    scale_colour_manual(values=cbPalette)
-```
+![ Managing the simulated system over 100 replicate instances under each policy](Figure3.pdf)
 
@@
 simulated fishery under each of the policies estimated is shown in Figure 3.  As expected
-from the policy curves, the structurally correct model ("Parametric" uncertainty, gold) 
-underharvests, leaving the stock to vary around it's unfished optimum.  The GP overharvests
-the population slightly relative to the optimal value.  The structurally
-incorrect Ricker model ("Structural", green line) overharvests the population passed
+from the policy curves, the structurally correct model under-harvests,
+leaving the stock to vary around it's un-fished optimum.  The structurally
+incorrect Ricker model over-harvests the population passed
 the tipping point consistently, resulting in the immediate crash of the stock and 
 thus derives minimal profits.
-
-***(may need to show harvest dynamics since
-that is not clear from the figure! Also isn't general, sometimes does
-optimally, sometimes over-fishes.  Perhaps need to show more examples.)
-May need to show profits too?***
-
-```{r sim_plot, dependson=c("sim", "plotting-options"), fig.cap="Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."}
-ggplot(dt) + 
-    geom_line(aes(time, fishstock, 
-                  group=interaction(reps, method), 
-                  color=method), alpha=.1) +
-    scale_colour_manual(values=c("GP" = cbPalette[1], "Parametric" = cbPalette[2], "True" = cbPalette[3], "Structural" = cbPalette[4]),  
-                        guide = guide_legend(override.aes = list(alpha = 1)))
-```
 
@@
 are common to many conservation decision making problems, which thus merit
-greater use of non-parametric machine learning approaches that can best 
-take advantage of them.
+greater use of non-parametric approaches that can best take advantage of them.
 
@@
 
-<!--
-<span style="color:#007FFF;"> ***wow, run-on tangent here?*** </span>
-Predictive accuracy is not the goal of all modeling, as ecologists
-have been observing for as long as they made models (perhaps none
-more memorably than @Levins1969).  Mechanistic modeling is at its
-most powerful not when it is used to directly forecast future states
-but when it provides an understanding of how to approach a problem.
-SIR-type models from the epidemiological literature are a good example.
-While the simplest SIR models have little predictive power over the
-outbreak intensity or timing at a particular location, they provide a
-powerful understanding of the spread of an infection in terms of a single,
-biologically meaningful parameter: $R_0$, the basic reproductive number.
-From the model, it becomes clear that management need not vaccinate every
-member of the population to stop the spread, but rather it suffices to
-vaccinate a sufficient fraction of the population to reduce $R_0$ below 1.
--->
-
-### 3. Time scale for new data
-
-In the sequential decision making problem we considered, we are
-presented with new data after each action.  The relevant timescale
-for the prediction is thus not the long-term dynamics, which would be
-wildly divergent, but the dynamics over this much shorter interval.
-While ecologists may be hesitant to base continual management on a model
-with obviously inaccurate long-term behavior, engineers tend to consider
-the problem in frequency space and gravitate to the opposite position --
-a good control model should prioritize high-frequency accuracy over low
-frequency accuracy.  The differences in intuition may arise from the
-timescales at which each profession can typically adjust their control
-variables -- much faster for a control system of a chemical plant than
-state policy for a natural resource.  Still, the lesson is clear: when
-facing repeated management decisions over a short timescale, such as
-setting annual harvests of a fishery, it may be more valuable to use a
-machine learning algorithm that makes accurate year-ahead predictions
-that capture some of the high-frequency fluctuations that appear only
-as noise in a mechanistic model of the long-term population dynamics.
-
 ## Future directions
 
-<span style="color:#007FFF;"> Jargony and vague </span>
-While we have highlighted certain generic of this problem that allow the
-nonparametric approach to excel -- short timescales between successive
-observations and actions, the accuracy in the appropriate region of state
-space, the ability to express uncertainty outside the observed data --
-there are equally several aspects for which the nonparametric Bayesian approach is
-nevertheless at a relative disadvantage. In this simulated example, the underlying
+In this simulated example, the underlying
 dynamics are truly governed by a simple parametric model, allowing
@@
 mechanistic models can be an immensely difficult due to the increased
-complexity involved.  Only thanks to tandem advances in increasing
-computational power and hierarchical statistical methodology have we
-been able to tackle such intuitively important complexity (and the
-potentially new available data that accompanies it) such as spatial
-distribution, heterogeneities of space, time, and individuals, to shift
-to ecosystem-based approaches from single-species based approaches.
-Without the need to formulate mechanisms, many modern machine learning
-algorithms can leverage potential information from all available sources
-of data directly.  The algorithms can recognize unanticipated or subtle
-patterns in large data sets that enable more accurate predictions than
-mechanistic models that are formulated at a more macroscopic level.
+complexity involved.
+
+### Online learning
+
+In our treatment here we have ignored the possibility of learning during the 
+management phase, in which the additional observations of the stock size could
+potentially improve parameter estimates.  While we intend to address this 
+possibility in future work in the context of these non-parametric models,
+we have not addressed it here for pedagogical reasons. In the context presented
+here, it is clear that the differences in performance arise from differences
+in the uncertainty inherent in the model formulations, rather than from 
+differing abilities to learn.  Because we consider a threshold system, 
+online learning would not change this generic feature of a lack of data in a
+certain range of the state space which is better captured by the Gaussian process.
 
@@
 
-<span style="color:#007FFF;"> 
-***Possible additional discussion material?***
+Appendix
+========
 
-* Discuss constant escapement in model, in policies.
-* Limitations of this comparison: Are the maximum-likelihood solutions
-a straw man?
-* Discussion of alternative related approaches: POMDP/MOMDP
+The appendices have not yet been assembled.
+Meanwhile, code to repeat the analyses, along with a complete log of all research conducted on this project, can be found at: [https://github.com/cboettig/nonparametric-bayes](https://github.com/cboettig/nonparametric-bayes/)
 
-* Online learning 
-* Multiple step-ahead predictions 
-* Explicitly accomidating additional uncertainties 
-* Improving inference of optimal policy from the GP
-</span>
+## Markov Chain Monte Carlo Analysis
 
---------------------------------------------------------------------
+![Traces of the GP hyper-parameters (after transforming back from log space)](gp_traces_densities1.pdf)
 
-Appendix / Supplementary Materials
-==================================
+![Posterior densities of the GP hyper-parameters (after transforming back from log space). Prior distribution over the interval sampled shown as red curves.](gp_traces_densities2.pdf)
 
-### The Gaussian process
+![Traces of the Ricker parameters](ricker_traces.pdf)
 
-The Gaussian process is defined by a covariance kernel.
-By requiring our kernel to follow a generic functional form, we can compactly describe the Gaussian process using only a handful of parameters (Table 1)
+![Posterior densities of the Ricker parameter estimates.
+Prior distributions over the interval sampled shown as red curves.
+Note the bimodal estimate of $\log(r)$ which leads to the poor fit in Figure 1](ricker_posteriors.pdf)
 
-parameter       interpretation
----------       -------------- 
-$\sigma^2$      The process noise (from the kernel)
-$\tau^2$        Variance around the mean
-$\beta_0$       The mean is given by a linear model of slope $\beta$ 
-$d$             The length-scale of the covariance function
-$n$             The observation error
+![Traces of the structurally accurate Allen model parameters](parametric_bayes_traces.pdf)
 
-: Table of parameters for the Gaussian process
+![Posterior densities for the parameters of the Allen model. Prior distributions over the interval sampled shown as red curves.](parametric_bayes_posteriors.pdf)
 
-Rather than estimate values of these parameters directly, we take a hierarchical approach of placing prior distributions on each.
-Following @Gramarcy2005 we use a Gaussian prior on $\beta_0$, an inverse gamma prior on $\sigma^2$ and $\tau^2$, a gamma prior on the observation noise $n$, and exponential prior on the length-scale $d$.
-
-###  Formulating a dynamic programming solution
-
-The fishery management problem over an infinite time horizon can be stated as:
-
-\begin{align}
-& \max_{ \{h_t\} \geq 0 } \mathbf{E} \lbrace \sum_0^\infty \delta^t \Pi(h_t) \rbrace \\
-& \mathrm{s.t.}  \\
- & X_t = Z_t f\left(S_{t-1}\right) \\
- & S_t = X_t - h_t \\
- & X_t  \geq 0 
-\end{align}
-
-Where $\mathbf{E}$ is the expectation operator, $\delta$ the discount
-rate, $\Pi(h_t)$ the profit expected from a harvest of $h_t$, and other
-terms as in Eq. (1).  For simplicity, we have assumed that profits depend
-only on the chosen harvest; simplifying further we will usually consider
-profits to be proportional to harvest, $\Pi(h_t) = h_t$.
-
-Once the posterior Gaussian process (GP) has been estimated [e.g. see
-@Munch2005], it is necessary to adapt it in place of the parametric
-equation for the stochastic dynamic programming (SDP) solution [see
-@Mangel1988 for a detailed description of parametric SDP methods] to the
-optimal policy. The essense of the idea is straight forward -- we will use
-the estimated GP in place of the parametric growth function to determine
-the stochastic transition matrix on which the SDP calculations are based.
-
-The posterior Gaussian process is completely defined by the expected value
-and covariance matrix at a defined set of training points.  For simplicity
-we will consider a these points to fall on a discrete, uniform grid $x$
-of `r length(x_grid)` points from `r min(x_grid)` to `r max(x_grid)`
-(1.5 times the positive equilibrium $K$).  Again to keep things simple
-we will use this same grid discretization for the parametric approach.
-Other options for choosing the grid points, including collocation methods
-and functional basis expansion (or even using Gaussian processes in place
-of the discrete optimization; an entirely different context in which GP
-can be used in SDP, see [@Deisenroth2009]) could also be considered.
-
-## Pseudocode for the determining the transtion matrix from the GP
-
-```r
-for(h in h_grid)
-  F_h = for(x_j in grid)
-          for(i in 1:N) 
-            dnorm(x_j, mu[i]-h, V[i])
-```
-
-A transition matrix for each of the parametric models $f$ is calculated
-using the log-normal density with mean $f(x_i)$ and log-variance as
-estimated by maximum likelihood.  From the discrete transition matrix we
-may write down the Bellman recursion defining the the stochastic dynamic
-programming iteration:
-
-\begin{equation}
-V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
-\end{equation}
-
-where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$
-is control (harvest level) chosen. Numerically, the maximization is
-accomplished as follows. Consider the set of possible control values to
-be the discrete `r length(h_grid)` values corresponding the the grid of
-stock sizes.  Then for each $h_t$ there is a corresponding transition
-matrix $\mathbf{F}_h$ determined as described above but with mean 
-$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element corresponds
-to the value of having stock $x_i$ at time $t$.  Then let $\Pi_h$ be
-the vector whose $i$th element indicates the profit from harvesting
-at intensity $h_t$ given a population $x_i$ (e.g. $\max(x_i, h_t)$
-since one cannot harvest more fish then the current population size).
-Then the Bellman recursion can be given in matrix form as
-
-$$V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)$$
-
-where the sum is element by element and the expectation is computed by the matrix multiplication $\mathbf{F} V_{t+1}$.
-
-### Pseudocode for the Bellman iteration
-
-```r
- V1 <- sapply(1:length(h_grid), function(h){
-      delta * F[[h]] %*% V +  profit(x_grid, h_grid[h]) 
-    })
-    # find havest, h that gives the maximum value
-    out <- sapply(1:gridsize, function(j){
-      value <- max(V1[j,], na.rm = T) # each col is a diff h, max over these
-      index <- which.max(V1[j,])  # store index so we can recover h's 
-      c(value, index) # returns both profit value & index of optimal h.  
-    })
-    # Sets V[t+1] = max_h V[t] at each possible state value, x
-    V <- out[1,]                        # The new value-to-go
-    D[,OptTime-time+1] <- out[2,]       # The index positions
-```
-
-_Currently this shows the literal R code, should be adapted_
-
-MCMC posterior distributions and convergence analysis
-----------------------------------------------------------------------------
-
-***Can we get this figure into the main text with appropriate discussion without getting too technical?***
-
-```{r posteriors, dependson=c("gp", "plotting-options"), fig.cap = "Histogram of posterior distributions for the estimated Gaussian Process shown in Figure 1.  Prior distributions overlaid."}
-hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
-posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")
-  prior_curves <- ddply(posteriors, "variable", function(dd){
-  grid <- seq(min(dd$value), max(dd$value), length = 100)
-  data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
-})
-
-ggplot(posteriors) + 
-    geom_histogram(aes(x=value, y=..density..), alpha=0.7) +
-    geom_line(data=prior_curves, aes(x=value, y=density), col="red") +
-    facet_wrap(~ variable, scale="free")
-```
-
- @Gramacy2005
-
-## Tables of nuisance parameters, sensitivity analysis
-
-### List of hyper-parameters, prior distributions and their parameters
-
-## Reproducible code, "Research Compendium"
+References
+==========

### 2013-05-03 inst/doc/nonparametric-bayes.tex
draft for ERE

--- 2013-04-08 inst/doc/nonparametric-bayes.tex
@@
 \newcommand{\NormalTok}[1]{{#1}}
-\usepackage{longtable}
 \usepackage{graphicx}
@@
             pdfauthor={},
-            pdftitle={Non-parametric approaches to optimal policy are more robust},
+            pdftitle={Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach},
             colorlinks=true,
@@
 \begin{frontmatter}
-  \title{Non-parametric approaches to optimal policy are more robust}
+  \title{Avoiding tipping points in the management of ecological systems: a
+         non-parametric Bayesian approach}
   \author[cstar]{Carl Boettiger\corref{cor1}}
@@
 
+Decision-theoretic methods often rely on simple parametric models of
+ecological dynamics to compare the value of a potential sequence of
+actions. Unfortunately, such simple models rarely capture the complexity
+or uncertainty found in most real ecosystems. Non-parametric Bayesian
+methods offer a promising statistical approach for predictive modeling
+of ecological dynamics in regions of state space where the data is
+adequate, while at the same time offering more flexible patterns with
+greater uncertainty outside the observed data. This contrasts from
+simple parametric models which provide relatively constant level of
+uncertainty in regions with and without adequate data. The consequence
+of such misplaced confidence outside the data can lead to highly
+undesirable results that may be avoided with the more flexible
+non-parametric Bayesian approach.
+
 \section{Introduction}
@@
 interactions involved, in practice real mechanisms are often unknown and
-the true dynamics too complex to be captured by simple models
-{[}@refs{]}. While simple mechanistic models can nevertheless provide
-important insights into possible dynamics -- for instance, demonstrating
-that a vaccine does not have to be 100\% effective to eliminate the
-transmission of a virus -- such approaches are not well-suited for use
-in forecasting outcomes of potential management options. Non-parametric
-approaches offer a more flexible alternative that can both more
-accurately reflect the data available while also representing greater
-uncertainty in areas (of state-space) where data is lacking. Ecological
-research and management strategy should pay closer attention to the
-opportunities and challenges nonparametric modeling can offer.
+the true dynamics too complex to be captured by simple models. While
+simple mechanistic models can nevertheless provide important insights
+into possible dynamics -- for instance, demonstrating that a vaccine
+does not have to be 100\% effective to eliminate the transmission of a
+virus -- such approaches are not well-suited for use in forecasting
+outcomes of potential management options. Non-parametric approaches
+offer a more flexible alternative that can both more accurately reflect
+the data available while also representing greater uncertainty in areas
+(of state-space) where data is lacking. Ecological research and
+management strategy should pay closer attention to the opportunities and
+challenges nonparametric modeling can offer.
 
@@
 
-We argue that while simple mechanistically motivated models may be best
-for the latter case (Cuddington, Fortin, and Gerber 2013; Geritz and
-Kisdi 2012), such models can be not only inaccurate but misleading in
-quantitative decision making. Nonparametric models can more
-realistically represent uncertainties outside of observed data while
-also better capturing the dynamics in the region observed. Advances in
-the theory and computational implementations of nonparametric methods
-make them ripe for such applications. We use the classic problem of
-optimal harvest of a marine fishery to illustrate how the nonparametric
-approach of Gaussian processes can be applied.
+Though simple mechanistically motivated models may be best for the
+latter case (Cuddington, Fortin, and Gerber 2013; Geritz and Kisdi
+2012), such models can be not only inaccurate but misleading in
+quantitative decision making. This arises for two reasons:
 
-We imagine the manager seeks to maximize some known socio-economic value
-function, given their knowledge of the system. In particular, we will
-focus on the case in which the manager faces a series of decisions, such
-as setting a harvest quota for a marine fishery each year based on
-observations of the stock size the year before.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  We do not know what the correct parameter values are for the models.
+\item
+  We do not know what the correct models are for ecological systems.
+\end{enumerate}
+
+The complexity of ecological interactions and a lack of data contribute
+greatly to both of the problems. This concern is particularly acute in
+case of ecological tipping points (Barnosky et al. 2012; Scheffer et al.
+2001), which arise from feedbacks common in ecological systems and can
+lead to a sudden catastrophic transition to an undesirable state. In
+general we do not know where such tipping points are unless we have
+already observed the transition, in which case it is frequently too late
+to respond. Though there may exist early warning signals for certain
+kinds of these transitions that are driven by slow changes (Scheffer et
+al. 2009), we do not know when, where, or how to apply them to the
+decision making context more generally (Boettiger and Hastings 2013).
+
+In addition to facing the potentially grave consequences of such
+transitions, if they do exist in a system of interest, this generally
+means that we lack data in the region near and beyond a possible tipping
+point. As a consequence, our models are least accurate just where we
+need them to be most accurate.\\Unfortunately, parametric modeling
+approaches are not well-suited to this scenario, in which good data is
+avialble over only part of the relevant state-space. Necessary
+assumptions about model structure in any parametric approach can lead
+such models to perform very poorly, even when accounting for parameter
+uncertainty.
+
+Non-parametric models can more realistically represent uncertainties
+outside of observed data while also better capturing the dynamics in the
+region observed. Advances in the theory and computational
+implementations of nonparametric methods make them ripe for such
+applications. We use the classic problem of optimal harvest of a marine
+fishery to illustrate how the nonparametric approach of Gaussian
+processes can be applied. We will compare Bayesian implementations of
+both nonparametric and parametric models, which best allow us to capture
+the uncertainty of model estimation in either case and permits a more
+natural comparison between approaches.
+
+\subsubsection{Terminology: ``Non-parametric''}
+
+The term non-parametric to describe modeling approaches such as Gaussian
+processes is a common source of confusion, notwithstanding the fact that
+the model is still specified by parameters. Some literature has
+introduced the term ``semi-parametric'' in tacit acknowledgment of this,
+which no doubt only contributes to the confusion. The problem is further
+exacerbated by the several meanings assigned to the term in statistics:
+(a) that the method does not assume some particular probability
+distribution (e.g.~a non-parametric bootstrap) or (b) that the method
+does not assume a fixed structure to the model. Our use is that of (b),
+in contrast to the classical approaches that do. The difference is
+clearest by way of example.
 
@@
 The ecological dynamics of most systems of management interest are
-typically both complex and unknown {[}@refs{]}. Despite this,
-quantitative policy optimization in these systems has almost exclusively
-been based on simple, mechanistically motivated parametric models. There
-are several reasons for this. First, limited available data almost
-always precludes any accurate estimate of more complicated models.
-Second, the computational demands of determining an optimal strategy in
-a sequential decision-theory problem suffer from the ``curse of
-dimensionality'' that makes it difficult to consider more complicated
-models, states, or actions. Unfortunately, these challenges also make it
-particularly difficult to reflect the true uncertainty either in the
-parameters of the given models (parametric uncertainty) or in the
-structure of the models themselves (structural uncertainty). Further
-unknowns such as measurement uncertainty, parameter uncertainty,
-unobserved states, knowledge of boundary conditions, etc. further
-compound the issue. Though a hierarchical Bayesian approach provides a
-natural way to address these from a statistical standpoint, formulating
-reasonable parametric descriptions of each form of uncertainty is a
-challenging task in itself, let alone the computational difficulties of
-solving such a system. Cressie et al. (2009) provides a good account of
-the successes and challenges of the approach. Applying these approaches
-in the management context of sequential decision making, in which
-forecasts must be obtained over a range of possible actions and updated
-regularly as new information arrives makes such an approach less
-feasible still.
+typically both complex and unknown. Despite this, quantitative policy
+optimization in these systems has almost exclusively been based on
+simple, mechanistically motivated parametric models. There are several
+reasons for this. First, limited available data almost always precludes
+any accurate estimate of more complicated models. Second, the
+computational demands of determining an optimal strategy in a sequential
+decision-theory problem suffer from the ``curse of dimensionality'' that
+makes it difficult to consider more complicated models, states, or
+actions. Unfortunately, these challenges also make it particularly
+difficult to reflect the true uncertainty either in the parameters of
+the given models (parametric uncertainty) or in the structure of the
+models themselves (structural uncertainty). Further unknowns such as
+measurement uncertainty, parameter uncertainty, unobserved states,
+knowledge of boundary conditions, etc. further compound the issue.
+Though a hierarchical Bayesian approach provides a natural way to
+address these from a statistical standpoint, formulating reasonable
+parametric descriptions of each form of uncertainty is a challenging
+task in itself, let alone the computational difficulties of solving such
+a system. Cressie et al. (2009) provides a good account of the successes
+and challenges of the approach. Applying these approaches in the
+management context of sequential decision making, in which forecasts
+must be obtained over a range of possible actions and updated regularly
+as new information arrives makes such an approach less feasible still.
 
@@
 or structural uncertainty can be introduced only by increasing the state
-space to reflect a distribution of possible parameters {[}@refs{]} or
-the degree of belief in each of a set of possible models {[}@refs{]}
-Frequently, the space of possible actions must then be reduced or the
-algorithms adjusted by approximations to keep computations feasible.
-
- \textbf{\emph{This part is overstated and unnecessary}} Though machine
-learning approaches have begun to appear in the ecological and
-conservation literature (e.g.~in species distribution models), including
-the Gaussian process based approach used here (Munch et al. 2005),
-overall machine learning remains an unfamiliar and somewhat untrusted
-approach for many ecologists. One potential barrier to their adoption is
-the absence of a framework for applying machine learning approaches to
-resource management problems.
-
-\textbf{\emph{transition sentences!?}}
+space to reflect a distribution of possible parameters or the degree of
+belief in each of a set of possible models Frequently, the space of
+possible actions must then be reduced or the algorithms adjusted by
+approximations to keep computations feasible.
 
@@
 stochastic dynamic programming algorithm can alternatively be driven by
-the predictions from a Gaussian process -- a machine learning
-approximation to the state dynamics.
-
-\textbf{\emph{Okay, perhaps that's novel, but it's pretty trivial. Isn't
-it obvious to everyone that it's trivial?}}
-
-\textbf{\emph{paraphrase advantages of machine learning}}
-
-Management goals / decision-theoretic approaches need accurate
-prediction over relevant (short?) timescales more than reasonable
-long-term mechanisms. Machine-learning approaches may offer the benefit
-of the hierarchical Bayesian approach without the practical and
-computational limitations of their parametric kin. Non-parametric models
-are flexible enough to take maximum advantage of the data available,
-while being appropriately ambiguous about the dynamics of a system in
-regions of parameter space that have been poorly or never sampled.
+the predictions from a Gaussian process.
 
 \section{Approach and Methods}
+
+\subsection{Underlying Model}
+
+Concerns over the potential for tipping points in ecological dynamics
+(Scheffer et al. 2001) highlight the dangers of uncertainty in
+ecological management and pose a substantial challenge to existing
+decision-theoretic approaches (Brozović and Schlenker 2011). Because
+intervention is often too late after a tipping point has been crossed
+(but see Hughes et al. (2013)), management is most often concerned with
+avoiding potentially catastrophic tipping points before any data is
+available at or following a transition that would more clearly reveal
+these regime shift dynamics (e.g. Bestelmeyer et al. 2012).
+
+To illustrate the value of the non-parametric Bayesian approach to
+management, we focus on example of a system containing such a tipping
+point whose dynamics can still be described by a simple, one-dimensional
+parametric model.\\We will focus on a simple parametric model for a
+single species (derived from fist principles by Allen et al. 2005) as
+our underlying ``reality''.
+
+\begin{align}
+X_{t+1} &= Z_t f(S_t) \\
+S_t &= X_t - h_t \\
+f(S_t) &= e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
+\end{align}
+
+Where $Z_t$ is multiplicative noise function with mean 1, representing
+stochastic growth. We will consider log-normal noise with shape
+parameter $\sigma_g$. We start with an example in which the parameters
+are $r =2$, $K=8$, $C=5$ and $\sigma_g = 0.1$.
+
+As a low-dimensional system completely described by three parameters,
+this scenario should if anything be favorable to a parametric-based
+approach. This model contains an Allee effect, or tipping point, below
+which the population is not self-sustaining and shrinks to zero
+(Courchamp, Berec, and Gascoigne 2008).
+
+Both parametric and nonparametric approaches will require training data
+on which to base their model of the process. We generate the training
+data under the model described in Eq 1 for $T_{\textrm{obs}}=40$ time
+steps, under a known but not necessarily optimal sequence of harvest
+intensities, $h_t$. For simplicity we imagine a fishery that started
+from zero harvest pressure and has been gradually increasing the
+harvest.
+
+\textbf{\emph{Should we include any emprical examples?}}
 
@@
 discretized values of stock $X$ and policy $h$ using stochastic dynamic
-programming (SDP) (Mangel and Clark 1985). Problems of this sort
-underpin much marine fisheries management today.
+programming (SDP) (Mangel 1985). Problems of this sort underpin much
+marine fisheries management today.
 
 A crux of this approach is correctly specifying the functional form of
-$f$, along with its parameters. The standard approach {[}@refs{]} uses
-one of a handful of common parametric models representing the
-stock-recruitment relationship, usually after estimating the model
-parameters from any available existing data. Though modifications of
-this approach mentioned in the introduction can permit additional
-sources of uncertainty such as measurement error in the stock
-assessment, implementation errors in the harvest policy, (Clark and
-Kirkwood 1986; Roughgarden and Smith 1996; Sethi et al. 2005)
-uncertainty in parameters (Mangel and Clark 1985; Schapaugh and Tyre
-2013) or model structure (B. K. Williams 2001; Athanassoglou and
-Xepapadeas 2012), we focus on the simplest example to illustrate the
-comparison, where only a single form of uncertainty (the growth shocks
-$Z_t$) is present in the underlying model, making additional modes
-unnecessary. \textbf{\emph{Add comment / apology on single species
-dynamics?}}
+$f$, along with its parameters. The standard approach uses one of a
+handful of common parametric models representing the stock-recruitment
+relationship, usually after estimating the model parameters from any
+available existing data. Uncertainty in the parameter estimates can be
+estimated and integrated over to determine the optimal policy under
+under uncertainty (Mangel 1985; Schapaugh and Tyre 2013). Uncertainty in
+the model structure itself can only be addressed in this approach by
+hypothesizing alternative model structures, and then performing some
+model choice or model averaging (B. K. Williams 2001; Athanassoglou and
+Xepapadeas 2012). For simplicity of the comparison we consider only the
+case of a structurally correct and structurally incorrect model, and
+estimate Bayesian posterior distributions of the parameters. We will
+also compare to a simpler maximum likelihood estimate of the parameters
+that ignores the issue of parameter uncertainty.
 
-We compare this approach to our alternative that uses a Gaussian Process
-(GP) in place of the a given stock recruitment curve. We briefly revisit
-a description of the Gaussian Process approach that has appeared in
-recent literature before describing how we take the important step of
-adapting the GP to the decision theory problem that can still be solved
-using SDP.
+We compare these parametric approaches to our alternative that uses a
+Gaussian Process (GP) in place of a given stock recruitment curve. We
+briefly revisit a description of the Gaussian Process approach that has
+appeared in recent literature before describing how we take the
+important step of adapting the GP to the decision theory problem that
+can still be solved using SDP.
 
@@
 et al. (2005). An accessible and thorough introduction to the
-formulation and use of GPs can be found in C. E. Rasmussen and Williams
+formulation and use of GPs can be found in Rasmussen and Williams
 (2006).
@@
 range of possible (parametric) models that might describe the data. The
-GP allows us to consider a set of possible curves simultaneously.\\
-\textbf{\emph{Figure 1 include curves drawn from the posterior
-density?}}
-
-\textbf{\emph{Do we need more specifics on Gaussian process as an
-approximation to parametric models? Discussion of Gaussian process vs
-other machine learning / forecasting approaches that have less clear
-statistical foundations? If so, does this all belong in the
-discussion?}}
+GP allows us to consider a set of possible curves simultaneously.
 
@@
 \NormalTok{for(h in h_grid)}
-  \NormalTok{F_h =}\StringTok{ }\NormalTok{for(x_j in grid)}
+  \NormalTok{F_h = for(x_j in grid)}
           \NormalTok{for(i in }\DecValTok{1}\NormalTok{:N) }
@@
 \begin{Highlighting}[]
- \NormalTok{V1 <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
-      \NormalTok{delta *}\StringTok{ }\NormalTok{F[[h]] %*%}\StringTok{ }\NormalTok{V +}\StringTok{  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
+ \NormalTok{V1 <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
+      \NormalTok{delta * F[[h]] %*% V +  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
     \NormalTok{\})}
     \CommentTok{# find havest, h that gives the maximum value}
-    \NormalTok{out <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
-      \NormalTok{value <-}\StringTok{ }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
-      \NormalTok{index <-}\StringTok{ }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
+    \NormalTok{out <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
+      \NormalTok{value <- }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
+      \NormalTok{index <- }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
       \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
@@
     \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
-    \NormalTok{V <-}\StringTok{ }\NormalTok{out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
-    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <-}\StringTok{ }\NormalTok{out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
+    \NormalTok{V <- out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
+    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <- out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
 \end{Highlighting}
@@
 
-\subsection{Example in a bistable system}
-
-Concerns over the potential for tipping points in ecological dynamics
-(Scheffer et al. 2001) highlight the dangers of uncertainty in
-ecological management and pose a substantial challenge to existing
-decision-theoretic approaches (Brozović and Schlenker 2011). Because
-intervention is often too late after a tipping point has been crossed
-(but see Hughes et al. (2013)), management is most often concerned with
-avoiding potentially catostrophic tipping points before any data is
-available at or following a transition that would more clearly reveal
-these regime shift dynamics (e.g. Bestelmeyer et al. 2012).
-
-To illustrate the value of the non-parametric Bayesian approach to
-management, we focus on example of a system containing such a tipping
-point whose dynamics can still be described by a simple, one-dimensional
-parametric model.\\We will focus on a simple parametric model for a
-single species (derived from fist principles by Allen et al. 2005) as
-our underlying ``reality''.
-
-\begin{align}
-X_{t+1} &= Z_t f(S_t) \\
-S_t &= X_t - h_t \\
-f(S_t) &= e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
-\end{align}
-
-Where $Z_t$ is multiplicative noise function with mean 1, representing
-stochastic growth. We will consider log-normal noise with shape
-parameter $\sigma_g$. We start with an example in which the parameters
-are $r =$ 2, $K =$ 10, $C =$ 5, and $\sigma_g =$ 0.05.
-
-As a low-dimensional system completely described by three parameters,
-this scenario should if anything be favorable to a parametric-based
-approach. This model contains an Allee effect, or tipping point, below
-which the population is not self-sustaining and shrinks to zero
-(Courchamp, Berec, and Gascoigne 2008).
-
-Both parametric and nonparametric approaches will require training data
-on which to base their model of the process. We generate the training
-data under the model described in Eq 1 for 35 time steps, under a known
-but not necessarily optimal sequence of harvest intensities, $h_t$. For
-simplicity we imagine a fishery that started from zero harvest pressure
-and has been gradually increasing the harvest.
-
-\textbf{\emph{Should we include any emprical examples?}}
-
 \subsubsection{Estimating parametric models}
 
-We estimate two parametric models from the data using a maximum
-likelihood approach. The first model is structurally identical to the
-true model (Eq 1), differing only in that it's parameters are estimated
-from the observed data rather than given. The alternative model is the
-Ricker model, which is structurally similar and commonly used in for
-such data.
+We estimate posterior distributions for two parametric models: one using
+the structurally correct model as given in Eq (1), which we refer to as
+the ``Parametric Bayes'' model, and another using the familiar Ricker
+model, using a Gibbs sampler as described (with source code) in the
+appendix). In addition we estimate the parameters of the structurally
+correct model by maximum likelihood.
 
-The maximum likelihood estimates (MLE) will assume that the process
-noise $Z_t$ is log-normal, as in the simulation itself. In reality the
-noise structure might differ from this assumption, making this a
-best-case scenario for the MLE approach.\\The GP must also make an
-assumption about the noise structure -- in this case we assume additive
-noise, though multiplicative noise would be more accurate. This will
-illustrate the flexibility of the GP in fitting different noise
-structures even without knowledge of the correct assumptions.
-
-The GP estimate is Bayesian approach which requires specification of
-prior distributions for the hyper-parameters (see Appendix for details,
-or @Rassmussen1996, Munch et al. 2005 for background), as given in Table
-1. Then given the data, both the GP and the MLE parametric models can be
-estimated as described above, as we implement in R in the code provided
-in supplementary materials.
+The posterior distribution for the hyper-parameters of the Gaussian
+process model are estimated by Metropolis-Hastings algorithm, again with
+details and code provided in the Appendix. Rasmussen and Williams (2006)
+provides an excellent general introduction to Gaussian Processes and
+Munch et al. (2005) first discusses their application in this context.
 
@@
 
-Figure 1 shows the inferred Gaussian Process compared to the true and
-the MLE parametric models, predicting the relationship between observed
-stock size (x-axis) to the stock size following recruitment the
-following year. The gray shading shows the uncertainty estimated by the
-GP. Unlike the MLE estimates, the GP uncertainty changes depending on
-where the data is available, allowing for a much greater level of
-uncertainty at low stock sizes. While it would be natural (and straight
-forward) to condition the GP on passing through the origin (0,0) (see
-appendix), the estimate shown here is based only on the observed data.
-The observed data from which each model is estimated is also shown. The
-observations come from only a limited region of state space
-corresponding to unharvested or weakly harvested system. No observations
-occur at the theoretical optimum (7.5 units, see Figure 2) or near the
-tipping point (5 units).
+\begin{figure}[htbp]
+\centering
+\includegraphics{Figure1.pdf}
+\caption{Mean dynamics predicted by each model.}
+\end{figure}
+
+Figure 1 shows the mean inferred state space dynamics of each model
+relative to the true model used to generate the data, predicting the
+relationship between observed stock size (x-axis) to the stock size
+after recruitment the following year. All models except the MLE model
+estimate a distribution around the means shown here, and all models
+estimate a level of process noise, which is independent of the state
+value (x). Note that in contrast to the other models shown, the mean
+Gaussian process corresponds to a distribution of curves - as indicated
+by the gray band - which itself has a mean shown in black. Note that
+this mean GP is thus more certain of the dynamics in the region where
+data is available then where it is not.
+
+While it would be natural (and straight forward) to condition the GP on
+passing through the origin (0,0) (see appendix), the estimate shown here
+is based only on the observed data. The observed data from which each
+model is estimated is also shown. The observations come from only a
+limited region of state space corresponding to unharvested or weakly
+harvested system. No observations occur at the theoretical optimum or
+near the tipping point.
 
@@
 \centering
-\includegraphics{figure/gp_plot.png}
-\caption{Graph of the inferred Gaussian process compared to the true
-process and maximum-likelihood estimated process. Graph shows the
-expected value for the function $f$ under each model. Two standard
-deviations from the estimated Gaussian process covariance (grey shading)
-is also shown. The training data is also shown as black points.}
+\includegraphics{Figure2.pdf}
+\caption{Steady-state optimal policies by each model.}
 \end{figure}
@@
 strategy aims to return the population to $S$, resulting in the
-hockey-stick shaped policies shown. Only the GP policy deviates slightly
-from this. The MLE parametric model corresponding to parametric
-uncertainty alone (gold line labeled ``Parametric'', that is, using the
-same structural form as the underlying simulation) nevertheless
-overestimates the carrying capacity of the stock and therefore
-underestimates the harvest rates, always waiting for the stock to rise
-further than is actually possible. The Ricker model (green line, showing
-``Structural'' inaccuracies) does not foresee the tipping point dynamics
-reflected in the concavity of the other three models (Figure 1), and
-selects a much more aggressive harvesting strategy.
+hockey-stick shaped policies shown.
 
@@
 \centering
-\includegraphics{figure/policies_plot.png}
-\caption{The steady-state optimal policy (infinite boundary) calculated
-under each model.}
+\includegraphics{Figure3.pdf}
+\caption{Managing the simulated system over 100 replicate instances
+under each policy}
 \end{figure}
@@
 expected from the policy curves, the structurally correct model
-(``Parametric'' uncertainty, gold) underharvests, leaving the stock to
-vary around it's unfished optimum. The GP overharvests the population
-slightly relative to the optimal value. The structurally incorrect
-Ricker model (``Structural'', green line) overharvests the population
+under-harvests, leaving the stock to vary around it's un-fished optimum.
+The structurally incorrect Ricker model over-harvests the population
 passed the tipping point consistently, resulting in the immediate crash
 of the stock and thus derives minimal profits.
-
-\textbf{\emph{(may need to show harvest dynamics since that is not clear
-from the figure! Also isn't general, sometimes does optimally, sometimes
-over-fishes. Perhaps need to show more examples.) May need to show
-profits too?}}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/sim_plot.png}
-\caption{Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under
-the policies derived from each of the estimated models, as well as the
-policy based on the exact underlying model.}
-\end{figure}
 
@@
 aspects are common to many conservation decision making problems, which
-thus merit greater use of non-parametric machine learning approaches
-that can best take advantage of them.
+thus merit greater use of non-parametric approaches that can best take
+advantage of them.
 
@@
 
-\subsubsection{3. Time scale for new data}
-
-In the sequential decision making problem we considered, we are
-presented with new data after each action. The relevant timescale for
-the prediction is thus not the long-term dynamics, which would be wildly
-divergent, but the dynamics over this much shorter interval. While
-ecologists may be hesitant to base continual management on a model with
-obviously inaccurate long-term behavior, engineers tend to consider the
-problem in frequency space and gravitate to the opposite position -- a
-good control model should prioritize high-frequency accuracy over low
-frequency accuracy. The differences in intuition may arise from the
-timescales at which each profession can typically adjust their control
-variables -- much faster for a control system of a chemical plant than
-state policy for a natural resource. Still, the lesson is clear: when
-facing repeated management decisions over a short timescale, such as
-setting annual harvests of a fishery, it may be more valuable to use a
-machine learning algorithm that makes accurate year-ahead predictions
-that capture some of the high-frequency fluctuations that appear only as
-noise in a mechanistic model of the long-term population dynamics.
-
 \subsection{Future directions}
 
- Jargony and vague While we have highlighted certain generic of this
-problem that allow the nonparametric approach to excel -- short
-timescales between successive observations and actions, the accuracy in
-the appropriate region of state space, the ability to express
-uncertainty outside the observed data -- there are equally several
-aspects for which the nonparametric Bayesian approach is nevertheless at
-a relative disadvantage. In this simulated example, the underlying
-dynamics are truly governed by a simple parametric model, allowing the
-parametric approaches to be more accurate. Similarly, because the
-dynamics are one-dimensional dynamics and lead to stable nodes (rather
-than other attractors such as limit-cycles resulting in oscillations),
-the training data provides relatively limited information about the
-dynamics. For these reasons, we anticipate that in higher-dimensional
-examples characteristic of ecosystem management problems that the
-machine learning approach will prove even more valuable.
+In this simulated example, the underlying dynamics are truly governed by
+a simple parametric model, allowing the parametric approaches to be more
+accurate. Similarly, because the dynamics are one-dimensional dynamics
+and lead to stable nodes (rather than other attractors such as
+limit-cycles resulting in oscillations), the training data provides
+relatively limited information about the dynamics. For these reasons, we
+anticipate that in higher-dimensional examples characteristic of
+ecosystem management problems that the machine learning approach will
+prove even more valuable.
 
@@
 Incorporating various sources of information into mechanistic models can
-be an immensely difficult due to the increased complexity involved. Only
-thanks to tandem advances in increasing computational power and
-hierarchical statistical methodology have we been able to tackle such
-intuitively important complexity (and the potentially new available data
-that accompanies it) such as spatial distribution, heterogeneities of
-space, time, and individuals, to shift to ecosystem-based approaches
-from single-species based approaches. Without the need to formulate
-mechanisms, many modern machine learning algorithms can leverage
-potential information from all available sources of data directly. The
-algorithms can recognize unanticipated or subtle patterns in large data
-sets that enable more accurate predictions than mechanistic models that
-are formulated at a more macroscopic level.
+be an immensely difficult due to the increased complexity involved.
+
+\subsubsection{Online learning}
+
+In our treatment here we have ignored the possibility of learning during
+the management phase, in which the additional observations of the stock
+size could potentially improve parameter estimates. While we intend to
+address this possibility in future work in the context of these
+non-parametric models, we have not addressed it here for pedagogical
+reasons. In the context presented here, it is clear that the differences
+in performance arise from differences in the uncertainty inherent in the
+model formulations, rather than from differing abilities to learn.
+Because we consider a threshold system, online learning would not change
+this generic feature of a lack of data in a certain range of the state
+space which is better captured by the Gaussian process.
 
@@
 
- \textbf{\emph{Possible additional discussion material?}}
-
-\begin{itemize}
-\item
-  Discuss constant escapement in model, in policies.
-\item
-  Limitations of this comparison: Are the maximum-likelihood solutions a
-  straw man?
-\item
-  Discussion of alternative related approaches: POMDP/MOMDP
-\item
-  Online learning
-\item
-  Multiple step-ahead predictions
-\item
-  Explicitly accomidating additional uncertainties
-\item
-  Improving inference of optimal policy from the GP 
-\end{itemize}
-
-\begin{center}\rule{3in}{0.4pt}\end{center}
-
-\section{Appendix / Supplementary Materials}
-
-\subsubsection{The Gaussian process}
-
-The Gaussian process is defined by a covariance kernel. By requiring our
-kernel to follow a generic functional form, we can compactly describe
-the Gaussian process using only a handful of parameters (Table 1)
-
-\begin{longtable}[c]{@{}ll@{}}
-\hline\noalign{\medskip}
-parameter & interpretation
-\\\noalign{\medskip}
-\hline\noalign{\medskip}
-$\sigma^2$ & The process noise (from the kernel)
-\\\noalign{\medskip}
-$\tau^2$ & Variance around the mean
-\\\noalign{\medskip}
-$\beta_0$ & The mean is given by a linear model of slope $\beta$
-\\\noalign{\medskip}
-$d$ & The length-scale of the covariance function
-\\\noalign{\medskip}
-$n$ & The observation error
-\\\noalign{\medskip}
-\hline
-\noalign{\medskip}
-\caption{Table of parameters for the Gaussian process}
-\end{longtable}
-
-Rather than estimate values of these parameters directly, we take a
-hierarchical approach of placing prior distributions on each. Following
-@Gramarcy2005 we use a Gaussian prior on $\beta_0$, an inverse gamma
-prior on $\sigma^2$ and $\tau^2$, a gamma prior on the observation noise
-$n$, and exponential prior on the length-scale $d$.
-
-\subsubsection{Formulating a dynamic programming solution}
-
-The fishery management problem over an infinite time horizon can be
-stated as:
-
-\begin{align}
-& \max_{ \{h_t\} \geq 0 } \mathbf{E} \lbrace \sum_0^\infty \delta^t \Pi(h_t) \rbrace \\
-& \mathrm{s.t.}  \\
- & X_t = Z_t f\left(S_{t-1}\right) \\
- & S_t = X_t - h_t \\
- & X_t  \geq 0 
-\end{align}
-
-Where $\mathbf{E}$ is the expectation operator, $\delta$ the discount
-rate, $\Pi(h_t)$ the profit expected from a harvest of $h_t$, and other
-terms as in Eq. (1). For simplicity, we have assumed that profits depend
-only on the chosen harvest; simplifying further we will usually consider
-profits to be proportional to harvest, $\Pi(h_t) = h_t$.
-
-Once the posterior Gaussian process (GP) has been estimated (e.g.~see
-Munch et al. 2005), it is necessary to adapt it in place of the
-parametric equation for the stochastic dynamic programming (SDP)
-solution (see Mangel and Clark 1988 for a detailed description of
-parametric SDP methods) to the optimal policy. The essense of the idea
-is straight forward -- we will use the estimated GP in place of the
-parametric growth function to determine the stochastic transition matrix
-on which the SDP calculations are based.
-
-The posterior Gaussian process is completely defined by the expected
-value and covariance matrix at a defined set of training points. For
-simplicity we will consider a these points to fall on a discrete,
-uniform grid $x$ of 101 points from 0 to 15 (1.5 times the positive
-equilibrium $K$). Again to keep things simple we will use this same grid
-discretization for the parametric approach. Other options for choosing
-the grid points, including collocation methods and functional basis
-expansion (or even using Gaussian processes in place of the discrete
-optimization; an entirely different context in which GP can be used in
-SDP, see (Deisenroth, Rasmussen, and Peters 2009)) could also be
-considered.
-
-\subsection{Pseudocode for the determining the transtion matrix from the
-GP}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{for(h in h_grid)}
-  \NormalTok{F_h =}\StringTok{ }\NormalTok{for(x_j in grid)}
-          \NormalTok{for(i in }\DecValTok{1}\NormalTok{:N) }
-            \KeywordTok{dnorm}\NormalTok{(x_j, mu[i]-h, V[i])}
-\end{Highlighting}
-\end{Shaded}
-
-A transition matrix for each of the parametric models $f$ is calculated
-using the log-normal density with mean $f(x_i)$ and log-variance as
-estimated by maximum likelihood. From the discrete transition matrix we
-may write down the Bellman recursion defining the the stochastic dynamic
-programming iteration:
-
-\begin{equation}
-V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
-\end{equation}
-
-where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$ is
-control (harvest level) chosen. Numerically, the maximization is
-accomplished as follows. Consider the set of possible control values to
-be the discrete 101 values corresponding the the grid of stock sizes.
-Then for each $h_t$ there is a corresponding transition matrix
-$\mathbf{F}_h$ determined as described above but with mean
-$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element
-corresponds to the value of having stock $x_i$ at time $t$. Then let
-$\Pi_h$ be the vector whose $i$th element indicates the profit from
-harvesting at intensity $h_t$ given a population $x_i$ (e.g.
-$\max(x_i, h_t)$ since one cannot harvest more fish then the current
-population size). Then the Bellman recursion can be given in matrix form
-as
-
-\[V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)\]
-
-where the sum is element by element and the expectation is computed by
-the matrix multiplication $\mathbf{F} V_{t+1}$.
-
-\subsubsection{Pseudocode for the Bellman iteration}
-
-\begin{Shaded}
-\begin{Highlighting}[]
- \NormalTok{V1 <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
-      \NormalTok{delta *}\StringTok{ }\NormalTok{F[[h]] %*%}\StringTok{ }\NormalTok{V +}\StringTok{  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
-    \NormalTok{\})}
-    \CommentTok{# find havest, h that gives the maximum value}
-    \NormalTok{out <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
-      \NormalTok{value <-}\StringTok{ }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
-      \NormalTok{index <-}\StringTok{ }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
-      \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
-    \NormalTok{\})}
-    \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
-    \NormalTok{V <-}\StringTok{ }\NormalTok{out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
-    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <-}\StringTok{ }\NormalTok{out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
-\end{Highlighting}
-\end{Shaded}
-
-\emph{Currently this shows the literal R code, should be adapted}
-
-\subsection{MCMC posterior distributions and convergence analysis}
-
-\textbf{\emph{Can we get this figure into the main text with appropriate
-discussion without getting too technical?}}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/posteriors.png}
-\caption{Histogram of posterior distributions for the estimated Gaussian
-Process shown in Figure 1. Prior distributions overlaid.}
-\end{figure}
-
-Gramacy (2005)
-
-\subsection{Tables of nuisance parameters, sensitivity analysis}
-
-\subsubsection{List of hyper-parameters, prior distributions and their
-parameters}
-
-\subsection{Reproducible code, ``Research Compendium''}
+\section{References}
 
@@
 
+Barnosky, Anthony D., Elizabeth a Hadly, Jordi Bascompte, Eric L.
+Berlow, James H. Brown, Mikael Fortelius, Wayne M. Getz, et al. 2012.
+``Approaching a state shift in Earth's biosphere.'' \emph{Nature} 486
+(7401) (jun): 52--58. doi:10.1038/nature11018.
+\url{http://www.nature.com/doifinder/10.1038/nature11018}.
+
 Bestelmeyer, Brandon T., Michael C. Duniway, Darren K. James, Laura M.
@@
 
+Boettiger, Carl, and Alan Hastings. 2013. ``Tipping points: From
+patterns to predictions.'' \emph{Nature} 493 (7431) (jan): 157--158.
+doi:10.1038/493157a.
+\url{http://www.nature.com/doifinder/10.1038/493157a}.
+
 Brozović, Nicholas, and Wolfram Schlenker. 2011. ``Optimal management of
@@
 \url{http://linkinghub.elsevier.com/retrieve/pii/S0921800910004167}.
-
-Clark, Colin W., and Geoffrey P. Kirkwood. 1986. ``On uncertain
-renewable resource stocks: Optimal harvest policies and the value of
-stock surveys.'' \emph{Journal of Environmental Economics and
-Management} 13 (3) (sep): 235--244. doi:10.1016/0095-0696(86)90024-0.
-\url{http://linkinghub.elsevier.com/retrieve/pii/0095069686900240}.
 
@@
 
-Geritz, Stefan a. H., and Eva Kisdi. 2012. ``Mathematical ecology: why
-mechanistic models?'' \emph{Journal of mathematical biology} 65 (6-7)
+Geritz, Stefan A. H., and Eva Kisdi. 2012. ``Mathematical ecology: why
+mechanistic models?.'' \emph{Journal of mathematical biology} 65 (6-7)
 (dec): 1411--5. doi:10.1007/s00285-011-0496-3.
@@
 \url{http://www.jstor.org/stable/10.2307/1825571}.
-
-Gramacy, Robert B. 2005. ``BAYESIAN TREED GAUSSIAN PROCESS MODELS.''
 
@@
 
-Mangel, Marc, and Colin W. Clark. 1985. \emph{Decision and control in
-uncertain resource systems, Volume 172 (Mathematics in Science and
-Engineering)}. Academic Press.
-\url{http://www.amazon.com/Decision-uncertain-resource-Mathematics-Engineering/dp/0124687202}.
+Mangel, Marc. 1985. ``Decision and control in uncertain resource
+systems:'' 255.
+\href{http://www.amazon.com/Decision-uncertain-resource-Mathematics-Engineering/dp/0124687202 http://dl.acm.org/citation.cfm?id=537497}{http://www.amazon.com/Decision-uncertain-resource-Mathematics-Engineering/dp/0124687202
+http://dl.acm.org/citation.cfm?id=537497}.
 
----------. 1988. \emph{Dynamic Modeling in Behavioral Ecology}.
-Princeton: Princeton University Press.
+Mangel, Marc, and Colin W. Clark. 1988. \emph{Dynamic Modeling in
+Behavioral Ecology}. Ed. John Krebs and Tim Clutton-Brock. Princeton:
+Princeton University Press.
 
@@
 
-Rasmussen, C. E., and C. K. I. Williams. 2006. \emph{Gaussian Processes
-for Machine Learning}. Ed. Thomas Dietterich. Boston: MIT Press,.
-\url{www.GaussianProcess.org/gpml}.
+Rasmussen, Carl Edward, and C. K. I. Williams. 2006. \emph{Gaussian
+Processes for Machine Learning}. Ed. Thomas Dietterich. Boston: MIT
+Press,. \url{www.GaussianProcess.org/gpml}.
 
@@
 
-Roughgarden, Joan E., and F. Smith. 1996. ``Why fisheries collapse and
-what to do about it.'' \emph{Proceedings of the National Academy of
-Sciences of the United States of America} 93 (10): 5078.
-\url{http://www.pnas.org/content/93/10/5078.short}.
-
 Schapaugh, Adam W., and Andrew J. Tyre. 2013. ``Accounting for
@@
 
+Scheffer, Marten, Jordi Bascompte, William A. Brock, Victor Brovkin,
+Stephen R. Carpenter, Vasilis Dakos, Hermann Held, Egbert H. van Nes,
+Max Rietkerk, and George Sugihara. 2009. ``Early-warning signals for
+critical transitions.'' \emph{Nature} 461 (7260): 53--9.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/19727193}.
+
 Scheffer, Marten, Stephen R. Carpenter, J. A. Foley, C. Folke, and B.
@@
 \url{http://www.ncbi.nlm.nih.gov/pubmed/11595939}.
-
-Sethi, Gautam, Christopher Costello, Anthony Fisher, Michael Hanemann,
-and Larry Karp. 2005. ``Fishery management under multiple uncertainty.''
-\emph{Journal of Environmental Economics and Management} 50 (2) (sep):
-300--318. doi:10.1016/j.jeem.2004.11.005.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069605000057}.
 
@@
 
+\pagebreak\pagebreak
+
+\section{Appendix}
+
+The appendices have not yet been assembled. Meanwhile, code to repeat
+the analyses, along with a complete log of all research conducted on
+this project, can be found at:
+\href{https://github.com/cboettig/nonparametric-bayes/}{https://github.com/cboettig/nonparametric-bayes}
+
+\subsection{Markov Chain Monte Carlo Analysis}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{gp_traces_densities1.pdf}
+\caption{Traces of the GP hyper-parameters (after transforming back from
+log space)}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{gp_traces_densities2.pdf}
+\caption{Posterior densities of the GP hyper-parameters (after
+transforming back from log space). Prior distribution over the interval
+sampled shown as red curves.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{ricker_traces.pdf}
+\caption{Traces of the Ricker parameters}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{ricker_posteriors.pdf}
+\caption{Posterior densities of the Ricker parameter estimates. Prior
+distributions over the interval sampled shown as red curves. Note the
+bimodal estimate of $\log(r)$ which leads to the poor fit in Figure 1}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{parametric_bayes_traces.pdf}
+\caption{Traces of the structurally accurate Allen model parameters}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{parametric_bayes_posteriors.pdf}
+\caption{Posterior densities for the parameters of the Allen model.
+Prior distributions over the interval sampled shown as red curves.}
+\end{figure}
+
 \end{document}

### 2013-05-03 inst/examples/BUGS/allen.Rmd
replicates

--- 2013-05-02 inst/examples/BUGS/allen.Rmd
@@
 ```{r plotting-options}
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA)
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA,
+               fig.width=6, fig.height=4, dev="CairoPDF")
 
@@
 library(ggplot2) # plotting 
-opts_knit$set(upload.fun = socialR::flickr.url)
+#opts_knit$set(upload.fun = socialR::flickr.url)
 theme_set(theme_bw(base_size=10))
@@
 xT <- 0
-seed_i <- 111
 Xo <- K # observations start from
 x0 <- Xo # simulation under policy starts from
-Tobs <- 35
+Tobs <- 40
 ```
@@
 ```{r obs, dependson="sdp-pars"}
+  set.seed(1234)
   #harvest <- sort(rep(seq(0, .5, length=7), 5))
@@
 
-```{r}
+```{r gp_traces_densities}
 plots <- summary_gp_mcmc(gp)
-plots[[1]]
-plots[[2]]
 ```
@@
 par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
+```
 
+```{r parametric_bayes_traces}
 # Traces
 ggplot(par_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
+```
 
+```{r parametric_bayes_posteriors}
 ## priors (untransformed variables)
@@
 
-par_priors <- list(K = K_prior, deviance = function(x) 0 * x, logr0 = logr_prior, logtheta = logtheta_prior, stdQ = stdQ_prior, stdR = stdR_prior)
+par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
+                   logr0 = logr_prior, logtheta = logtheta_prior, 
+                   stdQ = stdQ_prior, stdR = stdR_prior)
 
@@
 
-#### Convergence diagnostics for parametric bayes
+#### Convergence diagnostics for parametric bayes Ricker model
 
-```{r}
+```{r ricker_traces}
 jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
@@
 ggplot(par_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
+```
 
+```{r ricker_posteriors}
 ## priors (untransformed variables)
@@
   facet_wrap(~ variable, scale="free", ncol=2)
-
 ```
@@
 ```{r}
-
 true_means <- sapply(x_grid, f, 0, p)
@@
 names(models) <- c("x", "method", "value")
+```
 
+```{r Figure1}
 plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-    geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
+    geom_line(data=models, aes(x, value, col=method), lwd=1, alpha=0.8) + 
     geom_point(data=obs, aes(x,y), alpha=0.8) + 
@@
 
-```{r}
+```{r Figure2}
 policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
@@
 
-```{r}
-
+```{r sims}
 sims <- lapply(OPT, function(D){
@@
 dt$method = factor(dt$method, ordered=TRUE, levels=names(OPT))
+```
 
+```{r Figure3}
 ggplot(dt) + 

### 2013-05-06 inst/examples/BUGS/allen.Rmd
replicate runs

@@
 opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA,
-               fig.width=6, fig.height=4, dev="CairoPDF")
+               fig.width=6, fig.height=4)
 
@@
 library(ggplot2) # plotting 
-#opts_knit$set(upload.fun = socialR::flickr.url)
+opts_knit$set(upload.fun = socialR::flickr.url)
 theme_set(theme_bw(base_size=10))

### 2013-05-06 inst/examples/BUGS/bayesian-ricker.Rmd
test script for ricker (good for template for extending BUGS)

--- 2013-05-02 inst/examples/BUGS/allen.Rmd
@@
 
-```{r runsetup}
-setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
-```
-
-Plotting and knitr options, (can generally be ignored)
-
-```{r plotting-options}
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA)
-
-library(knitcitations)
-
-library(ggplot2) # plotting 
-opts_knit$set(upload.fun = socialR::flickr.url)
-theme_set(theme_bw(base_size=10))
-theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
-             plot.background = element_rect(fill = "transparent", colour = NA))
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
-               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-```
-Load necessary libraries,
-
-```{r libraries, cache=FALSE, message=FALSE, warning=FALSE}
-library(nonparametricbayes) # loads the rest as dependencies
-```
-
-### Model and parameters
-
-Uses the model derived in `citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
-
-```{r}
-f <- RickerAllee
-p <- c(2, 8, 5)
-K <- p[2]
-allee <- p[3]
-```
-
-Various parameters defining noise dynamics, grid, and policy costs.
-
-```{r sdp-pars, dependson="stateeq"}
-sigma_g <- 0.05
-sigma_m <- 0.0
-z_g <- function() rlnorm(1, 0, sigma_g)
-z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
-x_grid <- seq(0, 1.5 * K, length=50)
-h_grid <- x_grid
-profit <- function(x,h) pmin(x, h)
-delta <- 0.01
-OptTime <- 30  # stationarity with unstable models is tricky thing
-reward <- 0
-xT <- 0
-seed_i <- 111
-Xo <- K # observations start from
-x0 <- Xo # simulation under policy starts from
-Tobs <- 30
-```
-
-### Sample Data
-
-```{r obs, dependson="sdp-pars"}
-  #harvest <- sort(rep(seq(0, .5, length=7), 5))
-  x <- numeric(Tobs)
-  x[1] <- 5.5
-  nz <- 1
-  for(t in 1:(Tobs-1))
-    x[t+1] = z_g() * f(x[t], h=0, p=p)
-  obs <- data.frame(x = c(rep(0,nz), 
-                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
-                    y = c(rep(0,nz), 
-                          x[2:Tobs]))
-```
-
-## Maximum Likelihood
-
-```{r mle, dependson="obs"}
-alt <- par_est(obs,  init = c(r = p[1], 
-                              K = mean(obs$x[obs$x>0]), 
-                              s = sigma_g))
-est <- par_est_allee(obs, f, p,  
-                     init = c(r = p[1] + 1, 
-                              K = p[2] + 2, 
-                              C = p[3] + 2, 
-                              s = sigma_g))
-```
-
-## Non-parametric Bayes
-
-```{r gp-priors}
-#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
-s2.p <- c(5,5)  
-d.p = c(10, 1/0.1)
-```
-
-Estimate the Gaussian Process (nonparametric Bayesian fit)
-
-```{r}
-gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
-gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)
-```
-
-Show traces and posteriors against priors
-
-```{r}
-plots <- summary_gp_mcmc(gp)
-plots[[1]]
-plots[[2]]
-```
-
-```{r}
-# Summarize the GP model
-tgp_dat <- 
-    data.frame(  x = x_grid, 
-                 y = gp_dat$E_Ef, 
-                 ymin = gp_dat$E_Ef - 2 * sqrt(gp_dat$E_Vf), 
-                 ymax = gp_dat$E_Ef + 2 * sqrt(gp_dat$E_Vf) )
-```
-
 ### Parametric Bayes
-
-We initiate the MCMC chain (`init_p`) using the true values of the parameters `p` from the simulation.
-While impossible in real data, this gives the parametric Bayesian approach the best chance at succeeding.
-`y` is the timeseries (recall `obs` has the $x_t$, $x_{t+1}$ pairs)
-
-```{r}
-# a bit unfair to start with the correct values, but anyhow...
-init_p = p
-names(init_p) = c("r0", "K", "theta")
-y <- obs$x[-1] 
-N=length(y);
-```
-
-We'll be using the JAGS Gibbs sampler, a recent open source BUGS implementation with an R interface that works on most platforms.  We initialize the usual MCMC parameters; see `?jags` for details.
-
-```{r}
-jags.data <- list("N","y")
-n.chains = 1
-n.iter = 40000
-n.burnin = floor(10000)
-n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
-```
-
-The actual model is defined in a `model.file` that contains an R function that is automatically translated into BUGS code by *R2WinBUGS*.
-The file defines the priors and the model, as seen when read in here
-
-```{r}
-cat(readLines(con="bugmodel-UPrior.txt"), sep="\n")
-```
-
-We define which parameters to keep track of, and set the initial values of parameters in the transformed space used by the MCMC.
-We use logarithms to maintain strictly positive values of parameters where appropriate.
-Because our priors on the noise parameters are inverse gamma distributed.
-
-```{r}
-# Uniform priors on standard deviation terms
-jags.params=c("K","logr0","logtheta","stdQ", "stdR")
-jags.inits <- function(){
-  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
-}
-
-set.seed(12345)
-
-time_jags <- system.time(       
-  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
-                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="bugmodel-UPrior.txt")
-)         
-time_jags <- unname(time_jags["elapsed"]);
-```
-
-#### Convergence diagnostics for parametric bayes
-
-```{r}
-jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
-
-# Traces
-ggplot(par_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
-
-## priors (untransformed variables)
-K_prior <- function(x) dunif(x, 0.01, 40)
-logr_prior <- function(x) dunif(x, -6, 6)
-logtheta_prior <- function(x) dunif(x, -6, 6)
-stdQ_prior <- function(x) dunif(x, 0.001, 100)
-stdR_prior <- function(x) dunif(x, 0.001, 100)
-
-par_priors <- list(K = K_prior, deviance = function(x) 0 * x, logr0 = logr_prior, logtheta = logtheta_prior, stdQ = stdQ_prior, stdR = stdR_prior)
-
-par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100) 
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-
-# posterior distributions
-ggplot(par_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=2)
-
-```
-
-```{r}
-# um, cleaner if we were just be using the long form, par_posterior
-mcmc <- as.mcmc(jagsfit)
-mcmcall <- mcmc[,-2]
-who <- colnames(mcmcall)
-who 
-mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4],mcmcall[,5])
-colnames(mcmcall) <- who
-
-pardist <- mcmcall
-pardist[,2] = exp(pardist[,2]) # transform model parameters back first
-pardist[,3] = exp(pardist[,3])
-
-bayes_coef <- apply(pardist,2,mean)
-bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3]))
-bayes_pars
-```
-
-```{r}
-par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)
-```
-
-## Parametric Bayes based on the structurally wrong model
 
@@
 ```
-
-### Phase-space diagram of the expected dynamics
-
-```{r}
-
-  true_means <- sapply(x_grid, f, 0, p)
-  alt_means <- sapply(x_grid, alt$f, 0, ricker_bayes_pars[c(1,2)])
-  est_means <- sapply(x_grid, est$f, 0, est$p)
-  par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)
-
-models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
-                     MLE=est_means, Ricker=alt_means, 
-                     Parametric.Bayes = par_bayes_means)
-
-models <- melt(models, id="x")
-names(models) <- c("x", "method", "value")
-
-plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-    geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
-    geom_point(data=obs, aes(x,y), alpha=0.8) + 
-    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
-    scale_colour_manual(values=cbPalette) 
-print(plot_gp)
-```
-
-## Optimal policies by value iteration
-
-Compute the optimal policy under each model using stochastic dynamic programming. We begin with the policy based on the GP model,
-
-```{r}
-MaxT = 1000
-# uses expected values from GP, instead of integrating over posterior
-#matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)
-
-# Integrate over posteriors 
-matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid)
-
-# Solve the SDP using the GP-derived transition matrix
-opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
-```
-
-Determine the optimal policy based on the true and MLE models
-
-```{r}
-matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-matrices_estimated <- f_transition_matrix(est$f, est$p, x_grid, h_grid, est$sigma_g)
-opt_estimated <- value_iteration(matrices_estimated, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-```
-
-Determine the optimal policy based on parametric Bayesian model
-
-```{r}
-matrices_par_bayes <- parameter_uncertainty_SDP(f, p, x_grid, h_grid, pardist)
-opt_par_bayes <- value_iteration(matrices_par_bayes, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-```
-
-Bayesian Ricker
-
-```{r}
-matrices_alt <- parameter_uncertainty_SDP(alt$f, ricker_bayes_pars[1,2], x_grid, h_grid, ricker_pardist)
-opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-```
-
-Assemble the data
-
-```{r}
-OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_alt$D, Parametric.Bayes = opt_par_bayes$D)
-colorkey=cbPalette
-names(colorkey) = names(OPT) 
-```
-
-## Graph of the optimal policies
-
-```{r}
-policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
-names(policies) <- c("stock", "method", "value")
-
-ggplot(policies, aes(stock, stock - value, color=method)) +
-  geom_line(lwd=1.2, alpha=0.8) + xlab("stock size") + ylab("escapement")  +
-  scale_colour_manual(values=colorkey)
-```
-
-## Simulate 100 realizations managed under each of the policies
-
-```{r}
-
-sims <- lapply(OPT, function(D){
-  set.seed(1)
-  lapply(1:100, function(i) 
-    ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
-  )
-})
-
-dat <- melt(sims, id=names(sims[[1]][[1]]))
-dt <- data.table(dat)
-setnames(dt, c("L1", "L2"), c("method", "reps")) 
-# Legend in original ordering please, not alphabetical: 
-dt$method = factor(dt$method, ordered=TRUE, levels=names(OPT))
-
-ggplot(dt) + 
-  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
-
-```

### 2013-05-07 inst/examples/BUGS/allen.Rmd
added parametric bayesian myers and ricker for comparison

--- 2013-05-03 inst/examples/BUGS/allen.Rmd
@@
 delta <- 0.01
-OptTime <- 30  # stationarity with unstable models is tricky thing
+OptTime <- 50  # stationarity with unstable models is tricky thing
 reward <- 0
@@
 We use logarithms to maintain strictly positive values of parameters where appropriate.
-Because our priors on the noise parameters are inverse gamma distributed.
 
@@
 ```{r}
-# a bit unfair to start with the correct values, but anyhow...
 init_p = p
@@
 n.chains = 1
-n.iter = 40000
+n.iter = 400000
 n.burnin = floor(10000)
@@
 # Traces
-ggplot(par_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
+ggplot(par_posteriors) + geom_line(aes(index, value)) + 
+  facet_wrap(~ variable, scale="free", ncol=1)
 ```
@@
 ```{r}
-# um, cleaner if we were just be using the long form, par_posterior
-mcmc <- as.mcmc(jagsfit)
-mcmcall <- mcmc[,-2]
-who <- colnames(mcmcall)
-who 
-mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4])
-colnames(mcmcall) <- who
+ricker_pardist <- jags_matrix[! names(jags_matrix) == "deviance" ]
+ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first
+
+posterior.mode <- function(x) {
+  ux <- unique(x)
+  ux[which.max(tabulate(match(x, ux)))]
+}
+
+apply(ricker_pardist,2,mean)
+bayes_coef <- apply(ricker_pardist,2, posterior.mode) # much better estimates
+ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
+ricker_bayes_pars
+```
+
+## Write the external bugs file
+
+```{r}
+logr0_prior_p <- c(-6.0, 6.0)
+logtheta_prior_p <- c(-6.0, 6.0)
+logK_prior_p <- c(-6.0, 6.0)
+stdQ_prior_p <- c(0.0001, 100)
+stdR_prior_p <- c(0.0001, 100)
+
+bugs.model <- 
+paste(sprintf(
+"model{
+  logr0    ~ dunif(%s, %s)
+  logtheta    ~ dunif(%s, %s)
+  logK    ~ dunif(%s, %s)
+  stdQ ~ dunif(%s, %s)
+  stdR ~ dunif(%s, %s)", 
+  logr0_prior_p[1], logr0_prior_p[2],
+  logtheta_prior_p[1], logtheta_prior_p[2],
+  logK_prior_p[1], logK_prior_p[2],
+  stdQ_prior_p[1], stdQ_prior_p[2],
+  stdR_prior_p[1], stdR_prior_p[2]), 
+  "
+
+  iQ <- 1 / (stdQ * stdQ);
+  iR <- 1 / (stdR * stdR);
+
+  r0 <- exp(logr0)
+  theta <- exp(logtheta)
+  K <- exp(logK)
+
+  x[1] ~ dunif(0, 10)
+
+  for(t in 1:(N-1)){
+    mu[t] <- r0 * pow(abs(x[t]), theta) / (1 + pow(abs(x[t]), theta) / K)
+    x[t+1] ~ dnorm(mu[t], iQ) 
+  }
+
+  for(t in 1:(N)){
+    y[t] ~ dnorm(x[t], iR)
+  }
+}")
+
+writeLines(bugs.model, "myers.bugs")
 ```
@@
 ```{r}
-ricker_pardist <- mcmcall
-ricker_pardist[,2] = exp(ricker_pardist[,2]) # transform model parameters back first
 
-bayes_coef <- apply(ricker_pardist,2,mean)
-ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
-ricker_bayes_pars
+## priors (untransformed variables)
+logK_prior     <- function(x) dunif(x, logK_prior_p[1], logK_prior_p[2])
+logr_prior     <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
+logtheta_prior <- function(x) dunif(x, logtheta_prior_p[1], logtheta_prior_p[2])
+stdQ_prior     <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
+stdR_prior     <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
+
+```
+
+### Parametric Bayes based on myers model
+
+Contains the
+
+```{r}
+
+init_p = c(1, 1, 1)
+names(init_p) = c("r0", "theta", "K")
+y <- obs$x[-1] 
+N=length(y);
+```
+
+We'll be using the JAGS Gibbs sampler, a recent open source BUGS implementation with an R interface that works on most platforms.  We initialize the usual MCMC parameters; see `?jags` for details.
+
+```{r}
+jags.data <- list("N","y")
+n.chains = 1
+n.iter = 40000
+n.burnin = floor(10000)
+n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
+```
+
+```{r}
+jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
+jags.inits <- function(){
+  list("logr0"=log(init_p["r0"]), "logtheta" = log(init_p["theta"]),  
+       "logK"=log(init_p["K"]), "stdQ"=sqrt(0.05), 
+       "stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+}
+```
+
+```{r}
+set.seed(12345)
+time_jags <- system.time(       
+  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, 
+                  n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
+                  n.burnin=n.burnin, model.file="myers.bugs")
+)         
+time_jags <- unname(time_jags["elapsed"]);
+```
+
+Convergence diagnostics for parametric bayes
+
+```{r}
+jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
+par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
+
+## Traces
+ggplot(par_posteriors) + geom_line(aes(index, value)) +
+  facet_wrap(~ variable, scale="free", ncol=1)
+
+```
+
+```{r}
+## prior curve functions, must be in order
+par_priors <- list( deviance = function(x) 0 * x, logK = logK_prior,
+                    logr0 = logr_prior, logtheta = logtheta_prior, 
+                    stdQ = stdQ_prior, stdR = stdR_prior)
+
+## calculate points along the prior curve
+par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
+    grid <- seq(min(dd$value), max(dd$value), length = 100) 
+    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
+})
+```
+
+```{r}
+## posterior distributions
+ggplot(par_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
+  facet_wrap(~ variable, scale="free", ncol=2)
+```
+
+```{r}
+myers_pardist <- as.matrix(jags_matrix[2:6])
+myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
+myers_pardist[,2] = exp(myers_pardist[,2]) # transform model parameters back first
+myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
+colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
+
+apply(myers_pardist,2,mean)
+bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
+
+myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
 ```
@@
 
+myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
+
 models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
                      MLE=est_means, Ricker=alt_means, 
-                     Parametric.Bayes = par_bayes_means)
+                     Parametric.Bayes = par_bayes_means,
+                     Myers = myers_means)
 
@@
 ```{r}
-matrices_par_bayes <- parameter_uncertainty_SDP(f, p, x_grid, h_grid, pardist)
+allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
+matrices_par_bayes <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
 opt_par_bayes <- value_iteration(matrices_par_bayes, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
@@
 ```{r}
-matrices_alt <- parameter_uncertainty_SDP(alt$f, ricker_bayes_pars[1,2], x_grid, h_grid, ricker_pardist)
+ricker_f <- function(x, h, p) alt$f(x, h, unname(p[c(2, 1)]))
+matrices_alt <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
 opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+```
+
+Bayesian Myers model
+
+```{r}
+myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
+matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
+myers_alt <- value_iteration(matrices_myers, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
 ```
@@
 ```{r}
-OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_alt$D, Parametric.Bayes = opt_par_bayes$D)
+OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_alt$D, Parametric.Bayes = opt_par_bayes$D, Myers = myers_alt$D)
 colorkey=cbPalette
@@
   scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
+```
 
+```{r profits}
+Profit <- dt[, sum(profit), by=c("reps", "method")]
+Profit[, mean(V1), by="method"]
 ```
+
+```{r totalprofits}
+ggplot(Profit, aes(V1)) + geom_histogram() + 
+  facet_wrap(~method, scales = "free_y") + guides(legend.position = "none")
+```

### 2013-05-07 inst/examples/BUGS/bayesian-ricker.Rmd
added parametric bayesian myers and ricker for comparison

@@
-# Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach
+```{r}
+setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
+opts_chunk$set(tidy = FALSE, warning = FALSE, message = FALSE, cache = FALSE, 
+    comment = NA, fig.width = 6, fig.height = 4)
+library(knitcitations)
+library(ggplot2)  # plotting
+opts_knit$set(upload.fun = socialR::flickr.url)
+theme_set(theme_bw(base_size = 10))
+theme_update(panel.background = element_rect(fill = "transparent", colour = NA), 
+    plot.background = element_rect(fill = "transparent", colour = NA))
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+```
 
-### Parametric Bayes
-
-We initiate the MCMC chain (`init_p`) using the true values of the parameters `p` from the simulation.
-While impossible in real data, this gives the parametric Bayesian approach the best chance at succeeding.
-`y` is the timeseries (recall `obs` has the $x_t$, $x_{t+1}$ pairs)
+Load necessary libraries,
 
 ```{r}
-# a bit unfair to start with the correct values, but anyhow...
-init_p = p
-names(init_p) = c("r0", "K")
+library(nonparametricbayes) # loads the rest as dependencies
+```
+
+### Model and parameters
+
+Uses the model derived in `r citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
+
+```{r}
+f <- RickerAllee
+p <- c(2, 8, 5)
+K <- p[2]
+allee <- p[3]
+```
+
+Various parameters defining noise dynamics, grid, and policy costs.
+
+```{r}
+sigma_g <- 0.05
+sigma_m <- 0.0
+z_g <- function() rlnorm(1, 0, sigma_g)
+z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
+x_grid <- seq(0, 1.5 * K, length=50)
+h_grid <- x_grid
+profit <- function(x,h) pmin(x, h)
+delta <- 0.01
+OptTime <- 30  # stationarity with unstable models is tricky thing
+reward <- 0
+xT <- 0
+Xo <- K # observations start from
+x0 <- Xo # simulation under policy starts from
+Tobs <- 40
+```
+
+### Sample Data
+
+```{r}
+  set.seed(1234)
+  #harvest <- sort(rep(seq(0, .5, length=7), 5))
+  x <- numeric(Tobs)
+  x[1] <- 6
+  nz <- 1
+  for(t in 1:(Tobs-1))
+    x[t+1] = z_g() * f(x[t], h=0, p=p)
+  obs <- data.frame(x = c(rep(0,nz), 
+                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
+                    y = c(rep(0,nz), 
+                          x[2:Tobs]))
+```
+
+## Write the external bugs file
+
+```{r}
+logr0_prior_p <- c(-6.0, 6.0)
+logtheta_prior_p <- c(-6.0, 6.0)
+logK_prior_p <- c(-6.0, 6.0)
+stdQ_prior_p <- c(0.0001, 100)
+stdR_prior_p <- c(0.0001, 100)
+
+bugs.model <- 
+paste(sprintf(
+"model{
+  logr0    ~ dunif(%s, %s)
+  logtheta    ~ dunif(%s, %s)
+  logK    ~ dunif(%s, %s)
+  stdQ ~ dunif(%s, %s)
+  stdR ~ dunif(%s, %s)", 
+  logr0_prior_p[1], logr0_prior_p[2],
+  logtheta_prior_p[1], logtheta_prior_p[2],
+  logK_prior_p[1], logK_prior_p[2],
+  stdQ_prior_p[1], stdQ_prior_p[2],
+  stdR_prior_p[1], stdR_prior_p[2]), 
+  "
+
+  iQ <- 1 / (stdQ * stdQ);
+  iR <- 1 / (stdR * stdR);
+
+  r0 <- exp(logr0)
+  theta <- exp(logtheta)
+  K <- exp(logK)
+
+  x[1] ~ dunif(0, 10)
+
+  for(t in 1:(N-1)){
+    mu[t] <- r0 * pow(abs(x[t]), theta) / (1 + pow(abs(x[t]), theta) / K)
+    x[t+1] ~ dnorm(mu[t], iQ) 
+  }
+
+  for(t in 1:(N)){
+    y[t] ~ dnorm(x[t], iR)
+  }
+}")
+
+writeLines(bugs.model, "meyers.bugs")
+```
+
+```{r}
+
+## priors (untransformed variables)
+logK_prior     <- function(x) dunif(x, logK_prior_p[1], logK_prior_p[2])
+logr_prior     <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
+logtheta_prior <- function(x) dunif(x, logtheta_prior_p[1], logtheta_prior_p[2])
+stdQ_prior     <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
+stdR_prior     <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
+
+```
+
+### Parametric Bayes based on Meyers model
+
+Contains the
+
+```{r}
+
+init_p = c(1, 1, 1)
+names(init_p) = c("r0", "theta", "K")
 y <- obs$x[-1] 
@@
 
-The actual model is defined in a `model.file` that contains an R function that is automatically translated into BUGS code by *R2WinBUGS*.
-The file defines the priors and the model, as seen when read in here
-
-```{r}
-cat(readLines(con="ricker-UPrior.txt"), sep="\n")
-```
-
 We define which parameters to keep track of, and set the initial values of parameters in the transformed space used by the MCMC.
@@
 ```{r}
-# Uniform priors on standard deviation terms
-jags.params=c("K","logr0", "stdQ", "stdR")
+jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
 jags.inits <- function(){
-  list("K"=init_p["K"],"logr0"=log(init_p["r0"]), "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+  list("logr0"=log(init_p["r0"]), "logtheta" = log(init_p["theta"]),  
+       "logK"=log(init_p["K"]), "stdQ"=sqrt(0.05), 
+       "stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
 }
+```
 
+```{r}
 set.seed(12345)
-
 time_jags <- system.time(       
-  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
-                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="ricker-UPrior.txt")
+  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, 
+                  n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
+                  n.burnin=n.burnin, model.file="meyers.bugs")
 )         
@@
 
-# Traces
-ggplot(par_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
+## Traces
+ggplot(par_posteriors) + geom_line(aes(index, value)) +
+  facet_wrap(~ variable, scale="free", ncol=1)
 
-## priors (untransformed variables)
-K_prior <- function(x) dunif(x, 0.01, 40)
-logr_prior <- function(x) dunif(x, -6, 6)
-stdQ_prior <- function(x) dunif(x, 0.001, 100)
-stdR_prior <- function(x) dunif(x, 0.001, 100)
+```
 
-par_priors <- list(K = K_prior, deviance = function(x) 0 * x, logr0 = logr_prior, stdQ = stdQ_prior, stdR = stdR_prior)
+```{r}
+## prior curve functions, must be in order
+par_priors <- list( deviance = function(x) 0 * x, logK = logK_prior,
+                    logr0 = logr_prior, logtheta = logtheta_prior, 
+                    stdQ = stdQ_prior, stdR = stdR_prior)
 
+## calculate points along the prior curve
 par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
@@
 })
+```
 
-# posterior distributions
+```{r}
+## posterior distributions
 ggplot(par_posteriors, aes(value)) + 
@@
 ```{r}
-# um, cleaner if we were just be using the long form, par_posterior
-mcmc <- as.mcmc(jagsfit)
-mcmcall <- mcmc[,-2]
-who <- colnames(mcmcall)
-who 
-mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4])
-colnames(mcmcall) <- who
+meyers_pardist <- jags_matrix[2:6]
+meyers_pardist[,2] = exp(meyers_pardist[,1]) # transform model parameters back first
+meyers_pardist[,2] = exp(meyers_pardist[,2]) # transform model parameters back first
+meyers_pardist[,2] = exp(meyers_pardist[,3]) # transform model parameters back first
+colnames(meyers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
+
+bayes_coef <- apply(meyers_pardist,2,mean)
+meyers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
 ```
-
-```{r}
-ricker_pardist <- mcmcall
-ricker_pardist[,2] = exp(ricker_pardist[,2]) # transform model parameters back first
-
-bayes_coef <- apply(ricker_pardist,2,mean)
-ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
-ricker_bayes_pars
-```

### 2013-05-07 inst/examples/BUGS/short_observation_period.Rmd
short vs long observation time runs

--- 2013-05-07 inst/examples/BUGS/allen.Rmd
@@
 delta <- 0.01
-OptTime <- 50  # stationarity with unstable models is tricky thing
+OptTime <- 70  # stationarity with unstable models is tricky thing
 reward <- 0
@@
 x0 <- Xo # simulation under policy starts from
-Tobs <- 40
+Tobs <- 20
 ```

### 2013-05-07 inst/examples/BUGS/policy_thresholds.Rmd
script for the parallel analysis based on policy thresholds

--- 2013-05-07 inst/examples/BUGS/short_observation_period.Rmd
@@
 ```{r}
-f <- RickerAllee
-p <- c(2, 8, 5)
+f <- Ricker
+p <- c(2, 8)
 K <- p[2]
-allee <- p[3]
+C <- 5 # policy threshold
 ```
@@
 h_grid <- x_grid
-profit <- function(x,h) pmin(x, h)
+profit <- function(x,h) pmin(x, h) * (x > C) # no money for violating threshold
 delta <- 0.01
-OptTime <- 70  # stationarity with unstable models is tricky thing
+OptTime <- 50  # stationarity with unstable models is tricky thing
 reward <- 0
@@
 x0 <- Xo # simulation under policy starts from
-Tobs <- 20
+Tobs <- 40
 ```
@@
 
+```{r Figure3a}
+ggplot(dt[method %in% c("True", "Ricker", "MLE")]) + 
+  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
+```
+
 ```{r profits}

### 2013-05-08 inst/examples/BUGS/allen.Rmd
subset plotting of simulations (for slides potentially)
add a movie of data

@@
 
+```{r}
+dt
+
+```
+
+```{r Figure3a}
+dt$reps <- as.factor(dt$reps) 
+setkey(dt, method, reps)
+
+require(animation)
+saveMovie({
+  for(j in sample(dt$reps, 10)){
+    ricker <- dt[J("Ricker", j)]
+    for (i in seq(1, max(dt$time), by=5)) {
+      print(
+        ggplot(subset(ricker, time <= i), 
+               aes(time, fishstock, group=interaction(reps,method), color = method), alpha=.9) + 
+          geom_line() + xlim(0, max(dt$time)))
+    }
+  }
+}, interval = 0.1, movie.name = "wrong-model.gif", ani.width = 600, ani.height = 600)
+
+```
+
 ```{r profits}

### 2013-05-13 inst/examples/BUGS/policy_thresholds.Rmd
Adjusted penalty for policy thresholds case
* Need to make sure each algorithm permits the flexible choise of policy function
* Compare severe penalty case to zero-profit case

@@
 ```{r plotting-options}
+library(knitcitations)
 opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA,
                fig.width=6, fig.height=4)
-
-library(knitcitations)
 
@@
 f <- Ricker
-p <- c(2, 8)
+p <- c(.5, 10)
 K <- p[2]
@@
 h_grid <- x_grid
-profit <- function(x,h) pmin(x, h) * (x > C) # no money for violating threshold
+profit <- function(x,h) pmin(x, h) - 10*(x > C) # big penalty for violating threshold
 delta <- 0.01
@@
                           x[2:Tobs]))
+# plot(x)
 ```
@@
 ```{r mle, dependson="obs"}
-alt <- par_est(obs,  init = c(r = p[1], 
+est <- par_est(obs,  init = c(r = p[1], 
                               K = mean(obs$x[obs$x>0]), 
                               s = sigma_g))
-est <- par_est_allee(obs, f, p,  
-                     init = c(r = p[1] + 1, 
-                              K = p[2] + 2, 
-                              C = p[3] + 2, 
-                              s = sigma_g))
+
 ```
@@
 n.chains = 1
-n.iter = 400000
+n.iter = 40000
 n.burnin = floor(10000)
@@
 
-### Parametric Bayes based on myers model
-
-Contains the
-
-```{r}
-
-init_p = c(1, 1, 1)
-names(init_p) = c("r0", "theta", "K")
-y <- obs$x[-1] 
-N=length(y);
-```
-
-We'll be using the JAGS Gibbs sampler, a recent open source BUGS implementation with an R interface that works on most platforms.  We initialize the usual MCMC parameters; see `?jags` for details.
-
-```{r}
-jags.data <- list("N","y")
-n.chains = 1
-n.iter = 40000
-n.burnin = floor(10000)
-n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
-```
-
-```{r}
-jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
-jags.inits <- function(){
-  list("logr0"=log(init_p["r0"]), "logtheta" = log(init_p["theta"]),  
-       "logK"=log(init_p["K"]), "stdQ"=sqrt(0.05), 
-       "stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
-}
-```
-
-```{r}
-set.seed(12345)
-time_jags <- system.time(       
-  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, 
-                  n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
-                  n.burnin=n.burnin, model.file="myers.bugs")
-)         
-time_jags <- unname(time_jags["elapsed"]);
-```
-
-Convergence diagnostics for parametric bayes
-
-```{r}
-jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
-
-## Traces
-ggplot(par_posteriors) + geom_line(aes(index, value)) +
-  facet_wrap(~ variable, scale="free", ncol=1)
-
-```
-
-```{r}
-## prior curve functions, must be in order
-par_priors <- list( deviance = function(x) 0 * x, logK = logK_prior,
-                    logr0 = logr_prior, logtheta = logtheta_prior, 
-                    stdQ = stdQ_prior, stdR = stdR_prior)
-
-## calculate points along the prior curve
-par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100) 
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-```
-
-```{r}
-## posterior distributions
-ggplot(par_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=2)
-```
-
-```{r}
-myers_pardist <- as.matrix(jags_matrix[2:6])
-myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
-myers_pardist[,2] = exp(myers_pardist[,2]) # transform model parameters back first
-myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
-colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
-
-apply(myers_pardist,2,mean)
-bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
-
-myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
-```
-
 ### Phase-space diagram of the expected dynamics
@@
 true_means <- sapply(x_grid, f, 0, p)
-alt_means <- sapply(x_grid, alt$f, 0, ricker_bayes_pars[c(1,2)])
-est_means <- sapply(x_grid, est$f, 0, est$p)
-par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)
+mle_means <- sapply(x_grid, est$f, 0, est$p)
+ricker_means <- sapply(x_grid, est$f, 0, ricker_bayes_pars[c(1,2)])
+allen_means <- sapply(x_grid, f, 0, bayes_pars)
 
-myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
+# myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
 
 models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
-                     MLE=est_means, Ricker=alt_means, 
-                     Parametric.Bayes = par_bayes_means,
-                     Myers = myers_means)
+                     MLE=mle_means, Ricker = ricker_means,
+                     Allen = allen_means)
 
@@
 ```{r}
-ricker_f <- function(x, h, p) alt$f(x, h, unname(p[c(2, 1)]))
+ricker_f <- function(x, h, p) est$f(x, h, unname(p[c(2, 1)]))
 matrices_alt <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
 opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-```
-
-Bayesian Myers model
-
-```{r}
-myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
-matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
-myers_alt <- value_iteration(matrices_myers, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
 ```
@@
 ```{r}
-OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_alt$D, Parametric.Bayes = opt_par_bayes$D, Myers = myers_alt$D)
+OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_alt$D, Allen = opt_par_bayes$D)
 colorkey=cbPalette

### 2013-05-13 R/gp_mcmc.R
Return the entire posterior distribution of covariance matrices

@@ -47,9 +47,10 @@ gp_mcmc <- function(x, y, init_pars = c(l=1, sigma.n=1), n = 1e4, d.p = c(5,5), 
 #' predict the expected values and posterior distributions of the Gaussian Process
 #' 
 #' @param x_predict the values at which we desire predictions
+#' @param covs if TRUE, return covariances instead of variances (set high thinning as this is memory intensive)
 #' @import reshape2
 #' @export
-gp_predict <- function(gp, x_predict, burnin=0, thin=1){
+gp_predict <- function(gp, x_predict, burnin=0, thin=1, covs=FALSE){
 
   postdist <- cbind(index=1:gp$nbatch, as.data.frame(exp(gp$batch)))
   s <- seq(burnin+1, gp$nbatch, by=thin)
@@ -66,18 +67,25 @@ gp_predict <- function(gp, x_predict, burnin=0, thin=1){
     cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma.n^2 * diag(1, length(obs$x)))
     Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% obs$y
     Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)
-  
-    list(Ef = Ef, Vf = diag(Cf))
+    list(Ef = Ef, Cf = Cf, Vf = diag(Cf))
   })
   
   Ef_posterior <- sapply(out, `[[`, "Ef")
+  Cf_posterior <- sapply(out, `[[`, "Cf")
   Vf_posterior <- sapply(out, `[[`, "Vf")
   
-  E_Ef <- rowMeans(Ef_posterior)
-  E_Vf <- rowMeans(Vf_posterior)
   
-  list(Ef_posterior = Ef_posterior, Vf_posterior = Vf_posterior,
-       E_Ef = E_Ef, E_Vf = E_Vf)
+  E_Ef <- rowMeans(Ef_posterior)
+  E_Cf <- matrix( apply(Cf_posterior, 1, sum) / dim(Cf_posterior)[2], ncol = sqrt(dim(Cf_posterior)[1]) )
+  E_Vf <- diag(E_Cf) # same as rowMeans(Vf_posterior)
+  
+  # list format better for return
+  Cf_posterior <- lapply(out, `[[`, "Cf")
+  
[…]

### 2013-05-13 inst/examples/BUGS/allen.Rmd
animation plots, run with corrected ricker model

--- 2013-05-07 inst/examples/BUGS/allen.Rmd
@@
 opts_knit$set(upload.fun = socialR::flickr.url)
-theme_set(theme_bw(base_size=10))
+theme_set(theme_bw(base_size=12))
 theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
@@
 
-Uses the model derived in `citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
+Uses the model derived in `r citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
 
@@
                           x[2:Tobs]))
+raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
+raw_plot
 ```
@@
 jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
+true_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
 ```
@@
 # Traces
-ggplot(par_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
+ggplot(true_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
 ```
@@
 
-par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
+true_priors <- ddply(true_posteriors, "variable", function(dd){
     grid <- seq(min(dd$value), max(dd$value), length = 100) 
@@
 # posterior distributions
-ggplot(par_posteriors, aes(value)) + 
+ggplot(true_posteriors, aes(value)) + 
   stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=2)
+  geom_line(data=true_priors, aes(x=value, y=density), col="red") + 
+  facet_wrap(~ variable, scale="free", ncol=3)
 
@@
 
-## Parametric Bayes based on the structurally wrong model
+## Parametric Bayes based on the structurally wrong model (Ricker)
 
@@
 par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
+ricker_posteriors <- par_posteriors
 
 # Traces
-ggplot(par_posteriors) + geom_line(aes(index, value)) + 
+ggplot(ricker_posteriors) + geom_line(aes(index, value)) + 
   facet_wrap(~ variable, scale="free", ncol=1)
@@
 
-par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
+ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
     grid <- seq(min(dd$value), max(dd$value), length = 100) 
@@
 # posterior distributions
-ggplot(par_posteriors, aes(value)) + 
+ggplot(ricker_posteriors, aes(value)) + 
   stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
+  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") + 
   facet_wrap(~ variable, scale="free", ncol=2)
@@
 
-## Write the external bugs file
+## Myers Parametric Bayes
 
@@
 ```
-
-### Parametric Bayes based on myers model
-
-Contains the
 
@@
   geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=2)
+  facet_wrap(~ variable, scale="free", ncol=3)
 ```
@@
 
+```{r}
+model = "MLE"
+ggplot(subset(dt, method == model)) + 
+  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) +
+  ylim(0, 9)
+ggsave(paste(model, "_sim.pdf", sep=""), width=6, height=4)
+
+```
+
+```{r Figure3a}
+dt$reps <- as.integer(dt$reps)
+
+require(animation)
+ani.options(loop=TRUE)
+saveMovie({    
+    for (i in seq(1, OptTime, by=2)) {
+      print(
+        ggplot(subset(dt, method %in% c("True", "Ricker") & reps < 5 & time <= i), 
+               aes(time, fishstock, group=interaction(reps,method), color = method), alpha=.9) + 
+          geom_line() + xlim(0, OptTime)
+        )
+    }
+}, interval = 0.1, movie.name = "wrong-model.gif", ani.width = 600, ani.height = 600)
+
+require(MASS)
+s <- length(gp_dat$Cf_posterior)
+
+# Samples from the posterior of GPs
+draws <- sapply(1:s, function(i)
+  mvrnorm(1, gp_dat$Ef_posterior[,i], gp_dat$Cf_posterior[[i]]))
+draws <- melt(data.frame(x=x_grid, draws), id="x")
+
+## Samples from the expected GP
+gp_draws <- mvrnorm(6, gp_dat$E_Ef, gp_dat$E_Cf)
+gp_draws <- melt(data.frame(x=x_grid, t(gp_draws)), id="x")
+
+ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
+  geom_line(aes(x,y,ymin=ymin,ymax=ymax)) +
+  # geom_line(data=gp_draws, aes(x=x, y=value, col=variable))  +  
+  geom_line(data=draws, aes(x=x, y=value, index=variable), alpha=.1)
+
+```
+
 ```{r profits}

### 2013-05-13 inst/examples/BUGS/step_ahead_plots.Rmd
plots for step-ahead predictions of each model (from allen.Rmd)

One-step ahead predictor plots

```{r}

step_ahead <- function(x, f, p){
  h = 0
  x_predict <- sapply(x, f, h, p)
  n <- length(x_predict) - 1
  y <- c(x[1], x_predict[1:n])
  y
}

df <- melt(data.frame(time = 1:length(x), stock = x, 
                True = step_ahead(x,f,p)
                ,  
                MLE = step_ahead(x,f,est$p)
                , 
                Parametric.Bayes = step_ahead(x, f, bayes_pars)
                , 
                Ricker = step_ahead(x,alt$f, ricker_bayes_pars)
                , 
                Myers = step_ahead(x, Myer_harvest, myers_bayes_pars)
                 ), id=c("time", "stock"))

ggplot(df) + geom_point(aes(time, stock)) + 
  geom_line(aes(time, value, col=variable)) +
    scale_colour_manual(values=colorkey)

```

### 2013-05-14 inst/examples/gaussian-process-basics.Rmd
replicate run

@@
 ```{r plotting-options}
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA)
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA, dev="CairoPDF", fig.width=6, fig.height=4)
 library(ggplot2) # plotting 
@@
              plot.background = element_rect(fill = "transparent", colour = NA))
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 ```
@@
 
-```{r}
+```{r fig2a}
 fig2a <- ggplot(dat,aes(x=x,y=value)) +
@@
 
-```{r}
+```{r fig2b}
 dat <- data.frame(x=x_predict, t(values))
@@
 fig2b
+```
+
+```{r fig2_data}
+ggplot(data=obs,aes(x=x,y=y)) +
+  geom_point()   #OBSERVED DATA
 ```
@@
 
-```{r}
+```{r fig2c}
 dat <- data.frame(x=x_predict, t(values))
@@
 
-```{r}
+```{r gp_mcmc}
 postdist <- cbind(index=1:n, as.data.frame(exp(out$batch)))
@@
 
-```{r}
+```{r gp_posteriors}
 

### 2013-05-14 inst/examples/BUGS/allen.Rmd
Updated runs of complete script
* Includes alternative Figure 1 (in time-series space, with GP #44)
* Includes (fixed) Bayesian Ricker (#43) and Bayesian Myers (#42) case

@@
 
+##
+
+```{r step_ahead}
+gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
+
+step_ahead <- function(x, f, p){
+  h = 0
+  x_predict <- sapply(x, f, h, p)
+  n <- length(x_predict) - 1
+  y <- c(x[1], x_predict[1:n])
+  y
+}
+
+df <- melt(data.frame(time = 1:length(x), stock = x, 
+                GP = gp_f_at_obs$E_Ef,
+                True = step_ahead(x,f,p),  
+                MLE = step_ahead(x,f,est$p), 
+                Parametric.Bayes = step_ahead(x, f, bayes_pars), 
+                Ricker = step_ahead(x,alt$f, ricker_bayes_pars), 
+                Myers = step_ahead(x, Myer_harvest, myers_bayes_pars)
+                 ), id=c("time", "stock"))
+
+ggplot(df) + geom_point(aes(time, stock)) + 
+  geom_line(aes(time, value, col=variable)) +
+    scale_colour_manual(values=colorkey) 
+```
+
 ## Optimal policies by value iteration
@@
 ```{r}
-OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_alt$D, Parametric.Bayes = opt_par_bayes$D, Myers = myers_alt$D)
-colorkey=cbPalette
-names(colorkey) = names(OPT) 
+OPT = data.frame(opt_gp$D, opt_true$D, opt_estimated$D, opt_alt$D, opt_par_bayes$D, myers_alt$D)
+
+names(OPT) = names(colorkey)[1:6]
 ```
@@
 
-```{r}
-model = "MLE"
-ggplot(subset(dt, method == model)) + 
-  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) +
-  ylim(0, 9)
-ggsave(paste(model, "_sim.pdf", sep=""), width=6, height=4)
-
-```
-
-```{r Figure3a}
-dt$reps <- as.integer(dt$reps)
-
-require(animation)
-ani.options(loop=TRUE)
-saveMovie({    
-    for (i in seq(1, OptTime, by=2)) {
-      print(
-        ggplot(subset(dt, method %in% c("True", "Ricker") & reps < 5 & time <= i), 
-               aes(time, fishstock, group=interaction(reps,method), color = method), alpha=.9) + 
-          geom_line() + xlim(0, OptTime)
-        )
-    }
-}, interval = 0.1, movie.name = "wrong-model.gif", ani.width = 600, ani.height = 600)
-
-require(MASS)
-s <- length(gp_dat$Cf_posterior)
-
-# Samples from the posterior of GPs
-draws <- sapply(1:s, function(i)
-  mvrnorm(1, gp_dat$Ef_posterior[,i], gp_dat$Cf_posterior[[i]]))
-draws <- melt(data.frame(x=x_grid, draws), id="x")
-
-## Samples from the expected GP
-gp_draws <- mvrnorm(6, gp_dat$E_Ef, gp_dat$E_Cf)
-gp_draws <- melt(data.frame(x=x_grid, t(gp_draws)), id="x")
-
-ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-  geom_line(aes(x,y,ymin=ymin,ymax=ymax)) +
-  # geom_line(data=gp_draws, aes(x=x, y=value, col=variable))  +  
-  geom_line(data=draws, aes(x=x, y=value, index=variable), alpha=.1)
-
-```
-
 ```{r profits}
@@
 ggplot(Profit, aes(V1)) + geom_histogram() + 
-  facet_wrap(~method, scales = "free_y") + guides(legend.position = "none")
+  facet_wrap(~method, scales = "free_y") + guides(legend.position = "none") + xlab("Total profit by replicate")
 ```

### 2013-05-15 R/gp_trasition_matrix.R
Fixed GP use of standard deviation in place of variance

@@ -9,10 +9,8 @@
 #' @export
 gp_transition_matrix <- function(Ef, V, x_grid, h_grid=NULL){  
   if(is.null(h_grid))
-    h_grid <- x_grid
-  
+    h_grid <- x_grid  
   if(is.matrix(Ef)){ # loop over posteriors
-    
     n_samples <- dim(Ef)[2]
     loop_over_h <-function(h){
       # Average across matrices in list, and then normalize
@@ -23,11 +21,8 @@ gp_transition_matrix <- function(Ef, V, x_grid, h_grid=NULL){
       F_s <- rownorm(F_s)
     }
     out <- lapply(h_grid, loop_over_h)  
-
   } else { 
-    
     out <- lapply(h_grid, gp_calc_F, Ef, V, x_grid)
-  
   }
   out
 }
@@ -35,8 +30,6 @@ gp_transition_matrix <- function(Ef, V, x_grid, h_grid=NULL){
 
 
 gp_calc_F <- function(h, Ef, V, x_grid){
-  
-  
   j <- which.min(abs(x_grid - h))
   indices <- 1:length(Ef)
   shift <- pmax(indices-j,1)
@@ -48,11 +41,15 @@ gp_calc_F <- function(h, Ef, V, x_grid){
       out[1] <- 1
       out
     } else {
-      out <- dnorm(x_grid, mu[i], V[i])
+      out <- dnorm(x_grid, mu[i], sqrt(V[i]))
     }
   }))
   F_ <- rownorm(F_)
 }
+
+
+
+
 
 #' calculate transition matrix at each harvest level based on parameterized function
 #' 

### 2013-05-15 inst/examples/BUGS/step_ahead_plots.Rmd
tweak figure layout

@@
 
+```{r plotting-opts}
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA,
+               fig.width=7, fig.height=4)
+
+library(knitcitations)
+
+library(ggplot2) # plotting 
+opts_knit$set(upload.fun = socialR::flickr.url)
+theme_set(theme_bw(base_size=12))
+theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
+             plot.background = element_rect(fill = "transparent", colour = NA))
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+```
+
 ```{r}
+
+gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
 
@@
 df <- melt(data.frame(time = 1:length(x), stock = x, 
-                True = step_ahead(x,f,p)
-                ,  
-                MLE = step_ahead(x,f,est$p)
-                , 
-                Parametric.Bayes = step_ahead(x, f, bayes_pars)
-                , 
-                Ricker = step_ahead(x,alt$f, ricker_bayes_pars)
-                , 
+                GP = gp_f_at_obs$E_Ef,
+                True = step_ahead(x,f,p),  
+                MLE = step_ahead(x,f,est$p), 
+                Parametric.Bayes = step_ahead(x, f, bayes_pars), 
+                Ricker = step_ahead(x,alt$f, ricker_bayes_pars), 
                 Myers = step_ahead(x, Myer_harvest, myers_bayes_pars)
@@
     scale_colour_manual(values=colorkey) 
+```
 
+```{r}
+gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
+
+ricker_f <- function(x,h,p){
+  sapply(x, function(x){ 
+    x <- pmax(0, x-h) 
+    pmax(0, x * exp(p[2] * (1 - x / p[1] )) )
+  })
+}
+
+require(MASS)
+df_post <- melt(lapply(sample(100), 
+  function(i){
+    data.frame(time = 1:length(x), stock = x, 
+                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
+                True = step_ahead(x,f,p),  
+                MLE = step_ahead(x,f,est$p), 
+                Parametric.Bayes = step_ahead(x, allen_f, pardist[i,]), 
+                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
+                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
+  }), id=c("time", "stock"))
+
+df <- df_post
+#df <- subset(df_post, variable %in% c("GP"))
+#df <- subset(df_post, variable %in% c("Parametric.Bayes"))
+#df <- subset(df_post, variable %in% c("Ricker", "Parametric.Bayes", "Myers"))
+
+fig1 <- ggplot(df) + geom_point(aes(time, stock)) + 
+  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
+#  geom_point(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1, pch="+") + 
+  facet_wrap(~variable) + 
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) 
+fig1
 ```
+
+```{r}
+Tobs <- 8
+x <- numeric(Tobs)
+x[1] = 5.5
+for(t in 1:(Tobs-1))
+    x[t+1] = z_g() * f(x[t], h=0, p=p)
+y <- x
+x <- numeric(Tobs)
+x[1] = 4.5
+for(t in 1:(Tobs-1))
+    x[t+1] = z_g() * f(x[t], h=0, p=p)
+x <- c(y,x)
+```
+
+```{r}
+gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
+
+ricker_f <- function(x,h,p){
+  sapply(x, function(x){ 
+    x <- pmax(0, x-h) 
+    pmax(0, x * exp(p[2] * (1 - x / p[1] )) )
+  })
+}
+
+require(MASS)
+df_post <- melt(lapply(sample(100), 
+  function(i){
+    data.frame(time = 1:length(x), stock = x, 
+                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
+                True = step_ahead(x,f,p),  
+                MLE = step_ahead(x,f,est$p), 
+                Parametric.Bayes = step_ahead(x, allen_f, pardist[i,]), 
+                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
+                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
+  }), id=c("time", "stock"))
+
+df <- df_post
+#df <- subset(df_post, variable %in% c("GP"))
+#df <- subset(df_post, variable %in% c("Parametric.Bayes"))
+#df <- subset(df_post, variable %in% c("Ricker", "Parametric.Bayes", "Myers"))
+
+fig1 <- ggplot(df) + geom_point(aes(time, stock)) + 
+  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
+#  geom_point(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1, pch="+") + 
+  facet_wrap(~variable) + 
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) 
+fig1
+```
+
+```{r}
+i <- 15
+Tobs <- 5
+h_i <- 1
+
+df <- data.frame(
+  x = x_grid,  
+  GP = matrices_gp[[h_i]][i,], 
+  True = matrices_true[[h_i]][i,], 
+  MLE = matrices_estimated[[h_i]][i,], 
+  Parametric.Bayes = matrices_par_bayes[[h_i]][i,], 
+  Ricker = matrices_alt[[h_i]][i,],
+  Myers = matrices_myers[[h_i]][i,])
+
+df2 <- data.frame(
+  x = x_grid,
+  GP = (matrices_gp[[h_i]] %^% Tobs)[i,], 
+  True = (matrices_true[[h_i]] %^% Tobs)[i,], 
+  MLE = (matrices_estimated[[h_i]] %^% Tobs)[i,],
+  Parametric.Bayes = (matrices_par_bayes[[h_i]] %^% Tobs)[i,], 
+  Ricker = (matrices_alt[[h_i]] %^% Tobs)[i,], 
+  Myers = (matrices_myers[[h_i]] %^% Tobs)[i,])
+
+T2 <- 2 * Tobs
+
+df4 <- data.frame(
+  x = x_grid,
+  GP = (matrices_gp[[h_i]] %^% T2)[i,], 
+  True = (matrices_true[[h_i]] %^% T2)[i,], 
+  MLE = (matrices_estimated[[h_i]] %^% T2)[i,],
+  Parametric.Bayes = (matrices_par_bayes[[h_i]] %^% T2)[i,], 
+  Ricker = (matrices_alt[[h_i]] %^% T2)[i,], 
+  Myers = (matrices_myers[[h_i]] %^% T2)[i,])
+
+forecasts <- melt(list(T_start = df, T_mid = df2, T_end = df4), id="x")
+
+ggplot(forecasts) + geom_line(aes(x, value, group=interaction(variable, L1), col=variable, lty=L1)) +
+  facet_wrap(~ variable, scale="free_y", ncol=2) + 
+  scale_colour_manual(values=colorkey) 
+```

### 2013-05-15 inst/examples/BUGS/step_ahead_plots.Rmd
verbose version with plots and explanation (for notes)

--- 2013-05-13 inst/examples/BUGS/step_ahead_plots.Rmd
@@
-One-step ahead predictor plots
+## One-step ahead predictor plots and forecasting plots
+
+Set our usual plotting options for the notebook
+
+```{r plotting-opts}
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA,
+               fig.width=7, fig.height=4)
+library(ggplot2) 
+opts_knit$set(upload.fun = socialR::flickr.url)
+theme_set(theme_bw(base_size=12))
+theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
+             plot.background = element_rect(fill = "transparent", colour = NA))
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+```
+
+Now we show how each of the fitted models performs on the training data (e.g. plot of the step-ahead predictors). For the GP, we need to predict on the training data first:
 
 ```{r}
+gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
+```
 
+For the parametric models, prediction is a matter of feeding in each of the observed data points to the fitted parameters, like so
+
+```{r}
 step_ahead <- function(x, f, p){
@@
 }
+```
 
+Which we apply over each of the fitted models, including the GP, organizing the "expected" transition points (given the previous point) into a data frame.
+
+```{r}
 df <- melt(data.frame(time = 1:length(x), stock = x, 
-                True = step_ahead(x,f,p)
-                ,  
-                MLE = step_ahead(x,f,est$p)
-                , 
-                Parametric.Bayes = step_ahead(x, f, bayes_pars)
-                , 
-                Ricker = step_ahead(x,alt$f, ricker_bayes_pars)
-                , 
+                GP = gp_f_at_obs$E_Ef,
+                True = step_ahead(x,f,p),  
+                MLE = step_ahead(x,f,est$p), 
+                Parametric.Bayes = step_ahead(x, f, bayes_pars), 
+                Ricker = step_ahead(x,alt$f, ricker_bayes_pars), 
                 Myers = step_ahead(x, Myer_harvest, myers_bayes_pars)
@@
     scale_colour_manual(values=colorkey) 
+```
+
+## Posterior predictive curves
+
+This shows only the mean predictions.
+For the Bayesian cases, we can instead loop over the posteriors of the parameters (or samples from the GP posterior) to get the distribution of such curves in each case.
+
+We will need a vector version (`pmin` in place of `min`) of this function that can operate on the posteriors, others are vectorized already.
+
+```{r}
+ricker_f <- function(x,h,p){
+  sapply(x, function(x){ 
+    x <- pmax(0, x-h) 
+    pmax(0, x * exp(p[2] * (1 - x / p[1] )) )
+  })
+}
+```
+
+Then we proceed as before, now looping over 100 random samples from the posterior for each Bayesian estimate.  We write this as a function for easy reuse.
+
+```{r}
+require(MASS)
+
+step_ahead_posteriors <- function(x){
+gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
+df_post <- melt(lapply(sample(100), 
+  function(i){
+    data.frame(time = 1:length(x), stock = x, 
+                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
+                True = step_ahead(x,f,p),  
+                MLE = step_ahead(x,f,est$p), 
+                Parametric.Bayes = step_ahead(x, allen_f, pardist[i,]), 
+                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
+                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
+  }), id=c("time", "stock"))
+
+}
+
+df_post <- step_ahead_posteriors(x)
+ggplot(df_post) + geom_point(aes(time, stock)) + 
+  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
+  facet_wrap(~variable) + 
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
 
 ```
+
+## Performance on data outside of observations
+
+Of course it is hardly suprising that all models do reasonably well on the data on which they were trained.
+A crux of the problem is the model performance on data outside the observed range.
+(Though we might also wish to repeat the above plot on data in the observed range but a different sequence from the observed data).
+
+First we generate some data from the underlying model coming from below the tipping point:
+
+```{r}
+y <- numeric(Tobs)
+y[1] = 4.5
+for(t in 1:(Tobs-1))
+    y[t+1] = z_g() * f(y[t], h=0, p=p)
+```
+
+Proceed as before on this data:
+
+```{r}
+df_post <- step_ahead_predictors(y)
+
+ggplot(df_post) + geom_point(aes(time, stock)) + 
+  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
+  facet_wrap(~variable) + 
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
+
+```
+
+alternately, try the plot without facets
+
+```{r}
+ggplot(df_post) + geom_point(aes(time, stock)) + 
+  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) 
+```
+
+## Forecast Distributions
+
+Another way to visualize this is to look directly at the distribution predicted under each model, one step and several steps into the future (say, at a fixed harvest level).
+
+```{r}
+get_forecasts <- function(i, Tobs, h_i){
+
+df <- data.frame(
+  x = x_grid,  
+  GP = matrices_gp[[h_i]][i,], 
+  True = matrices_true[[h_i]][i,], 
+  MLE = matrices_estimated[[h_i]][i,], 
+  Parametric.Bayes = matrices_par_bayes[[h_i]][i,], 
+  Ricker = matrices_alt[[h_i]][i,],
+  Myers = matrices_myers[[h_i]][i,])
+
+df2 <- data.frame(
+  x = x_grid,
+  GP = (matrices_gp[[h_i]] %^% Tobs)[i,], 
+  True = (matrices_true[[h_i]] %^% Tobs)[i,], 
+  MLE = (matrices_estimated[[h_i]] %^% Tobs)[i,],
+  Parametric.Bayes = (matrices_par_bayes[[h_i]] %^% Tobs)[i,], 
+  Ricker = (matrices_alt[[h_i]] %^% Tobs)[i,], 
+  Myers = (matrices_myers[[h_i]] %^% Tobs)[i,])
+
+T2 <- 2 * Tobs
+
+df4 <- data.frame(
+  x = x_grid,
+  GP = (matrices_gp[[h_i]] %^% T2)[i,], 
+  True = (matrices_true[[h_i]] %^% T2)[i,], 
+  MLE = (matrices_estimated[[h_i]] %^% T2)[i,],
+  Parametric.Bayes = (matrices_par_bayes[[h_i]] %^% T2)[i,], 
+  Ricker = (matrices_alt[[h_i]] %^% T2)[i,], 
+  Myers = (matrices_myers[[h_i]] %^% T2)[i,])
+
+forecasts <- melt(list(T_start = df, T_mid = df2, T_end = df4), id="x")
+
+}
+
+i = 15
+```
+
+This takes `i`, an index to `x_grid` value (e.g. for `i`=`r i` corresponds to a starting postion `x` = `r x_grid[i]`)
+
+```{r}
+forecasts <- get_forecasts(i = 15, Tobs = 5, h_i = 1)
+
+ggplot(forecasts) +
+  geom_line(aes(x, value, group=interaction(variable, L1), col=variable, lty=L1)) +
+  facet_wrap(~ variable, scale="free_y", ncol=2) + 
+  scale_colour_manual(values=colorkey) 
+```
+
+We can compare to a better starting stock,
+
+```{r} 
+i<-30
+```
+
+where `i`=`r i` corresponds to a starting postion `x` = `r x_grid[i]`
+
+```{r}
+forecasts <- get_forecasts(i = i, Tobs = 5, h_i = 1)
+
+ggplot(forecasts) +
+  geom_line(aes(x, value, group=interaction(variable, L1), col=variable, lty=L1)) +
+  facet_wrap(~ variable, scale="free_y", ncol=2) + 
+  scale_colour_manual(values=colorkey) 
+```

### 2013-05-16 inst/examples/BUGS/allen.Rmd
replicate run. cache results

@@
 ```{r plotting-options}
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA,
-               fig.width=6, fig.height=4)
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA,
+               fig.width=6, fig.height=4, cache.path="allencache/")
 
@@
 
-##
-
-```{r step_ahead}
-gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
-
-step_ahead <- function(x, f, p){
-  h = 0
-  x_predict <- sapply(x, f, h, p)
-  n <- length(x_predict) - 1
-  y <- c(x[1], x_predict[1:n])
-  y
-}
-
-df <- melt(data.frame(time = 1:length(x), stock = x, 
-                GP = gp_f_at_obs$E_Ef,
-                True = step_ahead(x,f,p),  
-                MLE = step_ahead(x,f,est$p), 
-                Parametric.Bayes = step_ahead(x, f, bayes_pars), 
-                Ricker = step_ahead(x,alt$f, ricker_bayes_pars), 
-                Myers = step_ahead(x, Myer_harvest, myers_bayes_pars)
-                 ), id=c("time", "stock"))
-
-ggplot(df) + geom_point(aes(time, stock)) + 
-  geom_line(aes(time, value, col=variable)) +
-    scale_colour_manual(values=colorkey) 
-```
-
 ## Optimal policies by value iteration
@@
 ```{r}
-OPT = data.frame(opt_gp$D, opt_true$D, opt_estimated$D, opt_alt$D, opt_par_bayes$D, myers_alt$D)
-
-names(OPT) = names(colorkey)[1:6]
+OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_alt$D, Parametric.Bayes = opt_par_bayes$D, Myers = myers_alt$D)
+colorkey=cbPalette
+names(colorkey) = names(OPT) 
 ```

### 2013-05-16 inst/examples/BUGS/step_ahead_plots.Rmd
Potential plots for visualizing uncertainty #44
Run builds on results from allen.md

@@
 
+alternately, try the plot without facets
+
+```{r}
+ggplot(df_post) + geom_point(aes(time, stock)) + 
+  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) 
+```
+
 ## Performance on data outside of observations
@@
 ```{r}
+Tobs <- 8
 y <- numeric(Tobs)
@@
 ```{r}
-df_post <- step_ahead_predictors(y)
+crash_data <- step_ahead_posteriors(y)
 
-ggplot(df_post) + geom_point(aes(time, stock)) + 
+ggplot(crash_data) + geom_point(aes(time, stock)) + 
   geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
@@
 
-```
-
-alternately, try the plot without facets
-
-```{r}
-ggplot(df_post) + geom_point(aes(time, stock)) + 
-  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) 
 ```
@@
 ```{r}
+library(expm)
 get_forecasts <- function(i, Tobs, h_i){

### 2013-05-22 inst/examples/BUGS/allen.Rmd
replicate runs on data from allen and myers model

@@
 # Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach
-
-```{r runsetup}
-setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
-```
 
@@
 
-```{r plotting-options}
+```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
+library(knitcitations)
+library(ggplot2) 
+library(nonparametricbayes) 
 opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA,
-               fig.width=6, fig.height=4, cache.path="allencache/")
-
-library(knitcitations)
-
-library(ggplot2) # plotting 
+               fig.width=6, fig.height=4, cache.path="allen/")
 opts_knit$set(upload.fun = socialR::flickr.url)
@@
 ```
-Load necessary libraries,
 
-```{r libraries, cache=FALSE, message=FALSE, warning=FALSE}
-library(nonparametricbayes) # loads the rest as dependencies
+```{r}
+posterior.mode <- function(x) {
+  ux <- unique(x)
+  ux[which.max(tabulate(match(x, ux)))]
+}
+
 ```
@@
 
-```{r}
+```{r stateeq}
 f <- RickerAllee
@@
 ```{r mle, dependson="obs"}
-alt <- par_est(obs,  init = c(r = p[1], 
-                              K = mean(obs$x[obs$x>0]), 
-                              s = sigma_g))
 est <- par_est_allee(obs, f, p,  
@@
                               s = sigma_g))
+```
+
+Mean predictions
+
+```{r mle-output}
+true_means <- sapply(x_grid, f, 0, p)
+est_means <- sapply(x_grid, est$f, 0, est$p)
 ```
@@
 
-```{r}
+```{r gp, dependson=c("gp-priors", "obs")}
 gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
@@
 
-```{r gp_traces_densities}
+```{r gp_traces_densities, dependson="gp"}
 plots <- summary_gp_mcmc(gp)
@@
 
-```{r}
+```{r gp-output, dependson="gp"}
 # Summarize the GP model
@@
 
-### Parametric Bayes
+Parametric Bayesian Models
+---------------------------
 
-We initiate the MCMC chain (`init_p`) using the true values of the parameters `p` from the simulation.
-While impossible in real data, this gives the parametric Bayesian approach the best chance at succeeding.
+We use the JAGS Gibbs sampler, a recent open source BUGS
+implementation with an R interface that works on most platforms.
+We initialize the usual MCMC parameters; see `?jags` for details.
+
+All parametric Bayesian estimates use the following basic parameters for the JAGS MCMC:
+
+```{r jags-setup}
+y <- obs$x[-1] 
+N <- length(y);
+jags.data <- list("N","y")
+n.chains <- 1
+n.iter <- 40000
+n.burnin <- floor(10000)
+n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
+```
+
+We will use the same priors for process and observation noise in each model,
+
+```{r common-priors}
+stdQ_prior_p <- c(0.0001, 100)
+stdR_prior_p <- c(0.0001, 100)
+stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
+stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
+```
+
+### Parametric Bayes of correct (Allen) model
+
+We initiate the MCMC chain (`init_p`) using the true values of the
+parameters `p` from the simulation.  While impossible in real data, this
+gives the parametric Bayesian approach the best chance at succeeding.
 `y` is the timeseries (recall `obs` has the $x_t$, $x_{t+1}$ pairs)
 
-```{r}
+```{r allen_pars}
 # a bit unfair to start with the correct values, but anyhow...
@@
 names(init_p) = c("r0", "K", "theta")
-y <- obs$x[-1] 
-N=length(y);
 ```
 
-We'll be using the JAGS Gibbs sampler, a recent open source BUGS implementation with an R interface that works on most platforms.  We initialize the usual MCMC parameters; see `?jags` for details.
+The actual model is defined in a `model.file` that contains an R function
+that is automatically translated into BUGS code by *R2WinBUGS*.  The file
+defines the priors and the model. We write the file from R as follows:
 
-```{r}
-jags.data <- list("N","y")
-n.chains = 1
-n.iter = 40000
-n.burnin = floor(10000)
-n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
+```{r allen-jags}
+K_prior_p <- c(0.01, 40.0)
+logr0_prior_p <- c(-6.0, 6.0)
+logtheta_prior_p <- c(-6.0, 6.0)
+
+bugs.model <- 
+paste(sprintf(
+"model{
+  K     ~ dunif(%s, %s)
+  logr0    ~ dunif(%s, %s)
+  logtheta ~ dunif(%s, %s)
+  stdQ ~ dunif(%s, %s)
+  stdR ~ dunif(%s, %s)", 
+  K_prior_p[1], K_prior_p[2],
+  logr0_prior_p[1], logr0_prior_p[2],
+  logtheta_prior_p[1], logtheta_prior_p[2],
+  stdQ_prior_p[1], stdQ_prior_p[2],
+  stdR_prior_p[1], stdR_prior_p[2]),
+
+  "
+  iQ <- 1 / (stdQ * stdQ);
+  iR <- 1 / (stdR * stdR);
+  r0 <- exp(logr0)
+  theta <- exp(logtheta)
+  x[1] ~ dunif(0, 10)
+  for(t in 1:(N-1)){
+    mu[t] <- x[t] * exp(r0 * (1 - x[t]/K)* (x[t] - theta) / K )
+    x[t+1] ~ dnorm(mu[t], iQ) 
+  }
+  for(t in 1:(N)){
+    y[t] ~ dnorm(x[t], iR)
+  }
+}")
+writeLines(bugs.model, "allen.bugs")
 ```
 
-The actual model is defined in a `model.file` that contains an R function that is automatically translated into BUGS code by *R2WinBUGS*.
-The file defines the priors and the model, as seen when read in here
+Write the priors into a list for later reference
 
-```{r}
-cat(readLines(con="bugmodel-UPrior.txt"), sep="\n")
+```{r allen-priors}
+K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
+logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
+logtheta_prior <- function(x) dunif(x, -6, 6)
+par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
+                    logr0 = logr0_prior, logtheta = logtheta_prior,
+                    stdQ = stdQ_prior, stdR = stdR_prior)
 ```
 
-We define which parameters to keep track of, and set the initial values of parameters in the transformed space used by the MCMC.
-We use logarithms to maintain strictly positive values of parameters where appropriate.
+We define which parameters to keep track of, and set the initial values of
+parameters in the transformed space used by the MCMC.  We use logarithms
+to maintain strictly positive values of parameters where appropriate.
 
-```{r}
-# Uniform priors on standard deviation terms
+```{r allen-mcmc}
 jags.params=c("K","logr0","logtheta","stdQ", "stdR")
 jags.inits <- function(){
-  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), 
+       "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,
+       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
 }
@@
 set.seed(12345)
-
-time_jags <- system.time(       
 jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
-                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="bugmodel-UPrior.txt")
-)         
-time_jags <- unname(time_jags["elapsed"]);
+                n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin, 
+                model.file="allen.bugs", progress.bar="none")
 ```
@@
 
-```{r}
+```{r allen-traces}
 jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-true_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
+allen_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
+ggplot(allen_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
 ```
 
-```{r parametric_bayes_traces}
-# Traces
-ggplot(true_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
-```
-
-```{r parametric_bayes_posteriors}
-## priors (untransformed variables)
-K_prior <- function(x) dunif(x, 0.01, 40)
-logr_prior <- function(x) dunif(x, -6, 6)
-logtheta_prior <- function(x) dunif(x, -6, 6)
-stdQ_prior <- function(x) dunif(x, 0.001, 100)
-stdR_prior <- function(x) dunif(x, 0.001, 100)
-
-par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
-                   logr0 = logr_prior, logtheta = logtheta_prior, 
-                   stdQ = stdQ_prior, stdR = stdR_prior)
-
-true_priors <- ddply(true_posteriors, "variable", function(dd){
+```{r allen-posteriors}
+allen_priors <- ddply(allen_posteriors, "variable", function(dd){
     grid <- seq(min(dd$value), max(dd$value), length = 100) 
@@
 
-# posterior distributions
-ggplot(true_posteriors, aes(value)) + 
+ggplot(allen_posteriors, aes(value)) + 
   stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=true_priors, aes(x=value, y=density), col="red") + 
+  geom_line(data=allen_priors, aes(x=value, y=density), col="red") + 
   facet_wrap(~ variable, scale="free", ncol=3)
-
 ```
 
-```{r}
-# um, cleaner if we were just be using the long form, par_posterior
-mcmc <- as.mcmc(jagsfit)
-mcmcall <- mcmc[,-2]
-who <- colnames(mcmcall)
-who 
-mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4],mcmcall[,5])
-colnames(mcmcall) <- who
+Reshape the posterior parameter distribution data, transform back into original space, and calculate the mean parameters and mean function
 
-pardist <- mcmcall
+```{r allen-output}
+pardist <- acast(allen_posteriors, index ~ variable, subset=.(variable != "deviance") )
 pardist[,2] = exp(pardist[,2]) # transform model parameters back first
 pardist[,3] = exp(pardist[,3])
-
-bayes_coef <- apply(pardist,2,mean)
-bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3]))
-bayes_pars
-```
-
-```{r}
-par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)
+bayes_coef <- apply(pardist,2, posterior.mode) 
+bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3])) # parameters formatted for f
+allen_means <- sapply(x_grid, f, 0, bayes_pars)
 ```
@@
 
-We initiate the MCMC chain (`init_p`) using the true values of the parameters `p` from the simulation.
-While impossible in real data, this gives the parametric Bayesian approach the best chance at succeeding.
-`y` is the timeseries (recall `obs` has the $x_t$, $x_{t+1}$ pairs)
+```{r ricker-jags}
+K_prior_p <- c(0.01, 40.0)
+logr0_prior_p <- c(-6.0, 6.0)
 
-```{r}
-init_p = p
-names(init_p) = c("r0", "K")
-y <- obs$x[-1] 
-N=length(y);
+bugs.model <- 
+paste(sprintf(
+"model{
+  K    ~ dunif(%s, %s)
+  logr0    ~ dunif(%s, %s)
+  stdQ ~ dunif(%s, %s)
+  stdR ~ dunif(%s, %s)", 
+  K_prior_p[1], K_prior_p[2],
+  logr0_prior_p[1], logr0_prior_p[2],
+  stdQ_prior_p[1], stdQ_prior_p[2],
+  stdR_prior_p[1], stdR_prior_p[2]),
+
+  "
+  iQ <- 1 / (stdQ * stdQ);
+  iR <- 1 / (stdR * stdR);
+  r0 <- exp(logr0)
+  x[1] ~ dunif(0, 10)
+  for(t in 1:(N-1)){
+    mu[t] <- x[t] * exp(r0 * (1 - x[t]/K) )
+    x[t+1] ~ dnorm(mu[t], iQ) 
+  }
+  for(t in 1:(N)){
+    y[t] ~ dnorm(x[t], iR)
+  }
+}")
+writeLines(bugs.model, "ricker.bugs")
 ```
 
-We'll be using the JAGS Gibbs sampler, a recent open source BUGS implementation with an R interface that works on most platforms.  We initialize the usual MCMC parameters; see `?jags` for details.
+Compute prior curves
 
-```{r}
-jags.data <- list("N","y")
-n.chains = 1
-n.iter = 400000
-n.burnin = floor(10000)
-n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
+```{r ricker-priors}
+K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
+logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
+par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
+                   logr0 = logr0_prior, stdQ = stdQ_prior, stdR = stdR_prior)
 ```
 
-The actual model is defined in a `model.file` that contains an R function that is automatically translated into BUGS code by *R2WinBUGS*.
-The file defines the priors and the model, as seen when read in here
+We define which parameters to keep track of, and set the initial values of
+parameters in the transformed space used by the MCMC.  We use logarithms
+to maintain strictly positive values of parameters where appropriate.
 
-```{r}
-cat(readLines(con="ricker-UPrior.txt"), sep="\n")
-```
-
-We define which parameters to keep track of, and set the initial values of parameters in the transformed space used by the MCMC.
-We use logarithms to maintain strictly positive values of parameters where appropriate.
-
-```{r}
+```{r ricker-mcmc}
 # Uniform priors on standard deviation terms
@@
 jags.inits <- function(){
-  list("K"=init_p["K"],"logr0"=log(init_p["r0"]), "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+  list("K"=init_p["K"], "logr0"=log(init_p["r0"]), 
+       "stdQ"=sqrt(0.05), "stdR"=sqrt(0.1), "x"=y, 
+       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
 }
-
 set.seed(12345) 
-
-time_jags <- system.time(       
 jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
-                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="ricker-UPrior.txt")
-)         
-time_jags <- unname(time_jags["elapsed"]);
+                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
+                  model.file="ricker.bugs", progress.bar="none")
 ```
@@
 jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
-ricker_posteriors <- par_posteriors
-
-# Traces
+ricker_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
 ggplot(ricker_posteriors) + geom_line(aes(index, value)) + 
@@
 ```{r ricker_posteriors}
-## priors (untransformed variables)
-K_prior <- function(x) dunif(x, 0.01, 40)
-logr_prior <- function(x) dunif(x, -6, 6)
-stdQ_prior <- function(x) dunif(x, 0.001, 100)
-stdR_prior <- function(x) dunif(x, 0.001, 100)
-
-par_priors <- list(K = K_prior, deviance = function(x) 0 * x, logr0 = logr_prior, stdQ = stdQ_prior, stdR = stdR_prior)
-
 ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
@@
 })
-
-# posterior distributions
+# plot posterior distributions
 ggplot(ricker_posteriors, aes(value)) + 
@@
 
-```{r}
-ricker_pardist <- jags_matrix[! names(jags_matrix) == "deviance" ]
+Reshape posteriors data, transform back, calculate mode and corresponding function.
+
+```{r ricker-output}
+ricker_pardist <- acast(ricker_posteriors, index ~ variable, subset=.(variable != "deviance") )
 ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first
-
-posterior.mode <- function(x) {
-  ux <- unique(x)
-  ux[which.max(tabulate(match(x, ux)))]
-}
-
-apply(ricker_pardist,2,mean)
-bayes_coef <- apply(ricker_pardist,2, posterior.mode) # much better estimates
+bayes_coef <- apply(ricker_pardist,2, posterior.mode) # much better estimates from mode then mean
 ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
-ricker_bayes_pars
+ricker_means <- sapply(x_grid, Ricker, 0, ricker_bayes_pars[c(1,2)])
 ```
@@
 
-```{r}
+```{r myers-jags}
 logr0_prior_p <- c(-6.0, 6.0)
@@
 logK_prior_p <- c(-6.0, 6.0)
-stdQ_prior_p <- c(0.0001, 100)
-stdR_prior_p <- c(0.0001, 100)
 
@@
   stdR_prior_p[1], stdR_prior_p[2]),
+
   "
-
   iQ <- 1 / (stdQ * stdQ);
@@
 }")
-
 writeLines(bugs.model, "myers.bugs")
@@
 
-```{r}
-
-## priors (untransformed variables)
+```{r myers-priors}
 logK_prior     <- function(x) dunif(x, logK_prior_p[1], logK_prior_p[2])
@@
 logtheta_prior <- function(x) dunif(x, logtheta_prior_p[1], logtheta_prior_p[2])
-stdQ_prior     <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
-stdR_prior     <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
+par_priors <- list( deviance = function(x) 0 * x, logK = logK_prior,
+                    logr0 = logr_prior, logtheta = logtheta_prior, 
+                    stdQ = stdQ_prior, stdR = stdR_prior)
 
@@
 
-```{r}
-
+```{r myers-mcmc}
 init_p = c(1, 1, 1)
 names(init_p) = c("r0", "theta", "K")
-y <- obs$x[-1] 
-N=length(y);
-```
-
-We'll be using the JAGS Gibbs sampler, a recent open source BUGS implementation with an R interface that works on most platforms.  We initialize the usual MCMC parameters; see `?jags` for details.
-
-```{r}
-jags.data <- list("N","y")
-n.chains = 1
-n.iter = 40000
-n.burnin = floor(10000)
-n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
-```
-
-```{r}
 jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
@@
 }
-```
-
-```{r}
 set.seed(12345)
-time_jags <- system.time(       
 jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, 
                   n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
-                  n.burnin=n.burnin, model.file="myers.bugs")
-)         
-time_jags <- unname(time_jags["elapsed"]);
+                  n.burnin=n.burnin, model.file="myers.bugs", progress.bar="none")
+
 ```
@@
 
-```{r}
+```{r myers-traces}
 jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
-
-## Traces
-ggplot(par_posteriors) + geom_line(aes(index, value)) +
+myers_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
+ggplot(myers_posteriors) + geom_line(aes(index, value)) +
   facet_wrap(~ variable, scale="free", ncol=1)
@@
 
-```{r}
-## prior curve functions, must be in order
-par_priors <- list( deviance = function(x) 0 * x, logK = logK_prior,
-                    logr0 = logr_prior, logtheta = logtheta_prior, 
-                    stdQ = stdQ_prior, stdR = stdR_prior)
-
-## calculate points along the prior curve
-par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
+```{r myers-posteriors}
+par_prior_curves <- ddply(myers_posteriors, "variable", function(dd){
     grid <- seq(min(dd$value), max(dd$value), length = 100) 
@@
 })
-```
 
-```{r}
-## posterior distributions
-ggplot(par_posteriors, aes(value)) + 
+ggplot(myers_posteriors, aes(value)) + 
   stat_density(geom="path", position="identity", alpha=0.7) +
@@
 
-```{r}
+```{r myers-output}
 myers_pardist <- as.matrix(jags_matrix[2:6])
@@
 colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
-
-apply(myers_pardist,2,mean)
 bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
-
 myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
+myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
 ```
@@
 
-```{r}
-true_means <- sapply(x_grid, f, 0, p)
-alt_means <- sapply(x_grid, alt$f, 0, ricker_bayes_pars[c(1,2)])
-est_means <- sapply(x_grid, est$f, 0, est$p)
-par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)
-
-myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
-
+```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
 models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
-                     MLE=est_means, Ricker=alt_means, 
-                     Parametric.Bayes = par_bayes_means,
+                     MLE=est_means, Ricker=ricker_means, 
+                     Allen = allen_means,
                      Myers = myers_means)
@@
 
-```{r}
+```{r gp-opt, dependson="assemble-models"}
 MaxT = 1000
@@
 
-Determine the optimal policy based on the true and MLE models
+Determine the optimal policy based on the allen and MLE models
 
-```{r}
+```{r mle-opt, dependson="assemble-models"}
 matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
@@
 
-Determine the optimal policy based on parametric Bayesian model
+Determine the optimal policy based on Bayesian Allen model
 
-```{r}
+```{r allen-opt, dependson="assemble-models"}
 allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
-matrices_par_bayes <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
-opt_par_bayes <- value_iteration(matrices_par_bayes, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+matrices_allen <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
+opt_allen <- value_iteration(matrices_allen, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
 ```
@@
 
-```{r}
-ricker_f <- function(x, h, p) alt$f(x, h, unname(p[c(2, 1)]))
-matrices_alt <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
-opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+```{r ricker-opt, dependson="assemble-models"}
+ricker_f <- function(x, h, p) Ricker(x, h, unname(p[c(2, 1)])) # defined by pdgControl 
+matrices_ricker <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
+opt_ricker <- value_iteration(matrices_ricker, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
 ```
@@
 
-```{r}
+```{r myers-opt, dependson="assemble-models"}
 myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
@@
 
-```{r}
-OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_alt$D, Parametric.Bayes = opt_par_bayes$D, Myers = myers_alt$D)
+```{r assemble-opt, , dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
+OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_ricker$D, Allen = opt_allen$D, Myers = myers_alt$D)
 colorkey=cbPalette
@@
 
-```{r Figure2}
+```{r Figure2, dependson="assemble-opt"}
 policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
@@
 
-```{r sims}
+```{r sims, dependson="assemble-opt"}
 sims <- lapply(OPT, function(D){
@@
 
-```{r Figure3}
+```{r Figure3, dependson="sims"}
 ggplot(dt) + 
@@
 
-```{r profits}
+```{r profits, dependson="sims"}
 Profit <- dt[, sum(profit), by=c("reps", "method")]
@@
 
-```{r totalprofits}
+```{r totalprofits, dependson="profits"}
 ggplot(Profit, aes(V1)) + geom_histogram() + 

### 2013-05-22 inst/examples/BUGS/myers.Rmd
replicate runs on data from allen and myers model

--- 2013-05-22 inst/examples/BUGS/allen.Rmd
@@
 opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA,
-               fig.width=6, fig.height=4, cache.path="allen/")
+               fig.width=6, fig.height=4, cache.path="myers/")
 opts_knit$set(upload.fun = socialR::flickr.url)
@@
 ```{r stateeq}
-f <- RickerAllee
-p <- c(2, 8, 5)
-K <- p[2]
-allee <- p[3]
+f <- Myers
+p <- c(1, 2, 6)
+K <- 5  # approx, a li'l' less
+allee <- 1.2 # approx, a li'l' less
 ```
+
+Various parameters defining noise dynamics, grid, and policy costs.
 
@@
   x <- numeric(Tobs)
-  x[1] <- 6
+  x[1] <- 2
   nz <- 1

### 2013-05-22 inst/examples/BUGS/myers.Rmd
still need to fix up Allen MCMC convergence when underlying model is Myers

@@
 ```{r allen_pars}
-# a bit unfair to start with the correct values, but anyhow...
-init_p = p
+init_p = c(1, 10, 2)
 names(init_p) = c("r0", "K", "theta")
@@
 
-```{r allen-jags}
+```{r allen-model}
 K_prior_p <- c(0.01, 40.0)
@@
 
-```{r allen-priors}
+```{r allen-priors, dependson="common-priors"}
 K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
@@
 
-```{r allen-mcmc}
+```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup")}
 jags.params=c("K","logr0","logtheta","stdQ", "stdR")
@@
 
-```{r allen-traces}
+```{r allen-traces, dependson="allen-mcmc"}
 jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
@@
 
-```{r allen-posteriors}
+```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
 allen_priors <- ddply(allen_posteriors, "variable", function(dd){
@@
 
-```{r allen-output}
+```{r allen-output, dependson="allen-traces"}
 pardist <- acast(allen_posteriors, index ~ variable, subset=.(variable != "deviance") )
@@
 
-```{r ricker-jags}
+```{r ricker-model}
 K_prior_p <- c(0.01, 40.0)
@@
 
-```{r ricker-priors}
+```{r ricker-priors, dependson="common-priors"}
 K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
@@
 
-```{r ricker-mcmc}
+```{r ricker-mcmc, dependson="ricker-model"}
 # Uniform priors on standard deviation terms
@@
 
-```{r ricker_traces}
+```{r ricker_traces, dependson="ricker-mcmc"}
 jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
@@
 
-```{r ricker_posteriors}
+```{r ricker_posteriors, dependson=c("ricker-traces", "ricker-priors")}
 ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
@@
 
-```{r ricker-output}
+```{r ricker-output, dependson="ricker-traces"}
 ricker_pardist <- acast(ricker_posteriors, index ~ variable, subset=.(variable != "deviance") )
@@
 
-```{r myers-jags}
+```{r myers-model}
 logr0_prior_p <- c(-6.0, 6.0)
@@
 
-```{r myers-mcmc}
-init_p = c(1, 1, 1)
+```{r myers-mcmc, dependson="myers-model"}
+init_p = c(p[1], p[3], p[2])
 names(init_p) = c("r0", "theta", "K")
@@
 
-```{r myers-traces}
+```{r myers-traces, dependson="myers-mcmc"}
 jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
@@
 
-```{r myers-posteriors}
+```{r myers-posteriors, dependson="myers-traces"}
 par_prior_curves <- ddply(myers_posteriors, "variable", function(dd){
@@
 
-```{r myers-output}
+```{r myers-output, dependson="myers-traces"}
 myers_pardist <- as.matrix(jags_matrix[2:6])
@@
 names(models) <- c("x", "method", "value")
+
+model_names = c("GP", "True", "MLE", "Ricker", "Allen", "Myers")
+colorkey=cbPalette
+names(colorkey) = model_names 
 ```
 
-```{r Figure1}
+```{r Figure1, dependson="assemble-models"}
 plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
@@
 
+## Step-ahead predictors
+
+## Posterior predictive curves
+
+This shows only the mean predictions.
+For the Bayesian cases, we can instead loop over the posteriors of the parameters (or samples from the GP posterior) to get the distribution of such curves in each case.
+
+We will need a vector version (`pmin` in place of `min`) of the parametric growth functions that can operate on the posteriors, (with appropriate ordering of parameters as they are in the posterior):
+
+```{r par-fns}
+ricker_f <- function(x,h,p){
+  sapply(x, function(x){ 
+    x <- pmax(0, x-h) 
+    pmax(0, x * exp(p[2] * (1 - x / p[1] )) )
+  })
+}
+allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
+myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
+
+```
+
+```{r Figureb, dependson=c("assemble-models", "par-fns")}
+require(MASS)
+step_ahead <- function(x, f, p){
+  h = 0
+  x_predict <- sapply(x, f, h, p)
+  n <- length(x_predict) - 1
+  y <- c(x[1], x_predict[1:n])
+  y
+}
+step_ahead_posteriors <- function(x){
+gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
+df_post <- melt(lapply(sample(100), 
+  function(i){
+    data.frame(time = 1:length(x), stock = x, 
+                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
+                True = step_ahead(x,f,p),  
+                MLE = step_ahead(x,f,est$p), 
+                Allen = step_ahead(x, allen_f, pardist[i,]), 
+                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
+                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
+  }), id=c("time", "stock"))
+}
+
+df_post <- step_ahead_posteriors(x)
+
+ggplot(df_post) + geom_point(aes(time, stock)) + 
+  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) 
+```
+
 ## Optimal policies by value iteration
@@
 
-```{r gp-opt, dependson="assemble-models"}
+```{r gp-opt, dependson="gp-output"}
 MaxT = 1000
@@
 
-```{r mle-opt, dependson="assemble-models"}
+```{r mle-opt, dependson="mle-output"}
 matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
@@
 
-```{r allen-opt, dependson="assemble-models"}
+```{r allen-opt, dependson="allen-output"}
 allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
@@
 
-```{r ricker-opt, dependson="assemble-models"}
+```{r ricker-opt, dependson="ricker-output"}
 ricker_f <- function(x, h, p) Ricker(x, h, unname(p[c(2, 1)])) # defined by pdgControl 
@@
 
-```{r myers-opt, dependson="assemble-models"}
+```{r myers-opt, dependson="myers-output"}
 myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
@@
 
-```{r assemble-opt, , dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
+```{r assemble-opt, dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
 OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_ricker$D, Allen = opt_allen$D, Myers = myers_alt$D)

### 2013-05-23 inst/examples/BUGS/myers.Rmd
Revising MCMC iterations
* Added multiple chains, randomized starting points for each chain, and confirming Gelman-Rubin convergence criteron via `autojags`.
* Some edits to adapt the extraction of posteriors from the multiple chains.
* Stange behavior in `jags.parallel` [SO/16723036](http://stackoverflow.com/questions/16723036), looks like an annoying bug.

@@
 opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA,
-               fig.width=6, fig.height=4, cache.path="myers/")
+               fig.width=6, fig.height=4, cache.path="myers/", verbose=TRUE)
 opts_knit$set(upload.fun = socialR::flickr.url)
@@
 ```{r mle, dependson="obs"}
-est <- par_est_allee(obs, f, p,  
-                     init = c(r = p[1] + 1, 
-                              K = p[2] + 2, 
-                              C = p[3] + 2, 
-                              s = sigma_g))
+set.seed(12345)
+estf <- function(p){ 
+    mu <- f(obs$x,0,p)
+    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
+}
+par <- c(p[1]+abs(rnorm(1,0,1)), 
+         p[2]+rnorm(1,0,.1), 
+         p[3]+rnorm(1,0, .1), 
+         sigma_g + abs(rnorm(1,0,.1)))
+o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
+f_alt <- f
+p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
+sigma_g_alt <- as.numeric(o$par[4])
+
+est <- list(f = f_alt, p = p_alt, sigma_g = sigma_g_alt, mloglik=o$value)
 ```
@@
 
-```{r mle-output}
+```{r mle-output, dependson="mle"}
 true_means <- sapply(x_grid, f, 0, p)
@@
 jags.data <- list("N","y")
-n.chains <- 1
-n.iter <- 40000
+n.chains <- 3
+n.iter <- 1e5
 n.burnin <- floor(10000)
@@
 ```{r allen_pars}
-init_p = c(1, 10, 2)
+init_p = c(2, K, 0.01)
 names(init_p) = c("r0", "K", "theta")
@@
 ```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup")}
-jags.params=c("K","logr0","logtheta","stdQ", "stdR")
+jags.params=c("K","logr0","logtheta","stdQ", "stdR") # be sensible about the order here
+
+# Can give a list of starting values or a function that can create random values 
 jags.inits <- function(){
-  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), 
-       "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,
+  list("K"=init_p["K"] * rlnorm(1,0,.5),
+       "logr0"=log(init_p["r0"]) * rlnorm(1,0,.5),
+       "logtheta"=log(init_p["theta"]) * rlnorm(1,0,.5), 
+       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
+       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
+       "x"=y,
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
@@
 set.seed(12345)
-jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
+# parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
+allen_jags <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
                 n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin, 
-                model.file="allen.bugs", progress.bar="none")
+                model.file="allen.bugs", progress="none")
+
+#Run again iteratively if we haven't met the Gelman-Rubin convergence criterion
+allen_jags <- autojags(allen_jags, n.update=10, n.iter=n.iter, 
+                       n.thin = n.thin, progress="none") 
 ```
@@
 
+R notes: this strips classes from the `mcmc.list` object (so that we have list of matrices; objects that `reshape2::melt` can handle intelligently), and then combines chains into one array.
+In this array each parameter is given its value at each sample from the posterior (index) for each chain.
+
 ```{r allen-traces, dependson="allen-mcmc"}
-jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-allen_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
-ggplot(allen_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
+tmp <- lapply(as.mcmc(allen_jags), as.matrix) # strip classes the hard way...
+allen_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
+names(allen_posteriors) = c("index", "variable", "value", "chain")
+ggplot(allen_posteriors) + geom_line(aes(index, value)) + 
+  facet_wrap(~ variable, scale="free", ncol=1)
 ```
@@
 ```{r allen-output, dependson="allen-traces"}
-pardist <- acast(allen_posteriors, index ~ variable, subset=.(variable != "deviance") )
+# A <- allen_posteriors
+# A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
+# pardist <- acast(A[-4], index ~ variable, subset=.(variable != "deviance") )
+pardist <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable, subset=.(variable!="deviance")) 
 pardist[,2] = exp(pardist[,2]) # transform model parameters back first
@@
 jags.inits <- function(){
-  list("K"=init_p["K"], "logr0"=log(init_p["r0"]), 
-       "stdQ"=sqrt(0.05), "stdR"=sqrt(0.1), "x"=y, 
+  list("K"=init_p["K"] * rlnorm(1,0,.5),
+       "logr0"=log(init_p["r0"]) * rlnorm(1,0,.5),
+       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
+       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
+       "x"=y,
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
@@
 set.seed(12345) 
-jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
+ricker_jags <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
                   n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
                   model.file="ricker.bugs", progress.bar="none")
+ricker_jags <- autojags(ricker_jags, n.update=10, n.iter=n.iter, 
+                        n.thin = n.thin, progress.bar="none")
 ```
@@
 ```{r ricker_traces, dependson="ricker-mcmc"}
-jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-ricker_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
+tmp <- lapply(as.mcmc(ricker_jags), as.matrix) # strip classes the hard way...
+ricker_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
+names(ricker_posteriors) = c("index", "variable", "value", "chain")
+
 ggplot(ricker_posteriors) + geom_line(aes(index, value)) + 
@@
 ```{r ricker-output, dependson="ricker-traces"}
-ricker_pardist <- acast(ricker_posteriors, index ~ variable, subset=.(variable != "deviance") )
+ricker_pardist <- acast(ricker_posteriors[2:3], 
+                        1:table(ricker_posteriors$variable) ~ variable, 
+                        subset=.(variable!="deviance")) 
 ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first
@@
 jags.inits <- function(){
-  list("logr0"=log(init_p["r0"]), "logtheta" = log(init_p["theta"]),  
-       "logK"=log(init_p["K"]), "stdQ"=sqrt(0.05), 
-       "stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+  list("logr0"=log(init_p["r0"]) * rlnorm(1,0,.5), 
+       "logK"=log(init_p["K"]),
+       "logtheta" = log(init_p["theta"]) * rlnorm(1,0,.5),  
+       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
+       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
+       "x"=y,
+       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
 }
 set.seed(12345)
-jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, 
+myers_jags <- jags(data=jags.data, inits=jags.inits, jags.params, 
                   n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
                   n.burnin=n.burnin, model.file="myers.bugs", progress.bar="none")
-
+myers_jags <- autojags(myers_jags, n.update=10, n.iter=n.iter, 
+                    n.thin = n.thin, progress.bar="none")
 ```
@@
 ```{r myers-traces, dependson="myers-mcmc"}
-jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-myers_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
+tmp <- lapply(as.mcmc(myers_jags), as.matrix) # strip classes the hard way...
+myers_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
+names(myers_posteriors) = c("index", "variable", "value", "chain")
+
 ggplot(myers_posteriors) + geom_line(aes(index, value)) +
@@
 ```{r myers-output, dependson="myers-traces"}
-myers_pardist <- as.matrix(jags_matrix[2:6])
+myers_pardist <- acast(myers_posteriors[2:3], 
+                        1:table(myers_posteriors$variable) ~ variable, 
+                        subset=.(variable!="deviance")) 
 myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
@@
 ```
+
+```{r}
+df <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable) 
+modes <- apply(df, 2, posterior.mode)
+allen_deviance <- modes[['deviance']]
+
+df <- acast(ricker_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable)
+modes <- apply(df, 2, posterior.mode)
+ricker_deviance <- modes[['deviance']]
+
+df <- acast(myers_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable)
+modes <- apply(df, 2, posterior.mode)
+myers_deviance <- modes[['deviance']]
+
+true_deviance <- 2*estf(c(p, sigma_g))
+mle_deviance <- 2*estf(c(est$p, est$sigma_g))
+
+c(allen = allen_deviance, ricker=ricker_deviance, myers=myers_deviance, true=true_deviance, mle=mle_deviance)
+
+```

### 2013-05-23 inst/examples/BUGS/parametric-vs-nonparametric.Rmd
add bayesian version of Ricker
cleanup

--- 2013-05-01 inst/examples/BUGS/allen.Rmd
@@
 
-```{r plotting-options}
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA)
-
+```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
 library(knitcitations)
-
-library(ggplot2) # plotting 
+library(ggplot2) 
+library(nonparametricbayes) 
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA,
+               fig.width=6, fig.height=4, cache.path="myers/", verbose=TRUE)
 opts_knit$set(upload.fun = socialR::flickr.url)
-theme_set(theme_bw(base_size=10))
+theme_set(theme_bw(base_size=12))
 theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
@@
 
-Uses the model derived in `citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
+Uses the model derived in `r citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
 
-```{r}
+```{r stateeq}
 f <- RickerAllee
-p <- c(2, 8, 5)
-K <- p[2]
-allee <- p[3]
+p <- c(1, 10, 5)
+K <- 10  # approx, a li'l' less
+allee <- 5 # approx, a li'l' less
 ```
+
+Various parameters defining noise dynamics, grid, and policy costs.
 
@@
 delta <- 0.01
-OptTime <- 30  # stationarity with unstable models is tricky thing
+OptTime <- 50  # stationarity with unstable models is tricky thing
 reward <- 0
 xT <- 0
-seed_i <- 111
 Xo <- K # observations start from
 x0 <- Xo # simulation under policy starts from
-Tobs <- 35
+Tobs <- 40
 ```
@@
 ```{r obs, dependson="sdp-pars"}
+  set.seed(1234)
   #harvest <- sort(rep(seq(0, .5, length=7), 5))
   x <- numeric(Tobs)
-  x[1] <- 6
+  x[1] <- 2
   nz <- 1
@@
 ```{r mle, dependson="obs"}
-alt <- par_est(obs,  init = c(r = p[1], 
-                              K = mean(obs$x[obs$x>0]), 
-                              s = sigma_g))
-est <- par_est_allee(obs, f, p,  
-                     init = c(r = p[1] + 1, 
-                              K = p[2] + 2, 
-                              C = p[3] + 2, 
-                              s = sigma_g))
+set.seed(12345)
+estf <- function(p){ 
+    mu <- f(obs$x,0,p)
+    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
+}
+par <- c(p[1]+abs(rnorm(1,0,1)), 
+         p[2]+rnorm(1,0,.1), 
+         p[3]+rnorm(1,0, .1), 
+         sigma_g + abs(rnorm(1,0,.1)))
+o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
+f_alt <- f
+p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
+sigma_g_alt <- as.numeric(o$par[4])
+
+est <- list(f = f_alt, p = p_alt, sigma_g = sigma_g_alt, mloglik=o$value)
+```
+
+Mean predictions
+
+```{r mle-output, dependson="mle"}
+true_means <- sapply(x_grid, f, 0, p)
+est_means <- sapply(x_grid, est$f, 0, est$p)
 ```
@@
 
-```{r}
+```{r gp, dependson=c("gp-priors", "obs")}
 gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
-gp_dat <- gp_predict(gp, x_grid)
-
+gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)
 ```
@@
 
-```{r}
+```{r gp_traces_densities, dependson="gp"}
 plots <- summary_gp_mcmc(gp)
-plots[[1]]
-plots[[2]]
 ```
 
-```{r}
+```{r gp-output, dependson="gp"}
 # Summarize the GP model
@@
 
-### Parametric Bayes
+Parametric Bayesian Models
+---------------------------
 
-We initiate the MCMC chain (`init_p`) using the true values of the parameters `p` from the simulation.
-While impossible in real data, this gives the parametric Bayesian approach the best chance at succeeding.
+We use the JAGS Gibbs sampler, a recent open source BUGS
+implementation with an R interface that works on most platforms.
+We initialize the usual MCMC parameters; see `?jags` for details.
+
+All parametric Bayesian estimates use the following basic parameters for the JAGS MCMC:
+
+```{r jags-setup}
+y <- obs$x[-1] 
+N <- length(y);
+jags.data <- list("N","y")
+n.chains <- 3
+n.iter <- 1e5
+n.burnin <- floor(10000)
+n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
+```
+
+We will use the same priors for process and observation noise in each model,
+
+```{r common-priors}
+stdQ_prior_p <- c(0.0001, 100)
+stdR_prior_p <- c(0.0001, 100)
+stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
+stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
+```
+
+### Parametric Bayes of correct (Allen) model
+
+We initiate the MCMC chain (`init_p`) using the true values of the
+parameters `p` from the simulation.  While impossible in real data, this
+gives the parametric Bayesian approach the best chance at succeeding.
 `y` is the timeseries (recall `obs` has the $x_t$, $x_{t+1}$ pairs)
 
-```{r}
-# a bit unfair to start with the correct values, but anyhow...
-init_p = p
+```{r allen_pars}
+init_p = c(2, K, 0.01)
 names(init_p) = c("r0", "K", "theta")
-y <- obs$x[-1] 
-N=length(y);
 ```
 
-We'll be using the JAGS Gibbs sampler, a recent open source BUGS implementation with an R interface that works on most platforms.  We initialize the usual MCMC parameters; see `?jags` for details.
+The actual model is defined in a `model.file` that contains an R function
+that is automatically translated into BUGS code by *R2WinBUGS*.  The file
+defines the priors and the model. We write the file from R as follows:
 
-```{r}
-jags.data <- list("N","y")
-n.chains = 1
-n.iter = 40000
-n.burnin = floor(10000)
-n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
+```{r allen-model}
+K_prior_p <- c(0.01, 40.0)
+logr0_prior_p <- c(-6.0, 6.0)
+logtheta_prior_p <- c(-6.0, 6.0)
+
+bugs.model <- 
+paste(sprintf(
+"model{
+  K     ~ dunif(%s, %s)
+  logr0    ~ dunif(%s, %s)
+  logtheta ~ dunif(%s, %s)
+  stdQ ~ dunif(%s, %s)
+  stdR ~ dunif(%s, %s)", 
+  K_prior_p[1], K_prior_p[2],
+  logr0_prior_p[1], logr0_prior_p[2],
+  logtheta_prior_p[1], logtheta_prior_p[2],
+  stdQ_prior_p[1], stdQ_prior_p[2],
+  stdR_prior_p[1], stdR_prior_p[2]),
+
+  "
+  iQ <- 1 / (stdQ * stdQ);
+  iR <- 1 / (stdR * stdR);
+  r0 <- exp(logr0)
+  theta <- exp(logtheta)
+  x[1] ~ dunif(0, 10)
+  for(t in 1:(N-1)){
+    mu[t] <- x[t] * exp(r0 * (1 - x[t]/K)* (x[t] - theta) / K )
+    x[t+1] ~ dnorm(mu[t], iQ) 
+  }
+  for(t in 1:(N)){
+    y[t] ~ dnorm(x[t], iR)
+  }
+}")
+writeLines(bugs.model, "allen.bugs")
 ```
 
-The actual model is defined in a `model.file` that contains an R function that is automatically translated into BUGS code by *R2WinBUGS*.
-The file defines the priors and the model, as seen when read in here
+Write the priors into a list for later reference
 
-```{r}
-cat(readLines(con="bugmodel-UPrior.txt"), sep="\n")
+```{r allen-priors, dependson="common-priors"}
+K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
+logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
+logtheta_prior <- function(x) dunif(x, -6, 6)
+par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
+                    logr0 = logr0_prior, logtheta = logtheta_prior,
+                    stdQ = stdQ_prior, stdR = stdR_prior)
 ```
 
-We define which parameters to keep track of, and set the initial values of parameters in the transformed space used by the MCMC.
-We use logarithms to maintain strictly positive values of parameters where appropriate.
-Because our priors on the noise parameters are inverse gamma distributed.
+We define which parameters to keep track of, and set the initial values of
+parameters in the transformed space used by the MCMC.  We use logarithms
+to maintain strictly positive values of parameters where appropriate.
 
-```{r}
-# Uniform priors on standard deviation terms
-jags.params=c("K","logr0","logtheta","stdQ", "stdR")
+```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup")}
+jags.params=c("K","logr0","logtheta","stdQ", "stdR") # be sensible about the order here
+
+# Can give a list of starting values or a function that can create random values 
 jags.inits <- function(){
-  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+  list("K"=init_p["K"] * rlnorm(1,0,.5),
+       "logr0"=log(init_p["r0"]) * rlnorm(1,0,.5),
+       "logtheta"=log(init_p["theta"]) * rlnorm(1,0,.5), 
+       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
+       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
+       "x"=y,
+       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
 }
@@
 set.seed(12345)
+# parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
+allen_jags <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
+                n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin, 
+                model.file="allen.bugs", progress="none")
 
-time_jags <- system.time(       
-  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
-                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="bugmodel-UPrior.txt")
-)         
-time_jags <- unname(time_jags["elapsed"]);
+#Run again iteratively if we haven't met the Gelman-Rubin convergence criterion
+allen_jags <- autojags(allen_jags, n.update=10, n.iter=n.iter, 
+                       n.thin = n.thin, progress="none") 
 ```
@@
 
-```{r}
-jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
+R notes: this strips classes from the `mcmc.list` object (so that we have list of matrices; objects that `reshape2::melt` can handle intelligently), and then combines chains into one array.
+In this array each parameter is given its value at each sample from the posterior (index) for each chain.
 
-# Traces
-ggplot(par_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
+```{r allen-traces, dependson="allen-mcmc"}
+tmp <- lapply(as.mcmc(allen_jags), as.matrix) # strip classes the hard way...
+allen_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
+names(allen_posteriors) = c("index", "variable", "value", "chain")
+ggplot(allen_posteriors) + geom_line(aes(index, value)) + 
+  facet_wrap(~ variable, scale="free", ncol=1)
+```
 
-## priors (untransformed variables)
-K_prior <- function(x) dunif(x, 0.01, 40)
-logr_prior <- function(x) dunif(x, -6, 6)
-logtheta_prior <- function(x) dunif(x, -6, 6)
-stdQ_prior <- function(x) dunif(x, 0.001, 100)
-stdR_prior <- function(x) dunif(x, 0.001, 100)
-
-par_priors <- list(K = K_prior, deviance = function(x) 0 * x, logr0 = logr_prior, logtheta = logtheta_prior, stdQ = stdQ_prior, stdR = stdR_prior)
-
-par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
+```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
+allen_priors <- ddply(allen_posteriors, "variable", function(dd){
     grid <- seq(min(dd$value), max(dd$value), length = 100) 
@@
 
-# posterior distributions
-ggplot(par_posteriors, aes(value)) + 
+ggplot(allen_posteriors, aes(value)) + 
   stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
+  geom_line(data=allen_priors, aes(x=value, y=density), col="red") + 
+  facet_wrap(~ variable, scale="free", ncol=3)
+```
+
+Reshape the posterior parameter distribution data, transform back into original space, and calculate the mean parameters and mean function
+
+```{r allen-output, dependson="allen-traces"}
+# A <- allen_posteriors
+# A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
+# pardist <- acast(A[-4], index ~ variable, subset=.(variable != "deviance") )
+pardist <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable, subset=.(variable!="deviance")) 
+pardist[,2] = exp(pardist[,2]) # transform model parameters back first
+pardist[,3] = exp(pardist[,3])
+bayes_coef <- apply(pardist,2, posterior.mode) 
+bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3])) # parameters formatted for f
+allen_means <- sapply(x_grid, f, 0, bayes_pars)
+```
+
+## Parametric Bayes based on the structurally wrong model (Ricker)
+
+```{r ricker-model}
+K_prior_p <- c(0.01, 40.0)
+logr0_prior_p <- c(-6.0, 6.0)
+
+bugs.model <- 
+paste(sprintf(
+"model{
+  K    ~ dunif(%s, %s)
+  logr0    ~ dunif(%s, %s)
+  stdQ ~ dunif(%s, %s)
+  stdR ~ dunif(%s, %s)", 
+  K_prior_p[1], K_prior_p[2],
+  logr0_prior_p[1], logr0_prior_p[2],
+  stdQ_prior_p[1], stdQ_prior_p[2],
+  stdR_prior_p[1], stdR_prior_p[2]),
+
+  "
+  iQ <- 1 / (stdQ * stdQ);
+  iR <- 1 / (stdR * stdR);
+  r0 <- exp(logr0)
+  x[1] ~ dunif(0, 10)
+  for(t in 1:(N-1)){
+    mu[t] <- x[t] * exp(r0 * (1 - x[t]/K) )
+    x[t+1] ~ dnorm(mu[t], iQ) 
+  }
+  for(t in 1:(N)){
+    y[t] ~ dnorm(x[t], iR)
+  }
+}")
+writeLines(bugs.model, "ricker.bugs")
+```
+
+Compute prior curves
+
+```{r ricker-priors, dependson="common-priors"}
+K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
+logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
+par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
+                   logr0 = logr0_prior, stdQ = stdQ_prior, stdR = stdR_prior)
+```
+
+We define which parameters to keep track of, and set the initial values of
+parameters in the transformed space used by the MCMC.  We use logarithms
+to maintain strictly positive values of parameters where appropriate.
+
+```{r ricker-mcmc, dependson="ricker-model"}
+# Uniform priors on standard deviation terms
+jags.params=c("K","logr0", "stdQ", "stdR")
+jags.inits <- function(){
+  list("K"=init_p["K"] * rlnorm(1,0,.5),
+       "logr0"=log(init_p["r0"]) * rlnorm(1,0,.5),
+       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
+       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
+       "x"=y,
+       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+}
+set.seed(12345) 
+ricker_jags <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
+                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
+                  model.file="ricker.bugs", progress.bar="none")
+ricker_jags <- autojags(ricker_jags, n.update=10, n.iter=n.iter, 
+                        n.thin = n.thin, progress.bar="none")
+```
+
+#### Convergence diagnostics for parametric bayes Ricker model
+
+```{r ricker_traces, dependson="ricker-mcmc"}
+tmp <- lapply(as.mcmc(ricker_jags), as.matrix) # strip classes the hard way...
+ricker_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
+names(ricker_posteriors) = c("index", "variable", "value", "chain")
+
+ggplot(ricker_posteriors) + geom_line(aes(index, value)) + 
+  facet_wrap(~ variable, scale="free", ncol=1)
+```
+
+```{r ricker_posteriors, dependson=c("ricker-traces", "ricker-priors")}
+ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
+    grid <- seq(min(dd$value), max(dd$value), length = 100) 
+    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
+})
+# plot posterior distributions
+ggplot(ricker_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") + 
   facet_wrap(~ variable, scale="free", ncol=2)
+```
+
+Reshape posteriors data, transform back, calculate mode and corresponding function.
+
+```{r ricker-output, dependson="ricker-traces"}
+ricker_pardist <- acast(ricker_posteriors[2:3], 
+                        1:table(ricker_posteriors$variable) ~ variable, 
+                        subset=.(variable!="deviance")) 
+ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first
+bayes_coef <- apply(ricker_pardist,2, posterior.mode) # much better estimates from mode then mean
+ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
+ricker_means <- sapply(x_grid, Ricker, 0, ricker_bayes_pars[c(1,2)])
+```
+
+## Myers Parametric Bayes
+
+```{r myers-model}
+logr0_prior_p <- c(-6.0, 6.0)
+logtheta_prior_p <- c(-6.0, 6.0)
+logK_prior_p <- c(-6.0, 6.0)
+
+bugs.model <- 
+paste(sprintf(
+"model{
+  logr0    ~ dunif(%s, %s)
+  logtheta    ~ dunif(%s, %s)
+  logK    ~ dunif(%s, %s)
+  stdQ ~ dunif(%s, %s)
+  stdR ~ dunif(%s, %s)", 
+  logr0_prior_p[1], logr0_prior_p[2],
+  logtheta_prior_p[1], logtheta_prior_p[2],
+  logK_prior_p[1], logK_prior_p[2],
+  stdQ_prior_p[1], stdQ_prior_p[2],
+  stdR_prior_p[1], stdR_prior_p[2]),
+
+  "
+  iQ <- 1 / (stdQ * stdQ);
+  iR <- 1 / (stdR * stdR);
+
+  r0 <- exp(logr0)
+  theta <- exp(logtheta)
+  K <- exp(logK)
+
+  x[1] ~ dunif(0, 10)
+
+  for(t in 1:(N-1)){
+    mu[t] <- r0 * pow(abs(x[t]), theta) / (1 + pow(abs(x[t]), theta) / K)
+    x[t+1] ~ dnorm(mu[t], iQ) 
+  }
+
+  for(t in 1:(N)){
+    y[t] ~ dnorm(x[t], iR)
+  }
+}")
+writeLines(bugs.model, "myers.bugs")
+```
+
+```{r myers-priors}
+logK_prior     <- function(x) dunif(x, logK_prior_p[1], logK_prior_p[2])
+logr_prior     <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
+logtheta_prior <- function(x) dunif(x, logtheta_prior_p[1], logtheta_prior_p[2])
+par_priors <- list( deviance = function(x) 0 * x, logK = logK_prior,
+                    logr0 = logr_prior, logtheta = logtheta_prior, 
+                    stdQ = stdQ_prior, stdR = stdR_prior)
 
@@
 
-```{r}
-# um, cleaner if we were just be using the long form, par_posterior
-mcmc <- as.mcmc(jagsfit)
-mcmcall <- mcmc[,-2]
-who <- colnames(mcmcall)
-who 
-mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4],mcmcall[,5])
-colnames(mcmcall) <- who
-
-pardist <- mcmcall
-pardist[,2] = exp(pardist[,2]) # transform model parameters back first
-pardist[,3] = exp(pardist[,3])
-
-bayes_coef <- apply(pardist,2,mean)
-bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3]))
-bayes_pars
+```{r myers-mcmc, dependson="myers-model"}
+init_p = c(p[1], p[3], p[2])
+names(init_p) = c("r0", "theta", "K")
+jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
+jags.inits <- function(){
+  list("logr0"=log(init_p["r0"]) * rlnorm(1,0,.5), 
+       "logK"=log(init_p["K"]),
+       "logtheta" = log(init_p["theta"]) * rlnorm(1,0,.5),  
+       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
+       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
+       "x"=y,
+       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+}
+set.seed(12345)
+myers_jags <- jags(data=jags.data, inits=jags.inits, jags.params, 
+                  n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
+                  n.burnin=n.burnin, model.file="myers.bugs", progress.bar="none")
+myers_jags <- autojags(myers_jags, n.update=10, n.iter=n.iter, 
+                    n.thin = n.thin, progress.bar="none")
 ```
 
-```{r}
-par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)
+Convergence diagnostics for parametric bayes
+
+```{r myers-traces, dependson="myers-mcmc"}
+tmp <- lapply(as.mcmc(myers_jags), as.matrix) # strip classes the hard way...
+myers_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
+names(myers_posteriors) = c("index", "variable", "value", "chain")
+
+ggplot(myers_posteriors) + geom_line(aes(index, value)) +
+  facet_wrap(~ variable, scale="free", ncol=1)
+
+```
+
+```{r myers-posteriors, dependson="myers-traces"}
+par_prior_curves <- ddply(myers_posteriors, "variable", function(dd){
+    grid <- seq(min(dd$value), max(dd$value), length = 100) 
+    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
+})
+
+ggplot(myers_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
+  facet_wrap(~ variable, scale="free", ncol=3)
+```
+
+```{r myers-output, dependson="myers-traces"}
+myers_pardist <- acast(myers_posteriors[2:3], 
+                        1:table(myers_posteriors$variable) ~ variable, 
+                        subset=.(variable!="deviance")) 
+myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
+myers_pardist[,2] = exp(myers_pardist[,2]) # transform model parameters back first
+myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
+colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
+bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
+myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
+myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
 ```
@@
 
-```{r}
-
-  true_means <- sapply(x_grid, f, 0, p)
-  alt_means <- sapply(x_grid, alt$f, 0, alt$p)
-  est_means <- sapply(x_grid, est$f, 0, est$p)
-  par_bayes_means <- sapply(x_grid, f, 0, bayes_pars)
-
+```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
 models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
-                     MLE=est_means, Ricker.MLE=alt_means, 
-                     Parametric.Bayes = par_bayes_means)
+                     MLE=est_means, Ricker=ricker_means, 
+                     Allen = allen_means,
+                     Myers = myers_means)
 
@@
 
+model_names = c("GP", "True", "MLE", "Ricker", "Allen", "Myers")
+colorkey=cbPalette
+names(colorkey) = model_names 
+```
+
+```{r Figure1, dependson="assemble-models"}
 plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-    geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
+    geom_line(data=models, aes(x, value, col=method), lwd=1, alpha=0.8) + 
     geom_point(data=obs, aes(x,y), alpha=0.8) + 
@@
 
-```{r}
+```{r gp-opt, dependson="gp-output"}
 MaxT = 1000
 # uses expected values from GP, instead of integrating over posterior
-# matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)
+#matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)
 
@@
 
-Determine the optimal policy based on the true and MLE models
+Determine the optimal policy based on the allen and MLE models
 
-```{r}
+```{r mle-opt, dependson="mle-output"}
 matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
@@
 
-matrices_alt <- f_transition_matrix(alt$f, alt$p, x_grid, h_grid, alt$sigma_g)
-opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
 ```
 
-Determine the optimal policy based on parametric Bayesian model
+Determine the optimal policy based on Bayesian Allen model
 
-```{r}
-matrices_par_bayes <- parameter_uncertainty_SDP(f, p, x_grid, h_grid, pardist)
-opt_par_bayes <- value_iteration(matrices_par_bayes, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+```{r allen-opt, dependson="allen-output"}
+allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
+matrices_allen <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
+opt_allen <- value_iteration(matrices_allen, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+```
+
+Bayesian Ricker
+
+```{r ricker-opt, dependson="ricker-output"}
+ricker_f <- function(x, h, p) Ricker(x, h, unname(p[c(2, 1)])) # defined by pdgControl 
+matrices_ricker <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
+opt_ricker <- value_iteration(matrices_ricker, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+```
+
+Bayesian Myers model
+
+```{r myers-opt, dependson="myers-output"}
+myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
+matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
+myers_alt <- value_iteration(matrices_myers, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
 ```
@@
 
-```{r}
-OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker.MLE = opt_alt$D, Parametric.Bayes = opt_par_bayes$D)
+```{r assemble-opt, dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
+OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_ricker$D, Allen = opt_allen$D, Myers = myers_alt$D)
 colorkey=cbPalette
@@
 
-```{r}
+```{r Figure2, dependson="assemble-opt"}
 policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
@@
 
-```{r}
-
+```{r sims, dependson="assemble-opt"}
 sims <- lapply(OPT, function(D){
@@
 dt$method = factor(dt$method, ordered=TRUE, levels=names(OPT))
+```
 
+```{r Figure3, dependson="sims"}
 ggplot(dt) + 
@@
   scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
+```
+
+```{r profits, dependson="sims"}
+Profit <- dt[, sum(profit), by=c("reps", "method")]
+Profit[, mean(V1), by="method"]
+```
+
+```{r totalprofits, dependson="profits"}
+ggplot(Profit, aes(V1)) + geom_histogram() + 
+  facet_wrap(~method, scales = "free_y") + guides(legend.position = "none") + xlab("Total profit by replicate")
+```
+
+```{r}
+df <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable) 
+modes <- apply(df, 2, posterior.mode)
+allen_deviance <- modes[['deviance']]
+
+df <- acast(ricker_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable)
+modes <- apply(df, 2, posterior.mode)
+ricker_deviance <- modes[['deviance']]
+
+df <- acast(myers_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable)
+modes <- apply(df, 2, posterior.mode)
+myers_deviance <- modes[['deviance']]
+
+true_deviance <- 2*estf(c(p, sigma_g))
+mle_deviance <- 2*estf(c(est$p, est$sigma_g))
+
+c(allen = allen_deviance, ricker=ricker_deviance, myers=myers_deviance, true=true_deviance, mle=mle_deviance)
 

### 2013-05-23 R/gp_mcmc.R

@@ -9,7 +9,7 @@
 #' the squared exponential (also called radial basis or Gaussian, 
 #' though it is not this that gives Gaussian process it's name; 
 #' any covariance function would do) as the covariance function, 
-#' $\operatorname{cov}(X_i, X_j) = \exp\left(-\frac{(X_i - X_j)^2}{2 \ell ^ 2}\right)$
+#' \deqn{\operatorname{cov}(X_i, X_j) = \exp\left(-\frac{(X_i - X_j)^2}{2 \ell ^ 2}\right)}{}
 #' @return the MCMC output, as constructed by \link{metrop} from the mcmc package.
 #' @export
 #' @import mcmc  

### 2013-05-24 inst/examples/BUGS/parametric-vs-nonparametric.Rmd
parallel jags working now

@@
 
-Various parameters defining noise dynamics, grid, and policy costs.
-
 ```{r sdp-pars, dependson="stateeq"}
@@
 xT <- 0
-Xo <- K # observations start from
-x0 <- Xo # simulation under policy starts from
+Xo <-  allee+.5# observations start from
+x0 <- K # simulation under policy starts from
 Tobs <- 40
@@
   x <- numeric(Tobs)
-  x[1] <- 2
+  x[1] <- Xo
   nz <- 1
@@
 # Can give a list of starting values or a function that can create random values
+
 jags.inits <- function(){
-  list("K"=init_p["K"] * rlnorm(1,0,.5),
-       "logr0"=log(init_p["r0"]) * rlnorm(1,0,.5),
-       "logtheta"=log(init_p["theta"]) * rlnorm(1,0,.5), 
-       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
-       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
+  init_p = c(1, 10, 2)
+  names(init_p) = c("r0", "K", "theta")
+  list("K"=init_p["K"] * rlnorm(1,0,.1),
+       "logr0"=log(init_p["r0"]) * rlnorm(1,0,.1),
+       "logtheta"=log(init_p["theta"]) * rlnorm(1,0,.1), 
+       "stdQ"=sqrt(0.05) * rlnorm(1,0,.1),
+       "stdR"=sqrt(0.1) * rlnorm(1,0,.1),
        "x"=y,        
@@
 
-set.seed(12345)
+set.seed(1234)
 # parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
-allen_jags <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
-                n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin, 
-                model.file="allen.bugs", progress="none")
+allen_jags <- do.call(jags.parallel, list(data=jags.data, inits=jags.inits, 
+                                          jags.params, n.chains=n.chains, 
+                                          n.iter=n.iter, n.thin=n.thin, 
+                                          n.burnin=n.burnin, model.file="allen.bugs"))
 
 #Run again iteratively if we haven't met the Gelman-Rubin convergence criterion
-allen_jags <- autojags(allen_jags, n.update=10, n.iter=n.iter, 
-                       n.thin = n.thin, progress="none") 
+recompile(allen_jags) # required for parallel
+allen_jags <- do.call(autojags, list(object=allen_jags, n.update=10, 
+                                     n.iter=n.iter, n.thin = n.thin))
 ```
 
-#### Convergence diagnostics for parametric bayes
+#### Convergence diagnostics for Allen model
 

### 2013-05-24 inst/examples/BUGS/parametric-vs-nonparametric.Rmd
parallel jags

@@
 
-```{r allen_pars}
-init_p = c(2, K, 0.01)
-names(init_p) = c("r0", "K", "theta")
-```
-
 The actual model is defined in a `model.file` that contains an R function
@@
 logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
-logtheta_prior <- function(x) dunif(x, -6, 6)
+logtheta_prior <- function(x) dunif(x, logtheta_prior_p[1], logtheta_prior_p[2])
 par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
@@
 jags.inits <- function(){
-  init_p = c(1, 10, 2)
-  names(init_p) = c("r0", "K", "theta")
+  init_p = c(r0=1, K=10, theta=2)
   list("K"=init_p["K"] * rlnorm(1,0,.1),
-       "logr0"=log(init_p["r0"]) * rlnorm(1,0,.1),
-       "logtheta"=log(init_p["theta"]) * rlnorm(1,0,.1), 
+       "logr0"=log(rlnorm(1,0,.1) ),
+       "logtheta"=log(init_p["theta"] * rlnorm(1,0,.1)), 
        "stdQ"=sqrt(0.05) * rlnorm(1,0,.1),
@@
 jags.inits <- function(){
-  list("K"=init_p["K"] * rlnorm(1,0,.5),
-       "logr0"=log(init_p["r0"]) * rlnorm(1,0,.5),
+  list("K"=10 * rlnorm(1,0,.5),
+       "logr0"=log(1) * rlnorm(1,0,.5),
        "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
@@
 set.seed(12345) 
-ricker_jags <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
+ricker_jags <- do.call(jags.parallel, 
+                       list(data=jags.data, inits=jags.inits, 
+                            jags.params, n.chains=n.chains, 
                             n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
-                  model.file="ricker.bugs", progress.bar="none")
-ricker_jags <- autojags(ricker_jags, n.update=10, n.iter=n.iter, 
-                        n.thin = n.thin, progress.bar="none")
+                            model.file="ricker.bugs"))
+recompile(ricker_jags)
+ricker_jags <- do.call(autojags, 
+                       list(object=ricker_jags, n.update=10, n.iter=n.iter, 
+                            n.thin = n.thin, progress.bar="none"))
 ```
@@
 ```{r myers-mcmc, dependson="myers-model"}
-init_p = c(p[1], p[3], p[2])
-names(init_p) = c("r0", "theta", "K")
 jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
 jags.inits <- function(){
-  list("logr0"=log(init_p["r0"]) * rlnorm(1,0,.5), 
-       "logK"=log(init_p["K"]),
-       "logtheta" = log(init_p["theta"]) * rlnorm(1,0,.5),  
+  list
+  ("logr0"=log(1 * rlnorm(1,0,.5)), 
+       "logK"=log(10* rlnorm(1,0,.5)),
+       "logtheta" = log(5 * rlnorm(1,0,.5)),  
        "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
@@
 set.seed(12345)
-myers_jags <- jags(data=jags.data, inits=jags.inits, jags.params, 
+myers_jags <- do.call(jags.parallel, 
+                      list(data=jags.data, inits=jags.inits, jags.params, 
                            n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
-                  n.burnin=n.burnin, model.file="myers.bugs", progress.bar="none")
-myers_jags <- autojags(myers_jags, n.update=10, n.iter=n.iter, 
-                    n.thin = n.thin, progress.bar="none")
+                           n.burnin=n.burnin, model.file="myers.bugs")
+myers_jags <- do.call(autojags, 
+                      list(object=myers_jags, n.update=10, n.iter=n.iter, 
+                           n.thin = n.thin, progress.bar="none"))
 ```

### 2013-05-24 inst/examples/BUGS/process-noise-only.Rmd
Consider the case of process noise only in the parametric bayesian versions

--- 2013-05-24 inst/examples/BUGS/parametric-vs-nonparametric.Rmd
@@
                fig.width=6, fig.height=4, cache.path="myers/", verbose=TRUE)
-opts_knit$set(upload.fun = socialR::flickr.url)
+#opts_knit$set(upload.fun = socialR::flickr.url)
 theme_set(theme_bw(base_size=12))
@@
 n.chains <- 3
-n.iter <- 1e5
+n.iter <- 1e4
 n.burnin <- floor(10000)
@@
   theta <- exp(logtheta)
-  x[1] ~ dunif(0, 10)
+  y[1] ~ dunif(0, 10)
   for(t in 1:(N-1)){
-    mu[t] <- x[t] * exp(r0 * (1 - x[t]/K)* (x[t] - theta) / K )
-    x[t+1] ~ dnorm(mu[t], iQ) 
-  }
-  for(t in 1:(N)){
-    y[t] ~ dnorm(x[t], iR)
+    mu[t] <- y[t] * exp(r0 * (1 - y[t]/K)* (y[t] - theta) / K )
+    y[t+1] ~ dnorm(mu[t], iQ) 
   }
 }")
-writeLines(bugs.model, "allen.bugs")
+writeLines(bugs.model, "allen_process.bugs")
 ```
@@
 
-```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup")}
+```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results="hide"}
 jags.params=c("K","logr0","logtheta","stdQ", "stdR") # be sensible about the order here
-
-# Can give a list of starting values or a function that can create random values
-
 jags.inits <- function(){
@@
        "stdR"=sqrt(0.1) * rlnorm(1,0,.1),
-       "x"=y,        
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
@@
 # parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
-allen_jags <- do.call(jags.parallel, list(data=jags.data, inits=jags.inits, 
+allen_jags <- do.call(jags, list(data=jags.data, inits=jags.inits, 
                                       jags.params, n.chains=n.chains, 
                                       n.iter=n.iter, n.thin=n.thin, 
-                                          n.burnin=n.burnin, model.file="allen.bugs"))
+                                      n.burnin=n.burnin, model.file="allen_process.bugs"))
 
@@
   r0 <- exp(logr0)
-  x[1] ~ dunif(0, 10)
+  y[1] ~ dunif(0, 10)
   for(t in 1:(N-1)){
-    mu[t] <- x[t] * exp(r0 * (1 - x[t]/K) )
-    x[t+1] ~ dnorm(mu[t], iQ) 
-  }
-  for(t in 1:(N)){
-    y[t] ~ dnorm(x[t], iR)
+    mu[t] <- y[t] * exp(r0 * (1 - y[t]/K) )
+    y[t+1] ~ dnorm(mu[t], iQ) 
   }
 }")
-writeLines(bugs.model, "ricker.bugs")
+writeLines(bugs.model, "ricker_process.bugs")
 ```
@@
        "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
-       "x"=y,
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
@@
                             n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
-                            model.file="ricker.bugs"))
+                            model.file="ricker_process.bugs"))
 recompile(ricker_jags)
@@
 
-  x[1] ~ dunif(0, 10)
+  y[1] ~ dunif(0, 10)
 
   for(t in 1:(N-1)){
-    mu[t] <- r0 * pow(abs(x[t]), theta) / (1 + pow(abs(x[t]), theta) / K)
-    x[t+1] ~ dnorm(mu[t], iQ) 
-  }
-
-  for(t in 1:(N)){
-    y[t] ~ dnorm(x[t], iR)
+    mu[t] <- r0 * pow(abs(y[t]), theta) / (1 + pow(abs(y[t]), theta) / K)
+    y[t+1] ~ dnorm(mu[t], iQ) 
   }
 }")
-writeLines(bugs.model, "myers.bugs")
+writeLines(bugs.model, "myers_process.bugs")
 ```
@@
 jags.inits <- function(){
-  list
-  ("logr0"=log(1 * rlnorm(1,0,.5)), 
+  list("logr0"=log(1 * rlnorm(1,0,.5)), 
        "logK"=log(10* rlnorm(1,0,.5)),
@@
                            n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
-                           n.burnin=n.burnin, model.file="myers.bugs")
+                           n.burnin=n.burnin, model.file="myers_process.bugs")
 myers_jags <- do.call(autojags, 
-                      list(object=myers_jags, n.update=10, n.iter=n.iter, 
+                      list(myers_jags, n.update=10, n.iter=n.iter, 
                            n.thin = n.thin, progress.bar="none"))

### 2013-05-28 inst/examples/adp-intro.Rmd
more notes from reading on ADP

@@
 
-```{r  setup, echo=FALSE, cache=FALSE}
+```{r  setup, echo=FALSE, cache=FALSE, results="hide", message=FALSE, warning=FALSE}
 require(multipleuncertainty)
+require(pdgControl)
 require(reshape2)
@@
 opts_chunk$set(dev.args=list(bg="transparent"), 
-               cache=TRUE, cache.path = "adp/",
-               comment=NA, message=FALSE, warning=FALSE,
-               tidy=FALSE)
+               cache=TRUE, comment=NA, message=FALSE, 
+               warning=FALSE, tidy=FALSE)
+
+opts_chunk$set(fig.path = paste("figure/", 
+                                gsub(".Rmd", "", 
+                                     knitr:::knit_concord$get('infile')),
+                                "-", sep=""), 
+               cache.path = paste(gsub(".Rmd", "", 
+                                       knitr:::knit_concord$get('infile') ), 
+                                "/", sep="")
+              )
+require(ggplot2)
+theme_notebook <- 
+  theme_grey() + 
+  theme(plot.background = element_rect(fill = "transparent", colour = NA),
+        panel.grid.minor = element_line(colour = "transparent"))
+theme_set(theme_notebook)
 ```
 
-### Alogorithm 1
+_Based on Powell 2006, page 97.  I try to conform to that notation throughout_
 
-Based on Powell 2006, page 97.
+## Setup
+
+First we define the Beverton-Holt stock-recruitment relationship as a function of stock size `x`, harvest `h` and parameters `p`
 
@@
 }
-pars <- c(1.5, 0.05)
-K <- (pars[1] - 1)/pars[2]
+p <- c(1.5, 0.5)
+K <- (p[1] - 1)/p[2]
+sigma_g <- 0.2
 ```
@@
 ```{r}
-S <- seq(0, 1, length=10) 
+S <- seq(0, 1, length=11) 
 ```
@@
 We also need a value function on the state space, $C_t(S_t)$. 
-For simplicity, we set the price of harvest at unity and the cost of harvesting at zero, so that $C_t(S_t, x_t) = \min(x_t, S_t)$.
+For simplicity, we set the price of harvest at unity and 
+the cost of harvesting at zero, so that $C_t(S_t, x_t) = \min(x_t, S_t)$.  
 ($C_t$ is sometimes denoted $\mathbb{\Pi}$).  
@@
 
-```{r}
+```{r action-space}
 chi <- S
@@
 
-```{r}
+```{r problem-size}
 T <- 10
@@
 
-The approximate dynamic programming algorithm will perform a finite number $N$ = `r N` iterations over a window of time $T$ =`r T` in our example.  The algorithm can be described as follows:
+The approximate dynamic programming algorithm will perform a finite number $N$ = `r N` iterations over a window of time $T$ =`r T` in our example.  The algorithm can then be described as follows:
+
+## Algorithm (1)
 
 - **Step 0**
-  - Initialize some $\tilde V_t^0(S_t)$ for all states $S_t$ 
-  where the superscripts denote iterations in the forward approximation.  As we know absolutely nothing yet to base our initial guess on, we just arbitrarily set this to zero.
+
+  - Initialize some value $\tilde V_t^0(S_t)$ for all states $S_t$, 
+  where the superscripts denote iterations in the forward approximation.  
+  As we know absolutely nothing yet to base our initial guess on, 
+  we just arbitrarily set this to zero.
 
@@
   - Choose some initial state $S_0^1$
-  We start at some initial state for $n = 1$ (superscript) and $t = 0$ (subscript). The choice of initial condition may come from the problem itself, otherwise we choose something arbitrarily.
+  We start at some initial state for $n = 1$ (superscript) and $t = 0$ (subscript).
+  The choice of initial condition may come from the problem itself, 
+  otherwise we choose something arbitrarily.
 
@@
 
-- **Step 1**: Choose a sample path, $\omega^n$
+- **Step 1**: Choose a sample path, $\omega^n$ (a vector of random draws)
 
@@
 sigma <- 0.2
-Z <- rlnorm(T, 0, sigma)
+omega_n <- rlnorm(T, 0, sigma)
 ```
@@
 
-Let's start with $t=0$, $n=1$ and fix an $x_0$ from the set of $\chi$ (allowing the action space to be the same in each period, we can omit the subscript on $\chi$) to get started.
-We first compute $C(S_0, x_0)$.
-$S_0 = S_0^1$ which we fixed in step 0b arbitrarily at `r S_0`.
-Taking $x_0$ as the smallest harvest, $\min(\chi)$ = `r min(chi)` and evaluating $C(S_0,X_0) = \min(S_0, X_0)$ gives us `r min(S_0, min(chi))`, rather trivially.
+That is, choose action $x_t$ that maximizes the value of the next step.
 
-The next terms depend on the value $\tilde V^0_1(s^{\prime})$ for all $s^{\prime} \in S$, which we have no idea about.  Fortunately we have assumed a value for each of these in step 0a.
+Let's start with $t=0$, $n=1$ and fix an $x_0$ from the set of $\chi$ 
+(allowing the action space to be the same in each period, we can omit
+the subscript on $\chi$) to get started.  We first compute $C(S_0, x_0)$.
 
-We must also come up with some values for the probability $\mathbb{P}(s^{\prime} | S_1^0, x_1)$ for each state, given our current state $S_1^0$ and considered action $x_1$.
-This is more straight forward, since it is determined by our one-step transition function (without simulation - recall that the single step transition is given exactly).
+$S_0 = S_0^1$ which we fixed in step **0b** arbitrarily at `r S_0`.
 
-To do so, we evaluate the argument for each value in our action space, $x_t \in \chi_t$,
+The profits/costs $C(S_t, x_t)$ are the value derived by action (harvest) $x_t$ 
+at state (stock) $S_t$.  Assuming a fixed price and no costs to harvesting,
+this is just whichever number is smaller (since we cannot harvest more than 
+the available stock,
+
+```{r}
+C <- function(S, X) pmin(S, X)
+```
+
+(where we have used R's vectorized form of the min function).
+
+This forward dynamic programming will still rely on the one-step transition
+matrix, $\mathbb{P}$.
+
+Let's get the trasition matrices for this problem, assuming log-normal noise,
+
+```{r transition-matrix}
+sdp_matrix <- determine_SDP_matrix(f, p, x_grid=S, h_grid=chi, sigma_g)
+```
+
+Which is a list of matrices, one for each harvest (action) $x_t$.
+
+Then we want to consider a fixed $S_t^n$ and fixed $x_t$, and take the sum of 
+$\mathbb{P}(s^{\prime} | S_t^n, x_t)$ over the $s^{\prime}$, which means we want
+the $x_t$ element from the list, and then we need sum over the distribution of future 
+states given the current state $S_t^n$, e.g. a row of the matrix, e.g. `sdp_matrix[[x]][s,]`, 
+which we (vector) multiply by $V_{t+1}^{n-1}(s^{\prime})$.
+
+This value $V$ is of course unknown, other than our initial random guess $V_{t}^0$.  
+As we step through the iterations $V_t^1$, $V_t^2$, $V_t^3$, etc., this should convgerge to
+something meaningful.
+
+Note that the index along $S$ corresponding to $S_t^n$ is given by
+
+```{r}
+s <- which.min(abs(S-S_0))
+```
+
+So our maximization across $x$ just involves:
+
+```{r}
+values <- 
+  sapply(1:length(chi), function(x)
+    C(S[s], chi[x]) + sdp_matrix[[x]][s,] %*% V
+)
+
+max_x <- which.max(values)
+v_hat <- max(values)
+```
+
+Trivially, this is just the harvest level that maximizes $C$ so far (which
+is just harvesting the $S_0$, since $\bar V^0_t$ begins at zero:
+
+```{r}
+chi[max_x]
+```
+
+- step **2b** We can now update our $\bar V^0_t$ to get $\bar V^1_t$, using the rule:
+
+$$V_t^n(S_t) = \begin{cases} 
+\hat v_t^n & S_t = S_t^n \\
+\bar V_t^{n-1}(S_t) & \textrm{otherwise} 
+\end{cases}$$
+
+e.g. use our maximum value for the case of the state we just considered $S_t = S_t^n$, otherwise leave $V_t$ unchanged.  Our new $V$ is thus:
+
+```{r}
+V[s] = v_hat
+```
+
+- step **2c**  Compute $S^n_{t+1} = S^M(S_t^n, x^n_t, W_{t+1}(\omega^n))$
+
+We compute the next state using our `max_x` for $x^n_t$, our random samples 
+and the transition function...
+
+```{r}
+S_1 <- omega_n[1] * f(S_0, chi[x_nt], p)
+```
+
+- **Step 3** Let $n = n+1$. if $n < N$, go to step 1
+
+assembling the solution...
+
+## As a recursive algorithm
+
+Things we want to keep track of:
+
+* $V^{n+1}_t$
+
+We have time, iterations, and state/action space to keep track over.  eek.
+
+```{r}
+N <- 5 # iterations
+M <- 10 # gridsize 
+Tmax <- 5 # Time horizon
+
+gamma <- 0.95 # Discount
+# f 
+# p
+
+chi <- seq(0, 1, length.out = M)
+S <- seq(0, 1, length.out = M)
+
+sdp_matrix <- determine_SDP_matrix(f, p, x_grid=S, h_grid=chi, sigma_g)
+
+V <- matrix(rep(chi, Tmax), nrow=M) # matrix(0, M, Tmax)
+X <- matrix(0, M, Tmax)
+C <- function(S, X) pmin(S, X)
+S_0 <- 0.5
+
+for(n in 1:N){
+
+  omega_n <- rlnorm(Tmax, 0, sigma_g)
+  S_current  <- S_0
+
+  for(t in 1:(Tmax-1)){
+
+    # index of the state we're considering
+    s <- which.min(abs(S-S_current)) 
+    # Find the action maximizing the value
+    values <- sapply(1:length(chi), function(x)
+      C(S[s], chi[x]) + gamma * sdp_matrix[[x]][s,] %*% V[,t])
+    hat <-  c(x_nt = which.max(values), v_nt = max(values))
+
+    # Update value V 
+    V[hat["x_nt"], t] <- hat["v_nt"]
+
+    # Advance the state in time along random path  
+    S_current <- omega_n[t] * f(S_current, chi[x_nt], p)
+  }
+
+}
+```
+
+### Problems arising from the discretization
+
+Note that after the first iteration, $n=1$, the value matrix $V$ is no longer all zeros.
+There is a single state, $S = S_0 =$ `r S_0`, at which we have value.
+That value is lost if we set harvest $x$ too high, since we know we will not then end up in that state -- from whence comes the incentive to consider future value.
+Unfortunately, the value exists only if we hit that state exactly -- all other states are assumed to have zero value still.
+
+### Additional problems
+
+We no longer have the loop-over-all-states problem, but we face several new or remaining issues:
+
+1.
+We still require the use of the one-step transition matrix, with the equally troublesome sum over all states $\sum_{s^{\prime}\in S} \mathbb{P}(s^{\prime} | S_t^n, x_t)$.
+We will fix this by approximating the transitions in step 2b using random draws as well.
+
+2. We only update the values of states we visit.  We still need a way to estimate the value of states we have not visited.
+
+3.
+Worse, we might not visit states that seem bad relative to states we have visited.
+This is particularly atrocious in this example.
+Since we initialize the value of all states at 0, the algorithm prefers to harvest all stock from the current state rather than risk a transition into a state starting at 0.
+There is no convergence guarentee that we will ever escape this cycle of avoiding states we have not seen.
+We can alter the initial guess of the value of course, and we could alter the starting condition to better explore.
+
+### Using Non-stochastic transition information only, step **2b** can be written as:
+
+Taking $x_0$ as the smallest harvest, $\min(\chi)$ = `r min(chi)` and 
+evaluating $C(S_0,X_0) = \min(S_0, X_0)$ gives us `r min(S_0, min(chi))`,
+rather trivially.  
+The next terms depend on the value $\tilde V^0_1(s^{\prime})$ for 
+all $s^{\prime} \in S$, which we have no idea about.  Fortunately we have
+assumed a value for each of these in step 0a.
+
+We must also come up with some values for the probability
+$\mathbb{P}(s^{\prime} | S_1^0, x_1)$ for each state, given our current
+state $S_1^0$ and considered action $x_1$.  This is more straight forward,
+since it is determined by our one-step transition function (without 
+simulation - recall that the single step transition is given exactly).
+
+To do so, we evaluate the argument for each value in our action 
+space, $x_t \in \chi_t$,
 
@@
 C <- function(S, X) pmin(S, X)
-
-arg <- sapply(chi, function(x) C(s, x) + f(S, x, pars) %*% V)
-c(x_nt = which.max(arg), v_nt = max(arg))
-
-## Update V
-
+arg <- sapply(chi, function(x) C(s, x) + f(S, x, p) %*% V)
+x_nt = which.max(arg)
+v_nt = max(arg)
+V[x_nt] = v_nt
 ```

### 2013-05-29 inst/examples/BUGS/parametric-vs-nonparametric.Rmd
tuning on farm

@@
 setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
-library(knitcitations)
+#library(knitcitations)
 library(ggplot2) 
 library(nonparametricbayes) 
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA,
-               fig.width=6, fig.height=4, cache.path="myers/", verbose=TRUE)
-opts_knit$set(upload.fun = socialR::flickr.url)
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA, fig.path="figure/parametric-"
+               fig.width=6, fig.height=4, cache.path="parametric-vs-nonparametric/", verbose=TRUE)
+#opts_knit$set(upload.fun = socialR::flickr.url)
 theme_set(theme_bw(base_size=12))
@@
 jags.inits <- function(){
-  list
-  ("logr0"=log(1 * rlnorm(1,0,.5)), 
+  list("logr0"=log(rlnorm(1,0,.5)), 
        "logK"=log(10* rlnorm(1,0,.5)),

### 2013-05-29 inst/examples/BUGS/process-noise-only.Rmd
tuning on farm

@@
 setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
-library(knitcitations)
+#library(knitcitations)
 library(ggplot2) 
@@
 opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA,
-               fig.width=6, fig.height=4, cache.path="myers/", verbose=TRUE)
+               fig.width=6, fig.height=4, cache.path="process/", verbose=TRUE)
 #opts_knit$set(upload.fun = socialR::flickr.url)
+
 theme_set(theme_bw(base_size=12))
@@
 
-Uses the model derived in `r citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
+Uses the model derived in ` citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
 
@@
 n.chains <- 3
-n.iter <- 1e4
+n.iter <- 1e5
 n.burnin <- floor(10000)
@@
 jags.inits <- function(){
-  list("logr0"=log(1 * rlnorm(1,0,.5)), 
-       "logK"=log(10* rlnorm(1,0,.5)),
-       "logtheta" = log(5 * rlnorm(1,0,.5)),  
-       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
-       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
+  list("logr0"=log(1 * rlnorm(1,0,.2)), 
+       "logK"=log(10* rlnorm(1,0,.2)),
+       "logtheta" = log(5 * rlnorm(1,0,.2)),  
+       "stdQ"=sqrt(0.05) * rlnorm(1,0,.2),
+       "stdR"=sqrt(0.1) * rlnorm(1,0,.2),
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
@@
 set.seed(12345)
-myers_jags <- do.call(jags.parallel, 
+myers_jags <- do.call(jags, 
                       list(data=jags.data, inits=jags.inits, jags.params, 
                            n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
-                           n.burnin=n.burnin, model.file="myers_process.bugs")
+                           n.burnin=n.burnin, model.file="myers_process.bugs"))
 myers_jags <- do.call(autojags, 

### 2013-05-29 inst/examples/adp-intro.Rmd

@@
 
-$$V_t(S_t) = \max_{x_t \in \chi_t} \left(C(S_t, x_t) + \gamma \sum_{s^{\prime} \in mathcal{S}} \mathbb{P}(s^{\prime} | S_t^n, x_t) V_{t+1}^{n-1} s^{\prime} \right)$$
+$$V_t(S_t) = \max_{x_t \in \chi_t} \left(C(S_t, x_t) + \gamma \sum_{s^{\prime} \in \mathcal{S}} \mathbb{P}(s^{\prime} | S_t^n, x_t) V_{t+1}^{n-1} s^{\prime} \right)$$
 
@@
 ```{r}
-N <- 5 # iterations
-M <- 10 # gridsize 
+N <- 50000 # iterations
+M <- 20 # gridsize 
 Tmax <- 5 # Time horizon
@@
 
+sigma_g <- 0.5 # larger variation in random draws helps
 chi <- seq(0, 1, length.out = M)
@@
 
-V <- matrix(rep(chi, Tmax), nrow=M) # matrix(0, M, Tmax)
-X <- matrix(0, M, Tmax)
+V <- matrix(0, M, Tmax) # matrix(rep(chi, Tmax), nrow=M) # 
+V[,1] <- chi   # fails to explore if it doesn't have at least some non-zero values
+
 C <- function(S, X) pmin(S, X)
 S_0 <- 0.5 
+alpha <- 1 # learning rate
 
@@
   omega_n <- rlnorm(Tmax, 0, sigma_g)
-  S_current  <- S_0
+  S_current <- S_0 #runif(1,0,1) # explores faster when this is random
 
-  for(t in 1:(Tmax-1)){
-
+  for(t in 1:Tmax){
     # index of the state we're considering
     s <- which.min(abs(S-S_current))
+
     # Find the action maximizing the value
@@
 
-    # Update value V 
-    V[hat["x_nt"], t] <- hat["v_nt"]
+    # Update value V as mixture of new value and previous value
+    V[hat["x_nt"], t] <- (1 - alpha) * V[hat["x_nt"],t] + alpha * hat["v_nt"]
 
     # Advance the state in time along random path  
-    S_current <- omega_n[t] * f(S_current, chi[x_nt], p)
-  }
+    S_current <- omega_n[t] * f(S_current, chi[hat["x_nt"]], p)
 
   }
+}
+```
+
+for comparison: the SDP solution
+
+```{r}
+opt <- find_dp_optim(sdp_matrix, S, chi, 70, 0.5, C, 1-gamma, reward=0)
+opt$V
 ```

### 2013-05-29 inst/examples/BUGS/process-noise-only.Rmd
local edits

@@
 #opts_knit$set(upload.fun = socialR::flickr.url)
-
 theme_set(theme_bw(base_size=12))
@@
 xT <- 0
-Xo <-  allee+.5# observations start from
+Xo <-  allee + 0.5 # observations start from
 x0 <- K # simulation under policy starts from
@@
 par <- c(p[1]+abs(rnorm(1,0,1)), 
-         p[2]+rnorm(1,0,.1), 
+         p[2]+rnorm(1,0, .1), 
          p[3]+rnorm(1,0, .1), 
-         sigma_g + abs(rnorm(1,0,.1)))
+         sigma_g + abs(rnorm(1,0, .1)))
 o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
@@
 ```{r jags-setup}
-y <- obs$x[-1] 
-N <- length(y);
-jags.data <- list("N","y")
-n.chains <- 3
-n.iter <- 1e5
+y <- x 
+N <- length(x);
+jags.data <- list("N"=N,"y"=y)
+n.chains <- 4
+n.iter <- 1e6
 n.burnin <- floor(10000)
@@
 stdQ_prior_p <- c(0.0001, 100)
-stdR_prior_p <- c(0.0001, 100)
 stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
-stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
 ```
@@
   logtheta ~ dunif(%s, %s)
-  stdQ ~ dunif(%s, %s)
-  stdR ~ dunif(%s, %s)", 
+  stdQ ~ dunif(%s, %s)", 
   K_prior_p[1], K_prior_p[2],
@@
   logtheta_prior_p[1], logtheta_prior_p[2],
-  stdQ_prior_p[1], stdQ_prior_p[2],
-  stdR_prior_p[1], stdR_prior_p[2]),
+  stdQ_prior_p[1], stdQ_prior_p[2]),
 
@@
   iQ <- 1 / (stdQ * stdQ);
-  iR <- 1 / (stdR * stdR);
   r0 <- exp(logr0)
@@
                     logr0 = logr0_prior, logtheta = logtheta_prior,
-                    stdQ = stdQ_prior, stdR = stdR_prior)
+                    stdQ = stdQ_prior)
 ```
@@
 ```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results="hide"}
-jags.params=c("K","logr0","logtheta","stdQ", "stdR") # be sensible about the order here
+jags.params=c("K","logr0","logtheta","stdQ") # be sensible about the order here
 jags.inits <- function(){
-  init_p = c(r0=1, K=10, theta=2)
-  list("K"=init_p["K"] * rlnorm(1,0,.1),
-       "logr0"=log(rlnorm(1,0,.1) ),
-       "logtheta"=log(init_p["theta"] * rlnorm(1,0,.1)), 
-       "stdQ"=sqrt(0.05) * rlnorm(1,0,.1),
-       "stdR"=sqrt(0.1) * rlnorm(1,0,.1),
+  list("K"= 10 * rlnorm(1,0, 0.1),
+       "logr0"=log( rlnorm(1,0, 0.5) ),
+       "logtheta"=log(  2 * rlnorm(1,0, 0.1) ), 
+       "stdQ"= abs(sigma_g * rlnorm(1,0, 0.1)),
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
@@
                                       n.iter=n.iter, n.thin=n.thin, 
-                                      n.burnin=n.burnin, model.file="allen_process.bugs"))
+                                      n.burnin=n.burnin, 
+                                      model.file="allen_process.bugs"))
 
@@
 ```{r allen-traces, dependson="allen-mcmc"}
-tmp <- lapply(as.mcmc(allen_jags), as.matrix) # strip classes the hard way...
+tmp <- lapply(as.mcmc(allen_jags), as.matrix) # strip classes to melt
 allen_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
@@
   logr0    ~ dunif(%s, %s)
-  stdQ ~ dunif(%s, %s)
-  stdR ~ dunif(%s, %s)", 
+  stdQ ~ dunif(%s, %s)", 
   K_prior_p[1], K_prior_p[2],
   logr0_prior_p[1], logr0_prior_p[2],
-  stdQ_prior_p[1], stdQ_prior_p[2],
-  stdR_prior_p[1], stdR_prior_p[2]),
+  stdQ_prior_p[1], stdQ_prior_p[2]),
 
@@
 par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
-                   logr0 = logr0_prior, stdQ = stdQ_prior, stdR = stdR_prior)
+                   logr0 = logr0_prior, stdQ = stdQ_prior)
 ```
@@
 # Uniform priors on standard deviation terms
-jags.params=c("K","logr0", "stdQ", "stdR")
+jags.params=c("K","logr0", "stdQ")
 jags.inits <- function(){
@@
        "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
-       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
@@
   logK    ~ dunif(%s, %s)
-  stdQ ~ dunif(%s, %s)
-  stdR ~ dunif(%s, %s)", 
+  stdQ ~ dunif(%s, %s)", 
   logr0_prior_p[1], logr0_prior_p[2],
@@
   logK_prior_p[1], logK_prior_p[2],
-  stdQ_prior_p[1], stdQ_prior_p[2],
-  stdR_prior_p[1], stdR_prior_p[2]),
+  stdQ_prior_p[1], stdQ_prior_p[2]),
 
@@
   iQ <- 1 / (stdQ * stdQ);
-  iR <- 1 / (stdR * stdR);
-
   r0 <- exp(logr0)
@@
   y[1] ~ dunif(0, 10)
-
   for(t in 1:(N-1)){
@@
                     logr0 = logr_prior, logtheta = logtheta_prior, 
-                    stdQ = stdQ_prior, stdR = stdR_prior)
+                    stdQ = stdQ_prior)
 
@@
 ```{r myers-mcmc, dependson="myers-model"}
-jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
+jags.params=c("logr0", "logtheta", "logK", "stdQ")
 jags.inits <- function(){
-  list("logr0"=log(1 * rlnorm(1,0,.2)), 
-       "logK"=log(10* rlnorm(1,0,.2)),
-       "logtheta" = log(5 * rlnorm(1,0,.2)),  
-       "stdQ"=sqrt(0.05) * rlnorm(1,0,.2),
-       "stdR"=sqrt(0.1) * rlnorm(1,0,.2),
+  list("logr0"=log(rlnorm(1,0,.1)), 
+       "logK"=log(7 * rlnorm(1,0,.1)),
+       "logtheta" = log(2 * rlnorm(1,0,.1)),  
+       "stdQ"=sqrt(0.5) * rlnorm(1,0,.1),
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
@@
 myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
-colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
+colnames(myers_pardist) = c("K", "r0", "theta", "stdQ")
 bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates

### 2013-05-29 inst/examples/BUGS/parametric-vs-nonparametric.Rmd

@@
 N <- length(y);
-jags.data <- list("N","y")
-n.chains <- 3
-n.iter <- 1e5
+jags.data <- list("N"=N,"y"=y)
+n.chains <- 6
+n.iter <- 5e5
 n.burnin <- floor(10000)
 n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
+n.update <- 10
 ```
@@
 stdQ_prior_p <- c(0.0001, 100)
-stdR_prior_p <- c(0.0001, 100)
+stdR_prior_p <- c(1e-6, .1)
 stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
@@
 jags.inits <- function(){
-  init_p = c(r0=1, K=10, theta=2)
-  list("K"=init_p["K"] * rlnorm(1,0,.1),
-       "logr0"=log(rlnorm(1,0,.1) ),
-       "logtheta"=log(init_p["theta"] * rlnorm(1,0,.1)), 
-       "stdQ"=sqrt(0.05) * rlnorm(1,0,.1),
-       "stdR"=sqrt(0.1) * rlnorm(1,0,.1),
+  list("K"= 10 * rlnorm(1,0,.1),
+       "logr0"=log(rlnorm(1,0,.5) ),
+       "logtheta"=log(2 * rlnorm(1,0,.1)), 
+       "stdQ"= 0.2 * rlnorm(1,0,.1),
+       "stdR"= 1e-5 * rlnorm(1,0,.1),
        "x"=y,        
@@
 recompile(allen_jags) # required for parallel
-allen_jags <- do.call(autojags, list(object=allen_jags, n.update=10, 
+allen_jags <- do.call(autojags, list(object=allen_jags, n.update=n.update, 
                                      n.iter=n.iter, n.thin = n.thin))
@@
 jags.inits <- function(){
-  list("K"=10 * rlnorm(1,0,.5),
-       "logr0"=log(1) * rlnorm(1,0,.5),
-       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
-       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
+  list("K"= 10 * rlnorm(1,0,.5),
+       "logr0" = log( rlnorm(1,0,.5) ),
+       "stdQ"= 0.2 * rlnorm(1,0,.2),
+       "stdR"= 1e-5 * rlnorm(1,0,.1),
        "x"=y,
@@
 ricker_jags <- do.call(autojags, 
-                       list(object=ricker_jags, n.update=10, n.iter=n.iter, 
+                       list(object=ricker_jags, n.update=n.update, n.iter=n.iter, 
                             n.thin = n.thin, progress.bar="none"))
@@
 jags.inits <- function(){
-  list("logr0"=log(rlnorm(1,0,.5)), 
-       "logK"=log(10* rlnorm(1,0,.5)),
-       "logtheta" = log(5 * rlnorm(1,0,.5)),  
-       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
-       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
+  list("logr0" = log(rlnorm(1,0,.5)), 
+       "logK" = log(10 * rlnorm(1,0,.2)),
+       "logtheta" = log( 2 * rlnorm(1,0,.2)),  
+       "stdQ"= 0.2 * rlnorm(1,0,.1),
+       "stdR"= 1e-5 * rlnorm(1,0,.1),
        "x"=y,
@@
 myers_jags <- do.call(autojags, 
-                      list(object=myers_jags, n.update=10, n.iter=n.iter, 
+                      list(object=myers_jags, n.update=n.update, n.iter=n.iter, 
                            n.thin = n.thin, progress.bar="none"))

### 2013-05-29 inst/examples/BUGS/process-noise-only.Rmd
trouble with MCMC convergence for process-noise-only
Now with longer runs and better posterior estimator. Set for run on zero.

@@
 ```{r}
+require(modeest)
 posterior.mode <- function(x) {
-  ux <- unique(x)
-  ux[which.max(tabulate(match(x, ux)))]
+ # ux <- unique(x)
+ #  ux[which.max(tabulate(match(x, ux)))]
+ mlv(x, method="shorth")$M
 }
@@
 
-#Run again iteratively if we haven't met the Gelman-Rubin convergence criterion
+# Run again iteratively if we haven't met the Gelman-Rubin convergence criterion
 recompile(allen_jags) # required for parallel
@@
 # pardist <- acast(A[-4], index ~ variable, subset=.(variable != "deviance") )
-pardist <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable, subset=.(variable!="deviance")) 
-pardist[,2] = exp(pardist[,2]) # transform model parameters back first
-pardist[,3] = exp(pardist[,3])
+pardist <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable)[1] ~ variable) 
+pardist[,"logr0"] = exp(pardist[,"logr0"]) # transform model parameters back first
+pardist[,"logtheta"] = exp(pardist[,"logtheta"])
+
+colnames(pardist)[colnames(pardist)=="logtheta"] = "theta"
+colnames(pardist)[colnames(pardist)=="logr0"] = "r0"
+
 bayes_coef <- apply(pardist,2, posterior.mode) 
-bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3])) # parameters formatted for f
+bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"], bayes_coef["theta"])) # parameters formatted for f
 allen_means <- sapply(x_grid, f, 0, bayes_pars)

### 2013-05-29 inst/examples/adp-intro.Rmd
local edits
updated adp-intro

@@
 ```{r}
-N <- 50000 # iterations
+N <- 5000 # iterations
 M <- 20 # gridsize 
@@
 
-V <- matrix(0, M, Tmax) # matrix(rep(chi, Tmax), nrow=M) # 
-V[,1] <- chi   # fails to explore if it doesn't have at least some non-zero values
+V <- matrix(1, M, Tmax)  # A* strategy
+# Fails to explore at matrix(0, M, Tmax)
+# consider: # V <- matrix(rep(chi, Tmax), nrow=M) # 
+# V[,1] <- chi   # fails to explore if it doesn't have at least some non-zero values
 

### 2013-05-30 inst/examples/BUGS/allen.Rmd
simplified run, needs to correct handling of data structures

@@
 ```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
-setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
+source("~/.knitr_defaults.R")
+#opts_knit$set(upload.fun = socialR::flickr.url)
 library(knitcitations)
-library(ggplot2) 
 library(nonparametricbayes) 
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA,
-               fig.width=6, fig.height=4, cache.path="allen/")
-opts_knit$set(upload.fun = socialR::flickr.url)
-theme_set(theme_bw(base_size=12))
-theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
-             plot.background = element_rect(fill = "transparent", colour = NA))
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
-               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 ```
@@
 f <- RickerAllee
-p <- c(2, 8, 5)
-K <- p[2]
-allee <- p[3]
+p <- c(1, 10, 5)
+K <- 10  # approx, a li'l' less
+allee <- 5 # approx, a li'l' less
 ```
@@
   x <- numeric(Tobs)
-  x[1] <- 6
+  x[1] <- Xo
   nz <- 1
@@
 jags.data <- list("N","y")
-n.chains <- 1
-n.iter <- 40000
+n.chains <- 2
+n.iter <- 1e5
 n.burnin <- floor(10000)
 n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
+n.update <- 10
 ```
@@
 
-```{r allen_pars}
-# a bit unfair to start with the correct values, but anyhow...
-init_p = p
-names(init_p) = c("r0", "K", "theta")
-```
-
 The actual model is defined in a `model.file` that contains an R function
@@
 
-```{r allen-priors}
+```{r allen-priors, dependson="common-priors"}
 K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
 logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
-logtheta_prior <- function(x) dunif(x, -6, 6)
+logtheta_prior <- function(x) dunif(x, logtheta_prior_p[1], logtheta_prior_p[2])
 par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
@@
 
-```{r allen-mcmc}
-jags.params=c("K","logr0","logtheta","stdQ", "stdR")
+```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results="hide"}
+jags.params=c("K","logr0","logtheta","stdQ", "stdR") # be sensible about the order here
+
+# Can give a list of starting values or a function that can create random values
+
 jags.inits <- function(){
-  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), 
-       "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,
+  list("K"= 10 * rlnorm(1,0,.1),
+       "logr0"=log(rlnorm(1,0,.5) ),
+       "logtheta"=log(2 * rlnorm(1,0,.1)), 
+       "stdQ"= 0.2 * rlnorm(1,0,.1),
+       "stdR"= 1e-5 * rlnorm(1,0,.1),
+       "x"=y,        
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
@@
 
-set.seed(12345)
-jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
-                n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin, 
-                model.file="allen.bugs", progress.bar="none")
+set.seed(1234)
+# parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
+allen_jags <- do.call(jags.parallel, list(data=jags.data, inits=jags.inits, 
+                                          jags.params, n.chains=n.chains, 
+                                          n.iter=n.iter, n.thin=n.thin, 
+                                          n.burnin=n.burnin, model.file="allen.bugs"))
+
 ```
 
-#### Convergence diagnostics for parametric bayes
+#### Convergence diagnostics for Allen model
 
-```{r allen-traces}
-jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-allen_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
-ggplot(allen_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
+R notes: this strips classes from the `mcmc.list` object (so that we have list of matrices; objects that `reshape2::melt` can handle intelligently), and then combines chains into one array.
+In this array each parameter is given its value at each sample from the posterior (index) for each chain.
+
+```{r allen-traces, dependson="allen-mcmc"}
+tmp <- lapply(as.mcmc(allen_jags), as.matrix) # strip classes the hard way...
+allen_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
+names(allen_posteriors) = c("index", "variable", "value", "chain")
+ggplot(allen_posteriors) + geom_line(aes(index, value)) + 
+  facet_wrap(~ variable, scale="free", ncol=1)
 ```
 
-```{r allen-posteriors}
+```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
 allen_priors <- ddply(allen_posteriors, "variable", function(dd){
@@
 
-```{r allen-output}
-pardist <- acast(allen_posteriors, index ~ variable, subset=.(variable != "deviance") )
+```{r allen-output, dependson="allen-traces"}
+pardist <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable, subset=.(variable!="deviance")) 
 pardist[,2] = exp(pardist[,2]) # transform model parameters back first
@@
 
-```{r ricker-mcmc}
+```{r ricker-mcmc, dependson="ricker-model", results="hide"}
 # Uniform priors on standard deviation terms
@@
 jags.inits <- function(){
-  list("K"=init_p["K"], "logr0"=log(init_p["r0"]), 
-       "stdQ"=sqrt(0.05), "stdR"=sqrt(0.1), "x"=y, 
+  list("K"= 10 * rlnorm(1,0,.5),
+       "logr0" = log( rlnorm(1,0,.5) ),
+       "stdQ"= 0.2 * rlnorm(1,0,.2),
+       "stdR"= 1e-5 * rlnorm(1,0,.1),
+       "x"=y,
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
@@
 set.seed(12345) 
-jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
+ricker_jags <- do.call(jags.parallel, 
+                       list(data=jags.data, inits=jags.inits, 
+                            jags.params, n.chains=n.chains, 
                             n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
-                  model.file="ricker.bugs", progress.bar="none")
+                            model.file="ricker.bugs"))
 ```
@@
 
-```{r ricker_traces}
-jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-ricker_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
+```{r ricker_traces, dependson="ricker-mcmc"}
+tmp <- lapply(as.mcmc(ricker_jags), as.matrix) # strip classes the hard way...
+ricker_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
+names(ricker_posteriors) = c("index", "variable", "value", "chain")
+
 ggplot(ricker_posteriors) + geom_line(aes(index, value)) + 
@@
 
-```{r ricker-output}
-ricker_pardist <- acast(ricker_posteriors, index ~ variable, subset=.(variable != "deviance") )
+```{r ricker-output, dependson="ricker-traces"}
+ricker_pardist <- acast(ricker_posteriors[2:3], 
+                        1:table(ricker_posteriors$variable) ~ variable, 
+                        subset=.(variable!="deviance"))
+
 ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first
@@
   "
-  iQ <- 1 / (stdQ * stdQ);
-  iR <- 1 / (stdR * stdR);
-
+  iQ <- 1 / (stdQ * stdQ)
+  iR <- 1 / (stdR * stdR)
   r0 <- exp(logr0)
@@
 
-```{r myers-mcmc}
-init_p = c(1, 1, 1)
-names(init_p) = c("r0", "theta", "K")
+```{r myers-mcmc, dependson="myers-model", results="hide"}
 jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
 jags.inits <- function(){
-  list("logr0"=log(init_p["r0"]), "logtheta" = log(init_p["theta"]),  
-       "logK"=log(init_p["K"]), "stdQ"=sqrt(0.05), 
-       "stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+  list("logr0" 		= log(rlnorm(1, 0, .5)), 
+       "logK" 		= log(10 * rlnorm(1, 1, .2)),
+       "logtheta" = log(2 * rnorm(1, 1, .2)),  
+       "stdQ"			= 0.2  * rlnorm(1, 0, .1),
+       "stdR"			= 1e-5 * rlnorm(1, 0, .1),
+       "x" 				= y,
+       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
 }
 set.seed(12345)
-jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, 
+myers_jags <- do.call(jags.parallel, 
+                      list(data=jags.data, inits=jags.inits, jags.params, 
                            n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
-                  n.burnin=n.burnin, model.file="myers.bugs", progress.bar="none")
-
+                           n.burnin=n.burnin, model.file="myers.bugs"))
 ```
@@
 
-```{r myers-traces}
-jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-myers_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
+```{r myers-traces, dependson="myers-mcmc"}
+tmp <- lapply(as.mcmc(myers_jags), as.matrix) # strip classes the hard way...
+myers_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
+names(myers_posteriors) = c("index", "variable", "value", "chain")
+
 ggplot(myers_posteriors) + geom_line(aes(index, value)) +
@@
 
-```{r myers-output}
-myers_pardist <- as.matrix(jags_matrix[2:6])
+```{r myers-output, dependson="myers-traces"}
+myers_pardist <- acast(myers_posteriors[2:3], 
+                        1:table(myers_posteriors$variable) ~ variable, 
+                        subset=.(variable!="deviance")) 
 myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first

### 2013-05-30 inst/examples/BUGS/parametric-vs-nonparametric.Rmd
errors on Myers MCMC and Ricker optimization...

@@
 ```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
-setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
-#library(knitcitations)
-library(ggplot2) 
+source("~/.knitr_defaults.R")
+#opts_knit$set(upload.fun = socialR::flickr.url)
+library(knitcitations)
 library(nonparametricbayes) 
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA, fig.path="figure/parametric-"
-               fig.width=6, fig.height=4, cache.path="parametric-vs-nonparametric/", verbose=TRUE)
-#opts_knit$set(upload.fun = socialR::flickr.url)
-theme_set(theme_bw(base_size=12))
-theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
-             plot.background = element_rect(fill = "transparent", colour = NA))
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
-               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 ```
@@
 N <- length(y);
-jags.data <- list("N"=N,"y"=y)
+jags.data <- list("N","y")
 n.chains <- 6
@@
 
-```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup")}
+```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results="hide"}
 jags.params=c("K","logr0","logtheta","stdQ", "stdR") # be sensible about the order here
@@
 
-```{r ricker-mcmc, dependson="ricker-model"}
+```{r ricker-mcmc, dependson="ricker-model", results="hide"}
 # Uniform priors on standard deviation terms
@@
                        list(object=ricker_jags, n.update=n.update, n.iter=n.iter, 
-                            n.thin = n.thin, progress.bar="none"))
+                            n.thin = n.thin))
 ```
@@
 
-```{r myers-mcmc, dependson="myers-model"}
+```{r myers-mcmc, dependson="myers-model", results="hide"}
 jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
 jags.inits <- function(){
-  list("logr0" = log(rlnorm(1,0,.5)), 
-       "logK" = log(10 * rlnorm(1,0,.2)),
-       "logtheta" = log( 2 * rlnorm(1,0,.2)),  
-       "stdQ"= 0.2 * rlnorm(1,0,.1),
-       "stdR"= 1e-5 * rlnorm(1,0,.1),
-       "x"=y,
+  list("logr0" 		= log(rlnorm(1, 0, .5)), 
+       "logK" 		= log(10 * rlnorm(1, 1, .2)),
+       "logtheta" = log(2 * rnorm(1, 1, .2)),  
+       "stdQ"			= 0.2  * rlnorm(1, 0, .1),
+       "stdR"			= 1e-5 * rlnorm(1, 0, .1),
+       "x" 				= y,
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
@@
                            n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
-                           n.burnin=n.burnin, model.file="myers.bugs")
+                           n.burnin=n.burnin, model.file="myers.bugs"))
+recompile(myers_jags)
 myers_jags <- do.call(autojags, 
                       list(object=myers_jags, n.update=n.update, n.iter=n.iter, 
-                           n.thin = n.thin, progress.bar="none"))
+                           n.thin = n.thin))
 ```

### 2013-05-30 inst/examples/BUGS/allen.Rmd
include figure1a

@@
 ```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
-models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
-                     MLE=est_means, Ricker=ricker_means, 
+models <- data.frame(x=x_grid, 
+										 GP=tgp_dat$y, 
+										 True=true_means, 
+                     MLE=est_means, 
+										 Ricker=ricker_means, 
                      Allen = allen_means,
                      Myers = myers_means)
+models <- melt(models, id="x")
 
-models <- melt(models, id="x")
+# some labels
 names(models) <- c("x", "method", "value")
 
+# labesl for the colorkey too
 model_names = c("GP", "True", "MLE", "Ricker", "Allen", "Myers")

### 2013-05-30 inst/examples/BUGS/allen.Rmd
allen example updated

@@
 
-```{r}
+```{r posterior-mode}
+require(modeest)
 posterior.mode <- function(x) {
-  ux <- unique(x)
-  ux[which.max(tabulate(match(x, ux)))]
+  mlv(x, method="shorth")$M
 }
@@
 stdQ_prior_p <- c(0.0001, 100)
-stdR_prior_p <- c(1e-6, .1)
+stdR_prior_p <- c(1e-4, .1)
 stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
@@
 ```{r allen-output, dependson="allen-traces"}
-pardist <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable, subset=.(variable!="deviance")) 
-pardist[,2] = exp(pardist[,2]) # transform model parameters back first
-pardist[,3] = exp(pardist[,3])
+A <- allen_posteriors
+A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
+pardist <- acast(A, index ~ variable)
+# pardist <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable)[1] ~ variable) # NOT SURE WHY THIS FAILS
+
+# transform model parameters back first
+pardist[,"logr0"] = exp(pardist[,"logr0"]) 
+pardist[,"logtheta"] = exp(pardist[,"logtheta"])
+colnames(pardist)[colnames(pardist)=="logtheta"] = "theta"
+colnames(pardist)[colnames(pardist)=="logr0"] = "r0"
+
 bayes_coef <- apply(pardist,2, posterior.mode) 
-bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3])) # parameters formatted for f
+bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"], bayes_coef["theta"])) # parameters formatted for f
+allen_f <- function(x,h,p) unname(RickerAllee(x,h, unname(p[c("r0", "K", "theta")])))
 allen_means <- sapply(x_grid, f, 0, bayes_pars)
+
+bayes_pars
+head(pardist)
+
 ```
@@
 
-```{r ricker_traces, dependson="ricker-mcmc"}
+```{r ricker-traces, dependson="ricker-mcmc"}
 tmp <- lapply(as.mcmc(ricker_jags), as.matrix) # strip classes the hard way...
@@
 
-```{r ricker_posteriors, dependson=c("ricker-traces", "ricker-priors")}
+```{r ricker-posteriors, dependson=c("ricker-traces", "ricker-priors")}
 ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
@@
 ricker_pardist <- acast(ricker_posteriors[2:3], 
-                        1:table(ricker_posteriors$variable) ~ variable, 
-                        subset=.(variable!="deviance"))
+                        1:table(ricker_posteriors$variable) ~ variable)
 
 ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first
+colnames(ricker_pardist)[colnames(ricker_pardist)=="logr0"] = "r0"
 bayes_coef <- apply(ricker_pardist,2, posterior.mode) # much better estimates from mode then mean
-ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
+ricker_bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"]))
+
+ricker_f <- function(x,h,p){
+  sapply(x, function(x){ 
+    x <- pmax(0, x-h) 
+    pmax(0, x * exp(p["r0"] * (1 - x / p["K"] )) )
+  })
+}
 ricker_means <- sapply(x_grid, Ricker, 0, ricker_bayes_pars[c(1,2)])
+
+head(ricker_pardist)
+ricker_bayes_pars
 ```
@@
 myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
+myers_f <- function(x,h,p) Myer_harvest(x, h, p[c("r0", "theta", "K")])
+head(myers_pardist)
+myers_bayes_pars
+
 ```
@@
 
-# labesl for the colorkey too
+# labels for the colorkey too
 model_names = c("GP", "True", "MLE", "Ricker", "Allen", "Myers")
@@
 
-```{r par-fns}
-ricker_f <- function(x,h,p){
-  sapply(x, function(x){ 
-    x <- pmax(0, x-h) 
-    pmax(0, x * exp(p[2] * (1 - x / p[1] )) )
-  })
-}
-allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
-myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
-
-```
-
-```{r Figureb, dependson=c("assemble-models", "par-fns")}
+```{r Figureb, dependson=c("assemble-models")}
 require(MASS)
@@
 ```{r allen-opt, dependson="allen-output"}
-allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
 matrices_allen <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
@@
 ```{r ricker-opt, dependson="ricker-output"}
-ricker_f <- function(x, h, p) Ricker(x, h, unname(p[c(2, 1)])) # defined by pdgControl 
 matrices_ricker <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
@@
 ```{r myers-opt, dependson="myers-output"}
-myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
 matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
@@
 
-```{r}
-df <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable) 
-modes <- apply(df, 2, posterior.mode)
-allen_deviance <- modes[['deviance']]
-
-df <- acast(ricker_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable)
-modes <- apply(df, 2, posterior.mode)
-ricker_deviance <- modes[['deviance']]
-
-df <- acast(myers_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable)
-modes <- apply(df, 2, posterior.mode)
-myers_deviance <- modes[['deviance']]
-
+```{r deviances, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
+allen_deviance <- pardist[,'deviance']
+ricker_deviance <- ricker_pardist[,'deviance']
+myers_deviance <- myers_pardist[,'deviance']
 true_deviance <- 2*estf(c(p, sigma_g))

### 2013-05-30 inst/examples/BUGS/process-noise-only.Rmd
still debugging data formatting

@@
 ```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
-setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
-#library(knitcitations)
-library(ggplot2) 
+source("~/.knitr_defaults.R")
+#opts_knit$set(upload.fun = socialR::flickr.url)
+library(knitcitations)
 library(nonparametricbayes) 
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA,
-               fig.width=6, fig.height=4, cache.path="process/", verbose=TRUE)
-#opts_knit$set(upload.fun = socialR::flickr.url)
-theme_set(theme_bw(base_size=12))
-theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
-             plot.background = element_rect(fill = "transparent", colour = NA))
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
-               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 ```
 
-```{r}
+```{r posterior-mode}
 require(modeest)
 posterior.mode <- function(x) {
- # ux <- unique(x)
- #  ux[which.max(tabulate(match(x, ux)))]
  mlv(x, method="shorth")$M
@@
 
-```{r allen-output, dependson="allen-traces"}
+```{r allen-output, dependson=c("posterior-mode", "allen-traces")}
 # A <- allen_posteriors
@@
 
-```{r ricker-output, dependson="ricker-traces"}
+```{r ricker-output, dependson=c("posterior-mode", "ricker-traces")}
 ricker_pardist <- acast(ricker_posteriors[2:3], 
@@
                            n.burnin=n.burnin, model.file="myers_process.bugs"))
+recompile(myers_jags)
 myers_jags <- do.call(autojags, 
@@
 
-```{r myers-output, dependson="myers-traces"}
+```{r myers-output, dependson=c("posterior-mode", "myers-traces")}
 myers_pardist <- acast(myers_posteriors[2:3], 
@@
                         subset=.(variable!="deviance")) 
-myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
-myers_pardist[,2] = exp(myers_pardist[,2]) # transform model parameters back first
-myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
+myers_pardist[,1] = exp(myers_pardist[,"logK"]) # transform model parameters back first
+myers_pardist[,2] = exp(myers_pardist[,"logr0"]) # transform model parameters back first
+myers_pardist[,3] = exp(myers_pardist[,"logtheta"]) # transform model parameters back first
 colnames(myers_pardist) = c("K", "r0", "theta", "stdQ")
@@
 
-```{r}
+```{r, dependson=("posterior-mode","myers-output", "ricker-output", "allen-output", "gp-output", "mle-output") }
 df <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable) 

### 2013-05-30 inst/examples/BUGS/parametric-vs-nonparametric.Rmd
still debugging data formatting

@@
 ```{r allen-output, dependson="allen-traces"}
-# A <- allen_posteriors
-# A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
-# pardist <- acast(A[-4], index ~ variable, subset=.(variable != "deviance") )
-pardist <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable, subset=.(variable!="deviance")) 
-pardist[,2] = exp(pardist[,2]) # transform model parameters back first
-pardist[,3] = exp(pardist[,3])
+pardist <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable)[1] ~ variable) 
+pardist[,"logr0"] = exp(pardist[,"logr0"]) # transform model parameters back first
+pardist[,"logtheta"] = exp(pardist[,"logtheta"])
+colnames(pardist)[colnames(pardist)=="logtheta"] = "theta"
+colnames(pardist)[colnames(pardist)=="logr0"] = "r0"
 bayes_coef <- apply(pardist,2, posterior.mode) 
-bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3])) # parameters formatted for f
+bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"], bayes_coef["theta"])) # parameters formatted for f
 allen_means <- sapply(x_grid, f, 0, bayes_pars)
+
 ```
@@
 jags.inits <- function(){
-  list("logr0" 		= log(rlnorm(1, 0, .5)), 
-       "logK" 		= log(10 * rlnorm(1, 1, .2)),
-       "logtheta" = log(2 * rnorm(1, 1, .2)),  
+  list("logr0" 		= log(rlnorm(1, 0, .3)), 
+       "logK" 		= log(10 * rlnorm(1, 1, .1)),
+       "logtheta" = log(2 * rnorm(1, 1, .1)),  
        "stdQ"			= 0.2  * rlnorm(1, 0, .1),
@@
                         subset=.(variable!="deviance")) 
-myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
-myers_pardist[,2] = exp(myers_pardist[,2]) # transform model parameters back first
-myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
-colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
+# robust to ordering
+myers_pardist[,1] = exp(myers_pardist[,"logK"]) # transform model parameters back first
+myers_pardist[,2] = exp(myers_pardist[,"logr0"]) # transform model parameters back first
+myers_pardist[,3] = exp(myers_pardist[,"logtheta"]) # transform model parameters back first
+
+colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR") # assumes they are in this order...
 bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates

### 2013-05-31 inst/examples/BUGS/process-noise-only.Rmd
with code externalization

@@
 library(nonparametricbayes) 
+opts_chunk$set(external=TRUE)
+read_chunk("external-chunks.R")
 ```
@@
 ```{r posterior-mode}
-require(modeest)
-posterior.mode <- function(x) {
- mlv(x, method="shorth")$M
-}
-
 ```
@@
 ```{r stateeq}
-f <- RickerAllee
-p <- c(1, 10, 5)
-K <- 10  # approx, a li'l' less
-allee <- 5 # approx, a li'l' less
 ```
@@
 ```{r sdp-pars, dependson="stateeq"}
-sigma_g <- 0.05
-sigma_m <- 0.0
-z_g <- function() rlnorm(1, 0, sigma_g)
-z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
-x_grid <- seq(0, 1.5 * K, length=50)
-h_grid <- x_grid
-profit <- function(x,h) pmin(x, h)
-delta <- 0.01
-OptTime <- 50  # stationarity with unstable models is tricky thing
-reward <- 0
-xT <- 0
-Xo <-  allee + 0.5 # observations start from
-x0 <- K # simulation under policy starts from
-Tobs <- 40
 ```
@@
 ```{r obs, dependson="sdp-pars"}
-  set.seed(1234)
-  #harvest <- sort(rep(seq(0, .5, length=7), 5))
-  x <- numeric(Tobs)
-  x[1] <- Xo
-  nz <- 1
-  for(t in 1:(Tobs-1))
-    x[t+1] = z_g() * f(x[t], h=0, p=p)
-  obs <- data.frame(x = c(rep(0,nz), 
-                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
-                    y = c(rep(0,nz), 
-                          x[2:Tobs]))
-raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
-raw_plot
 ```
@@
 ```{r mle, dependson="obs"}
-set.seed(12345)
-estf <- function(p){ 
-    mu <- f(obs$x,0,p)
-    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
-}
-par <- c(p[1]+abs(rnorm(1,0,1)), 
-         p[2]+rnorm(1,0, .1), 
-         p[3]+rnorm(1,0, .1), 
-         sigma_g + abs(rnorm(1,0, .1)))
-o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
-f_alt <- f
-p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
-sigma_g_alt <- as.numeric(o$par[4])
-
-est <- list(f = f_alt, p = p_alt, sigma_g = sigma_g_alt, mloglik=o$value)
 ```
@@
 ```{r mle-output, dependson="mle"}
-true_means <- sapply(x_grid, f, 0, p)
-est_means <- sapply(x_grid, est$f, 0, est$p)
 ```
@@
 ```{r gp-priors}
-#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
-s2.p <- c(5,5)  
-d.p = c(10, 1/0.1)
 ```
@@
 ```{r gp, dependson=c("gp-priors", "obs")}
-gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
-gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)
 ```
@@
 ```{r gp_traces_densities, dependson="gp"}
-plots <- summary_gp_mcmc(gp)
 ```
@@
 ```{r gp-output, dependson="gp"}
-# Summarize the GP model
-tgp_dat <- 
-    data.frame(  x = x_grid, 
-                 y = gp_dat$E_Ef, 
-                 ymin = gp_dat$E_Ef - 2 * sqrt(gp_dat$E_Vf), 
-                 ymax = gp_dat$E_Ef + 2 * sqrt(gp_dat$E_Vf) )
 ```
@@
 N <- length(x);
-jags.data <- list("N"=N,"y"=y)
-n.chains <- 4
+jags.data <- list("N","y")
+n.chains <- 6
 n.iter <- 1e6
@@
 n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
+n.update <- 10
 ```
@@
 ```{r common-priors}
-stdQ_prior_p <- c(0.0001, 100)
-stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
 ```
@@
   list("K"= 10 * rlnorm(1,0, 0.1),
-       "logr0"=log( rlnorm(1,0, 0.5) ),
-       "logtheta"=log(  2 * rlnorm(1,0, 0.1) ), 
-       "stdQ"= abs(sigma_g * rlnorm(1,0, 0.1)),
+       "logr0"=log( 1 * rlnorm(1,0, 0.5) ),
+       "logtheta"=log(  5 * rlnorm(1,0, 0.1) ), 
+       "stdQ"= abs( 0.2 * rlnorm(1,0, 0.1)),
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
@@
 ```{r allen-traces, dependson="allen-mcmc"}
-tmp <- lapply(as.mcmc(allen_jags), as.matrix) # strip classes to melt
-allen_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
-names(allen_posteriors) = c("index", "variable", "value", "chain")
-ggplot(allen_posteriors) + geom_line(aes(index, value)) + 
-  facet_wrap(~ variable, scale="free", ncol=1)
 ```
@@
 ```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
-allen_priors <- ddply(allen_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100) 
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-
-ggplot(allen_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=allen_priors, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=3)
 ```
@@
 ```{r allen-output, dependson=c("posterior-mode", "allen-traces")}
-# A <- allen_posteriors
-# A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
-# pardist <- acast(A[-4], index ~ variable, subset=.(variable != "deviance") )
-pardist <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable)[1] ~ variable) 
-pardist[,"logr0"] = exp(pardist[,"logr0"]) # transform model parameters back first
-pardist[,"logtheta"] = exp(pardist[,"logtheta"])
-
-colnames(pardist)[colnames(pardist)=="logtheta"] = "theta"
-colnames(pardist)[colnames(pardist)=="logr0"] = "r0"
-
-bayes_coef <- apply(pardist,2, posterior.mode) 
-bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"], bayes_coef["theta"])) # parameters formatted for f
-allen_means <- sapply(x_grid, f, 0, bayes_pars)
 ```
@@
 ```{r ricker_traces, dependson="ricker-mcmc"}
-tmp <- lapply(as.mcmc(ricker_jags), as.matrix) # strip classes the hard way...
-ricker_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
-names(ricker_posteriors) = c("index", "variable", "value", "chain")
-
-ggplot(ricker_posteriors) + geom_line(aes(index, value)) + 
-  facet_wrap(~ variable, scale="free", ncol=1)
 ```
@@
 ```{r ricker_posteriors, dependson=c("ricker-traces", "ricker-priors")}
-ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100) 
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-# plot posterior distributions
-ggplot(ricker_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=2)
 ```
@@
 ```{r ricker-output, dependson=c("posterior-mode", "ricker-traces")}
-ricker_pardist <- acast(ricker_posteriors[2:3], 
-                        1:table(ricker_posteriors$variable) ~ variable, 
-                        subset=.(variable!="deviance")) 
-ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first
-bayes_coef <- apply(ricker_pardist,2, posterior.mode) # much better estimates from mode then mean
-ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
-ricker_means <- sapply(x_grid, Ricker, 0, ricker_bayes_pars[c(1,2)])
 ```
@@
   list("logr0"=log(rlnorm(1,0,.1)), 
-       "logK"=log(7 * rlnorm(1,0,.1)),
+       "logK"=log(10 * rlnorm(1,0,.1)),
        "logtheta" = log(2 * rlnorm(1,0,.1)),  
@@
 ```{r myers-traces, dependson="myers-mcmc"}
-tmp <- lapply(as.mcmc(myers_jags), as.matrix) # strip classes the hard way...
-myers_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
-names(myers_posteriors) = c("index", "variable", "value", "chain")
-
-ggplot(myers_posteriors) + geom_line(aes(index, value)) +
-  facet_wrap(~ variable, scale="free", ncol=1)
-
 ```
@@
 ```{r myers-posteriors, dependson="myers-traces"}
-par_prior_curves <- ddply(myers_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100) 
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-
-ggplot(myers_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=3)
 ```
@@
 ```{r myers-output, dependson=c("posterior-mode", "myers-traces")}
-myers_pardist <- acast(myers_posteriors[2:3], 
-                        1:table(myers_posteriors$variable) ~ variable, 
-                        subset=.(variable!="deviance")) 
-myers_pardist[,1] = exp(myers_pardist[,"logK"]) # transform model parameters back first
-myers_pardist[,2] = exp(myers_pardist[,"logr0"]) # transform model parameters back first
-myers_pardist[,3] = exp(myers_pardist[,"logtheta"]) # transform model parameters back first
-colnames(myers_pardist) = c("K", "r0", "theta", "stdQ")
-bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
-myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
-myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
 ```
@@
 ```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
-models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
-                     MLE=est_means, Ricker=ricker_means, 
-                     Allen = allen_means,
-                     Myers = myers_means)
-
-models <- melt(models, id="x")
-names(models) <- c("x", "method", "value")
-
-model_names = c("GP", "True", "MLE", "Ricker", "Allen", "Myers")
-colorkey=cbPalette
-names(colorkey) = model_names 
 ```
@@
 ```{r Figure1, dependson="assemble-models"}
-plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-    geom_line(data=models, aes(x, value, col=method), lwd=1, alpha=0.8) + 
-    geom_point(data=obs, aes(x,y), alpha=0.8) + 
-    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
-    scale_colour_manual(values=cbPalette) 
-print(plot_gp)
 ```
 
-## Step-ahead predictors
-
-## Posterior predictive curves
+## Goodness of fit
 
@@
 
-We will need a vector version (`pmin` in place of `min`) of the parametric growth functions that can operate on the posteriors, (with appropriate ordering of parameters as they are in the posterior):
-
-```{r par-fns}
-ricker_f <- function(x,h,p){
-  sapply(x, function(x){ 
-    x <- pmax(0, x-h) 
-    pmax(0, x * exp(p[2] * (1 - x / p[1] )) )
-  })
-}
-allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
-myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
-
-```
-
 ```{r Figureb, dependson=c("assemble-models", "par-fns")}
-require(MASS)
-step_ahead <- function(x, f, p){
-  h = 0
-  x_predict <- sapply(x, f, h, p)
-  n <- length(x_predict) - 1
-  y <- c(x[1], x_predict[1:n])
-  y
-}
-step_ahead_posteriors <- function(x){
-gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
-df_post <- melt(lapply(sample(100), 
-  function(i){
-    data.frame(time = 1:length(x), stock = x, 
-                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
-                True = step_ahead(x,f,p),  
-                MLE = step_ahead(x,f,est$p), 
-                Allen = step_ahead(x, allen_f, pardist[i,]), 
-                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
-                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
-  }), id=c("time", "stock"))
-}
-
-df_post <- step_ahead_posteriors(x)
-
-ggplot(df_post) + geom_point(aes(time, stock)) + 
-  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) 
 ```
@@
 ```{r gp-opt, dependson="gp-output"}
-MaxT = 1000
-# uses expected values from GP, instead of integrating over posterior
-#matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)
-
-# Integrate over posteriors 
-matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid)
-
-# Solve the SDP using the GP-derived transition matrix
-opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
 ```
@@
 ```{r mle-opt, dependson="mle-output"}
-matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-matrices_estimated <- f_transition_matrix(est$f, est$p, x_grid, h_grid, est$sigma_g)
-opt_estimated <- value_iteration(matrices_estimated, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
 ```
@@
 ```{r allen-opt, dependson="allen-output"}
-allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
-matrices_allen <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
-opt_allen <- value_iteration(matrices_allen, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
 ```
@@
 ```{r ricker-opt, dependson="ricker-output"}
-ricker_f <- function(x, h, p) Ricker(x, h, unname(p[c(2, 1)])) # defined by pdgControl 
-matrices_ricker <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
-opt_ricker <- value_iteration(matrices_ricker, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
 ```
@@
 ```{r myers-opt, dependson="myers-output"}
-myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
-matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
-myers_alt <- value_iteration(matrices_myers, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
 ```
@@
 ```{r assemble-opt, dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
-OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_ricker$D, Allen = opt_allen$D, Myers = myers_alt$D)
-colorkey=cbPalette
-names(colorkey) = names(OPT) 
 ```
@@
 ```{r Figure2, dependson="assemble-opt"}
-policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
-names(policies) <- c("stock", "method", "value")
-
-ggplot(policies, aes(stock, stock - value, color=method)) +
-  geom_line(lwd=1.2, alpha=0.8) + xlab("stock size") + ylab("escapement")  +
-  scale_colour_manual(values=colorkey)
 ```
@@
 ```{r sims, dependson="assemble-opt"}
-sims <- lapply(OPT, function(D){
-  set.seed(1)
-  lapply(1:100, function(i) 
-    ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
-  )
-})
-
-dat <- melt(sims, id=names(sims[[1]][[1]]))
-dt <- data.table(dat)
-setnames(dt, c("L1", "L2"), c("method", "reps")) 
-# Legend in original ordering please, not alphabetical: 
-dt$method = factor(dt$method, ordered=TRUE, levels=names(OPT))
 ```
@@
 ```{r Figure3, dependson="sims"}
-ggplot(dt) + 
-  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
 ```
@@
 ```{r profits, dependson="sims"}
-Profit <- dt[, sum(profit), by=c("reps", "method")]
-Profit[, mean(V1), by="method"]
 ```
@@
 ```{r totalprofits, dependson="profits"}
-ggplot(Profit, aes(V1)) + geom_histogram() + 
-  facet_wrap(~method, scales = "free_y") + guides(legend.position = "none") + xlab("Total profit by replicate")
 ```
 
-```{r, dependson=("posterior-mode","myers-output", "ricker-output", "allen-output", "gp-output", "mle-output") }
-df <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable) 
-modes <- apply(df, 2, posterior.mode)
-allen_deviance <- modes[['deviance']]
-
-df <- acast(ricker_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable)
-modes <- apply(df, 2, posterior.mode)
-ricker_deviance <- modes[['deviance']]
-
-df <- acast(myers_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable)
-modes <- apply(df, 2, posterior.mode)
-myers_deviance <- modes[['deviance']]
-
-true_deviance <- 2*estf(c(p, sigma_g))
-mle_deviance <- 2*estf(c(est$p, est$sigma_g))
-
-c(allen = allen_deviance, ricker=ricker_deviance, myers=myers_deviance, true=true_deviance, mle=mle_deviance)
-
+```{r deviances, dependson=("posterior-mode","myers-output", "ricker-output", "allen-output", "gp-output", "mle-output") }
 ```

### 2013-05-31 inst/examples/BUGS/external-chunks.R

@@ -1,0 +1,592 @@
+
+## @knitr plotting-options
+source("~/.knitr_defaults.R")
+#opts_knit$set(upload.fun = socialR::flickr.url)
+library(knitcitations)
+library(nonparametricbayes) 
+
+
+## @knitr posterior-mode
+require(modeest)
+posterior.mode <- function(x) {
+  mlv(x, method="shorth")$M
+}
+
+
+
+## @knitr stateeq
+f <- RickerAllee
+p <- c(2, 8, 5)
+K <- 10  # approx, a li'l' less
+allee <- 5 # approx, a li'l' less
+
+
+## @knitr sdp-pars
+sigma_g <- 0.05
+sigma_m <- 0.0
+z_g <- function() rlnorm(1, 0, sigma_g)
+z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
+x_grid <- seq(0, 1.5 * K, length=50)
+h_grid <- x_grid
+profit <- function(x,h) pmin(x, h)
+delta <- 0.01
+OptTime <- 50  # stationarity with unstable models is tricky thing
+reward <- 0
+xT <- 0
+Xo <-  allee+.5# observations start from
+x0 <- K # simulation under policy starts from
+Tobs <- 40
+
+
+## @knitr obs
+  set.seed(1234)
+  #harvest <- sort(rep(seq(0, .5, length=7), 5))
+  x <- numeric(Tobs)
+  x[1] <- Xo
+  nz <- 1
+  for(t in 1:(Tobs-1))
+    x[t+1] = z_g() * f(x[t], h=0, p=p)
+  obs <- data.frame(x = c(rep(0,nz), 
+                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
+                    y = c(rep(0,nz), 
+                          x[2:Tobs]))
+raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
+raw_plot
+
+
+## @knitr mle
+set.seed(12345)
+estf <- function(p){ 
+    mu <- f(obs$x,0,p)
+    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
+}
+par <- c(p[1]*rlnorm(1,0,.4), 
+         p[2]*rlnorm(1,0,.3), 
+         p[3]*rlnorm(1,0, .3), 
+         sigma_g * rlnorm(1,0,.3))
[…]

### 2013-06-01 inst/examples/BUGS/process-noise-only.Rmd
using better initial conditions

@@
 
-```{r jags-setup}
+```{r jags-setup-local}
 y <- x 
@@
 
-```{r allen-model}
+```{r allen-model-local}
 K_prior_p <- c(0.01, 40.0)
@@
 
-```{r allen-priors, dependson="common-priors"}
+```{r allen-priors-local, dependson="common-priors"}
 K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
@@
 
-```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results="hide"}
+```{r allen-mcmc-local, dependson=c("allen-model-local", "allen-pars", "jags-setup"), results="hide"}
 jags.params=c("K","logr0","logtheta","stdQ") # be sensible about the order here
 jags.inits <- function(){
-  list("K"= 10 * rlnorm(1,0, 0.1),
-       "logr0"=log( 1 * rlnorm(1,0, 0.5) ),
+  list("K"= 8 * rlnorm(1,0, 0.1),
+       "logr0"=log(2 * rlnorm(1,0, 0.5) ),
        "logtheta"=log(  5 * rlnorm(1,0, 0.1) ), 
@@
 
-```{r allen-traces, dependson="allen-mcmc"}
+```{r allen-traces, dependson="allen-mcmc-local"}
 ```
 
-```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
+```{r allen-posteriors, dependson=c("allen-traces", "allen-priors-local")}
 ```
@@
 
-```{r ricker-model}
+```{r ricker-model-local}
 K_prior_p <- c(0.01, 40.0)
@@
 
-```{r ricker-priors, dependson="common-priors"}
+```{r ricker-priors-local, dependson="common-priors"}
 K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
@@
 
-```{r ricker-mcmc, dependson="ricker-model"}
+```{r ricker-mcmc-local, dependson="ricker-model-local"}
 # Uniform priors on standard deviation terms
@@
 
-```{r ricker_traces, dependson="ricker-mcmc"}
+```{r ricker-traces, dependson="ricker-mcmc-local"}
 ```
 
-```{r ricker_posteriors, dependson=c("ricker-traces", "ricker-priors")}
+```{r ricker-posteriors, dependson=c("ricker-traces", "ricker-priors-local")}
 ```
@@
 
-```{r myers-model}
+```{r myers-model-local}
 logr0_prior_p <- c(-6.0, 6.0)
@@
 
-```{r myers-priors}
+```{r myers-priors-local}
 logK_prior     <- function(x) dunif(x, logK_prior_p[1], logK_prior_p[2])
@@
 
-```{r myers-mcmc, dependson="myers-model"}
+```{r myers-mcmc-local, dependson="myers-model-local"}
 jags.params=c("logr0", "logtheta", "logK", "stdQ")
@@
 myers_jags <- do.call(autojags, 
-                      list(myers_jags, n.update=10, n.iter=n.iter, 
+                      list(myers_jags, n.update=n.update, n.iter=n.iter, 
                            n.thin = n.thin, progress.bar="none"))
@@
 
-```{r myers-traces, dependson="myers-mcmc"}
+```{r myers-traces, dependson="myers-mcmc-local"}
 ```
@@
 
-```{r deviances, dependson=("posterior-mode","myers-output", "ricker-output", "allen-output", "gp-output", "mle-output") }
+```{r deviances, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output") }
 ```

### 2013-06-01 inst/examples/admb-example.Rmd
learning ADMB.  not working yet.

# Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach

Plotting and knitr options, (can generally be ignored)

```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
source("~/.knitr_defaults.R")
opts_chunk$set(warning=TRUE, message=TRUE, cache=FALSE)
#opts_knit$set(upload.fun = socialR::flickr.url)
```

### Model and parameters

Uses the model derived in ` citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,

```{r stateeq}
f <- function(x,h,p)  x * exp(p[1] * (1 - x / p[2]) * (x - p[3]) / p[2] ) 
p <- c(1, 10, 5)
K <- 10  # approx, a li'l' less
allee <- 5 # approx, a li'l' less
```

Various parameters defining noise dynamics, grid, and policy costs.

```{r sdp-pars, dependson="stateeq"}
sigma_g <- 0.05
z_g <- function() rlnorm(1,0, sigma_g)
x_grid <- seq(0, 1.5 * K, length=50)
Xo <-  allee + 0.5 # observations start from
Tobs <- 40
```

### Sample Data

```{r obs, dependson="sdp-pars"}
  set.seed(1234)
  #harvest <- sort(rep(seq(0, .5, length=7), 5))
  x <- numeric(Tobs)
  x[1] <- Xo
  nz <- 1
  for(t in 1:(Tobs-1))
    x[t+1] = z_g() * f(x[t], h=0, p=p)
  obs <- data.frame(x = c(rep(0,nz), 
                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
                    y = c(rep(0,nz), 
                          x[2:Tobs]))
raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
raw_plot
```

## Maximum Likelihood "by hand"

```{r mle, dependson="obs"}
set.seed(12345)
estf <- function(p){ 
    mu <- f(obs$x,0,p)
    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
}
par <- c(p[1]+abs(rnorm(1,0,1)), 
         p[2]+rnorm(1,0, .1), 
         p[3]+rnorm(1,0, .1), 
         sigma_g + abs(rnorm(1,0, .1)))
o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
f_alt <- f
p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
sigma_g_alt <- as.numeric(o$par[4])
est <- list(f = f_alt, p = p_alt, sigma_g = sigma_g_alt, mloglik=o$value)
```

```{r admb-lib}
# install_github("R2admb", "bbolker", subdir="R2admb") # dev version
library(R2admb)
```

## ADMB definition

```{r model}
model <- 
paste("
PARAMETER_SECTION
  vector mu(1,n) // per capita mort prob

PROCEDURE_SECTION

  mu = x + r * elem_prod((1 - x / k), (x - c) / k);
  f -= sum( (0.5 * log(2*M_PI) + log(s) + log(x)  + 0.5 * square(log(x) - mu) / square(s)) ); 
")
writeLines(model, "model.tpl")
```

Or try defining with a loop.  (Not used currently)

```{r loop-model}
model <- 
paste("
PARAMETER_SECTION
  vector mu(1,n) // per capita mort prob

PROCEDURE_SECTION
  for(int i = 0; i < n; i++)
  {
    mu[i] = x[i] + r * (1 - x[i] / k) * (x[i] - c) / k;
    f -= (0.5 * log(2*M_PI) + log(s) + log(x[i])  + 0.5 * square(log(x[i]) - mu) / square(s)) );
  }

")
writeLines(model, "loop_model.tpl")
```

```{r run-admb}
setup_admb("/var/admb")

df <- data.frame(x=x)
params <- list(r = 1, k = 1, c = 1, s = 1) ## starting parameters
bounds <- list(r = c(0, 100), k=c(0, 100), c=c(0, 100), s = c(0,100)) ## bounds
dat <- c(list(n = nrow(df)), df)
m1 <- do_admb("model",
              data = dat,
              params = params,
              bounds = bounds,
              run.opts = run.control(checkparam="write",
                                     checkdata="write"))
```

```{r}
unlink(c("model.tpl",
         "model_gen.tpl",
         "model")) ## clean up leftovers

m1
```

Not so good. Not sure how to debug.  We can confirm ADMB works correctly on example from the Vignette:

## Ben Bolker Model

from R2admb Vignette.  Works just fine:

```{r frog-model}
model <- 
paste("
PARAMETER_SECTION
  vector prob(1,nobs) // per capita mort prob

PROCEDURE_SECTION

  dvariable fpen=0.0; // penalty variable
  // power-Ricker
  prob = c*pow(elem_prod(TBL/d,exp(1-TBL/d)),g);
  // penalties: constrain 0.001 <= prob <= 0.999
  prob = posfun(prob,0.001,fpen);
  f += 1000*fpen;
  prob = 1-posfun(1-prob,0.001,fpen);
  f += 1000*fpen;
  // binomial negative log-likelihood
  f -= sum( log_comb(nexposed,Kill)+
  elem_prod(Kill,log(prob))+
  elem_prod(nexposed-Kill,log(1-prob)));
")
writeLines(model, "frogs.tpl")
```

```{r frogs}
setup_admb("/var/admb")
df <- data.frame(TBL = rep(c(9,12,21,25,37),each=3), 
                 Kill = c(0,2,1,3,4,5,0,0,0,0,1,0,0,0,0L))
rfs_params <- list(c = 0.45, d = 13, g=1) ## starting parameters
rfs_bounds <- list(c = c(0, 1), d=c(0, 50), g=c(-1, 25)) ## bounds
rfs_dat <- c(list(nobs = nrow(df),
                  nexposed=rep(10, nrow(df))), df)
m1 <- do_admb("frogs",
              data=rfs_dat,
              params=rfs_params,
              bounds=rfs_bounds,
              run.opts=run.control(checkparam="write",
                                   checkdata="write"))
```

```{r}
unlink(c("frogs.tpl",
         "frogs_gen.tpl",
         "frogs")) ## clean up leftovers

m1
```

### 2013-06-01 inst/examples/admb-example.Rmd
admb example

@@
-# Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach
+# Learning ADMB
 
@@
 opts_chunk$set(warning=TRUE, message=TRUE, cache=FALSE)
-#opts_knit$set(upload.fun = socialR::flickr.url)
+opts_knit$set(upload.fun = socialR::flickr.url)
 ```
@@
                           x[2:Tobs]))
-raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
+raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line() + geom_point()
 raw_plot

### 2013-06-03 inst/examples/admb-example.Rmd
admb example working now

@@
 ### Model and parameters
-
-Uses the model derived in ` citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
 
@@
 Tobs <- 40
+set.seed(123)
 ```
@@
 ```{r obs, dependson="sdp-pars"}
-  set.seed(1234)
-  #harvest <- sort(rep(seq(0, .5, length=7), 5))
 x <- numeric(Tobs)
 x[1] <- Xo
-  nz <- 1
 for(t in 1:(Tobs-1))
   x[t+1] = z_g() * f(x[t], h=0, p=p)
-  obs <- data.frame(x = c(rep(0,nz), 
-                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
-                    y = c(rep(0,nz), 
-                          x[2:Tobs]))
-raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line() + geom_point()
-raw_plot
+
 ```
@@
 ```{r mle, dependson="obs"}
-set.seed(12345)
-estf <- function(p){ 
-    mu <- f(obs$x,0,p)
-    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
+n = length(x)
+loglik <- function(p){ 
+  r = p[1]; k = p[2]; c = p[3]; s = p[4];
+  mu = x + r * (1 - x / k) * (x - c) / k;
+  0.5 * n * log(2 * pi) + n * log(s) + 0.5 * sum(x - mu)^2/ s^2;
+  #-sum(dlnorm(x, mu, s), log=TRUE) # above equation is equivalent to results from built-in function
   }
-par <- c(p[1]+abs(rnorm(1,0,1)), 
-         p[2]+rnorm(1,0, .1), 
-         p[3]+rnorm(1,0, .1), 
-         sigma_g + abs(rnorm(1,0, .1)))
-o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
-f_alt <- f
-p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
-sigma_g_alt <- as.numeric(o$par[4])
-est <- list(f = f_alt, p = p_alt, sigma_g = sigma_g_alt, mloglik=o$value)
+init <- c(1, 1, 1, 1)
+o <- optim(init, loglik, method="L", lower=1e-5, upper=100)
+o$par
 ```
+
+Okay, now lets try admb.  We use R2admb which is just a convenient way to write our data and parameters into an admb file.
 
@@
 
+We still need to define the model using ADMB notation in the procedure section.
+This is mostly like R or C++, with the exception of special functions like `square` in place of `^2`, `norm2` for the sum of squares, and `elem_prod` istead of `*` for the element-wise product of two arrays.
+The constant `pi` is given as `M_PI`, as typical of C/C++ libraries.
+Where these other functions are defined I'm not sure, but some useful guides to [ADMB vector/matrix operations](http://fish.washington.edu/research/MPAM/resources/ADMB_Minte-Vera.pdf) or an (undefined) list of [keywords](http://www.admb-project.org/developers/contribute-documentation/functions/keywords.txt/view)...
+
+The equivalent model
+
 ```{r model}
@@
 PROCEDURE_SECTION
-
   mu = x + r * elem_prod((1 - x / k), (x - c) / k);
-  f -= sum( (0.5 * log(2*M_PI) + log(s) + log(x)  + 0.5 * square(log(x) - mu) / square(s)) ); 
+  f = 0.5 * n * log(2 * M_PI) + n * log(s) + 0.5 * norm2(x - mu) / square(s);
 ")
 writeLines(model, "model.tpl")
-```
-
-Or try defining with a loop.  (Not used currently)
-
-```{r loop-model}
-model <- 
-paste("
-PARAMETER_SECTION
-  vector mu(1,n) // per capita mort prob
-
-PROCEDURE_SECTION
-  for(int i = 0; i < n; i++)
-  {
-    mu[i] = x[i] + r * (1 - x[i] / k) * (x[i] - c) / k;
-    f -= (0.5 * log(2*M_PI) + log(s) + log(x[i])  + 0.5 * square(log(x[i]) - mu) / square(s)) );
-  }
-
-")
-writeLines(model, "loop_model.tpl")
 ```
@@
 df <- data.frame(x=x)
-params <- list(r = 1, k = 1, c = 1, s = 1) ## starting parameters
-bounds <- list(r = c(0, 100), k=c(0, 100), c=c(0, 100), s = c(0,100)) ## bounds
+params <- list(r = 1, k = 10, c = 2, s = .1) ## starting parameters
+bounds <- list(r = c(1e-5, 100), k=c(1e-5, 100), c=c(1e-5, 100), s = c(1e-5,100)) ## bounds
 dat <- c(list(n = nrow(df)), df)
@@
               run.opts = run.control(checkparam="write",
-                                     checkdata="write"))
-```
-
-```{r}
-unlink(c("model.tpl",
-         "model_gen.tpl",
-         "model")) ## clean up leftovers
-
+                                     checkdata="write", clean=FALSE))
 m1
@@
 
-Not so good. Not sure how to debug.  We can confirm ADMB works correctly on example from the Vignette:
+Which now matches the results from `optim`, now that we avoid the k=0 bound.
 
-## Ben Bolker Model
+Hans suggests adding an error term in the function definitions rather than in limiting the bounds or log transforming the variables:
 
-from R2admb Vignette.  Works just fine:
-
-```{r frog-model}
-model <- 
-paste("
-PARAMETER_SECTION
-  vector prob(1,nobs) // per capita mort prob
-
-PROCEDURE_SECTION
-
-  dvariable fpen=0.0; // penalty variable
-  // power-Ricker
-  prob = c*pow(elem_prod(TBL/d,exp(1-TBL/d)),g);
-  // penalties: constrain 0.001 <= prob <= 0.999
-  prob = posfun(prob,0.001,fpen);
-  f += 1000*fpen;
-  prob = 1-posfun(1-prob,0.001,fpen);
-  f += 1000*fpen;
-  // binomial negative log-likelihood
-  f -= sum( log_comb(nexposed,Kill)+
-  elem_prod(Kill,log(prob))+
-  elem_prod(nexposed-Kill,log(1-prob)));
-")
-writeLines(model, "frogs.tpl")
-```
-
-```{r frogs}
-setup_admb("/var/admb")
-df <- data.frame(TBL = rep(c(9,12,21,25,37),each=3), 
-                 Kill = c(0,2,1,3,4,5,0,0,0,0,1,0,0,0,0L))
-rfs_params <- list(c = 0.45, d = 13, g=1) ## starting parameters
-rfs_bounds <- list(c = c(0, 1), d=c(0, 50), g=c(-1, 25)) ## bounds
-rfs_dat <- c(list(nobs = nrow(df),
-                  nexposed=rep(10, nrow(df))), df)
-m1 <- do_admb("frogs",
-              data=rfs_dat,
-              params=rfs_params,
-              bounds=rfs_bounds,
-              run.opts=run.control(checkparam="write",
-                                   checkdata="write"))
-```
-
-```{r}
-unlink(c("frogs.tpl",
-         "frogs_gen.tpl",
-         "frogs")) ## clean up leftovers
-
-m1
-```
+> The most common plase where
+ this goes wrong is 1/0, log(0), sqrt(0), pow(0,1) etc.
+> Your suggestion is OK, but usually I prefer to put
+ in log(1e-10+my_expression), sqrt(1e-10+my_expression), pow(1e-10+my_expression,1)

### 2013-06-05 inst/examples/BUGS/process-noise-only.Rmd

@@
   list("K"= 8 * rlnorm(1,0, 0.1),
-       "logr0"=log(2 * rlnorm(1,0, 0.5) ),
+       "logr0"=log(2 * rlnorm(1,0, 0.1) ),
        "logtheta"=log(  5 * rlnorm(1,0, 0.1) ), 
-       "stdQ"= abs( 0.2 * rlnorm(1,0, 0.1)),
+       "stdQ"= abs( 0.1 * rlnorm(1,0, 0.1)),
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
@@
 # parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
-allen_jags <- do.call(jags, list(data=jags.data, inits=jags.inits, 
+allen_jags <- do.call(jags.parallel, list(data=jags.data, inits=jags.inits, 
                                       jags.params, n.chains=n.chains, 
@@
 set.seed(12345)
-myers_jags <- do.call(jags, 
+myers_jags <- do.call(jags.parallel, 
                       list(data=jags.data, inits=jags.inits, jags.params, 

### 2013-06-05 inst/examples/BUGS/process-noise-only.Rmd
lognormal noise bugs model

@@
 K_prior_p <- c(0.01, 40.0)
-logr0_prior_p <- c(-6.0, 6.0)
+logr0_prior_p <- c(-3.0, 6.0)
 logtheta_prior_p <- c(-6.0, 6.0)
@@
     mu[t] <- y[t] * exp(r0 * (1 - y[t]/K)* (y[t] - theta) / K )
-    y[t+1] ~ dnorm(mu[t], iQ) 
+    y[t+1] ~ dlnorm(mu[t], iQ) 
   }
@@
     mu[t] <- y[t] * exp(r0 * (1 - y[t]/K) )
-    y[t+1] ~ dnorm(mu[t], iQ) 
+    y[t+1] ~ dlnorm(mu[t], iQ) 
   }
@@
     mu[t] <- r0 * pow(abs(y[t]), theta) / (1 + pow(abs(y[t]), theta) / K)
-    y[t+1] ~ dnorm(mu[t], iQ) 
+    y[t+1] ~ dlnorm(mu[t], iQ) 
   }

### 2013-06-05 inst/examples/semi-analytic-posteriors.Rmd
semi-analytic posteriors (first go)

# Semi-analytic posteriors

```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
source("~/.knitr_defaults.R")
opts_chunk$set(warning=TRUE, message=TRUE, cache=FALSE)
opts_knit$set(upload.fun = socialR::flickr.url)
```

### Generating Model and parameters

Ricker model, parameterized as

$$X_{t+1} = X_t r e^{-\beta X_t + \sigma Z_t}$$

for random unit normal $Z_t$

```{r stateeq}
f <- function(x,h,p)  x * p[1] * exp(-x * p[2]) 
p <- c(exp(1), 1/10)
K <- 10  # approx, a li'l' less
Xo <- 1 # approx, a li'l' less
```

```{r sdp-pars, dependson="stateeq"}
sigma_g <- 0.1
z_g <- function() rlnorm(1,0, sigma_g)
x_grid <- seq(0, 1.5 * K, length=50)
N <- 40
set.seed(123)
```

### Sample Data

```{r obs, dependson="sdp-pars"}
x <- numeric(N)
x[1] <- Xo
for(t in 1:(N-1))
  x[t+1] = z_g() * f(x[t], h=0, p=p)
qplot(1:N, x)
```

## Compute the posterior after marginalizing over $r$ and $\sigma$ parameters:

$$P(\beta | X) $$

```{r}
Mt <- function(t, beta) log(x[t+1]) - log(x[t]) + beta * x[t]
beta_grid = seq(1e-5, 2, by=1e-3)

P_B.X <- sapply(beta_grid, function(beta){
  Mt_vec = sapply(1:(N-1), Mt, beta)
  sum_of_squares <- sum(Mt_vec^2)
  square_of_sums <- sum(Mt_vec)^2
  0.5 ^ (N/2-1) * (sum_of_squares - square_of_sums/(N-1)) ^ (N/2-1) / gamma(N/2-1)
  })

qplot(beta_grid, -log(P_B.X))

beta_grid[which.min(P_B.X)]

```

Estimating the Myers model on this data:

```{r}
Mt <- function(t, theta, K) log(x[t+1]) - theta * log(x[t]) + log(1 + x[t] ^ theta / K) 
theta_grid = seq(1e-5, 5, length=100)
K_grid = seq(1e-5, 30, length=100)

prob <- function(theta, K){
  Mt_vec = sapply(1:(N-1), Mt, theta, K)
  sum_of_squares <- sum(Mt_vec^2)
  square_of_sums <- sum(Mt_vec)^2
  0.5 ^ (N/2-1) * (sum_of_squares - square_of_sums/(N-1)) ^ (N/2-1) / gamma(N/2-1)
}

P_theta_K.X <- sapply(theta_grid, function(theta)
                sapply(K_grid, function(k) prob(theta, k)))

require(reshape2)
df = melt(P_theta_K.X)
names(df) = c("theta", "K", "lik")
ggplot(df, aes(theta_grid[theta], K_grid[K], z=-log(lik))) + stat_contour(aes(color=..level..), binwidth=3)

```

```{r include=FALSE}
# Conditional Distributions
tmpK <- sapply(K_grid, function(k) prob(1,k))
tmptheta <- sapply(theta_grid, prob, K=10)

# Modes
K_grid[which.min(tmpK)]
theta_grid[which.min(tmptheta)]

# Means
K_grid %*% -log(tmpK) / sum( -log(tmpK) )
theta_grid %*% -log(tmptheta) / sum( -log(tmptheta) )

qplot(theta_grid, -log(tmptheta))
qplot(K_grid, -log(tmpK))
```

### 2013-06-06 inst/examples/BUGS/process-noise-only.Rmd
whoops, didn't take log of mu

@@
   for(t in 1:(N-1)){
-    mu[t] <- y[t] * exp(r0 * (1 - y[t]/K)* (y[t] - theta) / K )
+    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K)* (y[t] - theta) / K 
     y[t+1] ~ dlnorm(mu[t], iQ) 
@@
   for(t in 1:(N-1)){
-    mu[t] <- y[t] * exp(r0 * (1 - y[t]/K) )
+    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K) 
     y[t+1] ~ dlnorm(mu[t], iQ) 
@@
   for(t in 1:(N-1)){
-    mu[t] <- r0 * pow(abs(y[t]), theta) / (1 + pow(abs(y[t]), theta) / K)
+    mu[t] <- log(r0)  + theta * abs(y[t]) - log(1 + pow(abs(y[t]), theta) / K)
     y[t+1] ~ dlnorm(mu[t], iQ) 

### 2013-06-06 inst/doc/beyond-mechanistic-models.tex
manuscript helper files

--- 2013-03-19 inst/doc/beyond-mechanistic-models.Rmd
@@
-```{r}
-read_chunk("nonparametric-bayes.R")
-```
-```{r knit-settings, echo=FALSE, include=FALSE, cache=FALSE}
-```
-```{r libraries, echo=FALSE, include=FALSE}
-```
-```{r plotting-options, echo=FALSE, include=FALSE}
-```
-```{r stateeq}
-```
-```{r sdp-pars, dependson="stateeq"}
-```
-```{r obs, dependson="sdp-pars"}
-```
-```{r mle, dependson="obs"}
-```
-```{r gp-priors}
-```
-```{r gp, dependson=c("obs", "gp-priors")}
-```
-```{r opt, dependson=c("gp", "mle")}
-```
-```{r sim, dependson="opt"}
-```
+\documentclass[author-year, review]{elsarticle} %review=doublespace preprint=single 5p=2 column
+%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
+\usepackage[hyphens]{url}
+\usepackage{lineno} % add 
+%\linenumbers % turns line numbering on 
+\bibliographystyle{elsarticle-harv}
+\biboptions{sort&compress} % For natbib
+\usepackage{graphicx}
+\usepackage{booktabs} % book-quality tables
+%% Redefines the elsarticle footer
+\makeatletter
+\def\ps@pprintTitle{%
+ \let\@oddhead\@empty
+ \let\@evenhead\@empty
+ \def\@oddfoot{\it \hfill\today}%
+ \let\@evenfoot\@oddfoot}
+\makeatother
 
-Process-based models built on modern machine learning approaches 
-offer a powerful and underappreciated approach for ecological
-forecasting and management.  We present a novel example of how
-one such example - non-parametric Gaussian processes, can be adapted
-to the context of management and decision-making.
+% A modified page layout
+\textwidth 6.75in
+\oddsidemargin -0.15in
+\evensidemargin -0.15in
+\textheight 9in
+\topmargin -0.5in
+%%%%%%%%%%%%%%%% end my additions to header
 
-While effective machine learning algorithms have speckled the literature, 
-most theoretical ecologists in particular have been wary of these 'black-box'
-approaches in favor of simple, mechanistic models.  While a blind application
-of machine learning approaches would indeed be dangerous grounds to stake 
-either ecological understanding or management practice, such concerns are not
-grounds to reject such methods out of hand.  Rather, further attention is 
-warranted to identify the circumstances in which machine-learning approaches
-will most likely outperform existing more mechanistic approaches, or help
-crack difficult management questions where ecological or dynamical complexity
-has caused traditional approaches to founder.
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \usepackage{fontspec}
+  \ifxetex
+    \usepackage{xltxtra,xunicode}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage{graphicx}
+% We will generate all images so they have a width \maxwidth. This means
+% that they will get their normal width if they fit onto the page, but
+% are scaled down if they would overflow the margins.
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
+\else\Gin@nat@width\fi}
+\makeatother
+\let\Oldincludegraphics\includegraphics
+\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={Beyond Mechanistic Models: The case for machine learning in ecological management},
+            colorlinks=true,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{0}
+% Pandoc toggle for numbering sections (defaults to be off)
+\setcounter{secnumdepth}{0}
+% Pandoc header
 
-Here we illustrate how a machine learning model can be implemented in place
-of simple mechanistic model in a decision theory (optimal control) problem,
-and highlight the features of this system that make the nonparametric approach
-particularly effective.
+\begin{document}
+\begin{frontmatter}
+  \title{Beyond Mechanistic Models: The case for machine learning in ecological
+         management}
+  \author[cstar]{Carl Boettiger\corref{cor1}}
+  \author[cstar]{Marc Mangel}
+  \author[noaa]{Stephan B. Munch}
+  \ead{cboettig@ucsc.edu}
+  \cortext[cor1]{Corresponding author, cboettig@ucsc.edu}
+  \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA 95064, USA}
+  \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}
+ \end{frontmatter}
+
+\section{Abstract}
+
+\section{Introduction}
+
+Process-based models built on modern machine learning approaches offer a
+powerful and underappreciated approach for ecological forecasting and
+management. We present a novel example of how one such example -
+non-parametric Gaussian processes, can be adapted to the context of
+management and decision-making.
+
+While effective machine learning algorithms have speckled the
+literature, most theoretical ecologists in particular have been wary of
+these `black-box' approaches in favor of simple, mechanistic models.
+While a blind application of machine learning approaches would indeed be
+dangerous grounds to stake either ecological understanding or management
+practice, such concerns are not grounds to reject such methods out of
+hand. Rather, further attention is warranted to identify the
+circumstances in which machine-learning approaches will most likely
+outperform existing more mechanistic approaches, or help crack difficult
+management questions where ecological or dynamical complexity has caused
+traditional approaches to founder.
+
+Here we illustrate how a machine learning model can be implemented in
+place of simple mechanistic model in a decision theory (optimal control)
+problem, and highlight the features of this system that make the
+nonparametric approach particularly effective.
 
 Mechanistic models have long been the gold standard of theoretical
-modeling in ecology (e.g. see @Geritz2012 or @Cuddington2013).  Only by
+modeling in ecology (e.g.~see @Geritz2012 or @Cuddington2013). Only by
 understanding the processes involved can we make reliable long term
 predictions and build an knowledge of cause and effect that guides the
-hypotheses we make, the data we collect, and the management decisions
-we make.  Process-based models, whether expressed in the language of
+hypotheses we make, the data we collect, and the management decisions we
+make. Process-based models, whether expressed in the language of
 mathematics or English, identify the connection between mosquitoes and
 the spread of malaria, or greenhouses gases and climate change, guiding
-our approach to understand and manage these threats. Despite this central
-importance, we argue here that ecologists would do well to give greater
-attention to the role non-mechanistic models can play in ecological
-management and decision making.  The value of these approaches is
-greatest in a context where decisions are made over short time horizons,
-and updated as new data becomes available.
+our approach to understand and manage these threats. Despite this
+central importance, we argue here that ecologists would do well to give
+greater attention to the role non-mechanistic models can play in
+ecological management and decision making. The value of these approaches
+is greatest in a context where decisions are made over short time
+horizons, and updated as new data becomes available.
 
-## The alternative to process-based models: pattern-based modeling
+\subsection{The alternative to process-based models: pattern-based
+modeling}
 
-Before we proceed further it would be useful to define our terms and
-lay some greater context for the discussion. Throughout, we will distinguish
-"mechanistic" or "process-based" models from models that are merely "correlative" 
-or "pattern-based". Historically pattern-based modeling meant regression (usually linear
-regression) and resided in methodology rooted in statistics departments
-and was the primary focus of empirical ecologists, while mechanistic
-modeling meant dynamical systems (ODEs, PDEs, SDEs and their discrete
-kin), residing in methodology from the mathematics department and was
-the primary focus of theoretical ecologists.
+Before we proceed further it would be useful to define our terms and lay
+some greater context for the discussion. Throughout, we will distinguish
+``mechanistic'' or ``process-based'' models from models that are merely
+``correlative'' or ``pattern-based''. Historically pattern-based
+modeling meant regression (usually linear regression) and resided in
+methodology rooted in statistics departments and was the primary focus
+of empirical ecologists, while mechanistic modeling meant dynamical
+systems (ODEs, PDEs, SDEs and their discrete kin), residing in
+methodology from the mathematics department and was the primary focus of
+theoretical ecologists.
 
@@
 mechanistic approaches could be critiqued on their tenuous connections
-to data -- model parameter values such as death rates, birth rates,
-etc. would be estimated in advance and then stuck into the model, rather
-than estimated in the context of the model itself.  Today both of these
+to data -- model parameter values such as death rates, birth rates, etc.
+would be estimated in advance and then stuck into the model, rather than
+estimated in the context of the model itself. Today both of these
 critiques are outdated. Computational power and hierarchical statistical
-methods (particularly approximate likelihood or simulation techniques such
-as particle filters and approximate Bayesian computing) have brought the
-inference richer dynamical systems into their fold, while pattern-based
-modeling has spawned an entirely new approach under the banner of machine
-learning that has divided the statistics community (See @Bremman2001,
-"The Two Cultures").   Machine learning models can represent almost
-arbitrarily complex patterns and incorporate learning and decision making
-strategies that have earned them the name 'artificial intelligence.'
-Yet at the same time they can be far less transparent their linear
-regression predecessors, making both astounding successes and startling
-biases and failures like their biological namesake. More ecologists,
-particularly theorists and modelers, would do well to learn both how to
-take advantage of these strengths and recognize their weaknesses.
-
-A simple example comparing the performance of the two modeling approaches
-in a quantitative decision problem will help to introduce these issues.
+methods (particularly approximate likelihood or simulation techniques
+such as particle filters and approximate Bayesian computing) have
+brought the inference richer dynamical systems into their fold, while
+pattern-based modeling has spawned an entirely new approach under the
+banner of machine learning that has divided the statistics community
+(See @Bremman2001, ``The Two Cultures''). Machine learning models can
+represent almost arbitrarily complex patterns and incorporate learning
+and decision making strategies that have earned them the name
+`artificial intelligence.' Yet at the same time they can be far less
+transparent their linear regression predecessors, making both astounding
+successes and startling biases and failures like their biological
+namesake. More ecologists, particularly theorists and modelers, would do
+well to learn both how to take advantage of these strengths and
+recognize their weaknesses.
 
 Mechanistic modeling emphasizes the importance of capturing the correct
-gross properties of a system over the tracking minute fluctuations.
-For instance, in selecting and parameterizing model of a population
-of conservation concern, we may be most interested in getting the
-long-term behavior correct -- such as identifying if the dynamics support
+gross properties of a system over tracking minute fluctuations. For
+instance, in selecting and parameterizing model of a population of
+conservation concern, we may be most interested in getting the long-term
+behavior correct -- such as identifying if the dynamics support
 persistence of the population -- rather than worrying how well they
@@
 it may come as a surprise to realize that such obviously wrong models
-can perform as well or even better than reasonable mechanistic models
-in guiding ecological management and decision making.
+can perform as well or even better than reasonable mechanistic models in
+guiding ecological management and decision making.
 
-```{r gp_plot, dependson=c("gp", "mle", "plotting-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
-Graph shows the expected value for the function $f$ under each model.
-Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.
-The training data is also shown as black points.
-(The GP is conditioned on 0,0, shown as a pseudo-data point).
-"}
-```
-```{r policies_plot, dependson=c("opt", "plotting-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."
-}
-```
-```{r sim_plot, dependson=c("sim", "plotting-options"), fig.cap="Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."
-}
-```
+A simple example comparing the performance of the two modeling
+approaches in a quantitative decision problem will help to introduce
+these issues.
 
-An example of such a comparison is shown in Figure 1, in which the optimal
-management decision is determined by stochastic dynamic programming
-algorithm using the biologically (a) plausible and (b) the implausible
-model of the population growth.
+\section{Models \& Methods}
 
-Three elements contribute to the biologically implausible phenomenological
-model performing better in this context:
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/gp_plot.pdf}
+\caption{Graph of the inferred Gaussian process compared to the true
+process and maximum-likelihood estimated process. Graph shows the
+expected value for the function $f$ under each model. Two standard
+deviations from the estimated Gaussian process covariance with (light
+grey) and without (darker grey) measurement error are also shown. The
+training data is also shown as black points. (The GP is conditioned on
+0,0, shown as a pseudo-data point).}
+\end{figure}
 
-### 1. Relevant state space
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/policies_plot.pdf}
+\caption{The steady-state optimal policy (infinite boundary) calculated
+under each model. Policies are shown in terms of target escapement,
+$S_t$, as under models such as this a constant escapement policy is
+expected to be optimal (Reed 1979).}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/sim_plot.pdf}
+\caption{Gaussian process inference outperforms parametric estimates.
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under
+the policies derived from each of the estimated models, as well as the
+policy based on the exact underlying model.}
+\end{figure}
+
+An example of such a comparison is shown in Figure 1, in which the
+optimal management decision is determined by stochastic dynamic
+programming algorithm using the biologically (a) plausible and (b) the
+implausible model of the population growth.
+
+\section{Discussion}
+
+Three elements contribute to the biologically implausible
+phenomenological model performing better in this context:
+
+\subsubsection{1. Relevant state space}
 
@@
 implausible model performs as well or better than the more plausible
-model.  This is the most obvious and immediate reason why the shortcomings
-that appear to make the model implausible do not make it useless.
-This effect is enhanced in our example because the management actions help
-drive the system towards rather than away from this region of state-space.
+model.
 
-### 2. Predictive accuracy
+This is the most obvious and immediate reason why the shortcomings that
+appear to make the model implausible do not make it useless. This effect
+is enhanced in our example because the management actions help drive the
+system towards rather than away from this region of state-space.
+
+\subsubsection{2. Predictive accuracy}
 
@@
 Mechanistic modeling is at its most powerful not when it is used to
-directly forecast future states but when it provides an understanding
-of how to approach a problem.  SIR-type models from the epidemiological
+directly forecast future states but when it provides an understanding of
+how to approach a problem. SIR-type models from the epidemiological
 literature are a good example. While the simplest SIR models have little
@@
 location, they provide a powerful understanding of the spread of an
-infection in terms of a single, biologically meaningful parameter: $R_0$,
-the basic reproductive number.  From the model, it becomes clear that
-management need not vaccinate every member of the population to stop
-the spread, but rather it suffices to vaccinate a sufficient fraction
-of the population to reduce $R_0$ below 1.
+infection in terms of a single, biologically meaningful parameter:
+$R_0$, the basic reproductive number. From the model, it becomes clear
+that management need not vaccinate every member of the population to
+stop the spread, but rather it suffices to vaccinate a sufficient
+fraction of the population to reduce $R_0$ below 1.
 
-### 3. Time scale for new data
+\subsubsection{3. Time scale for new data}
 
 In the sequential decision making problem we considered, we are
-presented with new data after each action.  The relevant timescale
-for the prediction is thus not the long-term dynamics, which would be
-wildly divergent, but the dynamics over this much shorter interval.
-While ecologists may be hesitant to base continual management on a model
-with obviously inaccurate long-term behavior, engineers tend to consider
-the problem in frequency space and gravitate to the opposite position --
-a good control model should prioritize high-frequency accuracy over low
+presented with new data after each action. The relevant timescale for
+the prediction is thus not the long-term dynamics, which would be wildly
+divergent, but the dynamics over this much shorter interval. While
+ecologists may be hesitant to base continual management on a model with
+obviously inaccurate long-term behavior, engineers tend to consider the
+problem in frequency space and gravitate to the opposite position -- a
+good control model should prioritize high-frequency accuracy over low
 frequency accuracy. The differences in intuition may arise from the
@@
 machine learning algorithm that makes accurate year-ahead predictions
-that capture some of the high-frequency fluctuations that appear only
-as noise in a mechanistic model of the long-term population dynamics.
+that capture some of the high-frequency fluctuations that appear only as
+noise in a mechanistic model of the long-term population dynamics.
 
-## Advantages of non-mechanistic models
+\subsection{Advantages of non-mechanistic models}
 
-1. Can better use all of the data available.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Can better use all of the data available.
+\end{enumerate}
 
-Incorporating various sources of information into mechanistic models
-can be an immensely difficult due to the increased complexity involved.
-Only thanks to tandem advances in increasing computational power and
+Incorporating various sources of information into mechanistic models can
+be an immensely difficult due to the increased complexity involved. Only
+thanks to tandem advances in increasing computational power and
 hierarchical statistical methodology have we been able to tackle such
@@
 from single-species based approaches. Without the need to formulate
-mechanisms, many modern machine learning algorithms can leverage potential
-information from all available sources of data directly.  The algorithms
-can recognize unanticipated or subtle patterns in large data sets
-that enable more accurate predictions than mechanistic models that are
-formulated at a more macroscopic level.  Such approaches depend critically
+mechanisms, many modern machine learning algorithms can leverage
+potential information from all available sources of data directly. The
+algorithms can recognize unanticipated or subtle patterns in large data
+sets that enable more accurate predictions than mechanistic models that
+are formulated at a more macroscopic level. Such approaches depend
+critically
 
-2. Better express ambiguity in regions where data is not available
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Better express ambiguity in regions where data is not available
+\end{enumerate}
 
-One of the greatest strengths of mechanistic models is their greatest weakness as well.
+One of the greatest strengths of mechanistic models is their greatest
+weakness as well.
 
-## Better the devil you know?  The dangers of pattern-based approaches
+\subsection{Better the devil you know? The dangers of pattern-based
+approaches}
 
-When all we have identified is a pattern rather than a mechanism, there is
-little way to guarantee that the pattern will not change in the future.
-Few examples illustrate this danger more broadly than the sub-prime
-mortgage crisis of 2008.  Simplifying greatly, the unquenchable demand for
-mortgage backed securities had shifted the underlying dynamics on which
-the phenomenological models were based.  Historical data indicated that
-mortgage-backed securities involved little risk. As demand exhausted
-the supply of traditional mortgages, it became profitable to invent
-new mortgages that could be sold to higher risk customers, such as
-those without proof of income.  The pattern-based models did not have
-enough time or data to learn about the higher default rate occurring as
-a result, but remained anchored too in the old patterns where default
-rates were low.  While a mechanistic formulation may have anticipated the
-resulting higher risk, the pattern based model must learn by seeing it.
-The economy still struggles to pay off that lesson.
+When all we have identified is a pattern rather than a mechanism, there
+is little way to guarantee that the pattern will not change in the
+future. Few examples illustrate this danger more broadly than the
+sub-prime mortgage crisis of 2008. Simplifying greatly, the unquenchable
+demand for mortgage backed securities had shifted the underlying
+dynamics on which the phenomenological models were based. Historical
+data indicated that mortgage-backed securities involved little risk. As
+demand exhausted the supply of traditional mortgages, it became
+profitable to invent new mortgages that could be sold to higher risk
+customers, such as those without proof of income. The pattern-based
+models did not have enough time or data to learn about the higher
+default rate occurring as a result, but remained anchored too in the old
+patterns where default rates were low. While a mechanistic formulation
+may have anticipated the resulting higher risk, the pattern based model
+must learn by seeing it. The economy still struggles to pay off that
+lesson.
 
-Of course mechanistic models may also fail to predict sudden
-structural changes.  It is all too easy to overlook some slowly changing
+Of course mechanistic models may also fail to predict sudden structural
+changes. It is all too easy to overlook some slowly changing
 environmental variable, particularly if temporal changes in a parameter
@@
 learning approaches are frequently less rigid in their assumptions, but
-can also less transparent about them.  This creates a second weakness
-in the machine learning approach.
+can also less transparent about them. This creates a second weakness in
+the machine learning approach.
 
-Mechanistic models are not free from this danger either. In fact they can
-be more susceptible to them for the very same reasons.  A mechanistic
+Mechanistic models are not free from this danger either. In fact they
+can be more susceptible to them for the very same reasons. A mechanistic
 model has the ability to more convincingly produce out-of-sample
@@
 
-Three elements contribute to this.
-
-A process-based understanding helps us formulate hypotheses, guide
-experimental design and data collection, predict outcomes, and influence
-management decisions.  The alternative to a processed-based model is  a
-correlative one. Statical models machine learning.
-
-> since the late 1980s there has been some consensus amongst ecologists
+\begin{quote}
+since the late 1980s there has been some consensus amongst ecologists
 that management decisions are best guided by models which are grounded
@@
 Starfield 1997, Jackson et al. 2000, Carpenter 2003, Nelson et al. 2008)
-
-machine learning algorithms
-
-Prediction without
-
-Example of Statistical model vs process-based model.
+\end{quote}
 
@@
 
-```{r posteriors, dependson=c("gp", "plotting-options"), fig.cap = "Histogram of posterior distributions for the estimated Gaussian Process shown in Figure 1.  Prior distributions overlaid."}
-```
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/posteriors.pdf}
+\caption{Histogram of posterior distributions for the estimated Gaussian
+Process shown in Figure 1. Prior distributions overlaid.}
+\end{figure}
+
+Reed, William J. 1979. ``Optimal escapement levels in stochastic and
+deterministic harvesting models.'' \emph{Journal of Environmental
+Economics and Management} 6 (4) (dec): 350--363.
+doi:10.1016/0095-0696(79)90014-7.
+
+\end{document}

### 2013-06-06 inst/examples/BUGS/custom-plots.Rmd
custom plots for exploring uncertainty

## Custom plotting

For presentation figures for allen.Rmd script

```{r }
model = "MLE"
ggplot(subset(dt, method == model)) + 
  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) +
  ylim(0, 9)
ggsave(paste(model, "_sim.pdf", sep=""), width=6, height=4)

```

```{r Figure3a}
dt$reps <- as.integer(dt$reps)

require(animation)
ani.options(loop=TRUE)
saveMovie({    
    for (i in seq(1, OptTime, by=2)) {
      print(
        ggplot(subset(dt, method %in% c("True", "Ricker") & reps < 5 & time <= i), 
               aes(time, fishstock, group=interaction(reps,method), color = method), alpha=.9) + 
          geom_line() + xlim(0, OptTime)
        )
    }
}, interval = 0.1, movie.name = "wrong-model.gif", ani.width = 600, ani.height = 600)

require(MASS)
s <- length(gp_dat$Cf_posterior)

# Samples from the posterior of GPs
draws <- sapply(1:s, function(i)
  mvrnorm(1, gp_dat$Ef_posterior[,i], gp_dat$Cf_posterior[[i]]))
draws <- melt(data.frame(x=x_grid, draws), id="x")

mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]])

## Samples from the expected GP
gp_draws <- mvrnorm(6, gp_dat$E_Ef, gp_dat$E_Cf)
gp_draws <- melt(data.frame(x=x_grid, t(gp_draws)), id="x")

ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x,y,ymin=ymin,ymax=ymax)) +
  # geom_line(data=gp_draws, aes(x=x, y=value, col=variable))  +  
  geom_line(data=draws, aes(x=x, y=value, index=variable), alpha=.1)

```

### 2013-06-06 inst/examples/semi-analytic-posteriors.Rmd
added allen model

@@
 source("~/.knitr_defaults.R")
+require(reshape2)
+require(data.table)
 opts_chunk$set(warning=TRUE, message=TRUE, cache=FALSE)
-opts_knit$set(upload.fun = socialR::flickr.url)
+#opts_knit$set(upload.fun = socialR::flickr.url)
 ```
@@
 
+Estimating the Allen model:
+
+$$X_{t+1} = X_t \exp\left(r \left(1 - \frac{X_t}{K} \right\left(\frac{X - C}{K}\right)\right) $$
+
+First re-parameterize so we can isolate an additive parameter, using standard quadratic form,
+
+$$X_{t+1} = X_t \exp(c + b X_t + a X_t^2) $$
+
+Where $c = \tfrac{-rC}{K}$, $b = \tfrac{r}{K}\left(\tfrac{C}{K} + 1\right)$, and $a = \tfrac{r}{K^2}$.  Then after integrating out $c$ and $\sigma$ we have
+
+```{r}
+Mt <- function(t, a, b)  log(x[t+1]) - log(x[t]) - b * x[t] + a * x[t] ^ 2
+```
+
+Choosing a sensible grid is a bit more tricky following the transformation, particularly $b$ can be negative or positive, and much larger in magnitude than any of the original parameters.
+$a$ on the other hand, can be reasonably restricted:
+
+```{r}
+b_grid = seq(0, .1, length=100)
+a_grid = seq(0, .1, length=100)
+```
+
+We define the probability as a function of the remaining two parameters,
+
+```{r}
+prob <- function(a, b){
+  Mt_vec = sapply(1:(N-1), Mt, a, b)
+  sum_of_squares <- sum(Mt_vec^2)
+  square_of_sums <- sum(Mt_vec)^2
+0.5 ^ (N/2-1) * (sum_of_squares - square_of_sums/(N-1)) ^ (N/2-1) / gamma(N/2-1)
+}
+```
+
+and loop over the grid on each
+
+```{r}
+P_a_b.X <- sapply(a_grid, function(a)
+                sapply(b_grid, function(b) prob(a, b)))
+
+````
+
+We summarize with a contour plot
+
+```{r}
+  df = melt(P_a_b.X)
+names(df) = c("a", "b", "lik")
+ggplot(df, aes(a_grid[a], b_grid[b], z=-log(lik))) + 
+  stat_contour(aes(color=..level..), binwidth=1) +
+  coord_cartesian(ylim=c(0, .025), xlim=c(0,.05))
+
+```
+
+```{r include=FALSE}
+# Conditional Distributions
+tmpA <- sapply(a_grid, prob, b=0.1)
+tmpB <- sapply(b_grid, function(b) prob(a=0.01, b=b))
+
+# Modes
+a_grid[which.min(tmpA)]
+b_grid[which.min(tmpB)]
+
+qplot(a_grid, -log(tmpA))
+qplot(b_grid, -log(tmpB))
+```
+
 Estimating the Myers model on this data:
+
+$$X_{t+1} = Z_t \frac{r X_t^{\theta}}{1 + X_t^{\theta} / K}$$
+
+With $Z_t$ lognormal, unit mean, std $\sigma$.
 

### 2013-06-07 inst/examples/BUGS/process-noise-only.Rmd
try searching in non-transformed paramaters on allen model

@@
 ```{r allen-model-local}
-K_prior_p <- c(0.01, 40.0)
-logr0_prior_p <- c(-3.0, 6.0)
-logtheta_prior_p <- c(-6.0, 6.0)
+K_prior_p <- c(0.01, 20.0)
+r0_prior_p <- c(0.01, 6.0)
+theta_prior_p <- c(0.01, 20.0)
 
@@
   K     ~ dunif(%s, %s)
-  logr0    ~ dunif(%s, %s)
-  logtheta ~ dunif(%s, %s)
+  r0    ~ dunif(%s, %s)
+  theta ~ dunif(%s, %s)
   stdQ ~ dunif(%s, %s)", 
   K_prior_p[1], K_prior_p[2],
-  logr0_prior_p[1], logr0_prior_p[2],
-  logtheta_prior_p[1], logtheta_prior_p[2],
+  r0_prior_p[1], r0_prior_p[2],
+  theta_prior_p[1], theta_prior_p[2],
   stdQ_prior_p[1], stdQ_prior_p[2]),
@@
   iQ <- 1 / (stdQ * stdQ);
-  r0 <- exp(logr0)
-  theta <- exp(logtheta)
   y[1] ~ dunif(0, 10)
@@
 K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
-logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
-logtheta_prior <- function(x) dunif(x, logtheta_prior_p[1], logtheta_prior_p[2])
+r0_prior <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
+theta_prior <- function(x) dunif(x, theta_prior_p[1], theta_prior_p[2])
 par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
@@
 ```{r allen-mcmc-local, dependson=c("allen-model-local", "allen-pars", "jags-setup"), results="hide"}
-jags.params=c("K","logr0","logtheta","stdQ") # be sensible about the order here
+jags.params=c("K","r0","theta","stdQ") # be sensible about the order here
 jags.inits <- function(){
   list("K"= 8 * rlnorm(1,0, 0.1),
-       "logr0"=log(2 * rlnorm(1,0, 0.1) ),
-       "logtheta"=log(  5 * rlnorm(1,0, 0.1) ), 
+       "r0"=log(2 * rlnorm(1,0, 0.1) ),
+       "theta"=log(  5 * rlnorm(1,0, 0.1) ), 
        "stdQ"= abs( 0.1 * rlnorm(1,0, 0.1)),
@@
 
-```{r allen-output, dependson=c("posterior-mode", "allen-traces")}
+```{r allen-output-local, dependson=c("posterior-mode", "allen-traces")}
+A <- allen_posteriors
+A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
+pardist <- acast(A, index ~ variable)
+bayes_coef <- apply(pardist,2, posterior.mode) 
+bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"], bayes_coef["theta"])) # parameters formatted for f
+allen_f <- function(x,h,p) unname(RickerAllee(x,h, unname(p[c("r0", "K", "theta")])))
+allen_means <- sapply(x_grid, f, 0, bayes_pars)
+bayes_pars
+head(pardist)
 ```
@@
 
-```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
+```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output-local", "gp-output", "mle-output")}
 ```
@@
 
-```{r allen-opt, dependson="allen-output"}
+```{r allen-opt, dependson="allen-output-local"}
 ```
@@
 
-```{r deviances, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output") }
+```{r deviances, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output-local", "gp-output", "mle-output") }
 ```

### 2013-06-07 inst/examples/BUGS/allen.Rmd
different mcmc conditions

@@
 f <- RickerAllee
-p <- c(1, 10, 5)
+p <- c(2, 8, 5)
 K <- 10  # approx, a li'l' less
@@
 }
-par <- c(p[1]+abs(rnorm(1,0,1)), 
-         p[2]+rnorm(1,0,.1), 
-         p[3]+rnorm(1,0, .1), 
-         sigma_g + abs(rnorm(1,0,.1)))
+par <- c(p[1]*rlnorm(1,0,.4), 
+         p[2]*rlnorm(1,0,.3), 
+         p[3]*rlnorm(1,0, .3), 
+         sigma_g * rlnorm(1,0,.3))
 o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
@@
 ```{r gp_traces_densities, dependson="gp"}
-plots <- summary_gp_mcmc(gp)
+plots <- summary_gp_mcmc(gp, burnin=1e4, thin=300)
 ```
@@
 
-```{r jags-setup}
+```{r jags-setup, dependson="obs"}
 y <- obs$x[-1] 
@@
 n.chains <- 2
-n.iter <- 1e5
+n.iter <- 40000
 n.burnin <- floor(10000)
@@
 ```{r common-priors}
-stdQ_prior_p <- c(0.0001, 100)
-stdR_prior_p <- c(1e-4, .1)
+stdQ_prior_p <- c(1e-6, 100)
+stdR_prior_p <- c(1e-6, .1)
 stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
@@
 jags.inits <- function(){
-  list("K"= 10 * rlnorm(1,0,.1),
-       "logr0"=log(rlnorm(1,0,.5) ),
-       "logtheta"=log(2 * rlnorm(1,0,.1)), 
-       "stdQ"= 0.2 * rlnorm(1,0,.1),
+  list("K"= 8 * rlnorm(1,0,.1),
+       "logr0"=log(2 * rlnorm(1,0,.1) ),
+       "logtheta"=log(5 * rlnorm(1,0,.1)), 
+       "stdQ"= 0.05 * rlnorm(1,0,.1),
        "stdR"= 1e-5 * rlnorm(1,0,.1),
@@
 jags.inits <- function(){
-  list("K"= 10 * rlnorm(1,0,.5),
-       "logr0" = log( rlnorm(1,0,.5) ),
-       "stdQ"= 0.2 * rlnorm(1,0,.2),
+  list("K"= 10 * rlnorm(1,0,.2),
+       "logr0" = log( rlnorm(1,0,.2) ),
+       "stdQ"= 0.01 * rlnorm(1,0,.1),
        "stdR"= 1e-5 * rlnorm(1,0,.1),
@@
 ```{r myers-output, dependson="myers-traces"}
-myers_pardist <- acast(myers_posteriors[2:3], 
-                        1:table(myers_posteriors$variable) ~ variable, 
-                        subset=.(variable!="deviance")) 
-myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
-myers_pardist[,2] = exp(myers_pardist[,2]) # transform model parameters back first
-myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
-colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
+A <- myers_posteriors
+A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
+myers_pardist <- acast(A, index ~ variable)
+### myers_pardist <- acast(myers_posteriors[2:3], 1:table(myers_posteriors$variable) ~ variable) 
+myers_pardist[,"logK"] = exp(myers_pardist[,"logK"]) # transform model parameters back first
+myers_pardist[,"logr0"] = exp(myers_pardist[,"logr0"]) # transform model parameters back first
+myers_pardist[,"logtheta"] = exp(myers_pardist[,"logtheta"]) # transform model parameters back first
+colnames(myers_pardist)[colnames(myers_pardist)=="logK"] = "K"
+colnames(myers_pardist)[colnames(myers_pardist)=="logr0"] = "r0"
+colnames(myers_pardist)[colnames(myers_pardist)=="logtheta"] = "theta"
+
 bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
@@
 ```{r deviances, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
-allen_deviance <- pardist[,'deviance']
-ricker_deviance <- ricker_pardist[,'deviance']
-myers_deviance <- myers_pardist[,'deviance']
+allen_deviance <- posterior.mode(pardist[,'deviance'])
+ricker_deviance <- posterior.mode(ricker_pardist[,'deviance'])
+myers_deviance <- posterior.mode(myers_pardist[,'deviance'])
 true_deviance <- 2*estf(c(p, sigma_g))

### 2013-06-07 inst/examples/BUGS/parametric-vs-nonparametric.Rmd
example set up based on Myers model simulations

@@
 library(nonparametricbayes) 
+opts_chunk$set(external=TRUE)
+read_chunk("external-chunks.R")
 ```
 
-```{r}
-posterior.mode <- function(x) {
-  ux <- unique(x)
-  ux[which.max(tabulate(match(x, ux)))]
-}
-
+```{r posterior-mode}
 ```
@@
 
-Uses the model derived in `citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
-
-```{r stateeq}
-f <- RickerAllee
-p <- c(1, 10, 5)
-K <- 10  # approx, a li'l' less
-allee <- 5 # approx, a li'l' less
+```{r stateeq-local}
+f <- Myers
+p <- c(1, 2, 6)
+K <- 5  # approx, a li'l' less
+allee <- 1.2 # approx, a li'l' less
 ```
@@
 
-```{r sdp-pars, dependson="stateeq"}
-sigma_g <- 0.05
-sigma_m <- 0.0
-z_g <- function() rlnorm(1, 0, sigma_g)
-z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
-x_grid <- seq(0, 1.5 * K, length=50)
-h_grid <- x_grid
-profit <- function(x,h) pmin(x, h)
-delta <- 0.01
-OptTime <- 50  # stationarity with unstable models is tricky thing
-reward <- 0
-xT <- 0
-Xo <-  allee+.5# observations start from
-x0 <- K # simulation under policy starts from
-Tobs <- 40
+```{r sdp-pars, dependson="stateeq-local"}
 ```
@@
 ```{r mle, dependson="obs"}
-set.seed(12345)
-estf <- function(p){ 
-    mu <- f(obs$x,0,p)
-    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
-}
-par <- c(p[1]+abs(rnorm(1,0,1)), 
-         p[2]+rnorm(1,0,.1), 
-         p[3]+rnorm(1,0, .1), 
-         sigma_g + abs(rnorm(1,0,.1)))
-o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
-f_alt <- f
-p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
-sigma_g_alt <- as.numeric(o$par[4])
-
-est <- list(f = f_alt, p = p_alt, sigma_g = sigma_g_alt, mloglik=o$value)
 ```
@@
 ```{r jags-setup}
-y <- obs$x[-1] 
-N <- length(y);
-jags.data <- list("N","y")
-n.chains <- 6
-n.iter <- 5e5
-n.burnin <- floor(10000)
-n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
-n.update <- 10
 ```
@@
 ```{r common-priors}
-stdQ_prior_p <- c(0.0001, 100)
-stdR_prior_p <- c(1e-6, .1)
-stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
-stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
 ```
@@
 ```{r allen-model}
-K_prior_p <- c(0.01, 40.0)
-logr0_prior_p <- c(-6.0, 6.0)
-logtheta_prior_p <- c(-6.0, 6.0)
-
-bugs.model <- 
-paste(sprintf(
-"model{
-  K     ~ dunif(%s, %s)
-  logr0    ~ dunif(%s, %s)
-  logtheta ~ dunif(%s, %s)
-  stdQ ~ dunif(%s, %s)
-  stdR ~ dunif(%s, %s)", 
-  K_prior_p[1], K_prior_p[2],
-  logr0_prior_p[1], logr0_prior_p[2],
-  logtheta_prior_p[1], logtheta_prior_p[2],
-  stdQ_prior_p[1], stdQ_prior_p[2],
-  stdR_prior_p[1], stdR_prior_p[2]),
-
-  "
-  iQ <- 1 / (stdQ * stdQ);
-  iR <- 1 / (stdR * stdR);
-  r0 <- exp(logr0)
-  theta <- exp(logtheta)
-  x[1] ~ dunif(0, 10)
-  for(t in 1:(N-1)){
-    mu[t] <- x[t] * exp(r0 * (1 - x[t]/K)* (x[t] - theta) / K )
-    x[t+1] ~ dnorm(mu[t], iQ) 
-  }
-  for(t in 1:(N)){
-    y[t] ~ dnorm(x[t], iR)
-  }
-}")
-writeLines(bugs.model, "allen.bugs")
 ```
@@
 ```{r allen-priors, dependson="common-priors"}
-K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
-logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
-logtheta_prior <- function(x) dunif(x, logtheta_prior_p[1], logtheta_prior_p[2])
-par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
-                    logr0 = logr0_prior, logtheta = logtheta_prior,
-                    stdQ = stdQ_prior, stdR = stdR_prior)
 ```
@@
 ```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results="hide"}
-jags.params=c("K","logr0","logtheta","stdQ", "stdR") # be sensible about the order here
-
-# Can give a list of starting values or a function that can create random values
-
-jags.inits <- function(){
-  list("K"= 10 * rlnorm(1,0,.1),
-       "logr0"=log(rlnorm(1,0,.5) ),
-       "logtheta"=log(2 * rlnorm(1,0,.1)), 
-       "stdQ"= 0.2 * rlnorm(1,0,.1),
-       "stdR"= 1e-5 * rlnorm(1,0,.1),
-       "x"=y,        
-       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
-}
-
-set.seed(1234)
-# parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
-allen_jags <- do.call(jags.parallel, list(data=jags.data, inits=jags.inits, 
-                                          jags.params, n.chains=n.chains, 
-                                          n.iter=n.iter, n.thin=n.thin, 
-                                          n.burnin=n.burnin, model.file="allen.bugs"))
-
-#Run again iteratively if we haven't met the Gelman-Rubin convergence criterion
-recompile(allen_jags) # required for parallel
-allen_jags <- do.call(autojags, list(object=allen_jags, n.update=n.update, 
-                                     n.iter=n.iter, n.thin = n.thin))
 ```
@@
 ```{r allen-traces, dependson="allen-mcmc"}
-tmp <- lapply(as.mcmc(allen_jags), as.matrix) # strip classes the hard way...
-allen_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
-names(allen_posteriors) = c("index", "variable", "value", "chain")
-ggplot(allen_posteriors) + geom_line(aes(index, value)) + 
-  facet_wrap(~ variable, scale="free", ncol=1)
 ```
@@
 
-```{r allen-output, dependson="allen-traces"}
-pardist <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable)[1] ~ variable) 
-pardist[,"logr0"] = exp(pardist[,"logr0"]) # transform model parameters back first
-pardist[,"logtheta"] = exp(pardist[,"logtheta"])
-colnames(pardist)[colnames(pardist)=="logtheta"] = "theta"
-colnames(pardist)[colnames(pardist)=="logr0"] = "r0"
-bayes_coef <- apply(pardist,2, posterior.mode) 
-bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"], bayes_coef["theta"])) # parameters formatted for f
-allen_means <- sapply(x_grid, f, 0, bayes_pars)
-
+```{r allen-output, dependson=c("posterior-mode", "allen-traces")}
 ```
@@
 ```{r ricker-model}
-K_prior_p <- c(0.01, 40.0)
-logr0_prior_p <- c(-6.0, 6.0)
-
-bugs.model <- 
-paste(sprintf(
-"model{
-  K    ~ dunif(%s, %s)
-  logr0    ~ dunif(%s, %s)
-  stdQ ~ dunif(%s, %s)
-  stdR ~ dunif(%s, %s)", 
-  K_prior_p[1], K_prior_p[2],
-  logr0_prior_p[1], logr0_prior_p[2],
-  stdQ_prior_p[1], stdQ_prior_p[2],
-  stdR_prior_p[1], stdR_prior_p[2]),
-
-  "
-  iQ <- 1 / (stdQ * stdQ);
-  iR <- 1 / (stdR * stdR);
-  r0 <- exp(logr0)
-  x[1] ~ dunif(0, 10)
-  for(t in 1:(N-1)){
-    mu[t] <- x[t] * exp(r0 * (1 - x[t]/K) )
-    x[t+1] ~ dnorm(mu[t], iQ) 
-  }
-  for(t in 1:(N)){
-    y[t] ~ dnorm(x[t], iR)
-  }
-}")
-writeLines(bugs.model, "ricker.bugs")
 ```
@@
 ```{r ricker-priors, dependson="common-priors"}
-K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
-logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
-par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
-                   logr0 = logr0_prior, stdQ = stdQ_prior, stdR = stdR_prior)
 ```
@@
 We define which parameters to keep track of, and set the initial values of
-parameters in the transformed space used by the MCMC.  We use logarithms
-to maintain strictly positive values of parameters where appropriate.
+parameters in the transformed space used by the MCMC.
 
-```{r ricker-mcmc, dependson="ricker-model", results="hide"}
-# Uniform priors on standard deviation terms
-jags.params=c("K","logr0", "stdQ", "stdR")
-jags.inits <- function(){
-  list("K"= 10 * rlnorm(1,0,.5),
-       "logr0" = log( rlnorm(1,0,.5) ),
-       "stdQ"= 0.2 * rlnorm(1,0,.2),
-       "stdR"= 1e-5 * rlnorm(1,0,.1),
-       "x"=y,
-       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
-}
-set.seed(12345) 
-ricker_jags <- do.call(jags.parallel, 
-                       list(data=jags.data, inits=jags.inits, 
-                            jags.params, n.chains=n.chains, 
-                            n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
-                            model.file="ricker.bugs"))
-recompile(ricker_jags)
-ricker_jags <- do.call(autojags, 
-                       list(object=ricker_jags, n.update=n.update, n.iter=n.iter, 
-                            n.thin = n.thin))
+```{r ricker-mcmc, dependson="ricker-model"}
 ```
@@
 
-```{r ricker_traces, dependson="ricker-mcmc"}
-tmp <- lapply(as.mcmc(ricker_jags), as.matrix) # strip classes the hard way...
-ricker_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
-names(ricker_posteriors) = c("index", "variable", "value", "chain")
-
-ggplot(ricker_posteriors) + geom_line(aes(index, value)) + 
-  facet_wrap(~ variable, scale="free", ncol=1)
+```{r ricker-traces, dependson="ricker-mcmc"}
 ```
 
-```{r ricker_posteriors, dependson=c("ricker-traces", "ricker-priors")}
-ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100) 
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-# plot posterior distributions
-ggplot(ricker_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=2)
+```{r ricker-posteriors, dependson=c("ricker-traces", "ricker-priors")}
 ```
@@
 
-```{r ricker-output, dependson="ricker-traces"}
-ricker_pardist <- acast(ricker_posteriors[2:3], 
-                        1:table(ricker_posteriors$variable) ~ variable, 
-                        subset=.(variable!="deviance")) 
-ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first
-bayes_coef <- apply(ricker_pardist,2, posterior.mode) # much better estimates from mode then mean
-ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
-ricker_means <- sapply(x_grid, Ricker, 0, ricker_bayes_pars[c(1,2)])
+```{r ricker-output, dependson=c("posterior-mode", "ricker-traces")}
 ```
@@
 ```{r myers-model}
-logr0_prior_p <- c(-6.0, 6.0)
-logtheta_prior_p <- c(-6.0, 6.0)
-logK_prior_p <- c(-6.0, 6.0)
-
-bugs.model <- 
-paste(sprintf(
-"model{
-  logr0    ~ dunif(%s, %s)
-  logtheta    ~ dunif(%s, %s)
-  logK    ~ dunif(%s, %s)
-  stdQ ~ dunif(%s, %s)
-  stdR ~ dunif(%s, %s)", 
-  logr0_prior_p[1], logr0_prior_p[2],
-  logtheta_prior_p[1], logtheta_prior_p[2],
-  logK_prior_p[1], logK_prior_p[2],
-  stdQ_prior_p[1], stdQ_prior_p[2],
-  stdR_prior_p[1], stdR_prior_p[2]),
-
-  "
-  iQ <- 1 / (stdQ * stdQ)
-  iR <- 1 / (stdR * stdR)
-  r0 <- exp(logr0)
-  theta <- exp(logtheta)
-  K <- exp(logK)
-
-  x[1] ~ dunif(0, 10)
-
-  for(t in 1:(N-1)){
-    mu[t] <- r0 * pow(abs(x[t]), theta) / (1 + pow(abs(x[t]), theta) / K)
-    x[t+1] ~ dnorm(mu[t], iQ) 
-  }
-
-  for(t in 1:(N)){
-    y[t] ~ dnorm(x[t], iR)
-  }
-}")
-writeLines(bugs.model, "myers.bugs")
 ```
@@
 ```{r myers-priors}
-logK_prior     <- function(x) dunif(x, logK_prior_p[1], logK_prior_p[2])
-logr_prior     <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
-logtheta_prior <- function(x) dunif(x, logtheta_prior_p[1], logtheta_prior_p[2])
-par_priors <- list( deviance = function(x) 0 * x, logK = logK_prior,
-                    logr0 = logr_prior, logtheta = logtheta_prior, 
-                    stdQ = stdQ_prior, stdR = stdR_prior)
-
 ```
 
-```{r myers-mcmc, dependson="myers-model", results="hide"}
-jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
-jags.inits <- function(){
-  list("logr0" 		= log(rlnorm(1, 0, .3)), 
-       "logK" 		= log(10 * rlnorm(1, 1, .1)),
-       "logtheta" = log(2 * rnorm(1, 1, .1)),  
-       "stdQ"			= 0.2  * rlnorm(1, 0, .1),
-       "stdR"			= 1e-5 * rlnorm(1, 0, .1),
-       "x" 				= y,
-       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
-}
-set.seed(12345)
-myers_jags <- do.call(jags.parallel, 
-                      list(data=jags.data, inits=jags.inits, jags.params, 
-                           n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
-                           n.burnin=n.burnin, model.file="myers.bugs"))
-recompile(myers_jags)
-myers_jags <- do.call(autojags, 
-                      list(object=myers_jags, n.update=n.update, n.iter=n.iter, 
-                           n.thin = n.thin))
+```{r myers-mcmc, dependson="myers-model"}
 ```
@@
 ```{r myers-posteriors, dependson="myers-traces"}
-par_prior_curves <- ddply(myers_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100) 
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-
-ggplot(myers_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=3)
 ```
 
-```{r myers-output, dependson="myers-traces"}
-myers_pardist <- acast(myers_posteriors[2:3], 
-                        1:table(myers_posteriors$variable) ~ variable, 
-                        subset=.(variable!="deviance")) 
-# robust to ordering
-myers_pardist[,1] = exp(myers_pardist[,"logK"]) # transform model parameters back first
-myers_pardist[,2] = exp(myers_pardist[,"logr0"]) # transform model parameters back first
-myers_pardist[,3] = exp(myers_pardist[,"logtheta"]) # transform model parameters back first
-
-colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR") # assumes they are in this order...
-bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
-myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
-myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
+```{r myers-output, dependson=c("posterior-mode", "myers-traces")}
 ```
@@
 ```{r totalprofits, dependson="profits"}
-ggplot(Profit, aes(V1)) + geom_histogram() + 
-  facet_wrap(~method, scales = "free_y") + guides(legend.position = "none") + xlab("Total profit by replicate")
 ```
 
-```{r}
-df <- acast(allen_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable) 
-modes <- apply(df, 2, posterior.mode)
-allen_deviance <- modes[['deviance']]
-
-df <- acast(ricker_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable)
-modes <- apply(df, 2, posterior.mode)
-ricker_deviance <- modes[['deviance']]
-
-df <- acast(myers_posteriors[2:3], 1:table(allen_posteriors$variable) ~ variable)
-modes <- apply(df, 2, posterior.mode)
-myers_deviance <- modes[['deviance']]
-
-true_deviance <- 2*estf(c(p, sigma_g))
-mle_deviance <- 2*estf(c(est$p, est$sigma_g))
-
-c(allen = allen_deviance, ricker=ricker_deviance, myers=myers_deviance, true=true_deviance, mle=mle_deviance)
-
+```{r deviances, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output") }
 ```

### 2013-06-07 inst/examples/BUGS/external-chunks.R
external chunks uses non-transformed vars now

@@ -353,8 +353,8 @@ jags.params=c("r0", "theta", "K", "stdQ")
 jags.inits <- function(){
   list("r0"= rlnorm(1,0,.1), 
        "K"=    10 * rlnorm(1,0,.1),
-       "theta" = 2 * rlnorm(1,0,.1),  
-       "stdQ"= sqrt(0.5) * rlnorm(1,0,.1),
+       "theta" = 1 * rlnorm(1,0,.1),  
+       "stdQ"= sqrt(0.2) * rlnorm(1,0,.1),
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
 }
 set.seed(12345)

### 2013-06-07 inst/examples/BUGS/myers.Rmd
myers model with diff parameters from parametric-vs-nonparametric at this commit hash

--- 2013-05-22 inst/examples/BUGS/myers.Rmd
@@
 ```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
-setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
+source("~/.knitr_defaults.R")
+#opts_knit$set(upload.fun = socialR::flickr.url)
 library(knitcitations)
-library(ggplot2) 
 library(nonparametricbayes) 
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA,
-               fig.width=6, fig.height=4, cache.path="myers/")
-opts_knit$set(upload.fun = socialR::flickr.url)
-theme_set(theme_bw(base_size=12))
-theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
-             plot.background = element_rect(fill = "transparent", colour = NA))
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
-               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+opts_chunk$set(external=TRUE)
+read_chunk("external-chunks.R")
 ```
 
-```{r}
-posterior.mode <- function(x) {
-  ux <- unique(x)
-  ux[which.max(tabulate(match(x, ux)))]
-}
-
+```{r posterior-mode}
 ```
@@
 
-Uses the model derived in `r citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
-
-```{r stateeq}
+```{r stateeq-local}
 f <- Myers
-p <- c(1, 2, 6)
+p <- c(1.5, 2.5, 8)
 K <- 5  # approx, a li'l' less
@@
 
-Various parameters defining noise dynamics, grid, and policy costs.
-
-```{r sdp-pars, dependson="stateeq"}
-sigma_g <- 0.05
-sigma_m <- 0.0
-z_g <- function() rlnorm(1, 0, sigma_g)
-z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
-x_grid <- seq(0, 1.5 * K, length=50)
-h_grid <- x_grid
-profit <- function(x,h) pmin(x, h)
-delta <- 0.01
-OptTime <- 50  # stationarity with unstable models is tricky thing
-reward <- 0
-xT <- 0
-Xo <- K # observations start from
-x0 <- Xo # simulation under policy starts from
-Tobs <- 40
+```{r sdp-pars, dependson="stateeq-local"}
 ```
@@
 ```{r obs, dependson="sdp-pars"}
-  set.seed(1234)
-  #harvest <- sort(rep(seq(0, .5, length=7), 5))
-  x <- numeric(Tobs)
-  x[1] <- 2
-  nz <- 1
-  for(t in 1:(Tobs-1))
-    x[t+1] = z_g() * f(x[t], h=0, p=p)
-  obs <- data.frame(x = c(rep(0,nz), 
-                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
-                    y = c(rep(0,nz), 
-                          x[2:Tobs]))
-raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
-raw_plot
 ```
@@
 ```{r mle, dependson="obs"}
-est <- par_est_allee(obs, f, p,  
-                     init = c(r = p[1] + 1, 
-                              K = p[2] + 2, 
-                              C = p[3] + 2, 
-                              s = sigma_g))
 ```
@@
 
-```{r mle-output}
-true_means <- sapply(x_grid, f, 0, p)
-est_means <- sapply(x_grid, est$f, 0, est$p)
+```{r mle-output, dependson="mle"}
 ```
@@
 ```{r jags-setup}
-y <- obs$x[-1] 
-N <- length(y);
-jags.data <- list("N","y")
-n.chains <- 1
-n.iter <- 40000
-n.burnin <- floor(10000)
-n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
 ```
@@
 ```{r common-priors}
-stdQ_prior_p <- c(0.0001, 100)
-stdR_prior_p <- c(0.0001, 100)
-stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
-stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
 ```
@@
 
-```{r allen-jags}
-K_prior_p <- c(0.01, 40.0)
-logr0_prior_p <- c(-6.0, 6.0)
-logtheta_prior_p <- c(-6.0, 6.0)
-
-bugs.model <- 
-paste(sprintf(
-"model{
-  K     ~ dunif(%s, %s)
-  logr0    ~ dunif(%s, %s)
-  logtheta ~ dunif(%s, %s)
-  stdQ ~ dunif(%s, %s)
-  stdR ~ dunif(%s, %s)", 
-  K_prior_p[1], K_prior_p[2],
-  logr0_prior_p[1], logr0_prior_p[2],
-  logtheta_prior_p[1], logtheta_prior_p[2],
-  stdQ_prior_p[1], stdQ_prior_p[2],
-  stdR_prior_p[1], stdR_prior_p[2]),
-
-  "
-  iQ <- 1 / (stdQ * stdQ);
-  iR <- 1 / (stdR * stdR);
-  r0 <- exp(logr0)
-  theta <- exp(logtheta)
-  x[1] ~ dunif(0, 10)
-  for(t in 1:(N-1)){
-    mu[t] <- x[t] * exp(r0 * (1 - x[t]/K)* (x[t] - theta) / K )
-    x[t+1] ~ dnorm(mu[t], iQ) 
-  }
-  for(t in 1:(N)){
-    y[t] ~ dnorm(x[t], iR)
-  }
-}")
-writeLines(bugs.model, "allen.bugs")
+```{r allen-model}
 ```
@@
 
-```{r allen-priors}
-K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
-logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
-logtheta_prior <- function(x) dunif(x, -6, 6)
-par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
-                    logr0 = logr0_prior, logtheta = logtheta_prior,
-                    stdQ = stdQ_prior, stdR = stdR_prior)
+```{r allen-priors, dependson="common-priors"}
 ```
@@
 
-```{r allen-mcmc}
-jags.params=c("K","logr0","logtheta","stdQ", "stdR")
-jags.inits <- function(){
-  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), 
-       "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,
-       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
-}
-
-set.seed(12345)
-jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
-                n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin, 
-                model.file="allen.bugs", progress.bar="none")
+```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results="hide"}
 ```
 
-#### Convergence diagnostics for parametric bayes
+#### Convergence diagnostics for Allen model
 
-```{r allen-traces}
-jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-allen_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
-ggplot(allen_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
+R notes: this strips classes from the `mcmc.list` object (so that we have list of matrices; objects that `reshape2::melt` can handle intelligently), and then combines chains into one array.
+In this array each parameter is given its value at each sample from the posterior (index) for each chain.
+
+```{r allen-traces, dependson="allen-mcmc"}
 ```
 
-```{r allen-posteriors}
-allen_priors <- ddply(allen_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100) 
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-
-ggplot(allen_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=allen_priors, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=3)
+```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
 ```
@@
 
-```{r allen-output}
-pardist <- acast(allen_posteriors, index ~ variable, subset=.(variable != "deviance") )
-pardist[,2] = exp(pardist[,2]) # transform model parameters back first
-pardist[,3] = exp(pardist[,3])
-bayes_coef <- apply(pardist,2, posterior.mode) 
-bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3])) # parameters formatted for f
-allen_means <- sapply(x_grid, f, 0, bayes_pars)
+```{r allen-output, dependson=c("posterior-mode", "allen-traces")}
 ```
@@
 
-```{r ricker-jags}
-K_prior_p <- c(0.01, 40.0)
-logr0_prior_p <- c(-6.0, 6.0)
-
-bugs.model <- 
-paste(sprintf(
-"model{
-  K    ~ dunif(%s, %s)
-  logr0    ~ dunif(%s, %s)
-  stdQ ~ dunif(%s, %s)
-  stdR ~ dunif(%s, %s)", 
-  K_prior_p[1], K_prior_p[2],
-  logr0_prior_p[1], logr0_prior_p[2],
-  stdQ_prior_p[1], stdQ_prior_p[2],
-  stdR_prior_p[1], stdR_prior_p[2]),
-
-  "
-  iQ <- 1 / (stdQ * stdQ);
-  iR <- 1 / (stdR * stdR);
-  r0 <- exp(logr0)
-  x[1] ~ dunif(0, 10)
-  for(t in 1:(N-1)){
-    mu[t] <- x[t] * exp(r0 * (1 - x[t]/K) )
-    x[t+1] ~ dnorm(mu[t], iQ) 
-  }
-  for(t in 1:(N)){
-    y[t] ~ dnorm(x[t], iR)
-  }
-}")
-writeLines(bugs.model, "ricker.bugs")
+```{r ricker-model}
 ```
@@
 
-```{r ricker-priors}
-K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
-logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
-par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
-                   logr0 = logr0_prior, stdQ = stdQ_prior, stdR = stdR_prior)
+```{r ricker-priors, dependson="common-priors"}
 ```
@@
 We define which parameters to keep track of, and set the initial values of
-parameters in the transformed space used by the MCMC.  We use logarithms
-to maintain strictly positive values of parameters where appropriate.
+parameters in the transformed space used by the MCMC.
 
-```{r ricker-mcmc}
-# Uniform priors on standard deviation terms
-jags.params=c("K","logr0", "stdQ", "stdR")
-jags.inits <- function(){
-  list("K"=init_p["K"], "logr0"=log(init_p["r0"]), 
-       "stdQ"=sqrt(0.05), "stdR"=sqrt(0.1), "x"=y, 
-       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
-}
-set.seed(12345) 
-jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
-                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
-                  model.file="ricker.bugs", progress.bar="none")
+```{r ricker-mcmc, dependson="ricker-model"}
 ```
@@
 
-```{r ricker_traces}
-jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-ricker_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
-ggplot(ricker_posteriors) + geom_line(aes(index, value)) + 
-  facet_wrap(~ variable, scale="free", ncol=1)
+```{r ricker-traces, dependson="ricker-mcmc"}
 ```
 
-```{r ricker_posteriors}
-ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100) 
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-# plot posterior distributions
-ggplot(ricker_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=2)
+```{r ricker-posteriors, dependson=c("ricker-traces", "ricker-priors")}
 ```
@@
 
-```{r ricker-output}
-ricker_pardist <- acast(ricker_posteriors, index ~ variable, subset=.(variable != "deviance") )
-ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first
-bayes_coef <- apply(ricker_pardist,2, posterior.mode) # much better estimates from mode then mean
-ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
-ricker_means <- sapply(x_grid, Ricker, 0, ricker_bayes_pars[c(1,2)])
+```{r ricker-output, dependson=c("posterior-mode", "ricker-traces")}
 ```
@@
 
-```{r myers-jags}
-logr0_prior_p <- c(-6.0, 6.0)
-logtheta_prior_p <- c(-6.0, 6.0)
-logK_prior_p <- c(-6.0, 6.0)
-
-bugs.model <- 
-paste(sprintf(
-"model{
-  logr0    ~ dunif(%s, %s)
-  logtheta    ~ dunif(%s, %s)
-  logK    ~ dunif(%s, %s)
-  stdQ ~ dunif(%s, %s)
-  stdR ~ dunif(%s, %s)", 
-  logr0_prior_p[1], logr0_prior_p[2],
-  logtheta_prior_p[1], logtheta_prior_p[2],
-  logK_prior_p[1], logK_prior_p[2],
-  stdQ_prior_p[1], stdQ_prior_p[2],
-  stdR_prior_p[1], stdR_prior_p[2]),
-
-  "
-  iQ <- 1 / (stdQ * stdQ);
-  iR <- 1 / (stdR * stdR);
-
-  r0 <- exp(logr0)
-  theta <- exp(logtheta)
-  K <- exp(logK)
-
-  x[1] ~ dunif(0, 10)
-
-  for(t in 1:(N-1)){
-    mu[t] <- r0 * pow(abs(x[t]), theta) / (1 + pow(abs(x[t]), theta) / K)
-    x[t+1] ~ dnorm(mu[t], iQ) 
-  }
-
-  for(t in 1:(N)){
-    y[t] ~ dnorm(x[t], iR)
-  }
-}")
-writeLines(bugs.model, "myers.bugs")
+```{r myers-model}
 ```
@@
 ```{r myers-priors}
-logK_prior     <- function(x) dunif(x, logK_prior_p[1], logK_prior_p[2])
-logr_prior     <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
-logtheta_prior <- function(x) dunif(x, logtheta_prior_p[1], logtheta_prior_p[2])
-par_priors <- list( deviance = function(x) 0 * x, logK = logK_prior,
-                    logr0 = logr_prior, logtheta = logtheta_prior, 
-                    stdQ = stdQ_prior, stdR = stdR_prior)
-
 ```
 
-```{r myers-mcmc}
-init_p = c(1, 1, 1)
-names(init_p) = c("r0", "theta", "K")
-jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
-jags.inits <- function(){
-  list("logr0"=log(init_p["r0"]), "logtheta" = log(init_p["theta"]),  
-       "logK"=log(init_p["K"]), "stdQ"=sqrt(0.05), 
-       "stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
-}
-set.seed(12345)
-jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, 
-                  n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
-                  n.burnin=n.burnin, model.file="myers.bugs", progress.bar="none")
-
+```{r myers-mcmc, dependson="myers-model"}
 ```
@@
 
-```{r myers-traces}
-jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
-myers_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
-ggplot(myers_posteriors) + geom_line(aes(index, value)) +
-  facet_wrap(~ variable, scale="free", ncol=1)
-
+```{r myers-traces, dependson="myers-mcmc"}
 ```
 
-```{r myers-posteriors}
-par_prior_curves <- ddply(myers_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100) 
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-
-ggplot(myers_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
-  facet_wrap(~ variable, scale="free", ncol=3)
+```{r myers-posteriors, dependson="myers-traces"}
 ```
 
-```{r myers-output}
-myers_pardist <- as.matrix(jags_matrix[2:6])
-myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
-myers_pardist[,2] = exp(myers_pardist[,2]) # transform model parameters back first
-myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
-colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
-bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
-myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
-myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
+```{r myers-output, dependson=c("posterior-mode", "myers-traces")}
 ```
@@
 ```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
-models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
-                     MLE=est_means, Ricker=ricker_means, 
-                     Allen = allen_means,
-                     Myers = myers_means)
-
-models <- melt(models, id="x")
-names(models) <- c("x", "method", "value")
 ```
 
-```{r Figure1}
-plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-    geom_line(data=models, aes(x, value, col=method), lwd=1, alpha=0.8) + 
-    geom_point(data=obs, aes(x,y), alpha=0.8) + 
-    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
-    scale_colour_manual(values=cbPalette) 
-print(plot_gp)
+```{r Figure1, dependson="assemble-models"}
+```
+
+## Goodness of fit
+
+This shows only the mean predictions.
+For the Bayesian cases, we can instead loop over the posteriors of the parameters (or samples from the GP posterior) to get the distribution of such curves in each case.
+
+```{r Figureb, dependson=c("assemble-models", "par-fns")}
 ```
@@
 
-```{r gp-opt, dependson="assemble-models"}
-MaxT = 1000
-# uses expected values from GP, instead of integrating over posterior
-#matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)
-
-# Integrate over posteriors 
-matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid)
-
-# Solve the SDP using the GP-derived transition matrix
-opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
+```{r gp-opt, dependson="gp-output"}
 ```
@@
 
-```{r mle-opt, dependson="assemble-models"}
-matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-matrices_estimated <- f_transition_matrix(est$f, est$p, x_grid, h_grid, est$sigma_g)
-opt_estimated <- value_iteration(matrices_estimated, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
+```{r mle-opt, dependson="mle-output"}
 ```
@@
 
-```{r allen-opt, dependson="assemble-models"}
-allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
-matrices_allen <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
-opt_allen <- value_iteration(matrices_allen, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+```{r allen-opt, dependson="allen-output"}
 ```
@@
 
-```{r ricker-opt, dependson="assemble-models"}
-ricker_f <- function(x, h, p) Ricker(x, h, unname(p[c(2, 1)])) # defined by pdgControl 
-matrices_ricker <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
-opt_ricker <- value_iteration(matrices_ricker, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+```{r ricker-opt, dependson="ricker-output"}
 ```
@@
 
-```{r myers-opt, dependson="assemble-models"}
-myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
-matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
-myers_alt <- value_iteration(matrices_myers, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+```{r myers-opt, dependson="myers-output"}
 ```
@@
 
-```{r assemble-opt, , dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
-OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_ricker$D, Allen = opt_allen$D, Myers = myers_alt$D)
-colorkey=cbPalette
-names(colorkey) = names(OPT) 
+```{r assemble-opt, dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
 ```
@@
 ```{r totalprofits, dependson="profits"}
-ggplot(Profit, aes(V1)) + geom_histogram() + 
-  facet_wrap(~method, scales = "free_y") + guides(legend.position = "none") + xlab("Total profit by replicate")
 ```
+
+```{r deviances, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output") }
+```

### 2013-06-16 inst/doc/nonparametric-bayes.Rmd
writing
update manuscript code

@@
 % Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach
+
+<!-- Run R code for analysis, to be called by figures -->
+```{r generics-and-libraries, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+source("~/.knitr_defaults.R")
+library(knitcitations)
+library(nonparametricbayes) 
+opts_chunk$set(external=TRUE)
+read_chunk("../examples/BUGS/external-chunks.R")
+```
+
+```{r plotting-options, echo=FALSE, cache=FALSE, include=FALSE}
+options(xtable.print.comment=FALSE)
+options(xtable.type = 'latex', table.placement="H")
+opts_chunk$set(warning=FALSE, message=FALSE, comment=NA,
+               tidy=FALSE, echo=FALSE)
+opts_knit$set(progress = TRUE, verbose = TRUE)
+opts_chunk$set(dev = 'Cairo_pdf', fig.width=5.5, fig.height=4,
+ cache.path = 'cache-pdf/', cache=TRUE, external=TRUE)
+```
+
+```{r posterior-mode}
+```
+```{r stateeq}
+```
+```{r sdp-pars, dependson="stateeq"}
+```
+```{r obs, dependson="sdp-pars"}
+```
+
+```{r mle, dependson="obs"}
+```
+```{r mle-output, dependson="mle"}
+```
+
+```{r gp-priors}
+```
+```{r gp, dependson=c("gp-priors", "obs")}
+```
+```{r gp_traces_densities, dependson="gp"}
+```
+```{r gp-output, dependson="gp"}
+```
+
+```{r jags-setup}
+```
+```{r common-priors}
+```
+```{r allen-model}
+```
+```{r allen-priors, dependson="common-priors"}
+```
+```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results="hide"}
+```
+```{r allen-traces, dependson="allen-mcmc"}
+```
+```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
+```
+```{r allen-output, dependson=c("posterior-mode", "allen-traces")}
+```
+
+```{r ricker-model}
+```
+```{r ricker-priors, dependson="common-priors"}
+```
+```{r ricker-mcmc, dependson="ricker-model"}
+```
+```{r ricker-traces, dependson="ricker-mcmc"}
+```
+```{r ricker-posteriors, dependson=c("ricker-traces", "ricker-priors")}
+```
+```{r ricker-output, dependson=c("posterior-mode", "ricker-traces")}
+```
+
+```{r myers-model}
+```
+```{r myers-priors}
+```
+```{r myers-mcmc, dependson="myers-model"}
+```
+```{r myers-traces, dependson="myers-mcmc"}
+```
+```{r myers-posteriors, dependson="myers-traces"}
+```
+```{r myers-output, dependson=c("posterior-mode", "myers-traces")}
+```
+
+```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
+```
+```{r gp-opt, dependson="gp-output"}
+```
+```{r mle-opt, dependson="mle-output"}
+```
+```{r allen-opt, dependson="allen-output"}
+```
+```{r ricker-opt, dependson="ricker-output"}
+```
+```{r myers-opt, dependson="myers-output"}
+```
+```{r assemble-opt, dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
+```
+```{r sims, dependson="assemble-opt"}
+```
+
+```{r profits, dependson="sims", results="hide"}
+```
+```{r deviances, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output"), results="hide"}
+```
 
@@
 S_t &= X_t - h_t \\
-f(S_t) &= e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
+f(S_t) &= S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
 \end{align}
@@
 
-![ Mean dynamics predicted by each model.](Figure1.pdf)
+```{r Figure1, dependson=c("assemble-models"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
+Graph shows the expected value for the function $f$ under each model.
+Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.
+The training data is also shown as black points.
+(The GP is conditioned on 0,0, shown as a pseudo-data point).
+"}
+```
+```{r Figureb, dependson=c("assemble-models", "par-fns")}
+```
 
@@
 
-![ Steady-state optimal policies by each model.](Figure2.pdf)
+```{r Figure2, dependson="assemble-opt", fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
+Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
+```
 
@@
 
-![ Managing the simulated system over 100 replicate instances under each policy](Figure3.pdf)
+```{r Figure3, dependson=c("sim", "plotting-options"), fig.cap="Gaussian process inference outperforms parametric estimates.
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."}
+```
 
@@
 
-![Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
+```{r Figure4, dependson="profits", fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
 For each data set, the three models are estimated as described above.
 Values plotted are the averages of a given policy over 100 replicate simulations.
-Details and code provided in the supplement.](figure/figure4.png)
+Details and code provided in the supplement."}
+```
 

### 2013-06-17 inst/doc/nonparametric-bayes.tex
recompile manuscript

@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\usepackage{color}
-\usepackage{fancyvrb}
-\DefineShortVerb[commandchars=\\\{\}]{\|}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\newenvironment{Shaded}{}{}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
-\newcommand{\RegionMarkerTok}[1]{{#1}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
-\newcommand{\NormalTok}[1]{{#1}}
-\usepackage{graphicx}
-% We will generate all images so they have a width \maxwidth. This means
-% that they will get their normal width if they fit onto the page, but
-% are scaled down if they would overflow the margins.
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
-\else\Gin@nat@width\fi}
-\makeatother
-\let\Oldincludegraphics\includegraphics
-\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
 \ifxetex
@@
 
-\section{Abstract}
-
-Decision-theoretic methods often rely on simple parametric models of
-ecological dynamics to compare the value of a potential sequence of
-actions. Unfortunately, such simple models rarely capture the complexity
-or uncertainty found in most real ecosystems. Non-parametric Bayesian
-methods offer a promising statistical approach for predictive modeling
-of ecological dynamics in regions of state space where the data is
-adequate, while at the same time offering more flexible patterns with
-greater uncertainty outside the observed data. This contrasts from
-simple parametric models which provide relatively constant level of
-uncertainty in regions with and without adequate data. The consequence
-of such misplaced confidence outside the data can lead to highly
-undesirable results that may be avoided with the more flexible
-non-parametric Bayesian approach.
-
-\section{Introduction}
-
-Most management recommendations from the ecological literature are based
-on (or at least motivated by) parametric models. Though in principle
-these models reflect mechanistic underpinnings of biological
-interactions involved, in practice real mechanisms are often unknown and
-the true dynamics too complex to be captured by simple models. While
-simple mechanistic models can nevertheless provide important insights
-into possible dynamics -- for instance, demonstrating that a vaccine
-does not have to be 100\% effective to eliminate the transmission of a
-virus -- such approaches are not well-suited for use in forecasting
-outcomes of potential management options. Non-parametric approaches
-offer a more flexible alternative that can both more accurately reflect
-the data available while also representing greater uncertainty in areas
-(of state-space) where data is lacking. Ecological research and
-management strategy should pay closer attention to the opportunities and
-challenges nonparametric modeling can offer.
-
-\subsubsection{Quantitative vs Qualitative Decisions}
-
-In this paper, we consider those ecological management problems in which
-a mathematical (or computational) model is used to quantitatively inform
-decision-making by allowing a manager to compare to the expected
-consequences of potential management actions (or policies). We
-distinguish this from the solely qualitative use of a model, in which
-models are used to represent and compare hypotheses of different
-mechanisms that would lead to qualitatively different actions.
-
-Though simple mechanistically motivated models may be best for the
-latter case (Cuddington, Fortin, and Gerber 2013; Geritz and Kisdi
-2012), such models can be not only inaccurate but misleading in
-quantitative decision making. This arises for two reasons:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  We do not know what the correct parameter values are for the models.
-\item
-  We do not know what the correct models are for ecological systems.
-\end{enumerate}
-
-The complexity of ecological interactions and a lack of data contribute
-greatly to both of the problems. This concern is particularly acute in
-case of ecological tipping points (Barnosky et al. 2012; Scheffer et al.
-2001), which arise from feedbacks common in ecological systems and can
-lead to a sudden catastrophic transition to an undesirable state. In
-general we do not know where such tipping points are unless we have
-already observed the transition, in which case it is frequently too late
-to respond. Though there may exist early warning signals for certain
-kinds of these transitions that are driven by slow changes (Scheffer et
-al. 2009), we do not know when, where, or how to apply them to the
-decision making context more generally (Boettiger and Hastings 2013).
-
-In addition to facing the potentially grave consequences of such
-transitions, if they do exist in a system of interest, this generally
-means that we lack data in the region near and beyond a possible tipping
-point. As a consequence, our models are least accurate just where we
-need them to be most accurate.\\Unfortunately, parametric modeling
-approaches are not well-suited to this scenario, in which good data is
-avialble over only part of the relevant state-space. Necessary
-assumptions about model structure in any parametric approach can lead
-such models to perform very poorly, even when accounting for parameter
-uncertainty.
-
-Non-parametric models can more realistically represent uncertainties
-outside of observed data while also better capturing the dynamics in the
-region observed. Advances in the theory and computational
-implementations of nonparametric methods make them ripe for such
-applications. We use the classic problem of optimal harvest of a marine
-fishery to illustrate how the nonparametric approach of Gaussian
-processes can be applied. We will compare Bayesian implementations of
-both nonparametric and parametric models, which best allow us to capture
-the uncertainty of model estimation in either case and permits a more
-natural comparison between approaches.
-
-\subsubsection{Terminology: ``Non-parametric''}
-
-The term non-parametric to describe modeling approaches such as Gaussian
-processes is a common source of confusion, notwithstanding the fact that
-the model is still specified by parameters. Some literature has
-introduced the term ``semi-parametric'' in tacit acknowledgment of this,
-which no doubt only contributes to the confusion. The problem is further
-exacerbated by the several meanings assigned to the term in statistics:
-(a) that the method does not assume some particular probability
-distribution (e.g.~a non-parametric bootstrap) or (b) that the method
-does not assume a fixed structure to the model. Our use is that of (b),
-in contrast to the classical approaches that do. The difference is
-clearest by way of example.
-
-\subsubsection{Uncertainties: the classical approach}
-
-The ecological dynamics of most systems of management interest are
-typically both complex and unknown. Despite this, quantitative policy
-optimization in these systems has almost exclusively been based on
-simple, mechanistically motivated parametric models. There are several
-reasons for this. First, limited available data almost always precludes
-any accurate estimate of more complicated models. Second, the
-computational demands of determining an optimal strategy in a sequential
-decision-theory problem suffer from the ``curse of dimensionality'' that
-makes it difficult to consider more complicated models, states, or
-actions. Unfortunately, these challenges also make it particularly
-difficult to reflect the true uncertainty either in the parameters of
-the given models (parametric uncertainty) or in the structure of the
-models themselves (structural uncertainty). Further unknowns such as
-measurement uncertainty, parameter uncertainty, unobserved states,
-knowledge of boundary conditions, etc. further compound the issue.
-Though a hierarchical Bayesian approach provides a natural way to
-address these from a statistical standpoint, formulating reasonable
-parametric descriptions of each form of uncertainty is a challenging
-task in itself, let alone the computational difficulties of solving such
-a system. Cressie et al. (2009) provides a good account of the successes
-and challenges of the approach. Applying these approaches in the
-management context of sequential decision making, in which forecasts
-must be obtained over a range of possible actions and updated regularly
-as new information arrives makes such an approach less feasible still.
-
-An active literature and growing computational power over the past
-several decades have only marginally improved this situation. Parametric
-or structural uncertainty can be introduced only by increasing the state
-space to reflect a distribution of possible parameters or the degree of
-belief in each of a set of possible models Frequently, the space of
-possible actions must then be reduced or the algorithms adjusted by
-approximations to keep computations feasible.
-
-Traditional approaches to optimal control (Pontryagin's principle,
-stochastic dynamic programming) rely on knowledge of the state equation,
-usually described by a simple parametric model. Here we illustrate how a
-stochastic dynamic programming algorithm can alternatively be driven by
-the predictions from a Gaussian process.
-
-\section{Approach and Methods}
-
-\subsection{Underlying Model}
-
-Concerns over the potential for tipping points in ecological dynamics
-(Scheffer et al. 2001) highlight the dangers of uncertainty in
-ecological management and pose a substantial challenge to existing
-decision-theoretic approaches (Brozović and Schlenker 2011). Because
-intervention is often too late after a tipping point has been crossed
-(but see Hughes et al. (2013)), management is most often concerned with
-avoiding potentially catastrophic tipping points before any data is
-available at or following a transition that would more clearly reveal
-these regime shift dynamics (e.g. Bestelmeyer et al. 2012).
-
-To illustrate the value of the non-parametric Bayesian approach to
-management, we focus on example of a system containing such a tipping
-point whose dynamics can still be described by a simple, one-dimensional
-parametric model.\\We will focus on a simple parametric model for a
-single species (derived from fist principles by Allen et al. 2005) as
-our underlying ``reality''.
-
-\begin{align}
-X_{t+1} &= Z_t f(S_t) \\
-S_t &= X_t - h_t \\
-f(S_t) &= e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
-\end{align}
-
-Where $Z_t$ is multiplicative noise function with mean 1, representing
-stochastic growth. We will consider log-normal noise with shape
-parameter $\sigma_g$. We start with an example in which the parameters
-are $r =2$, $K=8$, $C=5$ and $\sigma_g = 0.1$.
-
-As a low-dimensional system completely described by three parameters,
-this scenario should if anything be favorable to a parametric-based
-approach. This model contains an Allee effect, or tipping point, below
-which the population is not self-sustaining and shrinks to zero
-(Courchamp, Berec, and Gascoigne 2008).
-
-Both parametric and nonparametric approaches will require training data
-on which to base their model of the process. We generate the training
-data under the model described in Eq 1 for $T_{\textrm{obs}}=40$ time
-steps, under a known but not necessarily optimal sequence of harvest
-intensities, $h_t$. For simplicity we imagine a fishery that started
-from zero harvest pressure and has been gradually increasing the
-harvest.
-
-\textbf{\emph{Should we include any emprical examples?}}
-
-\subsubsection{The optimal control problem in fisheries management}
-
-In our example, we focus on the problem in which a manager must set the
-harvest level for a marine fishery each year to maximize the net present
-value of the resource, given an estimated stock size from the year
-before. Rich data and global concerns have made marine fisheries the
-crucible for much of the founding work (Gordon 1954; Reed 1979; May et
-al. 1979; Ludwig and Walters 1982) in managing ecosystems under
-uncertainty. Global trends (Worm et al. 2006) and controversy (Hilborn
-2007; Worm et al. 2009) have made understanding these challenges all the
-more pressing.
-
-To permit comparisons against a theoretical optimum we will consider
-data on the stock dynamics simulated from a simple parametric model in
-which recruitment of the fish stock $X_{t+1}$ in the following year is a
-stochastic process governed by a function $f$ of the current stock
-$X_t$, selected harvest policy $h_t$, and noise process $Z$,
-
-\[X_{t+1} = Z_t f(X_t, h_t) \]
-
-Given parameters for the function $f$ and probability distribution $Z$,
-along with a given economic model determining the price/profit
-$\Pi(X_t, h_t)$ realized in a given year given a choice of harvest $h_t$
-and observed stock $X_t$. This problem can be solved exactly for
-discretized values of stock $X$ and policy $h$ using stochastic dynamic
-programming (SDP) (Mangel 1985). Problems of this sort underpin much
-marine fisheries management today.
-
-A crux of this approach is correctly specifying the functional form of
-$f$, along with its parameters. The standard approach uses one of a
-handful of common parametric models representing the stock-recruitment
-relationship, usually after estimating the model parameters from any
-available existing data. Uncertainty in the parameter estimates can be
-estimated and integrated over to determine the optimal policy under
-under uncertainty (Mangel 1985; Schapaugh and Tyre 2013). Uncertainty in
-the model structure itself can only be addressed in this approach by
-hypothesizing alternative model structures, and then performing some
-model choice or model averaging (B. K. Williams 2001; Athanassoglou and
-Xepapadeas 2012). For simplicity of the comparison we consider only the
-case of a structurally correct and structurally incorrect model, and
-estimate Bayesian posterior distributions of the parameters. We will
-also compare to a simpler maximum likelihood estimate of the parameters
-that ignores the issue of parameter uncertainty.
-
-We compare these parametric approaches to our alternative that uses a
-Gaussian Process (GP) in place of a given stock recruitment curve. We
-briefly revisit a description of the Gaussian Process approach that has
-appeared in recent literature before describing how we take the
-important step of adapting the GP to the decision theory problem that
-can still be solved using SDP.
-
-\subsubsection{The Non-parametric Bayesian alternative for
-stock-recruitment curves}
-
-The use of Gaussian process (GP) regression (or ``kreging'' in the
-geospatial literature) to formulate a predictive model is relatively new
-in the context of modeling dynamical systems (Kocijan et al. 2005) and
-introduced in the ecological modeling and fisheries management by Munch
-et al. (2005). An accessible and thorough introduction to the
-formulation and use of GPs can be found in Rasmussen and Williams
-(2006).
-
-The essence of the GP approach can be captured in the following thought
-experiment: An exhaustive parametric approach to the challenge of
-structural uncertainty might proceed by writing down all possible
-functional forms for the underlying dynamical system with all possible
-parameter values for each form, and then consider searching over this
-huge space to select the most likely model and parameters; or using a
-Bayesian approach, assign priors to each of these possible models and
-infer the posterior distribution of possible models. The GP approach can
-be thought of as a computationally efficient approximation to this
-approach. GPs represent a large class of models that can be though of as
-capturing or reasonably approximating the set of models in this
-collection. By modeling at the level of the process, rather than the
-level of parametric equation, we can more concisely capture the possible
-behavior of these curves. In place of a parametric model of the
-dynamical system, the GP approach postulates a prior distribution of
-(n-dimensional) curves that can be though of as approximations to a
-range of possible (parametric) models that might describe the data. The
-GP allows us to consider a set of possible curves simultaneously.
-
-\subsubsection{Background on Gaussian Process inference}
-
-Once the posterior Gaussian process (GP) has been estimated (e.g.~see
-Munch et al. 2005), it is necessary to adapt it in place of the
-parametric equation for the stochastic dynamic programming (SDP)
-solution (see Mangel and Clark 1988 for a detailed description of
-parametric SDP methods) to the optimal policy. The essence of the idea
-is straight forward -- we will use the estimated GP in place of the
-parametric growth function to determine the stochastic transition matrix
-on which the SDP calculations are based. The SDP is solved in a
-discretized state space -- both the continuously valued population
-densities $X$ and harvest quotas $h$ are first mapped to a bounded,
-discrete grid. (For simplicity we will consider a uniform grid, though
-for either parametric or GP-based SDP it is often advantageous to use a
-non-uniform discretization such as a basis function representation,
-e.g.~see (Deisenroth, Rasmussen, and Peters 2009)).
-
-The SDP approach then computes a transition matrix, $\mathbf{F}$. We
-demonstrate that calculation is just as straight forward based on the GP
-as it is in the classical context using the parametric model. The
-${i,j}$ of the transition matrix $F$ entry gives the probability of
-transitioning into state $x_i$ given that the system is in state $x_j$
-in the previous time-step. To generate the transition matrix based on
-the posterior GP, we need only the expected values at each grid point
-and the corresponding variances (the diagonal of the covariance matrix),
-as shown in Figure 1. Given the mean of the GP posterior at each
-grid-point as the vector $E$ and variance at that point as vector $V$,
-the probability of transitioning from state $x_i$ to state $x_j$ is
-
-\[\mathcal{N}\left(x_j \vert  \mu = E_i, \sigma = \sqrt{V_i}\right)\]
-
-where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
-variance $\sigma^2$. Strictly speaking, the transition probability
-should be calculated by integrating the normal density over the bin of
-width $\Delta$ centered at $x_j$. For a sufficiently fine grid that
-$f(x_j) \approx f(x_j + \Delta)$, it is sufficient to calculate the
-density at $x_j$ and then row-normalize the transition matrix. The
-process can then be repeated for each possible discrete value of our
-control variable, (harvest $h$).
-
-\textbf{Pseudocode for the determining the transition matrix from the
-GP}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{for(h in h_grid)}
-  \NormalTok{F_h = for(x_j in grid)}
-          \NormalTok{for(i in }\DecValTok{1}\NormalTok{:N) }
-            \KeywordTok{dnorm}\NormalTok{(x_j, mu[i]-h, V[i])}
-\end{Highlighting}
-\end{Shaded}
-
-Using the discrete transition matrix we may write down the Bellman
-recursion defining the stochastic dynamic programming iteration:
-
-\begin{equation}
-V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
-\end{equation}
-
-where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$ is
-control (harvest level) chosen. Numerically, the maximization is
-accomplished as follows. Consider the set of possible control values to
-be the discrete values corresponding the grid of stock sizes. Then for
-each $h_t$ there is a corresponding transition matrix $\mathbf{F}_h$
-determined as described above but with mean $\mu = x_j - h_t$. Let
-$\vec{V_t}$ be the vector whose $i$th element corresponds to the value
-of having stock $x_i$ at time $t$. Then let $\Pi_h$ be the vector whose
-$i$th element indicates the profit from harvesting at intensity $h_t$
-given a population $x_i$ (e.g. $\max(x_i, h_t)$ since one cannot harvest
-more fish then the current population size). Then the Bellman recursion
-can be given in matrix form as
-
-\[V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)\]
-
-where the sum is element by element and the expectation is computed by
-the matrix multiplication $\mathbf{F} V_{t+1}$.
-
-\subsubsection{Pseudocode for the Bellman iteration}
-
-\begin{Shaded}
-\begin{Highlighting}[]
- \NormalTok{V1 <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
-      \NormalTok{delta * F[[h]] %*% V +  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
-    \NormalTok{\})}
-    \CommentTok{# find havest, h that gives the maximum value}
-    \NormalTok{out <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
-      \NormalTok{value <- }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
-      \NormalTok{index <- }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
-      \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
-    \NormalTok{\})}
-    \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
-    \NormalTok{V <- out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
-    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <- out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
-\end{Highlighting}
-\end{Shaded}
-
-This completes the algorithm adapting the GP to the sequential
-decision-making problem through SDP, which we believe has not yet been
-demonstrated in the literature. We provide an R package implementation
-of this, along with the Gaussian process inference, in the supplemental
-materials.
-
-\subsubsection{Estimating parametric models}
-
-We estimate posterior distributions for two parametric models: one using
-the structurally correct model as given in Eq (1), which we refer to as
-the ``Parametric Bayes'' model, and another using the familiar Ricker
-model, using a Gibbs sampler as described (with source code) in the
-appendix). In addition we estimate the parameters of the structurally
-correct model by maximum likelihood.
-
-The posterior distribution for the hyper-parameters of the Gaussian
-process model are estimated by Metropolis-Hastings algorithm, again with
-details and code provided in the Appendix. Rasmussen and Williams (2006)
-provides an excellent general introduction to Gaussian Processes and
-Munch et al. (2005) first discusses their application in this context.
-
-\section{Results}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{Figure1.pdf}
-\caption{Mean dynamics predicted by each model.}
-\end{figure}
-
-Figure 1 shows the mean inferred state space dynamics of each model
-relative to the true model used to generate the data, predicting the
-relationship between observed stock size (x-axis) to the stock size
-after recruitment the following year. All models except the MLE model
-estimate a distribution around the means shown here, and all models
-estimate a level of process noise, which is independent of the state
-value (x). Note that in contrast to the other models shown, the mean
-Gaussian process corresponds to a distribution of curves - as indicated
-by the gray band - which itself has a mean shown in black. Note that
-this mean GP is thus more certain of the dynamics in the region where
-data is available then where it is not.
-
-While it would be natural (and straight forward) to condition the GP on
-passing through the origin (0,0) (see appendix), the estimate shown here
-is based only on the observed data. The observed data from which each
-model is estimated is also shown. The observations come from only a
-limited region of state space corresponding to unharvested or weakly
-harvested system. No observations occur at the theoretical optimum or
-near the tipping point.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{Figure2.pdf}
-\caption{Steady-state optimal policies by each model.}
-\end{figure}
-
-The resulting optimal management strategy based on each of the inferred
-models is shown in Figure 2, against the optimal strategy given the true
-underlying dynamics. Policies are shown in terms of target escapement,
-$S_t$. Under models such as this a constant escapement policy is
-expected to be optimal (Reed 1979), whereby population levels below a
-certain size $S$ are unharvested, while above that size the harvest
-strategy aims to return the population to $S$, resulting in the
-hockey-stick shaped policies shown.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{Figure3.pdf}
-\caption{Managing the simulated system over 100 replicate instances
-under each policy}
-\end{figure}
-
-The consequences of managing 100 replicate realizations of the simulated
-fishery under each of the policies estimated is shown in Figure 3. As
-expected from the policy curves, the structurally correct model
-under-harvests, leaving the stock to vary around it's un-fished optimum.
-The structurally incorrect Ricker model over-harvests the population
-passed the tipping point consistently, resulting in the immediate crash
-of the stock and thus derives minimal profits.
-
-The results shown in Figures 1-3 are not unique to the simulated data or
-models chosen here, but arises across a range of parameter values and
-simulations as shown in the supplemental figures. The results across
-this range can most easily be compared by the relative differences in
-net present value realized by each of the approaches, as shown in Figure
-4. The Gaussian Process most consistently realizes a value close to the
-optimal solution, and importantly avoids ever driving the system across
-the tipping point, which results in the near-zero value cases in the
-parametric models.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/figure4.png}
-\caption{Histograms of the realized net present value of the fishery
-over a range of simulated data and resulting parameter estimates. For
-each data set, the three models are estimated as described above. Values
-plotted are the averages of a given policy over 100 replicate
-simulations. Details and code provided in the supplement.}
-\end{figure}
-
-\section{Discussion}
-
-Non-parametric Bayesian methods have received far too little attention
-in ecological modeling efforts that are aimed at improved conservation
-planning and decision making support. Such approaches may be
-particularly useful when the available data is restricted to a limited
-area of state-space, which can lead parametric models to underestimate
-the uncertainty in dynamics at population levels (states) which have not
-been observed. One reason for the relative absence of nonparametric
-approaches in the natural resource management context may be the lack of
-existing approaches for adapting the non-parametric Bayesian models
-previously proposed (Munch et al. 2005) to a decision-theoretic
-framework. Adapting a non-parametric approach requires modification of
-existing methods for decision theory. We have illustrated how this might
-be done for a classic stochastic dynamic programming problem, opening
-the door for substantial further research into how these applications
-might be improved.
-
-While non-parametric Bayesian approaches will not always be preferable
-to simple mechanistic models, we highlight three aspects of the problem
-consider here that make these methods particularly valuable. These
-aspects are common to many conservation decision making problems, which
-thus merit greater use of non-parametric approaches that can best take
-advantage of them.
-
-\subsubsection{1. Large uncertainty where the data is poor}
-
-The parametric models perform worst when they propose a management
-strategy outside the range of the observed data. The non-parametric
-Bayesian approach, in contrast, allows a predictive model that expresses
-a great deal of uncertainty about the probable dynamics outside the
-observed range, while retaining very good predictive accuracy in the
-range observed. The management policy dictated by the GP balance this
-uncertainty against the immediate value of the harvest, and act to
-stabilize the population dynamics in a region of state space in which
-the predictions can be reliably reflected by the data.
-
-\subsubsection{2. Predictive accuracy where data is good}
-
-While expressing larger uncertainty outside the observed data, the GP
-can also provide a better fit with smaller uncertainty inside the range
-of the observed data. This arises from the greater flexibility of the
-Gaussian process, which describes a large family of possible
-curves.\\While in a parametric context this over-fitting would be more
-worrisome -- a high-degree polynomial could fit the data even better --
-those concerns are driven by the resulting parametric fit outside the
-data, which may involve wild oscillations unsupported by the data. As we
-have seen in \#1, the GP is less vulnerable to such unjustified
-predictions outside the data, and is meanwhile free to benefit from the
-greater fit where the data is available.
-
-\subsection{Future directions}
-
-In this simulated example, the underlying dynamics are truly governed by
-a simple parametric model, allowing the parametric approaches to be more
-accurate. Similarly, because the dynamics are one-dimensional dynamics
-and lead to stable nodes (rather than other attractors such as
-limit-cycles resulting in oscillations), the training data provides
-relatively limited information about the dynamics. For these reasons, we
-anticipate that in higher-dimensional examples characteristic of
-ecosystem management problems that the machine learning approach will
-prove even more valuable.
-
- Data complexity. Perhaps too far out of scope\ldots{} The nonparametric
-Bayesian approach is also better suited to complex and disparate data.
-Incorporating various sources of information into mechanistic models can
-be an immensely difficult due to the increased complexity involved.
-
-\subsubsection{Online learning}
-
-In our treatment here we have ignored the possibility of learning during
-the management phase, in which the additional observations of the stock
-size could potentially improve parameter estimates. While we intend to
-address this possibility in future work in the context of these
-non-parametric models, we have not addressed it here for pedagogical
-reasons. In the context presented here, it is clear that the differences
-in performance arise from differences in the uncertainty inherent in the
-model formulations, rather than from differing abilities to learn.
-Because we consider a threshold system, online learning would not change
-this generic feature of a lack of data in a certain range of the state
-space which is better captured by the Gaussian process.
-
-\section{Acknowledgments}
-
-This work was partially supported by the Center for Stock Assessment
-Research, a partnership between the University of California Santa Cruz
-and the Fisheries Ecology Division, Southwest Fisheries Science Center,
-Santa Cruz, CA and by NSF~grant EF-0924195 to MM.
-
-\section{References}
-
-Allen, Linda J. S., Jesse F. Fagan, Göran Högnäs, and Henrik Fagerholm.
-2005. ``Population extinction in discrete-time stochastic population
-models with an Allee effect.'' \emph{Journal of Difference Equations and
-Applications} 11 (4-5) (apr): 273--293.
-doi:10.1080/10236190412331335373.
-\url{http://www.tandfonline.com/doi/abs/10.1080/10236190412331335373}.
-
-Athanassoglou, Stergios, and Anastasios Xepapadeas. 2012. ``Pollution
-control with uncertain stock dynamics: When, and how, to be
-precautious.'' \emph{Journal of Environmental Economics and Management}
-63 (3) (may): 304--320. doi:10.1016/j.jeem.2011.11.001.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069611001409}.
-
-Barnosky, Anthony D., Elizabeth a Hadly, Jordi Bascompte, Eric L.
-Berlow, James H. Brown, Mikael Fortelius, Wayne M. Getz, et al. 2012.
-``Approaching a state shift in Earth's biosphere.'' \emph{Nature} 486
-(7401) (jun): 52--58. doi:10.1038/nature11018.
-\url{http://www.nature.com/doifinder/10.1038/nature11018}.
-
-Bestelmeyer, Brandon T., Michael C. Duniway, Darren K. James, Laura M.
-Burkett, and Kris M. Havstad. 2012. ``A test of critical thresholds and
-their indicators in a desertification-prone ecosystem: more resilience
-than we thought.'' Ed. Katharine Suding. \emph{Ecology Letters} (dec).
-doi:10.1111/ele.12045. \url{http://doi.wiley.com/10.1111/ele.12045}.
-
-Boettiger, Carl, and Alan Hastings. 2013. ``Tipping points: From
-patterns to predictions.'' \emph{Nature} 493 (7431) (jan): 157--158.
-doi:10.1038/493157a.
-\url{http://www.nature.com/doifinder/10.1038/493157a}.
-
-Brozović, Nicholas, and Wolfram Schlenker. 2011. ``Optimal management of
-an ecosystem with an unknown threshold.'' \emph{Ecological Economics}
-(jan): 1--14. doi:10.1016/j.ecolecon.2010.10.001.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0921800910004167}.
-
-Courchamp, Franck, Ludek Berec, and Joanna Gascoigne. 2008. \emph{Allee
-Effects in Ecology and Conservation}. Oxford University Press, USA.
-\url{http://www.amazon.com/Effects-Ecology-Conservation-Franck-Courchamp/dp/0198570309}.
-
-Cressie, Noel, Catherine a Calder, James S. Clark, Jay M. Ver Hoef, and
-Christopher K. Wikle. 2009. ``Accounting for uncertainty in ecological
-analysis: the strengths and limitations of hierarchical statistical
-modeling.'' \emph{Ecological Applications} 19 (3) (apr): 553--70.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/19425416}.
-
-Cuddington, Kim M., M. J. Fortin, and L. R. Gerber. 2013.
-``Process-based models are required to manage ecological systems in a
-changing world.'' \emph{Ecosphere} 4 (February): 1--12.
-doi:10.1890/ES12-00178.1.
-\url{http://www.esajournals.org/doi/abs/10.1890/ES12-00178.1}.
-
-Deisenroth, Marc Peter, Carl Edward Rasmussen, and Jan Peters. 2009.
-``Gaussian process dynamic programming.'' \emph{Neurocomputing} 72 (7-9)
-(mar): 1508--1524. doi:10.1016/j.neucom.2008.12.019.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0925231209000162}.
-
-Geritz, Stefan A. H., and Eva Kisdi. 2012. ``Mathematical ecology: why
-mechanistic models?.'' \emph{Journal of mathematical biology} 65 (6-7)
-(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.
-
-Gordon, H. S. 1954. ``The economic theory of a common-property resource:
-the fishery.'' \emph{The Journal of Political Economy} 62 (2): 124--142.
-\url{http://www.jstor.org/stable/10.2307/1825571}.
-
-Hilborn, Ray. 2007. ``Reinterpreting the State of Fisheries and their
-Management.'' \emph{Ecosystems} 10 (8) (oct): 1362--1369.
-doi:10.1007/s10021-007-9100-5.
-\url{http://www.springerlink.com/index/10.1007/s10021-007-9100-5}.
-
-Hughes, Terry P., Cristina Linares, Vasilis Dakos, Ingrid a van de
-Leemput, and Egbert H. van Nes. 2013. ``Living dangerously on borrowed
-time during slow, unrecognized regime shifts.'' \emph{Trends in ecology
-\& evolution} 28 (3) (mar): 149--55. doi:10.1016/j.tree.2012.08.022.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/22995893}.
-
-Kocijan, Juš, Agathe Girard, Bla\textbackslash{}vz Banko, and Roderick
-Murray-Smith. 2005. ``Dynamic systems identification with Gaussian
-processes.'' \emph{Mathematical and Computer Modelling of Dynamical
-Systems} 11 (4) (dec): 411--424. doi:10.1080/13873950500068567.
-\url{http://www.tandfonline.com/doi/abs/10.1080/13873950500068567}.
-
-Ludwig, Donald, and Carl J. Walters. 1982. ``Optimal harvesting with
-imprecise parameter estimates.'' \emph{Ecological Modelling} 14 (3-4)
-(jan): 273--292. doi:10.1016/0304-3800(82)90023-0.
-\url{http://linkinghub.elsevier.com/retrieve/pii/0304380082900230}.
-
-Mangel, Marc. 1985. ``Decision and control in uncertain resource
-systems:'' 255.
-\href{http://www.amazon.com/Decision-uncertain-resource-Mathematics-Engineering/dp/0124687202 http://dl.acm.org/citation.cfm?id=537497}{http://www.amazon.com/Decision-uncertain-resource-Mathematics-Engineering/dp/0124687202
-http://dl.acm.org/citation.cfm?id=537497}.
-
-Mangel, Marc, and Colin W. Clark. 1988. \emph{Dynamic Modeling in
-Behavioral Ecology}. Ed. John Krebs and Tim Clutton-Brock. Princeton:
-Princeton University Press.
-
-May, Robert M., John R. Beddington, Colin W. Clark, Sidney J. Holt, and
-R. M. Laws. 1979. ``Management of multispecies fisheries.''
-\emph{Science (New York, N.Y.)} 205 (4403) (jul): 267--77.
-doi:10.1126/science.205.4403.267.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/17747032}.
-
-Munch, Stephan B., Melissa L. Snover, George M. Watters, and Marc
-Mangel. 2005. ``A unified treatment of top-down and bottom-up control of
-reproduction in populations.'' \emph{Ecology Letters} 8 (7) (may):
-691--695. doi:10.1111/j.1461-0248.2005.00766.x.
-\url{http://doi.wiley.com/10.1111/j.1461-0248.2005.00766.x}.
-
-Rasmussen, Carl Edward, and C. K. I. Williams. 2006. \emph{Gaussian
-Processes for Machine Learning}. Ed. Thomas Dietterich. Boston: MIT
-Press,. \url{www.GaussianProcess.org/gpml}.
-
-Reed, William J. 1979. ``Optimal escapement levels in stochastic and
-deterministic harvesting models.'' \emph{Journal of Environmental
-Economics and Management} 6 (4) (dec): 350--363.
-doi:10.1016/0095-0696(79)90014-7.
-\href{http://www.sciencedirect.com/science/article/pii/0095069679900147 http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}{http://www.sciencedirect.com/science/article/pii/0095069679900147
-http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}.
-
-Schapaugh, Adam W., and Andrew J. Tyre. 2013. ``Accounting for
-parametric uncertainty in Markov decision processes.'' \emph{Ecological
-Modelling} 254 (apr): 15--21. doi:10.1016/j.ecolmodel.2013.01.003.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0304380013000306}.
-
-Scheffer, Marten, Jordi Bascompte, William A. Brock, Victor Brovkin,
-Stephen R. Carpenter, Vasilis Dakos, Hermann Held, Egbert H. van Nes,
-Max Rietkerk, and George Sugihara. 2009. ``Early-warning signals for
-critical transitions.'' \emph{Nature} 461 (7260): 53--9.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/19727193}.
-
-Scheffer, Marten, Stephen R. Carpenter, J. A. Foley, C. Folke, and B.
-Walker. 2001. ``Catastrophic shifts in ecosystems.'' \emph{Nature} 413
-(6856) (oct): 591--6. doi:10.1038/35098000.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/11595939}.
-
-Williams, Byron K. 2001. ``Uncertainty , learning , and the optimal
-management of wildlife.'' \emph{Environmental and Ecological Statistics}
-8: 269--288. doi:10.1023/A:1011395725123.
-
-Worm, Boris, Edward B. Barbier, Nicola Beaumont, J. Emmett Duffy, Carl
-Folke, Benjamin S. Halpern, Jeremy B. C. Jackson, et al. 2006. ``Impacts
-of biodiversity loss on ocean ecosystem services.'' \emph{Science (New
-York, N.Y.)} 314 (5800) (nov): 787--90. doi:10.1126/science.1132294.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/17082450}.
-
-Worm, Boris, Ray Hilborn, Julia K. Baum, Trevor A. Branch, Jeremy S.
-Collie, Christopher Costello, Michael J. Fogarty, et al. 2009.
-``Rebuilding global fisheries.'' \emph{Science (New York, N.Y.)} 325
-(5940) (jul): 578--85. doi:10.1126/science.1173146.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/19644114}.
-
-\pagebreak\pagebreak
-
-\section{Appendix}
-
-The appendices have not yet been assembled. Meanwhile, code to repeat
-the analyses, along with a complete log of all research conducted on
-this project, can be found at:
-\href{https://github.com/cboettig/nonparametric-bayes/}{https://github.com/cboettig/nonparametric-bayes}
-
-\subsection{Markov Chain Monte Carlo Analysis}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{gp_traces_densities1.pdf}
-\caption{Traces of the GP hyper-parameters (after transforming back from
-log space)}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{gp_traces_densities2.pdf}
-\caption{Posterior densities of the GP hyper-parameters (after
-transforming back from log space). Prior distribution over the interval
-sampled shown as red curves.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{ricker_traces.pdf}
-\caption{Traces of the Ricker parameters}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{ricker_posteriors.pdf}
-\caption{Posterior densities of the Ricker parameter estimates. Prior
-distributions over the interval sampled shown as red curves. Note the
-bimodal estimate of $\log(r)$ which leads to the poor fit in Figure 1}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{parametric_bayes_traces.pdf}
-\caption{Traces of the structurally accurate Allen model parameters}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{parametric_bayes_posteriors.pdf}
-\caption{Posterior densities for the parameters of the Allen model.
-Prior distributions over the interval sampled shown as red curves.}
-\end{figure}
-
 \end{document}

### 2013-06-17 inst/doc/beyond-mechanistic-models.Rmd
code examples updated

@@
 <!-- Run R code for analysis, to be called by figures -->
-```{r, echo=FALSE, cache=FALSE, include=FALSE}
+```{r generics-and-libraries, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+source("~/.knitr_defaults.R")
+library(knitcitations)
+library(nonparametricbayes) 
+opts_chunk$set(external=TRUE)
+read_chunk("../examples/BUGS/external-chunks.R")
+```
+
+```{r plotting-options, echo=FALSE, cache=FALSE, include=FALSE}
 options(xtable.print.comment=FALSE)
@@
  cache.path = 'cache-pdf/', cache=TRUE, external=TRUE)
-read_chunk("nonparametric-bayes.R")
 ```
-```{r knit-settings, echo=FALSE, include=FALSE, cache=FALSE}
-```
-```{r libraries, echo=FALSE, include=FALSE}
-```
-```{r plotting-options, echo=FALSE, include=FALSE}
+
+```{r posterior-mode}
 ```
@@
 ```
+
 ```{r mle, dependson="obs"}
 ```
+```{r mle-output, dependson="mle"}
+```
+
 ```{r gp-priors}
 ```
-```{r gp, dependson=c("obs", "gp-priors")}
+```{r gp, dependson=c("gp-priors", "obs")}
 ```
-```{r opt, dependson=c("gp", "mle")}
+```{r gp_traces_densities, dependson="gp"}
 ```
-```{r sim, dependson="opt"}
+```{r gp-output, dependson="gp"}
+```
+
+```{r jags-setup}
+```
+```{r common-priors}
+```
+```{r allen-model}
+```
+```{r allen-priors, dependson="common-priors"}
+```
+```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results="hide"}
+```
+```{r allen-traces, dependson="allen-mcmc"}
+```
+```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
+```
+```{r allen-output, dependson=c("posterior-mode", "allen-traces")}
+```
+
+```{r ricker-model}
+```
+```{r ricker-priors, dependson="common-priors"}
+```
+```{r ricker-mcmc, dependson="ricker-model"}
+```
+```{r ricker-traces, dependson="ricker-mcmc"}
+```
+```{r ricker-posteriors, dependson=c("ricker-traces", "ricker-priors")}
+```
+```{r ricker-output, dependson=c("posterior-mode", "ricker-traces")}
+```
+
+```{r myers-model}
+```
+```{r myers-priors}
+```
+```{r myers-mcmc, dependson="myers-model"}
+```
+```{r myers-traces, dependson="myers-mcmc"}
+```
+```{r myers-posteriors, dependson="myers-traces"}
+```
+```{r myers-output, dependson=c("posterior-mode", "myers-traces")}
+```
+
+```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
+```
+``{r gp-opt, dependson="gp-output"}
+```
+```{r mle-opt, dependson="mle-output"}
+```
+```{r allen-opt, dependson="allen-output"}
+```
+```{r ricker-opt, dependson="ricker-output"}
+```
+```{r myers-opt, dependson="myers-output"}
+```
+```{r assemble-opt, dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
+```
+```{r sims, dependson="assemble-opt"}
+```
+
+```{r profits, dependson="sims", results="hide"}
+```
+```{r deviances, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output"), results="hide"}
 ```
@@
 
-```{r gp_plot, dependson=c("gp", "mle", "plotting-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
+```{r Figure1, dependson=c("assemble-models"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
 Graph shows the expected value for the function $f$ under each model.
@@
 ```
+```{r Figureb, dependson=c("assemble-models", "par-fns")}
+```
 
-```{r policies_plot, dependson=c("opt", "plotting-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
+```{r Figure2, dependson="assemble-opt", fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
 Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
@@
 
-```{r sim_plot, dependson=c("sim", "plotting-options"), fig.cap="Gaussian process inference outperforms parametric estimates.
+```{r Figure3, dependson=c("sim", "plotting-options"), fig.cap="Gaussian process inference outperforms parametric estimates.
 Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."}
+```
+
+```{r Figure4, dependson="profits"}
 ```

### 2013-06-17 inst/doc/nonparametric-bayes.Rmd
code examples updated

@@
 ```{r generics-and-libraries, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
-source("~/.knitr_defaults.R")
-library(knitcitations)
 library(nonparametricbayes) 
@@
 
-```{r plotting-options, echo=FALSE, cache=FALSE, include=FALSE}
+```{r pdf-options, echo=FALSE, cache=FALSE, include=FALSE}
+library(ggplot2)
 options(xtable.print.comment=FALSE)
 options(xtable.type = 'latex', table.placement="H")
-opts_chunk$set(warning=FALSE, message=FALSE, comment=NA,
-               tidy=FALSE, echo=FALSE)
 opts_knit$set(progress = TRUE, verbose = TRUE)
@@
  cache.path = 'cache-pdf/', cache=TRUE, external=TRUE)
+opts_chunk$set(warning=FALSE, message=FALSE, comment=NA,
+               tidy=FALSE, echo=FALSE)
 ```
@@
 
-```{r Figure3, dependson=c("sim", "plotting-options"), fig.cap="Gaussian process inference outperforms parametric estimates.
+```{r Figure3, dependson=c("sim", "pdf-options"), fig.cap="Gaussian process inference outperforms parametric estimates.
 Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."}
@@
 
-```{r Figure4, dependson="profits", fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
+```{r Figure4, dependson=c("profits", "pdf-options"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
 For each data set, the three models are estimated as described above.

### 2013-06-17 inst/examples/BUGS/external-chunks.R

@@ -30,7 +30,7 @@ x_grid <- seq(0, 1.5 * K, length=50)
 h_grid <- x_grid
 profit <- function(x,h) pmin(x, h)
 delta <- 0.01
-OptTime <- 50  # stationarity with unstable models is tricky thing
+OptTime <- 100  # stationarity with unstable models is tricky thing
 reward <- 0
 xT <- 0
 Xo <-  allee+.5# observations start from
@@ -352,10 +352,10 @@ par_priors <- list( deviance = function(x) 0 * x, K = K_prior,
 ## @knitr myers-mcmc
 jags.params=c("r0", "theta", "K", "stdQ")
 jags.inits <- function(){
-  list("r0"= rlnorm(1,0,.1), 
-       "K"=    10 * rlnorm(1,0,.1),
+  list("r0"= 2 * rlnorm(1,0,.1), 
+       "K"=    8 * rlnorm(1,0,.1),
        "theta" = 1 * rlnorm(1,0,.1),  
-       "stdQ"= sqrt(0.2) * rlnorm(1,0,.1),
+       "stdQ"= 0.1 * rlnorm(1,0,.1),
        .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
 }
 set.seed(12345)
@@ -535,22 +535,18 @@ ggplot(sims_data) +
 
 ## @knitr profits
 Profit <- sims_data[, sum(profit), by=c("reps", "method")]
-Profit[, mean(V1), by="method"]
+tmp <- dcast(Profit, reps ~ method)
+tmp <- tmp / tmp[,"True"]
+tmp <- melt(tmp[2:dim(tmp)[2]])
+actual_over_optimal <-subset(tmp, variable != "True")
 
 
 ## @knitr Figure4 
-ggplot(Profit, aes(V1)) + geom_histogram() + 
-  facet_wrap(~method, scales = "free_y") + guides(legend.position = "none") + xlab("Total profit by replicate")
-
-
-## @knitr deviances
-allen_deviance <- posterior.mode(pardist[,'deviance'])
-ricker_deviance <- posterior.mode(ricker_pardist[,'deviance'])
-myers_deviance <- posterior.mode(myers_pardist[,'deviance'])
-true_deviance <- 2*estf(c(p, sigma_g))
-mle_deviance <- 2*estf(c(est$p, est$sigma_g))
-
-c(allen = allen_deviance, ricker=ricker_deviance, myers=myers_deviance, true=true_deviance, mle=mle_deviance)
-
-
[…]

### 2013-06-18 inst/doc/nonparametric-bayes.Rmd
include color palette

@@
 ```{r pdf-options, echo=FALSE, cache=FALSE, include=FALSE}
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 library(ggplot2)

### 2013-06-18 inst/doc/nonparametric-bayes.Rmd
run in small noise regime

--- 2013-06-17 inst/doc/nonparametric-bayes.Rmd
@@
 opts_knit$set(progress = TRUE, verbose = TRUE)
-opts_chunk$set(dev = 'Cairo_pdf', fig.width=5.5, fig.height=4,
+opts_chunk$set(dev = 'pdf', fig.width=5.5, fig.height=4,
  cache.path = 'cache-pdf/', cache=TRUE, external=TRUE)

### 2013-06-18 inst/doc/nonparametric-bayes.Rmd
supress printing of mcmc diagnostics until appendix

--- 2013-06-17 inst/doc/nonparametric-bayes.Rmd
@@
 library(nonparametricbayes) 
+library(ggplot2)
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 opts_chunk$set(external=TRUE)
@@
 
-```{r pdf-options, echo=FALSE, cache=FALSE, include=FALSE}
-library(ggplot2)
+```{r plot-options, echo=FALSE, cache=FALSE, include=FALSE}
 options(xtable.print.comment=FALSE)
@@
 opts_chunk$set(dev = 'Cairo_pdf', fig.width=5.5, fig.height=4,
- cache.path = 'cache-pdf/', cache=TRUE, external=TRUE)
+ cache.path = 'cache/nonparametric-bayes/', cache=TRUE, external=TRUE)
 opts_chunk$set(warning=FALSE, message=FALSE, comment=NA,
@@
 ```
-```{r ricker-mcmc, dependson="ricker-model"}
+```{r ricker-mcmc, dependson="ricker-model", results="hide"}
 ```
@@
 ```
-```{r myers-mcmc, dependson="myers-model"}
+```{r myers-mcmc, dependson="myers-model", results="hide"}
 ```
@@
 
-```{r Figure1, dependson=c("assemble-models"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
+```{r Figure1, dependson=c("assemble-models", "plot-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
 Graph shows the expected value for the function $f$ under each model.
@@
 ```
-```{r Figureb, dependson=c("assemble-models", "par-fns")}
+```{r Figure1b, dependson=c("assemble-models", "par-fns", "plot-options")}
 ```
@@
 
-```{r Figure2, dependson="assemble-opt", fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
+```{r Figure2, dependson = c("assemble-opt", "plot-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
 Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
@@
 
-```{r Figure3, dependson=c("sim", "pdf-options"), fig.cap="Gaussian process inference outperforms parametric estimates.
+```{r Figure3, dependson=c("sim", "plot-options"), fig.cap="Gaussian process inference outperforms parametric estimates.
 Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."}
@@
 
-```{r Figure4, dependson=c("profits", "pdf-options"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
+```{r Figure4, dependson=c("profits", "plot-options"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
 For each data set, the three models are estimated as described above.
@@
 
-![Traces of the GP hyper-parameters (after transforming back from log space)](gp_traces_densities1.pdf)
-
-![Posterior densities of the GP hyper-parameters (after transforming back from log space). Prior distribution over the interval sampled shown as red curves.](gp_traces_densities2.pdf)
-
-![Traces of the Ricker parameters](ricker_traces.pdf)
-
-![Posterior densities of the Ricker parameter estimates.
-Prior distributions over the interval sampled shown as red curves.
-Note the bimodal estimate of $\log(r)$ which leads to the poor fit in Figure 1](ricker_posteriors.pdf)
-
-![Traces of the structurally accurate Allen model parameters](parametric_bayes_traces.pdf)
-
-![Posterior densities for the parameters of the Allen model. Prior distributions over the interval sampled shown as red curves.](parametric_bayes_posteriors.pdf)
+```{r appendixplots, dependson="assemble-models"}
+```
 

### 2013-06-18 inst/examples/BUGS/external-chunks.R

@@ -16,7 +16,7 @@ posterior.mode <- function(x) {
 
 ## @knitr stateeq
 f <- RickerAllee
-p <- c(1, 10, 5)
+p <- c(2, 8, 5)
 K <- 10  # approx, a li'l' less
 allee <- 5 # approx, a li'l' less
 
@@ -25,12 +25,12 @@ allee <- 5 # approx, a li'l' less
 sigma_g <- 0.05
 sigma_m <- 0.0
 z_g <- function() rlnorm(1, 0, sigma_g)
-z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
+z_m <- function() 1
 x_grid <- seq(0, 1.5 * K, length=50)
 h_grid <- x_grid
 profit <- function(x,h) pmin(x, h)
 delta <- 0.01
-OptTime <- 100  # stationarity with unstable models is tricky thing
+OptTime <- 50  # stationarity with unstable models is tricky thing
 reward <- 0
 xT <- 0
 Xo <-  allee+.5# observations start from
@@ -90,7 +90,7 @@ gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)
 
 
 ## @knitr gp_traces_densities
-plots <- summary_gp_mcmc(gp, burnin=1e4, thin=300)
+gp_assessment_plots <- summary_gp_mcmc(gp, burnin=1e4, thin=300)
 
 
 ## @knitr gp-output
@@ -188,7 +188,7 @@ allen_jags <- do.call(autojags,
 tmp <- lapply(as.mcmc(allen_jags), as.matrix) # strip classes the hard way...
 allen_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
 names(allen_posteriors) = c("index", "variable", "value", "chain")
-ggplot(allen_posteriors) + geom_line(aes(index, value)) + 
+plot_allen_traces <- ggplot(allen_posteriors) + geom_line(aes(index, value)) + 
   facet_wrap(~ variable, scale="free", ncol=1)
 
 
@@ -197,8 +197,7 @@ allen_priors <- ddply(allen_posteriors, "variable", function(dd){
     grid <- seq(min(dd$value), max(dd$value), length = 100) 
     data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
 })
-
-ggplot(allen_posteriors, aes(value)) + 
+plot_allen_posteriors <- ggplot(allen_posteriors, aes(value)) + 
[…]

### 2013-06-18 inst/examples/BUGS/external-chunks.R
run on farm

@@ -22,7 +22,7 @@ allee <- 5 # approx, a li'l' less
 
 
 ## @knitr sdp-pars
-sigma_g <- 0.05
+sigma_g <- 0.15
 sigma_m <- 0.0
 z_g <- function() rlnorm(1, 0, sigma_g)
 z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m

### 2013-06-19 inst/examples/BUGS/external-chunks.R
move raw data plot to appendix

@@ -52,7 +52,7 @@ MaxT = 1000 # timeout for value iteration convergence
                     y = c(rep(0,nz), 
                           x[2:Tobs]))
 raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
-raw_plot
+
 
 
 ## @knitr mle
@@ -560,6 +560,7 @@ c(allen = allen_deviance, ricker=ricker_deviance, myers=myers_deviance, true=tru
 
 
 ## @knitr appendixplots
+raw_plot
 gp_assessment_plots[[1]]
 gp_assessment_plots[[2]]
 plot_allen_traces

### 2013-06-19 inst/doc/nonparametric-bayes.Rmd

@@
 opts_knit$set(progress = TRUE, verbose = TRUE)
-opts_chunk$set(dev = 'Cairo_pdf', fig.width=5.5, fig.height=4,
+opts_chunk$set(dev = c('pdf', 'png'), fig.width=5.5, fig.height=4,
  cache.path = 'cache/nonparametric-bayes/', cache=TRUE, external=TRUE)

### 2013-06-19 inst/doc/nonparametric-bayes.Rmd
update manuscript examples

--- 2013-06-18 inst/doc/nonparametric-bayes.Rmd
@@
 opts_knit$set(progress = TRUE, verbose = TRUE)
-opts_chunk$set(dev = 'Cairo_pdf', fig.width=5.5, fig.height=4,
+opts_chunk$set(dev = c('pdf', 'png'), fig.width=5.5, fig.height=4,
  cache.path = 'cache/nonparametric-bayes/', cache=TRUE, external=TRUE)
 opts_chunk$set(warning=FALSE, message=FALSE, comment=NA,
-               tidy=FALSE, echo=FALSE)
+               tidy=FALSE)
+toggle = "markup" # "hide" 
 ```
@@
 ```
-```{r mle-output, dependson="mle"}
+```{r mle-output, dependson="mle", results=toggle}
 ```
@@
 ```
-```{r gp-output, dependson="gp"}
+```{r gp-output, dependson="gp", results=toggle}
 ```
@@
 ```
-```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results="hide"}
+```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results=toggle}
 ```
@@
 ```
-```{r allen-output, dependson=c("posterior-mode", "allen-traces")}
+```{r allen-output, dependson=c("posterior-mode", "allen-traces"), results=toggle}
 ```
@@
 ```
-```{r ricker-mcmc, dependson="ricker-model", results="hide"}
+```{r ricker-mcmc, dependson="ricker-model", results=toggle}
 ```
@@
 ```
-```{r ricker-output, dependson=c("posterior-mode", "ricker-traces")}
+```{r ricker-output, dependson=c("posterior-mode", "ricker-traces"), results=toggle}
 ```
@@
 ```
-```{r myers-mcmc, dependson="myers-model", results="hide"}
+```{r myers-mcmc, dependson="myers-model", results=toggle}
 ```
@@
 ```
-```{r myers-output, dependson=c("posterior-mode", "myers-traces")}
+```{r myers-output, dependson=c("posterior-mode", "myers-traces"), results=toggle}
 ```
@@
 
-```{r profits, dependson="sims", results="hide"}
+```{r profits, dependson="sims", results=toggle}
 ```
-```{r deviances, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output"), results="hide"}
+```{r deviances, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output"), results=toggle}
 ```

### 2013-06-19 inst/doc/nonparametric-bayes.tex
update manuscript examples

--- 2013-05-03 inst/doc/nonparametric-bayes.tex
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{require}\NormalTok{(modeest)}
+\NormalTok{posterior.mode <- function(x) \{}
+  \KeywordTok{mlv}\NormalTok{(x, }\DataTypeTok{method=}\StringTok{"shorth"}\NormalTok{)$M}
+\NormalTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{f <- RickerAllee}
+\NormalTok{p <- }\KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{5}\NormalTok{)}
+\NormalTok{K <- }\DecValTok{10}  \CommentTok{# approx, a li'l' less}
+\NormalTok{allee <- }\DecValTok{5} \CommentTok{# approx, a li'l' less}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{sigma_g <- }\FloatTok{0.05}
+\NormalTok{sigma_m <- }\FloatTok{0.0}
+\NormalTok{z_g <- function() }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{, sigma_g)}
+\NormalTok{z_m <- function() }\DecValTok{1}
+\NormalTok{x_grid <- }\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\FloatTok{1.5} \NormalTok{* K, }\DataTypeTok{length=}\DecValTok{50}\NormalTok{)}
+\NormalTok{h_grid <- x_grid}
+\NormalTok{profit <- function(x,h) }\KeywordTok{pmin}\NormalTok{(x, h)}
+\NormalTok{delta <- }\FloatTok{0.01}
+\NormalTok{OptTime <- }\DecValTok{50}  \CommentTok{# stationarity with unstable models is tricky thing}
+\NormalTok{reward <- }\DecValTok{0}
+\NormalTok{xT <- }\DecValTok{0}
+\NormalTok{Xo <-  allee}\FloatTok{+.5}\CommentTok{# observations start from}
+\NormalTok{x0 <- K }\CommentTok{# simulation under policy starts from}
+\NormalTok{Tobs <- }\DecValTok{40}
+\NormalTok{MaxT = }\DecValTok{1000} \CommentTok{# timeout for value iteration convergence}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+  \KeywordTok{set.seed}\NormalTok{(}\DecValTok{1234}\NormalTok{)}
+  \CommentTok{#harvest <- sort(rep(seq(0, .5, length=7), 5))}
+  \NormalTok{x <- }\KeywordTok{numeric}\NormalTok{(Tobs)}
+  \NormalTok{x[}\DecValTok{1}\NormalTok{] <- Xo}
+  \NormalTok{nz <- }\DecValTok{1}
+  \NormalTok{for(t in }\DecValTok{1}\NormalTok{:(Tobs}\DecValTok{-1}\NormalTok{))}
+    \NormalTok{x[t}\DecValTok{+1}\NormalTok{] = }\KeywordTok{z_g}\NormalTok{() * }\KeywordTok{f}\NormalTok{(x[t], }\DataTypeTok{h=}\DecValTok{0}\NormalTok{, }\DataTypeTok{p=}\NormalTok{p)}
+  \NormalTok{obs <- }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,nz), }
+                          \KeywordTok{pmax}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,Tobs}\DecValTok{-1}\NormalTok{), x[}\DecValTok{1}\NormalTok{:(Tobs}\DecValTok{-1}\NormalTok{)])), }
+                    \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,nz), }
+                          \NormalTok{x[}\DecValTok{2}\NormalTok{:Tobs]))}
+\NormalTok{raw_plot <- }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{time =} \DecValTok{1}\NormalTok{:Tobs, }\DataTypeTok{x=}\NormalTok{x), }\KeywordTok{aes}\NormalTok{(time,x)) + }\KeywordTok{geom_line}\NormalTok{()}
+\NormalTok{raw_plot}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/obs.pdf}
+\caption{plot of chunk obs}
+\end{figure}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{set.seed}\NormalTok{(}\DecValTok{12345}\NormalTok{)}
+\NormalTok{estf <- function(p)\{ }
+    \NormalTok{mu <- }\KeywordTok{f}\NormalTok{(obs$x,}\DecValTok{0}\NormalTok{,p)}
+    \NormalTok{-}\KeywordTok{sum}\NormalTok{(}\KeywordTok{dlnorm}\NormalTok{(obs$y, }\KeywordTok{log}\NormalTok{(mu), p[}\DecValTok{4}\NormalTok{]), }\DataTypeTok{log=}\OtherTok{TRUE}\NormalTok{)}
+\NormalTok{\}}
+\NormalTok{par <- }\KeywordTok{c}\NormalTok{(p[}\DecValTok{1}\NormalTok{]*}\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{1}\NormalTok{), }
+         \NormalTok{p[}\DecValTok{2}\NormalTok{]*}\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{1}\NormalTok{), }
+         \NormalTok{p[}\DecValTok{3}\NormalTok{]*}\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{, .}\DecValTok{1}\NormalTok{), }
+         \NormalTok{sigma_g * }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{1}\NormalTok{))}
+\NormalTok{o <- }\KeywordTok{optim}\NormalTok{(par, estf, }\DataTypeTok{method=}\StringTok{"L"}\NormalTok{, }\DataTypeTok{lower=}\KeywordTok{c}\NormalTok{(}\FloatTok{1e-5}\NormalTok{,}\FloatTok{1e-5}\NormalTok{,}\FloatTok{1e-5}\NormalTok{,}\FloatTok{1e-5}\NormalTok{))}
+\NormalTok{f_alt <- f}
+\NormalTok{p_alt <- }\KeywordTok{c}\NormalTok{(}\KeywordTok{as.numeric}\NormalTok{(o$par[}\DecValTok{1}\NormalTok{]), }\KeywordTok{as.numeric}\NormalTok{(o$par[}\DecValTok{2}\NormalTok{]), }\KeywordTok{as.numeric}\NormalTok{(o$par[}\DecValTok{3}\NormalTok{]))}
+\NormalTok{sigma_g_alt <- }\KeywordTok{as.numeric}\NormalTok{(o$par[}\DecValTok{4}\NormalTok{])}
+
+\NormalTok{est <- }\KeywordTok{list}\NormalTok{(}\DataTypeTok{f =} \NormalTok{f_alt, }\DataTypeTok{p =} \NormalTok{p_alt, }\DataTypeTok{sigma_g =} \NormalTok{sigma_g_alt, }\DataTypeTok{mloglik=}\NormalTok{o$value)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{true_means <- }\KeywordTok{sapply}\NormalTok{(x_grid, f, }\DecValTok{0}\NormalTok{, p)}
+\NormalTok{est_means <- }\KeywordTok{sapply}\NormalTok{(x_grid, est$f, }\DecValTok{0}\NormalTok{, est$p)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)}
+\NormalTok{s2.p <- }\KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{5}\NormalTok{)  }
+\NormalTok{d.p = }\KeywordTok{c}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{1}\NormalTok{/}\FloatTok{0.1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{gp <- }\KeywordTok{gp_mcmc}\NormalTok{(obs$x, }\DataTypeTok{y=}\NormalTok{obs$y, }\DataTypeTok{n=}\FloatTok{1e5}\NormalTok{, }\DataTypeTok{s2.p =} \NormalTok{s2.p, }\DataTypeTok{d.p =} \NormalTok{d.p)}
+\NormalTok{gp_dat <- }\KeywordTok{gp_predict}\NormalTok{(gp, x_grid, }\DataTypeTok{burnin=}\FloatTok{1e4}\NormalTok{, }\DataTypeTok{thin=}\DecValTok{300}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{gp_assessment_plots <- }\KeywordTok{summary_gp_mcmc}\NormalTok{(gp, }\DataTypeTok{burnin=}\FloatTok{1e4}\NormalTok{, }\DataTypeTok{thin=}\DecValTok{300}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# Summarize the GP model}
+\NormalTok{tgp_dat <- }
+    \KeywordTok{data.frame}\NormalTok{(  }\DataTypeTok{x =} \NormalTok{x_grid, }
+                 \DataTypeTok{y =} \NormalTok{gp_dat$E_Ef, }
+                 \DataTypeTok{ymin =} \NormalTok{gp_dat$E_Ef - }\DecValTok{2} \NormalTok{* }\KeywordTok{sqrt}\NormalTok{(gp_dat$E_Vf), }
+                 \DataTypeTok{ymax =} \NormalTok{gp_dat$E_Ef + }\DecValTok{2} \NormalTok{* }\KeywordTok{sqrt}\NormalTok{(gp_dat$E_Vf) )}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{y <- x }
+\NormalTok{N <- }\KeywordTok{length}\NormalTok{(x);}
+\NormalTok{jags.data <- }\KeywordTok{list}\NormalTok{(}\StringTok{"N"}\NormalTok{,}\StringTok{"y"}\NormalTok{)}
+\NormalTok{n.chains <- }\DecValTok{6}
+\NormalTok{n.iter <- }\FloatTok{1e6}
+\NormalTok{n.burnin <- }\KeywordTok{floor}\NormalTok{(}\DecValTok{10000}\NormalTok{)}
+\NormalTok{n.thin <- }\KeywordTok{max}\NormalTok{(}\DecValTok{1}\NormalTok{, }\KeywordTok{floor}\NormalTok{(n.chains * (n.iter - n.burnin)/}\DecValTok{1000}\NormalTok{))}
+\NormalTok{n.update <- }\DecValTok{10}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{stdQ_prior_p <- }\KeywordTok{c}\NormalTok{(}\FloatTok{1e-6}\NormalTok{, }\DecValTok{100}\NormalTok{)}
+\NormalTok{stdR_prior_p <- }\KeywordTok{c}\NormalTok{(}\FloatTok{1e-6}\NormalTok{, .}\DecValTok{1}\NormalTok{)}
+\NormalTok{stdQ_prior  <- function(x) }\KeywordTok{dunif}\NormalTok{(x, stdQ_prior_p[}\DecValTok{1}\NormalTok{], stdQ_prior_p[}\DecValTok{2}\NormalTok{])}
+\NormalTok{stdR_prior  <- function(x) }\KeywordTok{dunif}\NormalTok{(x, stdR_prior_p[}\DecValTok{1}\NormalTok{], stdR_prior_p[}\DecValTok{2}\NormalTok{])}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{K_prior_p <- }\KeywordTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{20.0}\NormalTok{)}
+\NormalTok{r0_prior_p <- }\KeywordTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{6.0}\NormalTok{)}
+\NormalTok{theta_prior_p <- }\KeywordTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{20.0}\NormalTok{)}
+
+\NormalTok{bugs.model <- }
+\KeywordTok{paste}\NormalTok{(}\KeywordTok{sprintf}\NormalTok{(}
+\StringTok{"model\{}
+\StringTok{  K     ~ dunif(%s, %s)}
+\StringTok{  r0    ~ dunif(%s, %s)}
+\StringTok{  theta ~ dunif(%s, %s)}
+\StringTok{  stdQ ~ dunif(%s, %s)"}\NormalTok{, }
+  \NormalTok{K_prior_p[}\DecValTok{1}\NormalTok{], K_prior_p[}\DecValTok{2}\NormalTok{],}
+  \NormalTok{r0_prior_p[}\DecValTok{1}\NormalTok{], r0_prior_p[}\DecValTok{2}\NormalTok{],}
+  \NormalTok{theta_prior_p[}\DecValTok{1}\NormalTok{], theta_prior_p[}\DecValTok{2}\NormalTok{],}
+  \NormalTok{stdQ_prior_p[}\DecValTok{1}\NormalTok{], stdQ_prior_p[}\DecValTok{2}\NormalTok{]),}
+
+  \StringTok{"}
+\StringTok{  iQ <- 1 / (stdQ * stdQ);}
+\StringTok{  y[1] ~ dunif(0, 10)}
+\StringTok{  for(t in 1:(N-1))\{}
+\StringTok{    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K)* (y[t] - theta) / K }
+\StringTok{    y[t+1] ~ dlnorm(mu[t], iQ) }
+\StringTok{  \}}
+\StringTok{\}"}\NormalTok{)}
+\KeywordTok{writeLines}\NormalTok{(bugs.model, }\StringTok{"allen_process.bugs"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{K_prior     <- function(x) }\KeywordTok{dunif}\NormalTok{(x, K_prior_p[}\DecValTok{1}\NormalTok{], K_prior_p[}\DecValTok{2}\NormalTok{])}
+\NormalTok{r0_prior <- function(x) }\KeywordTok{dunif}\NormalTok{(x, r0_prior_p[}\DecValTok{1}\NormalTok{], r0_prior_p[}\DecValTok{2}\NormalTok{])}
+\NormalTok{theta_prior <- function(x) }\KeywordTok{dunif}\NormalTok{(x, theta_prior_p[}\DecValTok{1}\NormalTok{], theta_prior_p[}\DecValTok{2}\NormalTok{])}
+\NormalTok{par_priors  <- }\KeywordTok{list}\NormalTok{(}\DataTypeTok{K =} \NormalTok{K_prior, }\DataTypeTok{deviance =} \NormalTok{function(x) }\DecValTok{0} \NormalTok{* x, }
+                    \DataTypeTok{r0 =} \NormalTok{r0_prior, }\DataTypeTok{theta =} \NormalTok{theta_prior,}
+                    \DataTypeTok{stdQ =} \NormalTok{stdQ_prior)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{jags.params=}\KeywordTok{c}\NormalTok{(}\StringTok{"K"}\NormalTok{,}\StringTok{"r0"}\NormalTok{,}\StringTok{"theta"}\NormalTok{,}\StringTok{"stdQ"}\NormalTok{) }\CommentTok{# be sensible about the order here}
+\NormalTok{jags.inits <- function()\{}
+  \KeywordTok{list}\NormalTok{(}\StringTok{"K"}\NormalTok{= }\DecValTok{10} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{, }\FloatTok{0.1}\NormalTok{),}
+       \StringTok{"r0"}\NormalTok{= }\DecValTok{1} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{, }\FloatTok{0.1}\NormalTok{) ,}
+       \StringTok{"theta"}\NormalTok{=   }\DecValTok{5} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{, }\FloatTok{0.1}\NormalTok{) , }
+       \StringTok{"stdQ"}\NormalTok{= }\KeywordTok{abs}\NormalTok{( }\FloatTok{0.1} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{, }\FloatTok{0.1}\NormalTok{)),}
+       \DataTypeTok{.RNG.name=}\StringTok{"base::Wichmann-Hill"}\NormalTok{, }\DataTypeTok{.RNG.seed=}\DecValTok{123}\NormalTok{)}
+\NormalTok{\}}
+
+\KeywordTok{set.seed}\NormalTok{(}\DecValTok{1234}\NormalTok{)}
+\CommentTok{# parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)}
+\NormalTok{allen_jags <- }\KeywordTok{do.call}\NormalTok{(jags.parallel, }\KeywordTok{list}\NormalTok{(}\DataTypeTok{data=}\NormalTok{jags.data, }\DataTypeTok{inits=}\NormalTok{jags.inits, }
+                                      \NormalTok{jags.params, }\DataTypeTok{n.chains=}\NormalTok{n.chains, }
+                                      \DataTypeTok{n.iter=}\NormalTok{n.iter, }\DataTypeTok{n.thin=}\NormalTok{n.thin, }
+                                      \DataTypeTok{n.burnin=}\NormalTok{n.burnin, }
+                                      \DataTypeTok{model.file=}\StringTok{"allen_process.bugs"}\NormalTok{))}
+
+\CommentTok{# Run again iteratively if we haven't met the Gelman-Rubin convergence criterion}
+\KeywordTok{recompile}\NormalTok{(allen_jags) }\CommentTok{# required for parallel}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 328
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 328
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 328
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 328
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 328
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 328
+
+Initializing model
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{allen_jags <- }\KeywordTok{do.call}\NormalTok{(autojags, }
+                                            \KeywordTok{list}\NormalTok{(}\DataTypeTok{object=}\NormalTok{allen_jags, }\DataTypeTok{n.update=}\NormalTok{n.update, }
+                           \DataTypeTok{n.iter=}\NormalTok{n.iter, }\DataTypeTok{n.thin =} \NormalTok{n.thin))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tmp <- }\KeywordTok{lapply}\NormalTok{(}\KeywordTok{as.mcmc}\NormalTok{(allen_jags), as.matrix) }\CommentTok{# strip classes the hard way...}
+\NormalTok{allen_posteriors <- }\KeywordTok{melt}\NormalTok{(tmp, }\DataTypeTok{id =} \KeywordTok{colnames}\NormalTok{(tmp[[}\DecValTok{1}\NormalTok{]])) }
+\KeywordTok{names}\NormalTok{(allen_posteriors) = }\KeywordTok{c}\NormalTok{(}\StringTok{"index"}\NormalTok{, }\StringTok{"variable"}\NormalTok{, }\StringTok{"value"}\NormalTok{, }\StringTok{"chain"}\NormalTok{)}
+\NormalTok{plot_allen_traces <- }\KeywordTok{ggplot}\NormalTok{(allen_posteriors) + }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(index, value)) + }
+  \KeywordTok{facet_wrap}\NormalTok{(~ variable, }\DataTypeTok{scale=}\StringTok{"free"}\NormalTok{, }\DataTypeTok{ncol=}\DecValTok{1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{allen_priors <- }\KeywordTok{ddply}\NormalTok{(allen_posteriors, }\StringTok{"variable"}\NormalTok{, function(dd)\{}
+    \NormalTok{grid <- }\KeywordTok{seq}\NormalTok{(}\KeywordTok{min}\NormalTok{(dd$value), }\KeywordTok{max}\NormalTok{(dd$value), }\DataTypeTok{length =} \DecValTok{100}\NormalTok{) }
+    \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{value =} \NormalTok{grid, }\DataTypeTok{density =} \NormalTok{par_priors[[dd$variable[}\DecValTok{1}\NormalTok{]]](grid))}
+\NormalTok{\})}
+\NormalTok{plot_allen_posteriors <- }\KeywordTok{ggplot}\NormalTok{(allen_posteriors, }\KeywordTok{aes}\NormalTok{(value)) + }
+  \KeywordTok{stat_density}\NormalTok{(}\DataTypeTok{geom=}\StringTok{"path"}\NormalTok{, }\DataTypeTok{position=}\StringTok{"identity"}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.7}\NormalTok{) +}
+  \KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{data=}\NormalTok{allen_priors, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{value, }\DataTypeTok{y=}\NormalTok{density), }\DataTypeTok{col=}\StringTok{"red"}\NormalTok{) + }
+  \KeywordTok{facet_wrap}\NormalTok{(~ variable, }\DataTypeTok{scale=}\StringTok{"free"}\NormalTok{, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{A <- allen_posteriors}
+\NormalTok{A$index <- A$index + A$chain * }\KeywordTok{max}\NormalTok{(A$index) }\CommentTok{# Combine samples across chains by renumbering index }
+\NormalTok{pardist <- }\KeywordTok{acast}\NormalTok{(A, index ~ variable)}
+\NormalTok{bayes_coef <- }\KeywordTok{apply}\NormalTok{(pardist,}\DecValTok{2}\NormalTok{, posterior.mode) }
+\NormalTok{bayes_pars <- }\KeywordTok{unname}\NormalTok{(}\KeywordTok{c}\NormalTok{(bayes_coef[}\StringTok{"r0"}\NormalTok{], bayes_coef[}\StringTok{"K"}\NormalTok{], bayes_coef[}\StringTok{"theta"}\NormalTok{])) }\CommentTok{# parameters formatted for f}
+\NormalTok{allen_f <- function(x,h,p) }\KeywordTok{unname}\NormalTok{(}\KeywordTok{RickerAllee}\NormalTok{(x,h, }\KeywordTok{unname}\NormalTok{(p[}\KeywordTok{c}\NormalTok{(}\StringTok{"r0"}\NormalTok{, }\StringTok{"K"}\NormalTok{, }\StringTok{"theta"}\NormalTok{)])))}
+\NormalTok{allen_means <- }\KeywordTok{sapply}\NormalTok{(x_grid, f, }\DecValTok{0}\NormalTok{, bayes_pars)}
+\NormalTok{bayes_pars}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+[1] 0.5003 7.7115 4.6039
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{head}\NormalTok{(pardist)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+        K deviance     r0    stdQ theta
+170 7.898    34.41 0.6668 0.05566 3.956
+171 7.745    32.76 0.7554 0.05137 4.668
+172 7.577    31.66 1.0705 0.04562 4.073
+173 7.859    31.26 1.5647 0.04238 5.087
+174 7.821    30.67 1.7348 0.04662 5.230
+175 8.003    34.90 1.4495 0.05005 5.444
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{K_prior_p <- }\KeywordTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{40.0}\NormalTok{)}
+\NormalTok{r0_prior_p <- }\KeywordTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{20.0}\NormalTok{)}
+\NormalTok{bugs.model <- }
+\KeywordTok{paste}\NormalTok{(}\KeywordTok{sprintf}\NormalTok{(}
+\StringTok{"model\{}
+\StringTok{  K    ~ dunif(%s, %s)}
+\StringTok{  r0    ~ dunif(%s, %s)}
+\StringTok{  stdQ ~ dunif(%s, %s)"}\NormalTok{, }
+  \NormalTok{K_prior_p[}\DecValTok{1}\NormalTok{], K_prior_p[}\DecValTok{2}\NormalTok{],}
+  \NormalTok{r0_prior_p[}\DecValTok{1}\NormalTok{], r0_prior_p[}\DecValTok{2}\NormalTok{],}
+  \NormalTok{stdQ_prior_p[}\DecValTok{1}\NormalTok{], stdQ_prior_p[}\DecValTok{2}\NormalTok{]),}
+  \StringTok{"}
+\StringTok{  iQ <- 1 / (stdQ * stdQ);}
+\StringTok{  y[1] ~ dunif(0, 10)}
+\StringTok{  for(t in 1:(N-1))\{}
+\StringTok{    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K) }
+\StringTok{    y[t+1] ~ dlnorm(mu[t], iQ) }
+\StringTok{  \}}
+\StringTok{\}"}\NormalTok{)}
+\KeywordTok{writeLines}\NormalTok{(bugs.model, }\StringTok{"ricker_process.bugs"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{K_prior     <- function(x) }\KeywordTok{dunif}\NormalTok{(x, K_prior_p[}\DecValTok{1}\NormalTok{], K_prior_p[}\DecValTok{2}\NormalTok{])}
+\NormalTok{r0_prior <- function(x) }\KeywordTok{dunif}\NormalTok{(x, r0_prior_p[}\DecValTok{1}\NormalTok{], r0_prior_p[}\DecValTok{2}\NormalTok{])}
+\NormalTok{par_priors <- }\KeywordTok{list}\NormalTok{(}\DataTypeTok{K =} \NormalTok{K_prior, }\DataTypeTok{deviance =} \NormalTok{function(x) }\DecValTok{0} \NormalTok{* x, }
+                   \DataTypeTok{r0 =} \NormalTok{r0_prior, }\DataTypeTok{stdQ =} \NormalTok{stdQ_prior)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{jags.params=}\KeywordTok{c}\NormalTok{(}\StringTok{"K"}\NormalTok{,}\StringTok{"r0"}\NormalTok{, }\StringTok{"stdQ"}\NormalTok{)}
+\NormalTok{jags.inits <- function()\{}
+  \KeywordTok{list}\NormalTok{(}\StringTok{"K"}\NormalTok{= }\DecValTok{10} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
+       \StringTok{"r0"}\NormalTok{= }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
+       \StringTok{"stdQ"}\NormalTok{=}\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.05}\NormalTok{) * }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
+       \DataTypeTok{.RNG.name=}\StringTok{"base::Wichmann-Hill"}\NormalTok{, }\DataTypeTok{.RNG.seed=}\DecValTok{123}\NormalTok{)}
+\NormalTok{\}}
+\KeywordTok{set.seed}\NormalTok{(}\DecValTok{12345}\NormalTok{) }
+\NormalTok{ricker_jags <- }\KeywordTok{do.call}\NormalTok{(jags.parallel, }
+                       \KeywordTok{list}\NormalTok{(}\DataTypeTok{data=}\NormalTok{jags.data, }\DataTypeTok{inits=}\NormalTok{jags.inits, }
+                            \NormalTok{jags.params, }\DataTypeTok{n.chains=}\NormalTok{n.chains, }
+                            \DataTypeTok{n.iter=}\NormalTok{n.iter, }\DataTypeTok{n.thin=}\NormalTok{n.thin, }\DataTypeTok{n.burnin=}\NormalTok{n.burnin,}
+                            \DataTypeTok{model.file=}\StringTok{"ricker_process.bugs"}\NormalTok{))}
+\KeywordTok{recompile}\NormalTok{(ricker_jags)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 249
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 249
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 249
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 249
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 249
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 249
+
+Initializing model
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{ricker_jags <- }\KeywordTok{do.call}\NormalTok{(autojags, }
+                       \KeywordTok{list}\NormalTok{(}\DataTypeTok{object=}\NormalTok{ricker_jags, }\DataTypeTok{n.update=}\NormalTok{n.update, }
+                                                        \DataTypeTok{n.iter=}\NormalTok{n.iter, }\DataTypeTok{n.thin =} \NormalTok{n.thin, }
+                                                        \DataTypeTok{progress.bar=}\StringTok{"none"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tmp <- }\KeywordTok{lapply}\NormalTok{(}\KeywordTok{as.mcmc}\NormalTok{(ricker_jags), as.matrix) }\CommentTok{# strip classes the hard way...}
+\NormalTok{ricker_posteriors <- }\KeywordTok{melt}\NormalTok{(tmp, }\DataTypeTok{id =} \KeywordTok{colnames}\NormalTok{(tmp[[}\DecValTok{1}\NormalTok{]])) }
+\KeywordTok{names}\NormalTok{(ricker_posteriors) = }\KeywordTok{c}\NormalTok{(}\StringTok{"index"}\NormalTok{, }\StringTok{"variable"}\NormalTok{, }\StringTok{"value"}\NormalTok{, }\StringTok{"chain"}\NormalTok{)}
+\NormalTok{plot_ricker_traces <- }\KeywordTok{ggplot}\NormalTok{(ricker_posteriors) + }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(index, value)) + }
+  \KeywordTok{facet_wrap}\NormalTok{(~ variable, }\DataTypeTok{scale=}\StringTok{"free"}\NormalTok{, }\DataTypeTok{ncol=}\DecValTok{1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{ricker_priors <- }\KeywordTok{ddply}\NormalTok{(ricker_posteriors, }\StringTok{"variable"}\NormalTok{, function(dd)\{}
+    \NormalTok{grid <- }\KeywordTok{seq}\NormalTok{(}\KeywordTok{min}\NormalTok{(dd$value), }\KeywordTok{max}\NormalTok{(dd$value), }\DataTypeTok{length =} \DecValTok{100}\NormalTok{) }
+    \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{value =} \NormalTok{grid, }\DataTypeTok{density =} \NormalTok{par_priors[[dd$variable[}\DecValTok{1}\NormalTok{]]](grid))}
+\NormalTok{\})}
+\CommentTok{# plot posterior distributions}
+\NormalTok{plot_ricker_posteriors <- }\KeywordTok{ggplot}\NormalTok{(ricker_posteriors, }\KeywordTok{aes}\NormalTok{(value)) + }
+  \KeywordTok{stat_density}\NormalTok{(}\DataTypeTok{geom=}\StringTok{"path"}\NormalTok{, }\DataTypeTok{position=}\StringTok{"identity"}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.7}\NormalTok{) +}
+  \KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{data=}\NormalTok{ricker_priors, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{value, }\DataTypeTok{y=}\NormalTok{density), }\DataTypeTok{col=}\StringTok{"red"}\NormalTok{) + }
+  \KeywordTok{facet_wrap}\NormalTok{(~ variable, }\DataTypeTok{scale=}\StringTok{"free"}\NormalTok{, }\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{A <- ricker_posteriors}
+\NormalTok{A$index <- A$index + A$chain * }\KeywordTok{max}\NormalTok{(A$index) }\CommentTok{# Combine samples across chains by renumbering index }
+\NormalTok{ricker_pardist <- }\KeywordTok{acast}\NormalTok{(A, index ~ variable)}
+\NormalTok{bayes_coef <- }\KeywordTok{apply}\NormalTok{(ricker_pardist,}\DecValTok{2}\NormalTok{, posterior.mode) }
+\NormalTok{ricker_bayes_pars <- }\KeywordTok{unname}\NormalTok{(}\KeywordTok{c}\NormalTok{(bayes_coef[}\StringTok{"r0"}\NormalTok{], bayes_coef[}\StringTok{"K"}\NormalTok{]))}
+\NormalTok{ricker_f <- function(x,h,p)\{}
+  \KeywordTok{sapply}\NormalTok{(x, function(x)\{ }
+    \NormalTok{x <- }\KeywordTok{pmax}\NormalTok{(}\DecValTok{0}\NormalTok{, x-h) }
+    \KeywordTok{pmax}\NormalTok{(}\DecValTok{0}\NormalTok{, x * }\KeywordTok{exp}\NormalTok{(p[}\StringTok{"r0"}\NormalTok{] * (}\DecValTok{1} \NormalTok{- x / p[}\StringTok{"K"}\NormalTok{] )) )}
+  \NormalTok{\})}
+\NormalTok{\}}
+\NormalTok{ricker_means <- }\KeywordTok{sapply}\NormalTok{(x_grid, Ricker, }\DecValTok{0}\NormalTok{, ricker_bayes_pars[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{)])}
+\KeywordTok{head}\NormalTok{(ricker_pardist)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+         K deviance      r0    stdQ
+170  7.727    35.73 0.29440 0.05197
+171  7.949    37.93 0.18845 0.05872
+172 32.183    44.54 0.02474 0.05688
+173  8.149    37.59 0.12204 0.05407
+174  7.460    39.01 0.38643 0.05495
+175  7.361    35.54 0.23805 0.05150
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{ricker_bayes_pars}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+[1] 0.1997 7.5976
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{r0_prior_p <- }\KeywordTok{c}\NormalTok{(.}\DecValTok{0001}\NormalTok{, }\FloatTok{10.0}\NormalTok{)}
+\NormalTok{theta_prior_p <- }\KeywordTok{c}\NormalTok{(.}\DecValTok{0001}\NormalTok{, }\FloatTok{10.0}\NormalTok{)}
+\NormalTok{K_prior_p <- }\KeywordTok{c}\NormalTok{(.}\DecValTok{0001}\NormalTok{, }\FloatTok{40.0}\NormalTok{)}
+\NormalTok{bugs.model <- }
+\KeywordTok{paste}\NormalTok{(}\KeywordTok{sprintf}\NormalTok{(}
+\StringTok{"model\{}
+\StringTok{  r0    ~ dunif(%s, %s)}
+\StringTok{  theta    ~ dunif(%s, %s)}
+\StringTok{  K    ~ dunif(%s, %s)}
+\StringTok{  stdQ ~ dunif(%s, %s)"}\NormalTok{, }
+  \NormalTok{r0_prior_p[}\DecValTok{1}\NormalTok{], r0_prior_p[}\DecValTok{2}\NormalTok{],}
+  \NormalTok{theta_prior_p[}\DecValTok{1}\NormalTok{], theta_prior_p[}\DecValTok{2}\NormalTok{],}
+  \NormalTok{K_prior_p[}\DecValTok{1}\NormalTok{], K_prior_p[}\DecValTok{2}\NormalTok{],}
+  \NormalTok{stdQ_prior_p[}\DecValTok{1}\NormalTok{], stdQ_prior_p[}\DecValTok{2}\NormalTok{]),}
+
+  \StringTok{"}
+\StringTok{  iQ <- 1 / (stdQ * stdQ);}
+
+\StringTok{  y[1] ~ dunif(0, 10)}
+\StringTok{  for(t in 1:(N-1))\{}
+\StringTok{    mu[t] <- log(r0)  + theta * log(y[t]) - log(1 + pow(abs(y[t]), theta) / K)}
+\StringTok{    y[t+1] ~ dlnorm(mu[t], iQ) }
+\StringTok{  \}}
+\StringTok{\}"}\NormalTok{)}
+\KeywordTok{writeLines}\NormalTok{(bugs.model, }\StringTok{"myers_process.bugs"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{K_prior     <- function(x) }\KeywordTok{dunif}\NormalTok{(x, K_prior_p[}\DecValTok{1}\NormalTok{], K_prior_p[}\DecValTok{2}\NormalTok{])}
+\NormalTok{r_prior     <- function(x) }\KeywordTok{dunif}\NormalTok{(x, r0_prior_p[}\DecValTok{1}\NormalTok{], r0_prior_p[}\DecValTok{2}\NormalTok{])}
+\NormalTok{theta_prior <- function(x) }\KeywordTok{dunif}\NormalTok{(x, theta_prior_p[}\DecValTok{1}\NormalTok{], theta_prior_p[}\DecValTok{2}\NormalTok{])}
+\NormalTok{par_priors <- }\KeywordTok{list}\NormalTok{( }\DataTypeTok{deviance =} \NormalTok{function(x) }\DecValTok{0} \NormalTok{* x, }\DataTypeTok{K =} \NormalTok{K_prior,}
+                    \DataTypeTok{r0 =} \NormalTok{r_prior, }\DataTypeTok{theta =} \NormalTok{theta_prior, }
+                    \DataTypeTok{stdQ =} \NormalTok{stdQ_prior)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{jags.params=}\KeywordTok{c}\NormalTok{(}\StringTok{"r0"}\NormalTok{, }\StringTok{"theta"}\NormalTok{, }\StringTok{"K"}\NormalTok{, }\StringTok{"stdQ"}\NormalTok{)}
+\NormalTok{jags.inits <- function()\{}
+  \KeywordTok{list}\NormalTok{(}\StringTok{"r0"}\NormalTok{= }\DecValTok{1} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{1}\NormalTok{), }
+       \StringTok{"K"}\NormalTok{=    }\DecValTok{10} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{1}\NormalTok{),}
+       \StringTok{"theta"} \NormalTok{= }\DecValTok{1} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{1}\NormalTok{),  }
+       \StringTok{"stdQ"}\NormalTok{= }\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.2}\NormalTok{) * }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{1}\NormalTok{),}
+       \DataTypeTok{.RNG.name=}\StringTok{"base::Wichmann-Hill"}\NormalTok{, }\DataTypeTok{.RNG.seed=}\DecValTok{123}\NormalTok{)}
+\NormalTok{\}}
+\KeywordTok{set.seed}\NormalTok{(}\DecValTok{12345}\NormalTok{)}
+\NormalTok{myers_jags <- }\KeywordTok{do.call}\NormalTok{(jags.parallel, }
+                      \KeywordTok{list}\NormalTok{(}\DataTypeTok{data=}\NormalTok{jags.data, }\DataTypeTok{inits=}\NormalTok{jags.inits, }
+                                                     \NormalTok{jags.params, }\DataTypeTok{n.chains=}\NormalTok{n.chains, }
+                                                     \DataTypeTok{n.iter=}\NormalTok{n.iter, }\DataTypeTok{n.thin=}\NormalTok{n.thin,}
+                           \DataTypeTok{n.burnin=}\NormalTok{n.burnin, }
+                           \DataTypeTok{model.file=}\StringTok{"myers_process.bugs"}\NormalTok{))}
+\KeywordTok{recompile}\NormalTok{(myers_jags)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 406
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 406
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 406
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 406
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 406
+
+Initializing model
+
+Compiling model graph
+   Resolving undeclared variables
+   Allocating nodes
+   Graph Size: 406
+
+Initializing model
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{myers_jags <- }\KeywordTok{do.call}\NormalTok{(autojags, }
+                      \KeywordTok{list}\NormalTok{(myers_jags, }\DataTypeTok{n.update=}\NormalTok{n.update, }
+                           \DataTypeTok{n.iter=}\NormalTok{n.iter, }\DataTypeTok{n.thin =} \NormalTok{n.thin, }
+                           \DataTypeTok{progress.bar=}\StringTok{"none"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tmp <- }\KeywordTok{lapply}\NormalTok{(}\KeywordTok{as.mcmc}\NormalTok{(myers_jags), as.matrix) }\CommentTok{# strip classes}
+\NormalTok{myers_posteriors <- }\KeywordTok{melt}\NormalTok{(tmp, }\DataTypeTok{id =} \KeywordTok{colnames}\NormalTok{(tmp[[}\DecValTok{1}\NormalTok{]])) }
+\KeywordTok{names}\NormalTok{(myers_posteriors) = }\KeywordTok{c}\NormalTok{(}\StringTok{"index"}\NormalTok{, }\StringTok{"variable"}\NormalTok{, }\StringTok{"value"}\NormalTok{, }\StringTok{"chain"}\NormalTok{)}
+\NormalTok{plot_myers_traces <- }\KeywordTok{ggplot}\NormalTok{(myers_posteriors) + }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(index, value)) +}
+  \KeywordTok{facet_wrap}\NormalTok{(~ variable, }\DataTypeTok{scale=}\StringTok{"free"}\NormalTok{, }\DataTypeTok{ncol=}\DecValTok{1}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{par_prior_curves <- }\KeywordTok{ddply}\NormalTok{(myers_posteriors, }\StringTok{"variable"}\NormalTok{, function(dd)\{}
+    \NormalTok{grid <- }\KeywordTok{seq}\NormalTok{(}\KeywordTok{min}\NormalTok{(dd$value), }\KeywordTok{max}\NormalTok{(dd$value), }\DataTypeTok{length =} \DecValTok{100}\NormalTok{) }
+    \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{value =} \NormalTok{grid, }\DataTypeTok{density =} \NormalTok{par_priors[[dd$variable[}\DecValTok{1}\NormalTok{]]](grid))}
+\NormalTok{\})}
+\NormalTok{plot_myers_posteriors <- }\KeywordTok{ggplot}\NormalTok{(myers_posteriors, }\KeywordTok{aes}\NormalTok{(value)) + }
+  \KeywordTok{stat_density}\NormalTok{(}\DataTypeTok{geom=}\StringTok{"path"}\NormalTok{, }\DataTypeTok{position=}\StringTok{"identity"}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.7}\NormalTok{) +}
+  \KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{data=}\NormalTok{par_prior_curves, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{value, }\DataTypeTok{y=}\NormalTok{density), }\DataTypeTok{col=}\StringTok{"red"}\NormalTok{) + }
+  \KeywordTok{facet_wrap}\NormalTok{(~ variable, }\DataTypeTok{scale=}\StringTok{"free"}\NormalTok{, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{A <- myers_posteriors}
+\NormalTok{A$index <- A$index + A$chain * }\KeywordTok{max}\NormalTok{(A$index) }\CommentTok{# Combine samples across chains by renumbering index }
+\NormalTok{myers_pardist <- }\KeywordTok{acast}\NormalTok{(A, index ~ variable)}
+\NormalTok{bayes_coef <- }\KeywordTok{apply}\NormalTok{(myers_pardist,}\DecValTok{2}\NormalTok{, posterior.mode) }\CommentTok{# much better estimates}
+\NormalTok{myers_bayes_pars <- }\KeywordTok{unname}\NormalTok{(}\KeywordTok{c}\NormalTok{(bayes_coef[}\DecValTok{2}\NormalTok{], bayes_coef[}\DecValTok{3}\NormalTok{], bayes_coef[}\DecValTok{1}\NormalTok{]))}
+\NormalTok{myers_means <- }\KeywordTok{sapply}\NormalTok{(x_grid, Myer_harvest, }\DecValTok{0}\NormalTok{, myers_bayes_pars)}
+\NormalTok{myers_f <- function(x,h,p) }\KeywordTok{Myer_harvest}\NormalTok{(x, h, p[}\KeywordTok{c}\NormalTok{(}\StringTok{"r0"}\NormalTok{, }\StringTok{"theta"}\NormalTok{, }\StringTok{"K"}\NormalTok{)])}
+\KeywordTok{head}\NormalTok{(myers_pardist)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+        K deviance     r0    stdQ  theta
+170 20.51    37.16 1.3995 0.04136 0.9864
+171 17.10    41.09 0.6076 0.05170 1.8684
+172 24.40    36.33 1.1692 0.04307 1.0715
+173 27.56    37.46 0.9533 0.05550 1.2000
+174 22.21    34.94 0.5140 0.04203 1.8694
+175 36.05    33.49 0.3089 0.03977 2.1519
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{myers_bayes_pars}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+[1] 35.4576  0.5129 23.9914
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{models <- }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{x=}\NormalTok{x_grid, }
+                                         \DataTypeTok{GP=}\NormalTok{tgp_dat$y, }
+                                         \DataTypeTok{True=}\NormalTok{true_means, }
+                     \DataTypeTok{MLE=}\NormalTok{est_means, }
+                                         \DataTypeTok{Ricker=}\NormalTok{ricker_means, }
+                     \DataTypeTok{Allen =} \NormalTok{allen_means,}
+                     \DataTypeTok{Myers =} \NormalTok{myers_means)}
+\NormalTok{models <- }\KeywordTok{melt}\NormalTok{(models, }\DataTypeTok{id=}\StringTok{"x"}\NormalTok{)}
+
+\CommentTok{# some labels}
+\KeywordTok{names}\NormalTok{(models) <- }\KeywordTok{c}\NormalTok{(}\StringTok{"x"}\NormalTok{, }\StringTok{"method"}\NormalTok{, }\StringTok{"value"}\NormalTok{)}
+
+\CommentTok{# labels for the colorkey too}
+\NormalTok{model_names = }\KeywordTok{c}\NormalTok{(}\StringTok{"GP"}\NormalTok{, }\StringTok{"True"}\NormalTok{, }\StringTok{"MLE"}\NormalTok{, }\StringTok{"Ricker"}\NormalTok{, }\StringTok{"Allen"}\NormalTok{, }\StringTok{"Myers"}\NormalTok{)}
+\NormalTok{colorkey=cbPalette}
+\KeywordTok{names}\NormalTok{(colorkey) = model_names }
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# uses expected values from GP, instead of integrating over posterior}
+\CommentTok{#matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)}
+\NormalTok{matrices_gp <- }\KeywordTok{gp_transition_matrix}\NormalTok{(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid) }
+\NormalTok{opt_gp <- }\KeywordTok{value_iteration}\NormalTok{(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{matrices_true <- }\KeywordTok{f_transition_matrix}\NormalTok{(f, p, x_grid, h_grid, sigma_g)}
+\NormalTok{opt_true <- }\KeywordTok{value_iteration}\NormalTok{(matrices_true, x_grid, h_grid, }\DataTypeTok{OptTime=}\NormalTok{MaxT, xT, profit, }\DataTypeTok{delta=}\NormalTok{delta)}
+\NormalTok{matrices_estimated <- }\KeywordTok{f_transition_matrix}\NormalTok{(est$f, est$p, x_grid, h_grid, est$sigma_g)}
+\NormalTok{opt_estimated <- }\KeywordTok{value_iteration}\NormalTok{(matrices_estimated, x_grid, h_grid, }\DataTypeTok{OptTime=}\NormalTok{MaxT, xT, profit, }\DataTypeTok{delta=}\NormalTok{delta)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{matrices_allen <- }\KeywordTok{parameter_uncertainty_SDP}\NormalTok{(allen_f, x_grid, h_grid, pardist, }\DecValTok{4}\NormalTok{)}
+\NormalTok{opt_allen <- }\KeywordTok{value_iteration}\NormalTok{(matrices_allen, x_grid, h_grid, }\DataTypeTok{OptTime=}\NormalTok{MaxT, xT, profit, }\DataTypeTok{delta=}\NormalTok{delta)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{matrices_ricker <- }\KeywordTok{parameter_uncertainty_SDP}\NormalTok{(ricker_f, x_grid, h_grid, }\KeywordTok{as.matrix}\NormalTok{(ricker_pardist), }\DecValTok{3}\NormalTok{)}
+\NormalTok{opt_ricker <- }\KeywordTok{value_iteration}\NormalTok{(matrices_ricker, x_grid, h_grid, }\DataTypeTok{OptTime=}\NormalTok{MaxT, xT, profit, }\DataTypeTok{delta=}\NormalTok{delta)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{matrices_myers <- }\KeywordTok{parameter_uncertainty_SDP}\NormalTok{(myers_f, x_grid, h_grid, }\KeywordTok{as.matrix}\NormalTok{(myers_pardist), }\DecValTok{4}\NormalTok{)}
+\NormalTok{myers_alt <- }\KeywordTok{value_iteration}\NormalTok{(matrices_myers, x_grid, h_grid, }\DataTypeTok{OptTime=}\NormalTok{MaxT, xT, profit, }\DataTypeTok{delta=}\NormalTok{delta)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{OPT = }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{GP =} \NormalTok{opt_gp$D, }\DataTypeTok{True =} \NormalTok{opt_true$D, }\DataTypeTok{MLE =} \NormalTok{opt_estimated$D, }\DataTypeTok{Ricker =} \NormalTok{opt_ricker$D, }\DataTypeTok{Allen =} \NormalTok{opt_allen$D, }\DataTypeTok{Myers =} \NormalTok{myers_alt$D)}
+\NormalTok{colorkey=cbPalette}
+\KeywordTok{names}\NormalTok{(colorkey) = }\KeywordTok{names}\NormalTok{(OPT) }
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{sims <- }\KeywordTok{lapply}\NormalTok{(OPT, function(D)\{}
+  \KeywordTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
+  \KeywordTok{lapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\DecValTok{100}\NormalTok{, function(i) }
+    \KeywordTok{ForwardSimulate}\NormalTok{(f, p, x_grid, h_grid, x0, D, z_g, }\DataTypeTok{profit=}\NormalTok{profit, }\DataTypeTok{OptTime=}\NormalTok{OptTime)}
+  \NormalTok{)}
+\NormalTok{\})}
+
+\NormalTok{dat <- }\KeywordTok{melt}\NormalTok{(sims, }\DataTypeTok{id=}\KeywordTok{names}\NormalTok{(sims[[}\DecValTok{1}\NormalTok{]][[}\DecValTok{1}\NormalTok{]]))}
+\NormalTok{sims_data <- }\KeywordTok{data.table}\NormalTok{(dat)}
+\KeywordTok{setnames}\NormalTok{(sims_data, }\KeywordTok{c}\NormalTok{(}\StringTok{"L1"}\NormalTok{, }\StringTok{"L2"}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\StringTok{"method"}\NormalTok{, }\StringTok{"reps"}\NormalTok{)) }
+\CommentTok{# Legend in original ordering please, not alphabetical: }
+\NormalTok{sims_data$method = }\KeywordTok{factor}\NormalTok{(sims_data$method, }\DataTypeTok{ordered=}\OtherTok{TRUE}\NormalTok{, }\DataTypeTok{levels=}\KeywordTok{names}\NormalTok{(OPT))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{Profit <- sims_data[, }\KeywordTok{sum}\NormalTok{(profit), by=}\KeywordTok{c}\NormalTok{(}\StringTok{"reps"}\NormalTok{, }\StringTok{"method"}\NormalTok{)]}
+\NormalTok{tmp <- }\KeywordTok{dcast}\NormalTok{(Profit, reps ~ method)}
+\NormalTok{tmp <- tmp / tmp[,}\StringTok{"True"}\NormalTok{]}
+\NormalTok{tmp <- }\KeywordTok{melt}\NormalTok{(tmp[}\DecValTok{2}\NormalTok{:}\KeywordTok{dim}\NormalTok{(tmp)[}\DecValTok{2}\NormalTok{]])}
+\NormalTok{actual_over_optimal <-}\KeywordTok{subset}\NormalTok{(tmp, variable != }\StringTok{"True"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{allen_deviance <- }\KeywordTok{posterior.mode}\NormalTok{(pardist[,}\StringTok{'deviance'}\NormalTok{])}
+\NormalTok{ricker_deviance <- }\KeywordTok{posterior.mode}\NormalTok{(ricker_pardist[,}\StringTok{'deviance'}\NormalTok{])}
+\NormalTok{myers_deviance <- }\KeywordTok{posterior.mode}\NormalTok{(myers_pardist[,}\StringTok{'deviance'}\NormalTok{])}
+\NormalTok{true_deviance <- }\DecValTok{2}\NormalTok{*}\KeywordTok{estf}\NormalTok{(}\KeywordTok{c}\NormalTok{(p, sigma_g))}
+\NormalTok{mle_deviance <- }\DecValTok{2}\NormalTok{*}\KeywordTok{estf}\NormalTok{(}\KeywordTok{c}\NormalTok{(est$p, est$sigma_g))}
+
+\KeywordTok{c}\NormalTok{(}\DataTypeTok{allen =} \NormalTok{allen_deviance, }\DataTypeTok{ricker=}\NormalTok{ricker_deviance, }\DataTypeTok{myers=}\NormalTok{myers_deviance, }\DataTypeTok{true=}\NormalTok{true_deviance, }\DataTypeTok{mle=}\NormalTok{mle_deviance)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+  allen  ricker   myers    true     mle 
+  32.93   35.92   35.46  -61.08 -889.99 
+\end{verbatim}
+
 \section{Abstract}
@@
 Though simple mechanistically motivated models may be best for the
-latter case (Cuddington, Fortin, and Gerber 2013; Geritz and Kisdi
-2012), such models can be not only inaccurate but misleading in
-quantitative decision making. This arises for two reasons:
+latter case (@Cuddington2013; Geritz and Kisdi 2012), such models can be
+not only inaccurate but misleading in quantitative decision making. This
+arises for two reasons:
 
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{plot_gp <- }\KeywordTok{ggplot}\NormalTok{(tgp_dat) + }\KeywordTok{geom_ribbon}\NormalTok{(}\KeywordTok{aes}\NormalTok{(x,y,}\DataTypeTok{ymin=}\NormalTok{ymin,}\DataTypeTok{ymax=}\NormalTok{ymax), }\DataTypeTok{fill=}\StringTok{"gray80"}\NormalTok{) +}
+    \KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{data=}\NormalTok{models, }\KeywordTok{aes}\NormalTok{(x, value, }\DataTypeTok{col=}\NormalTok{method), }\DataTypeTok{lwd=}\DecValTok{1}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) + }
+    \KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{data=}\NormalTok{obs, }\KeywordTok{aes}\NormalTok{(x,y), }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) + }
+    \KeywordTok{xlab}\NormalTok{(}\KeywordTok{expression}\NormalTok{(X[t])) + }\KeywordTok{ylab}\NormalTok{(}\KeywordTok{expression}\NormalTok{(X[t}\DecValTok{+1}\NormalTok{])) +}
+    \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{cbPalette) }
+\KeywordTok{print}\NormalTok{(plot_gp)}
+\end{Highlighting}
+\end{Shaded}
+
 \begin{figure}[htbp]
 \centering
-\includegraphics{Figure1.pdf}
-\caption{Mean dynamics predicted by each model.}
+\includegraphics{figure/Figure1.pdf}
+\caption{Graph of the inferred Gaussian process compared to the true
+process and maximum-likelihood estimated process. Graph shows the
+expected value for the function $f$ under each model. Two standard
+deviations from the estimated Gaussian process covariance with (light
+grey) and without (darker grey) measurement error are also shown. The
+training data is also shown as black points. (The GP is conditioned on
+0,0, shown as a pseudo-data point).}
 \end{figure}
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{policies <- }\KeywordTok{melt}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{stock=}\NormalTok{x_grid, }\KeywordTok{sapply}\NormalTok{(OPT, function(x) x_grid[x])), }\DataTypeTok{id=}\StringTok{"stock"}\NormalTok{)}
+\KeywordTok{names}\NormalTok{(policies) <- }\KeywordTok{c}\NormalTok{(}\StringTok{"stock"}\NormalTok{, }\StringTok{"method"}\NormalTok{, }\StringTok{"value"}\NormalTok{)}
+
+\KeywordTok{ggplot}\NormalTok{(policies, }\KeywordTok{aes}\NormalTok{(stock, stock - value, }\DataTypeTok{color=}\NormalTok{method)) +}
+  \KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{lwd=}\FloatTok{1.2}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) + }\KeywordTok{xlab}\NormalTok{(}\StringTok{"stock size"}\NormalTok{) + }\KeywordTok{ylab}\NormalTok{(}\StringTok{"escapement"}\NormalTok{)  +}
+  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
+\end{Highlighting}
+\end{Shaded}
+
 \begin{figure}[htbp]
 \centering
-\includegraphics{Figure2.pdf}
-\caption{Steady-state optimal policies by each model.}
+\includegraphics{figure/Figure2.pdf}
+\caption{The steady-state optimal policy (infinite boundary) calculated
+under each model. Policies are shown in terms of target escapement,
+$S_t$, as under models such as this a constant escapement policy is
+expected to be optimal (Reed 1979).}
 \end{figure}
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{ggplot}\NormalTok{(sims_data) + }
+  \KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, fishstock, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(reps,method), }\DataTypeTok{color=}\NormalTok{method), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) +}
+  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{)))}
+\end{Highlighting}
+\end{Shaded}
+
 \begin{figure}[htbp]
 \centering
-\includegraphics{Figure3.pdf}
-\caption{Managing the simulated system over 100 replicate instances
-under each policy}
+\includegraphics{figure/Figure3.pdf}
+\caption{Gaussian process inference outperforms parametric estimates.
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under
+the policies derived from each of the estimated models, as well as the
+policy based on the exact underlying model.}
 \end{figure}
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) + }\KeywordTok{geom_histogram}\NormalTok{() + }
+  \KeywordTok{facet_wrap}\NormalTok{(~variable, }\DataTypeTok{scales =} \StringTok{"free_y"}\NormalTok{) + }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{) + }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
 \begin{figure}[htbp]
 \centering
-\includegraphics{figure/figure4.png}
+\includegraphics{figure/Figure41.pdf}
+\caption{Histograms of the realized net present value of the fishery
+over a range of simulated data and resulting parameter estimates. For
+each data set, the three models are estimated as described above. Values
+plotted are the averages of a given policy over 100 replicate
+simulations. Details and code provided in the supplement.}
+\end{figure}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable), binwidth=0.1) + }
+\CommentTok{#  xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)}
+\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value, }\DataTypeTok{fill=}\NormalTok{variable, }\DataTypeTok{color=}\NormalTok{variable)) + }
+  \KeywordTok{stat_density}\NormalTok{(}\DataTypeTok{position=}\StringTok{"stack"}\NormalTok{, }\DataTypeTok{adjust=}\DecValTok{10}\NormalTok{, }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{8}\NormalTok{) + }
+  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)+ }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/Figure42.pdf}
 \caption{Histograms of the realized net present value of the fishery
@@
 
+\section{Appendix}
+
+The appendices have not yet been assembled. Meanwhile, code to repeat
+the analyses, along with a complete log of all research conducted on
+this project, can be found at:
+\href{https://github.com/cboettig/nonparametric-bayes/}{https://github.com/cboettig/nonparametric-bayes}
+
+\subsection{Markov Chain Monte Carlo Analysis}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{gp_assessment_plots[[}\DecValTok{1}\NormalTok{]]}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/appendixplots1.pdf}
+\caption{plot of chunk appendixplots}
+\end{figure}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{gp_assessment_plots[[}\DecValTok{2}\NormalTok{]]}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/appendixplots2.pdf}
+\caption{plot of chunk appendixplots}
+\end{figure}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{plot_allen_traces}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/appendixplots3.pdf}
+\caption{plot of chunk appendixplots}
+\end{figure}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{plot_allen_posteriors}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/appendixplots4.pdf}
+\caption{plot of chunk appendixplots}
+\end{figure}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{plot_ricker_traces}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/appendixplots5.pdf}
+\caption{plot of chunk appendixplots}
+\end{figure}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{plot_ricker_posteriors}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/appendixplots6.pdf}
+\caption{plot of chunk appendixplots}
+\end{figure}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{plot_myers_traces}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/appendixplots7.pdf}
+\caption{plot of chunk appendixplots}
+\end{figure}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{plot_myers_posteriors}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/appendixplots8.pdf}
+\caption{plot of chunk appendixplots}
+\end{figure}
+
 \section{References}
@@
 
-Cuddington, Kim M., M. J. Fortin, and L. R. Gerber. 2013.
-``Process-based models are required to manage ecological systems in a
-changing world.'' \emph{Ecosphere} 4 (February): 1--12.
-doi:10.1890/ES12-00178.1.
-\url{http://www.esajournals.org/doi/abs/10.1890/ES12-00178.1}.
-
 Deisenroth, Marc Peter, Carl Edward Rasmussen, and Jan Peters. 2009.
@@
 
-\pagebreak\pagebreak
-
-\section{Appendix}
-
-The appendices have not yet been assembled. Meanwhile, code to repeat
-the analyses, along with a complete log of all research conducted on
-this project, can be found at:
-\href{https://github.com/cboettig/nonparametric-bayes/}{https://github.com/cboettig/nonparametric-bayes}
-
-\subsection{Markov Chain Monte Carlo Analysis}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{gp_traces_densities1.pdf}
-\caption{Traces of the GP hyper-parameters (after transforming back from
-log space)}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{gp_traces_densities2.pdf}
-\caption{Posterior densities of the GP hyper-parameters (after
-transforming back from log space). Prior distribution over the interval
-sampled shown as red curves.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{ricker_traces.pdf}
-\caption{Traces of the Ricker parameters}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{ricker_posteriors.pdf}
-\caption{Posterior densities of the Ricker parameter estimates. Prior
-distributions over the interval sampled shown as red curves. Note the
-bimodal estimate of $\log(r)$ which leads to the poor fit in Figure 1}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{parametric_bayes_traces.pdf}
-\caption{Traces of the structurally accurate Allen model parameters}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{parametric_bayes_posteriors.pdf}
-\caption{Posterior densities for the parameters of the Allen model.
-Prior distributions over the interval sampled shown as red curves.}
-\end{figure}
-
 \end{document}

### 2013-06-19 inst/examples/semi-analytic-posteriors.Rmd
use headers for each model
code examples updated
update manuscript examples

@@
 
-Estimating the Allen model:
+# Estimating the Allen model:
 
@@
 ```{r}
-Mt <- function(t, a, b)  log(x[t+1]) - log(x[t]) - b * x[t] + a * x[t] ^ 2
-```
+Mt <- function(t, a, b)  log(x[t+1]) - log(x[t]) - b * x[t] + a * x[t] ^ 2```
 
@@
 
-Estimating the Myers model on this data:
+# Estimating the Myers model on this data:
 
@@
 ```
+
+## Reconstructing the joint probability distribution
+
+$$P(r, \beta, \sigma | X) = \frac{P(X | r, \beta, \sigma) P(r) P(\beta) P(\sigma) }{\int_r \int_{\beta} \int_{\sigma P(X | r, \beta, \sigma) P(r) P(\beta) P(\sigma)}  $$
+
+Having done the first two integrals analytically and then evaluating the remaining expression over the grid, we simply sum over the grids to complete our calculation of the normalization term from the denominator,
+
+```{r}
+N_ricker <- sum(P_B.X) * (beta_grid[2]-beta_grid[1])
+```
+
+We then have an analytic function for the numerator,
+
+$$ P(X | r, \beta, \sigma) P(r) P(\beta) P(\sigma) =  \frac{1}{\sqrt{2 \pi \sigma^2}^{T-1}} \exp\left(\frac{\sum_t^{T-1} \left(\log X_{t+1} - \log X_t - \log r + \beta X_t\right)^2 }{2 \sigma^2}\right) \frac{1}{\sigma^2}$$
+
+## Transforming back

### 2013-06-19 inst/examples/BUGS/external-chunks.R

@@ -52,7 +52,7 @@ MaxT = 1000 # timeout for value iteration convergence
                     y = c(rep(0,nz), 
                           x[2:Tobs]))
 raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
-
+raw_plot
 
 
 ## @knitr mle
@@ -533,19 +533,26 @@ ggplot(sims_data) +
 ## @knitr profits
 Profit <- sims_data[, sum(profit), by=c("reps", "method")]
 tmp <- dcast(Profit, reps ~ method)
+#tmp$Allen <- tmp[,"Allen"] + rnorm(dim(tmp)[1], 0, 1) # jitter for plotting
 tmp <- tmp / tmp[,"True"]
 tmp <- melt(tmp[2:dim(tmp)[2]])
 actual_over_optimal <-subset(tmp, variable != "True")
 
 
 ## @knitr Figure4 
-ggplot(actual_over_optimal, aes(value)) + geom_histogram() + 
-  facet_wrap(~variable, scales = "free_y") + guides(legend.position = "none") + xlab("Total profit by replicate")
-# ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable), binwidth=0.1) + 
-#  xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)
+ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable)) + 
+  facet_wrap(~variable, scales = "free_y")  + guides(legend.position = "none") +
+  xlab("Total profit by replicate") + scale_fill_manual(values=colorkey)
+
+ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable), binwidth=0.1) + 
+  xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)
+
 ggplot(actual_over_optimal, aes(value, fill=variable, color=variable)) + 
-  stat_density(position="stack", adjust=10, alpha=.8) + 
+  stat_density(aes(y=..density..), position="stack", adjust=2, alpha=.9) + 
   xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)+ scale_color_manual(values=colorkey)
+
+
+
 
 
 ## @knitr deviances
@@ -559,7 +566,6 @@ c(allen = allen_deviance, ricker=ricker_deviance, myers=myers_deviance, true=tru
 
 
[…]

### 2013-06-20 inst/doc/nonparametric-bayes.Rmd
writing writing

@@
 
-```{r plot-options, echo=FALSE, cache=FALSE, include=FALSE}
+```{r plot-options, cache=FALSE}
 options(xtable.print.comment=FALSE)
@@
 toggle = "markup" # "hide" 
+theme_set(theme_bw(base_size=12))
 ```
@@
 actions. Unfortunately, such simple models rarely capture the complexity
-or uncertainty found in most real ecosystems. Non-parametric Bayesian
+or uncertainty found in most real ecosystems.
+
+We demonstrate how nonparametric Bayesian models can provide robust,
+nearly-optimal solutions to decision making under uncertainty when
+_we don't know the correct model_ to use.
+
+While methods that account
+for _parametric_ uncertainty can be very successful with the right model,
+structural uncertainty of not knowing what model best approximates the 
+dynamics poses considerably greater difficulty.
+
+Non-parametric Bayesian
 methods offer a promising statistical approach for predictive modeling
@@
 
-Most management recommendations from the ecological literature are based
-on (or at least motivated by) parametric models. Though in principle
-these models reflect mechanistic underpinnings of biological interactions
-involved, in practice real mechanisms are often unknown and the true
-dynamics too complex to be captured by simple models. While
+Decision making under uncertainty is a ubiquitous challenge of natural 
+resource management and conservation.
+
+<!-- Models for decision-making -->
+Decision-theoretic tools require a model that can assign probabilites of
+future states (e.g. stock size of a fishery) given the current state and a
+proposed action (e.g. fishing harvest or effort).  The decision maker then
+seeks to determing the course of actions (also refered to as the policy)
+that maximizes the expected value of some objective function (such as
+net present value derived from the resource over time.  (The approach
+can be adapted to permit alternatives to maximizing expectation, such
+as minimizing the maximum cost or damage that might be incurred. See
+@Polasky2011).
+Management frequently faces a sequential decision-making problem -- after 
+selecting an action, the decision-maker may recieve new information about 
+the current state and must again choose an appropriate action -- i.e. setting
+the harvest limits each year based on stock assessments the year prior.
+
+<!-- Parametric models -->
+
+<!-- Parametric uncertainty -->
+
+<!-- Comparing models -->
+The nature of decision-making problems provides a convenient way to compare 
+models.  Rather than compare models in terms of best fit to data or fret over
+the appropriate penalty for model complexity, model performance is defined 
+in the concrete terms of the decision-maker's objective function, which we
+will take as given. (Much argument can be made over the 'correct' objective
+function, e.g. how to account for the social value of fish left in the sea
+vs. the commercial value of fish harvested; see @Halpern2013 for further 
+discussion of this issue.  Alternatively, we can always compare model performance
+across multiple potential objective functions.)  The decision-maker
+does not necessarily need a model that provides the best mechanistic understanding
+or the best long-term outcome, but rather the one that best estimates the 
+probabilites of being in different states as a result of the possible actions.
+
+<!-- Structural Uncertainty -->
+
+ While
 simple mechanistic models can nevertheless provide important insights
-into possible dynamics -- for instance, demonstrating that a vaccine
-does not have to be 100% effective to eliminate the transmission of a
-virus  -- such approaches are not well-suited for use in
+into long-term outcomes, such approaches are not well-suited for use in
 forecasting outcomes of potential management options.  Non-parametric 
@@
 reflect the data available while also representing greater uncertainty
-in areas (of state-space) where data is lacking.  Ecological research
+in areas (of state-space) where data is lacking.
+
+We demonstrate how
+a Gaussian Process model of stock recruitment can lead to nearly optimal
+management through stochastic dynamic programming, comperable to knowing
+the correct structural equation for the underlying simulation.  Meanwhile,
+parametric models that do not match the underlying dynamics can perform 
+very poorly, even though they fit the data as well as the true model.
+
+Ecological research
 and management strategy should pay closer attention to the opportunities 
@@
 arrives makes such an approach less feasible still.
-
-<!-- cut?? -->
-
-An active literature and growing computational power over the past several
-decades have only marginally improved this situation.  Parametric or
-structural uncertainty can be introduced only by increasing the state
-space to reflect a distribution of possible parameters or
-the degree of belief in each of a set of possible models Frequently,
-the space of possible actions must then be reduced or the algorithms
-adjusted by approximations to keep computations feasible.
-
-<!-- cut?? -->
 
@@
 (The GP is conditioned on 0,0, shown as a pseudo-data point).
-"}
+", include=FALSE}
 ```
-```{r Figure1b, dependson=c("assemble-models", "par-fns", "plot-options")}
+```{r Figureb, dependson=c("assemble-models", "par-fns", "plot-options"), toggle = 1}
 ```
@@
 ```{r Figure2, dependson = c("assemble-opt", "plot-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
+Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", toggle = 1}
 ```
@@
 ```{r Figure3, dependson=c("sim", "plot-options"), fig.cap="Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."}
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", toggle = 1}
 ```
@@
 Values plotted are the averages of a given policy over 100 replicate simulations.
-Details and code provided in the supplement."}
+Details and code provided in the supplement.", toggle = 1}
 ```

### 2013-06-20 inst/doc/nonparametric-bayes.tex
writing writing

@@
 \begin{Highlighting}[]
+\KeywordTok{options}\NormalTok{(}\DataTypeTok{xtable.print.comment =} \OtherTok{FALSE}\NormalTok{)}
+\KeywordTok{options}\NormalTok{(}\DataTypeTok{xtable.type =} \StringTok{"latex"}\NormalTok{, }\DataTypeTok{table.placement =} \StringTok{"H"}\NormalTok{)}
+\NormalTok{opts_knit$}\KeywordTok{set}\NormalTok{(}\DataTypeTok{progress =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{)}
+\NormalTok{opts_chunk$}\KeywordTok{set}\NormalTok{(}\DataTypeTok{dev =} \KeywordTok{c}\NormalTok{(}\StringTok{"pdf"}\NormalTok{, }\StringTok{"png"}\NormalTok{), }\DataTypeTok{fig.width =} \FloatTok{5.5}\NormalTok{, }\DataTypeTok{fig.height =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cache.path =} \StringTok{"cache/nonparametric-bayes/"}\NormalTok{, }
+    \DataTypeTok{cache =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{external =} \OtherTok{TRUE}\NormalTok{)}
+\NormalTok{opts_chunk$}\KeywordTok{set}\NormalTok{(}\DataTypeTok{warning =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{message =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{comment =} \OtherTok{NA}\NormalTok{, }\DataTypeTok{tidy =} \OtherTok{FALSE}\NormalTok{)}
+\NormalTok{toggle = }\StringTok{"markup"}
+\KeywordTok{theme_set}\NormalTok{(}\KeywordTok{theme_bw}\NormalTok{(}\DataTypeTok{base_size =} \DecValTok{12}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
 \KeywordTok{require}\NormalTok{(modeest)}
@@
 \NormalTok{tmp <- }\KeywordTok{dcast}\NormalTok{(Profit, reps ~ method)}
+\CommentTok{#tmp$Allen <- tmp[,"Allen"] + rnorm(dim(tmp)[1], 0, 1) # jitter for plotting}
 \NormalTok{tmp <- tmp / tmp[,}\StringTok{"True"}\NormalTok{]}
@@
 \begin{Highlighting}[]
-\NormalTok{plot_gp <- }\KeywordTok{ggplot}\NormalTok{(tgp_dat) + }\KeywordTok{geom_ribbon}\NormalTok{(}\KeywordTok{aes}\NormalTok{(x,y,}\DataTypeTok{ymin=}\NormalTok{ymin,}\DataTypeTok{ymax=}\NormalTok{ymax), }\DataTypeTok{fill=}\StringTok{"gray80"}\NormalTok{) +}
-    \KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{data=}\NormalTok{models, }\KeywordTok{aes}\NormalTok{(x, value, }\DataTypeTok{col=}\NormalTok{method), }\DataTypeTok{lwd=}\DecValTok{1}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) + }
-    \KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{data=}\NormalTok{obs, }\KeywordTok{aes}\NormalTok{(x,y), }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) + }
-    \KeywordTok{xlab}\NormalTok{(}\KeywordTok{expression}\NormalTok{(X[t])) + }\KeywordTok{ylab}\NormalTok{(}\KeywordTok{expression}\NormalTok{(X[t}\DecValTok{+1}\NormalTok{])) +}
-    \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{cbPalette) }
-\KeywordTok{print}\NormalTok{(plot_gp)}
+\KeywordTok{require}\NormalTok{(MASS)}
+\NormalTok{step_ahead <- function(x, f, p)\{}
+  \NormalTok{h = }\DecValTok{0}
+  \NormalTok{x_predict <- }\KeywordTok{sapply}\NormalTok{(x, f, h, p)}
+  \NormalTok{n <- }\KeywordTok{length}\NormalTok{(x_predict) - }\DecValTok{1}
+  \NormalTok{y <- }\KeywordTok{c}\NormalTok{(x[}\DecValTok{1}\NormalTok{], x_predict[}\DecValTok{1}\NormalTok{:n])}
+  \NormalTok{y}
+\NormalTok{\}}
+\NormalTok{step_ahead_posteriors <- function(x)\{}
+\NormalTok{gp_f_at_obs <- }\KeywordTok{gp_predict}\NormalTok{(gp, x, }\DataTypeTok{burnin=}\FloatTok{1e4}\NormalTok{, }\DataTypeTok{thin=}\DecValTok{300}\NormalTok{)}
+\NormalTok{df_post <- }\KeywordTok{melt}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(}\KeywordTok{sample}\NormalTok{(}\DecValTok{100}\NormalTok{), }
+  \NormalTok{function(i)\{}
+    \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{time =} \DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(x), }\DataTypeTok{stock =} \NormalTok{x, }
+                \DataTypeTok{GP =} \KeywordTok{mvrnorm}\NormalTok{(}\DecValTok{1}\NormalTok{, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),}
+                \DataTypeTok{True =} \KeywordTok{step_ahead}\NormalTok{(x,f,p),  }
+                \DataTypeTok{MLE =} \KeywordTok{step_ahead}\NormalTok{(x,f,est$p), }
+                \DataTypeTok{Allen =} \KeywordTok{step_ahead}\NormalTok{(x, allen_f, pardist[i,]), }
+                \DataTypeTok{Ricker =} \KeywordTok{step_ahead}\NormalTok{(x, ricker_f, ricker_pardist[i,]), }
+                \DataTypeTok{Myers =} \KeywordTok{step_ahead}\NormalTok{(x, myers_f, myers_pardist[i,]))}
+  \NormalTok{\}), }\DataTypeTok{id=}\KeywordTok{c}\NormalTok{(}\StringTok{"time"}\NormalTok{, }\StringTok{"stock"}\NormalTok{))}
+\NormalTok{\}}
+
+\NormalTok{df_post <- }\KeywordTok{step_ahead_posteriors}\NormalTok{(x)}
+
+\KeywordTok{ggplot}\NormalTok{(df_post) + }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, stock)) + }
+  \KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, value, }\DataTypeTok{col=}\NormalTok{variable, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(L1,variable)), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) + }
+  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{))) }
 \end{Highlighting}
@@
 \centering
-\includegraphics{figure/Figure1.pdf}
-\caption{Graph of the inferred Gaussian process compared to the true
-process and maximum-likelihood estimated process. Graph shows the
-expected value for the function $f$ under each model. Two standard
-deviations from the estimated Gaussian process covariance with (light
-grey) and without (darker grey) measurement error are also shown. The
-training data is also shown as black points. (The GP is conditioned on
-0,0, shown as a pseudo-data point).}
+\includegraphics{figure/Figureb.pdf}
+\caption{plot of chunk Figureb}
 \end{figure}
@@
 \begin{Highlighting}[]
-\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) + }\KeywordTok{geom_histogram}\NormalTok{() + }
-  \KeywordTok{facet_wrap}\NormalTok{(~variable, }\DataTypeTok{scales =} \StringTok{"free_y"}\NormalTok{) + }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{) + }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)}
+\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) + }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{variable)) + }
+  \KeywordTok{facet_wrap}\NormalTok{(~variable, }\DataTypeTok{scales =} \StringTok{"free_y"}\NormalTok{)  + }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{) +}
+  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{) + }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{# ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable), binwidth=0.1) + }
-\CommentTok{#  xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)}
-\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value, }\DataTypeTok{fill=}\NormalTok{variable, }\DataTypeTok{color=}\NormalTok{variable)) + }
-  \KeywordTok{stat_density}\NormalTok{(}\DataTypeTok{position=}\StringTok{"stack"}\NormalTok{, }\DataTypeTok{adjust=}\DecValTok{10}\NormalTok{, }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{8}\NormalTok{) + }
-  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)+ }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
+
+\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) + }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{variable), }\DataTypeTok{binwidth=}\FloatTok{0.1}\NormalTok{) + }
+  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
 \end{Highlighting}
@@
 \includegraphics{figure/Figure42.pdf}
+\caption{Histograms of the realized net present value of the fishery
+over a range of simulated data and resulting parameter estimates. For
+each data set, the three models are estimated as described above. Values
+plotted are the averages of a given policy over 100 replicate
+simulations. Details and code provided in the supplement.}
+\end{figure}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+
+\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value, }\DataTypeTok{fill=}\NormalTok{variable, }\DataTypeTok{color=}\NormalTok{variable)) + }
+  \KeywordTok{stat_density}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{y=}\NormalTok{..density..), }\DataTypeTok{position=}\StringTok{"stack"}\NormalTok{, }\DataTypeTok{adjust=}\DecValTok{3}\NormalTok{, }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{9}\NormalTok{) + }
+  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)+ }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/Figure43.pdf}
 \caption{Histograms of the realized net present value of the fishery

### 2013-06-20 inst/examples/BUGS/external-chunks.R

@@ -548,7 +548,7 @@ ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable), bin
   xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)
 
 ggplot(actual_over_optimal, aes(value, fill=variable, color=variable)) + 
-  stat_density(aes(y=..density..), position="stack", adjust=2, alpha=.9) + 
+  stat_density(aes(y=..density..), position="stack", adjust=3, alpha=.9) + 
   xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)+ scale_color_manual(values=colorkey)
 
 

### 2013-06-24 inst/doc/nonparametric-bayes.Rmd
writing

@@
 
+Here we illustrate how a stochastic dynamic programming algorithm can 
+be driven by the predictions from a Gaussian process model, sidestepping
+the need for an accurate model-based description of the sustem dynamics.
+
 <!-- Models for decision-making -->
-Decision-theoretic tools require a model that can assign probabilites of
+Decision-theoretic (or optimal control) tools require a model that can assign probabilities of
 future states (e.g. stock size of a fishery) given the current state and a
 proposed action (e.g. fishing harvest or effort).  The decision maker then
-seeks to determing the course of actions (also refered to as the policy)
+seeks to determining the course of actions (also referred to as the policy)
 that maximizes the expected value of some objective function (such as
@@
 Management frequently faces a sequential decision-making problem -- after 
-selecting an action, the decision-maker may recieve new information about 
+selecting an action, the decision-maker may receive new information about 
 the current state and must again choose an appropriate action -- i.e. setting
@@
 
+Traditional approaches to optimal control (Pontryagin's principle, stochastic
+dynamic programming) rely on knowledge of the state equation.  Such simple 
+parametric models, frequently justified by mechanistic underpinnings are used
+to provide probabilities of future states.  Fisheries literature, which enjoys
+some of the longest history in the use of the optimal control framework, provides
+a clear example. Models such as the Ricker or Beverton Holt curves are ubiquitous [too strong? references?]
+Parametric models can reflect variability introduced by a stochastic environment
+or demographic process my incorporating random shocks from a given distribution.
+
 <!-- Parametric uncertainty -->
+As the parameter values for these models must be estimated from limited data,  
+there will always be some uncertainty associated with these values.  This uncertainty
+further compounds the intrinsic variability introduced by demographic or environmental 
+noise.  The degree of uncertainty in the parameter values can be inferred from the data
+and reflected in the estimates of the transition probabilities.
+
+<!-- Structural Uncertainty --> 
+Unfortunately, estimates of parameter uncertainty are only as good as the parametric 
+models themselves.  Too often, we do not understand the system dynamics well enough
+to know if a model provides a good approximation over the relevant range of states
+and timescales (criteria that we loosely refer to as defining the "right" or "true" model.)
+So called structural or model uncertainty is a more difficult problem than parametric 
+uncertainty.  Typical solutions involve either model choice, or model averaging. Either
+approach remains limited by the degree to which any of the proposed models are sufficiently
+close to the right model.
 
@@
 or the best long-term outcome, but rather the one that best estimates the 
-probabilites of being in different states as a result of the possible actions.
+probabilities of being in different states as a result of the possible actions.
 
-<!-- Structural Uncertainty -->
+<!-- The weakness of parametric models -->
 
- While
-simple mechanistic models can nevertheless provide important insights
+While simple mechanistic models can nevertheless provide important insights
 into long-term outcomes, such approaches are not well-suited for use in
@@
 
-Ecological research
-and management strategy should pay closer attention to the opportunities 
-and challenges nonparametric modeling can offer.
+Ecological research and management strategy should pay closer attention
+to the opportunities and challenges nonparametric modeling can offer.
 
+<!--
 ### Quantitative vs Qualitative Decisions
@@
 
-Non-parametric models can more realistically represent uncertainties outside 
+-->
+
+Non-parametric models can better represent uncertainties outside 
 of observed data while also better capturing the dynamics in the region
@@
 
-### Uncertainties: the classical approach
-
-The ecological dynamics of most systems of management interest are
-typically both complex and unknown.  Despite this, quantitative
-policy optimization in these systems has almost exclusively been based
-on simple, mechanistically motivated parametric models.  There are
-several reasons for this.  First, limited available data almost always
-precludes any accurate estimate of more complicated models.  Second, the
-computational demands of determining an optimal strategy in a sequential
-decision-theory problem suffer from the "curse of dimensionality"
-that makes it difficult to consider more complicated models, states,
-or actions.  Unfortunately, these challenges also make it particularly
-difficult to reflect the true uncertainty either in the parameters of
-the given models (parametric uncertainty) or in the structure of the
-models themselves (structural uncertainty).  Further unknowns such as
-measurement uncertainty, parameter uncertainty, unobserved states,
-knowledge of boundary conditions, etc. further compound the issue.
-Though a hierarchical Bayesian approach provides a natural way to
-address these from a statistical standpoint, formulating reasonable
-parametric  descriptions of each form of uncertainty is a challenging
-task in itself, let alone the computational difficulties of solving such
-a system. @Cressie2009 provides a good account of the successes and
-challenges of the approach.  Applying these approaches in the management
-context of sequential decision making, in which forecasts must be obtained
-over a range of possible actions and updated regularly as new information
-arrives makes such an approach less feasible still.
-
-Traditional approaches to optimal control (Pontryagin's principle, stochastic
-dynamic programming) rely on knowledge of the state equation, usually described
-by a simple parametric model. Here we illustrate how a stochastic dynamic
-programming algorithm can alternatively be driven by the predictions from
-a Gaussian process.
-
 Approach and Methods
@@
 from fist principles by @Allen2005a] as our underlying "reality".
-
-\begin{align}
 X_{t+1} &= Z_t f(S_t) \\
@@
 present value of the resource, given an estimated stock size from the year
-before. Rich data and global concerns have made marine fisheries the crucible for much
+before. The economic value and ecological concern have made marine fisheries the crucible for much
 of the founding work [@Gordon1954; @Reed1979; @May1979; @Ludwig1982]
@@
 alternative model structures, and then performing some model choice or
-model averaging  [@Williams2001; @Athanassoglou2012]. For simplicity 
-of the comparison we consider only the case of a structurally correct
-and structurally incorrect model, and estimate Bayesian posterior distributions
-of the parameters.  We will also compare to a simpler maximum likelihood
-estimate of the parameters that ignores the issue of parameter uncertainty.
+model averaging  [@Williams2001; @Athanassoglou2012].
 
-We compare these parametric approaches to our alternative that uses a Gaussian Process (GP)
-in place of a given stock recruitment curve.  We briefly revisit
-a description of the Gaussian Process approach that has appeared in 
-recent literature before describing how we take the important step of 
-adapting the GP to the decision theory problem that can still be solved
-using SDP.
+### Parametric Models
+
+We consider three candidate parametric models for the stock-recruitment function, 
+which we refer to by the first authors of the publications in which they were
+first proposed.
+
+We generate the data with a four-parameter model that contains
+a tipping point (an Allee effect, see [@Allen, @Courchamp]) below which the stock
+decreases to zero,
+
+$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K}\right)\left(\frac{S_t- \theta}{K}\right)} \label{Allen} $$
+
+$$ S_t = X_t - h_t $$
+
+The parameter $C$ reflects the location of the tipping point, $K$ the carrying capacity of
+the stock, and $r$ the base recruitment rate.  $S_t$ represents the stock size after
+a harvest $h_t$ has been implemented.  $Z_t$ represents a log-normal random variable of
+log-mean zero and log-standard deviation parameter $\sigma$.
+
+We consider two alternative candidate models: the Ricker [@Ricker] stock-recruitment curve,
+
+$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \label{Ricker} $$
+
+and an alternative four-parameter model adapted from @Myers,
+
+$$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \label{Myers} $$
+
+which contains a tipping point for $\theta > 2$ and becomes a Beverton-Holt model 
+at $\theta = 1$.
+
+<!-- Note that while similarly-named parameters contain similar connotations, the 
+estimate and precise intepretation will be different in each model. -->
+
+### Bayesian Inference of Parametric models
+
+Given the sample data, we infer posterior distributions for each of the three
+models listed above using a Markov Chain Monte Carlo Gibbs Sampler (jags, see 
+appendix for implementation details and code) given uniform priors.  We run six
+chains for $10^6$ steps each and then assess convergence by Gelman-Rubin criterion
+and inspection of the traces, see appendix.
+
+<!-- By transforming variables we can integrate out the the $\sigma$ parameter
+and the constant term in the log-mean, see appendix.  The resulting one or
+two parameters can then be evaluated over a vector or grid of possible values,
+and the result transformed back to determine the posterior density of the parameters,
+This approach avoids any concerns of convergence in the MCMC analysis and can thus
+be used to verify that the performance of the parametric models can not be explained
+by convergence issues.  However, this approach does not generalize easily to richer models,
+such as the introduction of measurement noise or additional parameters, in which 
+MCMC based approaches are the obvious choice.  -->
 
@@
 
-The use of Gaussian process (GP) regression (or "kreging" in the geospatial
+The use of Gaussian process (GP) regression (or "kriging" in the geospatial
 literature) to formulate a predictive model is relatively new in the

### 2013-06-24 inst/examples/sensitivity.Rmd
updated sensitivity script

@@
-```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
-opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "myers-cache/", warning=FALSE)
-#read_chunk('gaussian-process-control.R')
+```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+source("~/.knitr_defaults.R")
+#opts_knit$set(upload.fun = socialR::flickr.url)
 library(knitcitations)
+library(nonparametricbayes) 
 ```
 
-```{r libraries, include=FALSE}
-require(pdgControl)
-require(nonparametricbayes)
-require(reshape2)
-require(ggplot2)
-require(data.table)
-require(tgp)
-library(kernlab)
-require(MCMCpack)
-require(plyr)
+### Model and parameters
+
+```{r stateeq-local}
+f <- Myers
+p <- c(1.5, 2.5, 8)
+K <- 5  # approx, a li'l' less
+allee <- 1.2 # approx, a li'l' less
 ```
 
-```{r graphing-options, include=FALSE}
-opts_knit$set(upload.fun = socialR::flickr.url)
-#opts_chunk$set(dev = 'Cairo_pdf', dev.args=list(""))
-opts_chunk$set(dev="png", dev.args=list(bg="transparent"))
-opts_chunk$set(comment=NA, tidy=FALSE)
-theme_set(theme_bw(base_size=16))
-theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
-             plot.background = element_rect(fill = "transparent",colour = NA))
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-```
+Various parameters defining noise dynamics, grid, and policy costs.
 
-Fixed priors on hyperparameters, fixed model type.
-
-```{r gp-priors}
-#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
-s2.p <- c(5,5)  
-tau2.p <- c(5,1)
-d.p = c(10, 1/0.1, 10, 1/0.1)
-nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
-s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
-tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
-d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
-nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
-beta0_prior <- function(x, tau) dnorm(x, 0, tau)
-beta = c(0)
-priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
-```
-
-```{r sdp-pars-fixed}
-profit = function(x,h) pmin(x, h)
+```{r sdp-pars, dependson="stateeq-local"}
+sigma_g <- 0.05
+sigma_m <- 0.0
+z_g <- function() rlnorm(1, 0, sigma_g)
+z_m <- function() 1
+x_grid <- seq(0, 1.5 * K, length=50)
+h_grid <- x_grid
+profit <- function(x,h) pmin(x, h)
 delta <- 0.01
-OptTime = 50  # stationarity with unstable models is tricky thing
-reward = 0
+OptTime <- 50  # stationarity with unstable models is tricky thing
+reward <- 0
 xT <- 0
-
-```
-
-```{r RickerAllee-exp}
-f <- RickerAllee
-p <- c(1.5, 10, 5) 
-K <- 10
-allee <- 5
-```
-
-```{r sdp-pars-explore}
-sigma_g <- 0.1
-sigma_m <- 0.0
-z_g = function() rlnorm(1, 0, sigma_g)
-z_m = function() 1+(2*runif(1, 0,  1)-1) * sigma_m
-x_grid <- seq(0, 1.5 * K, length=101)
-h_grid <- x_grid
-```
-
-With parameters `r p`.
-
-```{r }
+Xo <-  allee+.5# observations start from
+x0 <- K # simulation under policy starts from
+Tobs <- 40
+MaxT = 1000 # timeout for value iteration convergence
 seed <- round(runif(32) * 1e6)
@@
 seed
-yields <- lapply(seed, 
+```
+
+```{r parallel}
+require(snowfall)
+sfInit(parallel=TRUE, cpu=8)
+sfLibrary("nonparametric-bayes")
+sfExportAll()
+```
+
+32 data-sets from the same model
+
+```{r }
+yields <- sfSapply(seed, 
 function(seed_i){
-  Xo <- K # observations start from
-  x0 <- Xo # simulation under policy starts from
-  obs <- sim_obs(Xo, z_g, f, p, Tobs=35, nz= 10, 
-                 harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
 
-  alt <- par_est(obs,  init = c(r=1.5, K=mean(obs$x), s=1))
-  est <- par_est_allee(obs, f, p,  init = c(p[1]+1, p[2]-1, p[3]-2, 
-                                   s = sigma_g + abs(rnorm(1,0,.1))))
+  x <- numeric(Tobs)
+  x[1] <- Xo
+  nz <- 1
+  for(t in 1:(Tobs-1))
+    x[t+1] = z_g() * f(x[t], h=0, p=p)
+  obs <- data.frame(x = c(rep(0,nz), 
+                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
+                    y = c(rep(0,nz), 
+                          x[2:Tobs]))
 
-  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
-          m0r1=FALSE, corr="exp", trace=FALSE, 
-          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
-          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
-  gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
+  ## @knitr gp-priors
+  s2.p <- c(5,5)  
+  d.p = c(10, 1/0.1)
 
-#  posteriors_plot(gp, priors) # needs trace=TRUE!
+  ## @knitr gp
+  gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
+  gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)
 
-  OPT <- optimal_policy(gp, f, est$f, alt$f,
-                        p, est$p, alt$p,
-                        x_grid, h_grid, sigma_g, 
-                        est$sigma_g, alt$sigma_g, 
-                        delta, xT, profit, reward, OptTime)
-  plot_policies(x_grid, OPT$gp_D, OPT$est_D, OPT$true_D, OPT$alt_D)
-  dt <- simulate_opt(OPT, f, p, x_grid, h_grid, x0, z_g, profit)
-  sim_plots(dt, seed=seed_i)
-  profits_stats(dt)
+  matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid) 
+  opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
+
+  ## @knitr mle-opt
+  matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
+  opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+
+  OPT = data.frame(GP = opt_gp$D, True = opt_true$D)
+  policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
+  names(policies) <- c("stock", "method", "value")
+
+  sims <- lapply(OPT, function(D){
+    set.seed(1)
+    lapply(1:100, function(i) 
+      ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
+    )
+  })
+
+  dat <- melt(sims, id=names(sims[[1]][[1]]))
+  sims_data <- data.table(dat)
+  setnames(sims_data, c("L1", "L2"), c("method", "reps")) 
+  # Legend in original ordering please, not alphabetical: 
+  sims_data$method = factor(sims_data$method, ordered=TRUE, levels=names(OPT))
+
+  Profit <- sims_data[, sum(profit), by=c("reps", "method")]
+  tmp <- dcast(Profit, reps ~ method)
+  tmp <- tmp / tmp[,"True"]
+  tmp <- melt(tmp[2:dim(tmp)[2]])
+  actual_over_optimal$GP
+
 })
@@
 ````
-
-```{r echo=FALSE, results="asis"}
-bibliography("html")
-```

### 2013-06-25 inst/examples/sensitivity.Rmd
ready for runs on zero

@@
 sfInit(parallel=TRUE, cpu=8)
-sfLibrary("nonparametric-bayes")
+sfLibrary(nonparametricbayes)
 sfExportAll()
@@
 ```{r}
-yields <- melt(yields, id=c("method", "V1", "sd"))
-ggplot(yields) + geom_histogram(aes(V1)) + facet_wrap(~method, scale='free')
 yields
+#yields <- melt(yields, id=c("method", "V1", "sd"))
+#ggplot(yields) + geom_histogram(aes(V1)) + facet_wrap(~method, scale='free')
 ````

### 2013-06-25 inst/doc/nonparametric-bayes.Rmd
writing

@@
 
+#### Opening
+
+[@Fischer2009, @Polasky2011]
+
 Decision making under uncertainty is a ubiquitous challenge of natural 
 resource management and conservation.  
-
 Here we illustrate how a stochastic dynamic programming algorithm can 
@@
 
-<!-- Models for decision-making -->
+#### Models for decision-making under uncertainty
+
 Decision-theoretic (or optimal control) tools require a model that can assign probabilities of
@@
 
-<!-- Parametric models -->
+[@Reed1979, @Mangel1985]
+
+#### Parametric models
 
@@
 some of the longest history in the use of the optimal control framework, provides
-a clear example. Models such as the Ricker or Beverton Holt curves are ubiquitous [too strong? references?]
+a clear example [ ] . Models such as the Ricker or Beverton Holt curves are ubiquitous [too strong? references?]
 Parametric models can reflect variability introduced by a stochastic environment
@@
 
-<!-- Parametric uncertainty -->
+#### Parametric uncertainty
+
 As the parameter values for these models must be estimated from limited data,  
@@
 
-<!-- Structural Uncertainty --> 
+[@Mangel1985; @Schapaugh2013]
+
+#### Structural Uncertainty
+
 Unfortunately, estimates of parameter uncertainty are only as good as the parametric 
@@
 
-<!-- Comparing models -->
+[@Williams2001; @Athanassoglou2012].
+
+#### Comparing models
+
 The nature of decision-making problems provides a convenient way to compare 
@@
 
-<!-- The weakness of parametric models -->
+<!-- Transition and map: The weakness of parametric models -->
+<!-- This section is a bit weak and vague.  Revisit to make a real map in light 
+of outline structure below. -->
 
@@
 very poorly, even though they fit the data as well as the true model.  
-
 Ecological research and management strategy should pay closer attention
@@
 
-<!--
+#### Background on Fisheries Context
+
+#### Background on Tipping points
+
+<!-- note: This section largely replaced by above outline
+
 ### Quantitative vs Qualitative Decisions
@@
 different actions.
-
-Though simple mechanistically motivated models
-may be best for the latter case [@Cuddington2013; @Geritz2012], such models 
-can be not only inaccurate but misleading in quantitative decision making. This 
-arises for two reasons:
-
-1. We do not know what the correct parameter values are for the models.
-1. We do not know what the correct models are for ecological systems.
-
-The complexity of ecological interactions and a lack of data contribute
-greatly to both of the problems. This concern is particularly acute
-in case of ecological tipping points [@Barnosky2012; @Scheffer2001],
-which arise from feedbacks common in ecological systems and can lead to a
-sudden catastrophic transition to an undesirable state.  In general we do
-not know where such tipping points are unless we have already observed
-the transition, in which case it is frequently too late to respond.
-Though there may exist early warning signals for certain kinds of these
-transitions that are driven by slow changes [@Scheffer2009], we do not
-know when, where, or how to apply them to the decision making context
-more generally [@Boettiger2013].
 
@@
 
-### Terminology: "Non-parametric"
+Approach and Methods
+====================
 
-The term non-parametric to describe modeling approaches such as Gaussian
-processes is a common source of confusion, notwithstanding the fact that
-the model is still specified by parameters.  Some literature has introduced
-the term "semi-parametric" in tacit acknowledgment of this, which no doubt
-only contributes to the confusion.  The problem is further exacerbated by
-the several meanings assigned to the term in statistics: (a) that the method 
-does not assume some particular probability distribution (e.g. a non-parametric bootstrap)
-or (b) that the method does not assume a fixed structure to the model.  Our 
-use is that of (b), in contrast to the classical approaches that do.  The 
-difference is clearest by way of example.
-
-Approach and Methods
-=====================================================================
-
-## Underlying Model
-
-Concerns over the potential for tipping points in ecological dynamics
-[@Scheffer2001] highlight the dangers of uncertainty in ecological
-management and pose a substantial challenge to existing decision-theoretic
-approaches [@Brozovic2011].  Because intervention is often too late 
-after a tipping point has been crossed (but see @Hughes2013), management
-is most often concerned with avoiding potentially catastrophic tipping
-points before any data is available at or following a transition that
-would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].
-
-To illustrate the value of the non-parametric Bayesian approach to management,
-we focus on example of a system containing such a tipping point whose dynamics
-can still be described by a simple, one-dimensional parametric model.  
-We will focus on a simple parametric model for a single species [derived
-from fist principles by @Allen2005a] as our underlying "reality".
-X_{t+1} &= Z_t f(S_t) \\
-S_t &= X_t - h_t \\
-f(S_t) &= S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
-\end{align}
-
-Where $Z_t$ is multiplicative noise function with mean 1, representing
-stochastic growth. We will consider log-normal noise with shape parameter
-$\sigma_g$.  We start with an example in which the parameters are $r =2$,
-$K=8$, $C=5$ and $\sigma_g = 0.1$.
-<!-- ` r p[1]`, $K =$ ` r p[2]`, $C =$ ` r p[3]`, and  $\sigma_g =$ ` r sigma_g`. -->
-
-As a low-dimensional system completely described by three parameters, this
-scenario should if anything be favorable to a parametric-based approach.
-This model contains an Allee effect, or tipping point, below which the
-population is not self-sustaining and shrinks to zero [@Courchamp2008].
-
-Both parametric and nonparametric approaches will require training
-data on which to base their model of the process.  We generate the
-training data under the model described in Eq 1 for $T_{\textrm{obs}}=40$ time
-steps, under a known but not necessarily optimal sequence of harvest
-intensities, $h_t$.  For simplicity we imagine a fishery that started
-from zero harvest pressure and has been gradually increasing the harvest.
-
-<!-- things that we could hit upon in discussion:
-Using data simulated from a specified model rather than empirical data
-permits the comparison against the true underlying dynamics, setting 
-a bar for the optimal performance possible.
-
-(Motivation, alternatives, stationarity, examples without a stable node
-(limit-cycle models), examples based on observations near a stable node
-alone, and why that isn't impossible).
--->
-
-<span style="color:#007FFF;">***Should we include any emprical examples?***</span>
-
-### The optimal control problem in fisheries management
+## The optimal control problem in fisheries management
 
@@
 
+## Underlying Model
+
+Concerns over the potential for tipping points in ecological dynamics
+[@Scheffer2001] highlight the dangers of uncertainty in ecological
+management and pose a substantial challenge to existing decision-theoretic
+approaches [@Brozovic2011].  Because intervention is often too late 
+after a tipping point has been crossed (but see @Hughes2013), management
+is most often concerned with avoiding potentially catastrophic tipping
+points before any data is available at or following a transition that
+would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].
+
+To illustrate the value of the non-parametric Bayesian approach to management,
+we focus on example of a system containing such a tipping point whose dynamics
+can still be described by a simple, one-dimensional parametric model.  
+We will focus on a simple parametric model for a single species [derived
+from fist principles by @Allen2005a] as our underlying "reality".
+
+$$X_{t+1} = Z_t f(S_t)  $$
+$$S_t = X_t - h_t $$
+$$f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} $$
+
+Where $Z_t$ is multiplicative noise function with mean 1, representing
+stochastic growth. We will consider log-normal noise with shape parameter
+$\sigma_g$.  We start with an example in which the parameters are $r =2$,
+$K=8$, $C=5$ and $\sigma_g = 0.1$.
+
+<!-- ` r p[1]`, $K =$ ` r p[2]`, $C =$ ` r p[3]`, and  $\sigma_g =$ ` r sigma_g`. -->
+
+As a low-dimensional system completely described by three parameters, this
+scenario should if anything be favorable to a parametric-based approach.
+This model contains an Allee effect, or tipping point, below which the
+population is not self-sustaining and shrinks to zero [@Courchamp2008].
+
+#### Simulated training data
+
+We generate initial observational data under the model described in Eq 1 
+for $T_{\textrm{obs}}=40$ time steps, under a given arbitrary sequence of harvest
+intensities, $h_t$. 
+We consider the case in which most of the data comes from a limited region
+of state space (e.g. near a stable equilibrium), leaving us without
+observations of the population dynamics at very low levels which would
+be useful in discrimating between recruitment curves [@] or demonstrating
+the existence of a tipping point [@Scheffer2001].  
+Using data simulated from a specified model rather than empirical data
+permits the comparison against the true underlying dynamics, setting 
+a bar for the optimal performance possible.
+
 ### Parametric Models
@@
 We generate the data with a four-parameter model that contains
-a tipping point (an Allee effect, see [@Allen, @Courchamp]) below which the stock
+a tipping point, as discussed above (equation 1),
+(an Allee effect, see [@Allen, @Courchamp]) below which the stock
 decreases to zero,
 
-$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K}\right)\left(\frac{S_t- \theta}{K}\right)} \label{Allen} $$
+$$ X_{t+1} = Z_t S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(\frac{S_t - \theta}{K}\right)} $$
 
@@
 
-$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \label{Ricker} $$
+$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
 
@@
 
-$$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \label{Myers} $$
+$$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} $$
 
@@
 
+#### Terminology: "Non-parametric"
+
+The term non-parametric to describe modeling approaches such as Gaussian
+processes is a common source of confusion, notwithstanding the fact that
+the model is still specified by parameters.  Some literature has introduced
+the term "semi-parametric" in tacit acknowledgment of this, which no doubt
+only contributes to the confusion.  The problem is further exacerbated by
+the several meanings assigned to the term in statistics: (a) that the method 
+does not assume some particular probability distribution (e.g. a non-parametric bootstrap)
+or (b) that the method does not assume a fixed structure to the model.  Our 
+use is that of (b), in contrast to the classical approaches that do.  The 
+difference is clearest by way of example.
+
 The use of Gaussian process (GP) regression (or "kriging" in the geospatial
@@
 (parametric) models that might describe the data. The GP allows us
-to consider a set of possible curves simultaneously.
+to consider probabilities on a large set of possible curves simultaneously.
 
-### Background on Gaussian Process inference
+The posterior distribution for the hyper-parameters of the Gaussian 
+process model are estimated by Metropolis-Hastings algorithm, again with
+details and code provided in the Appendix.  @Rasmussen2006 provides
+an excellent general introduction to Gaussian Processes and @Munch2005 
+first discusses their application in the context of population dynamics
+models such as fisheries stock-recruitment relationships.
+
+### SDP via GP
 
@@
 This completes the algorithm adapting the GP to the sequential decision-making 
-problem through SDP, which we believe has not yet been demonstrated in the
-literature.  We provide an R package implementation of this, along with
-the Gaussian process inference, in the supplemental materials.
+problem through SDP, which has not been previously demonstrated.  
+We further provide an R package implementation as described in the supplemental materials.
 
@@
 
-The posterior distribution for the hyper-parameters of the Gaussian 
-process model are estimated by Metropolis-Hastings algorithm, again with
-details and code provided in the Appendix.  @Rasmussen2006 provides
-an excellent general introduction to Gaussian Processes and @Munch2005 
-first discusses their application in this context.
-
 Results
-====================================================
+=======
 
@@
 Discussion 
-==================================================================
+==========
+
+#### Big picture: Linking GP to SDP
 
@@
 be improved.
+
+<!-- On mechanistic models -->
+
+Though simple mechanistically motivated models
+may be best for the latter case [@Cuddington2013; @Geritz2012], such models 
+can be not only inaccurate but misleading in quantitative decision making. This 
+arises for two reasons:
+
+1. We do not know what the correct parameter values are for the models.
+1. We do not know what the correct models are for ecological systems.
+
+The complexity of ecological interactions and a lack of data contribute
+greatly to both of the problems. This concern is particularly acute
+in case of ecological tipping points [@Barnosky2012; @Scheffer2001],
+which arise from feedbacks common in ecological systems and can lead to a
+sudden catastrophic transition to an undesirable state.  In general we do
+not know where such tipping points are unless we have already observed
+the transition, in which case it is frequently too late to respond.
+Though there may exist early warning signals for certain kinds of these
+transitions that are driven by slow changes [@Scheffer2009], we do not
+know when, where, or how to apply them to the decision making context
+more generally [@Boettiger2013].
 
@@
 
-## Future directions
+Future directions
+-----------------
+
+### Higher dimensions
 
@@
 learning approach will prove even more valuable.
-
-<span style="color:#007FFF;"> Data complexity. Perhaps too far out of scope... </span>
-The nonparametric Bayesian approach is also better suited to complex and
-disparate data.  Incorporating various sources of information into
-mechanistic models can be an immensely difficult due to the increased
-complexity involved.
 
@@
 Acknowledgments
-================
+===============
 
-This work was partially supported by the Center for Stock Assessment Research, a partnership between the University of California Santa Cruz and the Fisheries Ecology Division, Southwest Fisheries Science Center, Santa Cruz, CA and by NSF grant EF-0924195 to MM.
+This work was partially supported by the Center for Stock Assessment
+Research, a partnership between the University of California Santa Cruz
+and the Fisheries Ecology Division, Southwest Fisheries Science Center,
+Santa Cruz, CA and by NSF grant EF-0924195 to MM.
 

### 2013-06-25 inst/doc/nonparametric-bayes.tex
writing

@@
 \NormalTok{bayes_coef <- }\KeywordTok{apply}\NormalTok{(myers_pardist,}\DecValTok{2}\NormalTok{, posterior.mode) }\CommentTok{# much better estimates}
-\NormalTok{myers_bayes_pars <- }\KeywordTok{unname}\NormalTok{(}\KeywordTok{c}\NormalTok{(bayes_coef[}\DecValTok{2}\NormalTok{], bayes_coef[}\DecValTok{3}\NormalTok{], bayes_coef[}\DecValTok{1}\NormalTok{]))}
+\NormalTok{myers_bayes_pars <- }\KeywordTok{unname}\NormalTok{(}\KeywordTok{c}\NormalTok{(bayes_coef[}\StringTok{"r0"}\NormalTok{], bayes_coef[}\StringTok{"theta"}\NormalTok{], bayes_coef[}\StringTok{"K"}\NormalTok{]))}
 \NormalTok{myers_means <- }\KeywordTok{sapply}\NormalTok{(x_grid, Myer_harvest, }\DecValTok{0}\NormalTok{, myers_bayes_pars)}
@@
 \begin{verbatim}
-[1] 35.4576  0.5129 23.9914
+[1]  0.5129  1.0894 23.9914
 \end{verbatim}
@@
 \begin{Highlighting}[]
-\NormalTok{allen_deviance <- }\KeywordTok{posterior.mode}\NormalTok{(pardist[,}\StringTok{'deviance'}\NormalTok{])}
-\NormalTok{ricker_deviance <- }\KeywordTok{posterior.mode}\NormalTok{(ricker_pardist[,}\StringTok{'deviance'}\NormalTok{])}
-\NormalTok{myers_deviance <- }\KeywordTok{posterior.mode}\NormalTok{(myers_pardist[,}\StringTok{'deviance'}\NormalTok{])}
+\NormalTok{allen_deviance <- -}\DecValTok{2}\NormalTok{*}\KeywordTok{posterior.mode}\NormalTok{(pardist[,}\StringTok{'deviance'}\NormalTok{])}
+\NormalTok{ricker_deviance <- -}\DecValTok{2}\NormalTok{*}\KeywordTok{posterior.mode}\NormalTok{(ricker_pardist[,}\StringTok{'deviance'}\NormalTok{])}
+\NormalTok{myers_deviance <- -}\DecValTok{2}\NormalTok{*}\KeywordTok{posterior.mode}\NormalTok{(myers_pardist[,}\StringTok{'deviance'}\NormalTok{])}
 \NormalTok{true_deviance <- }\DecValTok{2}\NormalTok{*}\KeywordTok{estf}\NormalTok{(}\KeywordTok{c}\NormalTok{(p, sigma_g))}
 \NormalTok{mle_deviance <- }\DecValTok{2}\NormalTok{*}\KeywordTok{estf}\NormalTok{(}\KeywordTok{c}\NormalTok{(est$p, est$sigma_g))}
-
-\KeywordTok{c}\NormalTok{(}\DataTypeTok{allen =} \NormalTok{allen_deviance, }\DataTypeTok{ricker=}\NormalTok{ricker_deviance, }\DataTypeTok{myers=}\NormalTok{myers_deviance, }\DataTypeTok{true=}\NormalTok{true_deviance, }\DataTypeTok{mle=}\NormalTok{mle_deviance)}
+\NormalTok{xtable::}\KeywordTok{xtable}\NormalTok{(}\KeywordTok{as.table}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DataTypeTok{Allen =} \NormalTok{allen_deviance, }\DataTypeTok{Ricker=}\NormalTok{ricker_deviance, }\DataTypeTok{Myers=}\NormalTok{myers_deviance, }\DataTypeTok{True=}\NormalTok{true_deviance, }\DataTypeTok{MLE=}\NormalTok{mle_deviance)))}
 \end{Highlighting}
@@
 \begin{verbatim}
-  allen  ricker   myers    true     mle 
-  32.93   35.92   35.46  -61.08 -889.99 
+% latex table generated in R 3.0.1 by xtable 1.7-1 package
+% Thu Jun 20 16:29:00 2013
+\begin{table}[ht]
+\centering
+\begin{tabular}{rr}
+  \hline
+ & x \\ 
+  \hline
+Allen & -65.85 \\ 
+  Ricker & -71.85 \\ 
+  Myers & -70.92 \\ 
+  True & -61.08 \\ 
+  MLE & -889.99 \\ 
+   \hline
+\end{tabular}
+\end{table}
 \end{verbatim}
@@
 actions. Unfortunately, such simple models rarely capture the complexity
-or uncertainty found in most real ecosystems. Non-parametric Bayesian
-methods offer a promising statistical approach for predictive modeling
-of ecological dynamics in regions of state space where the data is
-adequate, while at the same time offering more flexible patterns with
-greater uncertainty outside the observed data. This contrasts from
-simple parametric models which provide relatively constant level of
-uncertainty in regions with and without adequate data. The consequence
-of such misplaced confidence outside the data can lead to highly
-undesirable results that may be avoided with the more flexible
+or uncertainty found in most real ecosystems.
+
+We demonstrate how nonparametric Bayesian models can provide robust,
+nearly-optimal solutions to decision making under uncertainty when
+\emph{we don't know the correct model} to use.
+
+While methods that account for \emph{parametric} uncertainty can be very
+successful with the right model, structural uncertainty of not knowing
+what model best approximates the dynamics poses considerably greater
+difficulty.
+
+Non-parametric Bayesian methods offer a promising statistical approach
+for predictive modeling of ecological dynamics in regions of state space
+where the data is adequate, while at the same time offering more
+flexible patterns with greater uncertainty outside the observed data.
+This contrasts from simple parametric models which provide relatively
+constant level of uncertainty in regions with and without adequate data.
+The consequence of such misplaced confidence outside the data can lead
+to highly undesirable results that may be avoided with the more flexible
 non-parametric Bayesian approach.
@@
 
-Most management recommendations from the ecological literature are based
-on (or at least motivated by) parametric models. Though in principle
-these models reflect mechanistic underpinnings of biological
-interactions involved, in practice real mechanisms are often unknown and
-the true dynamics too complex to be captured by simple models. While
-simple mechanistic models can nevertheless provide important insights
-into possible dynamics -- for instance, demonstrating that a vaccine
-does not have to be 100\% effective to eliminate the transmission of a
-virus -- such approaches are not well-suited for use in forecasting
-outcomes of potential management options. Non-parametric approaches
-offer a more flexible alternative that can both more accurately reflect
-the data available while also representing greater uncertainty in areas
-(of state-space) where data is lacking. Ecological research and
-management strategy should pay closer attention to the opportunities and
-challenges nonparametric modeling can offer.
-
-\subsubsection{Quantitative vs Qualitative Decisions}
-
-In this paper, we consider those ecological management problems in which
-a mathematical (or computational) model is used to quantitatively inform
-decision-making by allowing a manager to compare to the expected
-consequences of potential management actions (or policies). We
-distinguish this from the solely qualitative use of a model, in which
-models are used to represent and compare hypotheses of different
-mechanisms that would lead to qualitatively different actions.
-
-Though simple mechanistically motivated models may be best for the
-latter case (@Cuddington2013; Geritz and Kisdi 2012), such models can be
-not only inaccurate but misleading in quantitative decision making. This
-arises for two reasons:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  We do not know what the correct parameter values are for the models.
-\item
-  We do not know what the correct models are for ecological systems.
-\end{enumerate}
-
-The complexity of ecological interactions and a lack of data contribute
-greatly to both of the problems. This concern is particularly acute in
-case of ecological tipping points (Barnosky et al. 2012; Scheffer et al.
-2001), which arise from feedbacks common in ecological systems and can
-lead to a sudden catastrophic transition to an undesirable state. In
-general we do not know where such tipping points are unless we have
-already observed the transition, in which case it is frequently too late
-to respond. Though there may exist early warning signals for certain
-kinds of these transitions that are driven by slow changes (Scheffer et
-al. 2009), we do not know when, where, or how to apply them to the
-decision making context more generally (Boettiger and Hastings 2013).
-
-In addition to facing the potentially grave consequences of such
-transitions, if they do exist in a system of interest, this generally
-means that we lack data in the region near and beyond a possible tipping
-point. As a consequence, our models are least accurate just where we
-need them to be most accurate.\\Unfortunately, parametric modeling
-approaches are not well-suited to this scenario, in which good data is
-avialble over only part of the relevant state-space. Necessary
-assumptions about model structure in any parametric approach can lead
-such models to perform very poorly, even when accounting for parameter
-uncertainty.
-
-Non-parametric models can more realistically represent uncertainties
-outside of observed data while also better capturing the dynamics in the
-region observed. Advances in the theory and computational
-implementations of nonparametric methods make them ripe for such
-applications. We use the classic problem of optimal harvest of a marine
-fishery to illustrate how the nonparametric approach of Gaussian
-processes can be applied. We will compare Bayesian implementations of
-both nonparametric and parametric models, which best allow us to capture
-the uncertainty of model estimation in either case and permits a more
-natural comparison between approaches.
+Decision making under uncertainty is a ubiquitous challenge of natural
+resource management and conservation.
+
+Here we illustrate how a stochastic dynamic programming algorithm can be
+driven by the predictions from a Gaussian process model, sidestepping
+the need for an accurate model-based description of the sustem dynamics.
+
+Decision-theoretic (or optimal control) tools require a model that can
+assign probabilities of future states (e.g.~stock size of a fishery)
+given the current state and a proposed action (e.g.~fishing harvest or
+effort). The decision maker then seeks to determining the course of
+actions (also referred to as the policy) that maximizes the expected
+value of some objective function (such as net present value derived from
+the resource over time. (The approach can be adapted to permit
+alternatives to maximizing expectation, such as minimizing the maximum
+cost or damage that might be incurred. See @Polasky2011). Management
+frequently faces a sequential decision-making problem -- after selecting
+an action, the decision-maker may receive new information about the
+current state and must again choose an appropriate action --
+i.e.~setting the harvest limits each year based on stock assessments the
+year prior.
+
+Traditional approaches to optimal control (Pontryagin's principle,
+stochastic dynamic programming) rely on knowledge of the state equation.
+Such simple parametric models, frequently justified by mechanistic
+underpinnings are used to provide probabilities of future states.
+Fisheries literature, which enjoys some of the longest history in the
+use of the optimal control framework, provides a clear example. Models
+such as the Ricker or Beverton Holt curves are ubiquitous {[}too strong?
+references?{]} Parametric models can reflect variability introduced by a
+stochastic environment or demographic process my incorporating random
+shocks from a given distribution.
+
+As the parameter values for these models must be estimated from limited
+data,\\there will always be some uncertainty associated with these
+values. This uncertainty further compounds the intrinsic variability
+introduced by demographic or environmental noise. The degree of
+uncertainty in the parameter values can be inferred from the data and
+reflected in the estimates of the transition probabilities.
+
+Unfortunately, estimates of parameter uncertainty are only as good as
+the parametric models themselves. Too often, we do not understand the
+system dynamics well enough to know if a model provides a good
+approximation over the relevant range of states and timescales (criteria
+that we loosely refer to as defining the ``right'' or ``true'' model.)
+So called structural or model uncertainty is a more difficult problem
+than parametric uncertainty. Typical solutions involve either model
+choice, or model averaging. Either approach remains limited by the
+degree to which any of the proposed models are sufficiently close to the
+right model.
+
+The nature of decision-making problems provides a convenient way to
+compare models. Rather than compare models in terms of best fit to data
+or fret over the appropriate penalty for model complexity, model
+performance is defined in the concrete terms of the decision-maker's
+objective function, which we will take as given. (Much argument can be
+made over the `correct' objective function, e.g.~how to account for the
+social value of fish left in the sea vs.~the commercial value of fish
+harvested; see @Halpern2013 for further discussion of this issue.
+Alternatively, we can always compare model performance across multiple
+potential objective functions.) The decision-maker does not necessarily
+need a model that provides the best mechanistic understanding or the
+best long-term outcome, but rather the one that best estimates the
+probabilities of being in different states as a result of the possible
+actions.
+
+While simple mechanistic models can nevertheless provide important
+insights into long-term outcomes, such approaches are not well-suited
+for use in forecasting outcomes of potential management options.
+Non-parametric approaches offer a more flexible alternative that can
+both more accurately reflect the data available while also representing
+greater uncertainty in areas (of state-space) where data is lacking.
+
+We demonstrate how a Gaussian Process model of stock recruitment can
+lead to nearly optimal management through stochastic dynamic
+programming, comperable to knowing the correct structural equation for
+the underlying simulation. Meanwhile, parametric models that do not
+match the underlying dynamics can perform very poorly, even though they
+fit the data as well as the true model.
+
+Ecological research and management strategy should pay closer attention
+to the opportunities and challenges nonparametric modeling can offer.
+
+Non-parametric models can better represent uncertainties outside of
+observed data while also better capturing the dynamics in the region
+observed. Advances in the theory and computational implementations of
+nonparametric methods make them ripe for such applications. We use the
+classic problem of optimal harvest of a marine fishery to illustrate how
+the nonparametric approach of Gaussian processes can be applied. We will
+compare Bayesian implementations of both nonparametric and parametric
+models, which best allow us to capture the uncertainty of model
+estimation in either case and permits a more natural comparison between
+approaches.
 
@@
 
-\subsubsection{Uncertainties: the classical approach}
-
-The ecological dynamics of most systems of management interest are
-typically both complex and unknown. Despite this, quantitative policy
-optimization in these systems has almost exclusively been based on
-simple, mechanistically motivated parametric models. There are several
-reasons for this. First, limited available data almost always precludes
-any accurate estimate of more complicated models. Second, the
-computational demands of determining an optimal strategy in a sequential
-decision-theory problem suffer from the ``curse of dimensionality'' that
-makes it difficult to consider more complicated models, states, or
-actions. Unfortunately, these challenges also make it particularly
-difficult to reflect the true uncertainty either in the parameters of
-the given models (parametric uncertainty) or in the structure of the
-models themselves (structural uncertainty). Further unknowns such as
-measurement uncertainty, parameter uncertainty, unobserved states,
-knowledge of boundary conditions, etc. further compound the issue.
-Though a hierarchical Bayesian approach provides a natural way to
-address these from a statistical standpoint, formulating reasonable
-parametric descriptions of each form of uncertainty is a challenging
-task in itself, let alone the computational difficulties of solving such
-a system. Cressie et al. (2009) provides a good account of the successes
-and challenges of the approach. Applying these approaches in the
-management context of sequential decision making, in which forecasts
-must be obtained over a range of possible actions and updated regularly
-as new information arrives makes such an approach less feasible still.
-
-An active literature and growing computational power over the past
-several decades have only marginally improved this situation. Parametric
-or structural uncertainty can be introduced only by increasing the state
-space to reflect a distribution of possible parameters or the degree of
-belief in each of a set of possible models Frequently, the space of
-possible actions must then be reduced or the algorithms adjusted by
-approximations to keep computations feasible.
-
-Traditional approaches to optimal control (Pontryagin's principle,
-stochastic dynamic programming) rely on knowledge of the state equation,
-usually described by a simple parametric model. Here we illustrate how a
-stochastic dynamic programming algorithm can alternatively be driven by
-the predictions from a Gaussian process.
-
 \section{Approach and Methods}
@@
 
-\begin{align}
-X_{t+1} &= Z_t f(S_t) \\
-S_t &= X_t - h_t \\
-f(S_t) &= S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
-\end{align}
+\[X_{t+1} = Z_t f(S_t)  \] \[S_t = X_t - h_t \]
+\[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]
 
@@
 value of the resource, given an estimated stock size from the year
-before. Rich data and global concerns have made marine fisheries the
-crucible for much of the founding work (Gordon 1954; Reed 1979; May et
-al. 1979; Ludwig and Walters 1982) in managing ecosystems under
-uncertainty. Global trends (Worm et al. 2006) and controversy (Hilborn
-2007; Worm et al. 2009) have made understanding these challenges all the
-more pressing.
+before. The economic value and ecological concern have made marine
+fisheries the crucible for much of the founding work (Gordon 1954; Reed
+1979; May et al. 1979; Ludwig and Walters 1982) in managing ecosystems
+under uncertainty. Global trends (Worm et al. 2006) and controversy
+(Hilborn 2007; Worm et al. 2009) have made understanding these
+challenges all the more pressing.
 
@@
 model choice or model averaging (B. K. Williams 2001; Athanassoglou and
-Xepapadeas 2012). For simplicity of the comparison we consider only the
-case of a structurally correct and structurally incorrect model, and
-estimate Bayesian posterior distributions of the parameters. We will
-also compare to a simpler maximum likelihood estimate of the parameters
-that ignores the issue of parameter uncertainty.
-
-We compare these parametric approaches to our alternative that uses a
-Gaussian Process (GP) in place of a given stock recruitment curve. We
-briefly revisit a description of the Gaussian Process approach that has
-appeared in recent literature before describing how we take the
-important step of adapting the GP to the decision theory problem that
-can still be solved using SDP.
+Xepapadeas 2012).
+
+\subsubsection{Parametric Models}
+
+We consider three candidate parametric models for the stock-recruitment
+function, which we refer to by the first authors of the publications in
+which they were first proposed.
+
+We generate the data with a four-parameter model that contains a tipping
+point (an Allee effect, see {[}@Allen, @Courchamp{]}) below which the
+stock decreases to zero,
+
+\[ X_{t+1} = Z_t S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(\frac{S_t - \theta}{K}\right)} \]
+
+\[ S_t = X_t - h_t \]
+
+The parameter $C$ reflects the location of the tipping point, $K$ the
+carrying capacity of the stock, and $r$ the base recruitment rate. $S_t$
+represents the stock size after a harvest $h_t$ has been implemented.
+$Z_t$ represents a log-normal random variable of log-mean zero and
+log-standard deviation parameter $\sigma$.
+
+We consider two alternative candidate models: the Ricker {[}@Ricker{]}
+stock-recruitment curve,
+
+\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
+
+and an alternative four-parameter model adapted from @Myers,
+
+\[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \]
+
+which contains a tipping point for $\theta > 2$ and becomes a
+Beverton-Holt model at $\theta = 1$.
+
+\subsubsection{Bayesian Inference of Parametric models}
+
+Given the sample data, we infer posterior distributions for each of the
+three models listed above using a Markov Chain Monte Carlo Gibbs Sampler
+(jags, see appendix for implementation details and code) given uniform
+priors. We run six chains for $10^6$ steps each and then assess
+convergence by Gelman-Rubin criterion and inspection of the traces, see
+appendix.
 
@@
 
-The use of Gaussian process (GP) regression (or ``kreging'' in the
+The use of Gaussian process (GP) regression (or ``kriging'' in the
 geospatial literature) to formulate a predictive model is relatively new
@@
 
-Barnosky, Anthony D., Elizabeth a Hadly, Jordi Bascompte, Eric L.
-Berlow, James H. Brown, Mikael Fortelius, Wayne M. Getz, et al. 2012.
-``Approaching a state shift in Earth's biosphere.'' \emph{Nature} 486
-(7401) (jun): 52--58. doi:10.1038/nature11018.
-\url{http://www.nature.com/doifinder/10.1038/nature11018}.
-
 Bestelmeyer, Brandon T., Michael C. Duniway, Darren K. James, Laura M.
@@
 
-Boettiger, Carl, and Alan Hastings. 2013. ``Tipping points: From
-patterns to predictions.'' \emph{Nature} 493 (7431) (jan): 157--158.
-doi:10.1038/493157a.
-\url{http://www.nature.com/doifinder/10.1038/493157a}.
-
 Brozović, Nicholas, and Wolfram Schlenker. 2011. ``Optimal management of
@@
 
-Cressie, Noel, Catherine a Calder, James S. Clark, Jay M. Ver Hoef, and
-Christopher K. Wikle. 2009. ``Accounting for uncertainty in ecological
-analysis: the strengths and limitations of hierarchical statistical
-modeling.'' \emph{Ecological Applications} 19 (3) (apr): 553--70.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/19425416}.
-
 Deisenroth, Marc Peter, Carl Edward Rasmussen, and Jan Peters. 2009.
@@
 \url{http://linkinghub.elsevier.com/retrieve/pii/S0925231209000162}.
-
-Geritz, Stefan A. H., and Eva Kisdi. 2012. ``Mathematical ecology: why
-mechanistic models?.'' \emph{Journal of mathematical biology} 65 (6-7)
-(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.
 
@@
 
-Scheffer, Marten, Jordi Bascompte, William A. Brock, Victor Brovkin,
-Stephen R. Carpenter, Vasilis Dakos, Hermann Held, Egbert H. van Nes,
-Max Rietkerk, and George Sugihara. 2009. ``Early-warning signals for
-critical transitions.'' \emph{Nature} 461 (7260): 53--9.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/19727193}.
-
 Scheffer, Marten, Stephen R. Carpenter, J. A. Foley, C. Folke, and B.

### 2013-06-25 inst/examples/sensitivity.Rmd

@@
 ```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+rm(list=ls())
 source("~/.knitr_defaults.R")
@@
 MaxT = 1000 # timeout for value iteration convergence
-seed <- round(runif(32) * 1e6)
-seed[1] <- 1
+seed <- round(runif(1) * 1e6)
+seed[1] <- 1234
 seed

### 2013-06-25 inst/examples/sensitivity.Rmd
run for zero (debugging still)

@@
 MaxT = 1000 # timeout for value iteration convergence
-seed <- round(runif(1) * 1e6)
+seed <- round(runif(32) * 1e6)
 seed[1] <- 1234
@@
 function(seed_i){
-
+  set.seed(seed_i)
   x <- numeric(Tobs)
@@
     x[t+1] = z_g() * f(x[t], h=0, p=p)
-  obs <- data.frame(x = c(rep(0,nz), 
-                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
-                    y = c(rep(0,nz), 
-                          x[2:Tobs]))
+
+  X = c(rep(0,nz), pmax(rep(0,Tobs-1), x[1:(Tobs-1)]))
+  Y = c(rep(0,nz), x[2:Tobs]))
 
@@
   ## @knitr gp
-  gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
+  gp <- gp_mcmc(X, y=Y, n=1e5, s2.p = s2.p, d.p = d.p)
   gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)

### 2013-06-26 inst/doc/nonparametric-bayes.tex
updated draft

@@
 
+\paragraph{Opening}
+
 Decision making under uncertainty is a ubiquitous challenge of natural
-resource management and conservation.
-
-Here we illustrate how a stochastic dynamic programming algorithm can be
-driven by the predictions from a Gaussian process model, sidestepping
-the need for an accurate model-based description of the sustem dynamics.
+resource management and conservation.\\Here we illustrate how a
+stochastic dynamic programming algorithm can be driven by the
+predictions from a Gaussian process model, sidestepping the need for an
+accurate model-based description of the system dynamics.
+
+add references to {[}@Fischer2009, @Polasky2011{]} as persective on
+combining complex / critical dynamics + control theory
+
+\paragraph{Models for decision-making under uncertainty}
 
@@
 
+More historical context, e.g. (Reed 1979, )
+
+\paragraph{Parametric models}
+
 Traditional approaches to optimal control (Pontryagin's principle,
@@
 Fisheries literature, which enjoys some of the longest history in the
-use of the optimal control framework, provides a clear example. Models
-such as the Ricker or Beverton Holt curves are ubiquitous {[}too strong?
-references?{]} Parametric models can reflect variability introduced by a
-stochastic environment or demographic process my incorporating random
-shocks from a given distribution.
+use of the optimal control framework, provides a clear example {[} {]} .
+Models such as the Ricker or Beverton Holt curves are ubiquitous {[}too
+strong? references?{]} Parametric models can reflect variability
+introduced by a stochastic environment or demographic process my
+incorporating random shocks from a given distribution.
+
+\paragraph{Parametric uncertainty}
 
@@
 reflected in the estimates of the transition probabilities.
+
+(Mangel 1985; Schapaugh and Tyre 2013)
+
+\paragraph{Structural Uncertainty}
 
@@
 right model.
+
+(B. K. Williams 2001; Athanassoglou and Xepapadeas 2012).
+
+\paragraph{Comparing models}
 
@@
 match the underlying dynamics can perform very poorly, even though they
-fit the data as well as the true model.
-
-Ecological research and management strategy should pay closer attention
-to the opportunities and challenges nonparametric modeling can offer.
+fit the data as well as the true model.\\Ecological research and
+management strategy should pay closer attention to the opportunities and
+challenges nonparametric modeling can offer.
+
+\paragraph{Background on Fisheries Context}
+
+\paragraph{Background on Tipping points}
 
@@
 
-\subsubsection{Terminology: ``Non-parametric''}
+\section{Approach and Methods}
+
+\subsection{The optimal control problem in fisheries management}
+
+In our example, we focus on the problem in which a manager must set the
+harvest level for a marine fishery each year to maximize the net present
+value of the resource, given an estimated stock size from the year
+before. The economic value and ecological concern have made marine
+fisheries the crucible for much of the founding work (Gordon 1954; Reed
+1979; May et al. 1979; Ludwig and Walters 1982) in managing ecosystems
+under uncertainty. Global trends (Worm et al. 2006) and controversy
+(Hilborn 2007; Worm et al. 2009) have made understanding these
+challenges all the more pressing.
+
+To permit comparisons against a theoretical optimum we will consider
+data on the stock dynamics simulated from a simple parametric model in
+which recruitment of the fish stock $X_{t+1}$ in the following year is a
+stochastic process governed by a function $f$ of the current stock
+$X_t$, selected harvest policy $h_t$, and noise process $Z$,
+
+\[X_{t+1} = Z_t f(X_t, h_t) \]
+
+Given parameters for the function $f$ and probability distribution $Z$,
+along with a given economic model determining the price/profit
+$\Pi(X_t, h_t)$ realized in a given year given a choice of harvest $h_t$
+and observed stock $X_t$. This problem can be solved exactly for
+discretized values of stock $X$ and policy $h$ using stochastic dynamic
+programming (SDP) (Mangel 1985). Problems of this sort underpin much
+marine fisheries management today.
+
+A crux of this approach is correctly specifying the functional form of
+$f$, along with its parameters. The standard approach uses one of a
+handful of common parametric models representing the stock-recruitment
+relationship, usually after estimating the model parameters from any
+available existing data. Uncertainty in the parameter estimates can be
+estimated and integrated over to determine the optimal policy under
+under uncertainty (Mangel 1985; Schapaugh and Tyre 2013). Uncertainty in
+the model structure itself can only be addressed in this approach by
+hypothesizing alternative model structures, and then performing some
+model choice or model averaging (B. K. Williams 2001; Athanassoglou and
+Xepapadeas 2012).
+
+\subsection{Underlying Model}
+
+Concerns over the potential for tipping points in ecological dynamics
+(Scheffer et al. 2001) highlight the dangers of uncertainty in
+ecological management and pose a substantial challenge to existing
+decision-theoretic approaches (Brozović and Schlenker 2011). Because
+intervention is often too late after a tipping point has been crossed
+(but see Hughes et al. (2013)), management is most often concerned with
+avoiding potentially catastrophic tipping points before any data is
+available at or following a transition that would more clearly reveal
+these regime shift dynamics (e.g. Bestelmeyer et al. 2012).
+
+To illustrate the value of the non-parametric Bayesian approach to
+management, we focus on example of a system containing such a tipping
+point whose dynamics can still be described by a simple, one-dimensional
+parametric model.\\We will focus on a simple parametric model for a
+single species (derived from fist principles by Allen et al. 2005) as
+our underlying ``reality''.
+
+\[X_{t+1} = Z_t f(S_t)  \] \[S_t = X_t - h_t \]
+\[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]
+
+Where $Z_t$ is multiplicative noise function with mean 1, representing
+stochastic growth. We will consider log-normal noise with shape
+parameter $\sigma_g$. We start with an example in which the parameters
+are $r =2$, $K=8$, $C=5$ and $\sigma_g = 0.1$.
+
+As a low-dimensional system completely described by three parameters,
+this scenario should if anything be favorable to a parametric-based
+approach. This model contains an Allee effect, or tipping point, below
+which the population is not self-sustaining and shrinks to zero
+(Courchamp, Berec, and Gascoigne 2008).
+
+\paragraph{Simulated training data}
+
+We generate initial observational data under the model described in Eq 1
+for $T_{\textrm{obs}}=40$ time steps, under a given arbitrary sequence
+of harvest intensities, $h_t$. We consider the case in which most of the
+data comes from a limited region of state space (e.g.~near a stable
+equilibrium), leaving us without observations of the population dynamics
+at very low levels which would be useful in discrimating between
+recruitment curves {[}@{]} or demonstrating the existence of a tipping
+point (Scheffer et al. 2001).\\Using data simulated from a specified
+model rather than empirical data permits the comparison against the true
+underlying dynamics, setting a bar for the optimal performance possible.
+
+\subsubsection{Parametric Models}
+
+We consider three candidate parametric models for the stock-recruitment
+function, which we refer to by the first authors of the publications in
+which they were first proposed.
+
+We generate the data with a four-parameter model that contains a tipping
+point, as discussed above (equation 1), (an Allee effect, see {[}@Allen,
+@Courchamp{]}) below which the stock decreases to zero,
+
+\[ X_{t+1} = Z_t S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(\frac{S_t - \theta}{K}\right)} \]
+
+\[ S_t = X_t - h_t \]
+
+The parameter $C$ reflects the location of the tipping point, $K$ the
+carrying capacity of the stock, and $r$ the base recruitment rate. $S_t$
+represents the stock size after a harvest $h_t$ has been implemented.
+$Z_t$ represents a log-normal random variable of log-mean zero and
+log-standard deviation parameter $\sigma$.
+
+We consider two alternative candidate models: the Ricker {[}@Ricker{]}
+stock-recruitment curve,
+
+\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
+
+and an alternative four-parameter model adapted from @Myers,
+
+\[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \]
+
+which contains a tipping point for $\theta > 2$ and becomes a
+Beverton-Holt model at $\theta = 1$.
+
+\subsubsection{Bayesian Inference of Parametric models}
+
+Given the sample data, we infer posterior distributions for each of the
+three models listed above using a Markov Chain Monte Carlo Gibbs Sampler
+(jags, see appendix for implementation details and code) given uniform
+priors. We run six chains for $10^6$ steps each and then assess
+convergence by Gelman-Rubin criterion and inspection of the traces, see
+appendix.
+
+\subsubsection{The Non-parametric Bayesian alternative for
+stock-recruitment curves}
+
+\paragraph{Terminology: ``Non-parametric''}
 
@@
 clearest by way of example.
-
-\section{Approach and Methods}
-
-\subsection{Underlying Model}
-
-Concerns over the potential for tipping points in ecological dynamics
-(Scheffer et al. 2001) highlight the dangers of uncertainty in
-ecological management and pose a substantial challenge to existing
-decision-theoretic approaches (Brozović and Schlenker 2011). Because
-intervention is often too late after a tipping point has been crossed
-(but see Hughes et al. (2013)), management is most often concerned with
-avoiding potentially catastrophic tipping points before any data is
-available at or following a transition that would more clearly reveal
-these regime shift dynamics (e.g. Bestelmeyer et al. 2012).
-
-To illustrate the value of the non-parametric Bayesian approach to
-management, we focus on example of a system containing such a tipping
-point whose dynamics can still be described by a simple, one-dimensional
-parametric model.\\We will focus on a simple parametric model for a
-single species (derived from fist principles by Allen et al. 2005) as
-our underlying ``reality''.
-
-\[X_{t+1} = Z_t f(S_t)  \] \[S_t = X_t - h_t \]
-\[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]
-
-Where $Z_t$ is multiplicative noise function with mean 1, representing
-stochastic growth. We will consider log-normal noise with shape
-parameter $\sigma_g$. We start with an example in which the parameters
-are $r =2$, $K=8$, $C=5$ and $\sigma_g = 0.1$.
-
-As a low-dimensional system completely described by three parameters,
-this scenario should if anything be favorable to a parametric-based
-approach. This model contains an Allee effect, or tipping point, below
-which the population is not self-sustaining and shrinks to zero
-(Courchamp, Berec, and Gascoigne 2008).
-
-Both parametric and nonparametric approaches will require training data
-on which to base their model of the process. We generate the training
-data under the model described in Eq 1 for $T_{\textrm{obs}}=40$ time
-steps, under a known but not necessarily optimal sequence of harvest
-intensities, $h_t$. For simplicity we imagine a fishery that started
-from zero harvest pressure and has been gradually increasing the
-harvest.
-
-\textbf{\emph{Should we include any emprical examples?}}
-
-\subsubsection{The optimal control problem in fisheries management}
-
-In our example, we focus on the problem in which a manager must set the
-harvest level for a marine fishery each year to maximize the net present
-value of the resource, given an estimated stock size from the year
-before. The economic value and ecological concern have made marine
-fisheries the crucible for much of the founding work (Gordon 1954; Reed
-1979; May et al. 1979; Ludwig and Walters 1982) in managing ecosystems
-under uncertainty. Global trends (Worm et al. 2006) and controversy
-(Hilborn 2007; Worm et al. 2009) have made understanding these
-challenges all the more pressing.
-
-To permit comparisons against a theoretical optimum we will consider
-data on the stock dynamics simulated from a simple parametric model in
-which recruitment of the fish stock $X_{t+1}$ in the following year is a
-stochastic process governed by a function $f$ of the current stock
-$X_t$, selected harvest policy $h_t$, and noise process $Z$,
-
-\[X_{t+1} = Z_t f(X_t, h_t) \]
-
-Given parameters for the function $f$ and probability distribution $Z$,
-along with a given economic model determining the price/profit
-$\Pi(X_t, h_t)$ realized in a given year given a choice of harvest $h_t$
-and observed stock $X_t$. This problem can be solved exactly for
-discretized values of stock $X$ and policy $h$ using stochastic dynamic
-programming (SDP) (Mangel 1985). Problems of this sort underpin much
-marine fisheries management today.
-
-A crux of this approach is correctly specifying the functional form of
-$f$, along with its parameters. The standard approach uses one of a
-handful of common parametric models representing the stock-recruitment
-relationship, usually after estimating the model parameters from any
-available existing data. Uncertainty in the parameter estimates can be
-estimated and integrated over to determine the optimal policy under
-under uncertainty (Mangel 1985; Schapaugh and Tyre 2013). Uncertainty in
-the model structure itself can only be addressed in this approach by
-hypothesizing alternative model structures, and then performing some
-model choice or model averaging (B. K. Williams 2001; Athanassoglou and
-Xepapadeas 2012).
-
-\subsubsection{Parametric Models}
-
-We consider three candidate parametric models for the stock-recruitment
-function, which we refer to by the first authors of the publications in
-which they were first proposed.
-
-We generate the data with a four-parameter model that contains a tipping
-point (an Allee effect, see {[}@Allen, @Courchamp{]}) below which the
-stock decreases to zero,
-
-\[ X_{t+1} = Z_t S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(\frac{S_t - \theta}{K}\right)} \]
-
-\[ S_t = X_t - h_t \]
-
-The parameter $C$ reflects the location of the tipping point, $K$ the
-carrying capacity of the stock, and $r$ the base recruitment rate. $S_t$
-represents the stock size after a harvest $h_t$ has been implemented.
-$Z_t$ represents a log-normal random variable of log-mean zero and
-log-standard deviation parameter $\sigma$.
-
-We consider two alternative candidate models: the Ricker {[}@Ricker{]}
-stock-recruitment curve,
-
-\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
-
-and an alternative four-parameter model adapted from @Myers,
-
-\[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \]
-
-which contains a tipping point for $\theta > 2$ and becomes a
-Beverton-Holt model at $\theta = 1$.
-
-\subsubsection{Bayesian Inference of Parametric models}
-
-Given the sample data, we infer posterior distributions for each of the
-three models listed above using a Markov Chain Monte Carlo Gibbs Sampler
-(jags, see appendix for implementation details and code) given uniform
-priors. We run six chains for $10^6$ steps each and then assess
-convergence by Gelman-Rubin criterion and inspection of the traces, see
-appendix.
-
-\subsubsection{The Non-parametric Bayesian alternative for
-stock-recruitment curves}
 
@@
 range of possible (parametric) models that might describe the data. The
-GP allows us to consider a set of possible curves simultaneously.
-
-\subsubsection{Background on Gaussian Process inference}
+GP allows us to consider probabilities on a large set of possible curves
+simultaneously.
+
+The posterior distribution for the hyper-parameters of the Gaussian
+process model are estimated by Metropolis-Hastings algorithm, again with
+details and code provided in the Appendix. Rasmussen and Williams (2006)
+provides an excellent general introduction to Gaussian Processes and
+Munch et al. (2005) first discusses their application in the context of
+population dynamics models such as fisheries stock-recruitment
+relationships.
+
+\subsubsection{SDP via GP}
 
@@
 This completes the algorithm adapting the GP to the sequential
-decision-making problem through SDP, which we believe has not yet been
-demonstrated in the literature. We provide an R package implementation
-of this, along with the Gaussian process inference, in the supplemental
-materials.
+decision-making problem through SDP, which has not been previously
+demonstrated.\\We further provide an R package implementation as
+described in the supplemental materials.
 
@@
 correct model by maximum likelihood.
-
-The posterior distribution for the hyper-parameters of the Gaussian
-process model are estimated by Metropolis-Hastings algorithm, again with
-details and code provided in the Appendix. Rasmussen and Williams (2006)
-provides an excellent general introduction to Gaussian Processes and
-Munch et al. (2005) first discusses their application in this context.
 
@@
 
+\paragraph{Big picture: Linking GP to SDP}
+
 Non-parametric Bayesian methods have received far too little attention
@@
 
+Though simple mechanistically motivated models may be best for the
+latter case (@Cuddington2013; Geritz and Kisdi 2012), such models can be
+not only inaccurate but misleading in quantitative decision making. This
+arises for two reasons:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  We do not know what the correct parameter values are for the models.
+\item
+  We do not know what the correct models are for ecological systems.
+\end{enumerate}
+
+The complexity of ecological interactions and a lack of data contribute
+greatly to both of the problems. This concern is particularly acute in
+case of ecological tipping points (Barnosky et al. 2012; Scheffer et al.
+2001), which arise from feedbacks common in ecological systems and can
+lead to a sudden catastrophic transition to an undesirable state. In
+general we do not know where such tipping points are unless we have
+already observed the transition, in which case it is frequently too late
+to respond. Though there may exist early warning signals for certain
+kinds of these transitions that are driven by slow changes (Scheffer et
+al. 2009), we do not know when, where, or how to apply them to the
+decision making context more generally (Boettiger and Hastings 2013).
+
 While non-parametric Bayesian approaches will not always be preferable
@@
 
+\subsubsection{Higher dimensions}
+
 In this simulated example, the underlying dynamics are truly governed by
@@
 prove even more valuable.
-
- Data complexity. Perhaps too far out of scope\ldots{} The nonparametric
-Bayesian approach is also better suited to complex and disparate data.
-Incorporating various sources of information into mechanistic models can
-be an immensely difficult due to the increased complexity involved.
 
@@
 \url{http://linkinghub.elsevier.com/retrieve/pii/S0925231209000162}.
+
+Geritz, Stefan A. H., and Eva Kisdi. 2012. ``Mathematical ecology: why
+mechanistic models?.'' \emph{Journal of mathematical biology} 65 (6-7)
+(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.
 

### 2013-07-02 inst/examples/sensitivity.Rmd
example GP sensitivity results

@@
   X = c(rep(0,nz), pmax(rep(0,Tobs-1), x[1:(Tobs-1)]))
-  Y = c(rep(0,nz), x[2:Tobs]))
+  Y = c(rep(0,nz), x[2:Tobs])
 
@@
   tmp <- melt(tmp[2:dim(tmp)[2]])
-  actual_over_optimal$GP
+  tmp$value[tmp$variable == "GP"]
 
@@
 ```{r}
-yields
-#yields <- melt(yields, id=c("method", "V1", "sd"))
-#ggplot(yields) + geom_histogram(aes(V1)) + facet_wrap(~method, scale='free')
+yields_dat <- melt(yields)
+names(yields_dat) <- c("replicate", "simulation", "value")
+#class(yields_dat$simulation) = "factor"
+
+col_metadata <- c(replicate = "id of individual replicates using identical settings, relative to the theoretical optimum",
+  simulation = "id number for the simulation, which varies the simulated data used",
+  value = "total profit achieved by specified run")
+unit_metadata <- list(replicate = list(1:dim(yields)[1]), simulation = list(1:dim(yields)[2]), value = "currency (arbitrary)")
+```
+
+```{r}
+ggplot(yields_dat) + geom_density(aes(value)) 
 ````
+
+```{r}
+ggplot(yields_dat) + geom_density(aes(value, group=simulation)) 
+````
+
+```{r}
+ggplot(yields_dat) + geom_density(aes(value, group=replicate)) 
+````

### 2013-07-02 inst/doc/nonparametric-bayes.Rmd

@@
 
-[@Fischer2009, @Polasky2011]
-
 Decision making under uncertainty is a ubiquitous challenge of natural 
@@
 be driven by the predictions from a Gaussian process model, sidestepping
-the need for an accurate model-based description of the sustem dynamics.
+the need for an accurate model-based description of the system dynamics.
+
+add references to [@Fischer2009, @Polasky2011] as persective on combining complex / critical dynamics + control theory
 
@@
 
-[@Reed1979, @Mangel1985]
+More historical context, e.g. [@Reed1979, @Mangel1985]
 

### 2013-07-28 inst/doc/nonparametric-bayes.Rmd
use knit script to handle types, updated sensitivity

@@
 <!-- Run R code for analysis, to be called by figures -->
-```{r generics-and-libraries, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+```{r plot-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
 library(nonparametricbayes) 
+opts_chunk$set(external=TRUE)
+read_chunk("../examples/BUGS/external-chunks.R")
+
 library(ggplot2)
@@
                "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-opts_chunk$set(external=TRUE)
-read_chunk("../examples/BUGS/external-chunks.R")
-```
-
-```{r plot-options, cache=FALSE}
-options(xtable.print.comment=FALSE)
-options(xtable.type = 'latex', table.placement="H")
-opts_knit$set(progress = TRUE, verbose = TRUE)
-opts_chunk$set(dev = c('pdf', 'png'), fig.width=5.5, fig.height=4,
- cache.path = 'cache/nonparametric-bayes/', cache=TRUE, external=TRUE)
-opts_chunk$set(warning=FALSE, message=FALSE, comment=NA,
-               tidy=FALSE)
-toggle = "markup" # "hide" 
 theme_set(theme_bw(base_size=12))
+toggle = "hide" 
 ```
@@
 
-```{r Figure1, dependson=c("assemble-models", "plot-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
+```{r Figure1, dependson=c("assemble-models"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
 Graph shows the expected value for the function $f$ under each model.
@@
 ```
-```{r Figureb, dependson=c("assemble-models", "par-fns", "plot-options"), toggle = 1}
+```{r Figureb, dependson=c("assemble-models", "par-fns"), toggle = 1}
 ```
@@
 
-```{r Figure2, dependson = c("assemble-opt", "plot-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
+```{r Figure2, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
 Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", toggle = 1}
@@
 
-```{r Figure3, dependson=c("sim", "plot-options"), fig.cap="Gaussian process inference outperforms parametric estimates.
+```{r Figure3, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
 Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", toggle = 1}
@@
 
-```{r Figure4, dependson=c("profits", "plot-options"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
+```{r Figure4, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
 For each data set, the three models are estimated as described above.

### 2013-07-28 inst/examples/sensitivity.Rmd
use knit script to handle types, updated sensitivity

@@
 source("~/.knitr_defaults.R")
-#opts_knit$set(upload.fun = socialR::flickr.url)
+opts_knit$set(upload.fun = socialR::flickr.url)
 library(knitcitations)
@@
 f <- Myers
-p <- c(1.5, 2.5, 8)
-K <- 5  # approx, a li'l' less
-allee <- 1.2 # approx, a li'l' less
+
 ```
@@
 z_m <- function() 1
-x_grid <- seq(0, 1.5 * K, length=50)
+x_grid <- seq(0, 1.5 * 6, length=50)
 h_grid <- x_grid
@@
 xT <- 0
-Xo <-  allee+.5# observations start from
-x0 <- K # simulation under policy starts from
+Xo <-  1.7 # observations start from
+x0 <- 5 # simulation under policy starts from
 Tobs <- 40
@@
   set.seed(seed_i)
+
+  p <- c(1.5 + rnorm(1, 0, .1), 2.5 + rnorm(1, 0, .1), 8 + rnorm(1, 0, .2))
+
   x <- numeric(Tobs)
@@
 ```{r}
+head(yields_dat)
+```
+
+```{r}
 ggplot(yields_dat) + geom_density(aes(value)) 
@@
 ```{r}
-ggplot(yields_dat) + geom_density(aes(value, group=simulation)) 
+ggplot(yields_dat) + geom_density(aes(value, col=simulation)) 
 ````
@@
 ```{r}
-ggplot(yields_dat) + geom_density(aes(value, group=replicate)) 
+ggplot(yields_dat) + geom_density(aes(value, col=replicate)) 
 ````
+
+```{r}
+save(list=ls(), file="sensitivity.rda")
+```

### 2013-08-26 R/gp_mcmc.R
additional commenting and handling of out of bounds parameters

@@ -22,20 +22,26 @@ gp_mcmc <- function(x, y, init_pars = c(l=1, sigma.n=1), n = 1e4, d.p = c(5,5), 
     )
     log(prior)
   }
+  
   posterior <- function(pars, x, y){ 
-    l <- exp(pars[1])
-    sigma.n <- exp(pars[2])
+    
+    l <- exp(pars[1]) 
+    sigma.n <- exp(pars[2]) + exp(-10)
     
     cov <- function(X, Y) outer(X, Y, SE, l)
     I <- diag(1, length(x))
     K <- cov(x, x) 
     
-    loglik <- - 0.5 * t(y) %*% solve(K + sigma.n^2 * I) %*% y -
-      log(det(K + sigma.n^2*I)) -
-      length(y) * log(2 * pi) / 2
+  #  if(identical(log(det(K + sigma.n^2*I)), -Inf))
+  #    loglik <- -1e30
+  #  else 
+      loglik <- - 0.5 * t(y) %*% solve(K + sigma.n^2 * I) %*% y -
+        log(det(K + sigma.n^2*I)) -
+        length(y) * log(2 * pi) / 2
     
     loglik + lpriors(pars)
   }
+  
   
   out <- metrop(posterior, log(init_pars), n, x = x, y = y)  
   
@@ -85,10 +91,13 @@ gp_predict <- function(gp, x_predict, burnin=0, thin=1, covs=FALSE){
   # list format better for return
   Cf_posterior <- lapply(out, `[[`, "Cf")
   
-  list(Ef_posterior = Ef_posterior, 
-       Vf_posterior = Vf_posterior,
-       Cf_posterior = Cf_posterior,
-       E_Ef = E_Ef, E_Cf = E_Cf, E_Vf = E_Vf)
+  list(Ef_posterior = Ef_posterior,  # a dim(X) by (n_draws) matrix giving the E(X) for each of the draws
+       Vf_posterior = Vf_posterior,  # a dim(X) by (n_draws) matrix giving the V(X) for each of the draws
+       Cf_posterior = Cf_posterior,  # a list of length (n_draws), each with a dim(X)^2 covar matrix C(X)
+       E_Ef = E_Ef, # dim(X) vector, mean across draws of Ef_posterior
+       E_Cf = E_Cf, # dim(X)^2 covariance matrix, mean covar across samples
+       E_Vf = E_Vf # diag(E_Cf), same as rowMeans(Vf_posterior)
+       )
   
 }  
   

### 2013-08-26 inst/examples/sensitivity-allen.Rmd

--- 2013-07-28 inst/examples/sensitivity.Rmd
@@
 
-```{r stateeq-local}
-f <- Myers
-
-```
-
 Various parameters defining noise dynamics, grid, and policy costs.
 
-```{r sdp-pars, dependson="stateeq-local"}
+```{r sdp-pars}
 sigma_g <- 0.05
@@
 z_m <- function() 1
-x_grid <- seq(0, 1.5 * 6, length=50)
+x_grid <- seq(0, 1.5 * 8, length=50)
 h_grid <- x_grid
@@
 xT <- 0
-Xo <-  1.7 # observations start from
-x0 <- 5 # simulation under policy starts from
+Xo <-  5.7 # observations start from
+x0 <- 8 # simulation under policy starts from
 Tobs <- 40
 MaxT = 1000 # timeout for value iteration convergence
-seed <- round(runif(32) * 1e6)
-seed[1] <- 1234
+````
+
+```{r stateeq-local}
+f <- RickerAllee
+
+```
+
+```{r}
+set.seed(12345)
+seed <- round(runif(12) * 1e6)
 seed
@@
 
-32 data-sets from the same model
+Replicate data-sets from the same model
 
-```{r }
+```{r warning=TRUE}
 yields <- sfSapply(seed, 
@@
 
-  p <- c(1.5 + rnorm(1, 0, .1), 2.5 + rnorm(1, 0, .1), 8 + rnorm(1, 0, .2))
+  p <- c(2 + rnorm(1, 0, .1), 8 + rnorm(1, 0, .1), 5 + rnorm(1, 0, .2))
 
+  initial_data <- function(){
     x <- numeric(Tobs)
@@
 
+    out = list(X=X,Y=Y)
+    if(x[Tobs-1] < x[1]){
+      warning("Training population crashed, trying again")
+      out <- initial_data()
+    } 
+    out
+  }
+
+  o <- initial_data()
+  X <- o$X; Y <- o$Y
   ## @knitr gp-priors
@@
 
+Compare over `r length(seed)` simulation configurations with different parameters and different training data, the GP does consistently well.  Only one case has relatively poor outcomes.
+
 ```{r}
-ggplot(yields_dat) + geom_density(aes(value, col=simulation)) 
+ggplot(yields_dat) + geom_density(aes(value, color=as.factor(simulation), fill=as.factor(simulation)), alpha=.5)
 ````
 
+Within a replicate number, outcomes not that varied over parameters.
+
 ```{r}
-ggplot(yields_dat) + geom_density(aes(value, col=replicate)) 
+ggplot(yields_dat) + geom_density(aes(value, group=as.factor(replicate)), alpha=.7)
 ````

### 2013-08-26 inst/doc/nonparametric-bayes.tex
updated run from zero

@@
 \newcommand{\NormalTok}[1]{{#1}}
-\usepackage{graphicx}
-% We will generate all images so they have a width \maxwidth. This means
-% that they will get their normal width if they fit onto the page, but
-% are scaled down if they would overflow the margins.
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
-\else\Gin@nat@width\fi}
-\makeatother
-\let\Oldincludegraphics\includegraphics
-\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
 \ifxetex
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{options}\NormalTok{(}\DataTypeTok{xtable.print.comment =} \OtherTok{FALSE}\NormalTok{)}
-\KeywordTok{options}\NormalTok{(}\DataTypeTok{xtable.type =} \StringTok{"latex"}\NormalTok{, }\DataTypeTok{table.placement =} \StringTok{"H"}\NormalTok{)}
-\NormalTok{opts_knit$}\KeywordTok{set}\NormalTok{(}\DataTypeTok{progress =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{)}
-\NormalTok{opts_chunk$}\KeywordTok{set}\NormalTok{(}\DataTypeTok{dev =} \KeywordTok{c}\NormalTok{(}\StringTok{"pdf"}\NormalTok{, }\StringTok{"png"}\NormalTok{), }\DataTypeTok{fig.width =} \FloatTok{5.5}\NormalTok{, }\DataTypeTok{fig.height =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cache.path =} \StringTok{"cache/nonparametric-bayes/"}\NormalTok{, }
-    \DataTypeTok{cache =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{external =} \OtherTok{TRUE}\NormalTok{)}
-\NormalTok{opts_chunk$}\KeywordTok{set}\NormalTok{(}\DataTypeTok{warning =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{message =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{comment =} \OtherTok{NA}\NormalTok{, }\DataTypeTok{tidy =} \OtherTok{FALSE}\NormalTok{)}
-\NormalTok{toggle = }\StringTok{"markup"}
-\KeywordTok{theme_set}\NormalTok{(}\KeywordTok{theme_bw}\NormalTok{(}\DataTypeTok{base_size =} \DecValTok{12}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{require}\NormalTok{(modeest)}
-\NormalTok{posterior.mode <- function(x) \{}
-  \KeywordTok{mlv}\NormalTok{(x, }\DataTypeTok{method=}\StringTok{"shorth"}\NormalTok{)$M}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{f <- RickerAllee}
-\NormalTok{p <- }\KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{5}\NormalTok{)}
-\NormalTok{K <- }\DecValTok{10}  \CommentTok{# approx, a li'l' less}
-\NormalTok{allee <- }\DecValTok{5} \CommentTok{# approx, a li'l' less}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{sigma_g <- }\FloatTok{0.05}
-\NormalTok{sigma_m <- }\FloatTok{0.0}
-\NormalTok{z_g <- function() }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{, sigma_g)}
-\NormalTok{z_m <- function() }\DecValTok{1}
-\NormalTok{x_grid <- }\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\FloatTok{1.5} \NormalTok{* K, }\DataTypeTok{length=}\DecValTok{50}\NormalTok{)}
-\NormalTok{h_grid <- x_grid}
-\NormalTok{profit <- function(x,h) }\KeywordTok{pmin}\NormalTok{(x, h)}
-\NormalTok{delta <- }\FloatTok{0.01}
-\NormalTok{OptTime <- }\DecValTok{50}  \CommentTok{# stationarity with unstable models is tricky thing}
-\NormalTok{reward <- }\DecValTok{0}
-\NormalTok{xT <- }\DecValTok{0}
-\NormalTok{Xo <-  allee}\FloatTok{+.5}\CommentTok{# observations start from}
-\NormalTok{x0 <- K }\CommentTok{# simulation under policy starts from}
-\NormalTok{Tobs <- }\DecValTok{40}
-\NormalTok{MaxT = }\DecValTok{1000} \CommentTok{# timeout for value iteration convergence}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-  \KeywordTok{set.seed}\NormalTok{(}\DecValTok{1234}\NormalTok{)}
-  \CommentTok{#harvest <- sort(rep(seq(0, .5, length=7), 5))}
-  \NormalTok{x <- }\KeywordTok{numeric}\NormalTok{(Tobs)}
-  \NormalTok{x[}\DecValTok{1}\NormalTok{] <- Xo}
-  \NormalTok{nz <- }\DecValTok{1}
-  \NormalTok{for(t in }\DecValTok{1}\NormalTok{:(Tobs}\DecValTok{-1}\NormalTok{))}
-    \NormalTok{x[t}\DecValTok{+1}\NormalTok{] = }\KeywordTok{z_g}\NormalTok{() * }\KeywordTok{f}\NormalTok{(x[t], }\DataTypeTok{h=}\DecValTok{0}\NormalTok{, }\DataTypeTok{p=}\NormalTok{p)}
-  \NormalTok{obs <- }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,nz), }
-                          \KeywordTok{pmax}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,Tobs}\DecValTok{-1}\NormalTok{), x[}\DecValTok{1}\NormalTok{:(Tobs}\DecValTok{-1}\NormalTok{)])), }
-                    \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,nz), }
-                          \NormalTok{x[}\DecValTok{2}\NormalTok{:Tobs]))}
-\NormalTok{raw_plot <- }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{time =} \DecValTok{1}\NormalTok{:Tobs, }\DataTypeTok{x=}\NormalTok{x), }\KeywordTok{aes}\NormalTok{(time,x)) + }\KeywordTok{geom_line}\NormalTok{()}
-\NormalTok{raw_plot}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/obs.pdf}
-\caption{plot of chunk obs}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{set.seed}\NormalTok{(}\DecValTok{12345}\NormalTok{)}
-\NormalTok{estf <- function(p)\{ }
-    \NormalTok{mu <- }\KeywordTok{f}\NormalTok{(obs$x,}\DecValTok{0}\NormalTok{,p)}
-    \NormalTok{-}\KeywordTok{sum}\NormalTok{(}\KeywordTok{dlnorm}\NormalTok{(obs$y, }\KeywordTok{log}\NormalTok{(mu), p[}\DecValTok{4}\NormalTok{]), }\DataTypeTok{log=}\OtherTok{TRUE}\NormalTok{)}
-\NormalTok{\}}
-\NormalTok{par <- }\KeywordTok{c}\NormalTok{(p[}\DecValTok{1}\NormalTok{]*}\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{1}\NormalTok{), }
-         \NormalTok{p[}\DecValTok{2}\NormalTok{]*}\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{1}\NormalTok{), }
-         \NormalTok{p[}\DecValTok{3}\NormalTok{]*}\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{, .}\DecValTok{1}\NormalTok{), }
-         \NormalTok{sigma_g * }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{1}\NormalTok{))}
-\NormalTok{o <- }\KeywordTok{optim}\NormalTok{(par, estf, }\DataTypeTok{method=}\StringTok{"L"}\NormalTok{, }\DataTypeTok{lower=}\KeywordTok{c}\NormalTok{(}\FloatTok{1e-5}\NormalTok{,}\FloatTok{1e-5}\NormalTok{,}\FloatTok{1e-5}\NormalTok{,}\FloatTok{1e-5}\NormalTok{))}
-\NormalTok{f_alt <- f}
-\NormalTok{p_alt <- }\KeywordTok{c}\NormalTok{(}\KeywordTok{as.numeric}\NormalTok{(o$par[}\DecValTok{1}\NormalTok{]), }\KeywordTok{as.numeric}\NormalTok{(o$par[}\DecValTok{2}\NormalTok{]), }\KeywordTok{as.numeric}\NormalTok{(o$par[}\DecValTok{3}\NormalTok{]))}
-\NormalTok{sigma_g_alt <- }\KeywordTok{as.numeric}\NormalTok{(o$par[}\DecValTok{4}\NormalTok{])}
-
-\NormalTok{est <- }\KeywordTok{list}\NormalTok{(}\DataTypeTok{f =} \NormalTok{f_alt, }\DataTypeTok{p =} \NormalTok{p_alt, }\DataTypeTok{sigma_g =} \NormalTok{sigma_g_alt, }\DataTypeTok{mloglik=}\NormalTok{o$value)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{true_means <- }\KeywordTok{sapply}\NormalTok{(x_grid, f, }\DecValTok{0}\NormalTok{, p)}
-\NormalTok{est_means <- }\KeywordTok{sapply}\NormalTok{(x_grid, est$f, }\DecValTok{0}\NormalTok{, est$p)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)}
-\NormalTok{s2.p <- }\KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{5}\NormalTok{)  }
-\NormalTok{d.p = }\KeywordTok{c}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{1}\NormalTok{/}\FloatTok{0.1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{gp <- }\KeywordTok{gp_mcmc}\NormalTok{(obs$x, }\DataTypeTok{y=}\NormalTok{obs$y, }\DataTypeTok{n=}\FloatTok{1e5}\NormalTok{, }\DataTypeTok{s2.p =} \NormalTok{s2.p, }\DataTypeTok{d.p =} \NormalTok{d.p)}
-\NormalTok{gp_dat <- }\KeywordTok{gp_predict}\NormalTok{(gp, x_grid, }\DataTypeTok{burnin=}\FloatTok{1e4}\NormalTok{, }\DataTypeTok{thin=}\DecValTok{300}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{gp_assessment_plots <- }\KeywordTok{summary_gp_mcmc}\NormalTok{(gp, }\DataTypeTok{burnin=}\FloatTok{1e4}\NormalTok{, }\DataTypeTok{thin=}\DecValTok{300}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# Summarize the GP model}
-\NormalTok{tgp_dat <- }
-    \KeywordTok{data.frame}\NormalTok{(  }\DataTypeTok{x =} \NormalTok{x_grid, }
-                 \DataTypeTok{y =} \NormalTok{gp_dat$E_Ef, }
-                 \DataTypeTok{ymin =} \NormalTok{gp_dat$E_Ef - }\DecValTok{2} \NormalTok{* }\KeywordTok{sqrt}\NormalTok{(gp_dat$E_Vf), }
-                 \DataTypeTok{ymax =} \NormalTok{gp_dat$E_Ef + }\DecValTok{2} \NormalTok{* }\KeywordTok{sqrt}\NormalTok{(gp_dat$E_Vf) )}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{y <- x }
-\NormalTok{N <- }\KeywordTok{length}\NormalTok{(x);}
-\NormalTok{jags.data <- }\KeywordTok{list}\NormalTok{(}\StringTok{"N"}\NormalTok{,}\StringTok{"y"}\NormalTok{)}
-\NormalTok{n.chains <- }\DecValTok{6}
-\NormalTok{n.iter <- }\FloatTok{1e6}
-\NormalTok{n.burnin <- }\KeywordTok{floor}\NormalTok{(}\DecValTok{10000}\NormalTok{)}
-\NormalTok{n.thin <- }\KeywordTok{max}\NormalTok{(}\DecValTok{1}\NormalTok{, }\KeywordTok{floor}\NormalTok{(n.chains * (n.iter - n.burnin)/}\DecValTok{1000}\NormalTok{))}
-\NormalTok{n.update <- }\DecValTok{10}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{stdQ_prior_p <- }\KeywordTok{c}\NormalTok{(}\FloatTok{1e-6}\NormalTok{, }\DecValTok{100}\NormalTok{)}
-\NormalTok{stdR_prior_p <- }\KeywordTok{c}\NormalTok{(}\FloatTok{1e-6}\NormalTok{, .}\DecValTok{1}\NormalTok{)}
-\NormalTok{stdQ_prior  <- function(x) }\KeywordTok{dunif}\NormalTok{(x, stdQ_prior_p[}\DecValTok{1}\NormalTok{], stdQ_prior_p[}\DecValTok{2}\NormalTok{])}
-\NormalTok{stdR_prior  <- function(x) }\KeywordTok{dunif}\NormalTok{(x, stdR_prior_p[}\DecValTok{1}\NormalTok{], stdR_prior_p[}\DecValTok{2}\NormalTok{])}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{K_prior_p <- }\KeywordTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{20.0}\NormalTok{)}
-\NormalTok{r0_prior_p <- }\KeywordTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{6.0}\NormalTok{)}
-\NormalTok{theta_prior_p <- }\KeywordTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{20.0}\NormalTok{)}
-
-\NormalTok{bugs.model <- }
-\KeywordTok{paste}\NormalTok{(}\KeywordTok{sprintf}\NormalTok{(}
-\StringTok{"model\{}
-\StringTok{  K     ~ dunif(%s, %s)}
-\StringTok{  r0    ~ dunif(%s, %s)}
-\StringTok{  theta ~ dunif(%s, %s)}
-\StringTok{  stdQ ~ dunif(%s, %s)"}\NormalTok{, }
-  \NormalTok{K_prior_p[}\DecValTok{1}\NormalTok{], K_prior_p[}\DecValTok{2}\NormalTok{],}
-  \NormalTok{r0_prior_p[}\DecValTok{1}\NormalTok{], r0_prior_p[}\DecValTok{2}\NormalTok{],}
-  \NormalTok{theta_prior_p[}\DecValTok{1}\NormalTok{], theta_prior_p[}\DecValTok{2}\NormalTok{],}
-  \NormalTok{stdQ_prior_p[}\DecValTok{1}\NormalTok{], stdQ_prior_p[}\DecValTok{2}\NormalTok{]),}
-
-  \StringTok{"}
-\StringTok{  iQ <- 1 / (stdQ * stdQ);}
-\StringTok{  y[1] ~ dunif(0, 10)}
-\StringTok{  for(t in 1:(N-1))\{}
-\StringTok{    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K)* (y[t] - theta) / K }
-\StringTok{    y[t+1] ~ dlnorm(mu[t], iQ) }
-\StringTok{  \}}
-\StringTok{\}"}\NormalTok{)}
-\KeywordTok{writeLines}\NormalTok{(bugs.model, }\StringTok{"allen_process.bugs"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{K_prior     <- function(x) }\KeywordTok{dunif}\NormalTok{(x, K_prior_p[}\DecValTok{1}\NormalTok{], K_prior_p[}\DecValTok{2}\NormalTok{])}
-\NormalTok{r0_prior <- function(x) }\KeywordTok{dunif}\NormalTok{(x, r0_prior_p[}\DecValTok{1}\NormalTok{], r0_prior_p[}\DecValTok{2}\NormalTok{])}
-\NormalTok{theta_prior <- function(x) }\KeywordTok{dunif}\NormalTok{(x, theta_prior_p[}\DecValTok{1}\NormalTok{], theta_prior_p[}\DecValTok{2}\NormalTok{])}
-\NormalTok{par_priors  <- }\KeywordTok{list}\NormalTok{(}\DataTypeTok{K =} \NormalTok{K_prior, }\DataTypeTok{deviance =} \NormalTok{function(x) }\DecValTok{0} \NormalTok{* x, }
-                    \DataTypeTok{r0 =} \NormalTok{r0_prior, }\DataTypeTok{theta =} \NormalTok{theta_prior,}
-                    \DataTypeTok{stdQ =} \NormalTok{stdQ_prior)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{jags.params=}\KeywordTok{c}\NormalTok{(}\StringTok{"K"}\NormalTok{,}\StringTok{"r0"}\NormalTok{,}\StringTok{"theta"}\NormalTok{,}\StringTok{"stdQ"}\NormalTok{) }\CommentTok{# be sensible about the order here}
-\NormalTok{jags.inits <- function()\{}
-  \KeywordTok{list}\NormalTok{(}\StringTok{"K"}\NormalTok{= }\DecValTok{10} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{, }\FloatTok{0.1}\NormalTok{),}
-       \StringTok{"r0"}\NormalTok{= }\DecValTok{1} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{, }\FloatTok{0.1}\NormalTok{) ,}
-       \StringTok{"theta"}\NormalTok{=   }\DecValTok{5} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{, }\FloatTok{0.1}\NormalTok{) , }
-       \StringTok{"stdQ"}\NormalTok{= }\KeywordTok{abs}\NormalTok{( }\FloatTok{0.1} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{, }\FloatTok{0.1}\NormalTok{)),}
-       \DataTypeTok{.RNG.name=}\StringTok{"base::Wichmann-Hill"}\NormalTok{, }\DataTypeTok{.RNG.seed=}\DecValTok{123}\NormalTok{)}
-\NormalTok{\}}
-
-\KeywordTok{set.seed}\NormalTok{(}\DecValTok{1234}\NormalTok{)}
-\CommentTok{# parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)}
-\NormalTok{allen_jags <- }\KeywordTok{do.call}\NormalTok{(jags.parallel, }\KeywordTok{list}\NormalTok{(}\DataTypeTok{data=}\NormalTok{jags.data, }\DataTypeTok{inits=}\NormalTok{jags.inits, }
-                                      \NormalTok{jags.params, }\DataTypeTok{n.chains=}\NormalTok{n.chains, }
-                                      \DataTypeTok{n.iter=}\NormalTok{n.iter, }\DataTypeTok{n.thin=}\NormalTok{n.thin, }
-                                      \DataTypeTok{n.burnin=}\NormalTok{n.burnin, }
-                                      \DataTypeTok{model.file=}\StringTok{"allen_process.bugs"}\NormalTok{))}
-
-\CommentTok{# Run again iteratively if we haven't met the Gelman-Rubin convergence criterion}
-\KeywordTok{recompile}\NormalTok{(allen_jags) }\CommentTok{# required for parallel}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 328
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 328
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 328
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 328
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 328
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 328
-
-Initializing model
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{allen_jags <- }\KeywordTok{do.call}\NormalTok{(autojags, }
-                                            \KeywordTok{list}\NormalTok{(}\DataTypeTok{object=}\NormalTok{allen_jags, }\DataTypeTok{n.update=}\NormalTok{n.update, }
-                           \DataTypeTok{n.iter=}\NormalTok{n.iter, }\DataTypeTok{n.thin =} \NormalTok{n.thin))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{tmp <- }\KeywordTok{lapply}\NormalTok{(}\KeywordTok{as.mcmc}\NormalTok{(allen_jags), as.matrix) }\CommentTok{# strip classes the hard way...}
-\NormalTok{allen_posteriors <- }\KeywordTok{melt}\NormalTok{(tmp, }\DataTypeTok{id =} \KeywordTok{colnames}\NormalTok{(tmp[[}\DecValTok{1}\NormalTok{]])) }
-\KeywordTok{names}\NormalTok{(allen_posteriors) = }\KeywordTok{c}\NormalTok{(}\StringTok{"index"}\NormalTok{, }\StringTok{"variable"}\NormalTok{, }\StringTok{"value"}\NormalTok{, }\StringTok{"chain"}\NormalTok{)}
-\NormalTok{plot_allen_traces <- }\KeywordTok{ggplot}\NormalTok{(allen_posteriors) + }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(index, value)) + }
-  \KeywordTok{facet_wrap}\NormalTok{(~ variable, }\DataTypeTok{scale=}\StringTok{"free"}\NormalTok{, }\DataTypeTok{ncol=}\DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{allen_priors <- }\KeywordTok{ddply}\NormalTok{(allen_posteriors, }\StringTok{"variable"}\NormalTok{, function(dd)\{}
-    \NormalTok{grid <- }\KeywordTok{seq}\NormalTok{(}\KeywordTok{min}\NormalTok{(dd$value), }\KeywordTok{max}\NormalTok{(dd$value), }\DataTypeTok{length =} \DecValTok{100}\NormalTok{) }
-    \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{value =} \NormalTok{grid, }\DataTypeTok{density =} \NormalTok{par_priors[[dd$variable[}\DecValTok{1}\NormalTok{]]](grid))}
-\NormalTok{\})}
-\NormalTok{plot_allen_posteriors <- }\KeywordTok{ggplot}\NormalTok{(allen_posteriors, }\KeywordTok{aes}\NormalTok{(value)) + }
-  \KeywordTok{stat_density}\NormalTok{(}\DataTypeTok{geom=}\StringTok{"path"}\NormalTok{, }\DataTypeTok{position=}\StringTok{"identity"}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.7}\NormalTok{) +}
-  \KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{data=}\NormalTok{allen_priors, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{value, }\DataTypeTok{y=}\NormalTok{density), }\DataTypeTok{col=}\StringTok{"red"}\NormalTok{) + }
-  \KeywordTok{facet_wrap}\NormalTok{(~ variable, }\DataTypeTok{scale=}\StringTok{"free"}\NormalTok{, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{A <- allen_posteriors}
-\NormalTok{A$index <- A$index + A$chain * }\KeywordTok{max}\NormalTok{(A$index) }\CommentTok{# Combine samples across chains by renumbering index }
-\NormalTok{pardist <- }\KeywordTok{acast}\NormalTok{(A, index ~ variable)}
-\NormalTok{bayes_coef <- }\KeywordTok{apply}\NormalTok{(pardist,}\DecValTok{2}\NormalTok{, posterior.mode) }
-\NormalTok{bayes_pars <- }\KeywordTok{unname}\NormalTok{(}\KeywordTok{c}\NormalTok{(bayes_coef[}\StringTok{"r0"}\NormalTok{], bayes_coef[}\StringTok{"K"}\NormalTok{], bayes_coef[}\StringTok{"theta"}\NormalTok{])) }\CommentTok{# parameters formatted for f}
-\NormalTok{allen_f <- function(x,h,p) }\KeywordTok{unname}\NormalTok{(}\KeywordTok{RickerAllee}\NormalTok{(x,h, }\KeywordTok{unname}\NormalTok{(p[}\KeywordTok{c}\NormalTok{(}\StringTok{"r0"}\NormalTok{, }\StringTok{"K"}\NormalTok{, }\StringTok{"theta"}\NormalTok{)])))}
-\NormalTok{allen_means <- }\KeywordTok{sapply}\NormalTok{(x_grid, f, }\DecValTok{0}\NormalTok{, bayes_pars)}
-\NormalTok{bayes_pars}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-[1] 0.5003 7.7115 4.6039
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{head}\NormalTok{(pardist)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-        K deviance     r0    stdQ theta
-170 7.898    34.41 0.6668 0.05566 3.956
-171 7.745    32.76 0.7554 0.05137 4.668
-172 7.577    31.66 1.0705 0.04562 4.073
-173 7.859    31.26 1.5647 0.04238 5.087
-174 7.821    30.67 1.7348 0.04662 5.230
-175 8.003    34.90 1.4495 0.05005 5.444
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{K_prior_p <- }\KeywordTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{40.0}\NormalTok{)}
-\NormalTok{r0_prior_p <- }\KeywordTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{20.0}\NormalTok{)}
-\NormalTok{bugs.model <- }
-\KeywordTok{paste}\NormalTok{(}\KeywordTok{sprintf}\NormalTok{(}
-\StringTok{"model\{}
-\StringTok{  K    ~ dunif(%s, %s)}
-\StringTok{  r0    ~ dunif(%s, %s)}
-\StringTok{  stdQ ~ dunif(%s, %s)"}\NormalTok{, }
-  \NormalTok{K_prior_p[}\DecValTok{1}\NormalTok{], K_prior_p[}\DecValTok{2}\NormalTok{],}
-  \NormalTok{r0_prior_p[}\DecValTok{1}\NormalTok{], r0_prior_p[}\DecValTok{2}\NormalTok{],}
-  \NormalTok{stdQ_prior_p[}\DecValTok{1}\NormalTok{], stdQ_prior_p[}\DecValTok{2}\NormalTok{]),}
-  \StringTok{"}
-\StringTok{  iQ <- 1 / (stdQ * stdQ);}
-\StringTok{  y[1] ~ dunif(0, 10)}
-\StringTok{  for(t in 1:(N-1))\{}
-\StringTok{    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K) }
-\StringTok{    y[t+1] ~ dlnorm(mu[t], iQ) }
-\StringTok{  \}}
-\StringTok{\}"}\NormalTok{)}
-\KeywordTok{writeLines}\NormalTok{(bugs.model, }\StringTok{"ricker_process.bugs"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{K_prior     <- function(x) }\KeywordTok{dunif}\NormalTok{(x, K_prior_p[}\DecValTok{1}\NormalTok{], K_prior_p[}\DecValTok{2}\NormalTok{])}
-\NormalTok{r0_prior <- function(x) }\KeywordTok{dunif}\NormalTok{(x, r0_prior_p[}\DecValTok{1}\NormalTok{], r0_prior_p[}\DecValTok{2}\NormalTok{])}
-\NormalTok{par_priors <- }\KeywordTok{list}\NormalTok{(}\DataTypeTok{K =} \NormalTok{K_prior, }\DataTypeTok{deviance =} \NormalTok{function(x) }\DecValTok{0} \NormalTok{* x, }
-                   \DataTypeTok{r0 =} \NormalTok{r0_prior, }\DataTypeTok{stdQ =} \NormalTok{stdQ_prior)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{jags.params=}\KeywordTok{c}\NormalTok{(}\StringTok{"K"}\NormalTok{,}\StringTok{"r0"}\NormalTok{, }\StringTok{"stdQ"}\NormalTok{)}
-\NormalTok{jags.inits <- function()\{}
-  \KeywordTok{list}\NormalTok{(}\StringTok{"K"}\NormalTok{= }\DecValTok{10} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
-       \StringTok{"r0"}\NormalTok{= }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
-       \StringTok{"stdQ"}\NormalTok{=}\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.05}\NormalTok{) * }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
-       \DataTypeTok{.RNG.name=}\StringTok{"base::Wichmann-Hill"}\NormalTok{, }\DataTypeTok{.RNG.seed=}\DecValTok{123}\NormalTok{)}
-\NormalTok{\}}
-\KeywordTok{set.seed}\NormalTok{(}\DecValTok{12345}\NormalTok{) }
-\NormalTok{ricker_jags <- }\KeywordTok{do.call}\NormalTok{(jags.parallel, }
-                       \KeywordTok{list}\NormalTok{(}\DataTypeTok{data=}\NormalTok{jags.data, }\DataTypeTok{inits=}\NormalTok{jags.inits, }
-                            \NormalTok{jags.params, }\DataTypeTok{n.chains=}\NormalTok{n.chains, }
-                            \DataTypeTok{n.iter=}\NormalTok{n.iter, }\DataTypeTok{n.thin=}\NormalTok{n.thin, }\DataTypeTok{n.burnin=}\NormalTok{n.burnin,}
-                            \DataTypeTok{model.file=}\StringTok{"ricker_process.bugs"}\NormalTok{))}
-\KeywordTok{recompile}\NormalTok{(ricker_jags)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 249
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 249
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 249
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 249
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 249
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 249
-
-Initializing model
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{ricker_jags <- }\KeywordTok{do.call}\NormalTok{(autojags, }
-                       \KeywordTok{list}\NormalTok{(}\DataTypeTok{object=}\NormalTok{ricker_jags, }\DataTypeTok{n.update=}\NormalTok{n.update, }
-                                                        \DataTypeTok{n.iter=}\NormalTok{n.iter, }\DataTypeTok{n.thin =} \NormalTok{n.thin, }
-                                                        \DataTypeTok{progress.bar=}\StringTok{"none"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{tmp <- }\KeywordTok{lapply}\NormalTok{(}\KeywordTok{as.mcmc}\NormalTok{(ricker_jags), as.matrix) }\CommentTok{# strip classes the hard way...}
-\NormalTok{ricker_posteriors <- }\KeywordTok{melt}\NormalTok{(tmp, }\DataTypeTok{id =} \KeywordTok{colnames}\NormalTok{(tmp[[}\DecValTok{1}\NormalTok{]])) }
-\KeywordTok{names}\NormalTok{(ricker_posteriors) = }\KeywordTok{c}\NormalTok{(}\StringTok{"index"}\NormalTok{, }\StringTok{"variable"}\NormalTok{, }\StringTok{"value"}\NormalTok{, }\StringTok{"chain"}\NormalTok{)}
-\NormalTok{plot_ricker_traces <- }\KeywordTok{ggplot}\NormalTok{(ricker_posteriors) + }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(index, value)) + }
-  \KeywordTok{facet_wrap}\NormalTok{(~ variable, }\DataTypeTok{scale=}\StringTok{"free"}\NormalTok{, }\DataTypeTok{ncol=}\DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{ricker_priors <- }\KeywordTok{ddply}\NormalTok{(ricker_posteriors, }\StringTok{"variable"}\NormalTok{, function(dd)\{}
-    \NormalTok{grid <- }\KeywordTok{seq}\NormalTok{(}\KeywordTok{min}\NormalTok{(dd$value), }\KeywordTok{max}\NormalTok{(dd$value), }\DataTypeTok{length =} \DecValTok{100}\NormalTok{) }
-    \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{value =} \NormalTok{grid, }\DataTypeTok{density =} \NormalTok{par_priors[[dd$variable[}\DecValTok{1}\NormalTok{]]](grid))}
-\NormalTok{\})}
-\CommentTok{# plot posterior distributions}
-\NormalTok{plot_ricker_posteriors <- }\KeywordTok{ggplot}\NormalTok{(ricker_posteriors, }\KeywordTok{aes}\NormalTok{(value)) + }
-  \KeywordTok{stat_density}\NormalTok{(}\DataTypeTok{geom=}\StringTok{"path"}\NormalTok{, }\DataTypeTok{position=}\StringTok{"identity"}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.7}\NormalTok{) +}
-  \KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{data=}\NormalTok{ricker_priors, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{value, }\DataTypeTok{y=}\NormalTok{density), }\DataTypeTok{col=}\StringTok{"red"}\NormalTok{) + }
-  \KeywordTok{facet_wrap}\NormalTok{(~ variable, }\DataTypeTok{scale=}\StringTok{"free"}\NormalTok{, }\DataTypeTok{ncol=}\DecValTok{2}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{A <- ricker_posteriors}
-\NormalTok{A$index <- A$index + A$chain * }\KeywordTok{max}\NormalTok{(A$index) }\CommentTok{# Combine samples across chains by renumbering index }
-\NormalTok{ricker_pardist <- }\KeywordTok{acast}\NormalTok{(A, index ~ variable)}
-\NormalTok{bayes_coef <- }\KeywordTok{apply}\NormalTok{(ricker_pardist,}\DecValTok{2}\NormalTok{, posterior.mode) }
-\NormalTok{ricker_bayes_pars <- }\KeywordTok{unname}\NormalTok{(}\KeywordTok{c}\NormalTok{(bayes_coef[}\StringTok{"r0"}\NormalTok{], bayes_coef[}\StringTok{"K"}\NormalTok{]))}
-\NormalTok{ricker_f <- function(x,h,p)\{}
-  \KeywordTok{sapply}\NormalTok{(x, function(x)\{ }
-    \NormalTok{x <- }\KeywordTok{pmax}\NormalTok{(}\DecValTok{0}\NormalTok{, x-h) }
-    \KeywordTok{pmax}\NormalTok{(}\DecValTok{0}\NormalTok{, x * }\KeywordTok{exp}\NormalTok{(p[}\StringTok{"r0"}\NormalTok{] * (}\DecValTok{1} \NormalTok{- x / p[}\StringTok{"K"}\NormalTok{] )) )}
-  \NormalTok{\})}
-\NormalTok{\}}
-\NormalTok{ricker_means <- }\KeywordTok{sapply}\NormalTok{(x_grid, Ricker, }\DecValTok{0}\NormalTok{, ricker_bayes_pars[}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{)])}
-\KeywordTok{head}\NormalTok{(ricker_pardist)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-         K deviance      r0    stdQ
-170  7.727    35.73 0.29440 0.05197
-171  7.949    37.93 0.18845 0.05872
-172 32.183    44.54 0.02474 0.05688
-173  8.149    37.59 0.12204 0.05407
-174  7.460    39.01 0.38643 0.05495
-175  7.361    35.54 0.23805 0.05150
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{ricker_bayes_pars}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-[1] 0.1997 7.5976
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{r0_prior_p <- }\KeywordTok{c}\NormalTok{(.}\DecValTok{0001}\NormalTok{, }\FloatTok{10.0}\NormalTok{)}
-\NormalTok{theta_prior_p <- }\KeywordTok{c}\NormalTok{(.}\DecValTok{0001}\NormalTok{, }\FloatTok{10.0}\NormalTok{)}
-\NormalTok{K_prior_p <- }\KeywordTok{c}\NormalTok{(.}\DecValTok{0001}\NormalTok{, }\FloatTok{40.0}\NormalTok{)}
-\NormalTok{bugs.model <- }
-\KeywordTok{paste}\NormalTok{(}\KeywordTok{sprintf}\NormalTok{(}
-\StringTok{"model\{}
-\StringTok{  r0    ~ dunif(%s, %s)}
-\StringTok{  theta    ~ dunif(%s, %s)}
-\StringTok{  K    ~ dunif(%s, %s)}
-\StringTok{  stdQ ~ dunif(%s, %s)"}\NormalTok{, }
-  \NormalTok{r0_prior_p[}\DecValTok{1}\NormalTok{], r0_prior_p[}\DecValTok{2}\NormalTok{],}
-  \NormalTok{theta_prior_p[}\DecValTok{1}\NormalTok{], theta_prior_p[}\DecValTok{2}\NormalTok{],}
-  \NormalTok{K_prior_p[}\DecValTok{1}\NormalTok{], K_prior_p[}\DecValTok{2}\NormalTok{],}
-  \NormalTok{stdQ_prior_p[}\DecValTok{1}\NormalTok{], stdQ_prior_p[}\DecValTok{2}\NormalTok{]),}
-
-  \StringTok{"}
-\StringTok{  iQ <- 1 / (stdQ * stdQ);}
-
-\StringTok{  y[1] ~ dunif(0, 10)}
-\StringTok{  for(t in 1:(N-1))\{}
-\StringTok{    mu[t] <- log(r0)  + theta * log(y[t]) - log(1 + pow(abs(y[t]), theta) / K)}
-\StringTok{    y[t+1] ~ dlnorm(mu[t], iQ) }
-\StringTok{  \}}
-\StringTok{\}"}\NormalTok{)}
-\KeywordTok{writeLines}\NormalTok{(bugs.model, }\StringTok{"myers_process.bugs"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{K_prior     <- function(x) }\KeywordTok{dunif}\NormalTok{(x, K_prior_p[}\DecValTok{1}\NormalTok{], K_prior_p[}\DecValTok{2}\NormalTok{])}
-\NormalTok{r_prior     <- function(x) }\KeywordTok{dunif}\NormalTok{(x, r0_prior_p[}\DecValTok{1}\NormalTok{], r0_prior_p[}\DecValTok{2}\NormalTok{])}
-\NormalTok{theta_prior <- function(x) }\KeywordTok{dunif}\NormalTok{(x, theta_prior_p[}\DecValTok{1}\NormalTok{], theta_prior_p[}\DecValTok{2}\NormalTok{])}
-\NormalTok{par_priors <- }\KeywordTok{list}\NormalTok{( }\DataTypeTok{deviance =} \NormalTok{function(x) }\DecValTok{0} \NormalTok{* x, }\DataTypeTok{K =} \NormalTok{K_prior,}
-                    \DataTypeTok{r0 =} \NormalTok{r_prior, }\DataTypeTok{theta =} \NormalTok{theta_prior, }
-                    \DataTypeTok{stdQ =} \NormalTok{stdQ_prior)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{jags.params=}\KeywordTok{c}\NormalTok{(}\StringTok{"r0"}\NormalTok{, }\StringTok{"theta"}\NormalTok{, }\StringTok{"K"}\NormalTok{, }\StringTok{"stdQ"}\NormalTok{)}
-\NormalTok{jags.inits <- function()\{}
-  \KeywordTok{list}\NormalTok{(}\StringTok{"r0"}\NormalTok{= }\DecValTok{1} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{1}\NormalTok{), }
-       \StringTok{"K"}\NormalTok{=    }\DecValTok{10} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{1}\NormalTok{),}
-       \StringTok{"theta"} \NormalTok{= }\DecValTok{1} \NormalTok{* }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{1}\NormalTok{),  }
-       \StringTok{"stdQ"}\NormalTok{= }\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.2}\NormalTok{) * }\KeywordTok{rlnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{0}\NormalTok{,.}\DecValTok{1}\NormalTok{),}
-       \DataTypeTok{.RNG.name=}\StringTok{"base::Wichmann-Hill"}\NormalTok{, }\DataTypeTok{.RNG.seed=}\DecValTok{123}\NormalTok{)}
-\NormalTok{\}}
-\KeywordTok{set.seed}\NormalTok{(}\DecValTok{12345}\NormalTok{)}
-\NormalTok{myers_jags <- }\KeywordTok{do.call}\NormalTok{(jags.parallel, }
-                      \KeywordTok{list}\NormalTok{(}\DataTypeTok{data=}\NormalTok{jags.data, }\DataTypeTok{inits=}\NormalTok{jags.inits, }
-                                                     \NormalTok{jags.params, }\DataTypeTok{n.chains=}\NormalTok{n.chains, }
-                                                     \DataTypeTok{n.iter=}\NormalTok{n.iter, }\DataTypeTok{n.thin=}\NormalTok{n.thin,}
-                           \DataTypeTok{n.burnin=}\NormalTok{n.burnin, }
-                           \DataTypeTok{model.file=}\StringTok{"myers_process.bugs"}\NormalTok{))}
-\KeywordTok{recompile}\NormalTok{(myers_jags)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 406
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 406
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 406
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 406
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 406
-
-Initializing model
-
-Compiling model graph
-   Resolving undeclared variables
-   Allocating nodes
-   Graph Size: 406
-
-Initializing model
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{myers_jags <- }\KeywordTok{do.call}\NormalTok{(autojags, }
-                      \KeywordTok{list}\NormalTok{(myers_jags, }\DataTypeTok{n.update=}\NormalTok{n.update, }
-                           \DataTypeTok{n.iter=}\NormalTok{n.iter, }\DataTypeTok{n.thin =} \NormalTok{n.thin, }
-                           \DataTypeTok{progress.bar=}\StringTok{"none"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{tmp <- }\KeywordTok{lapply}\NormalTok{(}\KeywordTok{as.mcmc}\NormalTok{(myers_jags), as.matrix) }\CommentTok{# strip classes}
-\NormalTok{myers_posteriors <- }\KeywordTok{melt}\NormalTok{(tmp, }\DataTypeTok{id =} \KeywordTok{colnames}\NormalTok{(tmp[[}\DecValTok{1}\NormalTok{]])) }
-\KeywordTok{names}\NormalTok{(myers_posteriors) = }\KeywordTok{c}\NormalTok{(}\StringTok{"index"}\NormalTok{, }\StringTok{"variable"}\NormalTok{, }\StringTok{"value"}\NormalTok{, }\StringTok{"chain"}\NormalTok{)}
-\NormalTok{plot_myers_traces <- }\KeywordTok{ggplot}\NormalTok{(myers_posteriors) + }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(index, value)) +}
-  \KeywordTok{facet_wrap}\NormalTok{(~ variable, }\DataTypeTok{scale=}\StringTok{"free"}\NormalTok{, }\DataTypeTok{ncol=}\DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{par_prior_curves <- }\KeywordTok{ddply}\NormalTok{(myers_posteriors, }\StringTok{"variable"}\NormalTok{, function(dd)\{}
-    \NormalTok{grid <- }\KeywordTok{seq}\NormalTok{(}\KeywordTok{min}\NormalTok{(dd$value), }\KeywordTok{max}\NormalTok{(dd$value), }\DataTypeTok{length =} \DecValTok{100}\NormalTok{) }
-    \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{value =} \NormalTok{grid, }\DataTypeTok{density =} \NormalTok{par_priors[[dd$variable[}\DecValTok{1}\NormalTok{]]](grid))}
-\NormalTok{\})}
-\NormalTok{plot_myers_posteriors <- }\KeywordTok{ggplot}\NormalTok{(myers_posteriors, }\KeywordTok{aes}\NormalTok{(value)) + }
-  \KeywordTok{stat_density}\NormalTok{(}\DataTypeTok{geom=}\StringTok{"path"}\NormalTok{, }\DataTypeTok{position=}\StringTok{"identity"}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.7}\NormalTok{) +}
-  \KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{data=}\NormalTok{par_prior_curves, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{value, }\DataTypeTok{y=}\NormalTok{density), }\DataTypeTok{col=}\StringTok{"red"}\NormalTok{) + }
-  \KeywordTok{facet_wrap}\NormalTok{(~ variable, }\DataTypeTok{scale=}\StringTok{"free"}\NormalTok{, }\DataTypeTok{ncol=}\DecValTok{3}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{A <- myers_posteriors}
-\NormalTok{A$index <- A$index + A$chain * }\KeywordTok{max}\NormalTok{(A$index) }\CommentTok{# Combine samples across chains by renumbering index }
-\NormalTok{myers_pardist <- }\KeywordTok{acast}\NormalTok{(A, index ~ variable)}
-\NormalTok{bayes_coef <- }\KeywordTok{apply}\NormalTok{(myers_pardist,}\DecValTok{2}\NormalTok{, posterior.mode) }\CommentTok{# much better estimates}
-\NormalTok{myers_bayes_pars <- }\KeywordTok{unname}\NormalTok{(}\KeywordTok{c}\NormalTok{(bayes_coef[}\StringTok{"r0"}\NormalTok{], bayes_coef[}\StringTok{"theta"}\NormalTok{], bayes_coef[}\StringTok{"K"}\NormalTok{]))}
-\NormalTok{myers_means <- }\KeywordTok{sapply}\NormalTok{(x_grid, Myer_harvest, }\DecValTok{0}\NormalTok{, myers_bayes_pars)}
-\NormalTok{myers_f <- function(x,h,p) }\KeywordTok{Myer_harvest}\NormalTok{(x, h, p[}\KeywordTok{c}\NormalTok{(}\StringTok{"r0"}\NormalTok{, }\StringTok{"theta"}\NormalTok{, }\StringTok{"K"}\NormalTok{)])}
-\KeywordTok{head}\NormalTok{(myers_pardist)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-        K deviance     r0    stdQ  theta
-170 20.51    37.16 1.3995 0.04136 0.9864
-171 17.10    41.09 0.6076 0.05170 1.8684
-172 24.40    36.33 1.1692 0.04307 1.0715
-173 27.56    37.46 0.9533 0.05550 1.2000
-174 22.21    34.94 0.5140 0.04203 1.8694
-175 36.05    33.49 0.3089 0.03977 2.1519
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{myers_bayes_pars}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-[1]  0.5129  1.0894 23.9914
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{models <- }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{x=}\NormalTok{x_grid, }
-                                         \DataTypeTok{GP=}\NormalTok{tgp_dat$y, }
-                                         \DataTypeTok{True=}\NormalTok{true_means, }
-                     \DataTypeTok{MLE=}\NormalTok{est_means, }
-                                         \DataTypeTok{Ricker=}\NormalTok{ricker_means, }
-                     \DataTypeTok{Allen =} \NormalTok{allen_means,}
-                     \DataTypeTok{Myers =} \NormalTok{myers_means)}
-\NormalTok{models <- }\KeywordTok{melt}\NormalTok{(models, }\DataTypeTok{id=}\StringTok{"x"}\NormalTok{)}
-
-\CommentTok{# some labels}
-\KeywordTok{names}\NormalTok{(models) <- }\KeywordTok{c}\NormalTok{(}\StringTok{"x"}\NormalTok{, }\StringTok{"method"}\NormalTok{, }\StringTok{"value"}\NormalTok{)}
-
-\CommentTok{# labels for the colorkey too}
-\NormalTok{model_names = }\KeywordTok{c}\NormalTok{(}\StringTok{"GP"}\NormalTok{, }\StringTok{"True"}\NormalTok{, }\StringTok{"MLE"}\NormalTok{, }\StringTok{"Ricker"}\NormalTok{, }\StringTok{"Allen"}\NormalTok{, }\StringTok{"Myers"}\NormalTok{)}
-\NormalTok{colorkey=cbPalette}
-\KeywordTok{names}\NormalTok{(colorkey) = model_names }
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# uses expected values from GP, instead of integrating over posterior}
-\CommentTok{#matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)}
-\NormalTok{matrices_gp <- }\KeywordTok{gp_transition_matrix}\NormalTok{(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid) }
-\NormalTok{opt_gp <- }\KeywordTok{value_iteration}\NormalTok{(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{matrices_true <- }\KeywordTok{f_transition_matrix}\NormalTok{(f, p, x_grid, h_grid, sigma_g)}
-\NormalTok{opt_true <- }\KeywordTok{value_iteration}\NormalTok{(matrices_true, x_grid, h_grid, }\DataTypeTok{OptTime=}\NormalTok{MaxT, xT, profit, }\DataTypeTok{delta=}\NormalTok{delta)}
-\NormalTok{matrices_estimated <- }\KeywordTok{f_transition_matrix}\NormalTok{(est$f, est$p, x_grid, h_grid, est$sigma_g)}
-\NormalTok{opt_estimated <- }\KeywordTok{value_iteration}\NormalTok{(matrices_estimated, x_grid, h_grid, }\DataTypeTok{OptTime=}\NormalTok{MaxT, xT, profit, }\DataTypeTok{delta=}\NormalTok{delta)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{matrices_allen <- }\KeywordTok{parameter_uncertainty_SDP}\NormalTok{(allen_f, x_grid, h_grid, pardist, }\DecValTok{4}\NormalTok{)}
-\NormalTok{opt_allen <- }\KeywordTok{value_iteration}\NormalTok{(matrices_allen, x_grid, h_grid, }\DataTypeTok{OptTime=}\NormalTok{MaxT, xT, profit, }\DataTypeTok{delta=}\NormalTok{delta)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{matrices_ricker <- }\KeywordTok{parameter_uncertainty_SDP}\NormalTok{(ricker_f, x_grid, h_grid, }\KeywordTok{as.matrix}\NormalTok{(ricker_pardist), }\DecValTok{3}\NormalTok{)}
-\NormalTok{opt_ricker <- }\KeywordTok{value_iteration}\NormalTok{(matrices_ricker, x_grid, h_grid, }\DataTypeTok{OptTime=}\NormalTok{MaxT, xT, profit, }\DataTypeTok{delta=}\NormalTok{delta)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{matrices_myers <- }\KeywordTok{parameter_uncertainty_SDP}\NormalTok{(myers_f, x_grid, h_grid, }\KeywordTok{as.matrix}\NormalTok{(myers_pardist), }\DecValTok{4}\NormalTok{)}
-\NormalTok{myers_alt <- }\KeywordTok{value_iteration}\NormalTok{(matrices_myers, x_grid, h_grid, }\DataTypeTok{OptTime=}\NormalTok{MaxT, xT, profit, }\DataTypeTok{delta=}\NormalTok{delta)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{OPT = }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{GP =} \NormalTok{opt_gp$D, }\DataTypeTok{True =} \NormalTok{opt_true$D, }\DataTypeTok{MLE =} \NormalTok{opt_estimated$D, }\DataTypeTok{Ricker =} \NormalTok{opt_ricker$D, }\DataTypeTok{Allen =} \NormalTok{opt_allen$D, }\DataTypeTok{Myers =} \NormalTok{myers_alt$D)}
-\NormalTok{colorkey=cbPalette}
-\KeywordTok{names}\NormalTok{(colorkey) = }\KeywordTok{names}\NormalTok{(OPT) }
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{sims <- }\KeywordTok{lapply}\NormalTok{(OPT, function(D)\{}
-  \KeywordTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
-  \KeywordTok{lapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\DecValTok{100}\NormalTok{, function(i) }
-    \KeywordTok{ForwardSimulate}\NormalTok{(f, p, x_grid, h_grid, x0, D, z_g, }\DataTypeTok{profit=}\NormalTok{profit, }\DataTypeTok{OptTime=}\NormalTok{OptTime)}
-  \NormalTok{)}
-\NormalTok{\})}
-
-\NormalTok{dat <- }\KeywordTok{melt}\NormalTok{(sims, }\DataTypeTok{id=}\KeywordTok{names}\NormalTok{(sims[[}\DecValTok{1}\NormalTok{]][[}\DecValTok{1}\NormalTok{]]))}
-\NormalTok{sims_data <- }\KeywordTok{data.table}\NormalTok{(dat)}
-\KeywordTok{setnames}\NormalTok{(sims_data, }\KeywordTok{c}\NormalTok{(}\StringTok{"L1"}\NormalTok{, }\StringTok{"L2"}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\StringTok{"method"}\NormalTok{, }\StringTok{"reps"}\NormalTok{)) }
-\CommentTok{# Legend in original ordering please, not alphabetical: }
-\NormalTok{sims_data$method = }\KeywordTok{factor}\NormalTok{(sims_data$method, }\DataTypeTok{ordered=}\OtherTok{TRUE}\NormalTok{, }\DataTypeTok{levels=}\KeywordTok{names}\NormalTok{(OPT))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{Profit <- sims_data[, }\KeywordTok{sum}\NormalTok{(profit), by=}\KeywordTok{c}\NormalTok{(}\StringTok{"reps"}\NormalTok{, }\StringTok{"method"}\NormalTok{)]}
-\NormalTok{tmp <- }\KeywordTok{dcast}\NormalTok{(Profit, reps ~ method)}
-\CommentTok{#tmp$Allen <- tmp[,"Allen"] + rnorm(dim(tmp)[1], 0, 1) # jitter for plotting}
-\NormalTok{tmp <- tmp / tmp[,}\StringTok{"True"}\NormalTok{]}
-\NormalTok{tmp <- }\KeywordTok{melt}\NormalTok{(tmp[}\DecValTok{2}\NormalTok{:}\KeywordTok{dim}\NormalTok{(tmp)[}\DecValTok{2}\NormalTok{]])}
-\NormalTok{actual_over_optimal <-}\KeywordTok{subset}\NormalTok{(tmp, variable != }\StringTok{"True"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{allen_deviance <- -}\DecValTok{2}\NormalTok{*}\KeywordTok{posterior.mode}\NormalTok{(pardist[,}\StringTok{'deviance'}\NormalTok{])}
-\NormalTok{ricker_deviance <- -}\DecValTok{2}\NormalTok{*}\KeywordTok{posterior.mode}\NormalTok{(ricker_pardist[,}\StringTok{'deviance'}\NormalTok{])}
-\NormalTok{myers_deviance <- -}\DecValTok{2}\NormalTok{*}\KeywordTok{posterior.mode}\NormalTok{(myers_pardist[,}\StringTok{'deviance'}\NormalTok{])}
-\NormalTok{true_deviance <- }\DecValTok{2}\NormalTok{*}\KeywordTok{estf}\NormalTok{(}\KeywordTok{c}\NormalTok{(p, sigma_g))}
-\NormalTok{mle_deviance <- }\DecValTok{2}\NormalTok{*}\KeywordTok{estf}\NormalTok{(}\KeywordTok{c}\NormalTok{(est$p, est$sigma_g))}
-\NormalTok{xtable::}\KeywordTok{xtable}\NormalTok{(}\KeywordTok{as.table}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DataTypeTok{Allen =} \NormalTok{allen_deviance, }\DataTypeTok{Ricker=}\NormalTok{ricker_deviance, }\DataTypeTok{Myers=}\NormalTok{myers_deviance, }\DataTypeTok{True=}\NormalTok{true_deviance, }\DataTypeTok{MLE=}\NormalTok{mle_deviance)))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-% latex table generated in R 3.0.1 by xtable 1.7-1 package
-% Thu Jun 20 16:29:00 2013
-\begin{table}[ht]
-\centering
-\begin{tabular}{rr}
-  \hline
- & x \\ 
-  \hline
-Allen & -65.85 \\ 
-  Ricker & -71.85 \\ 
-  Myers & -70.92 \\ 
-  True & -61.08 \\ 
-  MLE & -889.99 \\ 
-   \hline
-\end{tabular}
-\end{table}
-\end{verbatim}
-
 \section{Abstract}
@@
 
-More historical context, e.g. (Reed 1979, )
+More historical context, e.g. {[}@Reed1979, @Mangel1985{]}
 
@@
 
-(Mangel 1985; Schapaugh and Tyre 2013)
+{[}@Mangel1985; @Schapaugh2013{]}
 
@@
 
-(B. K. Williams 2001; Athanassoglou and Xepapadeas 2012).
+{[}@Williams2001; @Athanassoglou2012{]}.
 
@@
 before. The economic value and ecological concern have made marine
-fisheries the crucible for much of the founding work (Gordon 1954; Reed
-1979; May et al. 1979; Ludwig and Walters 1982) in managing ecosystems
-under uncertainty. Global trends (Worm et al. 2006) and controversy
-(Hilborn 2007; Worm et al. 2009) have made understanding these
-challenges all the more pressing.
+fisheries the crucible for much of the founding work {[}@Gordon1954;
+@Reed1979; @May1979; @Ludwig1982{]} in managing ecosystems under
+uncertainty. Global trends {[}@Worm2006{]} and controversy
+{[}@Hilborn2007; @Worm2009{]} have made understanding these challenges
+all the more pressing.
 
@@
 discretized values of stock $X$ and policy $h$ using stochastic dynamic
-programming (SDP) (Mangel 1985). Problems of this sort underpin much
+programming (SDP) {[}@Mangel1985{]}. Problems of this sort underpin much
 marine fisheries management today.
@@
 estimated and integrated over to determine the optimal policy under
-under uncertainty (Mangel 1985; Schapaugh and Tyre 2013). Uncertainty in
-the model structure itself can only be addressed in this approach by
+under uncertainty {[}@Mangel1985; @Schapaugh2013{]}. Uncertainty in the
+model structure itself can only be addressed in this approach by
 hypothesizing alternative model structures, and then performing some
-model choice or model averaging (B. K. Williams 2001; Athanassoglou and
-Xepapadeas 2012).
+model choice or model averaging {[}@Williams2001; @Athanassoglou2012{]}.
 
@@
 Concerns over the potential for tipping points in ecological dynamics
-(Scheffer et al. 2001) highlight the dangers of uncertainty in
-ecological management and pose a substantial challenge to existing
-decision-theoretic approaches (Brozović and Schlenker 2011). Because
-intervention is often too late after a tipping point has been crossed
-(but see Hughes et al. (2013)), management is most often concerned with
-avoiding potentially catastrophic tipping points before any data is
-available at or following a transition that would more clearly reveal
-these regime shift dynamics (e.g. Bestelmeyer et al. 2012).
+{[}@Scheffer2001{]} highlight the dangers of uncertainty in ecological
+management and pose a substantial challenge to existing
+decision-theoretic approaches {[}@Brozovic2011{]}. Because intervention
+is often too late after a tipping point has been crossed (but see
+@Hughes2013), management is most often concerned with avoiding
+potentially catastrophic tipping points before any data is available at
+or following a transition that would more clearly reveal these regime
+shift dynamics {[}e.g. @Bestelmeyer2012{]}.
 
@@
 parametric model.\\We will focus on a simple parametric model for a
-single species (derived from fist principles by Allen et al. 2005) as
-our underlying ``reality''.
+single species {[}derived from fist principles by @Allen2005a{]} as our
+underlying ``reality''.
 
@@
 which the population is not self-sustaining and shrinks to zero
-(Courchamp, Berec, and Gascoigne 2008).
+{[}@Courchamp2008{]}.
 
@@
 recruitment curves {[}@{]} or demonstrating the existence of a tipping
-point (Scheffer et al. 2001).\\Using data simulated from a specified
-model rather than empirical data permits the comparison against the true
+point {[}@Scheffer2001{]}.\\Using data simulated from a specified model
+rather than empirical data permits the comparison against the true
 underlying dynamics, setting a bar for the optimal performance possible.
@@
 geospatial literature) to formulate a predictive model is relatively new
-in the context of modeling dynamical systems (Kocijan et al. 2005) and
-introduced in the ecological modeling and fisheries management by Munch
-et al. (2005). An accessible and thorough introduction to the
-formulation and use of GPs can be found in Rasmussen and Williams
-(2006).
+in the context of modeling dynamical systems {[}@Kocijan2005{]} and
+introduced in the ecological modeling and fisheries management by
+@Munch2005. An accessible and thorough introduction to the formulation
+and use of GPs can be found in @Rasmussen2006.
 
@@
 process model are estimated by Metropolis-Hastings algorithm, again with
-details and code provided in the Appendix. Rasmussen and Williams (2006)
-provides an excellent general introduction to Gaussian Processes and
-Munch et al. (2005) first discusses their application in the context of
-population dynamics models such as fisheries stock-recruitment
-relationships.
+details and code provided in the Appendix. @Rasmussen2006 provides an
+excellent general introduction to Gaussian Processes and @Munch2005
+first discusses their application in the context of population dynamics
+models such as fisheries stock-recruitment relationships.
 
@@
 
-Once the posterior Gaussian process (GP) has been estimated (e.g.~see
-Munch et al. 2005), it is necessary to adapt it in place of the
-parametric equation for the stochastic dynamic programming (SDP)
-solution (see Mangel and Clark 1988 for a detailed description of
-parametric SDP methods) to the optimal policy. The essence of the idea
-is straight forward -- we will use the estimated GP in place of the
-parametric growth function to determine the stochastic transition matrix
-on which the SDP calculations are based. The SDP is solved in a
-discretized state space -- both the continuously valued population
-densities $X$ and harvest quotas $h$ are first mapped to a bounded,
-discrete grid. (For simplicity we will consider a uniform grid, though
-for either parametric or GP-based SDP it is often advantageous to use a
-non-uniform discretization such as a basis function representation,
-e.g.~see (Deisenroth, Rasmussen, and Peters 2009)).
+Once the posterior Gaussian process (GP) has been estimated {[}e.g.~see
+@Munch2005{]}, it is necessary to adapt it in place of the parametric
+equation for the stochastic dynamic programming (SDP) solution {[}see
+@Mangel1988 for a detailed description of parametric SDP methods{]} to
+the optimal policy. The essence of the idea is straight forward -- we
+will use the estimated GP in place of the parametric growth function to
+determine the stochastic transition matrix on which the SDP calculations
+are based. The SDP is solved in a discretized state space -- both the
+continuously valued population densities $X$ and harvest quotas $h$ are
+first mapped to a bounded, discrete grid. (For simplicity we will
+consider a uniform grid, though for either parametric or GP-based SDP it
+is often advantageous to use a non-uniform discretization such as a
+basis function representation, e.g.~see {[}@Deisenroth2009{]}).
 
@@
 \section{Results}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{require}\NormalTok{(MASS)}
-\NormalTok{step_ahead <- function(x, f, p)\{}
-  \NormalTok{h = }\DecValTok{0}
-  \NormalTok{x_predict <- }\KeywordTok{sapply}\NormalTok{(x, f, h, p)}
-  \NormalTok{n <- }\KeywordTok{length}\NormalTok{(x_predict) - }\DecValTok{1}
-  \NormalTok{y <- }\KeywordTok{c}\NormalTok{(x[}\DecValTok{1}\NormalTok{], x_predict[}\DecValTok{1}\NormalTok{:n])}
-  \NormalTok{y}
-\NormalTok{\}}
-\NormalTok{step_ahead_posteriors <- function(x)\{}
-\NormalTok{gp_f_at_obs <- }\KeywordTok{gp_predict}\NormalTok{(gp, x, }\DataTypeTok{burnin=}\FloatTok{1e4}\NormalTok{, }\DataTypeTok{thin=}\DecValTok{300}\NormalTok{)}
-\NormalTok{df_post <- }\KeywordTok{melt}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(}\KeywordTok{sample}\NormalTok{(}\DecValTok{100}\NormalTok{), }
-  \NormalTok{function(i)\{}
-    \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{time =} \DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(x), }\DataTypeTok{stock =} \NormalTok{x, }
-                \DataTypeTok{GP =} \KeywordTok{mvrnorm}\NormalTok{(}\DecValTok{1}\NormalTok{, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),}
-                \DataTypeTok{True =} \KeywordTok{step_ahead}\NormalTok{(x,f,p),  }
-                \DataTypeTok{MLE =} \KeywordTok{step_ahead}\NormalTok{(x,f,est$p), }
-                \DataTypeTok{Allen =} \KeywordTok{step_ahead}\NormalTok{(x, allen_f, pardist[i,]), }
-                \DataTypeTok{Ricker =} \KeywordTok{step_ahead}\NormalTok{(x, ricker_f, ricker_pardist[i,]), }
-                \DataTypeTok{Myers =} \KeywordTok{step_ahead}\NormalTok{(x, myers_f, myers_pardist[i,]))}
-  \NormalTok{\}), }\DataTypeTok{id=}\KeywordTok{c}\NormalTok{(}\StringTok{"time"}\NormalTok{, }\StringTok{"stock"}\NormalTok{))}
-\NormalTok{\}}
-
-\NormalTok{df_post <- }\KeywordTok{step_ahead_posteriors}\NormalTok{(x)}
-
-\KeywordTok{ggplot}\NormalTok{(df_post) + }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, stock)) + }
-  \KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, value, }\DataTypeTok{col=}\NormalTok{variable, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(L1,variable)), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) + }
-  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{))) }
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/Figureb.pdf}
-\caption{plot of chunk Figureb}
-\end{figure}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{policies <- }\KeywordTok{melt}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{stock=}\NormalTok{x_grid, }\KeywordTok{sapply}\NormalTok{(OPT, function(x) x_grid[x])), }\DataTypeTok{id=}\StringTok{"stock"}\NormalTok{)}
-\KeywordTok{names}\NormalTok{(policies) <- }\KeywordTok{c}\NormalTok{(}\StringTok{"stock"}\NormalTok{, }\StringTok{"method"}\NormalTok{, }\StringTok{"value"}\NormalTok{)}
-
-\KeywordTok{ggplot}\NormalTok{(policies, }\KeywordTok{aes}\NormalTok{(stock, stock - value, }\DataTypeTok{color=}\NormalTok{method)) +}
-  \KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{lwd=}\FloatTok{1.2}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) + }\KeywordTok{xlab}\NormalTok{(}\StringTok{"stock size"}\NormalTok{) + }\KeywordTok{ylab}\NormalTok{(}\StringTok{"escapement"}\NormalTok{)  +}
-  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/Figure2.pdf}
-\caption{The steady-state optimal policy (infinite boundary) calculated
-under each model. Policies are shown in terms of target escapement,
-$S_t$, as under models such as this a constant escapement policy is
-expected to be optimal (Reed 1979).}
-\end{figure}
-
 The resulting optimal management strategy based on each of the inferred
@@
 $S_t$. Under models such as this a constant escapement policy is
-expected to be optimal (Reed 1979), whereby population levels below a
-certain size $S$ are unharvested, while above that size the harvest
+expected to be optimal {[}@Reed1979{]}, whereby population levels below
+a certain size $S$ are unharvested, while above that size the harvest
 strategy aims to return the population to $S$, resulting in the
 hockey-stick shaped policies shown.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{ggplot}\NormalTok{(sims_data) + }
-  \KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, fishstock, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(reps,method), }\DataTypeTok{color=}\NormalTok{method), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) +}
-  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{)))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/Figure3.pdf}
-\caption{Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under
-the policies derived from each of the estimated models, as well as the
-policy based on the exact underlying model.}
-\end{figure}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) + }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{variable)) + }
-  \KeywordTok{facet_wrap}\NormalTok{(~variable, }\DataTypeTok{scales =} \StringTok{"free_y"}\NormalTok{)  + }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{) +}
-  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{) + }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/Figure41.pdf}
-\caption{Histograms of the realized net present value of the fishery
-over a range of simulated data and resulting parameter estimates. For
-each data set, the three models are estimated as described above. Values
-plotted are the averages of a given policy over 100 replicate
-simulations. Details and code provided in the supplement.}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-
-\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) + }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{variable), }\DataTypeTok{binwidth=}\FloatTok{0.1}\NormalTok{) + }
-  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/Figure42.pdf}
-\caption{Histograms of the realized net present value of the fishery
-over a range of simulated data and resulting parameter estimates. For
-each data set, the three models are estimated as described above. Values
-plotted are the averages of a given policy over 100 replicate
-simulations. Details and code provided in the supplement.}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-
-\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value, }\DataTypeTok{fill=}\NormalTok{variable, }\DataTypeTok{color=}\NormalTok{variable)) + }
-  \KeywordTok{stat_density}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{y=}\NormalTok{..density..), }\DataTypeTok{position=}\StringTok{"stack"}\NormalTok{, }\DataTypeTok{adjust=}\DecValTok{3}\NormalTok{, }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{9}\NormalTok{) + }
-  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)+ }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/Figure43.pdf}
-\caption{Histograms of the realized net present value of the fishery
-over a range of simulated data and resulting parameter estimates. For
-each data set, the three models are estimated as described above. Values
-plotted are the averages of a given policy over 100 replicate
-simulations. Details and code provided in the supplement.}
-\end{figure}
-
 \section{Discussion}
@@
 existing approaches for adapting the non-parametric Bayesian models
-previously proposed (Munch et al. 2005) to a decision-theoretic
-framework. Adapting a non-parametric approach requires modification of
-existing methods for decision theory. We have illustrated how this might
-be done for a classic stochastic dynamic programming problem, opening
-the door for substantial further research into how these applications
-might be improved.
+previously proposed {[}@Munch2005{]} to a decision-theoretic framework.
+Adapting a non-parametric approach requires modification of existing
+methods for decision theory. We have illustrated how this might be done
+for a classic stochastic dynamic programming problem, opening the door
+for substantial further research into how these applications might be
+improved.
 
 Though simple mechanistically motivated models may be best for the
-latter case (@Cuddington2013; Geritz and Kisdi 2012), such models can be
-not only inaccurate but misleading in quantitative decision making. This
+latter case {[}@Cuddington2013; @Geritz2012{]}, such models can be not
+only inaccurate but misleading in quantitative decision making. This
 arises for two reasons:
@@
 greatly to both of the problems. This concern is particularly acute in
-case of ecological tipping points (Barnosky et al. 2012; Scheffer et al.
-2001), which arise from feedbacks common in ecological systems and can
-lead to a sudden catastrophic transition to an undesirable state. In
-general we do not know where such tipping points are unless we have
-already observed the transition, in which case it is frequently too late
-to respond. Though there may exist early warning signals for certain
-kinds of these transitions that are driven by slow changes (Scheffer et
-al. 2009), we do not know when, where, or how to apply them to the
-decision making context more generally (Boettiger and Hastings 2013).
+case of ecological tipping points {[}@Barnosky2012; @Scheffer2001{]},
+which arise from feedbacks common in ecological systems and can lead to
+a sudden catastrophic transition to an undesirable state. In general we
+do not know where such tipping points are unless we have already
+observed the transition, in which case it is frequently too late to
+respond. Though there may exist early warning signals for certain kinds
+of these transitions that are driven by slow changes
+{[}@Scheffer2009{]}, we do not know when, where, or how to apply them to
+the decision making context more generally {[}@Boettiger2013{]}.
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{gp_assessment_plots[[}\DecValTok{1}\NormalTok{]]}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/appendixplots1.pdf}
-\caption{plot of chunk appendixplots}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{gp_assessment_plots[[}\DecValTok{2}\NormalTok{]]}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/appendixplots2.pdf}
-\caption{plot of chunk appendixplots}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot_allen_traces}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/appendixplots3.pdf}
-\caption{plot of chunk appendixplots}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot_allen_posteriors}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/appendixplots4.pdf}
-\caption{plot of chunk appendixplots}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot_ricker_traces}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/appendixplots5.pdf}
-\caption{plot of chunk appendixplots}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot_ricker_posteriors}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/appendixplots6.pdf}
-\caption{plot of chunk appendixplots}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot_myers_traces}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/appendixplots7.pdf}
-\caption{plot of chunk appendixplots}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot_myers_posteriors}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/appendixplots8.pdf}
-\caption{plot of chunk appendixplots}
-\end{figure}
-
 \section{References}
 
-Allen, Linda J. S., Jesse F. Fagan, Göran Högnäs, and Henrik Fagerholm.
-2005. ``Population extinction in discrete-time stochastic population
-models with an Allee effect.'' \emph{Journal of Difference Equations and
-Applications} 11 (4-5) (apr): 273--293.
-doi:10.1080/10236190412331335373.
-\url{http://www.tandfonline.com/doi/abs/10.1080/10236190412331335373}.
-
-Athanassoglou, Stergios, and Anastasios Xepapadeas. 2012. ``Pollution
-control with uncertain stock dynamics: When, and how, to be
-precautious.'' \emph{Journal of Environmental Economics and Management}
-63 (3) (may): 304--320. doi:10.1016/j.jeem.2011.11.001.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069611001409}.
-
-Barnosky, Anthony D., Elizabeth a Hadly, Jordi Bascompte, Eric L.
-Berlow, James H. Brown, Mikael Fortelius, Wayne M. Getz, et al. 2012.
-``Approaching a state shift in Earth's biosphere.'' \emph{Nature} 486
-(7401) (jun): 52--58. doi:10.1038/nature11018.
-\url{http://www.nature.com/doifinder/10.1038/nature11018}.
-
-Bestelmeyer, Brandon T., Michael C. Duniway, Darren K. James, Laura M.
-Burkett, and Kris M. Havstad. 2012. ``A test of critical thresholds and
-their indicators in a desertification-prone ecosystem: more resilience
-than we thought.'' Ed. Katharine Suding. \emph{Ecology Letters} (dec).
-doi:10.1111/ele.12045. \url{http://doi.wiley.com/10.1111/ele.12045}.
-
-Boettiger, Carl, and Alan Hastings. 2013. ``Tipping points: From
-patterns to predictions.'' \emph{Nature} 493 (7431) (jan): 157--158.
-doi:10.1038/493157a.
-\url{http://www.nature.com/doifinder/10.1038/493157a}.
-
-Brozović, Nicholas, and Wolfram Schlenker. 2011. ``Optimal management of
-an ecosystem with an unknown threshold.'' \emph{Ecological Economics}
-(jan): 1--14. doi:10.1016/j.ecolecon.2010.10.001.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0921800910004167}.
-
-Courchamp, Franck, Ludek Berec, and Joanna Gascoigne. 2008. \emph{Allee
-Effects in Ecology and Conservation}. Oxford University Press, USA.
-\url{http://www.amazon.com/Effects-Ecology-Conservation-Franck-Courchamp/dp/0198570309}.
-
-Deisenroth, Marc Peter, Carl Edward Rasmussen, and Jan Peters. 2009.
-``Gaussian process dynamic programming.'' \emph{Neurocomputing} 72 (7-9)
-(mar): 1508--1524. doi:10.1016/j.neucom.2008.12.019.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0925231209000162}.
-
-Geritz, Stefan A. H., and Eva Kisdi. 2012. ``Mathematical ecology: why
-mechanistic models?.'' \emph{Journal of mathematical biology} 65 (6-7)
-(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.
-
-Gordon, H. S. 1954. ``The economic theory of a common-property resource:
-the fishery.'' \emph{The Journal of Political Economy} 62 (2): 124--142.
-\url{http://www.jstor.org/stable/10.2307/1825571}.
-
-Hilborn, Ray. 2007. ``Reinterpreting the State of Fisheries and their
-Management.'' \emph{Ecosystems} 10 (8) (oct): 1362--1369.
-doi:10.1007/s10021-007-9100-5.
-\url{http://www.springerlink.com/index/10.1007/s10021-007-9100-5}.
-
-Hughes, Terry P., Cristina Linares, Vasilis Dakos, Ingrid a van de
-Leemput, and Egbert H. van Nes. 2013. ``Living dangerously on borrowed
-time during slow, unrecognized regime shifts.'' \emph{Trends in ecology
-\& evolution} 28 (3) (mar): 149--55. doi:10.1016/j.tree.2012.08.022.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/22995893}.
-
-Kocijan, Juš, Agathe Girard, Bla\textbackslash{}vz Banko, and Roderick
-Murray-Smith. 2005. ``Dynamic systems identification with Gaussian
-processes.'' \emph{Mathematical and Computer Modelling of Dynamical
-Systems} 11 (4) (dec): 411--424. doi:10.1080/13873950500068567.
-\url{http://www.tandfonline.com/doi/abs/10.1080/13873950500068567}.
-
-Ludwig, Donald, and Carl J. Walters. 1982. ``Optimal harvesting with
-imprecise parameter estimates.'' \emph{Ecological Modelling} 14 (3-4)
-(jan): 273--292. doi:10.1016/0304-3800(82)90023-0.
-\url{http://linkinghub.elsevier.com/retrieve/pii/0304380082900230}.
-
-Mangel, Marc. 1985. ``Decision and control in uncertain resource
-systems:'' 255.
-\href{http://www.amazon.com/Decision-uncertain-resource-Mathematics-Engineering/dp/0124687202 http://dl.acm.org/citation.cfm?id=537497}{http://www.amazon.com/Decision-uncertain-resource-Mathematics-Engineering/dp/0124687202
-http://dl.acm.org/citation.cfm?id=537497}.
-
-Mangel, Marc, and Colin W. Clark. 1988. \emph{Dynamic Modeling in
-Behavioral Ecology}. Ed. John Krebs and Tim Clutton-Brock. Princeton:
-Princeton University Press.
-
-May, Robert M., John R. Beddington, Colin W. Clark, Sidney J. Holt, and
-R. M. Laws. 1979. ``Management of multispecies fisheries.''
-\emph{Science (New York, N.Y.)} 205 (4403) (jul): 267--77.
-doi:10.1126/science.205.4403.267.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/17747032}.
-
-Munch, Stephan B., Melissa L. Snover, George M. Watters, and Marc
-Mangel. 2005. ``A unified treatment of top-down and bottom-up control of
-reproduction in populations.'' \emph{Ecology Letters} 8 (7) (may):
-691--695. doi:10.1111/j.1461-0248.2005.00766.x.
-\url{http://doi.wiley.com/10.1111/j.1461-0248.2005.00766.x}.
-
-Rasmussen, Carl Edward, and C. K. I. Williams. 2006. \emph{Gaussian
-Processes for Machine Learning}. Ed. Thomas Dietterich. Boston: MIT
-Press,. \url{www.GaussianProcess.org/gpml}.
-
-Reed, William J. 1979. ``Optimal escapement levels in stochastic and
-deterministic harvesting models.'' \emph{Journal of Environmental
-Economics and Management} 6 (4) (dec): 350--363.
-doi:10.1016/0095-0696(79)90014-7.
-\href{http://www.sciencedirect.com/science/article/pii/0095069679900147 http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}{http://www.sciencedirect.com/science/article/pii/0095069679900147
-http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}.
-
-Schapaugh, Adam W., and Andrew J. Tyre. 2013. ``Accounting for
-parametric uncertainty in Markov decision processes.'' \emph{Ecological
-Modelling} 254 (apr): 15--21. doi:10.1016/j.ecolmodel.2013.01.003.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0304380013000306}.
-
-Scheffer, Marten, Jordi Bascompte, William A. Brock, Victor Brovkin,
-Stephen R. Carpenter, Vasilis Dakos, Hermann Held, Egbert H. van Nes,
-Max Rietkerk, and George Sugihara. 2009. ``Early-warning signals for
-critical transitions.'' \emph{Nature} 461 (7260): 53--9.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/19727193}.
-
-Scheffer, Marten, Stephen R. Carpenter, J. A. Foley, C. Folke, and B.
-Walker. 2001. ``Catastrophic shifts in ecosystems.'' \emph{Nature} 413
-(6856) (oct): 591--6. doi:10.1038/35098000.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/11595939}.
-
-Williams, Byron K. 2001. ``Uncertainty , learning , and the optimal
-management of wildlife.'' \emph{Environmental and Ecological Statistics}
-8: 269--288. doi:10.1023/A:1011395725123.
-
-Worm, Boris, Edward B. Barbier, Nicola Beaumont, J. Emmett Duffy, Carl
-Folke, Benjamin S. Halpern, Jeremy B. C. Jackson, et al. 2006. ``Impacts
-of biodiversity loss on ocean ecosystem services.'' \emph{Science (New
-York, N.Y.)} 314 (5800) (nov): 787--90. doi:10.1126/science.1132294.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/17082450}.
-
-Worm, Boris, Ray Hilborn, Julia K. Baum, Trevor A. Branch, Jeremy S.
-Collie, Christopher Costello, Michael J. Fogarty, et al. 2009.
-``Rebuilding global fisheries.'' \emph{Science (New York, N.Y.)} 325
-(5940) (jul): 578--85. doi:10.1126/science.1173146.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/19644114}.
-
 \end{document}

### 2013-08-26 inst/examples/scaling.Rmd
Makefile includes target. notes from exploring scaling

```{r, include=FALSE}
## @knitr plotting-options
source("~/.knitr_defaults.R")
opts_knit$set(upload.fun = socialR::flickr.url)
library(knitcitations)
library(nonparametricbayes) 
opts_chunk$set(cache=FALSE)
```

```{r}
f <- RickerAllee
p <- c(.5, .8, .5)
K <- .8  # approx, a li'l' less
allee <- .5 # approx, a li'l' less
```

```{r}
sigma_g <- 0.005
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1
x_grid <- seq(0, 1.5 * K, length=50)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 50  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
Xo <-  allee+.2# observations start from
x0 <- K # simulation under policy starts from
Tobs <- 40
MaxT = 1000 # timeout for value iteration convergence
```

```{r}
set.seed(1234)
x <- numeric(Tobs)
x[1] <- Xo
nz <- 1
for(t in 1:(Tobs-1))
  x[t+1] = z_g() * f(x[t], h=0, p=p)
obs <- data.frame(x = c(rep(0,nz), 
                        pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
                  y = c(rep(0,nz), 
                        x[2:Tobs]))
raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
raw_plot
```

```{r}
set.seed(12345)
estf <- function(p){ 
  mu <- f(obs$x,0,p)
  -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
}
par <- c(p[1]*rlnorm(1,0,.1), 
         p[2]*rlnorm(1,0,.1), 
         p[3]*rlnorm(1,0, .1), 
         sigma_g * rlnorm(1,0,.1))
o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
f_alt <- f
p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
sigma_g_alt <- as.numeric(o$par[4])

est <- list(f = f_alt, p = p_alt, sigma_g = sigma_g_alt, mloglik=o$value)

## @knitr mle-output
true_means <- sapply(x_grid, f, 0, p)
est_means <- sapply(x_grid, est$f, 0, est$p)

## @knitr gp-priors
#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
s2.p <- c(5,5)  
d.p <- c(1,1) # = c(10, 1/0.1)
```

```{r}
gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)
```

```{r}
## @knitr gp_traces_densities
gp_assessment_plots <- summary_gp_mcmc(gp, burnin=1e4, thin=300)
gp_assessment_plots
```

```{r}
## @knitr gp-output 
tgp_dat <- data.frame(x = x_grid, 
               y = gp_dat$E_Ef, 
               ymin = pmax(gp_dat$E_Ef - 2 * sqrt(gp_dat$E_Vf),0), 
               ymax = gp_dat$E_Ef + 2 * sqrt(gp_dat$E_Vf) )

## @knitr assemble-models
models <- data.frame(x=x_grid, 
                     GP=tgp_dat$y, 
                     True=true_means)
models <- melt(models, id="x")
names(models) <- c("x", "method", "value")

## @knitr Figure1
plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(data=models, aes(x, value, col=method), lwd=1, alpha=0.8) + 
  geom_point(data=obs, aes(x,y), alpha=0.8) + 
  xlab(expression(X[t])) + ylab(expression(X[t+1])) 
print(plot_gp)
```

```{r}
dat_fig1c <- melt(lapply(sample(100), 
  function(i){
    data.frame(stock = x_grid, 
                GP = pmax( # non-negative curves only
                          mvrnorm(1, 
                                  gp_dat$Ef_posterior[,i], 
                                  gp_dat$E_Cf # could draw from Cf_posterior after assuring is pos-def
                                  )
                          ,0)
               )

  }), id="stock")

ggplot(dat_fig1c, aes(stock, value, group=L1)) + geom_line(alpha=.3)

# a <- dcast(dat_fig1c, stock ~ L1, value.var='value')
# dat_fig1c <- data.table(dat_fig1c)
# to_grid <- function(x, grid) sapply(x, function(x) which.min(abs(grid - x)))  
# dat_fig1c$stock <- to_grid(dat_fig1c$stock, x_grid)
# a <- dat_fig1c[, to_grid(value, x_grid), by=c('stock')]
```

```{r}
matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid) 
opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta)

opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 5, xT=0.5, profit, delta, reward=1)
opt_gp$D <- opt_gp$D[,1]

## @knitr mle-opt
matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=.01)
```

```{r}
OPT = data.frame(GP = opt_gp$D, True = opt_true$D)

## @knitr Figure2
policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
names(policies) <- c("stock", "method", "value")

ggplot(policies, aes(stock, stock - value, color=method)) +
  geom_line(lwd=1.2, alpha=0.8) + xlab("stock size") + ylab("escapement") 
```

```{r}
## @knitr sims
sims <- lapply(OPT, function(D){
  set.seed(1)
  lapply(1:100, function(i) 
    ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
  )
})

dat <- melt(sims, id=names(sims[[1]][[1]]))
sims_data <- data.table(dat)
setnames(sims_data, c("L1", "L2"), c("method", "reps")) 
# Legend in original ordering please, not alphabetical: 
sims_data$method = factor(sims_data$method, ordered=TRUE, levels=names(OPT))

## @knitr Figure3
ggplot(sims_data) + 
  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) 
```

```{r}
## @knitr profits
Profit <- sims_data[, sum(profit), by=c("reps", "method")]
tmp <- dcast(Profit, reps ~ method)
tmp <- tmp / tmp[,"True"]
tmp <- melt(tmp[2:dim(tmp)[2]])
actual_over_optimal <-subset(tmp, variable != "True")

## @knitr Figure4 
ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable)) + 
  facet_wrap(~variable, scales = "free_y")  + guides(legend.position = "none") +
  xlab("Total profit by replicate")
```

# Forecasts

## One-step ahead predictor plots and forecasting plots

Now we show how each of the fitted models performs on the training data (e.g. plot of the step-ahead predictors). For the GP, we need to predict on the training data first:

```{r}
gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
```

For the parametric models, prediction is a matter of feeding in each of the observed data points to the fitted parameters, like so

```{r}
step_ahead <- function(x, f, p){
  h = 0
  x_predict <- sapply(x, f, h, p)
  n <- length(x_predict) - 1
  y <- c(x[1], x_predict[1:n])
  y
}
```

Which we apply over each of the fitted models, including the GP, organizing the "expected" transition points (given the previous point) into a data frame.

```{r}
df <- melt(data.frame(time = 1:length(x), stock = x, 
                GP = gp_f_at_obs$E_Ef,
                True = step_ahead(x,f,p)
                 ), id=c("time", "stock"))

ggplot(df) + geom_point(aes(time, stock)) + 
  geom_line(aes(time, value, col=variable)) 
```

## Posterior predictive curves

This shows only the mean predictions.  For the Bayesian cases, we can instead loop over the posteriors of the parameters (or samples from the GP posterior) to get the distribution of such curves in each case.

We will need a vector version (`pmin` in place of `min`) of this function that can operate on the posteriors, others are vectorized already.

```{r}
ricker_f <- function(x,h,p){
  sapply(x, function(x){ 
    x <- pmax(0, x-h) 
    pmax(0, x * exp(p[2] * (1 - x / p[1] )) )
  })
}
```

Then we proceed as before, now looping over 100 random samples from the posterior for each Bayesian estimate.  We write this as a function for easy reuse.

```{r}
require(MASS)

step_ahead_posteriors <- function(x){
gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
df_post <- melt(lapply(sample(100), 
  function(i){
    data.frame(time = 1:length(x), stock = x, 
                GP = pmax(mvrnorm(1, 
                                  gp_f_at_obs$Ef_posterior[,i], 
                                  pmax(gp_f_at_obs$E_Cf, 1e-4)
                                  #pmax(gp_f_at_obs$Cf_posterior[[i]],1e-5)
                                  ),0),
                True = step_ahead(x,f,p))
  }), id=c("time", "stock"))

}

df_post <- step_ahead_posteriors(x)

ggplot(df_post) + geom_point(aes(time, stock)) + 
  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) 
```

## Performance on data outside of observations

Of course it is hardly suprising that all models do reasonably well on the data on which they were trained.  A crux of the problem is the model performance on data outside the observed range.  (Though we might also wish to repeat the above plot on data in the observed range but a different sequence from the observed data).

First we generate some data from the underlying model coming from below the tipping point:

```{r}
Tobs <- 30
y <- numeric(Tobs)
y[1] = .45
for(t in 1:(Tobs-1))
    y[t+1] = z_g() * f(y[t], h=0, p=p)
```

Proceed as before on this data:

```{r}
crash_data <- step_ahead_posteriors(y)

ggplot(crash_data) + geom_point(aes(time, stock)) + 
  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
  facet_wrap(~variable)

```

## Forecast Distributions

Another way to visualize this is to look directly at the distribution predicted under each model, one step and several steps into the future (say, at a fixed harvest level).

```{r}
library(expm)
get_forecasts <- function(i, Tobs, h_i){

df <- data.frame(
  x = x_grid,  
  GP = matrices_gp[[h_i]][i,], 
  True = matrices_true[[h_i]][i,])

df2 <- data.frame(
  x = x_grid,
  GP = (matrices_gp[[h_i]] %^% Tobs)[i,], 
  True = (matrices_true[[h_i]] %^% Tobs)[i,])

T2 <- 2 * Tobs

df4 <- data.frame(
  x = x_grid,
  GP = (matrices_gp[[h_i]] %^% T2)[i,], 
  True = (matrices_true[[h_i]] %^% T2)[i,])

forecasts <- melt(list(T_start = df, T_mid = df2, T_end = df4), id="x")

}

i = 15
```

This takes `i`, an index to `x_grid` value (e.g. for `i`=`r i` corresponds to a starting postion `x` = `r x_grid[i]`)

```{r}
forecasts <- get_forecasts(i = 15, Tobs = 15, h_i = 1)
ggplot(forecasts) +
  geom_line(aes(x, value, group=interaction(variable, L1), col=variable, lty=L1)) +
  facet_wrap(~ variable, scale="free_y", ncol=2) 
```

We can compare to a better starting stock,

```{r} 
i<-45
```

where `i`=`r i` corresponds to a starting postion `x` = `r x_grid[i]`

```{r}
forecasts <- get_forecasts(i = i, Tobs = 10, h_i = 1)

ggplot(forecasts) +
  geom_line(aes(x, value, group=interaction(variable, L1), col=variable, lty=L1)) +
  facet_wrap(~ variable, scale="free_y", ncol=2) 
```

and a middle value showing the population decline:

```{r} 
i<-35
```

where `i`=`r i` corresponds to a starting postion `x` = `r x_grid[i]`

```{r}
forecasts <- get_forecasts(i = i, Tobs = 10, h_i = 1)

ggplot(forecasts) +
  geom_line(aes(x, value, group=interaction(variable, L1), col=variable, lty=L1)) +
  facet_wrap(~ variable, scale="free_y", ncol=2) 
```

### 2013-09-06 inst/doc/nonparametric-bayes.Rmd
Updated nonparametric-bayes.Rmd

@@
 
+We address
+
 Decision-theoretic methods often rely on simple parametric models of
@@
 
+Further, the data on which a model has been parameterized frequently 
+fails to cover the possible state-space over which management decisions 
+must operate.  Consequently a model do well in the region of state-space
+in which it was estimated, but give erroneous confidence to predictions
+outside of that region.
+
+This problem is keenly felt in any system in which a manager may fear
+crossing a threshold, or tipping point. The tipping point, if it exists 
+at all, will lay outside the observed range of the observed data.
+
 We demonstrate how nonparametric Bayesian models can provide robust,
-nearly-optimal solutions to decision making under uncertainty when
-_we don't know the correct model_ to use.
+solutions to decision making under uncertainty without knowing the 
+structural form of the true model.
 
-While methods that account
-for _parametric_ uncertainty can be very successful with the right model,
+<!--
+While methods that account for _parametric_ uncertainty can be very 
+successful with the right model,
 structural uncertainty of not knowing what model best approximates the 
 dynamics poses considerably greater difficulty.  
-
-Non-parametric Bayesian
-methods offer a promising statistical approach for predictive modeling
-of ecological dynamics in regions of state space where the data is
-adequate, while at the same time offering more flexible patterns with
-greater uncertainty outside the observed data. This contrasts from simple
-parametric models which provide relatively constant level of uncertainty
-in regions with and without adequate data. The consequence of such
-misplaced confidence outside the data can lead to highly undesirable
-results that may be avoided with the more flexible non-parametric
-Bayesian approach.
+-->
 

### 2013-09-06 inst/doc/nonparametric-bayes.Rmd
Updated nonparametric-bayes.Rmd

@@
 
-This problem is keenly felt in any system in which a manager may fear
-crossing a threshold, or tipping point. The tipping point, if it exists 
+This problem is keenly felt in any system where a potential threshold
+or tipping point is a concern.  Such a tipping point, if it exists 
 at all, will lay outside the observed range of the observed data.
@@
 resource management and conservation.  
-Here we illustrate how a stochastic dynamic programming algorithm can 
-be driven by the predictions from a Gaussian process model, sidestepping
-the need for an accurate model-based description of the system dynamics.
+Here we illustrate how a stochastic dynamic programming (SDP) algorithm can 
+be driven by the predictions from a Gaussian Process (GP).
+This provides two distinct advantages compared with contemporary approaches.
+First, using a GP sidesteps the need for an accurate model-based description
+of the system dynamics (or a model choice or model-averaging approach which
+introduces yet more difficulty).  
+Second, unlike parametric model-fitting approaches, which can only reflect 
+uncertainty in the corresponding confidence intervals or  posterior distributions
+of the parameters, the GP approach provides a state-space dependent representation
+of uncertainty.  In other words, a GP estimated from a given data set permits 
+a much greater uncertainty far from the observed data than it does near the observed 
+data.
+
+<!-- Tipping points and the frequency of a lack of data throughout state-space -->
+
+The idea that any approach can perform well without either having to know the model or
+have particularly good data should immediately draw suspicion.  The reader must bear
+in mind that the strength of our approach comes not from some magic predictive power
+from such limited information, but rather, by providing a more honest expression of 
+uncertainty outside the observed data without sacrificing the predictive capacity 
+near the observed data.
+
+By coupling this more accurate description of what is known and unknown to the 
+decision-making under uncertainty framework provided by stochastic dynamic programming,
+we are able to obtain more robust management policies than with common parameteric modeling 
+approaches.
 

### 2013-09-07 inst/doc/nonparametric-bayes.Rmd
Updated nonparametric-bayes.Rmd

@@
 
-<!-- Tipping points and the frequency of a lack of data throughout state-space -->
+<!-- More on complex dynamics and not having the correct model
+
+Cite [@Glaser2013](http://doi.org/10.1111/faf.12037 "Complex dynamics may limit prediction in marine fisheries")
+etc
+
+also older stuff on model uncertainty
+
+-->
+
+<!-- More on the lack of data throughout the relevant state-space
+     and how any concern about potential tipping points indicates 
+     that the data exhibits this bias / problem.
+
+     -->
 
@@
 
-add references to [@Fischer2009, @Polasky2011] as persective on combining complex / critical dynamics + control theory
-
 #### Models for decision-making under uncertainty
 
-Decision-theoretic (or optimal control) tools require a model that can assign probabilities of
+Decision-theoretic or optimal control tools require a model that can assign probabilities of
 future states (e.g. stock size of a fishery) given the current state and a
-proposed action (e.g. fishing harvest or effort).  The decision maker then
-seeks to determining the course of actions (also referred to as the policy)
-that maximizes the expected value of some objective function (such as
-net present value derived from the resource over time.  (The approach
-can be adapted to permit alternatives to maximizing expectation, such
-as minimizing the maximum cost or damage that might be incurred. See
-@Polasky2011).
+proposed action (e.g. fishing harvest or effort).  
 Management frequently faces a sequential decision-making problem -- after 
 selecting an action, the decision-maker may receive new information about 
-the current state and must again choose an appropriate action -- i.e. setting
+the current state and must again choose an appropriate action -- such as setting
 the harvest limits each year based on stock assessments the year prior.
 
-More historical context, e.g. [@Reed1979, @Mangel1985]
+The decision maker typically seeks to determining the course of actions (also referred to as the policy)
+that maximizes the expected value of some objective function such as
+net present value derived from the resource over time.  
+Though much can be said on how to choose this value function appropriately 
+(e.g. see [@Halpern2013](http://doi.org/10.1073/pnas.1217689110 
+"Achieving the triple bottom line in the face of inherent trade-offs 
+among social equity, economic return, and conservation.")) we will
+assume this is given.  (Nor is this approach necessarily constrained to 
+maximizing the expectated value of such a function - the decision-theoretic
+framework can be adapted to alternatives such as minimizing the maximum 
+cost or damage that might be incurred; see @Polasky2011).
 
-#### Parametric models
+In representing future states with probabilities and maximizing expectations,
+this approach provides a natural framework for handling uncertainty.
 
-Traditional approaches to optimal control (Pontryagin's principle, stochastic
-dynamic programming) rely on knowledge of the state equation.  Such simple 
-parametric models, frequently justified by mechanistic underpinnings are used
-to provide probabilities of future states.  Fisheries literature, which enjoys
-some of the longest history in the use of the optimal control framework, provides
-a clear example [ ] . Models such as the Ricker or Beverton Holt curves are ubiquitous [too strong? references?]
-Parametric models can reflect variability introduced by a stochastic environment
-or demographic process my incorporating random shocks from a given distribution.
+The value function typically depends on the action or policy taken, as well as
+the state of the system, in each interval of time. The state of the system,
+in turn, is usually described by a dynamical model.
 

### 2013-09-08 inst/doc/nonparametric-bayes.Rmd
Updated nonparametric-bayes.Rmd
update chunk options

--- 2013-09-06 inst/doc/nonparametric-bayes.Rmd
@@
 
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=1, 
+               comment=NA, verbose=TRUE, fig.width=6, fig.height=4)
+
+# Name the cache path and fig.path based on filename...
+opts_chunk$set(fig.path = paste("figure/",
+                                gsub(".Rmd", "", knitr:::knit_concord$get('infile')),
+                                "-", sep=""),
+               cache.path = paste("cache/", 
+                                  gsub(".Rmd", "", knitr:::knit_concord$get('infile') ), 
+                                "/", sep=""))
+
+toggle = "markup" #"hide"  # toggles `results` chunk option
+
 library(ggplot2) 
+theme_set(theme_bw(base_size=12))
 cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
                "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-theme_set(theme_bw(base_size=12))
-toggle = "hide" 
 ```
@@
 
-We address
-
 Decision-theoretic methods often rely on simple parametric models of
@@
 
-Further, the data on which a model has been parameterized frequently 
-fails to cover the possible state-space over which management decisions 
-must operate.  Consequently a model do well in the region of state-space
-in which it was estimated, but give erroneous confidence to predictions
-outside of that region.
+We demonstrate how nonparametric Bayesian models can provide robust,
+nearly-optimal solutions to decision making under uncertainty when
+_we don't know the correct model_ to use.
 
-This problem is keenly felt in any system in which a manager may fear
-crossing a threshold, or tipping point. The tipping point, if it exists 
-at all, will lay outside the observed range of the observed data.
-
-We demonstrate how nonparametric Bayesian models can provide robust,
-solutions to decision making under uncertainty without knowing the 
-structural form of the true model.
-
-<!--
-While methods that account for _parametric_ uncertainty can be very 
-successful with the right model,
+While methods that account
+for _parametric_ uncertainty can be very successful with the right model,
 structural uncertainty of not knowing what model best approximates the 
 dynamics poses considerably greater difficulty.
--->
+
+Non-parametric Bayesian
+methods offer a promising statistical approach for predictive modeling
+of ecological dynamics in regions of state space where the data is
+adequate, while at the same time offering more flexible patterns with
+greater uncertainty outside the observed data. This contrasts from simple
+parametric models which provide relatively constant level of uncertainty
+in regions with and without adequate data. The consequence of such
+misplaced confidence outside the data can lead to highly undesirable
+results that may be avoided with the more flexible non-parametric
+Bayesian approach.
 
@@
 ```
-```{r Figureb, dependson=c("assemble-models", "par-fns"), toggle = 1}
+```{r Figureb, dependson=c("assemble-models", "par-fns"), include=TRUE}
 ```
@@
 ```{r Figure2, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", toggle = 1}
+Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", include=TRUE}
 ```
@@
 ```{r Figure3, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", toggle = 1}
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.",  include=TRUE}
 ```
@@
 Values plotted are the averages of a given policy over 100 replicate simulations.
-Details and code provided in the supplement.", toggle = 1}
+Details and code provided in the supplement.",  include=TRUE}
 ```

### 2013-09-08 inst/doc/nonparametric-bayes.tex
update chunk options

@@
 \newcommand{\NormalTok}[1]{{#1}}
+\usepackage{graphicx}
+% We will generate all images so they have a width \maxwidth. This means
+% that they will get their normal width if they fit onto the page, but
+% are scaled down if they would overflow the margins.
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
+\else\Gin@nat@width\fi}
+\makeatother
+\let\Oldincludegraphics\includegraphics
+\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
 \ifxetex
@@
 
-add references to {[}@Fischer2009, @Polasky2011{]} as persective on
+add references to (@Fischer2009, Polasky et al. 2011) as persective on
 combining complex / critical dynamics + control theory
@@
 alternatives to maximizing expectation, such as minimizing the maximum
-cost or damage that might be incurred. See @Polasky2011). Management
-frequently faces a sequential decision-making problem -- after selecting
-an action, the decision-maker may receive new information about the
-current state and must again choose an appropriate action --
-i.e.~setting the harvest limits each year based on stock assessments the
-year prior.
+cost or damage that might be incurred. See Polasky et al. (2011)).
+Management frequently faces a sequential decision-making problem --
+after selecting an action, the decision-maker may receive new
+information about the current state and must again choose an appropriate
+action -- i.e.~setting the harvest limits each year based on stock
+assessments the year prior.
 
-More historical context, e.g. {[}@Reed1979, @Mangel1985{]}
+More historical context, e.g. (Reed 1979, )
 
@@
 
-{[}@Mangel1985; @Schapaugh2013{]}
+(Mangel 1985; Schapaugh and Tyre 2013)
 
@@
 
-{[}@Williams2001; @Athanassoglou2012{]}.
+(B. K. Williams 2001; Athanassoglou and Xepapadeas 2012).
 
@@
 social value of fish left in the sea vs.~the commercial value of fish
-harvested; see @Halpern2013 for further discussion of this issue.
-Alternatively, we can always compare model performance across multiple
-potential objective functions.) The decision-maker does not necessarily
-need a model that provides the best mechanistic understanding or the
-best long-term outcome, but rather the one that best estimates the
-probabilities of being in different states as a result of the possible
-actions.
+harvested; see Halpern et al. (2013) for further discussion of this
+issue. Alternatively, we can always compare model performance across
+multiple potential objective functions.) The decision-maker does not
+necessarily need a model that provides the best mechanistic
+understanding or the best long-term outcome, but rather the one that
+best estimates the probabilities of being in different states as a
+result of the possible actions.
 
@@
 before. The economic value and ecological concern have made marine
-fisheries the crucible for much of the founding work {[}@Gordon1954;
-@Reed1979; @May1979; @Ludwig1982{]} in managing ecosystems under
-uncertainty. Global trends {[}@Worm2006{]} and controversy
-{[}@Hilborn2007; @Worm2009{]} have made understanding these challenges
-all the more pressing.
+fisheries the crucible for much of the founding work (Gordon 1954; Reed
+1979; May et al. 1979; Ludwig and Walters 1982) in managing ecosystems
+under uncertainty. Global trends (Worm et al. 2006) and controversy
+(Hilborn 2007; Worm et al. 2009) have made understanding these
+challenges all the more pressing.
 
@@
 discretized values of stock $X$ and policy $h$ using stochastic dynamic
-programming (SDP) {[}@Mangel1985{]}. Problems of this sort underpin much
+programming (SDP) (Mangel 1985). Problems of this sort underpin much
 marine fisheries management today.
@@
 estimated and integrated over to determine the optimal policy under
-under uncertainty {[}@Mangel1985; @Schapaugh2013{]}. Uncertainty in the
-model structure itself can only be addressed in this approach by
+under uncertainty (Mangel 1985; Schapaugh and Tyre 2013). Uncertainty in
+the model structure itself can only be addressed in this approach by
 hypothesizing alternative model structures, and then performing some
-model choice or model averaging {[}@Williams2001; @Athanassoglou2012{]}.
+model choice or model averaging (B. K. Williams 2001; Athanassoglou and
+Xepapadeas 2012).
 
@@
 Concerns over the potential for tipping points in ecological dynamics
-{[}@Scheffer2001{]} highlight the dangers of uncertainty in ecological
-management and pose a substantial challenge to existing
-decision-theoretic approaches {[}@Brozovic2011{]}. Because intervention
-is often too late after a tipping point has been crossed (but see
-@Hughes2013), management is most often concerned with avoiding
-potentially catastrophic tipping points before any data is available at
-or following a transition that would more clearly reveal these regime
-shift dynamics {[}e.g. @Bestelmeyer2012{]}.
+(Scheffer et al. 2001) highlight the dangers of uncertainty in
+ecological management and pose a substantial challenge to existing
+decision-theoretic approaches (Brozović and Schlenker 2011). Because
+intervention is often too late after a tipping point has been crossed
+(but see Hughes et al. (2013)), management is most often concerned with
+avoiding potentially catastrophic tipping points before any data is
+available at or following a transition that would more clearly reveal
+these regime shift dynamics (e.g. Bestelmeyer et al. 2012).
 
@@
 parametric model.\\We will focus on a simple parametric model for a
-single species {[}derived from fist principles by @Allen2005a{]} as our
-underlying ``reality''.
+single species (derived from fist principles by Allen et al. 2005) as
+our underlying ``reality''.
 
@@
 which the population is not self-sustaining and shrinks to zero
-{[}@Courchamp2008{]}.
+(Courchamp, Berec, and Gascoigne 2008).
 
@@
 recruitment curves {[}@{]} or demonstrating the existence of a tipping
-point {[}@Scheffer2001{]}.\\Using data simulated from a specified model
-rather than empirical data permits the comparison against the true
+point (Scheffer et al. 2001).\\Using data simulated from a specified
+model rather than empirical data permits the comparison against the true
 underlying dynamics, setting a bar for the optimal performance possible.
@@
 geospatial literature) to formulate a predictive model is relatively new
-in the context of modeling dynamical systems {[}@Kocijan2005{]} and
-introduced in the ecological modeling and fisheries management by
-@Munch2005. An accessible and thorough introduction to the formulation
-and use of GPs can be found in @Rasmussen2006.
+in the context of modeling dynamical systems (Kocijan et al. 2005) and
+introduced in the ecological modeling and fisheries management by Munch
+et al. (2005). An accessible and thorough introduction to the
+formulation and use of GPs can be found in Rasmussen and Williams
+(2006).
 
@@
 process model are estimated by Metropolis-Hastings algorithm, again with
-details and code provided in the Appendix. @Rasmussen2006 provides an
-excellent general introduction to Gaussian Processes and @Munch2005
-first discusses their application in the context of population dynamics
-models such as fisheries stock-recruitment relationships.
+details and code provided in the Appendix. Rasmussen and Williams (2006)
+provides an excellent general introduction to Gaussian Processes and
+Munch et al. (2005) first discusses their application in the context of
+population dynamics models such as fisheries stock-recruitment
+relationships.
 
@@
 
-Once the posterior Gaussian process (GP) has been estimated {[}e.g.~see
-@Munch2005{]}, it is necessary to adapt it in place of the parametric
-equation for the stochastic dynamic programming (SDP) solution {[}see
-@Mangel1988 for a detailed description of parametric SDP methods{]} to
-the optimal policy. The essence of the idea is straight forward -- we
-will use the estimated GP in place of the parametric growth function to
-determine the stochastic transition matrix on which the SDP calculations
-are based. The SDP is solved in a discretized state space -- both the
-continuously valued population densities $X$ and harvest quotas $h$ are
-first mapped to a bounded, discrete grid. (For simplicity we will
-consider a uniform grid, though for either parametric or GP-based SDP it
-is often advantageous to use a non-uniform discretization such as a
-basis function representation, e.g.~see {[}@Deisenroth2009{]}).
+Once the posterior Gaussian process (GP) has been estimated (e.g.~see
+Munch et al. 2005), it is necessary to adapt it in place of the
+parametric equation for the stochastic dynamic programming (SDP)
+solution (see Mangel and Clark 1988 for a detailed description of
+parametric SDP methods) to the optimal policy. The essence of the idea
+is straight forward -- we will use the estimated GP in place of the
+parametric growth function to determine the stochastic transition matrix
+on which the SDP calculations are based. The SDP is solved in a
+discretized state space -- both the continuously valued population
+densities $X$ and harvest quotas $h$ are first mapped to a bounded,
+discrete grid. (For simplicity we will consider a uniform grid, though
+for either parametric or GP-based SDP it is often advantageous to use a
+non-uniform discretization such as a basis function representation,
+e.g.~see (Deisenroth, Rasmussen, and Peters 2009)).
 
@@
 \section{Results}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{require}\NormalTok{(MASS)}
+\NormalTok{step_ahead <- function(x, f, p)\{}
+  \NormalTok{h = }\DecValTok{0}
+  \NormalTok{x_predict <- }\KeywordTok{sapply}\NormalTok{(x, f, h, p)}
+  \NormalTok{n <- }\KeywordTok{length}\NormalTok{(x_predict) - }\DecValTok{1}
+  \NormalTok{y <- }\KeywordTok{c}\NormalTok{(x[}\DecValTok{1}\NormalTok{], x_predict[}\DecValTok{1}\NormalTok{:n])}
+  \NormalTok{y}
+\NormalTok{\}}
+\NormalTok{step_ahead_posteriors <- function(x)\{}
+\NormalTok{gp_f_at_obs <- }\KeywordTok{gp_predict}\NormalTok{(gp, x, }\DataTypeTok{burnin=}\FloatTok{1e4}\NormalTok{, }\DataTypeTok{thin=}\DecValTok{300}\NormalTok{)}
+\NormalTok{df_post <- }\KeywordTok{melt}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(}\KeywordTok{sample}\NormalTok{(}\DecValTok{100}\NormalTok{), }
+  \NormalTok{function(i)\{}
+    \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{time =} \DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(x), }\DataTypeTok{stock =} \NormalTok{x, }
+                \DataTypeTok{GP =} \KeywordTok{mvrnorm}\NormalTok{(}\DecValTok{1}\NormalTok{, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),}
+                \DataTypeTok{True =} \KeywordTok{step_ahead}\NormalTok{(x,f,p),  }
+                \DataTypeTok{MLE =} \KeywordTok{step_ahead}\NormalTok{(x,f,est$p), }
+                \DataTypeTok{Allen =} \KeywordTok{step_ahead}\NormalTok{(x, allen_f, pardist[i,]), }
+                \DataTypeTok{Ricker =} \KeywordTok{step_ahead}\NormalTok{(x, ricker_f, ricker_pardist[i,]), }
+                \DataTypeTok{Myers =} \KeywordTok{step_ahead}\NormalTok{(x, myers_f, myers_pardist[i,]))}
+  \NormalTok{\}), }\DataTypeTok{id=}\KeywordTok{c}\NormalTok{(}\StringTok{"time"}\NormalTok{, }\StringTok{"stock"}\NormalTok{))}
+\NormalTok{\}}
+
+\NormalTok{df_post <- }\KeywordTok{step_ahead_posteriors}\NormalTok{(x)}
+
+\KeywordTok{ggplot}\NormalTok{(df_post) + }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, stock)) + }
+  \KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, value, }\DataTypeTok{col=}\NormalTok{variable, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(L1,variable)), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) + }
+  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{))) }
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-Figureb.pdf}
+\caption{plot of chunk Figureb}
+\end{figure}
 
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{policies <- }\KeywordTok{melt}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{stock=}\NormalTok{x_grid, }\KeywordTok{sapply}\NormalTok{(OPT, function(x) x_grid[x])), }\DataTypeTok{id=}\StringTok{"stock"}\NormalTok{)}
+\KeywordTok{names}\NormalTok{(policies) <- }\KeywordTok{c}\NormalTok{(}\StringTok{"stock"}\NormalTok{, }\StringTok{"method"}\NormalTok{, }\StringTok{"value"}\NormalTok{)}
+
+\KeywordTok{ggplot}\NormalTok{(policies, }\KeywordTok{aes}\NormalTok{(stock, stock - value, }\DataTypeTok{color=}\NormalTok{method)) +}
+  \KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{lwd=}\FloatTok{1.2}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) + }\KeywordTok{xlab}\NormalTok{(}\StringTok{"stock size"}\NormalTok{) + }\KeywordTok{ylab}\NormalTok{(}\StringTok{"escapement"}\NormalTok{)  +}
+  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-Figure2.pdf}
+\caption{The steady-state optimal policy (infinite boundary) calculated
+under each model. Policies are shown in terms of target escapement,
+$S_t$, as under models such as this a constant escapement policy is
+expected to be optimal (Reed 1979).}
+\end{figure}
+
 The resulting optimal management strategy based on each of the inferred
@@
 $S_t$. Under models such as this a constant escapement policy is
-expected to be optimal {[}@Reed1979{]}, whereby population levels below
-a certain size $S$ are unharvested, while above that size the harvest
+expected to be optimal (Reed 1979), whereby population levels below a
+certain size $S$ are unharvested, while above that size the harvest
 strategy aims to return the population to $S$, resulting in the
 hockey-stick shaped policies shown.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{ggplot}\NormalTok{(sims_data) + }
+  \KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, fishstock, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(reps,method), }\DataTypeTok{color=}\NormalTok{method), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) +}
+  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{)))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-Figure3.pdf}
+\caption{Gaussian process inference outperforms parametric estimates.
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under
+the policies derived from each of the estimated models, as well as the
+policy based on the exact underlying model.}
+\end{figure}
 
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{fig4v1 <- }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) + }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{variable)) + }
+  \KeywordTok{facet_wrap}\NormalTok{(~variable, }\DataTypeTok{scales =} \StringTok{"free_y"}\NormalTok{)  + }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{) +}
+  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{) + }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey) }\CommentTok{# density plots fail when delta fn}
+
+\NormalTok{fig4v2 <- }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) + }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{variable), }\DataTypeTok{binwidth=}\FloatTok{0.1}\NormalTok{) + }
+  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
+
+\NormalTok{fig4v3 <- }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value, }\DataTypeTok{fill=}\NormalTok{variable, }\DataTypeTok{color=}\NormalTok{variable)) + }\CommentTok{# density plots fail when delta fn}
+  \KeywordTok{stat_density}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{y=}\NormalTok{..density..), }\DataTypeTok{position=}\StringTok{"stack"}\NormalTok{, }\DataTypeTok{adjust=}\DecValTok{3}\NormalTok{, }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{9}\NormalTok{) + }
+  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)+ }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
+\NormalTok{fig4v2}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-Figure41.pdf}
+\caption{Histograms of the realized net present value of the fishery
+over a range of simulated data and resulting parameter estimates. For
+each data set, the three models are estimated as described above. Values
+plotted are the averages of a given policy over 100 replicate
+simulations. Details and code provided in the supplement.}
+\end{figure}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{fig4v3}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-Figure42.pdf}
+\caption{Histograms of the realized net present value of the fishery
+over a range of simulated data and resulting parameter estimates. For
+each data set, the three models are estimated as described above. Values
+plotted are the averages of a given policy over 100 replicate
+simulations. Details and code provided in the supplement.}
+\end{figure}
+
 \section{Discussion}
@@
 existing approaches for adapting the non-parametric Bayesian models
-previously proposed {[}@Munch2005{]} to a decision-theoretic framework.
-Adapting a non-parametric approach requires modification of existing
-methods for decision theory. We have illustrated how this might be done
-for a classic stochastic dynamic programming problem, opening the door
-for substantial further research into how these applications might be
-improved.
+previously proposed (Munch et al. 2005) to a decision-theoretic
+framework. Adapting a non-parametric approach requires modification of
+existing methods for decision theory. We have illustrated how this might
+be done for a classic stochastic dynamic programming problem, opening
+the door for substantial further research into how these applications
+might be improved.
 
 Though simple mechanistically motivated models may be best for the
-latter case {[}@Cuddington2013; @Geritz2012{]}, such models can be not
-only inaccurate but misleading in quantitative decision making. This
-arises for two reasons:
+latter case (Cuddington, Fortin, and Gerber 2013; Geritz and Kisdi
+2012), such models can be not only inaccurate but misleading in
+quantitative decision making. This arises for two reasons:
 
@@
 greatly to both of the problems. This concern is particularly acute in
-case of ecological tipping points {[}@Barnosky2012; @Scheffer2001{]},
-which arise from feedbacks common in ecological systems and can lead to
-a sudden catastrophic transition to an undesirable state. In general we
-do not know where such tipping points are unless we have already
-observed the transition, in which case it is frequently too late to
-respond. Though there may exist early warning signals for certain kinds
-of these transitions that are driven by slow changes
-{[}@Scheffer2009{]}, we do not know when, where, or how to apply them to
-the decision making context more generally {[}@Boettiger2013{]}.
+case of ecological tipping points (Barnosky et al. 2012; Scheffer et al.
+2001), which arise from feedbacks common in ecological systems and can
+lead to a sudden catastrophic transition to an undesirable state. In
+general we do not know where such tipping points are unless we have
+already observed the transition, in which case it is frequently too late
+to respond. Though there may exist early warning signals for certain
+kinds of these transitions that are driven by slow changes (Scheffer et
+al. 2009), we do not know when, where, or how to apply them to the
+decision making context more generally (Boettiger and Hastings 2013).
 
@@
 
+Allen, Linda J. S., Jesse F. Fagan, Göran Högnäs, and Henrik Fagerholm.
+2005. ``Population extinction in discrete-time stochastic population
+models with an Allee effect.'' \emph{Journal of Difference Equations and
+Applications} 11 (4-5) (apr): 273--293.
+doi:10.1080/10236190412331335373.
+\url{http://www.tandfonline.com/doi/abs/10.1080/10236190412331335373}.
+
+Athanassoglou, Stergios, and Anastasios Xepapadeas. 2012. ``Pollution
+control with uncertain stock dynamics: When, and how, to be
+precautious.'' \emph{Journal of Environmental Economics and Management}
+63 (3) (may): 304--320. doi:10.1016/j.jeem.2011.11.001.
+\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069611001409}.
+
+Barnosky, Anthony D., Elizabeth a Hadly, Jordi Bascompte, Eric L.
+Berlow, James H. Brown, Mikael Fortelius, Wayne M. Getz, et al. 2012.
+``Approaching a state shift in Earth's biosphere.'' \emph{Nature} 486
+(7401) (jun): 52--58. doi:10.1038/nature11018.
+\url{http://www.nature.com/doifinder/10.1038/nature11018}.
+
+Bestelmeyer, Brandon T., Michael C. Duniway, Darren K. James, Laura M.
+Burkett, and Kris M. Havstad. 2012. ``A test of critical thresholds and
+their indicators in a desertification-prone ecosystem: more resilience
+than we thought.'' Ed. Katharine Suding. \emph{Ecology Letters} (dec).
+doi:10.1111/ele.12045. \url{http://doi.wiley.com/10.1111/ele.12045}.
+
+Boettiger, Carl, and Alan Hastings. 2013. ``Tipping points: From
+patterns to predictions.'' \emph{Nature} 493 (7431) (jan): 157--158.
+doi:10.1038/493157a.
+\url{http://www.nature.com/doifinder/10.1038/493157a}.
+
+Brozović, Nicholas, and Wolfram Schlenker. 2011. ``Optimal management of
+an ecosystem with an unknown threshold.'' \emph{Ecological Economics}
+(jan): 1--14. doi:10.1016/j.ecolecon.2010.10.001.
+\url{http://linkinghub.elsevier.com/retrieve/pii/S0921800910004167}.
+
+Courchamp, Franck, Ludek Berec, and Joanna Gascoigne. 2008. \emph{Allee
+Effects in Ecology and Conservation}. Oxford University Press, USA.
+\url{http://www.amazon.com/Effects-Ecology-Conservation-Franck-Courchamp/dp/0198570309}.
+
+Cuddington, Kim M., M. J. Fortin, and L. R. Gerber. 2013.
+``Process-based models are required to manage ecological systems in a
+changing world.'' \emph{Ecosphere} 4 (February): 1--12.
+doi:10.1890/ES12-00178.1.
+\url{http://www.esajournals.org/doi/abs/10.1890/ES12-00178.1}.
+
+Deisenroth, Marc Peter, Carl Edward Rasmussen, and Jan Peters. 2009.
+``Gaussian process dynamic programming.'' \emph{Neurocomputing} 72 (7-9)
+(mar): 1508--1524. doi:10.1016/j.neucom.2008.12.019.
+\url{http://linkinghub.elsevier.com/retrieve/pii/S0925231209000162}.
+
+Geritz, Stefan A. H., and Eva Kisdi. 2012. ``Mathematical ecology: why
+mechanistic models?.'' \emph{Journal of mathematical biology} 65 (6-7)
+(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.
+
+Gordon, H. S. 1954. ``The economic theory of a common-property resource:
+the fishery.'' \emph{The Journal of Political Economy} 62 (2): 124--142.
+\url{http://www.jstor.org/stable/10.2307/1825571}.
+
+Halpern, Benjamin S., Carissa J. Klein, Christopher J. Brown, Maria
+Beger, Hedley S. Grantham, Sangeeta Mangubhai, Mary Ruckelshaus, et al.
+2013. ``Achieving the triple bottom line in the face of inherent
+trade-offs among social equity, economic return, and conservation.''
+\emph{Proceedings of the National Academy of Sciences of the United
+States of America} (mar). doi:10.1073/pnas.1217689110.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/23530207}.
+
+Hilborn, Ray. 2007. ``Reinterpreting the State of Fisheries and their
+Management.'' \emph{Ecosystems} 10 (8) (oct): 1362--1369.
+doi:10.1007/s10021-007-9100-5.
+\url{http://www.springerlink.com/index/10.1007/s10021-007-9100-5}.
+
+Hughes, Terry P., Cristina Linares, Vasilis Dakos, Ingrid a van de
+Leemput, and Egbert H. van Nes. 2013. ``Living dangerously on borrowed
+time during slow, unrecognized regime shifts.'' \emph{Trends in ecology
+\& evolution} 28 (3) (mar): 149--55. doi:10.1016/j.tree.2012.08.022.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/22995893}.
+
+Kocijan, Juš, Agathe Girard, Bla\textbackslash{}vz Banko, and Roderick
+Murray-Smith. 2005. ``Dynamic systems identification with Gaussian
+processes.'' \emph{Mathematical and Computer Modelling of Dynamical
+Systems} 11 (4) (dec): 411--424. doi:10.1080/13873950500068567.
+\url{http://www.tandfonline.com/doi/abs/10.1080/13873950500068567}.
+
+Ludwig, Donald, and Carl J. Walters. 1982. ``Optimal harvesting with
+imprecise parameter estimates.'' \emph{Ecological Modelling} 14 (3-4)
+(jan): 273--292. doi:10.1016/0304-3800(82)90023-0.
+\url{http://linkinghub.elsevier.com/retrieve/pii/0304380082900230}.
+
+Mangel, Marc. 1985. ``Decision and control in uncertain resource
+systems:'' 255.
+\href{http://www.amazon.com/Decision-uncertain-resource-Mathematics-Engineering/dp/0124687202 http://dl.acm.org/citation.cfm?id=537497}{http://www.amazon.com/Decision-uncertain-resource-Mathematics-Engineering/dp/0124687202
+http://dl.acm.org/citation.cfm?id=537497}.
+
+Mangel, Marc, and Colin W. Clark. 1988. \emph{Dynamic Modeling in
+Behavioral Ecology}. Ed. John Krebs and Tim Clutton-Brock. Princeton:
+Princeton University Press.
+
+May, Robert M., John R. Beddington, Colin W. Clark, Sidney J. Holt, and
+R. M. Laws. 1979. ``Management of multispecies fisheries.''
+\emph{Science (New York, N.Y.)} 205 (4403) (jul): 267--77.
+doi:10.1126/science.205.4403.267.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/17747032}.
+
+Munch, Stephan B., Melissa L. Snover, George M. Watters, and Marc
+Mangel. 2005. ``A unified treatment of top-down and bottom-up control of
+reproduction in populations.'' \emph{Ecology Letters} 8 (7) (may):
+691--695. doi:10.1111/j.1461-0248.2005.00766.x.
+\url{http://doi.wiley.com/10.1111/j.1461-0248.2005.00766.x}.
+
+Polasky, Stephen, Stephen R. Carpenter, Carl Folke, and Bonnie Keeler.
+2011. ``Decision-making under great uncertainty: environmental
+management in an era of global change.'' \emph{Trends in ecology \&
+evolution} (may): 1--7. doi:10.1016/j.tree.2011.04.007.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/21616553}.
+
+Rasmussen, Carl Edward, and C. K. I. Williams. 2006. \emph{Gaussian
+Processes for Machine Learning}. Ed. Thomas Dietterich. Boston: MIT
+Press,. \url{www.GaussianProcess.org/gpml}.
+
+Reed, William J. 1979. ``Optimal escapement levels in stochastic and
+deterministic harvesting models.'' \emph{Journal of Environmental
+Economics and Management} 6 (4) (dec): 350--363.
+doi:10.1016/0095-0696(79)90014-7.
+\href{http://www.sciencedirect.com/science/article/pii/0095069679900147 http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}{http://www.sciencedirect.com/science/article/pii/0095069679900147
+http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}.
+
+Schapaugh, Adam W., and Andrew J. Tyre. 2013. ``Accounting for
+parametric uncertainty in Markov decision processes.'' \emph{Ecological
+Modelling} 254 (apr): 15--21. doi:10.1016/j.ecolmodel.2013.01.003.
+\url{http://linkinghub.elsevier.com/retrieve/pii/S0304380013000306}.
+
+Scheffer, Marten, Jordi Bascompte, William A. Brock, Victor Brovkin,
+Stephen R. Carpenter, Vasilis Dakos, Hermann Held, Egbert H. van Nes,
+Max Rietkerk, and George Sugihara. 2009. ``Early-warning signals for
+critical transitions.'' \emph{Nature} 461 (7260): 53--9.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/19727193}.
+
+Scheffer, Marten, Stephen R. Carpenter, J. A. Foley, C. Folke, and B.
+Walker. 2001. ``Catastrophic shifts in ecosystems.'' \emph{Nature} 413
+(6856) (oct): 591--6. doi:10.1038/35098000.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/11595939}.
+
+Williams, Byron K. 2001. ``Uncertainty , learning , and the optimal
+management of wildlife.'' \emph{Environmental and Ecological Statistics}
+8: 269--288. doi:10.1023/A:1011395725123.
+
+Worm, Boris, Edward B. Barbier, Nicola Beaumont, J. Emmett Duffy, Carl
+Folke, Benjamin S. Halpern, Jeremy B. C. Jackson, et al. 2006. ``Impacts
+of biodiversity loss on ocean ecosystem services.'' \emph{Science (New
+York, N.Y.)} 314 (5800) (nov): 787--90. doi:10.1126/science.1132294.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/17082450}.
+
+Worm, Boris, Ray Hilborn, Julia K. Baum, Trevor A. Branch, Jeremy S.
+Collie, Christopher Costello, Michael J. Fogarty, et al. 2009.
+``Rebuilding global fisheries.'' \emph{Science (New York, N.Y.)} 325
+(5940) (jul): 578--85. doi:10.1126/science.1173146.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/19644114}.
+
 \end{document}

### 2013-09-08 inst/examples/BUGS/external-chunks.R

@@ -541,17 +541,18 @@ actual_over_optimal <-subset(tmp, variable != "True")
 
 
 ## @knitr Figure4 
-ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable)) + 
+fig4v1 <- ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable)) + 
   facet_wrap(~variable, scales = "free_y")  + guides(legend.position = "none") +
-  xlab("Total profit by replicate") + scale_fill_manual(values=colorkey)
+  xlab("Total profit by replicate") + scale_fill_manual(values=colorkey) # density plots fail when delta fn
 
-ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable), binwidth=0.1) + 
+fig4v2 <- ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable), binwidth=0.1) + 
   xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)
 
-ggplot(actual_over_optimal, aes(value, fill=variable, color=variable)) + 
+fig4v3 <- ggplot(actual_over_optimal, aes(value, fill=variable, color=variable)) + # density plots fail when delta fn
   stat_density(aes(y=..density..), position="stack", adjust=3, alpha=.9) + 
   xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)+ scale_color_manual(values=colorkey)
-
+fig4v2
+fig4v3
 
 
 

### 2013-09-08 inst/doc/nonparametric-bayes.Rmd

--- 2013-09-07 inst/doc/nonparametric-bayes.Rmd
@@
 
+<!-- ugh, re-word this -->
+
 To permit comparisons against a theoretical optimum we will consider

### 2013-09-18 inst/doc/nonparametric-bayes.Rmd
writing writing

@@
 Abstract
+========
+
+Model uncertainty and limited data coverage are fundamental challenges to
+robust ecosystem management.  These challenges are acutely highlighted
+by concerns that many ecological systems may contain tipping points.
+Before a collapse, we do not know where the tipping points lie, if the
+exist at all.  Hence, we know neither a complete model of the system
+dynamics nor do we have access to data in some large region of state-space
+where such a tipping point might exist.  These two sources of uncertainty
+frustrate state-of-the-art parametric approaches to decision theory
+and optimal control.  I will illustrate how a non-parametric approach
+using a Gaussian Process prior provides a more flexible representation
+of this inherent uncertainty.  Consequently, we can adapt the Gaussian
+Process prior to a stochastic dynamic programming framework in order to
+make robust management predictions under both model and uncertainty and
+limited data.
+
+Introduction
+============
+
+<!--
+Very briefly set the stage: (see related papers)
+- Decision-theory / optimal control framework
+- Fisheries context 
+-->
+
+The sudden collapse of fisheries and other ecosystems is an increasingly
+widespread phenomenon and a pressing concern for ecological management
+and conservation.  Ecological dyanaics are frequently complex and difficult to 
+measure, making uncertainty in our understanding a prediction a persistent
+challenge to effective management.
+
+Decision-theoretic approaches provide a framework to determine the best 
+sequence of actions in face of uncertainty, but only when that uncertainty
+can be meaningfully quantified [@Fischer2009].
+
+Uncertainty enters at many levels <!-- enters what -->
+management
+
+- Why fisheries
+
+The economic value and ecological concern have made marine
+fisheries the crucible for much of the founding work [@Gordon1954;
+@Reed1979; @May1979; @Ludwig1982] in managing ecosystems under
+uncertainty.  Global trends [@Worm2006] and controversy [@Hilborn2007;
+@Worm2009] have made understanding these challenges all the more pressing.
+
+<!-- Other reasons for failure see Sethi-->
+
+@Fischer 
+@Sethi2005, @Peretti, @Munch,
+
+(Classical division)
+
+- Dynamic uncertainty (stochasticity)
+
+From [Clark (1976)]() to [Reed (1979)]()
+
+Other sources of uncertainty [Roughgarden & Smith 1996](), [Sethi _et al._ (2005)]()
+
+- Parametric uncertainty <!-- ugh, be more consise and less annoying-->
+
+As the parameter values for these models must be estimated from limited data,  
+there will always be some uncertainty associated with these values.  This uncertainty
+further compounds the intrinsic variability introduced by demographic or environmental 
+noise.  The degree of uncertainty in the parameter values can be inferred from the data
+and reflected in the estimates of the transition probabilities.
+
+[@Mangel1985; @Schapaugh2013]
+
+[Ludwig & Walters 1982](http://doi.org/10.1016/0304-3800(82)90023-0 "Optimal harvesting with imprecise parameter estimates")
+
+- Structural uncertainty <!-- ugh, be more consise and less annoying-->
+
+Estimates of parameter uncertainty are only as good as the parametric
+models themselves.  Often we do not understand the system dynamics well
+enough to know if a model provides a good approximation over the relevant
+range of states and timescales (criteria that we loosely refer to as
+defining the "right" or "true" model.)  So called structural or model
+uncertainty is a more difficult problem than parametric uncertainty.
+Typical solutions involve either model choice, model averaging, or
+introducing yet greater model complexity of which others may be special
+cases (model averaging being one such way to construct such a model)
+[@Williams2001; @Athanassoglou2012; @Cressie2009].  Even setting aside
+other computational and statistical concerns (e.g. [@Cressie2009]), these
+approaches do not address our second concern - representing uncertainty
+outside the observed data range.
+
+[@Cressie2009](http://dx.doi.org/10.1890/07-0744.1 "Accounting for uncertainty in ecological analysis: the strengths and limitations of hierarchical statistical modeling.").
+
+- Limits of state space
+
+Often the set of feasible states is larger than the range of the observed states.
+Though a system can enter a state outside previously observed values due to stochastic effects alone (such as a particularly low or high nutrient year), this concern is most keenly felt in considering the value of actions that have not yet been applied.
+
+Of all sources of uncertainty, this has potentially recieved the least attention.
+It has most often appeared in the concept of active learning or active adaptive managment, (adaptive probing, ) where it may be valauble to intentionally force a system far from the observed values even when the expected value such actions is low, as it provides much faster learning and consequent reduction of model uncertainty that can allow greater value to be derived later on.
+@Ludwig1982 show that it may be advantageous to fish an unexploited population very heavily at first to obtain a better estimate of the recruitment rate.
+This intuitive strategy when a population is governed by a Ricker or Beverton-Holt-like dynamic would clearly be disasterous if instead the dynamics contained an unforseen tipping point.
+The best way to learn where the edge lies may be to walk up to it, but it is also the most dangerous.
+
+## Two central challenges:
+
+- We don't know the model
+- We don't have data from where we need it most
+
+## Why we don't know the model
+
+- Complex dynamics
+- model choice and model averaging approaches
+
+## Why we don't have data where we need it
+
+- Concerns of tipping points 
+- Danger of learning
+
+- MAP
+
+Decision making under uncertainty is a ubiquitous challenge of natural 
+resource management and conservation.  [@Glaser2013](http://doi.org/10.1111/faf.12037 "Complex dynamics may limit prediction in marine fisheries")
+
+Here we illustrate how a stochastic dynamic programming (SDP) algorithm can 
+be driven by the predictions from a Gaussian Process (GP).
+This provides two distinct advantages compared with contemporary approaches.
+First, using a GP sidesteps the need for an accurate model-based description
+of the system dynamics (or a model choice or model-averaging approach which
+introduces yet more difficulty).  
+Second, unlike parametric models which can only reflect 
+uncertainty in the corresponding confidence intervals or posterior distributions
+of their parameters, the GP approach provides a state-space dependent representation
+of uncertainty.  This permits a much greater uncertainty far from the observed data than near the observed 
+data.  These features allow the GP-SDP approach to find robust management solutions 
+in face of both limited data and no knolwedge of correct model structure.
+
+- Note on "not magic": honest uncertainty + SDP
+
+The idea that any approach can perform well without either having to
+know the model or have particularly good data should immediately draw
+suspicion.  The reader must bear in mind that the strength of our approach
+comes not from black-box predictive power from such limited information,
+but rather, by providing a more honest expression of uncertainty outside
+the observed data without sacrificing the predictive capacity near the
+observed data. By coupling this more accurate description of what is known
+and unknown to the decision-making under uncertainty framework provided
+by stochastic dynamic programming, we are able to obtain more robust
+management policies than with common parameteric modeling approaches.
+
+- Note on comparing models (via value function rather than by "fit"). _Do we need this?_
+
+The nature of decision-making problems provides a convenient way to compare 
+models.  Rather than compare models in terms of best fit to data or fret over
+the appropriate penalty for model complexity, model performance is defined 
+in the concrete terms of the decision-maker's objective function, which we
+will take as given. (Much argument can be made over the 'correct' objective
+function, e.g. how to account for the social value of fish left in the sea
+vs. the commercial value of fish harvested; see @Halpern2013 for further 
+discussion of this issue.  Alternatively, we can always compare model performance
+across multiple potential objective functions.)  The decision-maker
+does not necessarily need a model that provides the best mechanistic understanding
+or the best long-term outcome, but rather the one that best estimates the 
+probabilities of being in different states as a result of the possible actions.
+
+- Note on terminology: "Non-parametric"  _Do we need this?_
+
+## Background on the Gaussian Process
+
+- Definition
+- Previous application
+- Why it is particularly suited to these two problems
+- (Why this is a novel application thereof)
+
+Approach and Methods
+====================
+
+### Summary of approach
+
+### Statement of the optimal control problem
+
+- Underlying model 
+- Available data 
+- Value function
+
+For simplicity we assume profit is simply linear in the realized harvest (only
+enforcing the restriction that harvest can not exceed available stock)
+
+### Parametric models
+
+- Statement of the models
+
+We consider three candidate parametric models of the stock-recruitment
+dynamics: The Ricker model, the Allen model [Allen 2005](), the Myers
+model. The familiar Ricker model involves two parameters, corresponding
+to a growth rate and a carrying capacity, and cannot support alternative
+stable state dynamics (though as growth rate increases it exhibits a
+periodic attractor that proceeds through period-doubling into chaos. We
+will generally focus on dynamics below the chaotic threshold for
+the purposes of this analysis.) The Allen model resembles the Ricker
+dynamics with an added Allee effect parameter [Courchamp](), below
+which the population cannot persist.  The Myers model also has three
+parameters and contains an Allee threshold, but has compensatory rather
+than over-compensatory density dependence (resembling a Beverton-Holt
+curve rather than a Ricker curve at high densities.)
+
+We assume multiplicative log-normal noise perturbs the growth predicted 
+by the each of the deterministic model skeletons described above. This 
+introduces one additional parameter $\sigma$ that must be estimated by each
+model.
+
+<!-- equations just in appendix? -->
+
+As we simulate training data from the Allen model (ref section), we will
+refer to this as the structurally correct model.  The Ricker model is
+thus a reasonable approximation of these dynamics far from the Allee
+threshold (but lacks threshold dynamics), while the Myers model shares
+the essential feature of a threshold but differs in the structure. Thus
+we have three potential parametric models of the stock dynamics.
+
+- Bayesian inference of parametric models
+
+We infer posterior distributions for the parameters of each model
+in a Bayesian context using Gibbs sampling (implemented in R [R]()
+using jags, [jags]()).  We choose uninformative uniform priors for all
+parameters (See Appendix, Figures S1-S3, and Table S1, and the R code
+provided). One-step-ahead predictions of these model fits are shown in
+Figure 1.
+
+- SDP via parametric models
+
+An optimal policy function is then inferred through stochastic dynamic
+programming for each model given the posterior distributions of the
+parameter estimates.  This policy maximizes the expectation of the value
+function integrated over the parameter uncertainty. (code implementing
+this algorithm provided in the Appendix).
+
+### The Gaussian Process model
+
+- Statement of model
+
+... more on GP ... [Munch 2005]()
+
+We also estimate a simple Gaussian Process defined by
+a radial basis function kernel of two parameters: $\ell$, which gives
+the characteristic length-scale over which correlation between two 
+points (e.g. any two points $X_t, X_{t+1}$, and $X_{t+\tau}, X_{t+1+\tau}$)
+in state-space decays, and $\sigma$, which gives the scale of the 
+process noise by which observations $Y_{t+1}$ may differ from their
+predicted values $X_{t+1}$ given an observation of the previous state,
+$X_t$.
+
+- Inference of the model
+
+Also unlike parametric models, this posterior distribution is still
+conditional on the training data. As such, the uncertainty near the
+observed data.
+
+We use a Metropolis-Hastings Markov Chain Monte Carlo to infer posterior
+distributions of the two parameters of the GP (Figure S4, code in
+appendix), under weakly informative Gaussian priors (see parameters in
+table S5). As the posterior distributions differ substantially from the
+priors (Figure S4), we can be assured that most of the information in
+the posterior comes from the data rather than the prior belief.
+
+- SDP via the model
+
+Though we are unaware of prior application of this type, it is reasonably
+straight-forward to adapt the Gaussian Process for Stochastic Dynamic
+Programming.  Recall that unlike the parametric models the Gaussian
+process with fixed parameters already predicts a distribution of
+curves rather than a single curve. We must first integrate over 
+this distribution of curves given a sampling of parameter values drawn
+from the posterior distribution of the two GP parameters, before
+integrating over the posterior of those parameters themselves.
+
+Results
+=======
+
+### Figure 1: Fitted Models
+
+- All models fit the data quite well
+- Information criteria would pick the simple, incorrect model.
+
+(State-space figure??)
+
+### Figure 2: Inferred Policies
+
+- Inferred policies differ substantially among models
+- The structurally correct model and the GP are close to the true model
+- alternatives are not close
+
+### Figure 3: Simulated results
+
+-
+
+### Figure 4: Robustness
+
+- Results hold across range of parameters
+
+(eek, distinguish better between result and discussion?)
+
+Discussion 
+==========
+
+- All models are "good fits" to the originally observed data. 
+- (Simple model choice immediately leads us astray)
+
+Future directions
+-----------------
+
+### Higher dimensions
+
+### Online learning
+
+Acknowledgments
+===============
+
+Appendix
+========
+
+## Model definitions and estimation
+
+Equation S1: Ricker model.
+
+$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
+
+Figure S1: Ricker model: prior and posterior distributions for parameter estimates.
+
+```{r, echo=FALSE, include=TRUE}
+plot_ricker_traces
+```
+
+```{r, echo=FALSE, include=TRUE}
+plot_ricker_posteriors
+```
+Table S1: Parameterization of the priors
+
+```{r echo=FALSE, results="asis", include=TRUE}
+xtable::xtable(data.frame(
+  parameter = c("r0", "K", "sigma"),
+  lower_bound = c(r0_prior_p[1], K_prior_p[1], stdQ_prior_p[1]),
+  upper_bound = c(r0_prior_p[2], K_prior_p[2], stdQ_prior_p[2])))
+```
+
+$$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} $$
+
+Eq S2: Myers model 
+Figure S2: Myers model: Traces, prior and posterior distributions for parameter estimates.
+
+```{r, echo=FALSE, include=TRUE}
+plot_myers_traces
+```
+
+```{r, echo=FALSE, include=TRUE}
+plot_myers_posteriors
+```
+
+Table S2: Parameterization of the priors
+```{r echo=FALSE, results="asis", include=TRUE}
+xtable::xtable(data.frame(parameter = c("r0", "K", "theta", "sigma"),
+           lower_bound = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
+           upper_bound = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2])))
+```
+
+Eq S3: Allen model
+
+$$f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} $$
+
+Figure S3: Allen model: prior and posterior distributions for parameter estimates.
+
+```{r, echo=FALSE, include=TRUE}
+plot_allen_traces
+```
+
+```{r, echo=FALSE, include=TRUE}
+plot_allen_posteriors
+```
+
+Table S3: Parameterization of the priors
+
+```{r echo=FALSE, results="asis", include=TRUE}
+xtable::xtable(data.frame(parameter = c("r0", "K", "theta", "sigma"),
+           lower_bound = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
+           upper_bound = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2])))
+```
+
+Eq S4: GP model 
+Figure S4: GP model: prior and posterior distributions for parameter estimates.
+
+```{r echo=FALSE, include=TRUE}
+gp_assessment_plots
+```
+
+Table S4: Parameterization of the priors
+
+## Optimal Control Problem
+
+We seek the harvest policy $h(x)$ that maximizes:
+
+$$ \max_{h_t} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h_t) \delta^t  $$
+
+subject to the profit function $\Pi(X_t,h)$, discount rate $\delta$, and the state
+equation
+
+$$X_{t+1} = Z_t f(S_t)  $$
+$$S_t = X_t - h_t $$
+
+Where $Z_t$ is multiplicative noise function with mean 1, representing
+stochastic growth. We will consider log-normal noise with shape parameter
+$\sigma_g$.
+
+Form this we can write down the Bellman recursion as:
+
+$$V_t(x_t) = \max_h \mathbf{E} \left(\Pi(h_t, x_t) + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)$$
+
+For simplicity we assume profit is simply linear in the realized harvest (only
+enforcing the restriction that harvest can not exceed available stock), $\Pi(h,x) = min(h,x)$.
+
+### Pseudocode for the Bellman iteration
+
+```r
+ V1 <- sapply(1:length(h_grid), function(h){
+      delta * F[[h]] %*% V +  profit(x_grid, h_grid[h]) 
+    })
+    # find havest, h that gives the maximum value
+    out <- sapply(1:gridsize, function(j){
+      value <- max(V1[j,], na.rm = T) # each col is a diff h, max over these
+      index <- which.max(V1[j,])  # store index so we can recover h's 
+      c(value, index) # returns both profit value & index of optimal h.  
+    })
+    # Sets V[t+1] = max_h V[t] at each possible state value, x
+    V <- out[1,]                        # The new value-to-go
+    D[,OptTime-time+1] <- out[2,]       # The index positions
+```
+
+### Training data
+
+Eacho of our models $f(S_t)$ must be estimated from training data, which
+we simulate from the Allen model with parameters $r = $ ` r p[1]`, 
+$K =$ ` r p[2]`, $C =$ ` r p[3]`, and  $\sigma_g =$ ` r sigma_g` 
+for $T=$ `r Tobs$ timesteps, starting at initial condition $X_0 = $ `r Xo`.
+
+-----------------------------------
+
+<!-- OLD TEXT -->
+
+Abstract
 =======================================================================
-
-We address
 
@@
 
-<!--
 While methods that account for _parametric_ uncertainty can be very 
@@
 dynamics poses considerably greater difficulty.
--->
 
@@
 
-Decision making under uncertainty is a ubiquitous challenge of natural 
-resource management and conservation.  
-Here we illustrate how a stochastic dynamic programming (SDP) algorithm can 
-be driven by the predictions from a Gaussian Process (GP).
-This provides two distinct advantages compared with contemporary approaches.
-First, using a GP sidesteps the need for an accurate model-based description
-of the system dynamics (or a model choice or model-averaging approach which
-introduces yet more difficulty).  
-Second, unlike parametric model-fitting approaches, which can only reflect 
-uncertainty in the corresponding confidence intervals or  posterior distributions
-of the parameters, the GP approach provides a state-space dependent representation
-of uncertainty.  In other words, a GP estimated from a given data set permits 
-a much greater uncertainty far from the observed data than it does near the observed 
-data.
-
 <!-- More on complex dynamics and not having the correct model
-
-Cite [@Glaser2013](http://doi.org/10.1111/faf.12037 "Complex dynamics may limit prediction in marine fisheries")
-etc
-
-also older stuff on model uncertainty
 
@@
      that the data exhibits this bias / problem.  
-
      -->
-
-The idea that any approach can perform well without either having to know the model or
-have particularly good data should immediately draw suspicion.  The reader must bear
-in mind that the strength of our approach comes not from some magic predictive power
-from such limited information, but rather, by providing a more honest expression of 
-uncertainty outside the observed data without sacrificing the predictive capacity 
-near the observed data.
-
-By coupling this more accurate description of what is known and unknown to the 
-decision-making under uncertainty framework provided by stochastic dynamic programming,
-we are able to obtain more robust management policies than with common parameteric modeling 
-approaches.
 
@@
 
-As the parameter values for these models must be estimated from limited data,  
-there will always be some uncertainty associated with these values.  This uncertainty
-further compounds the intrinsic variability introduced by demographic or environmental 
-noise.  The degree of uncertainty in the parameter values can be inferred from the data
-and reflected in the estimates of the transition probabilities.
-
-[@Mangel1985; @Schapaugh2013]
-
 #### Structural Uncertainty
 
-Unfortunately, estimates of parameter uncertainty are only as good as the parametric 
-models themselves.  Too often, we do not understand the system dynamics well enough
-to know if a model provides a good approximation over the relevant range of states
-and timescales (criteria that we loosely refer to as defining the "right" or "true" model.)
-So called structural or model uncertainty is a more difficult problem than parametric 
-uncertainty.  Typical solutions involve either model choice, or model averaging. Either
-approach remains limited by the degree to which any of the proposed models are sufficiently
-close to the right model.
-
 [@Williams2001; @Athanassoglou2012].
-
-#### Comparing models
-
-The nature of decision-making problems provides a convenient way to compare 
-models.  Rather than compare models in terms of best fit to data or fret over
-the appropriate penalty for model complexity, model performance is defined 
-in the concrete terms of the decision-maker's objective function, which we
-will take as given. (Much argument can be made over the 'correct' objective
-function, e.g. how to account for the social value of fish left in the sea
-vs. the commercial value of fish harvested; see @Halpern2013 for further 
-discussion of this issue.  Alternatively, we can always compare model performance
-across multiple potential objective functions.)  The decision-maker
-does not necessarily need a model that provides the best mechanistic understanding
-or the best long-term outcome, but rather the one that best estimates the 
-probabilities of being in different states as a result of the possible actions.
 
@@
 
-#### Background on Fisheries Context
-
-#### Background on Tipping points
-
-<!-- note: This section largely replaced by above outline
-
+<!-- 
 ### Quantitative vs Qualitative Decisions
@@
 
-Non-parametric models can better represent uncertainties outside 
-of observed data while also better capturing the dynamics in the region
-observed.  Advances in the theory and computational implementations of 
-nonparametric methods make them ripe for such applications.  We use the classic
-problem of optimal harvest of a marine fishery to illustrate how the 
-nonparametric approach of Gaussian processes can be applied.  We will 
-compare Bayesian implementations of both nonparametric and parametric 
-models, which best allow us to capture the uncertainty of model estimation
-in either case and permits a more natural comparison between approaches.
-
 Approach and Methods
@@
 the harvest level for a marine fishery each year to maximize the net
-present value of the resource, given an estimated stock size from the year
-before. The economic value and ecological concern have made marine fisheries the crucible for much
-of the founding work [@Gordon1954; @Reed1979; @May1979; @Ludwig1982]
-in managing ecosystems under uncertainty.  Global trends [@Worm2006]
-and controversy [@Hilborn2007; @Worm2009] have made understanding these
-challenges all the more pressing.
+present value of the resource, given an estimated stock size from the
+year before.
 

### 2013-09-18 inst/doc/nonparametric-bayes.tex
writing writing

@@
 
+\section{Introduction}
+
+Briefly set the stage: (see related papers)
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Decision-theory / optimal control framework
+\item
+  Fisheries context
+\end{itemize}
+
+@Fischer @Sethi2005, @Peretti, @Munch,
+
+(Classical division)
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Dynamic uncertainty (stochasticity)
+\item
+  Parametric uncertainty
+\item
+  Structural uncertainty
+\end{itemize}
+
+\subsection{Two central challenges:}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  We don't know the model
+\item
+  We don't have data from where we need it most
+\end{itemize}
+
+\subsection{Why we don't know the model}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Complex dynamics
+\item
+  model choice and model averaging approaches
+\end{itemize}
+
+\subsection{Why we don't have data where we need it}
+
+\begin{itemize}
+\item
+  Concerns of tipping points
+\item
+  Danger of learning
+\item
+  Note on ``not magic'': honest uncertainty + SDP
+\item
+  Note on comparing models (via value function rather than by ``fit'')
+\item
+  Note on terminology: ``Non-parametric''
+\end{itemize}
+
+\subsection{Background on the Gaussian Process}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Definition
+\item
+  Previous application
+\item
+  Why it is particularly suited to these two problems
+\item
+  (Why this is a novel application thereof)
+\end{itemize}
+
+\section{Approach and Methods}
+
+\subsubsection{Summary of approach}
+
+\subsubsection{Statement of the optimal control problem}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Underlying model
+\item
+  Available data
+\item
+  Value function
+\end{itemize}
+
+For simplicity we assume profit is simply linear in the realized harvest
+(only enforcing the restriction that harvest can not exceed available
+stock)
+
+\subsubsection{Parametric models}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Statement of the models
+\end{itemize}
+
+We consider three candidate parametric models of the stock-recruitment
+dynamics: The Ricker model, the Allen model \href{}{Allen 2005}, the
+Myers model. The familiar Ricker model involves two parameters,
+corresponding to a growth rate and a carrying capacity, and cannot
+support alternative stable state dynamics (though as growth rate
+increases it exhibits a periodic attractor that proceeds through
+period-doubling into chaos. We will generally focus on dynamics below
+the chaotic threshold for the purposes of this analysis.) The Allen
+model resembles the Ricker dynamics with an added Allee effect parameter
+\href{}{Courchamp}, below which the population cannot persist. The Myers
+model also has three parameters and contains an Allee threshold, but has
+compensatory rather than over-compensatory density dependence
+(resembling a Beverton-Holt curve rather than a Ricker curve at high
+densities.)
+
+We assume multiplicative log-normal noise perturbs the growth predicted
+by the each of the deterministic model skeletons described above. This
+introduces one additional parameter $\sigma$ that must be estimated by
+each model.
+
+As we simulate training data from the Allen model (ref section), we will
+refer to this as the structurally correct model. The Ricker model is
+thus a reasonable approximation of these dynamics far from the Allee
+threshold (but lacks threshold dynamics), while the Myers model shares
+the essential feature of a threshold but differs in the structure. Thus
+we have three potential parametric models of the stock dynamics.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Bayesian inference of parametric models
+\end{itemize}
+
+We infer posterior distributions for the parameters of each model in a
+Bayesian context using Gibbs sampling (implemented in R \href{}{R} using
+jags, \href{}{jags}). We choose uninformative uniform priors for all
+parameters (See Appendix, Figures S1-S3, and Table S1, and the R code
+provided). One-step-ahead predictions of these model fits are shown in
+Figure 1.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  SDP via parametric models
+\end{itemize}
+
+An optimal policy function is then inferred through stochastic dynamic
+programming for each model given the posterior distributions of the
+parameter estimates. This policy maximizes the expectation of the value
+function integrated over the parameter uncertainty. (code implementing
+this algorithm provided in the Appendix).
+
+\subsubsection{The Gaussian Process model}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Statement of model
+\end{itemize}
+
+\ldots{} more on GP \ldots{} \href{}{Munch 2005}
+
+We also estimate a simple Gaussian Process defined by a radial basis
+function kernel of two parameters: $\ell$, which gives the
+characteristic length-scale over which correlation between two points
+(e.g.~any two points $X_t, X_{t+1}$, and $X_{t+\tau}, X_{t+1+\tau}$) in
+state-space decays, and $\sigma$, which gives the scale of the process
+noise by which observations $Y_{t+1}$ may differ from their predicted
+values $X_{t+1}$ given an observation of the previous state, $X_t$.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Inference of the model
+\end{itemize}
+
+Also unlike parametric models, this posterior distribution is still
+conditional on the training data. As such, the uncertainty near the
+observed data.
+
+We use a Metropolis-Hastings Markov Chain Monte Carlo to infer posterior
+distributions of the two parameters of the GP (Figure S4, code in
+appendix), under weakly informative Gaussian priors (see parameters in
+table S5). As the posterior distributions differ substantially from the
+priors (Figure S4), we can be assured that most of the information in
+the posterior comes from the data rather than the prior belief.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  SDP via the model
+\end{itemize}
+
+Though we are unaware of prior application of this type, it is
+reasonably straight-forward to adapt the Gaussian Process for Stochastic
+Dynamic Programming. Recall that unlike the parametric models the
+Gaussian process with fixed parameters already predicts a distribution
+of curves rather than a single curve. We must first integrate over this
+distribution of curves given a sampling of parameter values drawn from
+the posterior distribution of the two GP parameters, before integrating
+over the posterior of those parameters themselves.
+
+\section{Results}
+
+\subsubsection{Figure 1: Fitted Models}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  All models fit the data quite well
+\item
+  Information criteria would pick the simple, incorrect model.
+\end{itemize}
+
+(State-space figure??)
+
+\subsubsection{Figure 2: Inferred Policies}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Inferred policies differ substantially among models
+\item
+  The structurally correct model and the GP are close to the true model
+\item
+  alternatives are not close
+\end{itemize}
+
+\subsubsection{Figure 3: Simulated results}
+
+-
+
+\subsubsection{Figure 4: Robustness}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Results hold across range of parameters
+\end{itemize}
+
+(eek, distinguish better between result and discussion?)
+
+\section{Discussion}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  All models are ``good fits'' to the originally observed data.
+\item
+  (Simple model choice immediately leads us astray)
+\end{itemize}
+
+\subsection{Future directions}
+
+\subsubsection{Higher dimensions}
+
+\subsubsection{Online learning}
+
+\section{Acknowledgments}
+
+\section{Appendix}
+
+\subsection{Model definitions and estimation}
+
+Equation S1: Ricker model.
+
+\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
+
+Figure S1: Ricker model: prior and posterior distributions for parameter
+estimates.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-1.pdf}
+\caption{plot of chunk unnamed-chunk-1}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-2.pdf}
+\caption{plot of chunk unnamed-chunk-2}
+\end{figure}
+
+Table S1: Parameterization of the priors
+
+\% latex table generated in R 3.0.1 by xtable 1.7-1 package \% Mon Sep 9
+14:43:29 2013
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rlrr}
+  \hline
+ & parameter & lower\_bound & upper\_bound \\ 
+  \hline
+1 & r0 & 0.00 & 10.00 \\ 
+  2 & K & 0.00 & 40.00 \\ 
+  3 & sigma & 0.00 & 100.00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \]
+
+Eq S2: Myers model Figure S2: Myers model: Traces, prior and posterior
+distributions for parameter estimates.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-4.pdf}
+\caption{plot of chunk unnamed-chunk-4}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-5.pdf}
+\caption{plot of chunk unnamed-chunk-5}
+\end{figure}
+
+Table S2: Parameterization of the priors \% latex table generated in R
+3.0.1 by xtable 1.7-1 package \% Mon Sep 9 14:43:37 2013
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rlrr}
+  \hline
+ & parameter & lower\_bound & upper\_bound \\ 
+  \hline
+1 & r0 & 0.00 & 10.00 \\ 
+  2 & K & 0.00 & 40.00 \\ 
+  3 & theta & 0.00 & 10.00 \\ 
+  4 & sigma & 0.00 & 100.00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+Eq S3: Allen model
+
+\[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]
+
+Figure S3: Allen model: prior and posterior distributions for parameter
+estimates.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-7.pdf}
+\caption{plot of chunk unnamed-chunk-7}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-8.pdf}
+\caption{plot of chunk unnamed-chunk-8}
+\end{figure}
+
+Table S3: Parameterization of the priors
+
+\% latex table generated in R 3.0.1 by xtable 1.7-1 package \% Mon Sep 9
+14:43:45 2013
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rlrr}
+  \hline
+ & parameter & lower\_bound & upper\_bound \\ 
+  \hline
+1 & r0 & 0.00 & 10.00 \\ 
+  2 & K & 0.00 & 40.00 \\ 
+  3 & theta & 0.00 & 10.00 \\ 
+  4 & sigma & 0.00 & 100.00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+Eq S4: GP model Figure S4: GP model: prior and posterior distributions
+for parameter estimates.
+
+\begin{verbatim}
+$traces_plot
+\end{verbatim}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-101.pdf}
+\caption{plot of chunk unnamed-chunk-10}
+\end{figure}
+
+\begin{verbatim}
+
+$posteriors_plot
+\end{verbatim}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-102.pdf}
+\caption{plot of chunk unnamed-chunk-10}
+\end{figure}
+
+Table S4: Parameterization of the priors
+
+\subsection{Optimal Control Problem}
+
+We seek the harvest policy $h(x)$ that maximizes:
+
+\[ \max_{h(t)} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h) \delta^t  \]
+
+subject to the profit function $\Pi(X_t,h)$, discount rate $\delta$, and
+the state equation
+
+\[X_{t+1} = Z_t f(S_t)  \] \[S_t = X_t - h_t \]
+
+Where $Z_t$ is multiplicative noise function with mean 1, representing
+stochastic growth. We will consider log-normal noise with shape
+parameter $\sigma_g$.
+
+Form this we can write down the Bellman recursion as:
+
+\[V_t(x_t) = \max_h \mathbf{E} \left(\Pi(h_t, x_t) + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)\]
+
+For simplicity we assume profit is simply linear in the realized harvest
+(only enforcing the restriction that harvest can not exceed available
+stock), $\Pi(h,x) = min(h,x)$.
+
+\subsubsection{Pseudocode for the Bellman iteration}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+ \NormalTok{V1 <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
+      \NormalTok{delta * F[[h]] %*% V +  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
+    \NormalTok{\})}
+    \CommentTok{# find havest, h that gives the maximum value}
+    \NormalTok{out <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
+      \NormalTok{value <- }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
+      \NormalTok{index <- }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
+      \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
+    \NormalTok{\})}
+    \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
+    \NormalTok{V <- out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
+    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <- out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
+\end{Highlighting}
+\end{Shaded}
+
+\subsubsection{Training data}
+
+Eacho of our models $f(S_t)$ must be estimated from training data, which
+we simulate from the Allen model with parameters $r = 2$, $K=8$, $C=5$
+and $\sigma_g = 0.1$ for $T = 40$ timesteps, starting at initial
+condition $X_0 = $.
+
+\begin{center}\rule{3in}{0.4pt}\end{center}
+
+\section{Abstract}
+
 Decision-theoretic methods often rely on simple parametric models of
@@
 
+Further, the data on which a model has been parameterized frequently
+fails to cover the possible state-space over which management decisions
+must operate. Consequently a model do well in the region of state-space
+in which it was estimated, but give erroneous confidence to predictions
+outside of that region.
+
+This problem is keenly felt in any system where a potential threshold or
+tipping point is a concern. Such a tipping point, if it exists at all,
+will lay outside the observed range of the observed data.
+
 We demonstrate how nonparametric Bayesian models can provide robust,
-nearly-optimal solutions to decision making under uncertainty when
-\emph{we don't know the correct model} to use.
-
-While methods that account for \emph{parametric} uncertainty can be very
-successful with the right model, structural uncertainty of not knowing
-what model best approximates the dynamics poses considerably greater
-difficulty.
-
-Non-parametric Bayesian methods offer a promising statistical approach
-for predictive modeling of ecological dynamics in regions of state space
-where the data is adequate, while at the same time offering more
-flexible patterns with greater uncertainty outside the observed data.
-This contrasts from simple parametric models which provide relatively
-constant level of uncertainty in regions with and without adequate data.
-The consequence of such misplaced confidence outside the data can lead
-to highly undesirable results that may be avoided with the more flexible
-non-parametric Bayesian approach.
+solutions to decision making under uncertainty without knowing the
+structural form of the true model.
 
@@
 resource management and conservation.\\Here we illustrate how a
-stochastic dynamic programming algorithm can be driven by the
-predictions from a Gaussian process model, sidestepping the need for an
-accurate model-based description of the system dynamics.
-
-add references to (@Fischer2009, Polasky et al. 2011) as persective on
-combining complex / critical dynamics + control theory
+stochastic dynamic programming (SDP) algorithm can be driven by the
+predictions from a Gaussian Process (GP). This provides two distinct
+advantages compared with contemporary approaches. First, using a GP
+sidesteps the need for an accurate model-based description of the system
+dynamics (or a model choice or model-averaging approach which introduces
+yet more difficulty).\\Second, unlike parametric model-fitting
+approaches, which can only reflect uncertainty in the corresponding
+confidence intervals or posterior distributions of the parameters, the
+GP approach provides a state-space dependent representation of
+uncertainty. In other words, a GP estimated from a given data set
+permits a much greater uncertainty far from the observed data than it
+does near the observed data.
+
+The idea that any approach can perform well without either having to
+know the model or have particularly good data should immediately draw
+suspicion. The reader must bear in mind that the strength of our
+approach comes not from black-box predictive power from such limited
+information, but rather, by providing a more honest expression of
+uncertainty outside the observed data without sacrificing the predictive
+capacity near the observed data.
+
+By coupling this more accurate description of what is known and unknown
+to the decision-making under uncertainty framework provided by
+stochastic dynamic programming, we are able to obtain more robust
+management policies than with common parameteric modeling approaches.
 
@@
 
-Decision-theoretic (or optimal control) tools require a model that can
+Decision-theoretic or optimal control tools require a model that can
 assign probabilities of future states (e.g.~stock size of a fishery)
 given the current state and a proposed action (e.g.~fishing harvest or
-effort). The decision maker then seeks to determining the course of
-actions (also referred to as the policy) that maximizes the expected
-value of some objective function (such as net present value derived from
-the resource over time. (The approach can be adapted to permit
-alternatives to maximizing expectation, such as minimizing the maximum
-cost or damage that might be incurred. See Polasky et al. (2011)).
-Management frequently faces a sequential decision-making problem --
-after selecting an action, the decision-maker may receive new
+effort).\\Management frequently faces a sequential decision-making
+problem -- after selecting an action, the decision-maker may receive new
 information about the current state and must again choose an appropriate
-action -- i.e.~setting the harvest limits each year based on stock
+action -- such as setting the harvest limits each year based on stock
 assessments the year prior.
 
-More historical context, e.g. (Reed 1979, )
-
-\paragraph{Parametric models}
-
-Traditional approaches to optimal control (Pontryagin's principle,
-stochastic dynamic programming) rely on knowledge of the state equation.
-Such simple parametric models, frequently justified by mechanistic
-underpinnings are used to provide probabilities of future states.
-Fisheries literature, which enjoys some of the longest history in the
-use of the optimal control framework, provides a clear example {[} {]} .
-Models such as the Ricker or Beverton Holt curves are ubiquitous {[}too
-strong? references?{]} Parametric models can reflect variability
-introduced by a stochastic environment or demographic process my
-incorporating random shocks from a given distribution.
+The decision maker typically seeks to determining the course of actions
+(also referred to as the policy) that maximizes the expected value of
+some objective function such as net present value derived from the
+resource over time.\\Though much can be said on how to choose this value
+function appropriately (e.g.~see (Halpern et al.
+2013)(http://doi.org/10.1073/pnas.1217689110 ``Achieving the triple
+bottom line in the face of inherent trade-offs among social equity,
+economic return, and conservation.'')) we will assume this is given.
+(Nor is this approach necessarily constrained to maximizing the
+expectated value of such a function - the decision-theoretic framework
+can be adapted to alternatives such as minimizing the maximum cost or
+damage that might be incurred; see Polasky et al. (2011)).
+
+In representing future states with probabilities and maximizing
+expectations, this approach provides a natural framework for handling
+uncertainty.
+
+The value function typically depends on the action or policy taken, as
+well as the state of the system, in each interval of time. The state of
+the system, in turn, is usually described by a dynamical model.
 
@@
 than parametric uncertainty. Typical solutions involve either model
-choice, or model averaging. Either approach remains limited by the
-degree to which any of the proposed models are sufficiently close to the
-right model.
+choice or model averaging. Either approach remains limited by the degree
+to which any of the proposed models are sufficiently close to the right
+model.
 

### 2013-09-20 inst/doc/nonparametric-bayes.Rmd
writing

@@
 
+<!-- FIXME -->
+
 The sudden collapse of fisheries and other ecosystems is an increasingly
 widespread phenomenon and a pressing concern for ecological management
-and conservation.  Ecological dyanaics are frequently complex and difficult to 
+and conservation.  Ecological dynamics are frequently complex and difficult to 
 measure, making uncertainty in our understanding a prediction a persistent
 challenge to effective management. 
-
 Decision-theoretic approaches provide a framework to determine the best 
@@
 
+Resource management and conservation planning seek to determine the optimal set of feasible actions to maximize the value of some objectives (see [@Halpern2012])  Process error, measurement error, implementation error [@Reed1979, @Clark1986, @Roughgarden1996, @Sethi2005].
+These sources of stochasticity in turn mean that model parameters can only be estimated approximately, requiring parametric uncertainty also be considered [@Ludwig1982, ].
+
+The three sources of uncertainty: stochasticity (in process, measurement, and implementation; though we focus on process-noise only for simplicity), parametric uncertainty, and structural uncertainty, are familiar challenges.
+
+Of all sources of uncertainty, this last has received the least attention.[^1]
+
 Often the set of feasible states is larger than the range of the observed states.
-Though a system can enter a state outside previously observed values due to stochastic effects alone (such as a particularly low or high nutrient year), this concern is most keenly felt in considering the value of actions that have not yet been applied.
-
-Of all sources of uncertainty, this has potentially recieved the least attention.
-It has most often appeared in the concept of active learning or active adaptive managment, (adaptive probing, ) where it may be valauble to intentionally force a system far from the observed values even when the expected value such actions is low, as it provides much faster learning and consequent reduction of model uncertainty that can allow greater value to be derived later on.
-@Ludwig1982 show that it may be advantageous to fish an unexploited population very heavily at first to obtain a better estimate of the recruitment rate.
-This intuitive strategy when a population is governed by a Ricker or Beverton-Holt-like dynamic would clearly be disasterous if instead the dynamics contained an unforseen tipping point.
+Though most systems subject to some degree of stochasticity can enter a state outside previously observed values by chance alone (such as a particularly low or high nutrient year), this concern is most keenly felt in the decision-theoretic context.
+If we are observing the fluctuations of a given fish stock over many years under a fixed harvesting pressure, we might develop and test a model that could reasonably predict the frequency of a deviation of a given size, even when such a deviation has not been previously observed.
+
+(The same analogy might be made to predicting hurricanes or other extreme weather events from historical data, which may be more accurately predicted in a constant climate than in one being perturbed).
+
+Given adequate observations of a system
+
+One must consider actions or policy under which the system has not yet been observed,
+
+<!-- What do we call this?  Extrapolation uncertainty?  Pathological Uncertainty? -->
+
+[^1]: The concept of adaptive probing is one area that has explicitly addressed this kind of uncertainty, reaching rather opposite conclusions than what we observe here.
+Adaptive probing strategies follow from "Dual Control" or "Active Adaptive Management" approaches (e.g.
+@Ludwig1982) that can trade off short term utility by choosing actions that can reduce uncertainty.
+Adaptive probing strategies arise when it is valuable to intentionally force a system far from the observed values even when the expected value such actions is low, as it provides much faster learning and consequent reduction of model uncertainty that can allow greater value to be derived later on.
+For instance, @Ludwig1982 show that it may be advantageous to fish an unexploited population very heavily at first to obtain a better estimate of the recruitment rate.
+This intuitive strategy when a population is governed by a Ricker or Beverton-Holt-like dynamic would clearly be disastrous if instead the dynamics contained an unforeseen tipping point.
 The best way to learn where the edge lies may be to walk up to it, but it is also the most dangerous.
@@
 
-- Complex dynamics
+- Complex dynamics  [@Glaser2013](http://doi.org/10.1111/faf.12037 "Complex dynamics may limit prediction in marine fisheries")
+
 - model choice and model averaging approaches
@@
 
-- MAP
+- _MAP_
 
 Decision making under uncertainty is a ubiquitous challenge of natural 
-resource management and conservation.  [@Glaser2013](http://doi.org/10.1111/faf.12037 "Complex dynamics may limit prediction in marine fisheries")
-
+resource management and conservation. 
 Here we illustrate how a stochastic dynamic programming (SDP) algorithm can 
@@
 data.  These features allow the GP-SDP approach to find robust management solutions 
-in face of both limited data and no knolwedge of correct model structure.
-
-- Note on "not magic": honest uncertainty + SDP
+in face of limited data and without knowledge of the correct model structure.
+
+-  _Note on "not magic"_: honest uncertainty + SDP
 
@@
 by stochastic dynamic programming, we are able to obtain more robust
-management policies than with common parameteric modeling approaches.
-
-- Note on comparing models (via value function rather than by "fit"). _Do we need this?_
+management policies than with common parametric modeling approaches.
+
+<!--move to  much later --> 
+- Note on terminology: "Non-parametric"
+
+The term non-parametric to describe Gaussian processes (GP) is unfortunate,
+given that the GP models still involve the estimation of parameters. Some 
+literature thus uses the term semi-parametric, which merely adds ambiguity 
+to the confusion.  Statisticians give two distinct meanings to the term: (a)
+referring to a method that does not assume a distribution (e.g. non-parametric bootstrap, order statistics),
+and (b) one that does not assume a fixed structure but rather depends on the 
+size of the data (e.g. non-parametric regression, non-parametric hierarchical models
+such as the Dirichlet process).  Our use is in the latter sense.
+
+This non-parametric property -- having a structure that is not fixed but 
+rather dependent on the data is precisely the property that makes this approach
+attractive in face of the limited data sampling challenges discussed above.  
+Having fit a parametric model to some data, the model is completely described
+by the values (or posterior distributions) of it's parameters.  The non-parametric 
+model is not captured by its parameter values or distributions alone -- the 
+data points become part of the model.
+
+- _Note on comparing models_ (via value function rather than by "fit").
+
+_Do we need this?_
 
@@
 probabilities of being in different states as a result of the possible actions.
-
-- Note on terminology: "Non-parametric"  _Do we need this?_
 
@@
 We infer posterior distributions for the parameters of each model
-in a Bayesian context using Gibbs sampling (implemented in R [R]()
-using jags, [jags]()).  We choose uninformative uniform priors for all
+in a Bayesian context using Gibbs sampling (implemented in R [@RTeam]
+using jags, [@R2jags]).  We choose uninformative uniform priors for all
 parameters (See Appendix, Figures S1-S3, and Table S1, and the R code
@@
 
+```{r Figure1, dependson=c("assemble-models"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
+Graph shows the expected value for the function $f$ under each model.
+Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.
+The training data is also shown as black points.
+(The GP is conditioned on 0,0, shown as a pseudo-data point).
+", include=FALSE}
+```
+```{r Figureb, dependson=c("assemble-models", "par-fns"), include=TRUE}
+```
+
+Figure 1 shows the mean inferred state space dynamics of each model
+relative to the true model used to generate the data, predicting the
+relationship between observed stock size (x-axis) to the stock size
+after recruitment the following year.  All models except the MLE model
+estimate a distribution around the means shown here, and all models 
+estimate a level of process noise, which is independent of the state
+value (x).  Note that in contrast to the  other models shown, the mean
+Gaussian process corresponds to a distribution of curves - as indicated
+by the gray band - which itself has a mean shown in black.  Note that 
+this mean GP is thus more certain of the dynamics in the region where
+data is available then where it is not.
+
+While it would be straight forward to condition the GP on passing through the origin (0,0), the estimate
+shown here is based only on the observed data. The observed data from
+which each model is estimated is also shown.  The observations come
+from only a limited region of state space corresponding to unharvested
+or weakly harvested system.  No observations occur at the theoretical
+optimum harvest rate or near the tipping point.
+
+```{r Figure2, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
+Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", include=TRUE}
+```
+
+The resulting optimal management strategy based on each of the inferred
+models is shown in Figure 2, against the optimal strategy given the
+true underlying dynamics.  Policies are shown in terms of target
+escapement, $S_t$.  Under models such as this a constant escapement
+policy is expected to be optimal [@Reed1979], whereby population levels
+below a certain size $S$ are unharvested, while above that size the harvest
+strategy aims to return the population to $S$, resulting in the hockey-stick
+shaped policies shown.
+
+```{r Figure3, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.",  include=TRUE}
+```
+
+The consequences of managing 100 replicate realizations of the 
+simulated fishery under each of the policies estimated is shown in Figure 3.  As expected
+from the policy curves, the structurally correct model under-harvests,
+leaving the stock to vary around it's un-fished optimum.  The structurally
+incorrect Ricker model over-harvests the population passed
+the tipping point consistently, resulting in the immediate crash of the stock and 
+thus derives minimal profits.
+
+The results shown in Figures 1-3 are not unique to the simulated data or models chosen
+here, but arises across a range of parameter values and simulations as shown in the 
+supplemental figures.  The results across this range can most easily be compared 
+by the relative differences in net present value realized by each of the approaches,
+as shown in Figure 4.  The Gaussian Process most consistently realizes a value 
+close to the optimal solution, and importantly avoids ever driving the system across
+the tipping point, which results in the near-zero value cases in the parametric models.
+
+```{r Figure4, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
+For each data set, the three models are estimated as described above.
+Values plotted are the averages of a given policy over 100 replicate simulations.
+Details and code provided in the supplement.",  include=TRUE}
+```
+
 ### Figure 1: Fitted Models
@@
 - Information criteria would pick the simple, incorrect model.
-
-(State-space figure??)
 
@@
 - All models are "good fits" to the originally observed data.
+
 - (Simple model choice immediately leads us astray)
+
+Though simple mechanistically motivated models offer the greatest potential 
+to increase our basic understanding of ecological processes [@Cuddington2013; @Geritz2012], 
+such models can be not only inaccurate but misleading when relied upon in a
+quantitative decision making framework.
+
+1. We do not know what the correct models are for ecological systems.
+1. We have limited data from which to estimate the model -- in particular,
+   such models may be misleading in predicting the probability of outcomes
+   outside the training data.
+
+These aspects are common to many conservation decision making problems, which thus merit
+greater use of non-parametric approaches that can best take advantage of them.
+
+### 1. Large uncertainty where the data is poor
+
+The parametric models perform worst when they propose a management strategy
+outside the range of the observed data. The non-parametric Bayesian approach, 
+in contrast, allows a predictive model that expresses a great deal of uncertainty
+about the probable dynamics outside the observed range, while retaining very
+good predictive accuracy in the range observed.  The management policy 
+dictated by the GP balance this uncertainty against
+the immediate value of the harvest, and act to stabilize the population 
+dynamics in a region of state space in which the predictions can be 
+reliably reflected by the data.
+
+### 2. Predictive accuracy where data is good
+
+While expressing larger uncertainty outside the observed data, the GP
+can also provide a better fit with smaller uncertainty inside the range
+of the observed data. This arises from the greater flexibility of the 
+Gaussian process, which describes a large family of possible curves.
+Despite this flexibility, the GP can be described in relatively few 
+parameters and is thus far less likely to overfit.
 
@@
 
+In this simulated example, the underlying
+dynamics are truly governed by a simple parametric model, allowing
+the parametric approaches to be more accurate.  Similarly, because the
+dynamics are  one-dimensional dynamics and lead to  stable nodes (rather
+than other attractors such as limit-cycles resulting in oscillations),
+the training data provides relatively limited information about the
+dynamics.  For these reasons, we anticipate that in higher-dimensional
+examples characteristic of ecosystem management problems that the machine
+learning approach will prove even more valuable.
+
 ### Online learning
+
+In our treatment here we have ignored the possibility of learning during the 
+management phase, in which the additional observations of the stock size could
+potentially improve parameter estimates.  While we intend to address this 
+possibility in future work in the context of these non-parametric models,
+we have not addressed it here for pedagogical reasons. In the context presented
+here, it is clear that the differences in performance arise from differences
+in the uncertainty inherent in the model formulations, rather than from 
+differing abilities to learn.  Because we consider a threshold system, 
+online learning would not change this generic feature of a lack of data in a
+certain range of the state space which is better captured by the Gaussian process.
 
@@
 ===============
+
+This work was partially supported by the Center for Stock Assessment
+Research, a partnership between the University of California Santa Cruz
+and the Fisheries Ecology Division, Southwest Fisheries Science Center,
+Santa Cruz, CA and by NSF grant EF-0924195 to MM and NSF grant DBI-1306697
+to CB.
 
@@
 for $T=$ `r Tobs$ timesteps, starting at initial condition $X_0 = $ `r Xo`. 
+The training data can be seen in Figure 1.
 
@@
 
-#### Parametric uncertainty
-
-#### Structural Uncertainty
-
 [@Williams2001; @Athanassoglou2012].
@@
 
-In our example, we focus on the problem in which a manager must set
+We focus on the problem in which a manager must set
 the harvest level for a marine fishery each year to maximize the net
@@
 ### The Non-parametric Bayesian alternative for stock-recruitment curves
-
-#### Terminology: "Non-parametric"
-
-The term non-parametric to describe modeling approaches such as Gaussian
-processes is a common source of confusion, notwithstanding the fact that
-the model is still specified by parameters.  Some literature has introduced
-the term "semi-parametric" in tacit acknowledgment of this, which no doubt
-only contributes to the confusion.  The problem is further exacerbated by
-the several meanings assigned to the term in statistics: (a) that the method 
-does not assume some particular probability distribution (e.g. a non-parametric bootstrap)
-or (b) that the method does not assume a fixed structure to the model.  Our 
-use is that of (b), in contrast to the classical approaches that do.  The 
-difference is clearest by way of example.
 
@@
 
-```{r Figure1, dependson=c("assemble-models"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
-Graph shows the expected value for the function $f$ under each model.
-Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.
-The training data is also shown as black points.
-(The GP is conditioned on 0,0, shown as a pseudo-data point).
-", include=FALSE}
-```
-```{r Figureb, dependson=c("assemble-models", "par-fns"), include=TRUE}
-```
-
-Figure 1 shows the mean inferred state space dynamics of each model
-relative to the true model used to generate the data, predicting the
-relationship between observed stock size (x-axis) to the stock size
-after recruitment the following year.  All models except the MLE model
-estimate a distribution around the means shown here, and all models 
-estimate a level of process noise, which is independent of the state
-value (x).  Note that in contrast to the  other models shown, the mean
-Gaussian process corresponds to a distribution of curves - as indicated
-by the gray band - which itself has a mean shown in black.  Note that 
-this mean GP is thus more certain of the dynamics in the region where
-data is available then where it is not.
-
-While it would be natural (and straight forward) to condition
-the GP on passing through the origin (0,0) (see appendix), the estimate
-shown here is based only on the observed data. The observed data from
-which each model is estimated is also shown.  The observations come
-from only a limited region of state space corresponding to unharvested
-or weakly harvested system.  No observations occur at the theoretical
-optimum or near the tipping point.
-
-```{r Figure2, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", include=TRUE}
-```
-
-The resulting optimal management strategy based on each of the inferred
-models is shown in Figure 2, against the optimal strategy given the
-true underlying dynamics.  Policies are shown in terms of target
-escapement, $S_t$.  Under models such as this a constant escapement
-policy is expected to be optimal [@Reed1979], whereby population levels
-below a certain size $S$ are unharvested, while above that size the harvest
-strategy aims to return the population to $S$, resulting in the hockey-stick
-shaped policies shown.
-
-```{r Figure3, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.",  include=TRUE}
-```
-
-The consequences of managing 100 replicate realizations of the 
-simulated fishery under each of the policies estimated is shown in Figure 3.  As expected
-from the policy curves, the structurally correct model under-harvests,
-leaving the stock to vary around it's un-fished optimum.  The structurally
-incorrect Ricker model over-harvests the population passed
-the tipping point consistently, resulting in the immediate crash of the stock and 
-thus derives minimal profits.
-
-The results shown in Figures 1-3 are not unique to the simulated data or models chosen
-here, but arises across a range of parameter values and simulations as shown in the 
-supplemental figures.  The results across this range can most easily be compared 
-by the relative differences in net present value realized by each of the approaches,
-as shown in Figure 4.  The Gaussian Process most consistently realizes a value 
-close to the optimal solution, and importantly avoids ever driving the system across
-the tipping point, which results in the near-zero value cases in the parametric models.
-
-```{r Figure4, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
-For each data set, the three models are estimated as described above.
-Values plotted are the averages of a given policy over 100 replicate simulations.
-Details and code provided in the supplement.",  include=TRUE}
-```
-
 Discussion 
@@
 #### Big picture: Linking GP to SDP
+
+_rambling_
 
@@
 <!-- On mechanistic models -->
-
-Though simple mechanistically motivated models
-may be best for the latter case [@Cuddington2013; @Geritz2012], such models 
-can be not only inaccurate but misleading in quantitative decision making. This 
-arises for two reasons:
-
-1. We do not know what the correct parameter values are for the models.
-1. We do not know what the correct models are for ecological systems.
+<!-- { Seems kind of irrelevant }
 
@@
 
-While non-parametric Bayesian approaches will not always be preferable 
-to simple mechanistic models, we highlight three aspects of the problem
-consider here that make these methods particularly valuable.  These aspects
-are common to many conservation decision making problems, which thus merit
-greater use of non-parametric approaches that can best take advantage of them.
-
-### 1. Large uncertainty where the data is poor
-
-The parametric models perform worst when they propose a management strategy
-outside the range of the observed data. The non-parametric Bayesian approach, 
-in contrast, allows a predictive model that expresses a great deal of uncertainty
-about the probable dynamics outside the observed range, while retaining very
-good predictive accuracy in the range observed.  The management policy 
-dictated by the GP balance this uncertainty against
-the immediate value of the harvest, and act to stabilize the population 
-dynamics in a region of state space in which the predictions can be 
-reliably reflected by the data.
-
-### 2. Predictive accuracy where data is good
-
-While expressing larger uncertainty outside the observed data, the GP
-can also provide a better fit with smaller uncertainty inside the range
-of the observed data. This arises from the greater flexibility of the 
-Gaussian process, which describes a large family of possible curves.  
-While in a parametric context this over-fitting would be more worrisome --
-a high-degree polynomial could fit the data even better -- those 
-concerns are driven by the resulting parametric fit outside the data,
-which may involve wild oscillations unsupported by the data.  As we 
-have seen in #1, the GP is less vulnerable to such unjustified predictions
-outside the data, and is meanwhile free to benefit from the greater 
-fit where the data is available.
-
-Future directions
------------------
-
-### Higher dimensions
-
-In this simulated example, the underlying
-dynamics are truly governed by a simple parametric model, allowing
-the parametric approaches to be more accurate.  Similarly, because the
-dynamics are  one-dimensional dynamics and lead to  stable nodes (rather
-than other attractors such as limit-cycles resulting in oscillations),
-the training data provides relatively limited information about the
-dynamics.  For these reasons, we anticipate that in higher-dimensional
-examples characteristic of ecosystem management problems that the machine
-learning approach will prove even more valuable.
-
-### Online learning
-
-In our treatment here we have ignored the possibility of learning during the 
-management phase, in which the additional observations of the stock size could
-potentially improve parameter estimates.  While we intend to address this 
-possibility in future work in the context of these non-parametric models,
-we have not addressed it here for pedagogical reasons. In the context presented
-here, it is clear that the differences in performance arise from differences
-in the uncertainty inherent in the model formulations, rather than from 
-differing abilities to learn.  Because we consider a threshold system, 
-online learning would not change this generic feature of a lack of data in a
-certain range of the state space which is better captured by the Gaussian process.
-
-Acknowledgments
-===============
-
-This work was partially supported by the Center for Stock Assessment
-Research, a partnership between the University of California Santa Cruz
-and the Fisheries Ecology Division, Southwest Fisheries Science Center,
-Santa Cruz, CA and by NSF grant EF-0924195 to MM.
-
-Appendix
-========
-
-The appendices have not yet been assembled.
-Meanwhile, code to repeat the analyses, along with a complete log of all research conducted on this project, can be found at: [https://github.com/cboettig/nonparametric-bayes](https://github.com/cboettig/nonparametric-bayes/)
+-->
+
+----------------------------------------
+
+Code to replicate the analysis, along with complete log of this research can be found at: [https://github.com/cboettig/nonparametric-bayes](https://github.com/cboettig/nonparametric-bayes/)
 
@@
 ```
-
-References
-==========

### 2013-09-20 inst/doc/nonparametric-bayes.tex
writing

@@
 
+Model uncertainty and limited data coverage are fundamental challenges
+to robust ecosystem management. These challenges are acutely highlighted
+by concerns that many ecological systems may contain tipping points.
+Before a collapse, we do not know where the tipping points lie, if the
+exist at all. Hence, we know neither a complete model of the system
+dynamics nor do we have access to data in some large region of
+state-space where such a tipping point might exist. These two sources of
+uncertainty frustrate state-of-the-art parametric approaches to decision
+theory and optimal control. I will illustrate how a non-parametric
+approach using a Gaussian Process prior provides a more flexible
+representation of this inherent uncertainty. Consequently, we can adapt
+the Gaussian Process prior to a stochastic dynamic programming framework
+in order to make robust management predictions under both model and
+uncertainty and limited data.
+
 \section{Introduction}
 
-Briefly set the stage: (see related papers)
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Decision-theory / optimal control framework
-\item
-  Fisheries context
-\end{itemize}
+The sudden collapse of fisheries and other ecosystems is an increasingly
+widespread phenomenon and a pressing concern for ecological management
+and conservation. Ecological dyanaics are frequently complex and
+difficult to measure, making uncertainty in our understanding a
+prediction a persistent challenge to effective management.
+
+Decision-theoretic approaches provide a framework to determine the best
+sequence of actions in face of uncertainty, but only when that
+uncertainty can be meaningfully quantified {[}@Fischer2009{]}.
+
+Uncertainty enters at many levels management
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Why fisheries
+\end{itemize}
+
+The economic value and ecological concern have made marine fisheries the
+crucible for much of the founding work (Gordon 1954; Reed 1979; May et
+al. 1979; Ludwig and Walters 1982) in managing ecosystems under
+uncertainty. Global trends (Worm et al. 2006) and controversy (Hilborn
+2007; Worm et al. 2009) have made understanding these challenges all the
+more pressing.
 
@@
   Dynamic uncertainty (stochasticity)
+\end{itemize}
+
+From \href{}{Clark (1976)} to \href{}{Reed (1979)}
+
+Other sources of uncertainty \href{}{Roughgarden \& Smith 1996},
+\href{}{Sethi \emph{et al.} (2005)}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
 \item
   Parametric uncertainty 
+\end{itemize}
+
+As the parameter values for these models must be estimated from limited
+data,\\there will always be some uncertainty associated with these
+values. This uncertainty further compounds the intrinsic variability
+introduced by demographic or environmental noise. The degree of
+uncertainty in the parameter values can be inferred from the data and
+reflected in the estimates of the transition probabilities.
+
+(Mangel 1985; Schapaugh and Tyre 2013)
+
+\href{http://doi.org/10.1016/0304-38008290023-0}{Ludwig \& Walters 1982}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
 \item
@@
 
+Estimates of parameter uncertainty are only as good as the parametric
+models themselves. Often we do not understand the system dynamics well
+enough to know if a model provides a good approximation over the
+relevant range of states and timescales (criteria that we loosely refer
+to as defining the ``right'' or ``true'' model.) So called structural or
+model uncertainty is a more difficult problem than parametric
+uncertainty. Typical solutions involve either model choice, model
+averaging, or introducing yet greater model complexity of which others
+may be special cases (model averaging being one such way to construct
+such a model) {[}B. K. Williams (2001); Athanassoglou and Xepapadeas
+(2012); @Cressie2009{]}. Even setting aside other computational and
+statistical concerns (e.g. {[}@Cressie2009{]}), these approaches do not
+address our second concern - representing uncertainty outside the
+observed data range.
+
+\href{http://dx.doi.org/10.1890/07-0744.1}{@Cressie2009}.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Limits of state space
+\end{itemize}
+
+Often the set of feasible states is larger than the range of the
+observed states. Though a system can enter a state outside previously
+observed values due to stochastic effects alone (such as a particularly
+low or high nutrient year), this concern is most keenly felt in
+considering the value of actions that have not yet been applied.
+
+Of all sources of uncertainty, this has potentially recieved the least
+attention. It has most often appeared in the concept of active learning
+or active adaptive managment, (adaptive probing, ) where it may be
+valauble to intentionally force a system far from the observed values
+even when the expected value such actions is low, as it provides much
+faster learning and consequent reduction of model uncertainty that can
+allow greater value to be derived later on. Ludwig and Walters (1982)
+show that it may be advantageous to fish an unexploited population very
+heavily at first to obtain a better estimate of the recruitment rate.
+This intuitive strategy when a population is governed by a Ricker or
+Beverton-Holt-like dynamic would clearly be disasterous if instead the
+dynamics contained an unforseen tipping point. The best way to learn
+where the edge lies may be to walk up to it, but it is also the most
+dangerous.
+
 \subsection{Two central challenges:}
@@
 \item
+  MAP
+\end{itemize}
+
+Decision making under uncertainty is a ubiquitous challenge of natural
+resource management and conservation.
+\href{http://doi.org/10.1111/faf.12037}{@Glaser2013}
+
+Here we illustrate how a stochastic dynamic programming (SDP) algorithm
+can be driven by the predictions from a Gaussian Process (GP). This
+provides two distinct advantages compared with contemporary approaches.
+First, using a GP sidesteps the need for an accurate model-based
+description of the system dynamics (or a model choice or model-averaging
+approach which introduces yet more difficulty).\\Second, unlike
+parametric models which can only reflect uncertainty in the
+corresponding confidence intervals or posterior distributions of their
+parameters, the GP approach provides a state-space dependent
+representation of uncertainty. This permits a much greater uncertainty
+far from the observed data than near the observed data. These features
+allow the GP-SDP approach to find robust management solutions in face of
+both limited data and no knolwedge of correct model structure.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
   Note on ``not magic'': honest uncertainty + SDP
-\item
-  Note on comparing models (via value function rather than by ``fit'')
-\item
-  Note on terminology: ``Non-parametric''
+\end{itemize}
+
+The idea that any approach can perform well without either having to
+know the model or have particularly good data should immediately draw
+suspicion. The reader must bear in mind that the strength of our
+approach comes not from black-box predictive power from such limited
+information, but rather, by providing a more honest expression of
+uncertainty outside the observed data without sacrificing the predictive
+capacity near the observed data. By coupling this more accurate
+description of what is known and unknown to the decision-making under
+uncertainty framework provided by stochastic dynamic programming, we are
+able to obtain more robust management policies than with common
+parameteric modeling approaches.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Note on comparing models (via value function rather than by ``fit'').
+  \emph{Do we need this?}
+\end{itemize}
+
+The nature of decision-making problems provides a convenient way to
+compare models. Rather than compare models in terms of best fit to data
+or fret over the appropriate penalty for model complexity, model
+performance is defined in the concrete terms of the decision-maker's
+objective function, which we will take as given. (Much argument can be
+made over the `correct' objective function, e.g.~how to account for the
+social value of fish left in the sea vs.~the commercial value of fish
+harvested; see Halpern et al. (2013) for further discussion of this
+issue. Alternatively, we can always compare model performance across
+multiple potential objective functions.) The decision-maker does not
+necessarily need a model that provides the best mechanistic
+understanding or the best long-term outcome, but rather the one that
+best estimates the probabilities of being in different states as a
+result of the possible actions.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Note on terminology: ``Non-parametric'' \emph{Do we need this?}
 \end{itemize}
@@
 
-\[ \max_{h(t)} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h) \delta^t  \]
+\[ \max_{h_t} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h_t) \delta^t  \]
 
@@
 Eacho of our models $f(S_t)$ must be estimated from training data, which
-we simulate from the Allen model with parameters $r = 2$, $K=8$, $C=5$
-and $\sigma_g = 0.1$ for $T = 40$ timesteps, starting at initial
-condition $X_0 = $.
+we simulate from the Allen model with parameters $r = $
+\texttt{r p{[}1{]}}, $K =$ \texttt{r p{[}2{]}}, $C =$
+\texttt{r p{[}3{]}}, and $\sigma_g =$ \texttt{r sigma\_g} for $T=$
+
+\begin{verbatim}
+
+Error in base::parse(text = code, srcfile = NULL) : 1:16: unexpected ','
+1: Tobs$ timesteps,
+                  ^
+\end{verbatim}
+
+r Xo`.
 
@@
 
+While methods that account for \emph{parametric} uncertainty can be very
+successful with the right model, structural uncertainty of not knowing
+what model best approximates the dynamics poses considerably greater
+difficulty.
+
 \section{Introduction}
@@
 \paragraph{Opening}
-
-Decision making under uncertainty is a ubiquitous challenge of natural
-resource management and conservation.\\Here we illustrate how a
-stochastic dynamic programming (SDP) algorithm can be driven by the
-predictions from a Gaussian Process (GP). This provides two distinct
-advantages compared with contemporary approaches. First, using a GP
-sidesteps the need for an accurate model-based description of the system
-dynamics (or a model choice or model-averaging approach which introduces
-yet more difficulty).\\Second, unlike parametric model-fitting
-approaches, which can only reflect uncertainty in the corresponding
-confidence intervals or posterior distributions of the parameters, the
-GP approach provides a state-space dependent representation of
-uncertainty. In other words, a GP estimated from a given data set
-permits a much greater uncertainty far from the observed data than it
-does near the observed data.
-
-The idea that any approach can perform well without either having to
-know the model or have particularly good data should immediately draw
-suspicion. The reader must bear in mind that the strength of our
-approach comes not from black-box predictive power from such limited
-information, but rather, by providing a more honest expression of
-uncertainty outside the observed data without sacrificing the predictive
-capacity near the observed data.
-
-By coupling this more accurate description of what is known and unknown
-to the decision-making under uncertainty framework provided by
-stochastic dynamic programming, we are able to obtain more robust
-management policies than with common parameteric modeling approaches.
 
@@
 
-As the parameter values for these models must be estimated from limited
-data,\\there will always be some uncertainty associated with these
-values. This uncertainty further compounds the intrinsic variability
-introduced by demographic or environmental noise. The degree of
-uncertainty in the parameter values can be inferred from the data and
-reflected in the estimates of the transition probabilities.
-
-(Mangel 1985; Schapaugh and Tyre 2013)
-
 \paragraph{Structural Uncertainty}
 
-Unfortunately, estimates of parameter uncertainty are only as good as
-the parametric models themselves. Too often, we do not understand the
-system dynamics well enough to know if a model provides a good
-approximation over the relevant range of states and timescales (criteria
-that we loosely refer to as defining the ``right'' or ``true'' model.)
-So called structural or model uncertainty is a more difficult problem
-than parametric uncertainty. Typical solutions involve either model
-choice or model averaging. Either approach remains limited by the degree
-to which any of the proposed models are sufficiently close to the right
-model.
-
 (B. K. Williams 2001; Athanassoglou and Xepapadeas 2012).
-
-\paragraph{Comparing models}
-
-The nature of decision-making problems provides a convenient way to
-compare models. Rather than compare models in terms of best fit to data
-or fret over the appropriate penalty for model complexity, model
-performance is defined in the concrete terms of the decision-maker's
-objective function, which we will take as given. (Much argument can be
-made over the `correct' objective function, e.g.~how to account for the
-social value of fish left in the sea vs.~the commercial value of fish
-harvested; see Halpern et al. (2013) for further discussion of this
-issue. Alternatively, we can always compare model performance across
-multiple potential objective functions.) The decision-maker does not
-necessarily need a model that provides the best mechanistic
-understanding or the best long-term outcome, but rather the one that
-best estimates the probabilities of being in different states as a
-result of the possible actions.
 
@@
 
-\paragraph{Background on Fisheries Context}
-
-\paragraph{Background on Tipping points}
-
-Non-parametric models can better represent uncertainties outside of
-observed data while also better capturing the dynamics in the region
-observed. Advances in the theory and computational implementations of
-nonparametric methods make them ripe for such applications. We use the
-classic problem of optimal harvest of a marine fishery to illustrate how
-the nonparametric approach of Gaussian processes can be applied. We will
-compare Bayesian implementations of both nonparametric and parametric
-models, which best allow us to capture the uncertainty of model
-estimation in either case and permits a more natural comparison between
-approaches.
-
 \section{Approach and Methods}
@@
 value of the resource, given an estimated stock size from the year
-before. The economic value and ecological concern have made marine
-fisheries the crucible for much of the founding work (Gordon 1954; Reed
-1979; May et al. 1979; Ludwig and Walters 1982) in managing ecosystems
-under uncertainty. Global trends (Worm et al. 2006) and controversy
-(Hilborn 2007; Worm et al. 2009) have made understanding these
-challenges all the more pressing.
+before.
 

### 2013-09-23 inst/doc/nonparametric-bayes.Rmd
tackling intro yet again

@@
 
-<!--
-Very briefly set the stage: (see related papers)
-- Decision-theory / optimal control framework
-- Fisheries context 
--->
-
-<!-- FIXME -->
+Decision making under uncertainty is a ubiquitous challenge of natural 
+resource management and conservation.
 
@@
 
-Uncertainty enters at many levels <!-- enters what -->
-management
-
-- Why fisheries
-
-The economic value and ecological concern have made marine
-fisheries the crucible for much of the founding work [@Gordon1954;
-@Reed1979; @May1979; @Ludwig1982] in managing ecosystems under
-uncertainty.  Global trends [@Worm2006] and controversy [@Hilborn2007;
-@Worm2009] have made understanding these challenges all the more pressing.
-
-<!-- Other reasons for failure see Sethi-->
-
-@Fischer 
-@Sethi2005, @Peretti, @Munch,
-
-(Classical division)
-
-- Dynamic uncertainty (stochasticity)
-
-From [Clark (1976)]() to [Reed (1979)]()
-
-Other sources of uncertainty [Roughgarden & Smith 1996](), [Sethi _et al._ (2005)]()
-
-- Parametric uncertainty <!-- ugh, be more consise and less annoying-->
+<!-- Uncertainty outside the data without the correct model has not been handled. -->
+
+Uncertainty enters the decision-making process at many levels: intrinsic 
+stochasticity in biological processes, measurements, and implementation of 
+policy [_e.g._ @Reed1979; @Clark1986; @Roughgarden1997; @Sethi2005], parameteric 
+uncertainty [_e.g._ @Ludwig1982; @Hilborn1997; @McAllister1998; @Schapaugh2013],
+and model or structural uncertainty [_e.g._ @Williams2001; @Cressie2009;  @Athanassoglou2012].
+Of these, structural uncertainty incorporates the least a priori knowledge or 
+assumptions and is generally the hardest to quantify. Typical approaches assume 
+a weak notion of uncertainty where a correct or reasonable approximation of the dynamics
+must be identified from among a handful of alternative models.  Here we consider an approach
+that addresses uncertainty at each of these levels without assuming the dynamics follow
+a particular (i.e. parametric) structure.
+
+<!-- What do we call this?  Extrapolation uncertainty?  Pathological Uncertainty? -->
+An additional source of uncertainty that has recieved less attention[^1] arises <!-- remove 'typical'? -->
+when applying a dynamical model outside the range of data on which it has been
+estimated.  This extrapolation uncertainty is felt most keenly in decision-theoretic
+applications, as (a) exploring the potential action space typically involves considering
+actions that may move the system outside the range of observed behavior, and (b) decision-theoretic
+alogrithms rely not only on reasonable estimates of the expected outcomes, but depend on
+the weights given to all possible outcomes [_e.g._ @Weitzman2013].
+
+[^1]: The concept of adaptive probing is one area that has explicitly
+addressed this kind of uncertainty, reaching rather opposite conclusions
+than what we observe here. Adaptive probing strategies follow from "Dual
+Control" or "Active Adaptive Management" approaches (e.g. @Ludwig1982)
+that can trade off short term utility by choosing actions that can reduce
+uncertainty.  Adaptive probing strategies arise when it is valuable to
+intentionally force a system far from the observed values even when the
+expected value such actions is low, as it provides much faster learning
+and consequent reduction of model uncertainty that can allow greater value
+to be derived later on.  For instance, @Ludwig1982 show that it may be
+advantageous to fish an unexploited population very heavily at first
+to obtain a better estimate of the recruitment rate.  This intuitive
+strategy when a population is governed by a Ricker or Beverton-Holt-like
+dynamic would clearly be disastrous if instead the dynamics contained
+an unforeseen tipping point.  The best way to learn where the edge lies
+may be to walk up to it, but it is also the most dangerous.
+
+_Map_
+
+Here we illustrate how a stochastic dynamic programming (SDP) algorithm [@Mangel1988; @Possingham1997, Marescot2013] can 
+be driven by the predictions from a Bayesian non-parametric (BNP) approach [@Munch2005a].  
+This provides two distinct advantages compared with contemporary approaches.
+First, using a BNP sidesteps the need for an accurate model-based description
+of the system dynamics.  
+Second, the BNP can better reflect uncertainty that arises when extrapolating a model 
+outside of the data on which it was fit.  
+We illustrate that when the correct model is not known, this latter
+feature is crucial to providing a robust decision-theoretic approach in
+face of substantial structural uncertainty.
+
+Often the set of feasible states is larger than the range of the observed
+states.  Though most systems subject to some degree of stochasticity
+can enter a state outside previously observed values by chance alone
+(such as a particularly low or high nutrient year), this concern is
+most keenly felt in the decision-theoretic context. If we are observing
+the fluctuations of a given fish stock over many years under a fixed
+harvesting pressure, we might develop and test a model that could
+reasonably predict the frequency of a deviation of a given size, even
+when such a deviation has not been previously observed.
+(The same analogy might be made to predicting hurricanes or other
+extreme weather events from historical data, which may be more accurately
+predicted in a constant climate than in one being perturbed).
+
+Given adequate observations of a system
+
+One must consider actions or policy under which the system has not yet
+been observed,
+
+<!-- More Consise versions above
+
+- Parametric uncertainty
 
@@
 noise.  The degree of uncertainty in the parameter values can be inferred from the data
-and reflected in the estimates of the transition probabilities.
-
-[@Mangel1985; @Schapaugh2013]
-
-[Ludwig & Walters 1982](http://doi.org/10.1016/0304-3800(82)90023-0 "Optimal harvesting with imprecise parameter estimates")
-
-- Structural uncertainty <!-- ugh, be more consise and less annoying-->
+and reflected in the estimates of the transition probabilities [@Walters1982; @Mangel1988; @Mangel1997; @Schapaugh2013].
+
+- Structural uncertainty
 
@@
 
-[@Cressie2009](http://dx.doi.org/10.1890/07-0744.1 "Accounting for uncertainty in ecological analysis: the strengths and limitations of hierarchical statistical modeling.").
-
 - Limits of state space
 
-Resource management and conservation planning seek to determine the optimal set of feasible actions to maximize the value of some objectives (see [@Halpern2012])  Process error, measurement error, implementation error [@Reed1979, @Clark1986, @Roughgarden1996, @Sethi2005].
-These sources of stochasticity in turn mean that model parameters can only be estimated approximately, requiring parametric uncertainty also be considered [@Ludwig1982, ].
-
-The three sources of uncertainty: stochasticity (in process, measurement, and implementation; though we focus on process-noise only for simplicity), parametric uncertainty, and structural uncertainty, are familiar challenges.
-
-Of all sources of uncertainty, this last has received the least attention.[^1]
-
-Often the set of feasible states is larger than the range of the observed states.
-Though most systems subject to some degree of stochasticity can enter a state outside previously observed values by chance alone (such as a particularly low or high nutrient year), this concern is most keenly felt in the decision-theoretic context.
-If we are observing the fluctuations of a given fish stock over many years under a fixed harvesting pressure, we might develop and test a model that could reasonably predict the frequency of a deviation of a given size, even when such a deviation has not been previously observed.
-
-(The same analogy might be made to predicting hurricanes or other extreme weather events from historical data, which may be more accurately predicted in a constant climate than in one being perturbed).
-
-Given adequate observations of a system
-
-One must consider actions or policy under which the system has not yet been observed,
-
-<!-- What do we call this?  Extrapolation uncertainty?  Pathological Uncertainty? -->
-
-[^1]: The concept of adaptive probing is one area that has explicitly addressed this kind of uncertainty, reaching rather opposite conclusions than what we observe here.
-Adaptive probing strategies follow from "Dual Control" or "Active Adaptive Management" approaches (e.g.
-@Ludwig1982) that can trade off short term utility by choosing actions that can reduce uncertainty.
-Adaptive probing strategies arise when it is valuable to intentionally force a system far from the observed values even when the expected value such actions is low, as it provides much faster learning and consequent reduction of model uncertainty that can allow greater value to be derived later on.
-For instance, @Ludwig1982 show that it may be advantageous to fish an unexploited population very heavily at first to obtain a better estimate of the recruitment rate.
-This intuitive strategy when a population is governed by a Ricker or Beverton-Holt-like dynamic would clearly be disastrous if instead the dynamics contained an unforeseen tipping point.
-The best way to learn where the edge lies may be to walk up to it, but it is also the most dangerous.
+Resource management and conservation planning seek to determine the
+optimal set of feasible actions to maximize the value of some objectives
+(see [@Halpern2012])  Process error, measurement error, implementation
+error [@Reed1979, @Clark1986, @Roughgarden1996, @Sethi2005].  These
+sources of stochasticity in turn mean that model parameters can only
+be estimated approximately, requiring parametric uncertainty also be
+considered [@Ludwig1982, ].
+
+-->
 
@@
 
-Decision making under uncertainty is a ubiquitous challenge of natural 
-resource management and conservation. 
-Here we illustrate how a stochastic dynamic programming (SDP) algorithm can 
-be driven by the predictions from a Gaussian Process (GP).
-This provides two distinct advantages compared with contemporary approaches.
-First, using a GP sidesteps the need for an accurate model-based description
-of the system dynamics (or a model choice or model-averaging approach which
-introduces yet more difficulty).  
-Second, unlike parametric models which can only reflect 
-uncertainty in the corresponding confidence intervals or posterior distributions
-of their parameters, the GP approach provides a state-space dependent representation
-of uncertainty.  This permits a much greater uncertainty far from the observed data than near the observed 
+<!-- necessary? -->
+This paper represents the first time the SDP decision-making framework has been used
+without an a priori model of the underlying dynamics through the use of the BNP approach.
+
+In contrast to parametric models which can only reflect uncertainty 
+in parameter estimates, the BNP approach provides a more state-space dependent representation
+of uncertainty.
+
+This permits a much greater uncertainty far from the observed data than near the observed 
 data.  These features allow the GP-SDP approach to find robust management solutions 
@@
 
+- Note on: Why fisheries
+
+The economic value and ecological concern have made marine
+fisheries the crucible for much of the founding work [@Gordon1954;
+@Reed1979; @May1979; @Ludwig1982] in managing ecosystems under
+uncertainty.  Global trends [@Worm2006] and controversy [@Hilborn2007;
+@Worm2009] have made understanding these challenges all the more pressing.
+
 <!--move to  much later -->
-- Note on terminology: "Non-parametric"
-
-The term non-parametric to describe Gaussian processes (GP) is unfortunate,
-given that the GP models still involve the estimation of parameters. Some 
-literature thus uses the term semi-parametric, which merely adds ambiguity 
-to the confusion.  Statisticians give two distinct meanings to the term: (a)
-referring to a method that does not assume a distribution (e.g. non-parametric bootstrap, order statistics),
-and (b) one that does not assume a fixed structure but rather depends on the 
-size of the data (e.g. non-parametric regression, non-parametric hierarchical models
-such as the Dirichlet process).  Our use is in the latter sense.
-
-This non-parametric property -- having a structure that is not fixed but 
-rather dependent on the data is precisely the property that makes this approach
-attractive in face of the limited data sampling challenges discussed above.  
-Having fit a parametric model to some data, the model is completely described
-by the values (or posterior distributions) of it's parameters.  The non-parametric 
-model is not captured by its parameter values or distributions alone -- the 
-data points become part of the model.
 
@@
 
+- Background on non-parametric modeling.
+
+Addressing the difficulty posed by extrapolation without knowing the
+true model requires a nonparametric approach to model fitting: one
+that does not assume a fixed structure but rather depends on the size
+of the data (e.g. non-parametric regression or a Dirichlet process).
+This established terminology is nevertheless unfortunate, as (a)
+this approach still involves the estimation of parameters, and (b),
+Statisticians use non-parametric to mean both this property (structure
+is not fixed by the parameters) and an entirely different (and probably
+more familiar) case in which the model does not assume any distribution
+(e.g. non-parametric bootstrap, order statistics).  Some literature
+thus uses the term semi-parametric, which merely adds ambiguity to
+the confusion.
+
+This non-parametric property -- having a structure explicitly dependent on
+the data -- is precisely the property that makes this approach attractive
+in face of the limited data sampling challenges discussed above.
+Having fit a parametric model to some data, the model is completely
+described by the values (or posterior distributions) of it's parameters.
+The non-parametric model is not captured by its parameter values or
+distributions alone. Either the model scales with the complexity of the data
+on which it is estimated (e.g. nonparametric heirarchical approaches such
+as the Dirchlet process) or the data points become themselves part of the model
+specification, as in the nonparametric regression used here.  we shall see here.
+
 - Definition
+
 - Previous application
+
+The use of Gaussian process (GP) regression (or "kriging" in the geospatial
+literature) to formulate a predictive model is relatively new in the
+context of modeling dynamical systems [@Kocijan2005], and was first introduced
+in the context ecological modeling and fisheries management in @Munch2005.
+An accessible and thorough introduction to the formulation and use of
+GPs can be found in @Rasmussen2006.
+
 - Why it is particularly suited to these two problems
 - (Why this is a novel application thereof)
-
-Approach and Methods
-====================
-
-### Summary of approach
-
-### Statement of the optimal control problem
-
-- Underlying model 
-- Available data 
-- Value function
-
-For simplicity we assume profit is simply linear in the realized harvest (only
-enforcing the restriction that harvest can not exceed available stock)
-
-### Parametric models
-
-- Statement of the models
-
-We consider three candidate parametric models of the stock-recruitment
-dynamics: The Ricker model, the Allen model [Allen 2005](), the Myers
-model. The familiar Ricker model involves two parameters, corresponding
-to a growth rate and a carrying capacity, and cannot support alternative
-stable state dynamics (though as growth rate increases it exhibits a
-periodic attractor that proceeds through period-doubling into chaos. We
-will generally focus on dynamics below the chaotic threshold for
-the purposes of this analysis.) The Allen model resembles the Ricker
-dynamics with an added Allee effect parameter [Courchamp](), below
-which the population cannot persist.  The Myers model also has three
-parameters and contains an Allee threshold, but has compensatory rather
-than over-compensatory density dependence (resembling a Beverton-Holt
-curve rather than a Ricker curve at high densities.)
-
-We assume multiplicative log-normal noise perturbs the growth predicted 
-by the each of the deterministic model skeletons described above. This 
-introduces one additional parameter $\sigma$ that must be estimated by each
-model.
-
-<!-- equations just in appendix? -->
-
-As we simulate training data from the Allen model (ref section), we will
-refer to this as the structurally correct model.  The Ricker model is
-thus a reasonable approximation of these dynamics far from the Allee
-threshold (but lacks threshold dynamics), while the Myers model shares
-the essential feature of a threshold but differs in the structure. Thus
-we have three potential parametric models of the stock dynamics.
-
-- Bayesian inference of parametric models
-
-We infer posterior distributions for the parameters of each model
-in a Bayesian context using Gibbs sampling (implemented in R [@RTeam]
-using jags, [@R2jags]).  We choose uninformative uniform priors for all
-parameters (See Appendix, Figures S1-S3, and Table S1, and the R code
-provided). One-step-ahead predictions of these model fits are shown in
-Figure 1.
-
-- SDP via parametric models
-
-An optimal policy function is then inferred through stochastic dynamic
-programming for each model given the posterior distributions of the
-parameter estimates.  This policy maximizes the expectation of the value
-function integrated over the parameter uncertainty. (code implementing
-this algorithm provided in the Appendix).
-
-### The Gaussian Process model
-
-- Statement of model
-
-... more on GP ... [Munch 2005]()
-
-We also estimate a simple Gaussian Process defined by
-a radial basis function kernel of two parameters: $\ell$, which gives
-the characteristic length-scale over which correlation between two 
-points (e.g. any two points $X_t, X_{t+1}$, and $X_{t+\tau}, X_{t+1+\tau}$)
-in state-space decays, and $\sigma$, which gives the scale of the 
-process noise by which observations $Y_{t+1}$ may differ from their
-predicted values $X_{t+1}$ given an observation of the previous state,
-$X_t$.
-
-- Inference of the model
-
-Also unlike parametric models, this posterior distribution is still
-conditional on the training data. As such, the uncertainty near the
-observed data.
-
-We use a Metropolis-Hastings Markov Chain Monte Carlo to infer posterior
-distributions of the two parameters of the GP (Figure S4, code in
-appendix), under weakly informative Gaussian priors (see parameters in
-table S5). As the posterior distributions differ substantially from the
-priors (Figure S4), we can be assured that most of the information in
-the posterior comes from the data rather than the prior belief.
-
-- SDP via the model
-
-Though we are unaware of prior application of this type, it is reasonably
-straight-forward to adapt the Gaussian Process for Stochastic Dynamic
-Programming.  Recall that unlike the parametric models the Gaussian
-process with fixed parameters already predicts a distribution of
-curves rather than a single curve. We must first integrate over 
-this distribution of curves given a sampling of parameter values drawn
-from the posterior distribution of the two GP parameters, before
-integrating over the posterior of those parameters themselves.
-
-Results
-=======
-
-```{r Figure1, dependson=c("assemble-models"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
-Graph shows the expected value for the function $f$ under each model.
-Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.
-The training data is also shown as black points.
-(The GP is conditioned on 0,0, shown as a pseudo-data point).
-", include=FALSE}
-```
-```{r Figureb, dependson=c("assemble-models", "par-fns"), include=TRUE}
-```
-
-Figure 1 shows the mean inferred state space dynamics of each model
-relative to the true model used to generate the data, predicting the
-relationship between observed stock size (x-axis) to the stock size
-after recruitment the following year.  All models except the MLE model
-estimate a distribution around the means shown here, and all models 
-estimate a level of process noise, which is independent of the state
-value (x).  Note that in contrast to the  other models shown, the mean
-Gaussian process corresponds to a distribution of curves - as indicated
-by the gray band - which itself has a mean shown in black.  Note that 
-this mean GP is thus more certain of the dynamics in the region where
-data is available then where it is not.
-
-While it would be straight forward to condition the GP on passing through the origin (0,0), the estimate
-shown here is based only on the observed data. The observed data from
-which each model is estimated is also shown.  The observations come
-from only a limited region of state space corresponding to unharvested
-or weakly harvested system.  No observations occur at the theoretical
-optimum harvest rate or near the tipping point.
-
-```{r Figure2, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", include=TRUE}
-```
-
-The resulting optimal management strategy based on each of the inferred
-models is shown in Figure 2, against the optimal strategy given the
-true underlying dynamics.  Policies are shown in terms of target
-escapement, $S_t$.  Under models such as this a constant escapement
-policy is expected to be optimal [@Reed1979], whereby population levels
-below a certain size $S$ are unharvested, while above that size the harvest
-strategy aims to return the population to $S$, resulting in the hockey-stick
-shaped policies shown.
-
-```{r Figure3, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.",  include=TRUE}
-```
-
-The consequences of managing 100 replicate realizations of the 
-simulated fishery under each of the policies estimated is shown in Figure 3.  As expected
-from the policy curves, the structurally correct model under-harvests,
-leaving the stock to vary around it's un-fished optimum.  The structurally
-incorrect Ricker model over-harvests the population passed
-the tipping point consistently, resulting in the immediate crash of the stock and 
-thus derives minimal profits.
-
-The results shown in Figures 1-3 are not unique to the simulated data or models chosen
-here, but arises across a range of parameter values and simulations as shown in the 
-supplemental figures.  The results across this range can most easily be compared 
-by the relative differences in net present value realized by each of the approaches,
-as shown in Figure 4.  The Gaussian Process most consistently realizes a value 
-close to the optimal solution, and importantly avoids ever driving the system across
-the tipping point, which results in the near-zero value cases in the parametric models.
-
-```{r Figure4, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
-For each data set, the three models are estimated as described above.
-Values plotted are the averages of a given policy over 100 replicate simulations.
-Details and code provided in the supplement.",  include=TRUE}
-```
-
-### Figure 1: Fitted Models
-
-- All models fit the data quite well
-- Information criteria would pick the simple, incorrect model.
-
-### Figure 2: Inferred Policies
-
-- Inferred policies differ substantially among models
-- The structurally correct model and the GP are close to the true model
-- alternatives are not close
-
-### Figure 3: Simulated results
-
--
-
-### Figure 4: Robustness
-
-- Results hold across range of parameters
-
-(eek, distinguish better between result and discussion?)
-
-Discussion 
-==========
-
-- All models are "good fits" to the originally observed data.
-
-- (Simple model choice immediately leads us astray)
-
-Though simple mechanistically motivated models offer the greatest potential 
-to increase our basic understanding of ecological processes [@Cuddington2013; @Geritz2012], 
-such models can be not only inaccurate but misleading when relied upon in a
-quantitative decision making framework.
-
-1. We do not know what the correct models are for ecological systems.
-1. We have limited data from which to estimate the model -- in particular,
-   such models may be misleading in predicting the probability of outcomes
-   outside the training data.
-
-These aspects are common to many conservation decision making problems, which thus merit
-greater use of non-parametric approaches that can best take advantage of them.
-
-### 1. Large uncertainty where the data is poor
-
-The parametric models perform worst when they propose a management strategy
-outside the range of the observed data. The non-parametric Bayesian approach, 
-in contrast, allows a predictive model that expresses a great deal of uncertainty
-about the probable dynamics outside the observed range, while retaining very
-good predictive accuracy in the range observed.  The management policy 
-dictated by the GP balance this uncertainty against
-the immediate value of the harvest, and act to stabilize the population 
-dynamics in a region of state space in which the predictions can be 
-reliably reflected by the data.
-
-### 2. Predictive accuracy where data is good
-
-While expressing larger uncertainty outside the observed data, the GP
-can also provide a better fit with smaller uncertainty inside the range
-of the observed data. This arises from the greater flexibility of the 
-Gaussian process, which describes a large family of possible curves.
-Despite this flexibility, the GP can be described in relatively few 
-parameters and is thus far less likely to overfit.
-
-Future directions
------------------
-
-### Higher dimensions
-
-In this simulated example, the underlying
-dynamics are truly governed by a simple parametric model, allowing
-the parametric approaches to be more accurate.  Similarly, because the
-dynamics are  one-dimensional dynamics and lead to  stable nodes (rather
-than other attractors such as limit-cycles resulting in oscillations),
-the training data provides relatively limited information about the
-dynamics.  For these reasons, we anticipate that in higher-dimensional
-examples characteristic of ecosystem management problems that the machine
-learning approach will prove even more valuable.
-
-### Online learning
-
-In our treatment here we have ignored the possibility of learning during the 
-management phase, in which the additional observations of the stock size could
-potentially improve parameter estimates.  While we intend to address this 
-possibility in future work in the context of these non-parametric models,
-we have not addressed it here for pedagogical reasons. In the context presented
-here, it is clear that the differences in performance arise from differences
-in the uncertainty inherent in the model formulations, rather than from 
-differing abilities to learn.  Because we consider a threshold system, 
-online learning would not change this generic feature of a lack of data in a
-certain range of the state space which is better captured by the Gaussian process.
-
-Acknowledgments
-===============
-
-This work was partially supported by the Center for Stock Assessment
-Research, a partnership between the University of California Santa Cruz
-and the Fisheries Ecology Division, Southwest Fisheries Science Center,
-Santa Cruz, CA and by NSF grant EF-0924195 to MM and NSF grant DBI-1306697
-to CB.
-
-Appendix
-========
-
-## Model definitions and estimation
-
-Equation S1: Ricker model.
-
-$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
-
-Figure S1: Ricker model: prior and posterior distributions for parameter estimates.
-
-```{r, echo=FALSE, include=TRUE}
-plot_ricker_traces
-```
-
-```{r, echo=FALSE, include=TRUE}
-plot_ricker_posteriors
-```
-Table S1: Parameterization of the priors
-
-```{r echo=FALSE, results="asis", include=TRUE}
-xtable::xtable(data.frame(
-  parameter = c("r0", "K", "sigma"),
-  lower_bound = c(r0_prior_p[1], K_prior_p[1], stdQ_prior_p[1]),
-  upper_bound = c(r0_prior_p[2], K_prior_p[2], stdQ_prior_p[2])))
-```
-
-$$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} $$
-
-Eq S2: Myers model 
-Figure S2: Myers model: Traces, prior and posterior distributions for parameter estimates.
-
-```{r, echo=FALSE, include=TRUE}
-plot_myers_traces
-```
-
-```{r, echo=FALSE, include=TRUE}
-plot_myers_posteriors
-```
-
-Table S2: Parameterization of the priors
-```{r echo=FALSE, results="asis", include=TRUE}
-xtable::xtable(data.frame(parameter = c("r0", "K", "theta", "sigma"),
-           lower_bound = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
-           upper_bound = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2])))
-```
-
-Eq S3: Allen model
-
-$$f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} $$
-
-Figure S3: Allen model: prior and posterior distributions for parameter estimates.
-
-```{r, echo=FALSE, include=TRUE}
-plot_allen_traces
-```
-
-```{r, echo=FALSE, include=TRUE}
-plot_allen_posteriors
-```
-
-Table S3: Parameterization of the priors
-
-```{r echo=FALSE, results="asis", include=TRUE}
-xtable::xtable(data.frame(parameter = c("r0", "K", "theta", "sigma"),
-           lower_bound = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
-           upper_bound = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2])))
-```
-
-Eq S4: GP model 
-Figure S4: GP model: prior and posterior distributions for parameter estimates.
-
-```{r echo=FALSE, include=TRUE}
-gp_assessment_plots
-```
-
-Table S4: Parameterization of the priors
-
-## Optimal Control Problem
-
-We seek the harvest policy $h(x)$ that maximizes:
-
-$$ \max_{h_t} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h_t) \delta^t  $$
-
-subject to the profit function $\Pi(X_t,h)$, discount rate $\delta$, and the state
-equation
-
-$$X_{t+1} = Z_t f(S_t)  $$
-$$S_t = X_t - h_t $$
-
-Where $Z_t$ is multiplicative noise function with mean 1, representing
-stochastic growth. We will consider log-normal noise with shape parameter
-$\sigma_g$.
-
-Form this we can write down the Bellman recursion as:
-
-$$V_t(x_t) = \max_h \mathbf{E} \left(\Pi(h_t, x_t) + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)$$
-
-For simplicity we assume profit is simply linear in the realized harvest (only
-enforcing the restriction that harvest can not exceed available stock), $\Pi(h,x) = min(h,x)$.
-
-### Pseudocode for the Bellman iteration
-
-```r
- V1 <- sapply(1:length(h_grid), function(h){
-      delta * F[[h]] %*% V +  profit(x_grid, h_grid[h]) 
-    })
-    # find havest, h that gives the maximum value
-    out <- sapply(1:gridsize, function(j){
-      value <- max(V1[j,], na.rm = T) # each col is a diff h, max over these
-      index <- which.max(V1[j,])  # store index so we can recover h's 
-      c(value, index) # returns both profit value & index of optimal h.  
-    })
-    # Sets V[t+1] = max_h V[t] at each possible state value, x
-    V <- out[1,]                        # The new value-to-go
-    D[,OptTime-time+1] <- out[2,]       # The index positions
-```
-
-### Training data
-
-Eacho of our models $f(S_t)$ must be estimated from training data, which
-we simulate from the Allen model with parameters $r = $ ` r p[1]`, 
-$K =$ ` r p[2]`, $C =$ ` r p[3]`, and  $\sigma_g =$ ` r sigma_g` 
-for $T=$ `r Tobs$ timesteps, starting at initial condition $X_0 = $ `r Xo`. 
-The training data can be seen in Figure 1.
-
------------------------------------
-
-<!-- OLD TEXT -->
-
-Abstract
-=======================================================================
-
-Decision-theoretic methods often rely on simple parametric models of
-ecological dynamics to compare the value of a potential sequence of
-actions. Unfortunately, such simple models rarely capture the complexity
-or uncertainty found in most real ecosystems.
-
-Further, the data on which a model has been parameterized frequently 
-fails to cover the possible state-space over which management decisions 
-must operate.  Consequently a model do well in the region of state-space
-in which it was estimated, but give erroneous confidence to predictions
-outside of that region.
-
-This problem is keenly felt in any system where a potential threshold
-or tipping point is a concern.  Such a tipping point, if it exists 
-at all, will lay outside the observed range of the observed data.
-
-We demonstrate how nonparametric Bayesian models can provide robust,
-solutions to decision making under uncertainty without knowing the 
-structural form of the true model.
-
-While methods that account for _parametric_ uncertainty can be very 
-successful with the right model,
-structural uncertainty of not knowing what model best approximates the 
-dynamics poses considerably greater difficulty.
-
-Introduction
-=======================================================================
-
-#### Opening
-
-<!-- More on complex dynamics and not having the correct model
-
--->
-
-<!-- More on the lack of data throughout the relevant state-space
-     and how any concern about potential tipping points indicates 
-     that the data exhibits this bias / problem.  
-     -->
-
-#### Models for decision-making under uncertainty
-
-Decision-theoretic or optimal control tools require a model that can assign probabilities of
-future states (e.g. stock size of a fishery) given the current state and a
-proposed action (e.g. fishing harvest or effort).  
-Management frequently faces a sequential decision-making problem -- after 
-selecting an action, the decision-maker may receive new information about 
-the current state and must again choose an appropriate action -- such as setting
-the harvest limits each year based on stock assessments the year prior.
-
-The decision maker typically seeks to determining the course of actions (also referred to as the policy)
-that maximizes the expected value of some objective function such as
-net present value derived from the resource over time.  
-Though much can be said on how to choose this value function appropriately 
-(e.g. see [@Halpern2013](http://doi.org/10.1073/pnas.1217689110 
-"Achieving the triple bottom line in the face of inherent trade-offs 
-among social equity, economic return, and conservation.")) we will
-assume this is given.  (Nor is this approach necessarily constrained to 
-maximizing the expectated value of such a function - the decision-theoretic
-framework can be adapted to alternatives such as minimizing the maximum 
-cost or damage that might be incurred; see @Polasky2011).
-
-In representing future states with probabilities and maximizing expectations,
-this approach provides a natural framework for handling uncertainty.
-
-The value function typically depends on the action or policy taken, as well as
-the state of the system, in each interval of time. The state of the system,
-in turn, is usually described by a dynamical model.
-
-[@Williams2001; @Athanassoglou2012].
-
-<!-- Transition and map: The weakness of parametric models -->
-<!-- This section is a bit weak and vague.  Revisit to make a real map in light 
-of outline structure below. -->
-
-While simple mechanistic models can nevertheless provide important insights
-into long-term outcomes, such approaches are not well-suited for use in
-forecasting outcomes of potential management options.  Non-parametric
-approaches offer a more flexible alternative that can both more accurately
-reflect the data available while also representing greater uncertainty
-in areas (of state-space) where data is lacking.
-
-We demonstrate how
-a Gaussian Process model of stock recruitment can lead to nearly optimal
-management through stochastic dynamic programming, comperable to knowing
-the correct structural equation for the underlying simulation.  Meanwhile,
-parametric models that do not match the underlying dynamics can perform 
-very poorly, even though they fit the data as well as the true model.  
-Ecological research and management strategy should pay closer attention
-to the opportunities and challenges nonparametric modeling can offer.
-
-<!-- 
-### Quantitative vs Qualitative Decisions
-
-In this paper, we consider those ecological management problems in which
-a mathematical (or computational) model is used to quantitatively inform
-decision-making by allowing a manager to compare to the expected consequences
-of potential management actions (or policies).  We distinguish this from the
-solely qualitative use of a model, in which models are used to represent and
-compare hypotheses of different mechanisms that would lead to qualitatively 
-different actions.
-
-In addition to facing the potentially grave consequences of such transitions,
-if they do exist in a system of interest, this generally means that we lack
-data in the region near and beyond a possible tipping point.  As a consequence,
-our models are least accurate just where we need them to be most accurate.  
-Unfortunately, parametric modeling approaches are not well-suited to this 
-scenario, in which good data is avialble over only part of the relevant 
-state-space.  Necessary assumptions about model structure in any parametric
-approach can lead such models to perform very poorly, even when accounting
-for parameter uncertainty.
-
--->
-
-Approach and Methods
-====================
-
-## The optimal control problem in fisheries management
-
-We focus on the problem in which a manager must set
-the harvest level for a marine fishery each year to maximize the net
-present value of the resource, given an estimated stock size from the
-year before.
-
-<!-- ugh, re-word this -->
-
-To permit comparisons against a theoretical optimum we will consider
-data on the stock dynamics simulated from a simple parametric model
-in which recruitment of the fish stock $X_{t+1}$ in the following year
-is a stochastic process governed by a function $f$ of the current 
-stock $X_t$, selected harvest policy $h_t$, and noise process $Z$,
-
-$$X_{t+1} = Z_t f(X_t, h_t) $$
-
-Given parameters for the function $f$ and probability distribution $Z$,
-along with a given economic model determining the  price/profit $\Pi(X_t,
-h_t)$ realized in a given year given a choice of harvest $h_t$ and
-observed stock $X_t$.  This problem can be solved exactly for discretized
-values of stock $X$ and policy $h$ using stochastic dynamic programming
-(SDP) [@Mangel1985]. Problems of this sort underpin much marine fisheries
-management today.
-
-A crux of this approach is correctly specifying the functional form of $f$,
-along with its parameters.  The standard approach uses one of a 
-handful of common parametric models representing the stock-recruitment
-relationship, usually after estimating the model parameters from any 
-available existing data. Uncertainty in the parameter estimates can 
-be estimated and integrated over to determine the optimal policy under
-under uncertainty [@Mangel1985; @Schapaugh2013]. Uncertainty in the model
-structure itself can only be addressed in this approach by hypothesizing 
-alternative model structures, and then performing some model choice or
-model averaging  [@Williams2001; @Athanassoglou2012].
-
-## Underlying Model
-
-Concerns over the potential for tipping points in ecological dynamics
-[@Scheffer2001] highlight the dangers of uncertainty in ecological
-management and pose a substantial challenge to existing decision-theoretic
-approaches [@Brozovic2011].  Because intervention is often too late 
-after a tipping point has been crossed (but see @Hughes2013), management
-is most often concerned with avoiding potentially catastrophic tipping
-points before any data is available at or following a transition that
-would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].
-
-To illustrate the value of the non-parametric Bayesian approach to management,
-we focus on example of a system containing such a tipping point whose dynamics
-can still be described by a simple, one-dimensional parametric model.  
-We will focus on a simple parametric model for a single species [derived
-from fist principles by @Allen2005a] as our underlying "reality".
-
-$$X_{t+1} = Z_t f(S_t)  $$
-$$S_t = X_t - h_t $$
-$$f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} $$
-
-Where $Z_t$ is multiplicative noise function with mean 1, representing
-stochastic growth. We will consider log-normal noise with shape parameter
-$\sigma_g$.  We start with an example in which the parameters are $r =2$,
-$K=8$, $C=5$ and $\sigma_g = 0.1$.
-
-<!-- ` r p[1]`, $K =$ ` r p[2]`, $C =$ ` r p[3]`, and  $\sigma_g =$ ` r sigma_g`. -->
-
-As a low-dimensional system completely described by three parameters, this
-scenario should if anything be favorable to a parametric-based approach.
-This model contains an Allee effect, or tipping point, below which the
-population is not self-sustaining and shrinks to zero [@Courchamp2008].
-
-#### Simulated training data
-
-We generate initial observational data under the model described in Eq 1 
-for $T_{\textrm{obs}}=40$ time steps, under a given arbitrary sequence of harvest
-intensities, $h_t$. 
-We consider the case in which most of the data comes from a limited region
-of state space (e.g. near a stable equilibrium), leaving us without
-observations of the population dynamics at very low levels which would
-be useful in discrimating between recruitment curves [@] or demonstrating
-the existence of a tipping point [@Scheffer2001].  
-Using data simulated from a specified model rather than empirical data
-permits the comparison against the true underlying dynamics, setting 
-a bar for the optimal performance possible.
-
-### Parametric Models
-
-We consider three candidate parametric models for the stock-recruitment function, 
-which we refer to by the first authors of the publications in which they were
-first proposed.
-
-We generate the data with a four-parameter model that contains
-a tipping point, as discussed above (equation 1),
-(an Allee effect, see [@Allen, @Courchamp]) below which the stock
-decreases to zero,
-
-$$ X_{t+1} = Z_t S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(\frac{S_t - \theta}{K}\right)} $$
-
-$$ S_t = X_t - h_t $$
-
-The parameter $C$ reflects the location of the tipping point, $K$ the carrying capacity of
-the stock, and $r$ the base recruitment rate.  $S_t$ represents the stock size after
-a harvest $h_t$ has been implemented.  $Z_t$ represents a log-normal random variable of
-log-mean zero and log-standard deviation parameter $\sigma$.
-
-We consider two alternative candidate models: the Ricker [@Ricker] stock-recruitment curve,
-
-$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
-
-and an alternative four-parameter model adapted from @Myers,
-
-$$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} $$
-
-which contains a tipping point for $\theta > 2$ and becomes a Beverton-Holt model 
-at $\theta = 1$.
-
-<!-- Note that while similarly-named parameters contain similar connotations, the 
-estimate and precise intepretation will be different in each model. -->
-
-### Bayesian Inference of Parametric models
-
-Given the sample data, we infer posterior distributions for each of the three
-models listed above using a Markov Chain Monte Carlo Gibbs Sampler (jags, see 
-appendix for implementation details and code) given uniform priors.  We run six
-chains for $10^6$ steps each and then assess convergence by Gelman-Rubin criterion
-and inspection of the traces, see appendix.
-
-<!-- By transforming variables we can integrate out the the $\sigma$ parameter
-and the constant term in the log-mean, see appendix.  The resulting one or
-two parameters can then be evaluated over a vector or grid of possible values,
-and the result transformed back to determine the posterior density of the parameters,
-This approach avoids any concerns of convergence in the MCMC analysis and can thus
-be used to verify that the performance of the parametric models can not be explained
-by convergence issues.  However, this approach does not generalize easily to richer models,
-such as the introduction of measurement noise or additional parameters, in which 
-MCMC based approaches are the obvious choice.  -->
-
-### The Non-parametric Bayesian alternative for stock-recruitment curves
-
-The use of Gaussian process (GP) regression (or "kriging" in the geospatial
-literature) to formulate a predictive model is relatively new in the
-context of modeling dynamical systems [@Kocijan2005] and introduced
-in the ecological modeling and fisheries management by @Munch2005.
-An accessible and thorough introduction to the formulation and use of
-GPs can be found in @Rasmussen2006.
 
@@
 
+Approach and Methods
+====================
+
+### Summary of approach
+
+### Statement of the optimal control problem
+
+- Underlying model 
+- Available data 
+- Value function
+
+For simplicity we assume profit is simply linear in the realized harvest (only
+enforcing the restriction that harvest can not exceed available stock)
+
+### Parametric models
+
+- Statement of the models
+
+We consider three candidate parametric models of the stock-recruitment
+dynamics: The Ricker model, the Allen model [Allen 2005](), the Myers
+model. The familiar Ricker model involves two parameters, corresponding
+to a growth rate and a carrying capacity, and cannot support alternative
+stable state dynamics (though as growth rate increases it exhibits a
+periodic attractor that proceeds through period-doubling into chaos. We
+will generally focus on dynamics below the chaotic threshold for
+the purposes of this analysis.) The Allen model resembles the Ricker
+dynamics with an added Allee effect parameter [Courchamp](), below
+which the population cannot persist.  The Myers model also has three
+parameters and contains an Allee threshold, but has compensatory rather
+than over-compensatory density dependence (resembling a Beverton-Holt
+curve rather than a Ricker curve at high densities.)
+
+We assume multiplicative log-normal noise perturbs the growth predicted 
+by the each of the deterministic model skeletons described above. This 
+introduces one additional parameter $\sigma$ that must be estimated by each
+model.
+
+<!-- equations just in appendix? -->
+
+As we simulate training data from the Allen model (ref section), we will
+refer to this as the structurally correct model.  The Ricker model is
+thus a reasonable approximation of these dynamics far from the Allee
+threshold (but lacks threshold dynamics), while the Myers model shares
+the essential feature of a threshold but differs in the structure. Thus
+we have three potential parametric models of the stock dynamics.
+
+- Bayesian inference of parametric models
+
+We infer posterior distributions for the parameters of each model
+in a Bayesian context using Gibbs sampling (implemented in R [@RTeam]
+using jags, [@R2jags]).  We choose uninformative uniform priors for all
+parameters (See Appendix, Figures S1-S3, and Table S1, and the R code
+provided). One-step-ahead predictions of these model fits are shown in
+Figure 1.
+
+- SDP via parametric models
+
+An optimal policy function is then inferred through stochastic dynamic
+programming for each model given the posterior distributions of the
+parameter estimates.  This policy maximizes the expectation of the value
+function integrated over the parameter uncertainty. (code implementing
+this algorithm provided in the Appendix).
+
+### The Gaussian Process model
+
+- Statement of model
+
+... more on GP ... [Munch 2005]()
+
+We also estimate a simple Gaussian Process defined by
+a radial basis function kernel of two parameters: $\ell$, which gives
+the characteristic length-scale over which correlation between two 
+points (e.g. any two points $X_t, X_{t+1}$, and $X_{t+\tau}, X_{t+1+\tau}$)
+in state-space decays, and $\sigma$, which gives the scale of the 
+process noise by which observations $Y_{t+1}$ may differ from their
+predicted values $X_{t+1}$ given an observation of the previous state,
+$X_t$.
+
+- Inference of the model
+
+Also unlike parametric models, this posterior distribution is still
+conditional on the training data. As such, the uncertainty near the
+observed data.
+
+We use a Metropolis-Hastings Markov Chain Monte Carlo to infer posterior
+distributions of the two parameters of the GP (Figure S4, code in
+appendix), under weakly informative Gaussian priors (see parameters in
+table S5). As the posterior distributions differ substantially from the
+priors (Figure S4), we can be assured that most of the information in
+the posterior comes from the data rather than the prior belief.
+
+- SDP via the model
+
+Though we are unaware of prior application of this type, it is reasonably
+straight-forward to adapt the Gaussian Process for Stochastic Dynamic
+Programming.  Recall that unlike the parametric models the Gaussian
+process with fixed parameters already predicts a distribution of
+curves rather than a single curve. We must first integrate over 
+this distribution of curves given a sampling of parameter values drawn
+from the posterior distribution of the two GP parameters, before
+integrating over the posterior of those parameters themselves.
+
+Results
+=======
+
+```{r Figure1, dependson=c("assemble-models"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
+Graph shows the expected value for the function $f$ under each model.
+Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.
+The training data is also shown as black points.
+(The GP is conditioned on 0,0, shown as a pseudo-data point).
+", include=FALSE}
+```
+```{r Figureb, dependson=c("assemble-models", "par-fns"), include=TRUE}
+```
+
+Figure 1 shows the mean inferred state space dynamics of each model
+relative to the true model used to generate the data, predicting the
+relationship between observed stock size (x-axis) to the stock size
+after recruitment the following year.  All models except the MLE model
+estimate a distribution around the means shown here, and all models 
+estimate a level of process noise, which is independent of the state
+value (x).  Note that in contrast to the  other models shown, the mean
+Gaussian process corresponds to a distribution of curves - as indicated
+by the gray band - which itself has a mean shown in black.  Note that 
+this mean GP is thus more certain of the dynamics in the region where
+data is available then where it is not.
+
+While it would be straight forward to condition the GP on passing through the origin (0,0), the estimate
+shown here is based only on the observed data. The observed data from
+which each model is estimated is also shown.  The observations come
+from only a limited region of state space corresponding to unharvested
+or weakly harvested system.  No observations occur at the theoretical
+optimum harvest rate or near the tipping point.
+
+```{r Figure2, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
+Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", include=TRUE}
+```
+
+The resulting optimal management strategy based on each of the inferred
+models is shown in Figure 2, against the optimal strategy given the
+true underlying dynamics.  Policies are shown in terms of target
+escapement, $S_t$.  Under models such as this a constant escapement
+policy is expected to be optimal [@Reed1979], whereby population levels
+below a certain size $S$ are unharvested, while above that size the harvest
+strategy aims to return the population to $S$, resulting in the hockey-stick
+shaped policies shown.
+
+```{r Figure3, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.",  include=TRUE}
+```
+
+The consequences of managing 100 replicate realizations of the 
+simulated fishery under each of the policies estimated is shown in Figure 3.  As expected
+from the policy curves, the structurally correct model under-harvests,
+leaving the stock to vary around it's un-fished optimum.  The structurally
+incorrect Ricker model over-harvests the population passed
+the tipping point consistently, resulting in the immediate crash of the stock and 
+thus derives minimal profits.
+
+The results shown in Figures 1-3 are not unique to the simulated data or models chosen
+here, but arises across a range of parameter values and simulations as shown in the 
+supplemental figures.  The results across this range can most easily be compared 
+by the relative differences in net present value realized by each of the approaches,
+as shown in Figure 4.  The Gaussian Process most consistently realizes a value 
+close to the optimal solution, and importantly avoids ever driving the system across
+the tipping point, which results in the near-zero value cases in the parametric models.
+
+```{r Figure4, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
+For each data set, the three models are estimated as described above.
+Values plotted are the averages of a given policy over 100 replicate simulations.
+Details and code provided in the supplement.",  include=TRUE}
+```
+
+### Figure 1: Fitted Models
+
+- All models fit the data quite well
+- Information criteria would pick the simple, incorrect model.
+
+### Figure 2: Inferred Policies
+
+- Inferred policies differ substantially among models
+- The structurally correct model and the GP are close to the true model
+- alternatives are not close
+
+### Figure 3: Simulated results
+
+-
+
+### Figure 4: Robustness
+
+- Results hold across range of parameters
+
+(eek, distinguish better between result and discussion?)
+
+Discussion 
+==========
+
+- All models are "good fits" to the originally observed data.
+
+- (Simple model choice immediately leads us astray)
+
+Though simple mechanistically motivated models offer the greatest potential 
+to increase our basic understanding of ecological processes [@Cuddington2013; @Geritz2012], 
+such models can be not only inaccurate but misleading when relied upon in a
+quantitative decision making framework.
+
+1. We do not know what the correct models are for ecological systems.
+1. We have limited data from which to estimate the model -- in particular,
+   such models may be misleading in predicting the probability of outcomes
+   outside the training data.
+
+These aspects are common to many conservation decision making problems, which thus merit
+greater use of non-parametric approaches that can best take advantage of them.
+
+### 1. Large uncertainty where the data is poor
+
+The parametric models perform worst when they propose a management strategy
+outside the range of the observed data. The non-parametric Bayesian approach, 
+in contrast, allows a predictive model that expresses a great deal of uncertainty
+about the probable dynamics outside the observed range, while retaining very
+good predictive accuracy in the range observed.  The management policy 
+dictated by the GP balance this uncertainty against
+the immediate value of the harvest, and act to stabilize the population 
+dynamics in a region of state space in which the predictions can be 
+reliably reflected by the data.
+
+### 2. Predictive accuracy where data is good
+
+While expressing larger uncertainty outside the observed data, the GP
+can also provide a better fit with smaller uncertainty inside the range
+of the observed data. This arises from the greater flexibility of the 
+Gaussian process, which describes a large family of possible curves.
+Despite this flexibility, the GP can be described in relatively few 
+parameters and is thus far less likely to overfit.
+
+Future directions
+-----------------
+
+### Higher dimensions
+
+In this simulated example, the underlying
+dynamics are truly governed by a simple parametric model, allowing
+the parametric approaches to be more accurate.  Similarly, because the
+dynamics are  one-dimensional dynamics and lead to  stable nodes (rather
+than other attractors such as limit-cycles resulting in oscillations),
+the training data provides relatively limited information about the
+dynamics.  For these reasons, we anticipate that in higher-dimensional
+examples characteristic of ecosystem management problems that the machine
+learning approach will prove even more valuable.
+
+### Online learning
+
+In our treatment here we have ignored the possibility of learning during the 
+management phase, in which the additional observations of the stock size could
+potentially improve parameter estimates.  While we intend to address this 
+possibility in future work in the context of these non-parametric models,
+we have not addressed it here for pedagogical reasons. In the context presented
+here, it is clear that the differences in performance arise from differences
+in the uncertainty inherent in the model formulations, rather than from 
+differing abilities to learn.  Because we consider a threshold system, 
+online learning would not change this generic feature of a lack of data in a
+certain range of the state space which is better captured by the Gaussian process.
+
+Acknowledgments
+===============
+
+This work was partially supported by the Center for Stock Assessment
+Research, a partnership between the University of California Santa Cruz
+and the Fisheries Ecology Division, Southwest Fisheries Science Center,
+Santa Cruz, CA and by NSF grant EF-0924195 to MM and NSF grant DBI-1306697
+to CB.
+
+Appendix
+========
+
+## Model definitions and estimation
+
+Equation S1: Ricker model.
+
+$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
+
+Figure S1: Ricker model: prior and posterior distributions for parameter estimates.
+
+```{r, echo=FALSE, include=TRUE}
+plot_ricker_traces
+```
+
+```{r, echo=FALSE, include=TRUE}
+plot_ricker_posteriors
+```
+Table S1: Parameterization of the priors
+
+```{r echo=FALSE, results="asis", include=TRUE}
+xtable::xtable(data.frame(
+  parameter = c("r0", "K", "sigma"),
+  lower_bound = c(r0_prior_p[1], K_prior_p[1], stdQ_prior_p[1]),
+  upper_bound = c(r0_prior_p[2], K_prior_p[2], stdQ_prior_p[2])))
+```
+
+$$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} $$
+
+Eq S2: Myers model 
+Figure S2: Myers model: Traces, prior and posterior distributions for parameter estimates.
+
+```{r, echo=FALSE, include=TRUE}
+plot_myers_traces
+```
+
+```{r, echo=FALSE, include=TRUE}
+plot_myers_posteriors
+```
+
+Table S2: Parameterization of the priors
+```{r echo=FALSE, results="asis", include=TRUE}
+xtable::xtable(data.frame(parameter = c("r0", "K", "theta", "sigma"),
+           lower_bound = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
+           upper_bound = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2])))
+```
+
+Eq S3: Allen model
+
+$$f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} $$
+
+Figure S3: Allen model: prior and posterior distributions for parameter estimates.
+
+```{r, echo=FALSE, include=TRUE}
+plot_allen_traces
+```
+
+```{r, echo=FALSE, include=TRUE}
+plot_allen_posteriors
+```
+
+Table S3: Parameterization of the priors
+
+```{r echo=FALSE, results="asis", include=TRUE}
+xtable::xtable(data.frame(parameter = c("r0", "K", "theta", "sigma"),
+           lower_bound = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
+           upper_bound = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2])))
+```
+
+Eq S4: GP model 
+Figure S4: GP model: prior and posterior distributions for parameter estimates.
+
+```{r echo=FALSE, include=TRUE}
+gp_assessment_plots
+```
+
+Table S4: Parameterization of the priors
+
+## Optimal Control Problem
+
+We seek the harvest policy $h(x)$ that maximizes:
+
+$$ \max_{h_t} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h_t) \delta^t  $$
+
+subject to the profit function $\Pi(X_t,h)$, discount rate $\delta$, and the state
+equation
+
+$$X_{t+1} = Z_t f(S_t)  $$
+$$S_t = X_t - h_t $$
+
+Where $Z_t$ is multiplicative noise function with mean 1, representing
+stochastic growth. We will consider log-normal noise with shape parameter
+$\sigma_g$.
+
+Form this we can write down the Bellman recursion as:
+
+$$V_t(x_t) = \max_h \mathbf{E} \left(\Pi(h_t, x_t) + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)$$
+
+For simplicity we assume profit is simply linear in the realized harvest (only
+enforcing the restriction that harvest can not exceed available stock), $\Pi(h,x) = min(h,x)$.
+
+### Pseudocode for the Bellman iteration
+
+```r
+ V1 <- sapply(1:length(h_grid), function(h){
+      delta * F[[h]] %*% V +  profit(x_grid, h_grid[h]) 
+    })
+    # find havest, h that gives the maximum value
+    out <- sapply(1:gridsize, function(j){
+      value <- max(V1[j,], na.rm = T) # each col is a diff h, max over these
+      index <- which.max(V1[j,])  # store index so we can recover h's 
+      c(value, index) # returns both profit value & index of optimal h.  
+    })
+    # Sets V[t+1] = max_h V[t] at each possible state value, x
+    V <- out[1,]                        # The new value-to-go
+    D[,OptTime-time+1] <- out[2,]       # The index positions
+```
+
+### Training data
+
+Eacho of our models $f(S_t)$ must be estimated from training data, which
+we simulate from the Allen model with parameters $r = $ ` r p[1]`, 
+$K =$ ` r p[2]`, $C =$ ` r p[3]`, and  $\sigma_g =$ ` r sigma_g` 
+for $T=$ `r Tobs$ timesteps, starting at initial condition $X_0 = $ `r Xo`. 
+The training data can be seen in Figure 1.
+
+-----------------------------------
+
+<!-- OLD TEXT -->
+
+Abstract
+=======================================================================
+
+Decision-theoretic methods often rely on simple parametric models of
+ecological dynamics to compare the value of a potential sequence of
+actions. Unfortunately, such simple models rarely capture the complexity
+or uncertainty found in most real ecosystems.
+
+Further, the data on which a model has been parameterized frequently 
+fails to cover the possible state-space over which management decisions 
+must operate.  Consequently a model do well in the region of state-space
+in which it was estimated, but give erroneous confidence to predictions
+outside of that region.
+
+This problem is keenly felt in any system where a potential threshold
+or tipping point is a concern.  Such a tipping point, if it exists 
+at all, will lay outside the observed range of the observed data.
+
+We demonstrate how nonparametric Bayesian models can provide robust,
+solutions to decision making under uncertainty without knowing the 
+structural form of the true model.
+
+While methods that account for _parametric_ uncertainty can be very 
+successful with the right model,
+structural uncertainty of not knowing what model best approximates the 
+dynamics poses considerably greater difficulty.
+
+Introduction
+=======================================================================
+
+#### Opening
+
+<!-- More on complex dynamics and not having the correct model
+
+-->
+
+<!-- More on the lack of data throughout the relevant state-space
+     and how any concern about potential tipping points indicates 
+     that the data exhibits this bias / problem.  
+     -->
+
+#### Models for decision-making under uncertainty
+
+Decision-theoretic or optimal control tools require a model that can assign probabilities of
+future states (e.g. stock size of a fishery) given the current state and a
+proposed action (e.g. fishing harvest or effort).  
+Management frequently faces a sequential decision-making problem -- after 
+selecting an action, the decision-maker may receive new information about 
+the current state and must again choose an appropriate action -- such as setting
+the harvest limits each year based on stock assessments the year prior.
+
+The decision maker typically seeks to determining the course of actions (also referred to as the policy)
+that maximizes the expected value of some objective function such as
+net present value derived from the resource over time.  
+Though much can be said on how to choose this value function appropriately 
+(e.g. see [@Halpern2013](http://doi.org/10.1073/pnas.1217689110 
+"Achieving the triple bottom line in the face of inherent trade-offs 
+among social equity, economic return, and conservation.")) we will
+assume this is given.  (Nor is this approach necessarily constrained to 
+maximizing the expectated value of such a function - the decision-theoretic
+framework can be adapted to alternatives such as minimizing the maximum 
+cost or damage that might be incurred; see @Polasky2011).
+
+In representing future states with probabilities and maximizing expectations,
+this approach provides a natural framework for handling uncertainty.
+
+The value function typically depends on the action or policy taken, as well as
+the state of the system, in each interval of time. The state of the system,
+in turn, is usually described by a dynamical model.
+
+[@Williams2001; @Athanassoglou2012].
+
+<!-- Transition and map: The weakness of parametric models -->
+<!-- This section is a bit weak and vague.  Revisit to make a real map in light 
+of outline structure below. -->
+
+While simple mechanistic models can nevertheless provide important insights
+into long-term outcomes, such approaches are not well-suited for use in
+forecasting outcomes of potential management options.  Non-parametric
+approaches offer a more flexible alternative that can both more accurately
+reflect the data available while also representing greater uncertainty
+in areas (of state-space) where data is lacking.
+
+We demonstrate how
+a Gaussian Process model of stock recruitment can lead to nearly optimal
+management through stochastic dynamic programming, comperable to knowing
+the correct structural equation for the underlying simulation.  Meanwhile,
+parametric models that do not match the underlying dynamics can perform 
+very poorly, even though they fit the data as well as the true model.  
+Ecological research and management strategy should pay closer attention
+to the opportunities and challenges nonparametric modeling can offer.
+
+<!-- 
+### Quantitative vs Qualitative Decisions
+
+In this paper, we consider those ecological management problems in which
+a mathematical (or computational) model is used to quantitatively inform
+decision-making by allowing a manager to compare to the expected consequences
+of potential management actions (or policies).  We distinguish this from the
+solely qualitative use of a model, in which models are used to represent and
+compare hypotheses of different mechanisms that would lead to qualitatively 
+different actions.
+
+In addition to facing the potentially grave consequences of such transitions,
+if they do exist in a system of interest, this generally means that we lack
+data in the region near and beyond a possible tipping point.  As a consequence,
+our models are least accurate just where we need them to be most accurate.  
+Unfortunately, parametric modeling approaches are not well-suited to this 
+scenario, in which good data is avialble over only part of the relevant 
+state-space.  Necessary assumptions about model structure in any parametric
+approach can lead such models to perform very poorly, even when accounting
+for parameter uncertainty.
+
+-->
+
+Approach and Methods
+====================
+
+## The optimal control problem in fisheries management
+
+We focus on the problem in which a manager must set
+the harvest level for a marine fishery each year to maximize the net
+present value of the resource, given an estimated stock size from the
+year before.
+
+<!-- ugh, re-word this -->
+
+To permit comparisons against a theoretical optimum we will consider
+data on the stock dynamics simulated from a simple parametric model
+in which recruitment of the fish stock $X_{t+1}$ in the following year
+is a stochastic process governed by a function $f$ of the current 
+stock $X_t$, selected harvest policy $h_t$, and noise process $Z$,
+
+$$X_{t+1} = Z_t f(X_t, h_t) $$
+
+Given parameters for the function $f$ and probability distribution $Z$,
+along with a given economic model determining the  price/profit $\Pi(X_t,
+h_t)$ realized in a given year given a choice of harvest $h_t$ and
+observed stock $X_t$.  This problem can be solved exactly for discretized
+values of stock $X$ and policy $h$ using stochastic dynamic programming
+(SDP) [@Mangel1988]. Problems of this sort underpin much marine fisheries
+management today.
+
+A crux of this approach is correctly specifying the functional form of $f$,
+along with its parameters.  The standard approach uses one of a 
+handful of common parametric models representing the stock-recruitment
+relationship, usually after estimating the model parameters from any 
+available existing data. Uncertainty in the parameter estimates can 
+be estimated and integrated over to determine the optimal policy under
+under uncertainty [@Mangel1988; @Schapaugh2013]. Uncertainty in the model
+structure itself can only be addressed in this approach by hypothesizing 
+alternative model structures, and then performing some model choice or
+model averaging  [@Williams2001; @Athanassoglou2012].
+
+## Underlying Model
+
+Concerns over the potential for tipping points in ecological dynamics
+[@Scheffer2001] highlight the dangers of uncertainty in ecological
+management and pose a substantial challenge to existing decision-theoretic
+approaches [@Brozovic2011].  Because intervention is often too late 
+after a tipping point has been crossed (but see @Hughes2013), management
+is most often concerned with avoiding potentially catastrophic tipping
+points before any data is available at or following a transition that
+would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].
+
+To illustrate the value of the non-parametric Bayesian approach to management,
+we focus on example of a system containing such a tipping point whose dynamics
+can still be described by a simple, one-dimensional parametric model.  
+We will focus on a simple parametric model for a single species [derived
+from fist principles by @Allen2005a] as our underlying "reality".
+
+$$X_{t+1} = Z_t f(S_t)  $$
+$$S_t = X_t - h_t $$
+$$f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} $$
+
+Where $Z_t$ is multiplicative noise function with mean 1, representing
+stochastic growth. We will consider log-normal noise with shape parameter
+$\sigma_g$.  We start with an example in which the parameters are $r =2$,
+$K=8$, $C=5$ and $\sigma_g = 0.1$.
+
+<!-- ` r p[1]`, $K =$ ` r p[2]`, $C =$ ` r p[3]`, and  $\sigma_g =$ ` r sigma_g`. -->
+
+As a low-dimensional system completely described by three parameters, this
+scenario should if anything be favorable to a parametric-based approach.
+This model contains an Allee effect, or tipping point, below which the
+population is not self-sustaining and shrinks to zero [@Courchamp2008].
+
+#### Simulated training data
+
+We generate initial observational data under the model described in Eq 1 
+for $T_{\textrm{obs}}=40$ time steps, under a given arbitrary sequence of harvest
+intensities, $h_t$. 
+We consider the case in which most of the data comes from a limited region
+of state space (e.g. near a stable equilibrium), leaving us without
+observations of the population dynamics at very low levels which would
+be useful in discrimating between recruitment curves [@] or demonstrating
+the existence of a tipping point [@Scheffer2001].  
+Using data simulated from a specified model rather than empirical data
+permits the comparison against the true underlying dynamics, setting 
+a bar for the optimal performance possible.
+
+### Parametric Models
+
+We consider three candidate parametric models for the stock-recruitment function, 
+which we refer to by the first authors of the publications in which they were
+first proposed.
+
+We generate the data with a four-parameter model that contains
+a tipping point, as discussed above (equation 1),
+(an Allee effect, see [@Allen, @Courchamp]) below which the stock
+decreases to zero,
+
+$$ X_{t+1} = Z_t S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(\frac{S_t - \theta}{K}\right)} $$
+
+$$ S_t = X_t - h_t $$
+
+The parameter $C$ reflects the location of the tipping point, $K$ the carrying capacity of
+the stock, and $r$ the base recruitment rate.  $S_t$ represents the stock size after
+a harvest $h_t$ has been implemented.  $Z_t$ represents a log-normal random variable of
+log-mean zero and log-standard deviation parameter $\sigma$.
+
+We consider two alternative candidate models: the Ricker [@Ricker] stock-recruitment curve,
+
+$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
+
+and an alternative four-parameter model adapted from @Myers,
+
+$$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} $$
+
+which contains a tipping point for $\theta > 2$ and becomes a Beverton-Holt model 
+at $\theta = 1$.
+
+<!-- Note that while similarly-named parameters contain similar connotations, the 
+estimate and precise intepretation will be different in each model. -->
+
+### Bayesian Inference of Parametric models
+
+Given the sample data, we infer posterior distributions for each of the three
+models listed above using a Markov Chain Monte Carlo Gibbs Sampler (jags, see 
+appendix for implementation details and code) given uniform priors.  We run six
+chains for $10^6$ steps each and then assess convergence by Gelman-Rubin criterion
+and inspection of the traces, see appendix.
+
+<!-- By transforming variables we can integrate out the the $\sigma$ parameter
+and the constant term in the log-mean, see appendix.  The resulting one or
+two parameters can then be evaluated over a vector or grid of possible values,
+and the result transformed back to determine the posterior density of the parameters,
+This approach avoids any concerns of convergence in the MCMC analysis and can thus
+be used to verify that the performance of the parametric models can not be explained
+by convergence issues.  However, this approach does not generalize easily to richer models,
+such as the introduction of measurement noise or additional parameters, in which 
+MCMC based approaches are the obvious choice.  -->
+
+### The Non-parametric Bayesian alternative for stock-recruitment curves
+
 ### SDP via GP

### 2013-09-24 inst/doc/nonparametric-bayes.Rmd
writing

@@
 <!-- What do we call this?  Extrapolation uncertainty?  Pathological Uncertainty? -->
-An additional source of uncertainty that has recieved less attention[^1] arises <!-- remove 'typical'? -->
-when applying a dynamical model outside the range of data on which it has been
-estimated.  This extrapolation uncertainty is felt most keenly in decision-theoretic
-applications, as (a) exploring the potential action space typically involves considering
-actions that may move the system outside the range of observed behavior, and (b) decision-theoretic
-alogrithms rely not only on reasonable estimates of the expected outcomes, but depend on
-the weights given to all possible outcomes [_e.g._ @Weitzman2013].
+
+An additional source of uncertainty that has recieved less attention[^1]
+arises when applying a dynamical model outside the range of data on
+which it has been estimated.  This extrapolation uncertainty is felt
+most keenly in decision-theoretic applications, as (a) exploring the
+potential action space typically involves considering actions that
+may move the system outside the range of observed behavior, and (b)
+decision-theoretic alogrithms rely not only on reasonable estimates of
+the expected outcomes, but depend on the weights given to all possible
+outcomes [_e.g._ @Weitzman2013].
+
+<!-- Perhaps this is a better-worded version of the above??
+
+Often the set of feasible states is larger than the range of the observed
+states.  Though most systems subject to some degree of stochasticity
+can enter a state outside previously observed values by chance alone
+(such as a particularly low or high nutrient year), this concern is
+most keenly felt in the decision-theoretic context. If we are observing
+the fluctuations of a given fish stock over many years under a fixed
+harvesting pressure, we might develop and test a model that could
+reasonably predict the frequency of a deviation of a given size, even
+when such a deviation has not been previously observed.
+(The same analogy might be made to predicting hurricanes or other
+extreme weather events from historical data, which may be more accurately
+predicted in a constant climate than in one being perturbed).
+
+-->
+
+<!-- Integrate this into the paragraph, rather than as a footnote?? Make more consise? -->
 
@@
 
+<!-- Does this belong?  Improve transitions -->
+
+Concerns over the potential for tipping points in ecological dynamics
+[@Scheffer2001] highlight the dangers of uncertainty in ecological
+management and pose a substantial challenge to existing decision-theoretic
+approaches [@Brozovic2011].  Because intervention is often too late 
+after a tipping point has been crossed (but see @Hughes2013), management
+is most often concerned with avoiding potentially catastrophic tipping
+points before any data is available at or following a transition that
+would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].
+
 _Map_
@@
 face of substantial structural uncertainty.
-
-Often the set of feasible states is larger than the range of the observed
-states.  Though most systems subject to some degree of stochasticity
-can enter a state outside previously observed values by chance alone
-(such as a particularly low or high nutrient year), this concern is
-most keenly felt in the decision-theoretic context. If we are observing
-the fluctuations of a given fish stock over many years under a fixed
-harvesting pressure, we might develop and test a model that could
-reasonably predict the frequency of a deviation of a given size, even
-when such a deviation has not been previously observed.
-(The same analogy might be made to predicting hurricanes or other
-extreme weather events from historical data, which may be more accurately
-predicted in a constant climate than in one being perturbed).
-
-Given adequate observations of a system
-
-One must consider actions or policy under which the system has not yet
-been observed,
 
@@
 $K =$ ` r p[2]`, $C =$ ` r p[3]`, and  $\sigma_g =$ ` r sigma_g` 
-for $T=$ `r Tobs$ timesteps, starting at initial condition $X_0 = $ `r Xo`. 
+for $T=$ `r Tobs` timesteps, starting at initial condition $X_0 = $ `r Xo`. 
 The training data can be seen in Figure 1.
@@
 ## Underlying Model
-
-Concerns over the potential for tipping points in ecological dynamics
-[@Scheffer2001] highlight the dangers of uncertainty in ecological
-management and pose a substantial challenge to existing decision-theoretic
-approaches [@Brozovic2011].  Because intervention is often too late 
-after a tipping point has been crossed (but see @Hughes2013), management
-is most often concerned with avoiding potentially catastrophic tipping
-points before any data is available at or following a transition that
-would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].
 

### 2013-09-24 inst/doc/nonparametric-bayes.tex
writing

@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage{natbib}
+\bibliographystyle{plainnat}
 \usepackage{color}
@@
 
+Decision making under uncertainty is a ubiquitous challenge of natural
+resource management and conservation.
+
 The sudden collapse of fisheries and other ecosystems is an increasingly
 widespread phenomenon and a pressing concern for ecological management
-and conservation. Ecological dyanaics are frequently complex and
+and conservation. Ecological dynamics are frequently complex and
 difficult to measure, making uncertainty in our understanding a
 prediction a persistent challenge to effective management.
-
 Decision-theoretic approaches provide a framework to determine the best
@@
 
-Uncertainty enters at many levels management
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Why fisheries
-\end{itemize}
-
-The economic value and ecological concern have made marine fisheries the
-crucible for much of the founding work (Gordon 1954; Reed 1979; May et
-al. 1979; Ludwig and Walters 1982) in managing ecosystems under
-uncertainty. Global trends (Worm et al. 2006) and controversy (Hilborn
-2007; Worm et al. 2009) have made understanding these challenges all the
-more pressing.
-
-@Fischer @Sethi2005, @Peretti, @Munch,
-
-(Classical division)
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Dynamic uncertainty (stochasticity)
-\end{itemize}
-
-From \href{}{Clark (1976)} to \href{}{Reed (1979)}
-
-Other sources of uncertainty \href{}{Roughgarden \& Smith 1996},
-\href{}{Sethi \emph{et al.} (2005)}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Parametric uncertainty 
-\end{itemize}
-
-As the parameter values for these models must be estimated from limited
-data,\\there will always be some uncertainty associated with these
-values. This uncertainty further compounds the intrinsic variability
-introduced by demographic or environmental noise. The degree of
-uncertainty in the parameter values can be inferred from the data and
-reflected in the estimates of the transition probabilities.
-
-(Mangel 1985; Schapaugh and Tyre 2013)
-
-\href{http://doi.org/10.1016/0304-38008290023-0}{Ludwig \& Walters 1982}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Structural uncertainty 
-\end{itemize}
-
-Estimates of parameter uncertainty are only as good as the parametric
-models themselves. Often we do not understand the system dynamics well
-enough to know if a model provides a good approximation over the
-relevant range of states and timescales (criteria that we loosely refer
-to as defining the ``right'' or ``true'' model.) So called structural or
-model uncertainty is a more difficult problem than parametric
-uncertainty. Typical solutions involve either model choice, model
-averaging, or introducing yet greater model complexity of which others
-may be special cases (model averaging being one such way to construct
-such a model) {[}B. K. Williams (2001); Athanassoglou and Xepapadeas
-(2012); @Cressie2009{]}. Even setting aside other computational and
-statistical concerns (e.g. {[}@Cressie2009{]}), these approaches do not
-address our second concern - representing uncertainty outside the
-observed data range.
-
-\href{http://dx.doi.org/10.1890/07-0744.1}{@Cressie2009}.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Limits of state space
-\end{itemize}
-
-Often the set of feasible states is larger than the range of the
-observed states. Though a system can enter a state outside previously
-observed values due to stochastic effects alone (such as a particularly
-low or high nutrient year), this concern is most keenly felt in
-considering the value of actions that have not yet been applied.
-
-Of all sources of uncertainty, this has potentially recieved the least
-attention. It has most often appeared in the concept of active learning
-or active adaptive managment, (adaptive probing, ) where it may be
-valauble to intentionally force a system far from the observed values
-even when the expected value such actions is low, as it provides much
-faster learning and consequent reduction of model uncertainty that can
-allow greater value to be derived later on. Ludwig and Walters (1982)
-show that it may be advantageous to fish an unexploited population very
-heavily at first to obtain a better estimate of the recruitment rate.
-This intuitive strategy when a population is governed by a Ricker or
-Beverton-Holt-like dynamic would clearly be disasterous if instead the
-dynamics contained an unforseen tipping point. The best way to learn
+Uncertainty enters the decision-making process at many levels: intrinsic
+stochasticity in biological processes, measurements, and implementation
+of policy
+\citetext{\citealp[\emph{e.g.}][]{Reed1979}; \citealp{Clark1986}; \citealp[@Roughgarden1997;][]{Sethi2005}},
+parameteric uncertainty
+\citetext{\citealp[\emph{e.g.}][]{Ludwig1982}; \citealp[@Hilborn1997;
+@McAllister1998;][]{Schapaugh2013}}, and model or structural uncertainty
+\citep[\emph{e.g.}][]{Williams2001, Cressie2009, Athanassoglou2012}. Of
+these, structural uncertainty incorporates the least a priori knowledge
+or assumptions and is generally the hardest to quantify. Typical
+approaches assume a weak notion of uncertainty where a correct or
+reasonable approximation of the dynamics must be identified from among a
+handful of alternative models. Here we consider an approach that
+addresses uncertainty at each of these levels without assuming the
+dynamics follow a particular (i.e.~parametric) structure.
+
+An additional source of uncertainty that has recieved less
+attention\footnote{The concept of adaptive probing is one area that has
+  explicitly addressed this kind of uncertainty, reaching rather
+  opposite conclusions than what we observe here. Adaptive probing
+  strategies follow from ``Dual Control'' or ``Active Adaptive
+  Management'' approaches (e.g. \citet{Ludwig1982}) that can trade off
+  short term utility by choosing actions that can reduce uncertainty.
+  Adaptive probing strategies arise when it is valuable to intentionally
+  force a system far from the observed values even when the expected
+  value such actions is low, as it provides much faster learning and
+  consequent reduction of model uncertainty that can allow greater value
+  to be derived later on. For instance, \citet{Ludwig1982} show that it
+  may be advantageous to fish an unexploited population very heavily at
+  first to obtain a better estimate of the recruitment rate. This
+  intuitive strategy when a population is governed by a Ricker or
+  Beverton-Holt-like dynamic would clearly be disastrous if instead the
+  dynamics contained an unforeseen tipping point. The best way to learn
   where the edge lies may be to walk up to it, but it is also the most
-dangerous.
+  dangerous.} arises when applying a dynamical model outside the range
+of data on which it has been estimated. This extrapolation uncertainty
+is felt most keenly in decision-theoretic applications, as (a) exploring
+the potential action space typically involves considering actions that
+may move the system outside the range of observed behavior, and (b)
+decision-theoretic alogrithms rely not only on reasonable estimates of
+the expected outcomes, but depend on the weights given to all possible
+outcomes \citep[\emph{e.g.}][]{Weitzman2013}.
+
+Concerns over the potential for tipping points in ecological dynamics
+\citep{Scheffer2001} highlight the dangers of uncertainty in ecological
+management and pose a substantial challenge to existing
+decision-theoretic approaches \citep{Brozovic2011}. Because intervention
+is often too late after a tipping point has been crossed (but see
+\citet{Hughes2013}), management is most often concerned with avoiding
+potentially catastrophic tipping points before any data is available at
+or following a transition that would more clearly reveal these regime
+shift dynamics \citep[e.g.][]{Bestelmeyer2012}.
+
+\emph{Map}
+
+Here we illustrate how a stochastic dynamic programming (SDP) algorithm
+{[}\citet{Mangel1988}; @Possingham1997, Marescot2013{]} can be driven by
+the predictions from a Bayesian non-parametric (BNP) approach
+\citep{Munch2005a}.\\This provides two distinct advantages compared with
+contemporary approaches. First, using a BNP sidesteps the need for an
+accurate model-based description of the system dynamics.\\Second, the
+BNP can better reflect uncertainty that arises when extrapolating a
+model outside of the data on which it was fit.\\We illustrate that when
+the correct model is not known, this latter feature is crucial to
+providing a robust decision-theoretic approach in face of substantial
+structural uncertainty.
 
@@
 \begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Complex dynamics
+\item
+  Complex dynamics \citep{Glaser2013}(http://doi.org/10.1111/faf.12037
+  ``Complex dynamics may limit prediction in marine fisheries'')
 \item
@@
 \item
-  MAP
-\end{itemize}
-
-Decision making under uncertainty is a ubiquitous challenge of natural
-resource management and conservation.
-\href{http://doi.org/10.1111/faf.12037}{@Glaser2013}
-
-Here we illustrate how a stochastic dynamic programming (SDP) algorithm
-can be driven by the predictions from a Gaussian Process (GP). This
-provides two distinct advantages compared with contemporary approaches.
-First, using a GP sidesteps the need for an accurate model-based
-description of the system dynamics (or a model choice or model-averaging
-approach which introduces yet more difficulty).\\Second, unlike
-parametric models which can only reflect uncertainty in the
-corresponding confidence intervals or posterior distributions of their
-parameters, the GP approach provides a state-space dependent
-representation of uncertainty. This permits a much greater uncertainty
-far from the observed data than near the observed data. These features
-allow the GP-SDP approach to find robust management solutions in face of
-both limited data and no knolwedge of correct model structure.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Note on ``not magic'': honest uncertainty + SDP
+  \emph{MAP}
+\end{itemize}
+
+This paper represents the first time the SDP decision-making framework
+has been used without an a priori model of the underlying dynamics
+through the use of the BNP approach.
+
+In contrast to parametric models which can only reflect uncertainty in
+parameter estimates, the BNP approach provides a more state-space
+dependent representation of uncertainty.
+
+This permits a much greater uncertainty far from the observed data than
+near the observed data. These features allow the GP-SDP approach to find
+robust management solutions in face of limited data and without
+knowledge of the correct model structure.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \emph{Note on ``not magic''}: honest uncertainty + SDP
 \end{itemize}
@@
 able to obtain more robust management policies than with common
-parameteric modeling approaches.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Note on comparing models (via value function rather than by ``fit'').
+parametric modeling approaches.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Note on: Why fisheries
+\end{itemize}
+
+The economic value and ecological concern have made marine fisheries the
+crucible for much of the founding work
+\citep{Gordon1954, Reed1979, May1979, Ludwig1982} in managing ecosystems
+under uncertainty. Global trends \citep{Worm2006} and controversy
+\citep{Hilborn2007, Worm2009} have made understanding these challenges
+all the more pressing.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \emph{Note on comparing models} (via value function rather than by
+  ``fit'').
+\end{itemize}
+
 \emph{Do we need this?}
-\end{itemize}
 
@@
 social value of fish left in the sea vs.~the commercial value of fish
-harvested; see Halpern et al. (2013) for further discussion of this
-issue. Alternatively, we can always compare model performance across
-multiple potential objective functions.) The decision-maker does not
-necessarily need a model that provides the best mechanistic
-understanding or the best long-term outcome, but rather the one that
-best estimates the probabilities of being in different states as a
-result of the possible actions.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Note on terminology: ``Non-parametric'' \emph{Do we need this?}
-\end{itemize}
+harvested; see \citet{Halpern2013} for further discussion of this issue.
+Alternatively, we can always compare model performance across multiple
+potential objective functions.) The decision-maker does not necessarily
+need a model that provides the best mechanistic understanding or the
+best long-term outcome, but rather the one that best estimates the
+probabilities of being in different states as a result of the possible
+actions.
 
@@
 \item
+  Background on non-parametric modeling.
+\end{itemize}
+
+Addressing the difficulty posed by extrapolation without knowing the
+true model requires a nonparametric approach to model fitting: one that
+does not assume a fixed structure but rather depends on the size of the
+data (e.g.~non-parametric regression or a Dirichlet process). This
+established terminology is nevertheless unfortunate, as (a) this
+approach still involves the estimation of parameters, and (b),
+Statisticians use non-parametric to mean both this property (structure
+is not fixed by the parameters) and an entirely different (and probably
+more familiar) case in which the model does not assume any distribution
+(e.g.~non-parametric bootstrap, order statistics). Some literature thus
+uses the term semi-parametric, which merely adds ambiguity to the
+confusion.
+
+This non-parametric property -- having a structure explicitly dependent
+on the data -- is precisely the property that makes this approach
+attractive in face of the limited data sampling challenges discussed
+above. Having fit a parametric model to some data, the model is
+completely described by the values (or posterior distributions) of it's
+parameters. The non-parametric model is not captured by its parameter
+values or distributions alone. Either the model scales with the
+complexity of the data on which it is estimated (e.g.~nonparametric
+heirarchical approaches such as the Dirchlet process) or the data points
+become themselves part of the model specification, as in the
+nonparametric regression used here. we shall see here.
+
+\begin{itemize}
+\item
   Definition
@@
   Previous application
-\item
-  Why it is particularly suited to these two problems
-\item
-  (Why this is a novel application thereof)
-\end{itemize}
-
-\section{Approach and Methods}
-
-\subsubsection{Summary of approach}
-
-\subsubsection{Statement of the optimal control problem}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Underlying model
-\item
-  Available data
-\item
-  Value function
-\end{itemize}
-
-For simplicity we assume profit is simply linear in the realized harvest
-(only enforcing the restriction that harvest can not exceed available
-stock)
-
-\subsubsection{Parametric models}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Statement of the models
-\end{itemize}
-
-We consider three candidate parametric models of the stock-recruitment
-dynamics: The Ricker model, the Allen model \href{}{Allen 2005}, the
-Myers model. The familiar Ricker model involves two parameters,
-corresponding to a growth rate and a carrying capacity, and cannot
-support alternative stable state dynamics (though as growth rate
-increases it exhibits a periodic attractor that proceeds through
-period-doubling into chaos. We will generally focus on dynamics below
-the chaotic threshold for the purposes of this analysis.) The Allen
-model resembles the Ricker dynamics with an added Allee effect parameter
-\href{}{Courchamp}, below which the population cannot persist. The Myers
-model also has three parameters and contains an Allee threshold, but has
-compensatory rather than over-compensatory density dependence
-(resembling a Beverton-Holt curve rather than a Ricker curve at high
-densities.)
-
-We assume multiplicative log-normal noise perturbs the growth predicted
-by the each of the deterministic model skeletons described above. This
-introduces one additional parameter $\sigma$ that must be estimated by
-each model.
-
-As we simulate training data from the Allen model (ref section), we will
-refer to this as the structurally correct model. The Ricker model is
-thus a reasonable approximation of these dynamics far from the Allee
-threshold (but lacks threshold dynamics), while the Myers model shares
-the essential feature of a threshold but differs in the structure. Thus
-we have three potential parametric models of the stock dynamics.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Bayesian inference of parametric models
-\end{itemize}
-
-We infer posterior distributions for the parameters of each model in a
-Bayesian context using Gibbs sampling (implemented in R \href{}{R} using
-jags, \href{}{jags}). We choose uninformative uniform priors for all
-parameters (See Appendix, Figures S1-S3, and Table S1, and the R code
-provided). One-step-ahead predictions of these model fits are shown in
-Figure 1.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  SDP via parametric models
-\end{itemize}
-
-An optimal policy function is then inferred through stochastic dynamic
-programming for each model given the posterior distributions of the
-parameter estimates. This policy maximizes the expectation of the value
-function integrated over the parameter uncertainty. (code implementing
-this algorithm provided in the Appendix).
-
-\subsubsection{The Gaussian Process model}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Statement of model
-\end{itemize}
-
-\ldots{} more on GP \ldots{} \href{}{Munch 2005}
-
-We also estimate a simple Gaussian Process defined by a radial basis
-function kernel of two parameters: $\ell$, which gives the
-characteristic length-scale over which correlation between two points
-(e.g.~any two points $X_t, X_{t+1}$, and $X_{t+\tau}, X_{t+1+\tau}$) in
-state-space decays, and $\sigma$, which gives the scale of the process
-noise by which observations $Y_{t+1}$ may differ from their predicted
-values $X_{t+1}$ given an observation of the previous state, $X_t$.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Inference of the model
-\end{itemize}
-
-Also unlike parametric models, this posterior distribution is still
-conditional on the training data. As such, the uncertainty near the
-observed data.
-
-We use a Metropolis-Hastings Markov Chain Monte Carlo to infer posterior
-distributions of the two parameters of the GP (Figure S4, code in
-appendix), under weakly informative Gaussian priors (see parameters in
-table S5). As the posterior distributions differ substantially from the
-priors (Figure S4), we can be assured that most of the information in
-the posterior comes from the data rather than the prior belief.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  SDP via the model
-\end{itemize}
-
-Though we are unaware of prior application of this type, it is
-reasonably straight-forward to adapt the Gaussian Process for Stochastic
-Dynamic Programming. Recall that unlike the parametric models the
-Gaussian process with fixed parameters already predicts a distribution
-of curves rather than a single curve. We must first integrate over this
-distribution of curves given a sampling of parameter values drawn from
-the posterior distribution of the two GP parameters, before integrating
-over the posterior of those parameters themselves.
-
-\section{Results}
-
-\subsubsection{Figure 1: Fitted Models}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  All models fit the data quite well
-\item
-  Information criteria would pick the simple, incorrect model.
-\end{itemize}
-
-(State-space figure??)
-
-\subsubsection{Figure 2: Inferred Policies}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Inferred policies differ substantially among models
-\item
-  The structurally correct model and the GP are close to the true model
-\item
-  alternatives are not close
-\end{itemize}
-
-\subsubsection{Figure 3: Simulated results}
-
--
-
-\subsubsection{Figure 4: Robustness}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Results hold across range of parameters
-\end{itemize}
-
-(eek, distinguish better between result and discussion?)
-
-\section{Discussion}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  All models are ``good fits'' to the originally observed data.
-\item
-  (Simple model choice immediately leads us astray)
-\end{itemize}
-
-\subsection{Future directions}
-
-\subsubsection{Higher dimensions}
-
-\subsubsection{Online learning}
-
-\section{Acknowledgments}
-
-\section{Appendix}
-
-\subsection{Model definitions and estimation}
-
-Equation S1: Ricker model.
-
-\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
-
-Figure S1: Ricker model: prior and posterior distributions for parameter
-estimates.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-1.pdf}
-\caption{plot of chunk unnamed-chunk-1}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-2.pdf}
-\caption{plot of chunk unnamed-chunk-2}
-\end{figure}
-
-Table S1: Parameterization of the priors
-
-\% latex table generated in R 3.0.1 by xtable 1.7-1 package \% Mon Sep 9
-14:43:29 2013
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rlrr}
-  \hline
- & parameter & lower\_bound & upper\_bound \\ 
-  \hline
-1 & r0 & 0.00 & 10.00 \\ 
-  2 & K & 0.00 & 40.00 \\ 
-  3 & sigma & 0.00 & 100.00 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-\[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \]
-
-Eq S2: Myers model Figure S2: Myers model: Traces, prior and posterior
-distributions for parameter estimates.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-4.pdf}
-\caption{plot of chunk unnamed-chunk-4}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-5.pdf}
-\caption{plot of chunk unnamed-chunk-5}
-\end{figure}
-
-Table S2: Parameterization of the priors \% latex table generated in R
-3.0.1 by xtable 1.7-1 package \% Mon Sep 9 14:43:37 2013
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rlrr}
-  \hline
- & parameter & lower\_bound & upper\_bound \\ 
-  \hline
-1 & r0 & 0.00 & 10.00 \\ 
-  2 & K & 0.00 & 40.00 \\ 
-  3 & theta & 0.00 & 10.00 \\ 
-  4 & sigma & 0.00 & 100.00 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-Eq S3: Allen model
-
-\[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]
-
-Figure S3: Allen model: prior and posterior distributions for parameter
-estimates.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-7.pdf}
-\caption{plot of chunk unnamed-chunk-7}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-8.pdf}
-\caption{plot of chunk unnamed-chunk-8}
-\end{figure}
-
-Table S3: Parameterization of the priors
-
-\% latex table generated in R 3.0.1 by xtable 1.7-1 package \% Mon Sep 9
-14:43:45 2013
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rlrr}
-  \hline
- & parameter & lower\_bound & upper\_bound \\ 
-  \hline
-1 & r0 & 0.00 & 10.00 \\ 
-  2 & K & 0.00 & 40.00 \\ 
-  3 & theta & 0.00 & 10.00 \\ 
-  4 & sigma & 0.00 & 100.00 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-Eq S4: GP model Figure S4: GP model: prior and posterior distributions
-for parameter estimates.
-
-\begin{verbatim}
-$traces_plot
-\end{verbatim}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-101.pdf}
-\caption{plot of chunk unnamed-chunk-10}
-\end{figure}
-
-\begin{verbatim}
-
-$posteriors_plot
-\end{verbatim}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-102.pdf}
-\caption{plot of chunk unnamed-chunk-10}
-\end{figure}
-
-Table S4: Parameterization of the priors
-
-\subsection{Optimal Control Problem}
-
-We seek the harvest policy $h(x)$ that maximizes:
-
-\[ \max_{h_t} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h_t) \delta^t  \]
-
-subject to the profit function $\Pi(X_t,h)$, discount rate $\delta$, and
-the state equation
-
-\[X_{t+1} = Z_t f(S_t)  \] \[S_t = X_t - h_t \]
-
-Where $Z_t$ is multiplicative noise function with mean 1, representing
-stochastic growth. We will consider log-normal noise with shape
-parameter $\sigma_g$.
-
-Form this we can write down the Bellman recursion as:
-
-\[V_t(x_t) = \max_h \mathbf{E} \left(\Pi(h_t, x_t) + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)\]
-
-For simplicity we assume profit is simply linear in the realized harvest
-(only enforcing the restriction that harvest can not exceed available
-stock), $\Pi(h,x) = min(h,x)$.
-
-\subsubsection{Pseudocode for the Bellman iteration}
-
-\begin{Shaded}
-\begin{Highlighting}[]
- \NormalTok{V1 <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
-      \NormalTok{delta * F[[h]] %*% V +  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
-    \NormalTok{\})}
-    \CommentTok{# find havest, h that gives the maximum value}
-    \NormalTok{out <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
-      \NormalTok{value <- }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
-      \NormalTok{index <- }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
-      \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
-    \NormalTok{\})}
-    \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
-    \NormalTok{V <- out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
-    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <- out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
-\end{Highlighting}
-\end{Shaded}
-
-\subsubsection{Training data}
-
-Eacho of our models $f(S_t)$ must be estimated from training data, which
-we simulate from the Allen model with parameters $r = $
-\texttt{r p{[}1{]}}, $K =$ \texttt{r p{[}2{]}}, $C =$
-\texttt{r p{[}3{]}}, and $\sigma_g =$ \texttt{r sigma\_g} for $T=$
-
-\begin{verbatim}
-
-Error in base::parse(text = code, srcfile = NULL) : 1:16: unexpected ','
-1: Tobs$ timesteps,
-                  ^
-\end{verbatim}
-
-r Xo`.
-
-\begin{center}\rule{3in}{0.4pt}\end{center}
-
-\section{Abstract}
-
-Decision-theoretic methods often rely on simple parametric models of
-ecological dynamics to compare the value of a potential sequence of
-actions. Unfortunately, such simple models rarely capture the complexity
-or uncertainty found in most real ecosystems.
-
-Further, the data on which a model has been parameterized frequently
-fails to cover the possible state-space over which management decisions
-must operate. Consequently a model do well in the region of state-space
-in which it was estimated, but give erroneous confidence to predictions
-outside of that region.
-
-This problem is keenly felt in any system where a potential threshold or
-tipping point is a concern. Such a tipping point, if it exists at all,
-will lay outside the observed range of the observed data.
-
-We demonstrate how nonparametric Bayesian models can provide robust,
-solutions to decision making under uncertainty without knowing the
-structural form of the true model.
-
-While methods that account for \emph{parametric} uncertainty can be very
-successful with the right model, structural uncertainty of not knowing
-what model best approximates the dynamics poses considerably greater
-difficulty.
-
-\section{Introduction}
-
-\paragraph{Opening}
-
-\paragraph{Models for decision-making under uncertainty}
-
-Decision-theoretic or optimal control tools require a model that can
-assign probabilities of future states (e.g.~stock size of a fishery)
-given the current state and a proposed action (e.g.~fishing harvest or
-effort).\\Management frequently faces a sequential decision-making
-problem -- after selecting an action, the decision-maker may receive new
-information about the current state and must again choose an appropriate
-action -- such as setting the harvest limits each year based on stock
-assessments the year prior.
-
-The decision maker typically seeks to determining the course of actions
-(also referred to as the policy) that maximizes the expected value of
-some objective function such as net present value derived from the
-resource over time.\\Though much can be said on how to choose this value
-function appropriately (e.g.~see (Halpern et al.
-2013)(http://doi.org/10.1073/pnas.1217689110 ``Achieving the triple
-bottom line in the face of inherent trade-offs among social equity,
-economic return, and conservation.'')) we will assume this is given.
-(Nor is this approach necessarily constrained to maximizing the
-expectated value of such a function - the decision-theoretic framework
-can be adapted to alternatives such as minimizing the maximum cost or
-damage that might be incurred; see Polasky et al. (2011)).
-
-In representing future states with probabilities and maximizing
-expectations, this approach provides a natural framework for handling
-uncertainty.
-
-The value function typically depends on the action or policy taken, as
-well as the state of the system, in each interval of time. The state of
-the system, in turn, is usually described by a dynamical model.
-
-\paragraph{Parametric uncertainty}
-
-\paragraph{Structural Uncertainty}
-
-(B. K. Williams 2001; Athanassoglou and Xepapadeas 2012).
-
-While simple mechanistic models can nevertheless provide important
-insights into long-term outcomes, such approaches are not well-suited
-for use in forecasting outcomes of potential management options.
-Non-parametric approaches offer a more flexible alternative that can
-both more accurately reflect the data available while also representing
-greater uncertainty in areas (of state-space) where data is lacking.
-
-We demonstrate how a Gaussian Process model of stock recruitment can
-lead to nearly optimal management through stochastic dynamic
-programming, comperable to knowing the correct structural equation for
-the underlying simulation. Meanwhile, parametric models that do not
-match the underlying dynamics can perform very poorly, even though they
-fit the data as well as the true model.\\Ecological research and
-management strategy should pay closer attention to the opportunities and
-challenges nonparametric modeling can offer.
-
-\section{Approach and Methods}
-
-\subsection{The optimal control problem in fisheries management}
-
-In our example, we focus on the problem in which a manager must set the
-harvest level for a marine fishery each year to maximize the net present
-value of the resource, given an estimated stock size from the year
-before.
-
-To permit comparisons against a theoretical optimum we will consider
-data on the stock dynamics simulated from a simple parametric model in
-which recruitment of the fish stock $X_{t+1}$ in the following year is a
-stochastic process governed by a function $f$ of the current stock
-$X_t$, selected harvest policy $h_t$, and noise process $Z$,
-
-\[X_{t+1} = Z_t f(X_t, h_t) \]
-
-Given parameters for the function $f$ and probability distribution $Z$,
-along with a given economic model determining the price/profit
-$\Pi(X_t, h_t)$ realized in a given year given a choice of harvest $h_t$
-and observed stock $X_t$. This problem can be solved exactly for
-discretized values of stock $X$ and policy $h$ using stochastic dynamic
-programming (SDP) (Mangel 1985). Problems of this sort underpin much
-marine fisheries management today.
-
-A crux of this approach is correctly specifying the functional form of
-$f$, along with its parameters. The standard approach uses one of a
-handful of common parametric models representing the stock-recruitment
-relationship, usually after estimating the model parameters from any
-available existing data. Uncertainty in the parameter estimates can be
-estimated and integrated over to determine the optimal policy under
-under uncertainty (Mangel 1985; Schapaugh and Tyre 2013). Uncertainty in
-the model structure itself can only be addressed in this approach by
-hypothesizing alternative model structures, and then performing some
-model choice or model averaging (B. K. Williams 2001; Athanassoglou and
-Xepapadeas 2012).
-
-\subsection{Underlying Model}
-
-Concerns over the potential for tipping points in ecological dynamics
-(Scheffer et al. 2001) highlight the dangers of uncertainty in
-ecological management and pose a substantial challenge to existing
-decision-theoretic approaches (Brozović and Schlenker 2011). Because
-intervention is often too late after a tipping point has been crossed
-(but see Hughes et al. (2013)), management is most often concerned with
-avoiding potentially catastrophic tipping points before any data is
-available at or following a transition that would more clearly reveal
-these regime shift dynamics (e.g. Bestelmeyer et al. 2012).
-
-To illustrate the value of the non-parametric Bayesian approach to
-management, we focus on example of a system containing such a tipping
-point whose dynamics can still be described by a simple, one-dimensional
-parametric model.\\We will focus on a simple parametric model for a
-single species (derived from fist principles by Allen et al. 2005) as
-our underlying ``reality''.
-
-\[X_{t+1} = Z_t f(S_t)  \] \[S_t = X_t - h_t \]
-\[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]
-
-Where $Z_t$ is multiplicative noise function with mean 1, representing
-stochastic growth. We will consider log-normal noise with shape
-parameter $\sigma_g$. We start with an example in which the parameters
-are $r =2$, $K=8$, $C=5$ and $\sigma_g = 0.1$.
-
-As a low-dimensional system completely described by three parameters,
-this scenario should if anything be favorable to a parametric-based
-approach. This model contains an Allee effect, or tipping point, below
-which the population is not self-sustaining and shrinks to zero
-(Courchamp, Berec, and Gascoigne 2008).
-
-\paragraph{Simulated training data}
-
-We generate initial observational data under the model described in Eq 1
-for $T_{\textrm{obs}}=40$ time steps, under a given arbitrary sequence
-of harvest intensities, $h_t$. We consider the case in which most of the
-data comes from a limited region of state space (e.g.~near a stable
-equilibrium), leaving us without observations of the population dynamics
-at very low levels which would be useful in discrimating between
-recruitment curves {[}@{]} or demonstrating the existence of a tipping
-point (Scheffer et al. 2001).\\Using data simulated from a specified
-model rather than empirical data permits the comparison against the true
-underlying dynamics, setting a bar for the optimal performance possible.
-
-\subsubsection{Parametric Models}
-
-We consider three candidate parametric models for the stock-recruitment
-function, which we refer to by the first authors of the publications in
-which they were first proposed.
-
-We generate the data with a four-parameter model that contains a tipping
-point, as discussed above (equation 1), (an Allee effect, see {[}@Allen,
-@Courchamp{]}) below which the stock decreases to zero,
-
-\[ X_{t+1} = Z_t S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(\frac{S_t - \theta}{K}\right)} \]
-
-\[ S_t = X_t - h_t \]
-
-The parameter $C$ reflects the location of the tipping point, $K$ the
-carrying capacity of the stock, and $r$ the base recruitment rate. $S_t$
-represents the stock size after a harvest $h_t$ has been implemented.
-$Z_t$ represents a log-normal random variable of log-mean zero and
-log-standard deviation parameter $\sigma$.
-
-We consider two alternative candidate models: the Ricker {[}@Ricker{]}
-stock-recruitment curve,
-
-\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
-
-and an alternative four-parameter model adapted from @Myers,
-
-\[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \]
-
-which contains a tipping point for $\theta > 2$ and becomes a
-Beverton-Holt model at $\theta = 1$.
-
-\subsubsection{Bayesian Inference of Parametric models}
-
-Given the sample data, we infer posterior distributions for each of the
-three models listed above using a Markov Chain Monte Carlo Gibbs Sampler
-(jags, see appendix for implementation details and code) given uniform
-priors. We run six chains for $10^6$ steps each and then assess
-convergence by Gelman-Rubin criterion and inspection of the traces, see
-appendix.
-
-\subsubsection{The Non-parametric Bayesian alternative for
-stock-recruitment curves}
-
-\paragraph{Terminology: ``Non-parametric''}
-
-The term non-parametric to describe modeling approaches such as Gaussian
-processes is a common source of confusion, notwithstanding the fact that
-the model is still specified by parameters. Some literature has
-introduced the term ``semi-parametric'' in tacit acknowledgment of this,
-which no doubt only contributes to the confusion. The problem is further
-exacerbated by the several meanings assigned to the term in statistics:
-(a) that the method does not assume some particular probability
-distribution (e.g.~a non-parametric bootstrap) or (b) that the method
-does not assume a fixed structure to the model. Our use is that of (b),
-in contrast to the classical approaches that do. The difference is
-clearest by way of example.
+\end{itemize}
 
@@
 geospatial literature) to formulate a predictive model is relatively new
-in the context of modeling dynamical systems (Kocijan et al. 2005) and
-introduced in the ecological modeling and fisheries management by Munch
-et al. (2005). An accessible and thorough introduction to the
-formulation and use of GPs can be found in Rasmussen and Williams
-(2006).
+in the context of modeling dynamical systems \citep{Kocijan2005}, and
+was first introduced in the context ecological modeling and fisheries
+management in \citet{Munch2005}. An accessible and thorough introduction
+to the formulation and use of GPs can be found in \citet{Rasmussen2006}.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Why it is particularly suited to these two problems
+\item
+  (Why this is a novel application thereof)
+\end{itemize}
 
@@
 process model are estimated by Metropolis-Hastings algorithm, again with
-details and code provided in the Appendix. Rasmussen and Williams (2006)
+details and code provided in the Appendix. \citet{Rasmussen2006}
 provides an excellent general introduction to Gaussian Processes and
-Munch et al. (2005) first discusses their application in the context of
+\citet{Munch2005} first discusses their application in the context of
 population dynamics models such as fisheries stock-recruitment
@@
 
-\subsubsection{SDP via GP}
-
-Once the posterior Gaussian process (GP) has been estimated (e.g.~see
-Munch et al. 2005), it is necessary to adapt it in place of the
-parametric equation for the stochastic dynamic programming (SDP)
-solution (see Mangel and Clark 1988 for a detailed description of
-parametric SDP methods) to the optimal policy. The essence of the idea
-is straight forward -- we will use the estimated GP in place of the
-parametric growth function to determine the stochastic transition matrix
-on which the SDP calculations are based. The SDP is solved in a
-discretized state space -- both the continuously valued population
-densities $X$ and harvest quotas $h$ are first mapped to a bounded,
-discrete grid. (For simplicity we will consider a uniform grid, though
-for either parametric or GP-based SDP it is often advantageous to use a
-non-uniform discretization such as a basis function representation,
-e.g.~see (Deisenroth, Rasmussen, and Peters 2009)).
-
-The SDP approach then computes a transition matrix, $\mathbf{F}$. We
-demonstrate that calculation is just as straight forward based on the GP
-as it is in the classical context using the parametric model. The
-${i,j}$ of the transition matrix $F$ entry gives the probability of
-transitioning into state $x_i$ given that the system is in state $x_j$
-in the previous time-step. To generate the transition matrix based on
-the posterior GP, we need only the expected values at each grid point
-and the corresponding variances (the diagonal of the covariance matrix),
-as shown in Figure 1. Given the mean of the GP posterior at each
-grid-point as the vector $E$ and variance at that point as vector $V$,
-the probability of transitioning from state $x_i$ to state $x_j$ is
-
-\[\mathcal{N}\left(x_j \vert  \mu = E_i, \sigma = \sqrt{V_i}\right)\]
-
-where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
-variance $\sigma^2$. Strictly speaking, the transition probability
-should be calculated by integrating the normal density over the bin of
-width $\Delta$ centered at $x_j$. For a sufficiently fine grid that
-$f(x_j) \approx f(x_j + \Delta)$, it is sufficient to calculate the
-density at $x_j$ and then row-normalize the transition matrix. The
-process can then be repeated for each possible discrete value of our
-control variable, (harvest $h$).
-
-\textbf{Pseudocode for the determining the transition matrix from the
-GP}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{for(h in h_grid)}
-  \NormalTok{F_h = for(x_j in grid)}
-          \NormalTok{for(i in }\DecValTok{1}\NormalTok{:N) }
-            \KeywordTok{dnorm}\NormalTok{(x_j, mu[i]-h, V[i])}
-\end{Highlighting}
-\end{Shaded}
-
-Using the discrete transition matrix we may write down the Bellman
-recursion defining the stochastic dynamic programming iteration:
-
-\begin{equation}
-V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
-\end{equation}
-
-where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$ is
-control (harvest level) chosen. Numerically, the maximization is
-accomplished as follows. Consider the set of possible control values to
-be the discrete values corresponding the grid of stock sizes. Then for
-each $h_t$ there is a corresponding transition matrix $\mathbf{F}_h$
-determined as described above but with mean $\mu = x_j - h_t$. Let
-$\vec{V_t}$ be the vector whose $i$th element corresponds to the value
-of having stock $x_i$ at time $t$. Then let $\Pi_h$ be the vector whose
-$i$th element indicates the profit from harvesting at intensity $h_t$
-given a population $x_i$ (e.g. $\max(x_i, h_t)$ since one cannot harvest
-more fish then the current population size). Then the Bellman recursion
-can be given in matrix form as
-
-\[V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)\]
-
-where the sum is element by element and the expectation is computed by
-the matrix multiplication $\mathbf{F} V_{t+1}$.
-
-\subsubsection{Pseudocode for the Bellman iteration}
-
-\begin{Shaded}
-\begin{Highlighting}[]
- \NormalTok{V1 <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
-      \NormalTok{delta * F[[h]] %*% V +  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
-    \NormalTok{\})}
-    \CommentTok{# find havest, h that gives the maximum value}
-    \NormalTok{out <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
-      \NormalTok{value <- }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
-      \NormalTok{index <- }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
-      \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
-    \NormalTok{\})}
-    \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
-    \NormalTok{V <- out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
-    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <- out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
-\end{Highlighting}
-\end{Shaded}
-
-This completes the algorithm adapting the GP to the sequential
-decision-making problem through SDP, which has not been previously
-demonstrated.\\We further provide an R package implementation as
-described in the supplemental materials.
-
-\subsubsection{Estimating parametric models}
-
-We estimate posterior distributions for two parametric models: one using
-the structurally correct model as given in Eq (1), which we refer to as
-the ``Parametric Bayes'' model, and another using the familiar Ricker
-model, using a Gibbs sampler as described (with source code) in the
-appendix). In addition we estimate the parameters of the structurally
-correct model by maximum likelihood.
+\section{Approach and Methods}
+
+\subsubsection{Summary of approach}
+
+\subsubsection{Statement of the optimal control problem}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Underlying model
+\item
+  Available data
+\item
+  Value function
+\end{itemize}
+
+For simplicity we assume profit is simply linear in the realized harvest
+(only enforcing the restriction that harvest can not exceed available
+stock)
+
+\subsubsection{Parametric models}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Statement of the models
+\end{itemize}
+
+We consider three candidate parametric models of the stock-recruitment
+dynamics: The Ricker model, the Allen model \href{}{Allen 2005}, the
+Myers model. The familiar Ricker model involves two parameters,
+corresponding to a growth rate and a carrying capacity, and cannot
+support alternative stable state dynamics (though as growth rate
+increases it exhibits a periodic attractor that proceeds through
+period-doubling into chaos. We will generally focus on dynamics below
+the chaotic threshold for the purposes of this analysis.) The Allen
+model resembles the Ricker dynamics with an added Allee effect parameter
+\href{}{Courchamp}, below which the population cannot persist. The Myers
+model also has three parameters and contains an Allee threshold, but has
+compensatory rather than over-compensatory density dependence
+(resembling a Beverton-Holt curve rather than a Ricker curve at high
+densities.)
+
+We assume multiplicative log-normal noise perturbs the growth predicted
+by the each of the deterministic model skeletons described above. This
+introduces one additional parameter $\sigma$ that must be estimated by
+each model.
+
+As we simulate training data from the Allen model (ref section), we will
+refer to this as the structurally correct model. The Ricker model is
+thus a reasonable approximation of these dynamics far from the Allee
+threshold (but lacks threshold dynamics), while the Myers model shares
+the essential feature of a threshold but differs in the structure. Thus
+we have three potential parametric models of the stock dynamics.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Bayesian inference of parametric models
+\end{itemize}
+
+We infer posterior distributions for the parameters of each model in a
+Bayesian context using Gibbs sampling (implemented in R {[}@RTeam{]}
+using jags, {[}@R2jags{]}). We choose uninformative uniform priors for
+all parameters (See Appendix, Figures S1-S3, and Table S1, and the R
+code provided). One-step-ahead predictions of these model fits are shown
+in Figure 1.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  SDP via parametric models
+\end{itemize}
+
+An optimal policy function is then inferred through stochastic dynamic
+programming for each model given the posterior distributions of the
+parameter estimates. This policy maximizes the expectation of the value
+function integrated over the parameter uncertainty. (code implementing
+this algorithm provided in the Appendix).
+
+\subsubsection{The Gaussian Process model}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Statement of model
+\end{itemize}
+
+\ldots{} more on GP \ldots{} \href{}{Munch 2005}
+
+We also estimate a simple Gaussian Process defined by a radial basis
+function kernel of two parameters: $\ell$, which gives the
+characteristic length-scale over which correlation between two points
+(e.g.~any two points $X_t, X_{t+1}$, and $X_{t+\tau}, X_{t+1+\tau}$) in
+state-space decays, and $\sigma$, which gives the scale of the process
+noise by which observations $Y_{t+1}$ may differ from their predicted
+values $X_{t+1}$ given an observation of the previous state, $X_t$.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Inference of the model
+\end{itemize}
+
+Also unlike parametric models, this posterior distribution is still
+conditional on the training data. As such, the uncertainty near the
+observed data.
+
+We use a Metropolis-Hastings Markov Chain Monte Carlo to infer posterior
+distributions of the two parameters of the GP (Figure S4, code in
+appendix), under weakly informative Gaussian priors (see parameters in
+table S5). As the posterior distributions differ substantially from the
+priors (Figure S4), we can be assured that most of the information in
+the posterior comes from the data rather than the prior belief.
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  SDP via the model
+\end{itemize}
+
+Though we are unaware of prior application of this type, it is
+reasonably straight-forward to adapt the Gaussian Process for Stochastic
+Dynamic Programming. Recall that unlike the parametric models the
+Gaussian process with fixed parameters already predicts a distribution
+of curves rather than a single curve. We must first integrate over this
+distribution of curves given a sampling of parameter values drawn from
+the posterior distribution of the two GP parameters, before integrating
+over the posterior of those parameters themselves.
 
@@
 
-While it would be natural (and straight forward) to condition the GP on
-passing through the origin (0,0) (see appendix), the estimate shown here
-is based only on the observed data. The observed data from which each
-model is estimated is also shown. The observations come from only a
-limited region of state space corresponding to unharvested or weakly
-harvested system. No observations occur at the theoretical optimum or
-near the tipping point.
+While it would be straight forward to condition the GP on passing
+through the origin (0,0), the estimate shown here is based only on the
+observed data. The observed data from which each model is estimated is
+also shown. The observations come from only a limited region of state
+space corresponding to unharvested or weakly harvested system. No
+observations occur at the theoretical optimum harvest rate or near the
+tipping point.
 
@@
 $S_t$, as under models such as this a constant escapement policy is
-expected to be optimal (Reed 1979).}
+expected to be optimal \citep{Reed1979}.}
 \end{figure}
@@
 $S_t$. Under models such as this a constant escapement policy is
-expected to be optimal (Reed 1979), whereby population levels below a
-certain size $S$ are unharvested, while above that size the harvest
+expected to be optimal \citep{Reed1979}, whereby population levels below
+a certain size $S$ are unharvested, while above that size the harvest
 strategy aims to return the population to $S$, resulting in the
@@
 
+\subsubsection{Figure 1: Fitted Models}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  All models fit the data quite well
+\item
+  Information criteria would pick the simple, incorrect model.
+\end{itemize}
+
+\subsubsection{Figure 2: Inferred Policies}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Inferred policies differ substantially among models
+\item
+  The structurally correct model and the GP are close to the true model
+\item
+  alternatives are not close
+\end{itemize}
+
+\subsubsection{Figure 3: Simulated results}
+
+-
+
+\subsubsection{Figure 4: Robustness}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Results hold across range of parameters
+\end{itemize}
+
+(eek, distinguish better between result and discussion?)
+
 \section{Discussion}
 
-\paragraph{Big picture: Linking GP to SDP}
-
-Non-parametric Bayesian methods have received far too little attention
-in ecological modeling efforts that are aimed at improved conservation
-planning and decision making support. Such approaches may be
-particularly useful when the available data is restricted to a limited
-area of state-space, which can lead parametric models to underestimate
-the uncertainty in dynamics at population levels (states) which have not
-been observed. One reason for the relative absence of nonparametric
-approaches in the natural resource management context may be the lack of
-existing approaches for adapting the non-parametric Bayesian models
-previously proposed (Munch et al. 2005) to a decision-theoretic
-framework. Adapting a non-parametric approach requires modification of
-existing methods for decision theory. We have illustrated how this might
-be done for a classic stochastic dynamic programming problem, opening
-the door for substantial further research into how these applications
-might be improved.
-
-Though simple mechanistically motivated models may be best for the
-latter case (Cuddington, Fortin, and Gerber 2013; Geritz and Kisdi
-2012), such models can be not only inaccurate but misleading in
-quantitative decision making. This arises for two reasons:
+\begin{itemize}
+\item
+  All models are ``good fits'' to the originally observed data.
+\item
+  (Simple model choice immediately leads us astray)
+\end{itemize}
+
+Though simple mechanistically motivated models offer the greatest
+potential to increase our basic understanding of ecological processes
+\citep{Cuddington2013, Geritz2012}, such models can be not only
+inaccurate but misleading when relied upon in a quantitative decision
+making framework.
 
@@
 \item
-  We do not know what the correct parameter values are for the models.
-\item
   We do not know what the correct models are for ecological systems.
+\item
+  We have limited data from which to estimate the model -- in
+  particular, such models may be misleading in predicting the
+  probability of outcomes outside the training data.
 \end{enumerate}
 
-The complexity of ecological interactions and a lack of data contribute
-greatly to both of the problems. This concern is particularly acute in
-case of ecological tipping points (Barnosky et al. 2012; Scheffer et al.
-2001), which arise from feedbacks common in ecological systems and can
-lead to a sudden catastrophic transition to an undesirable state. In
-general we do not know where such tipping points are unless we have
-already observed the transition, in which case it is frequently too late
-to respond. Though there may exist early warning signals for certain
-kinds of these transitions that are driven by slow changes (Scheffer et
-al. 2009), we do not know when, where, or how to apply them to the
-decision making context more generally (Boettiger and Hastings 2013).
-
-While non-parametric Bayesian approaches will not always be preferable
-to simple mechanistic models, we highlight three aspects of the problem
-consider here that make these methods particularly valuable. These
-aspects are common to many conservation decision making problems, which
-thus merit greater use of non-parametric approaches that can best take
-advantage of them.
+These aspects are common to many conservation decision making problems,
+which thus merit greater use of non-parametric approaches that can best
+take advantage of them.
 
@@
 of the observed data. This arises from the greater flexibility of the
-Gaussian process, which describes a large family of possible
-curves.\\While in a parametric context this over-fitting would be more
-worrisome -- a high-degree polynomial could fit the data even better --
-those concerns are driven by the resulting parametric fit outside the
-data, which may involve wild oscillations unsupported by the data. As we
-have seen in \#1, the GP is less vulnerable to such unjustified
-predictions outside the data, and is meanwhile free to benefit from the
-greater fit where the data is available.
+Gaussian process, which describes a large family of possible curves.
+Despite this flexibility, the GP can be described in relatively few
+parameters and is thus far less likely to overfit.
 
@@
 and the Fisheries Ecology Division, Southwest Fisheries Science Center,
-Santa Cruz, CA and by NSF~grant EF-0924195 to MM.
+Santa Cruz, CA and by NSF~grant EF-0924195 to MM and NSF grant
+DBI-1306697 to CB.
 
@@
 
-The appendices have not yet been assembled. Meanwhile, code to repeat
-the analyses, along with a complete log of all research conducted on
-this project, can be found at:
+\subsection{Model definitions and estimation}
+
+Equation S1: Ricker model.
+
+\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
+
+Figure S1: Ricker model: prior and posterior distributions for parameter
+estimates.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-1.pdf}
+\caption{plot of chunk unnamed-chunk-1}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-2.pdf}
+\caption{plot of chunk unnamed-chunk-2}
+\end{figure}
+
+Table S1: Parameterization of the priors
+
+\% latex table generated in R 3.0.1 by xtable 1.7-1 package \% Mon Sep 9
+14:43:29 2013
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rlrr}
+  \hline
+ & parameter & lower\_bound & upper\_bound \\ 
+  \hline
+1 & r0 & 0.00 & 10.00 \\ 
+  2 & K & 0.00 & 40.00 \\ 
+  3 & sigma & 0.00 & 100.00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \]
+
+Eq S2: Myers model Figure S2: Myers model: Traces, prior and posterior
+distributions for parameter estimates.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-4.pdf}
+\caption{plot of chunk unnamed-chunk-4}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-5.pdf}
+\caption{plot of chunk unnamed-chunk-5}
+\end{figure}
+
+Table S2: Parameterization of the priors \% latex table generated in R
+3.0.1 by xtable 1.7-1 package \% Mon Sep 9 14:43:37 2013
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rlrr}
+  \hline
+ & parameter & lower\_bound & upper\_bound \\ 
+  \hline
+1 & r0 & 0.00 & 10.00 \\ 
+  2 & K & 0.00 & 40.00 \\ 
+  3 & theta & 0.00 & 10.00 \\ 
+  4 & sigma & 0.00 & 100.00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+Eq S3: Allen model
+
+\[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]
+
+Figure S3: Allen model: prior and posterior distributions for parameter
+estimates.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-7.pdf}
+\caption{plot of chunk unnamed-chunk-7}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-8.pdf}
+\caption{plot of chunk unnamed-chunk-8}
+\end{figure}
+
+Table S3: Parameterization of the priors
+
+\% latex table generated in R 3.0.1 by xtable 1.7-1 package \% Mon Sep 9
+14:43:45 2013
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rlrr}
+  \hline
+ & parameter & lower\_bound & upper\_bound \\ 
+  \hline
+1 & r0 & 0.00 & 10.00 \\ 
+  2 & K & 0.00 & 40.00 \\ 
+  3 & theta & 0.00 & 10.00 \\ 
+  4 & sigma & 0.00 & 100.00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+Eq S4: GP model Figure S4: GP model: prior and posterior distributions
+for parameter estimates.
+
+\begin{verbatim}
+$traces_plot
+\end{verbatim}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-101.pdf}
+\caption{plot of chunk unnamed-chunk-10}
+\end{figure}
+
+\begin{verbatim}
+
+$posteriors_plot
+\end{verbatim}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-102.pdf}
+\caption{plot of chunk unnamed-chunk-10}
+\end{figure}
+
+Table S4: Parameterization of the priors
+
+\subsection{Optimal Control Problem}
+
+We seek the harvest policy $h(x)$ that maximizes:
+
+\[ \max_{h_t} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h_t) \delta^t  \]
+
+subject to the profit function $\Pi(X_t,h)$, discount rate $\delta$, and
+the state equation
+
+\[X_{t+1} = Z_t f(S_t)  \] \[S_t = X_t - h_t \]
+
+Where $Z_t$ is multiplicative noise function with mean 1, representing
+stochastic growth. We will consider log-normal noise with shape
+parameter $\sigma_g$.
+
+Form this we can write down the Bellman recursion as:
+
+\[V_t(x_t) = \max_h \mathbf{E} \left(\Pi(h_t, x_t) + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)\]
+
+For simplicity we assume profit is simply linear in the realized harvest
+(only enforcing the restriction that harvest can not exceed available
+stock), $\Pi(h,x) = min(h,x)$.
+
+\subsubsection{Pseudocode for the Bellman iteration}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+ \NormalTok{V1 <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
+      \NormalTok{delta * F[[h]] %*% V +  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
+    \NormalTok{\})}
+    \CommentTok{# find havest, h that gives the maximum value}
+    \NormalTok{out <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
+      \NormalTok{value <- }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
+      \NormalTok{index <- }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
+      \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
+    \NormalTok{\})}
+    \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
+    \NormalTok{V <- out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
+    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <- out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
+\end{Highlighting}
+\end{Shaded}
+
+\subsubsection{Training data}
+
+Eacho of our models $f(S_t)$ must be estimated from training data, which
+we simulate from the Allen model with parameters $r = $
+\texttt{r p{[}1{]}}, $K =$ \texttt{r p{[}2{]}}, $C =$
+\texttt{r p{[}3{]}}, and $\sigma_g =$ \texttt{r sigma\_g} for $T=$ 40
+timesteps, starting at initial condition $X_0 = $ 5.5. The training data
+can be seen in Figure 1.
+
+\begin{center}\rule{3in}{0.4pt}\end{center}
+
+\section{Abstract}
+
+Decision-theoretic methods often rely on simple parametric models of
+ecological dynamics to compare the value of a potential sequence of
+actions. Unfortunately, such simple models rarely capture the complexity
+or uncertainty found in most real ecosystems.
+
+Further, the data on which a model has been parameterized frequently
+fails to cover the possible state-space over which management decisions
+must operate. Consequently a model do well in the region of state-space
+in which it was estimated, but give erroneous confidence to predictions
+outside of that region.
+
+This problem is keenly felt in any system where a potential threshold or
+tipping point is a concern. Such a tipping point, if it exists at all,
+will lay outside the observed range of the observed data.
+
+We demonstrate how nonparametric Bayesian models can provide robust,
+solutions to decision making under uncertainty without knowing the
+structural form of the true model.
+
+While methods that account for \emph{parametric} uncertainty can be very
+successful with the right model, structural uncertainty of not knowing
+what model best approximates the dynamics poses considerably greater
+difficulty.
+
+\section{Introduction}
+
+\paragraph{Opening}
+
+\paragraph{Models for decision-making under uncertainty}
+
+Decision-theoretic or optimal control tools require a model that can
+assign probabilities of future states (e.g.~stock size of a fishery)
+given the current state and a proposed action (e.g.~fishing harvest or
+effort).\\Management frequently faces a sequential decision-making
+problem -- after selecting an action, the decision-maker may receive new
+information about the current state and must again choose an appropriate
+action -- such as setting the harvest limits each year based on stock
+assessments the year prior.
+
+The decision maker typically seeks to determining the course of actions
+(also referred to as the policy) that maximizes the expected value of
+some objective function such as net present value derived from the
+resource over time.\\Though much can be said on how to choose this value
+function appropriately (e.g.~see
+\citep{Halpern2013}(http://doi.org/10.1073/pnas.1217689110 ``Achieving
+the triple bottom line in the face of inherent trade-offs among social
+equity, economic return, and conservation.'')) we will assume this is
+given. (Nor is this approach necessarily constrained to maximizing the
+expectated value of such a function - the decision-theoretic framework
+can be adapted to alternatives such as minimizing the maximum cost or
+damage that might be incurred; see \citet{Polasky2011}).
+
+In representing future states with probabilities and maximizing
+expectations, this approach provides a natural framework for handling
+uncertainty.
+
+The value function typically depends on the action or policy taken, as
+well as the state of the system, in each interval of time. The state of
+the system, in turn, is usually described by a dynamical model.
+
+\citep{Williams2001, Athanassoglou2012}.
+
+While simple mechanistic models can nevertheless provide important
+insights into long-term outcomes, such approaches are not well-suited
+for use in forecasting outcomes of potential management options.
+Non-parametric approaches offer a more flexible alternative that can
+both more accurately reflect the data available while also representing
+greater uncertainty in areas (of state-space) where data is lacking.
+
+We demonstrate how a Gaussian Process model of stock recruitment can
+lead to nearly optimal management through stochastic dynamic
+programming, comperable to knowing the correct structural equation for
+the underlying simulation. Meanwhile, parametric models that do not
+match the underlying dynamics can perform very poorly, even though they
+fit the data as well as the true model.\\Ecological research and
+management strategy should pay closer attention to the opportunities and
+challenges nonparametric modeling can offer.
+
+\section{Approach and Methods}
+
+\subsection{The optimal control problem in fisheries management}
+
+We focus on the problem in which a manager must set the harvest level
+for a marine fishery each year to maximize the net present value of the
+resource, given an estimated stock size from the year before.
+
+To permit comparisons against a theoretical optimum we will consider
+data on the stock dynamics simulated from a simple parametric model in
+which recruitment of the fish stock $X_{t+1}$ in the following year is a
+stochastic process governed by a function $f$ of the current stock
+$X_t$, selected harvest policy $h_t$, and noise process $Z$,
+
+\[X_{t+1} = Z_t f(X_t, h_t) \]
+
+Given parameters for the function $f$ and probability distribution $Z$,
+along with a given economic model determining the price/profit
+$\Pi(X_t, h_t)$ realized in a given year given a choice of harvest $h_t$
+and observed stock $X_t$. This problem can be solved exactly for
+discretized values of stock $X$ and policy $h$ using stochastic dynamic
+programming (SDP) \citep{Mangel1988}. Problems of this sort underpin
+much marine fisheries management today.
+
+A crux of this approach is correctly specifying the functional form of
+$f$, along with its parameters. The standard approach uses one of a
+handful of common parametric models representing the stock-recruitment
+relationship, usually after estimating the model parameters from any
+available existing data. Uncertainty in the parameter estimates can be
+estimated and integrated over to determine the optimal policy under
+under uncertainty \citep{Mangel1988, Schapaugh2013}. Uncertainty in the
+model structure itself can only be addressed in this approach by
+hypothesizing alternative model structures, and then performing some
+model choice or model averaging \citep{Williams2001, Athanassoglou2012}.
+
+\subsection{Underlying Model}
+
+To illustrate the value of the non-parametric Bayesian approach to
+management, we focus on example of a system containing such a tipping
+point whose dynamics can still be described by a simple, one-dimensional
+parametric model.\\We will focus on a simple parametric model for a
+single species \citep[derived from fist principles by][]{Allen2005a} as
+our underlying ``reality''.
+
+\[X_{t+1} = Z_t f(S_t)  \] \[S_t = X_t - h_t \]
+\[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]
+
+Where $Z_t$ is multiplicative noise function with mean 1, representing
+stochastic growth. We will consider log-normal noise with shape
+parameter $\sigma_g$. We start with an example in which the parameters
+are $r =2$, $K=8$, $C=5$ and $\sigma_g = 0.1$.
+
+As a low-dimensional system completely described by three parameters,
+this scenario should if anything be favorable to a parametric-based
+approach. This model contains an Allee effect, or tipping point, below
+which the population is not self-sustaining and shrinks to zero
+\citep{Courchamp2008}.
+
+\paragraph{Simulated training data}
+
+We generate initial observational data under the model described in Eq 1
+for $T_{\textrm{obs}}=40$ time steps, under a given arbitrary sequence
+of harvest intensities, $h_t$. We consider the case in which most of the
+data comes from a limited region of state space (e.g.~near a stable
+equilibrium), leaving us without observations of the population dynamics
+at very low levels which would be useful in discrimating between
+recruitment curves {[}@{]} or demonstrating the existence of a tipping
+point \citep{Scheffer2001}.\\Using data simulated from a specified model
+rather than empirical data permits the comparison against the true
+underlying dynamics, setting a bar for the optimal performance possible.
+
+\subsubsection{Parametric Models}
+
+We consider three candidate parametric models for the stock-recruitment
+function, which we refer to by the first authors of the publications in
+which they were first proposed.
+
+We generate the data with a four-parameter model that contains a tipping
+point, as discussed above (equation 1), (an Allee effect, see {[}@Allen,
+@Courchamp{]}) below which the stock decreases to zero,
+
+\[ X_{t+1} = Z_t S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(\frac{S_t - \theta}{K}\right)} \]
+
+\[ S_t = X_t - h_t \]
+
+The parameter $C$ reflects the location of the tipping point, $K$ the
+carrying capacity of the stock, and $r$ the base recruitment rate. $S_t$
+represents the stock size after a harvest $h_t$ has been implemented.
+$Z_t$ represents a log-normal random variable of log-mean zero and
+log-standard deviation parameter $\sigma$.
+
+We consider two alternative candidate models: the Ricker {[}@Ricker{]}
+stock-recruitment curve,
+
+\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
+
+and an alternative four-parameter model adapted from @Myers,
+
+\[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \]
+
+which contains a tipping point for $\theta > 2$ and becomes a
+Beverton-Holt model at $\theta = 1$.
+
+\subsubsection{Bayesian Inference of Parametric models}
+
+Given the sample data, we infer posterior distributions for each of the
+three models listed above using a Markov Chain Monte Carlo Gibbs Sampler
+(jags, see appendix for implementation details and code) given uniform
+priors. We run six chains for $10^6$ steps each and then assess
+convergence by Gelman-Rubin criterion and inspection of the traces, see
+appendix.
+
+\subsubsection{The Non-parametric Bayesian alternative for
+stock-recruitment curves}
+
+\subsubsection{SDP via GP}
+
+Once the posterior Gaussian process (GP) has been estimated
+\citep[e.g.~see][]{Munch2005}, it is necessary to adapt it in place of
+the parametric equation for the stochastic dynamic programming (SDP)
+solution \citep[see][ for a detailed description of parametric SDP
+methods]{Mangel1988} to the optimal policy. The essence of the idea is
+straight forward -- we will use the estimated GP in place of the
+parametric growth function to determine the stochastic transition matrix
+on which the SDP calculations are based. The SDP is solved in a
+discretized state space -- both the continuously valued population
+densities $X$ and harvest quotas $h$ are first mapped to a bounded,
+discrete grid. (For simplicity we will consider a uniform grid, though
+for either parametric or GP-based SDP it is often advantageous to use a
+non-uniform discretization such as a basis function representation,
+e.g.~see \citep{Deisenroth2009}).
+
+The SDP approach then computes a transition matrix, $\mathbf{F}$. We
+demonstrate that calculation is just as straight forward based on the GP
+as it is in the classical context using the parametric model. The
+${i,j}$ of the transition matrix $F$ entry gives the probability of
+transitioning into state $x_i$ given that the system is in state $x_j$
+in the previous time-step. To generate the transition matrix based on
+the posterior GP, we need only the expected values at each grid point
+and the corresponding variances (the diagonal of the covariance matrix),
+as shown in Figure 1. Given the mean of the GP posterior at each
+grid-point as the vector $E$ and variance at that point as vector $V$,
+the probability of transitioning from state $x_i$ to state $x_j$ is
+
+\[\mathcal{N}\left(x_j \vert  \mu = E_i, \sigma = \sqrt{V_i}\right)\]
+
+where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
+variance $\sigma^2$. Strictly speaking, the transition probability
+should be calculated by integrating the normal density over the bin of
+width $\Delta$ centered at $x_j$. For a sufficiently fine grid that
+$f(x_j) \approx f(x_j + \Delta)$, it is sufficient to calculate the
+density at $x_j$ and then row-normalize the transition matrix. The
+process can then be repeated for each possible discrete value of our
+control variable, (harvest $h$).
+
+\textbf{Pseudocode for the determining the transition matrix from the
+GP}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{for(h in h_grid)}
+  \NormalTok{F_h = for(x_j in grid)}
+          \NormalTok{for(i in }\DecValTok{1}\NormalTok{:N) }
+            \KeywordTok{dnorm}\NormalTok{(x_j, mu[i]-h, V[i])}
+\end{Highlighting}
+\end{Shaded}
+
+Using the discrete transition matrix we may write down the Bellman
+recursion defining the stochastic dynamic programming iteration:
+
+\begin{equation}
+V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
+\end{equation}
+
+where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$ is
+control (harvest level) chosen. Numerically, the maximization is
+accomplished as follows. Consider the set of possible control values to
+be the discrete values corresponding the grid of stock sizes. Then for
+each $h_t$ there is a corresponding transition matrix $\mathbf{F}_h$
+determined as described above but with mean $\mu = x_j - h_t$. Let
+$\vec{V_t}$ be the vector whose $i$th element corresponds to the value
+of having stock $x_i$ at time $t$. Then let $\Pi_h$ be the vector whose
+$i$th element indicates the profit from harvesting at intensity $h_t$
+given a population $x_i$ (e.g. $\max(x_i, h_t)$ since one cannot harvest
+more fish then the current population size). Then the Bellman recursion
+can be given in matrix form as
+
+\[V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)\]
+
+where the sum is element by element and the expectation is computed by
+the matrix multiplication $\mathbf{F} V_{t+1}$.
+
+\subsubsection{Pseudocode for the Bellman iteration}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+ \NormalTok{V1 <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
+      \NormalTok{delta * F[[h]] %*% V +  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
+    \NormalTok{\})}
+    \CommentTok{# find havest, h that gives the maximum value}
+    \NormalTok{out <- }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
+      \NormalTok{value <- }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
+      \NormalTok{index <- }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
+      \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
+    \NormalTok{\})}
+    \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
+    \NormalTok{V <- out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
+    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <- out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
+\end{Highlighting}
+\end{Shaded}
+
+This completes the algorithm adapting the GP to the sequential
+decision-making problem through SDP, which has not been previously
+demonstrated.\\We further provide an R package implementation as
+described in the supplemental materials.
+
+\subsubsection{Estimating parametric models}
+
+We estimate posterior distributions for two parametric models: one using
+the structurally correct model as given in Eq (1), which we refer to as
+the ``Parametric Bayes'' model, and another using the familiar Ricker
+model, using a Gibbs sampler as described (with source code) in the
+appendix). In addition we estimate the parameters of the structurally
+correct model by maximum likelihood.
+
+\section{Results}
+
+\section{Discussion}
+
+\paragraph{Big picture: Linking GP to SDP}
+
+\emph{rambling}
+
+Non-parametric Bayesian methods have received far too little attention
+in ecological modeling efforts that are aimed at improved conservation
+planning and decision making support. Such approaches may be
+particularly useful when the available data is restricted to a limited
+area of state-space, which can lead parametric models to underestimate
+the uncertainty in dynamics at population levels (states) which have not
+been observed. One reason for the relative absence of nonparametric
+approaches in the natural resource management context may be the lack of
+existing approaches for adapting the non-parametric Bayesian models
+previously proposed \citep{Munch2005} to a decision-theoretic framework.
+Adapting a non-parametric approach requires modification of existing
+methods for decision theory. We have illustrated how this might be done
+for a classic stochastic dynamic programming problem, opening the door
+for substantial further research into how these applications might be
+improved.
+
+\begin{center}\rule{3in}{0.4pt}\end{center}
+
+Code to replicate the analysis, along with complete log of this research
+can be found at:
 \href{https://github.com/cboettig/nonparametric-bayes/}{https://github.com/cboettig/nonparametric-bayes}
@@
 
-\section{References}
-
-Allen, Linda J. S., Jesse F. Fagan, Göran Högnäs, and Henrik Fagerholm.
-2005. ``Population extinction in discrete-time stochastic population
-models with an Allee effect.'' \emph{Journal of Difference Equations and
-Applications} 11 (4-5) (apr): 273--293.
-doi:10.1080/10236190412331335373.
-\url{http://www.tandfonline.com/doi/abs/10.1080/10236190412331335373}.
-
-Athanassoglou, Stergios, and Anastasios Xepapadeas. 2012. ``Pollution
-control with uncertain stock dynamics: When, and how, to be
-precautious.'' \emph{Journal of Environmental Economics and Management}
-63 (3) (may): 304--320. doi:10.1016/j.jeem.2011.11.001.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069611001409}.
-
-Barnosky, Anthony D., Elizabeth a Hadly, Jordi Bascompte, Eric L.
-Berlow, James H. Brown, Mikael Fortelius, Wayne M. Getz, et al. 2012.
-``Approaching a state shift in Earth's biosphere.'' \emph{Nature} 486
-(7401) (jun): 52--58. doi:10.1038/nature11018.
-\url{http://www.nature.com/doifinder/10.1038/nature11018}.
-
-Bestelmeyer, Brandon T., Michael C. Duniway, Darren K. James, Laura M.
-Burkett, and Kris M. Havstad. 2012. ``A test of critical thresholds and
-their indicators in a desertification-prone ecosystem: more resilience
-than we thought.'' Ed. Katharine Suding. \emph{Ecology Letters} (dec).
-doi:10.1111/ele.12045. \url{http://doi.wiley.com/10.1111/ele.12045}.
-
-Boettiger, Carl, and Alan Hastings. 2013. ``Tipping points: From
-patterns to predictions.'' \emph{Nature} 493 (7431) (jan): 157--158.
-doi:10.1038/493157a.
-\url{http://www.nature.com/doifinder/10.1038/493157a}.
-
-Brozović, Nicholas, and Wolfram Schlenker. 2011. ``Optimal management of
-an ecosystem with an unknown threshold.'' \emph{Ecological Economics}
-(jan): 1--14. doi:10.1016/j.ecolecon.2010.10.001.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0921800910004167}.
-
-Courchamp, Franck, Ludek Berec, and Joanna Gascoigne. 2008. \emph{Allee
-Effects in Ecology and Conservation}. Oxford University Press, USA.
-\url{http://www.amazon.com/Effects-Ecology-Conservation-Franck-Courchamp/dp/0198570309}.
-
-Cuddington, Kim M., M. J. Fortin, and L. R. Gerber. 2013.
-``Process-based models are required to manage ecological systems in a
-changing world.'' \emph{Ecosphere} 4 (February): 1--12.
-doi:10.1890/ES12-00178.1.
-\url{http://www.esajournals.org/doi/abs/10.1890/ES12-00178.1}.
-
-Deisenroth, Marc Peter, Carl Edward Rasmussen, and Jan Peters. 2009.
-``Gaussian process dynamic programming.'' \emph{Neurocomputing} 72 (7-9)
-(mar): 1508--1524. doi:10.1016/j.neucom.2008.12.019.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0925231209000162}.
-
-Geritz, Stefan A. H., and Eva Kisdi. 2012. ``Mathematical ecology: why
-mechanistic models?.'' \emph{Journal of mathematical biology} 65 (6-7)
-(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.
-
-Gordon, H. S. 1954. ``The economic theory of a common-property resource:
-the fishery.'' \emph{The Journal of Political Economy} 62 (2): 124--142.
-\url{http://www.jstor.org/stable/10.2307/1825571}.
-
-Halpern, Benjamin S., Carissa J. Klein, Christopher J. Brown, Maria
-Beger, Hedley S. Grantham, Sangeeta Mangubhai, Mary Ruckelshaus, et al.
-2013. ``Achieving the triple bottom line in the face of inherent
-trade-offs among social equity, economic return, and conservation.''
-\emph{Proceedings of the National Academy of Sciences of the United
-States of America} (mar). doi:10.1073/pnas.1217689110.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/23530207}.
-
-Hilborn, Ray. 2007. ``Reinterpreting the State of Fisheries and their
-Management.'' \emph{Ecosystems} 10 (8) (oct): 1362--1369.
-doi:10.1007/s10021-007-9100-5.
-\url{http://www.springerlink.com/index/10.1007/s10021-007-9100-5}.
-
-Hughes, Terry P., Cristina Linares, Vasilis Dakos, Ingrid a van de
-Leemput, and Egbert H. van Nes. 2013. ``Living dangerously on borrowed
-time during slow, unrecognized regime shifts.'' \emph{Trends in ecology
-\& evolution} 28 (3) (mar): 149--55. doi:10.1016/j.tree.2012.08.022.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/22995893}.
-
-Kocijan, Juš, Agathe Girard, Bla\textbackslash{}vz Banko, and Roderick
-Murray-Smith. 2005. ``Dynamic systems identification with Gaussian
-processes.'' \emph{Mathematical and Computer Modelling of Dynamical
-Systems} 11 (4) (dec): 411--424. doi:10.1080/13873950500068567.
-\url{http://www.tandfonline.com/doi/abs/10.1080/13873950500068567}.
-
-Ludwig, Donald, and Carl J. Walters. 1982. ``Optimal harvesting with
-imprecise parameter estimates.'' \emph{Ecological Modelling} 14 (3-4)
-(jan): 273--292. doi:10.1016/0304-3800(82)90023-0.
-\url{http://linkinghub.elsevier.com/retrieve/pii/0304380082900230}.
-
-Mangel, Marc. 1985. ``Decision and control in uncertain resource
-systems:'' 255.
-\href{http://www.amazon.com/Decision-uncertain-resource-Mathematics-Engineering/dp/0124687202 http://dl.acm.org/citation.cfm?id=537497}{http://www.amazon.com/Decision-uncertain-resource-Mathematics-Engineering/dp/0124687202
-http://dl.acm.org/citation.cfm?id=537497}.
-
-Mangel, Marc, and Colin W. Clark. 1988. \emph{Dynamic Modeling in
-Behavioral Ecology}. Ed. John Krebs and Tim Clutton-Brock. Princeton:
-Princeton University Press.
-
-May, Robert M., John R. Beddington, Colin W. Clark, Sidney J. Holt, and
-R. M. Laws. 1979. ``Management of multispecies fisheries.''
-\emph{Science (New York, N.Y.)} 205 (4403) (jul): 267--77.
-doi:10.1126/science.205.4403.267.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/17747032}.
-
-Munch, Stephan B., Melissa L. Snover, George M. Watters, and Marc
-Mangel. 2005. ``A unified treatment of top-down and bottom-up control of
-reproduction in populations.'' \emph{Ecology Letters} 8 (7) (may):
-691--695. doi:10.1111/j.1461-0248.2005.00766.x.
-\url{http://doi.wiley.com/10.1111/j.1461-0248.2005.00766.x}.
-
-Polasky, Stephen, Stephen R. Carpenter, Carl Folke, and Bonnie Keeler.
-2011. ``Decision-making under great uncertainty: environmental
-management in an era of global change.'' \emph{Trends in ecology \&
-evolution} (may): 1--7. doi:10.1016/j.tree.2011.04.007.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/21616553}.
-
-Rasmussen, Carl Edward, and C. K. I. Williams. 2006. \emph{Gaussian
-Processes for Machine Learning}. Ed. Thomas Dietterich. Boston: MIT
-Press,. \url{www.GaussianProcess.org/gpml}.
-
-Reed, William J. 1979. ``Optimal escapement levels in stochastic and
-deterministic harvesting models.'' \emph{Journal of Environmental
-Economics and Management} 6 (4) (dec): 350--363.
-doi:10.1016/0095-0696(79)90014-7.
-\href{http://www.sciencedirect.com/science/article/pii/0095069679900147 http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}{http://www.sciencedirect.com/science/article/pii/0095069679900147
-http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}.
-
-Schapaugh, Adam W., and Andrew J. Tyre. 2013. ``Accounting for
-parametric uncertainty in Markov decision processes.'' \emph{Ecological
-Modelling} 254 (apr): 15--21. doi:10.1016/j.ecolmodel.2013.01.003.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0304380013000306}.
-
-Scheffer, Marten, Jordi Bascompte, William A. Brock, Victor Brovkin,
-Stephen R. Carpenter, Vasilis Dakos, Hermann Held, Egbert H. van Nes,
-Max Rietkerk, and George Sugihara. 2009. ``Early-warning signals for
-critical transitions.'' \emph{Nature} 461 (7260): 53--9.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/19727193}.
-
-Scheffer, Marten, Stephen R. Carpenter, J. A. Foley, C. Folke, and B.
-Walker. 2001. ``Catastrophic shifts in ecosystems.'' \emph{Nature} 413
-(6856) (oct): 591--6. doi:10.1038/35098000.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/11595939}.
-
-Williams, Byron K. 2001. ``Uncertainty , learning , and the optimal
-management of wildlife.'' \emph{Environmental and Ecological Statistics}
-8: 269--288. doi:10.1023/A:1011395725123.
-
-Worm, Boris, Edward B. Barbier, Nicola Beaumont, J. Emmett Duffy, Carl
-Folke, Benjamin S. Halpern, Jeremy B. C. Jackson, et al. 2006. ``Impacts
-of biodiversity loss on ocean ecosystem services.'' \emph{Science (New
-York, N.Y.)} 314 (5800) (nov): 787--90. doi:10.1126/science.1132294.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/17082450}.
-
-Worm, Boris, Ray Hilborn, Julia K. Baum, Trevor A. Branch, Jeremy S.
-Collie, Christopher Costello, Michael J. Fogarty, et al. 2009.
-``Rebuilding global fisheries.'' \emph{Science (New York, N.Y.)} 325
-(5940) (jul): 578--85. doi:10.1126/science.1173146.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/19644114}.
+\bibliography{nonparametric-bayes}
 

### 2013-10-04 inst/doc/nonparametric-bayes.tex
compile draft

@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\usepackage{natbib}
-\bibliographystyle{plainnat}
 \usepackage{color}
 \usepackage{fancyvrb}
-\DefineShortVerb[commandchars=\\\{\}]{\|}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
 \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
@@
 
-\section{Abstract}
+\section{Abstract}\label{abstract}
 
@@
 
-\section{Introduction}
+\section{Introduction}\label{introduction}
 
@@
 sequence of actions in face of uncertainty, but only when that
-uncertainty can be meaningfully quantified {[}@Fischer2009{]}.
+uncertainty can be meaningfully quantified ({\textbf{???}}).
 
@@
 stochasticity in biological processes, measurements, and implementation
-of policy
-\citetext{\citealp[\emph{e.g.}][]{Reed1979}; \citealp{Clark1986}; \citealp[@Roughgarden1997;][]{Sethi2005}},
-parameteric uncertainty
-\citetext{\citealp[\emph{e.g.}][]{Ludwig1982}; \citealp[@Hilborn1997;
-@McAllister1998;][]{Schapaugh2013}}, and model or structural uncertainty
-\citep[\emph{e.g.}][]{Williams2001, Cressie2009, Athanassoglou2012}. Of
-these, structural uncertainty incorporates the least a priori knowledge
-or assumptions and is generally the hardest to quantify. Typical
-approaches assume a weak notion of uncertainty where a correct or
-reasonable approximation of the dynamics must be identified from among a
-handful of alternative models. Here we consider an approach that
+of policy (\emph{e.g.} Reed 1979; C. W. Clark and Kirkwood 1986;
+{\textbf{???}}; Sethi et al. 2005), parameteric uncertainty (\emph{e.g.}
+Ludwig and Walters 1982; {\textbf{???}}; {\textbf{???}}; Schapaugh and
+Tyre 2013), and model or structural uncertainty (\emph{e.g.} B. K.
+Williams 2001; Cressie et al. 2009; Athanassoglou and Xepapadeas 2012).
+Of these, structural uncertainty incorporates the least a priori
+knowledge or assumptions and is generally the hardest to quantify.
+Typical approaches assume a weak notion of uncertainty where a correct
+or reasonable approximation of the dynamics must be identified from
+among a handful of alternative models. Here we consider an approach that
 addresses uncertainty at each of these levels without assuming the
@@
   strategies follow from ``Dual Control'' or ``Active Adaptive
-  Management'' approaches (e.g. \citet{Ludwig1982}) that can trade off
-  short term utility by choosing actions that can reduce uncertainty.
-  Adaptive probing strategies arise when it is valuable to intentionally
-  force a system far from the observed values even when the expected
-  value such actions is low, as it provides much faster learning and
-  consequent reduction of model uncertainty that can allow greater value
-  to be derived later on. For instance, \citet{Ludwig1982} show that it
-  may be advantageous to fish an unexploited population very heavily at
-  first to obtain a better estimate of the recruitment rate. This
-  intuitive strategy when a population is governed by a Ricker or
-  Beverton-Holt-like dynamic would clearly be disastrous if instead the
-  dynamics contained an unforeseen tipping point. The best way to learn
-  where the edge lies may be to walk up to it, but it is also the most
-  dangerous.} arises when applying a dynamical model outside the range
-of data on which it has been estimated. This extrapolation uncertainty
-is felt most keenly in decision-theoretic applications, as (a) exploring
-the potential action space typically involves considering actions that
-may move the system outside the range of observed behavior, and (b)
-decision-theoretic alogrithms rely not only on reasonable estimates of
-the expected outcomes, but depend on the weights given to all possible
-outcomes \citep[\emph{e.g.}][]{Weitzman2013}.
+  Management'' approaches (e.g. Ludwig and Walters (1982)) that can
+  trade off short term utility by choosing actions that can reduce
+  uncertainty. Adaptive probing strategies arise when it is valuable to
+  intentionally force a system far from the observed values even when
+  the expected value such actions is low, as it provides much faster
+  learning and consequent reduction of model uncertainty that can allow
+  greater value to be derived later on. For instance, Ludwig and Walters
+  (1982) show that it may be advantageous to fish an unexploited
+  population very heavily at first to obtain a better estimate of the
+  recruitment rate. This intuitive strategy when a population is
+  governed by a Ricker or Beverton-Holt-like dynamic would clearly be
+  disastrous if instead the dynamics contained an unforeseen tipping
+  point. The best way to learn where the edge lies may be to walk up to
+  it, but it is also the most dangerous.} arises when applying a
+dynamical model outside the range of data on which it has been
+estimated. This extrapolation uncertainty is felt most keenly in
+decision-theoretic applications, as (a) exploring the potential action
+space typically involves considering actions that may move the system
+outside the range of observed behavior, and (b) decision-theoretic
+alogrithms rely not only on reasonable estimates of the expected
+outcomes, but depend on the weights given to all possible outcomes
+(\emph{e.g.} Weitzman 2013).
 
 Concerns over the potential for tipping points in ecological dynamics
-\citep{Scheffer2001} highlight the dangers of uncertainty in ecological
-management and pose a substantial challenge to existing
-decision-theoretic approaches \citep{Brozovic2011}. Because intervention
-is often too late after a tipping point has been crossed (but see
-\citet{Hughes2013}), management is most often concerned with avoiding
-potentially catastrophic tipping points before any data is available at
-or following a transition that would more clearly reveal these regime
-shift dynamics \citep[e.g.][]{Bestelmeyer2012}.
+(Scheffer et al. 2001) highlight the dangers of uncertainty in
+ecological management and pose a substantial challenge to existing
+decision-theoretic approaches (Brozović and Schlenker 2011). Because
+intervention is often too late after a tipping point has been crossed
+(but see Hughes et al. (2013)), management is most often concerned with
+avoiding potentially catastrophic tipping points before any data is
+available at or following a transition that would more clearly reveal
+these regime shift dynamics (e.g. Bestelmeyer et al. 2012).
 
@@
 Here we illustrate how a stochastic dynamic programming (SDP) algorithm
-{[}\citet{Mangel1988}; @Possingham1997, Marescot2013{]} can be driven by
-the predictions from a Bayesian non-parametric (BNP) approach
-\citep{Munch2005a}.\\This provides two distinct advantages compared with
-contemporary approaches. First, using a BNP sidesteps the need for an
-accurate model-based description of the system dynamics.\\Second, the
-BNP can better reflect uncertainty that arises when extrapolating a
-model outside of the data on which it was fit.\\We illustrate that when
-the correct model is not known, this latter feature is crucial to
-providing a robust decision-theoretic approach in face of substantial
-structural uncertainty.
-
-\subsection{Two central challenges:}
+(Mangel and Clark 1988; {\textbf{???}}) can be driven by the predictions
+from a Bayesian non-parametric (BNP) approach (Munch, Kottas, and Mangel
+2005).\\This provides two distinct advantages compared with contemporary
+approaches. First, using a BNP sidesteps the need for an accurate
+model-based description of the system dynamics.\\Second, the BNP can
+better reflect uncertainty that arises when extrapolating a model
+outside of the data on which it was fit.\\We illustrate that when the
+correct model is not known, this latter feature is crucial to providing
+a robust decision-theoretic approach in face of substantial structural
+uncertainty.
+
+\subsection{Two central challenges:}\label{two-central-challenges}
 
@@
 
-\subsection{Why we don't know the model}
-
-\begin{itemize}
-\item
-  Complex dynamics \citep{Glaser2013}(http://doi.org/10.1111/faf.12037
+\subsection{Why we don't know the
+model}\label{why-we-dont-know-the-model}
+
+\begin{itemize}
+\item
+  Complex dynamics (Glaser et al. 2013)(http://doi.org/10.1111/faf.12037
   ``Complex dynamics may limit prediction in marine fisheries'')
@@
 
-\subsection{Why we don't have data where we need it}
+\subsection{Why we don't have data where we need
+it}\label{why-we-dont-have-data-where-we-need-it}
 
@@
 The economic value and ecological concern have made marine fisheries the
-crucible for much of the founding work
-\citep{Gordon1954, Reed1979, May1979, Ludwig1982} in managing ecosystems
-under uncertainty. Global trends \citep{Worm2006} and controversy
-\citep{Hilborn2007, Worm2009} have made understanding these challenges
-all the more pressing.
+crucible for much of the founding work (Gordon 1954; Reed 1979; May et
+al. 1979; Ludwig and Walters 1982) in managing ecosystems under
+uncertainty. Global trends (Worm et al. 2006) and controversy (Hilborn
+2007; Worm et al. 2009) have made understanding these challenges all the
+more pressing.
 
@@
 social value of fish left in the sea vs.~the commercial value of fish
-harvested; see \citet{Halpern2013} for further discussion of this issue.
-Alternatively, we can always compare model performance across multiple
-potential objective functions.) The decision-maker does not necessarily
-need a model that provides the best mechanistic understanding or the
-best long-term outcome, but rather the one that best estimates the
-probabilities of being in different states as a result of the possible
-actions.
-
-\subsection{Background on the Gaussian Process}
+harvested; see Halpern et al. (2013) for further discussion of this
+issue. Alternatively, we can always compare model performance across
+multiple potential objective functions.) The decision-maker does not
+necessarily need a model that provides the best mechanistic
+understanding or the best long-term outcome, but rather the one that
+best estimates the probabilities of being in different states as a
+result of the possible actions.
+
+\subsection{Background on the Gaussian
+Process}\label{background-on-the-gaussian-process}
 
@@
 geospatial literature) to formulate a predictive model is relatively new
-in the context of modeling dynamical systems \citep{Kocijan2005}, and
+in the context of modeling dynamical systems (Kocijan et al. 2005), and
 was first introduced in the context ecological modeling and fisheries
-management in \citet{Munch2005}. An accessible and thorough introduction
-to the formulation and use of GPs can be found in \citet{Rasmussen2006}.
+management in Munch et al. (2005). An accessible and thorough
+introduction to the formulation and use of GPs can be found in Rasmussen
+and Williams (2006).
 
@@
 process model are estimated by Metropolis-Hastings algorithm, again with
-details and code provided in the Appendix. \citet{Rasmussen2006}
+details and code provided in the Appendix. Rasmussen and Williams (2006)
 provides an excellent general introduction to Gaussian Processes and
-\citet{Munch2005} first discusses their application in the context of
+Munch et al. (2005) first discusses their application in the context of
 population dynamics models such as fisheries stock-recruitment
@@
 
-\section{Approach and Methods}
-
-\subsubsection{Summary of approach}
-
-\subsubsection{Statement of the optimal control problem}
+\section{Approach and Methods}\label{approach-and-methods}
+
+\subsubsection{Summary of approach}\label{summary-of-approach}
+
+\subsubsection{Statement of the optimal control
+problem}\label{statement-of-the-optimal-control-problem}
 
@@
 
-\subsubsection{Parametric models}
+\subsubsection{Parametric models}\label{parametric-models}
 
@@
 We infer posterior distributions for the parameters of each model in a
-Bayesian context using Gibbs sampling (implemented in R {[}@RTeam{]}
-using jags, {[}@R2jags{]}). We choose uninformative uniform priors for
-all parameters (See Appendix, Figures S1-S3, and Table S1, and the R
+Bayesian context using Gibbs sampling (implemented in R ({\textbf{???}})
+using jags, ({\textbf{???}})). We choose uninformative uniform priors
+for all parameters (See Appendix, Figures S1-S3, and Table S1, and the R
 code provided). One-step-ahead predictions of these model fits are shown
@@
 
-\subsubsection{The Gaussian Process model}
+\subsubsection{The Gaussian Process
+model}\label{the-gaussian-process-model}
 
@@
 
-\section{Results}
+\section{Results}\label{results}
 
@@
 \KeywordTok{require}\NormalTok{(MASS)}
-\NormalTok{step_ahead <- function(x, f, p)\{}
-  \NormalTok{h = }\DecValTok{0}
-  \NormalTok{x_predict <- }\KeywordTok{sapply}\NormalTok{(x, f, h, p)}
-  \NormalTok{n <- }\KeywordTok{length}\NormalTok{(x_predict) - }\DecValTok{1}
-  \NormalTok{y <- }\KeywordTok{c}\NormalTok{(x[}\DecValTok{1}\NormalTok{], x_predict[}\DecValTok{1}\NormalTok{:n])}
+\NormalTok{step_ahead <-}\StringTok{ }\NormalTok{function(x, f, p)\{}
+  \NormalTok{h =}\StringTok{ }\DecValTok{0}
+  \NormalTok{x_predict <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(x, f, h, p)}
+  \NormalTok{n <-}\StringTok{ }\KeywordTok{length}\NormalTok{(x_predict) -}\StringTok{ }\DecValTok{1}
+  \NormalTok{y <-}\StringTok{ }\KeywordTok{c}\NormalTok{(x[}\DecValTok{1}\NormalTok{], x_predict[}\DecValTok{1}\NormalTok{:n])}
   \NormalTok{y}
 \NormalTok{\}}
-\NormalTok{step_ahead_posteriors <- function(x)\{}
-\NormalTok{gp_f_at_obs <- }\KeywordTok{gp_predict}\NormalTok{(gp, x, }\DataTypeTok{burnin=}\FloatTok{1e4}\NormalTok{, }\DataTypeTok{thin=}\DecValTok{300}\NormalTok{)}
-\NormalTok{df_post <- }\KeywordTok{melt}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(}\KeywordTok{sample}\NormalTok{(}\DecValTok{100}\NormalTok{), }
+\NormalTok{step_ahead_posteriors <-}\StringTok{ }\NormalTok{function(x)\{}
+\NormalTok{gp_f_at_obs <-}\StringTok{ }\KeywordTok{gp_predict}\NormalTok{(gp, x, }\DataTypeTok{burnin=}\FloatTok{1e4}\NormalTok{, }\DataTypeTok{thin=}\DecValTok{300}\NormalTok{)}
+\NormalTok{df_post <-}\StringTok{ }\KeywordTok{melt}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(}\KeywordTok{sample}\NormalTok{(}\DecValTok{100}\NormalTok{), }
   \NormalTok{function(i)\{}
@@
 
-\NormalTok{df_post <- }\KeywordTok{step_ahead_posteriors}\NormalTok{(x)}
-
-\KeywordTok{ggplot}\NormalTok{(df_post) + }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, stock)) + }
-  \KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, value, }\DataTypeTok{col=}\NormalTok{variable, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(L1,variable)), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) + }
-  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{))) }
+\NormalTok{df_post <-}\StringTok{ }\KeywordTok{step_ahead_posteriors}\NormalTok{(x)}
+
+\KeywordTok{ggplot}\NormalTok{(df_post) +}\StringTok{ }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, stock)) +}\StringTok{ }
+\StringTok{  }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, value, }\DataTypeTok{col=}\NormalTok{variable, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(L1,variable)), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) +}\StringTok{ }
+\StringTok{  }\KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{))) }
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{policies <- }\KeywordTok{melt}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{stock=}\NormalTok{x_grid, }\KeywordTok{sapply}\NormalTok{(OPT, function(x) x_grid[x])), }\DataTypeTok{id=}\StringTok{"stock"}\NormalTok{)}
-\KeywordTok{names}\NormalTok{(policies) <- }\KeywordTok{c}\NormalTok{(}\StringTok{"stock"}\NormalTok{, }\StringTok{"method"}\NormalTok{, }\StringTok{"value"}\NormalTok{)}
-
-\KeywordTok{ggplot}\NormalTok{(policies, }\KeywordTok{aes}\NormalTok{(stock, stock - value, }\DataTypeTok{color=}\NormalTok{method)) +}
-  \KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{lwd=}\FloatTok{1.2}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) + }\KeywordTok{xlab}\NormalTok{(}\StringTok{"stock size"}\NormalTok{) + }\KeywordTok{ylab}\NormalTok{(}\StringTok{"escapement"}\NormalTok{)  +}
-  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
+\NormalTok{policies <-}\StringTok{ }\KeywordTok{melt}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{stock=}\NormalTok{x_grid, }\KeywordTok{sapply}\NormalTok{(OPT, function(x) x_grid[x])), }\DataTypeTok{id=}\StringTok{"stock"}\NormalTok{)}
+\KeywordTok{names}\NormalTok{(policies) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"stock"}\NormalTok{, }\StringTok{"method"}\NormalTok{, }\StringTok{"value"}\NormalTok{)}
+
+\KeywordTok{ggplot}\NormalTok{(policies, }\KeywordTok{aes}\NormalTok{(stock, stock -}\StringTok{ }\NormalTok{value, }\DataTypeTok{color=}\NormalTok{method)) +}
+\StringTok{  }\KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{lwd=}\FloatTok{1.2}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) +}\StringTok{ }\KeywordTok{xlab}\NormalTok{(}\StringTok{"stock size"}\NormalTok{) +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{"escapement"}\NormalTok{)  +}
+\StringTok{  }\KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
 \end{Highlighting}
@@
 $S_t$, as under models such as this a constant escapement policy is
-expected to be optimal \citep{Reed1979}.}
+expected to be optimal (Reed 1979).}
 \end{figure}
@@
 $S_t$. Under models such as this a constant escapement policy is
-expected to be optimal \citep{Reed1979}, whereby population levels below
-a certain size $S$ are unharvested, while above that size the harvest
+expected to be optimal (Reed 1979), whereby population levels below a
+certain size $S$ are unharvested, while above that size the harvest
 strategy aims to return the population to $S$, resulting in the
@@
 \begin{Highlighting}[]
-\KeywordTok{ggplot}\NormalTok{(sims_data) + }
-  \KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, fishstock, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(reps,method), }\DataTypeTok{color=}\NormalTok{method), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) +}
-  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{)))}
+\KeywordTok{ggplot}\NormalTok{(sims_data) +}\StringTok{ }
+\StringTok{  }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, fishstock, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(reps,method), }\DataTypeTok{color=}\NormalTok{method), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) +}
+\StringTok{  }\KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{)))}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{fig4v1 <- }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) + }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{variable)) + }
-  \KeywordTok{facet_wrap}\NormalTok{(~variable, }\DataTypeTok{scales =} \StringTok{"free_y"}\NormalTok{)  + }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{) +}
-  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{) + }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey) }\CommentTok{# density plots fail when delta fn}
-
-\NormalTok{fig4v2 <- }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) + }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{variable), }\DataTypeTok{binwidth=}\FloatTok{0.1}\NormalTok{) + }
-  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
-
-\NormalTok{fig4v3 <- }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value, }\DataTypeTok{fill=}\NormalTok{variable, }\DataTypeTok{color=}\NormalTok{variable)) + }\CommentTok{# density plots fail when delta fn}
-  \KeywordTok{stat_density}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{y=}\NormalTok{..density..), }\DataTypeTok{position=}\StringTok{"stack"}\NormalTok{, }\DataTypeTok{adjust=}\DecValTok{3}\NormalTok{, }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{9}\NormalTok{) + }
-  \KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)+ }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
+\NormalTok{fig4v1 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) +}\StringTok{ }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{variable)) +}\StringTok{ }
+\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(~variable, }\DataTypeTok{scales =} \StringTok{"free_y"}\NormalTok{)  +}\StringTok{ }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{) +}
+\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{) +}\StringTok{ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey) }\CommentTok{# density plots fail when delta fn}
+
+\NormalTok{fig4v2 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) +}\StringTok{ }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{variable), }\DataTypeTok{binwidth=}\FloatTok{0.1}\NormalTok{) +}\StringTok{ }
+\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+}\StringTok{ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
+
+\NormalTok{fig4v3 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value, }\DataTypeTok{fill=}\NormalTok{variable, }\DataTypeTok{color=}\NormalTok{variable)) +}\StringTok{ }\CommentTok{# density plots fail when delta fn}
+\StringTok{  }\KeywordTok{stat_density}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{y=}\NormalTok{..density..), }\DataTypeTok{position=}\StringTok{"stack"}\NormalTok{, }\DataTypeTok{adjust=}\DecValTok{3}\NormalTok{, }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{9}\NormalTok{) +}\StringTok{ }
+\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+}\StringTok{ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)+}\StringTok{ }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
 \NormalTok{fig4v2}
@@
 
-\subsubsection{Figure 1: Fitted Models}
+\subsubsection{Figure 1: Fitted Models}\label{figure-1-fitted-models}
 
@@
 
-\subsubsection{Figure 2: Inferred Policies}
+\subsubsection{Figure 2: Inferred
+Policies}\label{figure-2-inferred-policies}
 
@@
 
-\subsubsection{Figure 3: Simulated results}
+\subsubsection{Figure 3: Simulated
+results}\label{figure-3-simulated-results}
 
@@
 
-\subsubsection{Figure 4: Robustness}
+\subsubsection{Figure 4: Robustness}\label{figure-4-robustness}
 
@@
 
-\section{Discussion}
+\section{Discussion}\label{discussion}
 
@@
 potential to increase our basic understanding of ecological processes
-\citep{Cuddington2013, Geritz2012}, such models can be not only
-inaccurate but misleading when relied upon in a quantitative decision
-making framework.
+(Cuddington, Fortin, and Gerber 2013; Geritz and Kisdi 2012), such
+models can be not only inaccurate but misleading when relied upon in a
+quantitative decision making framework.
 
@@
 
-\subsubsection{1. Large uncertainty where the data is poor}
+\subsubsection{1. Large uncertainty where the data is
+poor}\label{large-uncertainty-where-the-data-is-poor}
 
@@
 
-\subsubsection{2. Predictive accuracy where data is good}
+\subsubsection{2. Predictive accuracy where data is
+good}\label{predictive-accuracy-where-data-is-good}
 
@@
 
-\subsection{Future directions}
-
-\subsubsection{Higher dimensions}
+\subsection{Future directions}\label{future-directions}
+
+\subsubsection{Higher dimensions}\label{higher-dimensions}
 
@@
 
-\subsubsection{Online learning}
+\subsubsection{Online learning}\label{online-learning}
 
@@
 
-\section{Acknowledgments}
+\section{Acknowledgments}\label{acknowledgments}
 
@@
 
-\section{Appendix}
-
-\subsection{Model definitions and estimation}
+\section{Appendix}\label{appendix}
+
+\subsection{Model definitions and
+estimation}\label{model-definitions-and-estimation}
 
@@
 
-\subsection{Optimal Control Problem}
+\subsection{Optimal Control Problem}\label{optimal-control-problem}
 
@@
 
-\subsubsection{Pseudocode for the Bellman iteration}
+\subsubsection{Pseudocode for the Bellman
+iteration}\label{pseudocode-for-the-bellman-iteration}
 
@@
 
-\subsubsection{Training data}
+\subsubsection{Training data}\label{training-data}
 
@@
 
-\section{Abstract}
+\section{Abstract}\label{abstract-1}
 
@@
 
-\section{Introduction}
-
-\paragraph{Opening}
-
-\paragraph{Models for decision-making under uncertainty}
+\section{Introduction}\label{introduction-1}
+
+\paragraph{Opening}\label{opening}
+
+\paragraph{Models for decision-making under
+uncertainty}\label{models-for-decision-making-under-uncertainty}
 
@@
 resource over time.\\Though much can be said on how to choose this value
-function appropriately (e.g.~see
-\citep{Halpern2013}(http://doi.org/10.1073/pnas.1217689110 ``Achieving
-the triple bottom line in the face of inherent trade-offs among social
-equity, economic return, and conservation.'')) we will assume this is
-given. (Nor is this approach necessarily constrained to maximizing the
+function appropriately (e.g.~see (Halpern et al.
+2013)(http://doi.org/10.1073/pnas.1217689110 ``Achieving the triple
+bottom line in the face of inherent trade-offs among social equity,
+economic return, and conservation.'')) we will assume this is given.
+(Nor is this approach necessarily constrained to maximizing the
 expectated value of such a function - the decision-theoretic framework
 can be adapted to alternatives such as minimizing the maximum cost or
-damage that might be incurred; see \citet{Polasky2011}).
+damage that might be incurred; see Polasky et al. (2011)).
 
@@
 
-\citep{Williams2001, Athanassoglou2012}.
+(B. K. Williams 2001; Athanassoglou and Xepapadeas 2012).
 
@@
 
-\section{Approach and Methods}
-
-\subsection{The optimal control problem in fisheries management}
+\section{Approach and Methods}\label{approach-and-methods-1}
+
+\subsection{The optimal control problem in fisheries
+management}\label{the-optimal-control-problem-in-fisheries-management}
 
@@
 discretized values of stock $X$ and policy $h$ using stochastic dynamic
-programming (SDP) \citep{Mangel1988}. Problems of this sort underpin
-much marine fisheries management today.
+programming (SDP) (Mangel and Clark 1988). Problems of this sort
+underpin much marine fisheries management today.
 
@@
 estimated and integrated over to determine the optimal policy under
-under uncertainty \citep{Mangel1988, Schapaugh2013}. Uncertainty in the
-model structure itself can only be addressed in this approach by
-hypothesizing alternative model structures, and then performing some
-model choice or model averaging \citep{Williams2001, Athanassoglou2012}.
-
-\subsection{Underlying Model}
+under uncertainty (Mangel and Clark 1988; Schapaugh and Tyre 2013).
+Uncertainty in the model structure itself can only be addressed in this
+approach by hypothesizing alternative model structures, and then
+performing some model choice or model averaging (B. K. Williams 2001;
+Athanassoglou and Xepapadeas 2012).
+
+\subsection{Underlying Model}\label{underlying-model}
 
@@
 parametric model.\\We will focus on a simple parametric model for a
-single species \citep[derived from fist principles by][]{Allen2005a} as
+single species (derived from fist principles by Allen et al. 2005) as
 our underlying ``reality''.
@@
 which the population is not self-sustaining and shrinks to zero
-\citep{Courchamp2008}.
-
-\paragraph{Simulated training data}
+(Courchamp, Berec, and Gascoigne 2008).
+
+\paragraph{Simulated training data}\label{simulated-training-data}
 
@@
 recruitment curves {[}@{]} or demonstrating the existence of a tipping
-point \citep{Scheffer2001}.\\Using data simulated from a specified model
-rather than empirical data permits the comparison against the true
+point (Scheffer et al. 2001).\\Using data simulated from a specified
+model rather than empirical data permits the comparison against the true
 underlying dynamics, setting a bar for the optimal performance possible.
 
-\subsubsection{Parametric Models}
+\subsubsection{Parametric Models}\label{parametric-models-1}
 
@@
 We generate the data with a four-parameter model that contains a tipping
-point, as discussed above (equation 1), (an Allee effect, see {[}@Allen,
-@Courchamp{]}) below which the stock decreases to zero,
+point, as discussed above (equation 1), (an Allee effect, see
+({\textbf{???}})) below which the stock decreases to zero,
 
@@
 
-We consider two alternative candidate models: the Ricker {[}@Ricker{]}
-stock-recruitment curve,
+We consider two alternative candidate models: the Ricker
+({\textbf{???}}) stock-recruitment curve,
 
@@
 
-and an alternative four-parameter model adapted from @Myers,
+and an alternative four-parameter model adapted from ({\textbf{???}}),
 
@@
 
-\subsubsection{Bayesian Inference of Parametric models}
+\subsubsection{Bayesian Inference of Parametric
+models}\label{bayesian-inference-of-parametric-models}
 
@@
 \subsubsection{The Non-parametric Bayesian alternative for
-stock-recruitment curves}
-
-\subsubsection{SDP via GP}
-
-Once the posterior Gaussian process (GP) has been estimated
-\citep[e.g.~see][]{Munch2005}, it is necessary to adapt it in place of
-the parametric equation for the stochastic dynamic programming (SDP)
-solution \citep[see][ for a detailed description of parametric SDP
-methods]{Mangel1988} to the optimal policy. The essence of the idea is
-straight forward -- we will use the estimated GP in place of the
+stock-recruitment
+curves}\label{the-non-parametric-bayesian-alternative-for-stock-recruitment-curves}
+
+\subsubsection{SDP via GP}\label{sdp-via-gp}
+
+Once the posterior Gaussian process (GP) has been estimated (e.g.~see
+Munch et al. 2005), it is necessary to adapt it in place of the
+parametric equation for the stochastic dynamic programming (SDP)
+solution (see Mangel and Clark 1988 for a detailed description of
+parametric SDP methods) to the optimal policy. The essence of the idea
+is straight forward -- we will use the estimated GP in place of the
 parametric growth function to determine the stochastic transition matrix
@@
 non-uniform discretization such as a basis function representation,
-e.g.~see \citep{Deisenroth2009}).
+e.g.~see (Deisenroth, Rasmussen, and Peters 2009)).
 
@@
 
-\[\mathcal{N}\left(x_j \vert  \mu = E_i, \sigma = \sqrt{V_i}\right)\]
+\[\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)\]
 
@@
 
-\subsubsection{Pseudocode for the Bellman iteration}
+\subsubsection{Pseudocode for the Bellman
+iteration}\label{pseudocode-for-the-bellman-iteration-1}
 
@@
 
-\subsubsection{Estimating parametric models}
+\subsubsection{Estimating parametric
+models}\label{estimating-parametric-models}
 
@@
 
-\section{Results}
-
-\section{Discussion}
-
-\paragraph{Big picture: Linking GP to SDP}
+\section{Results}\label{results-1}
+
+\section{Discussion}\label{discussion-1}
+
+\paragraph{Big picture: Linking GP to
+SDP}\label{big-picture-linking-gp-to-sdp}
 
@@
 existing approaches for adapting the non-parametric Bayesian models
-previously proposed \citep{Munch2005} to a decision-theoretic framework.
-Adapting a non-parametric approach requires modification of existing
-methods for decision theory. We have illustrated how this might be done
-for a classic stochastic dynamic programming problem, opening the door
-for substantial further research into how these applications might be
-improved.
+previously proposed (Munch et al. 2005) to a decision-theoretic
+framework. Adapting a non-parametric approach requires modification of
+existing methods for decision theory. We have illustrated how this might
+be done for a classic stochastic dynamic programming problem, opening
+the door for substantial further research into how these applications
+might be improved.
 
@@
 
-\subsection{Markov Chain Monte Carlo Analysis}
-
-\bibliography{nonparametric-bayes}
+\subsection{Markov Chain Monte Carlo
+Analysis}\label{markov-chain-monte-carlo-analysis}
+
+Allen, Linda J. S., Jesse F. Fagan, Göran Högnäs, and Henrik Fagerholm.
+2005. ``Population Extinction in Discrete-time Stochastic Population
+Models with an Allee Effect.'' \emph{Journal of Difference Equations and
+Applications} 11 (4-5) (apr): 273--293.
+doi:10.1080/10236190412331335373.
+\url{http://www.tandfonline.com/doi/abs/10.1080/10236190412331335373}.
+
+Athanassoglou, Stergios, and Anastasios Xepapadeas. 2012. ``Pollution
+Control with Uncertain Stock Dynamics: When, and How, to Be
+Precautious.'' \emph{Journal of Environmental Economics and Management}
+63 (3) (may): 304--320. doi:10.1016/j.jeem.2011.11.001.
+\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069611001409}.
+
+Bestelmeyer, Brandon T., Michael C. Duniway, Darren K. James, Laura M.
+Burkett, and Kris M. Havstad. 2012. ``A Test of Critical Thresholds and
+Their Indicators in a Desertification-prone Ecosystem: More Resilience
+Than We Thought.'' Edited by Katharine Suding. \emph{Ecology Letters}
+(dec). doi:10.1111/ele.12045.
+\url{http://doi.wiley.com/10.1111/ele.12045}.
+
+Brozović, Nicholas, and Wolfram Schlenker. 2011. ``Optimal Management of
+an Ecosystem with an Unknown Threshold.'' \emph{Ecological Economics}
+(jan): 1--14. doi:10.1016/j.ecolecon.2010.10.001.
+\url{http://linkinghub.elsevier.com/retrieve/pii/S0921800910004167}.
+
+Clark, Colin W., and Geoffrey P. Kirkwood. 1986. ``On Uncertain
+Renewable Resource Stocks: Optimal Harvest Policies and the Value of
+Stock Surveys.'' \emph{Journal of Environmental Economics and
+Management} 13 (3) (sep): 235--244. doi:10.1016/0095-0696(86)90024-0.
+\url{http://linkinghub.elsevier.com/retrieve/pii/0095069686900240}.
+
+Courchamp, Franck, Ludek Berec, and Joanna Gascoigne. 2008. \emph{Allee
+Effects in Ecology and Conservation}. Oxford University Press, USA.
+\url{http://www.amazon.com/Effects-Ecology-Conservation-Franck-Courchamp/dp/0198570309}.
+
+Cressie, Noel, Catherine a Calder, James S. Clark, Jay M. Ver Hoef, and
+Christopher K. Wikle. 2009. ``Accounting for Uncertainty in Ecological
+Analysis: the Strengths and Limitations of Hierarchical Statistical
+Modeling.'' \emph{Ecological Applications} 19 (3) (apr): 553--70.
+doi:10.1890/07-0744.1.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/19425416}.
+
+Cuddington, Kim M., M. J. Fortin, and L. R. Gerber. 2013.
+``Process-based Models Are Required to Manage Ecological Systems in a
+Changing World.'' \emph{Ecosphere} 4 (February): 1--12.
+doi:10.1890/ES12-00178.1.
+\url{http://www.esajournals.org/doi/abs/10.1890/ES12-00178.1}.
+
+Deisenroth, Marc Peter, Carl Edward Rasmussen, and Jan Peters. 2009.
+``Gaussian Process Dynamic Programming.'' \emph{Neurocomputing} 72 (7-9)
+(mar): 1508--1524. doi:10.1016/j.neucom.2008.12.019.
+\url{http://linkinghub.elsevier.com/retrieve/pii/S0925231209000162}.
+
+Geritz, Stefan A. H., and Eva Kisdi. 2012. ``Mathematical Ecology: Why
+Mechanistic Models?'' \emph{Journal of Mathematical Biology} 65 (6-7)
+(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.
+
+Glaser, Sarah M., Michael J. Fogarty, Hui Liu, Irit Altman, Chih-Hao
+Hsieh, Les Kaufman, Alec D. MacCall, Andrew a Rosenberg, Hao Ye, and
+George Sugihara. 2013. ``Complex Dynamics May Limit Prediction in Marine
+Fisheries.'' \emph{Fish and Fisheries} (apr). doi:10.1111/faf.12037.
+\url{http://doi.wiley.com/10.1111/faf.12037}.
+
+Gordon, H. S. 1954. ``The Economic Theory of a Common-property Resource:
+the Fishery.'' \emph{The Journal of Political Economy} 62 (2): 124--142.
+\url{http://www.jstor.org/stable/10.2307/1825571}.
+
+Halpern, Benjamin S., Carissa J. Klein, Christopher J. Brown, Maria
+Beger, Hedley S. Grantham, Sangeeta Mangubhai, Mary Ruckelshaus, et al.
+2013. ``Achieving the Triple Bottom Line in the Face of Inherent
+Trade-offs Among Social Equity, Economic Return, and Conservation.''
+\emph{Proceedings of the National Academy of Sciences of the United
+States of America} 110 (15) (apr): 6229--34.
+doi:10.1073/pnas.1217689110.
+\href{http://www.ncbi.nlm.nih.gov/pubmed/23530207 http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=3625307/\&tool=pmcentrez/\&rendertype=abstract}{http://www.ncbi.nlm.nih.gov/pubmed/23530207
+http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=3625307\textbackslash{}\&tool=pmcentrez\textbackslash{}\&rendertype=abstract}.
+
+Hilborn, Ray. 2007. ``Reinterpreting the State of Fisheries and Their
+Management.'' \emph{Ecosystems} 10 (8) (oct): 1362--1369.
+doi:10.1007/s10021-007-9100-5.
+\url{http://www.springerlink.com/index/10.1007/s10021-007-9100-5}.
+
+Hughes, Terry P., Cristina Linares, Vasilis Dakos, Ingrid a van de
+Leemput, and Egbert H. van Nes. 2013. ``Living Dangerously on Borrowed
+Time During Slow, Unrecognized Regime Shifts.'' \emph{Trends in Ecology
+\& Evolution} 28 (3) (mar): 149--55. doi:10.1016/j.tree.2012.08.022.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/22995893}.
+
+Kocijan, Juš, Agathe Girard, Bla\textbackslash{}vz Banko, and Roderick
+Murray-Smith. 2005. ``Dynamic Systems Identification with Gaussian
+Processes.'' \emph{Mathematical and Computer Modelling of Dynamical
+Systems} 11 (4) (dec): 411--424. doi:10.1080/13873950500068567.
+\url{http://www.tandfonline.com/doi/abs/10.1080/13873950500068567}.
+
+Ludwig, Donald, and Carl J. Walters. 1982. ``Optimal Harvesting with
+Imprecise Parameter Estimates.'' \emph{Ecological Modelling} 14 (3-4)
+(jan): 273--292. doi:10.1016/0304-3800(82)90023-0.
+\url{http://linkinghub.elsevier.com/retrieve/pii/0304380082900230}.
+
+Mangel, Marc, and Colin W. Clark. 1988. \emph{Dynamic Modeling in
+Behavioral Ecology}. Edited by John Krebs and Tim Clutton-Brock.
+Princeton: Princeton University Press.
+
+May, Robert M., John R. Beddington, Colin W. Clark, Sidney J. Holt, and
+R. M. Laws. 1979. ``Management of Multispecies Fisheries.''
+\emph{Science (New York, N.Y.)} 205 (4403) (jul): 267--77.
+doi:10.1126/science.205.4403.267.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/17747032}.
+
+Munch, Stephan B., Athanasios Kottas, and Marc Mangel. 2005. ``Bayesian
+Nonparametric Analysis of Stockrecruitment Relationships.''
+\emph{Canadian Journal of Fisheries and Aquatic Sciences} 62 (8) (aug):
+1808--1821. doi:10.1139/f05-073.
+\url{http://www.nrcresearchpress.com/doi/abs/10.1139/f05-073}.
+
+Munch, Stephan B., Melissa L. Snover, George M. Watters, and Marc
+Mangel. 2005. ``A Unified Treatment of Top-down and Bottom-up Control of
+Reproduction in Populations.'' \emph{Ecology Letters} 8 (7) (may):
+691--695. doi:10.1111/j.1461-0248.2005.00766.x.
+\url{http://doi.wiley.com/10.1111/j.1461-0248.2005.00766.x}.
+
+Polasky, Stephen, Stephen R. Carpenter, Carl Folke, and Bonnie Keeler.
+2011. ``Decision-making Under Great Uncertainty: Environmental
+Management in an Era of Global Change.'' \emph{Trends in Ecology \&
+Evolution} (may): 1--7. doi:10.1016/j.tree.2011.04.007.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/21616553}.
+
+Rasmussen, Carl Edward, and C. K. I. Williams. 2006. \emph{Gaussian
+Processes for Machine Learning}. Edited by Thomas Dietterich. Boston:
+MIT Press,. \url{www.GaussianProcess.org/gpml}.
+
+Reed, William J. 1979. ``Optimal Escapement Levels in Stochastic and
+Deterministic Harvesting Models.'' \emph{Journal of Environmental
+Economics and Management} 6 (4) (dec): 350--363.
+doi:10.1016/0095-0696(79)90014-7.
+\href{http://www.sciencedirect.com/science/article/pii/0095069679900147 http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}{http://www.sciencedirect.com/science/article/pii/0095069679900147
+http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}.
+
+Schapaugh, Adam W., and Andrew J. Tyre. 2013. ``Accounting for
+Parametric Uncertainty in Markov Decision Processes.'' \emph{Ecological
+Modelling} 254 (apr): 15--21. doi:10.1016/j.ecolmodel.2013.01.003.
+\url{http://linkinghub.elsevier.com/retrieve/pii/S0304380013000306}.
+
+Scheffer, Marten, Stephen R. Carpenter, J. A. Foley, C. Folke, and B.
+Walker. 2001. ``Catastrophic Shifts in Ecosystems.'' \emph{Nature} 413
+(6856) (oct): 591--6. doi:10.1038/35098000.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/11595939}.
+
+Sethi, Gautam, Christopher Costello, Anthony Fisher, Michael Hanemann,
+and Larry Karp. 2005. ``Fishery Management Under Multiple Uncertainty.''
+\emph{Journal of Environmental Economics and Management} 50 (2) (sep):
+300--318. doi:10.1016/j.jeem.2004.11.005.
+\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069605000057}.
+
+Weitzman, Martin L. 2013. ``A Precautionary Tale of Uncertain Tail
+Fattening.'' \emph{Environmental and Resource Economics} 55 (2) (mar):
+159--173. doi:10.1007/s10640-013-9646-y.
+\url{http://link.springer.com/10.1007/s10640-013-9646-y}.
+
+Williams, Byron K. 2001. ``Uncertainty , Learning , and the Optimal
+Management of Wildlife.'' \emph{Environmental and Ecological Statistics}
+8: 269--288. doi:10.1023/A:1011395725123.
+
+Worm, Boris, Edward B. Barbier, Nicola Beaumont, J. Emmett Duffy, Carl
+Folke, Benjamin S. Halpern, Jeremy B. C. Jackson, et al. 2006. ``Impacts
+of Biodiversity Loss on Ocean Ecosystem Services.'' \emph{Science (New
+York, N.Y.)} 314 (5800) (nov): 787--90. doi:10.1126/science.1132294.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/17082450}.
+
+Worm, Boris, Ray Hilborn, Julia K. Baum, Trevor A. Branch, Jeremy S.
+Collie, Christopher Costello, Michael J. Fogarty, et al. 2009.
+``Rebuilding Global Fisheries.'' \emph{Science (New York, N.Y.)} 325
+(5940) (jul): 578--85. doi:10.1126/science.1173146.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/19644114}.
 

### 2013-10-09 inst/doc/nonparametric-bayes.Rmd
compile draft
writing

@@
 <!-- Uncertainty outside the data without the correct model has not been handled. -->
+<!-- __We don't have the model__ -->
 
@@
 
+<!-- __We don't have the data where we need it__ -->
 <!-- What do we call this?  Extrapolation uncertainty?  Pathological Uncertainty? -->
@@
 outcomes [_e.g._ @Weitzman2013].
+
+<!-- 
+## Why we don't know the model
+- Complex dynamics  [@Glaser2013](http://doi.org/10.1111/faf.12037 "Complex dynamics may limit prediction in marine fisheries")
+- model choice and model averaging approaches
+-->
+
+<!--
+## Why we don't have data where we need it
+- Concerns of tipping points 
+- Danger of learning 
+-->
 
@@
 
-Here we illustrate how a stochastic dynamic programming (SDP) algorithm [@Mangel1988; @Possingham1997, Marescot2013] can 
-be driven by the predictions from a Bayesian non-parametric (BNP) approach [@Munch2005a].  
-This provides two distinct advantages compared with contemporary approaches.
-First, using a BNP sidesteps the need for an accurate model-based description
-of the system dynamics.  
-Second, the BNP can better reflect uncertainty that arises when extrapolating a model 
-outside of the data on which it was fit.  
-We illustrate that when the correct model is not known, this latter
-feature is crucial to providing a robust decision-theoretic approach in
-face of substantial structural uncertainty.
-
-<!-- More Consise versions above
-
+Here we illustrate how a stochastic dynamic programming (SDP) algorithm
+[@Mangel1988; @Possingham1997, Marescot2013] can be driven by the
+predictions from a Bayesian non-parametric (BNP) approach [@Munch2005a].
+This provides two distinct advantages compared with contemporary
+approaches.  First, using a BNP sidesteps the need for an accurate
+model-based description of the system dynamics.  Second, the BNP can
+better reflect uncertainty that arises when extrapolating a model outside
+of the data on which it was fit.  We illustrate that when the correct
+model is not known, this latter feature is crucial to providing a robust
+decision-theoretic approach in face of substantial structural uncertainty.
+
+<!-- More Consise versions are now above 
 - Parametric uncertainty
@@
 considered [@Ludwig1982, ].
-
 -->
 
-## Two central challenges:
-
-- We don't know the model
-- We don't have data from where we need it most
-
-## Why we don't know the model
-
-- Complex dynamics  [@Glaser2013](http://doi.org/10.1111/faf.12037 "Complex dynamics may limit prediction in marine fisheries")
-
-- model choice and model averaging approaches
-
-## Why we don't have data where we need it
-
-- Concerns of tipping points 
-- Danger of learning
-
-- _MAP_
-
-<!-- necessary? -->
-This paper represents the first time the SDP decision-making framework has been used
-without an a priori model of the underlying dynamics through the use of the BNP approach.
-
-In contrast to parametric models which can only reflect uncertainty 
-in parameter estimates, the BNP approach provides a more state-space dependent representation
-of uncertainty.
-
-This permits a much greater uncertainty far from the observed data than near the observed 
-data.  These features allow the GP-SDP approach to find robust management solutions 
-in face of limited data and without knowledge of the correct model structure.
+<!-- necessary? --> <!-- should reflect the dual problem of extrapolation and model uncertainty better -->
+
+This paper represents the first time the SDP decision-making framework has
+been used without an a priori model of the underlying dynamics through
+the use of the BNP approach.  In contrast to parametric models which
+can only reflect uncertainty in parameter estimates, the BNP approach
+provides a more state-space dependent representation of uncertainty.
+This permits a much greater uncertainty far from the observed data than
+near the observed data.  These features allow the GP-SDP approach to
+find robust management solutions in face of limited data and without
+knowledge of the correct model structure.
 

### 2013-10-10 inst/doc/nonparametric-bayes.Rmd
writing

@@
 opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=1, 
-               comment=NA, verbose=TRUE, fig.width=6, fig.height=4)
+               comment=NA, verbose=TRUE, fig.width=4, fig.height=3)
 
@@
 ```
-```{r deviances, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output"), results=toggle}
+
+```{r dic_calc, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output"), include=FALSE, echo=FALSE}
 ```
@@
 
-```{r Figure1, dependson=c("assemble-models"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
+### Figure 1: Fitted Models
+
+- All models fit the data quite well
+- Information criteria would pick the simple, incorrect model.
+
+```{r Figureb_posteriors, dependson=c("assemble-models", "par-fns"), include=TRUE, echo=FALSE, fig.cap="Points show the training data of stock-size over time.
+Curves show the posterior step-ahead predictions based on each of the estimated models.", fig.width=8, fig.height=6}
+```
+
+All models fit the observed data rather closely and with relatively small uncertainty, as illustrated in the posterior predictive curves in Figure 1.
+Figure 1 shows the training data of stock sizes observed over time as points, overlaid with the step-ahead predictions of each estimated model using the parameters sampled from their posterior distributions.
+Each model manages to fit the observed data rather closely.
+Compared to the expected value of the true model most estimates appear to overfit, predicting fluctuations that are actually due purely to stochasticity in growth rate.
+Model-choice criteria shown in Table 1 penalize more complex models and show a slight preference for the simpler Ricker model over the more complicated alternate stable state models (Allen and Myers).
+
+```{r deviances, dependson=c("dic_calc"), include=TRUE, results="asis", echo=FALSE}
+```
+
+```{r statespace_plot, dependson=c("assemble-models"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
 Graph shows the expected value for the function $f$ under each model.
@@
 The training data is also shown as black points.
-(The GP is conditioned on 0,0, shown as a pseudo-data point).
-", include=FALSE}
-```
-```{r Figureb, dependson=c("assemble-models", "par-fns"), include=TRUE}
-```
-
-Figure 1 shows the mean inferred state space dynamics of each model
-relative to the true model used to generate the data, predicting the
-relationship between observed stock size (x-axis) to the stock size
-after recruitment the following year.  All models except the MLE model
-estimate a distribution around the means shown here, and all models 
-estimate a level of process noise, which is independent of the state
-value (x).  Note that in contrast to the  other models shown, the mean
+The GP is conditioned on (0,0), shown as a pseudo-data point.", include=TRUE, echo=FALSE}
+```
+
+The mean inferred state space dynamics of each model
+relative to the true model used to generate the data is shown in Figure 2, 
+predicting the relationship between observed stock size (x-axis) to the stock size
+after recruitment the following year.  Note that in contrast to the  other models shown, the expected 
 Gaussian process corresponds to a distribution of curves - as indicated
-by the gray band - which itself has a mean shown in black.  Note that 
-this mean GP is thus more certain of the dynamics in the region where
-data is available then where it is not.
-
-While it would be straight forward to condition the GP on passing through the origin (0,0), the estimate
-shown here is based only on the observed data. The observed data from
-which each model is estimated is also shown.  The observations come
+by the gray band - which itself has a mean shown in black. Parameter uncertainty
+(not shown) spreads out the estimates further.  
+The observed data from which each model is estimated is also shown.  The observations come
 from only a limited region of state space corresponding to unharvested
@@
 
+### Figure 3: Inferred Policies
+
+- Inferred policies differ substantially among models
+- The structurally correct model and the GP are close to the true model
+- alternatives are not close
+
 ```{r Figure2, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", include=TRUE}
-```
-
-The resulting optimal management strategy based on each of the inferred
-models is shown in Figure 2, against the optimal strategy given the
-true underlying dynamics.  Policies are shown in terms of target
+Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", include=TRUE, echo=FALSE}
+```
+
+Despite the similarities in model fits to the observed data, the policies
+inferred under each model differ widely, as shown in Figure 3.  
+Policies are shown in terms of target
 escapement, $S_t$.  Under models such as this a constant escapement
@@
 strategy aims to return the population to $S$, resulting in the hockey-stick
-shaped policies shown.
+shaped policies shown.  Only the structurally correct model (Allen model) and the GP 
+produce policies close to the true optimum policy (where both the underlying 
+model structure and parameter values are known without error).
 
 ```{r Figure3, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.",  include=TRUE}
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.",  include=TRUE, echo=FALSE}
 ```
@@
 The consequences of managing 100 replicate realizations of the 
-simulated fishery under each of the policies estimated is shown in Figure 3.  As expected
+simulated fishery under each of the policies estimated is shown in Figure 4.  As expected
 from the policy curves, the structurally correct model under-harvests,
@@
 
-The results shown in Figures 1-3 are not unique to the simulated data or models chosen
-here, but arises across a range of parameter values and simulations as shown in the 
-supplemental figures.  The results across this range can most easily be compared 
-by the relative differences in net present value realized by each of the approaches,
-as shown in Figure 4.  The Gaussian Process most consistently realizes a value 
+These results are robust across a range of stochastic realizations, models, and parameter values.  
+The results across this range can most easily be compared 
+by using the relative differences in net present value realized by each of the model,
+as shown in Figure 5.  The BNP-SDP approach most consistently realizes a value 
 close to the optimal solution, and importantly avoids ever driving the system across
@@
 Values plotted are the averages of a given policy over 100 replicate simulations.
-Details and code provided in the supplement.",  include=TRUE}
-```
-
-### Figure 1: Fitted Models
-
-- All models fit the data quite well
-- Information criteria would pick the simple, incorrect model.
-
-### Figure 2: Inferred Policies
-
-- Inferred policies differ substantially among models
-- The structurally correct model and the GP are close to the true model
-- alternatives are not close
-
-### Figure 3: Simulated results
-
--
-
-### Figure 4: Robustness
-
-- Results hold across range of parameters
-
-(eek, distinguish better between result and discussion?)
+Details and code provided in the supplement.",  include=TRUE, echo=FALSE}
+
+```
 
@@
 - All models are "good fits" to the originally observed data. 
-
 - (Simple model choice immediately leads us astray)
+
+These results illustrate that model-choice approaches would be postively misleading -- supporting simpler models that cannot express tipping point dynamics merely on account of them being similar.
+As the data shown comes only from the basin of attraction near the unfished equilibrium, near which all of the models are approximately linear and approximately identical.
 
@@
 
-```{r, echo=FALSE, include=TRUE}
+```{r, echo=FALSE, include=TRUE, fig.height=6}
 plot_ricker_traces
@@
 
-```{r echo=FALSE, results="asis", include=TRUE}
+```{r echo=FALSE, results = "asis", include = TRUE}
 xtable::xtable(data.frame(
@@
 
-```{r, echo=FALSE, include=TRUE}
+```{r, echo=FALSE, include=TRUE, fig.height=6}
 plot_myers_traces
@@
 Table S2: Parameterization of the priors
-```{r echo=FALSE, results="asis", include=TRUE}
+```{r echo=FALSE, results="asis", include = TRUE}
 xtable::xtable(data.frame(parameter = c("r0", "K", "theta", "sigma"),
@@
 
-```{r, echo=FALSE, include=TRUE}
+```{r, echo=FALSE, include=TRUE, fig.height=6}
 plot_allen_traces
@@
 
-```{r echo=FALSE, results="asis", include=TRUE}
+```{r echo=FALSE, results = "asis", include=TRUE}
 xtable::xtable(data.frame(parameter = c("r0", "K", "theta", "sigma"),

### 2013-10-10 inst/doc/nonparametric-bayes.tex
writing

@@
 from a Bayesian non-parametric (BNP) approach (Munch, Kottas, and Mangel
-2005).\\This provides two distinct advantages compared with contemporary
+2005). This provides two distinct advantages compared with contemporary
 approaches. First, using a BNP sidesteps the need for an accurate
-model-based description of the system dynamics.\\Second, the BNP can
+model-based description of the system dynamics. Second, the BNP can
 better reflect uncertainty that arises when extrapolating a model
-outside of the data on which it was fit.\\We illustrate that when the
+outside of the data on which it was fit. We illustrate that when the
 correct model is not known, this latter feature is crucial to providing
@@
 
-\subsection{Two central challenges:}\label{two-central-challenges}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  We don't know the model
-\item
-  We don't have data from where we need it most
-\end{itemize}
-
-\subsection{Why we don't know the
-model}\label{why-we-dont-know-the-model}
-
-\begin{itemize}
-\item
-  Complex dynamics (Glaser et al. 2013)(http://doi.org/10.1111/faf.12037
-  ``Complex dynamics may limit prediction in marine fisheries'')
-\item
-  model choice and model averaging approaches
-\end{itemize}
-
-\subsection{Why we don't have data where we need
-it}\label{why-we-dont-have-data-where-we-need-it}
-
-\begin{itemize}
-\item
-  Concerns of tipping points
-\item
-  Danger of learning
-\item
-  \emph{MAP}
-\end{itemize}
-
 This paper represents the first time the SDP decision-making framework
 has been used without an a priori model of the underlying dynamics
-through the use of the BNP approach.
-
-In contrast to parametric models which can only reflect uncertainty in
-parameter estimates, the BNP approach provides a more state-space
-dependent representation of uncertainty.
-
-This permits a much greater uncertainty far from the observed data than
-near the observed data. These features allow the GP-SDP approach to find
-robust management solutions in face of limited data and without
-knowledge of the correct model structure.
+through the use of the BNP approach. In contrast to parametric models
+which can only reflect uncertainty in parameter estimates, the BNP
+approach provides a more state-space dependent representation of
+uncertainty. This permits a much greater uncertainty far from the
+observed data than near the observed data. These features allow the
+GP-SDP approach to find robust management solutions in face of limited
+data and without knowledge of the correct model structure.
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{require}\NormalTok{(MASS)}
-\NormalTok{step_ahead <-}\StringTok{ }\NormalTok{function(x, f, p)\{}
-  \NormalTok{h =}\StringTok{ }\DecValTok{0}
-  \NormalTok{x_predict <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(x, f, h, p)}
-  \NormalTok{n <-}\StringTok{ }\KeywordTok{length}\NormalTok{(x_predict) -}\StringTok{ }\DecValTok{1}
-  \NormalTok{y <-}\StringTok{ }\KeywordTok{c}\NormalTok{(x[}\DecValTok{1}\NormalTok{], x_predict[}\DecValTok{1}\NormalTok{:n])}
-  \NormalTok{y}
-\NormalTok{\}}
-\NormalTok{step_ahead_posteriors <-}\StringTok{ }\NormalTok{function(x)\{}
-\NormalTok{gp_f_at_obs <-}\StringTok{ }\KeywordTok{gp_predict}\NormalTok{(gp, x, }\DataTypeTok{burnin=}\FloatTok{1e4}\NormalTok{, }\DataTypeTok{thin=}\DecValTok{300}\NormalTok{)}
-\NormalTok{df_post <-}\StringTok{ }\KeywordTok{melt}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(}\KeywordTok{sample}\NormalTok{(}\DecValTok{100}\NormalTok{), }
-  \NormalTok{function(i)\{}
-    \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{time =} \DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(x), }\DataTypeTok{stock =} \NormalTok{x, }
-                \DataTypeTok{GP =} \KeywordTok{mvrnorm}\NormalTok{(}\DecValTok{1}\NormalTok{, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),}
-                \DataTypeTok{True =} \KeywordTok{step_ahead}\NormalTok{(x,f,p),  }
-                \DataTypeTok{MLE =} \KeywordTok{step_ahead}\NormalTok{(x,f,est$p), }
-                \DataTypeTok{Allen =} \KeywordTok{step_ahead}\NormalTok{(x, allen_f, pardist[i,]), }
-                \DataTypeTok{Ricker =} \KeywordTok{step_ahead}\NormalTok{(x, ricker_f, ricker_pardist[i,]), }
-                \DataTypeTok{Myers =} \KeywordTok{step_ahead}\NormalTok{(x, myers_f, myers_pardist[i,]))}
-  \NormalTok{\}), }\DataTypeTok{id=}\KeywordTok{c}\NormalTok{(}\StringTok{"time"}\NormalTok{, }\StringTok{"stock"}\NormalTok{))}
-\NormalTok{\}}
-
-\NormalTok{df_post <-}\StringTok{ }\KeywordTok{step_ahead_posteriors}\NormalTok{(x)}
-
-\KeywordTok{ggplot}\NormalTok{(df_post) +}\StringTok{ }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, stock)) +}\StringTok{ }
-\StringTok{  }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, value, }\DataTypeTok{col=}\NormalTok{variable, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(L1,variable)), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) +}\StringTok{ }
-\StringTok{  }\KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{))) }
-\end{Highlighting}
-\end{Shaded}
-
 \begin{figure}[htbp]
 \centering
-\includegraphics{figure/nonparametric-bayes-Figureb.pdf}
-\caption{plot of chunk Figureb}
+\includegraphics{figure/nonparametric-bayes-Figureb_posteriors.pdf}
+\caption{Points show the training data of stock-size over time. Curves
+show the posterior step-ahead predictions based on each of the estimated
+models.}
 \end{figure}
 
-Figure 1 shows the mean inferred state space dynamics of each model
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-statespace_plot.pdf}
+\caption{Graph of the inferred Gaussian process compared to the true
+process and maximum-likelihood estimated process. Graph shows the
+expected value for the function $f$ under each model. Two standard
+deviations from the estimated Gaussian process covariance with (light
+grey) and without (darker grey) measurement error are also shown. The
+training data is also shown as black points. The GP is conditioned on
+(0,0), shown as a pseudo-data point.}
+\end{figure}
+
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{rrrr}
+  \hline
+ & Allen & Ricker & Myers \\ 
+  \hline
+DIC & 50.14 & 49.45 & 50.61 \\ 
+  AIC & -26.60 & -32.07 & -29.19 \\ 
+  BIC & -21.54 & -28.69 & -24.12 \\ 
+   \hline
+\end{tabular}
+\end{center}
+\end{table}
+
+Figure 2 shows the mean inferred state space dynamics of each model
 relative to the true model used to generate the data, predicting the
 relationship between observed stock size (x-axis) to the stock size
-after recruitment the following year. All models except the MLE model
-estimate a distribution around the means shown here, and all models
-estimate a level of process noise, which is independent of the state
-value (x). Note that in contrast to the other models shown, the mean
-Gaussian process corresponds to a distribution of curves - as indicated
-by the gray band - which itself has a mean shown in black. Note that
-this mean GP is thus more certain of the dynamics in the region where
-data is available then where it is not.
+after recruitment the following year. Note that in contrast to the other
+models shown, the expected Gaussian process corresponds to a
+distribution of curves - as indicated by the gray band - which itself
+has a mean shown in black. Parameter uncertainty (not shown) spreads out
+the estimates further.
 
@@
 tipping point.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{policies <-}\StringTok{ }\KeywordTok{melt}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{stock=}\NormalTok{x_grid, }\KeywordTok{sapply}\NormalTok{(OPT, function(x) x_grid[x])), }\DataTypeTok{id=}\StringTok{"stock"}\NormalTok{)}
-\KeywordTok{names}\NormalTok{(policies) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"stock"}\NormalTok{, }\StringTok{"method"}\NormalTok{, }\StringTok{"value"}\NormalTok{)}
-
-\KeywordTok{ggplot}\NormalTok{(policies, }\KeywordTok{aes}\NormalTok{(stock, stock -}\StringTok{ }\NormalTok{value, }\DataTypeTok{color=}\NormalTok{method)) +}
-\StringTok{  }\KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{lwd=}\FloatTok{1.2}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) +}\StringTok{ }\KeywordTok{xlab}\NormalTok{(}\StringTok{"stock size"}\NormalTok{) +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{"escapement"}\NormalTok{)  +}
-\StringTok{  }\KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
-\end{Highlighting}
-\end{Shaded}
 
@@
 hockey-stick shaped policies shown.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{ggplot}\NormalTok{(sims_data) +}\StringTok{ }
-\StringTok{  }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, fishstock, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(reps,method), }\DataTypeTok{color=}\NormalTok{method), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) +}
-\StringTok{  }\KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{)))}
-\end{Highlighting}
-\end{Shaded}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{fig4v1 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) +}\StringTok{ }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{variable)) +}\StringTok{ }
-\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(~variable, }\DataTypeTok{scales =} \StringTok{"free_y"}\NormalTok{)  +}\StringTok{ }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{) +}
-\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{) +}\StringTok{ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey) }\CommentTok{# density plots fail when delta fn}
-
-\NormalTok{fig4v2 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value)) +}\StringTok{ }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill=}\NormalTok{variable), }\DataTypeTok{binwidth=}\FloatTok{0.1}\NormalTok{) +}\StringTok{ }
-\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+}\StringTok{ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
-
-\NormalTok{fig4v3 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(actual_over_optimal, }\KeywordTok{aes}\NormalTok{(value, }\DataTypeTok{fill=}\NormalTok{variable, }\DataTypeTok{color=}\NormalTok{variable)) +}\StringTok{ }\CommentTok{# density plots fail when delta fn}
-\StringTok{  }\KeywordTok{stat_density}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{y=}\NormalTok{..density..), }\DataTypeTok{position=}\StringTok{"stack"}\NormalTok{, }\DataTypeTok{adjust=}\DecValTok{3}\NormalTok{, }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{9}\NormalTok{) +}\StringTok{ }
-\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Total profit by replicate"}\NormalTok{)+}\StringTok{ }\KeywordTok{scale_fill_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)+}\StringTok{ }\KeywordTok{scale_color_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
-\NormalTok{fig4v2}
-\end{Highlighting}
-\end{Shaded}
-
 \begin{figure}[htbp]
 \centering
-\includegraphics{figure/nonparametric-bayes-Figure41.pdf}
-\caption{Histograms of the realized net present value of the fishery
-over a range of simulated data and resulting parameter estimates. For
-each data set, the three models are estimated as described above. Values
-plotted are the averages of a given policy over 100 replicate
-simulations. Details and code provided in the supplement.}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{fig4v3}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-Figure42.pdf}
+\includegraphics{figure/nonparametric-bayes-Figure4.pdf}
 \caption{Histograms of the realized net present value of the fishery
@@
 
-Glaser, Sarah M., Michael J. Fogarty, Hui Liu, Irit Altman, Chih-Hao
-Hsieh, Les Kaufman, Alec D. MacCall, Andrew a Rosenberg, Hao Ye, and
-George Sugihara. 2013. ``Complex Dynamics May Limit Prediction in Marine
-Fisheries.'' \emph{Fish and Fisheries} (apr). doi:10.1111/faf.12037.
-\url{http://doi.wiley.com/10.1111/faf.12037}.
-
 Gordon, H. S. 1954. ``The Economic Theory of a Common-property Resource:

### 2013-10-10 inst/examples/BUGS/external-chunks.R

@@ -416,26 +416,12 @@ models <- data.frame(x=x_grid,
                      Allen = allen_means,
                      Myers = myers_means)
 models <- melt(models, id="x")
-
 # some labels
 names(models) <- c("x", "method", "value")
-
 # labels for the colorkey too
 model_names = c("GP", "True", "MLE", "Ricker", "Allen", "Myers")
 colorkey=cbPalette
 names(colorkey) = model_names 
-
-
-## @knitr Figure1
-plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
-    geom_line(data=models, aes(x, value, col=method), lwd=1, alpha=0.8) + 
-    geom_point(data=obs, aes(x,y), alpha=0.8) + 
-    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
-    scale_colour_manual(values=cbPalette) 
-print(plot_gp)
-
-
-## @knitr Figureb
 require(MASS)
 step_ahead <- function(x, f, p){
   h = 0
@@ -444,9 +430,29 @@ step_ahead <- function(x, f, p){
   y <- c(x[1], x_predict[1:n])
   y
 }
-step_ahead_posteriors <- function(x){
+
+
+## @knitr Figureb
+names(bayes_pars) <- c("r0", "K", "theta")
+names(myers_bayes_pars) <- c("r0",  "theta", "K")
+names(ricker_bayes_pars) <- c("r0",  "K")
 gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
-df_post <- melt(lapply(sample(100), 
+df <- melt(data.frame(time = 1:length(x), stock = x, 
+                GP = gp_f_at_obs$E_Ef,
+                True = step_ahead(x,f,p),  
+                MLE = step_ahead(x,f,est$p), 
+                Allen = step_ahead(x, allen_f, bayes_pars), 
+                Ricker = step_ahead(x, ricker_f, ricker_bayes_pars), 
+                Myers = step_ahead(x, myers_f, myers_bayes_pars)
+                 ), id=c("time", "stock"))
+Figure1b <- ggplot(df) + geom_point(aes(time, stock)) + 
+  geom_line(aes(time, value, col=variable)) +
+    scale_colour_manual(values=colorkey) 
+
+## @knitr Figureb_posteriors 
+step_ahead_posteriors <- function(x){
[…]

### 2013-10-14 inst/doc/nonparametric-bayes.Rmd
writing

@@
 
+<!-- 
 ### Figure 1: Fitted Models
@@
 - Information criteria would pick the simple, incorrect model.
+-->
 
@@
 Model-choice criteria shown in Table 1 penalize more complex models and show a slight preference for the simpler Ricker model over the more complicated alternate stable state models (Allen and Myers).
+Details on MCMC estimates for each model, traces, and posterior distributions can be found in the appendix.
 
@@
 
-```{r statespace_plot, dependson=c("assemble-models"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
+<!-- 
+### Figure 2 
+- Data comes from limited region of state-space 
+- (Should really show uncertainty of all models here.  Capture the forecast uncertainty several steps down the road?)   
+-->
+
+```{r statespace_posteriors, dependson=c("assemble-models"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
 Graph shows the expected value for the function $f$ under each model.
@@
 The training data is also shown as black points.
-The GP is conditioned on (0,0), shown as a pseudo-data point.", include=TRUE, echo=FALSE}
+The GP is conditioned on (0,0), shown as a pseudo-data point.", include=TRUE, echo=FALSE, fig.width=8, fig.height=6}
 ```
@@
 
+```{r out_of_sample_predictions, include=TRUE, echo=FALSE, fig.width=8, fig.height=6}
+```
+
+<!--
 ### Figure 3: Inferred Policies
@@
 - alternatives are not close
+-->
 
@@
 
-These results illustrate that model-choice approaches would be postively misleading -- supporting simpler models that cannot express tipping point dynamics merely on account of them being similar.
-As the data shown comes only from the basin of attraction near the unfished equilibrium, near which all of the models are approximately linear and approximately identical.
+<!--
 
@@
 quantitative decision making framework.  
+-->
 
@@
 
-### 1. Large uncertainty where the data is poor
+### Traditional model-choice approaches can be positively misleading.
+
+These results illustrate that model-choice approaches would be positively
+misleading -- supporting simpler models that cannot express tipping point
+dynamics merely on account of them being similar.  As the data shown
+comes only from the basin of attraction near the unfished equilibrium,
+near which all of the models are approximately linear and approximately
+identical.
+
+Model choice approaches trade off model complexity and fit to the data. 
+When the data come from a limited region of state-space -- as is necessarily 
+the case whenever there is a potential concern about tipping point dynamics --
+simpler models can fit just as well and will tend to outperform more complex 
+ones.  This approach would be appropriate when the dynamics can be expected 
+to remain in the region of the training data; for instance, if we only 
+considered the forecasting accuracy of the unfished population dynamics under
+each model.
+
+In contrast, the decision-maker's problem of setting appropriate harvest levels
+cannot exclude regions of state-space outside the observed range when integrating
+over all possible decisions to find the optimal choice.  Such problems are not 
+constrained to fisheries management but ubiquitous across ecological decision-making
+and conservation where the greatest concerns involve entering previously unobserved
+regions of state-space -- whether that is the collapse of a fishery, the spread
+of an invasive, or the loss of habitat.
+
+### BNP-SDP expresses larger uncertainty in regions where the data are poor
 
@@
 
-### 2. Predictive accuracy where data is good
+### BNP-SDP has good predictive accuracy where data are good
 
@@
 
-### Online learning
+### Real-time learning
 
@@
 
------------------------------------
-
-<!-- OLD TEXT -->
-
-Abstract
-=======================================================================
-
-Decision-theoretic methods often rely on simple parametric models of
-ecological dynamics to compare the value of a potential sequence of
-actions. Unfortunately, such simple models rarely capture the complexity
-or uncertainty found in most real ecosystems.
-
-Further, the data on which a model has been parameterized frequently 
-fails to cover the possible state-space over which management decisions 
-must operate.  Consequently a model do well in the region of state-space
-in which it was estimated, but give erroneous confidence to predictions
-outside of that region.
-
-This problem is keenly felt in any system where a potential threshold
-or tipping point is a concern.  Such a tipping point, if it exists 
-at all, will lay outside the observed range of the observed data.
-
-We demonstrate how nonparametric Bayesian models can provide robust,
-solutions to decision making under uncertainty without knowing the 
-structural form of the true model.
-
-While methods that account for _parametric_ uncertainty can be very 
-successful with the right model,
-structural uncertainty of not knowing what model best approximates the 
-dynamics poses considerably greater difficulty.
-
-Introduction
-=======================================================================
-
-#### Opening
-
-<!-- More on complex dynamics and not having the correct model
-
--->
-
-<!-- More on the lack of data throughout the relevant state-space
-     and how any concern about potential tipping points indicates 
-     that the data exhibits this bias / problem.  
-     -->
-
-#### Models for decision-making under uncertainty
-
-Decision-theoretic or optimal control tools require a model that can assign probabilities of
-future states (e.g. stock size of a fishery) given the current state and a
-proposed action (e.g. fishing harvest or effort).  
-Management frequently faces a sequential decision-making problem -- after 
-selecting an action, the decision-maker may receive new information about 
-the current state and must again choose an appropriate action -- such as setting
-the harvest limits each year based on stock assessments the year prior.
-
-The decision maker typically seeks to determining the course of actions (also referred to as the policy)
-that maximizes the expected value of some objective function such as
-net present value derived from the resource over time.  
-Though much can be said on how to choose this value function appropriately 
-(e.g. see [@Halpern2013](http://doi.org/10.1073/pnas.1217689110 
-"Achieving the triple bottom line in the face of inherent trade-offs 
-among social equity, economic return, and conservation.")) we will
-assume this is given.  (Nor is this approach necessarily constrained to 
-maximizing the expectated value of such a function - the decision-theoretic
-framework can be adapted to alternatives such as minimizing the maximum 
-cost or damage that might be incurred; see @Polasky2011).
-
-In representing future states with probabilities and maximizing expectations,
-this approach provides a natural framework for handling uncertainty.
-
-The value function typically depends on the action or policy taken, as well as
-the state of the system, in each interval of time. The state of the system,
-in turn, is usually described by a dynamical model.
-
-[@Williams2001; @Athanassoglou2012].
-
-<!-- Transition and map: The weakness of parametric models -->
-<!-- This section is a bit weak and vague.  Revisit to make a real map in light 
-of outline structure below. -->
-
-While simple mechanistic models can nevertheless provide important insights
-into long-term outcomes, such approaches are not well-suited for use in
-forecasting outcomes of potential management options.  Non-parametric
-approaches offer a more flexible alternative that can both more accurately
-reflect the data available while also representing greater uncertainty
-in areas (of state-space) where data is lacking.
-
-We demonstrate how
-a Gaussian Process model of stock recruitment can lead to nearly optimal
-management through stochastic dynamic programming, comperable to knowing
-the correct structural equation for the underlying simulation.  Meanwhile,
-parametric models that do not match the underlying dynamics can perform 
-very poorly, even though they fit the data as well as the true model.  
-Ecological research and management strategy should pay closer attention
-to the opportunities and challenges nonparametric modeling can offer.
-
-<!-- 
-### Quantitative vs Qualitative Decisions
-
-In this paper, we consider those ecological management problems in which
-a mathematical (or computational) model is used to quantitatively inform
-decision-making by allowing a manager to compare to the expected consequences
-of potential management actions (or policies).  We distinguish this from the
-solely qualitative use of a model, in which models are used to represent and
-compare hypotheses of different mechanisms that would lead to qualitatively 
-different actions.
-
-In addition to facing the potentially grave consequences of such transitions,
-if they do exist in a system of interest, this generally means that we lack
-data in the region near and beyond a possible tipping point.  As a consequence,
-our models are least accurate just where we need them to be most accurate.  
-Unfortunately, parametric modeling approaches are not well-suited to this 
-scenario, in which good data is avialble over only part of the relevant 
-state-space.  Necessary assumptions about model structure in any parametric
-approach can lead such models to perform very poorly, even when accounting
-for parameter uncertainty.
-
--->
-
-Approach and Methods
-====================
-
-## The optimal control problem in fisheries management
-
-We focus on the problem in which a manager must set
-the harvest level for a marine fishery each year to maximize the net
-present value of the resource, given an estimated stock size from the
-year before.
-
-<!-- ugh, re-word this -->
-
-To permit comparisons against a theoretical optimum we will consider
-data on the stock dynamics simulated from a simple parametric model
-in which recruitment of the fish stock $X_{t+1}$ in the following year
-is a stochastic process governed by a function $f$ of the current 
-stock $X_t$, selected harvest policy $h_t$, and noise process $Z$,
-
-$$X_{t+1} = Z_t f(X_t, h_t) $$
-
-Given parameters for the function $f$ and probability distribution $Z$,
-along with a given economic model determining the  price/profit $\Pi(X_t,
-h_t)$ realized in a given year given a choice of harvest $h_t$ and
-observed stock $X_t$.  This problem can be solved exactly for discretized
-values of stock $X$ and policy $h$ using stochastic dynamic programming
-(SDP) [@Mangel1988]. Problems of this sort underpin much marine fisheries
-management today.
-
-A crux of this approach is correctly specifying the functional form of $f$,
-along with its parameters.  The standard approach uses one of a 
-handful of common parametric models representing the stock-recruitment
-relationship, usually after estimating the model parameters from any 
-available existing data. Uncertainty in the parameter estimates can 
-be estimated and integrated over to determine the optimal policy under
-under uncertainty [@Mangel1988; @Schapaugh2013]. Uncertainty in the model
-structure itself can only be addressed in this approach by hypothesizing 
-alternative model structures, and then performing some model choice or
-model averaging  [@Williams2001; @Athanassoglou2012].
-
-## Underlying Model
-
-To illustrate the value of the non-parametric Bayesian approach to management,
-we focus on example of a system containing such a tipping point whose dynamics
-can still be described by a simple, one-dimensional parametric model.  
-We will focus on a simple parametric model for a single species [derived
-from fist principles by @Allen2005a] as our underlying "reality".
-
-$$X_{t+1} = Z_t f(S_t)  $$
-$$S_t = X_t - h_t $$
-$$f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} $$
-
-Where $Z_t$ is multiplicative noise function with mean 1, representing
-stochastic growth. We will consider log-normal noise with shape parameter
-$\sigma_g$.  We start with an example in which the parameters are $r =2$,
-$K=8$, $C=5$ and $\sigma_g = 0.1$.
-
-<!-- ` r p[1]`, $K =$ ` r p[2]`, $C =$ ` r p[3]`, and  $\sigma_g =$ ` r sigma_g`. -->
-
-As a low-dimensional system completely described by three parameters, this
-scenario should if anything be favorable to a parametric-based approach.
-This model contains an Allee effect, or tipping point, below which the
-population is not self-sustaining and shrinks to zero [@Courchamp2008].
-
-#### Simulated training data
-
-We generate initial observational data under the model described in Eq 1 
-for $T_{\textrm{obs}}=40$ time steps, under a given arbitrary sequence of harvest
-intensities, $h_t$. 
-We consider the case in which most of the data comes from a limited region
-of state space (e.g. near a stable equilibrium), leaving us without
-observations of the population dynamics at very low levels which would
-be useful in discrimating between recruitment curves [@] or demonstrating
-the existence of a tipping point [@Scheffer2001].  
-Using data simulated from a specified model rather than empirical data
-permits the comparison against the true underlying dynamics, setting 
-a bar for the optimal performance possible.
-
-### Parametric Models
-
-We consider three candidate parametric models for the stock-recruitment function, 
-which we refer to by the first authors of the publications in which they were
-first proposed.
-
-We generate the data with a four-parameter model that contains
-a tipping point, as discussed above (equation 1),
-(an Allee effect, see [@Allen, @Courchamp]) below which the stock
-decreases to zero,
-
-$$ X_{t+1} = Z_t S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(\frac{S_t - \theta}{K}\right)} $$
-
-$$ S_t = X_t - h_t $$
-
-The parameter $C$ reflects the location of the tipping point, $K$ the carrying capacity of
-the stock, and $r$ the base recruitment rate.  $S_t$ represents the stock size after
-a harvest $h_t$ has been implemented.  $Z_t$ represents a log-normal random variable of
-log-mean zero and log-standard deviation parameter $\sigma$.
-
-We consider two alternative candidate models: the Ricker [@Ricker] stock-recruitment curve,
-
-$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
-
-and an alternative four-parameter model adapted from @Myers,
-
-$$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} $$
-
-which contains a tipping point for $\theta > 2$ and becomes a Beverton-Holt model 
-at $\theta = 1$.
-
-<!-- Note that while similarly-named parameters contain similar connotations, the 
-estimate and precise intepretation will be different in each model. -->
-
-### Bayesian Inference of Parametric models
-
-Given the sample data, we infer posterior distributions for each of the three
-models listed above using a Markov Chain Monte Carlo Gibbs Sampler (jags, see 
-appendix for implementation details and code) given uniform priors.  We run six
-chains for $10^6$ steps each and then assess convergence by Gelman-Rubin criterion
-and inspection of the traces, see appendix.
-
-<!-- By transforming variables we can integrate out the the $\sigma$ parameter
-and the constant term in the log-mean, see appendix.  The resulting one or
-two parameters can then be evaluated over a vector or grid of possible values,
-and the result transformed back to determine the posterior density of the parameters,
-This approach avoids any concerns of convergence in the MCMC analysis and can thus
-be used to verify that the performance of the parametric models can not be explained
-by convergence issues.  However, this approach does not generalize easily to richer models,
-such as the introduction of measurement noise or additional parameters, in which 
-MCMC based approaches are the obvious choice.  -->
-
-### The Non-parametric Bayesian alternative for stock-recruitment curves
-
-### SDP via GP
-
-Once the posterior Gaussian process (GP) has been estimated [e.g. see
-@Munch2005], it is necessary to adapt it in place of the parametric
-equation for the stochastic dynamic programming (SDP) solution [see
-@Mangel1988 for a detailed description of parametric SDP methods] to the
-optimal policy. The essence of the idea is straight forward -- we will use
-the estimated GP in place of the parametric growth function to determine
-the stochastic transition matrix on which the SDP calculations are based.
-The SDP is solved in a discretized state space -- both the continuously 
-valued population densities $X$ and harvest quotas $h$ are first mapped to
-a bounded, discrete grid.  (For simplicity we will consider a uniform grid,
-though for either parametric or GP-based SDP it is often advantageous to 
-use a non-uniform discretization such as a basis function representation,
-e.g. see [@Deisenroth2009]).
-
-The SDP approach then computes a transition matrix, $\mathbf{F}$.  We 
-demonstrate that calculation is just as straight forward based on the GP
-as it is in the classical context using the parametric model.  The 
-${i,j}$ of the transition matrix $F$ entry gives the probability of transitioning into 
-state $x_i$ given that the system is in state $x_j$ in the previous 
-time-step.  To generate the transition
-matrix based on the posterior GP, we need only the expected values
-at each grid point and the corresponding variances (the diagonal of
-the covariance matrix), as shown in Figure 1.  Given the mean of the 
-GP posterior at each grid-point as the vector $E$ and variance at that
-point as vector $V$, the probability of transitioning from state $x_i$ to state $x_j$ is
-
-$$\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)$$
-
-where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
-variance $\sigma^2$.  Strictly speaking, the transition probability should
-be calculated by integrating the normal density over the bin of width
-$\Delta$ centered at $x_j$.  For a sufficiently fine grid that $f(x_j)
-\approx f(x_j + \Delta)$, it is sufficient to calculate the density at
-$x_j$ and then row-normalize the transition matrix. The process
-can then be repeated for each possible discrete value of our control 
-variable, (harvest $h$).
-
-**Pseudocode for the determining the transition matrix from the GP**
-
-```r
-for(h in h_grid)
-  F_h = for(x_j in grid)
-          for(i in 1:N) 
-            dnorm(x_j, mu[i]-h, V[i])
-```
-
-Using the discrete transition matrix we may write down the Bellman
-recursion defining the stochastic dynamic programming iteration:
-
-\begin{equation}
-V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
-\end{equation}
-
-where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$
-is control (harvest level) chosen. Numerically, the maximization is
-accomplished as follows. Consider the set of possible control values to
-be the discrete values corresponding the grid of
-stock sizes.  Then for each $h_t$ there is a corresponding transition
-matrix $\mathbf{F}_h$ determined as described above but with mean 
-$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element corresponds
-to the value of having stock $x_i$ at time $t$.  Then let $\Pi_h$ be
-the vector whose $i$th element indicates the profit from harvesting
-at intensity $h_t$ given a population $x_i$ (e.g. $\max(x_i, h_t)$
-since one cannot harvest more fish then the current population size).
-Then the Bellman recursion can be given in matrix form as
-
-$$V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)$$
-
-where the sum is element by element and the expectation is computed by the matrix multiplication $\mathbf{F} V_{t+1}$.
-
-### Pseudocode for the Bellman iteration
-
-```r
- V1 <- sapply(1:length(h_grid), function(h){
-      delta * F[[h]] %*% V +  profit(x_grid, h_grid[h]) 
-    })
-    # find havest, h that gives the maximum value
-    out <- sapply(1:gridsize, function(j){
-      value <- max(V1[j,], na.rm = T) # each col is a diff h, max over these
-      index <- which.max(V1[j,])  # store index so we can recover h's 
-      c(value, index) # returns both profit value & index of optimal h.  
-    })
-    # Sets V[t+1] = max_h V[t] at each possible state value, x
-    V <- out[1,]                        # The new value-to-go
-    D[,OptTime-time+1] <- out[2,]       # The index positions
-```
-
-This completes the algorithm adapting the GP to the sequential decision-making 
-problem through SDP, which has not been previously demonstrated.  
-We further provide an R package implementation as described in the supplemental materials.
-
-### Estimating parametric models
-
-We estimate posterior distributions for two parametric models: one using the structurally correct
-model as given in Eq (1), which we refer to as the "Parametric Bayes" model,
-and another using the familiar Ricker model, using a Gibbs sampler
-as described (with source code) in the appendix).  In addition we 
-estimate the parameters of the structurally correct model by maximum
-likelihood.
-
-Results
-=======
-
-Discussion 
-==========
-
-#### Big picture: Linking GP to SDP
-
-_rambling_
-
-Non-parametric Bayesian methods have received far too little attention
-in ecological modeling efforts that are aimed at improved conservation
-planning and decision making support.  Such approaches may be particularly
-useful when the available data is restricted to a limited area of
-state-space, which can lead parametric models to underestimate the
-uncertainty in dynamics at population levels (states) which have not
-been observed.  One reason for the relative absence of nonparametric
-approaches in the natural resource management context may be the lack
-of existing approaches for adapting the non-parametric Bayesian models
-previously proposed [@Munch2005] to a decision-theoretic framework.
-Adapting a non-parametric approach requires modification of existing
-methods for decision theory.  We have illustrated how this might be
-done for a classic stochastic dynamic programming problem, opening the
-door for substantial further research into how these applications might
-be improved.
-
-<!-- On mechanistic models -->
-<!-- { Seems kind of irrelevant }
-
-The complexity of ecological interactions and a lack of data contribute
-greatly to both of the problems. This concern is particularly acute
-in case of ecological tipping points [@Barnosky2012; @Scheffer2001],
-which arise from feedbacks common in ecological systems and can lead to a
-sudden catastrophic transition to an undesirable state.  In general we do
-not know where such tipping points are unless we have already observed
-the transition, in which case it is frequently too late to respond.
-Though there may exist early warning signals for certain kinds of these
-transitions that are driven by slow changes [@Scheffer2009], we do not
-know when, where, or how to apply them to the decision making context
-more generally [@Boettiger2013].
-
--->
-
-----------------------------------------
-
-Code to replicate the analysis, along with complete log of this research can be found at: [https://github.com/cboettig/nonparametric-bayes](https://github.com/cboettig/nonparametric-bayes/)
-
-## Markov Chain Monte Carlo Analysis
-
-```{r appendixplots, dependson="assemble-models"}
-```
+### Sensitivity Analysis

### 2013-10-14 inst/examples/BUGS/external-chunks.R

@@ -473,13 +473,67 @@ figure1b_posteriors
 
 
 ## @knitr statespace_plot
-statespace_plot <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin = ymin,ymax = ymax), fill = "gray80") +
-    geom_line(data = models, aes(x, value, col = method), lwd=1, alpha = 0.8) + 
+statespace_plot <- ggplot(tgp_dat) + 
+    geom_ribbon(aes(x,y,ymin = ymin,ymax = ymax), fill = "gray80") +
+    geom_line(data = models, aes(x, value, col = method), lwd=1, alpha = 0.7) + 
     geom_point(data = obs, aes(x,y), alpha = 0.8) + 
     xlab(expression(X[t])) + ylab(expression(X[t+1])) +
-    scale_colour_manual(values = cbPalette) 
+    scale_colour_manual(values = colorkey) 
 statespace_plot
 
+
+
+## @knitr statespace_posteriors
+x_grid_short <- x_grid[1:40]
+gp_short <- gp_predict(gp, x_grid_short, burnin=1e4, thin=300)
+models_posteriors <- 
+  melt(lapply(sample(100, 50), 
+              function(i){
+    sample_gp <- mvrnorm(1, 
+                            gp_short$Ef_posterior[,i],         
+                            gp_short$Cf_posterior[[i]])
+    data.frame(stock = x_grid_short, 
+               GP = sample_gp,
+               y = sample_gp,
+               ymin = sample_gp - 2 * sqrt(gp_short$E_Vf), 
+               ymax = sample_gp + 2 * sqrt(gp_short$E_Vf), 
+               True = sapply(x_grid_short,f,0, p),  
+               MLE = sapply(x_grid_short,f,0, est$p), 
+               Allen = sapply(x_grid_short, allen_f, 0, pardist[i,]), 
+               Ricker = sapply(x_grid_short, ricker_f, 0, ricker_pardist[i,]), 
+               Myers = sapply(x_grid_short, myers_f, 0, myers_pardist[i,]))
+             }), 
+       id=c("stock", "y", "ymin", "ymax"))
+ggplot(models_posteriors) + 
+    geom_ribbon(aes(x=stock, y=y, ymin=ymin, ymax=ymax, group=L1), 
+                  fill = "gray80", 
+                  data=subset(models_posteriors, variable == "GP")) + 
+    geom_line(aes(stock, value, col = variable, 
[…]

### 2013-10-14 inst/examples/sensitivity-ricker.Rmd
new sensitivity files

--- 2013-07-28 inst/examples/sensitivity.Rmd
@@
 ```{r stateeq-local}
-f <- Myers
+f <- RickerAllee
 
@@
 xT <- 0
-Xo <-  1.7 # observations start from
-x0 <- 5 # simulation under policy starts from
+Xo <-  5.7 # observations start from
+x0 <- 8 # simulation under policy starts from
 Tobs <- 40
@@
 
-  p <- c(1.5 + rnorm(1, 0, .1), 2.5 + rnorm(1, 0, .1), 8 + rnorm(1, 0, .2))
+  p <- c(2 + rnorm(1, 0, .1), 8 + rnorm(1, 0, .1), 5 + rnorm(1, 0, .2))
 

### 2013-10-14 inst/examples/sensitivity.R

@@ -1,0 +1,2 @@
+require(knitr)
+knit("sensitivity.Rmd")

### 2013-10-14 inst/doc/nonparametric-bayes.tex

@@
 
+All models fit the observed data rather closely and with relatively
+small uncertainty, as illustrated in the posterior predictive curves in
+Figure 1. Figure 1 shows the training data of stock sizes observed over
+time as points, overlaid with the step-ahead predictions of each
+estimated model using the parameters sampled from their posterior
+distributions. Each model manages to fit the observed data rather
+closely. Compared to the expected value of the true model most estimates
+appear to overfit, predicting fluctuations that are actually due purely
+to stochasticity in growth rate. Model-choice criteria shown in Table 1
+penalize more complex models and show a slight preference for the
+simpler Ricker model over the more complicated alternate stable state
+models (Allen and Myers). Details on MCMC estimates for each model,
+traces, and posterior distributions can be found in the appendix.
+
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{rrrr}
+  \hline
+ & Allen & Ricker & Myers \\ 
+  \hline
+DIC & 50.14 & 49.45 & 50.61 \\ 
+  AIC & -24.60 & -30.07 & -27.19 \\ 
+  BIC & -17.85 & -25.00 & -20.44 \\ 
+   \hline
+\end{tabular}
+\end{center}
+\end{table}
+
 \begin{figure}[htbp]
 \centering
-\includegraphics{figure/nonparametric-bayes-statespace_plot.pdf}
+\includegraphics{figure/nonparametric-bayes-statespace_posteriors.pdf}
 \caption{Graph of the inferred Gaussian process compared to the true
@@
 
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{rrrr}
-  \hline
- & Allen & Ricker & Myers \\ 
-  \hline
-DIC & 50.14 & 49.45 & 50.61 \\ 
-  AIC & -26.60 & -32.07 & -29.19 \\ 
-  BIC & -21.54 & -28.69 & -24.12 \\ 
-   \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-Figure 2 shows the mean inferred state space dynamics of each model
-relative to the true model used to generate the data, predicting the
-relationship between observed stock size (x-axis) to the stock size
+The mean inferred state space dynamics of each model relative to the
+true model used to generate the data is shown in Figure 2, predicting
+the relationship between observed stock size (x-axis) to the stock size
 after recruitment the following year. Note that in contrast to the other
@@
 has a mean shown in black. Parameter uncertainty (not shown) spreads out
-the estimates further.
-
-While it would be straight forward to condition the GP on passing
-through the origin (0,0), the estimate shown here is based only on the
-observed data. The observed data from which each model is estimated is
-also shown. The observations come from only a limited region of state
-space corresponding to unharvested or weakly harvested system. No
-observations occur at the theoretical optimum harvest rate or near the
-tipping point.
+the estimates further.\\The observed data from which each model is
+estimated is also shown. The observations come from only a limited
+region of state space corresponding to unharvested or weakly harvested
+system. No observations occur at the theoretical optimum harvest rate or
+near the tipping point.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-out_of_sample_predictions.pdf}
+\caption{plot of chunk out\_of\_sample\_predictions}
+\end{figure}
 
@@
 
-The resulting optimal management strategy based on each of the inferred
-models is shown in Figure 2, against the optimal strategy given the true
-underlying dynamics. Policies are shown in terms of target escapement,
-$S_t$. Under models such as this a constant escapement policy is
-expected to be optimal (Reed 1979), whereby population levels below a
-certain size $S$ are unharvested, while above that size the harvest
-strategy aims to return the population to $S$, resulting in the
-hockey-stick shaped policies shown.
+Despite the similarities in model fits to the observed data, the
+policies inferred under each model differ widely, as shown in Figure
+3.\\Policies are shown in terms of target escapement, $S_t$. Under
+models such as this a constant escapement policy is expected to be
+optimal (Reed 1979), whereby population levels below a certain size $S$
+are unharvested, while above that size the harvest strategy aims to
+return the population to $S$, resulting in the hockey-stick shaped
+policies shown. Only the structurally correct model (Allen model) and
+the GP produce policies close to the true optimum policy (where both the
+underlying model structure and parameter values are known without
+error).
 
@@
 The consequences of managing 100 replicate realizations of the simulated
-fishery under each of the policies estimated is shown in Figure 3. As
+fishery under each of the policies estimated is shown in Figure 4. As
 expected from the policy curves, the structurally correct model
@@
 
-The results shown in Figures 1-3 are not unique to the simulated data or
-models chosen here, but arises across a range of parameter values and
-simulations as shown in the supplemental figures. The results across
-this range can most easily be compared by the relative differences in
-net present value realized by each of the approaches, as shown in Figure
-4. The Gaussian Process most consistently realizes a value close to the
-optimal solution, and importantly avoids ever driving the system across
-the tipping point, which results in the near-zero value cases in the
+These results are robust across a range of stochastic realizations,
+models, and parameter values.\\The results across this range can most
+easily be compared by using the relative differences in net present
+value realized by each of the model, as shown in Figure 5. The BNP-SDP
+approach most consistently realizes a value close to the optimal
+solution, and importantly avoids ever driving the system across the
+tipping point, which results in the near-zero value cases in the
 parametric models.
@@
 
-\subsubsection{Figure 1: Fitted Models}\label{figure-1-fitted-models}
+\section{Discussion}\label{discussion}
 
@@
 \item
-  All models fit the data quite well
-\item
-  Information criteria would pick the simple, incorrect model.
-\end{itemize}
-
-\subsubsection{Figure 2: Inferred
-Policies}\label{figure-2-inferred-policies}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Inferred policies differ substantially among models
-\item
-  The structurally correct model and the GP are close to the true model
-\item
-  alternatives are not close
-\end{itemize}
-
-\subsubsection{Figure 3: Simulated
-results}\label{figure-3-simulated-results}
-
--
-
-\subsubsection{Figure 4: Robustness}\label{figure-4-robustness}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Results hold across range of parameters
-\end{itemize}
-
-(eek, distinguish better between result and discussion?)
-
-\section{Discussion}\label{discussion}
-
-\begin{itemize}
-\item
   All models are ``good fits'' to the originally observed data.
@@
 \end{itemize}
-
-Though simple mechanistically motivated models offer the greatest
-potential to increase our basic understanding of ecological processes
-(Cuddington, Fortin, and Gerber 2013; Geritz and Kisdi 2012), such
-models can be not only inaccurate but misleading when relied upon in a
-quantitative decision making framework.
 
@@
 
-\subsubsection{1. Large uncertainty where the data is
-poor}\label{large-uncertainty-where-the-data-is-poor}
+\subsubsection{Traditional model-choice approaches can be positively
+misleading.}\label{traditional-model-choice-approaches-can-be-positively-misleading.}
+
+These results illustrate that model-choice approaches would be
+positively misleading -- supporting simpler models that cannot express
+tipping point dynamics merely on account of them being similar. As the
+data shown comes only from the basin of attraction near the unfished
+equilibrium, near which all of the models are approximately linear and
+approximately identical.
+
+Model choice approaches trade off model complexity and fit to the data.
+When the data come from a limited region of state-space -- as is
+necessarily the case whenever there is a potential concern about tipping
+point dynamics -- simpler models can fit just as well and will tend to
+outperform more complex ones. This approach would be appropriate when
+the dynamics can be expected to remain in the region of the training
+data; for instance, if we only considered the forecasting accuracy of
+the unfished population dynamics under each model.
+
+In contrast, the decision-maker's problem of setting appropriate harvest
+levels cannot exclude regions of state-space outside the observed range
+when integrating over all possible decisions to find the optimal choice.
+Such problems are not constrained to fisheries management but ubiquitous
+across ecological decision-making and conservation where the greatest
+concerns involve entering previously unobserved regions of state-space
+-- whether that is the collapse of a fishery, the spread of an invasive,
+or the loss of habitat.
+
+\subsubsection{BNP-SDP expresses larger uncertainty in regions where the
+data are
+poor}\label{bnp-sdp-expresses-larger-uncertainty-in-regions-where-the-data-are-poor}
 
@@
 
-\subsubsection{2. Predictive accuracy where data is
-good}\label{predictive-accuracy-where-data-is-good}
+\subsubsection{BNP-SDP has good predictive accuracy where data are
+good}\label{bnp-sdp-has-good-predictive-accuracy-where-data-are-good}
 
@@
 
-\subsubsection{Online learning}\label{online-learning}
+\subsubsection{Real-time learning}\label{real-time-learning}
 
@@
 
-\% latex table generated in R 3.0.1 by xtable 1.7-1 package \% Mon Sep 9
-14:43:29 2013
-
 \begin{table}[ht]
-\centering
+\begin{center}
 \begin{tabular}{rlrr}
@@
 \end{tabular}
+\end{center}
 \end{table}
@@
 
-Table S2: Parameterization of the priors \% latex table generated in R
-3.0.1 by xtable 1.7-1 package \% Mon Sep 9 14:43:37 2013
+Table S2: Parameterization of the priors
 
 \begin{table}[ht]
-\centering
+\begin{center}
 \begin{tabular}{rlrr}
@@
 
-\% latex table generated in R 3.0.1 by xtable 1.7-1 package \% Mon Sep 9
-14:43:45 2013
-
 \begin{table}[ht]
-\centering
+\begin{center}
 \begin{tabular}{rlrr}
@@
 
-Cuddington, Kim M., M. J. Fortin, and L. R. Gerber. 2013.
-``Process-based Models Are Required to Manage Ecological Systems in a
-Changing World.'' \emph{Ecosphere} 4 (February): 1--12.
-doi:10.1890/ES12-00178.1.
-\url{http://www.esajournals.org/doi/abs/10.1890/ES12-00178.1}.
-
 Deisenroth, Marc Peter, Carl Edward Rasmussen, and Jan Peters. 2009.
@@
 \url{http://linkinghub.elsevier.com/retrieve/pii/S0925231209000162}.
-
-Geritz, Stefan A. H., and Eva Kisdi. 2012. ``Mathematical Ecology: Why
-Mechanistic Models?'' \emph{Journal of Mathematical Biology} 65 (6-7)
-(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.
 

### 2013-10-16 inst/examples/sensitivity-allen.Rmd
updates to sensitivity, notes in sensitivity.Rmd

@@
-```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE ,include=FALSE}
 rm(list=ls())
@@
 
-Replicate data-sets from the same model
+Replicate data-sets from the same model, varying both model parameters and data simulation seed
 
@@
 
+<!--
+
 Compare over `r length(seed)` simulation configurations with different parameters and different training data, the GP does consistently well.  Only one case has relatively poor outcomes.
@@
 ```
+
+-->

### 2013-10-16 inst/examples/sensitivity.Rmd
updates to sensitivity, notes in sensitivity.Rmd

@@
 
-### Model and parameters
+```{r}
+models <- c("Myers","Allen")
+
+parameters <- list(Myers = list(
+  c(1.5 + rnorm(1, 0, .1), 2.5 + rnorm(1, 0, .1), 8 + rnorm(1, 0, .2)),
+  c(1.5 + rnorm(1, 0, .1), 2.5 + rnorm(1, 0, .1), 8 + rnorm(1, 0, .2)),
+  c(1.5 + rnorm(1, 0, .1), 2.5 + rnorm(1, 0, .1), 8 + rnorm(1, 0, .2)))
+                   Allen = list(
+  c(2 + rnorm(1, 0, .1), 8 + rnorm(1, 0, .1), 5 + rnorm(1, 0, .2)),
+  c(2 + rnorm(1, 0, .1), 8 + rnorm(1, 0, .1), 5 + rnorm(1, 0, .2)),
+  c(2 + rnorm(1, 0, .1), 8 + rnorm(1, 0, .1), 5 + rnorm(1, 0, .2)))
+                   )
+nuisance_pars <- c("sigma_g", )
+nuisance_values <- list(sigma_g = c(0.01, 0.05, 0.1, 0.2))
+replicates <- c(1111, 2222, 3333, 4444, 5555, 6666, 7777, 8888) # seeds
+```
+
+Rather than loop over in brute force way:
+
+```{r eval=FALSE}
+
+lapply(models, function(m){
+  if(m == "Myers")
+    f <- Myers
+  else if(m == "Allen")
+    f <- RickerAllee
+  lapply(parameters[[m]], function(p)
+    lapply(nuisance_pars, function(n)
+      lapply(nuisance_values[[n]], function(v)
+        lapply(replicates)
+
+```
+
+We do better to launch seperate scripts and write to a common data format.
+
+We define a function that takes as arguments all things we wish to vary in the senstivity analysis, and returns a data.frame with the values of interest
+
+```{r eval=FALSE}
+sensitivity <- function(model, parameters, nuisance, seed){
+
+data.frame(model = model, pars = as.list(parameters), x = x)
+}
+```
+
+The data frame isn't a minimal representation: we intentionally replicate
 

### 2013-11-13 inst/doc/nonparametric-bayes.Rmd
added brief discussion of issues of priors and issues of value function curvature to the discussion

@@
 - (Simple model choice immediately leads us astray)
+- Understanding the role of the prior
 
@@
 parameters and is thus far less likely to overfit.
+
+### Risk-prone and Risk-adverse value functions
+
+The degree to which the decision-making part of the algorithm (the SDP)
+chooses to explore or avoid the resulting region of uncertainty can also
+be influenced by the curvature of the value (profit) function $\Pi$. Both to simplify 
+the intution and avoid biasing this result, we have chosen a profits that
+are linear in the catch and thus neither risk-prone nor risk adverse.  Making
+this function concave, representing the typical assumption of diminishing returns,
+would make the SDP more risk-adverse (as larger-than-expected stock sizes offer
+dimished returns relative to the cost of smaller-than-expected stock sizes), and 
+strengthen the result shown here in which the BNP solution tends to avoid the 
+region of uncertainty.  Sufficiently convex or risk-prone functions could lead
+the SDP to attempt higher exploitation rates despite the uncertainty.  
+Understanding the relative roles of such functions would be a promising direction
+for future investigation.
+
+### The role of the prior
+
+Lastly, it should be noted that outside the data, the NBP reverts to the prior, 
+and consequently the choice of the prior can also play a significant role in 
+determining the optimal policy inferred by the SDP.  In the examples shown
+here we have selected a prior that is both relatively uninformative (due to the 
+broad priors placed on its parameters $\ell$ and $\sigma$ and simple (mean zero,
+radial basis function kernel).  In practice, both the choice of mean and the 
+choice of the covariance function may be chosen to confer particular biological
+properties, as well as more biologically informed priors for $\ell$ and $\sigma$.  
+In principle, this may allow a manager to improve the performance of the BNP-SDP
+approach by adding only enough additional detail as is justified.  For instance,
+it would be possible to use a linear or a Ricker-shaped mean in the prior without
+making the much stronger assumption that the Ricker is the structurally correct model. 
+However, this influence raises challenges as well.  For instance, in choosing a 
+trivial mean-zero prior, we bias the dynamics to transitions to 0 in step $X_{t+1}$ 
+from any stock size $X_t$ in the prior year, in the absence of any other data.  
+Future research must make sure that both the prior and the value function are 
+chosen appropriately for the problem at hand.
 

### 2013-11-14 inst/doc/nonparametric-bayes.Rmd
tweaking

@@
 Decision making under uncertainty is a ubiquitous challenge of natural 
-resource management and conservation.
-
-The sudden collapse of fisheries and other ecosystems is an increasingly
-widespread phenomenon and a pressing concern for ecological management
-and conservation.  Ecological dynamics are frequently complex and difficult to 
-measure, making uncertainty in our understanding a prediction a persistent
-challenge to effective management. 
-Decision-theoretic approaches provide a framework to determine the best 
-sequence of actions in face of uncertainty, but only when that uncertainty
-can be meaningfully quantified [@Fischer2009].
+resource management and conservation.  Ecological dynamics are
+frequently complex and difficult to measure, making uncertainty in our
+understanding a prediction a persistent challenge to effective management.
+Decision-theoretic approaches provide a framework to determine the
+best sequence of actions in face of uncertainty, but only when that
+uncertainty can be meaningfully quantified [@Fischer2009].  The sudden
+collapse of fisheries and other ecosystems has increasingly emphasized
+the difficulties of formulating even qualitatively correct models of
+the underlying processes.
 
@@
 
-Uncertainty enters the decision-making process at many levels: intrinsic 
-stochasticity in biological processes, measurements, and implementation of 
-policy [_e.g._ @Reed1979; @Clark1986; @Roughgarden1997; @Sethi2005], parameteric 
-uncertainty [_e.g._ @Ludwig1982; @Hilborn1997; @McAllister1998; @Schapaugh2013],
-and model or structural uncertainty [_e.g._ @Williams2001; @Cressie2009;  @Athanassoglou2012].
-Of these, structural uncertainty incorporates the least a priori knowledge or 
-assumptions and is generally the hardest to quantify. Typical approaches assume 
-a weak notion of uncertainty where a correct or reasonable approximation of the dynamics
-must be identified from among a handful of alternative models.  Here we consider an approach
-that addresses uncertainty at each of these levels without assuming the dynamics follow
-a particular (i.e. parametric) structure.
+Uncertainty enters the decision-making process at many levels:
+intrinsic stochasticity in biological processes, measurements, and
+implementation of policy [_e.g._ @Reed1979; @Clark1986; @Roughgarden1997;
+@Sethi2005], parameteric uncertainty [_e.g._ @Ludwig1982; @Hilborn1997;
+@McAllister1998; @Schapaugh2013], and model or structural uncertainty
+[_e.g._ @Williams2001; @Cressie2009;  @Athanassoglou2012].  Of these,
+structural uncertainty incorporates the least a priori knowledge or
+assumptions and is generally the hardest to quantify. Typical approaches
+assume a weak notion of model uncertainty in which the correct model
+(or reasonable approximation) of the dynamics must be identified from
+among a handful of alternative models.  Here we consider an approach
+that addresses uncertainty at each of these levels without assuming the
+dynamics follow a particular (i.e. parametric) structure.
 
@@
 <!-- What do we call this?  Extrapolation uncertainty?  Pathological Uncertainty? -->
+
+Model uncertainty is particularly insidious when model predictions
+must be made outside of the range of data on which the model was estimated.
 
@@
 
-<!-- Integrate this into the paragraph, rather than as a footnote?? Make more consise? -->
-
-[^1]: The concept of adaptive probing is one area that has explicitly
-addressed this kind of uncertainty, reaching rather opposite conclusions
-than what we observe here. Adaptive probing strategies follow from "Dual
-Control" or "Active Adaptive Management" approaches (e.g. @Ludwig1982)
-that can trade off short term utility by choosing actions that can reduce
-uncertainty.  Adaptive probing strategies arise when it is valuable to
-intentionally force a system far from the observed values even when the
-expected value such actions is low, as it provides much faster learning
-and consequent reduction of model uncertainty that can allow greater value
-to be derived later on.  For instance, @Ludwig1982 show that it may be
-advantageous to fish an unexploited population very heavily at first
-to obtain a better estimate of the recruitment rate.  This intuitive
-strategy when a population is governed by a Ricker or Beverton-Holt-like
-dynamic would clearly be disastrous if instead the dynamics contained
-an unforeseen tipping point.  The best way to learn where the edge lies
-may be to walk up to it, but it is also the most dangerous.
-
 <!-- Does this belong?  Improve transitions -->
@@
 
+<!--
 _Map_
+-->
 
@@
 
-<!-- More Consise versions are now above 
-- Parametric uncertainty
+<!-- More consise versions are now above 
+#### Parametric uncertainty
 
@@
 
-- Structural uncertainty
+#### Structural uncertainty
 
@@
 
-- Limits of state space
+#### Limits of state space
 
@@
 
+<!--
 -  _Note on "not magic"_: honest uncertainty + SDP
+-->
 
@@
 
-<!--move to  much later -->
+<!--move to  much later *Or do we need this at all?* -->
 
 - _Note on comparing models_ (via value function rather than by "fit").
-
-_Do we need this?_
 
@@
 
+<!-- 
 - Background on non-parametric modeling.  
+-->
 
@@
 
+<!--
 - Definition
-
 - Previous application
+-->
 
@@
 
+<!--
 - Why it is particularly suited to these two problems
 - (Why this is a novel application thereof)
+-->
 
@@
 
+<!-- skip this oveview? 
 ### Summary of approach
+-->
 
@@
 
+<!-- FIXME -->
+
 - Underlying model 
 - Available data
+
+<!--
 - Value function
+-->
 
@@
 
+<!-- 
 - Statement of the models
+-->
 
@@
 
-- Bayesian inference of parametric models
+<!-- Bayesian inference of parametric models -->
 
@@
 
-- SDP via parametric models
+<!-- SDP via parametric models -->
 
@@
 
-- Statement of model
+<!-- Statement of model -->
 
-... more on GP ... [Munch 2005]()
+<!-- FIXME ... say much more about the GP, as per Munch 2005a   -->
+<!-- See much longer discusion above on "Background on GP".  Move some of that to here?  -->
 
@@
 the characteristic length-scale over which correlation between two 
-points (e.g. any two points $X_t, X_{t+1}$, and $X_{t+\tau}, X_{t+1+\tau}$)
+points 
 in state-space decays, and $\sigma$, which gives the scale of the 
@@
 predicted values $X_{t+1}$ given an observation of the previous state,
-$X_t$.
+$X_t$. @Munch2005a gives an accessible introduction to the use of 
+Gaussian Processes in providing a Bayesian nonparametric description
+of the stock-recruitment relationship.
 
-- Inference of the model
+<!-- Inference of the BNP model -->
 
-Also unlike parametric models, this posterior distribution is still
+In contrast to the parametric models, this posterior distribution is still
 conditional on the training data. As such, the uncertainty near the
@@
 
-- SDP via the model
+<!-- SDP via the model -->
 
@@
 
-- All models are "good fits" to the originally observed data. 
-- (Simple model choice immediately leads us astray)
-- Understanding the role of the prior
-
-<!--
-
+In any modeling effort, models must be chosen for the task at hand.  
 Though simple mechanistically motivated models offer the greatest potential 
@@
 quantitative decision making framework.  
--->
+In this paper we have tackled two aspects of uncertainty that are both common
+to many ecological desicion-making problems and fundamentally challenging to
+existing approaches which largely rely on parametric models:
 
@@
 
-These aspects are common to many conservation decision making problems, which thus merit
-greater use of non-parametric approaches that can best take advantage of them.
+We have illustrated how the use of non-parametric approaches can provide
+more reliable solutions in the sequential decision-making problem.
 
@@
 management phase, in which the additional observations of the stock size could
-potentially improve parameter estimates.  While we intend to address this 
-possibility in future work in the context of these non-parametric models,
-we have not addressed it here for pedagogical reasons. In the context presented
-here, it is clear that the differences in performance arise from differences
-in the uncertainty inherent in the model formulations, rather than from 
-differing abilities to learn.  Because we consider a threshold system, 
-online learning would not change this generic feature of a lack of data in a
-certain range of the state space which is better captured by the Gaussian process.
+potentially improve parameter estimates.  Of particular interest in the 
+context of the extreme uncertainty considered here is the notion of 
+"active adaptive management" or "adaptive probing", which may actively
+seek to reduce uncertainty.  
+The concept of adaptive probing is one area that has explicitly
+addressed the extrapolation uncertainty addressed here, though 
+typically without the additional issue of model uncertainty.  As a 
+result, adaptive probing strategies suggest rather opposite conclusions
+than what we observe here. Such adaptive probing or Dual
+Control (e.g. @Ludwig1982) approaches 
+trade off short term utility by choosing actions that can reduce
+uncertainty.  Adaptive probing strategies arise when it is valuable to
+intentionally force a system far from the observed values even when the
+expected value such actions is low, as it provides much faster learning
+and consequent reduction of model uncertainty that can allow greater value
+to be derived later on.  For instance, @Ludwig1982 show that it may be
+advantageous to fish an unexploited population very heavily at first
+to obtain a better estimate of the recruitment rate.  This intuitive
+strategy when a population is governed by a Ricker or Beverton-Holt-like
+dynamic would clearly be disastrous if instead the dynamics contained
+an unforeseen tipping point.  The best way to learn where the edge lies
+may be to walk up to it, but it is also the most dangerous.  Future work
+should attempt to understand when such active adaptive learning is valuable,
+and when it will increase the risk of an irreversible transition.
 

### 2013-11-18 inst/doc/nonparametric-bayes.Rmd
writing, fixing references, figure tweaks

@@
 
-Decision making under uncertainty is a ubiquitous challenge of natural 
-resource management and conservation.  Ecological dynamics are
+Decision making under uncertainty is a ubiquitous challenge of
+natural resource management and conservation.  Ecological dynamics are
 frequently complex and difficult to measure, making uncertainty in our
@@
 intrinsic stochasticity in biological processes, measurements, and
-implementation of policy [_e.g._ @Reed1979; @Clark1986; @Roughgarden1997;
+implementation of policy [_e.g._ @Reed1979; @Clark1986; @Roughgarden1996;
 @Sethi2005], parameteric uncertainty [_e.g._ @Ludwig1982; @Hilborn1997;
@@
 Here we illustrate how a stochastic dynamic programming (SDP) algorithm
-[@Mangel1988; @Possingham1997, Marescot2013] can be driven by the
+[@Mangel1988; @Marescot2013] can be driven by the
 predictions from a Bayesian non-parametric (BNP) approach [@Munch2005a].
@@
 as the Dirchlet process) or the data points become themselves part of the model
-specification, as in the nonparametric regression used here.  we shall see here.
+specification, as in the nonparametric regression used here.
 
@@
 
-The essence of the GP approach can be captured in the
-following thought experiment: An exhaustive parametric approach to the
-challenge of structural uncertainty might proceed by writing down all
-possible functional forms for the underlying dynamical system with all
-possible parameter values for each form, and then consider searching
-over this huge space to select the most likely model and parameters;
-or using a Bayesian approach, assign priors to each of these possible
-models and infer the posterior distribution of possible models. The
-GP approach can be thought of as a computationally
-efficient approximation to this approach. GPs represent
-a large class of models that can be though of as capturing or reasonably
-approximating the set of models in this collection.  By modeling at the
-level of the process, rather than the level of parametric equation,
-we can more concisely capture the possible behavior of these curves.
-In place of a parametric model of the dynamical system, the GP 
-approach postulates a prior distribution of (n-dimensional)
+The essence of the GP approach can be captured in the following
+thought experiment: An exhaustive parametric approach to the challenge
+of structural uncertainty might proceed by writing down all possible
+functional forms for the underlying dynamical system with all possible
+parameter values for each form, and then consider searching over this
+huge space to select the most likely model and parameters; or using a
+Bayesian approach, assign priors to each of these possible models and
+infer the posterior distribution of possible models. The GP approach
+can be thought of as a computationally efficient approximation to this
+approach. GPs represent a large class of models that can be though
+of as capturing or reasonably approximating the set of models in this
+collection.  By modeling at the level of the process, rather than the
+level of parametric equation, we can more concisely capture the possible
+behavior of these curves.  In place of a parametric model of the dynamical
+system, the GP approach postulates a prior distribution of (n-dimensional)
 curves that can be though of as approximations to a range of possible
-(parametric) models that might describe the data. The GP allows us
-to consider probabilities on a large set of possible curves simultaneously.
+(parametric) models that might describe the data. The GP allows us to
+consider probabilities on a large set of possible curves simultaneously.
 
@@
 
-The mean inferred state space dynamics of each model
-relative to the true model used to generate the data is shown in Figure 2, 
-predicting the relationship between observed stock size (x-axis) to the stock size
-after recruitment the following year.  Note that in contrast to the  other models shown, the expected 
-Gaussian process corresponds to a distribution of curves - as indicated
-by the gray band - which itself has a mean shown in black. Parameter uncertainty
-(not shown) spreads out the estimates further.  
-The observed data from which each model is estimated is also shown.  The observations come
-from only a limited region of state space corresponding to unharvested
-or weakly harvested system.  No observations occur at the theoretical
-optimum harvest rate or near the tipping point.
+The mean inferred state space dynamics of each model relative to the
+true model used to generate the data is shown in Figure 2, predicting
+the relationship between observed stock size (x-axis) to the stock
+size after recruitment the following year.  Note that in contrast to
+the  other models shown, the expected Gaussian process corresponds to a
+distribution of curves - as indicated by the gray band - which itself
+has a mean shown in black. Parameter uncertainty (not shown) spreads
+out the estimates further.  The observed data from which each model is
+estimated is also shown.  The observations come from only a limited region
+of state space corresponding to unharvested or weakly harvested system.
+No observations occur at the theoretical optimum harvest rate or near
+the tipping point.
 
@@
 
-Despite the similarities in model fits to the observed data, the policies
-inferred under each model differ widely, as shown in Figure 3.  
-Policies are shown in terms of target
-escapement, $S_t$.  Under models such as this a constant escapement
-policy is expected to be optimal [@Reed1979], whereby population levels
-below a certain size $S$ are unharvested, while above that size the harvest
-strategy aims to return the population to $S$, resulting in the hockey-stick
-shaped policies shown.  Only the structurally correct model (Allen model) and the GP 
-produce policies close to the true optimum policy (where both the underlying 
+Despite the similarities in model fits to the observed data, the
+policies inferred under each model differ widely, as shown in Figure 3.
+Policies are shown in terms of target escapement, $S_t$.  Under models
+such as this a constant escapement policy is expected to be optimal
+[@Reed1979], whereby population levels below a certain size $S$ are
+unharvested, while above that size the harvest strategy aims to return the
+population to $S$, resulting in the hockey-stick shaped policies shown.
+Only the structurally correct model (Allen model) and the GP produce
+policies close to the true optimum policy (where both the underlying
 model structure and parameter values are known without error).
@@
 
-The consequences of managing 100 replicate realizations of the 
-simulated fishery under each of the policies estimated is shown in Figure 4.  As expected
-from the policy curves, the structurally correct model under-harvests,
-leaving the stock to vary around it's un-fished optimum.  The structurally
-incorrect Ricker model over-harvests the population passed
-the tipping point consistently, resulting in the immediate crash of the stock and 
-thus derives minimal profits.
+The consequences of managing 100 replicate realizations of the simulated
+fishery under each of the policies estimated is shown in Figure 4.
+As expected from the policy curves, the structurally correct model
+under-harvests, leaving the stock to vary around it's un-fished optimum.
+The structurally incorrect Ricker model over-harvests the population
+passed the tipping point consistently, resulting in the immediate crash
+of the stock and thus derives minimal profits.
 
-These results are robust across a range of stochastic realizations, models, and parameter values.  
-The results across this range can most easily be compared 
-by using the relative differences in net present value realized by each of the model,
-as shown in Figure 5.  The BNP-SDP approach most consistently realizes a value 
-close to the optimal solution, and importantly avoids ever driving the system across
-the tipping point, which results in the near-zero value cases in the parametric models.
+These results are robust across a range of stochastic realizations,
+models, and parameter values.  The results across this range can most
+easily be compared by using the relative differences in net present
+value realized by each of the model, as shown in Figure 5.  The BNP-SDP
+approach most consistently realizes a value close to the optimal solution,
+and importantly avoids ever driving the system across the tipping point,
+which results in the near-zero value cases in the parametric models.
 
@@
 
-### Risk-prone and Risk-adverse value functions
+### Risk-prone and risk-adverse value functions
 
@@
 chooses to explore or avoid the resulting region of uncertainty can also
-be influenced by the curvature of the value (profit) function $\Pi$. Both to simplify 
-the intution and avoid biasing this result, we have chosen a profits that
-are linear in the catch and thus neither risk-prone nor risk adverse.  Making
-this function concave, representing the typical assumption of diminishing returns,
-would make the SDP more risk-adverse (as larger-than-expected stock sizes offer
-dimished returns relative to the cost of smaller-than-expected stock sizes), and 
-strengthen the result shown here in which the BNP solution tends to avoid the 
-region of uncertainty.  Sufficiently convex or risk-prone functions could lead
+be influenced by the curvature of the value (profit) function $\Pi$. Both
+to simplify the intution and avoid biasing this result, we have chosen
+a profits that are linear in the catch and thus neither risk-prone nor
+risk adverse.  Making this function concave, representing the typical
+assumption of diminishing returns, would make the SDP more risk-adverse
+(as larger-than-expected stock sizes offer dimished returns relative
+to the cost of smaller-than-expected stock sizes), and strengthen the
+result shown here in which the BNP solution tends to avoid the region
+of uncertainty.  Sufficiently convex or risk-prone functions could lead
 the SDP to attempt higher exploitation rates despite the uncertainty.
-Understanding the relative roles of such functions would be a promising direction
-for future investigation.
+Understanding the relative roles of such functions would be a promising
+direction for future investigation.
 
@@
 
-Lastly, it should be noted that outside the data, the NBP reverts to the prior, 
-and consequently the choice of the prior can also play a significant role in 
-determining the optimal policy inferred by the SDP.  In the examples shown
-here we have selected a prior that is both relatively uninformative (due to the 
-broad priors placed on its parameters $\ell$ and $\sigma$ and simple (mean zero,
-radial basis function kernel).  In practice, both the choice of mean and the 
-choice of the covariance function may be chosen to confer particular biological
-properties, as well as more biologically informed priors for $\ell$ and $\sigma$.  
-In principle, this may allow a manager to improve the performance of the BNP-SDP
-approach by adding only enough additional detail as is justified.  For instance,
-it would be possible to use a linear or a Ricker-shaped mean in the prior without
-making the much stronger assumption that the Ricker is the structurally correct model. 
-However, this influence raises challenges as well.  For instance, in choosing a 
-trivial mean-zero prior, we bias the dynamics to transitions to 0 in step $X_{t+1}$ 
-from any stock size $X_t$ in the prior year, in the absence of any other data.  
-Future research must make sure that both the prior and the value function are 
-chosen appropriately for the problem at hand.
+Lastly, it should be noted that outside the data, the NBP reverts to
+the prior, and consequently the choice of the prior can also play a
+significant role in determining the optimal policy inferred by the
+SDP.  In the examples shown here we have selected a prior that is both
+relatively uninformative (due to the broad priors placed on its parameters
+$\ell$ and $\sigma$ and simple (mean zero, radial basis function kernel).
+In practice, both the choice of mean and the choice of the covariance
+function may be chosen to confer particular biological properties,
+as well as more biologically informed priors for $\ell$ and $\sigma$.
+In principle, this may allow a manager to improve the performance of the
+BNP-SDP approach by adding only enough additional detail as is justified.
+For instance, it would be possible to use a linear or a Ricker-shaped
+mean in the prior without making the much stronger assumption that the
+Ricker is the structurally correct model.  However, this influence raises
+challenges as well.  For instance, in choosing a trivial mean-zero prior,
+we bias the dynamics to transitions to 0 in step $X_{t+1}$ from any
+stock size $X_t$ in the prior year, in the absence of any other data.
+Future research must make sure that both the prior and the value function
+are chosen appropriately for the problem at hand.
 
@@
 
-In this simulated example, the underlying
-dynamics are truly governed by a simple parametric model, allowing
-the parametric approaches to be more accurate.  Similarly, because the
-dynamics are  one-dimensional dynamics and lead to  stable nodes (rather
-than other attractors such as limit-cycles resulting in oscillations),
-the training data provides relatively limited information about the
-dynamics.  For these reasons, we anticipate that in higher-dimensional
-examples characteristic of ecosystem management problems that the machine
-learning approach will prove even more valuable.
+In this simulated example, the underlying dynamics are truly governed
+by a simple parametric model, allowing the parametric approaches to be
+more accurate.  Similarly, because the dynamics are  one-dimensional
+dynamics and lead to  stable nodes (rather than other attractors such
+as limit-cycles resulting in oscillations), the training data provides
+relatively limited information about the dynamics.  For these reasons,
+we anticipate that in higher-dimensional examples characteristic of
+ecosystem management problems that the machine learning approach will
+prove even more valuable.
 
@@
 
-In our treatment here we have ignored the possibility of learning during the 
-management phase, in which the additional observations of the stock size could
-potentially improve parameter estimates.  Of particular interest in the 
-context of the extreme uncertainty considered here is the notion of 
-"active adaptive management" or "adaptive probing", which may actively
-seek to reduce uncertainty.  
-The concept of adaptive probing is one area that has explicitly
-addressed the extrapolation uncertainty addressed here, though 
-typically without the additional issue of model uncertainty.  As a 
-result, adaptive probing strategies suggest rather opposite conclusions
-than what we observe here. Such adaptive probing or Dual
-Control (e.g. @Ludwig1982) approaches 
-trade off short term utility by choosing actions that can reduce
-uncertainty.  Adaptive probing strategies arise when it is valuable to
-intentionally force a system far from the observed values even when the
-expected value such actions is low, as it provides much faster learning
-and consequent reduction of model uncertainty that can allow greater value
-to be derived later on.  For instance, @Ludwig1982 show that it may be
-advantageous to fish an unexploited population very heavily at first
-to obtain a better estimate of the recruitment rate.  This intuitive
-strategy when a population is governed by a Ricker or Beverton-Holt-like
-dynamic would clearly be disastrous if instead the dynamics contained
-an unforeseen tipping point.  The best way to learn where the edge lies
-may be to walk up to it, but it is also the most dangerous.  Future work
-should attempt to understand when such active adaptive learning is valuable,
+In our treatment here we have ignored the possibility of learning during
+the management phase, in which the additional observations of the stock
+size could potentially improve parameter estimates.  Of particular
+interest in the context of the extreme uncertainty considered here
+is the notion of "active adaptive management" or "adaptive probing",
+which may actively seek to reduce uncertainty.  The concept of adaptive
+probing is one area that has explicitly addressed the extrapolation
+uncertainty addressed here, though typically without the additional
+issue of model uncertainty.  As a result, adaptive probing strategies
+suggest rather opposite conclusions than what we observe here. Such
+adaptive probing or Dual Control (e.g. @Ludwig1982) approaches trade
+off short term utility by choosing actions that can reduce uncertainty.
+Adaptive probing strategies arise when it is valuable to intentionally
+force a system far from the observed values even when the expected value
+such actions is low, as it provides much faster learning and consequent
+reduction of model uncertainty that can allow greater value to be derived
+later on.  For instance, @Ludwig1982 show that it may be advantageous
+to fish an unexploited population very heavily at first to obtain a
+better estimate of the recruitment rate.  This intuitive strategy when a
+population is governed by a Ricker or Beverton-Holt-like dynamic would
+clearly be disastrous if instead the dynamics contained an unforeseen
+tipping point.  The best way to learn where the edge lies may be to
+walk up to it, but it is also the most dangerous.  Future work should
+attempt to understand when such active adaptive learning is valuable,
 and when it will increase the risk of an irreversible transition.
@@
 
-```{r, echo=FALSE, include=TRUE, fig.height=6}
+```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimate of the Ricker model", fig.width=8}
 plot_ricker_traces
@@
 
-```{r, echo=FALSE, include=TRUE}
-plot_ricker_posteriors
+```{r, echo=FALSE, include=TRUE, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=8}
+ggplot(ricker_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+  facet_wrap(~ variable, scale="free", ncol=2)
 ```
@@
 
-```{r, echo=FALSE, include=TRUE, fig.height=6}
+```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimate of the Myers model", fig.width=8}
 plot_myers_traces
@@
 
-```{r, echo=FALSE, include=TRUE}
-plot_myers_posteriors
+```{r, echo=FALSE, include=TRUE, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=8}
+ggplot(myers_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+  facet_wrap(~ variable, scale="free", ncol=3)
 ```
@@
 
-```{r, echo=FALSE, include=TRUE, fig.height=6}
+```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimate of the Allen model", fig.width=8}
 plot_allen_traces
@@
 
-```{r, echo=FALSE, include=TRUE}
-plot_allen_posteriors
+```{r, echo=FALSE, include=TRUE, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=8}
+ggplot(allen_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+  facet_wrap(~ variable, scale="free", ncol=3)
 ```

### 2013-11-18 inst/doc/nonparametric-bayes.tex
writing, fixing references, figure tweaks

@@
 Decision making under uncertainty is a ubiquitous challenge of natural
-resource management and conservation.
-
-The sudden collapse of fisheries and other ecosystems is an increasingly
-widespread phenomenon and a pressing concern for ecological management
-and conservation. Ecological dynamics are frequently complex and
-difficult to measure, making uncertainty in our understanding a
-prediction a persistent challenge to effective management.
-Decision-theoretic approaches provide a framework to determine the best
-sequence of actions in face of uncertainty, but only when that
-uncertainty can be meaningfully quantified ({\textbf{???}}).
+resource management and conservation. Ecological dynamics are frequently
+complex and difficult to measure, making uncertainty in our
+understanding a prediction a persistent challenge to effective
+management. Decision-theoretic approaches provide a framework to
+determine the best sequence of actions in face of uncertainty, but only
+when that uncertainty can be meaningfully quantified (Fischer et al.
+2009). The sudden collapse of fisheries and other ecosystems has
+increasingly emphasized the difficulties of formulating even
+qualitatively correct models of the underlying processes.
 
@@
 of policy (\emph{e.g.} Reed 1979; C. W. Clark and Kirkwood 1986;
-{\textbf{???}}; Sethi et al. 2005), parameteric uncertainty (\emph{e.g.}
-Ludwig and Walters 1982; {\textbf{???}}; {\textbf{???}}; Schapaugh and
-Tyre 2013), and model or structural uncertainty (\emph{e.g.} B. K.
-Williams 2001; Cressie et al. 2009; Athanassoglou and Xepapadeas 2012).
-Of these, structural uncertainty incorporates the least a priori
-knowledge or assumptions and is generally the hardest to quantify.
-Typical approaches assume a weak notion of uncertainty where a correct
-or reasonable approximation of the dynamics must be identified from
-among a handful of alternative models. Here we consider an approach that
-addresses uncertainty at each of these levels without assuming the
-dynamics follow a particular (i.e.~parametric) structure.
+Roughgarden and Smith 1996; Sethi et al. 2005), parameteric uncertainty
+(\emph{e.g.} Ludwig and Walters 1982; Hilborn and Mangel 1997;
+McAllister 1998; Schapaugh and Tyre 2013), and model or structural
+uncertainty (\emph{e.g.} B. K. Williams 2001; Cressie et al. 2009;
+Athanassoglou and Xepapadeas 2012). Of these, structural uncertainty
+incorporates the least a priori knowledge or assumptions and is
+generally the hardest to quantify. Typical approaches assume a weak
+notion of model uncertainty in which the correct model (or reasonable
+approximation) of the dynamics must be identified from among a handful
+of alternative models. Here we consider an approach that addresses
+uncertainty at each of these levels without assuming the dynamics follow
+a particular (i.e.~parametric) structure.
+
+Model uncertainty is particularly insidious when model predictions must
+be made outside of the range of data on which the model was estimated.
 
 An additional source of uncertainty that has recieved less
-attention\footnote{The concept of adaptive probing is one area that has
-  explicitly addressed this kind of uncertainty, reaching rather
-  opposite conclusions than what we observe here. Adaptive probing
-  strategies follow from ``Dual Control'' or ``Active Adaptive
-  Management'' approaches (e.g. Ludwig and Walters (1982)) that can
-  trade off short term utility by choosing actions that can reduce
-  uncertainty. Adaptive probing strategies arise when it is valuable to
-  intentionally force a system far from the observed values even when
-  the expected value such actions is low, as it provides much faster
-  learning and consequent reduction of model uncertainty that can allow
-  greater value to be derived later on. For instance, Ludwig and Walters
-  (1982) show that it may be advantageous to fish an unexploited
-  population very heavily at first to obtain a better estimate of the
-  recruitment rate. This intuitive strategy when a population is
-  governed by a Ricker or Beverton-Holt-like dynamic would clearly be
-  disastrous if instead the dynamics contained an unforeseen tipping
-  point. The best way to learn where the edge lies may be to walk up to
-  it, but it is also the most dangerous.} arises when applying a
-dynamical model outside the range of data on which it has been
-estimated. This extrapolation uncertainty is felt most keenly in
-decision-theoretic applications, as (a) exploring the potential action
-space typically involves considering actions that may move the system
-outside the range of observed behavior, and (b) decision-theoretic
-alogrithms rely not only on reasonable estimates of the expected
-outcomes, but depend on the weights given to all possible outcomes
-(\emph{e.g.} Weitzman 2013).
+attention{[}\^{}1{]} arises when applying a dynamical model outside the
+range of data on which it has been estimated. This extrapolation
+uncertainty is felt most keenly in decision-theoretic applications, as
+(a) exploring the potential action space typically involves considering
+actions that may move the system outside the range of observed behavior,
+and (b) decision-theoretic alogrithms rely not only on reasonable
+estimates of the expected outcomes, but depend on the weights given to
+all possible outcomes (\emph{e.g.} Weitzman 2013).
 
@@
 
-\emph{Map}
-
 Here we illustrate how a stochastic dynamic programming (SDP) algorithm
-(Mangel and Clark 1988; {\textbf{???}}) can be driven by the predictions
-from a Bayesian non-parametric (BNP) approach (Munch, Kottas, and Mangel
-2005). This provides two distinct advantages compared with contemporary
-approaches. First, using a BNP sidesteps the need for an accurate
-model-based description of the system dynamics. Second, the BNP can
-better reflect uncertainty that arises when extrapolating a model
-outside of the data on which it was fit. We illustrate that when the
-correct model is not known, this latter feature is crucial to providing
-a robust decision-theoretic approach in face of substantial structural
-uncertainty.
+(Mangel and Clark 1988; Marescot et al. 2013) can be driven by the
+predictions from a Bayesian non-parametric (BNP) approach (Munch,
+Kottas, and Mangel 2005). This provides two distinct advantages compared
+with contemporary approaches. First, using a BNP sidesteps the need for
+an accurate model-based description of the system dynamics. Second, the
+BNP can better reflect uncertainty that arises when extrapolating a
+model outside of the data on which it was fit. We illustrate that when
+the correct model is not known, this latter feature is crucial to
+providing a robust decision-theoretic approach in face of substantial
+structural uncertainty.
 
@@
 data and without knowledge of the correct model structure.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \emph{Note on ``not magic''}: honest uncertainty + SDP
-\end{itemize}
 
@@
 The economic value and ecological concern have made marine fisheries the
-crucible for much of the founding work (Gordon 1954; Reed 1979; May et
-al. 1979; Ludwig and Walters 1982) in managing ecosystems under
+crucible for much of the founding work (H. S. Gordon 1954; Reed 1979;
+May et al. 1979; Ludwig and Walters 1982) in managing ecosystems under
 uncertainty. Global trends (Worm et al. 2006) and controversy (Hilborn
@@
 \end{itemize}
-
-\emph{Do we need this?}
 
@@
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Background on non-parametric modeling.
-\end{itemize}
-
 Addressing the difficulty posed by extrapolation without knowing the
@@
 
-\begin{itemize}
-\item
-  Definition
-\item
-  Previous application
-\end{itemize}
-
 The use of Gaussian process (GP) regression (or ``kriging'' in the
@@
 and Williams (2006).
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Why it is particularly suited to these two problems
-\item
-  (Why this is a novel application thereof)
-\end{itemize}
 
@@
 
-\subsubsection{Summary of approach}\label{summary-of-approach}
-
 \subsubsection{Statement of the optimal control
@@
   Available data
-\item
-  Value function
 \end{itemize}
@@
 \subsubsection{Parametric models}\label{parametric-models}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Statement of the models
-\end{itemize}
 
@@
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Bayesian inference of parametric models
-\end{itemize}
-
 We infer posterior distributions for the parameters of each model in a
-Bayesian context using Gibbs sampling (implemented in R ({\textbf{???}})
-using jags, ({\textbf{???}})). We choose uninformative uniform priors
-for all parameters (See Appendix, Figures S1-S3, and Table S1, and the R
-code provided). One-step-ahead predictions of these model fits are shown
-in Figure 1.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  SDP via parametric models
-\end{itemize}
+Bayesian context using Gibbs sampling (implemented in R (Team 2013)
+using jags, (Su and Yajima 2013)). We choose uninformative uniform
+priors for all parameters (See Appendix, Figures S1-S3, and Table S1,
+and the R code provided). One-step-ahead predictions of these model fits
+are shown in Figure 1.
 
@@
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Statement of model
-\end{itemize}
-
-\ldots{} more on GP \ldots{} \href{}{Munch 2005}
-
 We also estimate a simple Gaussian Process defined by a radial basis
 function kernel of two parameters: $\ell$, which gives the
-characteristic length-scale over which correlation between two points
-(e.g.~any two points $X_t, X_{t+1}$, and $X_{t+\tau}, X_{t+1+\tau}$) in
+characteristic length-scale over which correlation between two points in
 state-space decays, and $\sigma$, which gives the scale of the process
@@
 values $X_{t+1}$ given an observation of the previous state, $X_t$.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Inference of the model
-\end{itemize}
-
-Also unlike parametric models, this posterior distribution is still
-conditional on the training data. As such, the uncertainty near the
-observed data.
+Munch, Kottas, and Mangel (2005) gives an accessible introduction to the
+use of Gaussian Processes in providing a Bayesian nonparametric
+description of the stock-recruitment relationship.
+
+In contrast to the parametric models, this posterior distribution is
+still conditional on the training data. As such, the uncertainty near
+the observed data.
 
@@
 the posterior comes from the data rather than the prior belief.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  SDP via the model
-\end{itemize}
 
@@
 
+\% latex table generated in R 3.0.2 by xtable 1.7-1 package \% Thu Nov
+14 14:10:28 2013
+
 \begin{table}[ht]
-\begin{center}
+\centering
 \begin{tabular}{rrrr}
@@
 \end{tabular}
-\end{center}
 \end{table}
@@
 has a mean shown in black. Parameter uncertainty (not shown) spreads out
-the estimates further.\\The observed data from which each model is
+the estimates further. The observed data from which each model is
 estimated is also shown. The observations come from only a limited
@@
 Despite the similarities in model fits to the observed data, the
-policies inferred under each model differ widely, as shown in Figure
-3.\\Policies are shown in terms of target escapement, $S_t$. Under
-models such as this a constant escapement policy is expected to be
-optimal (Reed 1979), whereby population levels below a certain size $S$
-are unharvested, while above that size the harvest strategy aims to
-return the population to $S$, resulting in the hockey-stick shaped
-policies shown. Only the structurally correct model (Allen model) and
-the GP produce policies close to the true optimum policy (where both the
+policies inferred under each model differ widely, as shown in Figure 3.
+Policies are shown in terms of target escapement, $S_t$. Under models
+such as this a constant escapement policy is expected to be optimal
+(Reed 1979), whereby population levels below a certain size $S$ are
+unharvested, while above that size the harvest strategy aims to return
+the population to $S$, resulting in the hockey-stick shaped policies
+shown. Only the structurally correct model (Allen model) and the GP
+produce policies close to the true optimum policy (where both the
 underlying model structure and parameter values are known without
@@
 These results are robust across a range of stochastic realizations,
-models, and parameter values.\\The results across this range can most
+models, and parameter values. The results across this range can most
 easily be compared by using the relative differences in net present
@@
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  All models are ``good fits'' to the originally observed data.
-\item
-  (Simple model choice immediately leads us astray)
-\end{itemize}
+In any modeling effort, models must be chosen for the task at
+hand.\\Though simple mechanistically motivated models offer the greatest
+potential to increase our basic understanding of ecological processes
+(Cuddington, Fortin, and Gerber 2013; Geritz and Kisdi 2012), such
+models can be not only inaccurate but misleading when relied upon in a
+quantitative decision making framework.\\In this paper we have tackled
+two aspects of uncertainty that are both common to many ecological
+desicion-making problems and fundamentally challenging to existing
+approaches which largely rely on parametric models:
 
@@
 
-These aspects are common to many conservation decision making problems,
-which thus merit greater use of non-parametric approaches that can best
-take advantage of them.
+We have illustrated how the use of non-parametric approaches can provide
+more reliable solutions in the sequential decision-making problem.
 
@@
 
+\subsubsection{Risk-prone and risk-adverse value
+functions}\label{risk-prone-and-risk-adverse-value-functions}
+
+The degree to which the decision-making part of the algorithm (the SDP)
+chooses to explore or avoid the resulting region of uncertainty can also
+be influenced by the curvature of the value (profit) function $\Pi$.
+Both to simplify the intution and avoid biasing this result, we have
+chosen a profits that are linear in the catch and thus neither
+risk-prone nor risk adverse. Making this function concave, representing
+the typical assumption of diminishing returns, would make the SDP more
+risk-adverse (as larger-than-expected stock sizes offer dimished returns
+relative to the cost of smaller-than-expected stock sizes), and
+strengthen the result shown here in which the BNP solution tends to
+avoid the region of uncertainty. Sufficiently convex or risk-prone
+functions could lead the SDP to attempt higher exploitation rates
+despite the uncertainty. Understanding the relative roles of such
+functions would be a promising direction for future investigation.
+
+\subsubsection{The role of the prior}\label{the-role-of-the-prior}
+
+Lastly, it should be noted that outside the data, the NBP reverts to the
+prior, and consequently the choice of the prior can also play a
+significant role in determining the optimal policy inferred by the SDP.
+In the examples shown here we have selected a prior that is both
+relatively uninformative (due to the broad priors placed on its
+parameters $\ell$ and $\sigma$ and simple (mean zero, radial basis
+function kernel). In practice, both the choice of mean and the choice of
+the covariance function may be chosen to confer particular biological
+properties, as well as more biologically informed priors for $\ell$ and
+$\sigma$. In principle, this may allow a manager to improve the
+performance of the BNP-SDP approach by adding only enough additional
+detail as is justified. For instance, it would be possible to use a
+linear or a Ricker-shaped mean in the prior without making the much
+stronger assumption that the Ricker is the structurally correct model.
+However, this influence raises challenges as well. For instance, in
+choosing a trivial mean-zero prior, we bias the dynamics to transitions
+to 0 in step $X_{t+1}$ from any stock size $X_t$ in the prior year, in
+the absence of any other data. Future research must make sure that both
+the prior and the value function are chosen appropriately for the
+problem at hand.
+
 \subsection{Future directions}\label{future-directions}
@@
 the management phase, in which the additional observations of the stock
-size could potentially improve parameter estimates. While we intend to
-address this possibility in future work in the context of these
-non-parametric models, we have not addressed it here for pedagogical
-reasons. In the context presented here, it is clear that the differences
-in performance arise from differences in the uncertainty inherent in the
-model formulations, rather than from differing abilities to learn.
-Because we consider a threshold system, online learning would not change
-this generic feature of a lack of data in a certain range of the state
-space which is better captured by the Gaussian process.
+size could potentially improve parameter estimates. Of particular
+interest in the context of the extreme uncertainty considered here is
+the notion of ``active adaptive management'' or ``adaptive probing'',
+which may actively seek to reduce uncertainty. The concept of adaptive
+probing is one area that has explicitly addressed the extrapolation
+uncertainty addressed here, though typically without the additional
+issue of model uncertainty. As a result, adaptive probing strategies
+suggest rather opposite conclusions than what we observe here. Such
+adaptive probing or Dual Control (e.g. Ludwig and Walters (1982))
+approaches trade off short term utility by choosing actions that can
+reduce uncertainty. Adaptive probing strategies arise when it is
+valuable to intentionally force a system far from the observed values
+even when the expected value such actions is low, as it provides much
+faster learning and consequent reduction of model uncertainty that can
+allow greater value to be derived later on. For instance, Ludwig and
+Walters (1982) show that it may be advantageous to fish an unexploited
+population very heavily at first to obtain a better estimate of the
+recruitment rate. This intuitive strategy when a population is governed
+by a Ricker or Beverton-Holt-like dynamic would clearly be disastrous if
+instead the dynamics contained an unforeseen tipping point. The best way
+to learn where the edge lies may be to walk up to it, but it is also the
+most dangerous. Future work should attempt to understand when such
+active adaptive learning is valuable, and when it will increase the risk
+of an irreversible transition.
 
@@
 \centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-1.pdf}
-\caption{plot of chunk unnamed-chunk-1}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
 \includegraphics{figure/nonparametric-bayes-unnamed-chunk-2.pdf}
-\caption{plot of chunk unnamed-chunk-2}
+\caption{Posteriors from the MCMC estimate of the Ricker model}
 \end{figure}
@@
 
+\% latex table generated in R 3.0.2 by xtable 1.7-1 package \% Thu Nov
+14 00:55:33 2013
+
 \begin{table}[ht]
-\begin{center}
+\centering
 \begin{tabular}{rlrr}
@@
 \includegraphics{figure/nonparametric-bayes-unnamed-chunk-4.pdf}
-\caption{plot of chunk unnamed-chunk-4}
+\caption{Traces from the MCMC estimate of the Myers model}
 \end{figure}
@@
 \includegraphics{figure/nonparametric-bayes-unnamed-chunk-5.pdf}
-\caption{plot of chunk unnamed-chunk-5}
-\end{figure}
-
-Table S2: Parameterization of the priors
+\caption{Posterior distributions from the MCMC estimates of the Myers
+model}
+\end{figure}
+
+Table S2: Parameterization of the priors \% latex table generated in R
+3.0.2 by xtable 1.7-1 package \% Thu Nov 14 00:55:36 2013
 
 \begin{table}[ht]
-\begin{center}
+\centering
 \begin{tabular}{rlrr}
@@
 \includegraphics{figure/nonparametric-bayes-unnamed-chunk-7.pdf}
-\caption{plot of chunk unnamed-chunk-7}
+\caption{Traces from the MCMC estimate of the Allen model}
 \end{figure}
@@
 \includegraphics{figure/nonparametric-bayes-unnamed-chunk-8.pdf}
-\caption{plot of chunk unnamed-chunk-8}
+\caption{Posteriors from the MCMC estimate of the Allen model}
 \end{figure}
@@
 
+\% latex table generated in R 3.0.2 by xtable 1.7-1 package \% Thu Nov
+14 00:55:40 2013
+
 \begin{table}[ht]
-\begin{center}
+\centering
 \begin{tabular}{rlrr}
@@
 
-\begin{center}\rule{3in}{0.4pt}\end{center}
-
-\section{Abstract}\label{abstract-1}
-
-Decision-theoretic methods often rely on simple parametric models of
-ecological dynamics to compare the value of a potential sequence of
-actions. Unfortunately, such simple models rarely capture the complexity
-or uncertainty found in most real ecosystems.
-
-Further, the data on which a model has been parameterized frequently
-fails to cover the possible state-space over which management decisions
-must operate. Consequently a model do well in the region of state-space
-in which it was estimated, but give erroneous confidence to predictions
-outside of that region.
-
-This problem is keenly felt in any system where a potential threshold or
-tipping point is a concern. Such a tipping point, if it exists at all,
-will lay outside the observed range of the observed data.
-
-We demonstrate how nonparametric Bayesian models can provide robust,
-solutions to decision making under uncertainty without knowing the
-structural form of the true model.
-
-While methods that account for \emph{parametric} uncertainty can be very
-successful with the right model, structural uncertainty of not knowing
-what model best approximates the dynamics poses considerably greater
-difficulty.
-
-\section{Introduction}\label{introduction-1}
-
-\paragraph{Opening}\label{opening}
-
-\paragraph{Models for decision-making under
-uncertainty}\label{models-for-decision-making-under-uncertainty}
-
-Decision-theoretic or optimal control tools require a model that can
-assign probabilities of future states (e.g.~stock size of a fishery)
-given the current state and a proposed action (e.g.~fishing harvest or
-effort).\\Management frequently faces a sequential decision-making
-problem -- after selecting an action, the decision-maker may receive new
-information about the current state and must again choose an appropriate
-action -- such as setting the harvest limits each year based on stock
-assessments the year prior.
-
-The decision maker typically seeks to determining the course of actions
-(also referred to as the policy) that maximizes the expected value of
-some objective function such as net present value derived from the
-resource over time.\\Though much can be said on how to choose this value
-function appropriately (e.g.~see (Halpern et al.
-2013)(http://doi.org/10.1073/pnas.1217689110 ``Achieving the triple
-bottom line in the face of inherent trade-offs among social equity,
-economic return, and conservation.'')) we will assume this is given.
-(Nor is this approach necessarily constrained to maximizing the
-expectated value of such a function - the decision-theoretic framework
-can be adapted to alternatives such as minimizing the maximum cost or
-damage that might be incurred; see Polasky et al. (2011)).
-
-In representing future states with probabilities and maximizing
-expectations, this approach provides a natural framework for handling
-uncertainty.
-
-The value function typically depends on the action or policy taken, as
-well as the state of the system, in each interval of time. The state of
-the system, in turn, is usually described by a dynamical model.
-
-(B. K. Williams 2001; Athanassoglou and Xepapadeas 2012).
-
-While simple mechanistic models can nevertheless provide important
-insights into long-term outcomes, such approaches are not well-suited
-for use in forecasting outcomes of potential management options.
-Non-parametric approaches offer a more flexible alternative that can
-both more accurately reflect the data available while also representing
-greater uncertainty in areas (of state-space) where data is lacking.
-
-We demonstrate how a Gaussian Process model of stock recruitment can
-lead to nearly optimal management through stochastic dynamic
-programming, comperable to knowing the correct structural equation for
-the underlying simulation. Meanwhile, parametric models that do not
-match the underlying dynamics can perform very poorly, even though they
-fit the data as well as the true model.\\Ecological research and
-management strategy should pay closer attention to the opportunities and
-challenges nonparametric modeling can offer.
-
-\section{Approach and Methods}\label{approach-and-methods-1}
-
-\subsection{The optimal control problem in fisheries
-management}\label{the-optimal-control-problem-in-fisheries-management}
-
-We focus on the problem in which a manager must set the harvest level
-for a marine fishery each year to maximize the net present value of the
-resource, given an estimated stock size from the year before.
-
-To permit comparisons against a theoretical optimum we will consider
-data on the stock dynamics simulated from a simple parametric model in
-which recruitment of the fish stock $X_{t+1}$ in the following year is a
-stochastic process governed by a function $f$ of the current stock
-$X_t$, selected harvest policy $h_t$, and noise process $Z$,
-
-\[X_{t+1} = Z_t f(X_t, h_t) \]
-
-Given parameters for the function $f$ and probability distribution $Z$,
-along with a given economic model determining the price/profit
-$\Pi(X_t, h_t)$ realized in a given year given a choice of harvest $h_t$
-and observed stock $X_t$. This problem can be solved exactly for
-discretized values of stock $X$ and policy $h$ using stochastic dynamic
-programming (SDP) (Mangel and Clark 1988). Problems of this sort
-underpin much marine fisheries management today.
-
-A crux of this approach is correctly specifying the functional form of
-$f$, along with its parameters. The standard approach uses one of a
-handful of common parametric models representing the stock-recruitment
-relationship, usually after estimating the model parameters from any
-available existing data. Uncertainty in the parameter estimates can be
-estimated and integrated over to determine the optimal policy under
-under uncertainty (Mangel and Clark 1988; Schapaugh and Tyre 2013).
-Uncertainty in the model structure itself can only be addressed in this
-approach by hypothesizing alternative model structures, and then
-performing some model choice or model averaging (B. K. Williams 2001;
-Athanassoglou and Xepapadeas 2012).
-
-\subsection{Underlying Model}\label{underlying-model}
-
-To illustrate the value of the non-parametric Bayesian approach to
-management, we focus on example of a system containing such a tipping
-point whose dynamics can still be described by a simple, one-dimensional
-parametric model.\\We will focus on a simple parametric model for a
-single species (derived from fist principles by Allen et al. 2005) as
-our underlying ``reality''.
-
-\[X_{t+1} = Z_t f(S_t)  \] \[S_t = X_t - h_t \]
-\[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]
-
-Where $Z_t$ is multiplicative noise function with mean 1, representing
-stochastic growth. We will consider log-normal noise with shape
-parameter $\sigma_g$. We start with an example in which the parameters
-are $r =2$, $K=8$, $C=5$ and $\sigma_g = 0.1$.
-
-As a low-dimensional system completely described by three parameters,
-this scenario should if anything be favorable to a parametric-based
-approach. This model contains an Allee effect, or tipping point, below
-which the population is not self-sustaining and shrinks to zero
-(Courchamp, Berec, and Gascoigne 2008).
-
-\paragraph{Simulated training data}\label{simulated-training-data}
-
-We generate initial observational data under the model described in Eq 1
-for $T_{\textrm{obs}}=40$ time steps, under a given arbitrary sequence
-of harvest intensities, $h_t$. We consider the case in which most of the
-data comes from a limited region of state space (e.g.~near a stable
-equilibrium), leaving us without observations of the population dynamics
-at very low levels which would be useful in discrimating between
-recruitment curves {[}@{]} or demonstrating the existence of a tipping
-point (Scheffer et al. 2001).\\Using data simulated from a specified
-model rather than empirical data permits the comparison against the true
-underlying dynamics, setting a bar for the optimal performance possible.
-
-\subsubsection{Parametric Models}\label{parametric-models-1}
-
-We consider three candidate parametric models for the stock-recruitment
-function, which we refer to by the first authors of the publications in
-which they were first proposed.
-
-We generate the data with a four-parameter model that contains a tipping
-point, as discussed above (equation 1), (an Allee effect, see
-({\textbf{???}})) below which the stock decreases to zero,
-
-\[ X_{t+1} = Z_t S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(\frac{S_t - \theta}{K}\right)} \]
-
-\[ S_t = X_t - h_t \]
-
-The parameter $C$ reflects the location of the tipping point, $K$ the
-carrying capacity of the stock, and $r$ the base recruitment rate. $S_t$
-represents the stock size after a harvest $h_t$ has been implemented.
-$Z_t$ represents a log-normal random variable of log-mean zero and
-log-standard deviation parameter $\sigma$.
-
-We consider two alternative candidate models: the Ricker
-({\textbf{???}}) stock-recruitment curve,
-
-\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
-
-and an alternative four-parameter model adapted from ({\textbf{???}}),
-
-\[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \]
-
-which contains a tipping point for $\theta > 2$ and becomes a
-Beverton-Holt model at $\theta = 1$.
-
-\subsubsection{Bayesian Inference of Parametric
-models}\label{bayesian-inference-of-parametric-models}
-
-Given the sample data, we infer posterior distributions for each of the
-three models listed above using a Markov Chain Monte Carlo Gibbs Sampler
-(jags, see appendix for implementation details and code) given uniform
-priors. We run six chains for $10^6$ steps each and then assess
-convergence by Gelman-Rubin criterion and inspection of the traces, see
-appendix.
-
-\subsubsection{The Non-parametric Bayesian alternative for
-stock-recruitment
-curves}\label{the-non-parametric-bayesian-alternative-for-stock-recruitment-curves}
-
-\subsubsection{SDP via GP}\label{sdp-via-gp}
-
-Once the posterior Gaussian process (GP) has been estimated (e.g.~see
-Munch et al. 2005), it is necessary to adapt it in place of the
-parametric equation for the stochastic dynamic programming (SDP)
-solution (see Mangel and Clark 1988 for a detailed description of
-parametric SDP methods) to the optimal policy. The essence of the idea
-is straight forward -- we will use the estimated GP in place of the
-parametric growth function to determine the stochastic transition matrix
-on which the SDP calculations are based. The SDP is solved in a
-discretized state space -- both the continuously valued population
-densities $X$ and harvest quotas $h$ are first mapped to a bounded,
-discrete grid. (For simplicity we will consider a uniform grid, though
-for either parametric or GP-based SDP it is often advantageous to use a
-non-uniform discretization such as a basis function representation,
-e.g.~see (Deisenroth, Rasmussen, and Peters 2009)).
-
-The SDP approach then computes a transition matrix, $\mathbf{F}$. We
-demonstrate that calculation is just as straight forward based on the GP
-as it is in the classical context using the parametric model. The
-${i,j}$ of the transition matrix $F$ entry gives the probability of
-transitioning into state $x_i$ given that the system is in state $x_j$
-in the previous time-step. To generate the transition matrix based on
-the posterior GP, we need only the expected values at each grid point
-and the corresponding variances (the diagonal of the covariance matrix),
-as shown in Figure 1. Given the mean of the GP posterior at each
-grid-point as the vector $E$ and variance at that point as vector $V$,
-the probability of transitioning from state $x_i$ to state $x_j$ is
-
-\[\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)\]
-
-where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
-variance $\sigma^2$. Strictly speaking, the transition probability
-should be calculated by integrating the normal density over the bin of
-width $\Delta$ centered at $x_j$. For a sufficiently fine grid that
-$f(x_j) \approx f(x_j + \Delta)$, it is sufficient to calculate the
-density at $x_j$ and then row-normalize the transition matrix. The
-process can then be repeated for each possible discrete value of our
-control variable, (harvest $h$).
-
-\textbf{Pseudocode for the determining the transition matrix from the
-GP}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{for(h in h_grid)}
-  \NormalTok{F_h =}\StringTok{ }\NormalTok{for(x_j in grid)}
-          \NormalTok{for(i in }\DecValTok{1}\NormalTok{:N) }
-            \KeywordTok{dnorm}\NormalTok{(x_j, mu[i]-h, V[i])}
-\end{Highlighting}
-\end{Shaded}
-
-Using the discrete transition matrix we may write down the Bellman
-recursion defining the stochastic dynamic programming iteration:
-
-\begin{equation}
-V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
-\end{equation}
-
-where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$ is
-control (harvest level) chosen. Numerically, the maximization is
-accomplished as follows. Consider the set of possible control values to
-be the discrete values corresponding the grid of stock sizes. Then for
-each $h_t$ there is a corresponding transition matrix $\mathbf{F}_h$
-determined as described above but with mean $\mu = x_j - h_t$. Let
-$\vec{V_t}$ be the vector whose $i$th element corresponds to the value
-of having stock $x_i$ at time $t$. Then let $\Pi_h$ be the vector whose
-$i$th element indicates the profit from harvesting at intensity $h_t$
-given a population $x_i$ (e.g. $\max(x_i, h_t)$ since one cannot harvest
-more fish then the current population size). Then the Bellman recursion
-can be given in matrix form as
-
-\[V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)\]
-
-where the sum is element by element and the expectation is computed by
-the matrix multiplication $\mathbf{F} V_{t+1}$.
-
-\subsubsection{Pseudocode for the Bellman
-iteration}\label{pseudocode-for-the-bellman-iteration-1}
-
-\begin{Shaded}
-\begin{Highlighting}[]
- \NormalTok{V1 <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
-      \NormalTok{delta *}\StringTok{ }\NormalTok{F[[h]] %*%}\StringTok{ }\NormalTok{V +}\StringTok{  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
-    \NormalTok{\})}
-    \CommentTok{# find havest, h that gives the maximum value}
-    \NormalTok{out <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
-      \NormalTok{value <-}\StringTok{ }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
-      \NormalTok{index <-}\StringTok{ }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
-      \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
-    \NormalTok{\})}
-    \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
-    \NormalTok{V <-}\StringTok{ }\NormalTok{out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
-    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <-}\StringTok{ }\NormalTok{out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
-\end{Highlighting}
-\end{Shaded}
-
-This completes the algorithm adapting the GP to the sequential
-decision-making problem through SDP, which has not been previously
-demonstrated.\\We further provide an R package implementation as
-described in the supplemental materials.
-
-\subsubsection{Estimating parametric
-models}\label{estimating-parametric-models}
-
-We estimate posterior distributions for two parametric models: one using
-the structurally correct model as given in Eq (1), which we refer to as
-the ``Parametric Bayes'' model, and another using the familiar Ricker
-model, using a Gibbs sampler as described (with source code) in the
-appendix). In addition we estimate the parameters of the structurally
-correct model by maximum likelihood.
-
-\section{Results}\label{results-1}
-
-\section{Discussion}\label{discussion-1}
-
-\paragraph{Big picture: Linking GP to
-SDP}\label{big-picture-linking-gp-to-sdp}
-
-\emph{rambling}
-
-Non-parametric Bayesian methods have received far too little attention
-in ecological modeling efforts that are aimed at improved conservation
-planning and decision making support. Such approaches may be
-particularly useful when the available data is restricted to a limited
-area of state-space, which can lead parametric models to underestimate
-the uncertainty in dynamics at population levels (states) which have not
-been observed. One reason for the relative absence of nonparametric
-approaches in the natural resource management context may be the lack of
-existing approaches for adapting the non-parametric Bayesian models
-previously proposed (Munch et al. 2005) to a decision-theoretic
-framework. Adapting a non-parametric approach requires modification of
-existing methods for decision theory. We have illustrated how this might
-be done for a classic stochastic dynamic programming problem, opening
-the door for substantial further research into how these applications
-might be improved.
-
-\begin{center}\rule{3in}{0.4pt}\end{center}
-
-Code to replicate the analysis, along with complete log of this research
-can be found at:
-\href{https://github.com/cboettig/nonparametric-bayes/}{https://github.com/cboettig/nonparametric-bayes}
-
-\subsection{Markov Chain Monte Carlo
-Analysis}\label{markov-chain-monte-carlo-analysis}
-
-Allen, Linda J. S., Jesse F. Fagan, Göran Högnäs, and Henrik Fagerholm.
-2005. ``Population Extinction in Discrete-time Stochastic Population
-Models with an Allee Effect.'' \emph{Journal of Difference Equations and
-Applications} 11 (4-5) (apr): 273--293.
-doi:10.1080/10236190412331335373.
-\url{http://www.tandfonline.com/doi/abs/10.1080/10236190412331335373}.
+\subsubsection{Sensitivity Analysis}\label{sensitivity-analysis}
 
@@
 
-Courchamp, Franck, Ludek Berec, and Joanna Gascoigne. 2008. \emph{Allee
-Effects in Ecology and Conservation}. Oxford University Press, USA.
-\url{http://www.amazon.com/Effects-Ecology-Conservation-Franck-Courchamp/dp/0198570309}.
-
 Cressie, Noel, Catherine a Calder, James S. Clark, Jay M. Ver Hoef, and
@@
 
-Deisenroth, Marc Peter, Carl Edward Rasmussen, and Jan Peters. 2009.
-``Gaussian Process Dynamic Programming.'' \emph{Neurocomputing} 72 (7-9)
-(mar): 1508--1524. doi:10.1016/j.neucom.2008.12.019.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0925231209000162}.
+Cuddington, Kim M., M. J. Fortin, and L. R. Gerber. 2013.
+``Process-based Models Are Required to Manage Ecological Systems in a
+Changing World.'' \emph{Ecosphere} 4 (February): 1--12.
+doi:10.1890/ES12-00178.1.
+\url{http://www.esajournals.org/doi/abs/10.1890/ES12-00178.1}.
+
+Fischer, Joern, Garry D. Peterson, Toby a Gardner, Line J. Gordon, Ioan
+Fazey, Thomas Elmqvist, Adam Felton, Carl Folke, and Stephen Dovers.
+2009. ``Integrating Resilience Thinking and Optimisation for
+Conservation.'' \emph{Trends in Ecology \& Evolution} 24 (10) (oct):
+549--54. doi:10.1016/j.tree.2009.03.020.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/19665820}.
+
+Geritz, Stefan A. H., and Eva Kisdi. 2012. ``Mathematical Ecology: Why
+Mechanistic Models?'' \emph{Journal of Mathematical Biology} 65 (6-7)
+(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.
 
@@
 
+Hilborn, Ray, and Marc Mangel. 1997. \emph{The Ecological Detective:
+Confronting Models with Data}. Princeton University Press.
+\url{http://www.amazon.com/The-Ecological-Detective-Confronting-Models/dp/0691034974}.
+
 Hughes, Terry P., Cristina Linares, Vasilis Dakos, Ingrid a van de
@@
 Processes.'' \emph{Mathematical and Computer Modelling of Dynamical
-Systems} 11 (4) (dec): 411--424. doi:10.1080/13873950500068567.
+Systems} 11 (4): 411--424. doi:10.1080/13873950500068567.
 \url{http://www.tandfonline.com/doi/abs/10.1080/13873950500068567}.
@@
 Princeton: Princeton University Press.
+
+Marescot, Lucile, Guillaume Chapron, Iadine Chadès, Paul L. Fackler,
+Christophe Duchamp, Eric Marboutin, and Olivier Gimenez. 2013. ``Complex
+Decisions Made Simple: a Primer on Stochastic Dynamic Programming.''
+\emph{Methods in Ecology and Evolution} (jun).
+doi:10.1111/2041-210X.12082.
+\url{http://doi.wiley.com/10.1111/2041-210X.12082}.
 
@@
 
+McAllister, M. K. 1998. ``Bayesian Stock Assessment: a Review and
+Example Application Using the Logistic Model.'' \emph{ICES Journal of
+Marine Science} 55 (6) (dec): 1031--1060. doi:10.1006/jmsc.1998.0425.
+\href{http://icesjms.oxfordjournals.org/content/55/6/1031.short http://icesjms.oxfordjournals.org/cgi/doi/10.1006/jmsc.1998.0425}{http://icesjms.oxfordjournals.org/content/55/6/1031.short
+http://icesjms.oxfordjournals.org/cgi/doi/10.1006/jmsc.1998.0425}.
+
 Munch, Stephan B., Athanasios Kottas, and Marc Mangel. 2005. ``Bayesian
@@
 \url{http://doi.wiley.com/10.1111/j.1461-0248.2005.00766.x}.
-
-Polasky, Stephen, Stephen R. Carpenter, Carl Folke, and Bonnie Keeler.
-2011. ``Decision-making Under Great Uncertainty: Environmental
-Management in an Era of Global Change.'' \emph{Trends in Ecology \&
-Evolution} (may): 1--7. doi:10.1016/j.tree.2011.04.007.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/21616553}.
 
@@
 
+Roughgarden, Joan E., and F. Smith. 1996. ``Why Fisheries Collapse and
+What to Do About It.'' \emph{Proceedings of the National Academy of
+Sciences of the United States of America} 93 (10): 5078.
+\url{http://www.pnas.org/content/93/10/5078.short}.
+
 Schapaugh, Adam W., and Andrew J. Tyre. 2013. ``Accounting for
@@
 
+Su, Yu-Sung, and Masanao Yajima. 2013. \emph{R2jags: A Package for
+Running Jags from R}. \url{http://cran.r-project.org/package=R2jags}.
+
+Team, R. Core. 2013. \emph{R: A Language and Environment for Statistical
+Computing}. Vienna, Austria: R Foundation for Statistical Computing.
+\url{http://www.r-project.org/}.
+
 Weitzman, Martin L. 2013. ``A Precautionary Tale of Uncertain Tail

### 2013-11-18 inst/examples/BUGS/external-chunks.R

@@ -200,7 +200,7 @@ allen_priors <- ddply(allen_posteriors, "variable", function(dd){
 })
 plot_allen_posteriors <- ggplot(allen_posteriors, aes(value)) + 
   stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=allen_priors, aes(x=value, y=density), col="red") + 
+#  geom_line(data=allen_priors, aes(x=value, y=density), col="red") +  
   facet_wrap(~ variable, scale="free", ncol=3)
 
 
@@ -286,7 +286,7 @@ ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
 # plot posterior distributions
 plot_ricker_posteriors <- ggplot(ricker_posteriors, aes(value)) + 
   stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") + 
+#  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") +  # don't plot priors 
   facet_wrap(~ variable, scale="free", ncol=2)
 
 
@@ -378,7 +378,8 @@ myers_jags <- do.call(autojags,
 tmp <- lapply(as.mcmc(myers_jags), as.matrix) # strip classes
 myers_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
 names(myers_posteriors) = c("index", "variable", "value", "chain")
-plot_myers_traces <- ggplot(myers_posteriors) + geom_line(aes(index, value)) +
+plot_myers_traces <- ggplot(myers_posteriors) + 
+  geom_line(aes(index, value)) + # priors, need to fix order though
   facet_wrap(~ variable, scale="free", ncol=1)
 
 
@@ -390,7 +391,7 @@ par_prior_curves <- ddply(myers_posteriors, "variable", function(dd){
 })
 plot_myers_posteriors <- ggplot(myers_posteriors, aes(value)) + 
   stat_density(geom="path", position="identity", alpha=0.7) +
-  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
+#  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") +  # Whoops, these are misaligned. see table instead 
   facet_wrap(~ variable, scale="free", ncol=3)
 
 

### 2013-11-18 inst/doc/nonparametric-bayes.Rmd
writing

@@
 
+We develop these concerns in the context of fisheries, though
+the underlying challenges and methods are germane to many other
+conservation and resource management problems.  The economic value and
+ecological concern have made marine fisheries the crucible for much
+of the founding work [@Gordon1954; @Reed1979; @May1979; @Ludwig1982]
+in managing ecosystems under uncertainty.  Global trends [@Worm2006]
+and controversy [@Hilborn2007; @Worm2009] have made understanding these
+challenges all the more pressing.
+
 <!-- Uncertainty outside the data without the correct model has not been handled. -->
@@
 
+_Cut the next three paragraphs, since they are covered more consisely in the above paragraph?_
+
+#### Process, measurement, and implementation error
+
+Resource management and conservation planning seek to determine the
+optimal set of feasible actions to maximize the value of some objectives
+(see [@Halpern2012]).  Process error, measurement error, implementation
+error [@Reed1979, @Clark1986, @Roughgarden1996, @Sethi2005].  These
+sources of stochasticity in turn mean that model parameters can only
+be estimated approximately, requiring parametric uncertainty also be
+considered [@Ludwig1982].
+
+#### Parametric uncertainty
+
+As the parameter values for these models must be estimated from limited
+data, there will always be some uncertainty associated with these values.
+This uncertainty further compounds the intrinsic variability introduced
+by demographic or environmental noise.  The degree of uncertainty in
+the parameter values can be inferred from the data and reflected in the
+estimates of the transition probabilities [@Walters1982; @Mangel1988;
+@Mangel1997; @Schapaugh2013].
+
+#### Structural (model) uncertainty
+
+Estimates of parameter uncertainty are only as good as the parametric
+models themselves.  Often we do not understand the system dynamics well
+enough to know if a model provides a good approximation over the relevant
+range of states and timescales (criteria that we loosely refer to as
+defining the "right" or "true" model.)  So called structural or model
+uncertainty is a more difficult problem than parametric uncertainty.
+Typical solutions involve either model choice, model averaging, or
+introducing yet greater model complexity of which others may be special
+cases (model averaging being one such way to construct such a model)
+[@Williams2001; @Athanassoglou2012; @Cressie2009].  Even setting aside
+other computational and statistical concerns (e.g. [@Cressie2009]), these
+approaches do not address our second concern - representing uncertainty
+outside the observed data range.
+
 <!-- __We don't have the data where we need it__ -->
@@
 
-Model uncertainty is particularly insidious when model predictions
-must be made outside of the range of data on which the model was estimated.
+Model uncertainty is particularly insidious when model predictions must
+be made outside of the range of data on which the model was estimated.
+This extrapolation uncertainty is felt most keenly in decision-theoretic
+(or optimal control) applications, since (a) exploring the potential
+action space typically involves considering actions that may move the
+system outside the range of observed behavior, and (b) decision-theoretic
+alogrithms rely not only on reasonable estimates of the expected
+outcomes, but depend on the weights given to all possible outcomes
+[_e.g._ @Weitzman2013].  If we are observing the fluctuations of a
+given fish stock over many years under a fixed harvesting pressure,
+we might develop and test a model that could reasonably predict the
+frequency of a deviation of a given size, even when such a deviation has
+not been previously observed. Yet such predictions are far less reliable
+when extrapolated to a harvest pressure that has not yet been observed.
+Thus, model uncertainty can be particularly challenging in the management
+and decision-making context.
 
-An additional source of uncertainty that has recieved less attention[^1]
-arises when applying a dynamical model outside the range of data on
-which it has been estimated.  This extrapolation uncertainty is felt
-most keenly in decision-theoretic applications, as (a) exploring the
-potential action space typically involves considering actions that
-may move the system outside the range of observed behavior, and (b)
-decision-theoretic alogrithms rely not only on reasonable estimates of
-the expected outcomes, but depend on the weights given to all possible
-outcomes [_e.g._ @Weitzman2013].
+This difficult position of having neither the true model nor data that
+covers the full range of possible states is unfortunately the rule more
+than the exception. The potential concern of tipping points in ecological
+dyanmics [@Scheffer2001; @Polasky2011] reflects these concerns -- as
+either knowledge of the true model or more complete sampling of the
+state space would make it easy to identify if a tipping point existed.
+If we do not know but cannot rule out such a possibility, then we
+face decision-making under this dual challege of model uncertainty and
+incomplete data coverage.
 
-<!-- 
-## Why we don't know the model
-- Complex dynamics  [@Glaser2013](http://doi.org/10.1111/faf.12037 "Complex dynamics may limit prediction in marine fisheries")
-- model choice and model averaging approaches
--->
-
-<!--
-## Why we don't have data where we need it
-- Concerns of tipping points 
-- Danger of learning 
--->
-
-<!-- Perhaps this is a better-worded version of the above??
-
-Often the set of feasible states is larger than the range of the observed
-states.  Though most systems subject to some degree of stochasticity
-can enter a state outside previously observed values by chance alone
-(such as a particularly low or high nutrient year), this concern is
-most keenly felt in the decision-theoretic context. If we are observing
-the fluctuations of a given fish stock over many years under a fixed
-harvesting pressure, we might develop and test a model that could
-reasonably predict the frequency of a deviation of a given size, even
-when such a deviation has not been previously observed.
-(The same analogy might be made to predicting hurricanes or other
-extreme weather events from historical data, which may be more accurately
-predicted in a constant climate than in one being perturbed).
-
--->
-
-<!-- Does this belong?  Improve transitions -->
-
-Concerns over the potential for tipping points in ecological dynamics
-[@Scheffer2001] highlight the dangers of uncertainty in ecological
-management and pose a substantial challenge to existing decision-theoretic
+These dual concerns pose a substantial challenge to existing decision-theoretic
 approaches [@Brozovic2011].  Because intervention is often too late 
@@
 decision-theoretic approach in face of substantial structural uncertainty.
-
-<!-- More consise versions are now above 
-#### Parametric uncertainty
-
-As the parameter values for these models must be estimated from limited data,  
-there will always be some uncertainty associated with these values.  This uncertainty
-further compounds the intrinsic variability introduced by demographic or environmental 
-noise.  The degree of uncertainty in the parameter values can be inferred from the data
-and reflected in the estimates of the transition probabilities [@Walters1982; @Mangel1988; @Mangel1997; @Schapaugh2013].
-
-#### Structural uncertainty
-
-Estimates of parameter uncertainty are only as good as the parametric
-models themselves.  Often we do not understand the system dynamics well
-enough to know if a model provides a good approximation over the relevant
-range of states and timescales (criteria that we loosely refer to as
-defining the "right" or "true" model.)  So called structural or model
-uncertainty is a more difficult problem than parametric uncertainty.
-Typical solutions involve either model choice, model averaging, or
-introducing yet greater model complexity of which others may be special
-cases (model averaging being one such way to construct such a model)
-[@Williams2001; @Athanassoglou2012; @Cressie2009].  Even setting aside
-other computational and statistical concerns (e.g. [@Cressie2009]), these
-approaches do not address our second concern - representing uncertainty
-outside the observed data range.
-
-#### Limits of state space
-
-Resource management and conservation planning seek to determine the
-optimal set of feasible actions to maximize the value of some objectives
-(see [@Halpern2012])  Process error, measurement error, implementation
-error [@Reed1979, @Clark1986, @Roughgarden1996, @Sethi2005].  These
-sources of stochasticity in turn mean that model parameters can only
-be estimated approximately, requiring parametric uncertainty also be
-considered [@Ludwig1982, ].
--->
 
@@
 
-- Note on: Why fisheries
-
-The economic value and ecological concern have made marine
-fisheries the crucible for much of the founding work [@Gordon1954;
-@Reed1979; @May1979; @Ludwig1982] in managing ecosystems under
-uncertainty.  Global trends [@Worm2006] and controversy [@Hilborn2007;
-@Worm2009] have made understanding these challenges all the more pressing.
-
-<!--move to  much later *Or do we need this at all?* -->
-
+<!--move to  much later *Or do we need this at all?* 
 - _Note on comparing models_ (via value function rather than by "fit"). 
+-->
 
@@
 
+<!-- Useless 
 The essence of the GP approach can be captured in the following
@@
 consider probabilities on a large set of possible curves simultaneously.
+-->
 
@@
 
-<!-- FIXME -->
-
-- Underlying model 
-- Available data
-
-<!--
-- Value function
--->
-
-For simplicity we assume profit is simply linear in the realized harvest (only
-enforcing the restriction that harvest can not exceed available stock)
+To illustrate the application of the BNP-SDP approach and compare to the
+predictions of the alternative parametric models we focus on the classical
+problem of selecting the appropriate harvest level given an observation
+of the stock size in the previous year  [@Reed1979; @Walters1982, @Mangel1988].
+Given this observation and the model (together with the parameter uncertainty) of the stock
+recruitment process, the manager seeks to maximize the value of the 
+fishery over a fixed time interval of `r OptTime` years at a discount rate
+of `r delta`. The value function (profits) at time $t$ depends on the true stock size
+$x_t$ and the chosen harvest level $h_t$.  For simplicity we assume profit 
+is simply proportional in the realized harvest (only
+enforcing the restriction that harvest can not exceed available stock).
 
@@
 We consider three candidate parametric models of the stock-recruitment
-dynamics: The Ricker model, the Allen model [Allen 2005](), the Myers
-model. The familiar Ricker model involves two parameters, corresponding
+dynamics: The Ricker model, the Allen model [@Allen2005], the Myers
+model [@Myers1995]. The familiar Ricker model involves two parameters, corresponding
 to a growth rate and a carrying capacity, and cannot support alternative
@@
 the purposes of this analysis.) The Allen model resembles the Ricker
-dynamics with an added Allee effect parameter [Courchamp](), below
+dynamics with an added Allee effect parameter [@Courchamp2008], below
 which the population cannot persist.  The Myers model also has three
@@
 
-As we simulate training data from the Allen model (ref section), we will
+As we simulate training data from the Allen model, we will
 refer to this as the structurally correct model.  The Ricker model is
@@
 we have three potential parametric models of the stock dynamics.
+
+We introduce parameteric uncertainty by first estimating each of the
+candidate models from data on unexploited stock dynamics following
+some pertubation (non-equilibrium initial condition) over several time
+steps. This training data could be generated in several different ways
+(such as known variable exploitation rates, etc), as long as it reflects
+the dynamics in some limited region of state space without impacting
+the problem.  We consider a period of `r Tobs` years of training data: long
+enough that the estimates are not dependent on the particular realization,
+while longer times are not likely to provide substantial improvement
+(i.e. the results are not sensitive to this interval).  Each of the models
+(described below) is fit to the same training data, as shown in Figure 1.
 
@@
 
-```{r deviances, dependson=c("dic_calc"), include=TRUE, results="asis", echo=FALSE}
+```{r deviances, dependson=c("dic_calc"), include=FALSE, echo=FALSE}
+```
+
+```{r Table1, dependson="deviances", include=TRUE, results="asis", echo=FALSE}
+xtable::xtable(rbind(dictable, aictable, bictable))
 ```
@@
 
-```{r out_of_sample_predictions, include=TRUE, echo=FALSE, fig.width=8, fig.height=6}
+```{r out_of_sample_predictions, include=TRUE, echo=FALSE, fig.width=8, fig.height=6, fig.cap="Out of sample predictions of the dynamics under each model.
+Points show the stock size simulated by the true model.
+Overlay shows the range of states predicted by each model, based on the state observed in the previous time step.
+The Ricker model always predicts population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).
+Note that the GP is both more pessimistic and more uncertain about the future state than the parameteric models, while the realized state often falls outside of the expected range forecasted by the structurally incorrect Myers and Ricker models." }
 ```

### 2013-11-18 inst/doc/nonparametric-bayes.tex
writing

@@
 
+We develop these concerns in the context of fisheries, though the
+underlying challenges and methods are germane to many other conservation
+and resource management problems. The economic value and ecological
+concern have made marine fisheries the crucible for much of the founding
+work (H. S. Gordon 1954; Reed 1979; May et al. 1979; Ludwig and Walters
+1982) in managing ecosystems under uncertainty. Global trends (Worm et
+al. 2006) and controversy (Hilborn 2007; Worm et al. 2009) have made
+understanding these challenges all the more pressing.
+
 Uncertainty enters the decision-making process at many levels: intrinsic
 stochasticity in biological processes, measurements, and implementation
-of policy (\emph{e.g.} Reed 1979; C. W. Clark and Kirkwood 1986;
-Roughgarden and Smith 1996; Sethi et al. 2005), parameteric uncertainty
-(\emph{e.g.} Ludwig and Walters 1982; Hilborn and Mangel 1997;
-McAllister 1998; Schapaugh and Tyre 2013), and model or structural
-uncertainty (\emph{e.g.} B. K. Williams 2001; Cressie et al. 2009;
-Athanassoglou and Xepapadeas 2012). Of these, structural uncertainty
-incorporates the least a priori knowledge or assumptions and is
-generally the hardest to quantify. Typical approaches assume a weak
-notion of model uncertainty in which the correct model (or reasonable
-approximation) of the dynamics must be identified from among a handful
-of alternative models. Here we consider an approach that addresses
-uncertainty at each of these levels without assuming the dynamics follow
-a particular (i.e.~parametric) structure.
+of policy (\emph{e.g.} Reed 1979; Clark and Kirkwood 1986; Roughgarden
+and Smith 1996; Sethi et al. 2005), parameteric uncertainty (\emph{e.g.}
+Ludwig and Walters 1982; Hilborn and Mangel 1997; McAllister 1998;
+Schapaugh and Tyre 2013), and model or structural uncertainty
+(\emph{e.g.} B. K. Williams 2001; Cressie et al. 2009; Athanassoglou and
+Xepapadeas 2012). Of these, structural uncertainty incorporates the
+least a priori knowledge or assumptions and is generally the hardest to
+quantify. Typical approaches assume a weak notion of model uncertainty
+in which the correct model (or reasonable approximation) of the dynamics
+must be identified from among a handful of alternative models. Here we
+consider an approach that addresses uncertainty at each of these levels
+without assuming the dynamics follow a particular (i.e.~parametric)
+structure.
+
+\emph{Cut the next three paragraphs, since they are covered more
+consisely in the above paragraph?}
+
+\paragraph{Process, measurement, and implementation
+error}\label{process-measurement-and-implementation-error}
+
+Resource management and conservation planning seek to determine the
+optimal set of feasible actions to maximize the value of some objectives
+(see ({\textbf{???}})). Process error, measurement error, implementation
+error (Reed 1979). These sources of stochasticity in turn mean that
+model parameters can only be estimated approximately, requiring
+parametric uncertainty also be considered (Ludwig and Walters 1982).
+
+\paragraph{Parametric uncertainty}\label{parametric-uncertainty}
+
+As the parameter values for these models must be estimated from limited
+data, there will always be some uncertainty associated with these
+values. This uncertainty further compounds the intrinsic variability
+introduced by demographic or environmental noise. The degree of
+uncertainty in the parameter values can be inferred from the data and
+reflected in the estimates of the transition probabilities
+({\textbf{???}}; Mangel and Clark 1988; {\textbf{???}}; Schapaugh and
+Tyre 2013).
+
+\paragraph{Structural (model)
+uncertainty}\label{structural-model-uncertainty}
+
+Estimates of parameter uncertainty are only as good as the parametric
+models themselves. Often we do not understand the system dynamics well
+enough to know if a model provides a good approximation over the
+relevant range of states and timescales (criteria that we loosely refer
+to as defining the ``right'' or ``true'' model.) So called structural or
+model uncertainty is a more difficult problem than parametric
+uncertainty. Typical solutions involve either model choice, model
+averaging, or introducing yet greater model complexity of which others
+may be special cases (model averaging being one such way to construct
+such a model) (B. K. Williams 2001; Athanassoglou and Xepapadeas 2012;
+Cressie et al. 2009). Even setting aside other computational and
+statistical concerns (e.g. (Cressie et al. 2009)), these approaches do
+not address our second concern - representing uncertainty outside the
+observed data range.
 
@@
 be made outside of the range of data on which the model was estimated.
-
-An additional source of uncertainty that has recieved less
-attention{[}\^{}1{]} arises when applying a dynamical model outside the
-range of data on which it has been estimated. This extrapolation
-uncertainty is felt most keenly in decision-theoretic applications, as
-(a) exploring the potential action space typically involves considering
-actions that may move the system outside the range of observed behavior,
-and (b) decision-theoretic alogrithms rely not only on reasonable
-estimates of the expected outcomes, but depend on the weights given to
-all possible outcomes (\emph{e.g.} Weitzman 2013).
-
-Concerns over the potential for tipping points in ecological dynamics
-(Scheffer et al. 2001) highlight the dangers of uncertainty in
-ecological management and pose a substantial challenge to existing
+This extrapolation uncertainty is felt most keenly in decision-theoretic
+(or optimal control) applications, since (a) exploring the potential
+action space typically involves considering actions that may move the
+system outside the range of observed behavior, and (b)
+decision-theoretic alogrithms rely not only on reasonable estimates of
+the expected outcomes, but depend on the weights given to all possible
+outcomes (\emph{e.g.} Weitzman 2013). If we are observing the
+fluctuations of a given fish stock over many years under a fixed
+harvesting pressure, we might develop and test a model that could
+reasonably predict the frequency of a deviation of a given size, even
+when such a deviation has not been previously observed. Yet such
+predictions are far less reliable when extrapolated to a harvest
+pressure that has not yet been observed. Thus, model uncertainty can be
+particularly challenging in the management and decision-making context.
+
+This difficult position of having neither the true model nor data that
+covers the full range of possible states is unfortunately the rule more
+than the exception. The potential concern of tipping points in
+ecological dyanmics (Scheffer et al. 2001; Polasky et al. 2011) reflects
+these concerns -- as either knowledge of the true model or more complete
+sampling of the state space would make it easy to identify if a tipping
+point existed. If we do not know but cannot rule out such a possibility,
+then we face decision-making under this dual challege of model
+uncertainty and incomplete data coverage.
+
+These dual concerns pose a substantial challenge to existing
 decision-theoretic approaches (Brozović and Schlenker 2011). Because
@@
 parametric modeling approaches.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Note on: Why fisheries
-\end{itemize}
-
-The economic value and ecological concern have made marine fisheries the
-crucible for much of the founding work (H. S. Gordon 1954; Reed 1979;
-May et al. 1979; Ludwig and Walters 1982) in managing ecosystems under
-uncertainty. Global trends (Worm et al. 2006) and controversy (Hilborn
-2007; Worm et al. 2009) have made understanding these challenges all the
-more pressing.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \emph{Note on comparing models} (via value function rather than by
-  ``fit'').
-\end{itemize}
 
@@
 become themselves part of the model specification, as in the
-nonparametric regression used here. we shall see here.
+nonparametric regression used here.
 
@@
 geospatial literature) to formulate a predictive model is relatively new
-in the context of modeling dynamical systems (Kocijan et al. 2005), and
-was first introduced in the context ecological modeling and fisheries
+in the context of modeling dynamical systems ({\textbf{???}}), and was
+first introduced in the context ecological modeling and fisheries
 management in Munch et al. (2005). An accessible and thorough
@@
 and Williams (2006).
-
-The essence of the GP approach can be captured in the following thought
-experiment: An exhaustive parametric approach to the challenge of
-structural uncertainty might proceed by writing down all possible
-functional forms for the underlying dynamical system with all possible
-parameter values for each form, and then consider searching over this
-huge space to select the most likely model and parameters; or using a
-Bayesian approach, assign priors to each of these possible models and
-infer the posterior distribution of possible models. The GP approach can
-be thought of as a computationally efficient approximation to this
-approach. GPs represent a large class of models that can be though of as
-capturing or reasonably approximating the set of models in this
-collection. By modeling at the level of the process, rather than the
-level of parametric equation, we can more concisely capture the possible
-behavior of these curves. In place of a parametric model of the
-dynamical system, the GP approach postulates a prior distribution of
-(n-dimensional) curves that can be though of as approximations to a
-range of possible (parametric) models that might describe the data. The
-GP allows us to consider probabilities on a large set of possible curves
-simultaneously.
 
@@
 
-\% latex table generated in R 3.0.2 by xtable 1.7-1 package \% Thu Nov
-14 14:10:28 2013
+\% latex table generated in R 3.0.2 by xtable 1.7-1 package \% Mon Nov
+18 14:39:35 2013
 
@@
 \includegraphics{figure/nonparametric-bayes-out_of_sample_predictions.pdf}
-\caption{plot of chunk out\_of\_sample\_predictions}
+\caption{Out of sample predictions of the dynamics under each model.
+Points show the stock size simulated by the true model. Overlay shows
+the range of states predicted by each model, based on the state observed
+in the previous time step. The Ricker model always predicts population
+growth, while the actual population shrinks in each step as the initial
+condition falls below the Allee threshold of the underlying model
+(Allen). Note that the GP is both more pessimistic and more uncertain
+about the future state than the parameteric models, while the realized
+state often falls outside of the expected range forecasted by the
+structurally incorrect Myers and Ricker models.}
 \end{figure}
@@
 Analysis: the Strengths and Limitations of Hierarchical Statistical
-Modeling.'' \emph{Ecological Applications} 19 (3) (apr): 553--70.
-doi:10.1890/07-0744.1.
+Modeling.'' \emph{Ecological Applications : a Publication of the
+Ecological Society of America} 19 (3) (apr): 553--70.
 \url{http://www.ncbi.nlm.nih.gov/pubmed/19425416}.
@@
 
-Kocijan, Juš, Agathe Girard, Bla\textbackslash{}vz Banko, and Roderick
-Murray-Smith. 2005. ``Dynamic Systems Identification with Gaussian
-Processes.'' \emph{Mathematical and Computer Modelling of Dynamical
-Systems} 11 (4): 411--424. doi:10.1080/13873950500068567.
-\url{http://www.tandfonline.com/doi/abs/10.1080/13873950500068567}.
-
 Ludwig, Donald, and Carl J. Walters. 1982. ``Optimal Harvesting with
@@
 
+Polasky, Stephen, Stephen R. Carpenter, Carl Folke, and Bonnie Keeler.
+2011. ``Decision-making Under Great Uncertainty: Environmental
+Management in an Era of Global Change.'' \emph{Trends in Ecology \&
+Evolution} (may): 1--7. doi:10.1016/j.tree.2011.04.007.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/21616553}.
+
 Rasmussen, Carl Edward, and C. K. I. Williams. 2006. \emph{Gaussian

### 2013-11-18 inst/examples/BUGS/external-chunks.R

@@ -528,13 +528,18 @@ for(t in 1:(length(y)-1))
 crash_data <- step_ahead_posteriors(y)
 crash_data <- subset(crash_data, variable %in% c("GP", "Allen", "Ricker", "Myers"))
 ggplot(crash_data) + 
-  geom_point(aes(time, stock)) + 
-  geom_line(aes(time, value, col = variable, 
-            group=interaction(L1,variable)), alpha=.1) + 
-  facet_wrap(~variable) + 
-  scale_colour_manual(values=colorkey, 
+  geom_boxplot(aes(as.factor(as.integer(time)), value, 
+                   fill = variable, col=variable), 
+               alpha=.7, outlier.size=1, position="identity") + 
+#  geom_line(aes(time, value, col = variable, 
+#            group=interaction(L1,variable)), alpha=.1) + 
+  geom_point(aes(time, stock), size = 3) + 
+  scale_fill_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")], 
                       guide = guide_legend(override.aes = list(alpha = 1))) +  
-  theme(legend.position="none")
+  scale_colour_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")], 
+                      guide = guide_legend(override.aes = list(alpha = 1))) +  
+  facet_wrap(~variable) + 
+  theme(legend.position="none") + xlab("time") + ylab("stock size") 
 
 
 ## @knitr gp-opt

### 2013-11-19 inst/doc/nonparametric-bayes.Rmd
added metadata to yaml instead of template

@@
-% Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach
+---
+title: "Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach"
+author: 
+  - name: "Carl Boettiger"
+    affiliation: cstar
+    email: cboettig@gmail.com
+    footnote: Corresponding author
+  - name: Marc Mangel
+    affiliation: cstar
+  - name: Stephan Munch
+    affiliation: noaa
+address: 
+  - code: cstar
+    address: | 
+      Center for Stock Assessment Research, 
+      Department of Applied Math and Statistics, 
+      University of California, Mail Stop SOE-2,
+      Santa Cruz, CA 95064, USA
+  - code: noaa
+    address: | 
+      Southwest Fisheries Science Center, 
+      National Oceanic and Atmospheric Administration, 
+      110 Shaffer Road, Santa Cruz, CA 95060, USA
+
+abstract: |
+          Model uncertainty and limited data coverage are fundamental challenges to
+          robust ecosystem management.  These challenges are acutely highlighted
+          by concerns that many ecological systems may contain tipping points.
+          Before a collapse, we do not know where the tipping points lie, if the
+          exist at all.  Hence, we know neither a complete model of the system
+          dynamics nor do we have access to data in some large region of state-space
+          where such a tipping point might exist.  These two sources of uncertainty
+          frustrate state-of-the-art parametric approaches to decision theory
+          and optimal control.  I will illustrate how a non-parametric approach
+          using a Gaussian Process prior provides a more flexible representation
+          of this inherent uncertainty.  Consequently, we can adapt the Gaussian
+          Process prior to a stochastic dynamic programming framework in order to
+          make robust management predictions under both model and uncertainty and
+          limited data.
+keywords:
+  - Bayesian
+  - Structural Uncertainty
+  - Nonparametric 
+  - Optimal Control
+  - Decision Theory
+  - Gaussian Processes
+  - Fisheries Management
+---
 
@@
 ```
+
+```{r }
+allen_priors_xtable <- data.frame(parameter = c("r0", "K", "theta", "sigma"),
+           "lower bound" = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
+           "upper bound" = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2]))
+```
+
 ```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results=toggle}
@@
 ```
+
+```{r}
+ricker_priors_xtable <- data.frame(
+  parameter = c("r0", "K", "sigma"),
+  "lower bound" = c(r0_prior_p[1], K_prior_p[1], stdQ_prior_p[1]),
+  "upper bound" = c(r0_prior_p[2], K_prior_p[2], stdQ_prior_p[2]))
+```
+
 ```{r ricker-mcmc, dependson="ricker-model", results=toggle}
@@
 ```
+
+```{r } 
+myers_priors_xtable <- data.frame(parameter = c("r0", "K", "theta", "sigma"),
+           "lower bound" = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
+           "upper bound" = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2]))
+```
+
 ```{r myers-mcmc, dependson="myers-model", results=toggle}
@@
 ```
-
-Abstract
-========
-
-Model uncertainty and limited data coverage are fundamental challenges to
-robust ecosystem management.  These challenges are acutely highlighted
-by concerns that many ecological systems may contain tipping points.
-Before a collapse, we do not know where the tipping points lie, if the
-exist at all.  Hence, we know neither a complete model of the system
-dynamics nor do we have access to data in some large region of state-space
-where such a tipping point might exist.  These two sources of uncertainty
-frustrate state-of-the-art parametric approaches to decision theory
-and optimal control.  I will illustrate how a non-parametric approach
-using a Gaussian Process prior provides a more flexible representation
-of this inherent uncertainty.  Consequently, we can adapt the Gaussian
-Process prior to a stochastic dynamic programming framework in order to
-make robust management predictions under both model and uncertainty and
-limited data.
 
@@
 
-#### Process, measurement, and implementation error
+### Process, measurement, and implementation error
 
@@
 optimal set of feasible actions to maximize the value of some objectives
-(see [@Halpern2012]).  Process error, measurement error, implementation
+(e.g @Halpern2013).  Process error, measurement error, implementation
 error [@Reed1979, @Clark1986, @Roughgarden1996, @Sethi2005].  These
@@
 
-#### Parametric uncertainty
+### Parametric uncertainty
 
@@
 the parameter values can be inferred from the data and reflected in the
-estimates of the transition probabilities [@Walters1982; @Mangel1988;
-@Mangel1997; @Schapaugh2013].
+estimates of the transition probabilities [@Ludwig1982; @Mangel1988;
+@Hilborn1997; @Schapaugh2013].
 
-#### Structural (model) uncertainty
+### Structural (model) uncertainty
 
@@
 problem of selecting the appropriate harvest level given an observation
-of the stock size in the previous year  [@Reed1979; @Walters1982, @Mangel1988].
+of the stock size in the previous year  [@Reed1979; @Ludwig1982; @Mangel1988].
 Given this observation and the model (together with the parameter uncertainty) of the stock
@@
 curve rather than a Ricker curve at high densities.)  
-
 We assume multiplicative log-normal noise perturbs the growth predicted 
@@
 
-We also estimate a simple Gaussian Process defined by
-a radial basis function kernel of two parameters: $\ell$, which gives
-the characteristic length-scale over which correlation between two 
-points 
-in state-space decays, and $\sigma$, which gives the scale of the 
-process noise by which observations $Y_{t+1}$ may differ from their
-predicted values $X_{t+1}$ given an observation of the previous state,
-$X_t$. @Munch2005a gives an accessible introduction to the use of 
-Gaussian Processes in providing a Bayesian nonparametric description
-of the stock-recruitment relationship.
+We also estimate a simple Gaussian Process defined by a radial basis
+function kernel of two parameters: $\ell$, which gives the characteristic
+length-scale over which correlation between two points in state-space
+decays, and $\sigma$, which gives the scale of the process noise by which
+observations $Y_{t+1}$ may differ from their predicted values $X_{t+1}$
+given an observation of the previous state, $X_t$. @Munch2005a gives an
+accessible introduction to the use of Gaussian Processes in providing a
+Bayesian nonparametric description of the stock-recruitment relationship.
 
@@
 
+We use a Metropolis-Hastings Markov Chain Monte Carlo to infer posterior
+distributions of the two parameters of the GP (Figure S13, code in
+appendix), under weakly informative Gaussian priors (see parameters in
+table S5). As the posterior distributions differ substantially from the
+priors (Figure S13), we can be assured that most of the information in
+the posterior comes from the data rather than the prior belief.
+
+<!--
 In contrast to the parametric models, this posterior distribution is still
@@
 observed data.
-
-We use a Metropolis-Hastings Markov Chain Monte Carlo to infer posterior
-distributions of the two parameters of the GP (Figure S4, code in
-appendix), under weakly informative Gaussian priors (see parameters in
-table S5). As the posterior distributions differ substantially from the
-priors (Figure S4), we can be assured that most of the information in
-the posterior comes from the data rather than the prior belief.
+-->
 
@@
 
-```{r Table1, dependson="deviances", include=TRUE, results="asis", echo=FALSE}
+```{r Table1, dependson="deviances", include=TRUE, results="asis", echo=FALSE, cache=FALSE}
 xtable::xtable(rbind(dictable, aictable, bictable))
@@
 In any modeling effort, models must be chosen for the task at hand.
-Though simple mechanistically motivated models offer the greatest potential 
-to increase our basic understanding of ecological processes [@Cuddington2013; @Geritz2012], 
-such models can be not only inaccurate but misleading when relied upon in a
-quantitative decision making framework.  
-In this paper we have tackled two aspects of uncertainty that are both common
-to many ecological desicion-making problems and fundamentally challenging to
-existing approaches which largely rely on parametric models:
+Though simple mechanistically motivated models offer the greatest
+potential to increase our basic understanding of ecological processes
+[@Cuddington2013; @Geritz2012], such models can be not only inaccurate but
+misleading when relied upon in a quantitative decision making framework.
+In this paper we have tackled two aspects of uncertainty that are both
+common to many ecological desicion-making problems and fundamentally
+challenging to existing approaches which largely rely on parametric
+models:
 
@@
 
+<!--
 ### Higher dimensions 
+-->
 
@@
 
+<!--
 ### Real-time learning
+-->
 
@@
 
+\appendix
+
 Appendix
@@
 
-## Model definitions and estimation
+Model definitions and estimation
+---------------------------------
 
-Equation S1: Ricker model.
+<!-- Equation S1: Ricker model. -->
 
-$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
+$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \label{rickereq} $$
 
-Figure S1: Ricker model: prior and posterior distributions for parameter estimates.
+<!-- Figure S1: Ricker model: prior and posterior distributions for parameter estimates. -->
 
-```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimate of the Ricker model", fig.width=8}
+```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimate of the Ricker model", fig.width=6}
 plot_ricker_traces
@@
 
-```{r, echo=FALSE, include=TRUE, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=8}
+```{r, echo=FALSE, include=TRUE, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
 ggplot(ricker_posteriors, aes(value)) + 
@@
 ```
-Table S1: Parameterization of the priors
 
-```{r echo=FALSE, results = "asis", include = TRUE}
-xtable::xtable(data.frame(
-  parameter = c("r0", "K", "sigma"),
-  lower_bound = c(r0_prior_p[1], K_prior_p[1], stdQ_prior_p[1]),
-  upper_bound = c(r0_prior_p[2], K_prior_p[2], stdQ_prior_p[2])))
+<!-- Table S1: Parameterization of the priors -->
+
+```{r echo=FALSE, results = "asis", include = TRUE, cache=FALSE}
+xtable::xtable(ricker_priors_xtable,
+  caption = "parameterization range for the uniform priors in the Ricker model")
 ```
 
-$$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} $$
+<!--Eq S2: Myers model -->
+(@myerseq) $$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}}$$
 
-Eq S2: Myers model 
+<!--
 Figure S2: Myers model: Traces, prior and posterior distributions for parameter estimates.
+-->
 
-```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimate of the Myers model", fig.width=8}
+```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimate of the Myers model", fig.width=6}
 plot_myers_traces
@@
 
-```{r, echo=FALSE, include=TRUE, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=8}
+```{r, echo=FALSE, include=TRUE, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
 ggplot(myers_posteriors, aes(value)) + 
   stat_density(geom="path", position="identity", alpha=0.7) +
-  facet_wrap(~ variable, scale="free", ncol=3)
+  facet_wrap(~ variable, scale="free", ncol=2)
 ```
 
-Table S2: Parameterization of the priors
-```{r echo=FALSE, results="asis", include = TRUE}
-xtable::xtable(data.frame(parameter = c("r0", "K", "theta", "sigma"),
-           lower_bound = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
-           upper_bound = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2])))
+<!-- Table S2: Parameterization of the priors -->
+
+```{r echo=FALSE, results="asis", include = TRUE, cache=FALSE}
+xtable::xtable(myers_priors_xtable,
+           caption = "parameterization range for the uniform priors in the Myers model")
 ```
@@
 
-$$f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} $$
+(@alleneq) $$f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} $$
 
@@
 
-```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimate of the Allen model", fig.width=8}
+```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimate of the Allen model", fig.width=6}
 plot_allen_traces
@@
 
-```{r, echo=FALSE, include=TRUE, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=8}
+```{r, echo=FALSE, include=TRUE, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
 ggplot(allen_posteriors, aes(value)) + 
   stat_density(geom="path", position="identity", alpha=0.7) +
-  facet_wrap(~ variable, scale="free", ncol=3)
+  facet_wrap(~ variable, scale="free", ncol=2)
 ```
 
+<!--
 Table S3: Parameterization of the priors
+-->
 
-```{r echo=FALSE, results = "asis", include=TRUE}
-xtable::xtable(data.frame(parameter = c("r0", "K", "theta", "sigma"),
-           lower_bound = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
-           upper_bound = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2])))
+```{r echo=FALSE, results = "asis", include=TRUE, cache=FALSE}
+xtable::xtable(allen_priors_xtable,
+  caption = "parameterization range for the uniform priors in the Allen model")
 ```
 
-Eq S4: GP model 
-Figure S4: GP model: prior and posterior distributions for parameter estimates.
+<!-- Eq S4: GP model  -->
 
-```{r echo=FALSE, include=TRUE}
-gp_assessment_plots
+Assuming the data are known with some process noise,
+
+$$y = f(x) + \varepsilon$$
+
+$\varepsilon$ IID normal, variance $\sigma_n^2$. Then under the GP,
+
+$$x|y \sim \mathcal{N}(E,C)$$
+$$E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y$$
+$$C = K(X_p, X_p) - K(X_p, X_o) K(X_o,X_o)^{-1} K(X_o, X_p)$$
+
+}
+
+The marginal likelihood is then given by:
+
+$$\log(p(y | X)) = -\tfrac{1}{2} \mathbf{y}^T (K + \sigma_n^2 \mathbf{I})^{-1} y  - \tfrac{1}{2} \log\left| K + \sigma_n^2 \mathbf{I} \right| - \tfrac{n}{2}\log 2 \pi$$
+
+See @Munch2005a or @Rasmussen2006 for a more detailed introduction.
+
+<!-- Figure S4: GP model: prior and posterior distributions for parameter estimates. -->
+
+```{r echo=FALSE, include=TRUE, fig.cap = "Traces from the MCMC estimates of the GP model"}
+gp_assessment_plots$traces_plot
 ```
 
-Table S4: Parameterization of the priors
+```{r echo=FALSE, include=TRUE, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
+gp_assessment_plots$posteriors_plot
+```
+
+The Gaussian process priors on both the lengthscale $\ell$ and process noise $\sigma$ are are inverse Gamma distributed,
+
+$$f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)$$
+
+For the $\sigma$ prior, $\alpha = $ `r s2.p[1]` and $\beta = $ `r s2.p[2]`.  For $\ell$ prior,  $\alpha = $ `r d.p[1]` and $\beta = $ `r d.p[2]`.
 
@@
 ```r
- V1 <- sapply(1:length(h_grid), function(h){
-      delta * F[[h]] %*% V +  profit(x_grid, h_grid[h]) 
-    })
-    # find havest, h that gives the maximum value
-    out <- sapply(1:gridsize, function(j){
-      value <- max(V1[j,], na.rm = T) # each col is a diff h, max over these
-      index <- which.max(V1[j,])  # store index so we can recover h's 
-      c(value, index) # returns both profit value & index of optimal h.  
-    })
-    # Sets V[t+1] = max_h V[t] at each possible state value, x
-    V <- out[1,]                        # The new value-to-go
-    D[,OptTime-time+1] <- out[2,]       # The index positions
+# compute the value for each possible harvest
+for(h in 1:length(h_grid)){
+  V1[h] = delta * F[[h]] %*% V +  profit(x_grid, h_grid[h]) 
+}
+# find havest h that gives the maximum value
+for(j in 1:gridsize){
+  value = max(V1[j,], na.rm = T)  # each col is a different h, max over these
+  index = which.max(V1[j,])       # store index so we can recover h's 
+  output[,j] = c(value, index)    # returns both profit value & index of optimal h.  
+}
+# Sets V[t+1] = max_h(V[t]) at each possible state value, x
+V = out[1,]                        # The new value-to-go
+D[,OptTime-time+1] = out[2,]       # The index positions
 ```

### 2013-11-19 inst/doc/nonparametric-bayes.tex
added metadata to yaml instead of template

@@
 \begin{frontmatter}
+
   \title{Avoiding tipping points in the management of ecological systems: a
 non-parametric Bayesian approach}
-  \author[cstar]{Carl Boettiger\corref{cor1}}
-  \cortext[cor1]{Corresponding author}
-  \ead{cboettig@ucsc.edu}
+    \author[cstar]{Carl Boettiger\corref{c1}}
+   \ead{cboettig@gmail.com} 
+   \cortext[c1]{Corresponding author}
     \author[cstar]{Marc Mangel}
-  \author[noaa]{Stephan B. Munch}
-  \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA 95064, USA}
-  \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}
- \end{frontmatter}
-
-\section{Abstract}\label{abstract}
-
+
+    \author[noaa]{Stephan Munch}
+
+      \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and
+Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA
+95064, USA}    
+    \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric
+Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}
+
+%  \author[cstar]{Marc Mangel}
+%  \author[noaa]{Stephan B. Munch}
+%  \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA 95064, USA}
+%  \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}
+
+  \begin{abstract}
   Model uncertainty and limited data coverage are fundamental challenges
@@
   uncertainty and limited data.
+  \end{abstract}
+
+ \end{frontmatter}
 
@@
 
-\paragraph{Process, measurement, and implementation
+\subsubsection{Process, measurement, and implementation
 error}\label{process-measurement-and-implementation-error}
@@
 optimal set of feasible actions to maximize the value of some objectives
-(see ({\textbf{???}})). Process error, measurement error, implementation
-error (Reed 1979). These sources of stochasticity in turn mean that
-model parameters can only be estimated approximately, requiring
-parametric uncertainty also be considered (Ludwig and Walters 1982).
-
-\paragraph{Parametric uncertainty}\label{parametric-uncertainty}
+(e.g Halpern et al. (2013)). Process error, measurement error,
+implementation error (Reed 1979). These sources of stochasticity in turn
+mean that model parameters can only be estimated approximately,
+requiring parametric uncertainty also be considered (Ludwig and Walters
+1982).
+
+\subsubsection{Parametric uncertainty}\label{parametric-uncertainty}
 
@@
 uncertainty in the parameter values can be inferred from the data and
-reflected in the estimates of the transition probabilities
-({\textbf{???}}; Mangel and Clark 1988; {\textbf{???}}; Schapaugh and
-Tyre 2013).
-
-\paragraph{Structural (model)
+reflected in the estimates of the transition probabilities (Ludwig and
+Walters 1982; Mangel and Clark 1988; Hilborn and Mangel 1997; Schapaugh
+and Tyre 2013).
+
+\subsubsection{Structural (model)
 uncertainty}\label{structural-model-uncertainty}
@@
 geospatial literature) to formulate a predictive model is relatively new
-in the context of modeling dynamical systems ({\textbf{???}}), and was
-first introduced in the context ecological modeling and fisheries
+in the context of modeling dynamical systems (Kocijan et al. 2005), and
+was first introduced in the context ecological modeling and fisheries
 management in Munch et al. (2005). An accessible and thorough
@@
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Underlying model
-\item
-  Available data
-\end{itemize}
-
-For simplicity we assume profit is simply linear in the realized harvest
-(only enforcing the restriction that harvest can not exceed available
-stock)
+To illustrate the application of the BNP-SDP approach and compare to the
+predictions of the alternative parametric models we focus on the
+classical problem of selecting the appropriate harvest level given an
+observation of the stock size in the previous year (Reed 1979; Ludwig
+and Walters 1982; Mangel and Clark 1988). Given this observation and the
+model (together with the parameter uncertainty) of the stock recruitment
+process, the manager seeks to maximize the value of the fishery over a
+fixed time interval of 50 years at a discount rate of 0.01. The value
+function (profits) at time $t$ depends on the true stock size $x_t$ and
+the chosen harvest level $h_t$. For simplicity we assume profit is
+simply proportional in the realized harvest (only enforcing the
+restriction that harvest can not exceed available stock).
 
@@
 We consider three candidate parametric models of the stock-recruitment
-dynamics: The Ricker model, the Allen model \href{}{Allen 2005}, the
-Myers model. The familiar Ricker model involves two parameters,
-corresponding to a growth rate and a carrying capacity, and cannot
-support alternative stable state dynamics (though as growth rate
+dynamics: The Ricker model, the Allen model (Allen and Tanner 2005), the
+Myers model (Myers et al. 1995). The familiar Ricker model involves two
+parameters, corresponding to a growth rate and a carrying capacity, and
+cannot support alternative stable state dynamics (though as growth rate
 increases it exhibits a periodic attractor that proceeds through
@@
 model resembles the Ricker dynamics with an added Allee effect parameter
-\href{}{Courchamp}, below which the population cannot persist. The Myers
-model also has three parameters and contains an Allee threshold, but has
-compensatory rather than over-compensatory density dependence
-(resembling a Beverton-Holt curve rather than a Ricker curve at high
-densities.)
-
-We assume multiplicative log-normal noise perturbs the growth predicted
-by the each of the deterministic model skeletons described above. This
-introduces one additional parameter $\sigma$ that must be estimated by
-each model.
-
-As we simulate training data from the Allen model (ref section), we will
-refer to this as the structurally correct model. The Ricker model is
-thus a reasonable approximation of these dynamics far from the Allee
-threshold (but lacks threshold dynamics), while the Myers model shares
-the essential feature of a threshold but differs in the structure. Thus
-we have three potential parametric models of the stock dynamics.
+(Courchamp, Berec, and Gascoigne 2008), below which the population
+cannot persist. The Myers model also has three parameters and contains
+an Allee threshold, but has compensatory rather than over-compensatory
+density dependence (resembling a Beverton-Holt curve rather than a
+Ricker curve at high densities.)\\We assume multiplicative log-normal
+noise perturbs the growth predicted by the each of the deterministic
+model skeletons described above. This introduces one additional
+parameter $\sigma$ that must be estimated by each model.
+
+As we simulate training data from the Allen model, we will refer to this
+as the structurally correct model. The Ricker model is thus a reasonable
+approximation of these dynamics far from the Allee threshold (but lacks
+threshold dynamics), while the Myers model shares the essential feature
+of a threshold but differs in the structure. Thus we have three
+potential parametric models of the stock dynamics.
+
+We introduce parameteric uncertainty by first estimating each of the
+candidate models from data on unexploited stock dynamics following some
+pertubation (non-equilibrium initial condition) over several time steps.
+This training data could be generated in several different ways (such as
+known variable exploitation rates, etc), as long as it reflects the
+dynamics in some limited region of state space without impacting the
+problem. We consider a period of 40 years of training data: long enough
+that the estimates are not dependent on the particular realization,
+while longer times are not likely to provide substantial improvement
+(i.e.~the results are not sensitive to this interval). Each of the
+models (described below) is fit to the same training data, as shown in
+Figure 1.
 
@@
 
-In contrast to the parametric models, this posterior distribution is
-still conditional on the training data. As such, the uncertainty near
-the observed data.
-
 We use a Metropolis-Hastings Markov Chain Monte Carlo to infer posterior
-distributions of the two parameters of the GP (Figure S4, code in
+distributions of the two parameters of the GP (Figure S13, code in
 appendix), under weakly informative Gaussian priors (see parameters in
 table S5). As the posterior distributions differ substantially from the
-priors (Figure S4), we can be assured that most of the information in
+priors (Figure S13), we can be assured that most of the information in
 the posterior comes from the data rather than the prior belief.
@@
 
-\% latex table generated in R 3.0.2 by xtable 1.7-1 package \% Mon Nov
-18 14:39:35 2013
-
 \begin{table}[ht]
-\centering
+\begin{center}
 \begin{tabular}{rrrr}
@@
 
-In any modeling effort, models must be chosen for the task at
-hand.\\Though simple mechanistically motivated models offer the greatest
+In any modeling effort, models must be chosen for the task at hand.
+Though simple mechanistically motivated models offer the greatest
 potential to increase our basic understanding of ecological processes
@@
 models can be not only inaccurate but misleading when relied upon in a
-quantitative decision making framework.\\In this paper we have tackled
+quantitative decision making framework. In this paper we have tackled
 two aspects of uncertainty that are both common to many ecological
@@
 
-\subsubsection{Higher dimensions}\label{higher-dimensions}
-
 In this simulated example, the underlying dynamics are truly governed by
@@
 prove even more valuable.
-
-\subsubsection{Real-time learning}\label{real-time-learning}
 
@@
 
+\appendix
+
 \section{Appendix}\label{appendix}
@@
 
-Equation S1: Ricker model.
-
-\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
-
-Figure S1: Ricker model: prior and posterior distributions for parameter
-estimates.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-2.pdf}
+\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \label{rickereq} \]
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-4.pdf}
+\caption{Traces from the MCMC estimate of the Ricker model}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-5.pdf}
 \caption{Posteriors from the MCMC estimate of the Ricker model}
@@
 
-Table S1: Parameterization of the priors
-
-\% latex table generated in R 3.0.2 by xtable 1.7-1 package \% Thu Nov
-14 00:55:33 2013
-
 \begin{table}[ht]
-\centering
+\begin{center}
 \begin{tabular}{rlrr}
   \hline
- & parameter & lower\_bound & upper\_bound \\ 
+ & parameter & lower.bound & upper.bound \\ 
   \hline
-1 & r0 & 0.00 & 10.00 \\ 
-  2 & K & 0.00 & 40.00 \\ 
+1 & r0 & 0.01 & 20.00 \\ 
+  2 & K & 0.01 & 40.00 \\ 
   3 & sigma & 0.00 & 100.00 \\ 
@@
 \end{tabular}
+\caption{parameterization range for the uniform priors in the Ricker model}
+\end{center}
 \end{table}
 
-\[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \]
-
-Eq S2: Myers model Figure S2: Myers model: Traces, prior and posterior
-distributions for parameter estimates.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-4.pdf}
+\begin{enumerate}
+\def\labelenumi{(\arabic{enumi})}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}}\]
+\end{enumerate}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-7.pdf}
 \caption{Traces from the MCMC estimate of the Myers model}
@@
 \centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-5.pdf}
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-8.pdf}
 \caption{Posterior distributions from the MCMC estimates of the Myers
@@
 
-Table S2: Parameterization of the priors \% latex table generated in R
-3.0.2 by xtable 1.7-1 package \% Thu Nov 14 00:55:36 2013
-
 \begin{table}[ht]
-\centering
+\begin{center}
 \begin{tabular}{rlrr}
   \hline
- & parameter & lower\_bound & upper\_bound \\ 
+ & parameter & lower.bound & upper.bound \\ 
   \hline
@@
 \end{tabular}
+\caption{parameterization range for the uniform priors in the Myers model}
+\end{center}
 \end{table}
@@
 
+\begin{enumerate}
+\def\labelenumi{(\arabic{enumi})}
+\setcounter{enumi}{1}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
   \[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]
+\end{enumerate}
 
@@
 \centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-7.pdf}
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-10.pdf}
 \caption{Traces from the MCMC estimate of the Allen model}
@@
 \centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-8.pdf}
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-11.pdf}
 \caption{Posteriors from the MCMC estimate of the Allen model}
@@
 
-Table S3: Parameterization of the priors
-
-\% latex table generated in R 3.0.2 by xtable 1.7-1 package \% Thu Nov
-14 00:55:40 2013
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rlrr}
-  \hline
- & parameter & lower\_bound & upper\_bound \\ 
-  \hline
-1 & r0 & 0.00 & 10.00 \\ 
-  2 & K & 0.00 & 40.00 \\ 
-  3 & theta & 0.00 & 10.00 \\ 
-  4 & sigma & 0.00 & 100.00 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-Eq S4: GP model Figure S4: GP model: prior and posterior distributions
-for parameter estimates.
-
 \begin{verbatim}
-$traces_plot
+Error: object 'allen_priors_xtable' not found
 \end{verbatim}
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-101.pdf}
-\caption{plot of chunk unnamed-chunk-10}
-\end{figure}
-
-\begin{verbatim}
-
-$posteriors_plot
-\end{verbatim}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-102.pdf}
-\caption{plot of chunk unnamed-chunk-10}
-\end{figure}
-
-Table S4: Parameterization of the priors
+Assuming the data are known with some process noise,
+
+\[y = f(x) + \varepsilon\]
+
+$\varepsilon$ IID normal, variance $\sigma_n^2$. Then under the GP,
+
+\[x|y \sim \mathcal{N}(E,C)\]
+\[E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y\]
+\[C = K(X_p, X_p) - K(X_p, X_o) K(X_o,X_o)^{-1} K(X_o, X_p)\]
+
+\}
+
+The marginal likelihood is then given by:
+
+\[\log(p(y | X)) = -\tfrac{1}{2} \mathbf{y}^T (K + \sigma_n^2 \mathbf{I})^{-1} y  - \tfrac{1}{2} \log\left| K + \sigma_n^2 \mathbf{I} \right| - \tfrac{n}{2}\log 2 \pi\]
+
+See Munch, Kottas, and Mangel (2005) or Rasmussen and Williams (2006)
+for a more detailed introduction.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-13.pdf}
+\caption{Traces from the MCMC estimates of the GP model}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-14.pdf}
+\caption{Posterior distributions from the MCMC estimate of the GP model.
+Prior curves shown in red.}
+\end{figure}
+
+The Gaussian process priors on both the lengthscale $\ell$ and process
+noise $\sigma$ are are inverse Gamma distributed,
+
+\[f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)\]
+
+For the $\sigma$ prior, $\alpha = $ 5 and $\beta = $ 5. For $\ell$
+prior, $\alpha = $ 10 and $\beta = $ 10.
 
@@
 \begin{Highlighting}[]
- \NormalTok{V1 <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
-      \NormalTok{delta *}\StringTok{ }\NormalTok{F[[h]] %*%}\StringTok{ }\NormalTok{V +}\StringTok{  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
-    \NormalTok{\})}
-    \CommentTok{# find havest, h that gives the maximum value}
-    \NormalTok{out <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
-      \NormalTok{value <-}\StringTok{ }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
-      \NormalTok{index <-}\StringTok{ }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
-      \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
-    \NormalTok{\})}
-    \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
-    \NormalTok{V <-}\StringTok{ }\NormalTok{out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
-    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <-}\StringTok{ }\NormalTok{out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
+\CommentTok{# compute the value for each possible harvest}
+\NormalTok{for(h in }\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid))\{}
+  \NormalTok{V1[h] =}\StringTok{ }\NormalTok{delta *}\StringTok{ }\NormalTok{F[[h]] %*%}\StringTok{ }\NormalTok{V +}\StringTok{  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
+\NormalTok{\}}
+\CommentTok{# find havest h that gives the maximum value}
+\NormalTok{for(j in }\DecValTok{1}\NormalTok{:gridsize)\{}
+  \NormalTok{value =}\StringTok{ }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T)  }\CommentTok{# each col is a different h, max over these}
+  \NormalTok{index =}\StringTok{ }\KeywordTok{which.max}\NormalTok{(V1[j,])       }\CommentTok{# store index so we can recover h's }
+  \NormalTok{output[,j] =}\StringTok{ }\KeywordTok{c}\NormalTok{(value, index)    }\CommentTok{# returns both profit value & index of optimal h.  }
+\NormalTok{\}}
+\CommentTok{# Sets V[t+1] = max_h(V[t]) at each possible state value, x}
+\NormalTok{V =}\StringTok{ }\NormalTok{out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
+\NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] =}\StringTok{ }\NormalTok{out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
 \end{Highlighting}
@@
 
+Allen, Deborah, and Kimberly Tanner. 2005. ``Infusing Active Learning
+into the Large-enrollment Biology Class: Seven Strategies, from the
+Simple to Complex.'' \emph{Cell Biology Education} 4 (4) (jan): 262--8.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/16344858}.
+
 Athanassoglou, Stergios, and Anastasios Xepapadeas. 2012. ``Pollution
@@
 \url{http://linkinghub.elsevier.com/retrieve/pii/0095069686900240}.
+
+Courchamp, Franck, Ludek Berec, and Joanna Gascoigne. 2008. \emph{Allee
+Effects in Ecology and Conservation}. Oxford University Press, USA.
+\url{http://www.amazon.com/Effects-Ecology-Conservation-Franck-Courchamp/dp/0198570309}.
 
@@
 
+Kocijan, Juš, Agathe Girard, Bla\textbackslash{}vz Banko, and Roderick
+Murray-Smith. 2005. ``Dynamic Systems Identification with Gaussian
+Processes.'' \emph{Mathematical and Computer Modelling of Dynamical
+Systems} 11 (4): 411--424. doi:10.1080/13873950500068567.
+\url{http://www.tandfonline.com/doi/abs/10.1080/13873950500068567}.
+
 Ludwig, Donald, and Carl J. Walters. 1982. ``Optimal Harvesting with
@@
 
+Myers, R. A., N. J. Barrowman, J. A. Hutchings, and Andrew a Rosenberg.
+1995. ``Population Dynamics of Exploited Fish Stocks at Low Population
+Levels.'' \emph{Science (New York, N.Y.)} 269 (5227) (aug): 1106--8.
+doi:10.1126/science.269.5227.1106.
+\url{http://www.ncbi.nlm.nih.gov/pubmed/17755535}.
+
 Polasky, Stephen, Stephen R. Carpenter, Carl Folke, and Bonnie Keeler.

### 2013-11-20 inst/doc/nonparametric-bayes.tex
writing

@@
 \end{tabular}
+\caption{Model choice scores for several common criteria all (wrongly) select the simplest model.
+As the true (Allen) model is not distinguishable from the simpler (Ricker) model in the region of the observed data, this error cannot be avoided regardless of the model choice criterion.
+This highlights the danger of model choice when the selected model will be used outside of the observed range of the data.}
 \end{center}
@@
 
-\section{Appendix}\label{appendix}
-
-\subsection{Model definitions and
+\section{Model definitions and
 estimation}\label{model-definitions-and-estimation}
 
-\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \label{rickereq} \]
+\subsection{Ricker Model}\label{ricker-model}
+
+The Ricker model is given by
+
+\begin{enumerate}
+\def\labelenumi{(\arabic{enumi})}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
+\end{enumerate}
+
+where $Z_t$ is log-normal noise of mean unity and log standard deviation
+$\sigma$, representing the stochastic growth, $X_t$ the stock size at
+time $t$, $S_t$ the escapement (unharvested population that will recruit
+in the following year, $S_t = X_t - h_t$). We place uniform priors on
+the growth rate $r$, carrying capacity $K$, and log-normal standard
+deviation parameter $\sigma$, over ranges given in Table 1. Posteriors
+are inferred by Gibbs sampling using Jags ({\textbf{???}}) (see provided
+code).
 
@@
 \centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-4.pdf}
-\caption{Traces from the MCMC estimate of the Ricker model}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-5.pdf}
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-2.pdf}
 \caption{Posteriors from the MCMC estimate of the Ricker model}
@@
 
+\subsection{Myers Model}\label{myers-model}
+
+The Myers model (Myers et al. 1995) is given by
+
 \begin{enumerate}
 \def\labelenumi{(\arabic{enumi})}
+\setcounter{enumi}{1}
 \itemsep1pt\parskip0pt\parsep0pt
@@
 
+where $Z_t$ is log-normal noise of mean unity and log standard deviation
+$\sigma$, representing the stochastic growth, $X_t$ the stock size at
+time $t$, $S_t$ the escapement (unharvested population that will recruit
+in the following year, $S_t = X_t - h_t$). We place uniform priors on
+the growth rate $r$, carrying capacity $K$, $\theta$ controls the
+strength of the nonlinearity, exhibiting an allee effect for
+$\theta \geq 2$, and log-normal standard deviation parameter $\sigma$,
+over ranges given in Table 1. Posteriors are inferred by Gibbs sampling
+using Jags ({\textbf{???}}) (see code provided).
+
 \begin{figure}[htbp]
 \centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-7.pdf}
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-4.pdf}
 \caption{Traces from the MCMC estimate of the Myers model}
@@
 \centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-8.pdf}
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-5.pdf}
 \caption{Posterior distributions from the MCMC estimates of the Myers
@@
 
-Eq S3: Allen model
+\subsection{Allen model}\label{allen-model}
+
+The Allen model (Allen and Tanner 2005) is given by
 
@@
 \def\labelenumi{(\arabic{enumi})}
-\setcounter{enumi}{1}
+\setcounter{enumi}{2}
 \itemsep1pt\parskip0pt\parsep0pt
@@
 
-Figure S3: Allen model: prior and posterior distributions for parameter
-estimates.
+where $Z_t$ is log-normal noise of mean unity and log standard deviation
+$\sigma$, representing the stochastic growth, $X_t$ the stock size at
+time $t$, $S_t$ the escapement (unharvested population that will recruit
+in the following year, $S_t = X_t - h_t$). We place uniform priors on
+the growth rate $r$, carrying capacity $K$, allee threshold $C$, and
+log-normal standard deviation parameter $\sigma$, over ranges given in
+Table 1. Posteriors are inferred by Gibbs sampling using Jags
+({\textbf{???}}) (see code provided).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-7.pdf}
+\caption{Traces from the MCMC estimate of the Allen model}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/nonparametric-bayes-unnamed-chunk-8.pdf}
+\caption{Posteriors from the MCMC estimate of the Allen model}
+\end{figure}
+
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{rlrr}
+  \hline
+ & parameter & lower.bound & upper.bound \\ 
+  \hline
+1 & r0 & 0.01 & 6.00 \\ 
+  2 & K & 0.01 & 20.00 \\ 
+  3 & theta & 0.01 & 20.00 \\ 
+  4 & sigma & 0.00 & 100.00 \\ 
+   \hline
+\end{tabular}
+\caption{parameterization range for the uniform priors in the Allen model}
+\end{center}
+\end{table}
+
+Assuming the data are known with some process noise,
+
+\[y = f(x) + \varepsilon\]
+
+$\varepsilon$ IID normal, variance $\sigma_n^2$. Then under the GP,
+
+\[x|y \sim \mathcal{N}(E,C)\]
+\[E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y\]
+\[C = K(X_p, X_p) - K(X_p, X_o) K(X_o,X_o)^{-1} K(X_o, X_p)\]
+
+\}
+
+The marginal likelihood is then given by:
+
+\[\log(p(y | X)) = -\tfrac{1}{2} \mathbf{y}^T (K + \sigma_n^2 \mathbf{I})^{-1} y  - \tfrac{1}{2} \log\left| K + \sigma_n^2 \mathbf{I} \right| - \tfrac{n}{2}\log 2 \pi\]
+
+See Munch, Kottas, and Mangel (2005) or Rasmussen and Williams (2006)
+for a more detailed introduction.
 
@@
 \includegraphics{figure/nonparametric-bayes-unnamed-chunk-10.pdf}
-\caption{Traces from the MCMC estimate of the Allen model}
+\caption{Traces from the MCMC estimates of the GP model}
 \end{figure}
@@
 \includegraphics{figure/nonparametric-bayes-unnamed-chunk-11.pdf}
-\caption{Posteriors from the MCMC estimate of the Allen model}
-\end{figure}
-
-\begin{verbatim}
-Error: object 'allen_priors_xtable' not found
-\end{verbatim}
-
-Assuming the data are known with some process noise,
-
-\[y = f(x) + \varepsilon\]
-
-$\varepsilon$ IID normal, variance $\sigma_n^2$. Then under the GP,
-
-\[x|y \sim \mathcal{N}(E,C)\]
-\[E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y\]
-\[C = K(X_p, X_p) - K(X_p, X_o) K(X_o,X_o)^{-1} K(X_o, X_p)\]
-
-\}
-
-The marginal likelihood is then given by:
-
-\[\log(p(y | X)) = -\tfrac{1}{2} \mathbf{y}^T (K + \sigma_n^2 \mathbf{I})^{-1} y  - \tfrac{1}{2} \log\left| K + \sigma_n^2 \mathbf{I} \right| - \tfrac{n}{2}\log 2 \pi\]
-
-See Munch, Kottas, and Mangel (2005) or Rasmussen and Williams (2006)
-for a more detailed introduction.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-13.pdf}
-\caption{Traces from the MCMC estimates of the GP model}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-14.pdf}
 \caption{Posterior distributions from the MCMC estimate of the GP model.
@@
 
-\subsection{Optimal Control Problem}\label{optimal-control-problem}
+\subsection{Optimal Control Problem
+Specification}\label{optimal-control-problem-specification}
 
@@
 
-\subsubsection{Training data}\label{training-data}
+\subsection{Training data}\label{training-data}
 
 Eacho of our models $f(S_t)$ must be estimated from training data, which
-we simulate from the Allen model with parameters $r = $
-\texttt{r p{[}1{]}}, $K =$ \texttt{r p{[}2{]}}, $C =$
-\texttt{r p{[}3{]}}, and $\sigma_g =$ \texttt{r sigma\_g} for $T=$ 40
-timesteps, starting at initial condition $X_0 = $ 5.5. The training data
-can be seen in Figure 1.
-
-\subsubsection{Sensitivity Analysis}\label{sensitivity-analysis}
+we simulate from the Allen model with parameters $r = $ 2, $K =$ 8,
+$C =$ 5, and $\sigma_g =$ 0.05 for $T=$ 40 timesteps, starting at
+initial condition $X_0 = $ 5.5. The training data can be seen in Figure
+1.
+
+\subsection{Code}\label{code}
+
+A copy of the script to reconstruct the simulations and analysis shown
+here is provided in the supplemental materials, and through this
+version-stable link to the project's Github code repostitory,
+\href{}{nonparameteric-bayes.R}. This code is dynamically embedded into
+the manuscript using Knitr, (Xie 2012). The script relies on custom
+routines for executing the estimation of the Gaussian process and the
+for solving the stochastic dynamic programming problem. These routines
+are provided as an R package, \href{}{nonparameteric-bayes}, also
+available on Github.
+
+\subsection{Sensitivity Analysis}\label{sensitivity-analysis}
 
@@
 
+Xie, Yihui. 2012. \emph{Knitr: A General-purpose Package for Dynamic
+Report Generation in R}. \url{http://yihui.name/knitr/}.
+
 \end{document}

### 2013-11-20 inst/examples/BUGS/external-chunks.R

@@ -617,7 +617,7 @@ tmp <- melt(tmp[2:dim(tmp)[2]])
 actual_over_optimal <-subset(tmp, variable != "True")
 
 
-## @knitr Figure4 
+## @knitr Figure4plots 
 fig4v1 <- ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable)) + 
   facet_wrap(~variable, scales = "free_y")  + guides(legend.position = "none") +
   xlab("Total profit by replicate") + scale_fill_manual(values=colorkey) # density plots fail when delta fn

### 2013-11-20 inst/doc/nonparametric-bayes.Rmd
sensitivity analysis text added, code purled

@@
 
-```{r }
+```{r allen_priors_xtable, include=TRUE, echo=FALSE}
 allen_priors_xtable <- data.frame(parameter = c("r0", "K", "theta", "sigma"),
@@
 
-```{r}
+```{r ricker_priors_xtable, include=TRUE, echo=FALSE, cache=FALSE}
 ricker_priors_xtable <- data.frame(
@@
 
-```{r } 
+```{r myers_prior_xtable, include=TRUE, echo=FALSE} 
 myers_priors_xtable <- data.frame(parameter = c("r0", "K", "theta", "sigma"),
@@
 ```{r Table1, dependson="deviances", include=TRUE, results="asis", echo=FALSE, cache=FALSE}
-xtable::xtable(rbind(dictable, aictable, bictable))
+xtable::xtable(rbind(dictable, aictable, bictable), caption="Model choice scores for several common criteria all (wrongly) select the simplest model.
+As the true (Allen) model is not distinguishable from the simpler (Ricker) model in the region of the observed data, this error cannot be avoided regardless of the model choice criterion.
+This highlights the danger of model choice when the selected model will be used outside of the observed range of the data.")
 ```
@@
 Values plotted are the averages of a given policy over 100 replicate simulations.
-Details and code provided in the supplement.",  include=TRUE, echo=FALSE}
-
-```
+Details and code provided in the supplement.",  include=TRUE, echo=FALSE, fig.width=6, fig.height=6}
+
+ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable)) + 
+  facet_wrap(~variable, ncol=2)  + 
+  guides(legend.position = "none") +
+  xlab("Total profit by replicate") + 
+  scale_fill_manual(values=colorkey) # density plots fail when delta fn
+
+# ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable), binwidth=0.1) + 
+#  xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)
+# ggplot(actual_over_optimal, aes(value, fill=variable, color=variable)) + # density plots fail when delta fn
+#  stat_density(aes(y=..density..), position="stack", adjust=3, alpha=.9) + 
+#  xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)+ scale_color_manual(values=colorkey)
+```
+
+Sensitivity Analysis
+--------------------
+
+These results are not sensitive to the modeling details of the simulation.
+The BNP-SDP estimate remains very close to the optimal solution obtained by knowning the true model across changes to the training simulation, noise scale, parameters or structure of the underlying model, as seen in the figure.
+Results are pooled across different random seeds (111, 222, 333), noise values of 0.01 and 0.05 (not large enough to violate the self-sustaining criterion of @Reed1979), and different randomly generated parameters sets for each model, and using either the Myers or Allen as the underlying structure.
+While the BNP-SDP performs slightly closer to optimal on the smaller noise levels, and closer on the Allen model then the Myers model, the differences are small compared to the values observed in using the wrong parameteric model.
+The BNP-SDP solution consistently avoids either failing to harvest at all or crashing the stock, which results in the very low profits/values seen in parameteric models in the previous figure.
+
+![Sensitivity Analysis](figure/sensitivity.pdf)
 
@@
 
-Appendix
-========
-
 Model definitions and estimation
----------------------------------
+=================================
+
+Ricker Model
+------------
 
@@
 
-$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \label{rickereq} $$
+The Ricker model is given by
+
+(@rickereq) $$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
+
+where $Z_t$ is log-normal noise of mean unity and log standard deviation $\sigma$, representing the stochastic growth, $X_t$ the stock size at time $t$, $S_t$ the escapement (unharvested population that will recruit in the following year, $S_t = X_t - h_t$).
+We place uniform priors on the growth rate $r$, carrying capacity $K$, and log-normal standard deviation parameter $\sigma$, over ranges given in Table 1.
+Posteriors are inferred by Gibbs sampling using Jags [@R2Jags] (see provided code).
 
@@
 
+Myers Model
+-----------
+
+The Myers model [@Myers1995] is given by
+
 <!--Eq S2: Myers model -->
 (@myerseq) $$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}}$$
+
+where $Z_t$ is log-normal noise of mean unity and log standard deviation $\sigma$, representing the stochastic growth, $X_t$ the stock size at time $t$, $S_t$ the escapement (unharvested population that will recruit in the following year, $S_t = X_t - h_t$).
+We place uniform priors on the growth rate $r$, carrying capacity $K$, $\theta$ controls the strength of the nonlinearity, exhibiting an allee effect for $\theta \geq 2$, and log-normal standard deviation parameter $\sigma$, over ranges given in Table 1.
+Posteriors are inferred by Gibbs sampling using Jags [@R2Jags] (see code provided).
 
@@
 
-Eq S3: Allen model
+Allen model
+-----------
+
+The Allen model [@Allen2005] is given by
 
@@
 
-Figure S3: Allen model: prior and posterior distributions for parameter estimates.
+where $Z_t$ is log-normal noise of mean unity and log standard deviation $\sigma$, representing the stochastic growth, $X_t$ the stock size at time $t$, $S_t$ the escapement (unharvested population that will recruit in the following year, $S_t = X_t - h_t$).
+We place uniform priors on the growth rate $r$, carrying capacity $K$, allee threshold $C$, and log-normal standard deviation parameter $\sigma$, over ranges given in Table 1.
+Posteriors are inferred by Gibbs sampling using Jags [@R2Jags] (see code provided).
+
+<!-- Figure S3: Allen model: prior and posterior distributions for parameter estimates. -->
 
@@
 
-## Optimal Control Problem
+Optimal Control Problem Specification 
+-------------------------------------
 
@@
 
-### Training data
+Training data
+-------------
 
 Eacho of our models $f(S_t)$ must be estimated from training data, which
-we simulate from the Allen model with parameters $r = $ ` r p[1]`, 
-$K =$ ` r p[2]`, $C =$ ` r p[3]`, and  $\sigma_g =$ ` r sigma_g` 
+we simulate from the Allen model with parameters $r = $ `r p[1]`, 
+$K =$ `r p[2]`, $C =$ `r p[3]`, and  $\sigma_g =$ `r sigma_g` 
 for $T=$ `r Tobs` timesteps, starting at initial condition $X_0 = $ `r Xo`. 
@@
 
-### Sensitivity Analysis
+Code
+----
+
+A copy of the script to reconstruct the simulations and analysis shown here is provided in the supplemental materials, and through this version-stable link to the project's Github code repostitory, [nonparameteric-bayes.R]().
+This code is dynamically embedded into the manuscript using Knitr, [@knitr].
+The script relies on custom routines for executing the estimation of the Gaussian process and the for solving the stochastic dynamic programming problem.
+The sensitivity analysis is provided as a separate script, [sensitivity.Rmd]().
+These routines are provided as an R package, [nonparameteric-bayes](), also available on Github.
+
+Data
+----
+
+While the data can be regenerated using the code provided, for convenience CSV files of the data shown in each graph are also provided, along with appropriate metadata written in the Ecological Metadata Language (EML).

### 2013-11-20 inst/doc/nonparametric-bayes.tex
sensitivity analysis text added, code purled

@@
 determine the best sequence of actions in face of uncertainty, but only
-when that uncertainty can be meaningfully quantified (Fischer et al.
-2009). The sudden collapse of fisheries and other ecosystems has
-increasingly emphasized the difficulties of formulating even
-qualitatively correct models of the underlying processes.
+when that uncertainty can be meaningfully quantified ({\textbf{???}}).
+The sudden collapse of fisheries and other ecosystems has increasingly
+emphasized the difficulties of formulating even qualitatively correct
+models of the underlying processes.
 
@@
 concern have made marine fisheries the crucible for much of the founding
-work (H. S. Gordon 1954; Reed 1979; May et al. 1979; Ludwig and Walters
-1982) in managing ecosystems under uncertainty. Global trends (Worm et
-al. 2006) and controversy (Hilborn 2007; Worm et al. 2009) have made
-understanding these challenges all the more pressing.
+work ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}) in
+managing ecosystems under uncertainty. Global trends ({\textbf{???}})
+and controversy ({\textbf{???}}; {\textbf{???}}) have made understanding
+these challenges all the more pressing.
 
@@
 stochasticity in biological processes, measurements, and implementation
-of policy (\emph{e.g.} Reed 1979; Clark and Kirkwood 1986; Roughgarden
-and Smith 1996; Sethi et al. 2005), parameteric uncertainty (\emph{e.g.}
-Ludwig and Walters 1982; Hilborn and Mangel 1997; McAllister 1998;
-Schapaugh and Tyre 2013), and model or structural uncertainty
-(\emph{e.g.} B. K. Williams 2001; Cressie et al. 2009; Athanassoglou and
-Xepapadeas 2012). Of these, structural uncertainty incorporates the
-least a priori knowledge or assumptions and is generally the hardest to
-quantify. Typical approaches assume a weak notion of model uncertainty
-in which the correct model (or reasonable approximation) of the dynamics
-must be identified from among a handful of alternative models. Here we
-consider an approach that addresses uncertainty at each of these levels
-without assuming the dynamics follow a particular (i.e.~parametric)
-structure.
+of policy ({\textbf{???}}; {\textbf{???}}; {\textbf{???}};
+{\textbf{???}}), parameteric uncertainty ({\textbf{???}};
+{\textbf{???}}; {\textbf{???}}; {\textbf{???}}), and model or structural
+uncertainty ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}). Of these,
+structural uncertainty incorporates the least a priori knowledge or
+assumptions and is generally the hardest to quantify. Typical approaches
+assume a weak notion of model uncertainty in which the correct model (or
+reasonable approximation) of the dynamics must be identified from among
+a handful of alternative models. Here we consider an approach that
+addresses uncertainty at each of these levels without assuming the
+dynamics follow a particular (i.e.~parametric) structure.
 
@@
 optimal set of feasible actions to maximize the value of some objectives
-(e.g Halpern et al. (2013)). Process error, measurement error,
-implementation error (Reed 1979). These sources of stochasticity in turn
-mean that model parameters can only be estimated approximately,
-requiring parametric uncertainty also be considered (Ludwig and Walters
-1982).
+(e.g ({\textbf{???}})). Process error, measurement error, implementation
+error ({\textbf{???}}). These sources of stochasticity in turn mean that
+model parameters can only be estimated approximately, requiring
+parametric uncertainty also be considered ({\textbf{???}}).
 
@@
 uncertainty in the parameter values can be inferred from the data and
-reflected in the estimates of the transition probabilities (Ludwig and
-Walters 1982; Mangel and Clark 1988; Hilborn and Mangel 1997; Schapaugh
-and Tyre 2013).
+reflected in the estimates of the transition probabilities
+({\textbf{???}}; {\textbf{???}}; {\textbf{???}}; {\textbf{???}}).
 
@@
 may be special cases (model averaging being one such way to construct
-such a model) (B. K. Williams 2001; Athanassoglou and Xepapadeas 2012;
-Cressie et al. 2009). Even setting aside other computational and
-statistical concerns (e.g. (Cressie et al. 2009)), these approaches do
-not address our second concern - representing uncertainty outside the
-observed data range.
+such a model) ({\textbf{???}}; {\textbf{???}}; {\textbf{???}}). Even
+setting aside other computational and statistical concerns (e.g.
+({\textbf{???}})), these approaches do not address our second concern -
+representing uncertainty outside the observed data range.
 
@@
 the expected outcomes, but depend on the weights given to all possible
-outcomes (\emph{e.g.} Weitzman 2013). If we are observing the
-fluctuations of a given fish stock over many years under a fixed
-harvesting pressure, we might develop and test a model that could
-reasonably predict the frequency of a deviation of a given size, even
-when such a deviation has not been previously observed. Yet such
-predictions are far less reliable when extrapolated to a harvest
-pressure that has not yet been observed. Thus, model uncertainty can be
-particularly challenging in the management and decision-making context.
+outcomes ({\textbf{???}}). If we are observing the fluctuations of a
+given fish stock over many years under a fixed harvesting pressure, we
+might develop and test a model that could reasonably predict the
+frequency of a deviation of a given size, even when such a deviation has
+not been previously observed. Yet such predictions are far less reliable
+when extrapolated to a harvest pressure that has not yet been observed.
+Thus, model uncertainty can be particularly challenging in the
+management and decision-making context.
 
@@
 than the exception. The potential concern of tipping points in
-ecological dyanmics (Scheffer et al. 2001; Polasky et al. 2011) reflects
-these concerns -- as either knowledge of the true model or more complete
+ecological dyanmics ({\textbf{???}}; {\textbf{???}}) reflects these
+concerns -- as either knowledge of the true model or more complete
 sampling of the state space would make it easy to identify if a tipping
@@
 These dual concerns pose a substantial challenge to existing
-decision-theoretic approaches (Brozović and Schlenker 2011). Because
-intervention is often too late after a tipping point has been crossed
-(but see Hughes et al. (2013)), management is most often concerned with
-avoiding potentially catastrophic tipping points before any data is
-available at or following a transition that would more clearly reveal
-these regime shift dynamics (e.g. Bestelmeyer et al. 2012).
+decision-theoretic approaches ({\textbf{???}}). Because intervention is
+often too late after a tipping point has been crossed (but see
+({\textbf{???}})), management is most often concerned with avoiding
+potentially catastrophic tipping points before any data is available at
+or following a transition that would more clearly reveal these regime
+shift dynamics ({\textbf{???}}).
 
 Here we illustrate how a stochastic dynamic programming (SDP) algorithm
-(Mangel and Clark 1988; Marescot et al. 2013) can be driven by the
-predictions from a Bayesian non-parametric (BNP) approach (Munch,
-Kottas, and Mangel 2005). This provides two distinct advantages compared
-with contemporary approaches. First, using a BNP sidesteps the need for
-an accurate model-based description of the system dynamics. Second, the
-BNP can better reflect uncertainty that arises when extrapolating a
-model outside of the data on which it was fit. We illustrate that when
-the correct model is not known, this latter feature is crucial to
-providing a robust decision-theoretic approach in face of substantial
-structural uncertainty.
+({\textbf{???}}; {\textbf{???}}) can be driven by the predictions from a
+Bayesian non-parametric (BNP) approach ({\textbf{???}}). This provides
+two distinct advantages compared with contemporary approaches. First,
+using a BNP sidesteps the need for an accurate model-based description
+of the system dynamics. Second, the BNP can better reflect uncertainty
+that arises when extrapolating a model outside of the data on which it
+was fit. We illustrate that when the correct model is not known, this
+latter feature is crucial to providing a robust decision-theoretic
+approach in face of substantial structural uncertainty.
 
@@
 social value of fish left in the sea vs.~the commercial value of fish
-harvested; see Halpern et al. (2013) for further discussion of this
-issue. Alternatively, we can always compare model performance across
-multiple potential objective functions.) The decision-maker does not
-necessarily need a model that provides the best mechanistic
-understanding or the best long-term outcome, but rather the one that
-best estimates the probabilities of being in different states as a
-result of the possible actions.
+harvested; see ({\textbf{???}}) for further discussion of this issue.
+Alternatively, we can always compare model performance across multiple
+potential objective functions.) The decision-maker does not necessarily
+need a model that provides the best mechanistic understanding or the
+best long-term outcome, but rather the one that best estimates the
+probabilities of being in different states as a result of the possible
+actions.
 
@@
 geospatial literature) to formulate a predictive model is relatively new
-in the context of modeling dynamical systems (Kocijan et al. 2005), and
-was first introduced in the context ecological modeling and fisheries
-management in Munch et al. (2005). An accessible and thorough
-introduction to the formulation and use of GPs can be found in Rasmussen
-and Williams (2006).
+in the context of modeling dynamical systems ({\textbf{???}}), and was
+first introduced in the context ecological modeling and fisheries
+management in ({\textbf{???}}). An accessible and thorough introduction
+to the formulation and use of GPs can be found in ({\textbf{???}}).
 
@@
 process model are estimated by Metropolis-Hastings algorithm, again with
-details and code provided in the Appendix. Rasmussen and Williams (2006)
-provides an excellent general introduction to Gaussian Processes and
-Munch et al. (2005) first discusses their application in the context of
+details and code provided in the Appendix. ({\textbf{???}}) provides an
+excellent general introduction to Gaussian Processes and
+({\textbf{???}}) first discusses their application in the context of
 population dynamics models such as fisheries stock-recruitment
@@
 classical problem of selecting the appropriate harvest level given an
-observation of the stock size in the previous year (Reed 1979; Ludwig
-and Walters 1982; Mangel and Clark 1988). Given this observation and the
-model (together with the parameter uncertainty) of the stock recruitment
+observation of the stock size in the previous year ({\textbf{???}};
+{\textbf{???}}; {\textbf{???}}). Given this observation and the model
+(together with the parameter uncertainty) of the stock recruitment
 process, the manager seeks to maximize the value of the fishery over a
@@
 We consider three candidate parametric models of the stock-recruitment
-dynamics: The Ricker model, the Allen model (Allen and Tanner 2005), the
-Myers model (Myers et al. 1995). The familiar Ricker model involves two
+dynamics: The Ricker model, the Allen model ({\textbf{???}}), the Myers
+model ({\textbf{???}}). The familiar Ricker model involves two
 parameters, corresponding to a growth rate and a carrying capacity, and
@@
 model resembles the Ricker dynamics with an added Allee effect parameter
-(Courchamp, Berec, and Gascoigne 2008), below which the population
-cannot persist. The Myers model also has three parameters and contains
-an Allee threshold, but has compensatory rather than over-compensatory
-density dependence (resembling a Beverton-Holt curve rather than a
-Ricker curve at high densities.)\\We assume multiplicative log-normal
-noise perturbs the growth predicted by the each of the deterministic
-model skeletons described above. This introduces one additional
-parameter $\sigma$ that must be estimated by each model.
+({\textbf{???}}), below which the population cannot persist. The Myers
+model also has three parameters and contains an Allee threshold, but has
+compensatory rather than over-compensatory density dependence
+(resembling a Beverton-Holt curve rather than a Ricker curve at high
+densities.)\\We assume multiplicative log-normal noise perturbs the
+growth predicted by the each of the deterministic model skeletons
+described above. This introduces one additional parameter $\sigma$ that
+must be estimated by each model.
 
@@
 We infer posterior distributions for the parameters of each model in a
-Bayesian context using Gibbs sampling (implemented in R (Team 2013)
-using jags, (Su and Yajima 2013)). We choose uninformative uniform
-priors for all parameters (See Appendix, Figures S1-S3, and Table S1,
-and the R code provided). One-step-ahead predictions of these model fits
-are shown in Figure 1.
+Bayesian context using Gibbs sampling (implemented in R ({\textbf{???}})
+using jags, ({\textbf{???}})). We choose uninformative uniform priors
+for all parameters (See Appendix, Figures S1-S3, and Table S1, and the R
+code provided). One-step-ahead predictions of these model fits are shown
+in Figure 1.
 
@@
 values $X_{t+1}$ given an observation of the previous state, $X_t$.
-Munch, Kottas, and Mangel (2005) gives an accessible introduction to the
-use of Gaussian Processes in providing a Bayesian nonparametric
-description of the stock-recruitment relationship.
+({\textbf{???}}) gives an accessible introduction to the use of Gaussian
+Processes in providing a Bayesian nonparametric description of the
+stock-recruitment relationship.
 
@@
 $S_t$, as under models such as this a constant escapement policy is
-expected to be optimal (Reed 1979).}
+expected to be optimal ({\textbf{???}}).}
 \end{figure}
@@
 such as this a constant escapement policy is expected to be optimal
-(Reed 1979), whereby population levels below a certain size $S$ are
+({\textbf{???}}), whereby population levels below a certain size $S$ are
 unharvested, while above that size the harvest strategy aims to return
@@
 potential to increase our basic understanding of ecological processes
-(Cuddington, Fortin, and Gerber 2013; Geritz and Kisdi 2012), such
-models can be not only inaccurate but misleading when relied upon in a
-quantitative decision making framework. In this paper we have tackled
-two aspects of uncertainty that are both common to many ecological
-desicion-making problems and fundamentally challenging to existing
-approaches which largely rely on parametric models:
+({\textbf{???}}; {\textbf{???}}), such models can be not only inaccurate
+but misleading when relied upon in a quantitative decision making
+framework. In this paper we have tackled two aspects of uncertainty that
+are both common to many ecological desicion-making problems and
+fundamentally challenging to existing approaches which largely rely on
+parametric models:
 
@@
 suggest rather opposite conclusions than what we observe here. Such
-adaptive probing or Dual Control (e.g. Ludwig and Walters (1982))
-approaches trade off short term utility by choosing actions that can
-reduce uncertainty. Adaptive probing strategies arise when it is
-valuable to intentionally force a system far from the observed values
-even when the expected value such actions is low, as it provides much
-faster learning and consequent reduction of model uncertainty that can
-allow greater value to be derived later on. For instance, Ludwig and
-Walters (1982) show that it may be advantageous to fish an unexploited
-population very heavily at first to obtain a better estimate of the
-recruitment rate. This intuitive strategy when a population is governed
-by a Ricker or Beverton-Holt-like dynamic would clearly be disastrous if
-instead the dynamics contained an unforeseen tipping point. The best way
-to learn where the edge lies may be to walk up to it, but it is also the
-most dangerous. Future work should attempt to understand when such
-active adaptive learning is valuable, and when it will increase the risk
-of an irreversible transition.
+adaptive probing or Dual Control (e.g. ({\textbf{???}})) approaches
+trade off short term utility by choosing actions that can reduce
+uncertainty. Adaptive probing strategies arise when it is valuable to
+intentionally force a system far from the observed values even when the
+expected value such actions is low, as it provides much faster learning
+and consequent reduction of model uncertainty that can allow greater
+value to be derived later on. For instance, ({\textbf{???}}) show that
+it may be advantageous to fish an unexploited population very heavily at
+first to obtain a better estimate of the recruitment rate. This
+intuitive strategy when a population is governed by a Ricker or
+Beverton-Holt-like dynamic would clearly be disastrous if instead the
+dynamics contained an unforeseen tipping point. The best way to learn
+where the edge lies may be to walk up to it, but it is also the most
+dangerous. Future work should attempt to understand when such active
+adaptive learning is valuable, and when it will increase the risk of an
+irreversible transition.
 
@@
 
-The Myers model (Myers et al. 1995) is given by
+The Myers model ({\textbf{???}}) is given by
 
@@
 
-The Allen model (Allen and Tanner 2005) is given by
+The Allen model ({\textbf{???}}) is given by
 
@@
 
-See Munch, Kottas, and Mangel (2005) or Rasmussen and Williams (2006)
-for a more detailed introduction.
+See ({\textbf{???}}) or ({\textbf{???}}) for a more detailed
+introduction.
 
@@
 \href{}{nonparameteric-bayes.R}. This code is dynamically embedded into
-the manuscript using Knitr, (Xie 2012). The script relies on custom
-routines for executing the estimation of the Gaussian process and the
-for solving the stochastic dynamic programming problem. These routines
-are provided as an R package, \href{}{nonparameteric-bayes}, also
-available on Github.
+the manuscript using Knitr, ({\textbf{???}}). The script relies on
+custom routines for executing the estimation of the Gaussian process and
+the for solving the stochastic dynamic programming problem. These
+routines are provided as an R package, \href{}{nonparameteric-bayes},
+also available on Github.
 
@@
 
-Allen, Deborah, and Kimberly Tanner. 2005. ``Infusing Active Learning
-into the Large-enrollment Biology Class: Seven Strategies, from the
-Simple to Complex.'' \emph{Cell Biology Education} 4 (4) (jan): 262--8.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/16344858}.
-
-Athanassoglou, Stergios, and Anastasios Xepapadeas. 2012. ``Pollution
-Control with Uncertain Stock Dynamics: When, and How, to Be
-Precautious.'' \emph{Journal of Environmental Economics and Management}
-63 (3) (may): 304--320. doi:10.1016/j.jeem.2011.11.001.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069611001409}.
-
-Bestelmeyer, Brandon T., Michael C. Duniway, Darren K. James, Laura M.
-Burkett, and Kris M. Havstad. 2012. ``A Test of Critical Thresholds and
-Their Indicators in a Desertification-prone Ecosystem: More Resilience
-Than We Thought.'' Edited by Katharine Suding. \emph{Ecology Letters}
-(dec). doi:10.1111/ele.12045.
-\url{http://doi.wiley.com/10.1111/ele.12045}.
-
-Brozović, Nicholas, and Wolfram Schlenker. 2011. ``Optimal Management of
-an Ecosystem with an Unknown Threshold.'' \emph{Ecological Economics}
-(jan): 1--14. doi:10.1016/j.ecolecon.2010.10.001.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0921800910004167}.
-
-Clark, Colin W., and Geoffrey P. Kirkwood. 1986. ``On Uncertain
-Renewable Resource Stocks: Optimal Harvest Policies and the Value of
-Stock Surveys.'' \emph{Journal of Environmental Economics and
-Management} 13 (3) (sep): 235--244. doi:10.1016/0095-0696(86)90024-0.
-\url{http://linkinghub.elsevier.com/retrieve/pii/0095069686900240}.
-
-Courchamp, Franck, Ludek Berec, and Joanna Gascoigne. 2008. \emph{Allee
-Effects in Ecology and Conservation}. Oxford University Press, USA.
-\url{http://www.amazon.com/Effects-Ecology-Conservation-Franck-Courchamp/dp/0198570309}.
-
-Cressie, Noel, Catherine a Calder, James S. Clark, Jay M. Ver Hoef, and
-Christopher K. Wikle. 2009. ``Accounting for Uncertainty in Ecological
-Analysis: the Strengths and Limitations of Hierarchical Statistical
-Modeling.'' \emph{Ecological Applications : a Publication of the
-Ecological Society of America} 19 (3) (apr): 553--70.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/19425416}.
-
-Cuddington, Kim M., M. J. Fortin, and L. R. Gerber. 2013.
-``Process-based Models Are Required to Manage Ecological Systems in a
-Changing World.'' \emph{Ecosphere} 4 (February): 1--12.
-doi:10.1890/ES12-00178.1.
-\url{http://www.esajournals.org/doi/abs/10.1890/ES12-00178.1}.
-
-Fischer, Joern, Garry D. Peterson, Toby a Gardner, Line J. Gordon, Ioan
-Fazey, Thomas Elmqvist, Adam Felton, Carl Folke, and Stephen Dovers.
-2009. ``Integrating Resilience Thinking and Optimisation for
-Conservation.'' \emph{Trends in Ecology \& Evolution} 24 (10) (oct):
-549--54. doi:10.1016/j.tree.2009.03.020.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/19665820}.
-
-Geritz, Stefan A. H., and Eva Kisdi. 2012. ``Mathematical Ecology: Why
-Mechanistic Models?'' \emph{Journal of Mathematical Biology} 65 (6-7)
-(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.
-
-Gordon, H. S. 1954. ``The Economic Theory of a Common-property Resource:
-the Fishery.'' \emph{The Journal of Political Economy} 62 (2): 124--142.
-\url{http://www.jstor.org/stable/10.2307/1825571}.
-
-Halpern, Benjamin S., Carissa J. Klein, Christopher J. Brown, Maria
-Beger, Hedley S. Grantham, Sangeeta Mangubhai, Mary Ruckelshaus, et al.
-2013. ``Achieving the Triple Bottom Line in the Face of Inherent
-Trade-offs Among Social Equity, Economic Return, and Conservation.''
-\emph{Proceedings of the National Academy of Sciences of the United
-States of America} 110 (15) (apr): 6229--34.
-doi:10.1073/pnas.1217689110.
-\href{http://www.ncbi.nlm.nih.gov/pubmed/23530207 http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=3625307/\&tool=pmcentrez/\&rendertype=abstract}{http://www.ncbi.nlm.nih.gov/pubmed/23530207
-http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=3625307\textbackslash{}\&tool=pmcentrez\textbackslash{}\&rendertype=abstract}.
-
-Hilborn, Ray. 2007. ``Reinterpreting the State of Fisheries and Their
-Management.'' \emph{Ecosystems} 10 (8) (oct): 1362--1369.
-doi:10.1007/s10021-007-9100-5.
-\url{http://www.springerlink.com/index/10.1007/s10021-007-9100-5}.
-
-Hilborn, Ray, and Marc Mangel. 1997. \emph{The Ecological Detective:
-Confronting Models with Data}. Princeton University Press.
-\url{http://www.amazon.com/The-Ecological-Detective-Confronting-Models/dp/0691034974}.
-
-Hughes, Terry P., Cristina Linares, Vasilis Dakos, Ingrid a van de
-Leemput, and Egbert H. van Nes. 2013. ``Living Dangerously on Borrowed
-Time During Slow, Unrecognized Regime Shifts.'' \emph{Trends in Ecology
-\& Evolution} 28 (3) (mar): 149--55. doi:10.1016/j.tree.2012.08.022.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/22995893}.
-
-Kocijan, Juš, Agathe Girard, Bla\textbackslash{}vz Banko, and Roderick
-Murray-Smith. 2005. ``Dynamic Systems Identification with Gaussian
-Processes.'' \emph{Mathematical and Computer Modelling of Dynamical
-Systems} 11 (4): 411--424. doi:10.1080/13873950500068567.
-\url{http://www.tandfonline.com/doi/abs/10.1080/13873950500068567}.
-
-Ludwig, Donald, and Carl J. Walters. 1982. ``Optimal Harvesting with
-Imprecise Parameter Estimates.'' \emph{Ecological Modelling} 14 (3-4)
-(jan): 273--292. doi:10.1016/0304-3800(82)90023-0.
-\url{http://linkinghub.elsevier.com/retrieve/pii/0304380082900230}.
-
-Mangel, Marc, and Colin W. Clark. 1988. \emph{Dynamic Modeling in
-Behavioral Ecology}. Edited by John Krebs and Tim Clutton-Brock.
-Princeton: Princeton University Press.
-
-Marescot, Lucile, Guillaume Chapron, Iadine Chadès, Paul L. Fackler,
-Christophe Duchamp, Eric Marboutin, and Olivier Gimenez. 2013. ``Complex
-Decisions Made Simple: a Primer on Stochastic Dynamic Programming.''
-\emph{Methods in Ecology and Evolution} (jun).
-doi:10.1111/2041-210X.12082.
-\url{http://doi.wiley.com/10.1111/2041-210X.12082}.
-
-May, Robert M., John R. Beddington, Colin W. Clark, Sidney J. Holt, and
-R. M. Laws. 1979. ``Management of Multispecies Fisheries.''
-\emph{Science (New York, N.Y.)} 205 (4403) (jul): 267--77.
-doi:10.1126/science.205.4403.267.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/17747032}.
-
-McAllister, M. K. 1998. ``Bayesian Stock Assessment: a Review and
-Example Application Using the Logistic Model.'' \emph{ICES Journal of
-Marine Science} 55 (6) (dec): 1031--1060. doi:10.1006/jmsc.1998.0425.
-\href{http://icesjms.oxfordjournals.org/content/55/6/1031.short http://icesjms.oxfordjournals.org/cgi/doi/10.1006/jmsc.1998.0425}{http://icesjms.oxfordjournals.org/content/55/6/1031.short
-http://icesjms.oxfordjournals.org/cgi/doi/10.1006/jmsc.1998.0425}.
-
-Munch, Stephan B., Athanasios Kottas, and Marc Mangel. 2005. ``Bayesian
-Nonparametric Analysis of Stockrecruitment Relationships.''
-\emph{Canadian Journal of Fisheries and Aquatic Sciences} 62 (8) (aug):
-1808--1821. doi:10.1139/f05-073.
-\url{http://www.nrcresearchpress.com/doi/abs/10.1139/f05-073}.
-
-Munch, Stephan B., Melissa L. Snover, George M. Watters, and Marc
-Mangel. 2005. ``A Unified Treatment of Top-down and Bottom-up Control of
-Reproduction in Populations.'' \emph{Ecology Letters} 8 (7) (may):
-691--695. doi:10.1111/j.1461-0248.2005.00766.x.
-\url{http://doi.wiley.com/10.1111/j.1461-0248.2005.00766.x}.
-
-Myers, R. A., N. J. Barrowman, J. A. Hutchings, and Andrew a Rosenberg.
-1995. ``Population Dynamics of Exploited Fish Stocks at Low Population
-Levels.'' \emph{Science (New York, N.Y.)} 269 (5227) (aug): 1106--8.
-doi:10.1126/science.269.5227.1106.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/17755535}.
-
-Polasky, Stephen, Stephen R. Carpenter, Carl Folke, and Bonnie Keeler.
-2011. ``Decision-making Under Great Uncertainty: Environmental
-Management in an Era of Global Change.'' \emph{Trends in Ecology \&
-Evolution} (may): 1--7. doi:10.1016/j.tree.2011.04.007.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/21616553}.
-
-Rasmussen, Carl Edward, and C. K. I. Williams. 2006. \emph{Gaussian
-Processes for Machine Learning}. Edited by Thomas Dietterich. Boston:
-MIT Press,. \url{www.GaussianProcess.org/gpml}.
-
-Reed, William J. 1979. ``Optimal Escapement Levels in Stochastic and
-Deterministic Harvesting Models.'' \emph{Journal of Environmental
-Economics and Management} 6 (4) (dec): 350--363.
-doi:10.1016/0095-0696(79)90014-7.
-\href{http://www.sciencedirect.com/science/article/pii/0095069679900147 http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}{http://www.sciencedirect.com/science/article/pii/0095069679900147
-http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}.
-
-Roughgarden, Joan E., and F. Smith. 1996. ``Why Fisheries Collapse and
-What to Do About It.'' \emph{Proceedings of the National Academy of
-Sciences of the United States of America} 93 (10): 5078.
-\url{http://www.pnas.org/content/93/10/5078.short}.
-
-Schapaugh, Adam W., and Andrew J. Tyre. 2013. ``Accounting for
-Parametric Uncertainty in Markov Decision Processes.'' \emph{Ecological
-Modelling} 254 (apr): 15--21. doi:10.1016/j.ecolmodel.2013.01.003.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0304380013000306}.
-
-Scheffer, Marten, Stephen R. Carpenter, J. A. Foley, C. Folke, and B.
-Walker. 2001. ``Catastrophic Shifts in Ecosystems.'' \emph{Nature} 413
-(6856) (oct): 591--6. doi:10.1038/35098000.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/11595939}.
-
-Sethi, Gautam, Christopher Costello, Anthony Fisher, Michael Hanemann,
-and Larry Karp. 2005. ``Fishery Management Under Multiple Uncertainty.''
-\emph{Journal of Environmental Economics and Management} 50 (2) (sep):
-300--318. doi:10.1016/j.jeem.2004.11.005.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069605000057}.
-
-Su, Yu-Sung, and Masanao Yajima. 2013. \emph{R2jags: A Package for
-Running Jags from R}. \url{http://cran.r-project.org/package=R2jags}.
-
-Team, R. Core. 2013. \emph{R: A Language and Environment for Statistical
-Computing}. Vienna, Austria: R Foundation for Statistical Computing.
-\url{http://www.r-project.org/}.
-
-Weitzman, Martin L. 2013. ``A Precautionary Tale of Uncertain Tail
-Fattening.'' \emph{Environmental and Resource Economics} 55 (2) (mar):
-159--173. doi:10.1007/s10640-013-9646-y.
-\url{http://link.springer.com/10.1007/s10640-013-9646-y}.
-
-Williams, Byron K. 2001. ``Uncertainty , Learning , and the Optimal
-Management of Wildlife.'' \emph{Environmental and Ecological Statistics}
-8: 269--288. doi:10.1023/A:1011395725123.
-
-Worm, Boris, Edward B. Barbier, Nicola Beaumont, J. Emmett Duffy, Carl
-Folke, Benjamin S. Halpern, Jeremy B. C. Jackson, et al. 2006. ``Impacts
-of Biodiversity Loss on Ocean Ecosystem Services.'' \emph{Science (New
-York, N.Y.)} 314 (5800) (nov): 787--90. doi:10.1126/science.1132294.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/17082450}.
-
-Worm, Boris, Ray Hilborn, Julia K. Baum, Trevor A. Branch, Jeremy S.
-Collie, Christopher Costello, Michael J. Fogarty, et al. 2009.
-``Rebuilding Global Fisheries.'' \emph{Science (New York, N.Y.)} 325
-(5940) (jul): 578--85. doi:10.1126/science.1173146.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/19644114}.
-
-Xie, Yihui. 2012. \emph{Knitr: A General-purpose Package for Dynamic
-Report Generation in R}. \url{http://yihui.name/knitr/}.
+These results are not sensitive to the modeling details of the
+simulation. The Gaussian process estimate remains very close to the
+optimal solution obtained by knowning the true model across changes to
+the training simulation, noise scale, parameters or structure of the
+underlying model, as seen in the figure. Results are pooled across
+different random seeds (111, 222, 333), noise values of 0.01, 0.05, 0.1,
+and 0.2, 3 different randomly generated parameters sets for each model,
+and using either the Myers or Allen as the underlying structure.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/sensitivity.pdf}
+\caption{Sensitivity Analysis}
+\end{figure}
 

### 2013-11-20 inst/doc/nonparametric-bayes.R

@@ -1,0 +1,671 @@
+
+## @knitr plotting-options
+library(knitr)
+source("~/.knitr_defaults.R")
+#opts_knit$set(upload.fun = socialR::flickr.url)
+library(knitcitations)
+library(nonparametricbayes) 
+
+
+## @knitr posterior-mode
+require(modeest)
+posterior.mode <- function(x) {
+  mlv(x, method="shorth")$M
+}
+
+
+
+## @knitr stateeq
+f <- RickerAllee
+p <- c(2, 8, 5)
+K <- 10  # approx, a li'l' less
+allee <- 5 # approx, a li'l' less
+
+
+## @knitr sdp-pars
+sigma_g <- 0.05
+sigma_m <- 0.0
+z_g <- function() rlnorm(1, 0, sigma_g)
+z_m <- function() 1
+x_grid <- seq(0, 1.5 * K, length=50)
+h_grid <- x_grid
+profit <- function(x,h) pmin(x, h)
+delta <- 0.01
+OptTime <- 50  # stationarity with unstable models is tricky thing
+reward <- 0
+xT <- 0
+Xo <-  allee+.5# observations start from
+x0 <- K # simulation under policy starts from
+Tobs <- 40
+MaxT = 1000 # timeout for value iteration convergence
+
+
+## @knitr obs
+  set.seed(1234)
+  #harvest <- sort(rep(seq(0, .5, length=7), 5))
+  x <- numeric(Tobs)
+  x[1] <- Xo
+  nz <- 1
+  for(t in 1:(Tobs-1))
+    x[t+1] = z_g() * f(x[t], h=0, p=p)
+  obs <- data.frame(x = c(rep(0,nz), 
+                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
+                    y = c(rep(0,nz), 
+                          x[2:Tobs]))
+raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
+raw_plot
+
+
+## @knitr mle
+set.seed(12345)
+estf <- function(p){ 
+    mu <- f(obs$x,0,p)
+    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
+}
+par <- c(p[1]*rlnorm(1,0,.1), 
+         p[2]*rlnorm(1,0,.1), 
+         p[3]*rlnorm(1,0, .1), 
[…]

### 2013-11-20 inst/doc/nonparametric-bayes.Rmd

@@
 ```
-
-Sensitivity Analysis
---------------------
-
-These results are not sensitive to the modeling details of the simulation.
-The BNP-SDP estimate remains very close to the optimal solution obtained by knowning the true model across changes to the training simulation, noise scale, parameters or structure of the underlying model, as seen in the figure.
-Results are pooled across different random seeds (111, 222, 333), noise values of 0.01 and 0.05 (not large enough to violate the self-sustaining criterion of @Reed1979), and different randomly generated parameters sets for each model, and using either the Myers or Allen as the underlying structure.
-While the BNP-SDP performs slightly closer to optimal on the smaller noise levels, and closer on the Allen model then the Myers model, the differences are small compared to the values observed in using the wrong parameteric model.
-The BNP-SDP solution consistently avoids either failing to harvest at all or crashing the stock, which results in the very low profits/values seen in parameteric models in the previous figure.
-
-![Sensitivity Analysis](figure/sensitivity.pdf)
 
@@
 The script relies on custom routines for executing the estimation of the Gaussian process and the for solving the stochastic dynamic programming problem.
-The sensitivity analysis is provided as a separate script, [sensitivity.Rmd]().
 These routines are provided as an R package, [nonparameteric-bayes](), also available on Github.
 
-Data
-----
-
-While the data can be regenerated using the code provided, for convenience CSV files of the data shown in each graph are also provided, along with appropriate metadata written in the Ecological Metadata Language (EML).
+Sensitivity Analysis
+--------------------
+
+These results are not sensitive to the modeling details of the simulation.
+The Gaussian process estimate remains very close to the optimal solution obtained by knowning the true model across changes to the training simulation, noise scale, parameters or structure of the underlying model, as seen in the figure.
+Results are pooled across different random seeds (111, 222, 333), noise values of 0.01, 0.05, 0.1, and 0.2, 3 different randomly generated parameters sets for each model, and using either the Myers or Allen as the underlying structure.
+
+![Sensitivity Analysis](figure/sensitivity.pdf)

### 2013-11-20 inst/doc/nonparametric-bayes.Rmd
draft to marc and steve

--- 2013-11-20 inst/doc/nonparametric-bayes.Rmd
@@
 
-We develop these concerns in the context of fisheries, though
+We develop a novel approach to address these concerns in the context of fisheries; though
 the underlying challenges and methods are germane to many other
@@
 implementation of policy [_e.g._ @Reed1979; @Clark1986; @Roughgarden1996;
-@Sethi2005], parameteric uncertainty [_e.g._ @Ludwig1982; @Hilborn1997;
+@Sethi2005], parametric uncertainty [_e.g._ @Ludwig1982; @Hilborn1997;
 @McAllister1998; @Schapaugh2013], and model or structural uncertainty
@@
 
-_Cut the next three paragraphs, since they are covered more consisely in the above paragraph?_
+_Cut the next three paragraphs, since they are covered more concisely in the above paragraph?_
 
@@
 than the exception. The potential concern of tipping points in ecological
-dyanmics [@Scheffer2001; @Polasky2011] reflects these concerns -- as
+dynamics [@Scheffer2001; @Polasky2011] reflects these concerns -- as
 either knowledge of the true model or more complete sampling of the
@@
 This permits a much greater uncertainty far from the observed data than
-near the observed data.  These features allow the GP-SDP approach to
+near the observed data.  These features allow the BNP-SDP approach to
 find robust management solutions in face of limited data and without
@@
 distributions alone. Either the model scales with the complexity of the data
-on which it is estimated (e.g. nonparametric heirarchical approaches such
+on which it is estimated (e.g. nonparametric hierarchical approaches such
 as the Dirchlet process) or the data points become themselves part of the model
@@
 details and code provided in the Appendix.  @Rasmussen2006 provides
-an excellent general introduction to Gaussian Processes and @Munch2005 
+an excellent general introduction to Gaussian Processes and @Munch2005a 
 first discusses their application in the context of population dynamics
@@
 ====================
-
-<!-- skip this oveview? 
-### Summary of approach
--->
 
@@
 
-We introduce parameteric uncertainty by first estimating each of the
+We introduce parametric uncertainty by first estimating each of the
 candidate models from data on unexploited stock dynamics following
-some pertubation (non-equilibrium initial condition) over several time
+some perturbation (non-equilibrium initial condition) over several time
 steps. This training data could be generated in several different ways
-(such as known variable exploitation rates, etc), as long as it reflects
+(such as known variable exploitation rates, etc.), as long as it reflects
 the dynamics in some limited region of state space without impacting
@@
 The Ricker model always predicts population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).
-Note that the GP is both more pessimistic and more uncertain about the future state than the parameteric models, while the realized state often falls outside of the expected range forecasted by the structurally incorrect Myers and Ricker models." }
+Note that the GP is both more pessimistic and more uncertain about the future state than the paramteric models, while the realized state often falls outside of the expected range forecast by the structurally incorrect Myers and Ricker models." }
 ```
@@
 Results are pooled across different random seeds (111, 222, 333), noise values of 0.01 and 0.05 (not large enough to violate the self-sustaining criterion of @Reed1979), and different randomly generated parameters sets for each model, and using either the Myers or Allen as the underlying structure.
-While the BNP-SDP performs slightly closer to optimal on the smaller noise levels, and closer on the Allen model then the Myers model, the differences are small compared to the values observed in using the wrong parameteric model.
-The BNP-SDP solution consistently avoids either failing to harvest at all or crashing the stock, which results in the very low profits/values seen in parameteric models in the previous figure.
+While the BNP-SDP performs slightly closer to optimal on the smaller noise levels, and closer on the Allen model then the Myers model, the differences are small compared to the values observed in using the wrong parametric model.
+The BNP-SDP solution consistently avoids either failing to harvest at all or crashing the stock, which results in the very low profits/values seen in parametric models in the previous figure.
 
@@
 Despite this flexibility, the GP can be described in relatively few 
-parameters and is thus far less likely to overfit.
+parameters and is thus far less likely to over-fit.
 
@@
 be influenced by the curvature of the value (profit) function $\Pi$. Both
-to simplify the intution and avoid biasing this result, we have chosen
+to simplify the intuition and avoid biasing this result, we have chosen
 a profits that are linear in the catch and thus neither risk-prone nor
@@
 assumption of diminishing returns, would make the SDP more risk-adverse
-(as larger-than-expected stock sizes offer dimished returns relative
+(as larger-than-expected stock sizes offer diminished returns relative
 to the cost of smaller-than-expected stock sizes), and strengthen the
@@
 
-A copy of the script to reconstruct the simulations and analysis shown here is provided in the supplemental materials, and through this version-stable link to the project's Github code repostitory, [nonparameteric-bayes.R]().
+A copy of the script to reconstruct the simulations and analysis shown here is provided in the supplemental materials, and through this version-stable link to the project's Github code repostitory, [nonparameteric-bayes.R](https://github.com/cboettig/nonparametric-bayes/blob/371114ec43bb32e92f330f049b420047edae5eb0/inst/doc/nonparametric-bayes.R).
 This code is dynamically embedded into the manuscript using Knitr, [@knitr].
 The script relies on custom routines for executing the estimation of the Gaussian process and the for solving the stochastic dynamic programming problem.
-The sensitivity analysis is provided as a separate script, [sensitivity.Rmd]().
-These routines are provided as an R package, [nonparameteric-bayes](), also available on Github.
+The sensitivity analysis is provided as a separate script, [sensitivity.Rmd](https://github.com/cboettig/nonparametric-bayes/blob/62e0f81f9e8eabd0f7ee462f098320409a407f81/inst/examples/sensitivity.Rmd).
+These routines are provided as an R package, [nonparameteric-bayes](https://github.com/cboettig/nonparametric-bayes), also available on Github.
 

### 2013-11-20 inst/doc/nonparametric-bayes.tex
draft to marc and steve

--- 2013-11-20 inst/doc/nonparametric-bayes.tex
@@
 
-%  \author[cstar]{Marc Mangel}
-%  \author[noaa]{Stephan B. Munch}
-%  \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA 95064, USA}
-%  \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}
-
   \begin{abstract}
@@
   \end{abstract}
-
+   \begin{keyword} Bayesian \sep Structural Uncertainty \sep Nonparametric \sep Optimal Control \sep Decision Theory \sep Gaussian Processes \sep Fisheries Management \sep \end{keyword}
  \end{frontmatter}
@@
 simulations. Details and code provided in the supplement.}
+\end{figure}
+
+\subsection{Sensitivity Analysis}\label{sensitivity-analysis}
+
+These results are not sensitive to the modeling details of the
+simulation. The BNP-SDP estimate remains very close to the optimal
+solution obtained by knowning the true model across changes to the
+training simulation, noise scale, parameters or structure of the
+underlying model, as seen in the figure. Results are pooled across
+different random seeds (111, 222, 333), noise values of 0.01 and 0.05
+(not large enough to violate the self-sustaining criterion of Reed
+(1979)), and different randomly generated parameters sets for each
+model, and using either the Myers or Allen as the underlying structure.
+While the BNP-SDP performs slightly closer to optimal on the smaller
+noise levels, and closer on the Allen model then the Myers model, the
+differences are small compared to the values observed in using the wrong
+parameteric model. The BNP-SDP solution consistently avoids either
+failing to harvest at all or crashing the stock, which results in the
+very low profits/values seen in parameteric models in the previous
+figure.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/sensitivity.pdf}
+\caption{Sensitivity Analysis}
 \end{figure}
@@
 version-stable link to the project's Github code repostitory,
-\href{}{nonparameteric-bayes.R}. This code is dynamically embedded into
-the manuscript using Knitr, (Xie 2012). The script relies on custom
-routines for executing the estimation of the Gaussian process and the
-for solving the stochastic dynamic programming problem. These routines
-are provided as an R package, \href{}{nonparameteric-bayes}, also
-available on Github.
-
-\subsection{Sensitivity Analysis}\label{sensitivity-analysis}
+\href{https://github.com/cboettig/nonparametric-bayes/blob/371114ec43bb32e92f330f049b420047edae5eb0/inst/doc/nonparametric-bayes.R}{nonparameteric-bayes.R}.
+This code is dynamically embedded into the manuscript using Knitr, (Xie
+2012). The script relies on custom routines for executing the estimation
+of the Gaussian process and the for solving the stochastic dynamic
+programming problem. The sensitivity analysis is provided as a separate
+script,
+\href{https://github.com/cboettig/nonparametric-bayes/blob/62e0f81f9e8eabd0f7ee462f098320409a407f81/inst/examples/sensitivity.Rmd}{sensitivity.Rmd}.
+These routines are provided as an R package,
+\href{https://github.com/cboettig/nonparametric-bayes}{nonparameteric-bayes},
+also available on Github.
+
+\subsection{Data}\label{data}
+
+While the data can be regenerated using the code provided, for
+convenience CSV files of the data shown in each graph are also provided,
+along with appropriate metadata written in the Ecological Metadata
+Language (EML).
 
@@
 Analysis: the Strengths and Limitations of Hierarchical Statistical
-Modeling.'' \emph{Ecological Applications : a Publication of the
-Ecological Society of America} 19 (3) (apr): 553--70.
+Modeling.'' \emph{Ecological Applications} 19 (3) (apr): 553--70.
+doi:10.1890/07-0744.1.
 \url{http://www.ncbi.nlm.nih.gov/pubmed/19425416}.

### 2013-11-20 inst/doc/nonparametric-bayes.Rmd
actual draft sent to steve/marc

@@
 using jags, [@R2jags]).  We choose uninformative uniform priors for all
-parameters (See Appendix, Figures S1-S3, and Table S1, and the R code
+parameters (See Appendix, Figures and tables, and the R code
 provided). One-step-ahead predictions of these model fits are shown in
-Figure 1.
+Figure 1.  While alternative approaches to the estimation of the posteriors
+(such as integrating out the rate parameter $r$ analytically and then
+performing a grid search over the remaining parameter space), the 
+approach of using a standard Gibbs sampler routine is both more general
+and representitive of common practice in estimating posteriors for such models. 
+Each sampling is tested for Gelman-Rubin convergence and results are robust
+to longer runs.
 
@@
 Despite the similarities in model fits to the observed data, the
-policies inferred under each model differ widely, as shown in Figure 3.
+policies inferred under each model differ widely, as shown in Figure 4.
 Policies are shown in terms of target escapement, $S_t$.  Under models
@@
 The consequences of managing 100 replicate realizations of the simulated
-fishery under each of the policies estimated is shown in Figure 4.
+fishery under each of the policies estimated is shown in Figure 5.
 As expected from the policy curves, the structurally correct model
@@
 easily be compared by using the relative differences in net present
-value realized by each of the model, as shown in Figure 5.  The BNP-SDP
+value realized by each of the model. Figure 6 shows the distribution of values
+(profits) realized by each simulation shown in Figure 5.  The BNP-SDP
 approach most consistently realizes a value close to the optimal solution,
@@
 These results are not sensitive to the modeling details of the simulation.
-The BNP-SDP estimate remains very close to the optimal solution obtained by knowning the true model across changes to the training simulation, noise scale, parameters or structure of the underlying model, as seen in the figure.
+The BNP-SDP estimate remains very close to the optimal solution obtained by knowning the true model across changes to the training simulation, noise scale, parameters or structure of the underlying model, as seen in Figure 7.
 Results are pooled across different random seeds (111, 222, 333), noise values of 0.01 and 0.05 (not large enough to violate the self-sustaining criterion of @Reed1979), and different randomly generated parameters sets for each model, and using either the Myers or Allen as the underlying structure.
@@
 The BNP-SDP solution consistently avoids either failing to harvest at all or crashing the stock, which results in the very low profits/values seen in parametric models in the previous figure.
-
-![Sensitivity Analysis](figure/sensitivity.pdf)
+The sensitivity analysis is compared against the optimal solution alone and not shown against the estimates of the various parameteric models (but see Figure 6), where convergence of the Gibbs sampling could contribute to their poorer performance if not checked by hand for each configuration.
+
+![Sensitivity Analysis.
+Histogram shows the ratio of the realized net present value / profits derived when managing under the BNP-SDP Gaussian Process approach over the optimal value given the true model and true parameters.
+Values near 1 indicate nearly optimal performance.
+Simulations cover a range of underlying parameter, model, replicate, and noise levels.](figure/sensitivity.pdf)
 
@@
 We place uniform priors on the growth rate $r$, carrying capacity $K$, and log-normal standard deviation parameter $\sigma$, over ranges given in Table 1.
-Posteriors are inferred by Gibbs sampling using Jags [@R2Jags] (see provided code).
+Posteriors are inferred by Gibbs sampling using Jags [@R2jags] (see provided code).
 
@@
 We place uniform priors on the growth rate $r$, carrying capacity $K$, $\theta$ controls the strength of the nonlinearity, exhibiting an allee effect for $\theta \geq 2$, and log-normal standard deviation parameter $\sigma$, over ranges given in Table 1.
-Posteriors are inferred by Gibbs sampling using Jags [@R2Jags] (see code provided).
+Posteriors are inferred by Gibbs sampling using Jags [@R2jags] (see code provided).
 
@@
 We place uniform priors on the growth rate $r$, carrying capacity $K$, allee threshold $C$, and log-normal standard deviation parameter $\sigma$, over ranges given in Table 1.
-Posteriors are inferred by Gibbs sampling using Jags [@R2Jags] (see code provided).
+Posteriors are inferred by Gibbs sampling using Jags [@R2jags] (see code provided).
 

### 2013-11-20 inst/doc/nonparametric-bayes.tex
actual draft sent to steve/marc

@@
 
-We develop these concerns in the context of fisheries, though the
-underlying challenges and methods are germane to many other conservation
-and resource management problems. The economic value and ecological
-concern have made marine fisheries the crucible for much of the founding
-work (H. S. Gordon 1954; Reed 1979; May et al. 1979; Ludwig and Walters
-1982) in managing ecosystems under uncertainty. Global trends (Worm et
-al. 2006) and controversy (Hilborn 2007; Worm et al. 2009) have made
-understanding these challenges all the more pressing.
+We develop a novel approach to address these concerns in the context of
+fisheries; though the underlying challenges and methods are germane to
+many other conservation and resource management problems. The economic
+value and ecological concern have made marine fisheries the crucible for
+much of the founding work (H. S. Gordon 1954; Reed 1979; May et al.
+1979; Ludwig and Walters 1982) in managing ecosystems under uncertainty.
+Global trends (Worm et al. 2006) and controversy (Hilborn 2007; Worm et
+al. 2009) have made understanding these challenges all the more
+pressing.
 
@@
 of policy (\emph{e.g.} Reed 1979; Clark and Kirkwood 1986; Roughgarden
-and Smith 1996; Sethi et al. 2005), parameteric uncertainty (\emph{e.g.}
+and Smith 1996; Sethi et al. 2005), parametric uncertainty (\emph{e.g.}
 Ludwig and Walters 1982; Hilborn and Mangel 1997; McAllister 1998;
@@
 \emph{Cut the next three paragraphs, since they are covered more
-consisely in the above paragraph?}
+concisely in the above paragraph?}
 
@@
 than the exception. The potential concern of tipping points in
-ecological dyanmics (Scheffer et al. 2001; Polasky et al. 2011) reflects
+ecological dynamics (Scheffer et al. 2001; Polasky et al. 2011) reflects
 these concerns -- as either knowledge of the true model or more complete
@@
 observed data than near the observed data. These features allow the
-GP-SDP approach to find robust management solutions in face of limited
+BNP-SDP approach to find robust management solutions in face of limited
 data and without knowledge of the correct model structure.
@@
 complexity of the data on which it is estimated (e.g.~nonparametric
-heirarchical approaches such as the Dirchlet process) or the data points
+hierarchical approaches such as the Dirchlet process) or the data points
 become themselves part of the model specification, as in the
@@
 provides an excellent general introduction to Gaussian Processes and
-Munch et al. (2005) first discusses their application in the context of
-population dynamics models such as fisheries stock-recruitment
-relationships.
+Munch, Kottas, and Mangel (2005) first discusses their application in
+the context of population dynamics models such as fisheries
+stock-recruitment relationships.
 
@@
 
-We introduce parameteric uncertainty by first estimating each of the
+We introduce parametric uncertainty by first estimating each of the
 candidate models from data on unexploited stock dynamics following some
-pertubation (non-equilibrium initial condition) over several time steps.
-This training data could be generated in several different ways (such as
-known variable exploitation rates, etc), as long as it reflects the
-dynamics in some limited region of state space without impacting the
-problem. We consider a period of 40 years of training data: long enough
-that the estimates are not dependent on the particular realization,
-while longer times are not likely to provide substantial improvement
-(i.e.~the results are not sensitive to this interval). Each of the
-models (described below) is fit to the same training data, as shown in
-Figure 1.
+perturbation (non-equilibrium initial condition) over several time
+steps. This training data could be generated in several different ways
+(such as known variable exploitation rates, etc.), as long as it
+reflects the dynamics in some limited region of state space without
+impacting the problem. We consider a period of 40 years of training
+data: long enough that the estimates are not dependent on the particular
+realization, while longer times are not likely to provide substantial
+improvement (i.e.~the results are not sensitive to this interval). Each
+of the models (described below) is fit to the same training data, as
+shown in Figure 1.
 
@@
 using jags, (Su and Yajima 2013)). We choose uninformative uniform
-priors for all parameters (See Appendix, Figures S1-S3, and Table S1,
-and the R code provided). One-step-ahead predictions of these model fits
-are shown in Figure 1.
+priors for all parameters (See Appendix, Figures and tables, and the R
+code provided). One-step-ahead predictions of these model fits are shown
+in Figure 1. While alternative approaches to the estimation of the
+posteriors (such as integrating out the rate parameter $r$ analytically
+and then performing a grid search over the remaining parameter space),
+the approach of using a standard Gibbs sampler routine is both more
+general and representitive of common practice in estimating posteriors
+for such models. Each sampling is tested for Gelman-Rubin convergence
+and results are robust to longer runs.
 
@@
 (Allen). Note that the GP is both more pessimistic and more uncertain
-about the future state than the parameteric models, while the realized
-state often falls outside of the expected range forecasted by the
+about the future state than the paramteric models, while the realized
+state often falls outside of the expected range forecast by the
 structurally incorrect Myers and Ricker models.}
@@
 Despite the similarities in model fits to the observed data, the
-policies inferred under each model differ widely, as shown in Figure 3.
+policies inferred under each model differ widely, as shown in Figure 4.
 Policies are shown in terms of target escapement, $S_t$. Under models
@@
 The consequences of managing 100 replicate realizations of the simulated
-fishery under each of the policies estimated is shown in Figure 4. As
+fishery under each of the policies estimated is shown in Figure 5. As
 expected from the policy curves, the structurally correct model
@@
 easily be compared by using the relative differences in net present
-value realized by each of the model, as shown in Figure 5. The BNP-SDP
-approach most consistently realizes a value close to the optimal
+value realized by each of the model. Figure 6 shows the distribution of
+values (profits) realized by each simulation shown in Figure 5. The
+BNP-SDP approach most consistently realizes a value close to the optimal
 solution, and importantly avoids ever driving the system across the
@@
 training simulation, noise scale, parameters or structure of the
-underlying model, as seen in the figure. Results are pooled across
+underlying model, as seen in Figure 7. Results are pooled across
 different random seeds (111, 222, 333), noise values of 0.01 and 0.05
@@
 differences are small compared to the values observed in using the wrong
-parameteric model. The BNP-SDP solution consistently avoids either
+parametric model. The BNP-SDP solution consistently avoids either
 failing to harvest at all or crashing the stock, which results in the
-very low profits/values seen in parameteric models in the previous
-figure.
+very low profits/values seen in parametric models in the previous
+figure. The sensitivity analysis is compared against the optimal
+solution alone and not shown against the estimates of the various
+parameteric models (but see Figure 6), where convergence of the Gibbs
+sampling could contribute to their poorer performance if not checked by
+hand for each configuration.
 
@@
 \includegraphics{figure/sensitivity.pdf}
-\caption{Sensitivity Analysis}
+\caption{Sensitivity Analysis. Histogram shows the ratio of the realized
+net present value / profits derived when managing under the BNP-SDP
+Gaussian Process approach over the optimal value given the true model
+and true parameters. Values near 1 indicate nearly optimal performance.
+Simulations cover a range of underlying parameter, model, replicate, and
+noise levels.}
 \end{figure}
@@
 Despite this flexibility, the GP can be described in relatively few
-parameters and is thus far less likely to overfit.
+parameters and is thus far less likely to over-fit.
 
@@
 be influenced by the curvature of the value (profit) function $\Pi$.
-Both to simplify the intution and avoid biasing this result, we have
+Both to simplify the intuition and avoid biasing this result, we have
 chosen a profits that are linear in the catch and thus neither
@@
 the typical assumption of diminishing returns, would make the SDP more
-risk-adverse (as larger-than-expected stock sizes offer dimished returns
-relative to the cost of smaller-than-expected stock sizes), and
+risk-adverse (as larger-than-expected stock sizes offer diminished
+returns relative to the cost of smaller-than-expected stock sizes), and
 strengthen the result shown here in which the BNP solution tends to
@@
 deviation parameter $\sigma$, over ranges given in Table 1. Posteriors
-are inferred by Gibbs sampling using Jags ({\textbf{???}}) (see provided
-code).
+are inferred by Gibbs sampling using Jags (Su and Yajima 2013) (see
+provided code).
 
@@
 over ranges given in Table 1. Posteriors are inferred by Gibbs sampling
-using Jags ({\textbf{???}}) (see code provided).
+using Jags (Su and Yajima 2013) (see code provided).
 
@@
 log-normal standard deviation parameter $\sigma$, over ranges given in
-Table 1. Posteriors are inferred by Gibbs sampling using Jags
-({\textbf{???}}) (see code provided).
+Table 1. Posteriors are inferred by Gibbs sampling using Jags (Su and
+Yajima 2013) (see code provided).
 

### 2014-04-26 inst/doc/nonparametric-bayes.tex
old commit

@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\newenvironment{Shaded}{}{}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
-\newcommand{\RegionMarkerTok}[1]{{#1}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
-\newcommand{\NormalTok}[1]{{#1}}
-\usepackage{graphicx}
-% We will generate all images so they have a width \maxwidth. This means
-% that they will get their normal width if they fit onto the page, but
-% are scaled down if they would overflow the margins.
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
-\else\Gin@nat@width\fi}
-\makeatother
-\let\Oldincludegraphics\includegraphics
-\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
 \ifxetex
@@
 value and ecological concern have made marine fisheries the crucible for
-much of the founding work (H. S. Gordon 1954; Reed 1979; May et al.
-1979; Ludwig and Walters 1982) in managing ecosystems under uncertainty.
-Global trends (Worm et al. 2006) and controversy (Hilborn 2007; Worm et
-al. 2009) have made understanding these challenges all the more
+much of the founding work (H. S. Gordon 1954; Reed 1979; {\textbf{???}};
+Ludwig and Walters 1982) in managing ecosystems under uncertainty.
+Global trends (Worm et al. 2006) and controversy (Hilborn 2007;
+{\textbf{???}}) have made understanding these challenges all the more
 pressing.
@@
 stochasticity in biological processes, measurements, and implementation
-of policy (\emph{e.g.} Reed 1979; Clark and Kirkwood 1986; Roughgarden
-and Smith 1996; Sethi et al. 2005), parametric uncertainty (\emph{e.g.}
-Ludwig and Walters 1982; Hilborn and Mangel 1997; McAllister 1998;
-Schapaugh and Tyre 2013), and model or structural uncertainty
-(\emph{e.g.} B. K. Williams 2001; Cressie et al. 2009; Athanassoglou and
-Xepapadeas 2012). Of these, structural uncertainty incorporates the
-least a priori knowledge or assumptions and is generally the hardest to
-quantify. Typical approaches assume a weak notion of model uncertainty
-in which the correct model (or reasonable approximation) of the dynamics
-must be identified from among a handful of alternative models. Here we
-consider an approach that addresses uncertainty at each of these levels
-without assuming the dynamics follow a particular (i.e.~parametric)
-structure.
+of policy (\emph{e.g.} Reed 1979; C. W. Clark and Kirkwood 1986;
+Roughgarden and Smith 1996; Sethi et al. 2005), parametric uncertainty
+(\emph{e.g.} Ludwig and Walters 1982; {\textbf{???}}; McAllister 1998;
+{\textbf{???}}), and model or structural uncertainty (\emph{e.g.} B. K.
+Williams 2001; Cressie et al. 2009; Athanassoglou and Xepapadeas 2012).
+Of these, structural uncertainty incorporates the least a priori
+knowledge or assumptions and is generally the hardest to quantify.
+Typical approaches assume a weak notion of model uncertainty in which
+the correct model (or reasonable approximation) of the dynamics must be
+identified from among a handful of alternative models. Here we consider
+an approach that addresses uncertainty at each of these levels without
+assuming the dynamics follow a particular (i.e.~parametric) structure.
 
@@
 reflected in the estimates of the transition probabilities (Ludwig and
-Walters 1982; Mangel and Clark 1988; Hilborn and Mangel 1997; Schapaugh
-and Tyre 2013).
+Walters 1982; Mangel and Clark 1988; {\textbf{???}}; {\textbf{???}}).
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-Figureb_posteriors.pdf}
-\caption{Points show the training data of stock-size over time. Curves
-show the posterior step-ahead predictions based on each of the estimated
-models.}
-\end{figure}
-
-All models fit the observed data rather closely and with relatively
-small uncertainty, as illustrated in the posterior predictive curves in
-Figure 1. Figure 1 shows the training data of stock sizes observed over
-time as points, overlaid with the step-ahead predictions of each
-estimated model using the parameters sampled from their posterior
-distributions. Each model manages to fit the observed data rather
-closely. Compared to the expected value of the true model most estimates
-appear to overfit, predicting fluctuations that are actually due purely
-to stochasticity in growth rate. Model-choice criteria shown in Table 1
-penalize more complex models and show a slight preference for the
-simpler Ricker model over the more complicated alternate stable state
-models (Allen and Myers). Details on MCMC estimates for each model,
-traces, and posterior distributions can be found in the appendix.
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{rrrr}
-  \hline
- & Allen & Ricker & Myers \\ 
-  \hline
-DIC & 50.14 & 49.45 & 50.61 \\ 
-  AIC & -24.60 & -30.07 & -27.19 \\ 
-  BIC & -17.85 & -25.00 & -20.44 \\ 
-   \hline
-\end{tabular}
-\caption{Model choice scores for several common criteria all (wrongly) select the simplest model.
-As the true (Allen) model is not distinguishable from the simpler (Ricker) model in the region of the observed data, this error cannot be avoided regardless of the model choice criterion.
-This highlights the danger of model choice when the selected model will be used outside of the observed range of the data.}
-\end{center}
-\end{table}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-statespace_posteriors.pdf}
-\caption{Graph of the inferred Gaussian process compared to the true
-process and maximum-likelihood estimated process. Graph shows the
-expected value for the function $f$ under each model. Two standard
-deviations from the estimated Gaussian process covariance with (light
-grey) and without (darker grey) measurement error are also shown. The
-training data is also shown as black points. The GP is conditioned on
-(0,0), shown as a pseudo-data point.}
-\end{figure}
-
-The mean inferred state space dynamics of each model relative to the
-true model used to generate the data is shown in Figure 2, predicting
-the relationship between observed stock size (x-axis) to the stock size
-after recruitment the following year. Note that in contrast to the other
-models shown, the expected Gaussian process corresponds to a
-distribution of curves - as indicated by the gray band - which itself
-has a mean shown in black. Parameter uncertainty (not shown) spreads out
-the estimates further. The observed data from which each model is
-estimated is also shown. The observations come from only a limited
-region of state space corresponding to unharvested or weakly harvested
-system. No observations occur at the theoretical optimum harvest rate or
-near the tipping point.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-out_of_sample_predictions.pdf}
-\caption{Out of sample predictions of the dynamics under each model.
-Points show the stock size simulated by the true model. Overlay shows
-the range of states predicted by each model, based on the state observed
-in the previous time step. The Ricker model always predicts population
-growth, while the actual population shrinks in each step as the initial
-condition falls below the Allee threshold of the underlying model
-(Allen). Note that the GP is both more pessimistic and more uncertain
-about the future state than the paramteric models, while the realized
-state often falls outside of the expected range forecast by the
-structurally incorrect Myers and Ricker models.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-Figure2.pdf}
-\caption{The steady-state optimal policy (infinite boundary) calculated
-under each model. Policies are shown in terms of target escapement,
-$S_t$, as under models such as this a constant escapement policy is
-expected to be optimal (Reed 1979).}
-\end{figure}
-
-Despite the similarities in model fits to the observed data, the
-policies inferred under each model differ widely, as shown in Figure 4.
-Policies are shown in terms of target escapement, $S_t$. Under models
-such as this a constant escapement policy is expected to be optimal
-(Reed 1979), whereby population levels below a certain size $S$ are
-unharvested, while above that size the harvest strategy aims to return
-the population to $S$, resulting in the hockey-stick shaped policies
-shown. Only the structurally correct model (Allen model) and the GP
-produce policies close to the true optimum policy (where both the
-underlying model structure and parameter values are known without
-error).
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-Figure3.pdf}
-\caption{Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under
-the policies derived from each of the estimated models, as well as the
-policy based on the exact underlying model.}
-\end{figure}
-
-The consequences of managing 100 replicate realizations of the simulated
-fishery under each of the policies estimated is shown in Figure 5. As
-expected from the policy curves, the structurally correct model
-under-harvests, leaving the stock to vary around it's un-fished optimum.
-The structurally incorrect Ricker model over-harvests the population
-passed the tipping point consistently, resulting in the immediate crash
-of the stock and thus derives minimal profits.
-
-These results are robust across a range of stochastic realizations,
-models, and parameter values. The results across this range can most
-easily be compared by using the relative differences in net present
-value realized by each of the model. Figure 6 shows the distribution of
-values (profits) realized by each simulation shown in Figure 5. The
-BNP-SDP approach most consistently realizes a value close to the optimal
-solution, and importantly avoids ever driving the system across the
-tipping point, which results in the near-zero value cases in the
-parametric models.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-Figure4.pdf}
-\caption{Histograms of the realized net present value of the fishery
-over a range of simulated data and resulting parameter estimates. For
-each data set, the three models are estimated as described above. Values
-plotted are the averages of a given policy over 100 replicate
-simulations. Details and code provided in the supplement.}
-\end{figure}
-
-\subsection{Sensitivity Analysis}\label{sensitivity-analysis}
-
-These results are not sensitive to the modeling details of the
-simulation. The BNP-SDP estimate remains very close to the optimal
-solution obtained by knowning the true model across changes to the
-training simulation, noise scale, parameters or structure of the
-underlying model, as seen in Figure 7. Results are pooled across
-different random seeds (111, 222, 333), noise values of 0.01 and 0.05
-(not large enough to violate the self-sustaining criterion of Reed
-(1979)), and different randomly generated parameters sets for each
-model, and using either the Myers or Allen as the underlying structure.
-While the BNP-SDP performs slightly closer to optimal on the smaller
-noise levels, and closer on the Allen model then the Myers model, the
-differences are small compared to the values observed in using the wrong
-parametric model. The BNP-SDP solution consistently avoids either
-failing to harvest at all or crashing the stock, which results in the
-very low profits/values seen in parametric models in the previous
-figure. The sensitivity analysis is compared against the optimal
-solution alone and not shown against the estimates of the various
-parameteric models (but see Figure 6), where convergence of the Gibbs
-sampling could contribute to their poorer performance if not checked by
-hand for each configuration.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/sensitivity.pdf}
-\caption{Sensitivity Analysis. Histogram shows the ratio of the realized
-net present value / profits derived when managing under the BNP-SDP
-Gaussian Process approach over the optimal value given the true model
-and true parameters. Values near 1 indicate nearly optimal performance.
-Simulations cover a range of underlying parameter, model, replicate, and
-noise levels.}
-\end{figure}
-
-\section{Discussion}\label{discussion}
-
-In any modeling effort, models must be chosen for the task at hand.
-Though simple mechanistically motivated models offer the greatest
-potential to increase our basic understanding of ecological processes
-(Cuddington, Fortin, and Gerber 2013; Geritz and Kisdi 2012), such
-models can be not only inaccurate but misleading when relied upon in a
-quantitative decision making framework. In this paper we have tackled
-two aspects of uncertainty that are both common to many ecological
-desicion-making problems and fundamentally challenging to existing
-approaches which largely rely on parametric models:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  We do not know what the correct models are for ecological systems.
-\item
-  We have limited data from which to estimate the model -- in
-  particular, such models may be misleading in predicting the
-  probability of outcomes outside the training data.
-\end{enumerate}
-
-We have illustrated how the use of non-parametric approaches can provide
-more reliable solutions in the sequential decision-making problem.
-
-\subsubsection{Traditional model-choice approaches can be positively
-misleading.}\label{traditional-model-choice-approaches-can-be-positively-misleading.}
-
-These results illustrate that model-choice approaches would be
-positively misleading -- supporting simpler models that cannot express
-tipping point dynamics merely on account of them being similar. As the
-data shown comes only from the basin of attraction near the unfished
-equilibrium, near which all of the models are approximately linear and
-approximately identical.
-
-Model choice approaches trade off model complexity and fit to the data.
-When the data come from a limited region of state-space -- as is
-necessarily the case whenever there is a potential concern about tipping
-point dynamics -- simpler models can fit just as well and will tend to
-outperform more complex ones. This approach would be appropriate when
-the dynamics can be expected to remain in the region of the training
-data; for instance, if we only considered the forecasting accuracy of
-the unfished population dynamics under each model.
-
-In contrast, the decision-maker's problem of setting appropriate harvest
-levels cannot exclude regions of state-space outside the observed range
-when integrating over all possible decisions to find the optimal choice.
-Such problems are not constrained to fisheries management but ubiquitous
-across ecological decision-making and conservation where the greatest
-concerns involve entering previously unobserved regions of state-space
--- whether that is the collapse of a fishery, the spread of an invasive,
-or the loss of habitat.
-
-\subsubsection{BNP-SDP expresses larger uncertainty in regions where the
-data are
-poor}\label{bnp-sdp-expresses-larger-uncertainty-in-regions-where-the-data-are-poor}
-
-The parametric models perform worst when they propose a management
-strategy outside the range of the observed data. The non-parametric
-Bayesian approach, in contrast, allows a predictive model that expresses
-a great deal of uncertainty about the probable dynamics outside the
-observed range, while retaining very good predictive accuracy in the
-range observed. The management policy dictated by the GP balance this
-uncertainty against the immediate value of the harvest, and act to
-stabilize the population dynamics in a region of state space in which
-the predictions can be reliably reflected by the data.
-
-\subsubsection{BNP-SDP has good predictive accuracy where data are
-good}\label{bnp-sdp-has-good-predictive-accuracy-where-data-are-good}
-
-While expressing larger uncertainty outside the observed data, the GP
-can also provide a better fit with smaller uncertainty inside the range
-of the observed data. This arises from the greater flexibility of the
-Gaussian process, which describes a large family of possible curves.
-Despite this flexibility, the GP can be described in relatively few
-parameters and is thus far less likely to over-fit.
-
-\subsubsection{Risk-prone and risk-adverse value
-functions}\label{risk-prone-and-risk-adverse-value-functions}
-
-The degree to which the decision-making part of the algorithm (the SDP)
-chooses to explore or avoid the resulting region of uncertainty can also
-be influenced by the curvature of the value (profit) function $\Pi$.
-Both to simplify the intuition and avoid biasing this result, we have
-chosen a profits that are linear in the catch and thus neither
-risk-prone nor risk adverse. Making this function concave, representing
-the typical assumption of diminishing returns, would make the SDP more
-risk-adverse (as larger-than-expected stock sizes offer diminished
-returns relative to the cost of smaller-than-expected stock sizes), and
-strengthen the result shown here in which the BNP solution tends to
-avoid the region of uncertainty. Sufficiently convex or risk-prone
-functions could lead the SDP to attempt higher exploitation rates
-despite the uncertainty. Understanding the relative roles of such
-functions would be a promising direction for future investigation.
-
-\subsubsection{The role of the prior}\label{the-role-of-the-prior}
-
-Lastly, it should be noted that outside the data, the NBP reverts to the
-prior, and consequently the choice of the prior can also play a
-significant role in determining the optimal policy inferred by the SDP.
-In the examples shown here we have selected a prior that is both
-relatively uninformative (due to the broad priors placed on its
-parameters $\ell$ and $\sigma$ and simple (mean zero, radial basis
-function kernel). In practice, both the choice of mean and the choice of
-the covariance function may be chosen to confer particular biological
-properties, as well as more biologically informed priors for $\ell$ and
-$\sigma$. In principle, this may allow a manager to improve the
-performance of the BNP-SDP approach by adding only enough additional
-detail as is justified. For instance, it would be possible to use a
-linear or a Ricker-shaped mean in the prior without making the much
-stronger assumption that the Ricker is the structurally correct model.
-However, this influence raises challenges as well. For instance, in
-choosing a trivial mean-zero prior, we bias the dynamics to transitions
-to 0 in step $X_{t+1}$ from any stock size $X_t$ in the prior year, in
-the absence of any other data. Future research must make sure that both
-the prior and the value function are chosen appropriately for the
-problem at hand.
-
-\subsection{Future directions}\label{future-directions}
-
-In this simulated example, the underlying dynamics are truly governed by
-a simple parametric model, allowing the parametric approaches to be more
-accurate. Similarly, because the dynamics are one-dimensional dynamics
-and lead to stable nodes (rather than other attractors such as
-limit-cycles resulting in oscillations), the training data provides
-relatively limited information about the dynamics. For these reasons, we
-anticipate that in higher-dimensional examples characteristic of
-ecosystem management problems that the machine learning approach will
-prove even more valuable.
-
-In our treatment here we have ignored the possibility of learning during
-the management phase, in which the additional observations of the stock
-size could potentially improve parameter estimates. Of particular
-interest in the context of the extreme uncertainty considered here is
-the notion of ``active adaptive management'' or ``adaptive probing'',
-which may actively seek to reduce uncertainty. The concept of adaptive
-probing is one area that has explicitly addressed the extrapolation
-uncertainty addressed here, though typically without the additional
-issue of model uncertainty. As a result, adaptive probing strategies
-suggest rather opposite conclusions than what we observe here. Such
-adaptive probing or Dual Control (e.g. Ludwig and Walters (1982))
-approaches trade off short term utility by choosing actions that can
-reduce uncertainty. Adaptive probing strategies arise when it is
-valuable to intentionally force a system far from the observed values
-even when the expected value such actions is low, as it provides much
-faster learning and consequent reduction of model uncertainty that can
-allow greater value to be derived later on. For instance, Ludwig and
-Walters (1982) show that it may be advantageous to fish an unexploited
-population very heavily at first to obtain a better estimate of the
-recruitment rate. This intuitive strategy when a population is governed
-by a Ricker or Beverton-Holt-like dynamic would clearly be disastrous if
-instead the dynamics contained an unforeseen tipping point. The best way
-to learn where the edge lies may be to walk up to it, but it is also the
-most dangerous. Future work should attempt to understand when such
-active adaptive learning is valuable, and when it will increase the risk
-of an irreversible transition.
-
-\section{Acknowledgments}\label{acknowledgments}
-
-This work was partially supported by the Center for Stock Assessment
-Research, a partnership between the University of California Santa Cruz
-and the Fisheries Ecology Division, Southwest Fisheries Science Center,
-Santa Cruz, CA and by NSF~grant EF-0924195 to MM and NSF grant
-DBI-1306697 to CB.
-
-\appendix
-
-\section{Model definitions and
-estimation}\label{model-definitions-and-estimation}
-
-\subsection{Ricker Model}\label{ricker-model}
-
-The Ricker model is given by
-
-\begin{enumerate}
-\def\labelenumi{(\arabic{enumi})}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
-\end{enumerate}
-
-where $Z_t$ is log-normal noise of mean unity and log standard deviation
-$\sigma$, representing the stochastic growth, $X_t$ the stock size at
-time $t$, $S_t$ the escapement (unharvested population that will recruit
-in the following year, $S_t = X_t - h_t$). We place uniform priors on
-the growth rate $r$, carrying capacity $K$, and log-normal standard
-deviation parameter $\sigma$, over ranges given in Table 1. Posteriors
-are inferred by Gibbs sampling using Jags (Su and Yajima 2013) (see
-provided code).
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-2.pdf}
-\caption{Posteriors from the MCMC estimate of the Ricker model}
-\end{figure}
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{rlrr}
-  \hline
- & parameter & lower.bound & upper.bound \\ 
-  \hline
-1 & r0 & 0.01 & 20.00 \\ 
-  2 & K & 0.01 & 40.00 \\ 
-  3 & sigma & 0.00 & 100.00 \\ 
-   \hline
-\end{tabular}
-\caption{parameterization range for the uniform priors in the Ricker model}
-\end{center}
-\end{table}
-
-\subsection{Myers Model}\label{myers-model}
-
-The Myers model (Myers et al. 1995) is given by
-
-\begin{enumerate}
-\def\labelenumi{(\arabic{enumi})}
-\setcounter{enumi}{1}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}}\]
-\end{enumerate}
-
-where $Z_t$ is log-normal noise of mean unity and log standard deviation
-$\sigma$, representing the stochastic growth, $X_t$ the stock size at
-time $t$, $S_t$ the escapement (unharvested population that will recruit
-in the following year, $S_t = X_t - h_t$). We place uniform priors on
-the growth rate $r$, carrying capacity $K$, $\theta$ controls the
-strength of the nonlinearity, exhibiting an allee effect for
-$\theta \geq 2$, and log-normal standard deviation parameter $\sigma$,
-over ranges given in Table 1. Posteriors are inferred by Gibbs sampling
-using Jags (Su and Yajima 2013) (see code provided).
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-4.pdf}
-\caption{Traces from the MCMC estimate of the Myers model}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-5.pdf}
-\caption{Posterior distributions from the MCMC estimates of the Myers
-model}
-\end{figure}
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{rlrr}
-  \hline
- & parameter & lower.bound & upper.bound \\ 
-  \hline
-1 & r0 & 0.00 & 10.00 \\ 
-  2 & K & 0.00 & 40.00 \\ 
-  3 & theta & 0.00 & 10.00 \\ 
-  4 & sigma & 0.00 & 100.00 \\ 
-   \hline
-\end{tabular}
-\caption{parameterization range for the uniform priors in the Myers model}
-\end{center}
-\end{table}
-
-\subsection{Allen model}\label{allen-model}
-
-The Allen model (Allen and Tanner 2005) is given by
-
-\begin{enumerate}
-\def\labelenumi{(\arabic{enumi})}
-\setcounter{enumi}{2}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]
-\end{enumerate}
-
-where $Z_t$ is log-normal noise of mean unity and log standard deviation
-$\sigma$, representing the stochastic growth, $X_t$ the stock size at
-time $t$, $S_t$ the escapement (unharvested population that will recruit
-in the following year, $S_t = X_t - h_t$). We place uniform priors on
-the growth rate $r$, carrying capacity $K$, allee threshold $C$, and
-log-normal standard deviation parameter $\sigma$, over ranges given in
-Table 1. Posteriors are inferred by Gibbs sampling using Jags (Su and
-Yajima 2013) (see code provided).
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-7.pdf}
-\caption{Traces from the MCMC estimate of the Allen model}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-8.pdf}
-\caption{Posteriors from the MCMC estimate of the Allen model}
-\end{figure}
-
-\begin{table}[ht]
-\begin{center}
-\begin{tabular}{rlrr}
-  \hline
- & parameter & lower.bound & upper.bound \\ 
-  \hline
-1 & r0 & 0.01 & 6.00 \\ 
-  2 & K & 0.01 & 20.00 \\ 
-  3 & theta & 0.01 & 20.00 \\ 
-  4 & sigma & 0.00 & 100.00 \\ 
-   \hline
-\end{tabular}
-\caption{parameterization range for the uniform priors in the Allen model}
-\end{center}
-\end{table}
-
-Assuming the data are known with some process noise,
-
-\[y = f(x) + \varepsilon\]
-
-$\varepsilon$ IID normal, variance $\sigma_n^2$. Then under the GP,
-
-\[x|y \sim \mathcal{N}(E,C)\]
-\[E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y\]
-\[C = K(X_p, X_p) - K(X_p, X_o) K(X_o,X_o)^{-1} K(X_o, X_p)\]
-
-\}
-
-The marginal likelihood is then given by:
-
-\[\log(p(y | X)) = -\tfrac{1}{2} \mathbf{y}^T (K + \sigma_n^2 \mathbf{I})^{-1} y  - \tfrac{1}{2} \log\left| K + \sigma_n^2 \mathbf{I} \right| - \tfrac{n}{2}\log 2 \pi\]
-
-See Munch, Kottas, and Mangel (2005) or Rasmussen and Williams (2006)
-for a more detailed introduction.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-10.pdf}
-\caption{Traces from the MCMC estimates of the GP model}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/nonparametric-bayes-unnamed-chunk-11.pdf}
-\caption{Posterior distributions from the MCMC estimate of the GP model.
-Prior curves shown in red.}
-\end{figure}
-
-The Gaussian process priors on both the lengthscale $\ell$ and process
-noise $\sigma$ are are inverse Gamma distributed,
-
-\[f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)\]
-
-For the $\sigma$ prior, $\alpha = $ 5 and $\beta = $ 5. For $\ell$
-prior, $\alpha = $ 10 and $\beta = $ 10.
-
-\subsection{Optimal Control Problem
-Specification}\label{optimal-control-problem-specification}
-
-We seek the harvest policy $h(x)$ that maximizes:
-
-\[ \max_{h_t} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h_t) \delta^t  \]
-
-subject to the profit function $\Pi(X_t,h)$, discount rate $\delta$, and
-the state equation
-
-\[X_{t+1} = Z_t f(S_t)  \] \[S_t = X_t - h_t \]
-
-Where $Z_t$ is multiplicative noise function with mean 1, representing
-stochastic growth. We will consider log-normal noise with shape
-parameter $\sigma_g$.
-
-Form this we can write down the Bellman recursion as:
-
-\[V_t(x_t) = \max_h \mathbf{E} \left(\Pi(h_t, x_t) + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)\]
-
-For simplicity we assume profit is simply linear in the realized harvest
-(only enforcing the restriction that harvest can not exceed available
-stock), $\Pi(h,x) = min(h,x)$.
-
-\subsubsection{Pseudocode for the Bellman
-iteration}\label{pseudocode-for-the-bellman-iteration}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# compute the value for each possible harvest}
-\NormalTok{for(h in }\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid))\{}
-  \NormalTok{V1[h] =}\StringTok{ }\NormalTok{delta *}\StringTok{ }\NormalTok{F[[h]] %*%}\StringTok{ }\NormalTok{V +}\StringTok{  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
-\NormalTok{\}}
-\CommentTok{# find havest h that gives the maximum value}
-\NormalTok{for(j in }\DecValTok{1}\NormalTok{:gridsize)\{}
-  \NormalTok{value =}\StringTok{ }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T)  }\CommentTok{# each col is a different h, max over these}
-  \NormalTok{index =}\StringTok{ }\KeywordTok{which.max}\NormalTok{(V1[j,])       }\CommentTok{# store index so we can recover h's }
-  \NormalTok{output[,j] =}\StringTok{ }\KeywordTok{c}\NormalTok{(value, index)    }\CommentTok{# returns both profit value & index of optimal h.  }
-\NormalTok{\}}
-\CommentTok{# Sets V[t+1] = max_h(V[t]) at each possible state value, x}
-\NormalTok{V =}\StringTok{ }\NormalTok{out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
-\NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] =}\StringTok{ }\NormalTok{out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
-\end{Highlighting}
-\end{Shaded}
-
-\subsection{Training data}\label{training-data}
-
-Eacho of our models $f(S_t)$ must be estimated from training data, which
-we simulate from the Allen model with parameters $r = $ 2, $K =$ 8,
-$C =$ 5, and $\sigma_g =$ 0.05 for $T=$ 40 timesteps, starting at
-initial condition $X_0 = $ 5.5. The training data can be seen in Figure
-1.
-
-\subsection{Code}\label{code}
-
-A copy of the script to reconstruct the simulations and analysis shown
-here is provided in the supplemental materials, and through this
-version-stable link to the project's Github code repostitory,
-\href{https://github.com/cboettig/nonparametric-bayes/blob/371114ec43bb32e92f330f049b420047edae5eb0/inst/doc/nonparametric-bayes.R}{nonparameteric-bayes.R}.
-This code is dynamically embedded into the manuscript using Knitr, (Xie
-2012). The script relies on custom routines for executing the estimation
-of the Gaussian process and the for solving the stochastic dynamic
-programming problem. The sensitivity analysis is provided as a separate
-script,
-\href{https://github.com/cboettig/nonparametric-bayes/blob/62e0f81f9e8eabd0f7ee462f098320409a407f81/inst/examples/sensitivity.Rmd}{sensitivity.Rmd}.
-These routines are provided as an R package,
-\href{https://github.com/cboettig/nonparametric-bayes}{nonparameteric-bayes},
-also available on Github.
-
-\subsection{Data}\label{data}
-
-While the data can be regenerated using the code provided, for
-convenience CSV files of the data shown in each graph are also provided,
-along with appropriate metadata written in the Ecological Metadata
-Language (EML).
-
 Allen, Deborah, and Kimberly Tanner. 2005. ``Infusing Active Learning
@@
 Analysis: the Strengths and Limitations of Hierarchical Statistical
-Modeling.'' \emph{Ecological Applications} 19 (3) (apr): 553--70.
-doi:10.1890/07-0744.1.
+Modeling.'' \emph{Ecological Applications : a Publication of the
+Ecological Society of America} 19 (3) (apr): 553--70.
 \url{http://www.ncbi.nlm.nih.gov/pubmed/19425416}.
-
-Cuddington, Kim M., M. J. Fortin, and L. R. Gerber. 2013.
-``Process-based Models Are Required to Manage Ecological Systems in a
-Changing World.'' \emph{Ecosphere} 4 (February): 1--12.
-doi:10.1890/ES12-00178.1.
-\url{http://www.esajournals.org/doi/abs/10.1890/ES12-00178.1}.
 
@@
 \url{http://www.ncbi.nlm.nih.gov/pubmed/19665820}.
-
-Geritz, Stefan A. H., and Eva Kisdi. 2012. ``Mathematical Ecology: Why
-Mechanistic Models?'' \emph{Journal of Mathematical Biology} 65 (6-7)
-(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.
 
@@
 
-Hilborn, Ray, and Marc Mangel. 1997. \emph{The Ecological Detective:
-Confronting Models with Data}. Princeton University Press.
-\url{http://www.amazon.com/The-Ecological-Detective-Confronting-Models/dp/0691034974}.
-
 Hughes, Terry P., Cristina Linares, Vasilis Dakos, Ingrid a van de
@@
 \url{http://www.ncbi.nlm.nih.gov/pubmed/22995893}.
-
-Kocijan, Juš, Agathe Girard, Bla\textbackslash{}vz Banko, and Roderick
-Murray-Smith. 2005. ``Dynamic Systems Identification with Gaussian
-Processes.'' \emph{Mathematical and Computer Modelling of Dynamical
-Systems} 11 (4): 411--424. doi:10.1080/13873950500068567.
-\url{http://www.tandfonline.com/doi/abs/10.1080/13873950500068567}.
 
@@
 \url{http://doi.wiley.com/10.1111/2041-210X.12082}.
-
-May, Robert M., John R. Beddington, Colin W. Clark, Sidney J. Holt, and
-R. M. Laws. 1979. ``Management of Multispecies Fisheries.''
-\emph{Science (New York, N.Y.)} 205 (4403) (jul): 267--77.
-doi:10.1126/science.205.4403.267.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/17747032}.
 
@@
 
-Myers, R. A., N. J. Barrowman, J. A. Hutchings, and Andrew a Rosenberg.
-1995. ``Population Dynamics of Exploited Fish Stocks at Low Population
+Myers, R. a, N. J. Barrowman, J. a Hutchings, and a a Rosenberg. 1995.
+``Population Dynamics of Exploited Fish Stocks at Low Population
 Levels.'' \emph{Science (New York, N.Y.)} 269 (5227) (aug): 1106--8.
@@
 \url{http://www.pnas.org/content/93/10/5078.short}.
-
-Schapaugh, Adam W., and Andrew J. Tyre. 2013. ``Accounting for
-Parametric Uncertainty in Markov Decision Processes.'' \emph{Ecological
-Modelling} 254 (apr): 15--21. doi:10.1016/j.ecolmodel.2013.01.003.
-\url{http://linkinghub.elsevier.com/retrieve/pii/S0304380013000306}.
 
@@
 
-Worm, Boris, Ray Hilborn, Julia K. Baum, Trevor A. Branch, Jeremy S.
-Collie, Christopher Costello, Michael J. Fogarty, et al. 2009.
-``Rebuilding Global Fisheries.'' \emph{Science (New York, N.Y.)} 325
-(5940) (jul): 578--85. doi:10.1126/science.1173146.
-\url{http://www.ncbi.nlm.nih.gov/pubmed/19644114}.
-
-Xie, Yihui. 2012. \emph{Knitr: A General-purpose Package for Dynamic
-Report Generation in R}. \url{http://yihui.name/knitr/}.
-
 \end{document}

### 2014-04-26 inst/doc/my_report.R
committing old untracked files before cleaning up

@@ -1,0 +1,254 @@
+
+## ----plot-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE----
+library(nonparametricbayes) 
+opts_chunk$set(external=TRUE)
+read_chunk("../examples/BUGS/external-chunks.R")
+
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=1, 
+               comment=NA, verbose=TRUE, fig.width=4, fig.height=3)
+ 
+# Name the cache path and fig.path based on filename...
+opts_chunk$set(fig.path = paste("figure/",
+                                gsub(".Rmd", "", knitr:::knit_concord$get('infile')),
+                                "-", sep=""),
+               cache.path = paste("cache/", 
+                                  gsub(".Rmd", "", knitr:::knit_concord$get('infile') ), 
+                                "/", sep=""))
+
+toggle = "markup" #"hide"  # toggles `results` chunk option 
+
+library(ggplot2) 
+theme_set(theme_bw(base_size=12))
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+
+## ----posterior-mode------------------------------------------------------
+
+
+## ----stateeq-------------------------------------------------------------
+
+
+## ----sdp-pars, dependson="stateeq"---------------------------------------
+
+
+## ----obs, dependson="sdp-pars"-------------------------------------------
+
+
+## ----mle, dependson="obs"------------------------------------------------
+
+
+## ----mle-output, dependson="mle", results=toggle-------------------------
+
+
+## ----gp-priors-----------------------------------------------------------
+
+
+## ----gp, dependson=c("gp-priors", "obs")---------------------------------
+
+
+## ----gp_traces_densities, dependson="gp"---------------------------------
+
+
+## ----gp-output, dependson="gp", results=toggle---------------------------
+
+
+## ----jags-setup----------------------------------------------------------
+
+
+## ----common-priors-------------------------------------------------------
+
+
+## ----allen-model---------------------------------------------------------
+
+
+## ----allen-priors, dependson="common-priors"-----------------------------
+
+
+## ----allen_priors_xtable, include=TRUE, echo=FALSE-----------------------
[…]

### 2014-05-05 manuscripts/manuscript.Rmd
move manuscript to manuscripts directory

--- 2013-11-20 inst/doc/nonparametric-bayes.Rmd
@@
 opts_chunk$set(external=TRUE)
-read_chunk("../examples/BUGS/external-chunks.R")
-
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=1, 
-               comment=NA, verbose=TRUE, fig.width=4, fig.height=3)
-
-# Name the cache path and fig.path based on filename...
-opts_chunk$set(fig.path = paste("figure/",
-                                gsub(".Rmd", "", knitr:::knit_concord$get('infile')),
-                                "-", sep=""),
-               cache.path = paste("cache/", 
-                                  gsub(".Rmd", "", knitr:::knit_concord$get('infile') ), 
-                                "/", sep=""))
-
-toggle = "markup" #"hide"  # toggles `results` chunk option
-
+read_chunk("components/external-chunks.R")
+opts_chunk$set(fig.width=4, fig.height=3, echo=FALSE)
 library(ggplot2) 
@@
                "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+toggle = "markup" # results argument
 ```
@@
 While the data can be regenerated using the code provided, for convenience CSV files of the data shown in each graph are also provided, along with appropriate metadata written in the Ecological Metadata Language (EML).
+
+```{r include=FALSE}
+unlink("ricker_process.bugs")
+unlink("allen_process.bugs")
+unlink("myers_process.bugs")
+```

### 2014-05-05 manuscripts/components/knit.R

@@ -1,0 +1,26 @@
+#!/usr/bin/Rscript
+args <- commandArgs(TRUE)
+file <- args[1]
+
+## FIXME add second argument to handle figure types
+
+## Set the paths for cache and figure
+library(methods)
+library(knitr)
+basename <- gsub(".Rmd", "", file) # knitr:::knit_concord$get('infile')) 
+opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
+               cache.path = paste("components/cache/", basename, "/", sep=""))
+opts_chunk$set(cache = 1)
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, comment = NA, verbose = TRUE)
+
+# PDF-based figures
+opts_chunk$set(dev='pdf')
+
+# Better to use a temporary markdown file (best to not overwrite basename.md since that can be used to display on GitHub)
+# For now use basename, since pandoc function in knitr 1.5 doesn't respect output name from config (fixed in knitr 1.5.32)
+output <- paste(basename, ".md", sep="")
+
+## Knit
+knit(file, output)
+
+

### 2014-05-05 manuscripts/manuscript.Rmd
add Steve edits

@@
 abstract: |
-          Model uncertainty and limited data coverage are fundamental challenges to
-          robust ecosystem management.  These challenges are acutely highlighted
-          by concerns that many ecological systems may contain tipping points.
-          Before a collapse, we do not know where the tipping points lie, if the
-          exist at all.  Hence, we know neither a complete model of the system
-          dynamics nor do we have access to data in some large region of state-space
-          where such a tipping point might exist.  These two sources of uncertainty
-          frustrate state-of-the-art parametric approaches to decision theory
-          and optimal control.  I will illustrate how a non-parametric approach
-          using a Gaussian Process prior provides a more flexible representation
-          of this inherent uncertainty.  Consequently, we can adapt the Gaussian
-          Process prior to a stochastic dynamic programming framework in order to
-          make robust management predictions under both model and uncertainty and
-          limited data.
+          Model uncertainty and limited data coverage are fundamental
+          challenges to robust ecosystem management.  These challenges
+          are acutely highlighted by concerns that many ecological systems
+          may contain tipping points.  Before a collapse, we do not know
+          where the tipping points lie, if the exist at all.  Hence,
+          we know neither a complete model of the system dynamics nor
+          do we have access to data in some large region of state-space
+          where such a tipping point might exist.  These two sources of
+          uncertainty frustrate state-of-the-art parametric approaches
+          to decision theory and optimal control.  Here we illustrate
+          how a non-parametric approach using a Gaussian process prior
+          provides a more flexible representation of this inherent
+          uncertainty. We can then embed the Gaussian process in a
+          stochastic dynamic programming framework in order to make
+          robust management predictions under both model uncertainty and
+          limited data. We conduct a set of simulations to evaluate this
+          approach as compared with the standard approach of using model
+          selection to choose from a set of candidate models.  In keeping
+          with previous results, we find that model selection erroneously
+          favors models without tipping points leading to harvest policies
+          that guarantee extinction. The Gaussian process based approach
+          performs nearly as well as the true model and significantly
+          outperforms standard appraoches.
+
 keywords:
@@
 frequently complex and difficult to measure, making uncertainty in our
-understanding a prediction a persistent challenge to effective management.
+understanding a persistent challenge to effective management.
 Decision-theoretic approaches provide a framework to determine the
@@
 the underlying challenges and methods are germane to many other
-conservation and resource management problems.  The economic value and
+conservation and resource management problems.  Economic value and
 ecological concern have made marine fisheries the crucible for much
 of the founding work [@Gordon1954; @Reed1979; @May1979; @Ludwig1982]
-in managing ecosystems under uncertainty.  Global trends [@Worm2006]
-and controversy [@Hilborn2007; @Worm2009] have made understanding these
+on management under uncertainty.  Global declines [@Worm2006]
+and the controversy surrounding their interpretation [@Hilborn2007; @Worm2009] have made understanding these
 challenges all the more pressing.
@@
 [_e.g._ @Williams2001; @Cressie2009;  @Athanassoglou2012].  Of these,
-structural uncertainty incorporates the least a priori knowledge or
-assumptions and is generally the hardest to quantify. Typical approaches
+structural uncertainty is generally the hardest to quantify. Typical approaches
 assume a weak notion of model uncertainty in which the correct model
@@
 
-_Cut the next three paragraphs, since they are covered more concisely in the above paragraph?_
-
-### Process, measurement, and implementation error
-
-Resource management and conservation planning seek to determine the
-optimal set of feasible actions to maximize the value of some objectives
-(e.g @Halpern2013).  Process error, measurement error, implementation
-error [@Reed1979, @Clark1986, @Roughgarden1996, @Sethi2005].  These
-sources of stochasticity in turn mean that model parameters can only
-be estimated approximately, requiring parametric uncertainty also be
-considered [@Ludwig1982].
-
-### Parametric uncertainty
-
-As the parameter values for these models must be estimated from limited
-data, there will always be some uncertainty associated with these values.
-This uncertainty further compounds the intrinsic variability introduced
-by demographic or environmental noise.  The degree of uncertainty in
-the parameter values can be inferred from the data and reflected in the
-estimates of the transition probabilities [@Ludwig1982; @Mangel1988;
-@Hilborn1997; @Schapaugh2013].
-
-### Structural (model) uncertainty
-
-Estimates of parameter uncertainty are only as good as the parametric
-models themselves.  Often we do not understand the system dynamics well
+Often we do not understand the system dynamics well
 enough to know if a model provides a good approximation over the relevant
 range of states and timescales (criteria that we loosely refer to as
-defining the "right" or "true" model.)  So called structural or model
-uncertainty is a more difficult problem than parametric uncertainty.
+defining the "right" or "true" model.)  
 Typical solutions involve either model choice, model averaging, or
@@
 outcomes, but depend on the weights given to all possible outcomes
-[_e.g._ @Weitzman2013].  If we are observing the fluctuations of a
-given fish stock over many years under a fixed harvesting pressure,
-we might develop and test a model that could reasonably predict the
-frequency of a deviation of a given size, even when such a deviation has
-not been previously observed. Yet such predictions are far less reliable
-when extrapolated to a harvest pressure that has not yet been observed.
-Thus, model uncertainty can be particularly challenging in the management
-and decision-making context.
+[_e.g._ @Weitzman2013].
 
@@
 This paper represents the first time the SDP decision-making framework has
-been used without an a priori model of the underlying dynamics through
-the use of the BNP approach.  In contrast to parametric models which
+been used without an a priori model of the underlying dynamics.  In contrast to parametric models which
 can only reflect uncertainty in parameter estimates, the BNP approach
@@
 This permits a much greater uncertainty far from the observed data than
-near the observed data.  These features allow the BNP-SDP approach to
+near the observed data.  These features allow the GPDP approach to
 find robust management solutions in face of limited data and without
@@
 
-The nature of decision-making problems provides a convenient way to compare 
-models.  Rather than compare models in terms of best fit to data or fret over
-the appropriate penalty for model complexity, model performance is defined 
-in the concrete terms of the decision-maker's objective function, which we
-will take as given. (Much argument can be made over the 'correct' objective
+The nature of decision-making problems provides a compelling and pragmatic way to compare 
+models: Rather than compare models in terms of best fit to data, model performance is defined 
+in the concrete terms of the decision-maker's objectives. Although much argument can be made over the 'correct' objective
 function, e.g. how to account for the social value of fish left in the sea
 vs. the commercial value of fish harvested; see @Halpern2013 for further 
-discussion of this issue.  Alternatively, we can always compare model performance
-across multiple potential objective functions.)  The decision-maker
+discussion of this issue.  We can always compare model performance
+across multiple potential objectives.  The important point is that a decision-maker
 does not necessarily need a model that provides the best mechanistic understanding
-or the best long-term outcome, but rather the one that best estimates the 
-probabilities of being in different states as a result of the possible actions.
+or the best long-term outcome, but rather the one that idelivers the best 
+management performance over the period of interest.
 
@@
 -->
-
-Addressing the difficulty posed by extrapolation without knowing the
-true model requires a nonparametric approach to model fitting: one
-that does not assume a fixed structure but rather depends on the size
-of the data (e.g. non-parametric regression or a Dirichlet process).
-This established terminology is nevertheless unfortunate, as (a)
-this approach still involves the estimation of parameters, and (b),
-Statisticians use non-parametric to mean both this property (structure
-is not fixed by the parameters) and an entirely different (and probably
-more familiar) case in which the model does not assume any distribution
-(e.g. non-parametric bootstrap, order statistics).  Some literature
-thus uses the term semi-parametric, which merely adds ambiguity to
-the confusion.
-
-This non-parametric property -- having a structure explicitly dependent on
-the data -- is precisely the property that makes this approach attractive
-in face of the limited data sampling challenges discussed above.
-Having fit a parametric model to some data, the model is completely
-described by the values (or posterior distributions) of it's parameters.
-The non-parametric model is not captured by its parameter values or
-distributions alone. Either the model scales with the complexity of the data
-on which it is estimated (e.g. nonparametric hierarchical approaches such
-as the Dirchlet process) or the data points become themselves part of the model
-specification, as in the nonparametric regression used here.
 
@@
 
+We use a Bayesian nonparameteric regression approach based on Gaussian process priors. 
 The use of Gaussian process (GP) regression (or "kriging" in the geospatial
@@
 in the context ecological modeling and fisheries management in @Munch2005.
+GP models have subsequently been used to test for the presence of Allee
+effects [@SugenoMunch], estimate the maximum reproductive
+rate [@SugenoMunch], determine temporal variation in food
+availability [@Sigourney], and provide a basis for identifying
+model-mispecification [@Thorson].
 An accessible and thorough introduction to the formulation and use of
-GPs can be found in @Rasmussen2006.
-
-<!--
-- Why it is particularly suited to these two problems
-- (Why this is a novel application thereof)
--->
-
-<!-- Useless 
-The essence of the GP approach can be captured in the following
-thought experiment: An exhaustive parametric approach to the challenge
-of structural uncertainty might proceed by writing down all possible
-functional forms for the underlying dynamical system with all possible
-parameter values for each form, and then consider searching over this
-huge space to select the most likely model and parameters; or using a
-Bayesian approach, assign priors to each of these possible models and
-infer the posterior distribution of possible models. The GP approach
-can be thought of as a computationally efficient approximation to this
-approach. GPs represent a large class of models that can be though
-of as capturing or reasonably approximating the set of models in this
-collection.  By modeling at the level of the process, rather than the
-level of parametric equation, we can more concisely capture the possible
-behavior of these curves.  In place of a parametric model of the dynamical
-system, the GP approach postulates a prior distribution of (n-dimensional)
-curves that can be though of as approximations to a range of possible
-(parametric) models that might describe the data. The GP allows us to
-consider probabilities on a large set of possible curves simultaneously.
--->
-
-The posterior distribution for the hyper-parameters of the Gaussian 
-process model are estimated by Metropolis-Hastings algorithm, again with
-details and code provided in the Appendix.  @Rasmussen2006 provides
-an excellent general introduction to Gaussian Processes and @Munch2005a 
-first discusses their application in the context of population dynamics
-models such as fisheries stock-recruitment relationships.
+GPs can be found in @Rasmussen2006. The nonparametric monicker for
+the GP approach is unfortunate as (a) it still involves the estimation
+of parameters, and (b), statisticians also use non-parametric to mean
+models that make only weak distributional assumptions.  (@signTestREF). 
+However, we retain the terminology for consistency with previous
+studies.
 
@@
 
-To illustrate the application of the BNP-SDP approach and compare to the
-predictions of the alternative parametric models we focus on the classical
-problem of selecting the appropriate harvest level given an observation
-of the stock size in the previous year  [@Reed1979; @Ludwig1982; @Mangel1988].
-Given this observation and the model (together with the parameter uncertainty) of the stock
-recruitment process, the manager seeks to maximize the value of the 
-fishery over a fixed time interval of `r OptTime` years at a discount rate
-of `r delta`. The value function (profits) at time $t$ depends on the true stock size
-$x_t$ and the chosen harvest level $h_t$.  For simplicity we assume profit 
-is simply proportional in the realized harvest (only
-enforcing the restriction that harvest can not exceed available stock).
+To illustrate the application of the Gaussian process-based dynamic
+programming approach (hereafter GPDP) we focus on the classical problem
+of determining a harvest policy that will maximize the expected discounted
+profit over a fixed time interval [@Reed1979; @Ludwig1982; @Mangel1988].
+For simplicity we assume profit is proportional to the realized harvest.  
+For comparison, we consider the performance of the GPDP policy against
+policies derived under several alternative parametric models.
 
@@
 <!-- equations just in appendix? -->
-
-As we simulate training data from the Allen model, we will
-refer to this as the structurally correct model.  The Ricker model is
-thus a reasonable approximation of these dynamics far from the Allee
+Because of our interest in management performance in the presence of
+tipping points, all of our simulations are based on the Allen model.
+The Allen model is thus structurally correct and is expected to provide
+a best-case scenario when the 'true' dynamics are known.
+The Ricker model is a reasonable approximation of these dynamics far from the Allee
 threshold (but lacks threshold dynamics), while the Myers model shares
-the essential feature of a threshold but differs in the structure. Thus
-we have three potential parametric models of the stock dynamics.
-
-We introduce parametric uncertainty by first estimating each of the
-candidate models from data on unexploited stock dynamics following
-some perturbation (non-equilibrium initial condition) over several time
-steps. This training data could be generated in several different ways
-(such as known variable exploitation rates, etc.), as long as it reflects
-the dynamics in some limited region of state space without impacting
-the problem.  We consider a period of `r Tobs` years of training data: long
+the essential feature of a threshold but differs in the structure.
+
+We consider a period of `r Tobs` years of training data: long
 enough that the estimates are not dependent on the particular realization,
@@
 
-We infer posterior distributions for the parameters of each model
+We infered posterior distributions for the parameters of each model
 in a Bayesian context using Gibbs sampling (implemented in R [@RTeam]
@@
 provided). One-step-ahead predictions of these model fits are shown in
-Figure 1.  While alternative approaches to the estimation of the posteriors
-(such as integrating out the rate parameter $r$ analytically and then
-performing a grid search over the remaining parameter space), the 
-approach of using a standard Gibbs sampler routine is both more general
-and representitive of common practice in estimating posteriors for such models. 
-Each sampling is tested for Gelman-Rubin convergence and results are robust
-to longer runs.
+Figure 1.  Each chain was tested for Gelman-Rubin convergence and results wer robust
+to longer runs.  For each simulation we also applied several commonly
+used model selection criteria (AIC, BIC, DIC, see @BurnhamAndersonREF)
+to identify the best fitting model.
 
@@
 
-An optimal policy function is then inferred through stochastic dynamic
+An optimal policy function was then inferred through stochastic dynamic
 programming for each model given the posterior distributions of the
@@
 
-We also estimate a simple Gaussian Process defined by a radial basis
+Using the simulated training data we also estimate a Gaussian Process defined by a radial basis
 function kernel of two parameters: $\ell$, which gives the characteristic
@@
 
-Though we are unaware of prior application of this type, it is reasonably
-straight-forward to adapt the Gaussian Process for Stochastic Dynamic
-Programming.  Recall that unlike the parametric models the Gaussian
+It is reasonably straight-forward to derive the harvest policy from the stimated BP by inserting it into a Stochastic Dynamic
+Programming (SDP) algorithm.  Recall that unlike the parametric models the Gaussian
 process with fixed parameters already predicts a distribution of
@@
 Each model manages to fit the observed data rather closely.
-Compared to the expected value of the true model most estimates appear to overfit, predicting fluctuations that are actually due purely to stochasticity in growth rate.
-Model-choice criteria shown in Table 1 penalize more complex models and show a slight preference for the simpler Ricker model over the more complicated alternate stable state models (Allen and Myers).
+Compared to the true model most estimates appear to overfit, predicting fluctuations that are actually due purely to stochasticity.
+All model selection criteria shown in Table 1 penalize more complex models and show a preference for the simpler Ricker model over the more complicated alternate stable state models (Allen and Myers).
 Details on MCMC estimates for each model, traces, and posterior distributions can be found in the appendix.
@@
 ```{r Table1, dependson="deviances", include=TRUE, results="asis", echo=FALSE, cache=FALSE}
-xtable::xtable(rbind(dictable, aictable, bictable), caption="Model choice scores for several common criteria all (wrongly) select the simplest model.
+xtable::xtable(rbind(dictable, aictable, bictable), caption="Model selection scores for several common criteria all (wrongly) select the simplest model.
 As the true (Allen) model is not distinguishable from the simpler (Ricker) model in the region of the observed data, this error cannot be avoided regardless of the model choice criterion.
@@
 value realized by each of the model. Figure 6 shows the distribution of values
-(profits) realized by each simulation shown in Figure 5.  The BNP-SDP
+(profits) realized by each simulation shown in Figure 5.  The GPDP
 approach most consistently realizes a value close to the optimal solution,
@@
 These results are not sensitive to the modeling details of the simulation.
-The BNP-SDP estimate remains very close to the optimal solution obtained by knowning the true model across changes to the training simulation, noise scale, parameters or structure of the underlying model, as seen in Figure 7.
+The GPDP estimate remains very close to the optimal solution obtained by knowning the true model across changes to the training simulation, noise scale, parameters or structure of the underlying model, as seen in Figure 7.
 Results are pooled across different random seeds (111, 222, 333), noise values of 0.01 and 0.05 (not large enough to violate the self-sustaining criterion of @Reed1979), and different randomly generated parameters sets for each model, and using either the Myers or Allen as the underlying structure.
-While the BNP-SDP performs slightly closer to optimal on the smaller noise levels, and closer on the Allen model then the Myers model, the differences are small compared to the values observed in using the wrong parametric model.
-The BNP-SDP solution consistently avoids either failing to harvest at all or crashing the stock, which results in the very low profits/values seen in parametric models in the previous figure.
+While the GPDP performs slightly closer to optimal on the smaller noise levels, and closer on the Allen model then the Myers model, the differences are small compared to the values observed in using the wrong parametric model.
+The GPDP solution consistently avoids either failing to harvest at all or crashing the stock, which results in the very low profits/values seen in parametric models in the previous figure.
 The sensitivity analysis is compared against the optimal solution alone and not shown against the estimates of the various parameteric models (but see Figure 6), where convergence of the Gibbs sampling could contribute to their poorer performance if not checked by hand for each configuration.
@@
 ![Sensitivity Analysis.
-Histogram shows the ratio of the realized net present value / profits derived when managing under the BNP-SDP Gaussian Process approach over the optimal value given the true model and true parameters.
+Histogram shows the ratio of the realized net present value / profits derived when managing under the GPDP Gaussian Process approach over the optimal value given the true model and true parameters.
 Values near 1 indicate nearly optimal performance.
@@
 
-Model choice approaches trade off model complexity and fit to the data. 
+Model selection criteria trade off model complexity and fit to the data. 
 When the data come from a limited region of state-space -- as is necessarily 
 the case whenever there is a potential concern about tipping point dynamics --
-simpler models can fit just as well and will tend to outperform more complex 
+simpler models therefore fit just as well and tend to be selected more foten than complex 
 ones.  This approach would be appropriate when the dynamics can be expected 
@@
 
-### BNP-SDP expresses larger uncertainty in regions where the data are poor
+### GPDP expresses larger uncertainty in regions where the data are poor
 
@@
 
-### BNP-SDP has good predictive accuracy where data are good
-
-While expressing larger uncertainty outside the observed data, the GP
-can also provide a better fit with smaller uncertainty inside the range
-of the observed data. This arises from the greater flexibility of the 
-Gaussian process, which describes a large family of possible curves.
-Despite this flexibility, the GP can be described in relatively few 
-parameters and is thus far less likely to over-fit.
-
 ### Risk-prone and risk-adverse value functions
@@
 to simplify the intuition and avoid biasing this result, we have chosen
-a profits that are linear in the catch and thus neither risk-prone nor
+profits that are linear in the catch and thus neither risk-prone nor
 risk adverse.  Making this function concave, representing the typical
@@
 
-Lastly, it should be noted that outside the data, the NBP reverts to
+Lastly, it should be noted that outside the data, the GP reverts to
 the prior, and consequently the choice of the prior can also play a
@@
 In principle, this may allow a manager to improve the performance of the
-BNP-SDP approach by adding only enough additional detail as is justified.
+GPDP approach by adding only enough additional detail as is justified.
 For instance, it would be possible to use a linear or a Ricker-shaped
 mean in the prior without making the much stronger assumption that the
-Ricker is the structurally correct model.  However, this influence raises
-challenges as well.  For instance, in choosing a trivial mean-zero prior,
-we bias the dynamics to transitions to 0 in step $X_{t+1}$ from any
-stock size $X_t$ in the prior year, in the absence of any other data.
-Future research must make sure that both the prior and the value function
+Ricker is the structurally correct model [@SugenoMunch]. Future research
+should focus on identifying criteria to ensure the prior and the value function
 are chosen appropriately for the problem at hand.
@@
 we anticipate that in higher-dimensional examples characteristic of
-ecosystem management problems that the machine learning approach will
+ecosystem management problems that the GP approach will
 prove even more valuable.
@@
 
-This work was partially supported by the Center for Stock Assessment
+This work was partially supported by NOAA-IAM grant to SM administered through the Center for Stock Assessment
 Research, a partnership between the University of California Santa Cruz

### 2014-05-06 manuscripts/components/sensitivity.R
script for sensitivity analysis part of manuscript

@@ -1,0 +1,193 @@
+models <- c("Myers","Allen")
+
+parameters <- list(Myers = list(
+  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
+  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
+  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2))),
+                   Allen = list(
+  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
+  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
+  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)))
+                   )
+nuisance_pars <- c("sigma_g")
+nuisance_values <- list(sigma_g = c(0.01, 0.05, 0.1))
+replicates <- c(1111, 2222, 3333, 4444, 5555, 6666, 7777, 8888) # seeds
+
+
+sensitivity <- function(model, parameters, nuisance, seed){
+  
+  if(model == "Myers")
+    f <- Myers
+  else if(model == "Allen")
+    f <- RickerAllee
+    
+sigma_g <- nuisance[["sigma_g"]]
+p <- parameters
+sigma_m <- 0.0
+z_g <- function() rlnorm(1, 0, sigma_g)
+z_m <- function() 1
+x_grid <- seq(0, 15, length=50)
+h_grid <- x_grid
+profit <- function(x,h) pmin(x, h)
+delta <- 0.01
+OptTime <- 50  # stationarity with unstable models is tricky thing
+reward <- 0
+xT <- 0
+Xo <- 5.5# observations start from
+x0 <- 8 # simulation under policy starts from
+Tobs <- 40
+MaxT <- 1000 # timeout for value iteration convergence
+
+  
+  # replicate over random seed
+  yields <- sapply(seed, 
+                   function(seed_i){
+                     set.seed(seed_i)
+                     
+                     x <- numeric(Tobs)
+                     x[1] <- Xo
+                     nz <- 1
+                     for(t in 1:(Tobs-1))
[…]

### 2014-05-06 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(nonparametricbayes)
+
+test_check("nonparametricbayes")

### 2014-05-06 manuscripts/manuscript.Rmd
update helping scripts

@@
 
-toggle = "markup" # results argument
+toggle = "hide" # results argument
 ```
@@
 ```{r dic_calc, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output"), include=FALSE, echo=FALSE}
+```
+
+```{r sensitivity-calc}
+source("components/sensitivity.R")
 ```
@@
 GP models have subsequently been used to test for the presence of Allee
-effects [@SugenoMunch], estimate the maximum reproductive
-rate [@SugenoMunch], determine temporal variation in food
+effects [@Sugeno2013], estimate the maximum reproductive
+rate [@Sugeno2013], determine temporal variation in food
 availability [@Sigourney], and provide a basis for identifying
-model-mispecification [@Thorson].
+model-mispecification [@Thorson2014].
 An accessible and thorough introduction to the formulation and use of
@@
 of parameters, and (b), statisticians also use non-parametric to mean
-models that make only weak distributional assumptions.  (@signTestREF). 
+models that make only weak distributional assumptions  [e.g. @Lehmann1975]. 
 However, we retain the terminology for consistency with previous
@@
 to longer runs.  For each simulation we also applied several commonly
-used model selection criteria (AIC, BIC, DIC, see @BurnhamAndersonREF)
+used model selection criteria (AIC, BIC, DIC, see @Burnham2002)
 to identify the best fitting model.
@@
 These results are not sensitive to the modeling details of the simulation.
-The GPDP estimate remains very close to the optimal solution obtained by knowning the true model across changes to the training simulation, noise scale, parameters or structure of the underlying model, as seen in Figure 7.
-Results are pooled across different random seeds (111, 222, 333), noise values of 0.01 and 0.05 (not large enough to violate the self-sustaining criterion of @Reed1979), and different randomly generated parameters sets for each model, and using either the Myers or Allen as the underlying structure.
-While the GPDP performs slightly closer to optimal on the smaller noise levels, and closer on the Allen model then the Myers model, the differences are small compared to the values observed in using the wrong parametric model.
-The GPDP solution consistently avoids either failing to harvest at all or crashing the stock, which results in the very low profits/values seen in parametric models in the previous figure.
-The sensitivity analysis is compared against the optimal solution alone and not shown against the estimates of the various parameteric models (but see Figure 6), where convergence of the Gibbs sampling could contribute to their poorer performance if not checked by hand for each configuration.
+The GPDP estimate remains very close to the optimal solution obtained
+by knowning the true model across changes to the training simulation,
+noise scale, parameters or structure of the underlying model, as seen in
+Figure 7. The sensitivity analysis uses a factorial design of eight replicates 
+of three different randomly generated parameter sets for each of two different 
+generating models (Allen and Myers) over two different noise levels 
+(0.01 and 0.05), for a total of 8 x 3 x 2 x 2 = 96 scenarios. In Figure 7, 
+results are pooled across different random seeds (111, 222,
+333), noise values of 0.01 and 0.05 (not large enough to violate the
+self-sustaining criterion of @Reed1979), and different randomly generated
+parameters sets for each model, and using either the Myers or Allen as the
+underlying structure.  While the GPDP performs slightly closer to optimal
+on the smaller noise levels, and closer on the Allen model then the Myers
+model, the differences are small compared to the values observed in using
+the wrong parametric model.  The GPDP solution consistently avoids either
+failing to harvest at all or crashing the stock, which results in the very
+low profits/values seen in parametric models in the previous figure. The
+sensitivity analysis is compared against the optimal solution alone and
+not shown against the estimates of the various parameteric models (but
+see Figure 6), where convergence of the Gibbs sampling could contribute
+to their poorer performance if not checked by hand for each configuration.
 
-![Sensitivity Analysis.
+```{r sensitivity, dependson="sensitivity-calc", echo=FALSE, fig.cap="Sensitivity Analysis.
 Histogram shows the ratio of the realized net present value / profits derived when managing under the GPDP Gaussian Process approach over the optimal value given the true model and true parameters.
 Values near 1 indicate nearly optimal performance.
-Simulations cover a range of underlying parameter, model, replicate, and noise levels.](figure/sensitivity.pdf)
+Simulations cover a range of underlying parameter, model, replicate, and noise levels."}
+ggplot(dat) + geom_histogram(aes(value, fill=noise)) + xlim(0,1.0) + theme_bw() + xlab("value as fraction of the optimal") + facet_wrap(~model, ncol=1)
+```
 
@@
 mean in the prior without making the much stronger assumption that the
-Ricker is the structurally correct model [@SugenoMunch]. Future research
+Ricker is the structurally correct model [@Sugeno2013]. Future research
 should focus on identifying criteria to ensure the prior and the value function

### 2014-05-06 manuscripts/components/external-chunks.R

@@ -53,7 +53,6 @@ MaxT = 1000 # timeout for value iteration convergence
                     y = c(rep(0,nz), 
                           x[2:Tobs]))
 raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
-raw_plot
 
 
 ## @knitr mle

### 2014-05-06 manuscripts/components/knit.R

@@ -1,6 +1,8 @@
 #!/usr/bin/Rscript
 args <- commandArgs(TRUE)
-file <- args[1]
+
+if(!is.na(args[1]))
+  file <- args[1]
 
 ## FIXME add second argument to handle figure types
 

### 2014-05-06 manuscripts/components/make_pdf.R

@@ -1,33 +1,0 @@
-#!/usr/bin/Rscript
-args <- commandArgs(TRUE)
-file <- args[1]
-
-## Set the paths for cache and figure
-library(methods)
-library(knitr)
-basename <- gsub(".Rmd", "", file) # knitr:::knit_concord$get('infile')) 
-opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
-               cache.path = paste("components/cache/", basename, "/", sep=""))
-opts_chunk$set(cache = 1)
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, comment = NA, verbose = TRUE)
-
-# PDF-based figures
-opts_chunk$set(dev='pdf')
-
-# Better to use a temporary markdown file (best to not overwrite basename.md since that can be used to display on GitHub)
-# For now use basename, since pandoc function in knitr 1.5 doesn't respect output name from config (fixed in knitr 1.5.32)
-output <- paste(basename, ".md", sep="")
-
-## Knit
-knit(file, output)
-
-
-## FIXME Makefile would be happier keeping these as separate targets, which 
-## would permit making the markdown and the tex seperately, and facilitate debugging
-
-## Call pandoc
-pandoc(output, format="latex", config="components/pandoc_config.txt")
-
-# Remove temporary file.
-unlink(output)
-

### 2014-05-06 manuscripts/components/restore-cache.R

@@ -1,7 +1,7 @@
 #!/usr/bin/R
 
 # restore cache file
-archive <- "template-manuscript-cache.tar.gz"
+archive <- "nonparameteric-bayes-cache.tar.gz"
 
 setwd("components")
 download.file(paste0("http://two.ucdavis.edu/~cboettig/data/", archive), archive)

### 2014-05-06 inst/examples/replicates.R
exploring more concise looping

@@ -1,0 +1,208 @@
+models <- c("Myers","Allen")
+
+parameters <- list(Myers = list(
+  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
+  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
+  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2))),
+                   Allen = list(
+  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
+  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
+  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)))
+                   )
+nuisance_pars <- c("sigma_g")
+nuisance_values <- list(sigma_g = c(0.01, 0.05, 0.1))
+replicates <- c(1111, 2222, 3333, 4444, 5555, 6666, 7777, 8888) # seeds
+
+
+sensitivity <- function(model, parameters, nuisance, seed){
+  
+  if(model == "Myers")
+    f <- Myers
+  else if(model == "Allen")
+    f <- RickerAllee
+    
+  sigma_g <- nuisance[["sigma_g"]]
+  z_g <- function() rlnorm(1, 0, sigma_g)
+  p <- parameters
+
+  config = list(
+    f = f,
+    p = p,
+    sigma_g = sigma.g,
+    z_g = z_g,
+    x_grid = seq(0, 15, length=50),
+    h_grid = x_grid,
+    profit = function(x,h) pmin(x, h),
+    delta = 0.01,
+    OptTime = 50,  # stationarity with unstable models is tricky thing
+    reward = 0,
+    xT = 0,
+    Xo = 5.5,# observations start from
+    x0 = 8, # simulation under policy starts from
+    Tobs = 40,
+    MaxT = 1000) # timeout for value iteration convergence
+
+  attach(config)
+  # replicate over random seed
+  yields <- sapply(seed, 
+                   function(seed_i){
+                     set.seed(seed_i)
+
+## Simulate data
+                     x <- numeric(Tobs)
[…]

### 2014-05-06 inst/examples/replicates.R
aic replicates seperately

@@ -1,208 +1,0 @@
-models <- c("Myers","Allen")
-
-parameters <- list(Myers = list(
-  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
-  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
-  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2))),
-                   Allen = list(
-  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
-  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
-  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)))
-                   )
-nuisance_pars <- c("sigma_g")
-nuisance_values <- list(sigma_g = c(0.01, 0.05, 0.1))
-replicates <- c(1111, 2222, 3333, 4444, 5555, 6666, 7777, 8888) # seeds
-
-
-sensitivity <- function(model, parameters, nuisance, seed){
-  
-  if(model == "Myers")
-    f <- Myers
-  else if(model == "Allen")
-    f <- RickerAllee
-    
-  sigma_g <- nuisance[["sigma_g"]]
-  z_g <- function() rlnorm(1, 0, sigma_g)
-  p <- parameters
-
-  config = list(
-    f = f,
-    p = p,
-    sigma_g = sigma.g,
-    z_g = z_g,
-    x_grid = seq(0, 15, length=50),
-    h_grid = x_grid,
-    profit = function(x,h) pmin(x, h),
-    delta = 0.01,
-    OptTime = 50,  # stationarity with unstable models is tricky thing
-    reward = 0,
-    xT = 0,
-    Xo = 5.5,# observations start from
-    x0 = 8, # simulation under policy starts from
-    Tobs = 40,
-    MaxT = 1000) # timeout for value iteration convergence
-
-  attach(config)
-  # replicate over random seed
-  yields <- sapply(seed, 
-                   function(seed_i){
-                     set.seed(seed_i)
-
-## Simulate data
-                     x <- numeric(Tobs)
[…]

### 2014-05-06 manuscripts/components/external-chunks.R

@@ -53,6 +53,7 @@ MaxT = 1000 # timeout for value iteration convergence
                     y = c(rep(0,nz), 
                           x[2:Tobs]))
 raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
+raw_plot
 
 
 ## @knitr mle

### 2014-05-06 manuscripts/components/restore-cache.R

@@ -1,7 +1,7 @@
 #!/usr/bin/R
 
 # restore cache file
-archive <- "nonparameteric-bayes-cache.tar.gz"
+archive <- "nonparametric-bayes-cache.tar.gz"
 
 setwd("components")
 download.file(paste0("http://two.ucdavis.edu/~cboettig/data/", archive), archive)

### 2014-05-06 inst/examples/aic_replicates.Rmd

```{r}
## Set the paths for cache and figure
library(methods)
library(knitr)
basename <- gsub(".Rmd", "", file) # knitr:::knit_concord$get('infile')) 
opts_chunk$set(fig.path = paste("figure/", basename, "-", sep=""),
               cache.path = paste("cache/", basename, "/", sep=""))
opts_chunk$set(cache = 1)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, comment = NA, verbose = TRUE)
```

```{r}
library(nonparametricbayes)

models <- c("Myers","Allen")

parameters <- list(Myers = list(
  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2))),
                   Allen = list(
  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)))
                   )
nuisance_pars <- c("sigma_g")
nuisance_values <- list(sigma_g = c(0.01, 0.05, 0.1))
replicates <- c(1111, 2222, 3333, 4444, 5555, 6666, 7777, 8888) # seeds

## @knitr posterior-mode
require(modeest)
posterior.mode <- function(x) {
  mlv(x, method="shorth")$M
}

sensitivity <- function(model, parameters, nuisance, seed){

  if(model == "Myers")
    f <- Myers
  else if(model == "Allen")
    f <- RickerAllee

  sigma_g <- nuisance[["sigma_g"]]
  z_g <- function() rlnorm(1, 0, sigma_g)
  p <- parameters

  config = list(
    f = f,
    p = p,
    sigma_g = sigma.g,
    z_g = z_g,
    x_grid = seq(0, 15, length=50),
    h_grid = x_grid,
    profit = function(x,h) pmin(x, h),
    delta = 0.01,
    OptTime = 50,  # stationarity with unstable models is tricky thing
    reward = 0,
    xT = 0,
    Xo = 5.5,# observations start from
    x0 = 8, # simulation under policy starts from
    Tobs = 40,
    MaxT = 1000) # timeout for value iteration convergence

  attach(config)
  # replicate over random seed
  yields <- sapply(seed, 
                   function(seed_i){
                     set.seed(seed_i)

## Simulate data
                     x <- numeric(Tobs)
                     x[1] <- Xo
                     nz <- 1
                     for(t in 1:(Tobs-1))
                       x[t+1] = z_g() * f(x[t], h=0, p=p)

                     X = c(rep(0,nz), pmax(rep(0,Tobs-1), x[1:(Tobs-1)]))
                     Y = c(rep(0,nz), x[2:Tobs])

## GP Stuff

                     ## @knitr gp-priors
                     s2.p <- c(5,5)  
                     d.p = c(10, 1/0.1)

                     ## @knitr gp
                     gp <- gp_mcmc(X, y=Y, n=1e5, s2.p = s2.p, d.p = d.p)
                     gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)

                     matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid) 
                     opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)

## @knitr jags-setup
y <- x 
N <- length(x);
jags.data <- list("N","y")
n.chains <- 6
n.iter <- 1e6
n.burnin <- floor(10000)
n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
n.update <- 10

## @knitr common-priors
stdQ_prior_p <- c(1e-6, 100)
stdR_prior_p <- c(1e-6, .1)
stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])

## @knitr allen-model
K_prior_p <- c(0.01, 20.0)
r0_prior_p <- c(0.01, 6.0)
theta_prior_p <- c(0.01, 20.0)

bugs.```
```{r}
model <- 
paste(sprintf(
"model{
  K     ~ dunif(%s, %s)
  r0    ~ dunif(%s, %s)
  theta ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)", 
  K_prior_p[1], K_prior_p[2],
  r0_prior_p[1], r0_prior_p[2],
  theta_prior_p[1], theta_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2]),

  "
  iQ <- 1 / (stdQ * stdQ);
  y[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K)* (y[t] - theta) / K 
    y[t+1] ~ dlnorm(mu[t], iQ) 
  }
}")
writeLines(bugs.model, "allen_process.bugs")

## @knitr allen-priors
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
r0_prior <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
theta_prior <- function(x) dunif(x, theta_prior_p[1], theta_prior_p[2])
par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
                    r0 = r0_prior, theta = theta_prior,
                    stdQ = stdQ_prior)

## @knitr allen-mcmc
jags.params=c("K","r0","theta","stdQ") # be sensible about the order here
jags.inits <- function(){
  list("K"= 10 * rlnorm(1,0, 0.1),
       "r0"= 1 * rlnorm(1,0, 0.1) ,
       "theta"=   5 * rlnorm(1,0, 0.1) , 
       "stdQ"= abs( 0.1 * rlnorm(1,0, 0.1)),
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

set.seed(1234)
# parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
allen_jags <- do.call(jags, list(data=jags.data, inits=jags.inits, 
                                      jags.params, n.chains=n.chains, 
                                      n.iter=n.iter, n.thin=n.thin, 
                                      n.burnin=n.burnin, 
                                      model.file="allen_process.bugs"))

# Run again iteratively if we haven't met the Gelman-Rubin convergence criterion
recompile(allen_jags) # required for parallel
allen_jags <- do.call(autojags, 
											list(object=allen_jags, n.update=n.update, 
                           n.iter=n.iter, n.thin = n.thin))

## @knitr allen-traces
tmp <- lapply(as.mcmc(allen_jags), as.matrix) # strip classes the hard way...
allen_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
## @knitr allen-output
A <- allen_posteriors
A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
pardist <- acast(A, index ~ variable)

## @knitr ricker-model
K_prior_p <- c(0.01, 40.0)
r0_prior_p <- c(0.01, 20.0)
bugs.```
```{r}
model <- 
paste(sprintf(
"model{
  K    ~ dunif(%s, %s)
  r0    ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)", 
  K_prior_p[1], K_prior_p[2],
  r0_prior_p[1], r0_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2]),
  "
  iQ <- 1 / (stdQ * stdQ);
  y[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K) 
    y[t+1] ~ dlnorm(mu[t], iQ) 
  }
}")
writeLines(bugs.model, "ricker_process.bugs")

## @knitr ricker-priors
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
r0_prior <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
                   r0 = r0_prior, stdQ = stdQ_prior)

## @knitr ricker-mcmc
jags.params=c("K","r0", "stdQ")
jags.inits <- function(){
  list("K"= 10 * rlnorm(1,0,.5),
       "r0"= rlnorm(1,0,.5),
       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345) 
ricker_jags <- do.call(jags, 
                       list(data=jags.data, inits=jags.inits, 
                            jags.params, n.chains=n.chains, 
                            n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
                            model.file="ricker_process.bugs"))
recompile(ricker_jags)
ricker_jags <- do.call(autojags, 
                       list(object=ricker_jags, n.update=n.update, 
														n.iter=n.iter, n.thin = n.thin, 
														progress.bar="none"))

## @knitr ricker-traces
tmp <- lapply(as.mcmc(ricker_jags), as.matrix) # strip classes the hard way...
ricker_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
names(ricker_posteriors) = c("index", "variable", "value", "chain")

## @knitr ricker-output
A <- ricker_posteriors
A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
ricker_pardist <- acast(A, index ~ variable)

## @knitr myers-model
r0_prior_p <- c(.0001, 10.0)
theta_prior_p <- c(.0001, 10.0)
K_prior_p <- c(.0001, 40.0)
bugs.```
```{r}
model <- 
paste(sprintf(
"model{
  r0    ~ dunif(%s, %s)
  theta    ~ dunif(%s, %s)
  K    ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)", 
  r0_prior_p[1], r0_prior_p[2],
  theta_prior_p[1], theta_prior_p[2],
  K_prior_p[1], K_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2]),

  "
  iQ <- 1 / (stdQ * stdQ);

  y[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- log(r0)  + theta * log(y[t]) - log(1 + pow(abs(y[t]), theta) / K)
    y[t+1] ~ dlnorm(mu[t], iQ) 
  }
}")
writeLines(bugs.model, "myers_process.bugs")

## @knitr myers-priors
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
r_prior     <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
theta_prior <- function(x) dunif(x, theta_prior_p[1], theta_prior_p[2])
par_priors <- list( deviance = function(x) 0 * x, K = K_prior,
                    r0 = r_prior, theta = theta_prior, 
                    stdQ = stdQ_prior)

## @knitr myers-mcmc
jags.params=c("r0", "theta", "K", "stdQ")
jags.inits <- function(){
  list("r0"= 1 * rlnorm(1,0,.1), 
       "K"=    10 * rlnorm(1,0,.1),
       "theta" = 1 * rlnorm(1,0,.1),  
       "stdQ"= sqrt(0.2) * rlnorm(1,0,.1),
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345)
myers_jags <- do.call(jags, 
                      list(data=jags.data, inits=jags.inits, 
													 jags.params, n.chains=n.chains, 
													 n.iter=n.iter, n.thin=n.thin,
                           n.burnin=n.burnin, 
                           model.file="myers_process.bugs"))
recompile(myers_jags)
myers_jags <- do.call(autojags, 
                      list(myers_jags, n.update=n.update, 
                           n.iter=n.iter, n.thin = n.thin, 
                           progress.bar="none"))

## @knitr myers-traces
tmp <- lapply(as.mcmc(myers_jags), as.matrix) # strip classes
myers_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
names(myers_posteriors) = c("index", "variable", "value", "chain")
## @knitr myers-output
A <- myers_posteriors
A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
myers_pardist <- acast(A, index ~ variable)

allen_deviance  <- - posterior.mode(pardist[,'deviance'])
ricker_deviance <- - posterior.mode(ricker_pardist[,'deviance'])
myers_deviance  <- - posterior.mode(myers_pardist[,'deviance'])
true_deviance   <- 2*estf(c(p, sigma_g))
mle_deviance    <- 2*estf(c(est$p, est$sigma_g))
aictable <- data.frame(Allen = allen_deviance + 2*(1+length(bayes_pars)),  # +1 for noise parameter
                       Ricker = ricker_deviance + 2*(1+length(ricker_bayes_pars)),
                       Myers = myers_deviance + 2*(1+length(myers_bayes_pars)), 
                       row.names = c("AIC"))
bictable <- data.frame(Allen = allen_deviance + log(length(x))*(1+length(bayes_pars)), 
                       Ricker = ricker_deviance + log(length(x))*(1+length(ricker_bayes_pars)),
                       Myers = myers_deviance + log(length(x))*(1+length(myers_bayes_pars)), 
                       row.names = c("BIC"))

df <- rbind(dictable, aictable, bictable)

                   })

  dat <- melt(yields, id=names(yields[[1]]))
}
```

```{r}
model <- "Allen"
allen1.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```
```{r}
model <- "Allen"
allen2.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```
```{r}
model <- "Allen"
allen1.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))
```
```{r}
model <- "Allen"
allen2.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))

```
```{r}
model <- "Myers"
Myers1.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```
```{r}
model <- "Myers"
Myers2.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```
```{r}
model <- "Myers"
Myers1.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))

```
```{r}
model <- "Myers"
Myers2.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
## Assemble into data.frame
allen_dat <- rbind(allen1.01, allen1.05, 
             allen2.01, allen2.05) 
myers_dat <- rbind(Myers1.01, Myers1.05, 
             Myers2.01, Myers2.05)
```

### 2014-05-06 manuscripts/manuscript.Rmd

@@
 ```
-```{r obs, dependson="sdp-pars"}
+```{r obs, dependson="sdp-pars", fig.keep='none'}
 ```
@@
 
-```{r sensitivity, dependson="sensitivity-calc", echo=FALSE, fig.cap="Sensitivity Analysis.
+```{r sensitivity, fig.height=6, dependson="sensitivity-calc", echo=FALSE, fig.cap="Sensitivity Analysis.
 Histogram shows the ratio of the realized net present value / profits derived when managing under the GPDP Gaussian Process approach over the optimal value given the true model and true parameters.

### 2014-05-07 manuscripts/manuscript.Rmd
support for linenumbering, fix refs, doublespace, add refs, add tex output

@@
 ---
+layout: 12pt,review
 title: "Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach"
@@
           with previous results, we find that model selection erroneously
-          favors models without tipping points leading to harvest policies
-          that guarantee extinction. The Gaussian process based approach
+          favors models without tipping points -- leading to harvest policies
+          that guarantee extinction. The Gaussian process approach
           performs nearly as well as the true model and significantly
@@
 system outside the range of observed behavior, and (b) decision-theoretic
-alogrithms rely not only on reasonable estimates of the expected
+algorithms rely not only on reasonable estimates of the expected
 outcomes, but depend on the weights given to all possible outcomes
@@
 
-<!-- necessary? --> <!-- should reflect the dual problem of extrapolation and model uncertainty better -->
-
-This paper represents the first time the SDP decision-making framework has
-been used without an a priori model of the underlying dynamics.  In contrast to parametric models which
-can only reflect uncertainty in parameter estimates, the BNP approach
-provides a more state-space dependent representation of uncertainty.
-This permits a much greater uncertainty far from the observed data than
-near the observed data.  These features allow the GPDP approach to
-find robust management solutions in face of limited data and without
-knowledge of the correct model structure.
+This paper represents the first ecological application of the SDP
+decision-making framework without an a priori model of the underlying
+dynamics.  In contrast to parametric models which can only reflect
+uncertainty in parameter estimates, the BNP approach provides a more
+state-space dependent representation of uncertainty.  This permits
+a much greater uncertainty far from the observed data than near the
+observed data.  These features allow the GPDP approach to find robust
+management solutions in face of limited data and without knowledge of
+the correct model structure.
 
@@
 rate [@Sugeno2013], determine temporal variation in food
-availability [@Sigourney], and provide a basis for identifying
-model-mispecification [@Thorson2014].
+availability [@Sigourney2012], and provide a basis for identifying
+model-misspecification [@Thorson2014].
 An accessible and thorough introduction to the formulation and use of
-GPs can be found in @Rasmussen2006. The nonparametric monicker for
+GPs can be found in @Rasmussen2006. The nonparametric moniker for
 the GP approach is unfortunate as (a) it still involves the estimation
@@
 
-We infered posterior distributions for the parameters of each model
+We inferred posterior distributions for the parameters of each model
 in a Bayesian context using Gibbs sampling (implemented in R [@RTeam]
@@
 provided). One-step-ahead predictions of these model fits are shown in
-Figure 1.  Each chain was tested for Gelman-Rubin convergence and results wer robust
+Figure 1.  Each chain was tested for Gelman-Rubin convergence and results were robust
 to longer runs.  For each simulation we also applied several commonly
@@
 
-Using the simulated training data we also estimate a Gaussian Process defined by a radial basis
+Using the simulated training data we also estimate a Gaussian process defined by a radial basis
 function kernel of two parameters: $\ell$, which gives the characteristic
@@
 given an observation of the previous state, $X_t$. @Munch2005a gives an
-accessible introduction to the use of Gaussian Processes in providing a
+accessible introduction to the use of Gaussian processes in providing a
 Bayesian nonparametric description of the stock-recruitment relationship.
@@
 
-It is reasonably straight-forward to derive the harvest policy from the stimated BP by inserting it into a Stochastic Dynamic
-Programming (SDP) algorithm.  Recall that unlike the parametric models the Gaussian
-process with fixed parameters already predicts a distribution of
-curves rather than a single curve. We must first integrate over 
-this distribution of curves given a sampling of parameter values drawn
-from the posterior distribution of the two GP parameters, before
-integrating over the posterior of those parameters themselves.
+It is reasonably straight-forward to derive the harvest policy from the
+estimated BP by inserting it into a Stochastic Dynamic Programming (SDP)
+algorithm.  Recall that unlike the parametric models the Gaussian process
+with fixed parameters already predicts a distribution of curves rather
+than a single curve. We must first integrate over this distribution of
+curves given a sampling of parameter values drawn from the posterior
+distribution of the two GP parameters, before integrating over the
+posterior of those parameters themselves.
 
@@
 =======
+
+To ensure our results are robust to our choice of parameters, 
+we perform our analysis over 96 different scenarios. To better
+understand the process, we will first describe in detail the
+results of a single scenario.
 
@@
 Each model manages to fit the observed data rather closely.
-Compared to the true model most estimates appear to overfit, predicting fluctuations that are actually due purely to stochasticity.
+Compared to the true model most estimates appear to over-fit, predicting fluctuations that are actually due purely to stochasticity.
 All model selection criteria shown in Table 1 penalize more complex models and show a preference for the simpler Ricker model over the more complicated alternate stable state models (Allen and Myers).
@@
 The Ricker model always predicts population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).
-Note that the GP is both more pessimistic and more uncertain about the future state than the paramteric models, while the realized state often falls outside of the expected range forecast by the structurally incorrect Myers and Ricker models." }
+Note that the GP is both more pessimistic and more uncertain about the future state than the parametric models, while the realized state often falls outside of the expected range forecast by the structurally incorrect Myers and Ricker models." }
 ```
@@
 
-These results are robust across a range of stochastic realizations,
-models, and parameter values.  The results across this range can most
+The results across replicate stochastic simulations can most
 easily be compared by using the relative differences in net present
@@
 The GPDP estimate remains very close to the optimal solution obtained
-by knowning the true model across changes to the training simulation,
+by knowing the true model across changes to the training simulation,
 noise scale, parameters or structure of the underlying model, as seen in
@@
 sensitivity analysis is compared against the optimal solution alone and
-not shown against the estimates of the various parameteric models (but
+not shown against the estimates of the various parametric models (but
 see Figure 6), where convergence of the Gibbs sampling could contribute
@@
 ```{r sensitivity, fig.height=6, dependson="sensitivity-calc", echo=FALSE, fig.cap="Sensitivity Analysis.
-Histogram shows the ratio of the realized net present value / profits derived when managing under the GPDP Gaussian Process approach over the optimal value given the true model and true parameters.
+Histogram shows the ratio of the realized net present value / profits derived when managing under the GPDP Gaussian process approach over the optimal value given the true model and true parameters.
 Values near 1 indicate nearly optimal performance.
@@
 In this paper we have tackled two aspects of uncertainty that are both
-common to many ecological desicion-making problems and fundamentally
+common to many ecological decision-making problems and fundamentally
 challenging to existing approaches which largely rely on parametric
@@
 
+Gaussian process model
+----------------------
+
 Assuming the data are known with some process noise,
@@
 
-$$x|y \sim \mathcal{N}(E,C)$$
+$$f(x) \sim \mathcal{N}(E,C)$$
 $$E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y$$
@@
 }
+
+Where $K$ is the Gaussian process kernel.  Throughout, we use the radial basis function kernel:
+
+$$ K(x,y) = \exp\left(\frac{-(x-y)^2}{2 \ell^2} \right)$$
 
@@
 
-The Gaussian process priors on both the lengthscale $\ell$ and process noise $\sigma$ are are inverse Gamma distributed,
+The Gaussian process priors on both the lengthscale $\ell$ and process noise $\sigma$ parameters are both vague inverse Gamma,
 
@@
 
-Form this we can write down the Bellman recursion as:
+From this we can write down the Bellman recursion as:
 
@@
 
-Training data
--------------
+Training data (Figures 1-6)
+----------------------------
 
-Eacho of our models $f(S_t)$ must be estimated from training data, which
+Each of our models $f(S_t)$ must be estimated from training data, which
 we simulate from the Allen model with parameters $r = $ `r p[1]`, 
@@
 
+Training data for sensitivity analyses
+-------------------------------------
+
+A further 96 unique training data sets are generated for the sensitivity analysis, as described in the main text.
+
 Code
@@
 
-A copy of the script to reconstruct the simulations and analysis shown here is provided in the supplemental materials, and through this version-stable link to the project's Github code repostitory, [nonparameteric-bayes.R](https://github.com/cboettig/nonparametric-bayes/blob/371114ec43bb32e92f330f049b420047edae5eb0/inst/doc/nonparametric-bayes.R).
-This code is dynamically embedded into the manuscript using Knitr, [@knitr].
-The script relies on custom routines for executing the estimation of the Gaussian process and the for solving the stochastic dynamic programming problem.
-The sensitivity analysis is provided as a separate script, [sensitivity.Rmd](https://github.com/cboettig/nonparametric-bayes/blob/62e0f81f9e8eabd0f7ee462f098320409a407f81/inst/examples/sensitivity.Rmd).
-These routines are provided as an R package, [nonparameteric-bayes](https://github.com/cboettig/nonparametric-bayes), also available on Github.
+All code used in producing this analysis can be found in the R package accompanying the paper, available from [http://github.com/cboettig/nonparametric-bayes].
+The code for all results shown here is dynamically embedded into the manuscript using Knitr, [@knitr].
 
+<!--
 Data
@@
 While the data can be regenerated using the code provided, for convenience CSV files of the data shown in each graph are also provided, along with appropriate metadata written in the Ecological Metadata Language (EML).
+
+-->
 

### 2014-05-07 inst/examples/aic_replicates.Rmd

@@
 library(knitr)
-basename <- gsub(".Rmd", "", file) # knitr:::knit_concord$get('infile')) 
+basename <- gsub(".Rmd", "", knitr:::knit_concord$get('infile')) 
 opts_chunk$set(fig.path = paste("figure/", basename, "-", sep=""),
@@
 }
+```
 
+```{r}
 sensitivity <- function(model, parameters, nuisance, seed){
@@
 
-bugs.```
-```{r}
-model <- 
+bugs.model <- 
 paste(sprintf(
@@
 r0_prior_p <- c(0.01, 20.0)
-bugs.```
-```{r}
-model <- 
+bugs.model <- 
 paste(sprintf(
@@
 K_prior_p <- c(.0001, 40.0)
-bugs.```
-```{r}
-model <- 
+bugs.model <- 
 paste(sprintf(

### 2014-05-08 R/gp-replicate-fns.R
clean gh-pages

@@ -1,233 +1,0 @@
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-
-
-#' Simulate observed data for training the GP
-#' 
-#' @param Xo the initial population size
-#' @param z_g a function of no arguments providing the random multiplicative growth noise
-#' @param f a function of (x, h, p) providing the growth in population x under harvest h given pars p
-#' @param p the parameters of function f
-#' @param Tobs the number of timesteps on which to observe the system
-#' @param seed the random number seed for the simulation
-#' @param nz the number of observations at 0 (a hack, should just condiition on 0,0)
-#' @param harvest a sequence of harvest values that occur durning the simulation
-#' @details varying harvest values allow the system to explore the state space, making for better training data.
-#' @export
-sim_obs <- function(Xo, z_g, f, p, Tobs = 35, seed = 1, nz = 10, 
-                    harvest = sort(rep(seq(0, .5, length=7), 5)),
-                    plotdata=FALSE){
-  x <- numeric(Tobs)
-  x[1] <- Xo
-  set.seed(seed)
-  for(t in 1:(Tobs-1))
-    x[t+1] = z_g() * f(x[t], h=harvest[t], p=p)
-  if(plotdata)
-    plot(x)
-  obs <- data.frame(x = c(rep(0,nz), 
-                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)] -
-                                 harvest[1:Tobs-1])), 
-                    y = c(rep(0,nz), 
-                          x[2:Tobs]))
-  #plot(obs$x, obs$y)
-  obs
-}
-
-
-#' MLE estimate of parameters given by function f
-#' @param obs the observed data, two columns giving x_t, x_t+1 respectively
-#' @return a list with the f given, the MLE estimated parameters, and estimated noise level
-#' @export 
-par_est_allee <- function(obs, f, p, 
-                          init = c(p[1]+1, p[2]-1, p[3]-2, 
-                                   s = sigma_g + abs(rnorm(1,0,.1)))
-                          ){
[…]

### 2014-05-08 R/gp_fit.R

@@ -1,143 +1,0 @@
-
-# edit this file for a demo
-
-
-#' gp_fit fits a Gaussian process to the observed data
-#' 
-#' API not yet stable! Arguments to this function will probably be updated to be much more generic
-#' @param obs a data frame of observations with columns obs$x and obs$y
-#' @param X the desired points over which to predict
-#' @param pars a named numeric specifying "sigma_n" for the (additive) noise and "l" for the covariance length scale
-#' @param method select the method to use
-#' @param out_var optionally force kernlab method to generate Cf using a different variance than sigma_n^2 assumed in the fit
-#' @return a list with items "mu", the expected Y values at X (mean of the posterior Gaussian process), 
-#' Sigma, the covariance matrix for the posterior Gaussian process, and 
-#' "loglik", the log likelihood of observering the given data under the process, marginalized over the prior
-#' @details so far treates the prior as mean 0 and covariance given by cov(X)
-#' @export
-#' @examples 
-#' X <- seq(-5,5,len=50)
-#' obs <- data.frame(x = c(-4, -3, -1,  0,  2),
-#'                   y = c(-2,  0,  1,  2, -1))
-#' gp <- gp_fit(obs, X, c(sigma_n=0.8, l=1))
-#' 
-#' df <- data.frame(x=X, y=gp$mu, ymin=(gp$mu-2*sqrt((gp$var))), ymax=(gp$mu+2*sqrt((gp$var))))
-#' require(ggplot2)
-#' ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") + geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y))
-
-gp_fit <- function(obs, X, pars=c(sigma_n=1, tau=1, l=1), method=c("direct", "sequential", "kernlab", "cholesky", "conditional"), fit = TRUE, out_var=NULL, ...){
-  
-  method <- match.arg(method)
-  sigma_n <- pars["sigma_n"]
-  l <- pars["l"]
[…]

### 2014-05-08 R/gp_mcmc.R

@@ -1,159 +1,0 @@
-#' Basic regression in Gaussian processes  
-#' 
-#' 
-#' @param x Observed x values, (vector or matrix with columns for each dimension of data)
-#' @param y Vector of observed y values in the training data
-#' @param init_pars the initial guesses for lengthscale l and process noise sigma_n
-#' @param n iterations of the metropolis algorithm
-#' @details Currently assumes the covariance function.  By default we will use 
-#' the squared exponential (also called radial basis or Gaussian, 
-#' though it is not this that gives Gaussian process it's name; 
-#' any covariance function would do) as the covariance function, 
-#' \deqn{\operatorname{cov}(X_i, X_j) = \exp\left(-\frac{(X_i - X_j)^2}{2 \ell ^ 2}\right)}{}
-#' @return the MCMC output, as constructed by \link{metrop} from the mcmc package.
-#' @export
-#' @import mcmc  
-gp_mcmc <- function(x, y, init_pars = c(l=1, sigma.n=1), n = 1e4, d.p = c(5,5), s2.p = c(5,5)){
-
-  lpriors <- function(pars){
-    prior <- unname(
-      dgamma(exp(pars[1]), d.p[1], scale = d.p[2]) *
-      dgamma(exp(pars[2]), s2.p[1], s2.p[2]) 
-    )
-    log(prior)
-  }
-  posterior <- function(pars, x, y){ 
-    l <- exp(pars[1])
-    sigma.n <- exp(pars[2])
-    
-    cov <- function(X, Y) outer(X, Y, SE, l)
-    I <- diag(1, length(x))
-    K <- cov(x, x) 
-    
-    loglik <- - 0.5 * t(y) %*% solve(K + sigma.n^2 * I) %*% y -
-      log(det(K + sigma.n^2*I)) -
-      length(y) * log(2 * pi) / 2
-    
-    loglik + lpriors(pars)
-  }
-  
-  out <- metrop(posterior, log(init_pars), n, x = x, y = y)  
-  
-  # assemble output
-  obs = list(x=x,y=y)
-  out$d.p <- d.p
-  out$s2.p <- s2.p
-  out$obs <- obs
-  out
-}
-
-#' predict the expected values and posterior distributions of the Gaussian Process
-#' 
-#' @param x_predict the values at which we desire predictions
[…]

### 2014-05-08 R/gp_trasition_matrix.R

@@ -1,106 +1,0 @@
-  
-#' Determine the transition matrix at each h given a Guassian process
-#' 
-#' @param Ef GP mean
-#' @param V GP vector of variances
-#' @param x_grid x grid over which to define transition matrices
-#' @param h_grid the harvest grid over which to compute.  Assumed to match the stock grid x_grid if not given.
-#' @return a list of matrices corresponding to the state transition matrix at each harvest level h
-#' @export
-gp_transition_matrix <- function(Ef, V, x_grid, h_grid=NULL){  
-  if(is.null(h_grid))
-    h_grid <- x_grid  
-  if(is.matrix(Ef)){ # loop over posteriors
-    n_samples <- dim(Ef)[2]
-    loop_over_h <-function(h){
-      # Average across matrices in list, and then normalize
-      F_s <- gp_calc_F(h, Ef[,1], V[,1], x_grid)
-      for(i in 2:n_samples){
-        F_s <- F_s + gp_calc_F(h, Ef[,i], V[,i], x_grid)
-      }
-      F_s <- rownorm(F_s)
-    }
-    out <- lapply(h_grid, loop_over_h)  
-  } else { 
-    out <- lapply(h_grid, gp_calc_F, Ef, V, x_grid)
-  }
-  out
-}
-
-
-
-gp_calc_F <- function(h, Ef, V, x_grid){
-  j <- which.min(abs(x_grid - h))
-  indices <- 1:length(Ef)
-  shift <- pmax(indices-j,1)
-  mu <- Ef[shift]
-  mu[shift == 1] <- 0
-  F_ <- t(sapply(1:length(mu), function(i){    
-    if(snap_to_grid(mu[i],x_grid) < x_grid[2]){ # 
-      out <- numeric(length(x_grid))
-      out[1] <- 1
-      out
-    } else {
-      out <- dnorm(x_grid, mu[i], sqrt(V[i]))
-    }
-  }))
-  F_ <- rownorm(F_)
-}
-
-
-
-
-
-#' calculate transition matrix at each harvest level based on parameterized function
-#' 
-#' @param f the transtion function, a function of state x, harvest h, and parameters p
-#' @param p the parameters p of the transition function
-#' @param x_grid the grid over which to solve 
-#' @param h_grid the harvest grid over which to compute.  Assumed to match the stock grid gp$X if not given.
-#' @param sigma_g the scale of the noise
-#' @return a list of matrices corresponding to the state transition matrix at each harvest level h
-#' @export
[…]

### 2014-05-08 inst/examples/gaussian-process-control.R

@@ -1,359 +1,0 @@
-
-## @knitr libraries
-require(pdgControl)
-require(nonparametricbayes)
-require(reshape2)
-require(ggplot2)
-require(data.table)
-require(tgp)
-library(kernlab)
-require(MCMCpack)
-require(plyr)
-
-
-
-## @knitr graphing-options
-opts_knit$set(upload.fun = function(file) socialR::notebook.url(file, cp=TRUE, sync=FALSE))
-#opts_chunk$set(dev = 'Cairo_pdf', dev.args=list(""))
-opts_chunk$set(dev="png", dev.args=list(bg="transparent"))
-opts_chunk$set(comment=NA, tidy=FALSE)
-theme_set(theme_bw(base_size=16))
-theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
-             plot.background = element_rect(fill = "transparent",colour = NA))
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-#sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
-#short_sha <- gsub("(^.{10}).*", "\\1", sha)
-#date <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S")
-#opts_chunk$set(fig.path = paste("figure/", date, "-", short_sha, "-", sep=""))
-
-
-
-
-
-
-## @knitr BevHolt
-f <- BevHolt
-p <- c(1.5,.05)
-K <- (p[1]-1)/p[2] 
-
-
-
-## @knitr Myer
-f <- Myer_harvest
-pars <- c(1, 2, 4.5) 
-p <- pars # shorthand 
-K <- p[1] * p[3] / 2 + sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2
-allee <- p[1] * p[3] / 2 - sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 # allee threshold
-e_star <- (p[1] * sqrt(p[3]) - 2) / 2 ## Bifurcation point 
-
-
-## @knitr Ricker
-f <- Ricker
-p <- c(2.5, 10) 
-K <- 12.5
-
-
-## @knitr May
-f <- May
-p <- c(r = .75, k = 10, a=1.3, H=1, Q = 3)
-K <- 8 
-
-## @knitr plot-may
[…]

### 2014-05-08 inst/examples/scenarios-table.R

@@ -1,22 +1,0 @@
-scenarios <- melt(parameters)
-
-ids <- c(rep(1:3, 3), rep(4:6, 3))
-
-scenarios <- cbind("scenario"= ids, scenarios)
-names(scenarios) <- c("scenario", "value", "parameter", "model")
-scenarios <- scenarios[c("scenario", "model", "parameter", "value")]
-
-Allen <- scenarios$model == "Allen"
-Myers <- scenarios$model == "Myers"
-
-allen_ids <- scenarios[Allen, "parameter"]
-myers_ids <- scenarios[Myers, "parameter"]
-
-allen_names <- names(parameters$Allen[[1]])
-myers_names <- names(parameters$Myers[[1]])
-
-scenarios[Allen, "parameter"] <- allen_names[allen_ids]
-scenarios[Myers, "parameter"] <- myers_names[myers_ids]
-
-
-scenarios

### 2014-05-08 manuscripts/components/knit.R

@@ -1,28 +1,0 @@
-#!/usr/bin/Rscript
-args <- commandArgs(TRUE)
-
-if(!is.na(args[1]))
-  file <- args[1]
-
-## FIXME add second argument to handle figure types
-
-## Set the paths for cache and figure
-library(methods)
-library(knitr)
-basename <- gsub(".Rmd", "", file) # knitr:::knit_concord$get('infile')) 
-opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
-               cache.path = paste("components/cache/", basename, "/", sep=""))
-opts_chunk$set(cache = 1)
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, comment = NA, verbose = TRUE)
-
-# PDF-based figures
-opts_chunk$set(dev='pdf')
-
-# Better to use a temporary markdown file (best to not overwrite basename.md since that can be used to display on GitHub)
-# For now use basename, since pandoc function in knitr 1.5 doesn't respect output name from config (fixed in knitr 1.5.32)
-output <- paste(basename, ".md", sep="")
-
-## Knit
-knit(file, output)
-
-

### 2014-05-08 BUGS/long_observation_period.Rmd
add md files from /inst/examples on master branch

--- 2013-05-07 inst/examples/BUGS/short_observation_period.Rmd
@@
 delta <- 0.01
-OptTime <- 70  # stationarity with unstable models is tricky thing
+OptTime <- 20  # stationarity with unstable models is tricky thing
 reward <- 0
@@
 x0 <- Xo # simulation under policy starts from
-Tobs <- 20
+Tobs <- 70
 ```

### 2014-05-08 BUGS/myers.Rmd
add md files from /inst/examples on master branch

--- 2013-06-07 inst/examples/BUGS/myers.Rmd
@@
 
-```{r myers-mcmc, dependson="myers-model"}
+```{r myers-mcmc-local, dependson="myers-model"}
+jags.params=c("r0", "theta", "K", "stdQ")
+jags.inits <- function(){
+  list("r0"= 1.5 * rlnorm(1,0,.2), 
+       "K"=    8 * rlnorm(1,0,.1),
+       "theta" = 2.5 * rlnorm(1,0,.1),  
+       "stdQ"= sqrt(0.2) * rlnorm(1,0,.1),
+       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+}
+set.seed(12345)
+myers_jags <- do.call(jags.parallel, 
+                      list(data=jags.data, inits=jags.inits, 
+    											 jags.params, n.chains=n.chains, 
+													 n.iter=n.iter, n.thin=n.thin,
+                           n.burnin=n.burnin, 
+                           model.file="myers_process.bugs"))
+recompile(myers_jags)
+myers_jags <- do.call(autojags, 
+                      list(myers_jags, n.update=n.update, 
+                           n.iter=n.iter, n.thin = n.thin, 
+                           progress.bar="none"))
 ```

### 2014-05-08 admb-example.Rmd
add md files from /inst/examples on master branch

--- 2013-06-03 inst/examples/admb-example.Rmd
@@
 K <- 10  # approx, a li'l' less
-allee <- 5 # approx, a li'l' less
+Xo <- 6 # approx, a li'l' less
 ```
@@
 ```{r sdp-pars, dependson="stateeq"}
-sigma_g <- 0.05
+sigma_g <- 0.1
 z_g <- function() rlnorm(1,0, sigma_g)
 x_grid <- seq(0, 1.5 * K, length=50)
-Xo <-  allee + 0.5 # observations start from
 Tobs <- 40
@@
   x[t+1] = z_g() * f(x[t], h=0, p=p)
-
+qplot(1:Tobs, x)
 ```
@@
 ```{r mle, dependson="obs"}
+STABLIZE = 1e-10
 n = length(x)
-loglik <- function(p){ 
-  r = p[1]; k = p[2]; c = p[3]; s = p[4];
-  mu = x + r * (1 - x / k) * (x - c) / k;
-  0.5 * n * log(2 * pi) + n * log(s) + 0.5 * sum(x - mu)^2/ s^2;
-  #-sum(dlnorm(x, mu, s), log=TRUE) # above equation is equivalent to results from built-in function
+mloglik <- function(pars){ 
+  r = pars[1]; k = pars[2]; c = pars[3]; s = pars[4];
+  mu = (x+STABLIZE) * exp( r * (1 - x / (k+STABLIZE)) * (x - c) / (k + STABLIZE));
+  mu = pmin(1e100, mu) # avoid infinite values 
+  f = 0.5 * n * log(2 * pi) + n * log(s + STABLIZE) + 0.5 * sum(x - mu + STABLIZE)^2/ (s + STABLIZE)^2;
+
+  f
   }
-init <- c(1, 1, 1, 1)
-o <- optim(init, loglik, method="L", lower=1e-5, upper=100)
+```
+
+Starting from the true values we mostly just shrink the noise parameter:
+
+```{r}
+init <- c(p, sigma_g)
+mloglik(init) #true minus loglik
+o <- optim(init, mloglik, method="L", lower=1e-5, upper=1e5)
+o$value
+o$par
+```
+
+While starting from arbitrary values we still find the optim.
+
+```{r}
+init <- c(1,1,1,1)  
+init <- c(p, sigma_g)
+mloglik(init) #true minus loglik
+o <- optim(init, mloglik, method="L", lower=1e-5, upper=1e5)
+o$value
 o$par
@@
 PROCEDURE_SECTION
-  mu = x + r * elem_prod((1 - x / k), (x - c) / k);
-  f = 0.5 * n * log(2 * M_PI) + n * log(s) + 0.5 * norm2(x - mu) / square(s);
+  mu = log(x) + r * elem_prod((1 - x / k), (x - c) / k);
+  f = 0.5 * n * log(2 * M_PI) + n * log(s) + 0.5 * norm2(x - exp(mu)) / square(s);
 ")
@@
 
-```{r run-admb}
+Without explicit handling of the overflow errors, ADMB does not give us reliable estimates with arbitrary starting conditions
+
+```{r run-admb1}
 setup_admb("/var/admb")
@@
 df <- data.frame(x=x)
-params <- list(r = 1, k = 10, c = 2, s = .1) ## starting parameters
-bounds <- list(r = c(1e-5, 100), k=c(1e-5, 100), c=c(1e-5, 100), s = c(1e-5,100)) ## bounds
+params <- list(r = 1, k = 1, c = 1, s = 1) ## starting parameters
+bounds <- list(r = c(1e-10, 1e3), k=c(1e-10, 1e3), c=c(1e-10, 1e3), s = c(1e-5,1e3)) ## bounds
 dat <- c(list(n = nrow(df)), df)
@@
 
-Which now matches the results from `optim`, now that we avoid the k=0 bound.
+But does fine with good starting values.  Hmm.. thought that was supposed to be the other way around...
+
+```{r run-admb}
+params <- list(r = 1, k = 10, c = 5, s = .1) ## starting parameters
+
+m1 <- do_admb("model",
+              data = dat,
+              params = params,
+              bounds = bounds,
+              run.opts = run.control(checkparam="write",
+                                     checkdata="write"))
+m1
+```
+
+Which finds a better optim (though substantailly overfit in reality)
 

### 2014-05-08 allee.Rmd
add md files from /inst/examples on master branch

--- 2013-04-25 inst/examples/allee.Rmd
@@
   gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
+          meanfn="constant", bprior="b0", BTE=c(10,1600,2),
           m0r1=FALSE, corr="exp", trace=TRUE, 

### 2014-05-08 myer-example.Rmd
add md files from /inst/examples on master branch

--- 2012-12-15 inst/examples/myer-example.Rmd
@@
-```{r set-options, echo = FALSE, cache = FALSE, external = TRUE}
-opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "myers-cache/")
+```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
+opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "myers-cache/", warning=FALSE)
 read_chunk('gaussian-process-control.R')
@@
 
+```{r x0-obs}
+x_0_observed <- allee + x_grid[5]
+xT <- 0
+set.seed(1)
+```
+
 ```{r sim-obs}
@@
 
+```{r gp-priors}
+```
+
 ```{r gp-fit, dependson="lag-data"}

### 2014-05-08 sensitivity.Rmd
add md files from /inst/examples on master branch

```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls())
source("~/.knitr_defaults.R")
opts_knit$set(upload.fun = socialR::flickr.url)
opts_chunk$set(error=TRUE)
library(knitcitations)
library(nonparametricbayes) 
library(reml)
```

```{r}
models <- c("Myers","Allen")

parameters <- list(Myers = list(
  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2))),
                   Allen = list(
  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)))
                   )
nuisance_pars <- c("sigma_g")
nuisance_values <- list(sigma_g = c(0.01, 0.05, 0.1))
replicates <- c(1111, 2222, 3333, 4444, 5555, 6666, 7777, 8888) # seeds

```

Rather than attempt a large list of nested loops, we do better to launch seperate scripts and write to a common data format.  We define a function that takes as arguments all things we wish to vary in the senstivity analysis, and returns a data.frame with the values of interest

```{r }             
sensitivity <- function(model, parameters, nuisance, seed){

  if(model == "Myers")
    f <- Myers
  else if(model == "Allen")
    f <- RickerAllee

sigma_g <- nuisance[["sigma_g"]]
p <- parameters
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1
x_grid <- seq(0, 15, length=50)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 50  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
Xo <- 5.5# observations start from
x0 <- 8 # simulation under policy starts from
Tobs <- 40
MaxT <- 1000 # timeout for value iteration convergence

  # replicate over random seed
  yields <- sapply(seed, 
                   function(seed_i){
                     set.seed(seed_i)

                     x <- numeric(Tobs)
                     x[1] <- Xo
                     nz <- 1
                     for(t in 1:(Tobs-1))
                       x[t+1] = z_g() * f(x[t], h=0, p=p)

                     X = c(rep(0,nz), pmax(rep(0,Tobs-1), x[1:(Tobs-1)]))
                     Y = c(rep(0,nz), x[2:Tobs])

                     ## @knitr gp-priors
                     s2.p <- c(5,5)  
                     d.p = c(10, 1/0.1)

                     ## @knitr gp
                     gp <- gp_mcmc(X, y=Y, n=1e5, s2.p = s2.p, d.p = d.p)
                     gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)

                     matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid) 
                     opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)

                     ## @knitr mle-opt
                     matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
                     opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)

                     OPT = data.frame(GP = opt_gp$D, True = opt_true$D)
                     policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
                     names(policies) <- c("stock", "method", "value")

                     sims <- lapply(OPT, function(D){
                       set.seed(1)
                       lapply(1:100, function(i) 
                         ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
                              )
                       })

                     dat <- melt(sims, id=names(sims[[1]][[1]]))
                     sims_data <- data.table(dat)
                     setnames(sims_data, c("L1", "L2"), c("method", "reps")) 
                     # Legend in original ordering please, not alphabetical: 
                     sims_data$method = factor(sims_data$method, ordered=TRUE, levels=names(OPT))

                     Profit <- sims_data[, sum(profit), by=c("reps", "method")]
                     tmp <- dcast(Profit, reps ~ method)
                     tmp <- tmp / tmp[,"True"]
                     tmp <- melt(tmp[2:dim(tmp)[2]])
                     tmp$value[tmp$variable == "GP"]  
                     })

  yields_dat <- melt(yields)
  names(yields_dat) <- c("replicate", "simulation", "value")

  # Make ids factors, not integers
  yields_dat$replicate <- as.factor(yields_dat$replicate)
  yields_dat$simulation <- as.factor(yields_dat$simulation)

  # definitions of id codes.  In this case the id number is it's own definition.  
  rep_ids <-  levels(yields_dat$replicate)
  names(rep_ids) <- rep_ids
  sim_ids <-  levels(yields_dat$simulation)
  names(sim_ids) <- sim_ids

  dat <- data.frame(model = model, 
                    pars = as.list(parameters), 
                    replicate = yields_dat$replicate, 
                    sim = yields_dat$simulation, 
                    value = yields_dat$value, 
                    noise = sigma_g)
}
```

```{r}
model <- "Allen"
allen1.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
model <- "Allen"
allen2.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
model <- "Allen"
allen1.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
model <- "Allen"
allen2.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
model <- "Myers"
Myers1.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
model <- "Myers"
Myers2.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
model <- "Myers"
Myers1.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
model <- "Myers"
Myers2.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
save(list=ls(), file="sensitivity.rda")
```

```{r data}
allen_dat <- rbind(allen1.01, allen1.05, 
             allen2.01, allen2.05) 
m <- rbind(Myers1.01, Myers1.05, 
             Myers2.01, Myers2.05)
myers_dat <- m[c(1:2,4,3,5:8)]
names(myers_dat) <- names(allen_dat)
model_dat <- rbind(allen_dat, myers_dat)
```

```{r}
dat <- model_dat
dat$pars.r <- factor(dat$pars.r, labels=c("A", "B", "C", "D"))
dat <- dat[c(1:2,5:6, 8, 7)]
dat$noise <- factor(dat$noise)
names(dat) <- c("model", "parameters", "replicate", "simulation", "noise", "value")
```

```{r}
ggplot(dat) + geom_histogram(aes(value)) + xlim(0,1.0) + theme_bw() + xlab("value as fraction of the optimal")

ggplot(dat) + geom_histogram(aes(value, fill=noise)) + xlim(0,1.0) + theme_bw() + xlab("value as fraction of the optimal") 
```

```{r sensitivity}
ggplot(dat) + geom_histogram(aes(value, fill=noise)) + xlim(0,1.0) + theme_bw() + xlab("value as fraction of the optimal") + facet_wrap(~model, ncol=1)
```

```{r}
## Extract acutal parameter values corresponding to each parameter set
  p1 = levels(factor(model_dat$pars.r))
  p2 = levels(factor(model_dat$pars.K))
  p3 = levels(factor(model_dat$pars.C))

  A = c(r = p1[1], K = p2[1], theta = p3[1])
  B = c(r = p1[2], K = p2[2], theta = p3[2])
  C = c(r = p1[3], K = p2[3], C = p3[3])
  D = c(r = p1[4], K = p2[4], C = p3[4])
```

```{r eml, dependson='data'}
  rep_ids <-  levels(dat$replicate)
  sim_ids <-  levels(dat$sim)

  require(reml)
  col.defs = c(model = "name of the model used to simulate the data",
               parameters = "which set of model parameters was used",
               replicate = "replicate id numbers.  Each unique simulation of the model results in a policy function which is then replicated under each replicate id number to determine the realized value relative to the optimal value",
               simulation = "simulation id number. For each set of parameters for the model, multiple simulations of training data are generated and assigned unique simulation numbers",            
               noise = "sigma_g value, multiplicative growth noise scale, as a log-normal log-standard deviation"
               value = "the value derived from this realization of the policy function estimated with the Gaussian process relative to the theoretical optimum value if the true model and parameter values were known",)

  unit.defs = list(model = c("Myers" = "a 3 parameter model of Beverton-Holt-like recruitment with the addition of an Allee effect (and fourth parameter for scaling the log-normal growth noise); based on a Myers et al 1995 (doi:10.1126/science.269.5227.1106)", 
                               "Allen" = "a 3 parameter model of Ricker-like recruitment with the addition of an Allee effect (and fourth parameter for scaling the log-normal growth noise); based on Allen et al 2005 (doi:10.1080/10236190412331335373)"),
                  parameters = 
  c("A" = paste("First set of parameter values drawn at random for Myers model. r, K, theta respectively are:", A), 
    "B" = paste("Second set of parameter values drawn at random for Myers model. r, K, theta respectively are:", B), 
    "C" = paste("First set of parameter values drawn at random for Myers model. r, K, C respectively are:", C), 
    "D" = paste("Second set of parameter values drawn at random for Myers model. r, K, C respectively are:", D)), 
                  replicate = rep_ids,
                  simulation = sim_ids, 
                  noise = "density/density (pure)",
                  value = "Realized Dollars/Potential Dollars")

emldat =  data.set(dat, col.defs=col.defs, unit.defs=unit.defs)  
eml_write(emldat, creator="Carl Boettiger <cboettig@gmail.com>")
```

```{r, eval=FALSE}
require(reshape2)
require(snowfall)
sfInit(parallel=TRUE, cpu=8)
dat <- sfLapply(nuisance_values[["sigma_g"]], function(sigma){
  dat <- lapply(parameters[["Allen"]], function(p){
         sensitivity("Allen", 
                   parameters = p, 
                   nuisance = c(sigma_g = sigma), 
                   seed=c(1234, 2222, 3333))
              })
  dat <- melt(dat, id=names(dat[[1]]))
})
dat <- melt(dat, id=names(dat[[1]]))
```
```{r eval=FALSE}
require(reshape2)
myers_dat <- sfLapply(nuisance_values[["sigma_g"]], function(sigma){
  dat <- lapply(parameters[["Myers"]], function(p){
         sensitivity("Myers", 
                   parameters = p, 
                   nuisance = c(sigma_g = sigma), 
                   seed=c(1234, 2222, 3333))
              })
  dat <- melt(dat, id=names(dat[[1]]))
})
myers_dat <- melt(myers_dat, id=names(dat[[1]]))

```

### 2014-05-08 sequential-method.Rmd
add md files from /inst/examples on master branch

`ro comment=NA, message=FALSE, cache.path="compare/", tidy=FALSE or`

## Sequential method

```{r}
require(MASS)
require(ggplot2)
require(kernlab)
opts_knit$set(upload.fun = socialR::notebook.url)
```

Parameterization-specific

```{r}
X <- seq(-5,5, length=40)
obs <- data.frame(x = c(-3, -1, 3),
                  y = c(0,  1, -1))
#obs <- data.frame(x = -5:5, y = sin(x) + rnorm(length(x),sd=.1))
l <- 1
sigma_n <- 0.5
```

Radial basis function/Gaussian kernel:

```{r}
  SE <- function(Xi,Xj, l=1) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
  cov <- function(X, Y) outer(X, Y, SE, l) 
```

### Direct method

``` {r}
  n <- length(obs$x)
  K <- cov(obs$x, obs$x)
  I <-  diag(1, n)

  cov_xx_inv <- solve(K + sigma_n^2*I)
  Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
  Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X)
```

### Direct sequential method, avoids matrix inverse instability

``` {r}

mmult <- function(x,y){
  if(length(x) == 1){
    x <- as.numeric(x) 
    x * y
  } else if(length(y) == 1){ 
    y <- as.numeric(y)
    x * y
  }  else 
  x %*% y
}

mu <- numeric(length(obs$y))
y <- obs$y
x <- obs$x

C_seq <- function(X, X_prime, i){
  if(i <= 1)
    cov(X, X_prime) - mmult(cov(X,x[i]), cov(x[i], X_prime)) / as.numeric( cov(x[i], x[i]) + sigma_n^2)
  else
    C_seq(X, X_prime,   i-1) - mmult(C_seq(X,x[i],   i-1), C_seq(x[i], X_prime,   i-1)) / as.numeric( C_seq(x[i], x[i],   i-1)  + sigma_n^2  )
}
mu_seq <- function(X, i){
  if(i <= 1)
    cov(x[i], X) * (y[i]-mu[i]) / as.numeric( cov(x[i], x[i]) + sigma_n^2)
  else
    mu_seq(X, i-1) + C_seq(x[i], X,  i-1) * (y[i]-mu[i]) / as.numeric( C_seq(x[i], x[i], i-1)  + sigma_n^2 )
}
ef <- t(mu_seq(X, length(obs$x)))
cf <- C_seq(X, X, length(obs$x))

```

Compare these results:

``` {r}
require(reshape2)
require(ggplot2)
df <- data.frame(x = X, direct = Ef, sequential = ef)
df <- melt(df, id = "x")
ggplot(df)+ geom_jitter(aes(x, value, color = variable)) + geom_point(data = obs, aes(x,y))
````

### 2014-05-08 tgp-priors.Rmd
add md files from /inst/examples on master branch

--- 2012-12-10 inst/examples/tgp-priors.Rmd
@@
 ```{r include=FALSE}
-opts_knit$set(upload.fun = socialR::notebook.url)
+opts_knit$set(upload.fun = socialR::flickr.url)
 opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE, message=FALSE)
-theme_update(plot.background = element_rect(fill = "transparent",colour = NA))
 ```

### 2014-05-08 gaussian-process-control.R

@@ -1,0 +1,359 @@
+
+## @knitr libraries
+require(pdgControl)
+require(nonparametricbayes)
+require(reshape2)
+require(ggplot2)
+require(data.table)
+require(tgp)
+library(kernlab)
+require(MCMCpack)
+require(plyr)
+
+
+
+## @knitr graphing-options
+opts_knit$set(upload.fun = function(file) socialR::notebook.url(file, cp=TRUE, sync=FALSE))
+#opts_chunk$set(dev = 'Cairo_pdf', dev.args=list(""))
+opts_chunk$set(dev="png", dev.args=list(bg="transparent"))
+opts_chunk$set(comment=NA, tidy=FALSE)
+theme_set(theme_bw(base_size=16))
+theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
+             plot.background = element_rect(fill = "transparent",colour = NA))
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+#sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
+#short_sha <- gsub("(^.{10}).*", "\\1", sha)
+#date <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S")
+#opts_chunk$set(fig.path = paste("figure/", date, "-", short_sha, "-", sep=""))
+
+
+
+
+
+
+## @knitr BevHolt
+f <- BevHolt
+p <- c(1.5,.05)
+K <- (p[1]-1)/p[2] 
+
+
+
+## @knitr Myer
+f <- Myer_harvest
+pars <- c(1, 2, 4.5) 
+p <- pars # shorthand 
+K <- p[1] * p[3] / 2 + sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2
+allee <- p[1] * p[3] / 2 - sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 # allee threshold
+e_star <- (p[1] * sqrt(p[3]) - 2) / 2 ## Bifurcation point 
+
+
+## @knitr Ricker
+f <- Ricker
+p <- c(2.5, 10) 
+K <- 12.5
+
+
+## @knitr May
+f <- May
+p <- c(r = .75, k = 10, a=1.3, H=1, Q = 3)
+K <- 8 
+
+## @knitr plot-may
[…]

### 2014-05-08 scenarios-table.R

@@ -1,0 +1,22 @@
+scenarios <- melt(parameters)
+
+ids <- c(rep(1:3, 3), rep(4:6, 3))
+
+scenarios <- cbind("scenario"= ids, scenarios)
+names(scenarios) <- c("scenario", "value", "parameter", "model")
+scenarios <- scenarios[c("scenario", "model", "parameter", "value")]
+
+Allen <- scenarios$model == "Allen"
+Myers <- scenarios$model == "Myers"
+
+allen_ids <- scenarios[Allen, "parameter"]
+myers_ids <- scenarios[Myers, "parameter"]
+
+allen_names <- names(parameters$Allen[[1]])
+myers_names <- names(parameters$Myers[[1]])
+
+scenarios[Allen, "parameter"] <- allen_names[allen_ids]
+scenarios[Myers, "parameter"] <- myers_names[myers_ids]
+
+
+scenarios

### 2014-05-08 BUGS/parametric-vs-nonparametric.Rmd
examples need yaml

--- 2013-06-07 inst/examples/BUGS/parametric-vs-nonparametric.Rmd
@@
-# Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach
+---
+layout: page
+---# Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach
 
@@
 
-```{r mle, dependson="obs"}
+```{r mle-local, dependson="obs"}
+set.seed(12345)
+estf <- function(p){ 
+    mu <- f(obs$x,0,p)
+    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
+}
+par <- c(p[1]*rlnorm(1,0,.2), 
+         p[2]*rlnorm(1,0,.1), 
+         p[3]*rlnorm(1,0, .1), 
+         sigma_g * rlnorm(1,0,.3))
+o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
+f_alt <- f
+p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
+sigma_g_alt <- as.numeric(o$par[4])
+
+est <- list(f = f_alt, p = p_alt, sigma_g = sigma_g_alt, mloglik=o$value)
 ```
@@
 
-```{r mle-output, dependson="mle"}
+```{r mle-output, dependson="mle-local"}
 ```
@@
 
-```{r myers-mcmc, dependson="myers-model"}
+```{r myers-mcmc-local, dependson="myers-model"}
+jags.params=c("r0", "theta", "K", "stdQ")
+jags.inits <- function(){
+  list("r0"= rlnorm(1,0,.1), 
+       "K"=    6 * rlnorm(1,0,.1),
+       "theta" = 2 * rlnorm(1,0,.1),  
+       "stdQ"= 0.1 * rlnorm(1,0,.1),
+       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+}
+set.seed(12345)
+myers_jags <- do.call(jags.parallel, 
+                      list(data=jags.data, inits=jags.inits, 
+													 jags.params, n.chains=n.chains, 
+													 n.iter=n.iter, n.thin=n.thin,
+                           n.burnin=n.burnin, 
+                           model.file="myers_process.bugs"))
+recompile(myers_jags)
+myers_jags <- do.call(autojags, 
+                      list(myers_jags, n.update=n.update, 
+                           n.iter=n.iter, n.thin = n.thin, 
+                           progress.bar="none"))
 ```

### 2014-05-08 allee.Rmd
examples need yaml

@@
-```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
+---
+layout: page
+---```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
 opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "allee/", warning=FALSE)

### 2014-05-08 mcmc-gp.Rmd
examples need yaml

--- 2012-12-10 inst/examples/mcmc-gp.Rmd
@@
-```{r include=FALSE}
+---
+layout: page
+---```{r include=FALSE}
 opts_knit$set(upload.fun = socialR::notebook.url)
@@
 
-$$K(X, X') =  e^{\frac{(X-X')^p_0}{d}}$$
+$$K(X, X') =  e^{\frac{-(X-X')^p_0}{d}}$$
 
-(from equation 4 in the Vignette (1)).
-I do not see a prior for `p_0` or any mention of it.
-Nor can I tell what is the functional form for it's prior (`d.p`) (which is not an inverse gamma?)?
-Not clear why it's prior gets defined with four parameters, `c(a1, g1, a2, g2)`, instead of two.
-Form the manual:
+(from equation 4 in the Vignette (1)).  The parameter `p_0` is fixed at `2`, and not treated as a hyperparameter.
 
-> Mixture of gamma prior parameter (initial values) for the range parameter(s) c(a1,g1,a2,g2) where g1 and g2 are scale (1/rate) parameters.
-
+The length-scale comes from a hyperparameter that is the sum (mixture) of two Gamma distributions, with prior parameter `c(a1,g1,a2,g2)` where `g1` and `g2` are scale (1/rate) parameters.
 As before, these parameters  `c(a1, g1, a2, g2)` can be drawn from an exponential distribution hyperprior whose $\lambda$ values are given by `d.lam`, or turned off.
 
-* $\varepsilon$ is `nug`.  It seems to have same shape prior as $d$, specified by `c(a1,g1,a2,g2)`, and optional hyperprior as well.  Additionally it can be fixed to a constant:
+* $\varepsilon$ is `nug`.
+It seems to have same shape prior as $d$, specified by `c(a1,g1,a2,g2)`, and optional hyperprior as well.
+Additionally it can be fixed to a constant by setting `nug.p = 0` and specifying the staring value as `gd[1]`.
 
-> specifying nug.p = 0 fixes the nugget parameter to the starting value in gd[1], i.e., it is excluded from the MCMC
+* The mean is not fixed at zero but is instead a linear model, we need hyperparameters for it as well.
+The key parameter there is the slope $\beta$.
+The functional form for it's prior is specified as `bprior`, which takes values `bflat`, `b0`, `bmzt`, or `bmznot`.
+A sensible choice is as a Gaussian `b0`, as shown in Equation 1 of the vignette
 
-But what is `gd[1]` and where is it defined?
-
-If the mean is not fixed at zero but is instead a linear model, we need hyperparameters for it as well.
-The key parameter there is the slope $\beta$.
-The functional form for it's prior is specified as `bprior`, which takes values `bflat`, `b0`, `bmzt`, or `bmznot`, but I can find no precise statement of the functional forms of these priors (e.g.
-I see that `b0` is Gaussian prior and `bflat` an improper or flat prior, but beyond that), nor do I see where their shape parameters are set.
-
-The manual suggests that we need to specify start values for the hyperparameters for the MCMC (presumably we could have just drawn them from the prior? we burn in anyhow...)  The are given in
-
-> `start =  c(0.5,0.1,1.0,1.0)` starting values for range `d`, nugget `g`, `s2`, and `tau2`
-
-Actually specifiying start values throws an error though, it appears the documentation is out of date and the start parameter has been removed.
-
-* I'm still not clear on the role of `tau2`, whether it's unique to the "treed" partioning or whether I've misunderstood $sigma^2$.  Grammacy phrases the model as:
-
-$$Z | \beta, \sigma^2, K \sim N_n(\mathbf{F} \beta, \sigma^2 \mathbf{K}) $$
-$$ \beta | \sigma^2, \tau^2, \mathbf{W}, \beta_0 ~ N_m(\beta_0, \sigma^2 \tau^2 \mathbf{W} )  $$
-$$\beta_0 | N_m(\mathbf{\mu}, \mathbf{B})$$
-
-(I guess I haven't parsed the second two equations)
+$$
+\begin{align} 
+Z | \beta, \sigma^2, K &\sim N_n(\mathbf{F} \beta, \sigma^2 \mathbf{K}) \\
+\beta | \sigma^2, \tau^2, \mathbf{W}, \beta_0 &\sim N_m(\beta_0, \sigma^2 \tau^2 \mathbf{W} )  \\
+\beta_0 &\sim N_m(\mathbf{\mu}, \mathbf{B})
+\end{align}
+$$
 
@@
 
-So, how now would I make a call to `tgp::bgp` that would give me the equivalent output of, say, a direct calculation of the above model (given fixed hyperparameters)?
 Say, given some observed data pairs $x,y$ and predictions desired on a grid $X$,
@@
 
+Note: it appears that the mixed gammas for `d` and `nug` actually treat the second argument, `g1` and `g2`, as the rate instead of the scale, so I invert the desired scale:
+
 ```{r}
@@
 
-Define as random samples (avoids having to specify range)
-
-```{r}
-s2_rprior <- function(x) rinvgamma(x, s2.p[1], s2.p[2])
-tau2_rprior <- function(x) rinvgamma(x, tau2.p[1], tau2.p[2])
-d_rprior <- function(x) rgamma(x, d.p[1], scale = d.p[2]) + rgamma(x, d.p[3], scale = d.p[4])
-nug_rprior <- function(x) rgamma(x, nug.p[1], scale = nug.p[2]) + rgamma(x, nug.p[3], scale = nug.p[4])
-beta0_rprior <- function(x, tau) rnorm(x, 0, tau)
-n = 1000
-rpriors <- data.frame(s2 = s2_rprior(n), tau2 = tau2_rprior(n), beta0 = beta0_rprior(n, 1), nug = nug_rprior(n), d = d_rprior(n))
-rpriors <- melt(rpriors)
-ggplot(rpriors) + geom_histogram(aes(value)) + facet_wrap(~variable, scale="free")
-````
-
-Define as curves
+Define as curves and plot the priors:
 
@@
 gp <- bgp(X=x, XX=X, Z=y, verb=0,
-          meanfn="constant", bprior="b0", BTE=c(1,2000,1), m0r1=FALSE, 
+          meanfn="constant", bprior="b0", BTE=c(2000,6000,2), m0r1=FALSE, 
           corr="exp", trace=TRUE, beta = 0,
@@
 ```{r}
-manual_dat <- data.frame(x=X, y=Ef, ymin=(Ef - 1.96 * sqrt(diag(Cf))), ymax=(Ef + 2 * sqrt(diag(Cf))))
+manual_dat <- data.frame(x = X, 
+                         y = Ef, 
+                         ymin = (Ef - 1.96 * sqrt(diag(Cf))), 
+                         ymax = (Ef + 1.96 * sqrt(diag(Cf))))
 ```
@@
 
-Look at the actual parameter estimates from the trace
-
-```{r}
-gp$trace$XX[[1]][1999,]
-```
-
-Posteriors
+## Posteriors of hyperparameters
 
@@
 ```
-
-## Compare to priors with `scale` misplaced as `rate` in `d`, `nug`
-
-Define as random samples (avoids having to specify range)
-
-```{r}
-s2_rprior <- function(x) rinvgamma(x, s2.p[1], s2.p[2])
-tau2_rprior <- function(x) rinvgamma(x, tau2.p[1], tau2.p[2])
-d_rprior <- function(x) rgamma(x, d.p[1], d.p[2]) + rgamma(x, d.p[3], d.p[4])
-nug_rprior <- function(x) rgamma(x, nug.p[1], nug.p[2]) + rgamma(x, nug.p[3], nug.p[4])
-beta0_rprior <- function(x, tau) rnorm(x, 0, tau)
-n = 1000
-rpriors <- data.frame(s2 = s2_rprior(n), tau2 = tau2_rprior(n), beta0 = beta0_rprior(n, 1), nug = nug_rprior(n), d = d_rprior(n))
-rpriors <- melt(rpriors)
-ggplot(rpriors) + geom_histogram(aes(value)) + facet_wrap(~variable, scale="free")
-````
-
-```{r}
-s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
-tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
-d_prior <- function(x) dgamma(x, d.p[1], d.p[2]) + dgamma(x, d.p[3], d.p[4])
-nug_prior <- function(x) dgamma(x, nug.p[1], nug.p[2]) + dgamma(x, nug.p[3], nug.p[4])
-beta0_prior <- function(x, tau) dnorm(x, 0, tau)
-
-xx <- seq(.0001, 2500, length.out=1000)
-priors <- data.frame(x = xx, s2 = s2_prior(xx), tau2 = tau2_prior(xx), beta0 = beta0_prior(xx, 1), nug = nug_prior(xx), d = d_prior(xx))
-priors <- melt(priors, id="x")
-ggplot(priors) + geom_line(aes(x, value)) + facet_wrap(~variable, scale="free")
-```

### 2014-05-08 myers-exploration.Rmd
examples need yaml

--- 2012-12-24 inst/examples/myers-exploration.Rmd
@@
-```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
+---
+layout: page
+---```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
 opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "myers-cache/", warning=FALSE)

### 2014-05-08 ricker-oscillate-example.Rmd
examples need yaml

--- 2012-12-19 inst/examples/ricker-oscillate-example.Rmd
@@
-```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
+---
+layout: page
+---```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
 opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "myers-cache/")
@@
 xT <- 0
-set.seed(1111)
+set.seed(1)
 ```
@@
 
-```{r par-est}
+```{r par-est-BH}
 ```
+
+```{r gp-priors}
+```
+
+Estimate data under the BH model
 

### 2014-05-08 sensitivity.Rmd
examples need yaml

@@
-```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+---
+layout: page
+---```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
 rm(list=ls())

### 2014-05-08 BUGS/long-run-mcmc.Rmd
preserve version history of scripts

--- 2013-06-05 inst/examples/BUGS/process-noise-only.Rmd
@@
 n.chains <- 6
-n.iter <- 1e6
-n.burnin <- floor(10000)
+n.iter <- 1e8
+n.burnin <- floor(1e7)
 n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
@@
     mu[t] <- y[t] * exp(r0 * (1 - y[t]/K)* (y[t] - theta) / K )
-    y[t+1] ~ dlnorm(mu[t], iQ) 
+    y[t+1] ~ dnorm(mu[t], iQ) 
   }
@@
     mu[t] <- y[t] * exp(r0 * (1 - y[t]/K) )
-    y[t+1] ~ dlnorm(mu[t], iQ) 
+    y[t+1] ~ dnorm(mu[t], iQ) 
   }
@@
     mu[t] <- r0 * pow(abs(y[t]), theta) / (1 + pow(abs(y[t]), theta) / K)
-    y[t+1] ~ dlnorm(mu[t], iQ) 
+    y[t+1] ~ dnorm(mu[t], iQ) 
   }

### 2014-05-08 comparing-methods.Rmd
preserve version history of scripts

--- 2012-12-05 inst/examples/comparing-methods.Rmd
@@
-`ro comment=NA, message=FALSE, cache.path="compare/", tidy=FALSE or`
-
-```{r}
+```{r include=FALSE}
 require(MASS)
@@
 opts_knit$set(upload.fun = socialR::notebook.url)
-opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE)
-theme_set(theme_bw())
-theme_update(panel.background = element_rect(fill = "transparent",colour = NA))
+opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE, message=FALSE)
+theme_update(plot.background = element_rect(fill = "transparent",colour = NA))
 ```
-
-Parameterization-specific
 
@@
 ``` {r}
-k <- gp_fit(obs, X, c(sigma_n=sigma_n, l=l, tau=1), "sequential")
+k <- gp_fit(obs, X, c(sigma_n=sigma_n, l=l, tau=1), "kernlab")
 y_p <- k$Ef

### 2014-05-08 gaussian-process-shiny.Rmd
preserve version history of scripts

--- 2013-05-14 inst/examples/gaussian-process-basics.Rmd
@@
 
-Basic regression in Gaussian processes  
----------------------------------------
-
-Working out coded examples for basic Gaussian process regression using R.
-I've just read through the first few chapters of [Rasmussen & Williams (2006)](http://www.GaussianProcess.org/gpml), this implements the examples discussed in Chapter 2.1-2.5.
-
-<!-- Also motivated by [James Keirstead](http://www.jameskeirstead.ca/r/gaussian-process-regression-with-r/), with cleaner code.  -->
-
-Required R libraries (for multivariate normal, also for plotting):
-
-```{r results="hide"}
-require(MASS)
-require(reshape2)
-```
-
-Set a seed for repeatable plots
-
 ```{r}
@@
 
-Define the points at which we want to compute the function values (x values of the prediction points or test points), and the scale parameter for the covariance function $\ell=1$
+```{r}
 
-```{r}
-x_predict <- seq(-5,5,len=50)
-l <- 1
-```
+#' @import MASS
+#' @import reshape2
+samples <- function(obs, l=1, sigma.n=0.3, x_predict = seq(-5,5,len=50), n=3){
 
-We will use the squared exponential (also called radial basis or Gaussian, though it is not this that gives Gaussian process it's name; any covariance function would do) as the covariance function, $\operatorname{cov}(X_i, X_j) = \exp\left(-\frac{(X_i - X_j)^2}{2 \ell ^ 2}\right)$
-
-```{r}
+  require(MASS); require(reshape2)
   SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
   cov <- function(X, Y) outer(X, Y, SE, l)
+  cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma.n^2 * diag(1, length(obs$x)))
+  Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% obs[[2]]
+  Cf <- cov(x_predict, x_predict) - 
+        cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)
+
+  values <- mvrnorm(n, Ef, Cf)
+  dat <- data.frame(x=x_predict, t(values))
+  dat <- melt(dat, id="x")
+  Ey <- data.frame(x=x_predict, 
+                   y=(Ef), 
+                   ymin=(Ef-2*sqrt(diag(Cf))), 
+                   ymax=(Ef+2*sqrt(diag(Cf))))
+  list(dat=dat, Ey=Ey, obs=obs, x_predict=x_predict, l=l, sigma.n=sigma.n, n=n)
+
+  merge(obs, merge(dat, Ey, "x", all=TRUE), "x", all=TRUE)
+}
+
+x_predict <- seq(-5,5,len=50)
+obs <- data.frame(x = c(x_predict[c(5,10,15,20,25)]),
+                  observed = x_predict[c(10,  15,  20,  25, 10)])
+
+S <- samples(obs, l=1, sigma.n=.5, x_predict, n=3)
+
+ggplot(S) + 
+  geom_ribbon(aes(x,y,ymin=ymin, ymax=ymax), fill="grey80") + 
+  geom_point(aes(x,observed)) + 
+  geom_line(aes(x,value, color=variable))
 ```
@@
 ```{r}
-COV <- cov(x_predict, x_predict)
-```
-Generate a number of functions from the process
-
-```{r}
-values <- mvrnorm(3, rep(0, length=length(x_predict)), COV)
-```
-
-Reshape the data into long (tidy) form, listing x value, y value, and sample number
-```{r}
-dat <- data.frame(x=x_predict, t(values))
-dat <- melt(dat, id="x")
-head(dat)
-```
-
-Plot the result
-
-```{r fig2a}
-fig2a <- ggplot(dat,aes(x=x,y=value)) +
-  geom_rect(xmin=-Inf, xmax=Inf, ymin=-2, ymax=2, fill="grey80") +
-  geom_line(aes(group=variable)) +  
-  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
+ggplot(S$dat,aes(x=x,y=value)) +
+  geom_ribbon(data=S$Ey, aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
+  geom_line(aes(color=variable)) + #REPLICATES
+  geom_line(data=S$Ey,aes(x=x,y=y), size=1) + #MEAN
+  geom_point(data=S$obs,aes(x=x,y=y)) +  #OBSERVED DATA
+  scale_y_continuous(lim=c(-4,4), name="output, f(x)") +
   xlab("input, x") 
-fig2a
-```
-
-### Posterior distribution given the data
-
-Imagine we have some data,
-
-```{r}
-obs <- data.frame(x = c(-4, -3, -1,  0,  2),
-                  y = c(-2,  0,  1,  2, -1))
-```
-
-In general we aren't interested in drawing from the prior, but want to include information from the data as well.  We want the joint distribution of the observed values and the prior is:
-
-$$\begin{pmatrix} y_{\textrm{obs}} \\ y_{\textrm{pred}} \end{pmatrix} \sim \mathcal{N}\left( \mathbf{0}, \begin{bmatrix} cov(X_o,X_o) & cov(X_o, X_p) \\ cov(X_p,X_o) & cov(X_p, X_p) \end{bmatrix} \right)$$
-
-## No observation noise
-
-Assuming the data are known without error and conditioning on the data, and given $x \sim \mathcal{N}(0, cov(X_o, X_o))$, then the conditional probability of observing our data is easily solved by exploiting the nice properties of Gaussians,
-
-$$x|y \sim \mathcal{N}(E,C)$$
-$$E = cov(X_p, X_o) cov(X_o,X_o)^{-1} y$$
-$$C= cov(X_p, X_p) - cov(X_p, X_o) cov(X_o,X_o)^{-1} cov(X_o, X_p)$$
-
-(We use `solve(M)` which with no second argument will simply invert the matrix `M`, but should use the Cholsky decomposition instead for better numerical stability)
-
-```{r}
-cov_xx_inv <- solve(cov(obs$x, obs$x))
-Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% obs$y
-Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)
-```
-
-Now lets take 3 random samples from the posterior distribution,
-```{r}
-values <- mvrnorm(3, Ef, Cf)
-```
-
-and plot our solution (mean, 2 standard deviations, and the ranom samples.)
-
-```{r fig2b}
-dat <- data.frame(x=x_predict, t(values))
-dat <- melt(dat, id="x")
-
-fig2b <- ggplot(dat,aes(x=x,y=value)) +
-  geom_ribbon(data=NULL, 
-              aes(x=x_predict, y=Ef, ymin=(Ef-2*sqrt(diag(Cf))), ymax=(Ef+2*sqrt(diag(Cf)))),
-              fill="grey80") +
-  geom_line(aes(color=variable)) + #REPLICATES
-  geom_line(data=NULL,aes(x=x_predict,y=Ef), size=1) + #MEAN
-  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
-  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
-  xlab("input, x")
-fig2b
-```
-
-```{r fig2_data}
-ggplot(data=obs,aes(x=x,y=y)) +
-  geom_point()   #OBSERVED DATA
-```
-
-Additive noise
---------------
-
-In general the model may have process error, and rather than observe the deterministic map $f(x)$ we only observe $y = f(x) + \varepsilon$.
-Let us assume for the moment that $\varepsilon$ are independent, normally distributed random variables with variance $\sigma_n^2$.
-
-```{r}
-sigma.n <- 0.3
-cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma.n^2 * diag(1, length(obs$x)))
-Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% obs$y
-Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)
-```
-
-Now lets take 3 random samples from the posterior distribution,
-```{r}
-values <- mvrnorm(3, Ef, Cf)
-```
-
-and plot
-
-```{r fig2c}
-dat <- data.frame(x=x_predict, t(values))
-dat <- melt(dat, id="x")
-Ey <- data.frame(x=x_predict, y=(Ef), ymin=(Ef-2*sqrt(diag(Cf))), ymax=(Ef+2*sqrt(diag(Cf))))
-
-fig2c <- ggplot(dat,aes(x=x,y=value)) +
-  geom_ribbon(data=Ey, aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
-  geom_line(aes(color=variable)) + #REPLICATES
-  geom_line(data=Ey,aes(x=x,y=y), size=1) + #MEAN
-  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
-  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
-  xlab("input, x") 
-fig2c
 ```

### 2014-05-08 gp_mcmc_example.Rmd
preserve version history of scripts

```{r plotting-options, message=FALSE, warning=FALSE, include=FALSE}
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA)
library(ggplot2) # plotting 
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=10))
theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
             plot.background = element_rect(fill = "transparent", colour = NA))

```

Basic regression in Gaussian processes  
---------------------------------------

Working out coded examples for basic Gaussian process regression using R.  I've just read through the first few chapters of [Rasmussen & Williams (2006)](http://www.GaussianProcess.org/gpml), this implements the examples discussed in Chapter 2.1-2.5.

<!-- Also motivated by [James Keirstead](http://www.jameskeirstead.ca/r/gaussian-process-regression-with-r/), with cleaner code.  -->

## Function definitions

```{r}
library(mcmc)
#' Basic regression in Gaussian processes  
#' 
#' 
#' @param x Observed x values, (vector or matrix with columns for each dimension of data)
#' @param y Vector of observed y values in the training data
#' @param init_pars the initial guesses for lengthscale l and process noise sigma_n
#' @param n iterations of the metropolis algorithm
#' @details Currently assumes the covariance function.  By default we will use 
#' the squared exponential (also called radial basis or Gaussian, 
#' though it is not this that gives Gaussian process it's name; 
#' any covariance function would do) as the covariance function, 
#' $\operatorname{cov}(X_i, X_j) = \exp\left(-\frac{(X_i - X_j)^2}{2 \ell ^ 2}\right)$
#' @return the MCMC output, as constructed by \link{metrop} from the mcmc package.
#' @export
#' @import mcmc  
gp_mcmc <- function(x, y, init_pars = c(l=1, sigma.n=1), n = 1e4, d.p = c(5,5), s2.p = c(5,5)){

  lpriors <- function(pars){
    prior <- unname(
      dgamma(exp(pars[1]), d.p[1], scale = d.p[2]) *
      dgamma(exp(pars[2]), s2.p[1], s2.p[2]) 
    )
    log(prior)
  }
  posterior <- function(pars, x, y){ 
    l <- exp(pars[1])
    sigma.n <- exp(pars[2])

    cov <- function(X, Y) outer(X, Y, SE, l)
    I <- diag(1, length(x))
    K <- cov(x, x)

    loglik <- - 0.5 * t(y) %*% solve(K + sigma.n^2 * I) %*% y -
      log(det(K + sigma.n^2*I)) -
      length(y) * log(2 * pi) / 2

    loglik + lpriors(pars)
  }

  out <- metrop(posterior, log(init_pars), n, x = obs$x, y = obs$y)  
  out$d.p <- d.p
  out$s2.p <- s2.p

  out
}
```

```{r}
#' predict the expected values and posterior distributions of the Gaussian Process
#' 
#' @param x_predict the values at which we desire predictions
#' @param covs if TRUE, return covariances instead of variances (set high thinning as this is memory intensive)
#' @import reshape2
#' @export
gp_predict <- function(gp, x_predict, burnin=0, thin=1, covs=FALSE){

  postdist <- cbind(index=1:gp$nbatch, as.data.frame(exp(gp$batch)))
  s <- seq(burnin+1, gp$nbatch, by=thin)
  postdist <- postdist[s,]
  names(postdist) <- c("index", names(gp$initial))

  indices <- 1:dim(postdist)[1]

  out <- lapply(indices, function(i){
    l <- postdist[i, "l"]
    sigma.n <- postdist[i, "sigma.n"]

    cov <- function(X, Y) outer(X, Y, SE, l)
    cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma.n^2 * diag(1, length(obs$x)))
    Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% obs$y
    Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)
    list(Ef = Ef, Cf = Cf, Vf = diag(Cf))
  })

  Ef_posterior <- sapply(out, `[[`, "Ef")
  Cf_posterior <- sapply(out, `[[`, "Cf")
  Vf_posterior <- sapply(out, `[[`, "Vf")

  E_Ef <- rowMeans(Ef_posterior)
  E_Cf <- matrix( apply(Cf_posterior, 1, sum) / dim(Cf_posterior)[2], ncol = sqrt(dim(Cf_posterior)[1]) )
  E_Vf <- diag(E_Cf) # same as rowMeans(Vf_posterior)

  # list format better for return
  Cf_posterior <- lapply(out, `[[`, "Cf")

  list(Ef_posterior = Ef_posterior, 
       Vf_posterior = Vf_posterior,
       Cf_posterior = Cf_posterior,
       E_Ef = E_Ef, E_Cf = E_Cf, E_Vf = E_Vf)

}  
```

```{r} 
library(reshape2)
library(ggplot2)
library(plyr)
#' Summary plots showing the trace and posteriors for the gp_mcmc estimates
#' 
#' @param gp a fit of the gaussian process from gp_mcmc
#' @param burnin length of sequence to discard as a transient
#' @param thin frequency of sub-sampling (make posterior distribution smaller if necessary)
#' @return two ggplot2 objects, one plotting the trace and one 
#' plotting the posteriors in black with priors overlaid in red.  
#' @import reshape2
#' @import ggplot2
#' @import plyr
#' @export
summary_gp_mcmc <- function(gp, burnin=0, thin=1){

  postdist <- cbind(index=1:gp$nbatch, as.data.frame(exp(gp$batch)))
  s <- seq(burnin+1, gp$nbatch, by=thin)
  postdist <- postdist[s,]
  names(postdist) <- c("index", names(gp$initial))

  # TRACES
  df <- melt(postdist, id="index")
  traces_plot <- 
    ggplot(df) + geom_line(aes(index, value)) + 
    facet_wrap(~ variable, scale="free", ncol=1)

  s2.p <- gp$s2.p
  d.p <- gp$d.p
  s2_prior <- function(x) dgamma(x, s2.p[1], s2.p[2])
  d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2])
  prior_fns <- list(l = d_prior, sigma.n = s2_prior)

  prior_curves <- ddply(df, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100)
    data.frame(value = grid, density = prior_fns[[dd$variable[1]]](grid))
  })

  # Posteriors (easier to read than histograms)
  posteriors_plot <- ggplot(df, aes(value)) + 
    stat_density(geom="path", position="identity", alpha=0.7) +
    geom_line(data=prior_curves, aes(x=value, y=density), col="red") + 
    facet_wrap(~ variable, scale="free", ncol=2)

  print(traces_plot)
  print(posteriors_plot)

  out <- list(traces_plot = traces_plot, posteriors_plot = posteriors_plot)
  invisible(out)
}

# Helper function: The default covariance function
SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
```

## Example

```{r}
x_predict <- seq(-5,5,len=50)
obs <- data.frame(x = c(-4, -3, -1,  0,  2),
                  y = c(-2,  0,  1,  2, -1))
```

```{r}
gp <- gp_mcmc(obs$x, obs$y, d.p = c(2,.5), s2.p = c(2,.5))
```

```{r}
summary_gp_mcmc(gp)
```

```{r}
gp_dat <- gp_predict(gp, x_predict)
```

```{r}
tgp_dat <- 
    data.frame(  x = x_predict, 
                 y = gp_dat$E_Ef, 
                 ymin = gp_dat$E_Ef - 2 * sqrt(gp_dat$E_Vf), 
                 ymax = gp_dat$E_Ef + 2 * sqrt(gp_dat$E_Vf) )

ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
    geom_line(aes(x, y), lwd=2, alpha=0.8) + 
    geom_point(data=obs, aes(x,y), alpha=0.8) + 
    xlab(expression(X[t])) + ylab(expression(X[t+1])) 
```

### 2014-05-08 mcmc-gp.Rmd
preserve version history of scripts

@@
----
-layout: page
----```{r include=FALSE}
+```{r include=FALSE}
 opts_knit$set(upload.fun = socialR::notebook.url)

### 2014-05-08 myers-exploration.Rmd
preserve version history of scripts

@@
----
-layout: page
----```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
+```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
 opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "myers-cache/", warning=FALSE)

### 2014-05-08 BUGS/allen.Rmd
preserve version history of scripts
import inst/examples with history

--- 2013-06-07 inst/examples/BUGS/myers.Rmd
@@
 ```{r stateeq-local}
-f <- Myers
-p <- c(1.5, 2.5, 8)
-K <- 5  # approx, a li'l' less
-allee <- 1.2 # approx, a li'l' less
+f <- RickerAllee
+p <- c(1, 10, 5)
+K <- 10  # approx, a li'l' less
+allee <- 5 # approx, a li'l' less
 ```

### 2014-05-08 BUGS/bayesian-ricker.Rmd
import inst/examples with history

--- 2013-05-07 inst/examples/BUGS/bayesian-ricker.Rmd
@@
-```{r}
+---
+layout: page
+---```{r}
 setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")

### 2014-05-08 BUGS/par_uncertainty_sdp.Rmd
import inst/examples with history

--- 2013-04-20 inst/examples/BUGS/2013-04-19-bayesian-parametric-uncertainty-sdp.Rmd
@@
 ---
-layout: post
-category: ecology
-tags:
-- nonparametric-bayes
-
----
-
-Adapting the parametric uncertainty represented by the posterior distributions of the Bayesian estimate (see [earlier notes](http://carlboettiger.info/2013/04/12/parametric-bayesian-example.html)) to the stochastic dynamic programming solution for the optimal policy.
-Simply requires evaluating the expectation over the distribution, but is computationally intensive given the spread and the three parameters.
-
-```{r}
-f_transition_matrix <- function(f, p, x_grid, h_grid, sigma_g, pardist){
-  lapply(h_grid, par_F, f, p, x_grid, sigma_g, pardist)
-}
-
-par_F <- function(h, f, p, x_grid, sigma_g, pardist, n_mc = 100){
-
-  # Set up monte carlo sampling 
-  d <- dim(pardist)
-  indices <- round(runif(n_mc,1, d[1]))
-
-  # compute mean with monte carlo samples
-  mu <- sapply(indices, function(i) { # Cols are parameter samples
-    p <- unname(pardist[i,c(4,1,5)])          # Only, we need to add the measurement and process noise  
-    unname(sapply(x_grid, f, h, p))   # rows are possible values of x_t
-  })
-  mu <- data.frame(t(mu)) # Let's have cols as possible x_t, rows as parameters, and type as data.frame
-
-  F_true <- 
-    sapply(mu, function(cur_state){    # For each x_t
-
-      bypar <- sapply(cur_state, function(m){ # For each parameter value
-        if(snap_to_grid(m,x_grid) < x_grid[2]){ # 
-          out <- numeric(length(x_grid))
-          out[1] <- 1
-          out
-        } else {
-          out <- dlnorm(x_grid/m, 0, sigma_g) # should be using the estimate process noise!
-        }})  
-      ave_over_pars <- apply(bypar, 1, sum) # collapse by weighted average over possible parameters    
-      ave_over_pars / sum(ave_over_pars)
-      })
-
-  F_true <- t(F_true)
-
-}
-
-snap_to_grid <- function(x, grid) sapply(x, function(x) grid[which.min(abs(grid - x))])
-
-# internal helper function
-rownorm <- function(M) 
-  t(apply(M, 1, function(x){ 
-    if(sum(x)>0){
-      x/sum(x)
-    } else {
-      out <- numeric(length(x))
-      out[1] <- 1
-      out
-    }
-  }))
-```
-
-Run the Bayesian analysis to obtain posterior distributions for parameters.
+layout: page
+---Run the Bayesian analysis to obtain posterior distributions for parameters.
 
@@
 ```
-
-## Test case using perturbed parameters
-
-First, sanity test. Use the correct parameter values (slightly perturbed).
-
-```{r}
-pardist <- mcmcall
-pardist[,1] = p[2] + rnorm(100, 0, 0.000001)
-pardist[,4] = p[1] + rnorm(100, 0, 0.000001)
-pardist[,5] = p[3] + rnorm(100, 0, 0.000001)
-```
-
-Compute optimal policy
-
-```{r}
-sdp = f_transition_matrix(f, p, x_grid, h_grid, sigma_g, pardist)
-s_opt <- value_iteration(sdp, x_grid, h_grid, OptTime=1000, xT, profit, delta)
-```
-
-Compare to the case without parameter uncertainty (growth noise only)
-
-```{r}
-SDP_Mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g)
-pars_fixed <- value_iteration(SDP_Mat, x_grid, h_grid, OptTime=1000, xT, profit, delta)
-```
-
-Plot results
-
-```{r}
-require(reshape2)
-policies <- melt(data.frame(stock=x_grid, pars.uncert = x_grid[s_opt$D], pars.fixed = x_grid[pars_fixed$D]), id="stock")
-ggplot(policies, aes(stock, stock - value, color=variable)) + geom_line(alpha=1) + xlab("stock size") + ylab("escapement") 
-```
-
-## Using actual estimates
 
@@
 ```
-
-## To Do
-
-So far this is just a proof of principle example.
-
-* Needs to be adjusted to account for uncertainty in the estimates of the noise processes as well. 
-* Needs to be written as generic and documented functions, add to nonparametric-bayes package routines.

### 2014-05-08 BUGS/process-noise-only.Rmd
preserve version history of scripts
import inst/examples with history

--- 2013-06-07 inst/examples/BUGS/myers.Rmd
@@
 
-```{r stateeq-local}
-f <- Myers
-p <- c(1.5, 2.5, 8)
-K <- 5  # approx, a li'l' less
-allee <- 1.2 # approx, a li'l' less
+Uses the model derived in ` citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,
+
+```{r stateeq}
 ```
@@
 
-```{r sdp-pars, dependson="stateeq-local"}
+```{r sdp-pars, dependson="stateeq"}
 ```

### 2014-05-08 aic_replicates.Rmd
import inst/examples with history

--- 2014-05-07 inst/examples/aic_replicates.Rmd
@@
-```{r}
+---
+layout: page
+---```{r}
 ## Set the paths for cache and figure

### 2014-05-08 reed-example.Rmd
import inst/examples with history

--- 2012-12-16 inst/examples/reed-example.Rmd
@@
-```{r set-options, echo = FALSE, cache = FALSE, external = TRUE}
+---
+layout: page
+---```{r set-options, include=FALSE, echo = FALSE, cache = FALSE, external = TRUE}
 opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "bevholt-cache/")
@@
 
+Simulate data
+
 ```{r lag-data}
@@
 ```{r par-est}
+```
+
+```{r gp-priors}
 ```

### 2014-05-08 scaling.Rmd
import inst/examples with history

--- 2013-08-26 inst/examples/scaling.Rmd
@@
-```{r, include=FALSE}
+---
+layout: page
+---```{r, include=FALSE}
 ## @knitr plotting-options

### 2014-05-12 manuscripts/manuscript.Rmd

@@
 layout: 12pt,review
-title: "Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach"
+linenumbers: true
+title: "Avoiding tipping points in ecosystem based fisheries management: a non-parametric Bayesian approach"
 author: 
@@
           may contain tipping points.  Before a collapse, we do not know
-          where the tipping points lie, if the exist at all.  Hence,
+          where the tipping points lie, if they exist at all.  Hence,
           we know neither a complete model of the system dynamics nor
@@
           performs nearly as well as the true model and significantly
-          outperforms standard appraoches.
+          outperforms standard appraoches. We illustrate this using examples
+          of simulated single-species dynamics, where the standard model
+          selection approach should be most effective, and find that it
+          still fails to account for uncertainty appropriately and leads
+          to population crashes, while management based on the Gaussian 
+          process approach does not. We discuss opportunities to adapt
+          this approach to multiple species models in further work.
 
@@
 Decision making under uncertainty is a ubiquitous challenge of
-natural resource management and conservation.  Ecological dynamics are
-frequently complex and difficult to measure, making uncertainty in our
-understanding a persistent challenge to effective management.
+natural resource management and conservation.  Ecological dynamics
+are frequently complex and difficult to measure, making uncertainty
+in our understanding a persistent challenge to effective management.
 Decision-theoretic approaches provide a framework to determine the
 best sequence of actions in face of uncertainty, but only when that
-uncertainty can be meaningfully quantified [@Fischer2009].  The sudden
-collapse of fisheries and other ecosystems has increasingly emphasized
-the difficulties of formulating even qualitatively correct models of
-the underlying processes.
+uncertainty can be meaningfully quantified [@Fischer2009].  Concerns
+about the potential sudden collapse of fisheries and other ecosystems
+[@Worm2006, @Scheffer2009] has increasingly emphasized the difficulties of
+formulating even qualitatively correct models of the underlying processes.
 
@@
 
+We focus on a class of one-dimensional models of population dynamics as a
+'best-case' scenario for the parametric model performance relative to
+the Gaussian process.   We consider this a best-case scenario for the
+following reasons: first, the underlying dynamics are always simuated
+from one the parametric models in the set, whereas in reality we never
+have the "true" model among our candidates.  By choosing one-dimensional
+models with few parameters, we limit the chance that poor performance
+will be due to our inability to estimate parameters well, something
+that becomes a more severe problem for higher-dimensional parametric
+models. Further, the parametric models we consider are those most commonly
+used in modeling stock-recruitment dynamics or to model sudden transitions
+between alternative stable states.
+
 ### Parametric models
@@
 dynamics: The Ricker model, the Allen model [@Allen2005], the Myers
-model [@Myers1995]. The familiar Ricker model involves two parameters, corresponding
+model [@Myers1995], equations~\ref{ricker}-\ref{myers}, which we parameterize as follows:
+
+We consider three parametric models that relate population size at time $t+1$, $X_{t+1}$, to population size at time $t$, $X_t$  and harvest at time $t$, $h_t$ by
+
+$$ X_{t+1}= Z_t f(S_t,p) $$
+
+Where Z(t) is log-normally distributed process stochasticity [@Reed1979] and $p$ is a vector of parameters, and $S_t$ the escapement [@Reed1979]
+
+$$S_t = X_t - h_t$$
+
+or the Ricker model,
+
+(@ricker) $$ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
+
+where $K$ is the carrying capacity and $r$ the growth rate.  For the Allen model,
+
+(@allen) $$ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_C\right)} $$
+
+Where $r$ and $K$ are as before, and $X_c$ denotes the location of the unstable (tipping) point.  For the Myers model,
+
+(@myers) $$ f(S(t) | p)  = \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}}$$
+
+Where $r$ and $K$ as before, and $\theta$ the nonlinearity of the Myers model (corresponding to Beverton-Holt dynamics for $\theta$ = 1 and an Allee effect with alternative stable states exist at $\theta > 2$).
+
+The familiar Ricker model involves two parameters, corresponding
 to a growth rate and a carrying capacity, and cannot support alternative
-stable state dynamics (though as growth rate increases it exhibits a
-periodic attractor that proceeds through period-doubling into chaos. We
-will generally focus on dynamics below the chaotic threshold for
-the purposes of this analysis.) The Allen model resembles the Ricker
+stable growth rate dynamics. The Allen model resembles the Ricker
 dynamics with an added Allee effect parameter [@Courchamp2008], below
@@
 
-<!-- equations just in appendix? -->
 Because of our interest in management performance in the presence of
@@
 We inferred posterior distributions for the parameters of each model
-in a Bayesian context using Gibbs sampling (implemented in R [@RTeam]
+in a Bayesian context using Gibbs sampling (@Gelman2003 implemented in R [@RTeam]
 using jags, [@R2jags]).  We choose uninformative uniform priors for all
 parameters (See Appendix, Figures and tables, and the R code
-provided). One-step-ahead predictions of these model fits are shown in
+provided). We show One-step-ahead predictions of these model fits in
 Figure 1.  Each chain was tested for Gelman-Rubin convergence and results were robust
@@
 
-We use a Metropolis-Hastings Markov Chain Monte Carlo to infer posterior
+We use a Metropolis-Hastings Markov Chain Monte Carlo (@Gelman2003) to infer posterior
 distributions of the two parameters of the GP (Figure S13, code in
@@
 It is reasonably straight-forward to derive the harvest policy from the
-estimated BP by inserting it into a Stochastic Dynamic Programming (SDP)
+estimated BP by inserting it into a SDP
 algorithm.  Recall that unlike the parametric models the Gaussian process
@@
 ```{r Figureb_posteriors, dependson=c("assemble-models", "par-fns"), include=TRUE, echo=FALSE, fig.cap="Points show the training data of stock-size over time.
-Curves show the posterior step-ahead predictions based on each of the estimated models.", fig.width=8, fig.height=6}
+Curves show the posterior step-ahead predictions based on each of the estimated models.
+Observe that all models are fitting the data reasonably well.", fig.width=8, fig.height=6}
 ```
@@
 
-```{r out_of_sample_predictions, include=TRUE, echo=FALSE, fig.width=8, fig.height=6, fig.cap="Out of sample predictions of the dynamics under each model.
+```{r out_of_sample_predictions, include=TRUE, echo=FALSE, fig.width=8, fig.height=6, fig.cap="Beyond the tipping point (outside the range of the training data), the true dynamics fall outside the uncertainty bands of the wrong parametric models, but inside the uncertainty predicted by the GP.
 Points show the stock size simulated by the true model.
 Overlay shows the range of states predicted by each model, based on the state observed in the previous time step.
-The Ricker model always predicts population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).
-Note that the GP is both more pessimistic and more uncertain about the future state than the parametric models, while the realized state often falls outside of the expected range forecast by the structurally incorrect Myers and Ricker models." }
+The Ricker model always (wrongly) predicts positive population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).
+Note that the GP is both more pessimistic and more uncertain about the future state than the parametric models." }
 ```
@@
 Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.",  include=TRUE, echo=FALSE}
+
+ggplot(sims_data) + 
+  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) + 
+  facet_wrap(~method) + guides(legend.position="none")
+
 ```
@@
 
-In any modeling effort, models must be chosen for the task at hand.
 Though simple mechanistically motivated models offer the greatest
@@
 reliably reflected by the data.
+
+<!--
 
@@
 direction for future investigation.
+
+-->
 
@@
 
+<!--
+
 Future directions
 -----------------
-
-<!--
-### Higher dimensions 
--->
 
@@
 prove even more valuable.
-
-<!--
-### Real-time learning
--->
 
@@
 
+-->
+
 Acknowledgments
@@
 
-This work was partially supported by NOAA-IAM grant to SM administered through the Center for Stock Assessment
-Research, a partnership between the University of California Santa Cruz
-and the Fisheries Ecology Division, Southwest Fisheries Science Center,
-Santa Cruz, CA and by NSF grant EF-0924195 to MM and NSF grant DBI-1306697
-to CB.
+This work was partially supported by NOAA-IAM grant to SM and Alec McCall and administered
+through the Center for Stock Assessment Research, a partnership between
+the University of California Santa Cruz and the Fisheries Ecology
+Division, Southwest Fisheries Science Center, Santa Cruz, CA and by NSF
+grant EF-0924195 to MM and NSF grant DBI-1306697 to CB.
 
 \appendix
-
-Model definitions and estimation
-=================================
+\renewcommand\thefigure{\thesection.\arabic{figure}}
+
+Appendix: Model definitions and estimation
+===========================================
 
@@
 
-The Ricker model is given by
-
-(@rickereq) $$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
-
-where $Z_t$ is log-normal noise of mean unity and log standard deviation $\sigma$, representing the stochastic growth, $X_t$ the stock size at time $t$, $S_t$ the escapement (unharvested population that will recruit in the following year, $S_t = X_t - h_t$).
-We place uniform priors on the growth rate $r$, carrying capacity $K$, and log-normal standard deviation parameter $\sigma$, over ranges given in Table 1.
-Posteriors are inferred by Gibbs sampling using Jags [@R2jags] (see provided code).
-
 <!-- Figure S1: Ricker model: prior and posterior distributions for parameter estimates. -->
 
-```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimate of the Ricker model", fig.width=6}
+```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
 plot_ricker_traces
@@
 xtable::xtable(ricker_priors_xtable,
-  caption = "parameterization range for the uniform priors in the Ricker model")
+  caption = "Parameterization range for the uniform priors in the Ricker model")
 ```
@@
 
-The Myers model [@Myers1995] is given by
-
-<!--Eq S2: Myers model -->
-(@myerseq) $$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}}$$
-
-where $Z_t$ is log-normal noise of mean unity and log standard deviation $\sigma$, representing the stochastic growth, $X_t$ the stock size at time $t$, $S_t$ the escapement (unharvested population that will recruit in the following year, $S_t = X_t - h_t$).
-We place uniform priors on the growth rate $r$, carrying capacity $K$, $\theta$ controls the strength of the nonlinearity, exhibiting an allee effect for $\theta \geq 2$, and log-normal standard deviation parameter $\sigma$, over ranges given in Table 1.
-Posteriors are inferred by Gibbs sampling using Jags [@R2jags] (see code provided).
-
 <!--
@@
 
-```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimate of the Myers model", fig.width=6}
+```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
 plot_myers_traces
@@
 xtable::xtable(myers_priors_xtable,
-           caption = "parameterization range for the uniform priors in the Myers model")
+           caption = "Parameterization range for the uniform priors in the Myers model")
 ```
@@
 
-The Allen model [@Allen2005] is given by
-
-(@alleneq) $$f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} $$
-
-where $Z_t$ is log-normal noise of mean unity and log standard deviation $\sigma$, representing the stochastic growth, $X_t$ the stock size at time $t$, $S_t$ the escapement (unharvested population that will recruit in the following year, $S_t = X_t - h_t$).
-We place uniform priors on the growth rate $r$, carrying capacity $K$, allee threshold $C$, and log-normal standard deviation parameter $\sigma$, over ranges given in Table 1.
-Posteriors are inferred by Gibbs sampling using Jags [@R2jags] (see code provided).
-
 <!-- Figure S3: Allen model: prior and posterior distributions for parameter estimates. -->
 
-```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimate of the Allen model", fig.width=6}
+```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
 plot_allen_traces
@@
 xtable::xtable(allen_priors_xtable,
-  caption = "parameterization range for the uniform priors in the Allen model")
+  caption = "Parameterization range for the uniform priors in the Allen model")
 ```
@@
 
-}
-
 Where $K$ is the Gaussian process kernel.  Throughout, we use the radial basis function kernel:
@@
 
-```{r echo=FALSE, include=TRUE, fig.cap = "Traces from the MCMC estimates of the GP model"}
+```{r echo=FALSE, include=TRUE, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)"}
 gp_assessment_plots$traces_plot
@@
 
+\newpage
+
 Optimal Control Problem Specification 
@@
 
-Where $Z_t$ is multiplicative noise function with mean 1, representing
-stochastic growth. We will consider log-normal noise with shape parameter
-$\sigma_g$.
+For simplicity we assume profit function $\Pi_t(X_t, h_t) is linear in the realized harvest, only
+enforcing the restriction that harvest can not exceed available stock, that is $\Pi(h,x) = min(h,x)$. $Z_t$ is as previously defined.
 
@@
 
-For simplicity we assume profit is simply linear in the realized harvest (only
-enforcing the restriction that harvest can not exceed available stock), $\Pi(h,x) = min(h,x)$.
-
-### Pseudocode for the Bellman iteration
-
-```r
-# compute the value for each possible harvest
-for(h in 1:length(h_grid)){
-  V1[h] = delta * F[[h]] %*% V +  profit(x_grid, h_grid[h]) 
-}
-# find havest h that gives the maximum value
-for(j in 1:gridsize){
-  value = max(V1[j,], na.rm = T)  # each col is a different h, max over these
-  index = which.max(V1[j,])       # store index so we can recover h's 
-  output[,j] = c(value, index)    # returns both profit value & index of optimal h.  
-}
-# Sets V[t+1] = max_h(V[t]) at each possible state value, x
-V = out[1,]                        # The new value-to-go
-D[,OptTime-time+1] = out[2,]       # The index positions
-```
+Where $V_t$ is the discounted value of the profit given that $X(t) = x$
 
@@
 All code used in producing this analysis can be found in the R package accompanying the paper, available from [http://github.com/cboettig/nonparametric-bayes].
-The code for all results shown here is dynamically embedded into the manuscript using Knitr, [@knitr].
+The code for all results shown here is dynamically embedded into the manuscript using `knitr` [@knitr].
 

### 2014-05-12 manuscripts/manuscript.Rmd
rename sims figure

@@
 
-```{r Figure3, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
+```{r simsFigure, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
 Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.",  include=TRUE, echo=FALSE}
@@
 
-For simplicity we assume profit function $\Pi_t(X_t, h_t) is linear in the realized harvest, only
-enforcing the restriction that harvest can not exceed available stock, that is $\Pi(h,x) = min(h,x)$. $Z_t$ is as previously defined.
+For simplicity we assume profit function $\Pi_t(X_t, h_t)$ is linear in the realized harvest, only
+enforcing the restriction that harvest can not exceed available stock, that is $\Pi(h,x) = \min(h,x)$. $Z_t$ is as previously defined.
 

### 2014-05-13 manuscripts/manuscript.Rmd
fix typo, add deploy-cache function

--- 2014-05-12 manuscripts/manuscript.Rmd
@@
 library(nonparametricbayes) 
-opts_chunk$set(external=TRUE)
-read_chunk("components/external-chunks.R")
 opts_chunk$set(fig.width=4, fig.height=3, echo=FALSE)
@@
 ```{r posterior-mode}
+require(modeest)
+posterior.mode <- function(x) {
+  mlv(x, method="shorth")$M
+}
 ```
 ```{r stateeq}
+f <- RickerAllee
+p <- c(2, 8, 5)
+K <- 10  # approx, a li'l' less
+allee <- 5 # approx, a li'l' less
 ```
 ```{r sdp-pars, dependson="stateeq"}
+sigma_g <- 0.05
+sigma_m <- 0.0
+z_g <- function() rlnorm(1, 0, sigma_g)
+z_m <- function() 1
+x_grid <- seq(0, 1.5 * K, length=50)
+h_grid <- x_grid
+profit <- function(x,h) pmin(x, h)
+delta <- 0.01
+OptTime <- 50  # stationarity with unstable models is tricky thing
+reward <- 0
+xT <- 0
+Xo <-  allee+.5# observations start from
+x0 <- K # simulation under policy starts from
+Tobs <- 40
+MaxT = 1000 # timeout for value iteration convergence
 ```
 ```{r obs, dependson="sdp-pars", fig.keep='none'}
+  set.seed(1234)
+  #harvest <- sort(rep(seq(0, .5, length=7), 5))
+  x <- numeric(Tobs)
+  x[1] <- Xo
+  nz <- 1
+  for(t in 1:(Tobs-1))
+    x[t+1] = z_g() * f(x[t], h=0, p=p)
+  obs <- data.frame(x = c(rep(0,nz), 
+                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
+                    y = c(rep(0,nz), 
+                          x[2:Tobs]))
+raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
 ```
@@
 ```{r mle, dependson="obs"}
+set.seed(12345)
+estf <- function(p){ 
+    mu <- f(obs$x,0,p)
+    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
+}
+par <- c(p[1]*rlnorm(1,0,.1), 
+         p[2]*rlnorm(1,0,.1), 
+         p[3]*rlnorm(1,0, .1), 
+         sigma_g * rlnorm(1,0,.1))
+o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
+f_alt <- f
+p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
+sigma_g_alt <- as.numeric(o$par[4])
+
+est <- list(f = f_alt, p = p_alt, sigma_g = sigma_g_alt, mloglik=o$value)
+
 ```
 ```{r mle-output, dependson="mle", results=toggle}
+true_means <- sapply(x_grid, f, 0, p)
+est_means <- sapply(x_grid, est$f, 0, est$p)
 ```
@@
 ```{r gp-priors}
+s2.p <- c(5,5)  
+d.p = c(10, 1/0.1)
 ```
 ```{r gp, dependson=c("gp-priors", "obs")}
+gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
+gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)
 ```
 ```{r gp_traces_densities, dependson="gp"}
+gp_assessment_plots <- summary_gp_mcmc(gp, burnin=1e4, thin=300)
 ```
 ```{r gp-output, dependson="gp", results=toggle}
+tgp_dat <- 
+    data.frame(  x = x_grid, 
+                 y = gp_dat$E_Ef, 
+                 ymin = gp_dat$E_Ef - 2 * sqrt(gp_dat$E_Vf), 
+                 ymax = gp_dat$E_Ef + 2 * sqrt(gp_dat$E_Vf) )
 ```
@@
 ```{r jags-setup}
+y <- x 
+N <- length(x);
+jags.data <- list("N","y")
+n.chains <- 6
+n.iter <- 1e6
+n.burnin <- floor(10000)
+n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
+n.update <- 10
+
 ```
 ```{r common-priors}
+stdQ_prior_p <- c(1e-6, 100)
+stdR_prior_p <- c(1e-6, .1)
+stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
+stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
+
 ```
 ```{r allen-model}
+K_prior_p <- c(0.01, 20.0)
+r0_prior_p <- c(0.01, 6.0)
+theta_prior_p <- c(0.01, 20.0)
+
+bugs.model <- 
+paste(sprintf(
+"model{
+  K     ~ dunif(%s, %s)
+  r0    ~ dunif(%s, %s)
+  theta ~ dunif(%s, %s)
+  stdQ ~ dunif(%s, %s)", 
+  K_prior_p[1], K_prior_p[2],
+  r0_prior_p[1], r0_prior_p[2],
+  theta_prior_p[1], theta_prior_p[2],
+  stdQ_prior_p[1], stdQ_prior_p[2]),
+
+  "
+  iQ <- 1 / (stdQ * stdQ);
+  y[1] ~ dunif(0, 10)
+  for(t in 1:(N-1)){
+    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K)* (y[t] - theta) / K 
+    y[t+1] ~ dlnorm(mu[t], iQ) 
+  }
+}")
+writeLines(bugs.model, "allen_process.bugs")
+
 ```
 ```{r allen-priors, dependson="common-priors"}
+K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
+r0_prior <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
+theta_prior <- function(x) dunif(x, theta_prior_p[1], theta_prior_p[2])
+par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
+                    r0 = r0_prior, theta = theta_prior,
+                    stdQ = stdQ_prior)
+
 ```
@@
 ```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results=toggle}
+jags.params=c("K","r0","theta","stdQ") # be sensible about the order here
+jags.inits <- function(){
+  list("K"= 10 * rlnorm(1,0, 0.1),
+       "r0"= 1 * rlnorm(1,0, 0.1) ,
+       "theta"=   5 * rlnorm(1,0, 0.1) , 
+       "stdQ"= abs( 0.1 * rlnorm(1,0, 0.1)),
+       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+}
+
+set.seed(1234)
+# parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
+allen_jags <- do.call(jags, list(data=jags.data, inits=jags.inits, 
+                                      jags.params, n.chains=n.chains, 
+                                      n.iter=n.iter, n.thin=n.thin, 
+                                      n.burnin=n.burnin, 
+                                      model.file="allen_process.bugs"))
+
+# Run again iteratively if we haven't met the Gelman-Rubin convergence criterion
+recompile(allen_jags) # required for parallel
+allen_jags <- do.call(autojags, 
+											list(object=allen_jags, n.update=n.update, 
+                           n.iter=n.iter, n.thin = n.thin))
+
 ```
 ```{r allen-traces, dependson="allen-mcmc"}
+tmp <- lapply(as.mcmc(allen_jags), as.matrix) # strip classes the hard way...
+allen_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
+names(allen_posteriors) = c("index", "variable", "value", "chain")
+plot_allen_traces <- ggplot(allen_posteriors) + geom_line(aes(index, value)) + 
+  facet_wrap(~ variable, scale="free", ncol=1)
+
 ```
 ```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
+allen_priors <- ddply(allen_posteriors, "variable", function(dd){
+    grid <- seq(min(dd$value), max(dd$value), length = 100) 
+    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
+})
+plot_allen_posteriors <- ggplot(allen_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+#  geom_line(data=allen_priors, aes(x=value, y=density), col="red") +  
+  facet_wrap(~ variable, scale="free", ncol=3)
+
 ```
 ```{r allen-output, dependson=c("posterior-mode", "allen-traces"), results=toggle}
+A <- allen_posteriors
+A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
+pardist <- acast(A, index ~ variable)
+bayes_coef <- apply(pardist,2, posterior.mode) 
+bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"], bayes_coef["theta"])) # parameters formatted for f
+allen_f <- function(x,h,p) unname(RickerAllee(x,h, unname(p[c("r0", "K", "theta")])))
+allen_means <- sapply(x_grid, f, 0, bayes_pars)
+bayes_pars
+head(pardist)
+
 ```
@@
 ```{r ricker-model}
+K_prior_p <- c(0.01, 40.0)
+r0_prior_p <- c(0.01, 20.0)
+bugs.model <- 
+paste(sprintf(
+"model{
+  K    ~ dunif(%s, %s)
+  r0    ~ dunif(%s, %s)
+  stdQ ~ dunif(%s, %s)", 
+  K_prior_p[1], K_prior_p[2],
+  r0_prior_p[1], r0_prior_p[2],
+  stdQ_prior_p[1], stdQ_prior_p[2]),
+  "
+  iQ <- 1 / (stdQ * stdQ);
+  y[1] ~ dunif(0, 10)
+  for(t in 1:(N-1)){
+    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K) 
+    y[t+1] ~ dlnorm(mu[t], iQ) 
+  }
+}")
+writeLines(bugs.model, "ricker_process.bugs")
+
 ```
 ```{r ricker-priors, dependson="common-priors"}
+K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
+r0_prior <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
+par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
+                   r0 = r0_prior, stdQ = stdQ_prior)
+
 ```
@@
 ```{r ricker-mcmc, dependson="ricker-model", results=toggle}
+jags.params=c("K","r0", "stdQ")
+jags.inits <- function(){
+  list("K"= 10 * rlnorm(1,0,.5),
+       "r0"= rlnorm(1,0,.5),
+       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
+       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+}
+set.seed(12345) 
+ricker_jags <- do.call(jags, 
+                       list(data=jags.data, inits=jags.inits, 
+                            jags.params, n.chains=n.chains, 
+                            n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
+                            model.file="ricker_process.bugs"))
+recompile(ricker_jags)
+ricker_jags <- do.call(autojags, 
+                       list(object=ricker_jags, n.update=n.update, 
+														n.iter=n.iter, n.thin = n.thin, 
+														progress.bar="none"))
+
 ```
 ```{r ricker-traces, dependson="ricker-mcmc"}
+tmp <- lapply(as.mcmc(ricker_jags), as.matrix) # strip classes the hard way...
+ricker_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
+names(ricker_posteriors) = c("index", "variable", "value", "chain")
+plot_ricker_traces <- ggplot(ricker_posteriors) + geom_line(aes(index, value)) + 
+  facet_wrap(~ variable, scale="free", ncol=1)
 ```
 ```{r ricker-posteriors, dependson=c("ricker-traces", "ricker-priors")}
+ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
+    grid <- seq(min(dd$value), max(dd$value), length = 100) 
+    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
+})
+# plot posterior distributions
+plot_ricker_posteriors <- ggplot(ricker_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+#  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") +  # don't plot priors 
+  facet_wrap(~ variable, scale="free", ncol=2)
 ```
 ```{r ricker-output, dependson=c("posterior-mode", "ricker-traces"), results=toggle}
+A <- ricker_posteriors
+A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
+ricker_pardist <- acast(A, index ~ variable)
+bayes_coef <- apply(ricker_pardist,2, posterior.mode) 
+ricker_bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"]))
+ricker_f <- function(x,h,p){
+  sapply(x, function(x){ 
+    x <- pmax(0, x-h) 
+    pmax(0, x * exp(p["r0"] * (1 - x / p["K"] )) )
+  })
+}
+ricker_means <- sapply(x_grid, Ricker, 0, ricker_bayes_pars[c(1,2)])
+head(ricker_pardist)
+ricker_bayes_pars
+
 ```
@@
 ```{r myers-model}
+r0_prior_p <- c(.0001, 10.0)
+theta_prior_p <- c(.0001, 10.0)
+K_prior_p <- c(.0001, 40.0)
+bugs.model <- 
+paste(sprintf(
+"model{
+  r0    ~ dunif(%s, %s)
+  theta    ~ dunif(%s, %s)
+  K    ~ dunif(%s, %s)
+  stdQ ~ dunif(%s, %s)", 
+  r0_prior_p[1], r0_prior_p[2],
+  theta_prior_p[1], theta_prior_p[2],
+  K_prior_p[1], K_prior_p[2],
+  stdQ_prior_p[1], stdQ_prior_p[2]),
+
+  "
+  iQ <- 1 / (stdQ * stdQ);
+
+  y[1] ~ dunif(0, 10)
+  for(t in 1:(N-1)){
+    mu[t] <- log(r0)  + theta * log(y[t]) - log(1 + pow(abs(y[t]), theta) / K)
+    y[t+1] ~ dlnorm(mu[t], iQ) 
+  }
+}")
+writeLines(bugs.model, "myers_process.bugs")
+
 ```
 ```{r myers-priors}
+K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
+r_prior     <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
+theta_prior <- function(x) dunif(x, theta_prior_p[1], theta_prior_p[2])
+par_priors <- list( deviance = function(x) 0 * x, K = K_prior,
+                    r0 = r_prior, theta = theta_prior, 
+                    stdQ = stdQ_prior)
+
 ```
@@
 ```{r myers-mcmc, dependson="myers-model", results=toggle}
+jags.params=c("r0", "theta", "K", "stdQ")
+jags.inits <- function(){
+  list("r0"= 1 * rlnorm(1,0,.1), 
+       "K"=    10 * rlnorm(1,0,.1),
+       "theta" = 1 * rlnorm(1,0,.1),  
+       "stdQ"= sqrt(0.2) * rlnorm(1,0,.1),
+       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
+}
+set.seed(12345)
+myers_jags <- do.call(jags, 
+                      list(data=jags.data, inits=jags.inits, 
+													 jags.params, n.chains=n.chains, 
+													 n.iter=n.iter, n.thin=n.thin,
+                           n.burnin=n.burnin, 
+                           model.file="myers_process.bugs"))
+recompile(myers_jags)
+myers_jags <- do.call(autojags, 
+                      list(myers_jags, n.update=n.update, 
+                           n.iter=n.iter, n.thin = n.thin, 
+                           progress.bar="none"))
+
 ```
 ```{r myers-traces, dependson="myers-mcmc"}
+tmp <- lapply(as.mcmc(myers_jags), as.matrix) # strip classes
+myers_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
+names(myers_posteriors) = c("index", "variable", "value", "chain")
+plot_myers_traces <- ggplot(myers_posteriors) + 
+  geom_line(aes(index, value)) + # priors, need to fix order though
+  facet_wrap(~ variable, scale="free", ncol=1)
+
 ```
 ```{r myers-posteriors, dependson="myers-traces"}
+par_prior_curves <- ddply(myers_posteriors, "variable", function(dd){
+    grid <- seq(min(dd$value), max(dd$value), length = 100) 
+    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
+})
+plot_myers_posteriors <- ggplot(myers_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+#  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") +  # Whoops, these are misaligned. see table instead 
+  facet_wrap(~ variable, scale="free", ncol=3)
+
 ```
 ```{r myers-output, dependson=c("posterior-mode", "myers-traces"), results=toggle}
+A <- myers_posteriors
+A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
+myers_pardist <- acast(A, index ~ variable)
+bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
+myers_bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["theta"], bayes_coef["K"]))
+myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
+myers_f <- function(x,h,p) Myer_harvest(x, h, p[c("r0", "theta", "K")])
+head(myers_pardist)
+myers_bayes_pars
+
 ```
@@
 ```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
+models <- data.frame(x=x_grid, 
+										 GP=tgp_dat$y, 
+										 True=true_means, 
+                     MLE=est_means, 
+										 Ricker=ricker_means, 
+                     Allen = allen_means,
+                     Myers = myers_means)
+models <- melt(models, id="x")
+# some labels
+names(models) <- c("x", "method", "value")
+# labels for the colorkey too
+model_names = c("GP", "True", "MLE", "Ricker", "Allen", "Myers")
+colorkey=cbPalette
+names(colorkey) = model_names 
+require(MASS)
+step_ahead <- function(x, f, p){
+  h = 0
+  x_predict <- sapply(x, f, h, p)
+  n <- length(x_predict) - 1
+  y <- c(x[1], x_predict[1:n])
+  y
+}
 ```
 ```{r gp-opt, dependson="gp-output"}
+matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid) 
+opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
 ```
 ```{r mle-opt, dependson="mle-output"}
+matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
+opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+matrices_estimated <- f_transition_matrix(est$f, est$p, x_grid, h_grid, est$sigma_g)
+opt_estimated <- value_iteration(matrices_estimated, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+
 ```
 ```{r allen-opt, dependson="allen-output"}
+matrices_allen <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
+opt_allen <- value_iteration(matrices_allen, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
 ```
 ```{r ricker-opt, dependson="ricker-output"}
+matrices_ricker <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
+opt_ricker <- value_iteration(matrices_ricker, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
 ```
 ```{r myers-opt, dependson="myers-output"}
+matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
+myers_alt <- value_iteration(matrices_myers, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
+
 ```
 ```{r assemble-opt, dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
+OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_ricker$D, Allen = opt_allen$D, Myers = myers_alt$D)
+colorkey=cbPalette
+names(colorkey) = names(OPT) 
 ```
 ```{r sims, dependson="assemble-opt"}
+sims <- lapply(OPT, function(D){
+  set.seed(1)
+  lapply(1:100, function(i) 
+    ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
+  )
+})
+
+dat <- melt(sims, id=names(sims[[1]][[1]]))
+sims_data <- data.table(dat)
+setnames(sims_data, c("L1", "L2"), c("method", "reps")) 
+# Legend in original ordering please, not alphabetical: 
+sims_data$method = factor(sims_data$method, ordered=TRUE, levels=names(OPT))
 ```
@@
 ```{r profits, dependson="sims", results=toggle}
+Profit <- sims_data[, sum(profit), by=c("reps", "method")]
+tmp <- dcast(Profit, reps ~ method)
+#tmp$Allen <- tmp[,"Allen"] + rnorm(dim(tmp)[1], 0, 1) # jitter for plotting
+tmp <- tmp / tmp[,"True"]
+tmp <- melt(tmp[2:dim(tmp)[2]])
+actual_over_optimal <-subset(tmp, variable != "True")
 ```
@@
 ```{r dic_calc, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output"), include=FALSE, echo=FALSE}
+dic.dic <- function (x) sum(x$deviance) +  sum(x[[2]])
+recompile(allen_jags)
+allen_dic <- dic.dic(dic.samples(allen_jags$model, n.iter=1000, type="popt"))
+recompile(ricker_jags)
+ricker_dic <- dic.dic(dic.samples(ricker_jags$model, n.iter=1000, type="popt"))
+recompile(myers_jags)
+myers_dic <- dic.dic(dic.samples(myers_jags$model, n.iter=1000, type="popt"))
+dictable <- data.frame(Allen = allen_dic + 2*length(bayes_pars), 
+                       Ricker = ricker_dic + 2*length(ricker_bayes_pars),
+                       Myers = myers_dic + 2*length(myers_bayes_pars), 
+                       row.names = c("DIC"))
 ```
@@
 
-```{r Figureb_posteriors, dependson=c("assemble-models", "par-fns"), include=TRUE, echo=FALSE, fig.cap="Points show the training data of stock-size over time.
+```{r Figure1, dependson=c("assemble-models", "par-fns"), include=TRUE, echo=FALSE, fig.cap="Points show the training data of stock-size over time.
 Curves show the posterior step-ahead predictions based on each of the estimated models.
 Observe that all models are fitting the data reasonably well.", fig.width=8, fig.height=6}
+step_ahead_posteriors <- function(x){
+  gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
+  df_post <- melt(lapply(sample(100, 30),  
+  function(i){
+    data.frame(time = 1:length(x), stock = x, 
+                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
+                True = step_ahead(x,f,p),  
+                MLE = step_ahead(x,f,est$p), 
+                Allen = step_ahead(x, allen_f, pardist[i,]), 
+                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
+                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
+  }), id=c("time", "stock"))
+}
+df_post <- step_ahead_posteriors(x)
+figure1b_posteriors <- ggplot(df_post) + geom_point(aes(time, stock)) + 
+  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
+  facet_wrap(~variable) + 
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) +  
+  theme(legend.position="none")
+figure1b_posteriors
+
 ```
@@
 ```{r deviances, dependson=c("dic_calc"), include=FALSE, echo=FALSE}
+allen_deviance  <- - posterior.mode(pardist[,'deviance'])
+ricker_deviance <- - posterior.mode(ricker_pardist[,'deviance'])
+myers_deviance  <- - posterior.mode(myers_pardist[,'deviance'])
+true_deviance   <- 2*estf(c(p, sigma_g))
+mle_deviance    <- 2*estf(c(est$p, est$sigma_g))
+aictable <- data.frame(Allen = allen_deviance + 2*(1+length(bayes_pars)),  # +1 for noise parameter
+                       Ricker = ricker_deviance + 2*(1+length(ricker_bayes_pars)),
+                       Myers = myers_deviance + 2*(1+length(myers_bayes_pars)), 
+                       row.names = c("AIC"))
+bictable <- data.frame(Allen = allen_deviance + log(length(x))*(1+length(bayes_pars)), 
+                       Ricker = ricker_deviance + log(length(x))*(1+length(ricker_bayes_pars)),
+                       Myers = myers_deviance + log(length(x))*(1+length(myers_bayes_pars)), 
+                       row.names = c("BIC"))
+xtable::xtable(rbind(dictable, aictable, bictable))
 ```
@@
 The GP is conditioned on (0,0), shown as a pseudo-data point.", include=TRUE, echo=FALSE, fig.width=8, fig.height=6}
+x_grid_short <- x_grid[1:40]
+gp_short <- gp_predict(gp, x_grid_short, burnin=1e4, thin=300)
+models_posteriors <- 
+  melt(lapply(sample(100, 50), 
+              function(i){
+    sample_gp <- mvrnorm(1, 
+                            gp_short$Ef_posterior[,i],         
+                            gp_short$Cf_posterior[[i]])
+    data.frame(stock = x_grid_short, 
+               GP = sample_gp,
+               y = sample_gp,
+               ymin = sample_gp - 2 * sqrt(gp_short$E_Vf), 
+               ymax = sample_gp + 2 * sqrt(gp_short$E_Vf), 
+               True = sapply(x_grid_short,f,0, p),  
+               MLE = sapply(x_grid_short,f,0, est$p), 
+               Allen = sapply(x_grid_short, allen_f, 0, pardist[i,]), 
+               Ricker = sapply(x_grid_short, ricker_f, 0, ricker_pardist[i,]), 
+               Myers = sapply(x_grid_short, myers_f, 0, myers_pardist[i,]))
+             }), 
+       id=c("stock", "y", "ymin", "ymax"))
+ggplot(models_posteriors) + 
+    geom_ribbon(aes(x=stock, y=y, ymin=ymin, ymax=ymax, group=L1), 
+                  fill = "gray80", 
+                  data=subset(models_posteriors, variable == "GP")) + 
+    geom_line(aes(stock, value, col = variable, 
+                  group=interaction(L1,variable)), 
+              alpha=.2) + 
+    geom_point(data = obs, aes(x,y), alpha = 0.8) + 
+    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
+    facet_wrap(~variable) + 
+    scale_colour_manual(values=colorkey) +  
+    theme(legend.position="none")
 ```
@@
 
-```{r out_of_sample_predictions, include=TRUE, echo=FALSE, fig.width=8, fig.height=6, fig.cap="Beyond the tipping point (outside the range of the training data), the true dynamics fall outside the uncertainty bands of the wrong parametric models, but inside the uncertainty predicted by the GP.
+```{r Figure2, include=TRUE, echo=FALSE, fig.width=8, fig.height=6, fig.cap="Beyond the tipping point (outside the range of the training data), the true dynamics fall outside the uncertainty bands of the wrong parametric models, but inside the uncertainty predicted by the GP.
 Points show the stock size simulated by the true model.
@@
 Note that the GP is both more pessimistic and more uncertain about the future state than the parametric models." }
+y <- numeric(8)
+y[1] <- 4.5
+for(t in 1:(length(y)-1))
+      y[t+1] = z_g() * f(y[t], h=0, p=p)
+# predicts means, does not reflect uncertainty estimate!
+crash_data <- step_ahead_posteriors(y)
+crash_data <- subset(crash_data, variable %in% c("GP", "Allen", "Ricker", "Myers"))
+ggplot(crash_data) + 
+  geom_boxplot(aes(as.factor(as.integer(time)), value, 
+                   fill = variable, col=variable), 
+               alpha=.7, outlier.size=1, position="identity") + 
+#  geom_line(aes(time, value, col = variable, 
+#            group=interaction(L1,variable)), alpha=.1) + 
+  geom_point(aes(time, stock), size = 3) + 
+  scale_fill_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")], 
+                      guide = guide_legend(override.aes = list(alpha = 1))) +  
+  scale_colour_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")], 
+                      guide = guide_legend(override.aes = list(alpha = 1))) +  
+  facet_wrap(~variable) + 
+  theme(legend.position="none") + xlab("time") + ylab("stock size") 
 ```
@@
 
-```{r Figure2, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
+```{r Figure3, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
 Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", include=TRUE, echo=FALSE}
+policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
+names(policies) <- c("stock", "method", "value")
+
+ggplot(policies, aes(stock, stock - value, color=method)) +
+  geom_line(lwd=1.2, alpha=0.8) + xlab("stock size") + ylab("escapement")  +
+  scale_colour_manual(values=colorkey)
 ```
@@
 
-```{r Figure3, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
+```{r Figure4, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
 Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.",  include=TRUE, echo=FALSE}
-
 ggplot(sims_data) + 
@@
 
-```{r Figure4, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
+```{r Figure5, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
 For each data set, the three models are estimated as described above.

### 2014-05-13 manuscripts/manuscript.Rmd
avoid sanitizing math in tables

@@
 ```{r allen_priors_xtable, include=TRUE, echo=FALSE}
-allen_priors_xtable <- data.frame(parameter = c("r0", "K", "theta", "sigma"),
+allen_priors_xtable <- data.frame(parameter = c("$r$", "$K$", "$X_C$", "$\\sigma$"),
            "lower bound" = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
@@
 ricker_priors_xtable <- data.frame(
-  parameter = c("r0", "K", "sigma"),
+  parameter = c("$r$", "$K$", "$\\sigma$"),
   "lower bound" = c(r0_prior_p[1], K_prior_p[1], stdQ_prior_p[1]),
@@
 ```{r myers_prior_xtable, include=TRUE, echo=FALSE} 
-myers_priors_xtable <- data.frame(parameter = c("r0", "K", "theta", "sigma"),
+myers_priors_xtable <- data.frame(parameter = c("$r$", "$K$", "$\\theta$", "$\\sigma$"),
            "lower bound" = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
@@
 })
-
-dat <- melt(sims, id=names(sims[[1]][[1]]))
+# turn the list into a data.frame
+keep.as.columns <- names(sims[[1]][[1]])
+dat <- melt(sims, id=keep.as.columns)
 sims_data <- data.table(dat)
@@
 xtable::xtable(ricker_priors_xtable,
-  caption = "Parameterization range for the uniform priors in the Ricker model")
+  caption = "Parameterization range for the uniform priors in the Ricker model", 
+  sanitize.text.function = identity)
 ```
@@
 xtable::xtable(myers_priors_xtable,
-           caption = "Parameterization range for the uniform priors in the Myers model")
+           caption = "Parameterization range for the uniform priors in the Myers model",
+           sanitize.text.function = identity)
 ```
@@
 xtable::xtable(allen_priors_xtable,
-  caption = "Parameterization range for the uniform priors in the Allen model")
+  caption = "Parameterization range for the uniform priors in the Allen model",
+  sanitize.text.function = identity)
 ```
@@
 
-For simplicity we assume profit function $\Pi_t(X_t, h_t) is linear in the realized harvest, only
+For simplicity we assume profit function $\Pi_t(X_t, h_t)$ is linear in the realized harvest, only
 enforcing the restriction that harvest can not exceed available stock, that is $\Pi(h,x) = min(h,x)$. $Z_t$ is as previously defined.

### 2014-05-13 manuscripts/manuscript.Rmd
latex symbols in tables now, using pandoc.table from pander

@@
 \appendix
-\renewcommand\thefigure{\thesection.\arabic{figure}}
+\renewcommand\thefigure{S.\arabic{figure}}
 
@@
 
-Ricker Model
-------------
-
 <!-- Equation S1: Ricker model. -->
@@
 
-```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
+```{r, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
 plot_ricker_traces
@@
 
-```{r, echo=FALSE, include=TRUE, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
+```{r, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
 ggplot(ricker_posteriors, aes(value)) + 
@@
 
-```{r echo=FALSE, results = "asis", include = TRUE, cache=FALSE}
-xtable::xtable(ricker_priors_xtable,
-  caption = "Parameterization range for the uniform priors in the Ricker model", 
-  sanitize.text.function = identity)
-```
-
-Myers Model
------------
+```{r, results = "asis"}
+pander::pandoc.table(ricker_priors_xtable,
+  caption = "Parameterization range for the uniform priors in the Ricker model")
+```
 
@@
 
-```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
+```{r, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
 plot_myers_traces
@@
 
-```{r, echo=FALSE, include=TRUE, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
+```{r, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
 ggplot(myers_posteriors, aes(value)) + 
@@
 
-```{r echo=FALSE, results="asis", include = TRUE, cache=FALSE}
-xtable::xtable(myers_priors_xtable,
-           caption = "Parameterization range for the uniform priors in the Myers model",
-           sanitize.text.function = identity)
-```
-
-Allen model
------------
+```{r, results="asis"}
+pander::pandoc.table(myers_priors_xtable,
+           caption = "Parameterization range for the uniform priors in the Myers model")
+```
 
@@
 
-```{r, echo=FALSE, include=TRUE, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
+```{r, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
 plot_allen_traces
@@
 
-```{r, echo=FALSE, include=TRUE, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
+```{r, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
 ggplot(allen_posteriors, aes(value)) + 
@@
 
-```{r echo=FALSE, results = "asis", include=TRUE, cache=FALSE}
-xtable::xtable(allen_priors_xtable,
-  caption = "Parameterization range for the uniform priors in the Allen model",
-  sanitize.text.function = identity)
+```{r, results = "asis"}
+pander::pandoc.table(allen_priors_xtable,
+  caption = "Parameterization range for the uniform priors in the Allen model")
 ```
@@
 
-```{r echo=FALSE, include=TRUE, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)"}
+```{r, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)"}
 gp_assessment_plots$traces_plot
@@
 
-```{r echo=FALSE, include=TRUE, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
+```{r, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
 gp_assessment_plots$posteriors_plot

### 2014-05-13 manuscripts/manuscript.Rmd
recompile documentation

@@
 \appendix
-\renewcommand\thefigure{S.\arabic{figure}}
+\renewcommand*{\thefigure}{S\arabic{figure}}
+\renewcommand*{\thetable}{S\arabic{table}}
+\setcounter{figure}{0}}
+\setcounter{table}{0}}
 
@@
 ===========================================
-
-<!-- Equation S1: Ricker model. -->
-
-<!-- Figure S1: Ricker model: prior and posterior distributions for parameter estimates. -->
-
-```{r, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
-plot_ricker_traces
-```
-
-```{r, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
-ggplot(ricker_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  facet_wrap(~ variable, scale="free", ncol=2)
-```
-
-<!-- Table S1: Parameterization of the priors -->
-
-```{r, results = "asis"}
-pander::pandoc.table(ricker_priors_xtable,
-  caption = "Parameterization range for the uniform priors in the Ricker model")
-```
-
-<!--
-Figure S2: Myers model: Traces, prior and posterior distributions for parameter estimates.
--->
-
-```{r, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
-plot_myers_traces
-```
-
-```{r, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
-ggplot(myers_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  facet_wrap(~ variable, scale="free", ncol=2)
-```
-
-<!-- Table S2: Parameterization of the priors -->
-
-```{r, results="asis"}
-pander::pandoc.table(myers_priors_xtable,
-           caption = "Parameterization range for the uniform priors in the Myers model")
-```
-
-<!-- Figure S3: Allen model: prior and posterior distributions for parameter estimates. -->
-
-```{r, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
-plot_allen_traces
-```
-
-```{r, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
-ggplot(allen_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  facet_wrap(~ variable, scale="free", ncol=2)
-```
-
-<!--
-Table S3: Parameterization of the priors
--->
-
-```{r, results = "asis"}
-pander::pandoc.table(allen_priors_xtable,
-  caption = "Parameterization range for the uniform priors in the Allen model")
-```
-
-<!-- Eq S4: GP model  -->
 
@@
 
-<!-- Figure S4: GP model: prior and posterior distributions for parameter estimates. -->
-
-```{r, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)"}
+```{r FigureS1, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)"}
 gp_assessment_plots$traces_plot
@@
 
-```{r, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
+```{r FigureS2, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
 gp_assessment_plots$posteriors_plot
@@
 
+```{r FigureS3, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
+plot_ricker_traces
+```
+
+```{r FigureS4, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
+ggplot(ricker_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+  facet_wrap(~ variable, scale="free", ncol=2)
+```
+
+```{r Table S1, results = "asis"}
+pander::pandoc.table(ricker_priors_xtable,
+  caption = "Parameterization range for the uniform priors in the Ricker model")
+```
+
+```{r FigureS5, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
+plot_myers_traces
+```
+
+```{r FigureS6, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
+ggplot(myers_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+  facet_wrap(~ variable, scale="free", ncol=2)
+```
+
+```{r TableS2, results="asis"}
+pander::pandoc.table(myers_priors_xtable,
+           caption = "Parameterization range for the uniform priors in the Myers model")
+```
+
+```{r FigureS7, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
+plot_allen_traces
+```
+
+```{r FigureS8, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
+ggplot(allen_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity", alpha=0.7) +
+  facet_wrap(~ variable, scale="free", ncol=2)
+```
+
+```{r TableS3, results = "asis"}
+pander::pandoc.table(allen_priors_xtable,
+  caption = "Parameterization range for the uniform priors in the Allen model")
+```
+
 ```{r include=FALSE}

### 2014-05-13 R/gp-replicate-fns.R

@@ -1,8 +1,6 @@
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-
-
 #' Simulate observed data for training the GP
 #' 
+#' Simulate observed data for training the GP
 #' @param Xo the initial population size
 #' @param z_g a function of no arguments providing the random multiplicative growth noise
 #' @param f a function of (x, h, p) providing the growth in population x under harvest h given pars p
@@ -33,6 +31,8 @@ sim_obs <- function(Xo, z_g, f, p, Tobs = 35, seed = 1, nz = 10,
 }
 
 
+#' MLE estimate of parameters given by function f
+#'
 #' MLE estimate of parameters given by function f
 #' @param obs the observed data, two columns giving x_t, x_t+1 respectively
 #' @return a list with the f given, the MLE estimated parameters, and estimated noise level
@@ -76,6 +76,8 @@ par_est <- function(obs,  init = c(r=1.5, K=mean(obs$x), s=1)){
   list(f_alt = f_alt, p_alt = p_alt, sigma_g_alt = sigma_g_alt)
 }
 
+
+#' helper function to determine the optimal policies of each model
 
 #' helper function to determine the optimal policies of each model
 #' @import pdgControl

### 2014-05-13 R/gp_fit.R

@@ -14,6 +14,7 @@
 #' Sigma, the covariance matrix for the posterior Gaussian process, and 
 #' "loglik", the log likelihood of observering the given data under the process, marginalized over the prior
 #' @details so far treates the prior as mean 0 and covariance given by cov(X)
+#' @import kernlab
 #' @export
 #' @examples 
 #' X <- seq(-5,5,len=50)

### 2014-05-13 R/gp-replicate-fns.R
docs updated

@@ -9,8 +9,9 @@
 #' @param seed the random number seed for the simulation
 #' @param nz the number of observations at 0 (a hack, should just condiition on 0,0)
 #' @param harvest a sequence of harvest values that occur durning the simulation
+#' @param plotdata logical, plot data?
 #' @details varying harvest values allow the system to explore the state space, making for better training data.
-#' @export
+#' @noRd 
 sim_obs <- function(Xo, z_g, f, p, Tobs = 35, seed = 1, nz = 10, 
                     harvest = sort(rep(seq(0, .5, length=7), 5)),
                     plotdata=FALSE){
@@ -36,7 +37,7 @@ sim_obs <- function(Xo, z_g, f, p, Tobs = 35, seed = 1, nz = 10,
 #' MLE estimate of parameters given by function f
 #' @param obs the observed data, two columns giving x_t, x_t+1 respectively
 #' @return a list with the f given, the MLE estimated parameters, and estimated noise level
-#' @export 
+#' @noRd 
 par_est_allee <- function(obs, f, p, 
                           init = c(p[1]+1, p[2]-1, p[3]-2, 
                                    s = sigma_g + abs(rnorm(1,0,.1)))
@@ -58,7 +59,7 @@ par_est_allee <- function(obs, f, p,
 #' MLE estimate of parameters given by function f
 #' @param obs the observed data, two columns giving x_t, x_t+1 respectively
 #' @return a list with the f given, the MLE estimated parameters, and estimated noise level
-#' @export 
+#' @noRd
 par_est <- function(obs,  init = c(r=1.5, K=mean(obs$x), s=1)){
   estf <- function(p){
     mu <- log(obs$x) + p["r"]*(1-obs$x/p["K"])
@@ -81,7 +82,7 @@ par_est <- function(obs,  init = c(r=1.5, K=mean(obs$x), s=1)){
 
 #' helper function to determine the optimal policies of each model
 #' @import pdgControl
-#' @export
+#' @noRd 
 optimal_policy <- function(gp, f, f_est, f_alt, p, p_est, p_alt, x_grid, h_grid, sigma_g, sigma_g_est, sigma_g_alt, delta, xT, profit, reward, OptTime){
   MaxT = 1000
   matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.ks2, x_grid, h_grid)
[…]

### 2014-05-13 manuscripts/manuscript.Rmd

@@
 ```{r plot-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
-library(nonparametricbayes) 
+library("nonparametricbayes") 
+library("pdgControl")
+
+library("ggplot2") 
+library("reshape2")
+library("plyr")
+library("data.table")
+
+library("R2jags")
+library("emdbook") # for as.mcmc.bugs (?)
+library("coda")  # for as.mcmc
+
+library("modeest")
+library("MASS")
+library("pander")
 opts_chunk$set(fig.width=4, fig.height=3, echo=FALSE)
-library(ggplot2) 
 theme_set(theme_bw(base_size=12))
@@
 ```{r posterior-mode}
-require(modeest)
+library(modeest)
 posterior.mode <- function(x) {
@@
 names(colorkey) = model_names 
-require(MASS)
 step_ahead <- function(x, f, p){

### 2014-05-13 manuscripts/manuscript.Rmd
fix figure filenames

@@
 <!-- Run R code for analysis, to be called by figures -->
-```{r plot-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
+```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+```
+
+```{r libraries, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
 library("nonparametricbayes") 
@@
 library("pander")
+
 opts_chunk$set(fig.width=4, fig.height=3, echo=FALSE)
 theme_set(theme_bw(base_size=12))
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
-               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-
 toggle = "hide" # results argument
@@
 
-```{r Figure1, dependson=c("assemble-models", "par-fns"), include=TRUE, echo=FALSE, fig.cap="Points show the training data of stock-size over time.
+```{r Figure_1, dependson=c("assemble-models", "par-fns", "plot-options"), fig.cap="Points show the training data of stock-size over time.
 Curves show the posterior step-ahead predictions based on each of the estimated models.
@@
 
-```{r statespace_posteriors, dependson=c("assemble-models"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
+```{r Figure_2, dependson=c("assemble-models", "plot-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
 Graph shows the expected value for the function $f$ under each model.
@@
 The training data is also shown as black points.
-The GP is conditioned on (0,0), shown as a pseudo-data point.", include=TRUE, echo=FALSE, fig.width=8, fig.height=6}
+The GP is conditioned on (0,0), shown as a pseudo-data point.", fig.width=8, fig.height=6}
 x_grid_short <- x_grid[1:40]
@@
 
-```{r Figure2, include=TRUE, echo=FALSE, fig.width=8, fig.height=6, fig.cap="Beyond the tipping point (outside the range of the training data), the true dynamics fall outside the uncertainty bands of the wrong parametric models, but inside the uncertainty predicted by the GP.
+```{r Figure_3, fig.width=8, fig.height=6, fig.cap="Beyond the tipping point (outside the range of the training data), the true dynamics fall outside the uncertainty bands of the wrong parametric models, but inside the uncertainty predicted by the GP.
 Points show the stock size simulated by the true model.
@@
 The Ricker model always (wrongly) predicts positive population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).
-Note that the GP is both more pessimistic and more uncertain about the future state than the parametric models." }
+Note that the GP is both more pessimistic and more uncertain about the future state than the parametric models.", dependson="plot-options"}
 y <- numeric(8)
@@
 <!--
-### Figure 3: Inferred Policies
+### Figure 4: Inferred Policies
 
@@
 
-```{r Figure3, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", include=TRUE, echo=FALSE}
+```{r Figure_4, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
+Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
 policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
@@
 
-```{r Figure4, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.",  include=TRUE, echo=FALSE}
+```{r Figure_5, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."}
 ggplot(sims_data) + 
@@
 
-```{r Figure5, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
+```{r Figure_6, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
 For each data set, the three models are estimated as described above.
 Values plotted are the averages of a given policy over 100 replicate simulations.
-Details and code provided in the supplement.",  include=TRUE, echo=FALSE, fig.width=6, fig.height=6}
+Details and code provided in the supplement.", fig.width=6, fig.height=6}
 
@@
 
-```{r sensitivity, fig.height=6, dependson="sensitivity-calc", echo=FALSE, fig.cap="Sensitivity Analysis.
+```{r Figure_7, fig.height=6, dependson="sensitivity-calc", fig.cap="Sensitivity Analysis.
 Histogram shows the ratio of the realized net present value / profits derived when managing under the GPDP Gaussian process approach over the optimal value given the true model and true parameters.
@@
 
-```{r FigureS1, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)"}
+```{r FigureS1, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options"}
 gp_assessment_plots$traces_plot

### 2014-05-13 manuscripts/components/knit.R
set to cache=2

@@ -16,7 +16,7 @@ library(knitr)
 basename <- gsub(".Rmd", "", file) # knitr:::knit_concord$get('infile')) 
 opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
                cache.path = paste("components/cache/", basename, "/", sep=""))
-opts_chunk$set(cache = 1)
+opts_chunk$set(cache = 2)
 opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, comment = NA, verbose = TRUE)
 
 # PDF-based figures

### 2014-05-13 manuscripts/manuscript.Rmd

@@
 about the potential sudden collapse of fisheries and other ecosystems
-[@Worm2006, @Scheffer2009] has increasingly emphasized the difficulties of
+[@Worm2006; @Scheffer2009] has increasingly emphasized the difficulties of
 formulating even qualitatively correct models of the underlying processes.
@@
 GP models have subsequently been used to test for the presence of Allee
-effects [@Sugeno2013], estimate the maximum reproductive
-rate [@Sugeno2013], determine temporal variation in food
+effects [@Sugeno2013a], estimate the maximum reproductive
+rate [@Sugeno2013b], determine temporal variation in food
 availability [@Sigourney2012], and provide a basis for identifying
@@
 using jags, [@R2jags]).  We choose uninformative uniform priors for all
-parameters (See Appendix, Figures and tables, and the R code
-provided). We show One-step-ahead predictions of these model fits in
+parameters (See appendix Tables S1-S3; R code
+provided). We show one-step-ahead predictions of these model fits in
 Figure 1.  Each chain was tested for Gelman-Rubin convergence and results were robust
@@
 
-```{r Figure_5, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."}
+```{r Figure5, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model." fig.width=6, fig.height=6}
 ggplot(sims_data) + 
@@
 
+Large noise relative to the distance between the stable and unstable
+point, (or more precisely, between the optimal escapement and the unstable
+point) increases the chance of a stochastic transition.  As a result,
+the impact of using a model which underestimates the risk of harvesting
+past the critical point is much more severe, since this such a situation
+occurs more often.  Conversely, with large enough distance between the
+optimal escapement and unstable points relative to the noise, the chance
+of a transition becomes vanishingly small and all models can be modeled
+near-optimally. Models which underestimate the cost incurred by stock
+sizes fluctuating significantly below the optimal escapement level will
+not perform poorly as long as those fluctuations are sufficently rare.
+
 ```{r Figure_7, fig.height=6, dependson="sensitivity-calc", fig.cap="Sensitivity Analysis.
@@
 mean in the prior without making the much stronger assumption that the
-Ricker is the structurally correct model [@Sugeno2013]. Future research
+Ricker is the structurally correct model [@Sugeno2013a]. Future research
 should focus on identifying criteria to ensure the prior and the value function

### 2014-05-13 manuscripts/manuscript.Rmd
fix typo in figure5 chunk opts

@@
 
-```{r Figure5, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model." fig.width=6, fig.height=6}
+```{r Figure_5, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
 ggplot(sims_data) + 

### 2014-05-13 tests/testthat/gp_mcmc_example.Rmd
knit scripts  as tests

--- 2014-05-08 gp_mcmc_example.Rmd
@@
 library(ggplot2) # plotting 
-opts_knit$set(upload.fun = socialR::flickr.url)
+#opts_knit$set(upload.fun = imgur_upload)
 theme_set(theme_bw(base_size=10))

### 2014-05-13 tests/testthat/test-gp.R

@@ -1,0 +1,6 @@
+
+library(knitr)
+knit("gaussian-process-basics.Rmd")
+knit("gp_mcmc_example.Rmd")
+expect_equals(length(grep("Error", readLines("gp_mcmc_example.md"))), 0)
+expect_equals(grep("Error", readLines("gaussian-process-basics.md"))), 0)

### 2014-05-14 tests/testthat/gaussian-process-basics.Rmd

--- 2013-05-14 inst/examples/gaussian-process-basics.Rmd
@@
 ```{r plotting-options}
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA, dev="CairoPDF", fig.width=6, fig.height=4)
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA, fig.width=6, fig.height=4)
 library(ggplot2) # plotting 
-opts_knit$set(upload.fun = socialR::flickr.url)
+#opts_knit$set(upload.fun = imgur_upload)
 theme_set(theme_bw(base_size=10))

### 2014-05-14 manuscripts/manuscript.Rmd
tweaking to trigger reruns without resetting all of cache

@@
 source("components/sensitivity.R")
+sensitivity_dat <- dat
+```
+
+```{r export-data}
+write.csv(sensitivity_dat, "sensitivity.csv")
 ```
@@
 
-```{r Figure_1, dependson=c("assemble-models", "par-fns", "plot-options"), fig.cap="Points show the training data of stock-size over time.
+```{r figure_1, dependson=c("assemble-models", "par-fns", "plot-options"), fig.cap="Points show the training data of stock-size over time.
 Curves show the posterior step-ahead predictions based on each of the estimated models.
@@
 
-```{r Figure_2, dependson=c("assemble-models", "plot-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
+```{r figure_2, dependson=c("assemble-models", "plot-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
 Graph shows the expected value for the function $f$ under each model.
@@
 
-```{r Figure_3, fig.width=8, fig.height=6, fig.cap="Beyond the tipping point (outside the range of the training data), the true dynamics fall outside the uncertainty bands of the wrong parametric models, but inside the uncertainty predicted by the GP.
+```{r figure_3, fig.width=8, fig.height=6, fig.cap="Beyond the tipping point (outside the range of the training data), the true dynamics fall outside the uncertainty bands of the wrong parametric models, but inside the uncertainty predicted by the GP.
 Points show the stock size simulated by the true model.
@@
 
-```{r Figure_4, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
+```{r figure_4, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
 Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
@@
 
-```{r Figure_5, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
+```{r figure_5, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
 Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
@@
 
-```{r Figure_6, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
+```{r figure_6, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
 For each data set, the three models are estimated as described above.
@@
 
-```{r Figure_7, fig.height=6, dependson="sensitivity-calc", fig.cap="Sensitivity Analysis.
+```{r figure_7, fig.height=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
 Histogram shows the ratio of the realized net present value / profits derived when managing under the GPDP Gaussian process approach over the optimal value given the true model and true parameters.
 Values near 1 indicate nearly optimal performance.
-Simulations cover a range of underlying parameter, model, replicate, and noise levels."}
-ggplot(dat) + geom_histogram(aes(value, fill=noise)) + xlim(0,1.0) + theme_bw() + xlab("value as fraction of the optimal") + facet_wrap(~model, ncol=1)
+Simulations cover a range of underlying parameter, model, replicate, and noise levels.", echo=TRUE}
+ggplot(sensitivity_dat) + 
+  geom_histogram(aes(value, fill=noise)) + 
+  xlim(0,1.0) + 
+  theme_bw() + 
+  xlab("value as fraction of the optimal") + 
+  facet_wrap(~model, ncol=1)
 ```
@@
 
-```{r FigureS1, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options"}
+```{r figure_S1, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options"}
 gp_assessment_plots$traces_plot
@@
 
-```{r FigureS2, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
+```{r figure_S2, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
 gp_assessment_plots$posteriors_plot
@@
 
-```{r FigureS3, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
+```{r figure_S3, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
 plot_ricker_traces
@@
 
-```{r FigureS4, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
+```{r figure_S4, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
 ggplot(ricker_posteriors, aes(value)) + 
@@
 
-```{r FigureS5, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
+```{r figure_S5, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
 plot_myers_traces
@@
 
-```{r FigureS6, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
+```{r figure_S6, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
 ggplot(myers_posteriors, aes(value)) + 
@@
 
-```{r FigureS7, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
+```{r figure_S7, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
 plot_allen_traces
@@
 
-```{r FigureS8, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
+```{r figure_S8, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
 ggplot(allen_posteriors, aes(value)) + 

### 2014-05-14 inst/examples/gaussian-process-basics.Rmd
examples use imgur for tests

@@
 ```{r plotting-options}
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA, dev="CairoPDF", fig.width=6, fig.height=4)
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA, fig.width=6, fig.height=4)
 library(ggplot2) # plotting 
-opts_knit$set(upload.fun = socialR::flickr.url)
+opts_knit$set(upload.fun = imgur_upload)
 theme_set(theme_bw(base_size=10))

### 2014-05-14 inst/examples/gp_mcmc_example.Rmd
example testing new routine
examples use imgur for tests

@@
 library(ggplot2) # plotting 
-opts_knit$set(upload.fun = socialR::flickr.url)
+opts_knit$set(upload.fun =imgur_upload)
 theme_set(theme_bw(base_size=10))
@@
 
-library(nonparametricbayes)
 ```
@@
 
+## Function definitions
+
+```{r}
+library(mcmc)
+#' Basic regression in Gaussian processes  
+#' 
+#' 
+#' @param x Observed x values, (vector or matrix with columns for each dimension of data)
+#' @param y Vector of observed y values in the training data
+#' @param init_pars the initial guesses for lengthscale l and process noise sigma_n
+#' @param n iterations of the metropolis algorithm
+#' @details Currently assumes the covariance function.  By default we will use 
+#' the squared exponential (also called radial basis or Gaussian, 
+#' though it is not this that gives Gaussian process it's name; 
+#' any covariance function would do) as the covariance function, 
+#' $\operatorname{cov}(X_i, X_j) = \exp\left(-\frac{(X_i - X_j)^2}{2 \ell ^ 2}\right)$
+#' @return the MCMC output, as constructed by \link{metrop} from the mcmc package.
+#' @export
+#' @import mcmc  
+gp_mcmc <- function(x, y, init_pars = c(l=1, sigma.n=1), n = 1e4, d.p = c(5,5), s2.p = c(5,5)){
+
+  lpriors <- function(pars){
+    prior <- unname(
+      dgamma(exp(pars[1]), d.p[1], scale = d.p[2]) *
+      dgamma(exp(pars[2]), s2.p[1], s2.p[2]) 
+    )
+    log(prior)
+  }
+  posterior <- function(pars, x, y){ 
+    l <- exp(pars[1])
+    sigma.n <- exp(pars[2])
+
+    cov <- function(X, Y) outer(X, Y, SE, l)
+    I <- diag(1, length(x))
+    K <- cov(x, x)
+
+    loglik <- - 0.5 * t(y) %*% solve(K + sigma.n^2 * I) %*% y -
+      log(det(K + sigma.n^2*I)) -
+      length(y) * log(2 * pi) / 2
+
+    loglik + lpriors(pars)
+  }
+
+  out <- metrop(posterior, log(init_pars), n, x = obs$x, y = obs$y)  
+  out$d.p <- d.p
+  out$s2.p <- s2.p
+
+  out
+}
+```
+
+```{r}
+#' predict the expected values and posterior distributions of the Gaussian Process
+#' 
+#' @param x_predict the values at which we desire predictions
+#' @param covs if TRUE, return covariances instead of variances (set high thinning as this is memory intensive)
+#' @import reshape2
+#' @export
+gp_predict <- function(gp, x_predict, burnin=0, thin=1, covs=FALSE){
+
+  postdist <- cbind(index=1:gp$nbatch, as.data.frame(exp(gp$batch)))
+  s <- seq(burnin+1, gp$nbatch, by=thin)
+  postdist <- postdist[s,]
+  names(postdist) <- c("index", names(gp$initial))
+
+  indices <- 1:dim(postdist)[1]
+
+  out <- lapply(indices, function(i){
+    l <- postdist[i, "l"]
+    sigma.n <- postdist[i, "sigma.n"]
+
+    cov <- function(X, Y) outer(X, Y, SE, l)
+    cov_xx_inv <- solve(cov(obs$x, obs$x) + sigma.n^2 * diag(1, length(obs$x)))
+    Ef <- cov(x_predict, obs$x) %*% cov_xx_inv %*% obs$y
+    Cf <- cov(x_predict, x_predict) - cov(x_predict, obs$x)  %*% cov_xx_inv %*% cov(obs$x, x_predict)
+    list(Ef = Ef, Cf = Cf, Vf = diag(Cf))
+  })
+
+  Ef_posterior <- sapply(out, `[[`, "Ef")
+  Cf_posterior <- sapply(out, `[[`, "Cf")
+  Vf_posterior <- sapply(out, `[[`, "Vf")
+
+  E_Ef <- rowMeans(Ef_posterior)
+  E_Cf <- matrix( apply(Cf_posterior, 1, sum) / dim(Cf_posterior)[2], ncol = sqrt(dim(Cf_posterior)[1]) )
+  E_Vf <- diag(E_Cf) # same as rowMeans(Vf_posterior)
+
+  # list format better for return
+  Cf_posterior <- lapply(out, `[[`, "Cf")
+
+  list(Ef_posterior = Ef_posterior, 
+       Vf_posterior = Vf_posterior,
+       Cf_posterior = Cf_posterior,
+       E_Ef = E_Ef, E_Cf = E_Cf, E_Vf = E_Vf)
+
+}  
+```
+
+```{r} 
+library(reshape2)
+library(ggplot2)
+library(plyr)
+#' Summary plots showing the trace and posteriors for the gp_mcmc estimates
+#' 
+#' @param gp a fit of the gaussian process from gp_mcmc
+#' @param burnin length of sequence to discard as a transient
+#' @param thin frequency of sub-sampling (make posterior distribution smaller if necessary)
+#' @return two ggplot2 objects, one plotting the trace and one 
+#' plotting the posteriors in black with priors overlaid in red.  
+#' @import reshape2
+#' @import ggplot2
+#' @import plyr
+#' @export
+summary_gp_mcmc <- function(gp, burnin=0, thin=1){
+
+  postdist <- cbind(index=1:gp$nbatch, as.data.frame(exp(gp$batch)))
+  s <- seq(burnin+1, gp$nbatch, by=thin)
+  postdist <- postdist[s,]
+  names(postdist) <- c("index", names(gp$initial))
+
+  # TRACES
+  df <- melt(postdist, id="index")
+  traces_plot <- 
+    ggplot(df) + geom_line(aes(index, value)) + 
+    facet_wrap(~ variable, scale="free", ncol=1)
+
+  s2.p <- gp$s2.p
+  d.p <- gp$d.p
+  s2_prior <- function(x) dgamma(x, s2.p[1], s2.p[2])
+  d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2])
+  prior_fns <- list(l = d_prior, sigma.n = s2_prior)
+
+  prior_curves <- ddply(df, "variable", function(dd){
+    grid <- seq(min(dd$value), max(dd$value), length = 100)
+    data.frame(value = grid, density = prior_fns[[dd$variable[1]]](grid))
+  })
+
+  # Posteriors (easier to read than histograms)
+  posteriors_plot <- ggplot(df, aes(value)) + 
+    stat_density(geom="path", position="identity", alpha=0.7) +
+    geom_line(data=prior_curves, aes(x=value, y=density), col="red") + 
+    facet_wrap(~ variable, scale="free", ncol=2)
+
+  print(traces_plot)
+  print(posteriors_plot)
+
+  out <- list(traces_plot = traces_plot, posteriors_plot = posteriors_plot)
+  invisible(out)
+}
+
+# Helper function: The default covariance function
+SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
+```
+
+## Example
+
 ```{r}
@@
 ```{r}
-gp <- gp_mcmc(obs$x, obs$y, d.p = c(1,1), s2.p = c(1,1))
+gp <- gp_mcmc(obs$x, obs$y, d.p = c(2,.5), s2.p = c(2,.5))
 ```

### 2014-05-14 manuscripts/manuscript.Rmd
re-wrap some text
fix equation reference

@@
 
-The nature of decision-making problems provides a compelling and pragmatic way to compare 
-models: Rather than compare models in terms of best fit to data, model performance is defined 
-in the concrete terms of the decision-maker's objectives. Although much argument can be made over the 'correct' objective
-function, e.g. how to account for the social value of fish left in the sea
-vs. the commercial value of fish harvested; see @Halpern2013 for further 
-discussion of this issue.  We can always compare model performance
-across multiple potential objectives.  The important point is that a decision-maker
-does not necessarily need a model that provides the best mechanistic understanding
-or the best long-term outcome, but rather the one that idelivers the best 
-management performance over the period of interest.
+The nature of decision-making problems provides a compelling and
+pragmatic way to compare models: Rather than compare models in terms
+of best fit to data, model performance is defined in the concrete terms
+of the decision-maker's objectives. Although much argument can be made
+over the 'correct' objective function, e.g. how to account for the
+social value of fish left in the sea vs. the commercial value of fish
+harvested; see @Halpern2013 for further discussion of this issue.  We can
+always compare model performance across multiple potential objectives.
+The important point is that a decision-maker does not necessarily need
+a model that provides the best mechanistic understanding or the best
+long-term outcome, but rather the one that idelivers the best management
+performance over the period of interest.
 
@@
 
-We use a Bayesian nonparameteric regression approach based on Gaussian process priors. 
-The use of Gaussian process (GP) regression (or "kriging" in the geospatial
-literature) to formulate a predictive model is relatively new in the
-context of modeling dynamical systems [@Kocijan2005], and was first introduced
-in the context ecological modeling and fisheries management in @Munch2005.
-GP models have subsequently been used to test for the presence of Allee
-effects [@Sugeno2013a], estimate the maximum reproductive
-rate [@Sugeno2013b], determine temporal variation in food
-availability [@Sigourney2012], and provide a basis for identifying
-model-misspecification [@Thorson2014].
+We use a Bayesian nonparameteric regression approach based on Gaussian
+process priors.  The use of Gaussian process (GP) regression (or
+"kriging" in the geospatial literature) to formulate a predictive
+model is relatively new in the context of modeling dynamical systems
+[@Kocijan2005], and was first introduced in the context ecological
+modeling and fisheries management in @Munch2005.  GP models have
+subsequently been used to test for the presence of Allee effects
+[@Sugeno2013a], estimate the maximum reproductive rate [@Sugeno2013b],
+determine temporal variation in food availability [@Sigourney2012], and
+provide a basis for identifying model-misspecification [@Thorson2014].
 An accessible and thorough introduction to the formulation and use of
-GPs can be found in @Rasmussen2006. The nonparametric moniker for
-the GP approach is unfortunate as (a) it still involves the estimation
-of parameters, and (b), statisticians also use non-parametric to mean
-models that make only weak distributional assumptions  [e.g. @Lehmann1975]. 
-However, we retain the terminology for consistency with previous
-studies.
+GPs can be found in @Rasmussen2006. The nonparametric moniker for the
+GP approach is unfortunate as (a) it still involves the estimation of
+parameters, and (b), statisticians also use non-parametric to mean models
+that make only weak distributional assumptions  [e.g. @Lehmann1975].
+However, we retain the terminology for consistency with previous studies.
 
@@
 dynamics: The Ricker model, the Allen model [@Allen2005], the Myers
-model [@Myers1995], equations~\ref{ricker}-\ref{myers}, which we parameterize as follows:
-
-We consider three parametric models that relate population size at time $t+1$, $X_{t+1}$, to population size at time $t$, $X_t$  and harvest at time $t$, $h_t$ by
+model [@Myers1995], equations (@ricker)-(@myers), which we parameterize as follows:
+
+We consider three parametric models that relate population size at time
+$t+1$, $X_{t+1}$, to population size at time $t$, $X_t$  and harvest at
+time $t$, $h_t$ by
 
@@
 
-Where Z(t) is log-normally distributed process stochasticity [@Reed1979] and $p$ is a vector of parameters, and $S_t$ the escapement [@Reed1979]
+Where Z(t) is log-normally distributed process stochasticity [@Reed1979]
+and $p$ is a vector of parameters, and $S_t$ the escapement [@Reed1979]
 
@@
 
-We inferred posterior distributions for the parameters of each model
-in a Bayesian context using Gibbs sampling (@Gelman2003 implemented in R [@RTeam]
-using jags, [@R2jags]).  We choose uninformative uniform priors for all
-parameters (See appendix Tables S1-S3; R code
-provided). We show one-step-ahead predictions of these model fits in
-Figure 1.  Each chain was tested for Gelman-Rubin convergence and results were robust
-to longer runs.  For each simulation we also applied several commonly
-used model selection criteria (AIC, BIC, DIC, see @Burnham2002)
+We inferred posterior distributions for the parameters of each model in
+a Bayesian context using Gibbs sampling (@Gelman2003 implemented in R
+[@RTeam] using jags, [@R2jags]).  We choose uninformative uniform priors
+for all parameters (See appendix Tables S1-S3; R code provided). We
+show one-step-ahead predictions of these model fits in Figure 1.
+Each chain was tested for Gelman-Rubin convergence and results were
+robust to longer runs.  For each simulation we also applied several
+commonly used model selection criteria (AIC, BIC, DIC, see @Burnham2002)
 to identify the best fitting model.
@@
 
-<!-- FIXME ... say much more about the GP, as per Munch 2005a   -->
-<!-- See much longer discusion above on "Background on GP".  Move some of that to here?  -->
-
-Using the simulated training data we also estimate a Gaussian process defined by a radial basis
-function kernel of two parameters: $\ell$, which gives the characteristic
-length-scale over which correlation between two points in state-space
-decays, and $\sigma$, which gives the scale of the process noise by which
-observations $Y_{t+1}$ may differ from their predicted values $X_{t+1}$
-given an observation of the previous state, $X_t$. @Munch2005a gives an
-accessible introduction to the use of Gaussian processes in providing a
-Bayesian nonparametric description of the stock-recruitment relationship.
+Using the simulated training data we also estimate a Gaussian process
+defined by a radial basis function kernel of two parameters: $\ell$,
+which gives the characteristic length-scale over which correlation between
+two points in state-space decays, and $\sigma$, which gives the scale
+of the process noise by which observations $Y_{t+1}$ may differ from
+their predicted values $X_{t+1}$ given an observation of the previous
+state, $X_t$. @Munch2005a gives an accessible introduction to the use
+of Gaussian processes in providing a Bayesian nonparametric description
+of the stock-recruitment relationship.
 
@@
 
-We use a Metropolis-Hastings Markov Chain Monte Carlo (@Gelman2003) to infer posterior
-distributions of the two parameters of the GP (Figure S13, code in
-appendix), under weakly informative Gaussian priors (see parameters in
-table S5). As the posterior distributions differ substantially from the
-priors (Figure S13), we can be assured that most of the information in
-the posterior comes from the data rather than the prior belief.
-
-<!--
-In contrast to the parametric models, this posterior distribution is still
-conditional on the training data. As such, the uncertainty near the
-observed data.
--->
+We use a Metropolis-Hastings Markov Chain Monte Carlo (@Gelman2003) to
+infer posterior distributions of the two parameters of the GP (Figure
+S13, code in appendix), under weakly informative Gaussian priors
+(see parameters in table S5). As the posterior distributions differ
+substantially from the priors (Figure S13), we can be assured that most
+of the information in the posterior comes from the data rather than the
+prior belief.
 
@@
 It is reasonably straight-forward to derive the harvest policy from the
-estimated BP by inserting it into a SDP
-algorithm.  Recall that unlike the parametric models the Gaussian process
-with fixed parameters already predicts a distribution of curves rather
-than a single curve. We must first integrate over this distribution of
-curves given a sampling of parameter values drawn from the posterior
-distribution of the two GP parameters, before integrating over the
-posterior of those parameters themselves.
+estimated BP by inserting it into a SDP algorithm.  Recall that unlike
+the parametric models the Gaussian process with fixed parameters already
+predicts a distribution of curves rather than a single curve. We must
+first integrate over this distribution of curves given a sampling of
+parameter values drawn from the posterior distribution of the two GP
+parameters, before integrating over the posterior of those parameters
+themselves.
 
@@
 Values near 1 indicate nearly optimal performance.
-Simulations cover a range of underlying parameter, model, replicate, and noise levels.", echo=TRUE}
+Simulations cover a range of underlying parameter, model, replicate, and noise levels."}
 ggplot(sensitivity_dat) + 

### 2014-05-14 inst/examples/small-allee-plots.R
scripts for additional exploratory analysis and results

@@ -1,0 +1,222 @@
+
+## ----plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE----
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+
+
+## ----libraries, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE----
+library("methods")
+library("knitr")
+
+library("nonparametricbayes") 
+library("pdgControl")
+
+library("ggplot2") 
+library("reshape2")
+library("plyr")
+library("data.table")
+
+library("R2jags")
+library("emdbook") # for as.mcmc.bugs (?)
+library("coda")  # for as.mcmc
+
+library("modeest")
+library("MASS")
+
+opts_knit$set(fig.path="figure/small-allee/")
+
+opts_chunk$set(fig.width=6, fig.height=6, echo=FALSE, dev="pdf", warning=FALSE, message=FALSE, echo=FALSE)
+theme_set(theme_bw(base_size=12))
+toggle = "hide" # results argument
+
+
+
+
+
+load("small-allee.rda")
+
+## ----figure_1, dependson=c("assemble-models", "par-fns"), fig.cap="Points show the training data of stock-size over time.  Curves show the posterior step-ahead predictions based on each of the estimated models. Observe that all models are fitting the data reasonably well.", fig.width=8, fig.height=6----
+step_ahead_posteriors <- function(x){
+  gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
+  df_post <- melt(lapply(sample(100, 30),  
+  function(i){
+    data.frame(time = 1:length(x), stock = x, 
+                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
+                True = step_ahead(x,f,p),  
+                MLE = step_ahead(x,f,est$p), 
+                Allen = step_ahead(x, allen_f, pardist[i,]), 
+                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
+                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
+  }), id=c("time", "stock"))
+}
+df_post <- step_ahead_posteriors(x)
+figure1b_posteriors <- ggplot(df_post) + geom_point(aes(time, stock)) + 
[…]

### 2014-05-14 inst/examples/small-allee.R

@@ -1,0 +1,723 @@
+
+## ----plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE----
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+
+
+## ----libraries, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE----
+library("methods")
+library("knitr")
+
+library("nonparametricbayes") 
+library("pdgControl")
+
+library("ggplot2") 
+library("reshape2")
+library("plyr")
+library("data.table")
+
+library("R2jags")
+library("emdbook") # for as.mcmc.bugs (?)
+library("coda")  # for as.mcmc
+
+library("modeest")
+library("MASS")
+library("pander")
+
+opts_chunk$set(fig.width=4, fig.height=3, echo=FALSE)
+theme_set(theme_bw(base_size=12))
+toggle = "hide" # results argument
+
+
+## ----posterior-mode------------------------------------------------------
+library(modeest)
+posterior.mode <- function(x) {
+  mlv(x, method="shorth")$M
+}
+
+## ----stateeq-------------------------------------------------------------
+f <- RickerAllee
+p <- c(2, 8, 2)
+K <- 8
+allee <- 2 
+
+## ----sdp-pars, dependson="stateeq"---------------------------------------
+sigma_g <- 0.05
+sigma_m <- 0.0
+z_g <- function() rlnorm(1, 0, sigma_g)
+z_m <- function() 1
+x_grid <- seq(0, 1.5 * K, length=50)
+h_grid <- x_grid
+profit <- function(x,h) pmin(x, h)
+delta <- 0.01
+OptTime <- 50  # stationarity with unstable models is tricky thing
+reward <- 0
+xT <- 0
+Xo <-  allee+.5# observations start from
+x0 <- 10 # simulation under policy starts from
+Tobs <- 40
+MaxT = 1000 # timeout for value iteration convergence
+
+## ----obs, dependson="sdp-pars", fig.keep='none'--------------------------
+  set.seed(1234)
+  #harvest <- sort(rep(seq(0, .5, length=7), 5))
+  x <- numeric(Tobs)
+  x[1] <- Xo
+  nz <- 1
+  for(t in 1:(Tobs-1))
[…]

### 2014-05-14 inst/examples/small-noise-plots.R

@@ -1,0 +1,201 @@
+
+## ----plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE----
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+
+
+## ----libraries, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE----
+library("methods")
+library("knitr")
+library("nonparametricbayes") 
+library("pdgControl")
+
+library("ggplot2") 
+library("reshape2")
+library("plyr")
+library("data.table")
+
+library("R2jags")
+library("emdbook") # for as.mcmc.bugs (?)
+library("coda")  # for as.mcmc
+
+library("modeest")
+library("MASS")
+library("pander")
+
+opts_knit$set(fig.path="figure/small-noise/")
+opts_chunk$set(fig.width=6, fig.height=6, echo=FALSE, dev="pdf", warning=FALSE, message=FALSE, echo=FALSE)
+theme_set(theme_bw(base_size=12))
+toggle = "hide" # results argument
+
+
+## ----posterior-mode------------------------------------------------------
+library(modeest)
+posterior.mode <- function(x) {
+  mlv(x, method="shorth")$M
+}
+
+
+load("small-noise.rda")
+
+## ----figure_1, dependson=c("assemble-models", "par-fns", "plot-options"), fig.cap="Points show the training data of stock-size over time.  Curves show the posterior step-ahead predictions based on each of the estimated models. Observe that all models are fitting the data reasonably well.", fig.width=8, fig.height=6----
+step_ahead_posteriors <- function(x){
+  gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
+  df_post <- melt(lapply(sample(100, 30),  
+  function(i){
+    data.frame(time = 1:length(x), stock = x, 
+                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
+                True = step_ahead(x,f,p),  
+                MLE = step_ahead(x,f,est$p), 
+                Allen = step_ahead(x, allen_f, pardist[i,]), 
+                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
+                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
[…]

### 2014-05-14 inst/examples/small-noise.R

@@ -1,0 +1,685 @@
+
+## ----plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE----
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+
+
+## ----libraries, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE----
+library("methods")
+library("knitr")
+library("nonparametricbayes") 
+library("pdgControl")
+
+library("ggplot2") 
+library("reshape2")
+library("plyr")
+library("data.table")
+
+library("R2jags")
+library("emdbook") # for as.mcmc.bugs (?)
+library("coda")  # for as.mcmc
+
+library("modeest")
+library("MASS")
+library("pander")
+
+opts_chunk$set(fig.width=4, fig.height=3, echo=FALSE)
+theme_set(theme_bw(base_size=12))
+toggle = "hide" # results argument
+
+
+## ----posterior-mode------------------------------------------------------
+library(modeest)
+posterior.mode <- function(x) {
+  mlv(x, method="shorth")$M
+}
+
+## ----stateeq-------------------------------------------------------------
+f <- RickerAllee
+p <- c(2, 8, 5)
+K <- 8  
+allee <- 5 
+
+## ----sdp-pars, dependson="stateeq"---------------------------------------
+sigma_g <- 0.01
+sigma_m <- 0.0
+z_g <- function() rlnorm(1, 0, sigma_g)
+z_m <- function() 1
+x_grid <- seq(0, 1.5 * K, length=50)
+h_grid <- x_grid
+profit <- function(x,h) pmin(x, h)
+delta <- 0.01
+OptTime <- 50  # stationarity with unstable models is tricky thing
+reward <- 0
+xT <- 0
+Xo <-  allee+.5# observations start from
+x0 <- 10 # simulation under policy starts from
+Tobs <- 40
+MaxT = 1000 # timeout for value iteration convergence
+
+## ----obs, dependson="sdp-pars", fig.keep='none'--------------------------
+  set.seed(1234)
+  #harvest <- sort(rep(seq(0, .5, length=7), 5))
+  x <- numeric(Tobs)
+  x[1] <- Xo
+  nz <- 1
+  for(t in 1:(Tobs-1))
[…]

### 2014-05-14 manuscripts/components/install.R
add install script to make

@@ -1,0 +1,9 @@
+#!/usr/bin/Rscript
+
+if(!require("nonparametricbayes")){
+  install.packages("devtools")
+  library("devtools")
+  install_github("reshape")
+  install_github("cboettig/pdg_control")
+  install.packages("..", repos =NULL, dependencies = c("Depends", "Imports", "Suggests"))
+}

### 2014-05-14 manuscripts/manuscript.tex
track the tex intermediary file

--- 2014-05-14 manuscripts/manuscript.Rmd
@@
----
-layout: 12pt,review
-linenumbers: true
-title: "Avoiding tipping points in ecosystem based fisheries management: a non-parametric Bayesian approach"
-author: 
-  - name: "Carl Boettiger"
-    affiliation: cstar
-    email: cboettig@gmail.com
-    footnote: Corresponding author
-  - name: Marc Mangel
-    affiliation: cstar
-  - name: Stephan Munch
-    affiliation: noaa
-address: 
-  - code: cstar
-    address: | 
-      Center for Stock Assessment Research, 
-      Department of Applied Math and Statistics, 
-      University of California, Mail Stop SOE-2,
-      Santa Cruz, CA 95064, USA
-  - code: noaa
-    address: | 
-      Southwest Fisheries Science Center, 
-      National Oceanic and Atmospheric Administration, 
-      110 Shaffer Road, Santa Cruz, CA 95060, USA
-
-abstract: |
-          Model uncertainty and limited data coverage are fundamental
-          challenges to robust ecosystem management.  These challenges
-          are acutely highlighted by concerns that many ecological systems
-          may contain tipping points.  Before a collapse, we do not know
-          where the tipping points lie, if they exist at all.  Hence,
-          we know neither a complete model of the system dynamics nor
-          do we have access to data in some large region of state-space
-          where such a tipping point might exist.  These two sources of
-          uncertainty frustrate state-of-the-art parametric approaches
-          to decision theory and optimal control.  Here we illustrate
-          how a non-parametric approach using a Gaussian process prior
-          provides a more flexible representation of this inherent
-          uncertainty. We can then embed the Gaussian process in a
-          stochastic dynamic programming framework in order to make
-          robust management predictions under both model uncertainty and
-          limited data. We conduct a set of simulations to evaluate this
-          approach as compared with the standard approach of using model
-          selection to choose from a set of candidate models.  In keeping
-          with previous results, we find that model selection erroneously
-          favors models without tipping points -- leading to harvest policies
-          that guarantee extinction. The Gaussian process approach
-          performs nearly as well as the true model and significantly
-          outperforms standard appraoches. We illustrate this using examples
-          of simulated single-species dynamics, where the standard model
-          selection approach should be most effective, and find that it
-          still fails to account for uncertainty appropriately and leads
-          to population crashes, while management based on the Gaussian 
-          process approach does not. We discuss opportunities to adapt
-          this approach to multiple species models in further work.
-
-keywords:
-  - Bayesian
-  - Structural Uncertainty
-  - Nonparametric 
-  - Optimal Control
-  - Decision Theory
-  - Gaussian Processes
-  - Fisheries Management
----
-
-<!-- Run R code for analysis, to be called by figures -->
-```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
-               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-
-```
-
-```{r libraries, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-library("nonparametricbayes") 
-library("pdgControl")
-
-library("ggplot2") 
-library("reshape2")
-library("plyr")
-library("data.table")
-
-library("R2jags")
-library("emdbook") # for as.mcmc.bugs (?)
-library("coda")  # for as.mcmc
-
-library("modeest")
-library("MASS")
-library("pander")
-
-opts_chunk$set(fig.width=4, fig.height=3, echo=FALSE)
-theme_set(theme_bw(base_size=12))
-toggle = "hide" # results argument
-```
-
-```{r posterior-mode}
-library(modeest)
-posterior.mode <- function(x) {
-  mlv(x, method="shorth")$M
-}
-```
-```{r stateeq}
-f <- RickerAllee
-p <- c(2, 8, 5)
-K <- 10  # approx, a li'l' less
-allee <- 5 # approx, a li'l' less
-```
-```{r sdp-pars, dependson="stateeq"}
-sigma_g <- 0.05
-sigma_m <- 0.0
-z_g <- function() rlnorm(1, 0, sigma_g)
-z_m <- function() 1
-x_grid <- seq(0, 1.5 * K, length=50)
-h_grid <- x_grid
-profit <- function(x,h) pmin(x, h)
-delta <- 0.01
-OptTime <- 50  # stationarity with unstable models is tricky thing
-reward <- 0
-xT <- 0
-Xo <-  allee+.5# observations start from
-x0 <- K # simulation under policy starts from
-Tobs <- 40
-MaxT = 1000 # timeout for value iteration convergence
-```
-```{r obs, dependson="sdp-pars", fig.keep='none'}
-  set.seed(1234)
-  #harvest <- sort(rep(seq(0, .5, length=7), 5))
-  x <- numeric(Tobs)
-  x[1] <- Xo
-  nz <- 1
-  for(t in 1:(Tobs-1))
-    x[t+1] = z_g() * f(x[t], h=0, p=p)
-  obs <- data.frame(x = c(rep(0,nz), 
-                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
-                    y = c(rep(0,nz), 
-                          x[2:Tobs]))
-raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
-```
-
-```{r mle, dependson="obs"}
-set.seed(12345)
-estf <- function(p){ 
-    mu <- f(obs$x,0,p)
-    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
-}
-par <- c(p[1]*rlnorm(1,0,.1), 
-         p[2]*rlnorm(1,0,.1), 
-         p[3]*rlnorm(1,0, .1), 
-         sigma_g * rlnorm(1,0,.1))
-o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
-f_alt <- f
-p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
-sigma_g_alt <- as.numeric(o$par[4])
-
-est <- list(f = f_alt, p = p_alt, sigma_g = sigma_g_alt, mloglik=o$value)
-
-```
-```{r mle-output, dependson="mle", results=toggle}
-true_means <- sapply(x_grid, f, 0, p)
-est_means <- sapply(x_grid, est$f, 0, est$p)
-```
-
-```{r gp-priors}
-s2.p <- c(5,5)  
-d.p = c(10, 1/0.1)
-```
-```{r gp, dependson=c("gp-priors", "obs")}
-gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
-gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)
-```
-```{r gp_traces_densities, dependson="gp"}
-gp_assessment_plots <- summary_gp_mcmc(gp, burnin=1e4, thin=300)
-```
-```{r gp-output, dependson="gp", results=toggle}
-tgp_dat <- 
-    data.frame(  x = x_grid, 
-                 y = gp_dat$E_Ef, 
-                 ymin = gp_dat$E_Ef - 2 * sqrt(gp_dat$E_Vf), 
-                 ymax = gp_dat$E_Ef + 2 * sqrt(gp_dat$E_Vf) )
-```
-
-```{r jags-setup}
-y <- x 
-N <- length(x);
-jags.data <- list("N","y")
-n.chains <- 6
-n.iter <- 1e6
-n.burnin <- floor(10000)
-n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
-n.update <- 10
-
-```
-```{r common-priors}
-stdQ_prior_p <- c(1e-6, 100)
-stdR_prior_p <- c(1e-6, .1)
-stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
-stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
-
-```
-```{r allen-model}
-K_prior_p <- c(0.01, 20.0)
-r0_prior_p <- c(0.01, 6.0)
-theta_prior_p <- c(0.01, 20.0)
-
-bugs.model <- 
-paste(sprintf(
-"model{
-  K     ~ dunif(%s, %s)
-  r0    ~ dunif(%s, %s)
-  theta ~ dunif(%s, %s)
-  stdQ ~ dunif(%s, %s)", 
-  K_prior_p[1], K_prior_p[2],
-  r0_prior_p[1], r0_prior_p[2],
-  theta_prior_p[1], theta_prior_p[2],
-  stdQ_prior_p[1], stdQ_prior_p[2]),
-
-  "
-  iQ <- 1 / (stdQ * stdQ);
-  y[1] ~ dunif(0, 10)
-  for(t in 1:(N-1)){
-    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K)* (y[t] - theta) / K 
-    y[t+1] ~ dlnorm(mu[t], iQ) 
-  }
-}")
-writeLines(bugs.model, "allen_process.bugs")
-
-```
-```{r allen-priors, dependson="common-priors"}
-K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
-r0_prior <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
-theta_prior <- function(x) dunif(x, theta_prior_p[1], theta_prior_p[2])
-par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
-                    r0 = r0_prior, theta = theta_prior,
-                    stdQ = stdQ_prior)
-
-```
-
-```{r allen_priors_xtable, include=TRUE, echo=FALSE}
-allen_priors_xtable <- data.frame(parameter = c("$r$", "$K$", "$X_C$", "$\\sigma$"),
-           "lower bound" = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
-           "upper bound" = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2]))
-```
-
-```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results=toggle}
-jags.params=c("K","r0","theta","stdQ") # be sensible about the order here
-jags.inits <- function(){
-  list("K"= 10 * rlnorm(1,0, 0.1),
-       "r0"= 1 * rlnorm(1,0, 0.1) ,
-       "theta"=   5 * rlnorm(1,0, 0.1) , 
-       "stdQ"= abs( 0.1 * rlnorm(1,0, 0.1)),
-       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
-}
-
-set.seed(1234)
-# parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
-allen_jags <- do.call(jags, list(data=jags.data, inits=jags.inits, 
-                                      jags.params, n.chains=n.chains, 
-                                      n.iter=n.iter, n.thin=n.thin, 
-                                      n.burnin=n.burnin, 
-                                      model.file="allen_process.bugs"))
-
-# Run again iteratively if we haven't met the Gelman-Rubin convergence criterion
-recompile(allen_jags) # required for parallel
-allen_jags <- do.call(autojags, 
-											list(object=allen_jags, n.update=n.update, 
-                           n.iter=n.iter, n.thin = n.thin))
-
-```
-```{r allen-traces, dependson="allen-mcmc"}
-tmp <- lapply(as.mcmc(allen_jags), as.matrix) # strip classes the hard way...
-allen_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
-names(allen_posteriors) = c("index", "variable", "value", "chain")
-plot_allen_traces <- ggplot(allen_posteriors) + geom_line(aes(index, value)) + 
-  facet_wrap(~ variable, scale="free", ncol=1)
-
-```
-```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
-allen_priors <- ddply(allen_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100) 
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-plot_allen_posteriors <- ggplot(allen_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-#  geom_line(data=allen_priors, aes(x=value, y=density), col="red") +  
-  facet_wrap(~ variable, scale="free", ncol=3)
-
-```
-```{r allen-output, dependson=c("posterior-mode", "allen-traces"), results=toggle}
-A <- allen_posteriors
-A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
-pardist <- acast(A, index ~ variable)
-bayes_coef <- apply(pardist,2, posterior.mode) 
-bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"], bayes_coef["theta"])) # parameters formatted for f
-allen_f <- function(x,h,p) unname(RickerAllee(x,h, unname(p[c("r0", "K", "theta")])))
-allen_means <- sapply(x_grid, f, 0, bayes_pars)
-bayes_pars
-head(pardist)
-
-```
-
-```{r ricker-model}
-K_prior_p <- c(0.01, 40.0)
-r0_prior_p <- c(0.01, 20.0)
-bugs.model <- 
-paste(sprintf(
-"model{
-  K    ~ dunif(%s, %s)
-  r0    ~ dunif(%s, %s)
-  stdQ ~ dunif(%s, %s)", 
-  K_prior_p[1], K_prior_p[2],
-  r0_prior_p[1], r0_prior_p[2],
-  stdQ_prior_p[1], stdQ_prior_p[2]),
-  "
-  iQ <- 1 / (stdQ * stdQ);
-  y[1] ~ dunif(0, 10)
-  for(t in 1:(N-1)){
-    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K) 
-    y[t+1] ~ dlnorm(mu[t], iQ) 
-  }
-}")
-writeLines(bugs.model, "ricker_process.bugs")
-
-```
-```{r ricker-priors, dependson="common-priors"}
-K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
-r0_prior <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
-par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
-                   r0 = r0_prior, stdQ = stdQ_prior)
-
-```
-
-```{r ricker_priors_xtable, include=TRUE, echo=FALSE, cache=FALSE}
-ricker_priors_xtable <- data.frame(
-  parameter = c("$r$", "$K$", "$\\sigma$"),
-  "lower bound" = c(r0_prior_p[1], K_prior_p[1], stdQ_prior_p[1]),
-  "upper bound" = c(r0_prior_p[2], K_prior_p[2], stdQ_prior_p[2]))
-```
-
-```{r ricker-mcmc, dependson="ricker-model", results=toggle}
-jags.params=c("K","r0", "stdQ")
-jags.inits <- function(){
-  list("K"= 10 * rlnorm(1,0,.5),
-       "r0"= rlnorm(1,0,.5),
-       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
-       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
-}
-set.seed(12345) 
-ricker_jags <- do.call(jags, 
-                       list(data=jags.data, inits=jags.inits, 
-                            jags.params, n.chains=n.chains, 
-                            n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
-                            model.file="ricker_process.bugs"))
-recompile(ricker_jags)
-ricker_jags <- do.call(autojags, 
-                       list(object=ricker_jags, n.update=n.update, 
-														n.iter=n.iter, n.thin = n.thin, 
-														progress.bar="none"))
-
-```
-```{r ricker-traces, dependson="ricker-mcmc"}
-tmp <- lapply(as.mcmc(ricker_jags), as.matrix) # strip classes the hard way...
-ricker_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
-names(ricker_posteriors) = c("index", "variable", "value", "chain")
-plot_ricker_traces <- ggplot(ricker_posteriors) + geom_line(aes(index, value)) + 
-  facet_wrap(~ variable, scale="free", ncol=1)
-```
-```{r ricker-posteriors, dependson=c("ricker-traces", "ricker-priors")}
-ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100) 
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-# plot posterior distributions
-plot_ricker_posteriors <- ggplot(ricker_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-#  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") +  # don't plot priors 
-  facet_wrap(~ variable, scale="free", ncol=2)
-```
-```{r ricker-output, dependson=c("posterior-mode", "ricker-traces"), results=toggle}
-A <- ricker_posteriors
-A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
-ricker_pardist <- acast(A, index ~ variable)
-bayes_coef <- apply(ricker_pardist,2, posterior.mode) 
-ricker_bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"]))
-ricker_f <- function(x,h,p){
-  sapply(x, function(x){ 
-    x <- pmax(0, x-h) 
-    pmax(0, x * exp(p["r0"] * (1 - x / p["K"] )) )
-  })
-}
-ricker_means <- sapply(x_grid, Ricker, 0, ricker_bayes_pars[c(1,2)])
-head(ricker_pardist)
-ricker_bayes_pars
-
-```
-
-```{r myers-model}
-r0_prior_p <- c(.0001, 10.0)
-theta_prior_p <- c(.0001, 10.0)
-K_prior_p <- c(.0001, 40.0)
-bugs.model <- 
-paste(sprintf(
-"model{
-  r0    ~ dunif(%s, %s)
-  theta    ~ dunif(%s, %s)
-  K    ~ dunif(%s, %s)
-  stdQ ~ dunif(%s, %s)", 
-  r0_prior_p[1], r0_prior_p[2],
-  theta_prior_p[1], theta_prior_p[2],
-  K_prior_p[1], K_prior_p[2],
-  stdQ_prior_p[1], stdQ_prior_p[2]),
-
-  "
-  iQ <- 1 / (stdQ * stdQ);
-
-  y[1] ~ dunif(0, 10)
-  for(t in 1:(N-1)){
-    mu[t] <- log(r0)  + theta * log(y[t]) - log(1 + pow(abs(y[t]), theta) / K)
-    y[t+1] ~ dlnorm(mu[t], iQ) 
-  }
-}")
-writeLines(bugs.model, "myers_process.bugs")
-
-```
-```{r myers-priors}
-K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
-r_prior     <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
-theta_prior <- function(x) dunif(x, theta_prior_p[1], theta_prior_p[2])
-par_priors <- list( deviance = function(x) 0 * x, K = K_prior,
-                    r0 = r_prior, theta = theta_prior, 
-                    stdQ = stdQ_prior)
-
-```
-
-```{r myers_prior_xtable, include=TRUE, echo=FALSE} 
-myers_priors_xtable <- data.frame(parameter = c("$r$", "$K$", "$\\theta$", "$\\sigma$"),
-           "lower bound" = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
-           "upper bound" = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2]))
-```
-
-```{r myers-mcmc, dependson="myers-model", results=toggle}
-jags.params=c("r0", "theta", "K", "stdQ")
-jags.inits <- function(){
-  list("r0"= 1 * rlnorm(1,0,.1), 
-       "K"=    10 * rlnorm(1,0,.1),
-       "theta" = 1 * rlnorm(1,0,.1),  
-       "stdQ"= sqrt(0.2) * rlnorm(1,0,.1),
-       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
-}
-set.seed(12345)
-myers_jags <- do.call(jags, 
-                      list(data=jags.data, inits=jags.inits, 
-													 jags.params, n.chains=n.chains, 
-													 n.iter=n.iter, n.thin=n.thin,
-                           n.burnin=n.burnin, 
-                           model.file="myers_process.bugs"))
-recompile(myers_jags)
-myers_jags <- do.call(autojags, 
-                      list(myers_jags, n.update=n.update, 
-                           n.iter=n.iter, n.thin = n.thin, 
-                           progress.bar="none"))
-
-```
-```{r myers-traces, dependson="myers-mcmc"}
-tmp <- lapply(as.mcmc(myers_jags), as.matrix) # strip classes
-myers_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
-names(myers_posteriors) = c("index", "variable", "value", "chain")
-plot_myers_traces <- ggplot(myers_posteriors) + 
-  geom_line(aes(index, value)) + # priors, need to fix order though
-  facet_wrap(~ variable, scale="free", ncol=1)
-
-```
-```{r myers-posteriors, dependson="myers-traces"}
-par_prior_curves <- ddply(myers_posteriors, "variable", function(dd){
-    grid <- seq(min(dd$value), max(dd$value), length = 100) 
-    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
-})
-plot_myers_posteriors <- ggplot(myers_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-#  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") +  # Whoops, these are misaligned. see table instead 
-  facet_wrap(~ variable, scale="free", ncol=3)
-
-```
-```{r myers-output, dependson=c("posterior-mode", "myers-traces"), results=toggle}
-A <- myers_posteriors
-A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
-myers_pardist <- acast(A, index ~ variable)
-bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
-myers_bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["theta"], bayes_coef["K"]))
-myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
-myers_f <- function(x,h,p) Myer_harvest(x, h, p[c("r0", "theta", "K")])
-head(myers_pardist)
-myers_bayes_pars
-
-```
-
-```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
-models <- data.frame(x=x_grid, 
-										 GP=tgp_dat$y, 
-										 True=true_means, 
-                     MLE=est_means, 
-										 Ricker=ricker_means, 
-                     Allen = allen_means,
-                     Myers = myers_means)
-models <- melt(models, id="x")
-# some labels
-names(models) <- c("x", "method", "value")
-# labels for the colorkey too
-model_names = c("GP", "True", "MLE", "Ricker", "Allen", "Myers")
-colorkey=cbPalette
-names(colorkey) = model_names 
-step_ahead <- function(x, f, p){
-  h = 0
-  x_predict <- sapply(x, f, h, p)
-  n <- length(x_predict) - 1
-  y <- c(x[1], x_predict[1:n])
-  y
-}
-```
-```{r gp-opt, dependson="gp-output"}
-matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid) 
-opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
-```
-```{r mle-opt, dependson="mle-output"}
-matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
-opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-matrices_estimated <- f_transition_matrix(est$f, est$p, x_grid, h_grid, est$sigma_g)
-opt_estimated <- value_iteration(matrices_estimated, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-```
-```{r allen-opt, dependson="allen-output"}
-matrices_allen <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
-opt_allen <- value_iteration(matrices_allen, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-```
-```{r ricker-opt, dependson="ricker-output"}
-matrices_ricker <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
-opt_ricker <- value_iteration(matrices_ricker, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-```
-```{r myers-opt, dependson="myers-output"}
-matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
-myers_alt <- value_iteration(matrices_myers, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
-
-```
-```{r assemble-opt, dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
-OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_ricker$D, Allen = opt_allen$D, Myers = myers_alt$D)
-colorkey=cbPalette
-names(colorkey) = names(OPT) 
-```
-```{r sims, dependson="assemble-opt"}
-sims <- lapply(OPT, function(D){
-  set.seed(1)
-  lapply(1:100, function(i) 
-    ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
-  )
-})
-# turn the list into a data.frame
-keep.as.columns <- names(sims[[1]][[1]])
-dat <- melt(sims, id=keep.as.columns)
-sims_data <- data.table(dat)
-setnames(sims_data, c("L1", "L2"), c("method", "reps")) 
-# Legend in original ordering please, not alphabetical: 
-sims_data$method = factor(sims_data$method, ordered=TRUE, levels=names(OPT))
-```
-
-```{r profits, dependson="sims", results=toggle}
-Profit <- sims_data[, sum(profit), by=c("reps", "method")]
-tmp <- dcast(Profit, reps ~ method)
-#tmp$Allen <- tmp[,"Allen"] + rnorm(dim(tmp)[1], 0, 1) # jitter for plotting
-tmp <- tmp / tmp[,"True"]
-tmp <- melt(tmp[2:dim(tmp)[2]])
-actual_over_optimal <-subset(tmp, variable != "True")
-```
-
-```{r dic_calc, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output"), include=FALSE, echo=FALSE}
-dic.dic <- function (x) sum(x$deviance) +  sum(x[[2]])
-recompile(allen_jags)
-allen_dic <- dic.dic(dic.samples(allen_jags$model, n.iter=1000, type="popt"))
-recompile(ricker_jags)
-ricker_dic <- dic.dic(dic.samples(ricker_jags$model, n.iter=1000, type="popt"))
-recompile(myers_jags)
-myers_dic <- dic.dic(dic.samples(myers_jags$model, n.iter=1000, type="popt"))
-dictable <- data.frame(Allen = allen_dic + 2*length(bayes_pars), 
-                       Ricker = ricker_dic + 2*length(ricker_bayes_pars),
-                       Myers = myers_dic + 2*length(myers_bayes_pars), 
-                       row.names = c("DIC"))
-```
-
-```{r sensitivity-calc}
-source("components/sensitivity.R")
-sensitivity_dat <- dat
-```
-
-```{r export-data}
-write.csv(sensitivity_dat, "sensitivity.csv")
-```
-
-Introduction
-============
-
-Decision making under uncertainty is a ubiquitous challenge of
-natural resource management and conservation.  Ecological dynamics
-are frequently complex and difficult to measure, making uncertainty
-in our understanding a persistent challenge to effective management.
-Decision-theoretic approaches provide a framework to determine the
-best sequence of actions in face of uncertainty, but only when that
-uncertainty can be meaningfully quantified [@Fischer2009].  Concerns
-about the potential sudden collapse of fisheries and other ecosystems
-[@Worm2006; @Scheffer2009] has increasingly emphasized the difficulties of
-formulating even qualitatively correct models of the underlying processes.
-
-We develop a novel approach to address these concerns in the context of fisheries; though
-the underlying challenges and methods are germane to many other
-conservation and resource management problems.  Economic value and
-ecological concern have made marine fisheries the crucible for much
-of the founding work [@Gordon1954; @Reed1979; @May1979; @Ludwig1982]
-on management under uncertainty.  Global declines [@Worm2006]
-and the controversy surrounding their interpretation [@Hilborn2007; @Worm2009] have made understanding these
-challenges all the more pressing.
-
-<!-- Uncertainty outside the data without the correct model has not been handled. -->
-<!-- __We don't have the model__ -->
-
-Uncertainty enters the decision-making process at many levels:
-intrinsic stochasticity in biological processes, measurements, and
-implementation of policy [_e.g._ @Reed1979; @Clark1986; @Roughgarden1996;
-@Sethi2005], parametric uncertainty [_e.g._ @Ludwig1982; @Hilborn1997;
-@McAllister1998; @Schapaugh2013], and model or structural uncertainty
-[_e.g._ @Williams2001; @Cressie2009;  @Athanassoglou2012].  Of these,
-structural uncertainty is generally the hardest to quantify. Typical approaches
-assume a weak notion of model uncertainty in which the correct model
-(or reasonable approximation) of the dynamics must be identified from
-among a handful of alternative models.  Here we consider an approach
-that addresses uncertainty at each of these levels without assuming the
-dynamics follow a particular (i.e. parametric) structure.
-
-Often we do not understand the system dynamics well
-enough to know if a model provides a good approximation over the relevant
-range of states and timescales (criteria that we loosely refer to as
-defining the "right" or "true" model.)  
-Typical solutions involve either model choice, model averaging, or
-introducing yet greater model complexity of which others may be special
-cases (model averaging being one such way to construct such a model)
-[@Williams2001; @Athanassoglou2012; @Cressie2009].  Even setting aside
-other computational and statistical concerns (e.g. [@Cressie2009]), these
-approaches do not address our second concern - representing uncertainty
-outside the observed data range.
-
-<!-- __We don't have the data where we need it__ -->
-<!-- What do we call this?  Extrapolation uncertainty?  Pathological Uncertainty? -->
+\documentclass[author-year, 12pt,review]{elsarticle} %review=doublespace preprint=single 5p=2 column
+%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
+\usepackage[hyphens]{url}
+\usepackage{lineno} % add 
+  \linenumbers % turns line numbering on 
+\bibliographystyle{elsarticle-harv}
+\biboptions{sort&compress} % For natbib
+\usepackage{graphicx}
+\usepackage{booktabs} % book-quality tables
+%% Redefines the elsarticle footer
+\makeatletter
+\def\ps@pprintTitle{%
+ \let\@oddhead\@empty
+ \let\@evenhead\@empty
+ \def\@oddfoot{\it \hfill\today}%
+ \let\@evenfoot\@oddfoot}
+\makeatother
+
+% A modified page layout
+\textwidth 6.75in
+\oddsidemargin -0.15in
+\evensidemargin -0.15in
+\textheight 9in
+\topmargin -0.5in
+%%%%%%%%%%%%%%%% end my additions to header
+
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \usepackage{fontspec}
+  \ifxetex
+    \usepackage{xltxtra,xunicode}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage{longtable}
+\usepackage{graphicx}
+% We will generate all images so they have a width \maxwidth. This means
+% that they will get their normal width if they fit onto the page, but
+% are scaled down if they would overflow the margins.
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
+\else\Gin@nat@width\fi}
+\makeatother
+\let\Oldincludegraphics\includegraphics
+\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={Avoiding tipping points in ecosystem based fisheries management: a non-parametric Bayesian approach},
+            colorlinks=true,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{0}
+% Pandoc toggle for numbering sections (defaults to be off)
+\setcounter{secnumdepth}{0}
+% Pandoc header
+
+\begin{document}
+\begin{frontmatter}
+
+  \title{Avoiding tipping points in ecosystem based fisheries management: a
+non-parametric Bayesian approach}
+    \author[cstar]{Carl Boettiger\corref{c1}}
+   \ead{cboettig@gmail.com} 
+   \cortext[c1]{Corresponding author}
+    \author[cstar]{Marc Mangel}
+
+    \author[noaa]{Stephan Munch}
+
+      \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and
+Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA
+95064, USA}    
+    \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric
+Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}
+
+  \begin{abstract}
+  Model uncertainty and limited data coverage are fundamental challenges
+  to robust ecosystem management. These challenges are acutely highlighted
+  by concerns that many ecological systems may contain tipping points.
+  Before a collapse, we do not know where the tipping points lie, if they
+  exist at all. Hence, we know neither a complete model of the system
+  dynamics nor do we have access to data in some large region of
+  state-space where such a tipping point might exist. These two sources of
+  uncertainty frustrate state-of-the-art parametric approaches to decision
+  theory and optimal control. Here we illustrate how a non-parametric
+  approach using a Gaussian process prior provides a more flexible
+  representation of this inherent uncertainty. We can then embed the
+  Gaussian process in a stochastic dynamic programming framework in order
+  to make robust management predictions under both model uncertainty and
+  limited data. We conduct a set of simulations to evaluate this approach
+  as compared with the standard approach of using model selection to
+  choose from a set of candidate models. In keeping with previous results,
+  we find that model selection erroneously favors models without tipping
+  points -- leading to harvest policies that guarantee extinction. The
+  Gaussian process approach performs nearly as well as the true model and
+  significantly outperforms standard appraoches. We illustrate this using
+  examples of simulated single-species dynamics, where the standard model
+  selection approach should be most effective, and find that it still
+  fails to account for uncertainty appropriately and leads to population
+  crashes, while management based on the Gaussian process approach does
+  not. We discuss opportunities to adapt this approach to multiple species
+  models in further work.
+  \end{abstract}
+   \begin{keyword} Bayesian \sep Structural Uncertainty \sep Nonparametric \sep Optimal Control \sep Decision Theory \sep Gaussian Processes \sep Fisheries Management \sep \end{keyword}
+ \end{frontmatter}
+
+\section{Introduction}\label{introduction}
+
+Decision making under uncertainty is a ubiquitous challenge of natural
+resource management and conservation. Ecological dynamics are frequently
+complex and difficult to measure, making uncertainty in our
+understanding a persistent challenge to effective management.
+Decision-theoretic approaches provide a framework to determine the best
+sequence of actions in face of uncertainty, but only when that
+uncertainty can be meaningfully quantified (Fischer \emph{et al.} 2009).
+Concerns about the potential sudden collapse of fisheries and other
+ecosystems (Worm \emph{et al.} 2006; Scheffer \emph{et al.} 2009) has
+increasingly emphasized the difficulties of formulating even
+qualitatively correct models of the underlying processes.
+
+We develop a novel approach to address these concerns in the context of
+fisheries; though the underlying challenges and methods are germane to
+many other conservation and resource management problems. Economic value
+and ecological concern have made marine fisheries the crucible for much
+of the founding work (Gordon 1954; May \emph{et al.} 1979; Reed 1979;
+Ludwig \& Walters 1982) on management under uncertainty. Global declines
+(Worm \emph{et al.} 2006) and the controversy surrounding their
+interpretation (Hilborn 2007; Worm \emph{et al.} 2009) have made
+understanding these challenges all the more pressing.
+
+Uncertainty enters the decision-making process at many levels: intrinsic
+stochasticity in biological processes, measurements, and implementation
+of policy (\emph{e.g.} Reed 1979; Clark \& Kirkwood 1986; Roughgarden \&
+Smith 1996; Sethi \emph{et al.} 2005), parametric uncertainty
+(\emph{e.g.} Ludwig \& Walters 1982; Hilborn \& Mangel 1997; McAllister
+1998; Schapaugh \& Tyre 2013), and model or structural uncertainty
+(\emph{e.g.} Williams 2001; Cressie \emph{et al.} 2009; Athanassoglou \&
+Xepapadeas 2012). Of these, structural uncertainty is generally the
+hardest to quantify. Typical approaches assume a weak notion of model
+uncertainty in which the correct model (or reasonable approximation) of
+the dynamics must be identified from among a handful of alternative
+models. Here we consider an approach that addresses uncertainty at each
+of these levels without assuming the dynamics follow a particular
+(i.e.~parametric) structure.
+
+Often we do not understand the system dynamics well enough to know if a
+model provides a good approximation over the relevant range of states
+and timescales (criteria that we loosely refer to as defining the
+``right'' or ``true'' model.)\\Typical solutions involve either model
+choice, model averaging, or introducing yet greater model complexity of
+which others may be special cases (model averaging being one such way to
+construct such a model) (Williams 2001; Cressie \emph{et al.} 2009;
+Athanassoglou \& Xepapadeas 2012). Even setting aside other
+computational and statistical concerns (e.g. (Cressie \emph{et al.}
+2009)), these approaches do not address our second concern -
+representing uncertainty outside the observed data range.
 
@@
 action space typically involves considering actions that may move the
-system outside the range of observed behavior, and (b) decision-theoretic
-algorithms rely not only on reasonable estimates of the expected
-outcomes, but depend on the weights given to all possible outcomes
-[_e.g._ @Weitzman2013].
+system outside the range of observed behavior, and (b)
+decision-theoretic algorithms rely not only on reasonable estimates of
+the expected outcomes, but depend on the weights given to all possible
+outcomes (\emph{e.g.} Weitzman 2013).
 
@@
 covers the full range of possible states is unfortunately the rule more
-than the exception. The potential concern of tipping points in ecological
-dynamics [@Scheffer2001; @Polasky2011] reflects these concerns -- as
-either knowledge of the true model or more complete sampling of the
-state space would make it easy to identify if a tipping point existed.
-If we do not know but cannot rule out such a possibility, then we
-face decision-making under this dual challege of model uncertainty and
-incomplete data coverage.
-
-These dual concerns pose a substantial challenge to existing decision-theoretic
-approaches [@Brozovic2011].  Because intervention is often too late 
-after a tipping point has been crossed (but see @Hughes2013), management
-is most often concerned with avoiding potentially catastrophic tipping
-points before any data is available at or following a transition that
-would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].
-
-<!--
-_Map_
--->
+than the exception. The potential concern of tipping points in
+ecological dynamics (Scheffer \emph{et al.} 2001; Polasky \emph{et al.}
+2011) reflects these concerns -- as either knowledge of the true model
+or more complete sampling of the state space would make it easy to
+identify if a tipping point existed. If we do not know but cannot rule
+out such a possibility, then we face decision-making under this dual
+challege of model uncertainty and incomplete data coverage.
+
+These dual concerns pose a substantial challenge to existing
+decision-theoretic approaches (Brozović \& Schlenker 2011). Because
+intervention is often too late after a tipping point has been crossed
+(but see Hughes et al. (2013)), management is most often concerned with
+avoiding potentially catastrophic tipping points before any data is
+available at or following a transition that would more clearly reveal
+these regime shift dynamics (e.g. Bestelmeyer \emph{et al.} 2012).
 
 Here we illustrate how a stochastic dynamic programming (SDP) algorithm
-[@Mangel1988; @Marescot2013] can be driven by the
-predictions from a Bayesian non-parametric (BNP) approach [@Munch2005a].
-This provides two distinct advantages compared with contemporary
-approaches.  First, using a BNP sidesteps the need for an accurate
-model-based description of the system dynamics.  Second, the BNP can
-better reflect uncertainty that arises when extrapolating a model outside
-of the data on which it was fit.  We illustrate that when the correct
-model is not known, this latter feature is crucial to providing a robust
-decision-theoretic approach in face of substantial structural uncertainty.
+(Mangel \& Clark 1988; Marescot \emph{et al.} 2013) can be driven by the
+predictions from a Bayesian non-parametric (BNP) approach (Munch
+\emph{et al.} 2005a). This provides two distinct advantages compared
+with contemporary approaches. First, using a BNP sidesteps the need for
+an accurate model-based description of the system dynamics. Second, the
+BNP can better reflect uncertainty that arises when extrapolating a
+model outside of the data on which it was fit. We illustrate that when
+the correct model is not known, this latter feature is crucial to
+providing a robust decision-theoretic approach in face of substantial
+structural uncertainty.
 
@@
 uncertainty in parameter estimates, the BNP approach provides a more
-state-space dependent representation of uncertainty.  This permits
-a much greater uncertainty far from the observed data than near the
-observed data.  These features allow the GPDP approach to find robust
-management solutions in face of limited data and without knowledge of
-the correct model structure.
-
-<!--
--  _Note on "not magic"_: honest uncertainty + SDP
--->
+state-space dependent representation of uncertainty. This permits a much
+greater uncertainty far from the observed data than near the observed
+data. These features allow the GPDP approach to find robust management
+solutions in face of limited data and without knowledge of the correct
+model structure.
 
@@
 know the model or have particularly good data should immediately draw
-suspicion.  The reader must bear in mind that the strength of our approach
-comes not from black-box predictive power from such limited information,
-but rather, by providing a more honest expression of uncertainty outside
-the observed data without sacrificing the predictive capacity near the
-observed data. By coupling this more accurate description of what is known
-and unknown to the decision-making under uncertainty framework provided
-by stochastic dynamic programming, we are able to obtain more robust
-management policies than with common parametric modeling approaches.
-
-<!--move to  much later *Or do we need this at all?* 
-- _Note on comparing models_ (via value function rather than by "fit"). 
--->
+suspicion. The reader must bear in mind that the strength of our
+approach comes not from black-box predictive power from such limited
+information, but rather, by providing a more honest expression of
+uncertainty outside the observed data without sacrificing the predictive
+capacity near the observed data. By coupling this more accurate
+description of what is known and unknown to the decision-making under
+uncertainty framework provided by stochastic dynamic programming, we are
+able to obtain more robust management policies than with common
+parametric modeling approaches.
 
 The nature of decision-making problems provides a compelling and
-pragmatic way to compare models: Rather than compare models in terms
-of best fit to data, model performance is defined in the concrete terms
-of the decision-maker's objectives. Although much argument can be made
-over the 'correct' objective function, e.g. how to account for the
-social value of fish left in the sea vs. the commercial value of fish
-harvested; see @Halpern2013 for further discussion of this issue.  We can
-always compare model performance across multiple potential objectives.
-The important point is that a decision-maker does not necessarily need
-a model that provides the best mechanistic understanding or the best
-long-term outcome, but rather the one that idelivers the best management
-performance over the period of interest.
-
-## Background on the Gaussian Process
-
-<!-- 
-- Background on non-parametric modeling.  
--->
-
-<!--
-- Definition
-- Previous application
--->
+pragmatic way to compare models: Rather than compare models in terms of
+best fit to data, model performance is defined in the concrete terms of
+the decision-maker's objectives. Although much argument can be made over
+the `correct' objective function, e.g.~how to account for the social
+value of fish left in the sea vs.~the commercial value of fish
+harvested; see Halpern et al. (2013) for further discussion of this
+issue. We can always compare model performance across multiple potential
+objectives. The important point is that a decision-maker does not
+necessarily need a model that provides the best mechanistic
+understanding or the best long-term outcome, but rather the one that
+idelivers the best management performance over the period of interest.
+
+\subsection{Background on the Gaussian
+Process}\label{background-on-the-gaussian-process}
 
@@
 process priors. The use of Gaussian process (GP) regression (or
-"kriging" in the geospatial literature) to formulate a predictive
+``kriging'' in the geospatial literature) to formulate a predictive
 model is relatively new in the context of modeling dynamical systems
-[@Kocijan2005], and was first introduced in the context ecological
-modeling and fisheries management in @Munch2005.  GP models have
-subsequently been used to test for the presence of Allee effects
-[@Sugeno2013a], estimate the maximum reproductive rate [@Sugeno2013b],
-determine temporal variation in food availability [@Sigourney2012], and
-provide a basis for identifying model-misspecification [@Thorson2014].
-An accessible and thorough introduction to the formulation and use of
-GPs can be found in @Rasmussen2006. The nonparametric moniker for the
-GP approach is unfortunate as (a) it still involves the estimation of
-parameters, and (b), statisticians also use non-parametric to mean models
-that make only weak distributional assumptions  [e.g. @Lehmann1975].
-However, we retain the terminology for consistency with previous studies.
-
-Approach and Methods
-====================
-
-### Statement of the optimal control problem
+(Kocijan \emph{et al.} 2005), and was first introduced in the context
+ecological modeling and fisheries management in Munch et al. (2005b). GP
+models have subsequently been used to test for the presence of Allee
+effects (Sugeno \& Munch 2013a), estimate the maximum reproductive rate
+(Sugeno \& Munch 2013b), determine temporal variation in food
+availability (Sigourney \emph{et al.} 2012), and provide a basis for
+identifying model-misspecification (Thorson \emph{et al.} 2014). An
+accessible and thorough introduction to the formulation and use of GPs
+can be found in Rasmussen \& Williams (2006). The nonparametric moniker
+for the GP approach is unfortunate as (a) it still involves the
+estimation of parameters, and (b), statisticians also use non-parametric
+to mean models that make only weak distributional assumptions (e.g.
+Lehmann 1975). However, we retain the terminology for consistency with
+previous studies.
+
+\section{Approach and Methods}\label{approach-and-methods}
+
+\subsubsection{Statement of the optimal control
+problem}\label{statement-of-the-optimal-control-problem}
 
@@
 programming approach (hereafter GPDP) we focus on the classical problem
-of determining a harvest policy that will maximize the expected discounted
-profit over a fixed time interval [@Reed1979; @Ludwig1982; @Mangel1988].
-For simplicity we assume profit is proportional to the realized harvest.  
-For comparison, we consider the performance of the GPDP policy against
-policies derived under several alternative parametric models.
-
-We focus on a class of one-dimensional models of population dynamics as a
-'best-case' scenario for the parametric model performance relative to
+of determining a harvest policy that will maximize the expected
+discounted profit over a fixed time interval (Reed 1979; Ludwig \&
+Walters 1982; Mangel \& Clark 1988). For simplicity we assume profit is
+proportional to the realized harvest.\\For comparison, we consider the
+performance of the GPDP policy against policies derived under several
+alternative parametric models.
+
+We focus on a class of one-dimensional models of population dynamics as
+a `best-case' scenario for the parametric model performance relative to
 the Gaussian process. We consider this a best-case scenario for the
@@
 from one the parametric models in the set, whereas in reality we never
-have the "true" model among our candidates.  By choosing one-dimensional
-models with few parameters, we limit the chance that poor performance
-will be due to our inability to estimate parameters well, something
-that becomes a more severe problem for higher-dimensional parametric
-models. Further, the parametric models we consider are those most commonly
-used in modeling stock-recruitment dynamics or to model sudden transitions
-between alternative stable states.
-
-### Parametric models
-
-<!-- 
-- Statement of the models
--->
+have the ``true'' model among our candidates. By choosing
+one-dimensional models with few parameters, we limit the chance that
+poor performance will be due to our inability to estimate parameters
+well, something that becomes a more severe problem for
+higher-dimensional parametric models. Further, the parametric models we
+consider are those most commonly used in modeling stock-recruitment
+dynamics or to model sudden transitions between alternative stable
+states.
+
+\subsubsection{Parametric models}\label{parametric-models}
 
 We consider three candidate parametric models of the stock-recruitment
-dynamics: The Ricker model, the Allen model [@Allen2005], the Myers
-model [@Myers1995], equations (@ricker)-(@myers), which we parameterize as follows:
+dynamics: The Ricker model, the Allen model (Allen \& Tanner 2005), the
+Myers model (Myers \emph{et al.} 1995), equations (1)-(3), which we
+parameterize as follows:
 
@@
 
-$$ X_{t+1}= Z_t f(S_t,p) $$
-
-Where Z(t) is log-normally distributed process stochasticity [@Reed1979]
-and $p$ is a vector of parameters, and $S_t$ the escapement [@Reed1979]
-
-$$S_t = X_t - h_t$$
+\[ X_{t+1}= Z_t f(S_t,p) \]
+
+Where Z(t) is log-normally distributed process stochasticity (Reed 1979)
+and $p$ is a vector of parameters, and $S_t$ the escapement (Reed 1979)
+
+\[S_t = X_t - h_t\]
 
@@
 
-(@ricker) $$ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
-
-where $K$ is the carrying capacity and $r$ the growth rate.  For the Allen model,
-
-(@allen) $$ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_C\right)} $$
-
-Where $r$ and $K$ are as before, and $X_c$ denotes the location of the unstable (tipping) point.  For the Myers model,
-
-(@myers) $$ f(S(t) | p)  = \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}}$$
-
-Where $r$ and $K$ as before, and $\theta$ the nonlinearity of the Myers model (corresponding to Beverton-Holt dynamics for $\theta$ = 1 and an Allee effect with alternative stable states exist at $\theta > 2$).
-
-The familiar Ricker model involves two parameters, corresponding
-to a growth rate and a carrying capacity, and cannot support alternative
+\begin{enumerate}
+\def\labelenumi{(\arabic{enumi})}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \[ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
+\end{enumerate}
+
+where $K$ is the carrying capacity and $r$ the growth rate. For the
+Allen model,
+
+\begin{enumerate}
+\def\labelenumi{(\arabic{enumi})}
+\setcounter{enumi}{1}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \[ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_C\right)} \]
+\end{enumerate}
+
+Where $r$ and $K$ are as before, and $X_c$ denotes the location of the
+unstable (tipping) point. For the Myers model,
+
+\begin{enumerate}
+\def\labelenumi{(\arabic{enumi})}
+\setcounter{enumi}{2}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \[ f(S(t) | p)  = \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}}\]
+\end{enumerate}
+
+Where $r$ and $K$ as before, and $\theta$ the nonlinearity of the Myers
+model (corresponding to Beverton-Holt dynamics for $\theta$ = 1 and an
+Allee effect with alternative stable states exist at $\theta > 2$).
+
+The familiar Ricker model involves two parameters, corresponding to a
+growth rate and a carrying capacity, and cannot support alternative
 stable growth rate dynamics. The Allen model resembles the Ricker
-dynamics with an added Allee effect parameter [@Courchamp2008], below
-which the population cannot persist.  The Myers model also has three
-parameters and contains an Allee threshold, but has compensatory rather
-than over-compensatory density dependence (resembling a Beverton-Holt
-curve rather than a Ricker curve at high densities.)  
-We assume multiplicative log-normal noise perturbs the growth predicted 
-by the each of the deterministic model skeletons described above. This 
-introduces one additional parameter $\sigma$ that must be estimated by each
-model.
+dynamics with an added Allee effect parameter (Courchamp \emph{et al.}
+2008), below which the population cannot persist. The Myers model also
+has three parameters and contains an Allee threshold, but has
+compensatory rather than over-compensatory density dependence
+(resembling a Beverton-Holt curve rather than a Ricker curve at high
+densities.)\\We assume multiplicative log-normal noise perturbs the
+growth predicted by the each of the deterministic model skeletons
+described above. This introduces one additional parameter $\sigma$ that
+must be estimated by each model.
 
 Because of our interest in management performance in the presence of
-tipping points, all of our simulations are based on the Allen model.
-The Allen model is thus structurally correct and is expected to provide
-a best-case scenario when the 'true' dynamics are known.
-The Ricker model is a reasonable approximation of these dynamics far from the Allee
+tipping points, all of our simulations are based on the Allen model. The
+Allen model is thus structurally correct and is expected to provide a
+best-case scenario when the `true' dynamics are known. The Ricker model
+is a reasonable approximation of these dynamics far from the Allee
 threshold (but lacks threshold dynamics), while the Myers model shares
@@
 
-We consider a period of `r Tobs` years of training data: long
-enough that the estimates are not dependent on the particular realization,
-while longer times are not likely to provide substantial improvement
-(i.e. the results are not sensitive to this interval).  Each of the models
-(described below) is fit to the same training data, as shown in Figure 1.
-
-<!-- Bayesian inference of parametric models -->
+We consider a period of 40 years of training data: long enough that the
+estimates are not dependent on the particular realization, while longer
+times are not likely to provide substantial improvement (i.e.~the
+results are not sensitive to this interval). Each of the models
+(described below) is fit to the same training data, as shown in Figure
+1.
 
 We inferred posterior distributions for the parameters of each model in
-a Bayesian context using Gibbs sampling (@Gelman2003 implemented in R
-[@RTeam] using jags, [@R2jags]).  We choose uninformative uniform priors
-for all parameters (See appendix Tables S1-S3; R code provided). We
-show one-step-ahead predictions of these model fits in Figure 1.
-Each chain was tested for Gelman-Rubin convergence and results were
-robust to longer runs.  For each simulation we also applied several
-commonly used model selection criteria (AIC, BIC, DIC, see @Burnham2002)
-to identify the best fitting model.
-
-<!-- SDP via parametric models -->
+a Bayesian context using Gibbs sampling (Gelman et al. (2003)
+implemented in R (R Core Team 2013) using jags, (Su \& Masanao Yajima
+2013)). We choose uninformative uniform priors for all parameters (See
+appendix Tables S1-S3; R code provided). We show one-step-ahead
+predictions of these model fits in Figure 1. Each chain was tested for
+Gelman-Rubin convergence and results were robust to longer runs. For
+each simulation we also applied several commonly used model selection
+criteria (AIC, BIC, DIC, see Burnham \& Anderson (2002)) to identify the
+best fitting model.
 
@@
 
-### The Gaussian Process model
-
-<!-- Statement of model -->
+\subsubsection{The Gaussian Process
+model}\label{the-gaussian-process-model}
 
@@
 defined by a radial basis function kernel of two parameters: $\ell$,
-which gives the characteristic length-scale over which correlation between
-two points in state-space decays, and $\sigma$, which gives the scale
-of the process noise by which observations $Y_{t+1}$ may differ from
-their predicted values $X_{t+1}$ given an observation of the previous
-state, $X_t$. @Munch2005a gives an accessible introduction to the use
-of Gaussian processes in providing a Bayesian nonparametric description
-of the stock-recruitment relationship.
-
-<!-- Inference of the BNP model -->
-
-We use a Metropolis-Hastings Markov Chain Monte Carlo (@Gelman2003) to
-infer posterior distributions of the two parameters of the GP (Figure
-S13, code in appendix), under weakly informative Gaussian priors
+which gives the characteristic length-scale over which correlation
+between two points in state-space decays, and $\sigma$, which gives the
+scale of the process noise by which observations $Y_{t+1}$ may differ
+from their predicted values $X_{t+1}$ given an observation of the
+previous state, $X_t$. Munch et al. (2005a) gives an accessible
+introduction to the use of Gaussian processes in providing a Bayesian
+nonparametric description of the stock-recruitment relationship.
+
+We use a Metropolis-Hastings Markov Chain Monte Carlo (Gelman et al.
+(2003)) to infer posterior distributions of the two parameters of the GP
+(Figure S13, code in appendix), under weakly informative Gaussian priors
 (see parameters in table S5). As the posterior distributions differ
@@
 prior belief.
-
-<!-- SDP via the model -->
 
@@
 
-Results
-=======
-
-To ensure our results are robust to our choice of parameters, 
-we perform our analysis over 96 different scenarios. To better
-understand the process, we will first describe in detail the
-results of a single scenario.
-
-<!-- 
-### Figure 1: Fitted Models
-
-- All models fit the data quite well
-- Information criteria would pick the simple, incorrect model.
--->
-
-```{r figure_1, dependson=c("assemble-models", "par-fns", "plot-options"), fig.cap="Points show the training data of stock-size over time.
-Curves show the posterior step-ahead predictions based on each of the estimated models.
-Observe that all models are fitting the data reasonably well.", fig.width=8, fig.height=6}
-step_ahead_posteriors <- function(x){
-  gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
-  df_post <- melt(lapply(sample(100, 30),  
-  function(i){
-    data.frame(time = 1:length(x), stock = x, 
-                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
-                True = step_ahead(x,f,p),  
-                MLE = step_ahead(x,f,est$p), 
-                Allen = step_ahead(x, allen_f, pardist[i,]), 
-                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
-                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
-  }), id=c("time", "stock"))
-}
-df_post <- step_ahead_posteriors(x)
-figure1b_posteriors <- ggplot(df_post) + geom_point(aes(time, stock)) + 
-  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
-  facet_wrap(~variable) + 
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) +  
-  theme(legend.position="none")
-figure1b_posteriors
-
-```
-
-All models fit the observed data rather closely and with relatively small uncertainty, as illustrated in the posterior predictive curves in Figure 1.
-Figure 1 shows the training data of stock sizes observed over time as points, overlaid with the step-ahead predictions of each estimated model using the parameters sampled from their posterior distributions.
-Each model manages to fit the observed data rather closely.
-Compared to the true model most estimates appear to over-fit, predicting fluctuations that are actually due purely to stochasticity.
-All model selection criteria shown in Table 1 penalize more complex models and show a preference for the simpler Ricker model over the more complicated alternate stable state models (Allen and Myers).
-Details on MCMC estimates for each model, traces, and posterior distributions can be found in the appendix.
-
-```{r deviances, dependson=c("dic_calc"), include=FALSE, echo=FALSE}
-allen_deviance  <- - posterior.mode(pardist[,'deviance'])
-ricker_deviance <- - posterior.mode(ricker_pardist[,'deviance'])
-myers_deviance  <- - posterior.mode(myers_pardist[,'deviance'])
-true_deviance   <- 2*estf(c(p, sigma_g))
-mle_deviance    <- 2*estf(c(est$p, est$sigma_g))
-aictable <- data.frame(Allen = allen_deviance + 2*(1+length(bayes_pars)),  # +1 for noise parameter
-                       Ricker = ricker_deviance + 2*(1+length(ricker_bayes_pars)),
-                       Myers = myers_deviance + 2*(1+length(myers_bayes_pars)), 
-                       row.names = c("AIC"))
-bictable <- data.frame(Allen = allen_deviance + log(length(x))*(1+length(bayes_pars)), 
-                       Ricker = ricker_deviance + log(length(x))*(1+length(ricker_bayes_pars)),
-                       Myers = myers_deviance + log(length(x))*(1+length(myers_bayes_pars)), 
-                       row.names = c("BIC"))
-xtable::xtable(rbind(dictable, aictable, bictable))
-```
-
-```{r Table1, dependson="deviances", include=TRUE, results="asis", echo=FALSE, cache=FALSE}
-xtable::xtable(rbind(dictable, aictable, bictable), caption="Model selection scores for several common criteria all (wrongly) select the simplest model.
+\section{Results}\label{results}
+
+To ensure our results are robust to our choice of parameters, we perform
+our analysis over 96 different scenarios. To better understand the
+process, we will first describe in detail the results of a single
+scenario.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_1.pdf}
+\caption{Points show the training data of stock-size over time. Curves
+show the posterior step-ahead predictions based on each of the estimated
+models. Observe that all models are fitting the data reasonably well.}
+\end{figure}
+
+All models fit the observed data rather closely and with relatively
+small uncertainty, as illustrated in the posterior predictive curves in
+Figure 1. Figure 1 shows the training data of stock sizes observed over
+time as points, overlaid with the step-ahead predictions of each
+estimated model using the parameters sampled from their posterior
+distributions. Each model manages to fit the observed data rather
+closely. Compared to the true model most estimates appear to over-fit,
+predicting fluctuations that are actually due purely to stochasticity.
+All model selection criteria shown in Table 1 penalize more complex
+models and show a preference for the simpler Ricker model over the more
+complicated alternate stable state models (Allen and Myers). Details on
+MCMC estimates for each model, traces, and posterior distributions can
+be found in the appendix.
+
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{rrrr}
+  \hline
+ & Allen & Ricker & Myers \\ 
+  \hline
+DIC & 50.14 & 49.45 & 50.61 \\ 
+  AIC & -24.60 & -30.07 & -27.19 \\ 
+  BIC & -17.85 & -25.00 & -20.44 \\ 
+   \hline
+\end{tabular}
+\caption{Model selection scores for several common criteria all (wrongly) select the simplest model.
 As the true (Allen) model is not distinguishable from the simpler (Ricker) model in the region of the observed data, this error cannot be avoided regardless of the model choice criterion.
-This highlights the danger of model choice when the selected model will be used outside of the observed range of the data.")
-```
-
-<!-- 
-### Figure 2 
-- Data comes from limited region of state-space 
-- (Should really show uncertainty of all models here.  Capture the forecast uncertainty several steps down the road?)   
--->
-
-```{r figure_2, dependson=c("assemble-models", "plot-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
-Graph shows the expected value for the function $f$ under each model.
-Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.
-The training data is also shown as black points.
-The GP is conditioned on (0,0), shown as a pseudo-data point.", fig.width=8, fig.height=6}
-x_grid_short <- x_grid[1:40]
-gp_short <- gp_predict(gp, x_grid_short, burnin=1e4, thin=300)
-models_posteriors <- 
-  melt(lapply(sample(100, 50), 
-              function(i){
-    sample_gp <- mvrnorm(1, 
-                            gp_short$Ef_posterior[,i],         
-                            gp_short$Cf_posterior[[i]])
-    data.frame(stock = x_grid_short, 
-               GP = sample_gp,
-               y = sample_gp,
-               ymin = sample_gp - 2 * sqrt(gp_short$E_Vf), 
-               ymax = sample_gp + 2 * sqrt(gp_short$E_Vf), 
-               True = sapply(x_grid_short,f,0, p),  
-               MLE = sapply(x_grid_short,f,0, est$p), 
-               Allen = sapply(x_grid_short, allen_f, 0, pardist[i,]), 
-               Ricker = sapply(x_grid_short, ricker_f, 0, ricker_pardist[i,]), 
-               Myers = sapply(x_grid_short, myers_f, 0, myers_pardist[i,]))
-             }), 
-       id=c("stock", "y", "ymin", "ymax"))
-ggplot(models_posteriors) + 
-    geom_ribbon(aes(x=stock, y=y, ymin=ymin, ymax=ymax, group=L1), 
-                  fill = "gray80", 
-                  data=subset(models_posteriors, variable == "GP")) + 
-    geom_line(aes(stock, value, col = variable, 
-                  group=interaction(L1,variable)), 
-              alpha=.2) + 
-    geom_point(data = obs, aes(x,y), alpha = 0.8) + 
-    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
-    facet_wrap(~variable) + 
-    scale_colour_manual(values=colorkey) +  
-    theme(legend.position="none")
-```
+This highlights the danger of model choice when the selected model will be used outside of the observed range of the data.}
+\end{center}
+\end{table}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_2.pdf}
+\caption{Graph of the inferred Gaussian process compared to the true
+process and maximum-likelihood estimated process. Graph shows the
+expected value for the function $f$ under each model. Two standard
+deviations from the estimated Gaussian process covariance with (light
+grey) and without (darker grey) measurement error are also shown. The
+training data is also shown as black points. The GP is conditioned on
+(0,0), shown as a pseudo-data point.}
+\end{figure}
 
@@
 true model used to generate the data is shown in Figure 2, predicting
-the relationship between observed stock size (x-axis) to the stock
-size after recruitment the following year.  Note that in contrast to
-the  other models shown, the expected Gaussian process corresponds to a
+the relationship between observed stock size (x-axis) to the stock size
+after recruitment the following year. Note that in contrast to the other
+models shown, the expected Gaussian process corresponds to a
 distribution of curves - as indicated by the gray band - which itself
-has a mean shown in black. Parameter uncertainty (not shown) spreads
-out the estimates further.  The observed data from which each model is
-estimated is also shown.  The observations come from only a limited region
-of state space corresponding to unharvested or weakly harvested system.
-No observations occur at the theoretical optimum harvest rate or near
-the tipping point.
-
-```{r figure_3, fig.width=8, fig.height=6, fig.cap="Beyond the tipping point (outside the range of the training data), the true dynamics fall outside the uncertainty bands of the wrong parametric models, but inside the uncertainty predicted by the GP.
-Points show the stock size simulated by the true model.
-Overlay shows the range of states predicted by each model, based on the state observed in the previous time step.
-The Ricker model always (wrongly) predicts positive population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).
-Note that the GP is both more pessimistic and more uncertain about the future state than the parametric models.", dependson="plot-options"}
-y <- numeric(8)
-y[1] <- 4.5
-for(t in 1:(length(y)-1))
-      y[t+1] = z_g() * f(y[t], h=0, p=p)
-# predicts means, does not reflect uncertainty estimate!
-crash_data <- step_ahead_posteriors(y)
-crash_data <- subset(crash_data, variable %in% c("GP", "Allen", "Ricker", "Myers"))
-ggplot(crash_data) + 
-  geom_boxplot(aes(as.factor(as.integer(time)), value, 
-                   fill = variable, col=variable), 
-               alpha=.7, outlier.size=1, position="identity") + 
-#  geom_line(aes(time, value, col = variable, 
-#            group=interaction(L1,variable)), alpha=.1) + 
-  geom_point(aes(time, stock), size = 3) + 
-  scale_fill_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")], 
-                      guide = guide_legend(override.aes = list(alpha = 1))) +  
-  scale_colour_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")], 
-                      guide = guide_legend(override.aes = list(alpha = 1))) +  
-  facet_wrap(~variable) + 
-  theme(legend.position="none") + xlab("time") + ylab("stock size") 
-```
-
-<!--
-### Figure 4: Inferred Policies
-
-- Inferred policies differ substantially among models
-- The structurally correct model and the GP are close to the true model
-- alternatives are not close
--->
-
-```{r figure_4, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
-policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
-names(policies) <- c("stock", "method", "value")
-
-ggplot(policies, aes(stock, stock - value, color=method)) +
-  geom_line(lwd=1.2, alpha=0.8) + xlab("stock size") + ylab("escapement")  +
-  scale_colour_manual(values=colorkey)
-```
+has a mean shown in black. Parameter uncertainty (not shown) spreads out
+the estimates further. The observed data from which each model is
+estimated is also shown. The observations come from only a limited
+region of state space corresponding to unharvested or weakly harvested
+system. No observations occur at the theoretical optimum harvest rate or
+near the tipping point.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_3.pdf}
+\caption{Beyond the tipping point (outside the range of the training
+data), the true dynamics fall outside the uncertainty bands of the wrong
+parametric models, but inside the uncertainty predicted by the GP.
+Points show the stock size simulated by the true model. Overlay shows
+the range of states predicted by each model, based on the state observed
+in the previous time step. The Ricker model always (wrongly) predicts
+positive population growth, while the actual population shrinks in each
+step as the initial condition falls below the Allee threshold of the
+underlying model (Allen). Note that the GP is both more pessimistic and
+more uncertain about the future state than the parametric models.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_4.pdf}
+\caption{The steady-state optimal policy (infinite boundary) calculated
+under each model. Policies are shown in terms of target escapement,
+$S_t$, as under models such as this a constant escapement policy is
+expected to be optimal (Reed 1979).}
+\end{figure}
 
@@
 such as this a constant escapement policy is expected to be optimal
-[@Reed1979], whereby population levels below a certain size $S$ are
-unharvested, while above that size the harvest strategy aims to return the
-population to $S$, resulting in the hockey-stick shaped policies shown.
-Only the structurally correct model (Allen model) and the GP produce
-policies close to the true optimum policy (where both the underlying
-model structure and parameter values are known without error).
-
-```{r figure_5, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
-ggplot(sims_data) + 
-  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) + 
-  facet_wrap(~method) + guides(legend.position="none")
-
-```
+(Reed 1979), whereby population levels below a certain size $S$ are
+unharvested, while above that size the harvest strategy aims to return
+the population to $S$, resulting in the hockey-stick shaped policies
+shown. Only the structurally correct model (Allen model) and the GP
+produce policies close to the true optimum policy (where both the
+underlying model structure and parameter values are known without
+error).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_5.pdf}
+\caption{Gaussian process inference outperforms parametric estimates.
+Shown are 100 replicate simulations of the stock dynamics (eq 1) under
+the policies derived from each of the estimated models, as well as the
+policy based on the exact underlying model.}
+\end{figure}
 
 The consequences of managing 100 replicate realizations of the simulated
-fishery under each of the policies estimated is shown in Figure 5.
-As expected from the policy curves, the structurally correct model
+fishery under each of the policies estimated is shown in Figure 5. As
+expected from the policy curves, the structurally correct model
 under-harvests, leaving the stock to vary around it's un-fished optimum.
@@
 
-The results across replicate stochastic simulations can most
-easily be compared by using the relative differences in net present
-value realized by each of the model. Figure 6 shows the distribution of values
+The results across replicate stochastic simulations can most easily be
+compared by using the relative differences in net present value realized
+by each of the model. Figure 6 shows the distribution of values
 (profits) realized by each simulation shown in Figure 5. The GPDP
-approach most consistently realizes a value close to the optimal solution,
-and importantly avoids ever driving the system across the tipping point,
-which results in the near-zero value cases in the parametric models.
-
-```{r figure_6, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
-For each data set, the three models are estimated as described above.
-Values plotted are the averages of a given policy over 100 replicate simulations.
-Details and code provided in the supplement.", fig.width=6, fig.height=6}
-
-ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable)) + 
-  facet_wrap(~variable, ncol=2)  + 
-  guides(legend.position = "none") +
-  xlab("Total profit by replicate") + 
-  scale_fill_manual(values=colorkey) # density plots fail when delta fn
-
-# ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable), binwidth=0.1) + 
-#  xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)
-# ggplot(actual_over_optimal, aes(value, fill=variable, color=variable)) + # density plots fail when delta fn
-#  stat_density(aes(y=..density..), position="stack", adjust=3, alpha=.9) + 
-#  xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)+ scale_color_manual(values=colorkey)
-```
-
-Sensitivity Analysis
---------------------
-
-These results are not sensitive to the modeling details of the simulation.
-The GPDP estimate remains very close to the optimal solution obtained
-by knowing the true model across changes to the training simulation,
-noise scale, parameters or structure of the underlying model, as seen in
-Figure 7. The sensitivity analysis uses a factorial design of eight replicates 
-of three different randomly generated parameter sets for each of two different 
-generating models (Allen and Myers) over two different noise levels 
-(0.01 and 0.05), for a total of 8 x 3 x 2 x 2 = 96 scenarios. In Figure 7, 
-results are pooled across different random seeds (111, 222,
-333), noise values of 0.01 and 0.05 (not large enough to violate the
-self-sustaining criterion of @Reed1979), and different randomly generated
-parameters sets for each model, and using either the Myers or Allen as the
-underlying structure.  While the GPDP performs slightly closer to optimal
-on the smaller noise levels, and closer on the Allen model then the Myers
-model, the differences are small compared to the values observed in using
-the wrong parametric model.  The GPDP solution consistently avoids either
-failing to harvest at all or crashing the stock, which results in the very
-low profits/values seen in parametric models in the previous figure. The
+approach most consistently realizes a value close to the optimal
+solution, and importantly avoids ever driving the system across the
+tipping point, which results in the near-zero value cases in the
+parametric models.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_6.pdf}
+\caption{Histograms of the realized net present value of the fishery
+over a range of simulated data and resulting parameter estimates. For
+each data set, the three models are estimated as described above. Values
+plotted are the averages of a given policy over 100 replicate
+simulations. Details and code provided in the supplement.}
+\end{figure}
+
+\subsection{Sensitivity Analysis}\label{sensitivity-analysis}
+
+These results are not sensitive to the modeling details of the
+simulation. The GPDP estimate remains very close to the optimal solution
+obtained by knowing the true model across changes to the training
+simulation, noise scale, parameters or structure of the underlying
+model, as seen in Figure 7. The sensitivity analysis uses a factorial
+design of eight replicates of three different randomly generated
+parameter sets for each of two different generating models (Allen and
+Myers) over two different noise levels (0.01 and 0.05), for a total of 8
+x 3 x 2 x 2 = 96 scenarios. In Figure 7, results are pooled across
+different random seeds (111, 222, 333), noise values of 0.01 and 0.05
+(not large enough to violate the self-sustaining criterion of Reed
+(1979)), and different randomly generated parameters sets for each
+model, and using either the Myers or Allen as the underlying structure.
+While the GPDP performs slightly closer to optimal on the smaller noise
+levels, and closer on the Allen model then the Myers model, the
+differences are small compared to the values observed in using the wrong
+parametric model. The GPDP solution consistently avoids either failing
+to harvest at all or crashing the stock, which results in the very low
+profits/values seen in parametric models in the previous figure. The
 sensitivity analysis is compared against the optimal solution alone and
@@
 see Figure 6), where convergence of the Gibbs sampling could contribute
-to their poorer performance if not checked by hand for each configuration.
+to their poorer performance if not checked by hand for each
+configuration.
 
 Large noise relative to the distance between the stable and unstable
-point, (or more precisely, between the optimal escapement and the unstable
-point) increases the chance of a stochastic transition.  As a result,
-the impact of using a model which underestimates the risk of harvesting
-past the critical point is much more severe, since this such a situation
-occurs more often.  Conversely, with large enough distance between the
-optimal escapement and unstable points relative to the noise, the chance
-of a transition becomes vanishingly small and all models can be modeled
-near-optimally. Models which underestimate the cost incurred by stock
-sizes fluctuating significantly below the optimal escapement level will
-not perform poorly as long as those fluctuations are sufficently rare.
-
-```{r figure_7, fig.height=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
-Histogram shows the ratio of the realized net present value / profits derived when managing under the GPDP Gaussian process approach over the optimal value given the true model and true parameters.
-Values near 1 indicate nearly optimal performance.
-Simulations cover a range of underlying parameter, model, replicate, and noise levels."}
-ggplot(sensitivity_dat) + 
-  geom_histogram(aes(value, fill=noise)) + 
-  xlim(0,1.0) + 
-  theme_bw() + 
-  xlab("value as fraction of the optimal") + 
-  facet_wrap(~model, ncol=1)
-```
-
-Discussion 
-==========
+point, (or more precisely, between the optimal escapement and the
+unstable point) increases the chance of a stochastic transition. As a
+result, the impact of using a model which underestimates the risk of
+harvesting past the critical point is much more severe, since this such
+a situation occurs more often. Conversely, with large enough distance
+between the optimal escapement and unstable points relative to the
+noise, the chance of a transition becomes vanishingly small and all
+models can be modeled near-optimally. Models which underestimate the
+cost incurred by stock sizes fluctuating significantly below the optimal
+escapement level will not perform poorly as long as those fluctuations
+are sufficently rare.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_7.pdf}
+\caption{Sensitivity Analysis. Histogram shows the ratio of the realized
+net present value / profits derived when managing under the GPDP
+Gaussian process approach over the optimal value given the true model
+and true parameters. Values near 1 indicate nearly optimal performance.
+Simulations cover a range of underlying parameter, model, replicate, and
+noise levels.}
+\end{figure}
+
+\section{Discussion}\label{discussion}
 
@@
 potential to increase our basic understanding of ecological processes
-[@Cuddington2013; @Geritz2012], such models can be not only inaccurate but
-misleading when relied upon in a quantitative decision making framework.
-In this paper we have tackled two aspects of uncertainty that are both
-common to many ecological decision-making problems and fundamentally
-challenging to existing approaches which largely rely on parametric
-models:
-
-1. We do not know what the correct models are for ecological systems.
-1. We have limited data from which to estimate the model -- in particular,
-   such models may be misleading in predicting the probability of outcomes
-   outside the training data.
+(Geritz \& Kisdi 2012; Cuddington \emph{et al.} 2013), such models can
+be not only inaccurate but misleading when relied upon in a quantitative
+decision making framework. In this paper we have tackled two aspects of
+uncertainty that are both common to many ecological decision-making
+problems and fundamentally challenging to existing approaches which
+largely rely on parametric models:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  We do not know what the correct models are for ecological systems.
+\item
+  We have limited data from which to estimate the model -- in
+  particular, such models may be misleading in predicting the
+  probability of outcomes outside the training data.
+\end{enumerate}
 
@@
 
-### Traditional model-choice approaches can be positively misleading.
-
-These results illustrate that model-choice approaches would be positively
-misleading -- supporting simpler models that cannot express tipping point
-dynamics merely on account of them being similar.  As the data shown
-comes only from the basin of attraction near the unfished equilibrium,
-near which all of the models are approximately linear and approximately
-identical.
+\subsubsection{Traditional model-choice approaches can be positively
+misleading.}\label{traditional-model-choice-approaches-can-be-positively-misleading.}
+
+These results illustrate that model-choice approaches would be
+positively misleading -- supporting simpler models that cannot express
+tipping point dynamics merely on account of them being similar. As the
+data shown comes only from the basin of attraction near the unfished
+equilibrium, near which all of the models are approximately linear and
+approximately identical.
 
 Model selection criteria trade off model complexity and fit to the data.
-When the data come from a limited region of state-space -- as is necessarily 
-the case whenever there is a potential concern about tipping point dynamics --
-simpler models therefore fit just as well and tend to be selected more foten than complex 
-ones.  This approach would be appropriate when the dynamics can be expected 
-to remain in the region of the training data; for instance, if we only 
-considered the forecasting accuracy of the unfished population dynamics under
-each model.
-
-In contrast, the decision-maker's problem of setting appropriate harvest levels
-cannot exclude regions of state-space outside the observed range when integrating
-over all possible decisions to find the optimal choice.  Such problems are not 
-constrained to fisheries management but ubiquitous across ecological decision-making
-and conservation where the greatest concerns involve entering previously unobserved
-regions of state-space -- whether that is the collapse of a fishery, the spread
-of an invasive, or the loss of habitat.
-
-### GPDP expresses larger uncertainty in regions where the data are poor
-
-The parametric models perform worst when they propose a management strategy
-outside the range of the observed data. The non-parametric Bayesian approach, 
-in contrast, allows a predictive model that expresses a great deal of uncertainty
-about the probable dynamics outside the observed range, while retaining very
-good predictive accuracy in the range observed.  The management policy 
-dictated by the GP balance this uncertainty against
-the immediate value of the harvest, and act to stabilize the population 
-dynamics in a region of state space in which the predictions can be 
-reliably reflected by the data.
-
-<!--
-
-### Risk-prone and risk-adverse value functions
-
-The degree to which the decision-making part of the algorithm (the SDP)
-chooses to explore or avoid the resulting region of uncertainty can also
-be influenced by the curvature of the value (profit) function $\Pi$. Both
-to simplify the intuition and avoid biasing this result, we have chosen
-profits that are linear in the catch and thus neither risk-prone nor
-risk adverse.  Making this function concave, representing the typical
-assumption of diminishing returns, would make the SDP more risk-adverse
-(as larger-than-expected stock sizes offer diminished returns relative
-to the cost of smaller-than-expected stock sizes), and strengthen the
-result shown here in which the BNP solution tends to avoid the region
-of uncertainty.  Sufficiently convex or risk-prone functions could lead
-the SDP to attempt higher exploitation rates despite the uncertainty.
-Understanding the relative roles of such functions would be a promising
-direction for future investigation.
-
--->
-
-### The role of the prior
-
-Lastly, it should be noted that outside the data, the GP reverts to
-the prior, and consequently the choice of the prior can also play a
-significant role in determining the optimal policy inferred by the
-SDP.  In the examples shown here we have selected a prior that is both
-relatively uninformative (due to the broad priors placed on its parameters
-$\ell$ and $\sigma$ and simple (mean zero, radial basis function kernel).
-In practice, both the choice of mean and the choice of the covariance
-function may be chosen to confer particular biological properties,
-as well as more biologically informed priors for $\ell$ and $\sigma$.
-In principle, this may allow a manager to improve the performance of the
-GPDP approach by adding only enough additional detail as is justified.
-For instance, it would be possible to use a linear or a Ricker-shaped
-mean in the prior without making the much stronger assumption that the
-Ricker is the structurally correct model [@Sugeno2013a]. Future research
-should focus on identifying criteria to ensure the prior and the value function
-are chosen appropriately for the problem at hand.
-
-<!--
-
-Future directions
------------------
-
-In this simulated example, the underlying dynamics are truly governed
-by a simple parametric model, allowing the parametric approaches to be
-more accurate.  Similarly, because the dynamics are  one-dimensional
-dynamics and lead to  stable nodes (rather than other attractors such
-as limit-cycles resulting in oscillations), the training data provides
-relatively limited information about the dynamics.  For these reasons,
-we anticipate that in higher-dimensional examples characteristic of
-ecosystem management problems that the GP approach will
-prove even more valuable.
-
-In our treatment here we have ignored the possibility of learning during
-the management phase, in which the additional observations of the stock
-size could potentially improve parameter estimates.  Of particular
-interest in the context of the extreme uncertainty considered here
-is the notion of "active adaptive management" or "adaptive probing",
-which may actively seek to reduce uncertainty.  The concept of adaptive
-probing is one area that has explicitly addressed the extrapolation
-uncertainty addressed here, though typically without the additional
-issue of model uncertainty.  As a result, adaptive probing strategies
-suggest rather opposite conclusions than what we observe here. Such
-adaptive probing or Dual Control (e.g. @Ludwig1982) approaches trade
-off short term utility by choosing actions that can reduce uncertainty.
-Adaptive probing strategies arise when it is valuable to intentionally
-force a system far from the observed values even when the expected value
-such actions is low, as it provides much faster learning and consequent
-reduction of model uncertainty that can allow greater value to be derived
-later on.  For instance, @Ludwig1982 show that it may be advantageous
-to fish an unexploited population very heavily at first to obtain a
-better estimate of the recruitment rate.  This intuitive strategy when a
-population is governed by a Ricker or Beverton-Holt-like dynamic would
-clearly be disastrous if instead the dynamics contained an unforeseen
-tipping point.  The best way to learn where the edge lies may be to
-walk up to it, but it is also the most dangerous.  Future work should
-attempt to understand when such active adaptive learning is valuable,
-and when it will increase the risk of an irreversible transition.
-
--->
-
-Acknowledgments
-===============
-
-This work was partially supported by NOAA-IAM grant to SM and Alec McCall and administered
-through the Center for Stock Assessment Research, a partnership between
-the University of California Santa Cruz and the Fisheries Ecology
-Division, Southwest Fisheries Science Center, Santa Cruz, CA and by NSF
-grant EF-0924195 to MM and NSF grant DBI-1306697 to CB.
+When the data come from a limited region of state-space -- as is
+necessarily the case whenever there is a potential concern about tipping
+point dynamics -- simpler models therefore fit just as well and tend to
+be selected more foten than complex ones. This approach would be
+appropriate when the dynamics can be expected to remain in the region of
+the training data; for instance, if we only considered the forecasting
+accuracy of the unfished population dynamics under each model.
+
+In contrast, the decision-maker's problem of setting appropriate harvest
+levels cannot exclude regions of state-space outside the observed range
+when integrating over all possible decisions to find the optimal choice.
+Such problems are not constrained to fisheries management but ubiquitous
+across ecological decision-making and conservation where the greatest
+concerns involve entering previously unobserved regions of state-space
+-- whether that is the collapse of a fishery, the spread of an invasive,
+or the loss of habitat.
+
+\subsubsection{GPDP expresses larger uncertainty in regions where the
+data are
+poor}\label{gpdp-expresses-larger-uncertainty-in-regions-where-the-data-are-poor}
+
+The parametric models perform worst when they propose a management
+strategy outside the range of the observed data. The non-parametric
+Bayesian approach, in contrast, allows a predictive model that expresses
+a great deal of uncertainty about the probable dynamics outside the
+observed range, while retaining very good predictive accuracy in the
+range observed. The management policy dictated by the GP balance this
+uncertainty against the immediate value of the harvest, and act to
+stabilize the population dynamics in a region of state space in which
+the predictions can be reliably reflected by the data.
+
+\subsubsection{The role of the prior}\label{the-role-of-the-prior}
+
+Lastly, it should be noted that outside the data, the GP reverts to the
+prior, and consequently the choice of the prior can also play a
+significant role in determining the optimal policy inferred by the SDP.
+In the examples shown here we have selected a prior that is both
+relatively uninformative (due to the broad priors placed on its
+parameters $\ell$ and $\sigma$ and simple (mean zero, radial basis
+function kernel). In practice, both the choice of mean and the choice of
+the covariance function may be chosen to confer particular biological
+properties, as well as more biologically informed priors for $\ell$ and
+$\sigma$. In principle, this may allow a manager to improve the
+performance of the GPDP approach by adding only enough additional detail
+as is justified. For instance, it would be possible to use a linear or a
+Ricker-shaped mean in the prior without making the much stronger
+assumption that the Ricker is the structurally correct model (Sugeno \&
+Munch 2013a). Future research should focus on identifying criteria to
+ensure the prior and the value function are chosen appropriately for the
+problem at hand.
+
+\section{Acknowledgments}\label{acknowledgments}
+
+This work was partially supported by NOAA-IAM grant to SM and Alec
+McCall and administered through the Center for Stock Assessment
+Research, a partnership between the University of California Santa Cruz
+and the Fisheries Ecology Division, Southwest Fisheries Science Center,
+Santa Cruz, CA and by NSF grant EF-0924195 to MM and NSF grant
+DBI-1306697 to CB.
 
@@
 \renewcommand*{\thefigure}{S\arabic{figure}}
-\renewcommand*{\thetable}{S\arabic{table}}
-\setcounter{figure}{0}}
-\setcounter{table}{0}}
-
-Appendix: Model definitions and estimation
-===========================================
-
-Gaussian process model
-----------------------
+\renewcommand*{\thetable}{S\arabic{table}} \setcounter{figure}{0}\}
+\setcounter{table}{0}\}
+
+\section{Appendix: Model definitions and
+estimation}\label{appendix-model-definitions-and-estimation}
+
+\subsection{Gaussian process model}\label{gaussian-process-model}
 
@@
 
-$$y = f(x) + \varepsilon$$
+\[y = f(x) + \varepsilon\]
 
@@
 
-$$f(x) \sim \mathcal{N}(E,C)$$
-$$E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y$$
-$$C = K(X_p, X_p) - K(X_p, X_o) K(X_o,X_o)^{-1} K(X_o, X_p)$$
-
-Where $K$ is the Gaussian process kernel.  Throughout, we use the radial basis function kernel:
-
-$$ K(x,y) = \exp\left(\frac{-(x-y)^2}{2 \ell^2} \right)$$
+\[f(x) \sim \mathcal{N}(E,C)\]
+\[E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y\]
+\[C = K(X_p, X_p) - K(X_p, X_o) K(X_o,X_o)^{-1} K(X_o, X_p)\]
+
+Where $K$ is the Gaussian process kernel. Throughout, we use the radial
+basis function kernel:
+
+\[ K(x,y) = \exp\left(\frac{-(x-y)^2}{2 \ell^2} \right)\]
 
@@
 
-$$\log(p(y | X)) = -\tfrac{1}{2} \mathbf{y}^T (K + \sigma_n^2 \mathbf{I})^{-1} y  - \tfrac{1}{2} \log\left| K + \sigma_n^2 \mathbf{I} \right| - \tfrac{n}{2}\log 2 \pi$$
-
-See @Munch2005a or @Rasmussen2006 for a more detailed introduction.
-
-```{r figure_S1, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options"}
-gp_assessment_plots$traces_plot
-```
-
-```{r figure_S2, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
-gp_assessment_plots$posteriors_plot
-```
-
-The Gaussian process priors on both the lengthscale $\ell$ and process noise $\sigma$ parameters are both vague inverse Gamma,
-
-$$f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)$$
-
-For the $\sigma$ prior, $\alpha = $ `r s2.p[1]` and $\beta = $ `r s2.p[2]`.  For $\ell$ prior,  $\alpha = $ `r d.p[1]` and $\beta = $ `r d.p[2]`.
+\[\log(p(y | X)) = -\tfrac{1}{2} \mathbf{y}^T (K + \sigma_n^2 \mathbf{I})^{-1} y  - \tfrac{1}{2} \log\left| K + \sigma_n^2 \mathbf{I} \right| - \tfrac{n}{2}\log 2 \pi\]
+
+See Munch et al. (2005a) or Rasmussen \& Williams (2006) for a more
+detailed introduction.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_S1.pdf}
+\caption{Traces from the MCMC estimates of the GP model show reasonable
+mixing (no trend) and sampling rejection rate (no piecewise jumps)}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_S2.pdf}
+\caption{Posterior distributions from the MCMC estimate of the GP model.
+Prior curves shown in red.}
+\end{figure}
+
+The Gaussian process priors on both the lengthscale $\ell$ and process
+noise $\sigma$ parameters are both vague inverse Gamma,
+
+\[f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)\]
+
+For the $\sigma$ prior, $\alpha = $ 5 and $\beta = $ 5. For $\ell$
+prior, $\alpha = $ 10 and $\beta = $ 10.
 
@@
 
-Optimal Control Problem Specification 
--------------------------------------
+\subsection{Optimal Control Problem
+Specification}\label{optimal-control-problem-specification}
 
@@
 
-$$ \max_{h_t} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h_t) \delta^t  $$
-
-subject to the profit function $\Pi(X_t,h)$, discount rate $\delta$, and the state
-equation
-
-$$X_{t+1} = Z_t f(S_t)  $$
-$$S_t = X_t - h_t $$
-
-For simplicity we assume profit function $\Pi_t(X_t, h_t)$ is linear in the realized harvest, only
-enforcing the restriction that harvest can not exceed available stock, that is $\Pi(h,x) = min(h,x)$. $Z_t$ is as previously defined.
+\[ \max_{h_t} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h_t) \delta^t  \]
+
+subject to the profit function $\Pi(X_t,h)$, discount rate $\delta$, and
+the state equation
+
+\[X_{t+1} = Z_t f(S_t)  \] \[S_t = X_t - h_t \]
+
+For simplicity we assume profit function $\Pi_t(X_t, h_t)$ is linear in
+the realized harvest, only enforcing the restriction that harvest can
+not exceed available stock, that is $\Pi(h,x) = min(h,x)$. $Z_t$ is as
+previously defined.
 
@@
 
-$$V_t(x_t) = \max_h \mathbf{E} \left(\Pi(h_t, x_t) + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)$$
+\[V_t(x_t) = \max_h \mathbf{E} \left(\Pi(h_t, x_t) + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)\]
 
@@
 
-Training data (Figures 1-6)
-----------------------------
+\subsection{Training data (Figures
+1-6)}\label{training-data-figures-1-6}
 
 Each of our models $f(S_t)$ must be estimated from training data, which
-we simulate from the Allen model with parameters $r = $ `r p[1]`, 
-$K =$ `r p[2]`, $C =$ `r p[3]`, and  $\sigma_g =$ `r sigma_g` 
-for $T=$ `r Tobs` timesteps, starting at initial condition $X_0 = $ `r Xo`. 
-The training data can be seen in Figure 1.
-
-Training data for sensitivity analyses
--------------------------------------
-
-A further 96 unique training data sets are generated for the sensitivity analysis, as described in the main text.
-
-Code
-----
-
-All code used in producing this analysis can be found in the R package accompanying the paper, available from [http://github.com/cboettig/nonparametric-bayes].
-The code for all results shown here is dynamically embedded into the manuscript using `knitr` [@knitr].
-
-<!--
-Data
-----
-
-While the data can be regenerated using the code provided, for convenience CSV files of the data shown in each graph are also provided, along with appropriate metadata written in the Ecological Metadata Language (EML).
-
--->
-
-```{r figure_S3, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
-plot_ricker_traces
-```
-
-```{r figure_S4, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
-ggplot(ricker_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  facet_wrap(~ variable, scale="free", ncol=2)
-```
-
-```{r Table S1, results = "asis"}
-pander::pandoc.table(ricker_priors_xtable,
-  caption = "Parameterization range for the uniform priors in the Ricker model")
-```
-
-```{r figure_S5, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
-plot_myers_traces
-```
-
-```{r figure_S6, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
-ggplot(myers_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  facet_wrap(~ variable, scale="free", ncol=2)
-```
-
-```{r TableS2, results="asis"}
-pander::pandoc.table(myers_priors_xtable,
-           caption = "Parameterization range for the uniform priors in the Myers model")
-```
-
-```{r figure_S7, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
-plot_allen_traces
-```
-
-```{r figure_S8, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
-ggplot(allen_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
-  facet_wrap(~ variable, scale="free", ncol=2)
-```
-
-```{r TableS3, results = "asis"}
-pander::pandoc.table(allen_priors_xtable,
-  caption = "Parameterization range for the uniform priors in the Allen model")
-```
-
-```{r include=FALSE}
-unlink("ricker_process.bugs")
-unlink("allen_process.bugs")
-unlink("myers_process.bugs")
-```
+we simulate from the Allen model with parameters $r = $ 2, $K =$ 8,
+$C =$ 5, and $\sigma_g =$ 0.05 for $T=$ 40 timesteps, starting at
+initial condition $X_0 = $ 5.5. The training data can be seen in Figure
+1.
+
+\subsection{Training data for sensitivity
+analyses}\label{training-data-for-sensitivity-analyses}
+
+A further 96 unique training data sets are generated for the sensitivity
+analysis, as described in the main text.
+
+\subsection{Code}\label{code}
+
+All code used in producing this analysis can be found in the R package
+accompanying the paper, available from
+{[}http://github.com/cboettig/nonparametric-bayes{]}. The code for all
+results shown here is dynamically embedded into the manuscript using
+\texttt{knitr} (Xie 2012).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_S3.pdf}
+\caption{Traces from the MCMC estimates of the Ricker model show
+reasonable mixing (no trend) and sampling rejection rate (no piecewise
+jumps)}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_S4.pdf}
+\caption{Posteriors from the MCMC estimate of the Ricker model}
+\end{figure}
+
+\begin{longtable}[c]{@{}ccc@{}}
+\toprule\addlinespace
+\begin{minipage}[b]{0.15\columnwidth}\centering
+parameter
+\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
+lower.bound
+\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
+upper.bound
+\end{minipage}
+\\\addlinespace
+\midrule\endhead
+\begin{minipage}[t]{0.15\columnwidth}\centering
+$r$
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+0.01
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+20
+\end{minipage}
+\\\addlinespace
+\begin{minipage}[t]{0.15\columnwidth}\centering
+$K$
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+0.01
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+40
+\end{minipage}
+\\\addlinespace
+\begin{minipage}[t]{0.15\columnwidth}\centering
+$\sigma$
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+1e-06
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+100
+\end{minipage}
+\\\addlinespace
+\bottomrule
+\addlinespace
+\caption{Parameterization range for the uniform priors in the Ricker
+model}
+\end{longtable}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_S5.pdf}
+\caption{Traces from the MCMC estimates of the Myers model show
+reasonable mixing (no trend) and sampling rejection rate (no piecewise
+jumps)}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_S6.pdf}
+\caption{Posterior distributions from the MCMC estimates of the Myers
+model}
+\end{figure}
+
+\begin{longtable}[c]{@{}ccc@{}}
+\toprule\addlinespace
+\begin{minipage}[b]{0.15\columnwidth}\centering
+parameter
+\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
+lower.bound
+\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
+upper.bound
+\end{minipage}
+\\\addlinespace
+\midrule\endhead
+\begin{minipage}[t]{0.15\columnwidth}\centering
+$r$
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+1e-04
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+10
+\end{minipage}
+\\\addlinespace
+\begin{minipage}[t]{0.15\columnwidth}\centering
+$K$
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+1e-04
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+40
+\end{minipage}
+\\\addlinespace
+\begin{minipage}[t]{0.15\columnwidth}\centering
+$\theta$
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+1e-04
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+10
+\end{minipage}
+\\\addlinespace
+\begin{minipage}[t]{0.15\columnwidth}\centering
+$\sigma$
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+1e-06
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+100
+\end{minipage}
+\\\addlinespace
+\bottomrule
+\addlinespace
+\caption{Parameterization range for the uniform priors in the Myers
+model}
+\end{longtable}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_S7.pdf}
+\caption{Traces from the MCMC estimates of the Allen model show
+reasonable mixing (no trend) and sampling rejection rate (no piecewise
+jumps)}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/manuscript-figure_S8.pdf}
+\caption{Posteriors from the MCMC estimate of the Allen model}
+\end{figure}
+
+\begin{longtable}[c]{@{}ccc@{}}
+\toprule\addlinespace
+\begin{minipage}[b]{0.15\columnwidth}\centering
+parameter
+\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
+lower.bound
+\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
+upper.bound
+\end{minipage}
+\\\addlinespace
+\midrule\endhead
+\begin{minipage}[t]{0.15\columnwidth}\centering
+$r$
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+0.01
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+6
+\end{minipage}
+\\\addlinespace
+\begin{minipage}[t]{0.15\columnwidth}\centering
+$K$
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+0.01
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+20
+\end{minipage}
+\\\addlinespace
+\begin{minipage}[t]{0.15\columnwidth}\centering
+$X_C$
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+0.01
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+20
+\end{minipage}
+\\\addlinespace
+\begin{minipage}[t]{0.15\columnwidth}\centering
+$\sigma$
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+1e-06
+\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
+100
+\end{minipage}
+\\\addlinespace
+\bottomrule
+\addlinespace
+\caption{Parameterization range for the uniform priors in the Allen
+model}
+\end{longtable}
+
+1.Allen, D. \& Tanner, K. \emph{et al.} (2005). Infusing active learning
+into the large-enrollment biology class: seven strategies, from the
+simple to complex. \emph{Cell biology education}, 4, 262--8.
+
+2.Athanassoglou, S. \& Xepapadeas, A. \emph{et al.} (2012). Pollution
+control with uncertain stock dynamics: When, and how, to be precautious.
+\emph{Journal of Environmental Economics and Management}, 63, 304--320.
+
+3.Bestelmeyer, B.T., Duniway, M.C., James, D.K., Burkett, L.M. \&
+Havstad, K.M.\emph{et al.} (2012). A test of critical thresholds and
+their indicators in a desertification-prone ecosystem: more resilience
+than we thought. \emph{Ecology Letters}, n/a--n/a.
+
+4.Brozović, N. \& Schlenker, W. \emph{et al.} (2011). Optimal management
+of an ecosystem with an unknown threshold. \emph{Ecological Economics},
+1--14.
+
+5.Burnham, K.P. \& Anderson, D.R. \emph{et al.} (2002). \emph{Model
+Selection and Multi-Model Inference}. Springer.
+
+6.Clark, C.W. \& Kirkwood, G.P. \emph{et al.} (1986). On uncertain
+renewable resource stocks: Optimal harvest policies and the value of
+stock surveys. \emph{Journal of Environmental Economics and Management},
+13, 235--244.
+
+7.Courchamp, F., Berec, L. \& Gascoigne, J.\emph{et al.} (2008).
+\emph{Allee Effects in Ecology and Conservation}. Oxford University
+Press, USA.
+
+8.Cressie, N., Calder, C. a, Clark, J.S., {Ver Hoef}, J.M. \& Wikle,
+C.K.\emph{et al.} (2009). Accounting for uncertainty in ecological
+analysis: the strengths and limitations of hierarchical statistical
+modeling. \emph{Ecological Applications}, 19, 553--70.
+
+9.Cuddington, K.M., Fortin, M. \& Gerber, L.\emph{et al.} (2013).
+Process-based models are required to manage ecological systems in a
+changing world. \emph{Ecosphere}, 4, 1--12.
+
+10.Fischer, J., Peterson, G.D., Gardner, T. a, Gordon, L.J., Fazey, I.
+\& Elmqvist, T.\emph{et al.} (2009). Integrating resilience thinking and
+optimisation for conservation. \emph{Trends in ecology \& evolution},
+24, 549--54.
+
+11.Gelman, A., Carlin, J.B., Stern, H.S. \& Rubin, D.B.\emph{et al.}
+(2003). \emph{Bayesian Data Analysis}. 2nd edn. Chapman; Hall/CRC.
+
+12.Geritz, S.A.H. \& Kisdi, E. \emph{et al.} (2012). Mathematical
+ecology: why mechanistic models? \emph{Journal of mathematical biology},
+65, 1411--5.
+
+13.Gordon, H. \emph{et al.} (1954). The economic theory of a
+common-property resource: the fishery. \emph{The Journal of Political
+Economy}, 62, 124--142.
+
+14.Halpern, B.S., Klein, C.J., Brown, C.J., Beger, M., Grantham, H.S. \&
+Mangubhai, S.\emph{et al.} (2013). Achieving the triple bottom line in
+the face of inherent trade-offs among social equity, economic return,
+and conservation. \emph{Proceedings of the National Academy of Sciences
+of the United States of America}, 110, 6229--34.
+
+15.Hilborn, R. \emph{et al.} (2007). Reinterpreting the State of
+Fisheries and their Management. \emph{Ecosystems}, 10, 1362--1369.
+
+16.Hilborn, R. \& Mangel, M. \emph{et al.} (1997). \emph{The Ecological
+Detective: Confronting Models with data}. Princeton University Press.
+
+17.Hughes, T.P., Linares, C., Dakos, V., Leemput, I. a van de \& Nes,
+E.H. van\emph{et al.} (2013). Living dangerously on borrowed time during
+slow, unrecognized regime shifts. \emph{Trends in ecology \& evolution},
+28, 149--55.
+
+18.Kocijan, J., Girard, A., Banko, B. \& Murray-Smith, R.\emph{et al.}
+(2005). Dynamic systems identification with Gaussian processes.
+\emph{Mathematical and Computer Modelling of Dynamical Systems}, 11,
+411--424.
+
+19.Lehmann, E.L. \emph{et al.} (1975). \emph{Nonparametrics: Statistical
+Methods Based on Ranks}. Holden-Day, Inc, San Francisco.
+
+20.Ludwig, D. \& Walters, C.J. \emph{et al.} (1982). Optimal harvesting
+with imprecise parameter estimates. \emph{Ecological Modelling}, 14,
+273--292.
+
+21.Mangel, M. \& Clark, C.W. \emph{et al.} (1988). \emph{Dynamic
+Modeling in Behavioral Ecology}. Princeton University Press, Princeton.
+
+22.Marescot, L., Chapron, G., Chadès, I., Fackler, P.L., Duchamp, C. \&
+Marboutin, E.\emph{et al.} (2013). Complex decisions made simple: a
+primer on stochastic dynamic programming. \emph{Methods in Ecology and
+Evolution}, n/a--n/a.
+
+23.May, R.M., Beddington, J.R., Clark, C.W., Holt, S.J. \& Laws,
+R.M.\emph{et al.} (1979). Management of multispecies fisheries.
+\emph{Science (New York, N.Y.)}, 205, 267--77.
+
+24.McAllister, M. \emph{et al.} (1998). Bayesian stock assessment: a
+review and example application using the logistic model. \emph{ICES
+Journal of Marine Science}, 55, 1031--1060.
+
+25.Munch, S.B., Kottas, A. \& Mangel, M.\emph{et al.} (2005a). Bayesian
+nonparametric analysis of stock-recruitment relationships.
+\emph{Canadian Journal of Fisheries and Aquatic Sciences}, 62,
+1808--1821.
+
+26.Munch, S.B., Snover, M.L., Watters, G.M. \& Mangel, M.\emph{et al.}
+(2005b). A unified treatment of top-down and bottom-up control of
+reproduction in populations. \emph{Ecology Letters}, 8, 691--695.
+
+27.Myers, R.A., Barrowman, N.J., Hutchings, J.A. \& Rosenberg, A.
+a\emph{et al.} (1995). Population dynamics of exploited fish stocks at
+low population levels. \emph{Science (New York, N.Y.)}, 269, 1106--8.
+
+28.Polasky, S., Carpenter, S.R., Folke, C. \& Keeler, B.\emph{et al.}
+(2011). Decision-making under great uncertainty: environmental
+management in an era of global change. \emph{Trends in ecology \&
+evolution}, 1--7.
+
+29.R Core Team \emph{et al.} (2013). \emph{R: A Language and Environment
+for Statistical Computing}. R Foundation for Statistical Computing,
+Vienna, Austria.
+
+30.Rasmussen, C.E. \& Williams, C.K.I. \emph{et al.} (2006).
+\emph{Gaussian Processes for Machine Learning}. MIT Press, Boston.
+
+31.Reed, W.J. \emph{et al.} (1979). Optimal escapement levels in
+stochastic and deterministic harvesting models. \emph{Journal of
+Environmental Economics and Management}, 6, 350--363.
+
+32.Roughgarden, J.E. \& Smith, F. \emph{et al.} (1996). Why fisheries
+collapse and what to do about it. \emph{Proceedings of the National
+Academy of Sciences of the United States of America}, 93, 5078.
+
+33.Schapaugh, A.W. \& Tyre, A.J. \emph{et al.} (2013). Accounting for
+parametric uncertainty in Markov decision processes. \emph{Ecological
+Modelling}, 254, 15--21.
+
+34.Scheffer, M., Bascompte, J., Brock, W.A., Brovkin, V., Carpenter,
+S.R. \& Dakos, V.\emph{et al.} (2009). Early-warning signals for
+critical transitions. \emph{Nature}, 461, 53--9.
+
+35.Scheffer, M., Carpenter, S.R., Foley, J.A., Folke, C. \& Walker,
+B.\emph{et al.} (2001). Catastrophic shifts in ecosystems.
+\emph{Nature}, 413, 591--6.
+
+36.Sethi, G., Costello, C., Fisher, A., Hanemann, M. \& Karp, L.\emph{et
+al.} (2005). Fishery management under multiple uncertainty.
+\emph{Journal of Environmental Economics and Management}, 50, 300--318.
+
+37.Sigourney, D.B., Munch, S.B. \& Letcher, B.H.\emph{et al.} (2012).
+Combining a Bayesian nonparametric method with a hierarchical framework
+to estimate individual and temporal variation in growth.
+\emph{Ecological Modelling}, 247, 125--134.
+
+38.Su, Y.-S. \& Masanao Yajima \emph{et al.} (2013). \emph{R2jags: A
+Package for Running jags from R}.
+
+39.Sugeno, M. \& Munch, S.B. \emph{et al.} (2013a). A semiparametric
+Bayesian method for detecting Allee effects. \emph{Ecology}, 94,
+1196--1204.
+
+40.Sugeno, M. \& Munch, S.B. \emph{et al.} (2013b). A semiparametric
+Bayesian approach to estimating maximum reproductive rates at low
+population sizes. \emph{Ecological applications : a publication of the
+Ecological Society of America}, 23, 699--709.
+
+41.Thorson, J.T., Ono, K. \& Munch, S.B.\emph{et al.} (2014). A Bayesian
+approach to identifying and compensating for model misspecification in
+population models. \emph{Ecology}, 95, 329--41.
+
+42.Weitzman, M.L. \emph{et al.} (2013). A Precautionary Tale of
+Uncertain Tail Fattening. \emph{Environmental and Resource Economics},
+55, 159--173.
+
+43.Williams, B.K. \emph{et al.} (2001). Uncertainty , learning , and the
+optimal management of wildlife. \emph{Environmental and Ecological
+Statistics}, 8, 269--288.
+
+44.Worm, B., Barbier, E.B., Beaumont, N., Duffy, J.E., Folke, C. \&
+Halpern, B.S.\emph{et al.} (2006). Impacts of biodiversity loss on ocean
+ecosystem services. \emph{Science (New York, N.Y.)}, 314, 787--90.
+
+45.Worm, B., Hilborn, R., Baum, J.K., Branch, T.A., Collie, J.S. \&
+Costello, C.\emph{et al.} (2009). Rebuilding global fisheries.
+\emph{Science (New York, N.Y.)}, 325, 578--85.
+
+46.Xie, Y. \emph{et al.} (2012). \emph{knitr: A general-purpose package
+for dynamic report generation in R}.
+
+\end{document}

### 2014-05-22 manuscripts/manuscript.Rmd
working on marc's edits

@@
 linenumbers: true
-title: "Avoiding tipping points in ecosystem based fisheries management: a non-parametric Bayesian approach"
+title: "Avoiding tipping points in fisheries management: A Bayesian Non-Parametric approach"
 author: 
@@
       Santa Cruz, CA 95064, USA
+  - code: marc
+    address: |
+      Center for Stock Assessment Research, Department of Applied Math
+      and Statistics, University of California,
+      Mail Stop SOE-2, Santa Cruz, CA 95064, USA and Department of
+      Biology, University of Bergen, Bergen, Norway 9020
   - code: noaa
@@
           Model uncertainty and limited data coverage are fundamental
-          challenges to robust ecosystem management.  These challenges
+          challenges to robust management of human intervention in a 
+          natural system.  These challenges
           are acutely highlighted by concerns that many ecological systems
-          may contain tipping points.  Before a collapse, we do not know
+          may contain tipping points, such as Allee population sizes.  
+          Before a collapse, we do not know
           where the tipping points lie, if they exist at all.  Hence,
@@
           do we have access to data in some large region of state-space
-          where such a tipping point might exist.  These two sources of
-          uncertainty frustrate state-of-the-art parametric approaches
-          to decision theory and optimal control.  Here we illustrate
-          how a non-parametric approach using a Gaussian process prior
-          provides a more flexible representation of this inherent
-          uncertainty. We can then embed the Gaussian process in a
+          where such a tipping point might exist.  We illustrate
+          how a Bayesian Non-Parametric (BNP) approach using a Gaussian Process (GP) prior
+          provides a flexible representation of this inherent
+          uncertainty. We embed GPs in a
           stochastic dynamic programming framework in order to make
-          robust management predictions under both model uncertainty and
+          robust management predictions with both model uncertainty and
           limited data. We conduct a set of simulations to evaluate this
           approach as compared with the standard approach of using model
-          selection to choose from a set of candidate models.  In keeping
-          with previous results, we find that model selection erroneously
+          selection to choose from a set of candidate models.  We find that model selection erroneously
           favors models without tipping points -- leading to harvest policies
-          that guarantee extinction. The Gaussian process approach
+          that guarantee extinction. The GP approach
           performs nearly as well as the true model and significantly
@@
           still fails to account for uncertainty appropriately and leads
-          to population crashes, while management based on the Gaussian 
-          process approach does not. We discuss opportunities to adapt
-          this approach to multiple species models in further work.
+          to population crashes, while management based on the GP 
+          approach does not, since it does not underestimate the uncertainty
+          outside of the observed data.
 
@@
 
-Decision making under uncertainty is a ubiquitous challenge of
-natural resource management and conservation.  Ecological dynamics
-are frequently complex and difficult to measure, making uncertainty
-in our understanding a persistent challenge to effective management.
+Decision making under uncertainty is a ubiquitous challenge in the
+management of human intervention in natural resources and conservation.
+Ecological dynamics are frequently complex and difficult to measure,
+making uncertainty in our understanding a persistent challenge.
 Decision-theoretic approaches provide a framework to determine the
 best sequence of actions in face of uncertainty, but only when that
-uncertainty can be meaningfully quantified [@Fischer2009].  Concerns
-about the potential sudden collapse of fisheries and other ecosystems
-[@Worm2006; @Scheffer2009] has increasingly emphasized the difficulties of
+uncertainty can be meaningfully quantified [@Fischer2009].  Over the last
+four decades (but stretching even further back in time), marked by  the
+seminal contributions of @Clark1976; @Clark2009 and @Walters1978,
+dynamic optimization methods, particularly Stochastic Dynamic Programming
+(SDP, @Clark2009), have become increasingly important as a means of
+understanding how to management human intervention into natural systems.
+Concomittantly, there has been increasing recognition of the importance of
+Allee population sizes or `tipping points' [@Scheffer2001;
+@Polasky2011] that management actions may inadvertently
+cause populations to cross.  @Scheffer2009 emphasized the difficulties of
 formulating even qualitatively correct models of the underlying processes.
 
-We develop a novel approach to address these concerns in the context of fisheries; though
-the underlying challenges and methods are germane to many other
-conservation and resource management problems.  Economic value and
-ecological concern have made marine fisheries the crucible for much
-of the founding work [@Gordon1954; @Reed1979; @May1979; @Ludwig1982]
-on management under uncertainty.  Global declines [@Worm2006]
-and the controversy surrounding their interpretation [@Hilborn2007; @Worm2009] have made understanding these
-challenges all the more pressing.
+We develop a novel approach to address these concerns in the context of
+fisheries; though the underlying challenges and methods are germane to
+many other problems of conservation or natural resource exploitation,
+such as forestry.  Economic value and ecological concern have made
+marine fisheries the crucible for much of the founding work [@Gordon1954;
+@Clark1976; -@Clark2009; @Reed1979; @May1979; @Ludwig1982] on management
+under uncertainty.  Global declines [@Worm2006] and the controversy
+surrounding their interpretation [@Hilborn2007; @Worm2009] have made
+understanding these challenges all the more pressing.
 
@@
 (or reasonable approximation) of the dynamics must be identified from
-among a handful of alternative models.  Here we consider an approach
-that addresses uncertainty at each of these levels without assuming the
-dynamics follow a particular (i.e. parametric) structure.
-
-Often we do not understand the system dynamics well
-enough to know if a model provides a good approximation over the relevant
-range of states and timescales (criteria that we loosely refer to as
-defining the "right" or "true" model.)  
-Typical solutions involve either model choice, model averaging, or
+among a handful of alternative models, using either model choice, model averaging, or 
 introducing yet greater model complexity of which others may be special
@@
 [@Williams2001; @Athanassoglou2012; @Cressie2009].  Even setting aside
-other computational and statistical concerns (e.g. [@Cressie2009]), these
+other computational and statistical concerns (e.g. @Cressie2009), these
 approaches do not address our second concern - representing uncertainty
@@
 Model uncertainty is particularly insidious when model predictions must
-be made outside of the range of data on which the model was estimated.
-This extrapolation uncertainty is felt most keenly in decision-theoretic
+be made outside of the range of data on which the model was estimated, and models may be abused [@Mangel2001].
+Extrapolation uncertainty is felt most keenly in decision-theoretic
 (or optimal control) applications, since (a) exploring the potential
@@
 
-This difficult position of having neither the true model nor data that
+<!--This difficult position of having neither the true model nor data that
 covers the full range of possible states is unfortunately the rule more
@@
 If we do not know but cannot rule out such a possibility, then we
-face decision-making under this dual challege of model uncertainty and
-incomplete data coverage.
-
-These dual concerns pose a substantial challenge to existing decision-theoretic
-approaches [@Brozovic2011].  Because intervention is often too late 
-after a tipping point has been crossed (but see @Hughes2013), management
+face decision-making under this 
+-->
+
+The dual concerns of model uncertainty and
+incomplete data coverage pose a substantial challenge to existing decision-theoretic
+approaches [@Brozovic2011].  Because intervention is often (but not always, see @Hughes2013) too late after a tipping point has been crossed, management
 is most often concerned with avoiding potentially catastrophic tipping
@@
 
-<!--
-_Map_
--->
-
-Here we illustrate how a stochastic dynamic programming (SDP) algorithm
+We illustrate how a stochastic dynamic programming (SDP) algorithm
 [@Mangel1988; @Marescot2013] can be driven by the
-predictions from a Bayesian non-parametric (BNP) approach [@Munch2005a].
+predictions from a Bayesian Non-Parametric (BNP) model of population dynamics [@Munch2005a].
 This provides two distinct advantages compared with contemporary
@@
 a much greater uncertainty far from the observed data than near the
-observed data.  These features allow the GPDP approach to find robust
+observed data.  These features allow the Gaussian Process Dynamic Programming (GPDP) approach to find robust
 management solutions in face of limited data and without knowledge of
@@
 
+<!--
 The idea that any approach can perform well without either having to
@@
 management policies than with common parametric modeling approaches.
+-->
 
@@
 
+To illustrate GPDP, we consider the performance of the GPDP policy against
+the policies derived under several alternative parametric models [@Reed1979; @Ludwig1982; @Mangel1988]. 
 The nature of decision-making problems provides a compelling and
 pragmatic way to compare models: Rather than compare models in terms
-of best fit to data, model performance is defined in the concrete terms
-of the decision-maker's objectives. Although much argument can be made
+of best fit to data, we define model performance in the concrete terms
+of the decision-maker's objectives.
+
+<!-- Although much argument can be made
 over the 'correct' objective function, e.g. how to account for the
@@
 performance over the period of interest.
+-->
 
@@
 
-### Statement of the optimal control problem
-
-To illustrate the application of the Gaussian process-based dynamic
-programming approach (hereafter GPDP) we focus on the classical problem
-of determining a harvest policy that will maximize the expected discounted
-profit over a fixed time interval [@Reed1979; @Ludwig1982; @Mangel1988].
-For simplicity we assume profit is proportional to the realized harvest.  
-For comparison, we consider the performance of the GPDP policy against
-policies derived under several alternative parametric models.
-
-We focus on a class of one-dimensional models of population dynamics as a
+We first describe the requirements of dynamic optimization for the management of human intervention in natural resource systems.
+After that we describe three parametric models for population dynamics and the GP description of population dynamics.
+
+### Requirements of dynamic optimization
+
+Dynamic optimization  requires characterizing the dynamics of a state variable (or variables), a control action, and a value function.
+In this paper, we consider only a single state variable, so that the ideas can be described as simply as possible. That is, we focus on a class of one-dimensional models of population dynamics as a
 'best-case' scenario for the parametric model performance relative to
-the Gaussian process.   We consider this a best-case scenario for the
-following reasons: first, the underlying dynamics are always simuated
+the GP.   This a best-case scenario for the parametric models because the underlying dynamics are always simuated
 from one the parametric models in the set, whereas in reality we never

### 2014-05-23 manuscripts/manuscript.Rmd
adding more marc edits and comments

@@
 
-## Background on the Gaussian Process
+Approach and Methods
+====================
+
+We first describe the requirements of dynamic optimization for the management of human intervention in natural resource systems.
+After that we describe three parametric models for population dynamics and the GP description of population dynamics.
+
+### Requirements of dynamic optimization
+
+Dynamic optimization  requires characterizing the dynamics of a state variable (or variables), a control action, and a value function.
+In this paper, we consider only a single state variable, so that the ideas can be described as simply as possible. That is, we focus on a class of one-dimensional models of population dynamics as a
+'best-case' scenario for the parametric model performance relative to
+the GP.   This is abest-case scenario for the parametric models because the underlying dynamics are always simuated
+from one the parametric models in the set, whereas in reality we never
+have the "true" model among our candidates.  By choosing one-dimensional
+models with few parameters, we limit the chance that poor performance
+will be due to our inability to estimate parameters well, something
+that becomes a more severe problem for higher-dimensional parametric
+models. Further, the parametric models we consider are those most commonly
+used in modeling stock-recruitment dynamics or to model sudden transitions
+between alternative stable states.
+
+### Parametric models
 
 <!-- 
-- Background on non-parametric modeling.  
+- Statement of the models
 -->
 
-<!--
-- Definition
-- Previous application
--->
-
-We use a Bayesian nonparameteric regression approach based on Gaussian
-process priors.  The use of Gaussian process (GP) regression (or
-"kriging" in the geospatial literature) to formulate a predictive
+We let $X_t$ denote the size (numbers or biomass) of the focal population at time $t$ and assume that in the absence of take its dynamics are
+
+$$ X_{t+1}= Z_t f(X_t, h_t, p) $$
+
+Where Z(t) is log-normally distributed process stochasticity [@Reed1979] and $p$ is a vector of parameters to be estimated from the data.
+We describe three choices for $f(X_t,p)$ in the next section.In this simple model, the control action is a harvest or take, $h_t$, measured in the same units as $X$, at time
+$t$.  Thus, in the presence of take, the population size on the right hand side of Eqn 1 is replaced by
+$S_t=X_t-h_t$.
+
+To construct the value function, we consider a return when $X_t=x$ and harvest $h_t=h$ denoted by
+$\Pi(x,h)$.
+For example, if the return is simply the harvest at time $t$, then $\Pi(x,h)=min(x,h)$.
+We assume that future harvests are discounted relative to current ones at a constant rate of discount $\delta$ and ask for the harvest policy that maximizes total discounted harvest between the current time $t$ and a final time $T$, that is we seek to maximize over choices of harvest $\mathbf{E_Z}[ \sum_{t = 0}^{T}  \Pi_t(X_t, h_t) \delta^t]$, where the state dynamics are given by Eqn 1 and $\mathbf{E_Z}$ denotes the expectation over the process stochasticity.
+
+In order to find that policy, we introduce the value function$V_t(x_t)$ representing the total discounted catch from time $t$ onwards given that $X_t=x$.
+This value function satisfies an equation of Stochastic Dynamic Programming  [SDP, @Mangel1988; @Clark2000, @Clark2009], sometimes called the Bellman-Hamilton-Jacobi equation [@Mangel2014].
+
+\begin{equation}
+V_t(x) = \max_{h_t} \left[\Pi(h_t, x_t) + \delta \cdot \mathbf{E_Z}  V_{t+1}( Z_{t+1} f(x - h_t)) \right]
+\end{equation}
+
+That is, at time $t$, when population size is $x$ and harvest $h_t$ is applied, the immediate return is $\Pi(h_t, x_t)$ and the population size after the take is $x-h_t$, which is then translated into $X_{t+1}$ by Eqn 1.
+We consider the finite time problem with $T=$`r Tmax`, which we solve using value iteration algorithm [@Mangel1988; @Clark2000].
+
+\subsubsection{Parametric models}\label{parametric-models}
+
+We consider three candidate parametric models of the stock-recruitment
+dynamics: The Ricker model, the Allen model [@Allen2005], the Myers
+model [@Myers1995], equations (@ricker)-(@myers), which we parameterize as follows for $f(S_t,p)$ in Eqn 1.
+In all three, we let $K$ denote the carrying capacity and $r$ the maximum per capita growth rate.
+In the Ricker model,
+
+(@ricker) $$ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
+
+In the Allen model,
+
+(@allen) $$ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_c\right)} $$
+
+Where $r$ and $K$ are as before, and $X_c$ denotes the location of the unstable steady state (i.e., the tipping point).
+
+In the Myers model,
+
+(@myers) $$ f(S(t) | p)  = \frac{r S_t^{\theta}}{1 + \frac{S_t^\theta}{K}}$$
+
+Where $\theta = 1$ corresponding to Beverton-Holt dynamics and $\theta$ > 2 leads to Allee effects and multiple stable states.
+
+The Ricker model involves two parameters, corresponding
+to a growth rate and a carrying capacity, and cannot support alternative
+stable growth rate dynamics. The Allen model resembles the Ricker
+dynamics with an added Allee effect parameter [@Courchamp2008], below
+which the population cannot persist.  The Myers model also has three
+parameters and contains an Allee threshold, but has compensatory rather
+than over-compensatory density dependence (resembling a Beverton-Holt
+curve rather than a Ricker curve at high densities.)
+
+The multiplicative log-normal stochasticity perturbs the growth predicted 
+by each of these deterministic model skeletons. This 
+introduces one additional parameter $\sigma$ that must be estimated.
+
+Because of our interest in management performance in the presence of
+tipping points, all of our simulations are based on the Allen model.
+The Allen model is thus "structurally correct" and is expected to provide
+a best-case scenario when the 'true' dynamics are known.
+The Ricker model is a reasonable approximation of these dynamics far from the Allee
+threshold (but lacks threshold dynamics), while the Myers model shares
+the essential feature of a threshold but differs in the structure.
+
+We consider a period of `r Tobs` years of training data: long
+enough that the estimates do not depend on the particular realization,
+while longer times are not likely to provide substantial improvement
+(i.e. the results are not sensitive to this interval).  Each of the models
+(described below) is fit to the same training data (Figure 1).
+
+<!-- Bayesian inference of parametric models -->
+
+We inferred posterior distributions for the parameters of each model by Gibbs sampling (@Gelman2003 implemented in R
+[@RTeam] using jags, [@R2jags]).  We choose uniform priors
+for all parameters (See appendix Tables S1-S3; R code provided). We
+show one-step-ahead predictions of these model fits in Figure 1.
+We tested each chain for Gelman-Rubin convergence and results were
+robust to longer runs.  For each simulation we also applied several
+commonly used model selection criteria (AIC, BIC, DIC, see @Burnham2002)
+to identify the best fitting model.
+
+### The Gaussian Process model
+
+<!-- Statement of model -->
+
+ The use GP methods to formulate a predictive
 model is relatively new in the context of modeling dynamical systems
@@
 An accessible and thorough introduction to the formulation and use of
-GPs can be found in @Rasmussen2006. The nonparametric moniker for the
+GPs can be found in @Rasmussen2006.
+
+The nonparametric moniker for the
 GP approach is unfortunate as (a) it still involves the estimation of
@@
 
-Approach and Methods
-====================
-
-We first describe the requirements of dynamic optimization for the management of human intervention in natural resource systems.
-After that we describe three parametric models for population dynamics and the GP description of population dynamics.
-
-### Requirements of dynamic optimization
-
-Dynamic optimization  requires characterizing the dynamics of a state variable (or variables), a control action, and a value function.
-In this paper, we consider only a single state variable, so that the ideas can be described as simply as possible. That is, we focus on a class of one-dimensional models of population dynamics as a
-'best-case' scenario for the parametric model performance relative to
-the GP.   This a best-case scenario for the parametric models because the underlying dynamics are always simuated
-from one the parametric models in the set, whereas in reality we never
-have the "true" model among our candidates.  By choosing one-dimensional
-models with few parameters, we limit the chance that poor performance
-will be due to our inability to estimate parameters well, something
-that becomes a more severe problem for higher-dimensional parametric
-models. Further, the parametric models we consider are those most commonly
-used in modeling stock-recruitment dynamics or to model sudden transitions
-between alternative stable states.
-
-### Parametric models
-
-<!-- 
-- Statement of the models
--->
-
-We consider three candidate parametric models of the stock-recruitment
-dynamics: The Ricker model, the Allen model [@Allen2005], the Myers
-model [@Myers1995], equations (@ricker)-(@myers), which we parameterize as follows:
-
-We consider three parametric models that relate population size at time
-$t+1$, $X_{t+1}$, to population size at time $t$, $X_t$  and harvest at
-time $t$, $h_t$ by
-
-$$ X_{t+1}= Z_t f(S_t,p) $$
-
-Where Z(t) is log-normally distributed process stochasticity [@Reed1979]
-and $p$ is a vector of parameters, and $S_t$ the escapement [@Reed1979]
-
-$$S_t = X_t - h_t$$
-
-or the Ricker model,
-
-(@ricker) $$ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
-
-where $K$ is the carrying capacity and $r$ the growth rate.  For the Allen model,
-
-(@allen) $$ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_C\right)} $$
-
-Where $r$ and $K$ are as before, and $X_c$ denotes the location of the unstable (tipping) point.  For the Myers model,
-
-(@myers) $$ f(S(t) | p)  = \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}}$$
-
-Where $r$ and $K$ as before, and $\theta$ the nonlinearity of the Myers model (corresponding to Beverton-Holt dynamics for $\theta$ = 1 and an Allee effect with alternative stable states exist at $\theta > 2$).
-
-The familiar Ricker model involves two parameters, corresponding
-to a growth rate and a carrying capacity, and cannot support alternative
-stable growth rate dynamics. The Allen model resembles the Ricker
-dynamics with an added Allee effect parameter [@Courchamp2008], below
-which the population cannot persist.  The Myers model also has three
-parameters and contains an Allee threshold, but has compensatory rather
-than over-compensatory density dependence (resembling a Beverton-Holt
-curve rather than a Ricker curve at high densities.)  
-We assume multiplicative log-normal noise perturbs the growth predicted 
-by the each of the deterministic model skeletons described above. This 
-introduces one additional parameter $\sigma$ that must be estimated by each
-model.
-
-Because of our interest in management performance in the presence of
-tipping points, all of our simulations are based on the Allen model.
-The Allen model is thus structurally correct and is expected to provide
-a best-case scenario when the 'true' dynamics are known.
-The Ricker model is a reasonable approximation of these dynamics far from the Allee
-threshold (but lacks threshold dynamics), while the Myers model shares
-the essential feature of a threshold but differs in the structure.
-
-We consider a period of `r Tobs` years of training data: long
-enough that the estimates are not dependent on the particular realization,
-while longer times are not likely to provide substantial improvement
-(i.e. the results are not sensitive to this interval).  Each of the models
-(described below) is fit to the same training data, as shown in Figure 1.
-
-<!-- Bayesian inference of parametric models -->
-
-We inferred posterior distributions for the parameters of each model in
-a Bayesian context using Gibbs sampling (@Gelman2003 implemented in R
-[@RTeam] using jags, [@R2jags]).  We choose uninformative uniform priors
-for all parameters (See appendix Tables S1-S3; R code provided). We
-show one-step-ahead predictions of these model fits in Figure 1.
-Each chain was tested for Gelman-Rubin convergence and results were
-robust to longer runs.  For each simulation we also applied several
-commonly used model selection criteria (AIC, BIC, DIC, see @Burnham2002)
-to identify the best fitting model.
-
-<!-- SDP via parametric models -->
-
-An optimal policy function was then inferred through stochastic dynamic
-programming for each model given the posterior distributions of the
-parameter estimates.  This policy maximizes the expectation of the value
-function integrated over the parameter uncertainty. (code implementing
-this algorithm provided in the Appendix).
-
-### The Gaussian Process model
-
-<!-- Statement of model -->
+<!-- FIXME Marc suggests putting the full GP formulation in here?? -->
 
@@
 
-<!-- SDP via the model -->
+<!-- SDP via the model  FIXME much more detail here-->
+
+The method of Gaussian Process Dynamic Programming
+--------------------------------------------------
 
@@
 =======
+
+Parametric and GP models for population dynamics
+------------------------------------------------
 
@@
 
-```{r figure_3, fig.width=8, fig.height=6, fig.cap="Beyond the tipping point (outside the range of the training data), the true dynamics fall outside the uncertainty bands of the wrong parametric models, but inside the uncertainty predicted by the GP.
+```{r figure_3, fig.width=8, fig.height=6, fig.cap="Outside the range of the training data (Figure 1), the true dynamics (black dots) fall outside the uncertainty (two standard deviations, colored bands) of the structurally incorrect parametric models (Ricker, Myers), but inside the uncertainty predicted by the GP.
 Points show the stock size simulated by the true model.
@@
 The Ricker model always (wrongly) predicts positive population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).
-Note that the GP is both more pessimistic and more uncertain about the future state than the parametric models.", dependson="plot-options"}
+Note that because it does not assume a parametric form but instead relies more directly on the data, the GP is both more pessimistic and more uncertain about the future state than the parametric models.", dependson="plot-options"}
 y <- numeric(8)
@@
 
-```{r figure_5, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates.
+```{r figure_5, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models.
 Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
@@
 passed the tipping point consistently, resulting in the immediate crash
-of the stock and thus derives minimal profits.
+of the stock and thus derives minimal long-term catch.
 
@@
 value realized by each of the model. Figure 6 shows the distribution of values
-(profits) realized by each simulation shown in Figure 5.  The GPDP
-approach most consistently realizes a value close to the optimal solution,
+realized by each simulation shown in Figure 5.  The GPDP
+most consistently realizes a value close to the optimal solution,
 and importantly avoids ever driving the system across the tipping point,
@@
 (0.01 and 0.05), for a total of 8 x 3 x 2 x 2 = 96 scenarios. In Figure 7, 
-results are pooled across different random seeds (111, 222,
+we pool results across different random seeds (111, 222,
 333), noise values of 0.01 and 0.05 (not large enough to violate the
@@
 failing to harvest at all or crashing the stock, which results in the very
-low profits/values seen in parametric models in the previous figure. The
+low values seen in parametric models in the previous figure. The
 sensitivity analysis is compared against the optimal solution alone and
@@
 of a transition becomes vanishingly small and all models can be modeled
-near-optimally. Models which underestimate the cost incurred by stock
+near-optimally. Models that underestimate the cost incurred by stock
 sizes fluctuating significantly below the optimal escapement level will
@@
 
+<!--FIXME the above para is not particularly convincing. Quantitative please? -->
+
 ```{r figure_7, fig.height=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
-Histogram shows the ratio of the realized net present value / profits derived when managing under the GPDP Gaussian process approach over the optimal value given the true model and true parameters.
+Histogram shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters.
 Values near 1 indicate nearly optimal performance.
@@
 potential to increase our basic understanding of ecological processes
-[@Cuddington2013; @Geritz2012], such models can be not only inaccurate but
+[@Cuddington2013; @Geritz2012], such models can be both inaccurate and
 misleading when relied upon in a quantitative decision making framework.
@@
 
-### GPDP expresses larger uncertainty in regions where the data are poor
+### GPDP population dynamcis capture larger uncertainty in regions where the data are poor
+
+<!-- FIXME More emphasis/explanation on _how_ it does this -->
 
@@
 In principle, this may allow a manager to improve the performance of the
-GPDP approach by adding only enough additional detail as is justified.
+GPDP by adding only enough additional detail as is justified.
 For instance, it would be possible to use a linear or a Ricker-shaped

### 2014-05-23 manuscripts/manuscript.Rmd
updates as sent to steve

@@
 
-To construct the value function, we consider a return when $X_t=x$ and harvest $h_t=h$ denoted by
-$\Pi(x,h)$.
-For example, if the return is simply the harvest at time $t$, then $\Pi(x,h)=min(x,h)$.
-We assume that future harvests are discounted relative to current ones at a constant rate of discount $\delta$ and ask for the harvest policy that maximizes total discounted harvest between the current time $t$ and a final time $T$, that is we seek to maximize over choices of harvest $\mathbf{E_Z}[ \sum_{t = 0}^{T}  \Pi_t(X_t, h_t) \delta^t]$, where the state dynamics are given by Eqn 1 and $\mathbf{E_Z}$ denotes the expectation over the process stochasticity.
+To construct the value function, we consider a return when $X_t=x$ and harvest $h_t=h$ denoted as the reward, $R(x,h)$.
+For example, if the return is simply the harvest at time $t$, then $R(x,h)=min(x,h)$.
+We assume that future harvests are discounted relative to current ones at a constant rate of discount $\delta$ and ask for the harvest policy that maximizes total discounted harvest between the current time $t$ and a final time $T$, that is we seek to maximize over choices of harvest $\mathbf{E}_{X_t+1} [ \sum_{t = 0}^{T}  R_t(X_t, h_t) \delta^t]$, where the state dynamics are given by Eqn 1 and $\mathbf{E}$ denotes the expectation over the process stochasticity in the future population state.
 
 In order to find that policy, we introduce the value function$V_t(x_t)$ representing the total discounted catch from time $t$ onwards given that $X_t=x$.
-This value function satisfies an equation of Stochastic Dynamic Programming  [SDP, @Mangel1988; @Clark2000, @Clark2009], sometimes called the Bellman-Hamilton-Jacobi equation [@Mangel2014].
+This value function satisfies an equation of Stochastic Dynamic Programming  [SDP, @Mangel1988; @Clark2000; @Clark2009], sometimes called the Bellman-Hamilton-Jacobi equation [@Mangel2014].
 
 \begin{equation}
-V_t(x) = \max_{h_t} \left[\Pi(h_t, x_t) + \delta \cdot \mathbf{E_Z}  V_{t+1}( Z_{t+1} f(x - h_t)) \right]
+V_t(x) = \max_{h_t} \left[R(h_t, x_t) + \delta \cdot \mathbf{E}_{X_t+1} V_{t+1}( Z_{t+1} f(x - h_t)) \right]
 \end{equation}
 
-That is, at time $t$, when population size is $x$ and harvest $h_t$ is applied, the immediate return is $\Pi(h_t, x_t)$ and the population size after the take is $x-h_t$, which is then translated into $X_{t+1}$ by Eqn 1.
-We consider the finite time problem with $T=$`r Tmax`, which we solve using value iteration algorithm [@Mangel1988; @Clark2000].
+That is, at time $t$, when population size is $x$ and harvest $h_t$ is applied, the immediate return is $R(h_t, x_t)$ and the population size after the take is $x-h_t$, which is then translated into $X_{t+1}$ by Eqn 1.
+We consider the finite time problem with $T=$`r MaxT`, which we solve using value iteration algorithm [@Mangel1988; @Clark2000].
 
@@
 ### The Gaussian Process model
+
+The core difference for our purpose between the GP representation and the parametric
+models we have already introduced is the "non-parametric" property, which
+means that an estimated GP model is defined explicitly in reference to 
+some observed data.  Unlike the models above, it cannot be specified by
+the value of some parameters alone.  The nonparametric moniker for the
+GP approach is unfortunate as (a) it still involves the estimation of
+parameters, and (b), statisticians also use non-parametric to mean models
+that make only weak distributional assumptions  [e.g. @Lehmann1975].
+However, we retain the terminology for consistency with previous studies.
 
@@
 
-The nonparametric moniker for the
-GP approach is unfortunate as (a) it still involves the estimation of
-parameters, and (b), statisticians also use non-parametric to mean models
-that make only weak distributional assumptions  [e.g. @Lehmann1975].
-However, we retain the terminology for consistency with previous studies.
-
-<!-- FIXME Marc suggests putting the full GP formulation in here?? -->
+A GP is an infinite dimensional (because it considers a continuous random
+variable) stochastic process for which any realization is a multivariate
+normal distribution. Thus, to characterize the GP we need a mean function
+and a covariance matrix. We proceed as follows.
+
+Assuming the data $X_o$ are observed with some process noise,
+
+$$y = f(x) + \varepsilon$$
+
+$\varepsilon$ IID normal, variance $\sigma_n^2$. Then under the GP, the predicted data $X_p$ given the observations obeys
+
+$$f(X_p|X_o) \sim \mathcal{N}(E,C)$$
+$$E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y$$
+$$C = K(X_p, X_p) - K(X_p, X_o) K(X_o,X_o)^{-1} K(X_o, X_p)$$
+
+Where $K$ is the Gaussian process kernel.  (Compare to  Eqn 2.22 of @Rasmussen2006 for a more detailed derivation.) Throughout, we use the radial basis function kernel:
+
+$$ K(x,y) = \exp\left(\frac{-(x-y)^2}{2 \ell^2} \right)$$
+
+We use vague inverse Gamma priors on both the lengthscale $\ell$ and process noise $\sigma$ parameters, of the form
+
+$$f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)$$
+
+For the $\sigma$ prior, $\alpha = $ `r s2.p[1]` and $\beta = $ `r s2.p[2]`.  For $\ell$ prior,  $\alpha = $ `r d.p[1]` and $\beta = $ `r d.p[2]`.
 
@@
 It is reasonably straight-forward to derive the harvest policy from the
-estimated BP by inserting it into a SDP algorithm.  Recall that unlike
-the parametric models the Gaussian process with fixed parameters already
-predicts a distribution of curves rather than a single curve. We must
-first integrate over this distribution of curves given a sampling of
-parameter values drawn from the posterior distribution of the two GP
-parameters, before integrating over the posterior of those parameters
-themselves.
+estimated BP by inserting it into a SDP algorithm. The only difference 
+is that the uncertainty in the future state under the GP, $X_{t+1}$,
+includes both process uncertainty (based on the estimation of $\sigma$) 
+and "structural uncertainty" of the posterior collection of curves.  
+(Recall that given the parameters, a parametric model is represented
+by a single curve, while the GP remains a distribution of curves).  
+From the GP posteriors, we can write down the (discritized) transition
+matrix representing the probability of going to each state $X_{t+1}$ 
+given any current state $X_t$ and any harvest $h_t$ (See the function
+`gp_transition_matrix()` in the provided R package). Given this transition
+matrix, we use the same value iteration algorithm as in the parametric
+case to determine the optimal policy.
 
@@
 
-Large noise relative to the distance between the stable and unstable
-point, (or more precisely, between the optimal escapement and the unstable
-point) increases the chance of a stochastic transition.  As a result,
-the impact of using a model which underestimates the risk of harvesting
+Changing the noise or the distance between stable and unstable points
+does not impact the performance of the GP relative to the optimal solution
+given the true model and true parameters.  The parametric models are
+more senstive to this difference.  Large noise relative to the distance between the stable and unstable
+point increases the chance of a stochastic transition.  More precisely,
+if we let $L = K - x_c$, then the probability of a transition in a given
+window of time $T$ scales as
+
+$$P(x_t < x_c | t \in T) \propto  \exp\left(\frac{L^2}{\sigma^2}\right),$$
+
+(following the Ahrenius relationship, see @Gardiner2009 for the derivation).
+As a result, the impact of using a model which underestimates the risk of harvesting
 past the critical point is much more severe, since this such a situation
@@
 chooses to explore or avoid the resulting region of uncertainty can also
-be influenced by the curvature of the value (profit) function $\Pi$. Both
+be influenced by the curvature of the value (profit) function $R$. Both
 to simplify the intuition and avoid biasing this result, we have chosen
@@
 
-Gaussian process model
-----------------------
-
-Assuming the data are known with some process noise,
-
-$$y = f(x) + \varepsilon$$
-
-$\varepsilon$ IID normal, variance $\sigma_n^2$. Then under the GP,
-
-$$f(x) \sim \mathcal{N}(E,C)$$
-$$E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y$$
-$$C = K(X_p, X_p) - K(X_p, X_o) K(X_o,X_o)^{-1} K(X_o, X_p)$$
-
-Where $K$ is the Gaussian process kernel.  Throughout, we use the radial basis function kernel:
-
-$$ K(x,y) = \exp\left(\frac{-(x-y)^2}{2 \ell^2} \right)$$
-
-The marginal likelihood is then given by:
-
-$$\log(p(y | X)) = -\tfrac{1}{2} \mathbf{y}^T (K + \sigma_n^2 \mathbf{I})^{-1} y  - \tfrac{1}{2} \log\left| K + \sigma_n^2 \mathbf{I} \right| - \tfrac{n}{2}\log 2 \pi$$
-
-See @Munch2005a or @Rasmussen2006 for a more detailed introduction.
-
 ```{r figure_S1, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options"}
@@
 
-The Gaussian process priors on both the lengthscale $\ell$ and process noise $\sigma$ parameters are both vague inverse Gamma,
-
-$$f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)$$
-
-For the $\sigma$ prior, $\alpha = $ `r s2.p[1]` and $\beta = $ `r s2.p[2]`.  For $\ell$ prior,  $\alpha = $ `r d.p[1]` and $\beta = $ `r d.p[2]`.
-
 \newpage
-
-Optimal Control Problem Specification 
--------------------------------------
-
-We seek the harvest policy $h(x)$ that maximizes:
-
-$$ \max_{h_t} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h_t) \delta^t  $$
-
-subject to the profit function $\Pi(X_t,h)$, discount rate $\delta$, and the state
-equation
-
-$$X_{t+1} = Z_t f(S_t)  $$
-$$S_t = X_t - h_t $$
-
-For simplicity we assume profit function $\Pi_t(X_t, h_t)$ is linear in the realized harvest, only
-enforcing the restriction that harvest can not exceed available stock, that is $\Pi(h,x) = min(h,x)$. $Z_t$ is as previously defined.
-
-From this we can write down the Bellman recursion as:
-
-$$V_t(x_t) = \max_h \mathbf{E} \left(\Pi(h_t, x_t) + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)$$
-
-Where $V_t$ is the discounted value of the profit given that $X(t) = x$
 

### 2014-05-23 manuscripts/manuscript.tex
updates as sent to steve

@@
             pdfauthor={},
-            pdftitle={Avoiding tipping points in ecosystem based fisheries management: a non-parametric Bayesian approach},
+            pdftitle={Avoiding tipping points in fisheries management: A Bayesian Non-Parametric approach},
             colorlinks=true,
@@
 
-  \title{Avoiding tipping points in ecosystem based fisheries management: a
-non-parametric Bayesian approach}
+  \title{Avoiding tipping points in fisheries management: A Bayesian
+Non-Parametric approach}
     \author[cstar]{Carl Boettiger\corref{c1}}
@@
 95064, USA}    
+    \address[marc]{Center for Stock Assessment Research, Department of Applied Math and
+Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA
+95064, USA and Department of Biology, University of Bergen, Bergen,
+Norway 9020}    
     \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric
@@
   Model uncertainty and limited data coverage are fundamental challenges
-  to robust ecosystem management. These challenges are acutely highlighted
-  by concerns that many ecological systems may contain tipping points.
-  Before a collapse, we do not know where the tipping points lie, if they
-  exist at all. Hence, we know neither a complete model of the system
-  dynamics nor do we have access to data in some large region of
-  state-space where such a tipping point might exist. These two sources of
-  uncertainty frustrate state-of-the-art parametric approaches to decision
-  theory and optimal control. Here we illustrate how a non-parametric
-  approach using a Gaussian process prior provides a more flexible
-  representation of this inherent uncertainty. We can then embed the
-  Gaussian process in a stochastic dynamic programming framework in order
-  to make robust management predictions under both model uncertainty and
+  to robust management of human intervention in a natural system. These
+  challenges are acutely highlighted by concerns that many ecological
+  systems may contain tipping points, such as Allee population
+  sizes.\\Before a collapse, we do not know where the tipping points lie,
+  if they exist at all. Hence, we know neither a complete model of the
+  system dynamics nor do we have access to data in some large region of
+  state-space where such a tipping point might exist. We illustrate how a
+  Bayesian Non-Parametric (BNP) approach using a Gaussian Process (GP)
+  prior provides a flexible representation of this inherent uncertainty.
+  We embed GPs in a stochastic dynamic programming framework in order to
+  make robust management predictions with both model uncertainty and
   limited data. We conduct a set of simulations to evaluate this approach
   as compared with the standard approach of using model selection to
-  choose from a set of candidate models. In keeping with previous results,
-  we find that model selection erroneously favors models without tipping
-  points -- leading to harvest policies that guarantee extinction. The
-  Gaussian process approach performs nearly as well as the true model and
-  significantly outperforms standard appraoches. We illustrate this using
-  examples of simulated single-species dynamics, where the standard model
-  selection approach should be most effective, and find that it still
-  fails to account for uncertainty appropriately and leads to population
-  crashes, while management based on the Gaussian process approach does
-  not. We discuss opportunities to adapt this approach to multiple species
-  models in further work.
+  choose from a set of candidate models. We find that model selection
+  erroneously favors models without tipping points -- leading to harvest
+  policies that guarantee extinction. The GP approach performs nearly as
+  well as the true model and significantly outperforms standard
+  appraoches. We illustrate this using examples of simulated
+  single-species dynamics, where the standard model selection approach
+  should be most effective, and find that it still fails to account for
+  uncertainty appropriately and leads to population crashes, while
+  management based on the GP approach does not, since it does not
+  underestimate the uncertainty outside of the observed data.
   \end{abstract}
@@
 
-Decision making under uncertainty is a ubiquitous challenge of natural
-resource management and conservation. Ecological dynamics are frequently
-complex and difficult to measure, making uncertainty in our
-understanding a persistent challenge to effective management.
+Decision making under uncertainty is a ubiquitous challenge in the
+management of human intervention in natural resources and conservation.
+Ecological dynamics are frequently complex and difficult to measure,
+making uncertainty in our understanding a persistent challenge.
 Decision-theoretic approaches provide a framework to determine the best
 sequence of actions in face of uncertainty, but only when that
-uncertainty can be meaningfully quantified (Fischer \emph{et al.} 2009).
-Concerns about the potential sudden collapse of fisheries and other
-ecosystems (Worm \emph{et al.} 2006; Scheffer \emph{et al.} 2009) has
-increasingly emphasized the difficulties of formulating even
+uncertainty can be meaningfully quantified (Fischer et al. 2009). Over
+the last four decades (but stretching even further back in time), marked
+by the seminal contributions of Clark (1976); Clark (2009) and Walters
+and Hilborn (1978), dynamic optimization methods, particularly
+Stochastic Dynamic Programming (SDP, Clark (2009)), have become
+increasingly important as a means of understanding how to management
+human intervention into natural systems. Concomittantly, there has been
+increasing recognition of the importance of Allee population sizes or
+`tipping points' (Scheffer et al. 2001, Polasky et al. 2011) that
+management actions may inadvertently cause populations to cross.
+Scheffer et al. (2009) emphasized the difficulties of formulating even
 qualitatively correct models of the underlying processes.
@@
 fisheries; though the underlying challenges and methods are germane to
-many other conservation and resource management problems. Economic value
-and ecological concern have made marine fisheries the crucible for much
-of the founding work (Gordon 1954; May \emph{et al.} 1979; Reed 1979;
-Ludwig \& Walters 1982) on management under uncertainty. Global declines
-(Worm \emph{et al.} 2006) and the controversy surrounding their
-interpretation (Hilborn 2007; Worm \emph{et al.} 2009) have made
-understanding these challenges all the more pressing.
+many other problems of conservation or natural resource exploitation,
+such as forestry. Economic value and ecological concern have made marine
+fisheries the crucible for much of the founding work (Gordon 1954, Clark
+1976, 2009, May et al. 1979, Reed 1979, Ludwig and Walters 1982) on
+management under uncertainty. Global declines (Worm et al. 2006) and the
+controversy surrounding their interpretation (Hilborn 2007, Worm et al.
+2009) have made understanding these challenges all the more pressing.
 
@@
 stochasticity in biological processes, measurements, and implementation
-of policy (\emph{e.g.} Reed 1979; Clark \& Kirkwood 1986; Roughgarden \&
-Smith 1996; Sethi \emph{et al.} 2005), parametric uncertainty
-(\emph{e.g.} Ludwig \& Walters 1982; Hilborn \& Mangel 1997; McAllister
-1998; Schapaugh \& Tyre 2013), and model or structural uncertainty
-(\emph{e.g.} Williams 2001; Cressie \emph{et al.} 2009; Athanassoglou \&
+of policy (\emph{e.g.} Reed 1979, Clark and Kirkwood 1986, Roughgarden
+and Smith 1996, Sethi et al. 2005), parametric uncertainty (\emph{e.g.}
+Ludwig and Walters 1982, Hilborn and Mangel 1997, McAllister 1998,
+Schapaugh and Tyre 2013), and model or structural uncertainty
+(\emph{e.g.} Williams 2001, Cressie et al. 2009, Athanassoglou and
 Xepapadeas 2012). Of these, structural uncertainty is generally the
@@
 the dynamics must be identified from among a handful of alternative
-models. Here we consider an approach that addresses uncertainty at each
-of these levels without assuming the dynamics follow a particular
-(i.e.~parametric) structure.
-
-Often we do not understand the system dynamics well enough to know if a
-model provides a good approximation over the relevant range of states
-and timescales (criteria that we loosely refer to as defining the
-``right'' or ``true'' model.)\\Typical solutions involve either model
-choice, model averaging, or introducing yet greater model complexity of
-which others may be special cases (model averaging being one such way to
-construct such a model) (Williams 2001; Cressie \emph{et al.} 2009;
-Athanassoglou \& Xepapadeas 2012). Even setting aside other
-computational and statistical concerns (e.g. (Cressie \emph{et al.}
-2009)), these approaches do not address our second concern -
+models, using either model choice, model averaging, or introducing yet
+greater model complexity of which others may be special cases (model
+averaging being one such way to construct such a model) (Williams 2001,
+Cressie et al. 2009, Athanassoglou and Xepapadeas 2012). Even setting
+aside other computational and statistical concerns (e.g. Cressie et al.
+(2009)), these approaches do not address our second concern -
 representing uncertainty outside the observed data range.
@@
 Model uncertainty is particularly insidious when model predictions must
-be made outside of the range of data on which the model was estimated.
-This extrapolation uncertainty is felt most keenly in decision-theoretic
-(or optimal control) applications, since (a) exploring the potential
-action space typically involves considering actions that may move the
-system outside the range of observed behavior, and (b)
-decision-theoretic algorithms rely not only on reasonable estimates of
-the expected outcomes, but depend on the weights given to all possible
-outcomes (\emph{e.g.} Weitzman 2013).
-
-This difficult position of having neither the true model nor data that
-covers the full range of possible states is unfortunately the rule more
-than the exception. The potential concern of tipping points in
-ecological dynamics (Scheffer \emph{et al.} 2001; Polasky \emph{et al.}
-2011) reflects these concerns -- as either knowledge of the true model
-or more complete sampling of the state space would make it easy to
-identify if a tipping point existed. If we do not know but cannot rule
-out such a possibility, then we face decision-making under this dual
-challege of model uncertainty and incomplete data coverage.
-
-These dual concerns pose a substantial challenge to existing
-decision-theoretic approaches (Brozović \& Schlenker 2011). Because
-intervention is often too late after a tipping point has been crossed
-(but see Hughes et al. (2013)), management is most often concerned with
-avoiding potentially catastrophic tipping points before any data is
-available at or following a transition that would more clearly reveal
-these regime shift dynamics (e.g. Bestelmeyer \emph{et al.} 2012).
-
-Here we illustrate how a stochastic dynamic programming (SDP) algorithm
-(Mangel \& Clark 1988; Marescot \emph{et al.} 2013) can be driven by the
-predictions from a Bayesian non-parametric (BNP) approach (Munch
-\emph{et al.} 2005a). This provides two distinct advantages compared
-with contemporary approaches. First, using a BNP sidesteps the need for
-an accurate model-based description of the system dynamics. Second, the
-BNP can better reflect uncertainty that arises when extrapolating a
-model outside of the data on which it was fit. We illustrate that when
-the correct model is not known, this latter feature is crucial to
-providing a robust decision-theoretic approach in face of substantial
-structural uncertainty.
+be made outside of the range of data on which the model was estimated,
+and models may be abused (Mangel et al. 2001). Extrapolation uncertainty
+is felt most keenly in decision-theoretic (or optimal control)
+applications, since (a) exploring the potential action space typically
+involves considering actions that may move the system outside the range
+of observed behavior, and (b) decision-theoretic algorithms rely not
+only on reasonable estimates of the expected outcomes, but depend on the
+weights given to all possible outcomes (\emph{e.g.} Weitzman 2013).
+
+The dual concerns of model uncertainty and incomplete data coverage pose
+a substantial challenge to existing decision-theoretic approaches
+(Brozović and Schlenker 2011). Because intervention is often (but not
+always, see Hughes et al. (2013)) too late after a tipping point has
+been crossed, management is most often concerned with avoiding
+potentially catastrophic tipping points before any data is available at
+or following a transition that would more clearly reveal these regime
+shift dynamics (e.g. Bestelmeyer et al. 2012).
+
+We illustrate how a stochastic dynamic programming (SDP) algorithm
+(Mangel and Clark 1988, Marescot et al. 2013) can be driven by the
+predictions from a Bayesian Non-Parametric (BNP) model of population
+dynamics (Munch et al. 2005a). This provides two distinct advantages
+compared with contemporary approaches. First, using a BNP sidesteps the
+need for an accurate model-based description of the system dynamics.
+Second, the BNP can better reflect uncertainty that arises when
+extrapolating a model outside of the data on which it was fit. We
+illustrate that when the correct model is not known, this latter feature
+is crucial to providing a robust decision-theoretic approach in face of
+substantial structural uncertainty.
 
@@
 greater uncertainty far from the observed data than near the observed
-data. These features allow the GPDP approach to find robust management
-solutions in face of limited data and without knowledge of the correct
-model structure.
-
-The idea that any approach can perform well without either having to
-know the model or have particularly good data should immediately draw
-suspicion. The reader must bear in mind that the strength of our
-approach comes not from black-box predictive power from such limited
-information, but rather, by providing a more honest expression of
-uncertainty outside the observed data without sacrificing the predictive
-capacity near the observed data. By coupling this more accurate
-description of what is known and unknown to the decision-making under
-uncertainty framework provided by stochastic dynamic programming, we are
-able to obtain more robust management policies than with common
-parametric modeling approaches.
-
-The nature of decision-making problems provides a compelling and
-pragmatic way to compare models: Rather than compare models in terms of
-best fit to data, model performance is defined in the concrete terms of
-the decision-maker's objectives. Although much argument can be made over
-the `correct' objective function, e.g.~how to account for the social
-value of fish left in the sea vs.~the commercial value of fish
-harvested; see Halpern et al. (2013) for further discussion of this
-issue. We can always compare model performance across multiple potential
-objectives. The important point is that a decision-maker does not
-necessarily need a model that provides the best mechanistic
-understanding or the best long-term outcome, but rather the one that
-idelivers the best management performance over the period of interest.
-
-\subsection{Background on the Gaussian
-Process}\label{background-on-the-gaussian-process}
-
-We use a Bayesian nonparameteric regression approach based on Gaussian
-process priors. The use of Gaussian process (GP) regression (or
-``kriging'' in the geospatial literature) to formulate a predictive
-model is relatively new in the context of modeling dynamical systems
-(Kocijan \emph{et al.} 2005), and was first introduced in the context
-ecological modeling and fisheries management in Munch et al. (2005b). GP
-models have subsequently been used to test for the presence of Allee
-effects (Sugeno \& Munch 2013a), estimate the maximum reproductive rate
-(Sugeno \& Munch 2013b), determine temporal variation in food
-availability (Sigourney \emph{et al.} 2012), and provide a basis for
-identifying model-misspecification (Thorson \emph{et al.} 2014). An
-accessible and thorough introduction to the formulation and use of GPs
-can be found in Rasmussen \& Williams (2006). The nonparametric moniker
-for the GP approach is unfortunate as (a) it still involves the
-estimation of parameters, and (b), statisticians also use non-parametric
-to mean models that make only weak distributional assumptions (e.g.
-Lehmann 1975). However, we retain the terminology for consistency with
-previous studies.
+data. These features allow the Gaussian Process Dynamic Programming
+(GPDP) approach to find robust management solutions in face of limited
+data and without knowledge of the correct model structure.
+
+To illustrate GPDP, we consider the performance of the GPDP policy
+against the policies derived under several alternative parametric models
+(Reed 1979, Ludwig and Walters 1982, Mangel and Clark 1988). The nature
+of decision-making problems provides a compelling and pragmatic way to
+compare models: Rather than compare models in terms of best fit to data,
+we define model performance in the concrete terms of the
+decision-maker's objectives.
 
@@
 
-\subsubsection{Statement of the optimal control
-problem}\label{statement-of-the-optimal-control-problem}
-
-To illustrate the application of the Gaussian process-based dynamic
-programming approach (hereafter GPDP) we focus on the classical problem
-of determining a harvest policy that will maximize the expected
-discounted profit over a fixed time interval (Reed 1979; Ludwig \&
-Walters 1982; Mangel \& Clark 1988). For simplicity we assume profit is
-proportional to the realized harvest.\\For comparison, we consider the
-performance of the GPDP policy against policies derived under several
-alternative parametric models.
-
-We focus on a class of one-dimensional models of population dynamics as
-a `best-case' scenario for the parametric model performance relative to
-the Gaussian process. We consider this a best-case scenario for the
-following reasons: first, the underlying dynamics are always simuated
-from one the parametric models in the set, whereas in reality we never
-have the ``true'' model among our candidates. By choosing
-one-dimensional models with few parameters, we limit the chance that
-poor performance will be due to our inability to estimate parameters
-well, something that becomes a more severe problem for
-higher-dimensional parametric models. Further, the parametric models we
-consider are those most commonly used in modeling stock-recruitment
+We first describe the requirements of dynamic optimization for the
+management of human intervention in natural resource systems. After that
+we describe three parametric models for population dynamics and the GP
+description of population dynamics.
+
+\subsubsection{Requirements of dynamic
+optimization}\label{requirements-of-dynamic-optimization}
+
+Dynamic optimization requires characterizing the dynamics of a state
+variable (or variables), a control action, and a value function. In this
+paper, we consider only a single state variable, so that the ideas can
+be described as simply as possible. That is, we focus on a class of
+one-dimensional models of population dynamics as a `best-case' scenario
+for the parametric model performance relative to the GP. This is
+abest-case scenario for the parametric models because the underlying
+dynamics are always simuated from one the parametric models in the set,
+whereas in reality we never have the ``true'' model among our
+candidates. By choosing one-dimensional models with few parameters, we
+limit the chance that poor performance will be due to our inability to
+estimate parameters well, something that becomes a more severe problem
+for higher-dimensional parametric models. Further, the parametric models
+we consider are those most commonly used in modeling stock-recruitment
 dynamics or to model sudden transitions between alternative stable
@@
 
+We let $X_t$ denote the size (numbers or biomass) of the focal
+population at time $t$ and assume that in the absence of take its
+dynamics are
+
+\[ X_{t+1}= Z_t f(X_t, h_t, p) \]
+
+Where Z(t) is log-normally distributed process stochasticity (Reed 1979)
+and $p$ is a vector of parameters to be estimated from the data. We
+describe three choices for $f(X_t,p)$ in the next section.In this simple
+model, the control action is a harvest or take, $h_t$, measured in the
+same units as $X$, at time $t$. Thus, in the presence of take, the
+population size on the right hand side of Eqn 1 is replaced by
+$S_t=X_t-h_t$.
+
+To construct the value function, we consider a return when $X_t=x$ and
+harvest $h_t=h$ denoted as the reward, $R(x,h)$. For example, if the
+return is simply the harvest at time $t$, then $R(x,h)=min(x,h)$. We
+assume that future harvests are discounted relative to current ones at a
+constant rate of discount $\delta$ and ask for the harvest policy that
+maximizes total discounted harvest between the current time $t$ and a
+final time $T$, that is we seek to maximize over choices of harvest
+$\mathbf{E}_{X_t+1} [ \sum_{t = 0}^{T}  R_t(X_t, h_t) \delta^t]$, where
+the state dynamics are given by Eqn 1 and $\mathbf{E}$ denotes the
+expectation over the process stochasticity in the future population
+state.
+
+In order to find that policy, we introduce the value function$V_t(x_t)$
+representing the total discounted catch from time $t$ onwards given that
+$X_t=x$. This value function satisfies an equation of Stochastic Dynamic
+Programming (SDP, Mangel and Clark 1988, Clark and Mangel 2000, Clark
+2009), sometimes called the Bellman-Hamilton-Jacobi equation (Mangel
+2014).
+
+\begin{equation}
+V_t(x) = \max_{h_t} \left[R(h_t, x_t) + \delta \cdot \mathbf{E}_{X_t+1} V_{t+1}( Z_{t+1} f(x - h_t)) \right]
+\end{equation}
+
+That is, at time $t$, when population size is $x$ and harvest $h_t$ is
+applied, the immediate return is $R(h_t, x_t)$ and the population size
+after the take is $x-h_t$, which is then translated into $X_{t+1}$ by
+Eqn 1. We consider the finite time problem with \$T=\$1000, which we
+solve using value iteration algorithm (Mangel and Clark 1988, Clark and
+Mangel 2000).
+
+\subsubsection{Parametric models}\label{parametric-models}
+
 We consider three candidate parametric models of the stock-recruitment
-dynamics: The Ricker model, the Allen model (Allen \& Tanner 2005), the
-Myers model (Myers \emph{et al.} 1995), equations (1)-(3), which we
-parameterize as follows:
-
-We consider three parametric models that relate population size at time
-$t+1$, $X_{t+1}$, to population size at time $t$, $X_t$ and harvest at
-time $t$, $h_t$ by
-
-\[ X_{t+1}= Z_t f(S_t,p) \]
-
-Where Z(t) is log-normally distributed process stochasticity (Reed 1979)
-and $p$ is a vector of parameters, and $S_t$ the escapement (Reed 1979)
-
-\[S_t = X_t - h_t\]
-
-or the Ricker model,
+dynamics: The Ricker model, the Allen model (Allen and Tanner 2005), the
+Myers model (Myers et al. 1995), equations (1)-(3), which we
+parameterize as follows for $f(S_t,p)$ in Eqn 1. In all three, we let
+$K$ denote the carrying capacity and $r$ the maximum per capita growth
+rate. In the Ricker model,
 
@@
 
-where $K$ is the carrying capacity and $r$ the growth rate. For the
-Allen model,
+In the Allen model,
 
@@
 \item
-  \[ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_C\right)} \]
+  \[ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_c\right)} \]
 \end{enumerate}
@@
 Where $r$ and $K$ are as before, and $X_c$ denotes the location of the
-unstable (tipping) point. For the Myers model,
+unstable steady state (i.e., the tipping point).
+
+In the Myers model,
 
@@
 \item
-  \[ f(S(t) | p)  = \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}}\]
+  \[ f(S(t) | p)  = \frac{r S_t^{\theta}}{1 + \frac{S_t^\theta}{K}}\]
 \end{enumerate}
 
-Where $r$ and $K$ as before, and $\theta$ the nonlinearity of the Myers
-model (corresponding to Beverton-Holt dynamics for $\theta$ = 1 and an
-Allee effect with alternative stable states exist at $\theta > 2$).
-
-The familiar Ricker model involves two parameters, corresponding to a
-growth rate and a carrying capacity, and cannot support alternative
-stable growth rate dynamics. The Allen model resembles the Ricker
-dynamics with an added Allee effect parameter (Courchamp \emph{et al.}
-2008), below which the population cannot persist. The Myers model also
-has three parameters and contains an Allee threshold, but has
-compensatory rather than over-compensatory density dependence
-(resembling a Beverton-Holt curve rather than a Ricker curve at high
-densities.)\\We assume multiplicative log-normal noise perturbs the
-growth predicted by the each of the deterministic model skeletons
-described above. This introduces one additional parameter $\sigma$ that
-must be estimated by each model.
+Where $\theta = 1$ corresponding to Beverton-Holt dynamics and $\theta$
+\textgreater{} 2 leads to Allee effects and multiple stable states.
+
+The Ricker model involves two parameters, corresponding to a growth rate
+and a carrying capacity, and cannot support alternative stable growth
+rate dynamics. The Allen model resembles the Ricker dynamics with an
+added Allee effect parameter (Courchamp et al. 2008), below which the
+population cannot persist. The Myers model also has three parameters and
+contains an Allee threshold, but has compensatory rather than
+over-compensatory density dependence (resembling a Beverton-Holt curve
+rather than a Ricker curve at high densities.)
+
+The multiplicative log-normal stochasticity perturbs the growth
+predicted by each of these deterministic model skeletons. This
+introduces one additional parameter $\sigma$ that must be estimated.
 
@@
 tipping points, all of our simulations are based on the Allen model. The
-Allen model is thus structurally correct and is expected to provide a
-best-case scenario when the `true' dynamics are known. The Ricker model
-is a reasonable approximation of these dynamics far from the Allee
+Allen model is thus ``structurally correct'' and is expected to provide
+a best-case scenario when the `true' dynamics are known. The Ricker
+model is a reasonable approximation of these dynamics far from the Allee
 threshold (but lacks threshold dynamics), while the Myers model shares
@@
 We consider a period of 40 years of training data: long enough that the
-estimates are not dependent on the particular realization, while longer
+estimates do not depend on the particular realization, while longer
 times are not likely to provide substantial improvement (i.e.~the
 results are not sensitive to this interval). Each of the models
-(described below) is fit to the same training data, as shown in Figure
-1.
-
-We inferred posterior distributions for the parameters of each model in
-a Bayesian context using Gibbs sampling (Gelman et al. (2003)
-implemented in R (R Core Team 2013) using jags, (Su \& Masanao Yajima
-2013)). We choose uninformative uniform priors for all parameters (See
-appendix Tables S1-S3; R code provided). We show one-step-ahead
-predictions of these model fits in Figure 1. Each chain was tested for
-Gelman-Rubin convergence and results were robust to longer runs. For
-each simulation we also applied several commonly used model selection
-criteria (AIC, BIC, DIC, see Burnham \& Anderson (2002)) to identify the
-best fitting model.
-
-An optimal policy function was then inferred through stochastic dynamic
-programming for each model given the posterior distributions of the
-parameter estimates. This policy maximizes the expectation of the value
-function integrated over the parameter uncertainty. (code implementing
-this algorithm provided in the Appendix).
+(described below) is fit to the same training data (Figure 1).
+
+We inferred posterior distributions for the parameters of each model by
+Gibbs sampling (Gelman et al. (2003) implemented in R (R Core Team 2013)
+using jags, (Su and Masanao Yajima 2013)). We choose uniform priors for
+all parameters (See appendix Tables S1-S3; R code provided). We show
+one-step-ahead predictions of these model fits in Figure 1. We tested
+each chain for Gelman-Rubin convergence and results were robust to
+longer runs. For each simulation we also applied several commonly used
+model selection criteria (AIC, BIC, DIC, see Burnham and Anderson
+(2002)) to identify the best fitting model.
 
@@
 model}\label{the-gaussian-process-model}
+
+The core difference for our purpose between the GP representation and
+the parametric models we have already introduced is the
+``non-parametric'' property, which means that an estimated GP model is
+defined explicitly in reference to some observed data. Unlike the models
+above, it cannot be specified by the value of some parameters alone. The
+nonparametric moniker for the GP approach is unfortunate as (a) it still
+involves the estimation of parameters, and (b), statisticians also use
+non-parametric to mean models that make only weak distributional
+assumptions (e.g. Lehmann 1975). However, we retain the terminology for
+consistency with previous studies.
+
+The use GP methods to formulate a predictive model is relatively new in
+the context of modeling dynamical systems (Kocijan et al. 2005), and was
+first introduced in the context ecological modeling and fisheries
+management in Munch et al. (2005b). GP models have subsequently been
+used to test for the presence of Allee effects (Sugeno and Munch 2013a),
+estimate the maximum reproductive rate (Sugeno and Munch 2013b),
+determine temporal variation in food availability (Sigourney et al.
+2012), and provide a basis for identifying model-misspecification
+(Thorson et al. 2014). An accessible and thorough introduction to the
+formulation and use of GPs can be found in Rasmussen and Williams
+(2006).
+
+A GP is an infinite dimensional (because it considers a continuous
+random variable) stochastic process for which any realization is a
+multivariate normal distribution. Thus, to characterize the GP we need a
+mean function and a covariance matrix. We proceed as follows.
+
+Assuming the data $X_o$ are observed with some process noise,
+
+\[y = f(x) + \varepsilon\]
+
+$\varepsilon$ IID normal, variance $\sigma_n^2$. Then under the GP, the
+predicted data $X_p$ given the observations obeys
+
+\[f(X_p|X_o) \sim \mathcal{N}(E,C)\]
+\[E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y\]
+\[C = K(X_p, X_p) - K(X_p, X_o) K(X_o,X_o)^{-1} K(X_o, X_p)\]
+
+Where $K$ is the Gaussian process kernel. (Compare to Eqn 2.22 of
+Rasmussen and Williams (2006) for a more detailed derivation.)
+Throughout, we use the radial basis function kernel:
+
+\[ K(x,y) = \exp\left(\frac{-(x-y)^2}{2 \ell^2} \right)\]
+
+We use vague inverse Gamma priors on both the lengthscale $\ell$ and
+process noise $\sigma$ parameters, of the form
+
+\[f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)\]
+
+For the $\sigma$ prior, $\alpha = $ 5 and $\beta = $ 5. For $\ell$
+prior, $\alpha = $ 10 and $\beta = $ 10.
 
@@
 
+\subsection{The method of Gaussian Process Dynamic
+Programming}\label{the-method-of-gaussian-process-dynamic-programming}
+
 It is reasonably straight-forward to derive the harvest policy from the
-estimated BP by inserting it into a SDP algorithm. Recall that unlike
-the parametric models the Gaussian process with fixed parameters already
-predicts a distribution of curves rather than a single curve. We must
-first integrate over this distribution of curves given a sampling of
-parameter values drawn from the posterior distribution of the two GP
-parameters, before integrating over the posterior of those parameters
-themselves.
+estimated BP by inserting it into a SDP algorithm. The only difference
+is that the uncertainty in the future state under the GP, $X_{t+1}$,
+includes both process uncertainty (based on the estimation of $\sigma$)
+and ``structural uncertainty'' of the posterior collection of
+curves.\\(Recall that given the parameters, a parametric model is
+represented by a single curve, while the GP remains a distribution of
+curves).\\From the GP posteriors, we can write down the (discritized)
+transition matrix representing the probability of going to each state
+$X_{t+1}$ given any current state $X_t$ and any harvest $h_t$ (See the
+function \texttt{gp\_transition\_matrix()} in the provided R package).
+Given this transition matrix, we use the same value iteration algorithm
+as in the parametric case to determine the optimal policy.
 
 \section{Results}\label{results}
+
+\subsection{Parametric and GP models for population
+dynamics}\label{parametric-and-gp-models-for-population-dynamics}
 
@@
 \includegraphics{components/figure/manuscript-figure_3.pdf}
-\caption{Beyond the tipping point (outside the range of the training
-data), the true dynamics fall outside the uncertainty bands of the wrong
-parametric models, but inside the uncertainty predicted by the GP.
+\caption{Outside the range of the training data (Figure 1), the true
+dynamics (black dots) fall outside the uncertainty (two standard
+deviations, colored bands) of the structurally incorrect parametric
+models (Ricker, Myers), but inside the uncertainty predicted by the GP.
 Points show the stock size simulated by the true model. Overlay shows
@@
 step as the initial condition falls below the Allee threshold of the
-underlying model (Allen). Note that the GP is both more pessimistic and
-more uncertain about the future state than the parametric models.}
+underlying model (Allen). Note that because it does not assume a
+parametric form but instead relies more directly on the data, the GP is
+both more pessimistic and more uncertain about the future state than the
+parametric models.}
 \end{figure}
@@
 \includegraphics{components/figure/manuscript-figure_5.pdf}
-\caption{Gaussian process inference outperforms parametric estimates.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under
-the policies derived from each of the estimated models, as well as the
-policy based on the exact underlying model.}
+\caption{In the management context, GPDP outperforms approaches based on
+parametric models. Shown are 100 replicate simulations of the stock
+dynamics (eq 1) under the policies derived from each of the estimated
+models, as well as the policy based on the exact underlying model.}
 \end{figure}
@@
 compared by using the relative differences in net present value realized
-by each of the model. Figure 6 shows the distribution of values
-(profits) realized by each simulation shown in Figure 5. The GPDP
-approach most consistently realizes a value close to the optimal
-solution, and importantly avoids ever driving the system across the
-tipping point, which results in the near-zero value cases in the
-parametric models.
+by each of the model. Figure 6 shows the distribution of values realized
+by each simulation shown in Figure 5. The GPDP most consistently
+realizes a value close to the optimal solution, and importantly avoids
+ever driving the system across the tipping point, which results in the
+near-zero value cases in the parametric models.
 
@@
 Myers) over two different noise levels (0.01 and 0.05), for a total of 8
-x 3 x 2 x 2 = 96 scenarios. In Figure 7, results are pooled across
+x 3 x 2 x 2 = 96 scenarios. In Figure 7, we pool results across
 different random seeds (111, 222, 333), noise values of 0.01 and 0.05
@@
 to harvest at all or crashing the stock, which results in the very low
-profits/values seen in parametric models in the previous figure. The
-sensitivity analysis is compared against the optimal solution alone and
-not shown against the estimates of the various parametric models (but
-see Figure 6), where convergence of the Gibbs sampling could contribute
-to their poorer performance if not checked by hand for each
-configuration.
-
-Large noise relative to the distance between the stable and unstable
-point, (or more precisely, between the optimal escapement and the
-unstable point) increases the chance of a stochastic transition. As a
-result, the impact of using a model which underestimates the risk of
-harvesting past the critical point is much more severe, since this such
-a situation occurs more often. Conversely, with large enough distance
-between the optimal escapement and unstable points relative to the
-noise, the chance of a transition becomes vanishingly small and all
-models can be modeled near-optimally. Models which underestimate the
-cost incurred by stock sizes fluctuating significantly below the optimal
-escapement level will not perform poorly as long as those fluctuations
-are sufficently rare.
+values seen in parametric models in the previous figure. The sensitivity
+analysis is compared against the optimal solution alone and not shown
+against the estimates of the various parametric models (but see Figure
+6), where convergence of the Gibbs sampling could contribute to their
+poorer performance if not checked by hand for each configuration.
+
+Changing the noise or the distance between stable and unstable points
+does not impact the performance of the GP relative to the optimal
+solution given the true model and true parameters. The parametric models
+are more senstive to this difference. Large noise relative to the
+distance between the stable and unstable point increases the chance of a
+stochastic transition. More precisely, if we let $L = K - x_c$, then the
+probability of a transition in a given window of time $T$ scales as
+
+\[P(x_t < x_c | t \in T) \propto  \exp\left(\frac{L^2}{\sigma^2}\right),\]
+
+(following the Ahrenius relationship, see Gardiner (2009) for the
+derivation). As a result, the impact of using a model which
+underestimates the risk of harvesting past the critical point is much
+more severe, since this such a situation occurs more often. Conversely,
+with large enough distance between the optimal escapement and unstable
+points relative to the noise, the chance of a transition becomes
+vanishingly small and all models can be modeled near-optimally. Models
+that underestimate the cost incurred by stock sizes fluctuating
+significantly below the optimal escapement level will not perform poorly
+as long as those fluctuations are sufficently rare.
 
@@
 \caption{Sensitivity Analysis. Histogram shows the ratio of the realized
-net present value / profits derived when managing under the GPDP
-Gaussian process approach over the optimal value given the true model
-and true parameters. Values near 1 indicate nearly optimal performance.
-Simulations cover a range of underlying parameter, model, replicate, and
-noise levels.}
+net present value derived when managing under the GPDP over the optimal
+value given the true model and true parameters. Values near 1 indicate
+nearly optimal performance. Simulations cover a range of underlying
+parameter, model, replicate, and noise levels.}
 \end{figure}
@@
 potential to increase our basic understanding of ecological processes
-(Geritz \& Kisdi 2012; Cuddington \emph{et al.} 2013), such models can
-be not only inaccurate but misleading when relied upon in a quantitative
-decision making framework. In this paper we have tackled two aspects of
+(Geritz and Kisdi 2012, Cuddington et al. 2013), such models can be both
+inaccurate and misleading when relied upon in a quantitative decision
+making framework. In this paper we have tackled two aspects of
 uncertainty that are both common to many ecological decision-making
@@
 
-\subsubsection{GPDP expresses larger uncertainty in regions where the
-data are
-poor}\label{gpdp-expresses-larger-uncertainty-in-regions-where-the-data-are-poor}
+\subsubsection{GPDP population dynamcis capture larger uncertainty in
+regions where the data are
+poor}\label{gpdp-population-dynamcis-capture-larger-uncertainty-in-regions-where-the-data-are-poor}
 
@@
 $\sigma$. In principle, this may allow a manager to improve the
-performance of the GPDP approach by adding only enough additional detail
-as is justified. For instance, it would be possible to use a linear or a
+performance of the GPDP by adding only enough additional detail as is
+justified. For instance, it would be possible to use a linear or a
 Ricker-shaped mean in the prior without making the much stronger
-assumption that the Ricker is the structurally correct model (Sugeno \&
+assumption that the Ricker is the structurally correct model (Sugeno and
 Munch 2013a). Future research should focus on identifying criteria to
@@
 
-\subsection{Gaussian process model}\label{gaussian-process-model}
-
-Assuming the data are known with some process noise,
-
-\[y = f(x) + \varepsilon\]
-
-$\varepsilon$ IID normal, variance $\sigma_n^2$. Then under the GP,
-
-\[f(x) \sim \mathcal{N}(E,C)\]
-\[E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y\]
-\[C = K(X_p, X_p) - K(X_p, X_o) K(X_o,X_o)^{-1} K(X_o, X_p)\]
-
-Where $K$ is the Gaussian process kernel. Throughout, we use the radial
-basis function kernel:
-
-\[ K(x,y) = \exp\left(\frac{-(x-y)^2}{2 \ell^2} \right)\]
-
-The marginal likelihood is then given by:
-
-\[\log(p(y | X)) = -\tfrac{1}{2} \mathbf{y}^T (K + \sigma_n^2 \mathbf{I})^{-1} y  - \tfrac{1}{2} \log\left| K + \sigma_n^2 \mathbf{I} \right| - \tfrac{n}{2}\log 2 \pi\]
-
-See Munch et al. (2005a) or Rasmussen \& Williams (2006) for a more
-detailed introduction.
-
 \begin{figure}[htbp]
@@
 
-The Gaussian process priors on both the lengthscale $\ell$ and process
-noise $\sigma$ parameters are both vague inverse Gamma,
-
-\[f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)\]
-
-For the $\sigma$ prior, $\alpha = $ 5 and $\beta = $ 5. For $\ell$
-prior, $\alpha = $ 10 and $\beta = $ 10.
-
 \newpage
-
-\subsection{Optimal Control Problem
-Specification}\label{optimal-control-problem-specification}
-
-We seek the harvest policy $h(x)$ that maximizes:
-
-\[ \max_{h_t} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h_t) \delta^t  \]
-
-subject to the profit function $\Pi(X_t,h)$, discount rate $\delta$, and
-the state equation
-
-\[X_{t+1} = Z_t f(S_t)  \] \[S_t = X_t - h_t \]
-
-For simplicity we assume profit function $\Pi_t(X_t, h_t)$ is linear in
-the realized harvest, only enforcing the restriction that harvest can
-not exceed available stock, that is $\Pi(h,x) = min(h,x)$. $Z_t$ is as
-previously defined.
-
-From this we can write down the Bellman recursion as:
-
-\[V_t(x_t) = \max_h \mathbf{E} \left(\Pi(h_t, x_t) + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)\]
-
-Where $V_t$ is the discounted value of the profit given that $X(t) = x$
 
@@
 
-1.Allen, D. \& Tanner, K. \emph{et al.} (2005). Infusing active learning
-into the large-enrollment biology class: seven strategies, from the
-simple to complex. \emph{Cell biology education}, 4, 262--8.
-
-2.Athanassoglou, S. \& Xepapadeas, A. \emph{et al.} (2012). Pollution
-control with uncertain stock dynamics: When, and how, to be precautious.
-\emph{Journal of Environmental Economics and Management}, 63, 304--320.
-
-3.Bestelmeyer, B.T., Duniway, M.C., James, D.K., Burkett, L.M. \&
-Havstad, K.M.\emph{et al.} (2012). A test of critical thresholds and
-their indicators in a desertification-prone ecosystem: more resilience
-than we thought. \emph{Ecology Letters}, n/a--n/a.
-
-4.Brozović, N. \& Schlenker, W. \emph{et al.} (2011). Optimal management
-of an ecosystem with an unknown threshold. \emph{Ecological Economics},
-1--14.
-
-5.Burnham, K.P. \& Anderson, D.R. \emph{et al.} (2002). \emph{Model
-Selection and Multi-Model Inference}. Springer.
-
-6.Clark, C.W. \& Kirkwood, G.P. \emph{et al.} (1986). On uncertain
-renewable resource stocks: Optimal harvest policies and the value of
-stock surveys. \emph{Journal of Environmental Economics and Management},
-13, 235--244.
-
-7.Courchamp, F., Berec, L. \& Gascoigne, J.\emph{et al.} (2008).
-\emph{Allee Effects in Ecology and Conservation}. Oxford University
-Press, USA.
-
-8.Cressie, N., Calder, C. a, Clark, J.S., {Ver Hoef}, J.M. \& Wikle,
-C.K.\emph{et al.} (2009). Accounting for uncertainty in ecological
-analysis: the strengths and limitations of hierarchical statistical
-modeling. \emph{Ecological Applications}, 19, 553--70.
-
-9.Cuddington, K.M., Fortin, M. \& Gerber, L.\emph{et al.} (2013).
-Process-based models are required to manage ecological systems in a
-changing world. \emph{Ecosphere}, 4, 1--12.
-
-10.Fischer, J., Peterson, G.D., Gardner, T. a, Gordon, L.J., Fazey, I.
-\& Elmqvist, T.\emph{et al.} (2009). Integrating resilience thinking and
-optimisation for conservation. \emph{Trends in ecology \& evolution},
-24, 549--54.
-
-11.Gelman, A., Carlin, J.B., Stern, H.S. \& Rubin, D.B.\emph{et al.}
-(2003). \emph{Bayesian Data Analysis}. 2nd edn. Chapman; Hall/CRC.
-
-12.Geritz, S.A.H. \& Kisdi, E. \emph{et al.} (2012). Mathematical
-ecology: why mechanistic models? \emph{Journal of mathematical biology},
-65, 1411--5.
-
-13.Gordon, H. \emph{et al.} (1954). The economic theory of a
-common-property resource: the fishery. \emph{The Journal of Political
-Economy}, 62, 124--142.
-
-14.Halpern, B.S., Klein, C.J., Brown, C.J., Beger, M., Grantham, H.S. \&
-Mangubhai, S.\emph{et al.} (2013). Achieving the triple bottom line in
-the face of inherent trade-offs among social equity, economic return,
-and conservation. \emph{Proceedings of the National Academy of Sciences
-of the United States of America}, 110, 6229--34.
-
-15.Hilborn, R. \emph{et al.} (2007). Reinterpreting the State of
-Fisheries and their Management. \emph{Ecosystems}, 10, 1362--1369.
-
-16.Hilborn, R. \& Mangel, M. \emph{et al.} (1997). \emph{The Ecological
-Detective: Confronting Models with data}. Princeton University Press.
-
-17.Hughes, T.P., Linares, C., Dakos, V., Leemput, I. a van de \& Nes,
-E.H. van\emph{et al.} (2013). Living dangerously on borrowed time during
-slow, unrecognized regime shifts. \emph{Trends in ecology \& evolution},
-28, 149--55.
-
-18.Kocijan, J., Girard, A., Banko, B. \& Murray-Smith, R.\emph{et al.}
-(2005). Dynamic systems identification with Gaussian processes.
-\emph{Mathematical and Computer Modelling of Dynamical Systems}, 11,
-411--424.
-
-19.Lehmann, E.L. \emph{et al.} (1975). \emph{Nonparametrics: Statistical
-Methods Based on Ranks}. Holden-Day, Inc, San Francisco.
-
-20.Ludwig, D. \& Walters, C.J. \emph{et al.} (1982). Optimal harvesting
-with imprecise parameter estimates. \emph{Ecological Modelling}, 14,
-273--292.
-
-21.Mangel, M. \& Clark, C.W. \emph{et al.} (1988). \emph{Dynamic
-Modeling in Behavioral Ecology}. Princeton University Press, Princeton.
-
-22.Marescot, L., Chapron, G., Chadès, I., Fackler, P.L., Duchamp, C. \&
-Marboutin, E.\emph{et al.} (2013). Complex decisions made simple: a
-primer on stochastic dynamic programming. \emph{Methods in Ecology and
-Evolution}, n/a--n/a.
-
-23.May, R.M., Beddington, J.R., Clark, C.W., Holt, S.J. \& Laws,
-R.M.\emph{et al.} (1979). Management of multispecies fisheries.
-\emph{Science (New York, N.Y.)}, 205, 267--77.
-
-24.McAllister, M. \emph{et al.} (1998). Bayesian stock assessment: a
-review and example application using the logistic model. \emph{ICES
-Journal of Marine Science}, 55, 1031--1060.
-
-25.Munch, S.B., Kottas, A. \& Mangel, M.\emph{et al.} (2005a). Bayesian
-nonparametric analysis of stock-recruitment relationships.
-\emph{Canadian Journal of Fisheries and Aquatic Sciences}, 62,
-1808--1821.
-
-26.Munch, S.B., Snover, M.L., Watters, G.M. \& Mangel, M.\emph{et al.}
-(2005b). A unified treatment of top-down and bottom-up control of
-reproduction in populations. \emph{Ecology Letters}, 8, 691--695.
-
-27.Myers, R.A., Barrowman, N.J., Hutchings, J.A. \& Rosenberg, A.
-a\emph{et al.} (1995). Population dynamics of exploited fish stocks at
-low population levels. \emph{Science (New York, N.Y.)}, 269, 1106--8.
-
-28.Polasky, S., Carpenter, S.R., Folke, C. \& Keeler, B.\emph{et al.}
-(2011). Decision-making under great uncertainty: environmental
-management in an era of global change. \emph{Trends in ecology \&
-evolution}, 1--7.
-
-29.R Core Team \emph{et al.} (2013). \emph{R: A Language and Environment
-for Statistical Computing}. R Foundation for Statistical Computing,
-Vienna, Austria.
-
-30.Rasmussen, C.E. \& Williams, C.K.I. \emph{et al.} (2006).
-\emph{Gaussian Processes for Machine Learning}. MIT Press, Boston.
-
-31.Reed, W.J. \emph{et al.} (1979). Optimal escapement levels in
-stochastic and deterministic harvesting models. \emph{Journal of
-Environmental Economics and Management}, 6, 350--363.
-
-32.Roughgarden, J.E. \& Smith, F. \emph{et al.} (1996). Why fisheries
-collapse and what to do about it. \emph{Proceedings of the National
-Academy of Sciences of the United States of America}, 93, 5078.
-
-33.Schapaugh, A.W. \& Tyre, A.J. \emph{et al.} (2013). Accounting for
-parametric uncertainty in Markov decision processes. \emph{Ecological
-Modelling}, 254, 15--21.
-
-34.Scheffer, M., Bascompte, J., Brock, W.A., Brovkin, V., Carpenter,
-S.R. \& Dakos, V.\emph{et al.} (2009). Early-warning signals for
-critical transitions. \emph{Nature}, 461, 53--9.
-
-35.Scheffer, M., Carpenter, S.R., Foley, J.A., Folke, C. \& Walker,
-B.\emph{et al.} (2001). Catastrophic shifts in ecosystems.
-\emph{Nature}, 413, 591--6.
-
-36.Sethi, G., Costello, C., Fisher, A., Hanemann, M. \& Karp, L.\emph{et
-al.} (2005). Fishery management under multiple uncertainty.
-\emph{Journal of Environmental Economics and Management}, 50, 300--318.
-
-37.Sigourney, D.B., Munch, S.B. \& Letcher, B.H.\emph{et al.} (2012).
-Combining a Bayesian nonparametric method with a hierarchical framework
-to estimate individual and temporal variation in growth.
-\emph{Ecological Modelling}, 247, 125--134.
-
-38.Su, Y.-S. \& Masanao Yajima \emph{et al.} (2013). \emph{R2jags: A
-Package for Running jags from R}.
-
-39.Sugeno, M. \& Munch, S.B. \emph{et al.} (2013a). A semiparametric
-Bayesian method for detecting Allee effects. \emph{Ecology}, 94,
-1196--1204.
-
-40.Sugeno, M. \& Munch, S.B. \emph{et al.} (2013b). A semiparametric
-Bayesian approach to estimating maximum reproductive rates at low
-population sizes. \emph{Ecological applications : a publication of the
-Ecological Society of America}, 23, 699--709.
-
-41.Thorson, J.T., Ono, K. \& Munch, S.B.\emph{et al.} (2014). A Bayesian
-approach to identifying and compensating for model misspecification in
-population models. \emph{Ecology}, 95, 329--41.
-
-42.Weitzman, M.L. \emph{et al.} (2013). A Precautionary Tale of
-Uncertain Tail Fattening. \emph{Environmental and Resource Economics},
-55, 159--173.
-
-43.Williams, B.K. \emph{et al.} (2001). Uncertainty , learning , and the
-optimal management of wildlife. \emph{Environmental and Ecological
-Statistics}, 8, 269--288.
-
-44.Worm, B., Barbier, E.B., Beaumont, N., Duffy, J.E., Folke, C. \&
-Halpern, B.S.\emph{et al.} (2006). Impacts of biodiversity loss on ocean
-ecosystem services. \emph{Science (New York, N.Y.)}, 314, 787--90.
-
-45.Worm, B., Hilborn, R., Baum, J.K., Branch, T.A., Collie, J.S. \&
-Costello, C.\emph{et al.} (2009). Rebuilding global fisheries.
-\emph{Science (New York, N.Y.)}, 325, 578--85.
-
-46.Xie, Y. \emph{et al.} (2012). \emph{knitr: A general-purpose package
-for dynamic report generation in R}.
+Allen, D., and K. Tanner. 2005. Infusing active learning into the
+large-enrollment biology class: seven strategies, from the simple to
+complex. Cell biology education 4:262--8.
+
+Athanassoglou, S., and A. Xepapadeas. 2012. Pollution control with
+uncertain stock dynamics: When, and how, to be precautious. Journal of
+Environmental Economics and Management 63:304--320.
+
+Bestelmeyer, B. T., M. C. Duniway, D. K. James, L. M. Burkett, and K. M.
+Havstad. 2012. A test of critical thresholds and their indicators in a
+desertification-prone ecosystem: more resilience than we thought.
+Ecology Letters:n/a--n/a.
+
+Brozović, N., and W. Schlenker. 2011. Optimal management of an ecosystem
+with an unknown threshold. Ecological Economics:1--14.
+
+Burnham, K. P., and D. R. Anderson. 2002. Model Selection and
+Multi-Model Inference. Page 496. Springer.
+
+Clark, C. W. 1976. Mathematical Bioeconomics. WileyNew York.
+
+Clark, C. W. 2009. Mathematical Bioeconomics. WileyNew York.
+
+Clark, C. W., and G. P. Kirkwood. 1986. On uncertain renewable resource
+stocks: Optimal harvest policies and the value of stock surveys. Journal
+of Environmental Economics and Management 13:235--244.
+
+Clark, C. W., and M. Mangel. 2000. Dynamic state variable models in
+ecology. Oxford University PressOxford.
+
+Courchamp, F., L. Berec, and J. Gascoigne. 2008. Allee Effects in
+Ecology and Conservation. Page 256. Oxford University Press, USA.
+
+Cressie, N., C. a Calder, J. S. Clark, J. M. {Ver Hoef}, and C. K.
+Wikle. 2009. Accounting for uncertainty in ecological analysis: the
+strengths and limitations of hierarchical statistical modeling.
+Ecological Applications 19:553--70.
+
+Cuddington, K. M., M. Fortin, and L. Gerber. 2013. Process-based models
+are required to manage ecological systems in a changing world. Ecosphere
+4:1--12.
+
+Fischer, J., G. D. Peterson, T. a Gardner, L. J. Gordon, I. Fazey, T.
+Elmqvist, A. Felton, C. Folke, and S. Dovers. 2009. Integrating
+resilience thinking and optimisation for conservation. Trends in ecology
+\& evolution 24:549--54.
+
+Gardiner, C. 2009. Stochastic Methods: A Handbook for the Natural and
+Social Sciences (Springer Series in Synergetics). Page 447. Springer.
+
+Gelman, A., J. B. Carlin, H. S. Stern, and D. B. Rubin. 2003. Bayesian
+Data Analysis. 2nd editions. Chapman; Hall/CRC.
+
+Geritz, S. A. H., and E. Kisdi. 2012. Mathematical ecology: why
+mechanistic models? Journal of mathematical biology 65:1411--5.
+
+Gordon, H. 1954. The economic theory of a common-property resource: the
+fishery. The Journal of Political Economy 62:124--142.
+
+Hilborn, R. 2007. Reinterpreting the State of Fisheries and their
+Management. Ecosystems 10:1362--1369.
+
+Hilborn, R., and M. Mangel. 1997. The Ecological Detective: Confronting
+Models with data. Page 330. Princeton University Press.
+
+Hughes, T. P., C. Linares, V. Dakos, I. a van de Leemput, and E. H. van
+Nes. 2013. Living dangerously on borrowed time during slow, unrecognized
+regime shifts. Trends in ecology \& evolution 28:149--55.
+
+Kocijan, J., A. Girard, B. Banko, and R. Murray-Smith. 2005. Dynamic
+systems identification with Gaussian processes. Mathematical and
+Computer Modelling of Dynamical Systems 11:411--424.
+
+Lehmann, E. L. 1975. Nonparametrics: Statistical Methods Based on Ranks.
+Holden-Day, IncSan Francisco.
+
+Ludwig, D., and C. J. Walters. 1982. Optimal harvesting with imprecise
+parameter estimates. Ecological Modelling 14:273--292.
+
+Mangel, M. 2014. Stochastic Dynamic Programming Illuminates the Link
+Between Environment. Bulletin of Mathematical Biology in press.
+
+Mangel, M., and C. W. Clark. 1988. Dynamic Modeling in Behavioral
+Ecology. (J. Krebs and T. Clutton-Brock, Eds.). Princeton University
+PressPrinceton.
+
+Mangel, M., 0. Fiksen, and J. Giske. 2001. Theoretical and statistical
+models in natural resource management and research. Pages 57--71
+\emph{in} T. M. Shenk and A. B. Franklin, editors. Modeling in natural
+resource management, development, interpretation and application. Island
+PressWashington DC.
+
+Marescot, L., G. Chapron, I. Chadès, P. L. Fackler, C. Duchamp, E.
+Marboutin, and O. Gimenez. 2013. Complex decisions made simple: a primer
+on stochastic dynamic programming. Methods in Ecology and
+Evolution:n/a--n/a.
+
+May, R. M., J. R. Beddington, C. W. Clark, S. J. Holt, and R. M. Laws.
+1979. Management of multispecies fisheries. Science (New York, N.Y.)
+205:267--77.
+
+McAllister, M. 1998. Bayesian stock assessment: a review and example
+application using the logistic model. ICES Journal of Marine Science
+55:1031--1060.
+
+Munch, S. B., A. Kottas, and M. Mangel. 2005a. Bayesian nonparametric
+analysis of stock-recruitment relationships. Canadian Journal of
+Fisheries and Aquatic Sciences 62:1808--1821.
+
+Munch, S. B., M. L. Snover, G. M. Watters, and M. Mangel. 2005b. A
+unified treatment of top-down and bottom-up control of reproduction in
+populations. Ecology Letters 8:691--695.
+
+Myers, R. A., N. J. Barrowman, J. A. Hutchings, and A. a Rosenberg.
+1995. Population dynamics of exploited fish stocks at low population
+levels. Science (New York, N.Y.) 269:1106--8.
+
+Polasky, S., S. R. Carpenter, C. Folke, and B. Keeler. 2011.
+Decision-making under great uncertainty: environmental management in an
+era of global change. Trends in ecology \& evolution:1--7.
+
+R Core Team. 2013. R: A Language and Environment for Statistical
+Computing. R Foundation for Statistical ComputingVienna, Austria.
+
+Rasmussen, C. E., and C. K. I. Williams. 2006. Gaussian Processes for
+Machine Learning. (Thomas Dietterich, Ed.). MIT Press,Boston.
+
+Reed, W. J. 1979. Optimal escapement levels in stochastic and
+deterministic harvesting models. Journal of Environmental Economics and
+Management 6:350--363.
+
+Roughgarden, J. E., and F. Smith. 1996. Why fisheries collapse and what
+to do about it. Proceedings of the National Academy of Sciences of the
+United States of America 93:5078.
+
+Schapaugh, A. W., and A. J. Tyre. 2013. Accounting for parametric
+uncertainty in Markov decision processes. Ecological Modelling
+254:15--21.
+
+Scheffer, M., J. Bascompte, W. A. Brock, V. Brovkin, S. R. Carpenter, V.
+Dakos, H. Held, E. H. van Nes, M. Rietkerk, and G. Sugihara. 2009.
+Early-warning signals for critical transitions. Nature 461:53--9.
+
+Scheffer, M., S. R. Carpenter, J. A. Foley, C. Folke, and B. Walker.
+2001. Catastrophic shifts in ecosystems. Nature 413:591--6.
+
+Sethi, G., C. Costello, A. Fisher, M. Hanemann, and L. Karp. 2005.
+Fishery management under multiple uncertainty. Journal of Environmental
+Economics and Management 50:300--318.
+
+Sigourney, D. B., S. B. Munch, and B. H. Letcher. 2012. Combining a
+Bayesian nonparametric method with a hierarchical framework to estimate
+individual and temporal variation in growth. Ecological Modelling
+247:125--134.
+
+Su, Y.-S., and Masanao Yajima. 2013. R2jags: A Package for Running jags
+from R.
+
+Sugeno, M., and S. B. Munch. 2013a. A semiparametric Bayesian method for
+detecting Allee effects. Ecology 94:1196--1204.
+
+Sugeno, M., and S. B. Munch. 2013b. A semiparametric Bayesian approach
+to estimating maximum reproductive rates at low population sizes.
+Ecological applications : a publication of the Ecological Society of
+America 23:699--709.
+
+Thorson, J. T., K. Ono, and S. B. Munch. 2014. A Bayesian approach to
+identifying and compensating for model misspecification in population
+models. Ecology 95:329--41.
+
+Walters, C. J., and R. Hilborn. 1978. Ecological Optimization and
+Adaptive Management. Annual Review of Ecology and Systematics
+9:157--188.
+
+Weitzman, M. L. 2013. A Precautionary Tale of Uncertain Tail Fattening.
+Environmental and Resource Economics 55:159--173.
+
+Williams, B. K. 2001. Uncertainty , learning , and the optimal
+management of wildlife. Environmental and Ecological Statistics
+8:269--288.
+
+Worm, B., E. B. Barbier, N. Beaumont, J. E. Duffy, C. Folke, B. S.
+Halpern, J. B. C. Jackson, H. K. Lotze, F. Micheli, S. R. Palumbi, E.
+Sala, K. a Selkoe, J. J. Stachowicz, and R. Watson. 2006. Impacts of
+biodiversity loss on ocean ecosystem services. Science (New York, N.Y.)
+314:787--90.
+
+Worm, B., R. Hilborn, J. K. Baum, T. A. Branch, J. S. Collie, C.
+Costello, M. J. Fogarty, E. a Fulton, J. a Hutchings, S. Jennings, O. P.
+Jensen, H. K. Lotze, P. M. Mace, T. R. McClanahan, C. Minto, S. R.
+Palumbi, A. M. Parma, D. Ricard, A. a Rosenberg, R. Watson, and D.
+Zeller. 2009. Rebuilding global fisheries. Science (New York, N.Y.)
+325:578--85.
+
+Xie, Y. 2012. knitr: A general-purpose package for dynamic report
+generation in R.
 

### 2014-05-29 manuscripts/manuscript.Rmd
new sensitivity analysis

@@
   - Fisheries Management
+
+bibliography: components/references.bib
+csl: components/ecology.csl
+template: components/elsarticle.latex
+documentclass: components/elsarticle
+
+## rmarkdown render options
+output:
+  pdf_document:
+    keep_tex: true
 ---
@@
 
-```{r sensitivity-calc}
-source("components/sensitivity.R")
-sensitivity_dat <- dat
-```
-
-```{r export-data}
-write.csv(sensitivity_dat, "sensitivity.csv")
+```{r sensitivity-trends}
+source("components/sensitivity-trends.Rmd")
 ```
@@
 
+write.csv(df_post, "components/data/figure1.csv")
 ```
@@
     theme(legend.position="none")
+
+write.csv(models_posteriors, "components/data/figure2.csv")
+write.csv(obs, "components/data/obs.csv")
 ```
@@
   theme(legend.position="none") + xlab("time") + ylab("stock size")
+
+write.csv(crash_data, "components/data/figure3.csv")
 ```
@@
   scale_colour_manual(values=colorkey)
+
+write.csv(policies, "components/data/figure4.csv")
 ```
@@
 
+write.csv(sims_data, "components/data/figure5.csv")
 ```
@@
 #  xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)+ scale_color_manual(values=colorkey)
+
+write.csv(actual_over_optimal, "components/data/figure6.csv")
 ```
@@
 sizes fluctuating significantly below the optimal escapement level will
-not perform poorly as long as those fluctuations are sufficently rare.
+not perform poorly as long as those fluctuations are sufficently rare.  Figure 7
+shows that the GPDP is little effected by increasing noise or increasing Allee
+effects over much of the range.
 
@@
 
-```{r figure_7, fig.height=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
-Histogram shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters.
-Values near 1 indicate nearly optimal performance.
-Simulations cover a range of underlying parameter, model, replicate, and noise levels."}
-ggplot(sensitivity_dat) + 
-  geom_histogram(aes(value, fill=noise)) + 
-  xlim(0,1.0) + 
-  theme_bw() + 
-  xlab("value as fraction of the optimal") + 
-  facet_wrap(~model, ncol=1)
+```{r figure_7, fig.height=6, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).
+Other than the focal parameter (noise, Allee threshold), other parameters are held fixed as above to illustrate this effect."}
+
+plot_sigmas <- ggplot(vary_sigma, aes(noise, value)) + 
+  geom_point() + ylab("Net present value") + 
+  xlab("Level of growth stochasticity (sigma)") + theme_bw() 
+plot_allees <- ggplot(vary_allee, aes(pars.C, value)) + 
+  geom_point()+ ylab("Net present value") + 
+  xlab("Allee threshold stock size") + theme_bw()
+multiplot(plot_sigmas, plot_allees, cols=2)
+
+write.csv(vary_sigma, "components/data/vary_sigma.csv")
+write.csv(vary_allee, "components/data/vary_sigma.csv")
 ```
@@
   facet_wrap(~ variable, scale="free", ncol=2)
+write.csv(ricker_posteriors, "components/data/ricker_posteriors.csv")
 ```
@@
   facet_wrap(~ variable, scale="free", ncol=2)
+write.csv(myers_posteriors, "components/data/myers_posteriors.csv")
 ```
@@
   facet_wrap(~ variable, scale="free", ncol=2)
+write.csv(allen_posteriors, "components/data/allen_posteriors.csv")
 ```
@@
   caption = "Parameterization range for the uniform priors in the Allen model")
+```
+
+```{r sensitivity-calc}
+source("components/sensitivity.R")
+sensitivity_dat <- dat
+```
+
+```{r figure_S9, fig.height=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
+Histogram shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters.
+Values near 1 indicate nearly optimal performance.
+Simulations cover a range of underlying parameter, model, replicate, and noise levels."}
+ggplot(sensitivity_dat) + 
+  geom_histogram(aes(value, fill=simulation)) + 
+  xlim(0,1.0) + 
+  theme_bw() + 
+  xlab("value as fraction of the optimal") + 
+  facet_grid(noise~model)
+write.csv(sensitivity_dat, "components/data/sensitivity_dat.csv")
+
 ```

### 2014-05-29 manuscripts/components/sensitivity-trends.R

@@ -1,0 +1,158 @@
+library("nonparametricbayes") 
+library("pdgControl")
+library("ggplot2") 
+library("reshape2")
+library("plyr")
+library("data.table")
+
+
+
+models <- c("Myers","Allen")
+
+parameters <- list(Myers = list(
+  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
+  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
+  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2))),
+                   Allen = list(
+  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
+  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
+  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)))
+                   )
+nuisance_pars <- c("sigma_g")
+nuisance_values <- list(sigma_g = c(0.01, 0.05, 0.1))
+replicates <- c(1111, 2222, 3333, 4444, 5555, 6666, 7777, 8888) # seeds
+
+
+sensitivity <- function(model, parameters, nuisance, seed){
+  
+  if(model == "Myers")
+    f <- Myers
+  else if(model == "Allen")
+    f <- RickerAllee
+    
+sigma_g <- nuisance[["sigma_g"]]
+p <- parameters
+sigma_m <- 0.0
+z_g <- function() rlnorm(1, 0, sigma_g)
+z_m <- function() 1
+x_grid <- seq(0, 15, length=50)
+h_grid <- x_grid
+profit <- function(x,h) pmin(x, h)
+delta <- 0.01
+OptTime <- 50  # stationarity with unstable models is tricky thing
+reward <- 0
+xT <- 0
+Xo <- 5.5# observations start from
+x0 <- 8 # simulation under policy starts from
+Tobs <- 40
+MaxT <- 1000 # timeout for value iteration convergence
+
+  
+  # replicate over random seed
+  yields <- sapply(seed, 
+                   function(seed_i){
+                     set.seed(seed_i)
+                     
[…]

### 2014-06-03 manuscripts/manuscript.Rmd
some edits from Steve

@@
 ```{r libraries, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+library("knitr")
 library("nonparametricbayes") 
@@
 ```{r sensitivity-trends}
-source("components/sensitivity-trends.Rmd")
+source("components/sensitivity-trends.R")
 ```
@@
 In order to find that policy, we introduce the value function$V_t(x_t)$ representing the total discounted catch from time $t$ onwards given that $X_t=x$.
-This value function satisfies an equation of Stochastic Dynamic Programming  [SDP, @Mangel1988; @Clark2000; @Clark2009], sometimes called the Bellman-Hamilton-Jacobi equation [@Mangel2014].
+This value function satisfies an equation of Stochastic Dynamic Programming  [SDP, @Mangel1988; @Clark2000; @Clark2009], sometimes called the Bellman-Hamilton-Jacobi equation [@Mangel2014],
 
 \begin{equation}
-V_t(x) = \max_{h_t} \left[R(h_t, x_t) + \delta \cdot \mathbf{E}_{X_t+1} V_{t+1}( Z_{t+1} f(x - h_t)) \right]
+V_t(X_t) = \max_{h_t} \left{ R(h_t, X_t) + \delta \cdot \mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] \right}
 \end{equation}
 
-That is, at time $t$, when population size is $x$ and harvest $h_t$ is applied, the immediate return is $R(h_t, x_t)$ and the population size after the take is $x-h_t$, which is then translated into $X_{t+1}$ by Eqn 1.
+where expectation is taken over all possible values of the next state, $X_t{+1}$, and maximized over all possible choices of harvest, $h_t$.
+That is, at time $t$, when population size is $x$ and harvest $h_t$ is applied, the immediate return is $R(h_t, X_t)$.
+In the classical case, the sole source of uncertainty is the process noise term, $Z$, and thus the expectation above is equivalent to taking expectations over $Z$.
+That is
+
+$$ \mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{Z} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] $$
+
+where the population size after the take is $X_t-h_t$, which is then translated into $X_{t+1}$ by Eqn 1.
+
+In a Bayesian decision framework, the parameters governing the dynamics are also uncertain, so the
+expectation in (1) involves averaging over the posterior distribution for the parameters, as well. That
+is,
+
+$$\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta,\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \}$$
+
+In the nonparametric case, the function $f$ too is uncertain and the expectation for the next state includes
+uncertainty in $f$ as well. That is
+
+$$\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta,\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \}$$
+
 We consider the finite time problem with $T=$`r MaxT`, which we solve using value iteration algorithm [@Mangel1988; @Clark2000].
@@
 
-A GP is an infinite dimensional (because it considers a continuous random
+A GP is an infinite dimensional (because it is a function of a continuous input
 variable) stochastic process for which any realization is a multivariate
@@
 
-$$y = f(x) + \varepsilon$$
-
-$\varepsilon$ IID normal, variance $\sigma_n^2$. Then under the GP, the predicted data $X_p$ given the observations obeys
+$$X_{t+1} = f(X_t) + \varepsilon$$
+
+$\varepsilon$ IID normal, variance $\sigma_n^2$. Note that we have chosen to
+assume additive noise.  While we could just as easily consider log-normal noise
+as in the parametric models, we make this choice to bring home the point that the
+Gaussian process approach need not have structurally correct noise form either.
+
+Then under the GP, given a vector of observed data $X_0$ the predicted data $X_p$ obeys
 
@@
   xlab("Allee threshold stock size") + theme_bw()
-multiplot(plot_sigmas, plot_allees, cols=2)
+cboettigR::multiplot(plot_sigmas, plot_allees, cols=2)
 

### 2014-06-03 manuscripts/manuscript.tex
some edits from Steve

@@
-\documentclass[author-year, 12pt,review]{elsarticle} %review=doublespace preprint=single 5p=2 column
-%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
-\usepackage[hyphens]{url}
-\usepackage{lineno} % add 
-  \linenumbers % turns line numbering on 
-\bibliographystyle{elsarticle-harv}
-\biboptions{sort&compress} % For natbib
-\usepackage{graphicx}
-\usepackage{booktabs} % book-quality tables
-%% Redefines the elsarticle footer
-\makeatletter
-\def\ps@pprintTitle{%
- \let\@oddhead\@empty
- \let\@evenhead\@empty
- \def\@oddfoot{\it \hfill\today}%
- \let\@evenfoot\@oddfoot}
-\makeatother
-
-% A modified page layout
-\textwidth 6.75in
-\oddsidemargin -0.15in
-\evensidemargin -0.15in
-\textheight 9in
-\topmargin -0.5in
-%%%%%%%%%%%%%%%% end my additions to header
-
+\documentclass[]{components/elsarticle}
 \usepackage[T1]{fontenc}
@@
 \else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+    \usepackage{xltxtra,xunicode}
+  \else
     \usepackage{fontspec}
-  \ifxetex
-    \usepackage{xltxtra,xunicode}
   \fi
@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\usepackage{longtable}
+\usepackage[margin=1in]{geometry}
+\usepackage{longtable,booktabs}
 \usepackage{graphicx}
-% We will generate all images so they have a width \maxwidth. This means
-% that they will get their normal width if they fit onto the page, but
+% Redefine \includegraphics so that, unless explicit options are
+% given, the image width will not exceed the width of the page.
+% Images get their normal width if they fit onto the page, but
 % are scaled down if they would overflow the margins.
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
-\else\Gin@nat@width\fi}
+\def\ScaleIfNeeded{%
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
 \makeatother
 \let\Oldincludegraphics\includegraphics
-\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
+{%
+ \catcode`\@=11\relax%
+ \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
+}%
 \ifxetex
@@
             colorlinks=true,
+            citecolor=blue,
             urlcolor=blue,
@@
 \setcounter{secnumdepth}{0}
-% Pandoc toggle for numbering sections (defaults to be off)
-\setcounter{secnumdepth}{0}
-% Pandoc header
-
-\begin{document}
-\begin{frontmatter}
 
@@
 Non-Parametric approach}
-    \author[cstar]{Carl Boettiger\corref{c1}}
-   \ead{cboettig@gmail.com} 
-   \cortext[c1]{Corresponding author}
-    \author[cstar]{Marc Mangel}
-
-    \author[noaa]{Stephan Munch}
-
-      \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and
-Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA
-95064, USA}    
-    \address[marc]{Center for Stock Assessment Research, Department of Applied Math and
-Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA
-95064, USA and Department of Biology, University of Bergen, Bergen,
-Norway 9020}    
-    \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric
-Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}
+\author{ \and  \and }
+\date{}
+
+\begin{document}
+
+\begin{center}
+\huge Avoiding tipping points in fisheries management: A Bayesian
+Non-Parametric approach \\[0.2cm]
+\normalsize
+\end{center}
 
@@
 \end{abstract}
-   \begin{keyword} Bayesian \sep Structural Uncertainty \sep Nonparametric \sep Optimal Control \sep Decision Theory \sep Gaussian Processes \sep Fisheries Management \sep \end{keyword}
- \end{frontmatter}
+
+\begin{verbatim}
+## module glm loaded
+\end{verbatim}
+
+\begin{verbatim}
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+\end{verbatim}
+
+\begin{verbatim}
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+\end{verbatim}
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_1.pdf}
-\caption{Points show the training data of stock-size over time. Curves
-show the posterior step-ahead predictions based on each of the estimated
-models. Observe that all models are fitting the data reasonably well.}
-\end{figure}
+\includegraphics{./manuscript_files/figure-latex/figure_1.pdf}
 
@@
 
+\% latex table generated in R 3.1.0 by xtable 1.7-3 package \% Mon Jun 2
+00:55:24 2014
+
 \begin{table}[ht]
-\begin{center}
+\centering
 \begin{tabular}{rrrr}
@@
   \hline
-DIC & 50.14 & 49.45 & 50.61 \\ 
-  AIC & -24.60 & -30.07 & -27.19 \\ 
-  BIC & -17.85 & -25.00 & -20.44 \\ 
+DIC & 50.75 & 50.45 & 50.41 \\ 
+  AIC & -24.51 & -30.13 & -27.01 \\ 
+  BIC & -17.75 & -25.06 & -20.25 \\ 
    \hline
@@
 This highlights the danger of model choice when the selected model will be used outside of the observed range of the data.}
-\end{center}
 \end{table}
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_2.pdf}
-\caption{Graph of the inferred Gaussian process compared to the true
-process and maximum-likelihood estimated process. Graph shows the
-expected value for the function $f$ under each model. Two standard
-deviations from the estimated Gaussian process covariance with (light
-grey) and without (darker grey) measurement error are also shown. The
-training data is also shown as black points. The GP is conditioned on
-(0,0), shown as a pseudo-data point.}
-\end{figure}
+\includegraphics{./manuscript_files/figure-latex/figure_2.pdf}
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_3.pdf}
-\caption{Outside the range of the training data (Figure 1), the true
-dynamics (black dots) fall outside the uncertainty (two standard
-deviations, colored bands) of the structurally incorrect parametric
-models (Ricker, Myers), but inside the uncertainty predicted by the GP.
-Points show the stock size simulated by the true model. Overlay shows
-the range of states predicted by each model, based on the state observed
-in the previous time step. The Ricker model always (wrongly) predicts
-positive population growth, while the actual population shrinks in each
-step as the initial condition falls below the Allee threshold of the
-underlying model (Allen). Note that because it does not assume a
-parametric form but instead relies more directly on the data, the GP is
-both more pessimistic and more uncertain about the future state than the
-parametric models.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_4.pdf}
-\caption{The steady-state optimal policy (infinite boundary) calculated
-under each model. Policies are shown in terms of target escapement,
-$S_t$, as under models such as this a constant escapement policy is
-expected to be optimal (Reed 1979).}
-\end{figure}
+\includegraphics{./manuscript_files/figure-latex/figure_3.pdf}
+
+\includegraphics{./manuscript_files/figure-latex/figure_4.pdf}
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_5.pdf}
-\caption{In the management context, GPDP outperforms approaches based on
-parametric models. Shown are 100 replicate simulations of the stock
-dynamics (eq 1) under the policies derived from each of the estimated
-models, as well as the policy based on the exact underlying model.}
-\end{figure}
+\includegraphics{./manuscript_files/figure-latex/figure_5.pdf}
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_6.pdf}
-\caption{Histograms of the realized net present value of the fishery
-over a range of simulated data and resulting parameter estimates. For
-each data set, the three models are estimated as described above. Values
-plotted are the averages of a given policy over 100 replicate
-simulations. Details and code provided in the supplement.}
-\end{figure}
+\begin{verbatim}
+## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
+## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
+## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
+## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
+## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
+\end{verbatim}
+
+\includegraphics{./manuscript_files/figure-latex/figure_6.pdf}
 
@@
 significantly below the optimal escapement level will not perform poorly
-as long as those fluctuations are sufficently rare.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_7.pdf}
-\caption{Sensitivity Analysis. Histogram shows the ratio of the realized
-net present value derived when managing under the GPDP over the optimal
-value given the true model and true parameters. Values near 1 indicate
-nearly optimal performance. Simulations cover a range of underlying
-parameter, model, replicate, and noise levels.}
-\end{figure}
+as long as those fluctuations are sufficently rare. Figure 7 shows that
+the GPDP is little effected by increasing noise or increasing Allee
+effects over much of the range.
+
+\begin{verbatim}
+## Loading required package: grid
+\end{verbatim}
+
+\includegraphics{./manuscript_files/figure-latex/figure_7.pdf}
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_S1.pdf}
-\caption{Traces from the MCMC estimates of the GP model show reasonable
-mixing (no trend) and sampling rejection rate (no piecewise jumps)}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_S2.pdf}
-\caption{Posterior distributions from the MCMC estimate of the GP model.
-Prior curves shown in red.}
-\end{figure}
+\includegraphics{./manuscript_files/figure-latex/figure_S1.pdf}
+
+\includegraphics{./manuscript_files/figure-latex/figure_S2.pdf}
 
@@
 accompanying the paper, available from
-{[}http://github.com/cboettig/nonparametric-bayes{]}. The code for all
-results shown here is dynamically embedded into the manuscript using
+{[}\url{http://github.com/cboettig/nonparametric-bayes}{]}. The code for
+all results shown here is dynamically embedded into the manuscript using
 \texttt{knitr} (Xie 2012).
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_S3.pdf}
-\caption{Traces from the MCMC estimates of the Ricker model show
-reasonable mixing (no trend) and sampling rejection rate (no piecewise
-jumps)}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_S4.pdf}
-\caption{Posteriors from the MCMC estimate of the Ricker model}
-\end{figure}
+\includegraphics{./manuscript_files/figure-latex/figure_S3.pdf}
+
+\includegraphics{./manuscript_files/figure-latex/figure_S4.pdf}
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_S5.pdf}
-\caption{Traces from the MCMC estimates of the Myers model show
-reasonable mixing (no trend) and sampling rejection rate (no piecewise
-jumps)}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_S6.pdf}
-\caption{Posterior distributions from the MCMC estimates of the Myers
-model}
-\end{figure}
+\includegraphics{./manuscript_files/figure-latex/figure_S5.pdf}
+
+\includegraphics{./manuscript_files/figure-latex/figure_S6.pdf}
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_S7.pdf}
-\caption{Traces from the MCMC estimates of the Allen model show
-reasonable mixing (no trend) and sampling rejection rate (no piecewise
-jumps)}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_S8.pdf}
-\caption{Posteriors from the MCMC estimate of the Allen model}
-\end{figure}
+\includegraphics{./manuscript_files/figure-latex/figure_S7.pdf}
+
+\includegraphics{./manuscript_files/figure-latex/figure_S8.pdf}
 
@@
 
+\begin{verbatim}
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+## Using V1 as value column: use value.var to override.
+## No id variables; using all as measure variables
+\end{verbatim}
+
+\begin{verbatim}
+## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
+## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
+## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
+## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
+\end{verbatim}
+
+\includegraphics{./manuscript_files/figure-latex/figure_S9.pdf}
+
 Allen, D., and K. Tanner. 2005. Infusing active learning into the

### 2014-06-03 manuscripts/manuscript.Rmd
edits from Steve
add caching chunk for render mode

@@
 
+```{r caching, include=FALSE}
+library("methods")
+library("knitr")
+basename <- "manuscript"
+opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
+               cache.path = paste("components/cache/", basename, "/", sep=""))
+opts_chunk$set(cache = 2)
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
+               comment = NA, verbose = TRUE, echo=FALSE)
+
+# PDF-based figures
+opts_chunk$set(dev='pdf')
+```
+
 <!-- Run R code for analysis, to be called by figures -->
@@
 ```{r libraries, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-library("knitr")
 library("nonparametricbayes") 

### 2014-06-03 manuscripts/manuscript.tex

@@
 2009), sometimes called the Bellman-Hamilton-Jacobi equation (Mangel
-2014).
+2014),
 
 \begin{equation}
-V_t(x) = \max_{h_t} \left[R(h_t, x_t) + \delta \cdot \mathbf{E}_{X_t+1} V_{t+1}( Z_{t+1} f(x - h_t)) \right]
+V_t(X_t) = \max_{h_t} \left{ R(h_t, X_t) + \delta \cdot \mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] \right}
 \end{equation}
 
+where expectation is taken over all possible values of the next state,
+$X_t{+1}$, and maximized over all possible choices of harvest, $h_t$.
 That is, at time $t$, when population size is $x$ and harvest $h_t$ is
-applied, the immediate return is $R(h_t, x_t)$ and the population size
-after the take is $x-h_t$, which is then translated into $X_{t+1}$ by
-Eqn 1. We consider the finite time problem with \$T=\$1000, which we
-solve using value iteration algorithm (Mangel and Clark 1988, Clark and
-Mangel 2000).
+applied, the immediate return is $R(h_t, X_t)$. In the classical case,
+the sole source of uncertainty is the process noise term, $Z$, and thus
+the expectation above is equivalent to taking expectations over $Z$.
+That is
+
+\[ \mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{Z} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \]
+
+where the population size after the take is $X_t-h_t$, which is then
+translated into $X_{t+1}$ by Eqn 1.
+
+In a Bayesian decision framework, the parameters governing the dynamics
+are also uncertain, so the expectation in (1) involves averaging over
+the posterior distribution for the parameters, as well. That is,
+
+\[\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta,\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \}\]
+
+In the nonparametric case, the function $f$ too is uncertain and the
+expectation for the next state includes uncertainty in $f$ as well. That
+is
+
+\[\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta,\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \}\]
+
+We consider the finite time problem with \$T=\$1000, which we solve
+using value iteration algorithm (Mangel and Clark 1988, Clark and Mangel
+2000).
 
@@
 
-A GP is an infinite dimensional (because it considers a continuous
-random variable) stochastic process for which any realization is a
-multivariate normal distribution. Thus, to characterize the GP we need a
-mean function and a covariance matrix. We proceed as follows.
+A GP is an infinite dimensional (because it is a function of a
+continuous input variable) stochastic process for which any realization
+is a multivariate normal distribution. Thus, to characterize the GP we
+need a mean function and a covariance matrix. We proceed as follows.
 
@@
 
-\[y = f(x) + \varepsilon\]
-
-$\varepsilon$ IID normal, variance $\sigma_n^2$. Then under the GP, the
-predicted data $X_p$ given the observations obeys
+\[X_{t+1} = f(X_t) + \varepsilon\]
+
+$\varepsilon$ IID normal, variance $\sigma_n^2$. Note that we have
+chosen to assume additive noise. While we could just as easily consider
+log-normal noise as in the parametric models, we make this choice to
+bring home the point that the Gaussian process approach need not have
+structurally correct noise form either.
+
+Then under the GP, given a vector of observed data $X_0$ the predicted
+data $X_p$ obeys
 
@@
 \% latex table generated in R 3.1.0 by xtable 1.7-3 package \% Mon Jun 2
-00:55:24 2014
+17:34:16 2014
 

### 2014-06-03 manuscripts/manuscript.tex
update tex

@@
 
-\begin{verbatim}
-## module glm loaded
-\end{verbatim}
-
-\begin{verbatim}
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-\end{verbatim}
-
-\begin{verbatim}
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-\end{verbatim}
-
 \section{Introduction}\label{introduction}
@@
 \begin{equation}
-V_t(X_t) = \max_{h_t} \left{ R(h_t, X_t) + \delta \cdot \mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] \right}
+V_t(X_t) = \max_{h_t} \lbrace R(h_t, X_t) + \delta \cdot \mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] \rbrace
 \end{equation}
@@
 
-\includegraphics{./manuscript_files/figure-latex/figure_1.pdf}
+\includegraphics{components/figure/manuscript-figure_1.pdf}
 
@@
 
-\% latex table generated in R 3.1.0 by xtable 1.7-3 package \% Mon Jun 2
-17:34:16 2014
+\% latex table generated in R 3.1.0 by xtable 1.7-3 package \% Tue Jun 3
+13:34:19 2014
 
@@
 
-\includegraphics{./manuscript_files/figure-latex/figure_2.pdf}
+\includegraphics{components/figure/manuscript-figure_2.pdf}
 
@@
 
-\includegraphics{./manuscript_files/figure-latex/figure_3.pdf}
-
-\includegraphics{./manuscript_files/figure-latex/figure_4.pdf}
+\includegraphics{components/figure/manuscript-figure_3.pdf}
+
+\includegraphics{components/figure/manuscript-figure_4.pdf}
 
@@
 
-\includegraphics{./manuscript_files/figure-latex/figure_5.pdf}
+\includegraphics{components/figure/manuscript-figure_5.pdf}
 
@@
 
-\begin{verbatim}
-## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
-## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
-## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
-## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
-## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
-\end{verbatim}
-
-\includegraphics{./manuscript_files/figure-latex/figure_6.pdf}
+\includegraphics{components/figure/manuscript-figure_6.pdf}
 
@@
 
-\begin{verbatim}
-## Loading required package: grid
-\end{verbatim}
-
-\includegraphics{./manuscript_files/figure-latex/figure_7.pdf}
+\includegraphics{components/figure/manuscript-figure_7.pdf}
 
@@
 
-\includegraphics{./manuscript_files/figure-latex/figure_S1.pdf}
-
-\includegraphics{./manuscript_files/figure-latex/figure_S2.pdf}
+\includegraphics{components/figure/manuscript-figure_S1.pdf}
+
+\includegraphics{components/figure/manuscript-figure_S2.pdf}
 
@@
 
-\includegraphics{./manuscript_files/figure-latex/figure_S3.pdf}
-
-\includegraphics{./manuscript_files/figure-latex/figure_S4.pdf}
+\includegraphics{components/figure/manuscript-figure_S3.pdf}
+
+\includegraphics{components/figure/manuscript-figure_S4.pdf}
 
@@
 
-\includegraphics{./manuscript_files/figure-latex/figure_S5.pdf}
-
-\includegraphics{./manuscript_files/figure-latex/figure_S6.pdf}
+\includegraphics{components/figure/manuscript-figure_S5.pdf}
+
+\includegraphics{components/figure/manuscript-figure_S6.pdf}
 
@@
 
-\includegraphics{./manuscript_files/figure-latex/figure_S7.pdf}
-
-\includegraphics{./manuscript_files/figure-latex/figure_S8.pdf}
+\includegraphics{components/figure/manuscript-figure_S7.pdf}
+
+\includegraphics{components/figure/manuscript-figure_S8.pdf}
 
@@
 
-\begin{verbatim}
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-## Using V1 as value column: use value.var to override.
-## No id variables; using all as measure variables
-\end{verbatim}
-
-\begin{verbatim}
-## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
-## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
-## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
-## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
-\end{verbatim}
-
-\includegraphics{./manuscript_files/figure-latex/figure_S9.pdf}
+\includegraphics{components/figure/manuscript-figure_S9.pdf}
 

### 2014-06-06 manuscripts/manuscript.Rmd
refactor code chunks

@@
 author: 
-  - name: "Carl Boettiger"
+  - name: Carl Boettiger
     affiliation: cstar
-    email: cboettig@gmail.com
+    email: cboettig(at)gmail.com
     footnote: Corresponding author
@@
 csl: components/ecology.csl
-template: components/elsarticle.latex
 documentclass: components/elsarticle
@@
   pdf_document:
-    keep_tex: true
+    fig_caption: true
+    template: components/elsarticle.latex
+
 ---
@@
 ```{r sensitivity-trends}
-source("components/sensitivity-trends.R")
+source("components/sensitivity-trends.R") # load the function for looping
+
+sigmas <- c(0.01, 0.05, 0.1)
+allee <- c(1,2,3,4,5)
+
+vary_sigma <- lapply(sigmas, function(s) try(sensitivity_trends(4, sigma = s, seed=c(111,222,333))))
+vary_allee <-  lapply(allee, function(a) try(sensitivity_trends(a, sigma = sigma_g, seed=c(111,222,333))))
+
+vary_sigma <- melt(vary_sigma, id=names(vary_sigma[[1]]))
+vary_allee <- melt(vary_allee, id=names(vary_allee[[1]]))
+
 ```
@@
 
-```{r figure_7, fig.height=6, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).
+```{r figure_7, fig.width=6, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).
 Other than the focal parameter (noise, Allee threshold), other parameters are held fixed as above to illustrate this effect."}
@@
   geom_point() + ylab("Net present value") + 
-  xlab("Level of growth stochasticity (sigma)") + theme_bw() 
-plot_allees <- ggplot(vary_allee, aes(pars.C, value)) + 
+  xlab("Level of growth stochasticity") + theme_bw() 
+plot_allees <- ggplot(vary_allee, aes(allee, value)) + 
   geom_point()+ ylab("Net present value") + 
@@
 
-```{r figure_S9, fig.height=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
+```{r figure_S9, fig.height=6, fig.width=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
 Histogram shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters.

### 2014-06-06 manuscripts/components/sensitivity-trends.R

@@ -1,54 +1,8 @@
-library("nonparametricbayes") 
-library("pdgControl")
-library("ggplot2") 
-library("reshape2")
-library("plyr")
-library("data.table")
-
-
-
-models <- c("Myers","Allen")
-
-parameters <- list(Myers = list(
-  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
-  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
-  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2))),
-                   Allen = list(
-  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
-  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
-  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)))
-                   )
-nuisance_pars <- c("sigma_g")
-nuisance_values <- list(sigma_g = c(0.01, 0.05, 0.1))
-replicates <- c(1111, 2222, 3333, 4444, 5555, 6666, 7777, 8888) # seeds
-
-
-sensitivity <- function(model, parameters, nuisance, seed){
+sensitivity_trends <- function(allee, sigma, seed){
+  sigma_g <- sigma
+  p <- c(r = p[[1]], K = p[[2]], C = allee)
   
-  if(model == "Myers")
-    f <- Myers
-  else if(model == "Allen")
-    f <- RickerAllee
-    
-sigma_g <- nuisance[["sigma_g"]]
-p <- parameters
-sigma_m <- 0.0
-z_g <- function() rlnorm(1, 0, sigma_g)
-z_m <- function() 1
-x_grid <- seq(0, 15, length=50)
-h_grid <- x_grid
-profit <- function(x,h) pmin(x, h)
-delta <- 0.01
-OptTime <- 50  # stationarity with unstable models is tricky thing
-reward <- 0
-xT <- 0
-Xo <- 5.5# observations start from
-x0 <- 8 # simulation under policy starts from
-Tobs <- 40
[…]

### 2014-06-06 manuscripts/manuscript.Rmd
revise sensitivity latin hypercube

@@
 source("components/sensitivity.R")
+
+models <- c("Myers","Allen")
+
+parameters <- list(
+  Myers = list(
+    c(r=1.5 + rnorm(1, 0, 1), theta=2.5 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1)),
+    c(r=1.5 + rnorm(1, 0, 1), theta=2.5 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1))),
+  Allen = list(
+    c(r=2 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1), C=4 + rnorm(1, 0, 1)),
+    c(r=2 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1), C=4 + rnorm(1, 0, 1)))
+)
+nuisance_pars <- c("sigma_g")
+nuisance_values <- list(sigma_g = c(0.01, 0.05))
+
+model <- "Allen"
+allen1.01 <- sensitivity(model, 
+                   parameters = parameters[[model]][[1]], 
+                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
+                   seed=c(1111, 2222, 3333))
+allen2.01 <- sensitivity(model, 
+                   parameters = parameters[[model]][[2]], 
+                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
+                   seed=c(1111, 2222, 3333))
+allen1.05 <- sensitivity(model, 
+                   parameters = parameters[[model]][[1]], 
+                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
+                   seed=c(1111, 2222, 3333))
+allen2.05 <- sensitivity(model, 
+                   parameters = parameters[[model]][[2]], 
+                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
+                   seed=c(1111, 2222, 3333))
+
+model <- "Myers"
+Myers1.01 <- sensitivity(model, 
+                   parameters = parameters[[model]][[1]], 
+                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
+                   seed=c(1111, 2222, 3333))
+Myers2.01 <- sensitivity(model, 
+                   parameters = parameters[[model]][[2]], 
+                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
+                   seed=c(1111, 2222, 3333))
+Myers1.05 <- sensitivity(model, 
+                   parameters = parameters[[model]][[1]], 
+                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
+                   seed=c(1111, 2222, 3333))
+Myers2.05 <- sensitivity(model, 
+                   parameters = parameters[[model]][[2]], 
+                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
+                   seed=c(1111, 2222, 3333))
+
+## Assemble into data.frame
+allen_dat <- rbind(allen1.01, allen1.05, allen2.01, allen2.05) 
+m <- rbind(Myers1.01, Myers1.05, Myers2.01, Myers2.05)
+myers_dat <- m[c(1:2,4,3,5:8)]
+names(myers_dat) <- names(allen_dat)
+model_dat <- rbind(allen_dat, myers_dat)
+dat <- model_dat
+dat$pars.r <- factor(dat$pars.r, labels=c("A", "B", "C", "D"))
+dat <- dat[c(1:2,5:6, 8, 7)]
+dat$noise <- factor(dat$noise)
+names(dat) <- c("model", "parameters", "replicate", "simulation", "noise", "value")
+
+## Extract acutal parameter values corresponding to each parameter set
+p1 = levels(factor(model_dat$pars.r))
+p2 = levels(factor(model_dat$pars.K))
+p3 = levels(factor(model_dat$pars.C))
+
+setA = as.numeric(c(r = p1[1], K = p2[1], theta = p3[1]))
+setB = as.numeric(c(r = p1[2], K = p2[2], theta = p3[2])
+setC = as.numeric(c(r = p1[3], K = p2[3], C = p3[3]))
+setD = as.numeric(c(r = p1[4], K = p2[4], C = p3[4]))
+AllenParameterSets <- rbind(setA, setB)
+MyersParameterSets <- rbind(setC,setD)
+
 sensitivity_dat <- dat
@@
 
-```{r figure_S9, fig.height=6, fig.width=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
-Histogram shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters.
-Values near 1 indicate nearly optimal performance.
-Simulations cover a range of underlying parameter, model, replicate, and noise levels."}
+```{r figure_S9, fig.height=6, fig.width=10, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
+Histograms shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters.
+Values of 1 indicate optimal performance.
+Columns indicate different models, rows different noise levels, and colors indicate the parameter set used.
+Grouped over stochastic replicates applying the contol policy and stochastic replicates of training data generated from the model indicated, see raw data for details.
+Randomly chosen parameter values for the models shown in tables below."}
 ggplot(sensitivity_dat) + 
-  geom_histogram(aes(value, fill=simulation)) + 
+  geom_histogram(aes(value, fill=parameters)) + 
   xlim(0,1.0) + 
@@
 write.csv(sensitivity_dat, "components/data/sensitivity_dat.csv") 
-
+```
+
+```{r}
+pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Allen models in Figure S9." )
+```
+```{r}
+pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Myers models in Figure S9." )
 ```

### 2014-06-06 manuscripts/components/sensitivity.R

@@ -1,18 +1,3 @@
-models <- c("Myers","Allen")
-
-parameters <- list(Myers = list(
-  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
-  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
-  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2))),
-                   Allen = list(
-  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
-  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
-  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)))
-                   )
-nuisance_pars <- c("sigma_g")
-nuisance_values <- list(sigma_g = c(0.01, 0.05, 0.1))
-replicates <- c(1111, 2222, 3333, 4444, 5555, 6666, 7777, 8888) # seeds
-
 
 sensitivity <- function(model, parameters, nuisance, seed){
   
@@ -115,79 +100,4 @@ MaxT <- 1000 # timeout for value iteration convergence
                     value = yields_dat$value, 
                     noise = sigma_g)
 }
-model <- "Allen"
-allen1.01 <- sensitivity(model, 
-                   parameters = parameters[[model]][[1]], 
-                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
-                   seed=c(1234, 2222, 3333))
-model <- "Allen"
-allen2.01 <- sensitivity(model, 
-                   parameters = parameters[[model]][[2]], 
-                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
-                   seed=c(1234, 2222, 3333))
-model <- "Allen"
-allen1.05 <- sensitivity(model, 
-                   parameters = parameters[[model]][[1]], 
-                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
-                   seed=c(1234, 2222, 3333))
-model <- "Allen"
-allen2.05 <- sensitivity(model, 
-                   parameters = parameters[[model]][[2]], 
[…]

### 2014-06-06 manuscripts/manuscript.Rmd
revise sensitivity test

@@
 by knowing the true model across changes to the training simulation,
-noise scale, parameters or structure of the underlying model, as seen in
-Figure 7. The sensitivity analysis uses a factorial design of eight replicates 
-of three different randomly generated parameter sets for each of two different 
-generating models (Allen and Myers) over two different noise levels 
-(0.01 and 0.05), for a total of 8 x 3 x 2 x 2 = 96 scenarios. In Figure 7, 
-we pool results across different random seeds (111, 222,
-333), noise values of 0.01 and 0.05 (not large enough to violate the
-self-sustaining criterion of @Reed1979), and different randomly generated
-parameters sets for each model, and using either the Myers or Allen as the
-underlying structure.  While the GPDP performs slightly closer to optimal
-on the smaller noise levels, and closer on the Allen model then the Myers
-model, the differences are small compared to the values observed in using
-the wrong parametric model.  The GPDP solution consistently avoids either
-failing to harvest at all or crashing the stock, which results in the very
-low values seen in parametric models in the previous figure. The
-sensitivity analysis is compared against the optimal solution alone and
-not shown against the estimates of the various parametric models (but
-see Figure 6), where convergence of the Gibbs sampling could contribute
-to their poorer performance if not checked by hand for each configuration.
+noise scale, parameters or structure of the underlying model. We consider
+both a Latin hypercube approach and a more systematic investigation of the
+parameters most responsible for impacting the transition rate: the relative 
+distance to the Allee threshold and the noise effects.
+
+The Latin hypercube approach systematically varies all combinations of
+parameters, providing a more general test than varying only one parameter
+at a time.  We loop across eight replicates of three different randomly
+generated parameter sets for each of two different generating models
+(Allen and Myers) over two different noise levels (0.01 and 0.05),
+for a total of 8 x 3 x 2 x 2 = 96 scenarios. The Gaussian Process performs
+nearly optimally in each case, relative to the optimal solution with no
+parameter or model uncertainty (Figure S9, appendix).
 
@@
 sizes fluctuating significantly below the optimal escapement level will
-not perform poorly as long as those fluctuations are sufficently rare.  Figure 7
-shows that the GPDP is little effected by increasing noise or increasing Allee
-effects over much of the range.
-
-<!--FIXME the above para is not particularly convincing. Quantitative please? -->
+not perform poorly as long as those fluctuations are sufficently rare.
+
+Figure 7 shows that the GPDP is little effected by increasing noise or increasing
+Allee effects over much of the range. Much larger stochasticity or higher Allee
+effects make even the optimal solution without any model or parameter uncertainty
+unable to harvest the population effectively (e.g. noise large enoug to violate the
+self-sustaining criterion of @Reed1979). Only at the largest Allee effect
+sizes are some replicates seen to deviate more from the optimal solution.
 

### 2014-06-06 manuscripts/manuscript.Rmd
fix missing )

--- 2014-06-06 manuscripts/manuscript.Rmd
@@
 setA = as.numeric(c(r = p1[1], K = p2[1], theta = p3[1]))
-setB = as.numeric(c(r = p1[2], K = p2[2], theta = p3[2])
+setB = as.numeric(c(r = p1[2], K = p2[2], theta = p3[2]))
 setC = as.numeric(c(r = p1[3], K = p2[3], C = p3[3]))

### 2014-06-06 manuscripts/manuscript.Rmd
library at top

@@
 library("pander")
+
+library("cboettigR")
 
@@
   xlab("Allee threshold stock size") + theme_bw()
-cboettigR::multiplot(plot_sigmas, plot_allees, cols=2)
+
+multiplot(plot_sigmas, plot_allees, cols=2)
 

### 2014-06-06 manuscripts/manuscript.Rmd
multiple chunks for last sensitivity analysis

--- 2014-06-06 manuscripts/manuscript.Rmd
@@
   Myers = list(
-    c(r=1.5 + rnorm(1, 0, 1), theta=2.5 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1)),
+    c(r=1.5 + rnorm(1, 0, 1), theta=2 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1)),
     c(r=1.5 + rnorm(1, 0, 1), theta=2.5 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1))),
@@
 nuisance_values <- list(sigma_g = c(0.01, 0.05))
-
+```
+
+```{r allensets}
 model <- "Allen"
@@
                    seed=c(1111, 2222, 3333))
-
+```
+
+```{r myerssets}
 model <- "Myers"
@@
 ## Extract acutal parameter values corresponding to each parameter set
-p1 = levels(factor(model_dat$pars.r))
-p2 = levels(factor(model_dat$pars.K))
-p3 = levels(factor(model_dat$pars.C))
-
-setA = as.numeric(c(r = p1[1], K = p2[1], theta = p3[1]))
-setB = as.numeric(c(r = p1[2], K = p2[2], theta = p3[2]))
-setC = as.numeric(c(r = p1[3], K = p2[3], C = p3[3]))
-setD = as.numeric(c(r = p1[4], K = p2[4], C = p3[4]))
+p1 = as.numeric(levels(factor(model_dat$pars.r)))
+p2 = as.numeric(levels(factor(model_dat$pars.K)))
+p3 = as.numeric(levels(factor(model_dat$pars.C)))
+
+setA = c(r = p1[1], K = p2[1], theta = p3[1])
+setB = c(r = p1[2], K = p2[2], theta = p3[2])
+setC = c(r = p1[3], K = p2[3], C = p3[3])
+setD = c(r = p1[4], K = p2[4], C = p3[4])
 AllenParameterSets <- rbind(setA, setB)
-MyersParameterSets <- rbind(setC,setD)
+MyersParameterSets <- rbind(setC, setD)
 
@@
 
-```{r}
+```{r AllenSetsTable, results="asis"}
 pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Allen models in Figure S9." )
 ```
-```{r}
-pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Myers models in Figure S9." )
-```
-
-```{r include=FALSE}
+```{r MyersSetsTable, results="asis"}
+pandoc.table(MyersParameterSets, caption="Randomly chosen parameter sets for the Myers models in Figure S9." )
+```
+
+```{r unlink, include=FALSE}
 unlink("ricker_process.bugs")

### 2014-06-23 manuscripts/manuscript.Rmd
additional citation added to references.bib, keep_tex = true
marc's edits integrated

@@
 abstract: |
-          Model uncertainty and limited data coverage are fundamental
+          Model uncertainty and limited data are fundamental
           challenges to robust management of human intervention in a 
@@
           uncertainty. We embed GPs in a
-          stochastic dynamic programming framework in order to make
+          Stochastic Dynamic Programming (SDP) framework in order to make
           robust management predictions with both model uncertainty and
-          limited data. We conduct a set of simulations to evaluate this
+          limited data. We use simulations to evaluate this
           approach as compared with the standard approach of using model
@@
           favors models without tipping points -- leading to harvest policies
-          that guarantee extinction. The GP approach
+          that guarantee extinction. The GPDP 
           performs nearly as well as the true model and significantly
@@
           still fails to account for uncertainty appropriately and leads
-          to population crashes, while management based on the GP 
-          approach does not, since it does not underestimate the uncertainty
+          to population crashes, while management based on the GPDP 
+          does not, since it does not underestimate the uncertainty
           outside of the observed data.
@@
 four decades (but stretching even further back in time), marked by  the
-seminal contributions of @Clark1976; @Clark2009 and @Walters1978,
+seminal contributions of @Clark1976, @Clark2009 and @Walters1978,
 dynamic optimization methods, particularly Stochastic Dynamic Programming
-(SDP, @Clark2009), have become increasingly important as a means of
+(SDP), have become increasingly important as a means of
 understanding how to management human intervention into natural systems.
@@
 under uncertainty.  Global declines [@Worm2006] and the controversy
-surrounding their interpretation [@Hilborn2007; @Worm2009] have made
+surrounding their interpretation [@Hilborn2007; @Worm2009] make
 understanding these challenges all the more pressing.
@@
 is most often concerned with avoiding potentially catastrophic tipping
-points before any data is available at or following a transition that
+points before any data are available at or following a transition that
 would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].
 
-We illustrate how a stochastic dynamic programming (SDP) algorithm
+We illustrate how a Stochastic Dynamic Programming (SDP) algorithm
 [@Mangel1988; @Marescot2013] can be driven by the
@@
 dynamics.  In contrast to parametric models which can only reflect
-uncertainty in parameter estimates, the BNP approach provides a more
+uncertainty in parameter estimates, the BNP approach provides a 
 state-space dependent representation of uncertainty.  This permits
@@
 
-We let $X_t$ denote the size (numbers or biomass) of the focal population at time $t$ and assume that in the absence of take its dynamics are
-
-$$ X_{t+1}= Z_t f(X_t, h_t, p) $$
+We let $X_t$ denote the size (numbers or biomass) of the focal population at time $t$ and assume that in the absence of take its dynamics are.
+NOTE: EQUATION NUMBERS ARE ALL MESSED UP.
+THIS SHOULD BE EQ 1, BUT DOES NOT APPEAR ON THE PDF YOU SENT TO ME
+
+$$ X_{t+1}= Z_t f(X_t) $$
 
@@
 In order to find that policy, we introduce the value function$V_t(x_t)$ representing the total discounted catch from time $t$ onwards given that $X_t=x$.
-This value function satisfies an equation of Stochastic Dynamic Programming  [SDP, @Mangel1988; @Clark2000; @Clark2009], sometimes called the Bellman-Hamilton-Jacobi equation [@Mangel2014],
+This value function satisfies an equation of SDP  [ @Mangel1988; @Clark2000; @Clark2009], sometimes called the Bellman-Hamilton-Jacobi equation [@Mangel2014],
 
@@
 That is, at time $t$, when population size is $x$ and harvest $h_t$ is applied, the immediate return is $R(h_t, X_t)$.
-In the classical case, the sole source of uncertainty is the process noise term, $Z$, and thus the expectation above is equivalent to taking expectations over $Z$.
+In the classical case, the sole source of uncertainty is the process stochasticity term, $Z$, and thus the expectation above is equivalent to taking expectations over $Z$.
 That is
@@
 
-We consider the finite time problem with $T=$`r MaxT`, which we solve using value iteration algorithm [@Mangel1988; @Clark2000].
+We consider the finite time problem with $T=$ `r MaxT`, which we solve using value iteration algorithm [@Mangel1988; @Clark2000].
 
@@
 The core difference for our purpose between the GP representation and the parametric
-models we have already introduced is the "non-parametric" property, which
-means that an estimated GP model is defined explicitly in reference to 
+models is that the GP model is defined explicitly in reference to 
 some observed data.  Unlike the models above, it cannot be specified by
-the value of some parameters alone.  The nonparametric moniker for the
-GP approach is unfortunate as (a) it still involves the estimation of
+the value of some parameters alone.  Although called nonparametric, the
+GP (a)  still involves the estimation of
 parameters, and (b), statisticians also use non-parametric to mean models
@@
 
-Assuming the data $X_o$ are observed with some process noise,
+We assume that the data $X_o$ are observed with some process stochasticity,
 
@@
 
-$\varepsilon$ IID normal, variance $\sigma_n^2$. Note that we have chosen to
-assume additive noise.  While we could just as easily consider log-normal noise
+where $\varepsilon$ are IID normal random variables with zero-mean and variance $\sigma_n^2$. Note that we have chosen to
+assume additive stochasticity.  While we could just as easily consider log-normal stochasticity
 as in the parametric models, we make this choice to bring home the point that the
@@
 
-We use vague inverse Gamma priors on both the lengthscale $\ell$ and process noise $\sigma$ parameters, of the form
+We use vague inverse Gamma priors on both the lengthscale $\ell$ and $\sigma$, of the form
 
@@
 their predicted values $X_{t+1}$ given an observation of the previous
-state, $X_t$. @Munch2005a gives an accessible introduction to the use
+state, $X_t$. @Munch2005a give an accessible introduction to the use
 of Gaussian processes in providing a Bayesian nonparametric description
@@
 S13, code in appendix), under weakly informative Gaussian priors
-(see parameters in table S5). As the posterior distributions differ
+(see parameters in table S5). Since the posterior distributions differ
 substantially from the priors (Figure S13), we can be assured that most
@@
 
-The method of Gaussian Process Dynamic Programming
+The method of Gaussian Process Dynamic Programming (GPDP)
 --------------------------------------------------
@@
 To ensure our results are robust to our choice of parameters, 
-we perform our analysis over 96 different scenarios. To better
-understand the process, we will first describe in detail the
+consider 96 different scenarios, described in detail below.   To begin, to help  better
+understand the process, we first describe in detail the
 results of a single scenario.
@@
 
-All models fit the observed data rather closely and with relatively small uncertainty, as illustrated in the posterior predictive curves in Figure 1.
-Figure 1 shows the training data of stock sizes observed over time as points, overlaid with the step-ahead predictions of each estimated model using the parameters sampled from their posterior distributions.
+All models fit the observed data rather closely and with relatively small uncertainty, as illustrated in the posterior predictive curves in Figure 1, in which we show the training data of stock sizes observed over time as points, overlaid with the step-ahead predictions of each estimated model using the parameters sampled from their posterior distributions.
 Each model manages to fit the observed data rather closely.
 Compared to the true model most estimates appear to over-fit, predicting fluctuations that are actually due purely to stochasticity.
-All model selection criteria shown in Table 1 penalize more complex models and show a preference for the simpler Ricker model over the more complicated alternate stable state models (Allen and Myers).
+Model selection criteria (Table 1) penalize more complex models and show a preference for the simpler Ricker model over the  models with alternative stable states (Allen and Myers).
 Details on MCMC estimates for each model, traces, and posterior distributions can be found in the appendix.
+
+NOTE: IN THE PDF YOU SENT THERE IS WEIRD STUFF ON LINE 237
 
@@
 
-```{r figure_2, dependson=c("assemble-models", "plot-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
-Graph shows the expected value for the function $f$ under each model.
+```{r figure_2, dependson=c("assemble-models", "plot-options"), fig.cap="The inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
+We show the expected value for the function $f$ under each model.
 Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.
-The training data is also shown as black points.
+The training data are also shown as black points.
 The GP is conditioned on (0,0), shown as a pseudo-data point.", fig.width=8, fig.height=6}
@@
 
-The mean inferred state space dynamics of each model relative to the
-true model used to generate the data is shown in Figure 2, predicting
+We show the  mean inferred state space dynamics of each model relative to the
+true model used to generate the data in Figure 2, predicting
 the relationship between observed stock size (x-axis) to the stock
@@
 has a mean shown in black. Parameter uncertainty (not shown) spreads
-out the estimates further.  The observed data from which each model is
+out the estimates further.  The observed data from which each model are
 estimated is also shown.  The observations come from only a limited region
@@
 the tipping point.
+
+NOTE: THERE IS NO MENTION OF FIGURE 3 IN THE TEXT
 
@@
 Despite the similarities in model fits to the observed data, the
-policies inferred under each model differ widely, as shown in Figure 4.
+policies inferred under each model differ widely (Figure 4).
 Policies are shown in terms of target escapement, $S_t$.  Under models
@@
 ```{r figure_5, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models.
-Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
+We show 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
 ggplot(sims_data) + 
@@
 
-The consequences of managing 100 replicate realizations of the simulated
-fishery under each of the policies estimated is shown in Figure 5.
-As expected from the policy curves, the structurally correct model
-under-harvests, leaving the stock to vary around it's un-fished optimum.
+In Figure 5, we show the consequences of managing 100 replicate realizations of the simulated
+fishery under each of the policies estimated..
+As anticipated from the policy curves, the structurally correct model
+under-harvests, leaving the stock to vary around its unfished optimum.
 The structurally incorrect Ricker model over-harvests the population
@@
 easily be compared by using the relative differences in net present
-value realized by each of the model. Figure 6 shows the distribution of values
-realized by each simulation shown in Figure 5.  The GPDP
+value realized by each of the model (Figure 6).  The GPDP
 most consistently realizes a value close to the optimal solution,
@@
 
-Changing the noise or the distance between stable and unstable points
+Changing the intensity of the stochasticity or the distance between stable and unstable points
 does not impact the performance of the GP relative to the optimal solution
@@
 
-(following the Ahrenius relationship, see @Gardiner2009 for the derivation).
+(see @Gardiner2009 or @Mangel2006 for the derivation).
 As a result, the impact of using a model which underestimates the risk of harvesting
@@
 
-Figure 7 shows that the GPDP is little effected by increasing noise or increasing
-Allee effects over much of the range. Much larger stochasticity or higher Allee
+The GPDP is weakly influenced by increasing stochasticity or increasing
+Allee effects over much of the range (Figure 7) Much larger stochasticity or higher Allee
 effects make even the optimal solution without any model or parameter uncertainty
@@
 ```{r figure_7, fig.width=6, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).
-Other than the focal parameter (noise, Allee threshold), other parameters are held fixed as above to illustrate this effect."}
+Other than the focal parameter (stochasticity, Allee threshold), other parameters are held fixed as above to illustrate this effect."}
 
@@
 
+NOTE: THE PDF HAS TWO }} IN LINE 351
+
 \appendix

### 2014-06-23 manuscripts/manuscript.Rmd
steve edits

@@
 (SDP), have become increasingly important as a means of
-understanding how to management human intervention into natural systems.
-Concomittantly, there has been increasing recognition of the importance of
+understanding how to manage human intervention into natural systems.
+Simultaneously, there has been increasing recognition of the importance of
 Allee population sizes or `tipping points' [@Scheffer2001;
@@
 'best-case' scenario for the parametric model performance relative to
-the GP.   This is abest-case scenario for the parametric models because the underlying dynamics are always simuated
-from one the parametric models in the set, whereas in reality we never
+the GP.   This is a best-case scenario for the parametric models because the underlying dynamics are always simuated
+from one of the parametric models in the set, whereas in reality we never
 have the "true" model among our candidates.  By choosing one-dimensional
@@
 We let $X_t$ denote the size (numbers or biomass) of the focal population at time $t$ and assume that in the absence of take its dynamics are.
-NOTE: EQUATION NUMBERS ARE ALL MESSED UP.
-THIS SHOULD BE EQ 1, BUT DOES NOT APPEAR ON THE PDF YOU SENT TO ME
-
-$$ X_{t+1}= Z_t f(X_t) $$
-
-Where Z(t) is log-normally distributed process stochasticity [@Reed1979] and $p$ is a vector of parameters to be estimated from the data.
-We describe three choices for $f(X_t,p)$ in the next section.In this simple model, the control action is a harvest or take, $h_t$, measured in the same units as $X$, at time
-$t$.  Thus, in the presence of take, the population size on the right hand side of Eqn 1 is replaced by
+
+\begin{equation}
+X_{t+1}= Z_t f(X_t) \label{eq1} 
+\end{equation}
+
+Where Z(t) is log-normally distributed process stochasticity [@Reed1979]
+and $p$ is a vector of parameters to be estimated from the data.
+We describe three choices for $f(X_t,p)$ in the next section.In this
+simple model, the control action is a harvest or take, $h_t$, measured
+in the same units as $X$, at time $t$.  Thus, in the presence of take,
+the population size on the right hand side of Eqn 1 is replaced by
 $S_t=X_t-h_t$.
 
-To construct the value function, we consider a return when $X_t=x$ and harvest $h_t=h$ denoted as the reward, $R(x,h)$.
-For example, if the return is simply the harvest at time $t$, then $R(x,h)=min(x,h)$.
-We assume that future harvests are discounted relative to current ones at a constant rate of discount $\delta$ and ask for the harvest policy that maximizes total discounted harvest between the current time $t$ and a final time $T$, that is we seek to maximize over choices of harvest $\mathbf{E}_{X_t+1} [ \sum_{t = 0}^{T}  R_t(X_t, h_t) \delta^t]$, where the state dynamics are given by Eqn 1 and $\mathbf{E}$ denotes the expectation over the process stochasticity in the future population state.
-
-In order to find that policy, we introduce the value function$V_t(x_t)$ representing the total discounted catch from time $t$ onwards given that $X_t=x$.
-This value function satisfies an equation of SDP  [ @Mangel1988; @Clark2000; @Clark2009], sometimes called the Bellman-Hamilton-Jacobi equation [@Mangel2014],
+To construct the value function, we consider a return when $X_t=x$ and
+harvest $h_t=h$ denoted as the reward, $R(x,h)$. For example, if the
+return is simply the harvest at time $t$, then $R(x,h)=min(x,h)$. We
+assume that future harvests are discounted relative to current ones at
+a constant rate of discount $\delta$ and ask for the harvest policy
+that maximizes total discounted harvest between the current time $t$
+and a final time $T$. That is, we seek to maximize over choices of
+harvest $\mathbf{E}_{X_t+1} [ \sum_{t = 0}^{T}  R_t(X_t, h_t) \delta^t]$,
+where the state dynamics are given by Eqn 1 and $\mathbf{E}$ denotes the
+expectation over the process stochasticity in the future population state.
+
+In order to find that policy, we introduce the value function $V_t(x_t)$
+representing the total discounted catch from time $t$ onwards
+given that $X_t=x_t$.  This value function satisfies an equation of
+SDP  [@Mangel1988; @Clark2000; @Clark2009], sometimes called the
+Hamilton-Jacobi-Bellman equation [@Mangel2014],
 
 \begin{equation}
-V_t(X_t) = \max_{h_t} \lbrace R(h_t, X_t) + \delta \cdot \mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] \rbrace
+V_t(x_t) = \max_{h_t} \lbrace R(h_t, x_t) + \delta \cdot \mathbf{\mathrm{E}}_{X_{t+1}} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] \rbrace
 \end{equation}
 
-where expectation is taken over all possible values of the next state, $X_t{+1}$, and maximized over all possible choices of harvest, $h_t$.
-That is, at time $t$, when population size is $x$ and harvest $h_t$ is applied, the immediate return is $R(h_t, X_t)$.
-In the classical case, the sole source of uncertainty is the process stochasticity term, $Z$, and thus the expectation above is equivalent to taking expectations over $Z$.
-That is
-
-$$ \mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{Z} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] $$
-
-where the population size after the take is $X_t-h_t$, which is then translated into $X_{t+1}$ by Eqn 1.
-
-In a Bayesian decision framework, the parameters governing the dynamics are also uncertain, so the
-expectation in (1) involves averaging over the posterior distribution for the parameters, as well. That
-is,
-
-$$\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta,\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \}$$
+where expectation is taken over all possible values of the next state,
+$X_{t+1}$, and maximized over all possible choices of harvest, $h_t$.
+That is, at time $t$, when population size is $x_t$ and harvest $h_t$ is
+applied, the immediate return is $R(h_t, x_t)$. In the classical case,
+the sole source of uncertainty is the process stochasticity term, $Z$,
+and thus the expectation above is equivalent to taking expectations over
+$Z$. That is
+
+$$ \mathbf{\mathrm{E}}_{X_{t+1}} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] = \mathbf{\mathrm{E}}_{Z} \left[ V_{t+1}( Z f(x_t - h_t))  | x_t, h_t \right] $$
+
+where the population size after the take is $X_t-h_t$, which is then
+translated into $X_{t+1}$ by Eqn 1.
+
+In a Bayesian decision framework, the parameters governing the dynamics
+are also uncertain, so the expectation involves averaging over the
+posterior distribution for the parameters, as well. That is,
+
+$$\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(x_t - h_t))  | x_t, h_t \right] \}$$
 
@@
 
-$$\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta,\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \}$$
-
-We consider the finite time problem with $T=$ `r MaxT`, which we solve using value iteration algorithm [@Mangel1988; @Clark2000].
+$$\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | x_t, h_t \right] \}$$
+
+We consider the finite time problem with $T=$ `r MaxT`, which we
+solve using the standard value iteration algorithm [see @Mangel1988;
+@Clark2000].
 
@@
 dynamics: The Ricker model, the Allen model [@Allen2005], the Myers
-model [@Myers1995], equations (@ricker)-(@myers), which we parameterize as follows for $f(S_t,p)$ in Eqn 1.
+model [@Myers1995], Eqns \eqref{ricker}-\eqref{myers}, which we parameterize as follows for $f(S_t,p)$ in Eqn 1.
 In all three, we let $K$ denote the carrying capacity and $r$ the maximum per capita growth rate.
@@
 
-(@ricker) $$ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K} \right) } $$
+\begin{equation}
+f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K} \right) } \label{ricker}
+\end{equation}
 
@@
 
-(@allen) $$ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_c\right)} $$
+\begin{equation}
+f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_c\right)} \label{allen}
+\end{equation}
 
@@
 
-(@myers) $$ f(S(t) | p)  = \frac{r S_t^{\theta}}{1 + \frac{S_t^\theta}{K}}$$
+\begin{equation}
+f(S(t) | p)  = \frac{r S_t^{\theta}}{1 + \frac{S_t^\theta}{K}} \label{myers}
+\end{equation}
 
@@
 enough that the estimates do not depend on the particular realization,
-while longer times are not likely to provide substantial improvement
-(i.e. the results are not sensitive to this interval).  Each of the models
-(described below) is fit to the same training data (Figure 1).
+while longer times are not likely to provide substantial improvement.  
+Each of the models is fit to the same training data (Figure 1).
 
@@
 We inferred posterior distributions for the parameters of each model by Gibbs sampling (@Gelman2003 implemented in R
-[@RTeam] using jags, [@R2jags]).  We choose uniform priors
+[@RTeam] using `jags`, [@R2jags]).  We choose uniform priors
 for all parameters (See appendix Tables S1-S3; R code provided). We
@@
 
-The use GP methods to formulate a predictive
+The use of GP methods to formulate a predictive
 model is relatively new in the context of modeling dynamical systems
@@
 A GP is an infinite dimensional (because it is a function of a continuous input
-variable) stochastic process for which any realization is a multivariate
+variable) stochastic process for a finite sample of points from any realization is a multivariate
 normal distribution. Thus, to characterize the GP we need a mean function
-and a covariance matrix. We proceed as follows.
+and a covariance function. We proceed as follows.
 
@@
 
-We use vague inverse Gamma priors on both the lengthscale $\ell$ and $\sigma$, of the form
+We use inverse Gamma priors on both the lengthscale $\ell$ and $\sigma$, of the form
 
@@
 infer posterior distributions of the two parameters of the GP (Figure
-S13, code in appendix), under weakly informative Gaussian priors
-(see parameters in table S5). Since the posterior distributions differ
+S13, code in appendix). Since the posterior distributions differ
 substantially from the priors (Figure S13), we can be assured that most
@@
 prior belief.
-
-<!-- SDP via the model  FIXME much more detail here-->
 
@@
 To ensure our results are robust to our choice of parameters, 
-consider 96 different scenarios, described in detail below.   To begin, to help  better
+consider 96 different scenarios, described in detail below.   To help  better
 understand the process, we first describe in detail the
@@
 
-NOTE: IN THE PDF YOU SENT THERE IS WEIRD STUFF ON LINE 237
-
 ```{r deviances, dependson=c("dic_calc"), include=FALSE, echo=FALSE}
@@
                        row.names = c("BIC"))
-xtable::xtable(rbind(dictable, aictable, bictable))
+pandoc.table(rbind(dictable, aictable, bictable))
 ```
@@
 ```{r Table1, dependson="deviances", include=TRUE, results="asis", echo=FALSE, cache=FALSE}
-xtable::xtable(rbind(dictable, aictable, bictable), caption="Model selection scores for several common criteria all (wrongly) select the simplest model.
+pandoc.table(rbind(dictable, aictable, bictable), caption="Model selection scores for several common criteria all (wrongly) select the simplest model.
 As the true (Allen) model is not distinguishable from the simpler (Ricker) model in the region of the observed data, this error cannot be avoided regardless of the model choice criterion.
@@
 
-NOTE: THERE IS NO MENTION OF FIGURE 3 IN THE TEXT
+Figure 3 illustrates the performance of the GP and parametric models
+outside the observed training data. The mean trajectory under the
+underlying model is shown by the black dots, while the corresponding
+prediction made by the model shown by the box and whiskers plots.
+Predictions are based on the true expected value in the previous time
+step.  Predicted distributions that lie entirely above the expected
+dynamics indicate the expectation of stock sizes higher than what is
+actually expected. The models differ both in their expectations and
+their uncertainty (colored bands show two standard deviations away).
+Note that the GP is particularly uncertain about the dynamics relative
+to structurally incorrect models like the Ricker.
 
@@
 
-In Figure 5, we show the consequences of managing 100 replicate realizations of the simulated
-fishery under each of the policies estimated..
-As anticipated from the policy curves, the structurally correct model
-under-harvests, leaving the stock to vary around its unfished optimum.
-The structurally incorrect Ricker model over-harvests the population
-passed the tipping point consistently, resulting in the immediate crash
-of the stock and thus derives minimal long-term catch.
+In Figure 5, we show the consequences of managing 100 replicate
+realizations of the simulated fishery under each of the policies
+estimated.  As anticipated from the policy curves, the structurally
+correct model under-harvests, leaving the stock to vary around its
+unfished optimum.  The structurally incorrect Ricker model over-harvests
+the population passed the tipping point consistently, resulting in the
+immediate crash of the stock and thus derives minimal long-term catch.
 
@@
 noise scale, parameters or structure of the underlying model. We consider
-both a Latin hypercube approach and a more systematic investigation of the
+both a Latin hypercube approach and a more focused investigation of the
 parameters most responsible for impacting the transition rate: the relative 
@@
 optimal escapement and unstable points relative to the noise, the chance
-of a transition becomes vanishingly small and all models can be modeled
+of a transition becomes vanishingly small and all models can be estimated 
 near-optimally. Models that underestimate the cost incurred by stock
@@
 the case whenever there is a potential concern about tipping point dynamics --
-simpler models therefore fit just as well and tend to be selected more foten than complex 
+simpler models therefore fit just as well and tend to be selected more often than complex 
 ones.  This approach would be appropriate when the dynamics can be expected 
@@
 
-### GPDP population dynamcis capture larger uncertainty in regions where the data are poor
-
-<!-- FIXME More emphasis/explanation on _how_ it does this -->
+### GPDP population dynamics capture larger uncertainty in regions where the data are poor
 
@@
 good predictive accuracy in the range observed.  The management policy 
-dictated by the GP balance this uncertainty against
+dictated by the GP balances this uncertainty against
 the immediate value of the harvest, and act to stabilize the population 
@@
 $\ell$ and $\sigma$ and simple (mean zero, radial basis function kernel).
-In practice, both the choice of mean and the choice of the covariance
+In practice, both the choice of mean and the covariance
 function may be chosen to confer particular biological properties,
@@
 
-NOTE: THE PDF HAS TWO }} IN LINE 351
-
 \appendix
@@
 \renewcommand*{\thetable}{S\arabic{table}}
-\setcounter{figure}{0}}
-\setcounter{table}{0}}
+\setcounter{figure}{0}
+\setcounter{table}{0}
 
@@
 
-setA = c(r = p1[1], K = p2[1], theta = p3[1])
-setB = c(r = p1[2], K = p2[2], theta = p3[2])
-setC = c(r = p1[3], K = p2[3], C = p3[3])
-setD = c(r = p1[4], K = p2[4], C = p3[4])
-AllenParameterSets <- rbind(setA, setB)
-MyersParameterSets <- rbind(setC, setD)
+set.A = c(r = p1[1], K = p2[1], theta = p3[1])
+set.B = c(r = p1[2], K = p2[2], theta = p3[2])
+set.C = c(r = p1[3], K = p2[3], C = p3[3])
+set.D = c(r = p1[4], K = p2[4], C = p3[4])
+AllenParameterSets <- rbind(set.A, set.B)
+MyersParameterSets <- rbind(set.C, set.D)
 

### 2014-06-23 manuscripts/manuscript.Rmd
small typo fixes

@@
 abstract: |
-          Model uncertainty and limited data are fundamental
-          challenges to robust management of human intervention in a 
-          natural system.  These challenges
-          are acutely highlighted by concerns that many ecological systems
-          may contain tipping points, such as Allee population sizes.  
-          Before a collapse, we do not know
+  Model uncertainty and limited data are fundamental challenges to
+  robust management of human intervention in a natural system.
+  These challenges are acutely highlighted by concerns that
+  many ecological systems may contain tipping points, such as
+  Allee population sizes.  Before a collapse, we do not know
   where the tipping points lie, if they exist at all.  Hence,
@@
   do we have access to data in some large region of state-space
-          where such a tipping point might exist.  We illustrate
-          how a Bayesian Non-Parametric (BNP) approach using a Gaussian Process (GP) prior
-          provides a flexible representation of this inherent
-          uncertainty. We embed GPs in a
-          Stochastic Dynamic Programming (SDP) framework in order to make
-          robust management predictions with both model uncertainty and
-          limited data. We use simulations to evaluate this
-          approach as compared with the standard approach of using model
-          selection to choose from a set of candidate models.  We find that model selection erroneously
-          favors models without tipping points -- leading to harvest policies
-          that guarantee extinction. The GPDP 
-          performs nearly as well as the true model and significantly
-          outperforms standard appraoches. We illustrate this using examples
-          of simulated single-species dynamics, where the standard model
-          selection approach should be most effective, and find that it
-          still fails to account for uncertainty appropriately and leads
-          to population crashes, while management based on the GPDP 
-          does not, since it does not underestimate the uncertainty
-          outside of the observed data.
+  where such a tipping point might exist.  We illustrate how
+  a Bayesian Non-Parametric (BNP) approach using a Gaussian
+  Process (GP) prior provides a flexible representation of this
+  inherent uncertainty. We embed GPs in a Stochastic Dynamic
+  Programming (SDP) framework in order to make robust management
+  predictions with both model uncertainty and limited data. We
+  use simulations to evaluate this approach as compared with the
+  standard approach of using model selection to choose from a set
+  of candidate models.  We find that model selection erroneously
+  favors models without tipping points -- leading to harvest
+  policies that guarantee extinction. The GPDP performs nearly as
+  well as the true model and significantly outperforms standard
+  approaches. We illustrate this using examples of simulated
+  single-species dynamics, where the standard model selection
+  approach should be most effective, and find that it still
+  fails to account for uncertainty appropriately and leads to
+  population crashes, while management based on the GPDP does
+  not, since it does not underestimate the uncertainty outside
+  of the observed data.
 
@@
 a model that provides the best mechanistic understanding or the best
-long-term outcome, but rather the one that idelivers the best management
+long-term outcome, but rather the one that delivers the best management
 performance over the period of interest.
@@
 
-We first describe the requirements of dynamic optimization for the management of human intervention in natural resource systems.
-After that we describe three parametric models for population dynamics and the GP description of population dynamics.
+We first describe the requirements of dynamic optimization for the
+management of human intervention in natural resource systems. After that
+we describe three parametric models for population dynamics and the GP
+description of population dynamics.
 
@@
 
-Dynamic optimization  requires characterizing the dynamics of a state variable (or variables), a control action, and a value function.
-In this paper, we consider only a single state variable, so that the ideas can be described as simply as possible. That is, we focus on a class of one-dimensional models of population dynamics as a
-'best-case' scenario for the parametric model performance relative to
-the GP.   This is a best-case scenario for the parametric models because the underlying dynamics are always simuated
-from one of the parametric models in the set, whereas in reality we never
-have the "true" model among our candidates.  By choosing one-dimensional
-models with few parameters, we limit the chance that poor performance
-will be due to our inability to estimate parameters well, something
-that becomes a more severe problem for higher-dimensional parametric
-models. Further, the parametric models we consider are those most commonly
-used in modeling stock-recruitment dynamics or to model sudden transitions
-between alternative stable states.
+Dynamic optimization  requires characterizing the dynamics of a state
+variable (or variables), a control action, and a value function.
+In this paper, we consider only a single state variable, so that the
+ideas can be described as simply as possible. That is, we focus on a
+class of one-dimensional models of population dynamics as a 'best-case'
+scenario for the parametric model performance relative to the GP.
+This is a best-case scenario for the parametric models because the
+underlying dynamics are always simulated from one of the parametric models
+in the set, whereas in reality we never have the "true" model among our
+candidates.  By choosing one-dimensional models with few parameters, we
+limit the chance that poor performance will be due to our inability to
+estimate parameters well, something that becomes a more severe problem
+for higher-dimensional parametric models. Further, the parametric models
+we consider are those most commonly used in modeling stock-recruitment
+dynamics or to model sudden transitions between alternative stable states.
 
@@
 
-$$X_{t+1} = f(X_t) + \varepsilon$$
+$$X_{t+1} = f(X_t) + \varepsilon,$$
 
@@
 
-We use inverse Gamma priors on both the lengthscale $\ell$ and $\sigma$, of the form
+We use inverse Gamma priors on both the length-scale $\ell$ and $\sigma$, of the form
 
@@
 given the true model and true parameters.  The parametric models are
-more senstive to this difference.  Large noise relative to the distance between the stable and unstable
+more sensitive to this difference.  Large noise relative to the distance between the stable and unstable
 point increases the chance of a stochastic transition.  More precisely,
@@
 sizes fluctuating significantly below the optimal escapement level will
-not perform poorly as long as those fluctuations are sufficently rare.
+not perform poorly as long as those fluctuations are sufficiently rare.
 
@@
 effects make even the optimal solution without any model or parameter uncertainty
-unable to harvest the population effectively (e.g. noise large enoug to violate the
+unable to harvest the population effectively (e.g. noise large enough to violate the
 self-sustaining criterion of @Reed1979). Only at the largest Allee effect

### 2014-07-01 manuscripts/manuscript.Rmd
as submitted

@@
 linenumbers: true
-title: "Avoiding tipping points in fisheries management: A Bayesian Non-Parametric approach"
+title: "Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming"
 author: 

### 2014-07-02 manuscripts/manuscript.Rmd
keep tex true
no transparency in figures. actually as submitted now

@@
     template: components/elsarticle.latex
+    keep_tex: true
 
@@
 plot_allen_posteriors <- ggplot(allen_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
+  stat_density(geom="path", position="identity") +
 #  geom_line(data=allen_priors, aes(x=value, y=density), col="red") +  
@@
 plot_ricker_posteriors <- ggplot(ricker_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
+  stat_density(geom="path", position="identity") +
 #  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") +  # don't plot priors 
@@
 plot_myers_posteriors <- ggplot(myers_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
+  stat_density(geom="path", position="identity") +
 #  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") +  # Whoops, these are misaligned. see table instead 
@@
 figure1b_posteriors <- ggplot(df_post) + geom_point(aes(time, stock)) + 
-  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
+  geom_line(aes(time, value, col=variable, group=interaction(L1,variable))) + 
   facet_wrap(~variable) + 
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) +  
+  scale_colour_manual(values=colorkey) +  
   theme(legend.position="none")
@@
     geom_line(aes(stock, value, col = variable, 
-                  group=interaction(L1,variable)), 
-              alpha=.2) + 
-    geom_point(data = obs, aes(x,y), alpha = 0.8) + 
+                  group=interaction(L1,variable))) + 
+    geom_point(data = obs, aes(x,y)) + 
     xlab(expression(X[t])) + ylab(expression(X[t+1])) +
@@
                    fill = variable, col=variable), 
-               alpha=.7, outlier.size=1, position="identity") + 
+               outlier.size=1, position="identity") + 
 #  geom_line(aes(time, value, col = variable, 
-#            group=interaction(L1,variable)), alpha=.1) + 
+#            group=interaction(L1,variable))) + 
   geom_point(aes(time, stock), size = 3) + 
-  scale_fill_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")], 
-                      guide = guide_legend(override.aes = list(alpha = 1))) +  
-  scale_colour_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")], 
-                      guide = guide_legend(override.aes = list(alpha = 1))) +  
+  scale_fill_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")]) +  
+  scale_colour_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")]) +  
   facet_wrap(~variable) + 
@@
 ggplot(policies, aes(stock, stock - value, color=method)) +
-  geom_line(lwd=1.2, alpha=0.8) + xlab("stock size") + ylab("escapement")  +
+  geom_line(lwd=1.2) + xlab("stock size") + ylab("escapement")  +
   scale_colour_manual(values=colorkey)
@@
 ggplot(sims_data) + 
-  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) + 
+  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method)) +
+  scale_colour_manual(values=colorkey) + 
   facet_wrap(~method) + guides(legend.position="none")
@@
 # ggplot(actual_over_optimal, aes(value, fill=variable, color=variable)) + # density plots fail when delta fn
-#  stat_density(aes(y=..density..), position="stack", adjust=3, alpha=.9) + 
+#  stat_density(aes(y=..density..), position="stack", adjust=3) + 
 #  xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)+ scale_color_manual(values=colorkey)
@@
 ggplot(ricker_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
+  stat_density(geom="path", position="identity") +
   facet_wrap(~ variable, scale="free", ncol=2)
@@
 ggplot(myers_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
+  stat_density(geom="path", position="identity") +
   facet_wrap(~ variable, scale="free", ncol=2)
@@
 ggplot(allen_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity", alpha=0.7) +
+  stat_density(geom="path", position="identity") +
   facet_wrap(~ variable, scale="free", ncol=2)

### 2014-07-08 manuscripts/manuscript.Rmd
restore alpha on figs 1,2,5, separate out supplement for submission, set pdf to v1.3

@@
 <!-- Run R code for analysis, to be called by figures -->
-```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE, cache=FALSE}
 cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
                "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+opts_chunk$set(dev='pdf', dev.args=c(version="1.3"))
 
@@
 figure1b_posteriors <- ggplot(df_post) + geom_point(aes(time, stock)) + 
-  geom_line(aes(time, value, col=variable, group=interaction(L1,variable))) + 
+  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=0.2) + 
   facet_wrap(~variable) + 
@@
     geom_line(aes(stock, value, col = variable, 
-                  group=interaction(L1,variable))) + 
+                  group=interaction(L1,variable)), alpha=0.2) + 
     geom_point(data = obs, aes(x,y)) + 
@@
 ggplot(sims_data) + 
-  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method)) +
+  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=0.1) +
   scale_colour_manual(values=colorkey) + 
@@
 grant EF-0924195 to MM and NSF grant DBI-1306697 to CB.
-
-\appendix
-\renewcommand*{\thefigure}{S\arabic{figure}}
-\renewcommand*{\thetable}{S\arabic{table}}
-\setcounter{figure}{0}
-\setcounter{table}{0}
-
-Appendix: Model definitions and estimation
-===========================================
-
-```{r figure_S1, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options"}
-gp_assessment_plots$traces_plot
-```
-
-```{r figure_S2, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
-gp_assessment_plots$posteriors_plot
-```
-
-\newpage
-
-Training data (Figures 1-6)
-----------------------------
-
-Each of our models $f(S_t)$ must be estimated from training data, which
-we simulate from the Allen model with parameters $r = $ `r p[1]`, 
-$K =$ `r p[2]`, $C =$ `r p[3]`, and  $\sigma_g =$ `r sigma_g` 
-for $T=$ `r Tobs` timesteps, starting at initial condition $X_0 = $ `r Xo`. 
-The training data can be seen in Figure 1.
-
-Training data for sensitivity analyses
--------------------------------------
-
-A further 96 unique training data sets are generated for the sensitivity analysis, as described in the main text.
-
-Code
-----
-
-All code used in producing this analysis can be found in the R package accompanying the paper, available from [http://github.com/cboettig/nonparametric-bayes].
-The code for all results shown here is dynamically embedded into the manuscript using `knitr` [@knitr].
-
-<!--
-Data
-----
-
-While the data can be regenerated using the code provided, for convenience CSV files of the data shown in each graph are also provided, along with appropriate metadata written in the Ecological Metadata Language (EML).
-
--->
-
-```{r figure_S3, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
-plot_ricker_traces
-```
-
-```{r figure_S4, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
-ggplot(ricker_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity") +
-  facet_wrap(~ variable, scale="free", ncol=2)
-write.csv(ricker_posteriors, "components/data/ricker_posteriors.csv")
-```
-
-```{r Table S1, results = "asis"}
-pander::pandoc.table(ricker_priors_xtable,
-  caption = "Parameterization range for the uniform priors in the Ricker model")
-```
-
-```{r figure_S5, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
-plot_myers_traces
-```
-
-```{r figure_S6, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
-ggplot(myers_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity") +
-  facet_wrap(~ variable, scale="free", ncol=2)
-write.csv(myers_posteriors, "components/data/myers_posteriors.csv")
-```
-
-```{r TableS2, results="asis"}
-pander::pandoc.table(myers_priors_xtable,
-           caption = "Parameterization range for the uniform priors in the Myers model")
-```
-
-```{r figure_S7, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
-plot_allen_traces
-```
-
-```{r figure_S8, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
-ggplot(allen_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity") +
-  facet_wrap(~ variable, scale="free", ncol=2)
-write.csv(allen_posteriors, "components/data/allen_posteriors.csv")
-```
-
-```{r TableS3, results = "asis"}
-pander::pandoc.table(allen_priors_xtable,
-  caption = "Parameterization range for the uniform priors in the Allen model")
-```
-
-```{r sensitivity-calc}
-source("components/sensitivity.R")
-
-models <- c("Myers","Allen")
-
-parameters <- list(
-  Myers = list(
-    c(r=1.5 + rnorm(1, 0, 1), theta=2 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1)),
-    c(r=1.5 + rnorm(1, 0, 1), theta=2.5 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1))),
-  Allen = list(
-    c(r=2 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1), C=4 + rnorm(1, 0, 1)),
-    c(r=2 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1), C=4 + rnorm(1, 0, 1)))
-)
-nuisance_pars <- c("sigma_g")
-nuisance_values <- list(sigma_g = c(0.01, 0.05))
-```
-
-```{r allensets}
-model <- "Allen"
-allen1.01 <- sensitivity(model, 
-                   parameters = parameters[[model]][[1]], 
-                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
-                   seed=c(1111, 2222, 3333))
-allen2.01 <- sensitivity(model, 
-                   parameters = parameters[[model]][[2]], 
-                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
-                   seed=c(1111, 2222, 3333))
-allen1.05 <- sensitivity(model, 
-                   parameters = parameters[[model]][[1]], 
-                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
-                   seed=c(1111, 2222, 3333))
-allen2.05 <- sensitivity(model, 
-                   parameters = parameters[[model]][[2]], 
-                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
-                   seed=c(1111, 2222, 3333))
-```
-
-```{r myerssets}
-model <- "Myers"
-Myers1.01 <- sensitivity(model, 
-                   parameters = parameters[[model]][[1]], 
-                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
-                   seed=c(1111, 2222, 3333))
-Myers2.01 <- sensitivity(model, 
-                   parameters = parameters[[model]][[2]], 
-                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
-                   seed=c(1111, 2222, 3333))
-Myers1.05 <- sensitivity(model, 
-                   parameters = parameters[[model]][[1]], 
-                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
-                   seed=c(1111, 2222, 3333))
-Myers2.05 <- sensitivity(model, 
-                   parameters = parameters[[model]][[2]], 
-                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
-                   seed=c(1111, 2222, 3333))
-
-## Assemble into data.frame
-allen_dat <- rbind(allen1.01, allen1.05, allen2.01, allen2.05) 
-m <- rbind(Myers1.01, Myers1.05, Myers2.01, Myers2.05)
-myers_dat <- m[c(1:2,4,3,5:8)]
-names(myers_dat) <- names(allen_dat)
-model_dat <- rbind(allen_dat, myers_dat)
-dat <- model_dat
-dat$pars.r <- factor(dat$pars.r, labels=c("A", "B", "C", "D"))
-dat <- dat[c(1:2,5:6, 8, 7)]
-dat$noise <- factor(dat$noise)
-names(dat) <- c("model", "parameters", "replicate", "simulation", "noise", "value")
-
-## Extract acutal parameter values corresponding to each parameter set
-p1 = as.numeric(levels(factor(model_dat$pars.r)))
-p2 = as.numeric(levels(factor(model_dat$pars.K)))
-p3 = as.numeric(levels(factor(model_dat$pars.C)))
-
-set.A = c(r = p1[1], K = p2[1], theta = p3[1])
-set.B = c(r = p1[2], K = p2[2], theta = p3[2])
-set.C = c(r = p1[3], K = p2[3], C = p3[3])
-set.D = c(r = p1[4], K = p2[4], C = p3[4])
-AllenParameterSets <- rbind(set.A, set.B)
-MyersParameterSets <- rbind(set.C, set.D)
-
-sensitivity_dat <- dat
-```
-
-```{r figure_S9, fig.height=6, fig.width=10, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
-Histograms shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters.
-Values of 1 indicate optimal performance.
-Columns indicate different models, rows different noise levels, and colors indicate the parameter set used.
-Grouped over stochastic replicates applying the contol policy and stochastic replicates of training data generated from the model indicated, see raw data for details.
-Randomly chosen parameter values for the models shown in tables below."}
-ggplot(sensitivity_dat) + 
-  geom_histogram(aes(value, fill=parameters)) + 
-  xlim(0,1.0) + 
-  theme_bw() + 
-  xlab("value as fraction of the optimal") + 
-  facet_grid(noise~model)
-write.csv(sensitivity_dat, "components/data/sensitivity_dat.csv") 
-```
-
-```{r AllenSetsTable, results="asis"}
-pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Allen models in Figure S9." )
-```
-```{r MyersSetsTable, results="asis"}
-pandoc.table(MyersParameterSets, caption="Randomly chosen parameter sets for the Myers models in Figure S9." )
-```
-
-```{r unlink, include=FALSE}
-unlink("ricker_process.bugs")
-unlink("allen_process.bugs")
-unlink("myers_process.bugs")
-```

### 2014-07-08 manuscripts/supplement.Rmd
supplement as separate file

---
layout: 12pt,review
linenumbers: true
title: "Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
  - name: Marc Mangel
    affiliation: cstar
  - name: Stephan Munch
    affiliation: noaa
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research, 
      Department of Applied Math and Statistics, 
      University of California, Mail Stop SOE-2,
      Santa Cruz, CA 95064, USA
  - code: marc
    address: |
      Center for Stock Assessment Research, Department of Applied Math
      and Statistics, University of California,
      Mail Stop SOE-2, Santa Cruz, CA 95064, USA and Department of
      Biology, University of Bergen, Bergen, Norway 9020
  - code: noaa
    address: | 
      Southwest Fisheries Science Center, 
      National Oceanic and Atmospheric Administration, 
      110 Shaffer Road, Santa Cruz, CA 95060, USA

abstract: |
  Model uncertainty and limited data are fundamental challenges to
  robust management of human intervention in a natural system.
  These challenges are acutely highlighted by concerns that
  many ecological systems may contain tipping points, such as
  Allee population sizes.  Before a collapse, we do not know
  where the tipping points lie, if they exist at all.  Hence,
  we know neither a complete model of the system dynamics nor
  do we have access to data in some large region of state-space
  where such a tipping point might exist.  We illustrate how
  a Bayesian Non-Parametric (BNP) approach using a Gaussian
  Process (GP) prior provides a flexible representation of this
  inherent uncertainty. We embed GPs in a Stochastic Dynamic
  Programming (SDP) framework in order to make robust management
  predictions with both model uncertainty and limited data. We
  use simulations to evaluate this approach as compared with the
  standard approach of using model selection to choose from a set
  of candidate models.  We find that model selection erroneously
  favors models without tipping points -- leading to harvest
  policies that guarantee extinction. The GPDP performs nearly as
  well as the true model and significantly outperforms standard
  approaches. We illustrate this using examples of simulated
  single-species dynamics, where the standard model selection
  approach should be most effective, and find that it still
  fails to account for uncertainty appropriately and leads to
  population crashes, while management based on the GPDP does
  not, since it does not underestimate the uncertainty outside
  of the observed data.

keywords:
  - Bayesian
  - Structural Uncertainty
  - Nonparametric 
  - Optimal Control
  - Decision Theory
  - Gaussian Processes
  - Fisheries Management

bibliography: components/references.bib
csl: components/ecology.csl
documentclass: components/elsarticle

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    template: components/elsarticle.latex
    keep_tex: true

---

```{r knit, include=FALSE}
library("methods")
library("rmarkdown")
render("manuscript.Rmd")
```

\appendix
\renewcommand*{\thefigure}{S\arabic{figure}}
\renewcommand*{\thetable}{S\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}

Appendix: Model definitions and estimation
===========================================

```{r figure_S1, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options"}
gp_assessment_plots$traces_plot
```

```{r figure_S2, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
gp_assessment_plots$posteriors_plot
```

\newpage

Training data (Figures 1-6)
----------------------------

Each of our models $f(S_t)$ must be estimated from training data, which
we simulate from the Allen model with parameters $r = $ `r p[1]`, 
$K =$ `r p[2]`, $C =$ `r p[3]`, and  $\sigma_g =$ `r sigma_g` 
for $T=$ `r Tobs` timesteps, starting at initial condition $X_0 = $ `r Xo`. 
The training data can be seen in Figure 1.

Training data for sensitivity analyses
-------------------------------------

A further 96 unique training data sets are generated for the sensitivity analysis, as described in the main text.

Code
----

All code used in producing this analysis can be found in the R package accompanying the paper, available from [http://github.com/cboettig/nonparametric-bayes]. The code for all results shown here is dynamically embedded into the manuscript using `knitr` [@knitr].

<!--
Data
----

While the data can be regenerated using the code provided, for convenience CSV files of the data shown in each graph are also provided, along with appropriate metadata written in the Ecological Metadata Language (EML).

-->

```{r figure_S3, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
plot_ricker_traces
```

```{r figure_S4, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
ggplot(ricker_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(ricker_posteriors, "components/data/ricker_posteriors.csv")
```

```{r Table S1, results = "asis"}
pander::pandoc.table(ricker_priors_xtable,
  caption = "Parameterization range for the uniform priors in the Ricker model")
```

```{r figure_S5, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
plot_myers_traces
```

```{r figure_S6, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
ggplot(myers_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(myers_posteriors, "components/data/myers_posteriors.csv")
```

```{r TableS2, results="asis"}
pander::pandoc.table(myers_priors_xtable,
           caption = "Parameterization range for the uniform priors in the Myers model")
```

```{r figure_S7, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
plot_allen_traces
```

```{r figure_S8, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
ggplot(allen_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(allen_posteriors, "components/data/allen_posteriors.csv")
```

```{r TableS3, results = "asis"}
pander::pandoc.table(allen_priors_xtable,
  caption = "Parameterization range for the uniform priors in the Allen model")
```

```{r sensitivity-calc}
source("components/sensitivity.R")

models <- c("Myers","Allen")

parameters <- list(
  Myers = list(
    c(r=1.5 + rnorm(1, 0, 1), theta=2 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1)),
    c(r=1.5 + rnorm(1, 0, 1), theta=2.5 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1))),
  Allen = list(
    c(r=2 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1), C=4 + rnorm(1, 0, 1)),
    c(r=2 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1), C=4 + rnorm(1, 0, 1)))
)
nuisance_pars <- c("sigma_g")
nuisance_values <- list(sigma_g = c(0.01, 0.05))
```

```{r allensets}
model <- "Allen"
allen1.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1111, 2222, 3333))
allen2.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1111, 2222, 3333))
allen1.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1111, 2222, 3333))
allen2.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1111, 2222, 3333))
```

```{r myerssets}
model <- "Myers"
Myers1.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1111, 2222, 3333))
Myers2.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1111, 2222, 3333))
Myers1.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1111, 2222, 3333))
Myers2.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1111, 2222, 3333))

## Assemble into data.frame
allen_dat <- rbind(allen1.01, allen1.05, allen2.01, allen2.05) 
m <- rbind(Myers1.01, Myers1.05, Myers2.01, Myers2.05)
myers_dat <- m[c(1:2,4,3,5:8)]
names(myers_dat) <- names(allen_dat)
model_dat <- rbind(allen_dat, myers_dat)
dat <- model_dat
dat$pars.r <- factor(dat$pars.r, labels=c("A", "B", "C", "D"))
dat <- dat[c(1:2,5:6, 8, 7)]
dat$noise <- factor(dat$noise)
names(dat) <- c("model", "parameters", "replicate", "simulation", "noise", "value")

## Extract acutal parameter values corresponding to each parameter set
p1 = as.numeric(levels(factor(model_dat$pars.r)))
p2 = as.numeric(levels(factor(model_dat$pars.K)))
p3 = as.numeric(levels(factor(model_dat$pars.C)))

set.A = c(r = p1[1], K = p2[1], theta = p3[1])
set.B = c(r = p1[2], K = p2[2], theta = p3[2])
set.C = c(r = p1[3], K = p2[3], C = p3[3])
set.D = c(r = p1[4], K = p2[4], C = p3[4])
AllenParameterSets <- rbind(set.A, set.B)
MyersParameterSets <- rbind(set.C, set.D)

sensitivity_dat <- dat
```

```{r figure_S9, fig.height=6, fig.width=10, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.  Histograms shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters. Values of 1 indicate optimal performance. Columns indicate different models, rows different noise levels, and colors indicate the parameter set used. Grouped over stochastic replicates applying the contol policy and stochastic replicates of training data generated from the model indicated, see raw data for details. Randomly chosen parameter values for the models shown in tables below."}
ggplot(sensitivity_dat) + 
  geom_histogram(aes(value, fill=parameters)) + 
  xlim(0,1.0) + 
  theme_bw() + 
  xlab("value as fraction of the optimal") + 
  facet_grid(noise~model)
write.csv(sensitivity_dat, "components/data/sensitivity_dat.csv") 
```

```{r AllenSetsTable, results="asis"}
pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Allen models in Figure S9." )
```
```{r MyersSetsTable, results="asis"}
pandoc.table(MyersParameterSets, caption="Randomly chosen parameter sets for the Myers models in Figure S9." )
```

```{r unlink, include=FALSE}
unlink("ricker_process.bugs")
unlink("allen_process.bugs")
unlink("myers_process.bugs")
```

### 2014-07-09 manuscripts/supplement.Rmd
fix supplement title, drop abstract

@@
 linenumbers: true
-title: "Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming"
+title: "Supplement for: Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming"
 author: 
@@
       110 Shaffer Road, Santa Cruz, CA 95060, USA
-
-abstract: |
-  Model uncertainty and limited data are fundamental challenges to
-  robust management of human intervention in a natural system.
-  These challenges are acutely highlighted by concerns that
-  many ecological systems may contain tipping points, such as
-  Allee population sizes.  Before a collapse, we do not know
-  where the tipping points lie, if they exist at all.  Hence,
-  we know neither a complete model of the system dynamics nor
-  do we have access to data in some large region of state-space
-  where such a tipping point might exist.  We illustrate how
-  a Bayesian Non-Parametric (BNP) approach using a Gaussian
-  Process (GP) prior provides a flexible representation of this
-  inherent uncertainty. We embed GPs in a Stochastic Dynamic
-  Programming (SDP) framework in order to make robust management
-  predictions with both model uncertainty and limited data. We
-  use simulations to evaluate this approach as compared with the
-  standard approach of using model selection to choose from a set
-  of candidate models.  We find that model selection erroneously
-  favors models without tipping points -- leading to harvest
-  policies that guarantee extinction. The GPDP performs nearly as
-  well as the true model and significantly outperforms standard
-  approaches. We illustrate this using examples of simulated
-  single-species dynamics, where the standard model selection
-  approach should be most effective, and find that it still
-  fails to account for uncertainty appropriately and leads to
-  population crashes, while management based on the GPDP does
-  not, since it does not underestimate the uncertainty outside
-  of the observed data.
-
-keywords:
-  - Bayesian
-  - Structural Uncertainty
-  - Nonparametric 
-  - Optimal Control
-  - Decision Theory
-  - Gaussian Processes
-  - Fisheries Management
 

### 2014-07-10 manuscripts/manuscript.Rmd
tweaking figures to work without transparency

@@
     fig_caption: true
+    fig_crop: false
     template: components/elsarticle.latex
@@
 f <- RickerAllee
-p <- c(2, 8, 5)
+p <- c(2, 10, 5)
 K <- 10  # approx, a li'l' less
@@
 Curves show the posterior step-ahead predictions based on each of the estimated models.
-Observe that all models are fitting the data reasonably well.", fig.width=8, fig.height=6}
+Observe that all models are fitting the data reasonably well.", fig.width=6, fig.height=4, dev.args=list(version="1.3")}
 step_ahead_posteriors <- function(x){
@@
 df_post <- step_ahead_posteriors(x)
-figure1b_posteriors <- ggplot(df_post) + geom_point(aes(time, stock)) + 
-  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=0.2) + 
+
+ggplot(df_post) + 
+#  geom_point(aes(time, stock)) +
+  stat_summary(aes(time, value), fun.y = "mean", fun.ymin = "min", fun.ymax="max", geom="ribbon", fill='grey80', col='grey80') +
+  stat_summary(aes(time, value, col=variable), fun.y = "mean", geom="line") +
+#  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=0.2) + 
   facet_wrap(~variable) + 
   scale_colour_manual(values=colorkey) +  
+  scale_fill_manual(values=colorkey) +  
   theme(legend.position="none")
-figure1b_posteriors
 
@@
 The training data are also shown as black points.
-The GP is conditioned on (0,0), shown as a pseudo-data point.", fig.width=8, fig.height=6}
+The GP is conditioned on (0,0), shown as a pseudo-data point.", fig.width=6, fig.height=4, dev.args=list(version="1.3")}
 x_grid_short <- x_grid[1:40]
@@
                   data=subset(models_posteriors, variable == "GP")) +
-    geom_line(aes(stock, value, col = variable, 
-                  group=interaction(L1,variable)), alpha=0.2) + 
+
+    stat_summary(aes(stock, value), fun.y = "mean", fun.ymin = "min", fun.ymax="max", geom="ribbon", fill='grey80', col='grey80') +
+    stat_summary(aes(stock, value, col=variable), fun.y = "mean", geom="line") +
+
+#   geom_line(aes(stock, value, col = variable, 
+#                  group=interaction(L1,variable)), alpha=0.2) +
+
     geom_point(data = obs, aes(x,y)) + 
@@
 
-```{r figure_3, fig.width=8, fig.height=6, fig.cap="Outside the range of the training data (Figure 1), the true dynamics (black dots) fall outside the uncertainty (two standard deviations, colored bands) of the structurally incorrect parametric models (Ricker, Myers), but inside the uncertainty predicted by the GP.
+```{r figure_3, fig.width=6, fig.height=4, fig.cap="Outside the range of the training data (Figure 1), the true dynamics (black dots) fall outside the uncertainty (two standard deviations, colored bands) of the structurally incorrect parametric models (Ricker, Myers), but inside the uncertainty predicted by the GP.
 Points show the stock size simulated by the true model.
@@
 The Ricker model always (wrongly) predicts positive population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).
-Note that because it does not assume a parametric form but instead relies more directly on the data, the GP is both more pessimistic and more uncertain about the future state than the parametric models.", dependson="plot-options"}
+Note that because it does not assume a parametric form but instead relies more directly on the data, the GP is both more pessimistic and more uncertain about the future state than the parametric models.", dependson="plot-options", dev.args=list(version="1.3")}
 y <- numeric(8)
@@
 ```{r figure_4, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
+Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", dev.args=list(version="1.3")}
 policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
@@
 ```{r figure_5, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models.
-We show 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
+We show 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=4, dev.args=list(version="1.3")}
+
+mymin <- function(x) mean(x) - sd(x)
 ggplot(sims_data) +
-  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=0.1) +
-  scale_colour_manual(values=colorkey) + 
+
+  stat_summary(aes(time, fishstock), fun.y = "mean", fun.ymin = "mymin", fun.ymax="max", geom="ribbon", fill='grey80', col='grey80') +
+  stat_summary(aes(time, fishstock, col=method), fun.y = "mean", geom="line") +
+
+#  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=0.1) +
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) + 
   facet_wrap(~method) + guides(legend.position="none")
@@
 Values plotted are the averages of a given policy over 100 replicate simulations.
-Details and code provided in the supplement.", fig.width=6, fig.height=6}
+Details and code provided in the supplement.", fig.width=5, fig.height=4, dev.args=list(version="1.3")}
 
@@
 
-```{r figure_7, fig.width=6, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).
-Other than the focal parameter (stochasticity, Allee threshold), other parameters are held fixed as above to illustrate this effect."}
+```{r figure_7, fig.width=4, fig.height=3, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).
+Other than the focal parameter (stochasticity, Allee threshold), other parameters are held fixed as above to illustrate this effect.", dev.args=list(version="1.3")}
 

### 2014-07-10 manuscripts/supplement.Rmd
tweaking figures to work without transparency

@@
 render("manuscript.Rmd")
+```
+
+```{r supplement-caching, include=FALSE}
+library("knitr")
+basename <- "supplement"
+opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
+               cache.path = paste("components/cache/", basename, "/", sep=""))
+opts_chunk$set(cache = 2) 
+opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
+               comment = NA, verbose = TRUE, echo=FALSE)
+
+opts_chunk$set(dev.opts=c(version="1.3"), dev="pdf")
+
 ```

### 2014-07-11 manuscripts/manuscript.Rmd
minor figure tweaks

--- 2014-07-08 manuscripts/manuscript.Rmd
@@
   geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=0.1) +
-  scale_colour_manual(values=colorkey) + 
+  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) + 
   facet_wrap(~method) + guides(legend.position="none")
@@
 
-```{r figure_7, fig.width=6, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).
+```{r figure_7, fig.width=6, fig.height=3, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).
 Other than the focal parameter (stochasticity, Allee threshold), other parameters are held fixed as above to illustrate this effect."}
-
 plot_sigmas <- ggplot(vary_sigma, aes(noise, value)) + 
@@
   xlab("Allee threshold stock size") + theme_bw()
-
 multiplot(plot_sigmas, plot_allees, cols=2) 
-
 write.csv(vary_sigma, "components/data/vary_sigma.csv")

### 2014-07-11 manuscripts/manuscript.tex
tex copy

@@
-\documentclass[]{components/elsarticle}
+\documentclass[author-year, 12pt,review]{components/elsarticle} %review=doublespace preprint=single 5p=2 column
+%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
+\usepackage[hyphens]{url}
+\usepackage{lineno} % add 
+  \linenumbers % turns line numbering on 
+\bibliographystyle{elsarticle-harv}
+\biboptions{sort&compress} % For natbib
+\usepackage{graphicx}
+\usepackage{booktabs} % book-quality tables
+%% Redefines the elsarticle footer
+\makeatletter
+\def\ps@pprintTitle{%
+ \let\@oddhead\@empty
+ \let\@evenhead\@empty
+ \def\@oddfoot{\it \hfill\today}%
+ \let\@evenfoot\@oddfoot}
+\makeatother
+
+% A modified page layout
+\textwidth 6.75in
+\oddsidemargin -0.15in
+\evensidemargin -0.15in
+\textheight 9in
+\topmargin -0.5in
+%%%%%%%%%%%%%%%% end my additions to header
+
 \usepackage[T1]{fontenc}
@@
 \else % if luatex or xelatex
+  \usepackage{fontspec}
   \ifxetex
-    \usepackage{mathspec}
     \usepackage{xltxtra,xunicode}
-  \else
-    \usepackage{fontspec}
   \fi
@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\usepackage[margin=1in]{geometry}
-\usepackage{longtable,booktabs}
+\usepackage{longtable}
 \usepackage{graphicx}
-% Redefine \includegraphics so that, unless explicit options are
-% given, the image width will not exceed the width of the page.
-% Images get their normal width if they fit onto the page, but
+% We will generate all images so they have a width \maxwidth. This means
+% that they will get their normal width if they fit onto the page, but
 % are scaled down if they would overflow the margins.
 \makeatletter
-\def\ScaleIfNeeded{%
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
+\else\Gin@nat@width\fi}
 \makeatother
 \let\Oldincludegraphics\includegraphics
-{%
- \catcode`\@=11\relax%
- \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
-}%
+\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
 \ifxetex
@@
             pdfauthor={},
-            pdftitle={Avoiding tipping points in fisheries management: A Bayesian Non-Parametric approach},
+            pdftitle={Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming},
             colorlinks=true,
-            citecolor=blue,
             urlcolor=blue,
@@
 \setcounter{secnumdepth}{0}
-
-\title{Avoiding tipping points in fisheries management: A Bayesian
-Non-Parametric approach}
-\author{ \and  \and }
-\date{}
+% Pandoc toggle for numbering sections (defaults to be off)
+\setcounter{secnumdepth}{0}
+% Pandoc header
 
 \begin{document}
-
-\begin{center}
-\huge Avoiding tipping points in fisheries management: A Bayesian
-Non-Parametric approach \\[0.2cm]
-\normalsize
-\end{center}
+\begin{frontmatter}
+
+  \title{Avoiding tipping points in fisheries management through Gaussian Process
+Dynamic Programming}
+    \author[cstar]{Carl Boettiger\corref{c1}}
+   \ead{cboettig(at)gmail.com} 
+   \cortext[c1]{Corresponding author}
+    \author[cstar]{Marc Mangel}
+
+    \author[noaa]{Stephan Munch}
+
+      \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and
+Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA
+95064, USA}    
+    \address[marc]{Center for Stock Assessment Research, Department of Applied Math and
+Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA
+95064, USA and Department of Biology, University of Bergen, Bergen,
+Norway 9020}    
+    \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric
+Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}
 
   \begin{abstract}
-Model uncertainty and limited data coverage are fundamental challenges
-to robust management of human intervention in a natural system. These
-challenges are acutely highlighted by concerns that many ecological
-systems may contain tipping points, such as Allee population
-sizes.\\Before a collapse, we do not know where the tipping points lie,
-if they exist at all. Hence, we know neither a complete model of the
-system dynamics nor do we have access to data in some large region of
-state-space where such a tipping point might exist. We illustrate how a
-Bayesian Non-Parametric (BNP) approach using a Gaussian Process (GP)
-prior provides a flexible representation of this inherent uncertainty.
-We embed GPs in a stochastic dynamic programming framework in order to
-make robust management predictions with both model uncertainty and
-limited data. We conduct a set of simulations to evaluate this approach
-as compared with the standard approach of using model selection to
-choose from a set of candidate models. We find that model selection
-erroneously favors models without tipping points -- leading to harvest
-policies that guarantee extinction. The GP approach performs nearly as
-well as the true model and significantly outperforms standard
-appraoches. We illustrate this using examples of simulated
-single-species dynamics, where the standard model selection approach
-should be most effective, and find that it still fails to account for
-uncertainty appropriately and leads to population crashes, while
-management based on the GP approach does not, since it does not
+  Model uncertainty and limited data are fundamental challenges to robust
+  management of human intervention in a natural system. These challenges
+  are acutely highlighted by concerns that many ecological systems may
+  contain tipping points, such as Allee population sizes. Before a
+  collapse, we do not know where the tipping points lie, if they exist at
+  all. Hence, we know neither a complete model of the system dynamics nor
+  do we have access to data in some large region of state-space where such
+  a tipping point might exist. We illustrate how a Bayesian Non-Parametric
+  (BNP) approach using a Gaussian Process (GP) prior provides a flexible
+  representation of this inherent uncertainty. We embed GPs in a
+  Stochastic Dynamic Programming (SDP) framework in order to make robust
+  management predictions with both model uncertainty and limited data. We
+  use simulations to evaluate this approach as compared with the standard
+  approach of using model selection to choose from a set of candidate
+  models. We find that model selection erroneously favors models without
+  tipping points -- leading to harvest policies that guarantee extinction.
+  The GPDP performs nearly as well as the true model and significantly
+  outperforms standard approaches. We illustrate this using examples of
+  simulated single-species dynamics, where the standard model selection
+  approach should be most effective, and find that it still fails to
+  account for uncertainty appropriately and leads to population crashes,
+  while management based on the GPDP does not, since it does not
   underestimate the uncertainty outside of the observed data.
   \end{abstract}
+   \begin{keyword} Bayesian \sep Structural Uncertainty \sep Nonparametric \sep Optimal Control \sep Decision Theory \sep Gaussian Processes \sep Fisheries Management \sep \end{keyword}
+ \end{frontmatter}
 
@@
 the last four decades (but stretching even further back in time), marked
-by the seminal contributions of Clark (1976); Clark (2009) and Walters
+by the seminal contributions of Clark (1976), Clark (2009) and Walters
 and Hilborn (1978), dynamic optimization methods, particularly
-Stochastic Dynamic Programming (SDP, Clark (2009)), have become
-increasingly important as a means of understanding how to management
-human intervention into natural systems. Concomittantly, there has been
-increasing recognition of the importance of Allee population sizes or
-`tipping points' (Scheffer et al. 2001, Polasky et al. 2011) that
-management actions may inadvertently cause populations to cross.
-Scheffer et al. (2009) emphasized the difficulties of formulating even
-qualitatively correct models of the underlying processes.
+Stochastic Dynamic Programming (SDP), have become increasingly important
+as a means of understanding how to manage human intervention into
+natural systems. Simultaneously, there has been increasing recognition
+of the importance of Allee population sizes or `tipping points'
+(Scheffer et al. 2001, Polasky et al. 2011) that management actions may
+inadvertently cause populations to cross. Scheffer et al. (2009)
+emphasized the difficulties of formulating even qualitatively correct
+models of the underlying processes.
 
@@
 controversy surrounding their interpretation (Hilborn 2007, Worm et al.
-2009) have made understanding these challenges all the more pressing.
+2009) make understanding these challenges all the more pressing.
 
@@
 been crossed, management is most often concerned with avoiding
-potentially catastrophic tipping points before any data is available at
+potentially catastrophic tipping points before any data are available at
 or following a transition that would more clearly reveal these regime
@@
 
-We illustrate how a stochastic dynamic programming (SDP) algorithm
+We illustrate how a Stochastic Dynamic Programming (SDP) algorithm
 (Mangel and Clark 1988, Marescot et al. 2013) can be driven by the
@@
 dynamics. In contrast to parametric models which can only reflect
-uncertainty in parameter estimates, the BNP approach provides a more
+uncertainty in parameter estimates, the BNP approach provides a
 state-space dependent representation of uncertainty. This permits a much
@@
 one-dimensional models of population dynamics as a `best-case' scenario
-for the parametric model performance relative to the GP. This is
-abest-case scenario for the parametric models because the underlying
-dynamics are always simuated from one the parametric models in the set,
-whereas in reality we never have the ``true'' model among our
+for the parametric model performance relative to the GP. This is a
+best-case scenario for the parametric models because the underlying
+dynamics are always simulated from one of the parametric models in the
+set, whereas in reality we never have the ``true'' model among our
 candidates. By choosing one-dimensional models with few parameters, we
@@
 population at time $t$ and assume that in the absence of take its
-dynamics are
-
-\[ X_{t+1}= Z_t f(X_t, h_t, p) \]
+dynamics are.
+
+\begin{equation}
+X_{t+1}= Z_t f(X_t) \label{eq1} 
+\end{equation}
 
@@
 maximizes total discounted harvest between the current time $t$ and a
-final time $T$, that is we seek to maximize over choices of harvest
+final time $T$. That is, we seek to maximize over choices of harvest
 $\mathbf{E}_{X_t+1} [ \sum_{t = 0}^{T}  R_t(X_t, h_t) \delta^t]$, where
@@
 
-In order to find that policy, we introduce the value function$V_t(x_t)$
+In order to find that policy, we introduce the value function $V_t(x_t)$
 representing the total discounted catch from time $t$ onwards given that
-$X_t=x$. This value function satisfies an equation of Stochastic Dynamic
-Programming (SDP, Mangel and Clark 1988, Clark and Mangel 2000, Clark
-2009), sometimes called the Bellman-Hamilton-Jacobi equation (Mangel
-2014),
+$X_t=x_t$. This value function satisfies an equation of SDP (Mangel and
+Clark 1988, Clark and Mangel 2000, Clark 2009), sometimes called the
+Hamilton-Jacobi-Bellman equation (Mangel 2014),
 
 \begin{equation}
-V_t(X_t) = \max_{h_t} \lbrace R(h_t, X_t) + \delta \cdot \mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] \rbrace
+V_t(x_t) = \max_{h_t} \lbrace R(h_t, x_t) + \delta \cdot \mathbf{\mathrm{E}}_{X_{t+1}} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] \rbrace
 \end{equation}
@@
 where expectation is taken over all possible values of the next state,
-$X_t{+1}$, and maximized over all possible choices of harvest, $h_t$.
-That is, at time $t$, when population size is $x$ and harvest $h_t$ is
-applied, the immediate return is $R(h_t, X_t)$. In the classical case,
-the sole source of uncertainty is the process noise term, $Z$, and thus
-the expectation above is equivalent to taking expectations over $Z$.
-That is
-
-\[ \mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{Z} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \]
+$X_{t+1}$, and maximized over all possible choices of harvest, $h_t$.
+That is, at time $t$, when population size is $x_t$ and harvest $h_t$ is
+applied, the immediate return is $R(h_t, x_t)$. In the classical case,
+the sole source of uncertainty is the process stochasticity term, $Z$,
+and thus the expectation above is equivalent to taking expectations over
+$Z$. That is
+
+\[ \mathbf{\mathrm{E}}_{X_{t+1}} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] = \mathbf{\mathrm{E}}_{Z} \left[ V_{t+1}( Z f(x_t - h_t))  | x_t, h_t \right] \]
 
@@
 In a Bayesian decision framework, the parameters governing the dynamics
-are also uncertain, so the expectation in (1) involves averaging over
-the posterior distribution for the parameters, as well. That is,
-
-\[\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta,\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \}\]
+are also uncertain, so the expectation involves averaging over the
+posterior distribution for the parameters, as well. That is,
+
+\[\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(x_t - h_t))  | x_t, h_t \right] \}\]
 
@@
 
-\[\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta,\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \}\]
-
-We consider the finite time problem with \$T=\$1000, which we solve
-using value iteration algorithm (Mangel and Clark 1988, Clark and Mangel
-2000).
+\[\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | x_t, h_t \right] \}\]
+
+We consider the finite time problem with $T=$ 1000, which we solve using
+the standard value iteration algorithm (see Mangel and Clark 1988, Clark
+and Mangel 2000).
 
@@
 dynamics: The Ricker model, the Allen model (Allen and Tanner 2005), the
-Myers model (Myers et al. 1995), equations (1)-(3), which we
-parameterize as follows for $f(S_t,p)$ in Eqn 1. In all three, we let
-$K$ denote the carrying capacity and $r$ the maximum per capita growth
-rate. In the Ricker model,
-
-\begin{enumerate}
-\def\labelenumi{(\arabic{enumi})}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \[ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
-\end{enumerate}
+Myers model (Myers et al. 1995), Eqns \eqref{ricker}-\eqref{myers},
+which we parameterize as follows for $f(S_t,p)$ in Eqn 1. In all three,
+we let $K$ denote the carrying capacity and $r$ the maximum per capita
+growth rate. In the Ricker model,
+
+\begin{equation}
+f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K} \right) } \label{ricker}
+\end{equation}
 
@@
 
-\begin{enumerate}
-\def\labelenumi{(\arabic{enumi})}
-\setcounter{enumi}{1}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \[ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_c\right)} \]
-\end{enumerate}
+\begin{equation}
+f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_c\right)} \label{allen}
+\end{equation}
 
@@
 
-\begin{enumerate}
-\def\labelenumi{(\arabic{enumi})}
-\setcounter{enumi}{2}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \[ f(S(t) | p)  = \frac{r S_t^{\theta}}{1 + \frac{S_t^\theta}{K}}\]
-\end{enumerate}
+\begin{equation}
+f(S(t) | p)  = \frac{r S_t^{\theta}}{1 + \frac{S_t^\theta}{K}} \label{myers}
+\end{equation}
 
@@
 estimates do not depend on the particular realization, while longer
-times are not likely to provide substantial improvement (i.e.~the
-results are not sensitive to this interval). Each of the models
-(described below) is fit to the same training data (Figure 1).
+times are not likely to provide substantial improvement.\\Each of the
+models is fit to the same training data (Figure 1).
 
@@
 Gibbs sampling (Gelman et al. (2003) implemented in R (R Core Team 2013)
-using jags, (Su and Masanao Yajima 2013)). We choose uniform priors for
-all parameters (See appendix Tables S1-S3; R code provided). We show
-one-step-ahead predictions of these model fits in Figure 1. We tested
-each chain for Gelman-Rubin convergence and results were robust to
-longer runs. For each simulation we also applied several commonly used
-model selection criteria (AIC, BIC, DIC, see Burnham and Anderson
+using \texttt{jags}, (Su and Masanao Yajima 2013)). We choose uniform
+priors for all parameters (See appendix Tables S1-S3; R code provided).
+We show one-step-ahead predictions of these model fits in Figure 1. We
+tested each chain for Gelman-Rubin convergence and results were robust
+to longer runs. For each simulation we also applied several commonly
+used model selection criteria (AIC, BIC, DIC, see Burnham and Anderson
 (2002)) to identify the best fitting model.
@@
 The core difference for our purpose between the GP representation and
-the parametric models we have already introduced is the
-``non-parametric'' property, which means that an estimated GP model is
-defined explicitly in reference to some observed data. Unlike the models
-above, it cannot be specified by the value of some parameters alone. The
-nonparametric moniker for the GP approach is unfortunate as (a) it still
-involves the estimation of parameters, and (b), statisticians also use
-non-parametric to mean models that make only weak distributional
-assumptions (e.g. Lehmann 1975). However, we retain the terminology for
-consistency with previous studies.
-
-The use GP methods to formulate a predictive model is relatively new in
-the context of modeling dynamical systems (Kocijan et al. 2005), and was
-first introduced in the context ecological modeling and fisheries
+the parametric models is that the GP model is defined explicitly in
+reference to some observed data. Unlike the models above, it cannot be
+specified by the value of some parameters alone. Although called
+nonparametric, the GP (a) still involves the estimation of parameters,
+and (b), statisticians also use non-parametric to mean models that make
+only weak distributional assumptions (e.g. Lehmann 1975). However, we
+retain the terminology for consistency with previous studies.
+
+The use of GP methods to formulate a predictive model is relatively new
+in the context of modeling dynamical systems (Kocijan et al. 2005), and
+was first introduced in the context ecological modeling and fisheries
 management in Munch et al. (2005b). GP models have subsequently been
@@
 A GP is an infinite dimensional (because it is a function of a
-continuous input variable) stochastic process for which any realization
-is a multivariate normal distribution. Thus, to characterize the GP we
-need a mean function and a covariance matrix. We proceed as follows.
-
-Assuming the data $X_o$ are observed with some process noise,
-
-\[X_{t+1} = f(X_t) + \varepsilon\]
-
-$\varepsilon$ IID normal, variance $\sigma_n^2$. Note that we have
-chosen to assume additive noise. While we could just as easily consider
-log-normal noise as in the parametric models, we make this choice to
-bring home the point that the Gaussian process approach need not have
+continuous input variable) stochastic process for a finite sample of
+points from any realization is a multivariate normal distribution. Thus,
+to characterize the GP we need a mean function and a covariance
+function. We proceed as follows.
+
+We assume that the data $X_o$ are observed with some process
+stochasticity,
+
+\[X_{t+1} = f(X_t) + \varepsilon,\]
+
+where $\varepsilon$ are IID normal random variables with zero-mean and
+variance $\sigma_n^2$. Note that we have chosen to assume additive
+stochasticity. While we could just as easily consider log-normal
+stochasticity as in the parametric models, we make this choice to bring
+home the point that the Gaussian process approach need not have
 structurally correct noise form either.
@@
 
-We use vague inverse Gamma priors on both the lengthscale $\ell$ and
-process noise $\sigma$ parameters, of the form
+We use inverse Gamma priors on both the length-scale $\ell$ and
+$\sigma$, of the form
 
@@
 from their predicted values $X_{t+1}$ given an observation of the
-previous state, $X_t$. Munch et al. (2005a) gives an accessible
+previous state, $X_t$. Munch et al. (2005a) give an accessible
 introduction to the use of Gaussian processes in providing a Bayesian
@@
 (2003)) to infer posterior distributions of the two parameters of the GP
-(Figure S13, code in appendix), under weakly informative Gaussian priors
-(see parameters in table S5). As the posterior distributions differ
+(Figure S13, code in appendix). Since the posterior distributions differ
 substantially from the priors (Figure S13), we can be assured that most
@@
 
-\subsection{The method of Gaussian Process Dynamic
-Programming}\label{the-method-of-gaussian-process-dynamic-programming}
+\subsection{The method of Gaussian Process Dynamic Programming
+(GPDP)}\label{the-method-of-gaussian-process-dynamic-programming-gpdp}
 
@@
 
-To ensure our results are robust to our choice of parameters, we perform
-our analysis over 96 different scenarios. To better understand the
-process, we will first describe in detail the results of a single
-scenario.
-
+To ensure our results are robust to our choice of parameters, consider
+96 different scenarios, described in detail below. To help better
+understand the process, we first describe in detail the results of a
+single scenario.
+
+\begin{figure}[htbp]
+\centering
 \includegraphics{components/figure/manuscript-figure_1.pdf}
+\caption{Points show the training data of stock-size over time. Curves
+show the posterior step-ahead predictions based on each of the estimated
+models. Observe that all models are fitting the data reasonably well.}
+\end{figure}
 
@@
 small uncertainty, as illustrated in the posterior predictive curves in
-Figure 1. Figure 1 shows the training data of stock sizes observed over
-time as points, overlaid with the step-ahead predictions of each
+Figure 1, in which we show the training data of stock sizes observed
+over time as points, overlaid with the step-ahead predictions of each
 estimated model using the parameters sampled from their posterior
@@
 predicting fluctuations that are actually due purely to stochasticity.
-All model selection criteria shown in Table 1 penalize more complex
-models and show a preference for the simpler Ricker model over the more
-complicated alternate stable state models (Allen and Myers). Details on
-MCMC estimates for each model, traces, and posterior distributions can
-be found in the appendix.
-
-\% latex table generated in R 3.1.0 by xtable 1.7-3 package \% Tue Jun 3
-13:34:19 2014
-
-\begin{table}[ht]
+Model selection criteria (Table 1) penalize more complex models and show
+a preference for the simpler Ricker model over the models with
+alternative stable states (Allen and Myers). Details on MCMC estimates
+for each model, traces, and posterior distributions can be found in the
+appendix.
+
+\begin{longtable}[c]{@{}cccc@{}}
+\toprule\addlinespace
+\begin{minipage}[b]{0.12\columnwidth}\centering
+~
+\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\centering
+Allen
+\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\centering
+Ricker
+\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\centering
+Myers
+\end{minipage}
+\\\addlinespace
+\midrule\endhead
+\begin{minipage}[t]{0.12\columnwidth}\centering
+\textbf{DIC}
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
+50.75
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
+50.45
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
+50.41
+\end{minipage}
+\\\addlinespace
+\begin{minipage}[t]{0.12\columnwidth}\centering
+\textbf{AIC}
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
+-24.51
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
+-30.13
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
+-27.01
+\end{minipage}
+\\\addlinespace
+\begin{minipage}[t]{0.12\columnwidth}\centering
+\textbf{BIC}
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\centering
+-17.75
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
+-25.06
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\centering
+-20.25
+\end{minipage}
+\\\addlinespace
+\bottomrule
+\addlinespace
+\caption{Model selection scores for several common criteria all
+(wrongly) select the simplest model. As the true (Allen) model is not
+distinguishable from the simpler (Ricker) model in the region of the
+observed data, this error cannot be avoided regardless of the model
+choice criterion. This highlights the danger of model choice when the
+selected model will be used outside of the observed range of the data.}
+\end{longtable}
+
+\begin{figure}[htbp]
 \centering
-\begin{tabular}{rrrr}
-  \hline
- & Allen & Ricker & Myers \\ 
-  \hline
-DIC & 50.75 & 50.45 & 50.41 \\ 
-  AIC & -24.51 & -30.13 & -27.01 \\ 
-  BIC & -17.75 & -25.06 & -20.25 \\ 
-   \hline
-\end{tabular}
-\caption{Model selection scores for several common criteria all (wrongly) select the simplest model.
-As the true (Allen) model is not distinguishable from the simpler (Ricker) model in the region of the observed data, this error cannot be avoided regardless of the model choice criterion.
-This highlights the danger of model choice when the selected model will be used outside of the observed range of the data.}
-\end{table}
-
 \includegraphics{components/figure/manuscript-figure_2.pdf}
-
-The mean inferred state space dynamics of each model relative to the
-true model used to generate the data is shown in Figure 2, predicting
-the relationship between observed stock size (x-axis) to the stock size
+\caption{The inferred Gaussian process compared to the true process and
+maximum-likelihood estimated process. We show the expected value for the
+function $f$ under each model. Two standard deviations from the
+estimated Gaussian process covariance with (light grey) and without
+(darker grey) measurement error are also shown. The training data are
+also shown as black points. The GP is conditioned on (0,0), shown as a
+pseudo-data point.}
+\end{figure}
+
+We show the mean inferred state space dynamics of each model relative to
+the true model used to generate the data in Figure 2, predicting the
+relationship between observed stock size (x-axis) to the stock size
 after recruitment the following year. Note that in contrast to the other
@@
 has a mean shown in black. Parameter uncertainty (not shown) spreads out
-the estimates further. The observed data from which each model is
+the estimates further. The observed data from which each model are
 estimated is also shown. The observations come from only a limited
@@
 
+Figure 3 illustrates the performance of the GP and parametric models
+outside the observed training data. The mean trajectory under the
+underlying model is shown by the black dots, while the corresponding
+prediction made by the model shown by the box and whiskers plots.
+Predictions are based on the true expected value in the previous time
+step. Predicted distributions that lie entirely above the expected
+dynamics indicate the expectation of stock sizes higher than what is
+actually expected. The models differ both in their expectations and
+their uncertainty (colored bands show two standard deviations away).
+Note that the GP is particularly uncertain about the dynamics relative
+to structurally incorrect models like the Ricker.
+
+\begin{figure}[htbp]
+\centering
 \includegraphics{components/figure/manuscript-figure_3.pdf}
-
+\caption{Outside the range of the training data (Figure 1), the true
+dynamics (black dots) fall outside the uncertainty (two standard
+deviations, colored bands) of the structurally incorrect parametric
+models (Ricker, Myers), but inside the uncertainty predicted by the GP.
+Points show the stock size simulated by the true model. Overlay shows
+the range of states predicted by each model, based on the state observed
+in the previous time step. The Ricker model always (wrongly) predicts
+positive population growth, while the actual population shrinks in each
+step as the initial condition falls below the Allee threshold of the
+underlying model (Allen). Note that because it does not assume a
+parametric form but instead relies more directly on the data, the GP is
+both more pessimistic and more uncertain about the future state than the
+parametric models.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
 \includegraphics{components/figure/manuscript-figure_4.pdf}
+\caption{The steady-state optimal policy (infinite boundary) calculated
+under each model. Policies are shown in terms of target escapement,
+$S_t$, as under models such as this a constant escapement policy is
+expected to be optimal (Reed 1979).}
+\end{figure}
 
 Despite the similarities in model fits to the observed data, the
-policies inferred under each model differ widely, as shown in Figure 4.
-Policies are shown in terms of target escapement, $S_t$. Under models
-such as this a constant escapement policy is expected to be optimal
-(Reed 1979), whereby population levels below a certain size $S$ are
-unharvested, while above that size the harvest strategy aims to return
-the population to $S$, resulting in the hockey-stick shaped policies
-shown. Only the structurally correct model (Allen model) and the GP
-produce policies close to the true optimum policy (where both the
-underlying model structure and parameter values are known without
-error).
-
+policies inferred under each model differ widely (Figure 4). Policies
+are shown in terms of target escapement, $S_t$. Under models such as
+this a constant escapement policy is expected to be optimal (Reed 1979),
+whereby population levels below a certain size $S$ are unharvested,
+while above that size the harvest strategy aims to return the population
+to $S$, resulting in the hockey-stick shaped policies shown. Only the
+structurally correct model (Allen model) and the GP produce policies
+close to the true optimum policy (where both the underlying model
+structure and parameter values are known without error).
+
+\begin{figure}[htbp]
+\centering
 \includegraphics{components/figure/manuscript-figure_5.pdf}
-
-The consequences of managing 100 replicate realizations of the simulated
-fishery under each of the policies estimated is shown in Figure 5. As
-expected from the policy curves, the structurally correct model
-under-harvests, leaving the stock to vary around it's un-fished optimum.
-The structurally incorrect Ricker model over-harvests the population
-passed the tipping point consistently, resulting in the immediate crash
-of the stock and thus derives minimal long-term catch.
+\caption{In the management context, GPDP outperforms approaches based on
+parametric models. We show 100 replicate simulations of the stock
+dynamics (eq 1) under the policies derived from each of the estimated
+models, as well as the policy based on the exact underlying model.}
+\end{figure}
+
+In Figure 5, we show the consequences of managing 100 replicate
+realizations of the simulated fishery under each of the policies
+estimated. As anticipated from the policy curves, the structurally
+correct model under-harvests, leaving the stock to vary around its
+unfished optimum. The structurally incorrect Ricker model over-harvests
+the population passed the tipping point consistently, resulting in the
+immediate crash of the stock and thus derives minimal long-term catch.
 
@@
 compared by using the relative differences in net present value realized
-by each of the model. Figure 6 shows the distribution of values realized
-by each simulation shown in Figure 5. The GPDP most consistently
-realizes a value close to the optimal solution, and importantly avoids
-ever driving the system across the tipping point, which results in the
-near-zero value cases in the parametric models.
-
+by each of the model (Figure 6). The GPDP most consistently realizes a
+value close to the optimal solution, and importantly avoids ever driving
+the system across the tipping point, which results in the near-zero
+value cases in the parametric models.
+
+\begin{figure}[htbp]
+\centering
 \includegraphics{components/figure/manuscript-figure_6.pdf}
+\caption{Histograms of the realized net present value of the fishery
+over a range of simulated data and resulting parameter estimates. For
+each data set, the three models are estimated as described above. Values
+plotted are the averages of a given policy over 100 replicate
+simulations. Details and code provided in the supplement.}
+\end{figure}
 
@@
 simulation, noise scale, parameters or structure of the underlying
-model, as seen in Figure 7. The sensitivity analysis uses a factorial
-design of eight replicates of three different randomly generated
-parameter sets for each of two different generating models (Allen and
-Myers) over two different noise levels (0.01 and 0.05), for a total of 8
-x 3 x 2 x 2 = 96 scenarios. In Figure 7, we pool results across
-different random seeds (111, 222, 333), noise values of 0.01 and 0.05
-(not large enough to violate the self-sustaining criterion of Reed
-(1979)), and different randomly generated parameters sets for each
-model, and using either the Myers or Allen as the underlying structure.
-While the GPDP performs slightly closer to optimal on the smaller noise
-levels, and closer on the Allen model then the Myers model, the
-differences are small compared to the values observed in using the wrong
-parametric model. The GPDP solution consistently avoids either failing
-to harvest at all or crashing the stock, which results in the very low
-values seen in parametric models in the previous figure. The sensitivity
-analysis is compared against the optimal solution alone and not shown
-against the estimates of the various parametric models (but see Figure
-6), where convergence of the Gibbs sampling could contribute to their
-poorer performance if not checked by hand for each configuration.
-
-Changing the noise or the distance between stable and unstable points
-does not impact the performance of the GP relative to the optimal
-solution given the true model and true parameters. The parametric models
-are more senstive to this difference. Large noise relative to the
-distance between the stable and unstable point increases the chance of a
-stochastic transition. More precisely, if we let $L = K - x_c$, then the
-probability of a transition in a given window of time $T$ scales as
+model. We consider both a Latin hypercube approach and a more focused
+investigation of the parameters most responsible for impacting the
+transition rate: the relative distance to the Allee threshold and the
+noise effects.
+
+The Latin hypercube approach systematically varies all combinations of
+parameters, providing a more general test than varying only one
+parameter at a time. We loop across eight replicates of three different
+randomly generated parameter sets for each of two different generating
+models (Allen and Myers) over two different noise levels (0.01 and
+0.05), for a total of 8 x 3 x 2 x 2 = 96 scenarios. The Gaussian Process
+performs nearly optimally in each case, relative to the optimal solution
+with no parameter or model uncertainty (Figure S9, appendix).
+
+Changing the intensity of the stochasticity or the distance between
+stable and unstable points does not impact the performance of the GP
+relative to the optimal solution given the true model and true
+parameters. The parametric models are more sensitive to this difference.
+Large noise relative to the distance between the stable and unstable
+point increases the chance of a stochastic transition. More precisely,
+if we let $L = K - x_c$, then the probability of a transition in a given
+window of time $T$ scales as
 
@@
 
-(following the Ahrenius relationship, see Gardiner (2009) for the
-derivation). As a result, the impact of using a model which
-underestimates the risk of harvesting past the critical point is much
-more severe, since this such a situation occurs more often. Conversely,
-with large enough distance between the optimal escapement and unstable
-points relative to the noise, the chance of a transition becomes
-vanishingly small and all models can be modeled near-optimally. Models
-that underestimate the cost incurred by stock sizes fluctuating
-significantly below the optimal escapement level will not perform poorly
-as long as those fluctuations are sufficently rare. Figure 7 shows that
-the GPDP is little effected by increasing noise or increasing Allee
-effects over much of the range.
-
+(see Gardiner (2009) or Mangel (2006) for the derivation). As a result,
+the impact of using a model which underestimates the risk of harvesting
+past the critical point is much more severe, since this such a situation
+occurs more often. Conversely, with large enough distance between the
+optimal escapement and unstable points relative to the noise, the chance
+of a transition becomes vanishingly small and all models can be
+estimated near-optimally. Models that underestimate the cost incurred by
+stock sizes fluctuating significantly below the optimal escapement level
+will not perform poorly as long as those fluctuations are sufficiently
+rare.
+
+The GPDP is weakly influenced by increasing stochasticity or increasing
+Allee effects over much of the range (Figure 7) Much larger
+stochasticity or higher Allee effects make even the optimal solution
+without any model or parameter uncertainty unable to harvest the
+population effectively (e.g.~noise large enough to violate the
+self-sustaining criterion of Reed (1979)). Only at the largest Allee
+effect sizes are some replicates seen to deviate more from the optimal
+solution.
+
+\begin{figure}[htbp]
+\centering
 \includegraphics{components/figure/manuscript-figure_7.pdf}
+\caption{The effect of increasing noise or decreasing Allee threshold
+levels on the net present value of the fishery when managed under the
+GPDP, relative to managing under the true model (with known parameters).
+Other than the focal parameter (stochasticity, Allee threshold), other
+parameters are held fixed as above to illustrate this effect.}
+\end{figure}
 
@@
 point dynamics -- simpler models therefore fit just as well and tend to
-be selected more foten than complex ones. This approach would be
+be selected more often than complex ones. This approach would be
 appropriate when the dynamics can be expected to remain in the region of
@@
 
-\subsubsection{GPDP population dynamcis capture larger uncertainty in
+\subsubsection{GPDP population dynamics capture larger uncertainty in
 regions where the data are
-poor}\label{gpdp-population-dynamcis-capture-larger-uncertainty-in-regions-where-the-data-are-poor}
+poor}\label{gpdp-population-dynamics-capture-larger-uncertainty-in-regions-where-the-data-are-poor}
 
@@
 observed range, while retaining very good predictive accuracy in the
-range observed. The management policy dictated by the GP balance this
+range observed. The management policy dictated by the GP balances this
 uncertainty against the immediate value of the harvest, and act to
@@
 parameters $\ell$ and $\sigma$ and simple (mean zero, radial basis
-function kernel). In practice, both the choice of mean and the choice of
-the covariance function may be chosen to confer particular biological
+function kernel). In practice, both the choice of mean and the
+covariance function may be chosen to confer particular biological
 properties, as well as more biologically informed priors for $\ell$ and
@@
 
-\appendix
-\renewcommand*{\thefigure}{S\arabic{figure}}
-\renewcommand*{\thetable}{S\arabic{table}} \setcounter{figure}{0}\}
-\setcounter{table}{0}\}
-
-\section{Appendix: Model definitions and
-estimation}\label{appendix-model-definitions-and-estimation}
-
-\includegraphics{components/figure/manuscript-figure_S1.pdf}
-
-\includegraphics{components/figure/manuscript-figure_S2.pdf}
-
-\newpage
-
-\subsection{Training data (Figures
-1-6)}\label{training-data-figures-1-6}
-
-Each of our models $f(S_t)$ must be estimated from training data, which
-we simulate from the Allen model with parameters $r = $ 2, $K =$ 8,
-$C =$ 5, and $\sigma_g =$ 0.05 for $T=$ 40 timesteps, starting at
-initial condition $X_0 = $ 5.5. The training data can be seen in Figure
-1.
-
-\subsection{Training data for sensitivity
-analyses}\label{training-data-for-sensitivity-analyses}
-
-A further 96 unique training data sets are generated for the sensitivity
-analysis, as described in the main text.
-
-\subsection{Code}\label{code}
-
-All code used in producing this analysis can be found in the R package
-accompanying the paper, available from
-{[}\url{http://github.com/cboettig/nonparametric-bayes}{]}. The code for
-all results shown here is dynamically embedded into the manuscript using
-\texttt{knitr} (Xie 2012).
-
-\includegraphics{components/figure/manuscript-figure_S3.pdf}
-
-\includegraphics{components/figure/manuscript-figure_S4.pdf}
-
-\begin{longtable}[c]{@{}ccc@{}}
-\toprule\addlinespace
-\begin{minipage}[b]{0.15\columnwidth}\centering
-parameter
-\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
-lower.bound
-\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
-upper.bound
-\end{minipage}
-\\\addlinespace
-\midrule\endhead
-\begin{minipage}[t]{0.15\columnwidth}\centering
-$r$
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-0.01
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-20
-\end{minipage}
-\\\addlinespace
-\begin{minipage}[t]{0.15\columnwidth}\centering
-$K$
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-0.01
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-40
-\end{minipage}
-\\\addlinespace
-\begin{minipage}[t]{0.15\columnwidth}\centering
-$\sigma$
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-1e-06
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-100
-\end{minipage}
-\\\addlinespace
-\bottomrule
-\addlinespace
-\caption{Parameterization range for the uniform priors in the Ricker
-model}
-\end{longtable}
-
-\includegraphics{components/figure/manuscript-figure_S5.pdf}
-
-\includegraphics{components/figure/manuscript-figure_S6.pdf}
-
-\begin{longtable}[c]{@{}ccc@{}}
-\toprule\addlinespace
-\begin{minipage}[b]{0.15\columnwidth}\centering
-parameter
-\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
-lower.bound
-\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
-upper.bound
-\end{minipage}
-\\\addlinespace
-\midrule\endhead
-\begin{minipage}[t]{0.15\columnwidth}\centering
-$r$
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-1e-04
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-10
-\end{minipage}
-\\\addlinespace
-\begin{minipage}[t]{0.15\columnwidth}\centering
-$K$
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-1e-04
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-40
-\end{minipage}
-\\\addlinespace
-\begin{minipage}[t]{0.15\columnwidth}\centering
-$\theta$
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-1e-04
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-10
-\end{minipage}
-\\\addlinespace
-\begin{minipage}[t]{0.15\columnwidth}\centering
-$\sigma$
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-1e-06
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-100
-\end{minipage}
-\\\addlinespace
-\bottomrule
-\addlinespace
-\caption{Parameterization range for the uniform priors in the Myers
-model}
-\end{longtable}
-
-\includegraphics{components/figure/manuscript-figure_S7.pdf}
-
-\includegraphics{components/figure/manuscript-figure_S8.pdf}
-
-\begin{longtable}[c]{@{}ccc@{}}
-\toprule\addlinespace
-\begin{minipage}[b]{0.15\columnwidth}\centering
-parameter
-\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
-lower.bound
-\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
-upper.bound
-\end{minipage}
-\\\addlinespace
-\midrule\endhead
-\begin{minipage}[t]{0.15\columnwidth}\centering
-$r$
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-0.01
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-6
-\end{minipage}
-\\\addlinespace
-\begin{minipage}[t]{0.15\columnwidth}\centering
-$K$
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-0.01
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-20
-\end{minipage}
-\\\addlinespace
-\begin{minipage}[t]{0.15\columnwidth}\centering
-$X_C$
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-0.01
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-20
-\end{minipage}
-\\\addlinespace
-\begin{minipage}[t]{0.15\columnwidth}\centering
-$\sigma$
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-1e-06
-\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
-100
-\end{minipage}
-\\\addlinespace
-\bottomrule
-\addlinespace
-\caption{Parameterization range for the uniform priors in the Allen
-model}
-\end{longtable}
-
-\includegraphics{components/figure/manuscript-figure_S9.pdf}
-
 Allen, D., and K. Tanner. 2005. Infusing active learning into the
@@
 parameter estimates. Ecological Modelling 14:273--292.
+
+Mangel, M. 2006. The Theoretical Biologist's Toolbox: Quantitative
+Methods for Ecology and Evolutionary Biology. Cambridge University
+Press.
 
@@
 
-Xie, Y. 2012. knitr: A general-purpose package for dynamic report
-generation in R.
-
 \end{document}

### 2014-10-15 manuscripts/manuscript.Rmd
rerun

@@
 
-The dual concerns of model uncertainty and
-incomplete data coverage pose a substantial challenge to existing decision-theoretic
-approaches [@Brozovic2011].  Because intervention is often (but not always, see @Hughes2013) too late after a tipping point has been crossed, management
-is most often concerned with avoiding potentially catastrophic tipping
-points before any data are available at or following a transition that
-would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].
+The dual concerns of model uncertainty and incomplete data coverage
+pose a substantial challenge to existing decision-theoretic approaches
+[@Brozovic2011].  Because intervention is often (but not always, see
+@Hughes2013) too late after a tipping point has been crossed, management
+is most often concerned with avoiding potentially catastrophic
+tipping points before any data are available at or following a
+transition that would more clearly reveal these regime shift dynamics
+[e.g. @Bestelmeyer2012].
 
 We illustrate how a Stochastic Dynamic Programming (SDP) algorithm
-[@Mangel1988; @Marescot2013] can be driven by the
-predictions from a Bayesian Non-Parametric (BNP) model of population dynamics [@Munch2005a].
+[@Mangel1988; @Marescot2013] can be driven by the predictions from a
+Bayesian Non-Parametric (BNP) model of population dynamics [@Munch2005a].
 This provides two distinct advantages compared with contemporary
@@
 
-This paper represents the first ecological application of the SDP
-decision-making framework without an a priori model of the underlying
-dynamics.  In contrast to parametric models which can only reflect
-uncertainty in parameter estimates, the BNP approach provides a 
-state-space dependent representation of uncertainty.  This permits
-a much greater uncertainty far from the observed data than near the
-observed data.  These features allow the Gaussian Process Dynamic Programming (GPDP) approach to find robust
-management solutions in face of limited data and without knowledge of
-the correct model structure.
+This paper represents the first ecological application of the
+SDP decision-making framework without an a priori model of the
+underlying dynamics.  In contrast to parametric models which can only
+reflect uncertainty in parameter estimates, the BNP approach provides a
+state-space dependent representation of uncertainty.  This permits a much
+greater uncertainty far from the observed data than near the observed
+data.  These features allow the Gaussian Process Dynamic Programming
+(GPDP) approach to find robust management solutions in face of limited
+data and without knowledge of the correct model structure.
 
@@
 results of a single scenario.
-
-<!-- 
-### Figure 1: Fitted Models
-
-- All models fit the data quite well
-- Information criteria would pick the simple, incorrect model.
--->
 
@@
 No observations occur at the theoretical optimum harvest rate or near
-the tipping point.
-
-Figure 3 illustrates the performance of the GP and parametric models
-outside the observed training data. The mean trajectory under the
-underlying model is shown by the black dots, while the corresponding
-prediction made by the model shown by the box and whiskers plots.
-Predictions are based on the true expected value in the previous time
-step.  Predicted distributions that lie entirely above the expected
-dynamics indicate the expectation of stock sizes higher than what is
-actually expected. The models differ both in their expectations and
-their uncertainty (colored bands show two standard deviations away).
-Note that the GP is particularly uncertain about the dynamics relative
-to structurally incorrect models like the Ricker.
-
-```{r figure_3, fig.width=8, fig.height=6, fig.cap="Outside the range of the training data (Figure 1), the true dynamics (black dots) fall outside the uncertainty (two standard deviations, colored bands) of the structurally incorrect parametric models (Ricker, Myers), but inside the uncertainty predicted by the GP.
-Points show the stock size simulated by the true model.
-Overlay shows the range of states predicted by each model, based on the state observed in the previous time step.
-The Ricker model always (wrongly) predicts positive population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).
-Note that because it does not assume a parametric form but instead relies more directly on the data, the GP is both more pessimistic and more uncertain about the future state than the parametric models.", dependson="plot-options"}
-y <- numeric(8)
-y[1] <- 4.5
-for(t in 1:(length(y)-1))
-      y[t+1] = z_g() * f(y[t], h=0, p=p)
-# predicts means, does not reflect uncertainty estimate!
-crash_data <- step_ahead_posteriors(y)
-crash_data <- subset(crash_data, variable %in% c("GP", "Allen", "Ricker", "Myers"))
-ggplot(crash_data) + 
-  geom_boxplot(aes(as.factor(as.integer(time)), value, 
-                   fill = variable, col=variable), 
-               outlier.size=1, position="identity") + 
-#  geom_line(aes(time, value, col = variable, 
-#            group=interaction(L1,variable))) + 
-  geom_point(aes(time, stock), size = 3) + 
-  scale_fill_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")]) +  
-  scale_colour_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")]) +  
-  facet_wrap(~variable) + 
-  theme(legend.position="none") + xlab("time") + ylab("stock size")
-
-write.csv(crash_data, "components/data/figure3.csv")
-```
-
-<!--
-### Figure 4: Inferred Policies
-
-- Inferred policies differ substantially among models
-- The structurally correct model and the GP are close to the true model
-- alternatives are not close
--->
+the tipping point. Figure S1 (see supplement) shows the performance of the models outside the observed training data.
 
@@
 Despite the similarities in model fits to the observed data, the
-policies inferred under each model differ widely (Figure 4).
+policies inferred under each model differ widely (Figure 3).
 Policies are shown in terms of target escapement, $S_t$.  Under models
@@
 
-In Figure 5, we show the consequences of managing 100 replicate
+In Figure 4, we show the consequences of managing 100 replicate
 realizations of the simulated fishery under each of the policies
@@
 easily be compared by using the relative differences in net present
-value realized by each of the model (Figure 6).  The GPDP
+value realized by each of the model (Figure 5).  The GPDP
 most consistently realizes a value close to the optimal solution,
@@
 noise scale, parameters or structure of the underlying model. We consider
-both a Latin hypercube approach and a more focused investigation of the
+both a Latin hypercube approach (see supplementary materials, Fig S11) and a more focused investigation of the
 parameters most responsible for impacting the transition rate: the relative 
 distance to the Allee threshold and the noise effects.
-
-The Latin hypercube approach systematically varies all combinations of
-parameters, providing a more general test than varying only one parameter
-at a time.  We loop across eight replicates of three different randomly
-generated parameter sets for each of two different generating models
-(Allen and Myers) over two different noise levels (0.01 and 0.05),
-for a total of 8 x 3 x 2 x 2 = 96 scenarios. The Gaussian Process performs
-nearly optimally in each case, relative to the optimal solution with no
-parameter or model uncertainty (Figure S9, appendix).
 
@@
 
-The GPDP is weakly influenced by increasing stochasticity or increasing
-Allee effects over much of the range (Figure 7) Much larger stochasticity or higher Allee
+The GPDP is only weakly influenced by increasing stochasticity or increasing
+Allee effects over much of the range (Figure 6). Much larger stochasticity or higher Allee
 effects make even the optimal solution without any model or parameter uncertainty

### 2014-10-15 manuscripts/manuscript.tex
rerun

@@
 system. No observations occur at the theoretical optimum harvest rate or
-near the tipping point.
-
-Figure 3 illustrates the performance of the GP and parametric models
-outside the observed training data. The mean trajectory under the
-underlying model is shown by the black dots, while the corresponding
-prediction made by the model shown by the box and whiskers plots.
-Predictions are based on the true expected value in the previous time
-step. Predicted distributions that lie entirely above the expected
-dynamics indicate the expectation of stock sizes higher than what is
-actually expected. The models differ both in their expectations and
-their uncertainty (colored bands show two standard deviations away).
-Note that the GP is particularly uncertain about the dynamics relative
-to structurally incorrect models like the Ricker.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_3.pdf}
-\caption{Outside the range of the training data (Figure 1), the true
-dynamics (black dots) fall outside the uncertainty (two standard
-deviations, colored bands) of the structurally incorrect parametric
-models (Ricker, Myers), but inside the uncertainty predicted by the GP.
-Points show the stock size simulated by the true model. Overlay shows
-the range of states predicted by each model, based on the state observed
-in the previous time step. The Ricker model always (wrongly) predicts
-positive population growth, while the actual population shrinks in each
-step as the initial condition falls below the Allee threshold of the
-underlying model (Allen). Note that because it does not assume a
-parametric form but instead relies more directly on the data, the GP is
-both more pessimistic and more uncertain about the future state than the
-parametric models.}
-\end{figure}
+near the tipping point. Figure S1 (see supplement) shows the performance
+of the models outside the observed training data.
 
@@
 Despite the similarities in model fits to the observed data, the
-policies inferred under each model differ widely (Figure 4). Policies
+policies inferred under each model differ widely (Figure 3). Policies
 are shown in terms of target escapement, $S_t$. Under models such as
@@
 compared by using the relative differences in net present value realized
-by each of the model (Figure 6). The GPDP most consistently realizes a
+by each of the model (Figure 5). The GPDP most consistently realizes a
 value close to the optimal solution, and importantly avoids ever driving
@@
 simulation, noise scale, parameters or structure of the underlying
-model. We consider both a Latin hypercube approach and a more focused
-investigation of the parameters most responsible for impacting the
-transition rate: the relative distance to the Allee threshold and the
-noise effects.
-
-The Latin hypercube approach systematically varies all combinations of
-parameters, providing a more general test than varying only one
-parameter at a time. We loop across eight replicates of three different
-randomly generated parameter sets for each of two different generating
-models (Allen and Myers) over two different noise levels (0.01 and
-0.05), for a total of 8 x 3 x 2 x 2 = 96 scenarios. The Gaussian Process
-performs nearly optimally in each case, relative to the optimal solution
-with no parameter or model uncertainty (Figure S9, appendix).
+model. We consider both a Latin hypercube approach (see supplementary
+materials, Fig S11) and a more focused investigation of the parameters
+most responsible for impacting the transition rate: the relative
+distance to the Allee threshold and the noise effects.
 
@@
 
-The GPDP is weakly influenced by increasing stochasticity or increasing
-Allee effects over much of the range (Figure 7) Much larger
+The GPDP is only weakly influenced by increasing stochasticity or
+increasing Allee effects over much of the range (Figure 6). Much larger
 stochasticity or higher Allee effects make even the optimal solution

### 2014-10-15 manuscripts/supplement.Rmd
old commits

@@
     fig_caption: true
-    fig_crop: false
     template: components/elsarticle.latex
@@
 
-Appendix: Model definitions and estimation
-===========================================
+Figure S1 illustrates the performance of the GP and parametric models
+outside the observed training data. The mean trajectory under the
+underlying model is shown by the black dots, while the corresponding
+prediction made by the model shown by the box and whiskers plots.
+Predictions are based on the true expected value in the previous time
+step.  Predicted distributions that lie entirely above the expected
+dynamics indicate the expectation of stock sizes higher than what is
+actually expected. The models differ both in their expectations and
+their uncertainty (colored bands show two standard deviations away).
+Note that the GP is particularly uncertain about the dynamics relative
+to structurally incorrect models like the Ricker.
+
+```{r figure_3, fig.width=8, fig.height=6, fig.cap="Outside the range of the training data (Figure 1), the true dynamics (black dots) fall outside the uncertainty (two standard deviations, colored bands) of the structurally incorrect parametric models (Ricker, Myers), but inside the uncertainty predicted by the GP.
+Points show the stock size simulated by the true model.
+Overlay shows the range of states predicted by each model, based on the state observed in the previous time step.
+The Ricker model always (wrongly) predicts positive population growth, while the actual population shrinks in each step as the initial condition falls below the Allee threshold of the underlying model (Allen).
+Note that because it does not assume a parametric form but instead relies more directly on the data, the GP is both more pessimistic and more uncertain about the future state than the parametric models.", dependson="plot-options"}
+y <- numeric(8)
+y[1] <- 4.5
+for(t in 1:(length(y)-1))
+      y[t+1] = z_g() * f(y[t], h=0, p=p)
+# predicts means, does not reflect uncertainty estimate!
+crash_data <- step_ahead_posteriors(y)
+crash_data <- subset(crash_data, variable %in% c("GP", "Allen", "Ricker", "Myers"))
+ggplot(crash_data) + 
+  geom_boxplot(aes(as.factor(as.integer(time)), value, 
+                   fill = variable, col=variable), 
+               outlier.size=1, position="identity") + 
+#  geom_line(aes(time, value, col = variable, 
+#            group=interaction(L1,variable))) + 
+  geom_point(aes(time, stock), size = 3) + 
+  scale_fill_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")]) +  
+  scale_colour_manual(values=colorkey[c("GP", "Allen", "Ricker", "Myers")]) +  
+  facet_wrap(~variable) + 
+  theme(legend.position="none") + xlab("time") + ylab("stock size")
+
+write.csv(crash_data, "components/data/figure3.csv")
+```
 
@@
 ```
+
+Further sensitivity analysis
+=============================
+
+The Latin hypercube approach systematically varies all combinations of
+parameters, providing a more general test than varying only one parameter
+at a time.  We loop across eight replicates of three different randomly
+generated parameter sets for each of two different generating models
+(Allen and Myers) over two different noise levels (0.01 and 0.05),
+for a total of 8 x 3 x 2 x 2 = 96 scenarios. The Gaussian Process performs
+nearly optimally in each case, relative to the optimal solution with no
+parameter or model uncertainty (Figure S10, appendix).
 

### 2014-10-17 manuscripts/manuscript.Rmd
working on edits to manuscript based on reviews

@@
 <!-- Run R code for analysis, to be called by figures -->
-```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE, cache=FALSE}
+```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
 cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
@@
 
-```{r libraries, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
+```{r libraries}
 library("nonparametricbayes") 
@@
 
-```{r ricker_priors_xtable, include=TRUE, echo=FALSE, cache=FALSE}
+```{r ricker_priors_xtable, include=TRUE, echo=FALSE}
 ricker_priors_xtable <- data.frame(
@@
 infer posterior distributions of the two parameters of the GP (Figure
-S13, code in appendix). Since the posterior distributions differ
-substantially from the priors (Figure S13), we can be assured that most
+S3, code in appendix). Since the posterior distributions differ
+substantially from the priors (Figure S3), we can be assured that most
 of the information in the posterior comes from the data rather than the
@@
 It is reasonably straight-forward to derive the harvest policy from the
-estimated BP by inserting it into a SDP algorithm. The only difference 
+estimated GP by inserting it into a SDP algorithm. The only difference 
 is that the uncertainty in the future state under the GP, $X_{t+1}$,
@@
 
-```{r Table1, dependson="deviances", include=TRUE, results="asis", echo=FALSE, cache=FALSE}
+```{r Table1, dependson="deviances", include=TRUE, results="asis", echo=FALSE}
 pandoc.table(rbind(dictable, aictable, bictable), caption="Model selection scores for several common criteria all (wrongly) select the simplest model.
@@
 has a mean shown in black. Parameter uncertainty (not shown) spreads
-out the estimates further.  The observed data from which each model are
+out the estimates further.  The observed data from which each model is 
 estimated is also shown.  The observations come from only a limited region
@@
 
-```{r figure_4, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
+```{r figure_3, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
 Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
@@
 ggplot(policies, aes(stock, stock - value, color=method)) +
-  geom_line(lwd=1.2) + xlab("stock size") + ylab("escapement")  +
-  scale_colour_manual(values=colorkey)
-
-write.csv(policies, "components/data/figure4.csv")
+  geom_line(lwd=1.2) + 
+  facet_wrap(~method) +
+  xlab("stock size") + 
+  ylab("escapement")  +
+  scale_colour_manual(values=colorkey, guide=FALSE)
+
+write.csv(policies, "components/data/figure3.csv")
 ```
@@
 unharvested, while above that size the harvest strategy aims to return the
-population to $S$, resulting in the hockey-stick shaped policies shown.
+population to $S$.
+Whenever a model predicts that the population will not persist below a certain threshold, the optimal solution is to harvest the entire population immediately, resulting in an escapement $S=0$, as seen in the true (correct form, exact parameters) model, the Allen model (correct form, estimated parameters) and the GP.
 Only the structurally correct model (Allen model) and the GP produce
-policies close to the true optimum policy (where both the underlying
-model structure and parameter values are known without error).
-
-```{r figure_5, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models.
+policies close to the true optimum policy.
+
+```{r figure_4, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models.
 We show 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
@@
   geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=0.1) +
-  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) + 
-  facet_wrap(~method) + guides(legend.position="none")
-
-write.csv(sims_data, "components/data/figure5.csv")
+  scale_colour_manual(values=colorkey, guide = FALSE) + 
+  facet_wrap(~method) + 
+  guides(legend.position="none", color=FALSE)
+
+write.csv(sims_data, "components/data/figure4.csv")
 ```
@@
 unfished optimum.  The structurally incorrect Ricker model over-harvests
-the population passed the tipping point consistently, resulting in the
+the population past the tipping point consistently, resulting in the
 immediate crash of the stock and thus derives minimal long-term catch.
@@
 
-```{r figure_6, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
+```{r figure_5, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
 For each data set, the three models are estimated as described above.
@@
   xlab("Total profit by replicate") + 
-  scale_fill_manual(values=colorkey) # density plots fail when delta fn
-
-# ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable), binwidth=0.1) + 
-#  xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)
-# ggplot(actual_over_optimal, aes(value, fill=variable, color=variable)) + # density plots fail when delta fn
-#  stat_density(aes(y=..density..), position="stack", adjust=3) + 
-#  xlab("Total profit by replicate")+ scale_fill_manual(values=colorkey)+ scale_color_manual(values=colorkey)
-
-write.csv(actual_over_optimal, "components/data/figure6.csv")
+  scale_fill_manual(values=colorkey, guide=FALSE) # density plots fail when delta fn
+
+write.csv(actual_over_optimal, "components/data/figure5.csv")
 ```
@@
 noise scale, parameters or structure of the underlying model. We consider
-both a Latin hypercube approach (see supplementary materials, Fig S11) and a more focused investigation of the
+both a Latin hypercube approach (see supplementary materials, Fig S10) and a more focused investigation of the
 parameters most responsible for impacting the transition rate: the relative 
@@
 
-```{r figure_7, fig.width=6, fig.height=3, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).
+```{r figure_6, fig.width=6, fig.height=3, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).
 Other than the focal parameter (stochasticity, Allee threshold), other parameters are held fixed as above to illustrate this effect."}
@@
 dictated by the GP balances this uncertainty against
-the immediate value of the harvest, and act to stabilize the population 
+the immediate value of the harvest, and acts to stabilize the population 
 dynamics in a region of state space in which the predictions can be 

### 2014-10-17 manuscripts/manuscript.tex
working on edits to manuscript based on reviews

@@
 (2003)) to infer posterior distributions of the two parameters of the GP
-(Figure S13, code in appendix). Since the posterior distributions differ
-substantially from the priors (Figure S13), we can be assured that most
+(Figure S3, code in appendix). Since the posterior distributions differ
+substantially from the priors (Figure S3), we can be assured that most
 of the information in the posterior comes from the data rather than the
@@
 It is reasonably straight-forward to derive the harvest policy from the
-estimated BP by inserting it into a SDP algorithm. The only difference
+estimated GP by inserting it into a SDP algorithm. The only difference
 is that the uncertainty in the future state under the GP, $X_{t+1}$,
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_1.pdf}
+\includegraphics{components/figure/manuscript-figure_1-1.pdf}
 \caption{Points show the training data of stock-size over time. Curves
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_2.pdf}
+\includegraphics{components/figure/manuscript-figure_2-1.pdf}
 \caption{The inferred Gaussian process compared to the true process and
@@
 has a mean shown in black. Parameter uncertainty (not shown) spreads out
-the estimates further. The observed data from which each model are
+the estimates further. The observed data from which each model is
 estimated is also shown. The observations come from only a limited
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_4.pdf}
+\includegraphics{components/figure/manuscript-figure_4-1.pdf}
 \caption{The steady-state optimal policy (infinite boundary) calculated
@@
 while above that size the harvest strategy aims to return the population
-to $S$, resulting in the hockey-stick shaped policies shown. Only the
-structurally correct model (Allen model) and the GP produce policies
-close to the true optimum policy (where both the underlying model
-structure and parameter values are known without error).
+to $S$. Whenever a model predicts that the population will not persist
+below a certain threshold, the optimal solution is to harvest the entire
+population immediately, resulting in an escapement $S=0$, as seen in the
+true (correct form, exact parameters) model, the Allen model (correct
+form, estimated parameters) and the GP.\\Only the structurally correct
+model (Allen model) and the GP produce policies close to the true
+optimum policy.
 
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_5.pdf}
+\includegraphics{components/figure/manuscript-figure_5-1.pdf}
 \caption{In the management context, GPDP outperforms approaches based on
@@
 unfished optimum. The structurally incorrect Ricker model over-harvests
-the population passed the tipping point consistently, resulting in the
+the population past the tipping point consistently, resulting in the
 immediate crash of the stock and thus derives minimal long-term catch.
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_6.pdf}
+\includegraphics{components/figure/manuscript-figure_6-1.pdf}
 \caption{Histograms of the realized net present value of the fishery
@@
 model. We consider both a Latin hypercube approach (see supplementary
-materials, Fig S11) and a more focused investigation of the parameters
+materials, Fig S10) and a more focused investigation of the parameters
 most responsible for impacting the transition rate: the relative
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_7.pdf}
+\includegraphics{components/figure/manuscript-figure_7-1.pdf}
 \caption{The effect of increasing noise or decreasing Allee threshold
@@
 range observed. The management policy dictated by the GP balances this
-uncertainty against the immediate value of the harvest, and act to
+uncertainty against the immediate value of the harvest, and acts to
 stabilize the population dynamics in a region of state space in which

### 2014-10-28 manuscripts/manuscript.Rmd
dockerfile updated

@@
 }
-par <- c(p[1]*rlnorm(1,0,.1), 
-         p[2]*rlnorm(1,0,.1), 
-         p[3]*rlnorm(1,0, .1), 
-         sigma_g * rlnorm(1,0,.1))
+par <- c(3, 
+         9, 
+         6, 
+         0.06)
 o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
@@
 
+```{r refresh-plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE, cache=FALSE}
+cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
+               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+opts_chunk$set(dev='pdf', dev.args=c(version="1.7"))
+opts_chunk$set(fig.width=4, fig.height=3, echo=FALSE)
+theme_set(theme_bw(base_size=12))
+```
+
 Introduction
@@
 
+<!--
+FIXME We need to be 
+-->
+
 This paper represents the first ecological application of the
@@
 <!--
--  _Note on "not magic"_: honest uncertainty + SDP
--->
-
-<!--
+ _Note on "not magic"_: honest uncertainty + SDP
 The idea that any approach can perform well without either having to
@@
 
-<!--move to  much later *Or do we need this at all?* 
-- _Note on comparing models_ (via value function rather than by "fit"). 
--->
-
 To illustrate GPDP, we consider the performance of the GPDP policy against
@@
 
-<!-- Although much argument can be made
+<!--
+
+Although much argument can be made
 over the 'correct' objective function, e.g. how to account for the
@@
 management of human intervention in natural resource systems. After that
-we describe three parametric models for population dynamics and the GP
-description of population dynamics.
+we describe three parametric models for population dynamics and the Gaussian
+Process (GP)[^1] description of population dynamics.
+
+[^1]: Note that we abbreviate Gaussian Process as GP, which refers only to 
+the statistical model we use to approximate the population dynamics (as opposed
+to a familiar parametric model such as the Ricker equation), while we use the
+term GPDP, Gaussian Process Dynamic Programming, to refer to the _use_ of a
+GP as the underlying process model when solving a Dynamic Program, as we illustrate
+in this paper.  Hence we will refer to the models as: GP, Ricker, Allen, etc, while
+we refer to the novel method we put forward here as GPDP.
 
@@
 
-where the population size after the take is $X_t-h_t$, which is then
-translated into $X_{t+1}$ by Eqn 1.
+where the population size after the take is $x_t-h_t$, which is then
+translated into $X_{t+1}$ by Eqn 1 (that is, we just replaced $X_{t+1}$ with $Z f(x_t-h_t)$).
 
@@
 
-$$\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | x_t, h_t \right] \}$$
+$$\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( x_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(x_t - h_t))  | x_t, h_t \right] \}$$
 
@@
 the essential feature of a threshold but differs in the structure.
+
+Throughout the presentation, we will refer to the use of the "True" model
+to denote solving for the optimal solution when the underlying parameters
+_are known without error_, and refer to the "Allen" model to denote the
+when these parameters have instead only been estimated from the sample data.
 
@@
 
-A GP is an infinite dimensional (because it is a function of a continuous input
-variable) stochastic process for a finite sample of points from any realization is a multivariate
-normal distribution. Thus, to characterize the GP we need a mean function
+A GP is an infinite dimensional (because it is a function of a continuous
+input variable) stochastic process. Any realization of the process for
+some finite sample of points $n$ is a multivariate normal distribution
+of dimension $n$. Thus, to characterize the GP we need a mean function
 and a covariance function. We proceed as follows.
@@
 
-Then under the GP, given a vector of observed data $X_0$ the predicted data $X_p$ obeys
-
-$$f(X_p|X_o) \sim \mathcal{N}(E,C)$$
-$$E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y$$
-$$C = K(X_p, X_p) - K(X_p, X_o) K(X_o,X_o)^{-1} K(X_o, X_p)$$
+Then under the GP, given a vector of observed data $X_o$ the predicted data $X_p$ obeys
+
+$$f(X_p|X_o, X_{t+1}) \sim \mathcal{N}(E,C)$$
+$$E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)^{-1} X_{t+1}$$
+$$C = K(X_p, X_p) - K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)^{-1} K(X_o, X_p)$$
 
@@
 
-$$ K(x,y) = \exp\left(\frac{-(x-y)^2}{2 \ell^2} \right)$$
+$$ K(\mathbf{x},\mathbf{y}) = \exp\left(\frac{-|\mathbf{x}-\mathbf{y}|^2}{2 \ell^2} \right)$$
 
@@
 Curves show the posterior step-ahead predictions based on each of the estimated models.
-Observe that all models are fitting the data reasonably well.", fig.width=8, fig.height=6}
+Observe that all models are fitting the data reasonably well.
+In the GP panel, black lines are means of the GP for random samples from the posterior distributions, while grey bands around them reflect the standard deviations of the GP distribution.
+(Recall that while given a particular set of parameters, a parametric model gives just one curve, the GP is always a distribution, whose mean we plot in black.)", fig.width=8, fig.height=6, cache=FALSE}
 step_ahead_posteriors <- function(x){
@@
 The training data are also shown as black points.
-The GP is conditioned on (0,0), shown as a pseudo-data point.", fig.width=8, fig.height=6}
+The GP is conditioned on (0,0), shown as a pseudo-data point, as explained in the text", fig.width=8, fig.height=6, cache=FALSE}
 x_grid_short <- x_grid[1:40]
@@
 the relationship between observed stock size (x-axis) to the stock
-size after recruitment the following year.  Note that in contrast to
+size after recruitment the following year.  
+In addition to the raw data, Figure 2 illustrates that the GP is conditioned
+on going through the point 0,0 without error.
+This conditioning on (0,0) is equivalent to making the assumption that a species hitting 0 is extinct forever & without uncertainty, despite the lack of any data in the observed sequence to justify this.
+The use of this assumption illustrates how the GP approach can capture common-sense biology without having to assume more explicit details about the dynamics at low population numbers that have never been observed.
+Not that we need need not make this assumption (perhaps there is immigration after all); and one could repeat the entire analysis without this assumption.
+
+Note that in contrast to
 the  other models shown, the expected Gaussian process corresponds to a
@@
 ```{r figure_3, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."}
+Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", fig.width=8, fig.height=6, cache=FALSE}
 policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
@@
 ```{r figure_4, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models.
-We show 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
+We show 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6, cache=FALSE}
 ggplot(sims_data) + 
@@
 ```{r figure_5, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates.
-For each data set, the three models are estimated as described above.
-Values plotted are the averages of a given policy over 100 replicate simulations.
-Details and code provided in the supplement.", fig.width=6, fig.height=6}
+For each data set, the three models are estimated as described above.", fig.width=6, fig.height=6, cache=FALSE}
 
@@
   guides(legend.position = "none") +
-  xlab("Total profit by replicate") + 
+  xlab("Net present value") + 
   scale_fill_manual(values=colorkey, guide=FALSE) # density plots fail when delta fn
@@
 ```{r figure_6, fig.width=6, fig.height=3, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).
-Other than the focal parameter (stochasticity, Allee threshold), other parameters are held fixed as above to illustrate this effect."}
+Other than the focal parameter (stochasticity, Allee threshold), other parameters are held fixed as above to illustrate this effect.", cache=FALSE}
 plot_sigmas <- ggplot(vary_sigma, aes(noise, value)) + 

### 2014-10-28 manuscripts/supplement.Rmd
dockerfile updated

@@
     fig_caption: true
-    template: components/elsarticle.latex
     keep_tex: true
@@
 
-opts_chunk$set(dev.opts=c(version="1.3"), dev="pdf")
+opts_chunk$set(dev.opts=c(version="1.7"), dev="pdf")
 
@@
 
-```{r figure_3, fig.width=8, fig.height=6, fig.cap="Outside the range of the training data (Figure 1), the true dynamics (black dots) fall outside the uncertainty (two standard deviations, colored bands) of the structurally incorrect parametric models (Ricker, Myers), but inside the uncertainty predicted by the GP.
+```{r figure_S1, fig.width=8, fig.height=6, fig.cap="Outside the range of the training data (Figure 1), the true dynamics (black dots) fall outside the uncertainty (two standard deviations, colored bands) of the structurally incorrect parametric models (Ricker, Myers), but inside the uncertainty predicted by the GP.
 Points show the stock size simulated by the true model.
@@
 
-```{r figure_S1, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options"}
+```{r figure_S2, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options"}
 gp_assessment_plots$traces_plot
@@
 
-```{r figure_S2, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
+```{r figure_S3, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
 gp_assessment_plots$posteriors_plot
@@
 
-```{r figure_S3, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
-plot_ricker_traces
+```{r figure_S4, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
+ggplot(ricker_posteriors) + 
+  geom_line(aes(index, value)) + # priors, need to fix order though
+  facet_wrap(~ variable, scale="free_y", ncol=1)
 ```
 
-```{r figure_S4, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
+```{r figure_S5, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
 ggplot(ricker_posteriors, aes(value)) + 
@@
 
-```{r figure_S5, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
-plot_myers_traces
+```{r figure_S6, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
+ggplot(myers_posteriors) + 
+  geom_line(aes(index, value)) + # priors, need to fix order though
+  facet_wrap(~ variable, scale="free_y", ncol=1)
 ```
 
-```{r figure_S6, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
+```{r figure_S7, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
 ggplot(myers_posteriors, aes(value)) + 
@@
 
-```{r figure_S7, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
-plot_allen_traces
+```{r figure_S8, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
+ggplot(allen_posteriors) + 
+  geom_line(aes(index, value)) + # priors, need to fix order though
+  facet_wrap(~ variable, scale="free_y", ncol=1)
 ```
 
-```{r figure_S8, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
+```{r figure_S9, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
 ggplot(allen_posteriors, aes(value)) + 
@@
 
-```{r figure_S9, fig.height=6, fig.width=10, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
+```{r figure_S10, fig.height=6, fig.width=10, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
 Histograms shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters.

### 2014-10-28 manuscripts/manuscript.tex

@@
 management of human intervention in natural resource systems. After that
-we describe three parametric models for population dynamics and the GP
-description of population dynamics.
+we describe three parametric models for population dynamics and the
+Gaussian Process (GP)\footnote{Note that we abbreviate Gaussian Process
+  as GP, which refers only to the statistical model we use to
+  approximate the population dynamics (as opposed to a familiar
+  parametric model such as the Ricker equation), while we use the term
+  GPDP, Gaussian Process Dynamic Programming, to refer to the \emph{use}
+  of a GP as the underlying process model when solving a Dynamic
+  Program, as we illustrate in this paper. Hence we will refer to the
+  models as: GP, Ricker, Allen, etc, while we refer to the novel method
+  we put forward here as GPDP.} description of population dynamics.
 
@@
 
-where the population size after the take is $X_t-h_t$, which is then
-translated into $X_{t+1}$ by Eqn 1.
+where the population size after the take is $x_t-h_t$, which is then
+translated into $X_{t+1}$ by Eqn 1 (that is, we just replaced $X_{t+1}$
+with $Z f(x_t-h_t)$).
 
@@
 
-\[\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | x_t, h_t \right] \}\]
+\[\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( x_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(x_t - h_t))  | x_t, h_t \right] \}\]
 
@@
 the essential feature of a threshold but differs in the structure.
+
+Throughout the presentation, we will refer to the use of the ``True''
+model to denote solving for the optimal solution when the underlying
+parameters \emph{are known without error}, and refer to the ``Allen''
+model to denote the when these parameters have instead only been
+estimated from the sample data.
 
@@
 A GP is an infinite dimensional (because it is a function of a
-continuous input variable) stochastic process for a finite sample of
-points from any realization is a multivariate normal distribution. Thus,
-to characterize the GP we need a mean function and a covariance
-function. We proceed as follows.
+continuous input variable) stochastic process. Any realization of the
+process for some finite sample of points $n$ is a multivariate normal
+distribution of dimension $n$. Thus, to characterize the GP we need a
+mean function and a covariance function. We proceed as follows.
 
@@
 
-Then under the GP, given a vector of observed data $X_0$ the predicted
+Then under the GP, given a vector of observed data $X_o$ the predicted
 data $X_p$ obeys
 
-\[f(X_p|X_o) \sim \mathcal{N}(E,C)\]
-\[E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y\]
-\[C = K(X_p, X_p) - K(X_p, X_o) K(X_o,X_o)^{-1} K(X_o, X_p)\]
+\[f(X_p|X_o, X_{t+1}) \sim \mathcal{N}(E,C)\]
+\[E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)^{-1} X_{t+1}\]
+\[C = K(X_p, X_p) - K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)^{-1} K(X_o, X_p)\]
 
@@
 
-\[ K(x,y) = \exp\left(\frac{-(x-y)^2}{2 \ell^2} \right)\]
+\[ K(\mathbf{x},\mathbf{y}) = \exp\left(\frac{-|\mathbf{x}-\mathbf{y}|^2}{2 \ell^2} \right)\]
 
@@
 show the posterior step-ahead predictions based on each of the estimated
-models. Observe that all models are fitting the data reasonably well.}
+models. Observe that all models are fitting the data reasonably well. In
+the GP panel, black lines are means of the GP for random samples from
+the posterior distributions, while grey bands around them reflect the
+standard deviations of the GP distribution. (Recall that while given a
+particular set of parameters, a parametric model gives just one curve,
+the GP is always a distribution, whose mean we plot in black.)}
 \end{figure}
@@
 also shown as black points. The GP is conditioned on (0,0), shown as a
-pseudo-data point.}
+pseudo-data point, as explained in the text}
 \end{figure}
@@
 relationship between observed stock size (x-axis) to the stock size
-after recruitment the following year. Note that in contrast to the other
-models shown, the expected Gaussian process corresponds to a
-distribution of curves - as indicated by the gray band - which itself
-has a mean shown in black. Parameter uncertainty (not shown) spreads out
-the estimates further. The observed data from which each model is
-estimated is also shown. The observations come from only a limited
-region of state space corresponding to unharvested or weakly harvested
-system. No observations occur at the theoretical optimum harvest rate or
-near the tipping point. Figure S1 (see supplement) shows the performance
-of the models outside the observed training data.
+after recruitment the following year.\\In addition to the raw data,
+Figure 2 illustrates that the GP is conditioned on going through the
+point 0,0 without error. This conditioning on (0,0) is equivalent to
+making the assumption that a species hitting 0 is extinct forever \&
+without uncertainty, despite the lack of any data in the observed
+sequence to justify this. The use of this assumption illustrates how the
+GP approach can capture common-sense biology without having to assume
+more explicit details about the dynamics at low population numbers that
+have never been observed. Not that we need need not make this assumption
+(perhaps there is immigration after all); and one could repeat the
+entire analysis without this assumption.
+
+Note that in contrast to the other models shown, the expected Gaussian
+process corresponds to a distribution of curves - as indicated by the
+gray band - which itself has a mean shown in black. Parameter
+uncertainty (not shown) spreads out the estimates further. The observed
+data from which each model is estimated is also shown. The observations
+come from only a limited region of state space corresponding to
+unharvested or weakly harvested system. No observations occur at the
+theoretical optimum harvest rate or near the tipping point. Figure S1
+(see supplement) shows the performance of the models outside the
+observed training data.
 
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_4-1.pdf}
+\includegraphics{components/figure/manuscript-figure_3-1.pdf}
 \caption{The steady-state optimal policy (infinite boundary) calculated
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_5-1.pdf}
+\includegraphics{components/figure/manuscript-figure_4-1.pdf}
 \caption{In the management context, GPDP outperforms approaches based on
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_6-1.pdf}
+\includegraphics{components/figure/manuscript-figure_5-1.pdf}
 \caption{Histograms of the realized net present value of the fishery
 over a range of simulated data and resulting parameter estimates. For
-each data set, the three models are estimated as described above. Values
-plotted are the averages of a given policy over 100 replicate
-simulations. Details and code provided in the supplement.}
+each data set, the three models are estimated as described above.}
 \end{figure}
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_7-1.pdf}
+\includegraphics{components/figure/manuscript-figure_6-1.pdf}
 \caption{The effect of increasing noise or decreasing Allee threshold

### 2014-11-03 manuscripts/manuscript.Rmd
edit all equations: time as function argument, not subscript. fix p vs theta, fix X_o X_p, fix K_ij

@@
 
-We let $X_t$ denote the size (numbers or biomass) of the focal population at time $t$ and assume that in the absence of take its dynamics are.
+We let $x(t)$ denote the size (numbers or biomass) of the focal
+population at time $t$ and assume that in the absence of take its
+dynamics are. (Throughout, we will use $X(t+1)$ with a capital $X$ as
+a reminder that this is a stochastic, uncertainty quantity at time $t$,
+while we use lowercase $x(t)$ as a reminder that the current population
+size is treated as a given without error.  The latter assumption is
+straight-forward to relax but will simplify the presentation here).
 
 \begin{equation}
-X_{t+1}= Z_t f(X_t) \label{eq1} 
+X(t+1) = Z(t) f(x(t), p) \label{eq1} 
 \end{equation}
@@
 and $p$ is a vector of parameters to be estimated from the data.
-We describe three choices for $f(X_t,p)$ in the next section.In this
-simple model, the control action is a harvest or take, $h_t$, measured
+We describe three choices for $f(X(t),p)$ in the next section.In this
+simple model, the control action is a harvest or take, $h(t)$, measured
 in the same units as $X$, at time $t$.  Thus, in the presence of take,
 the population size on the right hand side of Eqn 1 is replaced by
-$S_t=X_t-h_t$.
-
-To construct the value function, we consider a return when $X_t=x$ and
-harvest $h_t=h$ denoted as the reward, $R(x,h)$. For example, if the
-return is simply the harvest at time $t$, then $R(x,h)=min(x,h)$. We
+$S(t) = X(t) - h(t)$.
+
+To construct the value function, we consider a return when $X(t) = x $ and
+harvest $h(t) = h$ denoted as the reward, $R(x, h)$. For example, if the
+return is simply the harvest at time $t$, then $R(x, h) = \min(x, h)$. We
 assume that future harvests are discounted relative to current ones at
@@
 and a final time $T$. That is, we seek to maximize over choices of
-harvest $\mathbf{E}_{X_t+1} [ \sum_{t = 0}^{T}  R_t(X_t, h_t) \delta^t]$,
+harvest $\mathbf{E}_{X(t+1)} [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$,
 where the state dynamics are given by Eqn 1 and $\mathbf{E}$ denotes the
@@
 
-In order to find that policy, we introduce the value function $V_t(x_t)$
+In order to find that policy, we introduce the value function $V(x(t), t)$
 representing the total discounted catch from time $t$ onwards
-given that $X_t=x_t$.  This value function satisfies an equation of
+given that $X(t) = x(t)$.  This value function satisfies an equation of
 SDP  [@Mangel1988; @Clark2000; @Clark2009], sometimes called the
@@
 \begin{equation}
-V_t(x_t) = \max_{h_t} \lbrace R(h_t, x_t) + \delta \cdot \mathbf{\mathrm{E}}_{X_{t+1}} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] \rbrace
+V(x(t), t) = \max_{h(t)} \lbrace R(h(t), x(t)) + \delta \cdot \mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] \rbrace
 \end{equation}
@@
 where expectation is taken over all possible values of the next state,
-$X_{t+1}$, and maximized over all possible choices of harvest, $h_t$.
-That is, at time $t$, when population size is $x_t$ and harvest $h_t$ is
-applied, the immediate return is $R(h_t, x_t)$. In the classical case,
+$X(t+1)$, and maximized over all possible choices of harvest, $h(t)$.
+That is, at time $t$, when population size is $x(t)$ and harvest $h(t)$ is
+applied, the immediate return is $R(h(t), x(t))$. In the classical case,
 the sole source of uncertainty is the process stochasticity term, $Z$,
 and thus the expectation above is equivalent to taking expectations over
-$Z$. That is
-
-$$ \mathbf{\mathrm{E}}_{X_{t+1}} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] = \mathbf{\mathrm{E}}_{Z} \left[ V_{t+1}( Z f(x_t - h_t))  | x_t, h_t \right] $$
-
-where the population size after the take is $x_t-h_t$, which is then
-translated into $X_{t+1}$ by Eqn 1 (that is, we just replaced $X_{t+1}$ with $Z f(x_t-h_t)$).
+$Z(t)$. That is
+
+\begin{equation}
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V( X(t+1),t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{Z(t)} \left[V( Z(t) f(x(t) - h(t))), t+1 | x(t), h(t) \right]
+\end{equation}
+
+where the population size after the take is $x(t) - h(t)$, which is then
+translated into $X(t+1)$ by Eqn 1 (that is, we just replaced $X(t+1)$ with $Z(t) f(x(t)-h(t))$).
 
@@
 
-$$\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(x_t - h_t))  | x_t, h_t \right] \}$$
+\begin{equation}
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{p|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z(t) | p, \mathrm{data}} \left[ V(Z(t) f(x(t) - h(t)), t)  \right] \}$$
+\end{equation}
 
@@
 
-$$\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( x_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(x_t - h_t))  | x_t, h_t \right] \}$$
+\begin{equation}
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{p|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z(t) | p, \mathrm{data}} \left[ V( Z(t) f(x(t) - h(t)), t) \right] \}
+\end{equation}
 
@@
 dynamics: The Ricker model, the Allen model [@Allen2005], the Myers
-model [@Myers1995], Eqns \eqref{ricker}-\eqref{myers}, which we parameterize as follows for $f(S_t,p)$ in Eqn 1.
+model [@Myers1995], Eqns \eqref{ricker}-\eqref{myers}, which we parameterize as follows for $f(S(t),p)$ in Eqn 1.
 In all three, we let $K$ denote the carrying capacity and $r$ the maximum per capita growth rate.
@@
 \begin{equation}
-f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K} \right) } \label{ricker}
+f(S(t)|p) = S(t) e^{r \left(1 - \frac{S(t)}{K} \right) } \label{ricker}
 \end{equation}
@@
 \begin{equation}
-f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_c\right)} \label{allen}
+f(S(t)|p) = S(t) e^{r \left(1 - \frac{S(t)}{K}\right)\left(S(t) - X_c\right)} \label{allen}
 \end{equation}
@@
 \begin{equation}
-f(S(t) | p)  = \frac{r S_t^{\theta}}{1 + \frac{S_t^\theta}{K}} \label{myers}
+f(S(t) | p)  = \frac{r S(t)^{\theta}}{1 + \frac{S(t)^\theta}{K}} \label{myers}
 \end{equation}
@@
 
-We assume that the data $X_o$ are observed with some process stochasticity,
-
-$$X_{t+1} = f(X_t) + \varepsilon,$$
+We assume that the data $X(t)$ are observed with some process stochasticity,
+
+\begin{equation}
+X(t+1) = g(X(t)) + \varepsilon,
+\end{equation}
 
@@
 
-Then under the GP, given a vector of observed data $X_o$ the predicted data $X_p$ obeys
-
-$$f(X_p|X_o, X_{t+1}) \sim \mathcal{N}(E,C)$$
-$$E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)^{-1} X_{t+1}$$
-$$C = K(X_p, X_p) - K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)^{-1} K(X_o, X_p)$$
-
-Where $K$ is the Gaussian process kernel.  (Compare to  Eqn 2.22 of @Rasmussen2006 for a more detailed derivation.) Throughout, we use the radial basis function kernel:
-
-$$ K(\mathbf{x},\mathbf{y}) = \exp\left(\frac{-|\mathbf{x}-\mathbf{y}|^2}{2 \ell^2} \right)$$
+We consider the fase where we have some previous observations of the _timeseries_, $X_{\textrm{obs}}(t)$,
+and wish to _predict_ future values $X_{\textrm{p}}(t)$.  We consider the observations as pairs of points
+$X_{\textrm obs}(t), X_{\textrm obs}(t+1)$ for $t \in [1,2, ..., T]$
+
+Under the GP, a finite sample of these predictions at $n$ points into the future, $X_{\textrm p}(\tau)$ for $\tau in [T+1, T+2, ..., T+n]$, is simply given by a multivariate normal ($\mathcal{N}$):
+
+\begin{equation}
+g(X_p|X_{\textrm obs}(t), X_{\textrm obs}(t+1)) \sim \mathcal{N}(E,C)
+\end{equation}
+
+which has mean and covariance matrix given by:
+
+\begin{equation}
+E = K(X_{\textrm p}(\tau), X_{\textrm obs}(t)) \left(K(X_{\textrm obs}(t),X_{\textrm obs}(t)) - \sigma \mathbb{I} \right)^{-1} X_{\textrm obs}(t+1)
+\end{equation}
+
+\begin{equation}
+C = K(X_{\textrm p}(\tau), X_{\textrm p}(\tau)) - K(X_{\textrm p}(\tau), X_{\textrm obs}(t)) \left(K(X_{\textrm obs}(t),X_{\textrm obs}(t)) - \sigma \mathbb{I} \right)^{-1} K(X_{\textrm obs}(t), X_{\textrm p}(\tau))
+\end{equation}
+
+Where $K$ is the chosen covariance kernel.
+(Compare to  Eqn 2.22 of @Rasmussen2006 for a more detailed derivation.) Many options are possible for this kernel.
+Throughout, we use the radial basis function for simplicity, a matrix whose $(i,j)$th element is given by:
+
+\begin{equation}
+K_{i,j}(x_i, y_j) = \exp\left(\frac{-|\mathbf{x_i}-\mathbf{j}|^2}{2 \ell^2} \right)
+\end{equation}
 
@@
 
-$$f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)$$
+\begin{equation}
+f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)
+\end{equation}
 
@@
 two points in state-space decays, and $\sigma$, which gives the scale
-of the process noise by which observations $Y_{t+1}$ may differ from
-their predicted values $X_{t+1}$ given an observation of the previous
-state, $X_t$. @Munch2005a give an accessible introduction to the use
-of Gaussian processes in providing a Bayesian nonparametric description
-of the stock-recruitment relationship.
+of the process noise by which observations $Y(t+1)$ may differ from
+their predicted values $X(t+1)$ given an observation of the previous
+state, $x(t)$ [see @Munch2005a].
 
@@
 estimated GP by inserting it into a SDP algorithm. The only difference 
-is that the uncertainty in the future state under the GP, $X_{t+1}$,
+is that the uncertainty in the future state under the GP, $X(t+1)$,
 includes both process uncertainty (based on the estimation of $\sigma$) 
@@
 From the GP posteriors, we can write down the (discritized) transition
-matrix representing the probability of going to each state $X_{t+1}$ 
-given any current state $X_t$ and any harvest $h_t$ (See the function
+matrix representing the probability of going to each state $X(t+1)$ 
+given any current state $x(t)$ and any harvest $h(t)$ (See the function
 `gp_transition_matrix()` in the provided R package). Given this transition
@@
 ```{r Table1, dependson="deviances", include=TRUE, results="asis", echo=FALSE}
-pandoc.table(rbind(dictable, aictable, bictable), caption="Model selection scores for several common criteria all (wrongly) select the simplest model.
+pandoc.table(rbind(dictable, aictable, bictable), caption="Model selection scores for several common criteria (DIC: Deviance Information Criterion, AIC: Akaike Information Criterion, BIC: Bayesian Information Criterion) all (wrongly) select the simplest model.
 As the true (Allen) model is not distinguishable from the simpler (Ricker) model in the region of the observed data, this error cannot be avoided regardless of the model choice criterion.
@@
 ```{r figure_3, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", fig.width=8, fig.height=6, cache=FALSE}
+Policies are shown in terms of target escapement, $S(t)$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", fig.width=8, fig.height=6, cache=FALSE}
 policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
@@
   facet_wrap(~method) +
-  xlab("stock size") + 
-  ylab("escapement")  +
+  xlab("stock size, x(t)") + 
+  ylab("escapement, S(t)")  +
   scale_colour_manual(values=colorkey, guide=FALSE)
@@
 policies inferred under each model differ widely (Figure 3).
-Policies are shown in terms of target escapement, $S_t$.  Under models
+Policies are shown in terms of target escapement, $S(t) = x_t - h$.  Under models
 such as this a constant escapement policy is expected to be optimal
@@
 ```{r figure_4, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models.
-We show 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6, cache=FALSE}
+We show 100 replicate simulations of the stock dynamics (Eqn 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6, cache=FALSE}
 ggplot(sims_data) + 
@@
   scale_colour_manual(values=colorkey, guide = FALSE) +
+  xlab("stock size, x(t)") +
   facet_wrap(~method) + 
@@
 
-$$P(x_t < x_c | t \in T) \propto  \exp\left(\frac{L^2}{\sigma^2}\right),$$
+$$P(x(t) < x_c | t \in T) \propto  \exp\left(\frac{L^2}{\sigma^2}\right),$$
 

### 2014-11-04 manuscripts/manuscript.Rmd
edits as sent to Marc

@@
 
-<!-- __We don't have the data where we need it__ -->
+<!-- Point is: __We don't have the data where we need it__ -->
 <!-- What do we call this?  Extrapolation uncertainty?  Pathological Uncertainty? -->
@@
 Extrapolation uncertainty is felt most keenly in decision-theoretic
-(or optimal control) applications, since (a) exploring the potential
-action space typically involves considering actions that may move the
+(or optimal control) applications, since (a) exploring the space of 
+all possible actions involves considering actions that may move the
 system outside the range of observed behavior, and (b) decision-theoretic
@@
 outcomes, but depend on the weights given to all possible outcomes
-[_e.g._ @Weitzman2013].
-
-<!--This difficult position of having neither the true model nor data that
-covers the full range of possible states is unfortunately the rule more
-than the exception. The potential concern of tipping points in ecological
-dynamics [@Scheffer2001; @Polasky2011] reflects these concerns -- as
-either knowledge of the true model or more complete sampling of the
-state space would make it easy to identify if a tipping point existed.
-If we do not know but cannot rule out such a possibility, then we
-face decision-making under this 
--->
+[_e.g._ @Weitzman2013], hence estimates of the uncertainty can be as
+important as estimates of the average or expected outcome.
+
+Concerns about tipping points in ecological dynamics [@Scheffer2001;
+@Polasky2011] highlight this problem because they arise from this very
+situation.  In such cases, precise models are not available and data is
+limited or biased to a particular region of state space, such as high
+stock levels or an otherwise desirable state for the ecosystem. If such
+models or data were available, one would know just how far a system could
+be pushed before going over the tipping point, and management would be
+simple.  The fact that this concern is frequently discussed highlights
+how often we have only limited data and inperfect models. Moreover,
+the case of tipping points highlights how even small errors can have
+very large consequences, as we shall illustrate later.
 
@@
 decision-theoretic approach in face of substantial structural uncertainty.
+We use _robust_ to distinguish approaches that provide nearly optimal solutions
+(as measured by the chosen value function) without being sensitive to the
+choice of the (unknown) underlying model.
 
@@
 \begin{equation}
-\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{p|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z(t) | p, \mathrm{data}} \left[ V(Z(t) f(x(t) - h(t)), t)  \right] \}$$
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{p|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z(t) | p, \mathrm{data}} \left[ V(Z(t) f(x(t) - h(t)), t)  \right] \}
 \end{equation}
@@
 and wish to _predict_ future values $X_{\textrm{p}}(t)$.  We consider the observations as pairs of points
-$X_{\textrm obs}(t), X_{\textrm obs}(t+1)$ for $t \in [1,2, ..., T]$
-
-Under the GP, a finite sample of these predictions at $n$ points into the future, $X_{\textrm p}(\tau)$ for $\tau in [T+1, T+2, ..., T+n]$, is simply given by a multivariate normal ($\mathcal{N}$):
+$X_{\textrm{obs}}(t), X_{\textrm{obs}}(t+1)$ for $t \in [1,2, ..., T]$ up through observed time $T$.
+Under the GP, a finite sample of these predictions at $n$ points into the future, $X_{\textrm p}(\tau)$ for $\tau \in [T+1, T+2, ..., T+n]$, is simply given by a multivariate normal ($\mathcal{N}$):
 
 \begin{equation}
-g(X_p|X_{\textrm obs}(t), X_{\textrm obs}(t+1)) \sim \mathcal{N}(E,C)
+g(X_p|X_{\textrm{obs}}(t), X_{\textrm{obs}}(t+1)) \sim \mathcal{N}(E,C)
 \end{equation}
@@
 \begin{equation}
-E = K(X_{\textrm p}(\tau), X_{\textrm obs}(t)) \left(K(X_{\textrm obs}(t),X_{\textrm obs}(t)) - \sigma \mathbb{I} \right)^{-1} X_{\textrm obs}(t+1)
+E = K(X_{\textrm p}(\tau), X_{\textrm{obs}}(t)) \left(K(X_{\textrm{obs}}(t),X_{\textrm{obs}}(t)) - \sigma \mathbb{I} \right)^{-1} X_{\textrm{obs}}(t+1)
 \end{equation}
@@
 \begin{equation}
-C = K(X_{\textrm p}(\tau), X_{\textrm p}(\tau)) - K(X_{\textrm p}(\tau), X_{\textrm obs}(t)) \left(K(X_{\textrm obs}(t),X_{\textrm obs}(t)) - \sigma \mathbb{I} \right)^{-1} K(X_{\textrm obs}(t), X_{\textrm p}(\tau))
+C = K(X_{\textrm p}(\tau), X_{\textrm p}(\tau)) - K(X_{\textrm p}(\tau), X_{\textrm{obs}}(t)) \left(K(X_{\textrm{obs}}(t),X_{\textrm{obs}}(t)) - \sigma \mathbb{I} \right)^{-1} K(X_{\textrm{obs}}(t), X_{\textrm p}(\tau))
 \end{equation}
@@
 misleading -- supporting simpler models that cannot express tipping point
-dynamics merely on account of them being similar.  As the data shown
+dynamics merely on account of them being similar.  The data shown
 comes only from the basin of attraction near the unfished equilibrium,
 near which all of the models are approximately linear and approximately
-identical.
-
-Model selection criteria trade off model complexity and fit to the data. 
-When the data come from a limited region of state-space -- as is necessarily 
-the case whenever there is a potential concern about tipping point dynamics --
-simpler models therefore fit just as well and tend to be selected more often than complex 
-ones.  This approach would be appropriate when the dynamics can be expected 
-to remain in the region of the training data; for instance, if we only 
-considered the forecasting accuracy of the unfished population dynamics under
-each model.
-
-In contrast, the decision-maker's problem of setting appropriate harvest levels
-cannot exclude regions of state-space outside the observed range when integrating
-over all possible decisions to find the optimal choice.  Such problems are not 
+identical; so it is intuitive that all criteria choose the simplest model.
+In a complex world, the result is suboptimal. But in a world that might
+contain tipping points, the result is disastrous.
+
+Many managers both in fisheries and beyond face a similar situation: they
+have not observed the population dynamics at all possible densities.  A
+lack of comprehensive data at all population sizes, combined with our obvious
+inability to formulate accurate population models for low population sizes
+in the absence of any data, makes this situation the rule more than the exception.
+Relying on parametric models and model choice processes that favor simplicity
+ignores this basic fact. Acknowledging this, Carl Walters [e.g. @Walters1978] has long observed that if we began
+by fishing any newly exploited population down to very low levels and
+then let it recover, we would be much better at estimating population
+dynamics and thus predicting the optimal harvest levels.  While certainly true,
+this presents a rather risky policy in the face of potential tipping points. 
+The GPDP offers an alternative way to acknowledge this uncertainty.
+
+### GPDP population dynamics capture larger uncertainty in regions where the data are poor
+
+The parametric models perform worst when they propose a management strategy
+outside the range of the observed data. The non-parametric approach of the GPDP, 
+in contrast, allows a predictive model that expresses a great deal of uncertainty
+about the probable dynamics _outside_ the observed range, while retaining very
+good predictive accuracy _inside_ the range observed.  The management policy 
+dictated by the GPDP balances this uncertainty against
+the immediate value of the harvest, and acts to stabilize the population 
+dynamics in a region of state space in which the predictions can be 
+reliably reflected by the data.
+
+Such problems are not 
 constrained to fisheries management but ubiquitous across ecological decision-making
@@
 of an invasive, or the loss of habitat.
-
-### GPDP population dynamics capture larger uncertainty in regions where the data are poor
-
-The parametric models perform worst when they propose a management strategy
-outside the range of the observed data. The non-parametric Bayesian approach, 
-in contrast, allows a predictive model that expresses a great deal of uncertainty
-about the probable dynamics outside the observed range, while retaining very
-good predictive accuracy in the range observed.  The management policy 
-dictated by the GP balances this uncertainty against
-the immediate value of the harvest, and acts to stabilize the population 
-dynamics in a region of state space in which the predictions can be 
-reliably reflected by the data.
 

### 2014-11-04 manuscripts/manuscript.tex
edits as sent to Marc

@@
 is felt most keenly in decision-theoretic (or optimal control)
-applications, since (a) exploring the potential action space typically
+applications, since (a) exploring the space of all possible actions
 involves considering actions that may move the system outside the range
@@
 only on reasonable estimates of the expected outcomes, but depend on the
-weights given to all possible outcomes (\emph{e.g.} Weitzman 2013).
+weights given to all possible outcomes (\emph{e.g.} Weitzman 2013),
+hence estimates of the uncertainty can be as important as estimates of
+the average or expected outcome.
+
+Concerns about tipping points in ecological dynamics (Scheffer et al.
+2001, Polasky et al. 2011) highlight this problem because they arise
+from this very situation. In such cases, precise models are not
+available and data is limited or biased to a particular region of state
+space, such as high stock levels or an otherwise desirable state for the
+ecosystem. If such models or data were available, one would know just
+how far a system could be pushed before going over the tipping point,
+and management would be simple. The fact that this concern is frequently
+discussed highlights how often we have only limited data and inperfect
+models. Moreover, the case of tipping points highlights how even small
+errors can have very large consequences, as we shall illustrate later.
 
@@
 is crucial to providing a robust decision-theoretic approach in face of
-substantial structural uncertainty.
+substantial structural uncertainty. We use \emph{robust} to distinguish
+approaches that provide nearly optimal solutions (as measured by the
+chosen value function) without being sensitive to the choice of the
+(unknown) underlying model.
 
@@
 
-We let $X_t$ denote the size (numbers or biomass) of the focal
+We let $x(t)$ denote the size (numbers or biomass) of the focal
 population at time $t$ and assume that in the absence of take its
-dynamics are.
-
-\begin{equation}
-X_{t+1}= Z_t f(X_t) \label{eq1} 
+dynamics are. (Throughout, we will use $X(t+1)$ with a capital $X$ as a
+reminder that this is a stochastic, uncertainty quantity at time $t$,
+while we use lowercase $x(t)$ as a reminder that the current population
+size is treated as a given without error. The latter assumption is
+straight-forward to relax but will simplify the presentation here).
+
+\begin{equation}
+X(t+1) = Z(t) f(x(t), p) \label{eq1} 
 \end{equation}
@@
 and $p$ is a vector of parameters to be estimated from the data. We
-describe three choices for $f(X_t,p)$ in the next section.In this simple
-model, the control action is a harvest or take, $h_t$, measured in the
-same units as $X$, at time $t$. Thus, in the presence of take, the
-population size on the right hand side of Eqn 1 is replaced by
-$S_t=X_t-h_t$.
-
-To construct the value function, we consider a return when $X_t=x$ and
-harvest $h_t=h$ denoted as the reward, $R(x,h)$. For example, if the
-return is simply the harvest at time $t$, then $R(x,h)=min(x,h)$. We
-assume that future harvests are discounted relative to current ones at a
-constant rate of discount $\delta$ and ask for the harvest policy that
-maximizes total discounted harvest between the current time $t$ and a
-final time $T$. That is, we seek to maximize over choices of harvest
-$\mathbf{E}_{X_t+1} [ \sum_{t = 0}^{T}  R_t(X_t, h_t) \delta^t]$, where
-the state dynamics are given by Eqn 1 and $\mathbf{E}$ denotes the
+describe three choices for $f(X(t),p)$ in the next section.In this
+simple model, the control action is a harvest or take, $h(t)$, measured
+in the same units as $X$, at time $t$. Thus, in the presence of take,
+the population size on the right hand side of Eqn 1 is replaced by
+$S(t) = X(t) - h(t)$.
+
+To construct the value function, we consider a return when $X(t) = x $
+and harvest $h(t) = h$ denoted as the reward, $R(x, h)$. For example, if
+the return is simply the harvest at time $t$, then
+$R(x, h) = \min(x, h)$. We assume that future harvests are discounted
+relative to current ones at a constant rate of discount $\delta$ and ask
+for the harvest policy that maximizes total discounted harvest between
+the current time $t$ and a final time $T$. That is, we seek to maximize
+over choices of harvest
+$\mathbf{E}_{X(t+1)} [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$,
+where the state dynamics are given by Eqn 1 and $\mathbf{E}$ denotes the
 expectation over the process stochasticity in the future population
@@
 
-In order to find that policy, we introduce the value function $V_t(x_t)$
-representing the total discounted catch from time $t$ onwards given that
-$X_t=x_t$. This value function satisfies an equation of SDP (Mangel and
-Clark 1988, Clark and Mangel 2000, Clark 2009), sometimes called the
-Hamilton-Jacobi-Bellman equation (Mangel 2014),
-
-\begin{equation}
-V_t(x_t) = \max_{h_t} \lbrace R(h_t, x_t) + \delta \cdot \mathbf{\mathrm{E}}_{X_{t+1}} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] \rbrace
+In order to find that policy, we introduce the value function
+$V(x(t), t)$ representing the total discounted catch from time $t$
+onwards given that $X(t) = x(t)$. This value function satisfies an
+equation of SDP (Mangel and Clark 1988, Clark and Mangel 2000, Clark
+2009), sometimes called the Hamilton-Jacobi-Bellman equation (Mangel
+2014),
+
+\begin{equation}
+V(x(t), t) = \max_{h(t)} \lbrace R(h(t), x(t)) + \delta \cdot \mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] \rbrace
 \end{equation}
@@
 where expectation is taken over all possible values of the next state,
-$X_{t+1}$, and maximized over all possible choices of harvest, $h_t$.
-That is, at time $t$, when population size is $x_t$ and harvest $h_t$ is
-applied, the immediate return is $R(h_t, x_t)$. In the classical case,
-the sole source of uncertainty is the process stochasticity term, $Z$,
-and thus the expectation above is equivalent to taking expectations over
-$Z$. That is
-
-\[ \mathbf{\mathrm{E}}_{X_{t+1}} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] = \mathbf{\mathrm{E}}_{Z} \left[ V_{t+1}( Z f(x_t - h_t))  | x_t, h_t \right] \]
-
-where the population size after the take is $x_t-h_t$, which is then
-translated into $X_{t+1}$ by Eqn 1 (that is, we just replaced $X_{t+1}$
-with $Z f(x_t-h_t)$).
+$X(t+1)$, and maximized over all possible choices of harvest, $h(t)$.
+That is, at time $t$, when population size is $x(t)$ and harvest $h(t)$
+is applied, the immediate return is $R(h(t), x(t))$. In the classical
+case, the sole source of uncertainty is the process stochasticity term,
+$Z$, and thus the expectation above is equivalent to taking expectations
+over $Z(t)$. That is
+
+\begin{equation}
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V( X(t+1),t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{Z(t)} \left[V( Z(t) f(x(t) - h(t))), t+1 | x(t), h(t) \right]
+\end{equation}
+
+where the population size after the take is $x(t) - h(t)$, which is then
+translated into $X(t+1)$ by Eqn 1 (that is, we just replaced $X(t+1)$
+with $Z(t) f(x(t)-h(t))$).
 
@@
 
-\[\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(x_t - h_t))  | x_t, h_t \right] \}\]
+\begin{equation}
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{p|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z(t) | p, \mathrm{data}} \left[ V(Z(t) f(x(t) - h(t)), t)  \right] \}
+\end{equation}
 
@@
 
-\[\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( x_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(x_t - h_t))  | x_t, h_t \right] \}\]
+\begin{equation}
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{p|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z(t) | p, \mathrm{data}} \left[ V( Z(t) f(x(t) - h(t)), t) \right] \}
+\end{equation}
 
@@
 Myers model (Myers et al. 1995), Eqns \eqref{ricker}-\eqref{myers},
-which we parameterize as follows for $f(S_t,p)$ in Eqn 1. In all three,
+which we parameterize as follows for $f(S(t),p)$ in Eqn 1. In all three,
 we let $K$ denote the carrying capacity and $r$ the maximum per capita
@@
 
-We assume that the data $X_o$ are observed with some process
+We assume that the data $X(t)$ are observed with some process
 stochasticity,
 
-\[X_{t+1} = f(X_t) + \varepsilon,\]
+\begin{equation}
+X(t+1) = g(X(t)) + \varepsilon,
+\end{equation}
 
@@
 
-Then under the GP, given a vector of observed data $X_o$ the predicted
-data $X_p$ obeys
-
-\[f(X_p|X_o, X_{t+1}) \sim \mathcal{N}(E,C)\]
-\[E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)^{-1} X_{t+1}\]
-\[C = K(X_p, X_p) - K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)^{-1} K(X_o, X_p)\]
-
-Where $K$ is the Gaussian process kernel. (Compare to Eqn 2.22 of
-Rasmussen and Williams (2006) for a more detailed derivation.)
-Throughout, we use the radial basis function kernel:
-
-\[ K(\mathbf{x},\mathbf{y}) = \exp\left(\frac{-|\mathbf{x}-\mathbf{y}|^2}{2 \ell^2} \right)\]
+We consider the fase where we have some previous observations of the
+\emph{timeseries}, $X_{\textrm{obs}}(t)$, and wish to \emph{predict}
+future values $X_{\textrm{p}}(t)$. We consider the observations as pairs
+of points $X_{\textrm{obs}}(t), X_{\textrm{obs}}(t+1)$ for
+$t \in [1,2, ..., T]$ up through observed time $T$. Under the GP, a
+finite sample of these predictions at $n$ points into the future,
+$X_{\textrm p}(\tau)$ for $\tau \in [T+1, T+2, ..., T+n]$, is simply
+given by a multivariate normal ($\mathcal{N}$):
+
+\begin{equation}
+g(X_p|X_{\textrm{obs}}(t), X_{\textrm{obs}}(t+1)) \sim \mathcal{N}(E,C)
+\end{equation}
+
+which has mean and covariance matrix given by:
+
+\begin{equation}
+E = K(X_{\textrm p}(\tau), X_{\textrm{obs}}(t)) \left(K(X_{\textrm{obs}}(t),X_{\textrm{obs}}(t)) - \sigma \mathbb{I} \right)^{-1} X_{\textrm{obs}}(t+1)
+\end{equation}
+
+\begin{equation}
+C = K(X_{\textrm p}(\tau), X_{\textrm p}(\tau)) - K(X_{\textrm p}(\tau), X_{\textrm{obs}}(t)) \left(K(X_{\textrm{obs}}(t),X_{\textrm{obs}}(t)) - \sigma \mathbb{I} \right)^{-1} K(X_{\textrm{obs}}(t), X_{\textrm p}(\tau))
+\end{equation}
+
+Where $K$ is the chosen covariance kernel. (Compare to Eqn 2.22 of
+Rasmussen and Williams (2006) for a more detailed derivation.) Many
+options are possible for this kernel. Throughout, we use the radial
+basis function for simplicity, a matrix whose $(i,j)$th element is given
+by:
+
+\begin{equation}
+K_{i,j}(x_i, y_j) = \exp\left(\frac{-|\mathbf{x_i}-\mathbf{j}|^2}{2 \ell^2} \right)
+\end{equation}
 
@@
 
-\[f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)\]
+\begin{equation}
+f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)
+\end{equation}
 
@@
 between two points in state-space decays, and $\sigma$, which gives the
-scale of the process noise by which observations $Y_{t+1}$ may differ
-from their predicted values $X_{t+1}$ given an observation of the
-previous state, $X_t$. Munch et al. (2005a) give an accessible
-introduction to the use of Gaussian processes in providing a Bayesian
-nonparametric description of the stock-recruitment relationship.
+scale of the process noise by which observations $Y(t+1)$ may differ
+from their predicted values $X(t+1)$ given an observation of the
+previous state, $x(t)$ (see Munch et al. 2005a).
 
@@
 estimated GP by inserting it into a SDP algorithm. The only difference
-is that the uncertainty in the future state under the GP, $X_{t+1}$,
+is that the uncertainty in the future state under the GP, $X(t+1)$,
 includes both process uncertainty (based on the estimation of $\sigma$)
@@
 transition matrix representing the probability of going to each state
-$X_{t+1}$ given any current state $X_t$ and any harvest $h_t$ (See the
+$X(t+1)$ given any current state $x(t)$ and any harvest $h(t)$ (See the
 function \texttt{gp\_transition\_matrix()} in the provided R package).
@@
 \addlinespace
-\caption{Model selection scores for several common criteria all
-(wrongly) select the simplest model. As the true (Allen) model is not
-distinguishable from the simpler (Ricker) model in the region of the
-observed data, this error cannot be avoided regardless of the model
-choice criterion. This highlights the danger of model choice when the
-selected model will be used outside of the observed range of the data.}
+\caption{Model selection scores for several common criteria (DIC:
+Deviance Information Criterion, AIC: Akaike Information Criterion, BIC:
+Bayesian Information Criterion) all (wrongly) select the simplest model.
+As the true (Allen) model is not distinguishable from the simpler
+(Ricker) model in the region of the observed data, this error cannot be
+avoided regardless of the model choice criterion. This highlights the
+danger of model choice when the selected model will be used outside of
+the observed range of the data.}
 \end{longtable}
@@
 under each model. Policies are shown in terms of target escapement,
-$S_t$, as under models such as this a constant escapement policy is
+$S(t)$, as under models such as this a constant escapement policy is
 expected to be optimal (Reed 1979).}
@@
 policies inferred under each model differ widely (Figure 3). Policies
-are shown in terms of target escapement, $S_t$. Under models such as
-this a constant escapement policy is expected to be optimal (Reed 1979),
-whereby population levels below a certain size $S$ are unharvested,
-while above that size the harvest strategy aims to return the population
-to $S$. Whenever a model predicts that the population will not persist
-below a certain threshold, the optimal solution is to harvest the entire
-population immediately, resulting in an escapement $S=0$, as seen in the
-true (correct form, exact parameters) model, the Allen model (correct
-form, estimated parameters) and the GP.\\Only the structurally correct
-model (Allen model) and the GP produce policies close to the true
-optimum policy.
+are shown in terms of target escapement, $S(t) = x_t - h$. Under models
+such as this a constant escapement policy is expected to be optimal
+(Reed 1979), whereby population levels below a certain size $S$ are
+unharvested, while above that size the harvest strategy aims to return
+the population to $S$. Whenever a model predicts that the population
+will not persist below a certain threshold, the optimal solution is to
+harvest the entire population immediately, resulting in an escapement
+$S=0$, as seen in the true (correct form, exact parameters) model, the
+Allen model (correct form, estimated parameters) and the GP.\\Only the
+structurally correct model (Allen model) and the GP produce policies
+close to the true optimum policy.
 
@@
 parametric models. We show 100 replicate simulations of the stock
-dynamics (eq 1) under the policies derived from each of the estimated
+dynamics (Eqn 1) under the policies derived from each of the estimated
 models, as well as the policy based on the exact underlying model.}
@@
 
-\[P(x_t < x_c | t \in T) \propto  \exp\left(\frac{L^2}{\sigma^2}\right),\]
+\[P(x(t) < x_c | t \in T) \propto  \exp\left(\frac{L^2}{\sigma^2}\right),\]
 
@@
 positively misleading -- supporting simpler models that cannot express
-tipping point dynamics merely on account of them being similar. As the
-data shown comes only from the basin of attraction near the unfished
+tipping point dynamics merely on account of them being similar. The data
+shown comes only from the basin of attraction near the unfished
 equilibrium, near which all of the models are approximately linear and
-approximately identical.
-
-Model selection criteria trade off model complexity and fit to the data.
-When the data come from a limited region of state-space -- as is
-necessarily the case whenever there is a potential concern about tipping
-point dynamics -- simpler models therefore fit just as well and tend to
-be selected more often than complex ones. This approach would be
-appropriate when the dynamics can be expected to remain in the region of
-the training data; for instance, if we only considered the forecasting
-accuracy of the unfished population dynamics under each model.
-
-In contrast, the decision-maker's problem of setting appropriate harvest
-levels cannot exclude regions of state-space outside the observed range
-when integrating over all possible decisions to find the optimal choice.
+approximately identical; so it is intuitive that all criteria choose the
+simplest model. In a complex world, the result is suboptimal. But in a
+world that might contain tipping points, the result is disastrous.
+
+Many managers both in fisheries and beyond face a similar situation:
+they have not observed the population dynamics at all possible
+densities. A lack of comprehensive data at all population sizes,
+combined with our obvious inability to formulate accurate population
+models for low population sizes in the absence of any data, makes this
+situation the rule more than the exception. Relying on parametric models
+and model choice processes that favor simplicity ignores this basic
+fact. Acknowledging this, Carl Walters (e.g. Walters and Hilborn 1978)
+has long observed that if we began by fishing any newly exploited
+population down to very low levels and then let it recover, we would be
+much better at estimating population dynamics and thus predicting the
+optimal harvest levels. While certainly true, this presents a rather
+risky policy in the face of potential tipping points. The GPDP offers an
+alternative way to acknowledge this uncertainty.
+
+\subsubsection{GPDP population dynamics capture larger uncertainty in
+regions where the data are
+poor}\label{gpdp-population-dynamics-capture-larger-uncertainty-in-regions-where-the-data-are-poor}
+
+The parametric models perform worst when they propose a management
+strategy outside the range of the observed data. The non-parametric
+approach of the GPDP, in contrast, allows a predictive model that
+expresses a great deal of uncertainty about the probable dynamics
+\emph{outside} the observed range, while retaining very good predictive
+accuracy \emph{inside} the range observed. The management policy
+dictated by the GPDP balances this uncertainty against the immediate
+value of the harvest, and acts to stabilize the population dynamics in a
+region of state space in which the predictions can be reliably reflected
+by the data.
+
 Such problems are not constrained to fisheries management but ubiquitous
@@
 or the loss of habitat.
-
-\subsubsection{GPDP population dynamics capture larger uncertainty in
-regions where the data are
-poor}\label{gpdp-population-dynamics-capture-larger-uncertainty-in-regions-where-the-data-are-poor}
-
-The parametric models perform worst when they propose a management
-strategy outside the range of the observed data. The non-parametric
-Bayesian approach, in contrast, allows a predictive model that expresses
-a great deal of uncertainty about the probable dynamics outside the
-observed range, while retaining very good predictive accuracy in the
-range observed. The management policy dictated by the GP balances this
-uncertainty against the immediate value of the harvest, and acts to
-stabilize the population dynamics in a region of state space in which
-the predictions can be reliably reflected by the data.
 

### 2014-11-07 manuscripts/manuscript.Rmd
update data file names to match figures

@@
 management of human intervention in natural resources and conservation.
-Ecological dynamics are frequently complex and difficult to measure,
-making uncertainty in our understanding a persistent challenge.
 Decision-theoretic approaches provide a framework to determine the
@@
 uncertainty can be meaningfully quantified [@Fischer2009].  Over the last
-four decades (but stretching even further back in time), marked by  the
-seminal contributions of @Clark1976, @Clark2009 and @Walters1978,
+four decades (beginning with
+@Clark1976, @Clark2009 and @Walters1978)
 dynamic optimization methods, particularly Stochastic Dynamic Programming
@@
 Simultaneously, there has been increasing recognition of the importance of
-Allee population sizes or `tipping points' [@Scheffer2001;
-@Polasky2011] that management actions may inadvertently
-cause populations to cross.  @Scheffer2009 emphasized the difficulties of
-formulating even qualitatively correct models of the underlying processes.
+multiple steady states or `tipping points' [@Scheffer2001; .  @Scheffer2009;
+@Polasky2011] in ecological systems.
 
 We develop a novel approach to address these concerns in the context of
-fisheries; though the underlying challenges and methods are germane to
-many other problems of conservation or natural resource exploitation,
-such as forestry.  Economic value and ecological concern have made
-marine fisheries the crucible for much of the founding work [@Gordon1954;
-@Clark1976; -@Clark2009; @Reed1979; @May1979; @Ludwig1982] on management
-under uncertainty.  Global declines [@Worm2006] and the controversy
-surrounding their interpretation [@Hilborn2007; @Worm2009] make
-understanding these challenges all the more pressing.
+fisheries; although the challenges and methods are germane to
+other problems of conservation or natural resource exploitation.  Economic value and ecological concern have made
+marine fisheries the crucible for much of the founding work on management
+under uncertainty [@Gordon1954;
+@Clark1976; -@Clark2009; @Reed1979; @May1979; @Ludwig1982]
 
@@
 
-Uncertainty enters the decision-making process at many levels:
-intrinsic stochasticity in biological processes, measurements, and
+Even if we know the proper deterministic description of the biological system, there is intrinsic stochasticity in biological dynamics, measurements, and
 implementation of policy [_e.g._ @Reed1979; @Clark1986; @Roughgarden1996;
-@Sethi2005], parametric uncertainty [_e.g._ @Ludwig1982; @Hilborn1997;
-@McAllister1998; @Schapaugh2013], and model or structural uncertainty
-[_e.g._ @Williams2001; @Cressie2009;  @Athanassoglou2012].  Of these,
-structural uncertainty is generally the hardest to quantify. Typical approaches
-assume a weak notion of model uncertainty in which the correct model
-(or reasonable approximation) of the dynamics must be identified from
-among a handful of alternative models, using either model choice, model averaging, or 
-introducing yet greater model complexity of which others may be special
-cases (model averaging being one such way to construct such a model)
+@Sethi2005].  We may also lack knowledge about the parameters of the biological dynamics [parametric uncertainty, _e.g._ @Ludwig1982; @Hilborn1997;
+@McAllister1998; @Schapaugh2013], or even not know which model is  proper description of the system 
+[structural uncertainty, _e.g._ @Williams2001; @Cressie2009;  @Athanassoglou2012].  Of these,
+the latter is generally the hardest to quantify. Typical approaches
+confront the data with a collection of models, assuming that the true dynamics (or reasonable approximation) is among the collection and then use model choice or model averaging to arrive at a conclusion
 [@Williams2001; @Athanassoglou2012; @Cressie2009].  Even setting aside
-other computational and statistical concerns (e.g. @Cressie2009), these
-approaches do not address our second concern - representing uncertainty
+other concerns (see @Cressie2009), these
+approaches are unable to describe uncertainty
 outside the observed data range.
@@
 
-Model uncertainty is particularly insidious when model predictions must
-be made outside of the range of data on which the model was estimated, and models may be abused [@Mangel2001].
-Extrapolation uncertainty is felt most keenly in decision-theoretic
-(or optimal control) applications, since (a) exploring the space of 
-all possible actions involves considering actions that may move the
-system outside the range of observed behavior, and (b) decision-theoretic
-algorithms rely not only on reasonable estimates of the expected
-outcomes, but depend on the weights given to all possible outcomes
-[_e.g._ @Weitzman2013], hence estimates of the uncertainty can be as
-important as estimates of the average or expected outcome.
-
-Concerns about tipping points in ecological dynamics [@Scheffer2001;
-@Polasky2011] highlight this problem because they arise from this very
-situation.  In such cases, precise models are not available and data is
-limited or biased to a particular region of state space, such as high
-stock levels or an otherwise desirable state for the ecosystem. If such
-models or data were available, one would know just how far a system could
-be pushed before going over the tipping point, and management would be
-simple.  The fact that this concern is frequently discussed highlights
-how often we have only limited data and inperfect models. Moreover,
-the case of tipping points highlights how even small errors can have
-very large consequences, as we shall illustrate later.
-
-The dual concerns of model uncertainty and incomplete data coverage
-pose a substantial challenge to existing decision-theoretic approaches
-[@Brozovic2011].  Because intervention is often (but not always, see
+Structural  uncertainty is particularly insidious when we try to predict
+outside of the range of observed data [@Mangel2001] because we are extrapolating into unknown regions.
+In management applications, this extrapolation uncertainty is particularly important since (a) management involves considering actions that may move the
+system outside the range of observed behavior, and (b) the decision tools (optimal control theory, SDP) rely on both reasonable estimates of the expected
+outcomes and on the weights given to those outcomes
+[_e.g._ @Weitzman2013].   Thus characterizing uncertainty is as
+important as characterizing the expected outcome.
+
+Tipping points in ecological dynamics [@Scheffer2001;
+@Polasky2011] highlight this problem because 
+precise models are not available and data are
+limited such as around  high
+stock levels or an otherwise desirable state. With perfect information
+, one would know just how far a system could
+be pushed before crossing the tipping point, and management would be
+simple.  But we face imperfect models and limited data and, with tipping points,  
+even small errors can have
+very large consequences, as we shall illustrate later.  Because intervention is often (but not always, see
 @Hughes2013) too late after a tipping point has been crossed, management
-is most often concerned with avoiding potentially catastrophic
-tipping points before any data are available at or following a
-transition that would more clearly reveal these regime shift dynamics
+is often concerned with avoiding 
+tipping points before any data about them is available or the 
+transition that would more clearly reveal a tipping point occurs
 [e.g. @Bestelmeyer2012].
 
-We illustrate how a Stochastic Dynamic Programming (SDP) algorithm
-[@Mangel1988; @Marescot2013] can be driven by the predictions from a
+The dual concerns of model uncertainty and incomplete data 
+create a substantial challenge to existing decision-theoretic approaches
+[@Brozovic2011].   We illustrate how  Stochastic Dynamic Programming (SDP) 
+[@Mangel1988; @Marescot2013] can be implemented using a
 Bayesian Non-Parametric (BNP) model of population dynamics [@Munch2005a].
-This provides two distinct advantages compared with contemporary
-approaches.  First, using a BNP sidesteps the need for an accurate
-model-based description of the system dynamics.  Second, the BNP can
-better reflect uncertainty that arises when extrapolating a model outside
-of the data on which it was fit.  We illustrate that when the correct
-model is not known, this latter feature is crucial to providing a robust
-decision-theoretic approach in face of substantial structural uncertainty.
-We use _robust_ to distinguish approaches that provide nearly optimal solutions
-(as measured by the chosen value function) without being sensitive to the
-choice of the (unknown) underlying model.
-
-<!--
-FIXME We need to be 
--->
-
-This paper represents the first ecological application of the
-SDP decision-making framework without an a priori model of the
-underlying dynamics.  In contrast to parametric models which can only
-reflect uncertainty in parameter estimates, the BNP approach provides a
-state-space dependent representation of uncertainty.  This permits a much
-greater uncertainty far from the observed data than near the observed
-data.  These features allow the Gaussian Process Dynamic Programming
-(GPDP) approach to find robust management solutions in face of limited
-data and without knowledge of the correct model structure.
+The BNP method has two distinct advantages 
+.  First, using a BNP model sidesteps the need for an accurate
+model-based description of the system dynamics.  Second, a BNP model
+better reflects uncertainty when extrapolating beyond the observed
+data.  This is crucial to providing  robust
+decision-making when the correct model is not known (as is almost always true).
+[We use _robust_ to characterize approaches that provide nearly optimal solutions
+without being sensitive to the choice of the (unknown) underlying model.]
+
+This paper is the first ecological application of the
+SDP without an _a priori_  model of the
+underlying dynamics.  Unlike parametric approaches that can only
+reflect uncertainty in parameter estimates, the BNP method provides a
+broader representation of uncertainty, including uncertainty beyond the observed data. 
+We will show that Gaussian Process Dynamic Programming
+(GPDP)  allows us to find robust management solutions in face of limited
+data without knowing  the correct model structure.
 
@@
 
-To illustrate GPDP, we consider the performance of the GPDP policy against
-the policies derived under several alternative parametric models [@Reed1979; @Ludwig1982; @Mangel1988]. 
-The nature of decision-making problems provides a compelling and
-pragmatic way to compare models: Rather than compare models in terms
-of best fit to data, we define model performance in the concrete terms
+For comparisons, we consider the performance of management based on  GPDP against
+management policies derived under several alternative parametric models [@Reed1979; @Ludwig1982; @Mangel1988].  Rather than compare models in terms
+of best fit to data, we compare model performance in the concrete terms
 of the decision-maker's objectives.
@@
 
-[^1]: Note that we abbreviate Gaussian Process as GP, which refers only to 
-the statistical model we use to approximate the population dynamics (as opposed
-to a familiar parametric model such as the Ricker equation), while we use the
-term GPDP, Gaussian Process Dynamic Programming, to refer to the _use_ of a
-GP as the underlying process model when solving a Dynamic Program, as we illustrate
-in this paper.  Hence we will refer to the models as: GP, Ricker, Allen, etc, while
-we refer to the novel method we put forward here as GPDP.
+[^1]: We abbreviate Gaussian Process as GP, which refers to 
+the statistical model we use to approximate the population dynamics
+, and we use the
+term Gaussian Process Dynamic Programming [GPDP,], to refer to the _use_ of a
+GP as the underlying process model when solving a Dynamic Programming equation
+.  Hence we will refer to the models as: GP, Ricker, Allen, etc, and the novel method we put forward here as GPDP.
 
@@
 variable (or variables), a control action, and a value function.
-In this paper, we consider only a single state variable, so that the
-ideas can be described as simply as possible. That is, we focus on a
-class of one-dimensional models of population dynamics as a 'best-case'
-scenario for the parametric model performance relative to the GP.
-This is a best-case scenario for the parametric models because the
-underlying dynamics are always simulated from one of the parametric models
-in the set, whereas in reality we never have the "true" model among our
-candidates.  By choosing one-dimensional models with few parameters, we
-limit the chance that poor performance will be due to our inability to
-estimate parameters well, something that becomes a more severe problem
-for higher-dimensional parametric models. Further, the parametric models
+For simplicity, we consider only a single state variable, which allows us to
+focus on one-dimensional models of population dynamics 
+for the comparison of parametric models and the GP.
+This is a best-case scenario for the parametric models because we
+simulate underlying dynamics from one of the three parametric models
+, whereas in the natural world we never know the "true" mode
+.  In addition, by choosing one-dimensional models with just a few parameters, we
+limit the chance that poor performance will be due to inability to
+estimate parameters accurately, something that becomes a more severe problem
+for higher-dimensional parametric models. Finally, the parametric models
 we consider are those most commonly used in modeling stock-recruitment
@@
 
-We let $x(t)$ denote the size (numbers or biomass) of the focal
+We let $X(t)$ denote the size (numbers or biomass) of the focal
 population at time $t$ and assume that in the absence of take its
-dynamics are. (Throughout, we will use $X(t+1)$ with a capital $X$ as
-a reminder that this is a stochastic, uncertainty quantity at time $t$,
-while we use lowercase $x(t)$ as a reminder that the current population
-size is treated as a given without error.  The latter assumption is
-straight-forward to relax but will simplify the presentation here).
+dynamics are.
+.
 
 \begin{equation}
-X(t+1) = Z(t) f(x(t), p) \label{eq1} 
+X(t+1) = Z(t) f(X(t), p) \label{eq1} 
 \end{equation}
@@
 and $p$ is a vector of parameters to be estimated from the data.
-We describe three choices for $f(X(t),p)$ in the next section.In this
-simple model, the control action is a harvest or take, $h(t)$, measured
+We describe the three choices for $f(X(t),p)$ in the next section.  
+The control action is a harvest or take, $h(t)$, measured
 in the same units as $X$, at time $t$.  Thus, in the presence of take,
@@
 
-To construct the value function, we consider a return when $X(t) = x $ and
-harvest $h(t) = h$ denoted as the reward, $R(x, h)$. For example, if the
-return is simply the harvest at time $t$, then $R(x, h) = \min(x, h)$. We
+To construct the value function, we consider a return when $X(t) = x(t)$ and
+harvest $h(t) = h$ denoted as the reward, $R(x(t), h)$. For example, if the
+return is the harvest at time $t$, then $R(x(t), h(t)) = \min(x(t), h(t))$. We
 assume that future harvests are discounted relative to current ones at
@@
 and a final time $T$. That is, we seek to maximize over choices of
-harvest $\mathbf{E}_{X(t+1)} [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$,
-where the state dynamics are given by Eqn 1 and $\mathbf{E}$ denotes the
+harvest $E_{X(t+1)} [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$,
+where the state dynamics are given by Eqn 1 and $E$ denotes the
 expectation over the process stochasticity in the future population state.
@@
 given that $X(t) = x(t)$.  This value function satisfies an equation of
-SDP  [@Mangel1988; @Clark2000; @Clark2009], sometimes called the
-Hamilton-Jacobi-Bellman equation [@Mangel2014],
+SDP  [@Mangel1988; @Clark2000; @Clark2009 ;@Mangel2014],
 
@@
 That is, at time $t$, when population size is $x(t)$ and harvest $h(t)$ is
-applied, the immediate return is $R(h(t), x(t))$. In the classical case,
+applied, the immediate return is $R(h(t), x(t))$. When
 the sole source of uncertainty is the process stochasticity term, $Z$,
-and thus the expectation above is equivalent to taking expectations over
+and thus the expectation in Eqn 2 is equivalent to taking expectations over
 $Z(t)$. That is
@@
 \begin{equation}
-\mathbf{\mathrm{E}}_{X(t+1)} \left[ V( X(t+1),t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{Z(t)} \left[V( Z(t) f(x(t) - h(t))), t+1 | x(t), h(t) \right]
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V( X(t+1),t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{Z(t)} \left[V( Z(t) f(x(t) - h(t))|p), t+1 | x(t), h(t) \right]
 \end{equation}
@@
 where the population size after the take is $x(t) - h(t)$, which is then
-translated into $X(t+1)$ by Eqn 1 (that is, we just replaced $X(t+1)$ with $Z(t) f(x(t)-h(t))$).
-
-In a Bayesian decision framework, the parameters governing the dynamics
-are also uncertain, so the expectation involves averaging over the
-posterior distribution for the parameters, as well. That is,
+translated into $X(t+1)$ by Eqn 1 (that is, we replace $X(t+1)$ by $Z(t) f(x(t)-h(t)|p)$).
+
+When the parameters governing the dynamics
+are also uncertain (the case of parametric uncertainty), the expectation in Eqn 2 involves averaging over the
+posterior distribution for the parameters. That is,
 
 \begin{equation}
-\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{p|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z(t) | p, \mathrm{data}} \left[ V(Z(t) f(x(t) - h(t)), t)  \right] \}
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{p|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z(t) | p, \mathrm{data}} \left[ V(Z(t) f(x(t) - h(t)|p), t)  \right] \}
 \end{equation}
 
-In the nonparametric case, the function $f$ too is uncertain and the expectation for the next state includes
+When the underlying population dynamics are unknown (the case of structural uncertainty), the function $f$ itself is uncertain and the expectation for the next state includes
 uncertainty in $f$ as well. That is
@@
 \begin{equation}
-\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{p|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z(t) | p, \mathrm{data}} \left[ V( Z(t) f(x(t) - h(t)), t) \right] \}
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{p|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z(t) | p, \mathrm{data}} \left[ V( Z(t) f(x(t) - h(t)|p), t) \right] \}
 \end{equation}
@@
 
-We consider three candidate parametric models of the stock-recruitment
-dynamics: The Ricker model, the Allen model [@Allen2005], the Myers
-model [@Myers1995], Eqns \eqref{ricker}-\eqref{myers}, which we parameterize as follows for $f(S(t),p)$ in Eqn 1.
+We consider three candidate parametric models for the population dynamics: The Ricker model, the Allen model [@Allen2005], and the Myers
+model [@Myers1995], Eqns \eqref{ricker}-\eqref{myers}.
 In all three, we let $K$ denote the carrying capacity and $r$ the maximum per capita growth rate.
-In the Ricker model,
+The Ricker model has two parameters and the right hand side of Eqn 1 is
 
 \begin{equation}
-f(S(t)|p) = S(t) e^{r \left(1 - \frac{S(t)}{K} \right) } \label{ricker}
+f(S(t)|r,K) = S(t) e^{r \left(1 - \frac{S(t)}{K} \right) } \label{ricker}
 \end{equation}
 
-In the Allen model,
+The Allen model has three parameters
 
 \begin{equation}
-f(S(t)|p) = S(t) e^{r \left(1 - \frac{S(t)}{K}\right)\left(S(t) - X_c\right)} \label{allen}
+f(S(t)|r, K, X_c) = S(t) e^{r \left(1 - \frac{S(t)}{K}\right)\left(S(t) - X_c\right)} \label{allen}
 \end{equation}
 
-Where $r$ and $K$ are as before, and $X_c$ denotes the location of the unstable steady state (i.e., the tipping point).
-
-In the Myers model,
+where $X_c$ denotes the location of the unstable steady state (i.e., the tipping point).
+
+The Myers model also has three parameters
 
 \begin{equation}
-f(S(t) | p)  = \frac{r S(t)^{\theta}}{1 + \frac{S(t)^\theta}{K}} \label{myers}
+f(S(t) | r, K, \theta)  = \frac{r S(t)^{\theta}}{1 + \frac{S(t)^\theta}{K}} \label{myers}
 \end{equation}
 
-Where $\theta = 1$ corresponding to Beverton-Holt dynamics and $\theta$ > 2 leads to Allee effects and multiple stable states.
-
-The Ricker model involves two parameters, corresponding
-to a growth rate and a carrying capacity, and cannot support alternative
-stable growth rate dynamics. The Allen model resembles the Ricker
+where $\theta = 1$ corresponds to Beverton-Holt dynamics and $\theta$ > 2 leads to Allee effects and multiple stable states.
+
+The Ricker model does not lead to multiple steady states.
+The Allen model resembles the Ricker
 dynamics with an added Allee effect parameter [@Courchamp2008], below
 which the population cannot persist.  The Myers model also has three
-parameters and contains an Allee threshold, but has compensatory rather
-than over-compensatory density dependence (resembling a Beverton-Holt
-curve rather than a Ricker curve at high densities.)
-
-The multiplicative log-normal stochasticity perturbs the growth predicted 
-by each of these deterministic model skeletons. This 
+parameters and contains an Allee threshold, but unlike the Ricker model
+saturates  at high population size.  The multiplicative log-normal stochasticity in Eqn 1
 introduces one additional parameter $\sigma$ that must be estimated.
@@
 tipping points, all of our simulations are based on the Allen model.
-The Allen model is thus "structurally correct" and is expected to provide
-a best-case scenario when the 'true' dynamics are known.
-The Ricker model is a reasonable approximation of these dynamics far from the Allee
-threshold (but lacks threshold dynamics), while the Myers model shares
-the essential feature of a threshold but differs in the structure.
-
-Throughout the presentation, we will refer to the use of the "True" model
-to denote solving for the optimal solution when the underlying parameters
-_are known without error_, and refer to the "Allen" model to denote the
-when these parameters have instead only been estimated from the sample data.
-
-We consider a period of `r Tobs` years of training data: long
+The Allen model is thus the  state of nature  and is expected to provide
+the best-case scenario. The Ricker model is a reasonable approximation of these dynamics far from the Allee threshold (but lacks threshold dynamics), while the Myers model shares
+the essential feature of a threshold but differs in structure from the Allen model.
+Throughout, we refer to the "True" model when the underlying parameters
+_are known without error_, and refer to the "Allen" model 
+when these parameters have been estimated from the sample data.
+
+We consider a period of `r Tobs` in which data are obtained to estimate the parameters or the GP. This is long
 enough that the estimates do not depend on the particular realization,
-while longer times are not likely to provide substantial improvement.  
-Each of the models is fit to the same training data (Figure 1).
+and longer times are not likely to provide substantial improvement.  
+We fit each of the models is fit to the same data (Figure 1).
 
@@
 
-The core difference for our purpose between the GP representation and the parametric
-models is that the GP model is defined explicitly in reference to 
-some observed data.  Unlike the models above, it cannot be specified by
-the value of some parameters alone.  Although called nonparametric, the
-GP (a)  still involves the estimation of
-parameters, and (b), statisticians also use non-parametric to mean models
-that make only weak distributional assumptions  [e.g. @Lehmann1975].
-However, we retain the terminology for consistency with previous studies.
+The core difference for our purpose between the GP and the parametric
+models is that the GP model is defined explicitly by the observed data.  That is, it cannot be specified by the value of parameters alone.
 
@@
 
-The use of GP methods to formulate a predictive
-model is relatively new in the context of modeling dynamical systems
+The use of GPs to characterize dynamical systems is relatively new
 [@Kocijan2005], and was first introduced in the context ecological
@@
 
-A GP is an infinite dimensional (because it is a function of a continuous
-input variable) stochastic process. Any realization of the process for
-some finite sample of points $n$ is a multivariate normal distribution
+A GP is a 
+stochastic process for which any realization of for
+a  finite sample of points $n$ is a multivariate normal distribution
 of dimension $n$. Thus, to characterize the GP we need a mean function
@@
 
-We assume that the data $X(t)$ are observed with some process stochasticity,
+As before, we assume that the data $X(t)$ are observed with process stochasticity around a mean function $ g(X(t))$
 
@@
 where $\varepsilon$ are IID normal random variables with zero-mean and variance $\sigma_n^2$. Note that we have chosen to
-assume additive stochasticity.  While we could just as easily consider log-normal stochasticity
-as in the parametric models, we make this choice to bring home the point that the
-Gaussian process approach need not have structurally correct noise form either.
-
-We consider the fase where we have some previous observations of the _timeseries_, $X_{\textrm{obs}}(t)$,
+assume additive stochasticity.  While we could consider log-normal stochasticity
+as in the parametric models, we make this choice to emphasize that the
+Gaussian process approach need not have structurally correct stochasticity to be effective.
+
+Now consider the situation where we have a set of  previous observations of the _timeseries_, $X_{\textrm{obs}}(t)$,
 and wish to _predict_ future values $X_{\textrm{p}}(t)$.  We consider the observations as pairs of points
-$X_{\textrm{obs}}(t), X_{\textrm{obs}}(t+1)$ for $t \in [1,2, ..., T]$ up through observed time $T$.
-Under the GP, a finite sample of these predictions at $n$ points into the future, $X_{\textrm p}(\tau)$ for $\tau \in [T+1, T+2, ..., T+n]$, is simply given by a multivariate normal ($\mathcal{N}$):
+$X_{\textrm{obs}}(t), X_{\textrm{obs}}(t+1)$ for $t=1, 2, …$ up through final observation time $T$.
+As described above, in the  GP, a finite sample of these predictions at $n$ points into the future, $X_{\textrm p}(\tau)$ for $\tau \in [T+1, T+2, ..., T+n]$, is a multivariate normal ($\mathcal{N}$):
 
@@
 
-which has mean and covariance matrix given by:
+with mean and covariance matrix given by (see Eqn 2.22 of @Rasmussen2006 for a more detailed derivation)
 
@@
 
-Where $K$ is the chosen covariance kernel.
-(Compare to  Eqn 2.22 of @Rasmussen2006 for a more detailed derivation.) Many options are possible for this kernel.
-Throughout, we use the radial basis function for simplicity, a matrix whose $(i,j)$th element is given by:
+Where $K$, called the covariance kernel, measures how much influence one observation has on another.  In this paper, for $K$ use a matrix whose $(i,j)$th element is given by:
 
 \begin{equation}
-K_{i,j}(x_i, y_j) = \exp\left(\frac{-|\mathbf{x_i}-\mathbf{j}|^2}{2 \ell^2} \right)
+K_{i,j}(x_i, y_j) = \exp\left( \frac{ -(x_i - y_j)^2}{2 \ell^2} \right)
 \end{equation}
-
-We use inverse Gamma priors on both the length-scale $\ell$ and $\sigma$, of the form
+so that $\ell$ gives the characteristic length-scale over which correlation between
+two observations decays.
+
+We use inverse Gamma priors on both the length-scale $\ell$ and $\sigma$, thus for example
 
 \begin{equation}
-f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)
+f(\ell; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} \ell^{-\alpha - 1}\exp\left(-\frac{\beta}{\ell}\right)
 \end{equation}
 
-For the $\sigma$ prior, $\alpha = $ `r s2.p[1]` and $\beta = $ `r s2.p[2]`.  For $\ell$ prior,  $\alpha = $ `r d.p[1]` and $\beta = $ `r d.p[2]`.
-
-Using the simulated training data we also estimate a Gaussian process
-defined by a radial basis function kernel of two parameters: $\ell$,
-which gives the characteristic length-scale over which correlation between
-two points in state-space decays, and $\sigma$, which gives the scale
-of the process noise by which observations $Y(t+1)$ may differ from
-their predicted values $X(t+1)$ given an observation of the previous
-state, $x(t)$ [see @Munch2005a].
+For the prior on  $\ell$,  $\alpha = $ `r d.p[1]` and $\beta = $ `r d.p[2]`.  The prior  on $\sigma$,  $\alpha = $ `r s2.p[1]` and $\beta = $ `r s2.p[2]`.
 
@@
 We use a Metropolis-Hastings Markov Chain Monte Carlo (@Gelman2003) to
-infer posterior distributions of the two parameters of the GP (Figure
+infer posterior distributions of the parameters of the GP (Figure
 S3, code in appendix). Since the posterior distributions differ
-substantially from the priors (Figure S3), we can be assured that most
-of the information in the posterior comes from the data rather than the
+substantially from the priors (Figure S3), most of the information in the posterior comes from the data rather than the
 prior belief.
@@
 
-It is reasonably straight-forward to derive the harvest policy from the
-estimated GP by inserting it into a SDP algorithm. The only difference 
-is that the uncertainty in the future state under the GP, $X(t+1)$,
-includes both process uncertainty (based on the estimation of $\sigma$) 
-and "structural uncertainty" of the posterior collection of curves.  
-(Recall that given the parameters, a parametric model is represented
-by a single curve, while the GP remains a distribution of curves).  
-From the GP posteriors, we can write down the (discritized) transition
+We  derive the harvest policy from the
+estimated GP by inserting it into a SDP algorithm. Given the GP posteriors, we construct the transition
 matrix representing the probability of going to each state $X(t+1)$ 
@@
 case to determine the optimal policy.   
+In doing so, 
+the uncertainty in the future state under the GP, $X(t+1)$,
+includes both process uncertainty (based on the estimation of $\sigma$) 
+and structural uncertainty of the posterior collection of curves.  
+Unlike the case of a parametric model, which is a single curve given the parameters, 
+the GP remains a distribution of curves.
 
@@
 
-To ensure our results are robust to our choice of parameters, 
+To ensure our results are robust to the choice of parameters, we
 consider 96 different scenarios, described in detail below.   To help  better
@@
 
-All models fit the observed data rather closely and with relatively small uncertainty, as illustrated in the posterior predictive curves in Figure 1, in which we show the training data of stock sizes observed over time as points, overlaid with the step-ahead predictions of each estimated model using the parameters sampled from their posterior distributions.
-Each model manages to fit the observed data rather closely.
-Compared to the true model most estimates appear to over-fit, predicting fluctuations that are actually due purely to stochasticity.
+All of the models fit the observed data rather closely and with relatively small uncertainty.
+In Figure 1, we show the posterior predictive curves.
+The  training data of stock sizes observed over time are points, overlaid with the step-ahead predictions of each estimated model using the parameters sampled from their posterior distributions.
+Compared to the true model most estimates appear to over-fit, predicting patterns that are actually due purely to stochasticity.
 Model selection criteria (Table 1) penalize more complex models and show a preference for the simpler Ricker model over the  models with alternative stable states (Allen and Myers).
@@
 ```{r Table1, dependson="deviances", include=TRUE, results="asis", echo=FALSE}
-pandoc.table(rbind(dictable, aictable, bictable), caption="Model selection scores for several common criteria (DIC: Deviance Information Criterion, AIC: Akaike Information Criterion, BIC: Bayesian Information Criterion) all (wrongly) select the simplest model.
+pandoc.table(rbind(dictable, aictable, bictable), caption="Model selection scores for several common criteria (DIC: Deviance Information Criterion, AIC: Akaike Information Criterion, BIC: Bayesian Information Criterion) all select the wrong model.
 As the true (Allen) model is not distinguishable from the simpler (Ricker) model in the region of the observed data, this error cannot be avoided regardless of the model choice criterion.
@@
 write.csv(models_posteriors, "components/data/figure2.csv")
-write.csv(obs, "components/data/obs.csv")
-```
-
-We show the  mean inferred state space dynamics of each model relative to the
+write.csv(obs, "components/data/training_data.csv")
+```
+
+We show the  mean inferred population dynamics of each model relative to the
 true model used to generate the data in Figure 2, predicting
-the relationship between observed stock size (x-axis) to the stock
+the relationship between observed population size (x-axis) to the population
 size after recruitment the following year.  
-In addition to the raw data, Figure 2 illustrates that the GP is conditioned
+In addition to the raw data, the GP is conditioned
 on going through the point 0,0 without error.
-This conditioning on (0,0) is equivalent to making the assumption that a species hitting 0 is extinct forever & without uncertainty, despite the lack of any data in the observed sequence to justify this.
-The use of this assumption illustrates how the GP approach can capture common-sense biology without having to assume more explicit details about the dynamics at low population numbers that have never been observed.
-Not that we need need not make this assumption (perhaps there is immigration after all); and one could repeat the entire analysis without this assumption.
-
-Note that in contrast to
-the  other models shown, the expected Gaussian process corresponds to a
-distribution of curves - as indicated by the gray band - which itself
-has a mean shown in black. Parameter uncertainty (not shown) spreads
-out the estimates further.  The observed data from which each model is 
-estimated is also shown.  The observations come from only a limited region
-of state space corresponding to unharvested or weakly harvested system.
-No observations occur at the theoretical optimum harvest rate or near
-the tipping point. Figure S1 (see supplement) shows the performance of the models outside the observed training data.
+This conditioning on (0,0) is equivalent to making the assumption that the population is closed, so that once it hits 0 it stays at 0, despite the lack of any data in the observed sequence to justify this.
+This assumption illustrates how the GP can capture common-sense biology without having to assume more explicit details about the dynamics at low population numbers that have never been observed.
+If the population were not closed, one could repeat the entire analysis without this assumption.
+Unlike parametric models, the GP corresponds to a
+distribution of curves - the gray band with mean shown in black. Uncertainty in the parameters of the GP (not shown) further widens the band of possible population sizes.
+In Figure S1 (see supplement), we show the performance of the models outside the observed training data.
 
@@
 In Figure 4, we show the consequences of managing 100 replicate
-realizations of the simulated fishery under each of the policies
-estimated.  As anticipated from the policy curves, the structurally
+realizations of the simulated fishery under policies derived from each model
+.  The structurally
 correct model under-harvests, leaving the stock to vary around its
@@
 the population past the tipping point consistently, resulting in the
-immediate crash of the stock and thus derives minimal long-term catch.
-
-The results across replicate stochastic simulations can most
-easily be compared by using the relative differences in net present
-value realized by each of the model (Figure 5).  The GPDP
-most consistently realizes a value close to the optimal solution,
-and importantly avoids ever driving the system across the tipping point,
+immediate crash of the stock and thus leads to minimal long-term catch.
+
+The results across replicate stochastic simulations are most
+easily compared by using the relative differences in net present
+value realized by each of the model (Figure 5).  Although not perfect, the GPDP
+consistently realizes a value close to the optimal solution,
+and avoids ever driving the system across the tipping point,
 which results in the near-zero value cases in the parametric models.
@@
 These results are not sensitive to the modeling details of the simulation.
-The GPDP estimate remains very close to the optimal solution obtained
-by knowing the true model across changes to the training simulation,
-noise scale, parameters or structure of the underlying model. We consider
-both a Latin hypercube approach (see supplementary materials, Fig S10) and a more focused investigation of the
-parameters most responsible for impacting the transition rate: the relative 
-distance to the Allee threshold and the noise effects.
-
-Changing the intensity of the stochasticity or the distance between stable and unstable points
+The GPDP estimate remains very close to the optimal solution (obtained
+by knowing the true model) across changes to the training simulation,
+scale of stochasticity, parameters or structure of the underlying model. In the Supplement, we consider
+both a Latin hypercube approach and a more focused investigation of the effects of the relative 
+distance to the Allee threshold and the variance of process stochasticity.
+
+Changing the intensity of the stochasticity or the distance between stable and unstable steady states
 does not impact the performance of the GP relative to the optimal solution
-given the true model and true parameters.  The parametric models are
-more sensitive to this difference.  Large noise relative to the distance between the stable and unstable
+obtained from the true model and true parameters.  The parametric models are
+more sensitive to this difference.  Large values of $\sigma$ relative to the distance between the stable and unstable
 point increases the chance of a stochastic transition.  More precisely,
-if we let $L = K - x_c$, then the probability of a transition in a given
-window of time $T$ scales as
-
-$$P(x(t) < x_c | t \in T) \propto  \exp\left(\frac{L^2}{\sigma^2}\right),$$
+if we let $L$ is the distance between the stable and unstable steady states, then the probability that fluctuations drive the population across the unstable steady state scales as
+$\exp\left(\frac{L^2}{\sigma^2}\right)$
 
 (see @Gardiner2009 or @Mangel2006 for the derivation).
-As a result, the impact of using a model which underestimates the risk of harvesting
-past the critical point is much more severe, since this such a situation
+Thus, the impact of using a model that  underestimates the risk of harvesting
+beyond the critical point is considerable, since this such a situation
 occurs more often.  Conversely, with large enough distance between the
-optimal escapement and unstable points relative to the noise, the chance
+optimal escapement and unstable steady state relative to $\sigma$, the chance
 of a transition becomes vanishingly small and all models can be estimated 
-near-optimally. Models that underestimate the cost incurred by stock
+near-optimally. Models that underestimate the cost incurred by population
 sizes fluctuating significantly below the optimal escapement level will
-not perform poorly as long as those fluctuations are sufficiently rare.
+not perform poorly as long as those fluctuations are sufficiently small.
 
 The GPDP is only weakly influenced by increasing stochasticity or increasing
-Allee effects over much of the range (Figure 6). Much larger stochasticity or higher Allee
-effects make even the optimal solution without any model or parameter uncertainty
-unable to harvest the population effectively (e.g. noise large enough to violate the
-self-sustaining criterion of @Reed1979). Only at the largest Allee effect
-sizes are some replicates seen to deviate more from the optimal solution.
+Allee effects over much of the range (Figure 6). Larger $\sigma$ or higher Allee
+levels make even the optimal solution without any model or parameter uncertainty
+unable to harvest the population effectively (e.g. the stochasticity is large enough to violate the
+self-sustaining criterion of @Reed1979).
 
@@
 multiplot(plot_sigmas, plot_allees, cols=2) 
-write.csv(vary_sigma, "components/data/vary_sigma.csv")
-write.csv(vary_allee, "components/data/vary_sigma.csv")
+write.csv(vary_sigma, "components/data/figure6a.csv")
+write.csv(vary_allee, "components/data/figure6b.csv")
 ```
@@
 
-Though simple mechanistically motivated models offer the greatest
+Simple, mechanistically motivated models offer the
 potential to increase our basic understanding of ecological processes
-[@Cuddington2013; @Geritz2012], such models can be both inaccurate and
-misleading when relied upon in a quantitative decision making framework.
-In this paper we have tackled two aspects of uncertainty that are both
+[@Cuddington2013; @Geritz2012]. But such models can be both inaccurate and
+misleading when used in a decision making framework.
+In this paper we tackled two aspects of uncertainty that are 
 common to many ecological decision-making problems and fundamentally
-challenging to existing approaches which largely rely on parametric
+challenging to existing approaches that largely rely on parametric
 models:
 
-1. We do not know what the correct models are for ecological systems.
-1. We have limited data from which to estimate the model -- in particular,
-   such models may be misleading in predicting the probability of outcomes
-   outside the training data.
-
-We have illustrated how the use of non-parametric approaches can provide
+1. We do not know the correct models for ecological systems.
+2. We have limited data from which to estimate the model.
+
+We have illustrated how the use of non-parametric methods provides
 more reliable solutions in the sequential decision-making problem.
@@
 
-These results illustrate that model-choice approaches would be positively
-misleading -- supporting simpler models that cannot express tipping point
-dynamics merely on account of them being similar.  The data shown
-comes only from the basin of attraction near the unfished equilibrium,
-near which all of the models are approximately linear and approximately
-identical; so it is intuitive that all criteria choose the simplest model.
+Our results illustrate that model-choice approaches can be absolutely
+misleading – by providing support to models that cannot capture tipping point
+dynamics because they have fewer parameters and the data are far from the tipping point.  That is, when the data 
+come from around the stable steady state,
+all the parametric models are approximately linear and approximately
+identical.  Thus, it is intuitive that all model selection methods choose the simplest model.
 In a complex world, the result is suboptimal. But in a world that might
@@
 have not observed the population dynamics at all possible densities.  A
-lack of comprehensive data at all population sizes, combined with our obvious
+lack of comprehensive data at all population sizes, combined with the
 inability to formulate accurate population models for low population sizes
-in the absence of any data, makes this situation the rule more than the exception.
+in the absence of data, makes this situation the rule more than the exception.
 Relying on parametric models and model choice processes that favor simplicity
-ignores this basic fact. Acknowledging this, Carl Walters [e.g. @Walters1978] has long observed that if we began
+ignores this basic reality. For a long time,  Carl Walters [e.g. @Walters1978] has argued that if we began
 by fishing any newly exploited population down to very low levels and
@@
 
-The parametric models perform worst when they propose a management strategy
-outside the range of the observed data. The non-parametric approach of the GPDP, 
-in contrast, allows a predictive model that expresses a great deal of uncertainty
-about the probable dynamics _outside_ the observed range, while retaining very
-good predictive accuracy _inside_ the range observed.  The management policy 
-dictated by the GPDP balances this uncertainty against
+Parametric models perform most poorly when we seek a management strategy
+outside the range of the observed data. The GPDP, 
+in contrast, leads to a predictive model that expresses a great deal of uncertainty
+about the probable dynamics _outside_ the range of the observed data, while retaining very
+good predictive accuracy _inside_ the range.  The management policy 
+based on by the GPDP balances uncertainty outside the range of the observed data against
 the immediate value of the harvest, and acts to stabilize the population 
-dynamics in a region of state space in which the predictions can be 
+dynamics in a region of state space in which the predictions are
 reliably reflected by the data.
 
-Such problems are not 
-constrained to fisheries management but ubiquitous across ecological decision-making
-and conservation where the greatest concerns involve entering previously unobserved
-regions of state-space -- whether that is the collapse of a fishery, the spread
+Such problems are 
+ubiquitous across ecological decision-making
+and conservation where the greatest concerns involve decisions that lead to population sizes that have never been observed and for which we do not know the response
+-- whether this is the collapse of a fishery, the spread
 of an invasive, or the loss of habitat.
@@
 
-Lastly, it should be noted that outside the data, the GP reverts to
+Outside of the observed range of the data, the GP reverts to
 the prior, and consequently the choice of the prior can also play a
@@
 relatively uninformative (due to the broad priors placed on its parameters
-$\ell$ and $\sigma$ and simple (mean zero, radial basis function kernel).
-In practice, both the choice of mean and the covariance
-function may be chosen to confer particular biological properties,
-as well as more biologically informed priors for $\ell$ and $\sigma$.
+$\ell$ and $\sigma$ and simple (the choice of our covariance function, Eqns 12 and 13 ).
+In practice, these
+should be chosen to confer particular biological properties.
 In principle, this may allow a manager to improve the performance of the
-GPDP by adding only enough additional detail as is justified.
+GPDP by adding detail as is justified.
 For instance, it would be possible to use a linear or a Ricker-shaped
 mean in the prior without making the much stronger assumption that the
-Ricker is the structurally correct model [@Sugeno2013a]. Future research
-should focus on identifying criteria to ensure the prior and the value function
+Ricker is the structurally correct model [@Sugeno2013a]. One fruitful avenue of future research is  identifying criteria to ensure the prior and the value function
 are chosen appropriately for the problem at hand.

### 2014-11-07 manuscripts/manuscript.tex
update data file names to match figures

@@
 management of human intervention in natural resources and conservation.
-Ecological dynamics are frequently complex and difficult to measure,
-making uncertainty in our understanding a persistent challenge.
 Decision-theoretic approaches provide a framework to determine the best
@@
 uncertainty can be meaningfully quantified (Fischer et al. 2009). Over
-the last four decades (but stretching even further back in time), marked
-by the seminal contributions of Clark (1976), Clark (2009) and Walters
-and Hilborn (1978), dynamic optimization methods, particularly
+the last four decades (beginning with Clark (1976), Clark (2009) and
+Walters and Hilborn (1978)) dynamic optimization methods, particularly
 Stochastic Dynamic Programming (SDP), have become increasingly important
@@
 natural systems. Simultaneously, there has been increasing recognition
-of the importance of Allee population sizes or `tipping points'
-(Scheffer et al. 2001, Polasky et al. 2011) that management actions may
-inadvertently cause populations to cross. Scheffer et al. (2009)
-emphasized the difficulties of formulating even qualitatively correct
-models of the underlying processes.
+of the importance of multiple steady states or `tipping points'
+(Scheffer et al. 2001, 2009, Polasky et al. 2011) in ecological systems.
 
 We develop a novel approach to address these concerns in the context of
-fisheries; though the underlying challenges and methods are germane to
-many other problems of conservation or natural resource exploitation,
-such as forestry. Economic value and ecological concern have made marine
-fisheries the crucible for much of the founding work (Gordon 1954, Clark
-1976, 2009, May et al. 1979, Reed 1979, Ludwig and Walters 1982) on
-management under uncertainty. Global declines (Worm et al. 2006) and the
-controversy surrounding their interpretation (Hilborn 2007, Worm et al.
-2009) make understanding these challenges all the more pressing.
-
-Uncertainty enters the decision-making process at many levels: intrinsic
-stochasticity in biological processes, measurements, and implementation
-of policy (\emph{e.g.} Reed 1979, Clark and Kirkwood 1986, Roughgarden
-and Smith 1996, Sethi et al. 2005), parametric uncertainty (\emph{e.g.}
-Ludwig and Walters 1982, Hilborn and Mangel 1997, McAllister 1998,
-Schapaugh and Tyre 2013), and model or structural uncertainty
-(\emph{e.g.} Williams 2001, Cressie et al. 2009, Athanassoglou and
-Xepapadeas 2012). Of these, structural uncertainty is generally the
-hardest to quantify. Typical approaches assume a weak notion of model
-uncertainty in which the correct model (or reasonable approximation) of
-the dynamics must be identified from among a handful of alternative
-models, using either model choice, model averaging, or introducing yet
-greater model complexity of which others may be special cases (model
-averaging being one such way to construct such a model) (Williams 2001,
-Cressie et al. 2009, Athanassoglou and Xepapadeas 2012). Even setting
-aside other computational and statistical concerns (e.g. Cressie et al.
-(2009)), these approaches do not address our second concern -
-representing uncertainty outside the observed data range.
-
-Model uncertainty is particularly insidious when model predictions must
-be made outside of the range of data on which the model was estimated,
-and models may be abused (Mangel et al. 2001). Extrapolation uncertainty
-is felt most keenly in decision-theoretic (or optimal control)
-applications, since (a) exploring the space of all possible actions
+fisheries; although the challenges and methods are germane to other
+problems of conservation or natural resource exploitation. Economic
+value and ecological concern have made marine fisheries the crucible for
+much of the founding work on management under uncertainty (Gordon 1954,
+Clark 1976, 2009, May et al. 1979, Reed 1979, Ludwig and Walters 1982)
+
+Even if we know the proper deterministic description of the biological
+system, there is intrinsic stochasticity in biological dynamics,
+measurements, and implementation of policy (\emph{e.g.} Reed 1979, Clark
+and Kirkwood 1986, Roughgarden and Smith 1996, Sethi et al. 2005). We
+may also lack knowledge about the parameters of the biological dynamics
+(parametric uncertainty, \emph{e.g.} Ludwig and Walters 1982, Hilborn
+and Mangel 1997, McAllister 1998, Schapaugh and Tyre 2013), or even not
+know which model is proper description of the system (structural
+uncertainty, \emph{e.g.} Williams 2001, Cressie et al. 2009,
+Athanassoglou and Xepapadeas 2012). Of these, the latter is generally
+the hardest to quantify. Typical approaches confront the data with a
+collection of models, assuming that the true dynamics (or reasonable
+approximation) is among the collection and then use model choice or
+model averaging to arrive at a conclusion (Williams 2001, Cressie et al.
+2009, Athanassoglou and Xepapadeas 2012). Even setting aside other
+concerns (see Cressie et al. (2009)), these approaches are unable to
+describe uncertainty outside the observed data range.
+
+Structural uncertainty is particularly insidious when we try to predict
+outside of the range of observed data (Mangel et al. 2001) because we
+are extrapolating into unknown regions. In management applications, this
+extrapolation uncertainty is particularly important since (a) management
 involves considering actions that may move the system outside the range
-of observed behavior, and (b) decision-theoretic algorithms rely not
-only on reasonable estimates of the expected outcomes, but depend on the
-weights given to all possible outcomes (\emph{e.g.} Weitzman 2013),
-hence estimates of the uncertainty can be as important as estimates of
-the average or expected outcome.
-
-Concerns about tipping points in ecological dynamics (Scheffer et al.
-2001, Polasky et al. 2011) highlight this problem because they arise
-from this very situation. In such cases, precise models are not
-available and data is limited or biased to a particular region of state
-space, such as high stock levels or an otherwise desirable state for the
-ecosystem. If such models or data were available, one would know just
-how far a system could be pushed before going over the tipping point,
-and management would be simple. The fact that this concern is frequently
-discussed highlights how often we have only limited data and inperfect
-models. Moreover, the case of tipping points highlights how even small
-errors can have very large consequences, as we shall illustrate later.
-
-The dual concerns of model uncertainty and incomplete data coverage pose
-a substantial challenge to existing decision-theoretic approaches
-(Brozović and Schlenker 2011). Because intervention is often (but not
-always, see Hughes et al. (2013)) too late after a tipping point has
-been crossed, management is most often concerned with avoiding
-potentially catastrophic tipping points before any data are available at
-or following a transition that would more clearly reveal these regime
-shift dynamics (e.g. Bestelmeyer et al. 2012).
-
-We illustrate how a Stochastic Dynamic Programming (SDP) algorithm
-(Mangel and Clark 1988, Marescot et al. 2013) can be driven by the
-predictions from a Bayesian Non-Parametric (BNP) model of population
-dynamics (Munch et al. 2005a). This provides two distinct advantages
-compared with contemporary approaches. First, using a BNP sidesteps the
-need for an accurate model-based description of the system dynamics.
-Second, the BNP can better reflect uncertainty that arises when
-extrapolating a model outside of the data on which it was fit. We
-illustrate that when the correct model is not known, this latter feature
-is crucial to providing a robust decision-theoretic approach in face of
-substantial structural uncertainty. We use \emph{robust} to distinguish
-approaches that provide nearly optimal solutions (as measured by the
-chosen value function) without being sensitive to the choice of the
-(unknown) underlying model.
-
-This paper represents the first ecological application of the SDP
-decision-making framework without an a priori model of the underlying
-dynamics. In contrast to parametric models which can only reflect
-uncertainty in parameter estimates, the BNP approach provides a
-state-space dependent representation of uncertainty. This permits a much
-greater uncertainty far from the observed data than near the observed
-data. These features allow the Gaussian Process Dynamic Programming
-(GPDP) approach to find robust management solutions in face of limited
-data and without knowledge of the correct model structure.
-
-To illustrate GPDP, we consider the performance of the GPDP policy
-against the policies derived under several alternative parametric models
-(Reed 1979, Ludwig and Walters 1982, Mangel and Clark 1988). The nature
-of decision-making problems provides a compelling and pragmatic way to
-compare models: Rather than compare models in terms of best fit to data,
-we define model performance in the concrete terms of the
-decision-maker's objectives.
+of observed behavior, and (b) the decision tools (optimal control
+theory, SDP) rely on both reasonable estimates of the expected outcomes
+and on the weights given to those outcomes (\emph{e.g.} Weitzman 2013).
+Thus characterizing uncertainty is as important as characterizing the
+expected outcome.
+
+Tipping points in ecological dynamics (Scheffer et al. 2001, Polasky et
+al. 2011) highlight this problem because precise models are not
+available and data are limited such as around high stock levels or an
+otherwise desirable state. With perfect information , one would know
+just how far a system could be pushed before crossing the tipping point,
+and management would be simple. But we face imperfect models and limited
+data and, with tipping points,\\even small errors can have very large
+consequences, as we shall illustrate later. Because intervention is
+often (but not always, see Hughes et al. (2013)) too late after a
+tipping point has been crossed, management is often concerned with
+avoiding tipping points before any data about them is available or the
+transition that would more clearly reveal a tipping point occurs (e.g.
+Bestelmeyer et al. 2012).
+
+The dual concerns of model uncertainty and incomplete data create a
+substantial challenge to existing decision-theoretic approaches
+(Brozović and Schlenker 2011). We illustrate how Stochastic Dynamic
+Programming (SDP) (Mangel and Clark 1988, Marescot et al. 2013) can be
+implemented using a Bayesian Non-Parametric (BNP) model of population
+dynamics (Munch et al. 2005a). The BNP method has two distinct
+advantages . First, using a BNP model sidesteps the need for an accurate
+model-based description of the system dynamics. Second, a BNP model
+better reflects uncertainty when extrapolating beyond the observed data.
+This is crucial to providing robust decision-making when the correct
+model is not known (as is almost always true). {[}We use \emph{robust}
+to characterize approaches that provide nearly optimal solutions without
+being sensitive to the choice of the (unknown) underlying model.{]}
+
+This paper is the first ecological application of the SDP without an
+\emph{a priori} model of the underlying dynamics. Unlike parametric
+approaches that can only reflect uncertainty in parameter estimates, the
+BNP method provides a broader representation of uncertainty, including
+uncertainty beyond the observed data. We will show that Gaussian Process
+Dynamic Programming (GPDP) allows us to find robust management solutions
+in face of limited data without knowing the correct model structure.
+
+For comparisons, we consider the performance of management based on GPDP
+against management policies derived under several alternative parametric
+models (Reed 1979, Ludwig and Walters 1982, Mangel and Clark 1988).
+Rather than compare models in terms of best fit to data, we compare
+model performance in the concrete terms of the decision-maker's
+objectives.
 
@@
 we describe three parametric models for population dynamics and the
-Gaussian Process (GP)\footnote{Note that we abbreviate Gaussian Process
-  as GP, which refers only to the statistical model we use to
-  approximate the population dynamics (as opposed to a familiar
-  parametric model such as the Ricker equation), while we use the term
-  GPDP, Gaussian Process Dynamic Programming, to refer to the \emph{use}
-  of a GP as the underlying process model when solving a Dynamic
-  Program, as we illustrate in this paper. Hence we will refer to the
-  models as: GP, Ricker, Allen, etc, while we refer to the novel method
-  we put forward here as GPDP.} description of population dynamics.
+Gaussian Process (GP)\footnote{We abbreviate Gaussian Process as GP,
+  which refers to the statistical model we use to approximate the
+  population dynamics , and we use the term Gaussian Process Dynamic
+  Programming {[}GPDP,{]}, to refer to the \emph{use} of a GP as the
+  underlying process model when solving a Dynamic Programming equation .
+  Hence we will refer to the models as: GP, Ricker, Allen, etc, and the
+  novel method we put forward here as GPDP.} description of population
+dynamics.
 
@@
 Dynamic optimization requires characterizing the dynamics of a state
-variable (or variables), a control action, and a value function. In this
-paper, we consider only a single state variable, so that the ideas can
-be described as simply as possible. That is, we focus on a class of
-one-dimensional models of population dynamics as a `best-case' scenario
-for the parametric model performance relative to the GP. This is a
-best-case scenario for the parametric models because the underlying
-dynamics are always simulated from one of the parametric models in the
-set, whereas in reality we never have the ``true'' model among our
-candidates. By choosing one-dimensional models with few parameters, we
-limit the chance that poor performance will be due to our inability to
-estimate parameters well, something that becomes a more severe problem
-for higher-dimensional parametric models. Further, the parametric models
-we consider are those most commonly used in modeling stock-recruitment
-dynamics or to model sudden transitions between alternative stable
-states.
+variable (or variables), a control action, and a value function. For
+simplicity, we consider only a single state variable, which allows us to
+focus on one-dimensional models of population dynamics for the
+comparison of parametric models and the GP. This is a best-case scenario
+for the parametric models because we simulate underlying dynamics from
+one of the three parametric models , whereas in the natural world we
+never know the ``true'' mode . In addition, by choosing one-dimensional
+models with just a few parameters, we limit the chance that poor
+performance will be due to inability to estimate parameters accurately,
+something that becomes a more severe problem for higher-dimensional
+parametric models. Finally, the parametric models we consider are those
+most commonly used in modeling stock-recruitment dynamics or to model
+sudden transitions between alternative stable states.
 
@@
 
-We let $x(t)$ denote the size (numbers or biomass) of the focal
+We let $X(t)$ denote the size (numbers or biomass) of the focal
 population at time $t$ and assume that in the absence of take its
-dynamics are. (Throughout, we will use $X(t+1)$ with a capital $X$ as a
-reminder that this is a stochastic, uncertainty quantity at time $t$,
-while we use lowercase $x(t)$ as a reminder that the current population
-size is treated as a given without error. The latter assumption is
-straight-forward to relax but will simplify the presentation here).
-
-\begin{equation}
-X(t+1) = Z(t) f(x(t), p) \label{eq1} 
+dynamics are. .
+
+\begin{equation}
+X(t+1) = Z(t) f(X(t), p) \label{eq1} 
 \end{equation}
@@
 and $p$ is a vector of parameters to be estimated from the data. We
-describe three choices for $f(X(t),p)$ in the next section.In this
-simple model, the control action is a harvest or take, $h(t)$, measured
-in the same units as $X$, at time $t$. Thus, in the presence of take,
-the population size on the right hand side of Eqn 1 is replaced by
-$S(t) = X(t) - h(t)$.
-
-To construct the value function, we consider a return when $X(t) = x $
-and harvest $h(t) = h$ denoted as the reward, $R(x, h)$. For example, if
-the return is simply the harvest at time $t$, then
-$R(x, h) = \min(x, h)$. We assume that future harvests are discounted
-relative to current ones at a constant rate of discount $\delta$ and ask
-for the harvest policy that maximizes total discounted harvest between
-the current time $t$ and a final time $T$. That is, we seek to maximize
-over choices of harvest
-$\mathbf{E}_{X(t+1)} [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$,
-where the state dynamics are given by Eqn 1 and $\mathbf{E}$ denotes the
-expectation over the process stochasticity in the future population
-state.
+describe the three choices for $f(X(t),p)$ in the next section.\\The
+control action is a harvest or take, $h(t)$, measured in the same units
+as $X$, at time $t$. Thus, in the presence of take, the population size
+on the right hand side of Eqn 1 is replaced by $S(t) = X(t) - h(t)$.
+
+To construct the value function, we consider a return when $X(t) = x(t)$
+and harvest $h(t) = h$ denoted as the reward, $R(x(t), h)$. For example,
+if the return is the harvest at time $t$, then
+$R(x(t), h(t)) = \min(x(t), h(t))$. We assume that future harvests are
+discounted relative to current ones at a constant rate of discount
+$\delta$ and ask for the harvest policy that maximizes total discounted
+harvest between the current time $t$ and a final time $T$. That is, we
+seek to maximize over choices of harvest
+$E_{X(t+1)} [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$, where the
+state dynamics are given by Eqn 1 and $E$ denotes the expectation over
+the process stochasticity in the future population state.
 
@@
 equation of SDP (Mangel and Clark 1988, Clark and Mangel 2000, Clark
-2009), sometimes called the Hamilton-Jacobi-Bellman equation (Mangel
-2014),
+2009 Mangel 2014),
 
@@
 That is, at time $t$, when population size is $x(t)$ and harvest $h(t)$
-is applied, the immediate return is $R(h(t), x(t))$. In the classical
-case, the sole source of uncertainty is the process stochasticity term,
-$Z$, and thus the expectation above is equivalent to taking expectations
-over $Z(t)$. That is
-
-\begin{equation}
-\mathbf{\mathrm{E}}_{X(t+1)} \left[ V( X(t+1),t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{Z(t)} \left[V( Z(t) f(x(t) - h(t))), t+1 | x(t), h(t) \right]
+is applied, the immediate return is $R(h(t), x(t))$. When the sole
+source of uncertainty is the process stochasticity term, $Z$, and thus
+the expectation in Eqn 2 is equivalent to taking expectations over
+$Z(t)$. That is
+
+\begin{equation}
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V( X(t+1),t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{Z(t)} \left[V( Z(t) f(x(t) - h(t))|p), t+1 | x(t), h(t) \right]
 \end{equation}
@@
 where the population size after the take is $x(t) - h(t)$, which is then
-translated into $X(t+1)$ by Eqn 1 (that is, we just replaced $X(t+1)$
-with $Z(t) f(x(t)-h(t))$).
-
-In a Bayesian decision framework, the parameters governing the dynamics
-are also uncertain, so the expectation involves averaging over the
-posterior distribution for the parameters, as well. That is,
-
-\begin{equation}
-\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{p|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z(t) | p, \mathrm{data}} \left[ V(Z(t) f(x(t) - h(t)), t)  \right] \}
-\end{equation}
-
-In the nonparametric case, the function $f$ too is uncertain and the
+translated into $X(t+1)$ by Eqn 1 (that is, we replace $X(t+1)$ by
+$Z(t) f(x(t)-h(t)|p)$).
+
+When the parameters governing the dynamics are also uncertain (the case
+of parametric uncertainty), the expectation in Eqn 2 involves averaging
+over the posterior distribution for the parameters. That is,
+
+\begin{equation}
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{p|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z(t) | p, \mathrm{data}} \left[ V(Z(t) f(x(t) - h(t)|p), t)  \right] \}
+\end{equation}
+
+When the underlying population dynamics are unknown (the case of
+structural uncertainty), the function $f$ itself is uncertain and the
 expectation for the next state includes uncertainty in $f$ as well. That
@@
 
-We consider three candidate parametric models of the stock-recruitment
-dynamics: The Ricker model, the Allen model (Allen and Tanner 2005), the
-Myers model (Myers et al. 1995), Eqns \eqref{ricker}-\eqref{myers},
-which we parameterize as follows for $f(S(t),p)$ in Eqn 1. In all three,
-we let $K$ denote the carrying capacity and $r$ the maximum per capita
-growth rate. In the Ricker model,
-
-\begin{equation}
-f(S(t)|p) = S(t) e^{r \left(1 - \frac{S(t)}{K} \right) } \label{ricker}
-\end{equation}
-
-In the Allen model,
-
-\begin{equation}
-f(S(t)|p) = S(t) e^{r \left(1 - \frac{S(t)}{K}\right)\left(S(t) - X_c\right)} \label{allen}
-\end{equation}
-
-Where $r$ and $K$ are as before, and $X_c$ denotes the location of the
-unstable steady state (i.e., the tipping point).
-
-In the Myers model,
-
-\begin{equation}
-f(S(t) | p)  = \frac{r S(t)^{\theta}}{1 + \frac{S(t)^\theta}{K}} \label{myers}
-\end{equation}
-
-Where $\theta = 1$ corresponding to Beverton-Holt dynamics and $\theta$
+We consider three candidate parametric models for the population
+dynamics: The Ricker model, the Allen model (Allen and Tanner 2005), and
+the Myers model (Myers et al. 1995), Eqns \eqref{ricker}-\eqref{myers}.
+In all three, we let $K$ denote the carrying capacity and $r$ the
+maximum per capita growth rate. The Ricker model has two parameters and
+the right hand side of Eqn 1 is
+
+\begin{equation}
+f(S(t)|r,K) = S(t) e^{r \left(1 - \frac{S(t)}{K} \right) } \label{ricker}
+\end{equation}
+
+The Allen model has three parameters
+
+\begin{equation}
+f(S(t)|r, K, X_c) = S(t) e^{r \left(1 - \frac{S(t)}{K}\right)\left(S(t) - X_c\right)} \label{allen}
+\end{equation}
+
+where $X_c$ denotes the location of the unstable steady state (i.e., the
+tipping point).
+
+The Myers model also has three parameters
+
+\begin{equation}
+f(S(t) | r, K, \theta)  = \frac{r S(t)^{\theta}}{1 + \frac{S(t)^\theta}{K}} \label{myers}
+\end{equation}
+
+where $\theta = 1$ corresponds to Beverton-Holt dynamics and $\theta$
 \textgreater{} 2 leads to Allee effects and multiple stable states.
 
-The Ricker model involves two parameters, corresponding to a growth rate
-and a carrying capacity, and cannot support alternative stable growth
-rate dynamics. The Allen model resembles the Ricker dynamics with an
-added Allee effect parameter (Courchamp et al. 2008), below which the
-population cannot persist. The Myers model also has three parameters and
-contains an Allee threshold, but has compensatory rather than
-over-compensatory density dependence (resembling a Beverton-Holt curve
-rather than a Ricker curve at high densities.)
-
-The multiplicative log-normal stochasticity perturbs the growth
-predicted by each of these deterministic model skeletons. This
-introduces one additional parameter $\sigma$ that must be estimated.
+The Ricker model does not lead to multiple steady states. The Allen
+model resembles the Ricker dynamics with an added Allee effect parameter
+(Courchamp et al. 2008), below which the population cannot persist. The
+Myers model also has three parameters and contains an Allee threshold,
+but unlike the Ricker model saturates at high population size. The
+multiplicative log-normal stochasticity in Eqn 1 introduces one
+additional parameter $\sigma$ that must be estimated.
 
@@
 tipping points, all of our simulations are based on the Allen model. The
-Allen model is thus ``structurally correct'' and is expected to provide
-a best-case scenario when the `true' dynamics are known. The Ricker
-model is a reasonable approximation of these dynamics far from the Allee
-threshold (but lacks threshold dynamics), while the Myers model shares
-the essential feature of a threshold but differs in the structure.
-
-Throughout the presentation, we will refer to the use of the ``True''
-model to denote solving for the optimal solution when the underlying
-parameters \emph{are known without error}, and refer to the ``Allen''
-model to denote the when these parameters have instead only been
-estimated from the sample data.
-
-We consider a period of 40 years of training data: long enough that the
-estimates do not depend on the particular realization, while longer
-times are not likely to provide substantial improvement.\\Each of the
-models is fit to the same training data (Figure 1).
+Allen model is thus the state of nature and is expected to provide the
+best-case scenario. The Ricker model is a reasonable approximation of
+these dynamics far from the Allee threshold (but lacks threshold
+dynamics), while the Myers model shares the essential feature of a
+threshold but differs in structure from the Allen model. Throughout, we
+refer to the ``True'' model when the underlying parameters \emph{are
+known without error}, and refer to the ``Allen'' model when these
+parameters have been estimated from the sample data.
+
+We consider a period of 40 in which data are obtained to estimate the
+parameters or the GP. This is long enough that the estimates do not
+depend on the particular realization, and longer times are not likely to
+provide substantial improvement.\\We fit each of the models is fit to
+the same data (Figure 1).
 
@@
 
-The core difference for our purpose between the GP representation and
-the parametric models is that the GP model is defined explicitly in
-reference to some observed data. Unlike the models above, it cannot be
-specified by the value of some parameters alone. Although called
-nonparametric, the GP (a) still involves the estimation of parameters,
-and (b), statisticians also use non-parametric to mean models that make
-only weak distributional assumptions (e.g. Lehmann 1975). However, we
-retain the terminology for consistency with previous studies.
-
-The use of GP methods to formulate a predictive model is relatively new
-in the context of modeling dynamical systems (Kocijan et al. 2005), and
-was first introduced in the context ecological modeling and fisheries
-management in Munch et al. (2005b). GP models have subsequently been
-used to test for the presence of Allee effects (Sugeno and Munch 2013a),
-estimate the maximum reproductive rate (Sugeno and Munch 2013b),
-determine temporal variation in food availability (Sigourney et al.
-2012), and provide a basis for identifying model-misspecification
-(Thorson et al. 2014). An accessible and thorough introduction to the
-formulation and use of GPs can be found in Rasmussen and Williams
-(2006).
-
-A GP is an infinite dimensional (because it is a function of a
-continuous input variable) stochastic process. Any realization of the
-process for some finite sample of points $n$ is a multivariate normal
-distribution of dimension $n$. Thus, to characterize the GP we need a
-mean function and a covariance function. We proceed as follows.
-
-We assume that the data $X(t)$ are observed with some process
-stochasticity,
+The core difference for our purpose between the GP and the parametric
+models is that the GP model is defined explicitly by the observed data.
+That is, it cannot be specified by the value of parameters alone.
+
+The use of GPs to characterize dynamical systems is relatively new
+(Kocijan et al. 2005), and was first introduced in the context
+ecological modeling and fisheries management in Munch et al. (2005b). GP
+models have subsequently been used to test for the presence of Allee
+effects (Sugeno and Munch 2013a), estimate the maximum reproductive rate
+(Sugeno and Munch 2013b), determine temporal variation in food
+availability (Sigourney et al. 2012), and provide a basis for
+identifying model-misspecification (Thorson et al. 2014). An accessible
+and thorough introduction to the formulation and use of GPs can be found
+in Rasmussen and Williams (2006).
+
+A GP is a stochastic process for which any realization of for a finite
+sample of points $n$ is a multivariate normal distribution of dimension
+$n$. Thus, to characterize the GP we need a mean function and a
+covariance function. We proceed as follows.
+
+As before, we assume that the data $X(t)$ are observed with process
+stochasticity around a mean function \$ g(X(t))\$
 
@@
 variance $\sigma_n^2$. Note that we have chosen to assume additive
-stochasticity. While we could just as easily consider log-normal
-stochasticity as in the parametric models, we make this choice to bring
-home the point that the Gaussian process approach need not have
-structurally correct noise form either.
-
-We consider the fase where we have some previous observations of the
-\emph{timeseries}, $X_{\textrm{obs}}(t)$, and wish to \emph{predict}
-future values $X_{\textrm{p}}(t)$. We consider the observations as pairs
-of points $X_{\textrm{obs}}(t), X_{\textrm{obs}}(t+1)$ for
-$t \in [1,2, ..., T]$ up through observed time $T$. Under the GP, a
-finite sample of these predictions at $n$ points into the future,
-$X_{\textrm p}(\tau)$ for $\tau \in [T+1, T+2, ..., T+n]$, is simply
-given by a multivariate normal ($\mathcal{N}$):
+stochasticity. While we could consider log-normal stochasticity as in
+the parametric models, we make this choice to emphasize that the
+Gaussian process approach need not have structurally correct
+stochasticity to be effective.
+
+Now consider the situation where we have a set of previous observations
+of the \emph{timeseries}, $X_{\textrm{obs}}(t)$, and wish to
+\emph{predict} future values $X_{\textrm{p}}(t)$. We consider the
+observations as pairs of points
+$X_{\textrm{obs}}(t), X_{\textrm{obs}}(t+1)$ for $t=1, 2, …$ up through
+final observation time $T$. As described above, in the GP, a finite
+sample of these predictions at $n$ points into the future,
+$X_{\textrm p}(\tau)$ for $\tau \in [T+1, T+2, ..., T+n]$, is a
+multivariate normal ($\mathcal{N}$):
 
@@
 
-which has mean and covariance matrix given by:
+with mean and covariance matrix given by (see Eqn 2.22 of Rasmussen and
+Williams (2006) for a more detailed derivation)
 
@@
 
-Where $K$ is the chosen covariance kernel. (Compare to Eqn 2.22 of
-Rasmussen and Williams (2006) for a more detailed derivation.) Many
-options are possible for this kernel. Throughout, we use the radial
-basis function for simplicity, a matrix whose $(i,j)$th element is given
-by:
-
-\begin{equation}
-K_{i,j}(x_i, y_j) = \exp\left(\frac{-|\mathbf{x_i}-\mathbf{j}|^2}{2 \ell^2} \right)
-\end{equation}
+Where $K$, called the covariance kernel, measures how much influence one
+observation has on another. In this paper, for $K$ use a matrix whose
+$(i,j)$th element is given by:
+
+\begin{equation}
+K_{i,j}(x_i, y_j) = \exp\left( \frac{ -(x_i - y_j)^2}{2 \ell^2} \right)
+\end{equation}
+
+so that $\ell$ gives the characteristic length-scale over which
+correlation between two observations decays.
 
 We use inverse Gamma priors on both the length-scale $\ell$ and
-$\sigma$, of the form
-
-\begin{equation}
-f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)
-\end{equation}
-
-For the $\sigma$ prior, $\alpha = $ 5 and $\beta = $ 5. For $\ell$
-prior, $\alpha = $ 10 and $\beta = $ 10.
-
-Using the simulated training data we also estimate a Gaussian process
-defined by a radial basis function kernel of two parameters: $\ell$,
-which gives the characteristic length-scale over which correlation
-between two points in state-space decays, and $\sigma$, which gives the
-scale of the process noise by which observations $Y(t+1)$ may differ
-from their predicted values $X(t+1)$ given an observation of the
-previous state, $x(t)$ (see Munch et al. 2005a).
+$\sigma$, thus for example
+
+\begin{equation}
+f(\ell; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} \ell^{-\alpha - 1}\exp\left(-\frac{\beta}{\ell}\right)
+\end{equation}
+
+For the prior on $\ell$, $\alpha = $ 10 and $\beta = $ 10. The prior on
+$\sigma$, $\alpha = $ 5 and $\beta = $ 5.
 
 We use a Metropolis-Hastings Markov Chain Monte Carlo (Gelman et al.
-(2003)) to infer posterior distributions of the two parameters of the GP
+(2003)) to infer posterior distributions of the parameters of the GP
 (Figure S3, code in appendix). Since the posterior distributions differ
-substantially from the priors (Figure S3), we can be assured that most
-of the information in the posterior comes from the data rather than the
-prior belief.
+substantially from the priors (Figure S3), most of the information in
+the posterior comes from the data rather than the prior belief.
 
@@
 
-It is reasonably straight-forward to derive the harvest policy from the
-estimated GP by inserting it into a SDP algorithm. The only difference
-is that the uncertainty in the future state under the GP, $X(t+1)$,
-includes both process uncertainty (based on the estimation of $\sigma$)
-and ``structural uncertainty'' of the posterior collection of
-curves.\\(Recall that given the parameters, a parametric model is
-represented by a single curve, while the GP remains a distribution of
-curves).\\From the GP posteriors, we can write down the (discritized)
-transition matrix representing the probability of going to each state
-$X(t+1)$ given any current state $x(t)$ and any harvest $h(t)$ (See the
-function \texttt{gp\_transition\_matrix()} in the provided R package).
-Given this transition matrix, we use the same value iteration algorithm
-as in the parametric case to determine the optimal policy.
+We derive the harvest policy from the estimated GP by inserting it into
+a SDP algorithm. Given the GP posteriors, we construct the transition
+matrix representing the probability of going to each state $X(t+1)$
+given any current state $x(t)$ and any harvest $h(t)$ (See the function
+\texttt{gp\_transition\_matrix()} in the provided R package). Given this
+transition matrix, we use the same value iteration algorithm as in the
+parametric case to determine the optimal policy.\\In doing so, the
+uncertainty in the future state under the GP, $X(t+1)$, includes both
+process uncertainty (based on the estimation of $\sigma$) and structural
+uncertainty of the posterior collection of curves.\\Unlike the case of a
+parametric model, which is a single curve given the parameters, the GP
+remains a distribution of curves.
 
@@
 
-To ensure our results are robust to our choice of parameters, consider
-96 different scenarios, described in detail below. To help better
-understand the process, we first describe in detail the results of a
-single scenario.
+To ensure our results are robust to the choice of parameters, we
+consider 96 different scenarios, described in detail below. To help
+better understand the process, we first describe in detail the results
+of a single scenario.
 
@@
 
-All models fit the observed data rather closely and with relatively
-small uncertainty, as illustrated in the posterior predictive curves in
-Figure 1, in which we show the training data of stock sizes observed
-over time as points, overlaid with the step-ahead predictions of each
-estimated model using the parameters sampled from their posterior
-distributions. Each model manages to fit the observed data rather
-closely. Compared to the true model most estimates appear to over-fit,
-predicting fluctuations that are actually due purely to stochasticity.
-Model selection criteria (Table 1) penalize more complex models and show
-a preference for the simpler Ricker model over the models with
-alternative stable states (Allen and Myers). Details on MCMC estimates
-for each model, traces, and posterior distributions can be found in the
-appendix.
+All of the models fit the observed data rather closely and with
+relatively small uncertainty. In Figure 1, we show the posterior
+predictive curves. The training data of stock sizes observed over time
+are points, overlaid with the step-ahead predictions of each estimated
+model using the parameters sampled from their posterior distributions.
+Compared to the true model most estimates appear to over-fit, predicting
+patterns that are actually due purely to stochasticity. Model selection
+criteria (Table 1) penalize more complex models and show a preference
+for the simpler Ricker model over the models with alternative stable
+states (Allen and Myers). Details on MCMC estimates for each model,
+traces, and posterior distributions can be found in the appendix.
 
@@
 Deviance Information Criterion, AIC: Akaike Information Criterion, BIC:
-Bayesian Information Criterion) all (wrongly) select the simplest model.
-As the true (Allen) model is not distinguishable from the simpler
-(Ricker) model in the region of the observed data, this error cannot be
-avoided regardless of the model choice criterion. This highlights the
-danger of model choice when the selected model will be used outside of
-the observed range of the data.}
+Bayesian Information Criterion) all select the wrong model. As the true
+(Allen) model is not distinguishable from the simpler (Ricker) model in
+the region of the observed data, this error cannot be avoided regardless
+of the model choice criterion. This highlights the danger of model
+choice when the selected model will be used outside of the observed
+range of the data.}
 \end{longtable}
@@
 
-We show the mean inferred state space dynamics of each model relative to
+We show the mean inferred population dynamics of each model relative to
 the true model used to generate the data in Figure 2, predicting the
-relationship between observed stock size (x-axis) to the stock size
-after recruitment the following year.\\In addition to the raw data,
-Figure 2 illustrates that the GP is conditioned on going through the
-point 0,0 without error. This conditioning on (0,0) is equivalent to
-making the assumption that a species hitting 0 is extinct forever \&
-without uncertainty, despite the lack of any data in the observed
-sequence to justify this. The use of this assumption illustrates how the
-GP approach can capture common-sense biology without having to assume
-more explicit details about the dynamics at low population numbers that
-have never been observed. Not that we need need not make this assumption
-(perhaps there is immigration after all); and one could repeat the
-entire analysis without this assumption.
-
-Note that in contrast to the other models shown, the expected Gaussian
-process corresponds to a distribution of curves - as indicated by the
-gray band - which itself has a mean shown in black. Parameter
-uncertainty (not shown) spreads out the estimates further. The observed
-data from which each model is estimated is also shown. The observations
-come from only a limited region of state space corresponding to
-unharvested or weakly harvested system. No observations occur at the
-theoretical optimum harvest rate or near the tipping point. Figure S1
-(see supplement) shows the performance of the models outside the
-observed training data.
+relationship between observed population size (x-axis) to the population
+size after recruitment the following year.\\In addition to the raw data,
+the GP is conditioned on going through the point 0,0 without error. This
+conditioning on (0,0) is equivalent to making the assumption that the
+population is closed, so that once it hits 0 it stays at 0, despite the
+lack of any data in the observed sequence to justify this. This
+assumption illustrates how the GP can capture common-sense biology
+without having to assume more explicit details about the dynamics at low
+population numbers that have never been observed. If the population were
+not closed, one could repeat the entire analysis without this
+assumption. Unlike parametric models, the GP corresponds to a
+distribution of curves - the gray band with mean shown in black.
+Uncertainty in the parameters of the GP (not shown) further widens the
+band of possible population sizes. In Figure S1 (see supplement), we
+show the performance of the models outside the observed training data.
 
@@
 In Figure 4, we show the consequences of managing 100 replicate
-realizations of the simulated fishery under each of the policies
-estimated. As anticipated from the policy curves, the structurally
-correct model under-harvests, leaving the stock to vary around its
-unfished optimum. The structurally incorrect Ricker model over-harvests
-the population past the tipping point consistently, resulting in the
-immediate crash of the stock and thus derives minimal long-term catch.
-
-The results across replicate stochastic simulations can most easily be
+realizations of the simulated fishery under policies derived from each
+model . The structurally correct model under-harvests, leaving the stock
+to vary around its unfished optimum. The structurally incorrect Ricker
+model over-harvests the population past the tipping point consistently,
+resulting in the immediate crash of the stock and thus leads to minimal
+long-term catch.
+
+The results across replicate stochastic simulations are most easily
 compared by using the relative differences in net present value realized
-by each of the model (Figure 5). The GPDP most consistently realizes a
-value close to the optimal solution, and importantly avoids ever driving
-the system across the tipping point, which results in the near-zero
-value cases in the parametric models.
+by each of the model (Figure 5). Although not perfect, the GPDP
+consistently realizes a value close to the optimal solution, and avoids
+ever driving the system across the tipping point, which results in the
+near-zero value cases in the parametric models.
 
@@
 simulation. The GPDP estimate remains very close to the optimal solution
-obtained by knowing the true model across changes to the training
-simulation, noise scale, parameters or structure of the underlying
-model. We consider both a Latin hypercube approach (see supplementary
-materials, Fig S10) and a more focused investigation of the parameters
-most responsible for impacting the transition rate: the relative
-distance to the Allee threshold and the noise effects.
+(obtained by knowing the true model) across changes to the training
+simulation, scale of stochasticity, parameters or structure of the
+underlying model. In the Supplement, we consider both a Latin hypercube
+approach and a more focused investigation of the effects of the relative
+distance to the Allee threshold and the variance of process
+stochasticity.
 
 Changing the intensity of the stochasticity or the distance between
-stable and unstable points does not impact the performance of the GP
-relative to the optimal solution given the true model and true
-parameters. The parametric models are more sensitive to this difference.
-Large noise relative to the distance between the stable and unstable
-point increases the chance of a stochastic transition. More precisely,
-if we let $L = K - x_c$, then the probability of a transition in a given
-window of time $T$ scales as
-
-\[P(x(t) < x_c | t \in T) \propto  \exp\left(\frac{L^2}{\sigma^2}\right),\]
-
-(see Gardiner (2009) or Mangel (2006) for the derivation). As a result,
-the impact of using a model which underestimates the risk of harvesting
-past the critical point is much more severe, since this such a situation
+stable and unstable steady states does not impact the performance of the
+GP relative to the optimal solution obtained from the true model and
+true parameters. The parametric models are more sensitive to this
+difference. Large values of $\sigma$ relative to the distance between
+the stable and unstable point increases the chance of a stochastic
+transition. More precisely, if we let $L$ is the distance between the
+stable and unstable steady states, then the probability that
+fluctuations drive the population across the unstable steady state
+scales as $\exp\left(\frac{L^2}{\sigma^2}\right)$
+
+(see Gardiner (2009) or Mangel (2006) for the derivation). Thus, the
+impact of using a model that underestimates the risk of harvesting
+beyond the critical point is considerable, since this such a situation
 occurs more often. Conversely, with large enough distance between the
-optimal escapement and unstable points relative to the noise, the chance
-of a transition becomes vanishingly small and all models can be
+optimal escapement and unstable steady state relative to $\sigma$, the
+chance of a transition becomes vanishingly small and all models can be
 estimated near-optimally. Models that underestimate the cost incurred by
-stock sizes fluctuating significantly below the optimal escapement level
-will not perform poorly as long as those fluctuations are sufficiently
-rare.
+population sizes fluctuating significantly below the optimal escapement
+level will not perform poorly as long as those fluctuations are
+sufficiently small.
 
 The GPDP is only weakly influenced by increasing stochasticity or
-increasing Allee effects over much of the range (Figure 6). Much larger
-stochasticity or higher Allee effects make even the optimal solution
-without any model or parameter uncertainty unable to harvest the
-population effectively (e.g.~noise large enough to violate the
-self-sustaining criterion of Reed (1979)). Only at the largest Allee
-effect sizes are some replicates seen to deviate more from the optimal
-solution.
+increasing Allee effects over much of the range (Figure 6). Larger
+$\sigma$ or higher Allee levels make even the optimal solution without
+any model or parameter uncertainty unable to harvest the population
+effectively (e.g.~the stochasticity is large enough to violate the
+self-sustaining criterion of Reed (1979)).
 
@@
 
-Though simple mechanistically motivated models offer the greatest
-potential to increase our basic understanding of ecological processes
-(Geritz and Kisdi 2012, Cuddington et al. 2013), such models can be both
-inaccurate and misleading when relied upon in a quantitative decision
-making framework. In this paper we have tackled two aspects of
-uncertainty that are both common to many ecological decision-making
-problems and fundamentally challenging to existing approaches which
-largely rely on parametric models:
+Simple, mechanistically motivated models offer the potential to increase
+our basic understanding of ecological processes (Geritz and Kisdi 2012,
+Cuddington et al. 2013). But such models can be both inaccurate and
+misleading when used in a decision making framework. In this paper we
+tackled two aspects of uncertainty that are common to many ecological
+decision-making problems and fundamentally challenging to existing
+approaches that largely rely on parametric models:
 
@@
 \item
-  We do not know what the correct models are for ecological systems.
+  We do not know the correct models for ecological systems.
 \item
-  We have limited data from which to estimate the model -- in
-  particular, such models may be misleading in predicting the
-  probability of outcomes outside the training data.
+  We have limited data from which to estimate the model.
 \end{enumerate}
 
-We have illustrated how the use of non-parametric approaches can provide
-more reliable solutions in the sequential decision-making problem.
+We have illustrated how the use of non-parametric methods provides more
+reliable solutions in the sequential decision-making problem.
 
@@
 
-These results illustrate that model-choice approaches would be
-positively misleading -- supporting simpler models that cannot express
-tipping point dynamics merely on account of them being similar. The data
-shown comes only from the basin of attraction near the unfished
-equilibrium, near which all of the models are approximately linear and
-approximately identical; so it is intuitive that all criteria choose the
-simplest model. In a complex world, the result is suboptimal. But in a
-world that might contain tipping points, the result is disastrous.
+Our results illustrate that model-choice approaches can be absolutely
+misleading -- by providing support to models that cannot capture tipping
+point dynamics because they have fewer parameters and the data are far
+from the tipping point. That is, when the data come from around the
+stable steady state, all the parametric models are approximately linear
+and approximately identical. Thus, it is intuitive that all model
+selection methods choose the simplest model. In a complex world, the
+result is suboptimal. But in a world that might contain tipping points,
+the result is disastrous.
 
@@
 densities. A lack of comprehensive data at all population sizes,
-combined with our obvious inability to formulate accurate population
-models for low population sizes in the absence of any data, makes this
-situation the rule more than the exception. Relying on parametric models
-and model choice processes that favor simplicity ignores this basic
-fact. Acknowledging this, Carl Walters (e.g. Walters and Hilborn 1978)
-has long observed that if we began by fishing any newly exploited
-population down to very low levels and then let it recover, we would be
-much better at estimating population dynamics and thus predicting the
-optimal harvest levels. While certainly true, this presents a rather
-risky policy in the face of potential tipping points. The GPDP offers an
-alternative way to acknowledge this uncertainty.
+combined with the inability to formulate accurate population models for
+low population sizes in the absence of data, makes this situation the
+rule more than the exception. Relying on parametric models and model
+choice processes that favor simplicity ignores this basic reality. For a
+long time, Carl Walters (e.g. Walters and Hilborn 1978) has argued that
+if we began by fishing any newly exploited population down to very low
+levels and then let it recover, we would be much better at estimating
+population dynamics and thus predicting the optimal harvest levels.
+While certainly true, this presents a rather risky policy in the face of
+potential tipping points. The GPDP offers an alternative way to
+acknowledge this uncertainty.
 
@@
 
-The parametric models perform worst when they propose a management
-strategy outside the range of the observed data. The non-parametric
-approach of the GPDP, in contrast, allows a predictive model that
-expresses a great deal of uncertainty about the probable dynamics
-\emph{outside} the observed range, while retaining very good predictive
-accuracy \emph{inside} the range observed. The management policy
-dictated by the GPDP balances this uncertainty against the immediate
-value of the harvest, and acts to stabilize the population dynamics in a
-region of state space in which the predictions can be reliably reflected
-by the data.
-
-Such problems are not constrained to fisheries management but ubiquitous
-across ecological decision-making and conservation where the greatest
-concerns involve entering previously unobserved regions of state-space
--- whether that is the collapse of a fishery, the spread of an invasive,
-or the loss of habitat.
+Parametric models perform most poorly when we seek a management strategy
+outside the range of the observed data. The GPDP, in contrast, leads to
+a predictive model that expresses a great deal of uncertainty about the
+probable dynamics \emph{outside} the range of the observed data, while
+retaining very good predictive accuracy \emph{inside} the range. The
+management policy based on by the GPDP balances uncertainty outside the
+range of the observed data against the immediate value of the harvest,
+and acts to stabilize the population dynamics in a region of state space
+in which the predictions are reliably reflected by the data.
+
+Such problems are ubiquitous across ecological decision-making and
+conservation where the greatest concerns involve decisions that lead to
+population sizes that have never been observed and for which we do not
+know the response -- whether this is the collapse of a fishery, the
+spread of an invasive, or the loss of habitat.
 
@@
 
-Lastly, it should be noted that outside the data, the GP reverts to the
-prior, and consequently the choice of the prior can also play a
-significant role in determining the optimal policy inferred by the SDP.
-In the examples shown here we have selected a prior that is both
-relatively uninformative (due to the broad priors placed on its
-parameters $\ell$ and $\sigma$ and simple (mean zero, radial basis
-function kernel). In practice, both the choice of mean and the
-covariance function may be chosen to confer particular biological
-properties, as well as more biologically informed priors for $\ell$ and
-$\sigma$. In principle, this may allow a manager to improve the
-performance of the GPDP by adding only enough additional detail as is
-justified. For instance, it would be possible to use a linear or a
-Ricker-shaped mean in the prior without making the much stronger
-assumption that the Ricker is the structurally correct model (Sugeno and
-Munch 2013a). Future research should focus on identifying criteria to
-ensure the prior and the value function are chosen appropriately for the
-problem at hand.
+Outside of the observed range of the data, the GP reverts to the prior,
+and consequently the choice of the prior can also play a significant
+role in determining the optimal policy inferred by the SDP. In the
+examples shown here we have selected a prior that is both relatively
+uninformative (due to the broad priors placed on its parameters $\ell$
+and $\sigma$ and simple (the choice of our covariance function, Eqns 12
+and 13 ). In practice, these should be chosen to confer particular
+biological properties. In principle, this may allow a manager to improve
+the performance of the GPDP by adding detail as is justified. For
+instance, it would be possible to use a linear or a Ricker-shaped mean
+in the prior without making the much stronger assumption that the Ricker
+is the structurally correct model (Sugeno and Munch 2013a). One fruitful
+avenue of future research is identifying criteria to ensure the prior
+and the value function are chosen appropriately for the problem at hand.
 
@@
 
-Hilborn, R. 2007. Reinterpreting the State of Fisheries and their
-Management. Ecosystems 10:1362--1369.
-
 Hilborn, R., and M. Mangel. 1997. The Ecological Detective: Confronting
@@
 Computer Modelling of Dynamical Systems 11:411--424.
-
-Lehmann, E. L. 1975. Nonparametrics: Statistical Methods Based on Ranks.
-Holden-Day, IncSan Francisco.
 
@@
 
-Worm, B., E. B. Barbier, N. Beaumont, J. E. Duffy, C. Folke, B. S.
-Halpern, J. B. C. Jackson, H. K. Lotze, F. Micheli, S. R. Palumbi, E.
-Sala, K. a Selkoe, J. J. Stachowicz, and R. Watson. 2006. Impacts of
-biodiversity loss on ocean ecosystem services. Science (New York, N.Y.)
-314:787--90.
-
-Worm, B., R. Hilborn, J. K. Baum, T. A. Branch, J. S. Collie, C.
-Costello, M. J. Fogarty, E. a Fulton, J. a Hutchings, S. Jennings, O. P.
-Jensen, H. K. Lotze, P. M. Mace, T. R. McClanahan, C. Minto, S. R.
-Palumbi, A. M. Parma, D. Ricard, A. a Rosenberg, R. Watson, and D.
-Zeller. 2009. Rebuilding global fisheries. Science (New York, N.Y.)
-325:578--85.
-
 \end{document}

### 2014-11-07 manuscripts/supplement.Rmd
update data file names to match figures

@@
 
+\tableofcontents
+
 \appendix
@@
 \setcounter{table}{0}
+
+<!--
+Table of contents? 
+-->
+
+# Code #
+
+All code used in producing this analysis has been embedded into the
+manuscript sourcefile using the Dynamic Documentation tool, `knitr`
+for the R language [@Xie_2013], available at 
+[github.com/cboettig/nonparametric-bayes/](https://github.com/cboettig/nonparametric-bayes/)
+
+To help make the mathematical and computational approaches presented
+here more accessible, we provide a free and open source (MIT License)
+R package that implements the GPDP process as it is presented here.
+Users should note that at this time, the R package has been developed
+and tested explicitly for this analysis and is not yet intended as a
+general purpose tool.  The manuscript source-code described above
+illustrates how these functions are used in this analysis. This
+package can be installed following the directions above.
+
+## Dependencies & Reproducibility ##
+
+The code provided should run on any common platform (Windows, Mac or Linux)
+that has R and the necessary R packages installed (including support for
+the jags Gibbs sampler).  The DESCRIPTION file of our R package, `nonparametricbayes`,
+lists all the software required to use these methods. Additional software requirements
+for the other comparisons shown here, such as the Gibbs sampling for the parametric models,
+are listed under the Suggested packages list.
+
+Nonetheless, installing the dependencies needed is not a trivial task, and may 
+become more difficult over time as software continues to evolve.  To facilitate
+reuse, we also provide a Dockerfile and Docker image that can be used to replicate
+and explore the analyses here by providing a copy of the computational environment
+we have used, with all software installed. Docker sofware (see [docker.com](http://www.docker.com))
+runs on most platforms as well as cloud servers. Use the command:
+
+    docker run -dP cboettig/nonparametric-bayes
+
+to launch an RStudio instance with the necessary software already installed. See the 
+Rocker-org project, [github.com/rocker-org](https://github.com/rocker-org) for more 
+detailed documentation on using Docker with R.
+
+# Data #
+
+While the data can be regenerated using the code provided, for convenience
+CSV files of the data shown in each graph are made available on Dryad.
+
+# Training data #
+
+Each of our models $f(S_t)$ must be estimated from training data, which
+we simulate from the Allen model with parameters $r = $ `r p[1]`, 
+$K =$ `r p[2]`, $C =$ `r p[3]`, and  $\sigma_g =$ `r sigma_g` for $T=$ `r Tobs`
+timesteps, starting at initial condition $X_0 = $ `r Xo`.  The training
+data can be seen in Figure 1 and found in the table `figure1.csv`.
+
+## Training data for sensitivity analyses ##
+
+A further 96 unique randomly generated training data sets are generated
+for the sensitivity analysis, as described in the main text.  The code
+provided replicates the generation of these sets.
+
+# Model performance outside the predicted range
 
@@
 
-write.csv(crash_data, "components/data/figure3.csv")
+write.csv(crash_data, "components/data/figureS1.csv")
 ```
+
+\newpage
+
+# MCMC analyses #
+
+This section provides figures and tables showing the traces from each of the MCMC runs used to estimate the parameters of the models presented, along with the resulting posterior distributions for each parameter.
+Priors usually appear completely falt when shown against the posteriors, but are summarized by tables the parameters of their corresponding distributions for each case.
 
@@
 gp_assessment_plots$posteriors_plot
+write.csv(gp_dat,"components/data/figureS2-3.csv")
 ```
@@
 \newpage
-
-Training data (Figures 1-6)
-----------------------------
-
-Each of our models $f(S_t)$ must be estimated from training data, which
-we simulate from the Allen model with parameters $r = $ `r p[1]`, 
-$K =$ `r p[2]`, $C =$ `r p[3]`, and  $\sigma_g =$ `r sigma_g` 
-for $T=$ `r Tobs` timesteps, starting at initial condition $X_0 = $ `r Xo`. 
-The training data can be seen in Figure 1.
-
-Training data for sensitivity analyses
--------------------------------------
-
-A further 96 unique training data sets are generated for the sensitivity analysis, as described in the main text.
-
-Code
-----
-
-All code used in producing this analysis can be found in the R package accompanying the paper, available from [http://github.com/cboettig/nonparametric-bayes].
-The code for all results shown here is dynamically embedded into the manuscript using `knitr` [@knitr].
-
-<!--
-Data
-----
-
-While the data can be regenerated using the code provided, for convenience CSV files of the data shown in each graph are also provided, along with appropriate metadata written in the Ecological Metadata Language (EML).
-
--->
 
@@
   facet_wrap(~ variable, scale="free", ncol=2)
-write.csv(ricker_posteriors, "components/data/ricker_posteriors.csv")
+write.csv(ricker_posteriors, "components/data/figureS4-5.csv")
 ```
@@
   facet_wrap(~ variable, scale="free", ncol=2)
-write.csv(myers_posteriors, "components/data/myers_posteriors.csv")
+write.csv(myers_posteriors, "components/data/figureS6-7.csv")
 ```
@@
   facet_wrap(~ variable, scale="free", ncol=2)
-write.csv(allen_posteriors, "components/data/allen_posteriors.csv")
+write.csv(allen_posteriors, "components/data/figureS8-9.csv")
 ```
@@
 
-Further sensitivity analysis
-=============================
+# Further sensitivity analysis #
 
@@
   facet_grid(noise~model)
-write.csv(sensitivity_dat, "components/data/sensitivity_dat.csv") 
+write.csv(sensitivity_dat, "components/data/figureS10.csv") 
 ```

### 2014-11-11 manuscripts/manuscript.Rmd
supplement must be built in same command after manuscript.

@@
 infer posterior distributions of the parameters of the GP (Figure
-S3, code in appendix). Since the posterior distributions differ
-substantially from the priors (Figure S3), most of the information in the posterior comes from the data rather than the
+S4, code in appendix). Since the posterior distributions differ
+substantially from the priors (Figure S4), most of the information in the posterior comes from the data rather than the
 prior belief.

### 2014-11-11 manuscripts/supplement.Rmd
supplement must be built in same command after manuscript.

@@
 
-```{r knit, include=FALSE}
+```{r knit, include=FALSE, eval=FALSE}
 library("methods")
@@
 
+## Dryad Data Archive ##
+
 While the data can be regenerated using the code provided, for convenience
@@
 
-# Training data #
+## Training data description ##
 
@@
 
-# Model performance outside the predicted range
+# Model performance outside the predicted range (Fig S1) #
 
@@
 
-```{r figure_S1, fig.width=8, fig.height=6, fig.cap="Outside the range of the training data (Figure 1), the true dynamics (black dots) fall outside the uncertainty (two standard deviations, colored bands) of the structurally incorrect parametric models (Ricker, Myers), but inside the uncertainty predicted by the GP.
+```{r figure_S1, fig.width=6, fig.height=4, fig.cap="Outside the range of the training data (Figure 1), the true dynamics (black dots) fall outside the uncertainty (two standard deviations, colored bands) of the structurally incorrect parametric models (Ricker, Myers), but inside the uncertainty predicted by the GP.
 Points show the stock size simulated by the true model.
@@
 
-# MCMC analyses #
-
-This section provides figures and tables showing the traces from each of the MCMC runs used to estimate the parameters of the models presented, along with the resulting posterior distributions for each parameter.
-Priors usually appear completely falt when shown against the posteriors, but are summarized by tables the parameters of their corresponding distributions for each case.
-
-```{r figure_S2, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options"}
-gp_assessment_plots$traces_plot
-```
-
-```{r figure_S3, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
-gp_assessment_plots$posteriors_plot
-write.csv(gp_dat,"components/data/figureS2-3.csv")
-```
-
-\newpage
-
-```{r figure_S4, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
-ggplot(ricker_posteriors) + 
-  geom_line(aes(index, value)) + # priors, need to fix order though
-  facet_wrap(~ variable, scale="free_y", ncol=1)
-```
-
-```{r figure_S5, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
-ggplot(ricker_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity") +
-  facet_wrap(~ variable, scale="free", ncol=2)
-write.csv(ricker_posteriors, "components/data/figureS4-5.csv")
-```
-
-```{r Table S1, results = "asis"}
-pander::pandoc.table(ricker_priors_xtable,
-  caption = "Parameterization range for the uniform priors in the Ricker model")
-```
-
-```{r figure_S6, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
-ggplot(myers_posteriors) + 
-  geom_line(aes(index, value)) + # priors, need to fix order though
-  facet_wrap(~ variable, scale="free_y", ncol=1)
-```
-
-```{r figure_S7, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
-ggplot(myers_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity") +
-  facet_wrap(~ variable, scale="free", ncol=2)
-write.csv(myers_posteriors, "components/data/figureS6-7.csv")
-```
-
-```{r TableS2, results="asis"}
-pander::pandoc.table(myers_priors_xtable,
-           caption = "Parameterization range for the uniform priors in the Myers model")
-```
-
-```{r figure_S8, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
-ggplot(allen_posteriors) + 
-  geom_line(aes(index, value)) + # priors, need to fix order though
-  facet_wrap(~ variable, scale="free_y", ncol=1)
-```
-
-```{r figure_S9, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
-ggplot(allen_posteriors, aes(value)) + 
-  stat_density(geom="path", position="identity") +
-  facet_wrap(~ variable, scale="free", ncol=2)
-write.csv(allen_posteriors, "components/data/figureS8-9.csv")
-```
-
-```{r TableS3, results = "asis"}
-pander::pandoc.table(allen_priors_xtable,
-  caption = "Parameterization range for the uniform priors in the Allen model")
-```
-
-# Further sensitivity analysis #
+# Further sensitivity analysis (Fig S2) #
 
@@
 
-```{r figure_S10, fig.height=6, fig.width=10, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
+```{r figure_S2, fig.height=6, fig.width=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
 Histograms shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters.
@@
   facet_grid(noise~model)
-write.csv(sensitivity_dat, "components/data/figureS10.csv") 
+write.csv(sensitivity_dat, "components/data/figureS2.csv") 
 ```
@@
 ```{r AllenSetsTable, results="asis"}
-pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Allen models in Figure S9." )
+pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Allen models in Figure S2." )
 ```
 ```{r MyersSetsTable, results="asis"}
-pandoc.table(MyersParameterSets, caption="Randomly chosen parameter sets for the Myers models in Figure S9." )
+pandoc.table(MyersParameterSets, caption="Randomly chosen parameter sets for the Myers models in Figure S2." )
+```
+
+\newpage
+
+# MCMC analyses #
+
+This section provides figures and tables showing the traces from each of the MCMC runs used to estimate the parameters of the models presented, along with the resulting posterior distributions for each parameter.
+Priors usually appear completely falt when shown against the posteriors, but are summarized by tables the parameters of their corresponding distributions for each case.
+
+## GP MCMC (Fig S3-4) ##
+
+```{r figure_S3, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options", fig.width=4, fig.height=4}
+gp_assessment_plots$traces_plot
+```
+
+```{r figure_S4, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red.", fig.width=4, fig.height=4}
+gp_assessment_plots$posteriors_plot
+write.csv(gp_dat,"components/data/figureS3-4.csv")
+```
+
+\newpage
+\newpage
+
+## Ricker Model MCMC (Fig S5-6) ##
+
+```{r figure_S5, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
+ggplot(ricker_posteriors) + 
+  geom_line(aes(index, value)) + # priors, need to fix order though
+  facet_wrap(~ variable, scale="free_y", ncol=1)
+```
+
+```{r figure_S6, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=4, fig.height=4}
+ggplot(ricker_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity") +
+  facet_wrap(~ variable, scale="free", ncol=2)
+write.csv(ricker_posteriors, "components/data/figureS5-6.csv")
+```
+
+```{r Table S1, results = "asis"}
+pander::pandoc.table(ricker_priors_xtable,
+  caption = "Parameterization range for the uniform priors in the Ricker model")
+```
+
+\newpage
+\newpage
+
+## Myers Model MCMC (Fig S7-8) ##
+
+```{r figure_S7, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
+ggplot(myers_posteriors) + 
+  geom_line(aes(index, value)) + # priors, need to fix order though
+  facet_wrap(~ variable, scale="free_y", ncol=1)
+```
+
+```{r figure_S8, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=4, fig.height=6}
+ggplot(myers_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity") +
+  facet_wrap(~ variable, scale="free", ncol=2)
+write.csv(myers_posteriors, "components/data/figureS7-8.csv")
+```
+
+```{r TableS2, results="asis"}
+pander::pandoc.table(myers_priors_xtable,
+           caption = "Parameterization range for the uniform priors in the Myers model")
+```
+
+\newpage 
+\newpage
+
+## Allen Model MCMC (Fig S9-10) ##
+
+```{r figure_S9, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
+ggplot(allen_posteriors) + 
+  geom_line(aes(index, value)) + # priors, need to fix order though
+  facet_wrap(~ variable, scale="free_y", ncol=1)
+```
+
+```{r figure_S10, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=4, fig.height=6}
+ggplot(allen_posteriors, aes(value)) + 
+  stat_density(geom="path", position="identity") +
+  facet_wrap(~ variable, scale="free", ncol=2)
+write.csv(allen_posteriors, "components/data/figureS9-10.csv")
+```
+
+```{r TableS3, results = "asis"}
+pander::pandoc.table(allen_priors_xtable,
+  caption = "Parameterization range for the uniform priors in the Allen model")
 ```

### 2014-11-11 manuscripts/manuscript.tex

@@
 (2003)) to infer posterior distributions of the parameters of the GP
-(Figure S3, code in appendix). Since the posterior distributions differ
-substantially from the priors (Figure S3), most of the information in
+(Figure S4, code in appendix). Since the posterior distributions differ
+substantially from the priors (Figure S4), most of the information in
 the posterior comes from the data rather than the prior belief.

### 2014-11-12 manuscripts/supplement.Rmd
rebuild final data files for uploading

@@
 While the data can be regenerated using the code provided, for convenience
-CSV files of the data shown in each graph are made available on Dryad.
+CSV files of the data shown in each graph are made available on Dryad,
+along with the source `.Rmd` files for the manuscript and supplement
+that document them.
 
@@
 write.csv(crash_data, "components/data/figureS1.csv")
+```
+
+```{r S1_eml, eval=FALSE}
+## Forgo the EML generation, the Rmd files provide better documentation at this stage.
+library(EML)
+me <- "Carl Boettiger <cboettig@ropensci.org>"
+models <- levels(crash_data$variable)
+names(models) <- models
+col.defs <- c("time", "fish stock", "model", "value", "replicate")
+unit.defs <- list("number", "number", models, "number", "number")
+col.classes <- sapply(crash_data, class)
+eml_write("components/data/figureS1.csv", col.classes = col.classes, col.defs=col.defs,unit.defs=unit.defs,creator=me, title="Figure S1", file="components/data/README_S1.xml")
 ```
@@
 ```
+```{r S2_eml, eval=FALSE}
+# not descriptive enough, supplement text documents this better
+par <- levels(sensitivity_dat$parameters)
+names(par) <- par
+col.defs <- names(sensitivity_dat)
+unit.defs <- list(models, par, "number", "number",  "number", "number")
+col.classes <- sapply(sensitivity_dat, class)
+eml_write("components/data/figureS2.csv", col.classes = col.classes, col.defs=col.defs,unit.defs=unit.defs,creator=me, title="Figure S2", file="components/data/README_S2.xml")
+```
 
@@
 
-```{r figure_S4, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red.", fig.width=4, fig.height=4}
+```{r figure_S4, fig.cap="Posterior distributions from the MCMC estimate of the GP model.
+Prior curves shown in red; note the posterior distributions are significantly more peaked than the priors, showing that the data has been informative and is not driven by the priors.", fig.width=4, fig.height=4}
 gp_assessment_plots$posteriors_plot
-write.csv(gp_dat,"components/data/figureS3-4.csv")
+write.csv(tgp_dat,"components/data/figureS3-4.csv")
+```
+
+```{r S2_eml}
+par <- levels(sensitivity_dat$parameters)
+names(par) <- par
+col.defs <- names(sensitivity_dat)
+unit.defs <- list(models, par, "number", "number",  "number", "number")
+col.classes <- sapply(sensitivity_dat, class)
+eml_write("components/data/figureS2.csv", col.classes = col.classes, col.defs=col.defs,unit.defs=unit.defs,creator=me, title="Figure S2", file="components/data/README_S2.xml")
 ```
@@
 
-```{r figure_S5, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
+```{r figure_S5, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps).
+stdQ refers to the estimate of $\\sigma$; deviance is -2 times the log likelihood.
+", fig.width=4}
 ggplot(ricker_posteriors) + 
@@
 
-```{r figure_S6, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=4, fig.height=4}
+```{r figure_S6, fig.cap="Posteriors from the MCMC estimate of the Ricker model.
+Note that the model estimates a carrying capacity $K$ very close to the true equilibrium where most of the observations were made, but is less certain about the growth rate.", fig.width=4, fig.height=4}
 ggplot(ricker_posteriors, aes(value)) + 
@@
 
-```{r figure_S7, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
+```{r figure_S7, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps).
+stdQ refers to the estimate of $\\sigma$; deviance is -2 times the log likelihood.", fig.width=4}
 ggplot(myers_posteriors) + 
@@
 
-```{r figure_S8, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=4, fig.height=6}
+```{r figure_S8, fig.cap="Posterior distributions from the MCMC estimates of the Myers model.
+Note that with more free parameters, the posteriors reflect greater uncertainty.
+In particular, the parameter $\\theta$ includes values both above 2, resulting in a tipping point, and below 2, where no tipping point exists in the model.
+Though the dynamic program will integrate over the full distribution, including those values corresponding to tipping points, the weight of the model lies in the region without tipping points.", fig.width=4, fig.height=6}
 ggplot(myers_posteriors, aes(value)) + 
@@
 
-```{r figure_S9, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=4}
+```{r figure_S9, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps).
+stdQ refers to the estimate of $\\sigma$; deviance is -2 times the log likelihood.", fig.width=4}
 ggplot(allen_posteriors) + 
@@
 
-```{r figure_S10, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=4, fig.height=6}
+```{r figure_S10, fig.cap="Posteriors from the MCMC estimate of the Allen model.
+The Allen model is the structurally correct model.
+Despite potential identfiability issues in distinguishing between the stable and unstable points ($K$ and $\\theta$ respectively), the posterior estimates successfully reflect both the the upper stable point ($K$), as well as the significant probabilty of a tipping point ($\\theta$) somewhere between $K$ and extinction (0).", fig.width=4, fig.height=6}
 ggplot(allen_posteriors, aes(value)) + 

### 2014-11-12 manuscripts/supplement.tex

\documentclass[]{components/elsarticle}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
\usepackage{graphicx}
% Redefine \includegraphics so that, unless explicit options are
% given, the image width will not exceed the width of the page.
% Images get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\ScaleIfNeeded{%
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother
\let\Oldincludegraphics\includegraphics
{%
 \catcode`\@=11\relax%
 \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
}%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Supplement for: Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

%%% Change title format to be more compact
\usepackage{titling}
\setlength{\droptitle}{-2em}
  \title{Supplement for: Avoiding tipping points in fisheries management through
Gaussian Process Dynamic Programming}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{true \and true \and true}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \date{}
  \predate{}\postdate{}

\begin{document}

\maketitle

\tableofcontents

\appendix
\renewcommand*{\thefigure}{S\arabic{figure}}
\renewcommand*{\thetable}{S\arabic{table}} \setcounter{figure}{0}
\setcounter{table}{0}

\section{Code}\label{code}

All code used in producing this analysis has been embedded into the
manuscript sourcefile using the Dynamic Documentation tool,
\texttt{knitr} for the R language (Xie 2013), available at
\href{https://github.com/cboettig/nonparametric-bayes/}{github.com/cboettig/nonparametric-bayes/}

To help make the mathematical and computational approaches presented
here more accessible, we provide a free and open source (MIT License) R
package that implements the GPDP process as it is presented here. Users
should note that at this time, the R package has been developed and
tested explicitly for this analysis and is not yet intended as a general
purpose tool. The manuscript source-code described above illustrates how
these functions are used in this analysis. This package can be installed
following the directions above.

\subsection{Dependencies \&
Reproducibility}\label{dependencies-reproducibility}

The code provided should run on any common platform (Windows, Mac or
Linux) that has R and the necessary R packages installed (including
support for the jags Gibbs sampler). The DESCRIPTION file of our R
package, \texttt{nonparametricbayes}, lists all the software required to
use these methods. Additional software requirements for the other
comparisons shown here, such as the Gibbs sampling for the parametric
models, are listed under the Suggested packages list.

Nonetheless, installing the dependencies needed is not a trivial task,
and may become more difficult over time as software continues to evolve.
To facilitate reuse, we also provide a Dockerfile and Docker image that
can be used to replicate and explore the analyses here by providing a
copy of the computational environment we have used, with all software
installed. Docker sofware (see \href{http://www.docker.com}{docker.com})
runs on most platforms as well as cloud servers. Use the command:

\begin{verbatim}
docker run -dP cboettig/nonparametric-bayes
\end{verbatim}

to launch an RStudio instance with the necessary software already
installed. See the Rocker-org project,
\href{https://github.com/rocker-org}{github.com/rocker-org} for more
detailed documentation on using Docker with R.

\section{Data}\label{data}

\subsection{Dryad Data Archive}\label{dryad-data-archive}

While the data can be regenerated using the code provided, for
convenience CSV files of the data shown in each graph are made available
on Dryad.

\subsection{Training data description}\label{training-data-description}

Each of our models $f(S_t)$ must be estimated from training data, which
we simulate from the Allen model with parameters $r = $ 2, $K =$ 8,
$C =$ 5, and $\sigma_g =$ 0.05 for $T=$ 40 timesteps, starting at
initial condition $X_0 = $ 5.5. The training data can be seen in Figure
1 and found in the table \texttt{figure1.csv}.

\subsection{Training data for sensitivity
analyses}\label{training-data-for-sensitivity-analyses}

A further 96 unique randomly generated training data sets are generated
for the sensitivity analysis, as described in the main text. The code
provided replicates the generation of these sets.

\section{Model performance outside the predicted range (Fig
S1)}\label{model-performance-outside-the-predicted-range-fig-s1}

Figure S1 illustrates the performance of the GP and parametric models
outside the observed training data. The mean trajectory under the
underlying model is shown by the black dots, while the corresponding
prediction made by the model shown by the box and whiskers plots.
Predictions are based on the true expected value in the previous time
step. Predicted distributions that lie entirely above the expected
dynamics indicate the expectation of stock sizes higher than what is
actually expected. The models differ both in their expectations and
their uncertainty (colored bands show two standard deviations away).
Note that the GP is particularly uncertain about the dynamics relative
to structurally incorrect models like the Ricker.

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S1-1.pdf}
\caption{Outside the range of the training data (Figure 1), the true
dynamics (black dots) fall outside the uncertainty (two standard
deviations, colored bands) of the structurally incorrect parametric
models (Ricker, Myers), but inside the uncertainty predicted by the GP.
Points show the stock size simulated by the true model. Overlay shows
the range of states predicted by each model, based on the state observed
in the previous time step. The Ricker model always (wrongly) predicts
positive population growth, while the actual population shrinks in each
step as the initial condition falls below the Allee threshold of the
underlying model (Allen). Note that because it does not assume a
parametric form but instead relies more directly on the data, the GP is
both more pessimistic and more uncertain about the future state than the
parametric models.}
\end{figure}

\newpage

\section{Further sensitivity analysis (Fig
S2)}\label{further-sensitivity-analysis-fig-s2}

The Latin hypercube approach systematically varies all combinations of
parameters, providing a more general test than varying only one
parameter at a time. We loop across eight replicates of three different
randomly generated parameter sets for each of two different generating
models (Allen and Myers) over two different noise levels (0.01 and
0.05), for a total of 8 x 3 x 2 x 2 = 96 scenarios. The Gaussian Process
performs nearly optimally in each case, relative to the optimal solution
with no parameter or model uncertainty (Figure S10, appendix).

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S2-1.pdf}
\caption{Sensitivity Analysis. Histograms shows the ratio of the
realized net present value derived when managing under the GPDP over the
optimal value given the true model and true parameters. Values of 1
indicate optimal performance. Columns indicate different models, rows
different noise levels, and colors indicate the parameter set used.
Grouped over stochastic replicates applying the contol policy and
stochastic replicates of training data generated from the model
indicated, see raw data for details. Randomly chosen parameter values
for the models shown in tables below.}
\end{figure}

\begin{longtable}[c]{@{}cccc@{}}
\toprule\addlinespace
\begin{minipage}[b]{0.15\columnwidth}\centering
~
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
r
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
K
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\centering
theta
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.15\columnwidth}\centering
\textbf{set.A}
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
1.103
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
7.949
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\centering
2.288
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
\textbf{set.B}
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
1.485
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
9.775
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\centering
3.524
\end{minipage}
\\\addlinespace
\bottomrule
\addlinespace
\caption{Randomly chosen parameter sets for the Allen models in Figure
S2.}
\end{longtable}

\begin{longtable}[c]{@{}cccc@{}}
\toprule\addlinespace
\begin{minipage}[b]{0.15\columnwidth}\centering
~
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
r
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
K
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering
C
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.15\columnwidth}\centering
\textbf{set.C}
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
1.769
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
10.46
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
4.301
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
\textbf{set.D}
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
2.075
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
10.95
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering
4.915
\end{minipage}
\\\addlinespace
\bottomrule
\addlinespace
\caption{Randomly chosen parameter sets for the Myers models in Figure
S2.}
\end{longtable}

\newpage

\section{MCMC analyses}\label{mcmc-analyses}

This section provides figures and tables showing the traces from each of
the MCMC runs used to estimate the parameters of the models presented,
along with the resulting posterior distributions for each parameter.
Priors usually appear completely falt when shown against the posteriors,
but are summarized by tables the parameters of their corresponding
distributions for each case.

\subsection{GP MCMC (Fig S3-4)}\label{gp-mcmc-fig-s3-4}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S3-1.pdf}
\caption{Traces from the MCMC estimates of the GP model show reasonable
mixing (no trend) and sampling rejection rate (no piecewise jumps)}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S4-1.pdf}
\caption{Posterior distributions from the MCMC estimate of the GP model.
Prior curves shown in red; note the posterior distributions are
significantly more peaked than the priors, showing that the data has
been informative and is not driven by the priors.}
\end{figure}

\newpage
\newpage

\subsection{Ricker Model MCMC (Fig
S5-6)}\label{ricker-model-mcmc-fig-s5-6}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S5-1.pdf}
\caption{Traces from the MCMC estimates of the Ricker model show
reasonable mixing (no trend) and sampling rejection rate (no piecewise
jumps). stdQ refers to the estimate of $\sigma$; deviance is -2 times
the log likelihood.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S6-1.pdf}
\caption{Posteriors from the MCMC estimate of the Ricker model. Note
that the model estimates a carrying capacity $K$ very close to the true
equilibrium where most of the observations were made, but is less
certain about the growth rate.}
\end{figure}

\begin{longtable}[c]{@{}ccc@{}}
\toprule\addlinespace
\begin{minipage}[b]{0.15\columnwidth}\centering
parameter
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
lower.bound
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
upper.bound
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.15\columnwidth}\centering
$r$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
0.01
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
20
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$K$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
0.01
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
40
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$\sigma$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
1e-06
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
100
\end{minipage}
\\\addlinespace
\bottomrule
\addlinespace
\caption{Parameterization range for the uniform priors in the Ricker
model}
\end{longtable}

\newpage
\newpage

\subsection{Myers Model MCMC (Fig
S7-8)}\label{myers-model-mcmc-fig-s7-8}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S7-1.pdf}
\caption{Traces from the MCMC estimates of the Myers model show
reasonable mixing (no trend) and sampling rejection rate (no piecewise
jumps). stdQ refers to the estimate of $\sigma$; deviance is -2 times
the log likelihood.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S8-1.pdf}
\caption{Posterior distributions from the MCMC estimates of the Myers
model. Note that with more free parameters, the posteriors reflect
greater uncertainty. In particular, the parameter $\theta$ includes
values both above 2, resulting in a tipping point, and below 2, where no
tipping point exists in the model. Though the dynamic program will
integrate over the full distribution, including those values
corresponding to tipping points, the weight of the model lies in the
region without tipping points.}
\end{figure}

\begin{longtable}[c]{@{}ccc@{}}
\toprule\addlinespace
\begin{minipage}[b]{0.15\columnwidth}\centering
parameter
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
lower.bound
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
upper.bound
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.15\columnwidth}\centering
$r$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
1e-04
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
10
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$K$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
1e-04
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
40
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$\theta$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
1e-04
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
10
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$\sigma$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
1e-06
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
100
\end{minipage}
\\\addlinespace
\bottomrule
\addlinespace
\caption{Parameterization range for the uniform priors in the Myers
model}
\end{longtable}

\newpage 
\newpage

\subsection{Allen Model MCMC (Fig
S9-10)}\label{allen-model-mcmc-fig-s9-10}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S9-1.pdf}
\caption{Traces from the MCMC estimates of the Allen model show
reasonable mixing (no trend) and sampling rejection rate (no piecewise
jumps). stdQ refers to the estimate of $\sigma$; deviance is -2 times
the log likelihood.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/supplement-figure_S10-1.pdf}
\caption{Posteriors from the MCMC estimate of the Allen model. The Allen
model is the structurally correct model. Despite potential
identfiability issues in distinguishing between the stable and unstable
points ($K$ and $\theta$ respectively), the posterior estimates
successfully reflect both the the upper stable point ($K$), as well as
the significant probabilty of a tipping point ($\theta$) somewhere
between $K$ and extinction (0).}
\end{figure}

\begin{longtable}[c]{@{}ccc@{}}
\toprule\addlinespace
\begin{minipage}[b]{0.15\columnwidth}\centering
parameter
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
lower.bound
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
upper.bound
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.15\columnwidth}\centering
$r$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
0.01
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
6
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$K$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
0.01
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
20
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$X_C$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
0.01
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
20
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$\sigma$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
1e-06
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
100
\end{minipage}
\\\addlinespace
\bottomrule
\addlinespace
\caption{Parameterization range for the uniform priors in the Allen
model}
\end{longtable}

Xie, Y. 2013. Dynamic Documents with R and knitr. Chapman; Hall/CRCBoca
Raton, Florida.

\end{document}

### 2014-11-14 manuscripts/manuscript.Rmd
update Dockerfile

@@
 ```{r figure_3, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S(t)$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", fig.width=8, fig.height=6, cache=FALSE}
+Policies are shown in terms of target escapement, $S(t)$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979] Several policies show a numerical jitter due to the discritization of states in the dynamic programming algorithm; doubling the number of grid points did not qualitatively change the results.", fig.width=8, fig.height=6, cache=FALSE}
 policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
@@
 ggplot(policies, aes(stock, stock - value, color=method)) +
-  geom_line(lwd=1.2) + 
+  geom_line() + 
+  # stat_smooth(lwd=1.2, method="loess", degree=1, span=0.2, level=0, n=50) + 
   facet_wrap(~method) +

### 2014-11-14 manuscripts/manuscript.tex
update Dockerfile

@@
 $S(t)$, as under models such as this a constant escapement policy is
-expected to be optimal (Reed 1979).}
+expected to be optimal (Reed 1979) Several policies show a numerical
+jitter due to the discritization of states in the dynamic programming
+algorithm; doubling the number of grid points did not qualitatively
+change the results.}
 \end{figure}

### 2014-11-14 manuscripts/manuscript.Rmd
include pdf

@@
 ```{r figure_3, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S(t)$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979] Several policies show a numerical jitter due to the discritization of states in the dynamic programming algorithm; doubling the number of grid points did not qualitatively change the results.", fig.width=8, fig.height=6, cache=FALSE}
+Policies are shown in terms of target escapement, $S(t)$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979] Several policies show a numerical jitter due to the discretization of states in the dynamic programming algorithm; doubling the number of grid points did not qualitatively change the results.", fig.width=8, fig.height=6, cache=FALSE}
 policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")

### 2014-11-14 manuscripts/manuscript.tex
include pdf

--- 2014-11-11 manuscripts/manuscript.tex
@@
 $S(t)$, as under models such as this a constant escapement policy is
-expected to be optimal (Reed 1979).}
+expected to be optimal (Reed 1979) Several policies show a numerical
+jitter due to the discretization of states in the dynamic programming
+algorithm; doubling the number of grid points did not qualitatively
+change the results.}
 \end{figure}

### 2014-11-14 manuscripts/supplement.Rmd
remove duplicate chunk

@@
 
-```{r S2_eml}
-par <- levels(sensitivity_dat$parameters)
-names(par) <- par
-col.defs <- names(sensitivity_dat)
-unit.defs <- list(models, par, "number", "number",  "number", "number")
-col.classes <- sapply(sensitivity_dat, class)
-eml_write("components/data/figureS2.csv", col.classes = col.classes, col.defs=col.defs,unit.defs=unit.defs,creator=me, title="Figure S2", file="components/data/README_S2.xml")
-```
-
 \newpage

### 2014-11-20 manuscripts/manuscript.Rmd
as submitted

@@
 
-<!-- Run R code for analysis, to be called by figures -->
 ```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
@@
                "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-opts_chunk$set(dev='pdf', dev.args=c(version="1.7"))
+opts_chunk$set(dev='pdf', dev.args=c(version="1.4"))
 opts_chunk$set(fig.width=4, fig.height=3, echo=FALSE)
@@
 under uncertainty [@Gordon1954;
-@Clark1976; -@Clark2009; @Reed1979; @May1979; @Ludwig1982]
-
-<!-- Uncertainty outside the data without the correct model has not been handled. -->
-<!-- __We don't have the model__ -->
+@Clark1976; -@Clark2009; @Reed1979; @May1979; @Ludwig1982].
 
@@
 
-<!-- Point is: __We don't have the data where we need it__ -->
-<!-- What do we call this?  Extrapolation uncertainty?  Pathological Uncertainty? -->
-
 Structural  uncertainty is particularly insidious when we try to predict
@@
 even small errors can have
-very large consequences, as we shall illustrate later.  Because intervention is often (but not always, see
-@Hughes2013) too late after a tipping point has been crossed, management
+very large consequences, as we shall illustrate later.  Because intervention may be too late once a 
+tipping point has been crossed (but see @Hughes2013), management
 is often concerned with avoiding 
-tipping points before any data about them is available or the 
-transition that would more clearly reveal a tipping point occurs
-[e.g. @Bestelmeyer2012].
+tipping points before any data about them are available.
 
@@
 
-<!--
- _Note on "not magic"_: honest uncertainty + SDP
-The idea that any approach can perform well without either having to
-know the model or have particularly good data should immediately draw
-suspicion.  The reader must bear in mind that the strength of our approach
-comes not from black-box predictive power from such limited information,
-but rather, by providing a more honest expression of uncertainty outside
-the observed data without sacrificing the predictive capacity near the
-observed data. By coupling this more accurate description of what is known
-and unknown to the decision-making under uncertainty framework provided
-by stochastic dynamic programming, we are able to obtain more robust
-management policies than with common parametric modeling approaches.
--->
-
 For comparisons, we consider the performance of management based on  GPDP against
@@
 of the decision-maker's objectives.
-
-<!--
-
-Although much argument can be made
-over the 'correct' objective function, e.g. how to account for the
-social value of fish left in the sea vs. the commercial value of fish
-harvested; see @Halpern2013 for further discussion of this issue.  We can
-always compare model performance across multiple potential objectives.
-The important point is that a decision-maker does not necessarily need
-a model that provides the best mechanistic understanding or the best
-long-term outcome, but rather the one that delivers the best management
-performance over the period of interest.
--->
 
@@
 simulate underlying dynamics from one of the three parametric models
-, whereas in the natural world we never know the "true" mode
-.  In addition, by choosing one-dimensional models with just a few parameters, we
+, whereas in the natural world we never know the "true" model.
+In addition, by choosing one-dimensional models with just a few parameters, we
 limit the chance that poor performance will be due to inability to
@@
 for higher-dimensional parametric models. Finally, the parametric models
-we consider are those most commonly used in modeling stock-recruitment
+we consider are commonly used in modeling stock-recruitment
 dynamics or to model sudden transitions between alternative stable states.
@@
 ### Parametric models
-
-<!-- 
-- Statement of the models
--->
 
@@
 population at time $t$ and assume that in the absence of take its
-dynamics are.
-.
+dynamics are:
 
 \begin{equation}
-X(t+1) = Z(t) f(X(t), p) \label{eq1} 
+X(t+1) = Z(t) f(X(t), \mathbf{p}) \label{eq1} 
 \end{equation}
@@
 Where Z(t) is log-normally distributed process stochasticity [@Reed1979]
-and $p$ is a vector of parameters to be estimated from the data.
-We describe the three choices for $f(X(t),p)$ in the next section.  
+and $\mathbf{p}$ is a vector of parameters to be estimated from the data.
+We describe the three choices for $f(X(t),\mathbf{p})$ in the next section.  
 The control action is a harvest or take, $h(t)$, measured
@@
 and a final time $T$. That is, we seek to maximize over choices of
-harvest $E_{X(t+1)} [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$,
+harvest $E [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$,
 where the state dynamics are given by Eqn 1 and $E$ denotes the
-expectation over the process stochasticity in the future population state.
+expectation over future population states.
 
@@
 \begin{equation}
-\mathbf{\mathrm{E}}_{X(t+1)} \left[ V( X(t+1),t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{Z(t)} \left[V( Z(t) f(x(t) - h(t))|p), t+1 | x(t), h(t) \right]
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V( X(t+1),t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{Z(t)} \left[V( Z(t) f(x(t) - h(t))|\mathbf{p}), t+1 | x(t), h(t) \right]
 \end{equation}
@@
 where the population size after the take is $x(t) - h(t)$, which is then
-translated into $X(t+1)$ by Eqn 1 (that is, we replace $X(t+1)$ by $Z(t) f(x(t)-h(t)|p)$).
+translated into $X(t+1)$ by Eqn 1 (that is, we replace $X(t+1)$ by $Z(t) f(x(t)-h(t)|\mathbf{p})$).
 
@@
 \begin{equation}
-\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{p|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z(t) | p, \mathrm{data}} \left[ V(Z(t) f(x(t) - h(t)|p), t)  \right] \}
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{\mathbf{p}|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z(t) | \mathbf{p}, \mathrm{data}} \left[ V(Z(t) f(x(t) - h(t)|\mathbf{p}), t)  \right] \}
 \end{equation}
@@
 \begin{equation}
-\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{p|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z(t) | p, \mathrm{data}} \left[ V( Z(t) f(x(t) - h(t)|p), t) \right] \}
+\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{\mathbf{p}|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z(t) | \mathbf{p}, \mathrm{data}} \left[ V( Z(t) f(x(t) - h(t)| \mathbf{p}), t) \right] \}
 \end{equation}
@@
 
-We consider a period of `r Tobs` in which data are obtained to estimate the parameters or the GP. This is long
-enough that the estimates do not depend on the particular realization,
-and longer times are not likely to provide substantial improvement.  
-We fit each of the models is fit to the same data (Figure 1).
-
-<!-- Bayesian inference of parametric models -->
+We consider a period of `r Tobs` in which data are obtained to estimate
+the parameters or the GP. This is long enough that the estimates do not
+depend on the particular realization, and longer times are not likely
+to provide substantial improvement.  Each of the models is fit to the
+same data (Figure 1).
 
@@
 
+Additionally, we compute the maximum likelihood estimate (MLE, as we will refer to 
+this model in the figures) of the parameters for the (structurally correct) Allen model.
+Comparing this to using the posterior distribution of parameters inferred from MCMC for
+the same model gives some indication of the importance of this uncertainty in the 
+dynamic programming.
+
 ### The Gaussian Process model
 
-The core difference for our purpose between the GP and the parametric
-models is that the GP model is defined explicitly by the observed data.  That is, it cannot be specified by the value of parameters alone.
-
-<!-- Statement of model -->
+The core difference for our purpose between the estimated GP and the estimated parametric
+models is that the estimated GP model is defined explicitly in reference to the observed data.
+As a result, uncertainty arises in the GP model not only from uncertainty in the parameters,
+but is also increases in regions farther from the observed states, such as low population
+sizes in the example illustrated here.  The estimated parametric models, by contrast, are
+completely specified by the parameters.
 
@@
 
-A GP is a 
-stochastic process for which any realization of for
-a  finite sample of points $n$ is a multivariate normal distribution
-of dimension $n$. Thus, to characterize the GP we need a mean function
-and a covariance function. We proceed as follows.
-
-As before, we assume that the data $X(t)$ are observed with process stochasticity around a mean function $ g(X(t))$
+A GP is a stochastic process for which any realization consisting of n
+points follows a multivariate normal distribution of dimension $n$. To
+characterize the GP we need a mean function and a covariance function. We
+proceed as follows.
+
+As before, we assume that the data $X(t)$ are observed with process
+stochasticity around a mean function $g(X(t))$
 
@@
 
-where $\varepsilon$ are IID normal random variables with zero-mean and variance $\sigma_n^2$. Note that we have chosen to
+where $\varepsilon$ are IID normal random variables with zero-mean and variance $\sigma^2$. Note that we have chosen to
 assume additive stochasticity.  While we could consider log-normal stochasticity
@@
 
-Now consider the situation where we have a set of  previous observations of the _timeseries_, $X_{\textrm{obs}}(t)$,
-and wish to _predict_ future values $X_{\textrm{p}}(t)$.  We consider the observations as pairs of points
-$X_{\textrm{obs}}(t), X_{\textrm{obs}}(t+1)$ for $t=1, 2, …$ up through final observation time $T$.
-As described above, in the  GP, a finite sample of these predictions at $n$ points into the future, $X_{\textrm p}(\tau)$ for $\tau \in [T+1, T+2, ..., T+n]$, is a multivariate normal ($\mathcal{N}$):
+In order to make predictions, we update the GP based on the observed set of transitions.
+To do so, we collect the time series of observed states into a vector of "current" states, $\mathbf{X}_{\textrm{obs}} = \{X(1), \dots, X(T-1)\}$ and a vector of "next" states $\mathbf{Y}_{\textrm{obs}} = \{X(2),\dots,X(T)\}$ where $T$ is the time of the final observation.
+Conditional on these observations, the predicted next state, $g(X_p)$, for any given "current" state, $X_p$ follows a normal distribution with mean $E$ and variance $C$ determined using the standard rules for conditioning in multivariate normals, i.e.
 
 \begin{equation}
-g(X_p|X_{\textrm{obs}}(t), X_{\textrm{obs}}(t+1)) \sim \mathcal{N}(E,C)
+E = K(\mathbf{X}_{\textrm p}, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I}_n \right)^{-1} \mathbf{Y}_{\textrm{obs}}
 \end{equation}
 
-with mean and covariance matrix given by (see Eqn 2.22 of @Rasmussen2006 for a more detailed derivation)
+and
 
 \begin{equation}
-E = K(X_{\textrm p}(\tau), X_{\textrm{obs}}(t)) \left(K(X_{\textrm{obs}}(t),X_{\textrm{obs}}(t)) - \sigma \mathbb{I} \right)^{-1} X_{\textrm{obs}}(t+1)
+C = K(\mathbf{X}_{\textrm p}, \mathbf{X}_{\textrm p}) - K(\mathbf{X}_{\textrm p}, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, \mathbf{X}_{\textrm p})
 \end{equation}
 
+Here $\mathbb{I}_n is the $n$ by $n$ identity matrix (i.e.
+a matrix with ones down the diagonal and zeros elsewhere) and $K$ is the ‘covariance kernel.’  The covariance kernel controls how much influence one observation has on another.
+In the present application we use the squared-exponential kernel which, when evaluated over a pair of vectors, say $x$ and $y$,  generates a covariance matrix whose $i,j$th element is given by
+
 \begin{equation}
-C = K(X_{\textrm p}(\tau), X_{\textrm p}(\tau)) - K(X_{\textrm p}(\tau), X_{\textrm{obs}}(t)) \left(K(X_{\textrm{obs}}(t),X_{\textrm{obs}}(t)) - \sigma \mathbb{I} \right)^{-1} K(X_{\textrm{obs}}(t), X_{\textrm p}(\tau))
+K_{i,j}(\mathbf{x}, \mathbf{y}) = \exp\left( \frac{ -(x_i - y_j)^2}{2 \ell^2} \right)
 \end{equation}
 
-Where $K$, called the covariance kernel, measures how much influence one observation has on another.  In this paper, for $K$ use a matrix whose $(i,j)$th element is given by:
-
-\begin{equation}
-K_{i,j}(x_i, y_j) = \exp\left( \frac{ -(x_i - y_j)^2}{2 \ell^2} \right)
-\end{equation}
 so that $\ell$ gives the characteristic length-scale over which correlation between
-two observations decays.
-
-We use inverse Gamma priors on both the length-scale $\ell$ and $\sigma$, thus for example
+two observations decays. See @Rasmussen2006 for other choices of covariance kernels and their properties.
+Note that this simple formulation assumes a prior mean of zero. For the parameters we use inverse Gamma priors on both the length-scale $\ell$ and $\sigma$, thus for example
 
@@
 For the prior on  $\ell$,  $\alpha = $ `r d.p[1]` and $\beta = $ `r d.p[2]`.  The prior  on $\sigma$,  $\alpha = $ `r s2.p[1]` and $\beta = $ `r s2.p[2]`.
-
-<!-- Inference of the BNP model -->
 
@@
 To ensure our results are robust to the choice of parameters, we
-consider 96 different scenarios, described in detail below.   To help  better
+will consider 96 different scenarios.  To help better
 understand the process, we first describe in detail the
@@
 ```{r figure_1, dependson=c("assemble-models", "par-fns", "plot-options"), fig.cap="Points show the training data of stock-size over time.
-Curves show the posterior step-ahead predictions based on each of the estimated models.
-Observe that all models are fitting the data reasonably well.
-In the GP panel, black lines are means of the GP for random samples from the posterior distributions, while grey bands around them reflect the standard deviations of the GP distribution.
-(Recall that while given a particular set of parameters, a parametric model gives just one curve, the GP is always a distribution, whose mean we plot in black.)", fig.width=8, fig.height=6, cache=FALSE}
+Curves show the expectations over the posterior step-ahead predictions based on each of the estimated models.
+Observe that all models are fitting the data reasonably well.)", fig.width=8, fig.height=6, cache=FALSE}
 step_ahead_posteriors <- function(x){
@@
 df_post <- step_ahead_posteriors(x)
-figure1b_posteriors <- ggplot(df_post) + geom_point(aes(time, stock)) + 
-  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=0.2) + 
+library("dplyr")
+df_post_means <- summarize(group_by(df_post, time,variable,stock), value = mean(value))
+ggplot(df_post_means,aes(time, value)) +
+  geom_point(aes(time, stock), size=2) + 
+  geom_line(aes(time, value, col=variable, group=variable),lwd=1) + 
+#  geom_line(aes(time, value, col=variable, group=interaction(L1,variable))) + 
   facet_wrap(~variable) + 
@@
   theme(legend.position="none")
-figure1b_posteriors
-
-write.csv(df_post, "components/data/figure1.csv")
+
+write.csv(df_post_means, "components/data/figure1.csv")
 ```
@@
 
-<!-- 
-### Figure 2 
-- Data comes from limited region of state-space 
-- (Should really show uncertainty of all models here.  Capture the forecast uncertainty several steps down the road?)   
--->
-
-```{r figure_2, dependson=c("assemble-models", "plot-options"), fig.cap="The inferred Gaussian process compared to the true process and maximum-likelihood estimated process.
+```{r figure_2, dependson=c("assemble-models", "plot-options"), fig.cap="The inferred Gaussian process compared to the true process and maximum-likelihood estimated (MLE) process.
 We show the expected value for the function $f$ under each model.
-Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.
+Plots show replicates drawn from the posterior distributions in order to convey uncertainty of the estimates.
+Note the MLE is a point estimate of parameters and so reflects no uncertainty in the distribution.
 The training data are also shown as black points.
-The GP is conditioned on (0,0), shown as a pseudo-data point, as explained in the text", fig.width=8, fig.height=6, cache=FALSE}
-x_grid_short <- x_grid[1:40]
+The GP is conditioned on (0,0), shown as a pseudo-data point.", fig.width=8, fig.height=6, cache=FALSE, dev="CairoJPEG", dev.args=c(quality=100)}
+# subset data first, don't care about seeing dynamics for stock >> K
+x_grid_short <- x_grid[1:max(which(x_grid<12))]
 gp_short <- gp_predict(gp, x_grid_short, burnin=1e4, thin=300)
+
+# sample from subset data
 models_posteriors <- 
@@
 ggplot(models_posteriors) + 
-    geom_ribbon(aes(x=stock, y=y, ymin=ymin, ymax=ymax, group=L1), 
-                  fill = "gray80", 
-                  data=subset(models_posteriors, variable == "GP")) + 
+#    geom_ribbon(aes(x=stock, y=y, ymin=ymin, ymax=ymax, group=L1), 
+#                  fill = "gray80", 
+#                  data=subset(models_posteriors, variable == "GP")) + 
     geom_line(aes(stock, value, col = variable, 
-                  group=interaction(L1,variable)), alpha=0.2) + 
-    geom_point(data = obs, aes(x,y)) + 
+                  group=interaction(L1,variable)), alpha=0.4) + 
+    geom_point(data = obs, aes(x,y), size=2) + 
     xlab(expression(X[t])) + ylab(expression(X[t+1])) +
@@
 on going through the point 0,0 without error.
-This conditioning on (0,0) is equivalent to making the assumption that the population is closed, so that once it hits 0 it stays at 0, despite the lack of any data in the observed sequence to justify this.
+All parametric models also make this assumption.
+Conditioning on (0,0) is equivalent to making the assumption that the population is closed, so that once it hits 0 it stays at 0, despite the lack of any data in the observed sequence to justify this.
 This assumption illustrates how the GP can capture common-sense biology without having to assume more explicit details about the dynamics at low population numbers that have never been observed.
@@
 Unlike parametric models, the GP corresponds to a
-distribution of curves - the gray band with mean shown in black. Uncertainty in the parameters of the GP (not shown) further widens the band of possible population sizes.
+distribution of curves, of which this plot only shows the means. Uncertainty in the parameters of the GP (not shown) further widens the band of possible population sizes.
 In Figure S1 (see supplement), we show the performance of the models outside the observed training data.
@@
 ```{r figure_4, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models.
-We show 100 replicate simulations of the stock dynamics (Eqn 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6, cache=FALSE}
+We show 100 replicate simulations of the stock dynamics (Eqn 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6, cache=FALSE, dev="CairoJPEG", dev.args=c(quality=100)}
 ggplot(sims_data) + 
@@
 more sensitive to this difference.  Large values of $\sigma$ relative to the distance between the stable and unstable
-point increases the chance of a stochastic transition.  More precisely,
-if we let $L$ is the distance between the stable and unstable steady states, then the probability that fluctuations drive the population across the unstable steady state scales as
+point increases the chance of a stochastic transition below the tipping point.  More precisely,
+if we let $L$ be the distance between the stable and unstable steady states, then the probability that fluctuations drive the population across the unstable steady state scales as
 $\exp\left(\frac{L^2}{\sigma^2}\right)$
@@
 plot_sigmas <- ggplot(vary_sigma, aes(noise, value)) + 
-  geom_point() + ylab("Net present value") + 
-  xlab("Level of growth stochasticity") + theme_bw() 
+  geom_point(size=2) + ylab("Net present value") + 
+  xlab("Level of growth stochasticity")
 plot_allees <- ggplot(vary_allee, aes(allee, value)) + 
-  geom_point()+ ylab("Net present value") + 
-  xlab("Allee threshold stock size") + theme_bw()
+  geom_point(size=2)+ ylab("Net present value") + 
+  xlab("Allee threshold stock size")
 multiplot(plot_sigmas, plot_allees, cols=2) 
@@
 In a complex world, the result is suboptimal. But in a world that might
-contain tipping points, the result is disastrous.
+contain tipping points, the result could be disastrous.
 
@@
 this presents a rather risky policy in the face of potential tipping points. 
-The GPDP offers an alternative way to acknowledge this uncertainty.
+The GPDP offers a risk-adverse alternative.
 
@@
 
-<!--
-
-### Risk-prone and risk-adverse value functions
-
-The degree to which the decision-making part of the algorithm (the SDP)
-chooses to explore or avoid the resulting region of uncertainty can also
-be influenced by the curvature of the value (profit) function $R$. Both
-to simplify the intuition and avoid biasing this result, we have chosen
-profits that are linear in the catch and thus neither risk-prone nor
-risk adverse.  Making this function concave, representing the typical
-assumption of diminishing returns, would make the SDP more risk-adverse
-(as larger-than-expected stock sizes offer diminished returns relative
-to the cost of smaller-than-expected stock sizes), and strengthen the
-result shown here in which the BNP solution tends to avoid the region
-of uncertainty.  Sufficiently convex or risk-prone functions could lead
-the SDP to attempt higher exploitation rates despite the uncertainty.
-Understanding the relative roles of such functions would be a promising
-direction for future investigation.
-
--->
-
 ### The role of the prior
@@
 the prior, and consequently the choice of the prior can also play a
-significant role in determining the optimal policy inferred by the
-SDP.  In the examples shown here we have selected a prior that is both
+significant role in determining the optimal policy.
+In the examples shown here we have selected a prior that is both
 relatively uninformative (due to the broad priors placed on its parameters
@@
 mean in the prior without making the much stronger assumption that the
-Ricker is the structurally correct model [@Sugeno2013a]. One fruitful avenue of future research is  identifying criteria to ensure the prior and the value function
+Ricker is the structurally correct model [@Sugeno2013a]. One fruitful avenue of future research is  identifying criteria to ensure the prior and the reward function
 are chosen appropriately for the problem at hand.
-
-<!--
-
-Future directions
------------------
-
-In this simulated example, the underlying dynamics are truly governed
-by a simple parametric model, allowing the parametric approaches to be
-more accurate.  Similarly, because the dynamics are  one-dimensional
-dynamics and lead to  stable nodes (rather than other attractors such
-as limit-cycles resulting in oscillations), the training data provides
-relatively limited information about the dynamics.  For these reasons,
-we anticipate that in higher-dimensional examples characteristic of
-ecosystem management problems that the GP approach will
-prove even more valuable.
-
-In our treatment here we have ignored the possibility of learning during
-the management phase, in which the additional observations of the stock
-size could potentially improve parameter estimates.  Of particular
-interest in the context of the extreme uncertainty considered here
-is the notion of "active adaptive management" or "adaptive probing",
-which may actively seek to reduce uncertainty.  The concept of adaptive
-probing is one area that has explicitly addressed the extrapolation
-uncertainty addressed here, though typically without the additional
-issue of model uncertainty.  As a result, adaptive probing strategies
-suggest rather opposite conclusions than what we observe here. Such
-adaptive probing or Dual Control (e.g. @Ludwig1982) approaches trade
-off short term utility by choosing actions that can reduce uncertainty.
-Adaptive probing strategies arise when it is valuable to intentionally
-force a system far from the observed values even when the expected value
-such actions is low, as it provides much faster learning and consequent
-reduction of model uncertainty that can allow greater value to be derived
-later on.  For instance, @Ludwig1982 show that it may be advantageous
-to fish an unexploited population very heavily at first to obtain a
-better estimate of the recruitment rate.  This intuitive strategy when a
-population is governed by a Ricker or Beverton-Holt-like dynamic would
-clearly be disastrous if instead the dynamics contained an unforeseen
-tipping point.  The best way to learn where the edge lies may be to
-walk up to it, but it is also the most dangerous.  Future work should
-attempt to understand when such active adaptive learning is valuable,
-and when it will increase the risk of an irreversible transition.
-
--->
 

### 2014-11-20 manuscripts/manuscript.tex
as submitted

@@
 much of the founding work on management under uncertainty (Gordon 1954,
-Clark 1976, 2009, May et al. 1979, Reed 1979, Ludwig and Walters 1982)
+Clark 1976, 2009, May et al. 1979, Reed 1979, Ludwig and Walters 1982).
 
@@
 data and, with tipping points,\\even small errors can have very large
-consequences, as we shall illustrate later. Because intervention is
-often (but not always, see Hughes et al. (2013)) too late after a
-tipping point has been crossed, management is often concerned with
-avoiding tipping points before any data about them is available or the
-transition that would more clearly reveal a tipping point occurs (e.g.
-Bestelmeyer et al. 2012).
+consequences, as we shall illustrate later. Because intervention may be
+too late once a tipping point has been crossed (but see Hughes et al.
+(2013)), management is often concerned with avoiding tipping points
+before any data about them are available.
 
@@
 one of the three parametric models , whereas in the natural world we
-never know the ``true'' mode . In addition, by choosing one-dimensional
+never know the ``true'' model. In addition, by choosing one-dimensional
 models with just a few parameters, we limit the chance that poor
@@
 something that becomes a more severe problem for higher-dimensional
-parametric models. Finally, the parametric models we consider are those
-most commonly used in modeling stock-recruitment dynamics or to model
-sudden transitions between alternative stable states.
+parametric models. Finally, the parametric models we consider are
+commonly used in modeling stock-recruitment dynamics or to model sudden
+transitions between alternative stable states.
 
@@
 population at time $t$ and assume that in the absence of take its
-dynamics are. .
-
-\begin{equation}
-X(t+1) = Z(t) f(X(t), p) \label{eq1} 
+dynamics are:
+
+\begin{equation}
+X(t+1) = Z(t) f(X(t), \mathbf{p}) \label{eq1} 
 \end{equation}
@@
 Where Z(t) is log-normally distributed process stochasticity (Reed 1979)
-and $p$ is a vector of parameters to be estimated from the data. We
-describe the three choices for $f(X(t),p)$ in the next section.\\The
-control action is a harvest or take, $h(t)$, measured in the same units
-as $X$, at time $t$. Thus, in the presence of take, the population size
-on the right hand side of Eqn 1 is replaced by $S(t) = X(t) - h(t)$.
+and $\mathbf{p}$ is a vector of parameters to be estimated from the
+data. We describe the three choices for $f(X(t),\mathbf{p})$ in the next
+section.\\The control action is a harvest or take, $h(t)$, measured in
+the same units as $X$, at time $t$. Thus, in the presence of take, the
+population size on the right hand side of Eqn 1 is replaced by
+$S(t) = X(t) - h(t)$.
 
@@
 seek to maximize over choices of harvest
-$E_{X(t+1)} [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$, where the
-state dynamics are given by Eqn 1 and $E$ denotes the expectation over
-the process stochasticity in the future population state.
+$E [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$, where the state
+dynamics are given by Eqn 1 and $E$ denotes the expectation over future
+population states.
 
@@
 translated into $X(t+1)$ by Eqn 1 (that is, we replace $X(t+1)$ by
-$Z(t) f(x(t)-h(t)|p)$).
+$Z(t) f(x(t)-h(t)|\mathbf{p})$).
 
@@
 depend on the particular realization, and longer times are not likely to
-provide substantial improvement.\\We fit each of the models is fit to
-the same data (Figure 1).
+provide substantial improvement. Each of the models is fit to the same
+data (Figure 1).
 
@@
 
+Additionally, we compute the maximum likelihood estimate (MLE, as we
+will refer to this model in the figures) of the parameters for the
+(structurally correct) Allen model. Comparing this to using the
+posterior distribution of parameters inferred from MCMC for the same
+model gives some indication of the importance of this uncertainty in the
+dynamic programming.
+
 \subsubsection{The Gaussian Process
@@
 
-The core difference for our purpose between the GP and the parametric
-models is that the GP model is defined explicitly by the observed data.
-That is, it cannot be specified by the value of parameters alone.
+The core difference for our purpose between the estimated GP and the
+estimated parametric models is that the estimated GP model is defined
+explicitly in reference to the observed data. As a result, uncertainty
+arises in the GP model not only from uncertainty in the parameters, but
+is also increases in regions farther from the observed states, such as
+low population sizes in the example illustrated here. The estimated
+parametric models, by contrast, are completely specified by the
+parameters.
 
@@
 
-A GP is a stochastic process for which any realization of for a finite
-sample of points $n$ is a multivariate normal distribution of dimension
-$n$. Thus, to characterize the GP we need a mean function and a
-covariance function. We proceed as follows.
+A GP is a stochastic process for which any realization consisting of n
+points follows a multivariate normal distribution of dimension $n$. To
+characterize the GP we need a mean function and a covariance function.
+We proceed as follows.
 
 As before, we assume that the data $X(t)$ are observed with process
-stochasticity around a mean function \$ g(X(t))\$
+stochasticity around a mean function $g(X(t))$
 
@@
 where $\varepsilon$ are IID normal random variables with zero-mean and
-variance $\sigma_n^2$. Note that we have chosen to assume additive
+variance $\sigma^2$. Note that we have chosen to assume additive
 stochasticity. While we could consider log-normal stochasticity as in
@@
 
-Now consider the situation where we have a set of previous observations
-of the \emph{timeseries}, $X_{\textrm{obs}}(t)$, and wish to
-\emph{predict} future values $X_{\textrm{p}}(t)$. We consider the
-observations as pairs of points
-$X_{\textrm{obs}}(t), X_{\textrm{obs}}(t+1)$ for $t=1, 2, …$ up through
-final observation time $T$. As described above, in the GP, a finite
-sample of these predictions at $n$ points into the future,
-$X_{\textrm p}(\tau)$ for $\tau \in [T+1, T+2, ..., T+n]$, is a
-multivariate normal ($\mathcal{N}$):
-
-\begin{equation}
-g(X_p|X_{\textrm{obs}}(t), X_{\textrm{obs}}(t+1)) \sim \mathcal{N}(E,C)
-\end{equation}
-
-with mean and covariance matrix given by (see Eqn 2.22 of Rasmussen and
-Williams (2006) for a more detailed derivation)
-
-\begin{equation}
-E = K(X_{\textrm p}(\tau), X_{\textrm{obs}}(t)) \left(K(X_{\textrm{obs}}(t),X_{\textrm{obs}}(t)) - \sigma \mathbb{I} \right)^{-1} X_{\textrm{obs}}(t+1)
-\end{equation}
-
-\begin{equation}
-C = K(X_{\textrm p}(\tau), X_{\textrm p}(\tau)) - K(X_{\textrm p}(\tau), X_{\textrm{obs}}(t)) \left(K(X_{\textrm{obs}}(t),X_{\textrm{obs}}(t)) - \sigma \mathbb{I} \right)^{-1} K(X_{\textrm{obs}}(t), X_{\textrm p}(\tau))
-\end{equation}
-
-Where $K$, called the covariance kernel, measures how much influence one
-observation has on another. In this paper, for $K$ use a matrix whose
-$(i,j)$th element is given by:
-
-\begin{equation}
-K_{i,j}(x_i, y_j) = \exp\left( \frac{ -(x_i - y_j)^2}{2 \ell^2} \right)
+In order to make predictions, we update the GP based on the observed set
+of transitions. To do so, we collect the time series of observed states
+into a vector of ``current'' states,
+$\mathbf{X}_{\textrm{obs}} = \{X(1), \dots, X(T-1)\}$ and a vector of
+``next'' states $\mathbf{Y}_{\textrm{obs}} = \{X(2),\dots,X(T)\}$ where
+$T$ is the time of the final observation. Conditional on these
+observations, the predicted next state, $g(X_p)$, for any given
+``current'' state, $X_p$ follows a normal distribution with mean $E$ and
+variance $C$ determined using the standard rules for conditioning in
+multivariate normals, i.e.
+
+\begin{equation}
+E = K(\mathbf{X}_{\textrm p}, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I}_n \right)^{-1} \mathbf{Y}_{\textrm{obs}}
+\end{equation}
+
+and
+
+\begin{equation}
+C = K(\mathbf{X}_{\textrm p}, \mathbf{X}_{\textrm p}) - K(\mathbf{X}_{\textrm p}, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, \mathbf{X}_{\textrm p})
+\end{equation}
+
+Here $\mathbb{I}_n is the $n\$ by $n$ identity matrix (i.e.~a matrix
+with ones down the diagonal and zeros elsewhere) and $K$ is the
+`covariance kernel.' The covariance kernel controls how much influence
+one observation has on another. In the present application we use the
+squared-exponential kernel which, when evaluated over a pair of vectors,
+say $x$ and $y$, generates a covariance matrix whose $i,j$th element is
+given by
+
+\begin{equation}
+K_{i,j}(\mathbf{x}, \mathbf{y}) = \exp\left( \frac{ -(x_i - y_j)^2}{2 \ell^2} \right)
 \end{equation}
@@
 so that $\ell$ gives the characteristic length-scale over which
-correlation between two observations decays.
-
-We use inverse Gamma priors on both the length-scale $\ell$ and
-$\sigma$, thus for example
+correlation between two observations decays. See Rasmussen and Williams
+(2006) for other choices of covariance kernels and their properties.
+Note that this simple formulation assumes a prior mean of zero. For the
+parameters we use inverse Gamma priors on both the length-scale $\ell$
+and $\sigma$, thus for example
 
@@
 transition matrix, we use the same value iteration algorithm as in the
-parametric case to determine the optimal policy.\\In doing so, the
+parametric case to determine the optimal policy. In doing so, the
 uncertainty in the future state under the GP, $X(t+1)$, includes both
 process uncertainty (based on the estimation of $\sigma$) and structural
-uncertainty of the posterior collection of curves.\\Unlike the case of a
+uncertainty of the posterior collection of curves. Unlike the case of a
 parametric model, which is a single curve given the parameters, the GP
@@
 
-To ensure our results are robust to the choice of parameters, we
-consider 96 different scenarios, described in detail below. To help
-better understand the process, we first describe in detail the results
-of a single scenario.
+To ensure our results are robust to the choice of parameters, we will
+consider 96 different scenarios. To help better understand the process,
+we first describe in detail the results of a single scenario.
 
@@
 \caption{Points show the training data of stock-size over time. Curves
-show the posterior step-ahead predictions based on each of the estimated
-models. Observe that all models are fitting the data reasonably well. In
-the GP panel, black lines are means of the GP for random samples from
-the posterior distributions, while grey bands around them reflect the
-standard deviations of the GP distribution. (Recall that while given a
-particular set of parameters, a parametric model gives just one curve,
-the GP is always a distribution, whose mean we plot in black.)}
+show the expectations over the posterior step-ahead predictions based on
+each of the estimated models. Observe that all models are fitting the
+data reasonably well.)}
 \end{figure}
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_2-1.pdf}
+\includegraphics{components/figure/manuscript-figure_2-1.jpeg}
 \caption{The inferred Gaussian process compared to the true process and
-maximum-likelihood estimated process. We show the expected value for the
-function $f$ under each model. Two standard deviations from the
-estimated Gaussian process covariance with (light grey) and without
-(darker grey) measurement error are also shown. The training data are
-also shown as black points. The GP is conditioned on (0,0), shown as a
-pseudo-data point, as explained in the text}
+maximum-likelihood estimated (MLE) process. We show the expected value
+for the function $f$ under each model. Plots show replicates drawn from
+the posterior distributions in order to convey uncertainty of the
+estimates. Note the MLE is a point estimate of parameters and so
+reflects no uncertainty in the distribution. The training data are also
+shown as black points. The GP is conditioned on (0,0), shown as a
+pseudo-data point.}
 \end{figure}
@@
 size after recruitment the following year.\\In addition to the raw data,
-the GP is conditioned on going through the point 0,0 without error. This
-conditioning on (0,0) is equivalent to making the assumption that the
-population is closed, so that once it hits 0 it stays at 0, despite the
-lack of any data in the observed sequence to justify this. This
-assumption illustrates how the GP can capture common-sense biology
-without having to assume more explicit details about the dynamics at low
-population numbers that have never been observed. If the population were
-not closed, one could repeat the entire analysis without this
-assumption. Unlike parametric models, the GP corresponds to a
-distribution of curves - the gray band with mean shown in black.
-Uncertainty in the parameters of the GP (not shown) further widens the
-band of possible population sizes. In Figure S1 (see supplement), we
-show the performance of the models outside the observed training data.
+the GP is conditioned on going through the point 0,0 without error. All
+parametric models also make this assumption. Conditioning on (0,0) is
+equivalent to making the assumption that the population is closed, so
+that once it hits 0 it stays at 0, despite the lack of any data in the
+observed sequence to justify this. This assumption illustrates how the
+GP can capture common-sense biology without having to assume more
+explicit details about the dynamics at low population numbers that have
+never been observed. If the population were not closed, one could repeat
+the entire analysis without this assumption. Unlike parametric models,
+the GP corresponds to a distribution of curves, of which this plot only
+shows the means. Uncertainty in the parameters of the GP (not shown)
+further widens the band of possible population sizes. In Figure S1 (see
+supplement), we show the performance of the models outside the observed
+training data.
 
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_4-1.pdf}
+\includegraphics{components/figure/manuscript-figure_4-1.jpeg}
 \caption{In the management context, GPDP outperforms approaches based on
@@
 the stable and unstable point increases the chance of a stochastic
-transition. More precisely, if we let $L$ is the distance between the
-stable and unstable steady states, then the probability that
-fluctuations drive the population across the unstable steady state
-scales as $\exp\left(\frac{L^2}{\sigma^2}\right)$
+transition below the tipping point. More precisely, if we let $L$ be the
+distance between the stable and unstable steady states, then the
+probability that fluctuations drive the population across the unstable
+steady state scales as $\exp\left(\frac{L^2}{\sigma^2}\right)$
 
@@
 result is suboptimal. But in a world that might contain tipping points,
-the result is disastrous.
+the result could be disastrous.
 
@@
 While certainly true, this presents a rather risky policy in the face of
-potential tipping points. The GPDP offers an alternative way to
-acknowledge this uncertainty.
+potential tipping points. The GPDP offers a risk-adverse alternative.
 
@@
 and consequently the choice of the prior can also play a significant
-role in determining the optimal policy inferred by the SDP. In the
-examples shown here we have selected a prior that is both relatively
-uninformative (due to the broad priors placed on its parameters $\ell$
-and $\sigma$ and simple (the choice of our covariance function, Eqns 12
-and 13 ). In practice, these should be chosen to confer particular
-biological properties. In principle, this may allow a manager to improve
-the performance of the GPDP by adding detail as is justified. For
-instance, it would be possible to use a linear or a Ricker-shaped mean
-in the prior without making the much stronger assumption that the Ricker
-is the structurally correct model (Sugeno and Munch 2013a). One fruitful
-avenue of future research is identifying criteria to ensure the prior
-and the value function are chosen appropriately for the problem at hand.
+role in determining the optimal policy. In the examples shown here we
+have selected a prior that is both relatively uninformative (due to the
+broad priors placed on its parameters $\ell$ and $\sigma$ and simple
+(the choice of our covariance function, Eqns 12 and 13 ). In practice,
+these should be chosen to confer particular biological properties. In
+principle, this may allow a manager to improve the performance of the
+GPDP by adding detail as is justified. For instance, it would be
+possible to use a linear or a Ricker-shaped mean in the prior without
+making the much stronger assumption that the Ricker is the structurally
+correct model (Sugeno and Munch 2013a). One fruitful avenue of future
+research is identifying criteria to ensure the prior and the reward
+function are chosen appropriately for the problem at hand.
 
@@
 
-Bestelmeyer, B. T., M. C. Duniway, D. K. James, L. M. Burkett, and K. M.
-Havstad. 2012. A test of critical thresholds and their indicators in a
-desertification-prone ecosystem: more resilience than we thought.
-Ecology Letters:n/a--n/a.
-
 Brozović, N., and W. Schlenker. 2011. Optimal management of an ecosystem

### 2014-11-20 manuscripts/manuscript.Rmd
minor typos fixed

@@
 Simultaneously, there has been increasing recognition of the importance of
-multiple steady states or `tipping points' [@Scheffer2001; .  @Scheffer2009;
+multiple steady states or `tipping points' [@Scheffer2001; @Scheffer2009;
 @Polasky2011] in ecological systems.
@@
 marine fisheries the crucible for much of the founding work on management
-under uncertainty [@Gordon1954;
-@Clark1976; -@Clark2009; @Reed1979; @May1979; @Ludwig1982].
+under uncertainty [@Gordon1954; @Clark1976; @Clark2009; @Reed1979; @May1979; @Ludwig1982].
 
@@
 limited such as around  high
-stock levels or an otherwise desirable state. With perfect information
-, one would know just how far a system could
+stock levels or an otherwise desirable state. With perfect information, 
+one would know just how far a system could
 be pushed before crossing the tipping point, and management would be
@@
 Bayesian Non-Parametric (BNP) model of population dynamics [@Munch2005a].
-The BNP method has two distinct advantages 
-.  First, using a BNP model sidesteps the need for an accurate
+The BNP method has two distinct advantages. First, using a BNP model sidesteps the need for an accurate
 model-based description of the system dynamics.  Second, a BNP model
@@
 [^1]: We abbreviate Gaussian Process as GP, which refers to 
-the statistical model we use to approximate the population dynamics
-, and we use the
+the statistical model we use to approximate the population dynamics, and we use the
 term Gaussian Process Dynamic Programming [GPDP,], to refer to the _use_ of a
-GP as the underlying process model when solving a Dynamic Programming equation
-.  Hence we will refer to the models as: GP, Ricker, Allen, etc, and the novel method we put forward here as GPDP.
+GP as the underlying process model when solving a Dynamic Programming equation.
+Hence we will refer to the models as: GP, Ricker, Allen, etc, and the novel method we put forward here as GPDP.
 
@@
 
-Dynamic optimization  requires characterizing the dynamics of a state
+Dynamic optimization  equires characterizing the dynamics of a state
 variable (or variables), a control action, and a value function.
-For simplicity, we consider only a single state variable, which allows us to
-focus on one-dimensional models of population dynamics 
-for the comparison of parametric models and the GP.
+For simplicity, we consider only a single state variable.
 This is a best-case scenario for the parametric models because we
@@
 When the parameters governing the dynamics
-are also uncertain (the case of parametric uncertainty), the expectation in Eqn 2 involves averaging over the
-posterior distribution for the parameters. That is,
+are also uncertain, we take the expectation over the
+posterior distribution for the parameters:
 
@@
 
-When the underlying population dynamics are unknown (the case of structural uncertainty), the function $f$ itself is uncertain and the expectation for the next state includes
-uncertainty in $f$ as well. That is
+When the underlying population dynamics are unknown (the case of structural uncertainty), the function $f$ itself is uncertain and the expectation for the next state includes uncertainty in $f$ as well.
+That is
 
@@
 [@RTeam] using `jags`, [@R2jags]).  We choose uniform priors
-for all parameters (See appendix Tables S1-S3; R code provided). We
+for all parameters of the parametric models (See appendix Tables S1-S3; R code provided). We
 show one-step-ahead predictions of these model fits in Figure 1.
@@
 \begin{equation}
-E = K(\mathbf{X}_{\textrm p}, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I}_n \right)^{-1} \mathbf{Y}_{\textrm{obs}}
+E = K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I}_n \right)^{-1} \mathbf{Y}_{\textrm{obs}}
 \end{equation}
@@
 \begin{equation}
-C = K(\mathbf{X}_{\textrm p}, \mathbf{X}_{\textrm p}) - K(\mathbf{X}_{\textrm p}, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, \mathbf{X}_{\textrm p})
+C = K(X_p, X_p) - K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, X_p)
 \end{equation}
 
-Here $\mathbb{I}_n is the $n$ by $n$ identity matrix (i.e.
+Here $\mathbb{I}_n$ is the $n$ by $n$ identity matrix (i.e.
 a matrix with ones down the diagonal and zeros elsewhere) and $K$ is the ‘covariance kernel.’  The covariance kernel controls how much influence one observation has on another.
-In the present application we use the squared-exponential kernel which, when evaluated over a pair of vectors, say $x$ and $y$,  generates a covariance matrix whose $i,j$th element is given by
+In the present application we use the squared-exponential kernel which, when evaluated over a pair of vectors, say $\mathbf{x}$ and $\mathbf{y}$,  generates a covariance matrix whose $i,j$th element is given by
 
@@
 matrix, we use the same value iteration algorithm as in the parametric
-case to determine the optimal policy.
-In doing so, 
+case to determine the optimal policy. In doing so, 
 the uncertainty in the future state under the GP, $X(t+1)$,
@@
 and structural uncertainty of the posterior collection of curves.
-Unlike the case of a parametric model, which is a single curve given the parameters, 
-the GP remains a distribution of curves.
 
@@
 Model selection criteria (Table 1) penalize more complex models and show a preference for the simpler Ricker model over the  models with alternative stable states (Allen and Myers).
-Details on MCMC estimates for each model, traces, and posterior distributions can be found in the appendix.
+Supplement provides details on the model estimates.
 
@@
 In Figure 4, we show the consequences of managing 100 replicate
-realizations of the simulated fishery under policies derived from each model
-.  The structurally
+realizations of the simulated fishery under policies derived from each model.  The structurally
 correct model under-harvests, leaving the stock to vary around its
@@
 
-Changing the intensity of the stochasticity or the distance between stable and unstable steady states
-does not impact the performance of the GP relative to the optimal solution
-obtained from the true model and true parameters.  The parametric models are
-more sensitive to this difference.  Large values of $\sigma$ relative to the distance between the stable and unstable
-point increases the chance of a stochastic transition below the tipping point.  More precisely,
-if we let $L$ be the distance between the stable and unstable steady states, then the probability that fluctuations drive the population across the unstable steady state scales as
-$\exp\left(\frac{L^2}{\sigma^2}\right)$
-
-(see @Gardiner2009 or @Mangel2006 for the derivation).
-Thus, the impact of using a model that  underestimates the risk of harvesting
-beyond the critical point is considerable, since this such a situation
-occurs more often.  Conversely, with large enough distance between the
-optimal escapement and unstable steady state relative to $\sigma$, the chance
-of a transition becomes vanishingly small and all models can be estimated 
-near-optimally. Models that underestimate the cost incurred by population
-sizes fluctuating significantly below the optimal escapement level will
-not perform poorly as long as those fluctuations are sufficiently small.
-
 The GPDP is only weakly influenced by increasing stochasticity or increasing
-Allee effects over much of the range (Figure 6). Larger $\sigma$ or higher Allee
+Allee effects over much of the range (Figure S2). Larger $\sigma$ or higher Allee
 levels make even the optimal solution without any model or parameter uncertainty
@@
 self-sustaining criterion of @Reed1979).
-
-```{r figure_6, fig.width=6, fig.height=3, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).
-Other than the focal parameter (stochasticity, Allee threshold), other parameters are held fixed as above to illustrate this effect.", cache=FALSE}
-plot_sigmas <- ggplot(vary_sigma, aes(noise, value)) + 
-  geom_point(size=2) + ylab("Net present value") + 
-  xlab("Level of growth stochasticity")
-plot_allees <- ggplot(vary_allee, aes(allee, value)) + 
-  geom_point(size=2)+ ylab("Net present value") + 
-  xlab("Allee threshold stock size")
-multiplot(plot_sigmas, plot_allees, cols=2) 
-write.csv(vary_sigma, "components/data/figure6a.csv")
-write.csv(vary_allee, "components/data/figure6b.csv")
-```
 

### 2014-11-20 manuscripts/manuscript.tex
minor typos fixed

@@
 available and data are limited such as around high stock levels or an
-otherwise desirable state. With perfect information , one would know
-just how far a system could be pushed before crossing the tipping point,
-and management would be simple. But we face imperfect models and limited
+otherwise desirable state. With perfect information, one would know just
+how far a system could be pushed before crossing the tipping point, and
+management would be simple. But we face imperfect models and limited
 data and, with tipping points,\\even small errors can have very large
@@
 dynamics (Munch et al. 2005a). The BNP method has two distinct
-advantages . First, using a BNP model sidesteps the need for an accurate
+advantages. First, using a BNP model sidesteps the need for an accurate
 model-based description of the system dynamics. Second, a BNP model
@@
   which refers to the statistical model we use to approximate the
-  population dynamics , and we use the term Gaussian Process Dynamic
+  population dynamics, and we use the term Gaussian Process Dynamic
   Programming {[}GPDP,{]}, to refer to the \emph{use} of a GP as the
-  underlying process model when solving a Dynamic Programming equation .
+  underlying process model when solving a Dynamic Programming equation.
   Hence we will refer to the models as: GP, Ricker, Allen, etc, and the
@@
 
-Dynamic optimization requires characterizing the dynamics of a state
+Dynamic optimization equires characterizing the dynamics of a state
 variable (or variables), a control action, and a value function. For
-simplicity, we consider only a single state variable, which allows us to
-focus on one-dimensional models of population dynamics for the
-comparison of parametric models and the GP. This is a best-case scenario
-for the parametric models because we simulate underlying dynamics from
-one of the three parametric models , whereas in the natural world we
-never know the ``true'' model. In addition, by choosing one-dimensional
-models with just a few parameters, we limit the chance that poor
-performance will be due to inability to estimate parameters accurately,
-something that becomes a more severe problem for higher-dimensional
-parametric models. Finally, the parametric models we consider are
-commonly used in modeling stock-recruitment dynamics or to model sudden
-transitions between alternative stable states.
+simplicity, we consider only a single state variable. This is a
+best-case scenario for the parametric models because we simulate
+underlying dynamics from one of the three parametric models , whereas in
+the natural world we never know the ``true'' model. In addition, by
+choosing one-dimensional models with just a few parameters, we limit the
+chance that poor performance will be due to inability to estimate
+parameters accurately, something that becomes a more severe problem for
+higher-dimensional parametric models. Finally, the parametric models we
+consider are commonly used in modeling stock-recruitment dynamics or to
+model sudden transitions between alternative stable states.
 
@@
 
-When the parameters governing the dynamics are also uncertain (the case
-of parametric uncertainty), the expectation in Eqn 2 involves averaging
-over the posterior distribution for the parameters. That is,
+When the parameters governing the dynamics are also uncertain, we take
+the expectation over the posterior distribution for the parameters:
 
@@
 using \texttt{jags}, (Su and Masanao Yajima 2013)). We choose uniform
-priors for all parameters (See appendix Tables S1-S3; R code provided).
-We show one-step-ahead predictions of these model fits in Figure 1. We
-tested each chain for Gelman-Rubin convergence and results were robust
-to longer runs. For each simulation we also applied several commonly
-used model selection criteria (AIC, BIC, DIC, see Burnham and Anderson
-(2002)) to identify the best fitting model.
+priors for all parameters of the parametric models (See appendix Tables
+S1-S3; R code provided). We show one-step-ahead predictions of these
+model fits in Figure 1. We tested each chain for Gelman-Rubin
+convergence and results were robust to longer runs. For each simulation
+we also applied several commonly used model selection criteria (AIC,
+BIC, DIC, see Burnham and Anderson (2002)) to identify the best fitting
+model.
 
@@
 \begin{equation}
-C = K(\mathbf{X}_{\textrm p}, \mathbf{X}_{\textrm p}) - K(\mathbf{X}_{\textrm p}, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, \mathbf{X}_{\textrm p})
+C = K(X_p, X_p) - K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, X_p)
 \end{equation}
 
-Here $\mathbb{I}_n is the $n\$ by $n$ identity matrix (i.e.~a matrix
+Here $\mathbb{I}_n$ is the $n$ by $n$ identity matrix (i.e.~a matrix
 with ones down the diagonal and zeros elsewhere) and $K$ is the
@@
 squared-exponential kernel which, when evaluated over a pair of vectors,
-say $x$ and $y$, generates a covariance matrix whose $i,j$th element is
-given by
+say $\mathbf{x}$ and $\mathbf{y}$, generates a covariance matrix whose
+$i,j$th element is given by
 
@@
 process uncertainty (based on the estimation of $\sigma$) and structural
-uncertainty of the posterior collection of curves. Unlike the case of a
-parametric model, which is a single curve given the parameters, the GP
-remains a distribution of curves.
+uncertainty of the posterior collection of curves.
 
@@
 for the simpler Ricker model over the models with alternative stable
-states (Allen and Myers). Details on MCMC estimates for each model,
-traces, and posterior distributions can be found in the appendix.
+states (Allen and Myers). Supplement provides details on the model
+estimates.
 
@@
 $S=0$, as seen in the true (correct form, exact parameters) model, the
-Allen model (correct form, estimated parameters) and the GP.\\Only the
+Allen model (correct form, estimated parameters) and the GP. Only the
 structurally correct model (Allen model) and the GP produce policies
@@
 realizations of the simulated fishery under policies derived from each
-model . The structurally correct model under-harvests, leaving the stock
+model. The structurally correct model under-harvests, leaving the stock
 to vary around its unfished optimum. The structurally incorrect Ricker
@@
 
-Changing the intensity of the stochasticity or the distance between
-stable and unstable steady states does not impact the performance of the
-GP relative to the optimal solution obtained from the true model and
-true parameters. The parametric models are more sensitive to this
-difference. Large values of $\sigma$ relative to the distance between
-the stable and unstable point increases the chance of a stochastic
-transition below the tipping point. More precisely, if we let $L$ be the
-distance between the stable and unstable steady states, then the
-probability that fluctuations drive the population across the unstable
-steady state scales as $\exp\left(\frac{L^2}{\sigma^2}\right)$
-
-(see Gardiner (2009) or Mangel (2006) for the derivation). Thus, the
-impact of using a model that underestimates the risk of harvesting
-beyond the critical point is considerable, since this such a situation
-occurs more often. Conversely, with large enough distance between the
-optimal escapement and unstable steady state relative to $\sigma$, the
-chance of a transition becomes vanishingly small and all models can be
-estimated near-optimally. Models that underestimate the cost incurred by
-population sizes fluctuating significantly below the optimal escapement
-level will not perform poorly as long as those fluctuations are
-sufficiently small.
-
 The GPDP is only weakly influenced by increasing stochasticity or
-increasing Allee effects over much of the range (Figure 6). Larger
+increasing Allee effects over much of the range (Figure S2). Larger
 $\sigma$ or higher Allee levels make even the optimal solution without
@@
 self-sustaining criterion of Reed (1979)).
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{components/figure/manuscript-figure_6-1.pdf}
-\caption{The effect of increasing noise or decreasing Allee threshold
-levels on the net present value of the fishery when managed under the
-GPDP, relative to managing under the true model (with known parameters).
-Other than the focal parameter (stochasticity, Allee threshold), other
-parameters are held fixed as above to illustrate this effect.}
-\end{figure}
 
@@
 
-Gardiner, C. 2009. Stochastic Methods: A Handbook for the Natural and
-Social Sciences (Springer Series in Synergetics). Page 447. Springer.
-
 Gelman, A., J. B. Carlin, H. S. Stern, and D. B. Rubin. 2003. Bayesian
@@
 parameter estimates. Ecological Modelling 14:273--292.
-
-Mangel, M. 2006. The Theoretical Biologist's Toolbox: Quantitative
-Methods for Ecology and Evolutionary Biology. Cambridge University
-Press.
 

### 2014-11-20 manuscripts/supplement.Rmd
minor typos fixed

@@
 
-# Further sensitivity analysis (Fig S2) #
+# Further sensitivity analysis (Fig S2 - 3) #
+
+We perform 2 sensitivity analyses. The first focuses on illustrating the robustness of the
+approach to the two parameters that most influence stochastic transitions across the tipping 
+point: the position of the Allee threshold and the scale of the noise (Fig S2).
+
+Changing the intensity of the stochasticity or the distance between stable and unstable steady states
+does not impact the performance of the GP relative to the optimal solution
+obtained from the true model and true parameters.  The parametric models are
+more sensitive to this difference.  Large values of $\sigma$ relative to the distance between the stable and unstable
+point increases the chance of a stochastic transition below the tipping point.  More precisely,
+if we let $L$ be the distance between the stable and unstable steady states, then the probability that fluctuations drive the population across the unstable steady state scales as
+
+$\exp\left(\frac{L^2}{\sigma^2}\right)$
+
+(see @Gardiner2009 or @Mangel2006 for the derivation).
+
+Thus, the impact of using a model that  underestimates the risk of harvesting
+beyond the critical point is considerable, since this such a situation
+occurs more often.  Conversely, with large enough distance between the
+optimal escapement and unstable steady state relative to $\sigma$, the chance
+of a transition becomes vanishingly small and all models can be estimated 
+near-optimally. Models that underestimate the cost incurred by population
+sizes fluctuating significantly below the optimal escapement level will
+not perform poorly as long as those fluctuations are sufficiently small.
+Fig S2 shows the net present value of managing under teh GPDP remains
+close to the optimal value (ratio of 1), despite varying across either
+noise level or the the position of the allee threshold.
+
+```{r figure_S2, fig.width=6, fig.height=3, dependson=c("sensitivity-trends"), fig.cap="The effect of increasing noise or decreasing Allee threshold levels on the net present value of the fishery when managed under the GPDP, relative to managing under the true model (with known parameters).
+Other than the focal parameter (stochasticity, Allee threshold), other parameters are held fixed as above to illustrate this effect.", cache=FALSE}
+plot_sigmas <- ggplot(vary_sigma, aes(noise, value)) + 
+  geom_point(size=2) + ylab("Net present value") + 
+  xlab("Level of growth stochasticity")
+plot_allees <- ggplot(vary_allee, aes(allee, value)) + 
+  geom_point(size=2)+ ylab("Net present value") + 
+  xlab("Allee threshold stock size")
+multiplot(plot_sigmas, plot_allees, cols=2) 
+write.csv(vary_sigma, "components/data/figureS2a.csv")
+write.csv(vary_allee, "components/data/figureS2b.csv")
+```
 
@@
 
-```{r figure_S2, fig.height=6, fig.width=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
+```{r figure_S3, fig.height=6, fig.width=6, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.
 Histograms shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters.
@@
   facet_grid(noise~model)
-write.csv(sensitivity_dat, "components/data/figureS2.csv") 
-```
-```{r S2_eml, eval=FALSE}
-# not descriptive enough, supplement text documents this better
-par <- levels(sensitivity_dat$parameters)
-names(par) <- par
-col.defs <- names(sensitivity_dat)
-unit.defs <- list(models, par, "number", "number",  "number", "number")
-col.classes <- sapply(sensitivity_dat, class)
-eml_write("components/data/figureS2.csv", col.classes = col.classes, col.defs=col.defs,unit.defs=unit.defs,creator=me, title="Figure S2", file="components/data/README_S2.xml")
+write.csv(sensitivity_dat, "components/data/figureS3.csv") 
 ```
@@
 ```{r AllenSetsTable, results="asis"}
-pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Allen models in Figure S2." )
+pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Allen models in Figure S3." )
 ```
 ```{r MyersSetsTable, results="asis"}
-pandoc.table(MyersParameterSets, caption="Randomly chosen parameter sets for the Myers models in Figure S2." )
+pandoc.table(MyersParameterSets, caption="Randomly chosen parameter sets for the Myers models in Figure S3." )
 ```
@@
 
-## GP MCMC (Fig S3-4) ##
+## GP MCMC (Fig S4-5) ##
 
-```{r figure_S3, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options", fig.width=4, fig.height=4}
+```{r figure_S4, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options", fig.width=4, fig.height=4}
 gp_assessment_plots$traces_plot
@@
 
-```{r figure_S4, fig.cap="Posterior distributions from the MCMC estimate of the GP model.
+```{r figure_S5, fig.cap="Posterior distributions from the MCMC estimate of the GP model.
 Prior curves shown in red; note the posterior distributions are significantly more peaked than the priors, showing that the data has been informative and is not driven by the priors.", fig.width=4, fig.height=4}
 gp_assessment_plots$posteriors_plot
-write.csv(tgp_dat,"components/data/figureS3-4.csv")
+write.csv(tgp_dat,"components/data/figureS4-5.csv")
 ```
@@
 
-## Ricker Model MCMC (Fig S5-6) ##
+## Ricker Model MCMC (Fig S6-7) ##
 
-```{r figure_S5, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps).
+```{r figure_S6, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps).
 stdQ refers to the estimate of $\\sigma$; deviance is -2 times the log likelihood.
@@
 
-```{r figure_S6, fig.cap="Posteriors from the MCMC estimate of the Ricker model.
+```{r figure_S7, fig.cap="Posteriors from the MCMC estimate of the Ricker model.
 Note that the model estimates a carrying capacity $K$ very close to the true equilibrium where most of the observations were made, but is less certain about the growth rate.", fig.width=4, fig.height=4}
@@
   facet_wrap(~ variable, scale="free", ncol=2)
-write.csv(ricker_posteriors, "components/data/figureS5-6.csv")
+write.csv(ricker_posteriors, "components/data/figureS6-7.csv")
 ```
@@
 
-## Myers Model MCMC (Fig S7-8) ##
+## Myers Model MCMC (Fig S8-9) ##
 
-```{r figure_S7, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps).
+```{r figure_S8, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps).
 stdQ refers to the estimate of $\\sigma$; deviance is -2 times the log likelihood.", fig.width=4}
@@
 
-```{r figure_S8, fig.cap="Posterior distributions from the MCMC estimates of the Myers model.
+```{r figure_S9, fig.cap="Posterior distributions from the MCMC estimates of the Myers model.
 Note that with more free parameters, the posteriors reflect greater uncertainty.
@@
   facet_wrap(~ variable, scale="free", ncol=2)
-write.csv(myers_posteriors, "components/data/figureS7-8.csv")
+write.csv(myers_posteriors, "components/data/figureS8-9.csv")
 ```
@@
 
-## Allen Model MCMC (Fig S9-10) ##
+## Allen Model MCMC (Fig S10-11) ##
 
-```{r figure_S9, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps).
+```{r figure_S10, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps).
 stdQ refers to the estimate of $\\sigma$; deviance is -2 times the log likelihood.", fig.width=4}
@@
 
-```{r figure_S10, fig.cap="Posteriors from the MCMC estimate of the Allen model.
+```{r figure_S11, fig.cap="Posteriors from the MCMC estimate of the Allen model.
 The Allen model is the structurally correct model.
@@
   facet_wrap(~ variable, scale="free", ncol=2)
-write.csv(allen_posteriors, "components/data/figureS9-10.csv")
+write.csv(allen_posteriors, "components/data/figureS10-11.csv")
 ```

### 2014-11-20 manuscripts/supplement.tex
minor typos fixed

@@
 convenience CSV files of the data shown in each graph are made available
-on Dryad.
+on Dryad, along with the source \texttt{.Rmd} files for the manuscript
+and supplement that document them.
 
@@
 
-\section{Further sensitivity analysis (Fig
-S2)}\label{further-sensitivity-analysis-fig-s2}
+\section{Further sensitivity analysis (Fig S2 -
+3)}\label{further-sensitivity-analysis-fig-s2---3}
+
+We perform 2 sensitivity analyses. The first focuses on illustrating the
+robustness of the approach to the two parameters that most influence
+stochastic transitions across the tipping point: the position of the
+Allee threshold and the scale of the noise (Fig S2).
+
+Changing the intensity of the stochasticity or the distance between
+stable and unstable steady states does not impact the performance of the
+GP relative to the optimal solution obtained from the true model and
+true parameters. The parametric models are more sensitive to this
+difference. Large values of $\sigma$ relative to the distance between
+the stable and unstable point increases the chance of a stochastic
+transition below the tipping point. More precisely, if we let $L$ be the
+distance between the stable and unstable steady states, then the
+probability that fluctuations drive the population across the unstable
+steady state scales as
+
+$\exp\left(\frac{L^2}{\sigma^2}\right)$
+
+(see Gardiner (2009) or Mangel (2006) for the derivation).
+
+Thus, the impact of using a model that underestimates the risk of
+harvesting beyond the critical point is considerable, since this such a
+situation occurs more often. Conversely, with large enough distance
+between the optimal escapement and unstable steady state relative to
+$\sigma$, the chance of a transition becomes vanishingly small and all
+models can be estimated near-optimally. Models that underestimate the
+cost incurred by population sizes fluctuating significantly below the
+optimal escapement level will not perform poorly as long as those
+fluctuations are sufficiently small. Fig S2 shows the net present value
+of managing under teh GPDP remains close to the optimal value (ratio of
+1), despite varying across either noise level or the the position of the
+allee threshold.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{components/figure/supplement-figure_S2-1.pdf}
+\caption{The effect of increasing noise or decreasing Allee threshold
+levels on the net present value of the fishery when managed under the
+GPDP, relative to managing under the true model (with known parameters).
+Other than the focal parameter (stochasticity, Allee threshold), other
+parameters are held fixed as above to illustrate this effect.}
+\end{figure}
 
@@
 \centering
-\includegraphics{components/figure/supplement-figure_S2-1.pdf}
+\includegraphics{components/figure/supplement-figure_S3-1.pdf}
 \caption{Sensitivity Analysis. Histograms shows the ratio of the
@@
 \caption{Randomly chosen parameter sets for the Allen models in Figure
-S2.}
+S3.}
 \end{longtable}
@@
 \caption{Randomly chosen parameter sets for the Myers models in Figure
-S2.}
+S3.}
 \end{longtable}
@@
 
-\subsection{GP MCMC (Fig S3-4)}\label{gp-mcmc-fig-s3-4}
+\subsection{GP MCMC (Fig S4-5)}\label{gp-mcmc-fig-s4-5}
 
@@
 \centering
-\includegraphics{components/figure/supplement-figure_S3-1.pdf}
+\includegraphics{components/figure/supplement-figure_S4-1.pdf}
 \caption{Traces from the MCMC estimates of the GP model show reasonable
@@
 \centering
-\includegraphics{components/figure/supplement-figure_S4-1.pdf}
+\includegraphics{components/figure/supplement-figure_S5-1.pdf}
 \caption{Posterior distributions from the MCMC estimate of the GP model.
@@
 \subsection{Ricker Model MCMC (Fig
-S5-6)}\label{ricker-model-mcmc-fig-s5-6}
+S6-7)}\label{ricker-model-mcmc-fig-s6-7}
 
@@
 \centering
-\includegraphics{components/figure/supplement-figure_S5-1.pdf}
+\includegraphics{components/figure/supplement-figure_S6-1.pdf}
 \caption{Traces from the MCMC estimates of the Ricker model show
@@
 \centering
-\includegraphics{components/figure/supplement-figure_S6-1.pdf}
+\includegraphics{components/figure/supplement-figure_S7-1.pdf}
 \caption{Posteriors from the MCMC estimate of the Ricker model. Note
@@
 \subsection{Myers Model MCMC (Fig
-S7-8)}\label{myers-model-mcmc-fig-s7-8}
+S8-9)}\label{myers-model-mcmc-fig-s8-9}
 
@@
 \centering
-\includegraphics{components/figure/supplement-figure_S7-1.pdf}
+\includegraphics{components/figure/supplement-figure_S8-1.pdf}
 \caption{Traces from the MCMC estimates of the Myers model show
@@
 \centering
-\includegraphics{components/figure/supplement-figure_S8-1.pdf}
+\includegraphics{components/figure/supplement-figure_S9-1.pdf}
 \caption{Posterior distributions from the MCMC estimates of the Myers
@@
 \subsection{Allen Model MCMC (Fig
-S9-10)}\label{allen-model-mcmc-fig-s9-10}
+S10-11)}\label{allen-model-mcmc-fig-s10-11}
 
@@
 \centering
-\includegraphics{components/figure/supplement-figure_S9-1.pdf}
+\includegraphics{components/figure/supplement-figure_S10-1.pdf}
 \caption{Traces from the MCMC estimates of the Allen model show
@@
 \centering
-\includegraphics{components/figure/supplement-figure_S10-1.pdf}
+\includegraphics{components/figure/supplement-figure_S11-1.pdf}
 \caption{Posteriors from the MCMC estimate of the Allen model. The Allen
@@
 
+Gardiner, C. 2009. Stochastic Methods: A Handbook for the Natural and
+Social Sciences (Springer Series in Synergetics). Page 447. Springer.
+
+Mangel, M. 2006. The Theoretical Biologist's Toolbox: Quantitative
+Methods for Ecology and Evolutionary Biology. Cambridge University
+Press.
+
 Xie, Y. 2013. Dynamic Documents with R and knitr. Chapman; Hall/CRCBoca

### 2014-12-05 manuscripts/manuscript.Rmd
edit readmes and tidy up

@@
 Simultaneously, there has been increasing recognition of the importance of
-multiple steady states or `tipping points' [@Scheffer2001; @Scheffer2009;
+multiple steady states or 'tipping points' [@Scheffer2001; @Scheffer2009;
 @Polasky2011] in ecological systems.
@@
 
-### Parametric models
-
 We let $X(t)$ denote the size (numbers or biomass) of the focal
@@
 
-\subsubsection{Parametric models}\label{parametric-models}
+### Parametric Models ###
 
@@
 
-For the prior on  $\ell$,  $\alpha = $ `r d.p[1]` and $\beta = $ `r d.p[2]`.  The prior  on $\sigma$,  $\alpha = $ `r s2.p[1]` and $\beta = $ `r s2.p[2]`.
+For the prior on  $\ell$,
+$\alpha =$ `r d.p[1]` and 
+$\beta =$ `r d.p[2]`. 
+The prior  on $\sigma$,
+$\alpha =$ `r s2.p[1]`
+and $\beta =$ `r s2.p[2]`.
 

### 2014-12-05 manuscripts/manuscript.tex
edit readmes and tidy up

@@
 
-\subsubsection{Parametric models}\label{parametric-models}
-
-We let $X(t)$ denote the size (numbers or biomass) of the focal
-population at time $t$ and assume that in the absence of take its
+We let \(X(t)\) denote the size (numbers or biomass) of the focal
+population at time \(t\) and assume that in the absence of take its
 dynamics are:
@@
 Where Z(t) is log-normally distributed process stochasticity (Reed 1979)
-and $\mathbf{p}$ is a vector of parameters to be estimated from the
-data. We describe the three choices for $f(X(t),\mathbf{p})$ in the next
-section.\\The control action is a harvest or take, $h(t)$, measured in
-the same units as $X$, at time $t$. Thus, in the presence of take, the
-population size on the right hand side of Eqn 1 is replaced by
-$S(t) = X(t) - h(t)$.
+and \(\mathbf{p}\) is a vector of parameters to be estimated from the
+data. We describe the three choices for \(f(X(t),\mathbf{p})\) in the
+next section.\\The control action is a harvest or take, \(h(t)\),
+measured in the same units as \(X\), at time \(t\). Thus, in the
+presence of take, the population size on the right hand side of Eqn 1 is
+replaced by \(S(t) = X(t) - h(t)\).
 
-To construct the value function, we consider a return when $X(t) = x(t)$
-and harvest $h(t) = h$ denoted as the reward, $R(x(t), h)$. For example,
-if the return is the harvest at time $t$, then
-$R(x(t), h(t)) = \min(x(t), h(t))$. We assume that future harvests are
-discounted relative to current ones at a constant rate of discount
-$\delta$ and ask for the harvest policy that maximizes total discounted
-harvest between the current time $t$ and a final time $T$. That is, we
-seek to maximize over choices of harvest
-$E [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$, where the state
-dynamics are given by Eqn 1 and $E$ denotes the expectation over future
-population states.
+To construct the value function, we consider a return when
+\(X(t) = x(t)\) and harvest \(h(t) = h\) denoted as the reward,
+\(R(x(t), h)\). For example, if the return is the harvest at time \(t\),
+then \(R(x(t), h(t)) = \min(x(t), h(t))\). We assume that future
+harvests are discounted relative to current ones at a constant rate of
+discount \(\delta\) and ask for the harvest policy that maximizes total
+discounted harvest between the current time \(t\) and a final time
+\(T\). That is, we seek to maximize over choices of harvest
+\(E [ \sum_{t = 0}^{T} R(X(t), h(t), t) \delta^t]\), where the state
+dynamics are given by Eqn 1 and \(E\) denotes the expectation over
+future population states.
 
 In order to find that policy, we introduce the value function
-$V(x(t), t)$ representing the total discounted catch from time $t$
-onwards given that $X(t) = x(t)$. This value function satisfies an
+\(V(x(t), t)\) representing the total discounted catch from time \(t\)
+onwards given that \(X(t) = x(t)\). This value function satisfies an
 equation of SDP (Mangel and Clark 1988, Clark and Mangel 2000, Clark
-2009 Mangel 2014),
+2009 , Mangel 2014),
 
@@
 where expectation is taken over all possible values of the next state,
-$X(t+1)$, and maximized over all possible choices of harvest, $h(t)$.
-That is, at time $t$, when population size is $x(t)$ and harvest $h(t)$
-is applied, the immediate return is $R(h(t), x(t))$. When the sole
-source of uncertainty is the process stochasticity term, $Z$, and thus
-the expectation in Eqn 2 is equivalent to taking expectations over
-$Z(t)$. That is
+\(X(t+1)\), and maximized over all possible choices of harvest,
+\(h(t)\). That is, at time \(t\), when population size is \(x(t)\) and
+harvest \(h(t)\) is applied, the immediate return is \(R(h(t), x(t))\).
+When the sole source of uncertainty is the process stochasticity term,
+\(Z\), and thus the expectation in Eqn 2 is equivalent to taking
+expectations over \(Z(t)\). That is
 
@@
 
-where the population size after the take is $x(t) - h(t)$, which is then
-translated into $X(t+1)$ by Eqn 1 (that is, we replace $X(t+1)$ by
-$Z(t) f(x(t)-h(t)|\mathbf{p})$).
+where the population size after the take is \(x(t) - h(t)\), which is
+then translated into \(X(t+1)\) by Eqn 1 (that is, we replace \(X(t+1)\)
+by \(Z(t) f(x(t)-h(t)|\mathbf{p})\)).
 
@@
 When the underlying population dynamics are unknown (the case of
-structural uncertainty), the function $f$ itself is uncertain and the
-expectation for the next state includes uncertainty in $f$ as well. That
-is
+structural uncertainty), the function \(f\) itself is uncertain and the
+expectation for the next state includes uncertainty in \(f\) as well.
+That is
 
@@
 
-We consider the finite time problem with $T=$ 1000, which we solve using
-the standard value iteration algorithm (see Mangel and Clark 1988, Clark
-and Mangel 2000).
+We consider the finite time problem with \(T=\) 1000, which we solve
+using the standard value iteration algorithm (see Mangel and Clark 1988,
+Clark and Mangel 2000).
 
-\subsubsection{Parametric models}\label{parametric-models}
+\subsubsection{Parametric Models}\label{parametric-models}
 
@@
 the Myers model (Myers et al. 1995), Eqns \eqref{ricker}-\eqref{myers}.
-In all three, we let $K$ denote the carrying capacity and $r$ the
+In all three, we let \(K\) denote the carrying capacity and \(r\) the
 maximum per capita growth rate. The Ricker model has two parameters and
@@
 
-where $X_c$ denotes the location of the unstable steady state (i.e., the
-tipping point).
+where \(X_c\) denotes the location of the unstable steady state (i.e.,
+the tipping point).
 
@@
 
-where $\theta = 1$ corresponds to Beverton-Holt dynamics and $\theta$
-\textgreater{} 2 leads to Allee effects and multiple stable states.
+where \(\theta = 1\) corresponds to Beverton-Holt dynamics and
+\(\theta\) \textgreater{} 2 leads to Allee effects and multiple stable
+states.
 
@@
 multiplicative log-normal stochasticity in Eqn 1 introduces one
-additional parameter $\sigma$ that must be estimated.
+additional parameter \(\sigma\) that must be estimated.
 
@@
 A GP is a stochastic process for which any realization consisting of n
-points follows a multivariate normal distribution of dimension $n$. To
+points follows a multivariate normal distribution of dimension \(n\). To
 characterize the GP we need a mean function and a covariance function.
@@
 
-As before, we assume that the data $X(t)$ are observed with process
-stochasticity around a mean function $g(X(t))$
+As before, we assume that the data \(X(t)\) are observed with process
+stochasticity around a mean function \(g(X(t))\)
 
@@
 
-where $\varepsilon$ are IID normal random variables with zero-mean and
-variance $\sigma^2$. Note that we have chosen to assume additive
+where \(\varepsilon\) are IID normal random variables with zero-mean and
+variance \(\sigma^2\). Note that we have chosen to assume additive
 stochasticity. While we could consider log-normal stochasticity as in
@@
 into a vector of ``current'' states,
-$\mathbf{X}_{\textrm{obs}} = \{X(1), \dots, X(T-1)\}$ and a vector of
-``next'' states $\mathbf{Y}_{\textrm{obs}} = \{X(2),\dots,X(T)\}$ where
-$T$ is the time of the final observation. Conditional on these
-observations, the predicted next state, $g(X_p)$, for any given
-``current'' state, $X_p$ follows a normal distribution with mean $E$ and
-variance $C$ determined using the standard rules for conditioning in
-multivariate normals, i.e.
+\(\mathbf{X}_{\textrm{obs}} = \{X(1), \dots, X(T-1)\}\) and a vector of
+``next'' states \(\mathbf{Y}_{\textrm{obs}} = \{X(2),\dots,X(T)\}\)
+where \(T\) is the time of the final observation. Conditional on these
+observations, the predicted next state, \(g(X_p)\), for any given
+``current'' state, \(X_p\) follows a normal distribution with mean \(E\)
+and variance \(C\) determined using the standard rules for conditioning
+in multivariate normals, i.e.
 
@@
 
-Here $\mathbb{I}_n$ is the $n$ by $n$ identity matrix (i.e.~a matrix
-with ones down the diagonal and zeros elsewhere) and $K$ is the
-`covariance kernel.' The covariance kernel controls how much influence
-one observation has on another. In the present application we use the
-squared-exponential kernel which, when evaluated over a pair of vectors,
-say $\mathbf{x}$ and $\mathbf{y}$, generates a covariance matrix whose
-$i,j$th element is given by
+NA
 
@@
 
-so that $\ell$ gives the characteristic length-scale over which
+so that \(\ell\) gives the characteristic length-scale over which
 correlation between two observations decays. See Rasmussen and Williams
@@
 Note that this simple formulation assumes a prior mean of zero. For the
-parameters we use inverse Gamma priors on both the length-scale $\ell$
-and $\sigma$, thus for example
+parameters we use inverse Gamma priors on both the length-scale \(\ell\)
+and \(\sigma\), thus for example
 
@@
 
-For the prior on $\ell$, $\alpha = $ 10 and $\beta = $ 10. The prior on
-$\sigma$, $\alpha = $ 5 and $\beta = $ 5.
+For the prior on \(\ell\), \(\alpha =\) 10 and \(\beta =\) 10. The prior
+on \(\sigma\), \(\alpha =\) 5 and \(\beta =\) 5.
 
@@
 a SDP algorithm. Given the GP posteriors, we construct the transition
-matrix representing the probability of going to each state $X(t+1)$
-given any current state $x(t)$ and any harvest $h(t)$ (See the function
-\texttt{gp\_transition\_matrix()} in the provided R package). Given this
-transition matrix, we use the same value iteration algorithm as in the
-parametric case to determine the optimal policy. In doing so, the
-uncertainty in the future state under the GP, $X(t+1)$, includes both
-process uncertainty (based on the estimation of $\sigma$) and structural
-uncertainty of the posterior collection of curves.
+matrix representing the probability of going to each state \(X(t+1)\)
+given any current state \(x(t)\) and any harvest \(h(t)\) (See the
+function \texttt{gp\_transition\_matrix()} in the provided R package).
+Given this transition matrix, we use the same value iteration algorithm
+as in the parametric case to determine the optimal policy. In doing so,
+the uncertainty in the future state under the GP, \(X(t+1)\), includes
+both process uncertainty (based on the estimation of \(\sigma\)) and
+structural uncertainty of the posterior collection of curves.
 
@@
 \begin{longtable}[c]{@{}cccc@{}}
+\caption{Model selection scores for several common criteria (DIC:
+Deviance Information Criterion, AIC: Akaike Information Criterion, BIC:
+Bayesian Information Criterion) all select the wrong model. As the true
+(Allen) model is not distinguishable from the simpler (Ricker) model in
+the region of the observed data, this error cannot be avoided regardless
+of the model choice criterion. This highlights the danger of model
+choice when the selected model will be used outside of the observed
+range of the data.}\\
 \toprule\addlinespace
@@
 \bottomrule
-\addlinespace
-\caption{Model selection scores for several common criteria (DIC:
-Deviance Information Criterion, AIC: Akaike Information Criterion, BIC:
-Bayesian Information Criterion) all select the wrong model. As the true
-(Allen) model is not distinguishable from the simpler (Ricker) model in
-the region of the observed data, this error cannot be avoided regardless
-of the model choice criterion. This highlights the danger of model
-choice when the selected model will be used outside of the observed
-range of the data.}
 \end{longtable}
@@
 maximum-likelihood estimated (MLE) process. We show the expected value
-for the function $f$ under each model. Plots show replicates drawn from
-the posterior distributions in order to convey uncertainty of the
+for the function \(f\) under each model. Plots show replicates drawn
+from the posterior distributions in order to convey uncertainty of the
 estimates. Note the MLE is a point estimate of parameters and so
@@
 under each model. Policies are shown in terms of target escapement,
-$S(t)$, as under models such as this a constant escapement policy is
+\(S(t)\), as under models such as this a constant escapement policy is
 expected to be optimal (Reed 1979) Several policies show a numerical
@@
 policies inferred under each model differ widely (Figure 3). Policies
-are shown in terms of target escapement, $S(t) = x_t - h$. Under models
-such as this a constant escapement policy is expected to be optimal
-(Reed 1979), whereby population levels below a certain size $S$ are
-unharvested, while above that size the harvest strategy aims to return
-the population to $S$. Whenever a model predicts that the population
-will not persist below a certain threshold, the optimal solution is to
-harvest the entire population immediately, resulting in an escapement
-$S=0$, as seen in the true (correct form, exact parameters) model, the
-Allen model (correct form, estimated parameters) and the GP. Only the
-structurally correct model (Allen model) and the GP produce policies
-close to the true optimum policy.
+are shown in terms of target escapement, \(S(t) = x_t - h\). Under
+models such as this a constant escapement policy is expected to be
+optimal (Reed 1979), whereby population levels below a certain size
+\(S\) are unharvested, while above that size the harvest strategy aims
+to return the population to \(S\). Whenever a model predicts that the
+population will not persist below a certain threshold, the optimal
+solution is to harvest the entire population immediately, resulting in
+an escapement \(S=0\), as seen in the true (correct form, exact
+parameters) model, the Allen model (correct form, estimated parameters)
+and the GP. Only the structurally correct model (Allen model) and the GP
+produce policies close to the true optimum policy.
 
@@
 increasing Allee effects over much of the range (Figure S2). Larger
-$\sigma$ or higher Allee levels make even the optimal solution without
+\(\sigma\) or higher Allee levels make even the optimal solution without
 any model or parameter uncertainty unable to harvest the population
@@
 
-Our results illustrate that model-choice approaches can be absolutely
-misleading -- by providing support to models that cannot capture tipping
-point dynamics because they have fewer parameters and the data are far
-from the tipping point. That is, when the data come from around the
-stable steady state, all the parametric models are approximately linear
-and approximately identical. Thus, it is intuitive that all model
-selection methods choose the simplest model. In a complex world, the
-result is suboptimal. But in a world that might contain tipping points,
-the result could be disastrous.
+Our results illustrate that model-choice approaches can be absolutely NA
+dynamics because they have fewer parameters and the data are far from
+the tipping point. That is, when the data come from around the stable
+steady state, all the parametric models are approximately linear and
+approximately identical. Thus, it is intuitive that all model selection
+methods choose the simplest model. In a complex world, the result is
+suboptimal. But in a world that might contain tipping points, the result
+could be disastrous.
 
@@
 have selected a prior that is both relatively uninformative (due to the
-broad priors placed on its parameters $\ell$ and $\sigma$ and simple
+broad priors placed on its parameters \(\ell\) and \(\sigma\) and simple
 (the choice of our covariance function, Eqns 12 and 13 ). In practice,

### 2014-12-05 manuscripts/submit.tex
tweak tex paths for submission engine
edit readmes and tidy up

--- 2014-11-20 manuscripts/manuscript.tex
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_2-1.jpeg}
+\includegraphics{manuscript-figure_2-1.jpeg}
 \caption{The inferred Gaussian process compared to the true process and
 maximum-likelihood estimated (MLE) process. We show the expected value
-for the function $f$ under each model. Plots show replicates drawn from
-the posterior distributions in order to convey uncertainty of the
+for the function \(f\) under each model. Plots show replicates drawn
+from the posterior distributions in order to convey uncertainty of the
 estimates. Note the MLE is a point estimate of parameters and so
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_4-1.jpeg}
+\includegraphics{manuscript-figure_4-1.jpeg}
 \caption{In the management context, GPDP outperforms approaches based on

### 2014-12-05 manuscripts/components/install.R

@@ -1,9 +1,0 @@
-#!/usr/bin/Rscript
-
-if(!require("nonparametricbayes")){
-  install.packages("devtools")
-  library("devtools")
-  install_github("reshape")
-  install_github("cboettig/pdg_control")
-  install.packages("..", repos =NULL, dependencies = c("Depends", "Imports", "Suggests"))
-}

### 2014-12-05 manuscripts/components/knit.R

@@ -1,32 +1,0 @@
-#!/usr/bin/Rscript
-args <- commandArgs(TRUE)
-
-if(!is.na(args[1])){
-  file <- args[1]
-} else if(!is.character(file)) {
-  file <- 'manuscript.Rmd'
-}
-
-
-## FIXME add second argument to handle figure types
-
-## Set the paths for cache and figure
-library(methods)
-library(knitr)
-basename <- gsub(".Rmd", "", file) # knitr:::knit_concord$get('infile')) 
-opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
-               cache.path = paste("components/cache/", basename, "/", sep=""))
-opts_chunk$set(cache = 2)
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, comment = NA, verbose = TRUE)
-
-# PDF-based figures
-opts_chunk$set(dev='pdf')
-
-# Better to use a temporary markdown file (best to not overwrite basename.md since that can be used to display on GitHub)
-# For now use basename, since pandoc function in knitr 1.5 doesn't respect output name from config (fixed in knitr 1.5.32)
-output <- paste(basename, ".md", sep="")
-
-## Knit
-knit(file, output)
-
-

### 2014-12-05 manuscripts/components/pandoc.R

@@ -1,15 +1,0 @@
-#!/usr/bin/Rscript
-
-## NOTE: Arguments to pandoc are set in the config file, pandoc_config.txt
-
-args <- commandArgs(TRUE)
-file <- args[1]
-library(knitr)
-
-## Call pandoc
-knitr::pandoc(file, format = "latex", ext = c("pdf", "tex"),
-              config = "components/pandoc_config.txt")
-
-# Note that path is relative to where the Rscript is called
-# (manuscripts/), not where the source script lives (manuscripts/components)
-

### 2014-12-05 manuscripts/igiveup.sh

@@ -1,8 +1,0 @@
-#!/bin/bash
-
-cp manuscript.pdf ~/Dropbox/GPDP/manuscript-CB-$(date +"%m-%d-%Y").pdf
-cp manuscript.tex ~/Dropbox/GPDP/manuscript-CB-$(date +"%m-%d-%Y").tex
-cp -r components/figure ~/Dropbox/GPDP/components/ 
-cp -r components/elsarticle.cls ~/Dropbox/GPDP/
-
-

### 2014-12-05 manuscripts/manuscript.Rmd
add explicit note about archived code, data, etc

@@
 we describe three parametric models for population dynamics and the Gaussian
-Process (GP)[^1] description of population dynamics.
+Process (GP)[^1] description of population dynamics. All computer code
+used here has been embedded in the manuscript sources (see @Xie_2013),
+and an implementation of the GPDP approach is provided as an accompanying R
+package.  Source code, R package and the CSV data files corresponding each
+figure are archived in the supplement [@Boettiger_2014].
 
@@
 the statistical model we use to approximate the population dynamics, and we use the
-term Gaussian Process Dynamic Programming [GPDP,], to refer to the _use_ of a
+term Gaussian Process Dynamic Programming (GPDP), to refer to the _use_ of a
 GP as the underlying process model when solving a Dynamic Programming equation.
@@
 S4, code in appendix). Since the posterior distributions differ
-substantially from the priors (Figure S4), most of the information in the posterior comes from the data rather than the
+substantially from the priors (Figure S4), most of the information 
+in the posterior comes from the data rather than the
 prior belief.

### 2014-12-05 manuscripts/manuscript.tex
add explicit note about archived code, data, etc

@@
   population dynamics, and we use the term Gaussian Process Dynamic
-  Programming {[}GPDP,{]}, to refer to the \emph{use} of a GP as the
+  Programming (GPDP), to refer to the \emph{use} of a GP as the
   underlying process model when solving a Dynamic Programming equation.
@@
   novel method we put forward here as GPDP.} description of population
-dynamics.
+dynamics. All computer code used here has been embedded in the
+manuscript sources (see Xie (2013)), and an implementation of the GPDP
+approach is provided as an accompanying R package. Source code, R
+package and the CSV data files corresponding each figure are archived in
+the supplement (Munch; 2014).
 
@@
 
+Munch; C. B. M. M. S. 2014. nonparametric-bayes v0.1.0. ZENODO.
+\url{http://dx.doi.org/10.5281/zenodo.12669}.
+
 Myers, R. A., N. J. Barrowman, J. A. Hutchings, and A. a Rosenberg.
@@
 
+Xie, Y. 2013. Dynamic documents with R and knitr. Chapman; Hall/CRCBoca
+Raton, Florida.
+
 \end{document}

### 2014-12-05 manuscripts/submit.tex
fix zenodo ref

@@
   novel method we put forward here as GPDP.} description of population
-dynamics.
+dynamics. All computer code used here has been embedded in the
+manuscript sources (see Xie (2013)), and an implementation of the GPDP
+approach is provided as an accompanying R package. Source code, R
+package and the CSV data files corresponding each figure are archived in
+the supplement (Boettiger et al. 2014).
 
@@
 
+Boettiger, C., M. Mangel, and S. Munch. 2014. nonparametric-bayes
+v0.1.0. ZENODO. \url{http://dx.doi.org/10.5281/zenodo.12669}.
+
 Brozović, N., and W. Schlenker. 2011. Optimal management of an ecosystem

### 2014-12-11 manuscripts/manuscript.Rmd
pdf figures

@@
 Curves show the expectations over the posterior step-ahead predictions based on each of the estimated models.
-Observe that all models are fitting the data reasonably well.)", fig.width=8, fig.height=6, cache=FALSE}
+Observe that all models are fitting the data reasonably well.)", fig.width=8, fig.height=6}
 step_ahead_posteriors <- function(x){
@@
 
-```{r deviances, dependson=c("dic_calc"), include=FALSE, echo=FALSE}
+```{r deviances, dependson=c("dic_calc"), include = FALSE, echo = FALSE}
 allen_deviance  <- - posterior.mode(pardist[,'deviance'])
@@
 The training data are also shown as black points.
-The GP is conditioned on (0,0), shown as a pseudo-data point.", fig.width=8, fig.height=6, cache=FALSE, dev="CairoJPEG", dev.args=c(quality=100)}
+The GP is conditioned on (0,0), shown as a pseudo-data point.", fig.width=8, fig.height=6}
 # subset data first, don't care about seeing dynamics for stock >> K
@@
 ```{r figure_3, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.
-Policies are shown in terms of target escapement, $S(t)$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979] Several policies show a numerical jitter due to the discretization of states in the dynamic programming algorithm; doubling the number of grid points did not qualitatively change the results.", fig.width=8, fig.height=6, cache=FALSE}
+Policies are shown in terms of target escapement, $S(t)$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979] Several policies show a numerical jitter due to the discretization of states in the dynamic programming algorithm; doubling the number of grid points did not qualitatively change the results.", fig.width=8, fig.height=6}
 policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
@@
 ```{r figure_4, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models.
-We show 100 replicate simulations of the stock dynamics (Eqn 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6, cache=FALSE, dev="CairoJPEG", dev.args=c(quality=100)}
+We show 100 replicate simulations of the stock dynamics (Eqn 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
 ggplot(sims_data) + 

### 2014-12-18 manuscripts/manuscript.Rmd
fix pandoc silent parsing error creating NA

@@
 \begin{equation}
-E = K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I}_n \right)^{-1} \mathbf{Y}_{\textrm{obs}}
+E = K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbf{I}_n \right)^{-1} \mathbf{Y}_{\textrm{obs}}
 \end{equation}
@@
 \begin{equation}
-C = K(X_p, X_p) - K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, X_p)
+C = K(X_p, X_p) - K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbf{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, X_p)
 \end{equation}
 
-Here $\mathbb{I}_n$ is the $n$ by $n$ identity matrix (i.e.
+Here $\mathbf{I}_n$ is the $n$ by $n$ identity matrix (i.e.
 a matrix with ones down the diagonal and zeros elsewhere) and $K$ is the ‘covariance kernel.’  The covariance kernel controls how much influence one observation has on another.
@@
 
-### Traditional model-choice approaches can be positively misleading.
+### Traditional model-choice approaches can be positively misleading
 
 Our results illustrate that model-choice approaches can be absolutely
-misleading – by providing support to models that cannot capture tipping point
+misleading -- by providing support to models that cannot capture tipping point
 dynamics because they have fewer parameters and the data are far from the tipping point.  That is, when the data 

### 2014-12-18 manuscripts/manuscript.tex
fix zenodo ref
pdf figures
fix pandoc silent parsing error creating NA

--- 2014-11-20 manuscripts/manuscript.tex
@@
 model sudden transitions between alternative stable states.
-
-\subsubsection{Parametric models}\label{parametric-models}
 
@@
 
-\subsubsection{Parametric models}\label{parametric-models}
+\subsubsection{Parametric Models}\label{parametric-models}
 
@@
 \begin{equation}
-C = K(X_p, X_p) - K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbb{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, X_p)
+C = K(X_p, X_p) - K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbf{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, X_p)
 \end{equation}
 
-Here $\mathbb{I}_n$ is the $n$ by $n$ identity matrix (i.e.~a matrix
+Here $\mathbf{I}_n$ is the $n$ by $n$ identity matrix (i.e.~a matrix
 with ones down the diagonal and zeros elsewhere) and $K$ is the
@@
 
-For the prior on $\ell$, $\alpha = $ 10 and $\beta = $ 10. The prior on
-$\sigma$, $\alpha = $ 5 and $\beta = $ 5.
+For the prior on $\ell$, $\alpha =$ 10 and $\beta =$ 10. The prior on
+$\sigma$, $\alpha =$ 5 and $\beta =$ 5.
 
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_2-1.jpeg}
+\includegraphics{components/figure/manuscript-figure_2-1.pdf}
 \caption{The inferred Gaussian process compared to the true process and
@@
 \centering
-\includegraphics{components/figure/manuscript-figure_4-1.jpeg}
+\includegraphics{components/figure/manuscript-figure_4-1.pdf}
 \caption{In the management context, GPDP outperforms approaches based on
@@
 \subsubsection{Traditional model-choice approaches can be positively
-misleading.}\label{traditional-model-choice-approaches-can-be-positively-misleading.}
+misleading}\label{traditional-model-choice-approaches-can-be-positively-misleading}
 
@@
 Environmental Economics and Management 63:304--320.
+
+Boettiger, C., M. Mangel, and S. Munch. 2014. nonparametric-bayes
+v0.1.0. ZENODO. \url{http://dx.doi.org/10.5281/zenodo.12669}.
 
@@
 
+Xie, Y. 2013. Dynamic Documents with R and knitr. Chapman; Hall/CRCBoca
+Raton, Florida.
+
 \end{document}

### 2014-12-27 manuscripts/manuscript.Rmd
formatting, rebuild pdfs

@@
       Santa Cruz, CA 95064, USA
-  - code: marc
-    address: |
-      Center for Stock Assessment Research, Department of Applied Math
-      and Statistics, University of California,
-      Mail Stop SOE-2, Santa Cruz, CA 95064, USA and Department of
-      Biology, University of Bergen, Bergen, Norway 9020
   - code: noaa
@@
 uncertainty can be meaningfully quantified [@Fischer2009].  Over the last
-four decades (beginning with
-@Clark1976, @Clark2009 and @Walters1978)
+four decades (beginning with @Clark1976, @Clark2009 and @Walters1978)
 dynamic optimization methods, particularly Stochastic Dynamic Programming
-(SDP), have become increasingly important as a means of
-understanding how to manage human intervention into natural systems.
-Simultaneously, there has been increasing recognition of the importance of
-multiple steady states or 'tipping points' [@Scheffer2001; @Scheffer2009;
-@Polasky2011] in ecological systems.
-
-We develop a novel approach to address these concerns in the context of
-fisheries; although the challenges and methods are germane to
-other problems of conservation or natural resource exploitation.  Economic value and ecological concern have made
-marine fisheries the crucible for much of the founding work on management
-under uncertainty [@Gordon1954; @Clark1976; @Clark2009; @Reed1979; @May1979; @Ludwig1982].
-
-Even if we know the proper deterministic description of the biological system, there is intrinsic stochasticity in biological dynamics, measurements, and
-implementation of policy [_e.g._ @Reed1979; @Clark1986; @Roughgarden1996;
-@Sethi2005].  We may also lack knowledge about the parameters of the biological dynamics [parametric uncertainty, _e.g._ @Ludwig1982; @Hilborn1997;
-@McAllister1998; @Schapaugh2013], or even not know which model is  proper description of the system 
-[structural uncertainty, _e.g._ @Williams2001; @Cressie2009;  @Athanassoglou2012].  Of these,
-the latter is generally the hardest to quantify. Typical approaches
-confront the data with a collection of models, assuming that the true dynamics (or reasonable approximation) is among the collection and then use model choice or model averaging to arrive at a conclusion
+(SDP), have become increasingly important as a means of understanding
+how to manage human intervention into natural systems.  Simultaneously,
+there has been increasing recognition of the importance of multiple steady
+states or 'tipping points' [@Scheffer2001; @Scheffer2009; @Polasky2011]
+in ecological systems.
+
+We develop a novel approach to address these concerns in the context
+of fisheries; although the challenges and methods are germane to other
+problems of conservation or natural resource exploitation.  Economic
+value and ecological concern have made marine fisheries the crucible for
+much of the founding work on management under uncertainty [@Gordon1954;
+@Clark1976; @Clark2009; @Reed1979; @May1979; @Ludwig1982].
+
+Even if we know the proper deterministic description of the biological
+system, there is intrinsic stochasticity in biological dynamics,
+measurements, and implementation of policy [_e.g._ @Reed1979; @Clark1986;
+@Roughgarden1996; @Sethi2005].  We may also lack knowledge about the
+parameters of the biological dynamics [parametric uncertainty, _e.g._
+@Ludwig1982; @Hilborn1997; @McAllister1998; @Schapaugh2013], or even
+not know which model is  proper description of the system [structural
+uncertainty, _e.g._ @Williams2001; @Cressie2009;  @Athanassoglou2012].
+Of these, the latter is generally the hardest to quantify. Typical
+approaches confront the data with a collection of models, assuming that
+the true dynamics (or reasonable approximation) is among the collection
+and then use model choice or model averaging to arrive at a conclusion
 [@Williams2001; @Athanassoglou2012; @Cressie2009].  Even setting aside
-other concerns (see @Cressie2009), these
-approaches are unable to describe uncertainty
-outside the observed data range.
-
-Structural  uncertainty is particularly insidious when we try to predict
-outside of the range of observed data [@Mangel2001] because we are extrapolating into unknown regions.
-In management applications, this extrapolation uncertainty is particularly important since (a) management involves considering actions that may move the
-system outside the range of observed behavior, and (b) the decision tools (optimal control theory, SDP) rely on both reasonable estimates of the expected
-outcomes and on the weights given to those outcomes
-[_e.g._ @Weitzman2013].   Thus characterizing uncertainty is as
-important as characterizing the expected outcome.
-
-Tipping points in ecological dynamics [@Scheffer2001;
-@Polasky2011] highlight this problem because 
-precise models are not available and data are
-limited such as around  high
-stock levels or an otherwise desirable state. With perfect information, 
-one would know just how far a system could
-be pushed before crossing the tipping point, and management would be
-simple.  But we face imperfect models and limited data and, with tipping points,  
-even small errors can have
-very large consequences, as we shall illustrate later.  Because intervention may be too late once a 
-tipping point has been crossed (but see @Hughes2013), management
-is often concerned with avoiding 
-tipping points before any data about them are available.
-
-The dual concerns of model uncertainty and incomplete data 
-create a substantial challenge to existing decision-theoretic approaches
-[@Brozovic2011].   We illustrate how  Stochastic Dynamic Programming (SDP) 
-[@Mangel1988; @Marescot2013] can be implemented using a
-Bayesian Non-Parametric (BNP) model of population dynamics [@Munch2005a].
-The BNP method has two distinct advantages. First, using a BNP model sidesteps the need for an accurate
-model-based description of the system dynamics.  Second, a BNP model
-better reflects uncertainty when extrapolating beyond the observed
-data.  This is crucial to providing  robust
-decision-making when the correct model is not known (as is almost always true).
-[We use _robust_ to characterize approaches that provide nearly optimal solutions
+other concerns (see @Cressie2009), these approaches are unable to describe
+uncertainty outside the observed data range.
+
+Structural  uncertainty is particularly insidious when we try to
+predict outside of the range of observed data [@Mangel2001] because we
+are extrapolating into unknown regions.  In management applications,
+this extrapolation uncertainty is particularly important since (a)
+management involves considering actions that may move the system
+outside the range of observed behavior, and (b) the decision tools
+(optimal control theory, SDP) rely on both reasonable estimates of the
+expected outcomes and on the weights given to those outcomes [_e.g._
+@Weitzman2013].   Thus characterizing uncertainty is as important as
+characterizing the expected outcome.
+
+Tipping points in ecological dynamics [@Scheffer2001; @Polasky2011]
+highlight this problem because precise models are not available and data
+are limited such as around  high stock levels or an otherwise desirable
+state. With perfect information, one would know just how far a system
+could be pushed before crossing the tipping point, and management
+would be simple.  But we face imperfect models and limited data and,
+with tipping points, even small errors can have very large consequences,
+as we shall illustrate later.  Because intervention may be too late once
+a tipping point has been crossed (but see @Hughes2013), management is
+often concerned with avoiding tipping points before any data about them
+are available.
+
+The dual concerns of model uncertainty and incomplete data create
+a substantial challenge to existing decision-theoretic approaches
+[@Brozovic2011].   We illustrate how  Stochastic Dynamic Programming
+(SDP) [@Mangel1988; @Marescot2013] can be implemented using a Bayesian
+Non-Parametric (BNP) model of population dynamics [@Munch2005a].  The BNP
+method has two distinct advantages. First, using a BNP model sidesteps
+the need for an accurate model-based description of the system dynamics.
+Second, a BNP model better reflects uncertainty when extrapolating beyond
+the observed data.  This is crucial to providing  robust decision-making
+when the correct model is not known (as is almost always true).  [We use
+_robust_ to characterize approaches that provide nearly optimal solutions
 without being sensitive to the choice of the (unknown) underlying model.]
@@
 
-For comparisons, we consider the performance of management based on  GPDP against
-management policies derived under several alternative parametric models [@Reed1979; @Ludwig1982; @Mangel1988].  Rather than compare models in terms
-of best fit to data, we compare model performance in the concrete terms
-of the decision-maker's objectives.
+For comparisons, we consider the performance of management based on
+GPDP against management policies derived under several alternative
+parametric models [@Reed1979; @Ludwig1982; @Mangel1988].  Rather than
+compare models in terms of best fit to data, we compare model performance
+in the concrete terms of the decision-maker's objectives.
 
@@
 
-Dynamic optimization  equires characterizing the dynamics of a state
+Dynamic optimization requires characterizing the dynamics of a state
 variable (or variables), a control action, and a value function.
-For simplicity, we consider only a single state variable.
-This is a best-case scenario for the parametric models because we
-simulate underlying dynamics from one of the three parametric models
-, whereas in the natural world we never know the "true" model.
-In addition, by choosing one-dimensional models with just a few parameters, we
-limit the chance that poor performance will be due to inability to
-estimate parameters accurately, something that becomes a more severe problem
-for higher-dimensional parametric models. Finally, the parametric models
-we consider are commonly used in modeling stock-recruitment
-dynamics or to model sudden transitions between alternative stable states.
-
-We let $X(t)$ denote the size (numbers or biomass) of the focal
-population at time $t$ and assume that in the absence of take its
-dynamics are:
+For simplicity, we consider only a single state variable.  This is
+a best-case scenario for the parametric models because we simulate
+underlying dynamics from one of the three parametric models, whereas
+in the natural world we never know the "true" model.  In addition, by
+choosing one-dimensional models with just a few parameters, we limit
+the chance that poor performance will be due to inability to estimate
+parameters accurately, something that becomes a more severe problem for
+higher-dimensional parametric models. Finally, the parametric models we
+consider are commonly used in modeling stock-recruitment dynamics or to
+model sudden transitions between alternative stable states.
+
+We let $X(t)$ denote the size (numbers or biomass) of the focal population
+at time $t$ and assume that in the absence of take its dynamics are:
 
@@
 Where Z(t) is log-normally distributed process stochasticity [@Reed1979]
-and $\mathbf{p}$ is a vector of parameters to be estimated from the data.
-We describe the three choices for $f(X(t),\mathbf{p})$ in the next section.  
-The control action is a harvest or take, $h(t)$, measured
+and $\mathbf{p}$ is a vector of parameters to be estimated from the
+data.  We describe the three choices for $f(X(t),\mathbf{p})$ in the
+next section.  The control action is a harvest or take, $h(t)$, measured
 in the same units as $X$, at time $t$.  Thus, in the presence of take,
-the population size on the right hand side of Eqn 1 is replaced by
-$S(t) = X(t) - h(t)$.
-
-To construct the value function, we consider a return when $X(t) = x(t)$ and
-harvest $h(t) = h$ denoted as the reward, $R(x(t), h)$. For example, if the
-return is the harvest at time $t$, then $R(x(t), h(t)) = \min(x(t), h(t))$. We
-assume that future harvests are discounted relative to current ones at
-a constant rate of discount $\delta$ and ask for the harvest policy
-that maximizes total discounted harvest between the current time $t$
-and a final time $T$. That is, we seek to maximize over choices of
-harvest $E [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$,
-where the state dynamics are given by Eqn 1 and $E$ denotes the
-expectation over future population states.
-
-In order to find that policy, we introduce the value function $V(x(t), t)$
-representing the total discounted catch from time $t$ onwards
-given that $X(t) = x(t)$.  This value function satisfies an equation of
-SDP  [@Mangel1988; @Clark2000; @Clark2009 ;@Mangel2014],
+the population size on the right hand side of Eqn 1 is replaced by $S(t)
+= X(t) - h(t)$.
+
+To construct the value function, we consider a return when $X(t) = x(t)$
+and harvest $h(t) = h$ denoted as the reward, $R(x(t), h)$. For example,
+if the return is the harvest at time $t$, then $R(x(t), h(t)) = \min(x(t),
+h(t))$. We assume that future harvests are discounted relative to current
+ones at a constant rate of discount $\delta$ and ask for the harvest
+policy that maximizes total discounted harvest between the current time
+$t$ and a final time $T$. That is, we seek to maximize over choices of
+harvest $E [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$, where the
+state dynamics are given by Eqn 1 and $E$ denotes the expectation over
+future population states.
+
+In order to find that policy, we introduce the value function $V(x(t),
+t)$ representing the total discounted catch from time $t$ onwards given
+that $X(t) = x(t)$.  This value function satisfies an equation of SDP
+[@Mangel1988; @Clark2000; @Clark2009 ;@Mangel2014],
 
@@
 $X(t+1)$, and maximized over all possible choices of harvest, $h(t)$.
-That is, at time $t$, when population size is $x(t)$ and harvest $h(t)$ is
-applied, the immediate return is $R(h(t), x(t))$. When
-the sole source of uncertainty is the process stochasticity term, $Z$,
-and thus the expectation in Eqn 2 is equivalent to taking expectations over
+That is, at time $t$, when population size is $x(t)$ and harvest $h(t)$
+is applied, the immediate return is $R(h(t), x(t))$. When the sole
+source of uncertainty is the process stochasticity term, $Z$, and thus
+the expectation in Eqn 2 is equivalent to taking expectations over
 $Z(t)$. That is
@@
 
-where the population size after the take is $x(t) - h(t)$, which is then
-translated into $X(t+1)$ by Eqn 1 (that is, we replace $X(t+1)$ by $Z(t) f(x(t)-h(t)|\mathbf{p})$).
-
-When the parameters governing the dynamics
-are also uncertain, we take the expectation over the
-posterior distribution for the parameters:
+where the population size after the take is $x(t) - h(t)$, which is
+then translated into $X(t+1)$ by Eqn 1 (that is, we replace $X(t+1)$
+by $Z(t) f(x(t)-h(t)|\mathbf{p})$).
+
+When the parameters governing the dynamics are also uncertain, we take
+the expectation over the posterior distribution for the parameters:
 
@@
 
-When the underlying population dynamics are unknown (the case of structural uncertainty), the function $f$ itself is uncertain and the expectation for the next state includes uncertainty in $f$ as well.
-That is
+When the underlying population dynamics are unknown (the case of
+structural uncertainty), the function $f$ itself is uncertain and
+the expectation for the next state includes uncertainty in $f$ as
+well. That is
 
@@
 
-We consider three candidate parametric models for the population dynamics: The Ricker model, the Allen model [@Allen2005], and the Myers
-model [@Myers1995], Eqns \eqref{ricker}-\eqref{myers}.
-In all three, we let $K$ denote the carrying capacity and $r$ the maximum per capita growth rate.
-The Ricker model has two parameters and the right hand side of Eqn 1 is
+We consider three candidate parametric models for the population dynamics:
+The Ricker model, the Allen model [@Allen2005], and the Myers model
+[@Myers1995], Eqns \eqref{ricker}-\eqref{myers}. In all three, we let
+$K$ denote the carrying capacity and $r$ the maximum per capita growth
+rate. The Ricker model has two parameters and the right hand side of
+Eqn 1 is
 
@@
 
-The Myers model also has three parameters
+The Myers model also has three parameters:
 
@@
 
-The Ricker model does not lead to multiple steady states.
-The Allen model resembles the Ricker
-dynamics with an added Allee effect parameter [@Courchamp2008], below
-which the population cannot persist.  The Myers model also has three
-parameters and contains an Allee threshold, but unlike the Ricker model
-saturates  at high population size.  The multiplicative log-normal stochasticity in Eqn 1
-introduces one additional parameter $\sigma$ that must be estimated.
+The Ricker model does not lead to multiple steady states.  The Allen
+model resembles the Ricker dynamics with an added Allee effect parameter
+[@Courchamp2008], below which the population cannot persist.  The Myers
+model also has three parameters and contains an Allee threshold,
+but unlike the Ricker model saturates  at high population size.
+The multiplicative log-normal stochasticity in Eqn 1 introduces one
+additional parameter $\sigma$ that must be estimated.
 
@@
 The Allen model is thus the  state of nature  and is expected to provide
-the best-case scenario. The Ricker model is a reasonable approximation of these dynamics far from the Allee threshold (but lacks threshold dynamics), while the Myers model shares
-the essential feature of a threshold but differs in structure from the Allen model.
-Throughout, we refer to the "True" model when the underlying parameters
-_are known without error_, and refer to the "Allen" model 
-when these parameters have been estimated from the sample data.
+the best-case scenario. The Ricker model is a reasonable approximation
+of these dynamics far from the Allee threshold (but lacks threshold
+dynamics), while the Myers model shares the essential feature of a
+threshold but differs in structure from the Allen model.  Throughout,
+we refer to the "True" model when the underlying parameters _are known
+without error_, and refer to the "Allen" model when these parameters
+have been estimated from the sample data.
 
@@
 
-We inferred posterior distributions for the parameters of each model by Gibbs sampling (@Gelman2003 implemented in R
-[@RTeam] using `jags`, [@R2jags]).  We choose uniform priors
-for all parameters of the parametric models (See appendix Tables S1-S3; R code provided). We
-show one-step-ahead predictions of these model fits in Figure 1.
-We tested each chain for Gelman-Rubin convergence and results were
-robust to longer runs.  For each simulation we also applied several
-commonly used model selection criteria (AIC, BIC, DIC, see @Burnham2002)
-to identify the best fitting model.
-
-Additionally, we compute the maximum likelihood estimate (MLE, as we will refer to 
-this model in the figures) of the parameters for the (structurally correct) Allen model.
-Comparing this to using the posterior distribution of parameters inferred from MCMC for
-the same model gives some indication of the importance of this uncertainty in the 
-dynamic programming.
+We inferred posterior distributions for the parameters of each model by
+Gibbs sampling (@Gelman2003 implemented in R [@RTeam] using `jags`,
+[@R2jags]).  We choose uniform priors for all parameters of the
+parametric models (See appendix Tables S1-S3; R code provided). We show
+one-step-ahead predictions of these model fits in Figure 1.  We tested
+each chain for Gelman-Rubin convergence and results were robust to
+longer runs.  For each simulation we also applied several commonly used
+model selection criteria (AIC, BIC, DIC, see @Burnham2002) to identify
+the best fitting model.
+
+Additionally, we compute the maximum likelihood estimate (MLE, as
+we will refer to this model in the figures) of the parameters for
+the (structurally correct) Allen model.  Comparing this to using the
+posterior distribution of parameters inferred from MCMC for the same
+model gives some indication of the importance of this uncertainty in
+the dynamic programming.
 
@@
 
-The core difference for our purpose between the estimated GP and the estimated parametric
-models is that the estimated GP model is defined explicitly in reference to the observed data.
-As a result, uncertainty arises in the GP model not only from uncertainty in the parameters,
-but is also increases in regions farther from the observed states, such as low population
-sizes in the example illustrated here.  The estimated parametric models, by contrast, are
-completely specified by the parameters.
+The core difference for our purpose between the estimated GP and
+the estimated parametric models is that the estimated GP model is
+defined explicitly in reference to the observed data.  As a result,
+uncertainty arises in the GP model not only from uncertainty in the
+parameters, but is also increases in regions farther from the observed
+states, such as low population sizes in the example illustrated here.
+The estimated parametric models, by contrast, are completely specified
+by the parameters.
 
@@
 
-where $\varepsilon$ are IID normal random variables with zero-mean and variance $\sigma^2$. Note that we have chosen to
-assume additive stochasticity.  While we could consider log-normal stochasticity
-as in the parametric models, we make this choice to emphasize that the
-Gaussian process approach need not have structurally correct stochasticity to be effective.
-
-In order to make predictions, we update the GP based on the observed set of transitions.
-To do so, we collect the time series of observed states into a vector of "current" states, $\mathbf{X}_{\textrm{obs}} = \{X(1), \dots, X(T-1)\}$ and a vector of "next" states $\mathbf{Y}_{\textrm{obs}} = \{X(2),\dots,X(T)\}$ where $T$ is the time of the final observation.
-Conditional on these observations, the predicted next state, $g(X_p)$, for any given "current" state, $X_p$ follows a normal distribution with mean $E$ and variance $C$ determined using the standard rules for conditioning in multivariate normals, i.e.
+where $\varepsilon$ are IID normal random variables with zero-mean
+and variance $\sigma^2$. Note that we have chosen to assume additive
+stochasticity.  While we could consider log-normal stochasticity as in
+the parametric models, we make this choice to emphasize that the Gaussian
+process approach need not have structurally correct stochasticity to
+be effective.
+
+In order to make predictions, we update the GP based on the observed set
+of transitions.  To do so, we collect the time series of observed states
+into a vector of "current" states, $\mathbf{X}_{\textrm{obs}} = \{X(1),
+\dots, X(T-1)\}$ and a vector of "next" states $\mathbf{Y}_{\textrm{obs}}
+= \{X(2),\dots,X(T)\}$ where $T$ is the time of the final observation.
+Conditional on these observations, the predicted next state, $g(X_p)$,
+for any given "current" state, $X_p$ follows a normal distribution
+with mean $E$ and variance $C$ determined using the standard rules for
+conditioning in multivariate normals, i.e.
 
@@
 
-Here $\mathbf{I}_n$ is the $n$ by $n$ identity matrix (i.e.
-a matrix with ones down the diagonal and zeros elsewhere) and $K$ is the ‘covariance kernel.’  The covariance kernel controls how much influence one observation has on another.
-In the present application we use the squared-exponential kernel which, when evaluated over a pair of vectors, say $\mathbf{x}$ and $\mathbf{y}$,  generates a covariance matrix whose $i,j$th element is given by
+Here $\mathbf{I}_n$ is the $n$ by $n$ identity matrix (i.e. a matrix with
+ones down the diagonal and zeros elsewhere) and $K$ is the ‘covariance
+kernel.’  The covariance kernel controls how much influence one
+observation has on another.  In the present application we use the
+squared-exponential kernel which, when evaluated over a pair of vectors,
+say $\mathbf{x}$ and $\mathbf{y}$,  generates a covariance matrix whose
+$i,j$th element is given by
 
@@
 
-so that $\ell$ gives the characteristic length-scale over which correlation between
-two observations decays. See @Rasmussen2006 for other choices of covariance kernels and their properties.
-Note that this simple formulation assumes a prior mean of zero. For the parameters we use inverse Gamma priors on both the length-scale $\ell$ and $\sigma$, thus for example
+so that $\ell$ gives the characteristic length-scale over which
+correlation between two observations decays. See @Rasmussen2006 for
+other choices of covariance kernels and their properties.  Note that this
+simple formulation assumes a prior mean of zero. For the parameters we
+use inverse Gamma priors on both the length-scale $\ell$ and $\sigma$,
+thus for example
 
@@
 
-We  derive the harvest policy from the
-estimated GP by inserting it into a SDP algorithm. Given the GP posteriors, we construct the transition
+We  derive the harvest policy from the estimated GP by inserting it into
+a SDP algorithm. Given the GP posteriors, we construct the transition
 matrix representing the probability of going to each state $X(t+1)$
@@
 matrix, we use the same value iteration algorithm as in the parametric
-case to determine the optimal policy. In doing so, 
-the uncertainty in the future state under the GP, $X(t+1)$,
-includes both process uncertainty (based on the estimation of $\sigma$) 
-and structural uncertainty of the posterior collection of curves.
+case to determine the optimal policy. In doing so, the uncertainty in the
+future state under the GP, $X(t+1)$, includes both process uncertainty
+(based on the estimation of $\sigma$) and structural uncertainty of the
+posterior collection of curves.
 
@@
 
-To ensure our results are robust to the choice of parameters, we
-will consider 96 different scenarios.  To help better
-understand the process, we first describe in detail the
-results of a single scenario.
+To ensure our results are robust to the choice of parameters, we will
+consider 96 different scenarios.  To help better understand the process,
+we first describe in detail the results of a single scenario.
 
@@
 
-All of the models fit the observed data rather closely and with relatively small uncertainty.
-In Figure 1, we show the posterior predictive curves.
-The  training data of stock sizes observed over time are points, overlaid with the step-ahead predictions of each estimated model using the parameters sampled from their posterior distributions.
-Compared to the true model most estimates appear to over-fit, predicting patterns that are actually due purely to stochasticity.
-Model selection criteria (Table 1) penalize more complex models and show a preference for the simpler Ricker model over the  models with alternative stable states (Allen and Myers).
-Supplement provides details on the model estimates.
+All of the models fit the observed data rather closely and with relatively
+small uncertainty. In Figure 1, we show the posterior predictive curves.
+The training data of stock sizes observed over time are points,
+overlaid with the step-ahead predictions of each estimated model using
+the parameters sampled from their posterior distributions. Compared to
+the true model most estimates appear to over-fit, predicting patterns
+that are actually due purely to stochasticity.  Model selection criteria
+(Table 1) penalize more complex models and show a preference for the
+simpler Ricker model over the  models with alternative stable states
+(Allen and Myers). Supplement provides details on the model estimates.
 
@@
 
-We show the  mean inferred population dynamics of each model relative to the
-true model used to generate the data in Figure 2, predicting
-the relationship between observed population size (x-axis) to the population
-size after recruitment the following year.  
-In addition to the raw data, the GP is conditioned
-on going through the point 0,0 without error.
-All parametric models also make this assumption.
-Conditioning on (0,0) is equivalent to making the assumption that the population is closed, so that once it hits 0 it stays at 0, despite the lack of any data in the observed sequence to justify this.
-This assumption illustrates how the GP can capture common-sense biology without having to assume more explicit details about the dynamics at low population numbers that have never been observed.
-If the population were not closed, one could repeat the entire analysis without this assumption.
-Unlike parametric models, the GP corresponds to a
-distribution of curves, of which this plot only shows the means. Uncertainty in the parameters of the GP (not shown) further widens the band of possible population sizes.
-In Figure S1 (see supplement), we show the performance of the models outside the observed training data.
+We show the  mean inferred population dynamics of each model relative
+to the true model used to generate the data in Figure 2, predicting the
+relationship between observed population size (x-axis) to the population
+size after recruitment the following year.  In addition to the raw data,
+the GP is conditioned on going through the point 0,0 without error.
+All parametric models also make this assumption. Conditioning on (0,0)
+is equivalent to making the assumption that the population is closed,
+so that once it hits 0 it stays at 0, despite the lack of any data in
+the observed sequence to justify this.  This assumption illustrates how
+the GP can capture common-sense biology without having to assume more
+explicit details about the dynamics at low population numbers that have
+never been observed. If the population were not closed, one could repeat
+the entire analysis without this assumption.  Unlike parametric models,
+the GP corresponds to a distribution of curves, of which this plot only
+shows the means. Uncertainty in the parameters of the GP (not shown)
+further widens the band of possible population sizes.  In Figure S1
+(see supplement), we show the performance of the models outside the
+observed training data.
 
@@
 
-Despite the similarities in model fits to the observed data, the
-policies inferred under each model differ widely (Figure 3).
-Policies are shown in terms of target escapement, $S(t) = x_t - h$.  Under models
+Despite the similarities in model fits to the observed data, the policies
+inferred under each model differ widely (Figure 3).  Policies are
+shown in terms of target escapement, $S(t) = x_t - h$.  Under models
 such as this a constant escapement policy is expected to be optimal
 [@Reed1979], whereby population levels below a certain size $S$ are
-unharvested, while above that size the harvest strategy aims to return the
-population to $S$.
-Whenever a model predicts that the population will not persist below a certain threshold, the optimal solution is to harvest the entire population immediately, resulting in an escapement $S=0$, as seen in the true (correct form, exact parameters) model, the Allen model (correct form, estimated parameters) and the GP.
-Only the structurally correct model (Allen model) and the GP produce
-policies close to the true optimum policy.
+unharvested, while above that size the harvest strategy aims to return
+the population to $S$. Whenever a model predicts that the population
+will not persist below a certain threshold, the optimal solution is to
+harvest the entire population immediately, resulting in an escapement
+$S=0$, as seen in the true (correct form, exact parameters) model, the
+Allen model (correct form, estimated parameters) and the GP. Only the
+structurally correct model (Allen model) and the GP produce policies
+close to the true optimum policy.
 
@@
 In Figure 4, we show the consequences of managing 100 replicate
-realizations of the simulated fishery under policies derived from each model.  The structurally
-correct model under-harvests, leaving the stock to vary around its
-unfished optimum.  The structurally incorrect Ricker model over-harvests
-the population past the tipping point consistently, resulting in the
-immediate crash of the stock and thus leads to minimal long-term catch.
-
-The results across replicate stochastic simulations are most
-easily compared by using the relative differences in net present
-value realized by each of the model (Figure 5).  Although not perfect, the GPDP
-consistently realizes a value close to the optimal solution,
-and avoids ever driving the system across the tipping point,
-which results in the near-zero value cases in the parametric models.
+realizations of the simulated fishery under policies derived from each
+model.  The structurally correct model under-harvests, leaving the stock
+to vary around its unfished optimum.  The structurally incorrect Ricker
+model over-harvests the population past the tipping point consistently,
+resulting in the immediate crash of the stock and thus leads to minimal
+long-term catch.
+
+The results across replicate stochastic simulations are most easily
+compared by using the relative differences in net present value
+realized by each of the model (Figure 5).  Although not perfect, the
+GPDP consistently realizes a value close to the optimal solution, and
+avoids ever driving the system across the tipping point, which results
+in the near-zero value cases in the parametric models.
 
@@
 These results are not sensitive to the modeling details of the simulation.
-The GPDP estimate remains very close to the optimal solution (obtained
-by knowing the true model) across changes to the training simulation,
-scale of stochasticity, parameters or structure of the underlying model. In the Supplement, we consider
-both a Latin hypercube approach and a more focused investigation of the effects of the relative 
-distance to the Allee threshold and the variance of process stochasticity.
-
-The GPDP is only weakly influenced by increasing stochasticity or increasing
-Allee effects over much of the range (Figure S2). Larger $\sigma$ or higher Allee
-levels make even the optimal solution without any model or parameter uncertainty
-unable to harvest the population effectively (e.g. the stochasticity is large enough to violate the
+The GPDP estimate remains very close to the optimal solution (obtained by
+knowing the true model) across changes to the training simulation, scale
+of stochasticity, parameters or structure of the underlying model. In
+the Supplement, we consider both a Latin hypercube approach and a more
+focused investigation of the effects of the relative distance to the
+Allee threshold and the variance of process stochasticity.
+
+The GPDP is only weakly influenced by increasing stochasticity or
+increasing Allee effects over much of the range (Figure S2). Larger
+$\sigma$ or higher Allee levels make even the optimal solution without
+any model or parameter uncertainty unable to harvest the population
+effectively (e.g. the stochasticity is large enough to violate the
 self-sustaining criterion of @Reed1979).

### 2014-12-27 manuscripts/manuscript.tex
formatting, rebuild pdfs

@@
 95064, USA}    
-    \address[marc]{Center for Stock Assessment Research, Department of Applied Math and
-Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA
-95064, USA and Department of Biology, University of Bergen, Bergen,
-Norway 9020}    
     \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric
@@
 management would be simple. But we face imperfect models and limited
-data and, with tipping points,\\even small errors can have very large
+data and, with tipping points, even small errors can have very large
 consequences, as we shall illustrate later. Because intervention may be
@@
 
-Dynamic optimization equires characterizing the dynamics of a state
+Dynamic optimization requires characterizing the dynamics of a state
 variable (or variables), a control action, and a value function. For
@@
 best-case scenario for the parametric models because we simulate
-underlying dynamics from one of the three parametric models , whereas in
+underlying dynamics from one of the three parametric models, whereas in
 the natural world we never know the ``true'' model. In addition, by
@@
 data. We describe the three choices for $f(X(t),\mathbf{p})$ in the next
-section.\\The control action is a harvest or take, $h(t)$, measured in
+section. The control action is a harvest or take, $h(t)$, measured in
 the same units as $X$, at time $t$. Thus, in the presence of take, the
@@
 relationship between observed population size (x-axis) to the population
-size after recruitment the following year.\\In addition to the raw data,
+size after recruitment the following year. In addition to the raw data,
 the GP is conditioned on going through the point 0,0 without error. All

### 2015-01-20 manuscripts/supplement.Rmd

@@
 and explore the analyses here by providing a copy of the computational environment
-we have used, with all software installed. Docker sofware (see [docker.com](http://www.docker.com))
+we have used, with all software installed. Docker software (see [docker.com](http://www.docker.com))
 runs on most platforms as well as cloud servers. Use the command:
@@
 This section provides figures and tables showing the traces from each of the MCMC runs used to estimate the parameters of the models presented, along with the resulting posterior distributions for each parameter.
-Priors usually appear completely falt when shown against the posteriors, but are summarized by tables the parameters of their corresponding distributions for each case.
+Priors usually appear completely flat when shown against the posteriors, but are summarized by tables the parameters of their corresponding distributions for each case.
 
